### 2020-08-21 Chapter_0/04_Manuscript/99_PermeabilityModelAICs.tex
first commit

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Thu Jun  4 15:54:16 2020
\begin{tabular}{lrrrr}
  \toprule
Covariates & AIC & DeltaAIC & Weight & LogLik \\ 
  \midrule
cos(ta) + log(sl) + W + T + DTW + HB5 + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 & 90094.79 & 3.38 & 0.00 & -45034.39 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\ 
  cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\ 
  cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\ 
  cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\ 
  cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\ 
  cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\ 
  cos(ta) + log(sl) + W + T + HB5 & 90123.71 & 32.31 & 0.00 & -45050.85 \\ 
  cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\ 
  cos(ta) + log(sl) + W + T + S & 90129.36 & 37.96 & 0.00 & -45053.68 \\ 
  cos(ta) + log(sl) + W + T + DTR & 90130.11 & 38.71 & 0.00 & -45054.05 \\ 
  cos(ta) + log(sl) + W + T + P & 90131.75 & 40.35 & 0.00 & -45054.88 \\ 
  cos(ta) + log(sl) + W + T + RC & 90133.53 & 42.13 & 0.00 & -45053.77 \\ 
  cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\ 
  cos(ta) + log(sl) + W + HB5 & 90149.30 & 57.90 & 0.00 & -45065.65 \\ 
  cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\ 
  cos(ta) + log(sl) + W & 90153.70 & 62.30 & 0.00 & -45069.85 \\ 
  cos(ta) + log(sl) + W + DTR & 90156.32 & 64.92 & 0.00 & -45069.16 \\ 
  cos(ta) + log(sl) + W + P & 90157.23 & 65.83 & 0.00 & -45069.62 \\ 
  cos(ta) + log(sl) + W + RC & 90159.32 & 67.91 & 0.00 & -45068.66 \\ 
  cos(ta) + log(sl) + S & 90161.07 & 69.67 & 0.00 & -45073.54 \\ 
  cos(ta) + log(sl) + DTW & 90262.03 & 170.63 & 0.00 & -45124.01 \\ 
  cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\ 
  cos(ta) + log(sl) + HB5 & 90307.55 & 216.15 & 0.00 & -45146.78 \\ 
  cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\ 
  cos(ta) + log(sl) + P & 90316.62 & 225.22 & 0.00 & -45151.31 \\ 
  cos(ta) + log(sl) + RC & 90318.46 & 227.06 & 0.00 & -45150.23 \\ 
   \bottomrule
\end{tabular}

### 2020-08-21 Chapter_0/04_Manuscript/99_RoadsDescription.tex
first commit

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Tue Apr 21 10:00:46 2020
\begin{tabular}{lll}
  \toprule
Group & Subgroup & Description \\ 
  \midrule
Roads & motorway & A restricted access major divided highway, normally with 2 or more running lanes plus emergency hard shoulder. Equivalent to the Freeway, Autobahn, etc.. \\ 
  Roads & trunk & The most important roads in a country's system that aren't motorways. (Need not necessarily be a divided highway.) \\ 
  Roads & primary & The next most important roads in a country's system. (Often link larger towns.) \\ 
  Roads & secondary & The next most important roads in a country's system. (Often link towns.) \\ 
  Roads & tertiary & The next most important roads in a country's system. (Often link smaller towns and villages) \\ 
  Roads & unclassified & The least important through roads in a country's system  i.e. minor roads of a lower classification than tertiary, but which serve a purpose other than access to properties. (Often link villages and hamlets.) \\ 
  Roads & residential & Roads which serve as an access to housing, without function of connecting settlements. Often lined with housing. \\ 
  Roads & service & For access roads to, or within an industrial estate, camp site, business park, car park etc. Can be used in conjunction with service=* to indicate the type of usage and with access=* to indicate who can use it and in what circumstances. \\ 
  Link roads & motorway\_link & The link roads (sliproads/ramps) leading to/from a motorway from/to a motorway or lower class highway. Normally with the same motorway restrictions. \\ 
  Link roads & trunk\_link & The link roads (sliproads/ramps) leading to/from a trunk road from/to a trunk road or lower class highway. \\ 
  Link roads & primary\_link & The link roads (sliproads/ramps) leading to/from a primary road from/to a primary road or lower class highway. \\ 
  Link roads & secondary\_link & The link roads (sliproads/ramps) leading to/from a secondary road from/to a secondary road or lower class highway. \\ 
  Link roads & tertiary\_link & The link roads (sliproads/ramps) leading to/from a tertiary road from/to a tertiary road or lower class highway. \\ 
  Special road types & living\_street & For living streets, which are residential streets where pedestrians have legal priority over cars, speeds are kept very low and where children are allowed to play on the street. \\ 
  Special road types & pedestrian & For roads used mainly/exclusively for pedestrians in shopping and some residential areas which may allow access by motorised vehicles only for very limited periods of the day. To create a 'square' or 'plaza' create a closed way and tag as pedestrian and also with area=yes. \\ 
  Special road types & track & Roads for mostly agricultural or forestry uses. To describe the quality of a track, see tracktype=*. Note: Although tracks are often rough with unpaved surfaces, this tag is not describing the quality of a road but its use. Consequently, if you want to tag a general use road, use one of the general highway values instead of track. \\ 
  Special road types & bus\_guideway & A busway where the vehicle guided by the way (though not a railway) and is not suitable for other traffic. Please note: this is not a normal bus lane, use access=no, psv=yes instead! \\ 
  Special road types & escape & For runaway truck ramps, runaway truck lanes, emergency escape ramps, or truck arrester beds. It enables vehicles with braking failure to safely stop. \\ 
  Special road types & raceway & A course or track for (motor) racing \\ 
  Special road types & road & A road/way/street/motorway/etc. of unknown type. It can stand for anything ranging from a footpath to a motorway. This tag should only be used temporarily until the road/way/etc. has been properly surveyed. If you do know the road type, do not use this value, instead use one of the more specific highway=* values. \\ 
  Paths & footway & For designated footpaths; i.e., mainly/exclusively for pedestrians. This includes walking tracks and gravel paths. If bicycles are allowed as well, you can indicate this by adding a bicycle=yestag. Should not be used for paths where the primary or intended usage is unknown. Use highway=pedestrian for pedestrianised roads in shopping or residential areas and highway=track if it is usable by agricultural or similar vehicles. \\ 
  Paths & bridleway & For horse riders. Equivalent to highway=path + horse=designated. \\ 
  Paths & steps & For flights of steps (stairs) on footways. Use with step\_count=* to indicate the number of steps \\ 
  Paths & path & A non-specific path. Use highway=footway for paths mainly for walkers, highway=cycleway for one also usable by cyclists, highway=bridlewayfor ones available to horse riders as well as walkers and highway=track for ones which is passable by agriculture or similar vehicles. \\ 
  Cycleway & lane & A lane is a route that lies within the roadway \\ 
  Cycleway & opposite & Used on ways with oneway=yes where it is legally permitted to cycle in both directions. Used together with oneway:bicycle=no. \\ 
  Cycleway & opposite\_lane & Used on ways with oneway=yes that have a cycling lane going the opposite direction of normal traffic flow (a "contraflow" lane). Used together with oneway:bicycle=no. \\ 
  Cycleway & track & A track provides a route that is separated from traffic. In the United States, this term is often used to refer to bike lanes that are separated from lanes for cars by pavement buffers, bollards, parking lanes, and curbs. Note that a cycle track may alternatively be drawn as a separate way next to the road which is tagged as highway=cycleway. \\ 
  Cycleway & opposite\_track & Used on ways with oneway=yes that have a cycling track going the opposite direction of normal traffic flow. Used together with oneway:bicycle=no. \\ 
  Cycleway & share\_busway & There is a bus lane that cyclists are permitted to use. \\ 
  Cycleway & opposite\_share\_busway & Used on ways with oneway=yes that have a bus lane that cyclists are also permitted to use, and which go in the opposite direction to normal traffic flow (a "contraflow" bus lane). Used together with oneway:bicycle=no. \\ 
  Cycleway & shared\_lane & Cyclists share a lane with motor vehicles, but there are markings indicating that they should share the lane with motorists. In some places these markings are known as "sharrows" ('sharing arrows') and this is the tag to use for those. \\ 
  Busway & lane & Bus lane on both sides of the road. \\ 
   \bottomrule
\end{tabular}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript.tex
first commit

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[nomarkers, nolists]{endfloat}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Bound within Boundaries: How Well Do Protected Areas Match Movement
Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S,*} \and
  Dominik M. Behr\textsuperscript{1,2,*} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared first authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} dispersal, habitat selection, integrated step selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, least-cost corridors, Lycaon pictus, permeability surface,
protected areas, wildlife management

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
\noindent 1. Conserving and managing large portions of land to connect wildlife
reserves is increasingly used to maintain and restore connectivity among
wildlife populations. Boundaries of such conservation areas are often determined
based on expert opinion and socio-political constraints, yet the extent to which
they match species' movement corridors is rarely examined. This is mainly due to
a lack of data, particularly on wide-ranging movement behavior such as
dispersal. Nevertheless, empirically assessing the adequacy of protected areas
is key for the implementation of targeted management actions and efficient use
of limited conservation funds.

\noindent 2. Between 2011 and 2019, we collected high-resolution GPS movement
data on 16 dispersing African wild dog (\textit{Lycaon pictus}) coalitions from
a free-ranging population in the Kavango-Zambezi Transfrontier Conservation Area
(KAZA-TFCA). Spanning five countries and 520'000 km\textsuperscript{2} the
KAZA-TFCA is the world's largest transboundary conservation area and a prime
example for international conservation efforts. We used integrated step
selection analysis to estimate habitat preferences of dispersers and to create a
permeability surface for the entire KAZA-TFCA. We compared landscape
permeability across different regions within the KAZA-TFCA as well as outside
its boundaries. Lastly, we calculated least-cost paths and corridors to verify
that major movement routes were adequately encompassed within the KAZA-TFCA.

\noindent 3. Permeability within the boundaries of the KAZA-TFCA was more than
double compared to areas outside it. Furthermore, we observed a five-fold
permeability difference among the five KAZA-TFCA countries. We further showed
that major movement corridors of wild dogs run within the KAZA-TFCA, although
some minor routes remained outside formally protected areas.

\noindent 4. Differences in permeability were mainly caused by different degrees
of human activities across regions, which hampered dispersal. Rivers, swamps  or
open water also limited dispersal, while other landscape features had a limited
effect.

\noindent 5. \textit{Synthesis and Applications}: In this study, we showed how
pertinent dispersal data of a highly mobile species can be used to empirically
evaluate the adequacy of already-existing or planned protected areas.
Furthermore, observed regional differences in landscape permeability highlight
the need for a coordinated effort towards maintaining or restoring connectivity,
especially where international effort is required.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}
Connectivity among subpopulations is a crucial pre-requisite for many species to
thrive and persist \citep{Fahrig.2003}. Accordingly, preserving and protecting
movement corridors between wildlife reserves has become an utmost task for
conservation management \citep{Doerr.2011, Rudnick.2012}, resulting in an
ever-growing number of large and often transboundary protected areas. While
boundaries of such areas are often drawn according to expert opinion and
socio-political needs, subjective assessments have revealed deficiencies in the
past \citep{Clevenger.2002, Pullinger.2010}. Thus, an empirical assessment of
the adequacy of already-existing or planned protected areas using pertinent
animal movement data is paramount for targeted use of valuable and scarce
conservation funds \citep{Pullinger.2010}.

In recent years, a growing body of research has used animal relocation data to
identify movement corridors and assess connectivity at large scales (e.g.
\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). Identification of
potential corridors typically relies on the estimation of permeability surfaces,
which return the ease or willingness at which the focal species traverses a
specific landscape \citep{Sawyer.2011}. Such surfaces are created based on a
species' habitat preferences, which can be quantified using a suite of selection
functions \citep{Zeller.2012}. Specifically, habitat preferences are estimated
by comparing spatial covariates (e.g. environmental and anthropogenic) at
locations visited by the animal to the same spatial covariates at randomly
selected locations \citep{Zeller.2012}. Importantly, selection functions rely on
adequate landscape and relocation data that are representative of the process
being studied \citep{Diniz.2020}. For instance, relocation data collected on
dispersing individuals has been shown to outperform data collected on resident
individuals in the detection of large-scale movement corridors
\citep{Elliot.2014, Diniz.2020}. Nevertheless, dispersal data is inherently
difficult to collect and remains scarce in the connectivity literature
\citep{Vasudev.2015}. As such, most permeability surfaces upon which movement
corridors are identified are created using relocation data collected on resident
individuals. This introduces severe biases and substantially reduces the power
to reveal meaningful movement corridors, for dispersing individuals have
different needs and drives compared to resident individuals \citep{Elliot.2014,
Cozzi.2020}. Such biases have limited our ability to meaningfully assess the
effectiveness of protected areas in securing connectivity for their protected
species.

One initiative that aims at restoring and enhancing connectivity across large
scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), which
constitutes the world's largest transfrontier conservation area, spanning over
520'000 km\textsuperscript{2} and five countries (\url{www.kavangozambezi.org}).
While the KAZA-TFCA was originally set to facilitate movements of African
elephants (\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key
to the conservation of other wide-ranging species such as African wild dogs
(\textit{Lycaon pictus}; \citealp{Woodroffe.2012, Cozzi.2020}), lions
(\textit{Panthera leo}; \citealp{Elliot.2014, Cushman.2018}), and cheetahs
(\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
have attempted to assess the adequacy of the KAZA-TFCA using relevant global
positioning system (GPS) relocation data of its protected species at the
appropriate spatial scale \citep{Elliot.2014, Tshipa.2017}. Thus, how well the
boundaries of the KAZA-TFCA reflect natural movement patterns and dispersal
corridors of its most mobile protected species is virtually unknown.

Across the KAZA-TFCA, the African wild dog (\textit{Lycaon pictus}) represents a
highly mobile and endangered flagship species for conservation efforts. Once
widespread across the entire Sub-Saharan continent, wild dogs have been widely
extirpated through human persecution, habitat destruction, and disease outbreaks
\citep{Woodroffe.2012}. As a result, the species has become one of Africa's most
endangered large carnivores, and currently only survives in small, spatially
scattered subpopulations \citep{Woodroffe.2012}. Within these subpopulations,
wild dogs form cooperative breeding packs of up to thirty individuals
\citep{Creel.2002}, whose social structure is strongly governed by the process
of dispersal \citep{McNutt.1996, Behr.2020}. Both males and females disperse
from their natal pack, either alone or in same-sex dispersing coalitions, and
search for unrelated mates and a suitable territory to settle
\citep{McNutt.1996, Cozzi.2020, Behr.2020}. During dispersal, wild dogs can
cover several hundred kilometers \citep{Masenga.2016, Woodroffe.2019,
Cozzi.2020}. Despite the importance of dispersal for the long-term viability of
this species, little empirical information is available on habitat selection and
potential movement barriers during dispersal. The few studies that have
collected dispersal data have shown that dispersers quickly move over large
distances, avoid human-dominated landscapes and areas densely covered by trees,
but prefer proximity to water \citep{Masenga.2016, Woodroffe.2019, Oneill.2020,
Cozzi.2020}.

Here, we collected GPS relocation data on 16 dispersing wild dogs in as many
dispersing coalitions from a free-ranging population in northern Botswana and
analyzed it to assess the adequacy of the KAZA-TFCA in securing connectivity. We
estimated the degree of selection or avoidance for environmental and
anthropogenic landscape features and used the obtained habitat preferences to
predict a permeability surface spanning the entire KAZA-TFCA. We then
investigated how landscape permeability varies regionally and internationally
and compared permeability within and outside the KAZA-TFCA boundaries. Finally,
we calculated least-cost paths and corridors to identify major movement routes
and to verify that these are successfully covered by the KAZA-TFCA.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E;
\Cref{StudyArea}a) was outlined by a rectangular bounding box stretching over
1.3 Mio km\textsuperscript{2} and encompassing the entire KAZA-TFCA
(\Cref{StudyArea}b). The KAZA-TFCA lies in the basins of the Okavango and
Zambezi rivers and includes parts of Angola, Botswana, Namibia, Zimbabwe, and
Zambia. With a total area of over 520'000 km\textsuperscript{2} it constitutes
the earth's largest transboundary conservation area and is characterized by
diverse landscapes, including savanna, grassland, and dry or moist woodland
habitats. Rainfall in the study area is seasonal and lasts from November to
March. The KAZA-TFCA also comprises the Okavango Delta, which represents a
highly dynamic hydrological flood-pulsing system \citep{McNutt.1996,
Wolski.2017}. The extent of the flood in the delta greatly changes within and
between years depending on the amount of rain that descends from the catchment
areas in Angola and reaches the distal ends of the delta between July and August
(Figure S4). The flood drastically affects surrounding landscapes, so that
during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a
patchy conglomerate of swamps, open water, and islands, whereas these structures
run dry when the flood retracts to its minimum extent (ca. 5'000
km\textsuperscript{2}; \citealp{Wolski.2017}). Despite 36 national parks (NPs)
and other protected areas, there is considerable human influence in some regions
of the KAZA-TFCA, mainly originating from farms, high human density, and road
traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red dotted rectangle depicts
    the study area, which was confined by a bounding box encompassing the entire
    KAZA-TFCA. Gray areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
    area within which dispersing coalitions were collared. Since Game Reserves
    in Botswana virtually serve the same purpose as National Parks, we use the
    terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
We used a population of free-ranging African wild dogs inhabiting the Okavango
Delta in northern Botswana as a source population for dispersing individuals.
This population has been extensively studied since 1989 \citep{McNutt.1996,
Cozzi.2013, Cozzi.2020, Behr.2020}. Between 2011 and 2019, we systematically
collected GPS relocation data on 16 coalitions of dispersing African wild dogs
(7 female and 9 male coalitions). Candidate dispersing individuals were
identified based on criteria reported in \cite{Behr.2020}, immobilized according
to protocols described in \cite{Osofsky.1996}, and fitted with GPS/Satellite
radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while
still with their natal pack. All procedures were undertaken and supervised by a
Botswana-registered wildlife veterinarian. During dispersal, GPS collars were
programmed to record GPS relocation data every 4 hours and to regularly transmit
them via Iridium satellite system to a base station.

Because we were interested in dispersal behavior only, we discarded any GPS data
collected while individuals were still with their natal packs and after
settlement in a new territory \citep{Cozzi.2020}. We identified the exact time
of emigration and settlement based on direct field observations and through
visual inspection of the net squared displacement (NSD) metric. NSD quantifies
the Euclidean distance of a relocation to a reference point \citep{Borger.2012},
which in our case was the center of the dispersing coalition's natal home range.
Thus, dispersal was deemed to have started when a coalition had left its natal
home range and continued until the NSD metric remained stationary, implying that
the coalition had successfully settled (Figure S1). In total, we collected 4'169
GPS relocations during dispersal (Figure S2 \& Table S1), resulting in an
average of 260 locations per dispersing coalition (min = 37, max = 729). In our
analysis, we did not differentiate between male and female dispersing
coalitions, for previous research found little differences between sexes during
dispersal \citep{Woodroffe.2019, Cozzi.2020}.

\subsection{Spatial Covariates}
To investigate habitat preferences of dispersing wild dogs, we used a set of
geo-referenced covariates (\Cref{Covariates}) that we aggregated in the
categories \textit{land cover} (which included water cover, distance to water,
shrubs/grassland cover, and tree cover), \textit{protection status} (protected
vs. unprotected), and \textit{human pressure} (which included human influence,
presence of roads, and distance to roads). For each of these covariates we
prepared spatial raster layers from freely available online services or from
remotely sensed satellite imagery. To ensure a consistent resolution (i.e.
cell-size or grain) across covariates, we coarsened or interpolated all layers
to match a resolution of 250m x 250m. For further details on the preparation of
each covariate, see Appendix A.3. We performed processing and manipulation of
data as well as all spatial and statistical analyses using R, version 3.6.1
\citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of spatial covariates that we included in our models. We
    prepared all covariates for the entire study area but for better visibility
    we only plot them for the surroundings of the Okavango Delta. The white
    rectangle in each plot depicts the area within which dispersing coalitions
    were collared. (a) Averaged layer of all dynamic (binary) water maps. (b)
    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
    Anything that was not covered by trees or shrubs/grassland was deemed to be
    bare land. (d) Protection status of the area. (e) Human influence proxy
    composed of human density, farms, and roads. (f) Distance to nearest road
    (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsection{Habitat Selection Model}
We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
investigate dispersers' selection or avoidance for the above-mentioned spatial
covariates. That is, we paired each realized step (i.e. the connecting line
between two consecutive GPS relocations; \citealp{Turchin.1998}) with 24 random
steps. We generated random steps by sampling turning angles from a uniform
distribution U(\(-\pi, +\pi\)) and step lengths from a gamma distribution that
was fitted using realized steps \citep{Avgar.2016}. A realized step and its 24
associated random steps formed a stratum that received a unique identifier.
Along each step, we extracted the above-mentioned covariates (Table S3),
standardized extracted values using a z-score transformation, and checked for
correlation using Pearson's Correlation Coefficient \(r\). None of the
covariates were overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we
retained all of them for modeling. Our habitat selection model then assumed that
dispersing wild dogs assigned a selection score \(w(x)\) of the following
exponential form to each realized and random step \citep{Fortin.2005}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The selection score \(w(x)\) of a step depended on its associated
covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's preferences for
these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). To estimate habitat
preferences (i.e. the \(\beta\)'s), we used mixed effects conditional logistic
regression analysis as suggested by \cite{Muff.2020}. We implemented their
method using the R-package \textit{glmmTMB} \citep{Mollie.2017} and used
dispersing coalition ID to model random intercepts and slopes. We defined two
movement metrics, namely the cosine of the turning angle (\(cos(ta)\)) and the
log of the step length (\(log(sl)\)), as core covariates and ran stepwise
forward model selection based on Akaike's Information Criterion (AIC;
\citealp{Burnham.2002}) for all other covariates. The inclusion of movement
metrics served to reduce biases in estimated habitat preferences that may have
arisen due to movement preferences \citep{Avgar.2016}. To validate the
predictive power of the most parsimonious habitat selection model, we ran k-fold
cross-validation for case-control studies as described in \cite{Fortin.2009}
(details in Appendix A.5).

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell. Because our representation of water was dynamic (to
properly render the pulsing behavior of the Okavango Delta) we collapsed all
dynamic water maps into a single static map using areas that were covered by
water in at least 10\% of the cases. Using the resulting static map we also
calculated a layer returning the distance to water. To reduce the influence of
outliers in predicted permeability scores we followed \cite{Squires.2013} and
curtailed predicted scores between the 1\textsuperscript{st} and
99\textsuperscript{th} percentile of their original values. To compare
permeability across different regions, we normalized the permeability surface to
a range between 0 (most impermeable) and 1 (most permeable). We then determined
median permeability within and outside the KAZA-TFCA, within and outside
formally protected areas, and within each of the five KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify movement corridors of dispersing wild dogs, we specified source
points and calculated factorial least-cost paths (LCPs) as well as factorial
least-cost corridors (LCCs) among them \citep{Elliot.2014}. We generated source
points by overlaying the study area with a regular grid of points spaced at 100
km. We only considered those points that fell within protected areas \(>\) 700
km\textsuperscript{2}, which conforms with home-range requirements of African
wild dogs reported in \cite{Pomilia.2015}. Finally, we defined centroids as
source points for those protected areas \(>\) 700 km\textsuperscript{2} that
were not assigned any source points from the regular grid. In total, we
generated 68 source points, which resulted in 2'278 unique pairwise combinations
and therefore 2'278 unique LCPs and LCCs. We computed factorial LCPs and LCCs
between source points using the R-package \textit{gdistance} (for further
details see Appendix A.6). After computation, we tallied overlapping LCPs and
LCCs, respectively, into single connectivity maps.

\section{Results}
\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)} and
\textit{log(sl)} (\Cref{PermeabilityResults}a). Parameter estimates showed that
dispersing wild dogs moved in a directional and fast manner, as indicated by a
positive selection for small turning angles, i.e. high \(cos(ta)\) (\(\beta\) =
0.14; 95\% CI = 0.07 to 0.21) and longer steps, i.e. high \(log(sl)\) (\(\beta\)
= 0.06, 95\% CI = 0.02 to 0.09). Dispersers avoided moving through water
(\(\beta\) = -0.52, 95\% CI -0.77 to -0.26) but selected for locations in its
vicinity, although the latter effect was not significant (\(\beta\) = -0.32,
95\% CI = -0.72 to 0.08). Dispersers avoided areas that were densely covered by
trees (\(\beta\) = -0.31, CI = -0.46 to -0.15) and preferred areas covered by
shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42). Finally, dispersers
avoided areas that were influenced by humans (\(\beta\) = -0.41, 95\% CI = -0.78
to -0.05).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. (b) Results from the k-fold
    cross validation for case-control studies. The left graph shows rank
    frequencies of \textit{realized} steps according to predictions, whereas the
    right graph shows rank frequencies of \textit{randomly selected} steps
    according to predictions. \(\bar{r}_s\) indicates the mean correlation
    coefficient resulting from 100 repetitions of the k-fold cross validation.
    The blue smoothing line was fitted using a locally weighted polynomial
    regression and serves to aid the eye in detecting the trends. Correlation
    coefficients suggest that our prediction was significant and robust,
    evidenced by the fact that the confidence intervals of \(\bar{r}_{s,
    realized}\) and \(\bar{r}_{s, random}\) did overlap and by the fact that
    there was strong and significant correlation between ranks and associated
    frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

Results from the k-fold cross-validation suggested that our prediction was
significant and robust, as highlighted by the fact that the 95\%-CIs intervals
of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap
(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations (\Cref{PermeabilityResults}b).

\subsection{Permeability Surface}
Our prediction of landscape permeability revealed substantial differences across
regions in the study area (\Cref{PermeabilityMap}). Comparisons of median
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
the KAZA-TFCA is more than two times as high as permeability outside it.
Permeability varies by country, with a five-fold permeability difference among
them. Angola and Botswana are characterized by comparably highly permeable
landscapes, Zimbabwe and Zambia are relatively impermeable, and Namibia ranges
in between the two extremes (\Cref{PermeabilityComp}). Visual inspection of our
covariate layers indicated that high permeability in Angola and Botswana is
mainly caused by a combination of low human influences, low tree cover, high
shrubs/grassland cover, and a close distance to water. Although swamps,
wetlands, and permanent water themselves provide little permeability, their
surroundings act as strong attractants to dispersers. The low permeability that
characterizes Zambia and Zimbabwe, on the other hand, is mainly caused by
substantial human influences. Albeit the KAZA-TFCA covers most permeability
hot-spots, several highly permeable regions remain uncovered by its borders.
Across all countries, protected areas provide roughly double the permeability of
unprotected landscapes (\Cref{PermeabilityComp}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permeability was predicted by calculating selection scores \(w(x) =
    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each raster cell
    based on the raster cell's underlying covariates (\(x_i\)) and estimated
    habitat preferences (\(\beta_i\)). Areas that dispersers find easy to
    traverse are depicted in bright colors. Bold white lines delineate the
    borders of the KAZA-TFCA, whereas dashed white lines show country borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Comparison of median permeability (interquantile range in brackets)
  across countries, separated into areas within and outside the KAZA-TFCA, as
  well as within and outside formally protected areas. High values indicate high
  permeability, whereas low values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{tabular}{llllll}
    \toprule
    \multicolumn{1}{c}{} & \multicolumn{2}{c}{KAZA-TFCA} &
    \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{} \\
    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
    Country & Inside & Outside & Protected & Pastoral & Overall\\
    \midrule
    Angola & 0.36 (0.41) & 0.12 (0.32) & 0.36 (0.41) & 0.12 (0.33) & 0.20 (0.39)\\
    Botswana & 0.25 (0.30) & 0.15 (0.16) & 0.28 (0.35) & 0.15 (0.18) & 0.19 (0.25)\\
    Namibia & 0.22 (0.30) & 0.13 (0.18) & 0.24 (0.30) & 0.11 (0.15) & 0.16 (0.24)\\
    Zambia & 0.05 (0.09) & 0.03 (0.06) & 0.04 (0.10) & 0.03 (0.05) & 0.03 (0.07)\\
    Zimbabwe & 0.07 (0.16) & 0.06 (0.05) & 0.08 (0.17) & 0.05 (0.05) & 0.06 (0.07)\\
    \hline
    Overall & 0.16 (0.30) & 0.07 (0.15) & 0.15 (0.30) & 0.07 (0.15) & 0.10 (0.22)\\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Our least-cost analysis revealed three major movement corridors of which all
were well-contained within the KAZA-TFCA boundaries (\Cref{LeastCost}). One
major corridor runs SE-NW and connects the Okavango-Linyanti ecosystem in
Botswana with Luengue-Luiana NP in Angola. A second corridor runs W-E between
Chobe NP in Botswana and Zimbabwe's Hwange NP. A third major corridor runs
NE-SW, completely across unprotected areas, and connects Kafue NP in Zambia with
more central regions of the KAZA-TFCA. Several minor corridors branch off from
these three major corridors; these include a southward connection between the
Okavango-Linyanti and the Central Kalahari Game Reserve, a southwesterly
corridor connecting Luengue-Luiana NP with Namibia's Khaudum NP, and a
northeasterly extension of the Hwange corridor into Zimbabwe's Matusadona NP.
According to our predictions, the landscapes in the Okavango-Linyanti region are
the highest frequented dispersal routes within the KAZA-TFCA
(\Cref{LeastCost}b). Our model did not detect any significant direct corridors
between Zimbabwe and Zambia or Zambia and Angola, and only a very limited W-E
direct connection between the Okavango region and Namibia's Khaudum NP. Except
for the corridor into the Central Kalahari National Park, our model did not
detect any significant connectivity outside the boundaries of the KAZA-TFCA.
Furthermore, we found little to no direct connectivity between peripheral
points; that is, most paths and corridors connecting two adjacent peripheral
points run through more central regions before heading towards their destination
at the periphery (\Cref{LeastCost}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (light grey). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference, we also labeled some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

\section{Discussion}
We used GPS relocation data collected on dispersing African wild dogs to
investigate whether their main movement corridors are contained within the
boundaries of the world's largest transboundary conservation area, namely the
KAZA-TFCA. Our analysis suggests that the KAZA-TFCA indeed encompasses all major
corridors of African wild dogs, demonstrating the potential value of such an
initiative. We thus exemplified how pertinent dispersal data of a highly mobile
species can be used to assess the adequacy of already existing or planned
protected areas. Our approach is neither limited to the African wild dog, nor to
our study area and thus applicable to any study system. All covariates used
throughout this study are readily available on a global scale and many of them
are likely to be important determinants of movement behavior, landscape
permeability, and connectivity for other species \citep{Thurfjell.2014,
Zeller.2012}. Interestingly, our predicted network of least cost-paths and
corridors for African wild dogs shows surprising similarities to corridors of
dispersing lions inhabiting the same ecosystem \citep{Elliot.2014,
Cushman.2018}. This not only reinforces confidence in our own predictions but
also suggests potential synergies for the conservation of these two, and
possibly more, species. Expanding our analytical framework to additional species
will likely yield important insights on the consistency of inter-specific
movement corridors, thus highlighting areas that are exceptionally valuable for
the conservation of several species.

Our results emphasize that human influences constitute some of the main barriers
to connectivity among wild dog populations. This conforms to findings on
dispersing wild dogs from eastern Africa \citep{Masenga.2016, Oneill.2020} but
conflicts with findings from South Africa by \cite{DaviesMostert.2012}, who
reported a high willingness of dispersers to cross human-dominated landscapes.
We believe that such differences are due to the unavailability of alternative
routes through natural landscapes, which may have forced dispersers in South
Africa to cross human dominated landscapes despite a strong aversion to do so.
In this regard, our representation of dispersal corridors and the resulting
connectivity appear conservative, as dispersers may be able to make the best out
of a bad situation and cross landscapes characterized by considerably
unfavorable conditions \citep{Palomares.2000, Elliot.2014}. Nevertheless,
successful conservation of this species relies on policymakers' and local
authorities' willingness and ability to provide and conserve natural areas that
remain free from anthropogenic pressures. This is not only paramount in light of
increasing connectivity and facilitating dispersal, but also in terms of
reducing human-caused mortality during dispersal. In fact, previous studies have
shown that human-caused mortality represents a major threat to wild dogs'
ability to disperse \citep{Woodroffe.2019, Cozzi.2020}.

Besides human influences, we identified water as additional obstacle to
dispersal. This corroborates earlier studies showing that water bodies are
almost impenetrable to resident packs \citep{Abrahms.2017} and only infrequently
crossed by dispersing individuals \citep{Cozzi.2020}. An accurate and dynamic
representation of water is thus imperative and particularly relevant in seasonal
or flood-pulsing ecosystems such as the Okavango Delta.

Although dispersers avoided moving through water, they selected locations in its
vicinity. This preference may be caused by the occurence of prey close to water
\citep{Bonyongo.2005}. For the same reason, however, competitors such as lions,
spotted hyenas, and resident wild dogs may also use areas close to water
\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to move
into prey-poorer areas away from water. Given the influence that resident
conspecifics, competitors, and prey can have on dispersers \citep{Cozzi.2018,
Armansin.2019} future studies should strive to collect and incorporate intra-
and interspecific relationships into analyses of landscape connectivity.

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersing wild dogs gain access to more peripheral regions of the
KAZA-TFCA. It appears that the absence of human activities, the central position
within the KAZA-TFCA, and the presence of relatively impermeable water bodies
(e.g. Okavango Delta, Linyanti Swamp) funnel dispersal movements, resulting in a
highly frequented corridor. The key role of the Okavango-Linyanti region for
overall connectivity within the KAZA-TFCA thus calls for actions to secure its
protection status in the future. While the region is currently a Wildlife
Management Area, it has neither the status of a National Park nor that of a Game
Reserve. A similar case of non-formally protected but key dispersal landscape is
represented by the area south of Kafue NP in Zambia, for which a disruption of
its main and narrow dispersal corridor would result in considerable isolation of
its subpopulations. We also revealed a potential southwards corridor between the
Okavango-Linyanti ecosystem and the Central Kalahari National Park.
\cite{Elliot.2014} and \cite{Cushman.2018} identified a similar corridor for
dispersing lions, suggesting that upholding and protecting a link between those
ecosystems is pivotal. Some areas through which the corridor runs are neither
part of the KAZA-TFCA nor profit from any form of protection status. In fact,
human presence and activities along the national road that longitudinally
traverses this corridor may limit realized connectivity \citep{Cozzi.2020}.

Our approach of identifying movement corridors based on pre-defined start and
end points implicitly assumes that individuals know the end point of their
dispersal journey and that they have almost complete knowledge of associated
movement costs \citep{Panzacchi.2016}. Since dispersers often move into unknown
territory, this may not necessarily be the case \citep{Abrahms.2017,
Cozzi.2020}. However, specification of pre-defined end points might not be
necessary, as the parametrized iSSF model can be used as mechanistic movement
model to simulate dispersal events from known source points, yet without
restricting the domain of potential end points \citep{Signer.2017}.
Consequently, movement corridors would emerge more naturally as the result of a
myriad of simulated dispersal events. While a simulation-based approach is
conceptually straightforward, it requires a comprehensive mechanistic
understanding of dispersal movements, which is conditional on our ability to
collect additional dispersal data and adequately represent the landscape through
which individuals move.

Our work shows how dispersal data of a highly mobile species can be used t
identify movement corridors and to assess the adequacy of protected areas. In
our case, the predicted movement corridors of African wild dogs were well
contained within the boundaries of the world's largest transboundary
conservation area, namely the KAZA-TFCA, suggesting that it will significantly
contribute to the long-term viability of this species. Moreover, our
connectivity network allowed revealing potential dispersal hubs through which
dispersers gain access to more remote regions of the study area. Finally, our
investigations showed that human influence constitutes one of the main barriers
to dispersal and substantially reduces landscape connectivity. Successful
conservation of wide-ranging species, such as exemplified by the African wild
dog, will therefore be contingent on the willingness of local authorities,
policymakers, and land managers to preserve areas that remain free from human
strains. Ultimately, our work contributes to the growing field of connectivity
studies and provides and easily expandable framework for assessing the adequacy
of already-existing or planned protected areas.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modelling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing coalitions will be made available on dryad at
the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Prof. J. Fieberg, who consulted all statistical aspects of this work and P.
Wolski, from the Okavango Research Institute, who assisted us in generating
dynamic water maps. This study was funded by Basler Stiftung für Biologische
Forschung, Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic
Society, Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust
Foundation, Forschungkredit der Universität Zürich, and a Swiss National Science
Foundation Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-01-03.tex
first commit

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{epstopdf}
\usepackage[onehalfspacing]{setspace}
\usepackage[paper=a4paper,left=30mm,right=30mm,top=20mm,bottom=25mm]{geometry}
\usepackage[round]{natbib}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[labelfont=bf]{caption}
\usepackage{footnote}
\usepackage{tabularx}
\usepackage{gensymb}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{varwidth}
% \usepackage[nomarkers,figuresonly]{endfloat}
\usepackage{csvsimple}
\usepackage{arydshln}
\usepackage{titlesec}

% Define a new command to create a number inside a circle (for plots)
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=1pt] (char) {#1};}}

% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Increase the space between two rows in tables
\renewcommand{\arraystretch}{1.2}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Adjust the distance between title and paragraph
\RedeclareSectionCommand[
  beforeskip=-.5\baselineskip,
  afterskip=.5\baselineskip]{subsection}
\RedeclareSectionCommand[
  beforeskip=-.25\baselineskip,
  afterskip=.25\baselineskip]{subsubsection}

\doublespacing
\begin{document}
\pagenumbering{gobble}
\bibliographystyle{apalike}

\begin{titlepage}

% Define a new command to produce a horizontal line
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

% Center everything on the page
\center

%------------------------------------------------------------------------------
%	HEADING SECTIONS
%------------------------------------------------------------------------------

% Name of your university/college
\textsc{\LARGE  }\\[1.5cm]

% Major heading
\textsc{\Large Department of Evolutionary Biology and Environmental
Studies}\\[0.5cm]

% Minor heading
\textsc{\large Master's Thesis}\\[0.5cm]

%------------------------------------------------------------------------------
%	TITLE SECTION
%------------------------------------------------------------------------------
\doublespacing
\HRule \\[0.4cm]

% Title of the document
{\textsc{ \LARGE Running Wild: Movement Corridors of Dispersing African Wild
Dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation
Area}}\\[0.4cm]

\HRule \\[1.5cm]
\singlespacing

%------------------------------------------------------------------------------
%	AUTHOR SECTION
%------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large

% Main author
\emph{Author:}\\
David D. Hofmann\\
david.hofmann2@uzh.ch\\
12-728-846

\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large

% Supervisors
\emph{Supervisors:} \\
Dominik Behr\\
Ph.D. Gabriele Cozzi\\
Prof. Arpat Ozgul

\end{flushright}
\end{minipage}\\[2cm]

%------------------------------------------------------------------------------
%	DATE SECTION
%------------------------------------------------------------------------------

% Today's date
{\large \today}\\[5cm]

%------------------------------------------------------------------------------
%	LOGO SECTION
%------------------------------------------------------------------------------

% Logo of the university
\includegraphics{UZHVector.eps}\\[1cm]

%------------------------------------------------------------------------------

% Fill the rest of the page with whitespace
\vfill

\end{titlepage}

\begin{abstract}
The African wild dog \textit{(Lycaon pictus)} is an endangered species that
currently only survives in a few scattered subpopulations in Sub-Saharan Africa.
Long-term viability of the species thus requires the identification and
preservation of key dispersal corridors that connect such subpopulations. We
collected GPS data of 16 dispersing coalitions originating from a free-ranging
population in northern Botswana and used integrated step selection analysis to
create a permeability surface spanning over the entire Kavango-Zambezi
Transfrontier Conservation Area (KAZA-TFCA), the largest transboundary
conservation area on earth. Based on our permeability surface, we calculated
least-cost paths and least-cost corridors between selected locations within the
KAZA-TFCA. Our results show that dispersal corridors typically run through open
shrubs/grassland, and in the vicinity of water. Corridors rarely cross
waterbodies, areas with high human presence, or areas densely vegetated by
trees. Several dispersal corridors meet in the Linyanti-Chobe ecosystem, which
highlights the importance of northern Botswana for the conservation of this
charismatic species.
\end{abstract}
\newpage

\onehalfspacing
\tableofcontents
\doublespacing
\newpage

\pagenumbering{arabic}

\section{Introduction}
% Importance of Connectivity
Successful conservation of spatially fragmented subpopulations requires to
assess connectivity of habitats occupied by these subpopulations. In this
context, connectivity describes the degree to which the landscape facilitates or
impedes movement of the focal species between habitat patches
\citep{Taylor.1993, Clobert.2012}. Thus, connectivity is a prerequisite for
dispersal, which in turn promotes genetic exchange, reinforcement of small
populations, and colonization or recolonization of unoccupied habitats
\citep{Brown.1977, Hanski.1998, MacArthur.2001, Frankham.2002, Leigh.2012}. All
of these processes crucially contribute to the long-term viability of numerous
species, therefore emphasizing the importance of connectivity. However,
increasing anthropogenic pressure and ongoing loss and fragmentation of valuable
habitats erode connectivity on a worldwide scale \citep{Fahrig.2003,
Barnosky.2012}. As a result, management actions for the conservation of
endangered species increasingly aim to restore connectivity through the
identification and preservation of key movement corridors that link isolated
habitats \citep{Fahrig.2003, Heller.2009}. An unprecedented instance of an
initiative to restore connectivity on a large scale is the Kavango-Zambezi
Transfrontier Conservation Area (KAZA-TFCA), which traverses five African
countries and spans over 520,000km\textsuperscript{2}. While the primary purpose
of the KAZA-TFCA is to facilitate the migration of elephants across large
landscapes, numerous other wide-ranging species will likely benefit from its
planned protectorates too. Nevertheless, the extent to which such initiatives
are beneficial to certain species is rarely assessed, leaving open their true
impact on connectivity.

% African wild dogs
One of the species for which the degree to which it will benefit from the
KAZA-TFCA remains unknown is the African wild dog (\textit{Lycaon pictus}), a
keystone predator and flagship species for conservation efforts in the KAZA-TFCA
\citep{VanDerMeer.2016}. While once present across the entire Sub-Saharan
continent, the species has been widely extirpated through human persecution,
habitat destruction and disease outbreaks. As a result, the African wild dog has
become Africa's most endangered large carnivore that currently only survives in
small, spatially scattered subpopulations (see \Cref{StudyArea}a,
\cite{Woodroffe.2012}). Within these subpopulations, wild dogs form cooperative
breeding groups of up to thirty individuals in which a single dominant pair
monopolizes the majority of reproduction \citep{Frame.1979, Fuller.1992,
Creel.2002}, while subordinates of both sexes help to feed newborn pups
\citep{Malcolm.1982}. After reaching 15 months of age, male and female offspring
disperse from their natal pack in same-sex coalitions and search for unrelated
mates and a suitable territory to settle (\citeauthor{McNutt.1996}, 1996;
\citeauthor{Behr.2019}, unpublished). During such dispersal events, wild dogs
cover vast distances, sometimes crossing perilous areas that are strongly
influenced by humans (\citeauthor{DaviesMostert.2012}, 2012;
\citeauthor{Masenga.2016}, 2016; \citeauthor{Cozzi.2019}, in review). Despite
the importance of dispersal for the remaining wild dog subpopulations, little is
known about the movement corridors of dispersing individuals and connectivity
between subpopulations in the KAZA-TFCA in general.

% Connectivity Assessment
In recent years, there has been a growing body of literature that uses animal
movement data to identify movement corridors and thereby assess connectivity on
large scales \citep{Chetkiewicz.2006}. Revelation of potential movement
corridors between selected locations traditionally relies on the estimation of a
permeability surface (elsewhere referred to as resistance surface when
inverted), which depicts the ease or willingness at which the focal species
traverses a specific landscape. Latest and very powerful empirical methods to
predict permeability based on observed movements include \textit{resource
selection functions} (RSF, \cite{Boyce.2002}), \textit{step selection functions}
(SSF, \cite{Fortin.2005}) and \textit{path selection functions} (PSF,
\cite{Cushman.2010}). All methods have in common that they infer habitat
preferences of the focal species by comparing environmental covariates at
visited locations to environmental covariates at randomly selected locations
that are deemed available but presumably unused by the animal. Based on inferred
habitat preferences, the methods further assume that habitats preferred by the
focal species provide high permeability, whereas avoided habitats provide low
permeability. Regardless of the method chosen (i.e. RSF, SSF, or PSF), all
approaches require adequate relocation data that is representative for the
process being studied.

% Behavioral Modes
For connectivity studies in particular, the behavioral mode of the animal (e.g.
resident, dispersing) during data collection should be considered when
estimating landscape permeability and assessing movement corridors. However, few
studies have partitioned their movement data according to behavioral mode
\citep{Wilson.2012, Vasudev.2015}. As a consequence, most researchers estimated
permeability for dispersing individuals as if they had the same preferences as
resident individuals. This is a rather disputed assumption and recent research
suggests that such an approach seriously reduces the power to reveal meaningful
movement corridors \citep{Killeen.2014}. For instance, \cite{Elliot.2014}
demonstrated marked differences in predicted permeability surfaces for
dispersing lions in the KAZA-TFCA depending on whether permeability was assessed
using data collected during residence or dispersal. \cite{Jackson.2016} made
similar observations for African wild dogs and revealed higher permeability for
dispersers than for residents. Moreover, \cite{Abrahms.2017} found a
significantly greater overlap between predicted and observed corridors for three
dispersing wild dogs when predictions were based on information collected during
wild dog activity, rather then when ignoring the behavioral mode. Overall,
predicted movement corridors vastly differed depending on the type of data that
was being used \citep{Elliot.2014, Jackson.2016, Abrahms.2017}. Unfortunately,
data of dispersing wild dogs remains scarce and is collected opportunistically
at best \citep{DaviesMostert.2012, Masenga.2016, Jackson.2016, Abrahms.2017}.
For instance, the dispersers reported in \cite{DaviesMostert.2012} were not
fitted with global positioning system (GPS) collars, which is why the
researchers only recorded straight line dispersal-distances.
\cite{Masenga.2016}, on the other hand, were able to collect proper GPS data,
yet only on two dispersers. Furthermore, \cite{Jackson.2016} based their work on
occasional wild dog sightings and their distinction between residents and
dispersers was based purely on group size, assuming that any group with fewer
than nine members was dispersing. Finally, \cite{Abrahms.2017} used GPS
relocation data of resident individuals during phases of activity rather than
during true dispersal. In summary, none of these studies systematically
monitored dispersing individuals, which is likely owed to technological and
logistic limitations when monitoring such an elusive and wide-ranging species
\citep{Nathan.2001, Tesson.2013}. Yet, a better understanding of how the
landscape impedes or facilitates movement of dispersers is vital for improving
our understanding of connectivity and for conservation of the species in
general.

% Research Goal
Our study therefore had two aims: (1) to estimate landscape permeability for
dispersing wild dogs, and subsequently use this information to (2) assess
connectivity by revealing key dispersal corridors within the entire extent of
the KAZA-TFCA. For this purpose, we collected GPS data of 16 dispersing
coalitions originating from a free-ranging population of wild dogs in the
eastern section of the Okavango Delta in Botswana. We highlight key dispersal
corridors that provide connectivity between already established protected areas
within the KAZA-TFCA and we reveal potentially important dispersal corridors
outside the planned protectorates of the KAZA-TFCA. Our results have important
implications for the conservation and management of the endangered African wild
dog.

\newpage
\section{Methods}
\subsection{Study Area}
% Core Study Area
For the purpose of this study we differentiated between a \textit{core study
area}, where all dispersers originated from, and an \textit{extended study
area}, for which we conducted the connectivity analysis. The \textit{core study
area} was located in the eastern section of the Okavango Delta in northern
Botswana (\Cref{StudyArea}b, centered at -19\degree 25'30''S, 23\degree
37'30''E; elevation ca. 950m) and comprised parts of the Moremi Game Reserve and
adjacent wildlife management areas. The Okavango Delta is considered to be one
of the main strongholds of African wild dogs in southern Africa and likely acts
as a source for the recolonization of surrounding habitats
\citep{Woodroffe.2012, Cozzi.2013}. The Okavango Delta represents the earth's
largest inland delta and is characterized by diverse habitats, ranging from
permanent river systems, mopane woodland and acacia forests, to open grasslands
and seasonal swamps \citep{Broekhuis.2013}. Rainfall is periodic and lasts from
November to March, whereas the rest of the year remains dry \citep{McNutt.1996}.
Flooding of the delta greatly changes depending on the season and drastically
affects surrounding landscapes (see \Cref{FloodPulse} in
\Cref{Appendix:FloodPulse} for an illustration of a typical flood pulse). During
maximum extent, the delta becomes a patchy conglomerate of swamps, open water,
and islands. Because rainfalls only slowly descent from their catchment areas in
southern Angola into the Okavango Delta, the flood usually peaks at the height
of the dry season between August and September \citep{Wolski.2017}. Most of the
Okavango Delta's unique basin is under protection, such that human influence in
our core study area was mainly restricted to the surroundings of Maun and the
smaller villages of Khwai, Mababe, and Sankuyo.

% Large Study Area
Our \textit{extended study area} was confined by the extent of the KAZA-TFCA
(\Cref{StudyArea}a, centered at -17\degree 13'9''S, 23\degree 56'4''E; elevation
ca. 1,030m), which lies in the basins of the Okavango and Zambezi rivers and
embraces parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. With a total
area of over 520,000km\textsuperscript{2} the KAZA-TFCA contains several already
existing national parks (NPs) and other protected areas, and will eventually
become earth's largest transboundary conservation area. Since we considered the
entire rectangular extent of the KAZA-TFCA, our extended study area encompassed
a total of over 1.3 Mio km\textsuperscript{2}. Similar to our core study area,
the extended study area hosts a conglomerate of various landscapes, including
savanna, grassland, and dry or moist woodland patches. Furthermore, the
KAZA-TFCA encloses several wild dog subpopulations  that still reside in
Namibia, Botswana, and Zimbabwe (\Cref{StudyArea}a, \cite{Woodroffe.2012}).
Despite 36 NPs and numerous other protected areas, there is considerable human
influence in some regions of the KAZA-TFCA originating from farms, high human
density, and road traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red rectangle depicts our
    \textit{extended study area}, which was confined by the extent of the
    KAZA-TFCA. Purple areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The orange rectangle illustrates our
    \textit{core study area}, from which all dispersers departed. Yellow lines
    depict all dispersal trajectories that were recorded between the years 2011
    and 2019. Since game reserves in Botswana virtually serve the same purpose as
    National Parks, we will use the terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
% Collection
We collected GPS relocation data of 16 wild dog coalitions (i.e. 7 female and 9
male coalitions) that dispersed from the core study area between 2011 and 2019.
Potential dispersing individuals were immobilized according to protocols
described in \cite{Osofsky.1996} and fitted with GPS/Satellite radio collars
(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while they were still
with their natal pack. All required procedures were undertaken and supervised by
a Botswana-registered wildlife veterinarian. Immobilized individuals were
monitored and all were observed to successfully re-join their packs within one
hour after immobilization. GPS collars were programmed to record and transmit
GPS relocations via Iridium satellite system to a base station. The system thus
allowed remote-tracking of collared individuals, even where field conditions
were prohibitive or dispersal trajectories crossed international borders.
Because we were mainly interested in recording dispersal trajectories,
relocations were collected only every 24 hours during residence, but every 4
hours during dispersal. Dispersal and residence were strictly distinguished
using a combination of net squared displacement \citep{Borger.2012}, visual
inspection of trajectories, and observations of pack constellations in the field
(for details see \citeauthor{Cozzi.2019}, in review). Dispersal was deemed to
have started when an individual left its natal home range, and continued until
the individual formed a new pack or exhibited signs of resident behavior. Some
dispersers undertook exploratory movements prior to dispersal. Such movements
have been observed in other species as well and are reported to be very similar
to true dispersal \citep{Killeen.2014}. We therefore classified exploratory
movements as dispersal too. We did not attempt to distinguish sex in our
analysis, as previous research found little differences between males and
females during dispersal (\citeauthor{Woodroffe.2019}, 2019;
\citeauthor{Cozzi.2019}, in review). In total, we collected 4,691 GPS
relocations during dispersal (see \Cref{Trajectories}). On average, dispersal
events included three individuals (min = 1, max = 7), lasted for 41 days (min =
7, max = 133), and covered a mean cumulative distance of 370km (min = 105, max =
610). Detailed descriptions of the dispersal events can be found in
\citeauthor{Cozzi.2019} (in review). Although our final sample size of 16
dispersers was relatively small, we consider this an exceptional number for such
an endangered and elusive species.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Trajectories.pdf}
    \caption{Illustration of the dispersal trajectories that we recorded. Each
    color represents a different dispersal coalition. All coalitions departed
    from our core study area, which is indicated by the white square. The
    coalition dispersing towards the far east of the map covered over 360km in
    under 10 days.}
    \label{Trajectories}
  \end{center}
\end{figure}

\subsection{Covariates}
% General stuff
We used a set of geo-referenced covariates to investigate movement and habitat
preferences of dispersing wild dogs (see \Cref{Covariates}). Covariates fell
into one of the three categories \textit{land cover}, \textit{protection
status}, and \textit{human influence}. For each category we obtained spatial
raster layers from freely available online services or from remote sensed
satellite imagery. To ensure a consistent resolution (i.e. cell-size or grain)
across covariates, all layers were coarsened or interpolated to match a
resolution of 250m x 250m. Processing and manipulation of data, as well as all
spatial and statistical analyses were performed in R, version 3.6.1
\citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of covariates that we used in our models. Although all
    covariates were prepared for the extended study area, we present only an
    extract to maximize visibility. The white square in each plot depicts our
    core study area. (a) Averaged layer of all binary maps of water and dryland
    that we acquired using our remote sensing algorithm. (b) Percentage cover of
    trees. (c) Percentage cover of shrubs/grassland. Note that anything not
    covered by trees or shrubs/grassland was deemed to be bareland. (d)
    Protection status according to the Peace Parks Foundation. (e) Human
    influence proxy, including human density, farms, and roads. (f) Distance to
    nearest road (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsubsection{Land Cover}
% Overview
Land cover was represented by four layers: (1) a binary layer distinguishing
water and dryland, and (2) three continuous layers showing the percentage cover
by different vegetation types, together adding up to 100\%.

% Dynamic Watermapping
(1) For the purpose of this study, water included open waters, rivers, wetlands,
and swamps, whereas anything else was presumed to be dryland. Because flooding
of the Okavango Delta is highly variable, we decided to map the flood
dynamically for the extent shown in \Cref{Floodmapping}. To create dynamic
``floodmaps'', we applied an algorithm that was originally designed by the
\textit{Okavango Research Institute} (ORI) and is explained in detail by
\cite{Wolski.2017}. However, we rewrote and slightly amended the original Python
code in R and thus describe the underlying principles again and especially
indicate the discrepancies to the original algorithm.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
    \caption{Figures describing the floodmapping algorithm. (a) The colored
    polygons indicate permanent waters and permanent dryland. Below these
    polygons we extracted reflectance values of MODIS Terra Band 7 and used
    their repsective medians to calculate a classification threshold. (b)
    Example of a classified MODIS Terra Band 7 image after application of the
    threshold.}
    \label{Floodmapping}
  \end{center}
\end{figure}

% Algorithm
To implement the algorithm, we defined two sets of polygons located in the
region of the Okavango Delta (see \Cref{Floodmapping}). The first set consisted
of areas known to be permanent dryland, whereas the second set consisted of
permanent waters. Since we were unable to retrieve the original polygons used in
\cite{Wolski.2017}, we geo-referenced one of their illustrations and traced the
polygons there. After recreation of the polygons, we used the R-package
\textit{getSpatialData} \citep{Schwalb.2018} to download and pre-process all
relatively cloud-free MODIS Terra images (MCD43A4, \cite{Schaaf.2015}) available
for the period of our dispersal events. Assessment of cloud cover was based on
visual inspection of MODIS images on ORI's website \citep{ORI.2019}. The MCD43A4
MODIS product is particularly useful for dynamic floodmapping because it is
updated with new composite images every 8th day \citep{Wolski.2017}. After
download, we classified each MODIS image into a binary map of water (flood) and
dryland using a threshold that was identified as follows. First, we extracted
all reflectance values of MODIS Terra Band 7 within the water- and
dryland-polygons. Second, we computed histograms of water-reflectances and
dryland-reflectances and empirically verified that reflectances of the two
groups were sufficiently distinct. More specifically, we checked if
superimposing the two histograms resulted in a bimodal plot. This was said to be
achieved if the 99\textsuperscript{th} percentile of water-reflectances did not
severely exceed the 1\textsuperscript{st} percentile of dryland-reflectances
(\(p_{0.99, water} - \frac{10}{255} < p_{0.01, dryland}\)). Third, if peaks were
sufficiently different, we calculated a threshold (\(t\)) using \Cref{EQ1}:

\begin{equation}
\label{EQ1}
t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
\widetilde{p}_{water})
\end{equation}

\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
the median reflectances of water and dryland, respectively. We then classified
all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
all pixels with a value smaller than \(t\) as water.

% Dynamic Wetmask
Importantly, bimodality was not always achieved and in some cases no floodmap
could be calculated. In fact, it appears that non-bimodality caused the ORI
algorithm to fail since the end of 2018, such that no floodmaps have been
generated since then (ORI, personal comm.). We hypothesized that this was caused
by the application of static water-polygons that did not cover permanent waters
correctly anymore. We therefore revised the algorithm and allowed for a more
dynamic polygonization of water. That is, for each MODIS image we calculated new
water-polygons comprising areas that were covered by the flood in 99\% of the
floodmaps from the previous five years. All of the necessary floodmaps from
previous years were kindly provided to us by ORI. Using this slightly amended
approach, we were able to address some of the bimodality issues and to classify
several additional floodmaps for the period of our study. Because MODIS Terra
Band 7 had a resolution of 500m x 500m, we interpolated all maps to 250m x 250m.

% Validation
To validate and compare the performance of our own algorithm to the original
ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
images. To make sure that months were equally represented in the sampled dates,
we employed stratified sampling based on months (regardless of the year) and
randomly sampled four maps for each month. For the sampled dates we downloaded
and classified MODIS Terra Band 7 images and compared our classified images to
those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
created a difference map indicating false positives and false negatives and
computed the relative number of wrongly classified pixels. We achieved an
overall accuracy of 97\%, which presumably is an underestimate of the true
performance, as we introduced some errors when resampling the ORI-maps to our
reference raster.

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
    \caption{Validation procedure of our floodmapping algorithm. (a) Classified
    image that was provided to us by ORI. (b) Image for the same date but now
    classified using our own algorithm. (c) Difference image indicating false
    positives and false negatives in our own classification.}
    \label{FloodmappingValidation}
  \end{center}
\end{figure}

% Static Watermap
\noindent While water in the Okavango Delta was mapped dynamically, we used a
static watermap for its surroundings. This static map was based on Globeland's
land cover dataset \citep{Chen.2015}, which contains ten land cover categories.
From the original layer we only retained the categories \textit{wetland} and
\textit{waterbodies} and collectively reclassified them to \textit{water}.
Globeland had an original resolution of 30m x 30m, so we coarsened the layer to
250m x 250m using the mode of each 250m x 250m cell. Lastly, we improved river
representation using the MERIT Hydro dataset \citep{Yamazaki.2019}, from which
we added all rivers with a width of over 10m to our watermaps. The entire
process resulted in several hundred watermaps, of which 72 represented the
closest watermap to one of our GPS relocations.

(2) The three vegetation layers that we used in addition to the water/dryland
maps were derived from the MODIS Terra Vegetation Continuous Fields dataset
(MOD44B, \cite{Dimiceli.2015}). Combined, the layers added up to 100\% and
depicted the percentage cover of tree-vegetation (henceforth trees),
non-tree-vegetation (henceforth shrubs/grassland), and non-vegetated (henceforth
bareland). Since we wanted to make sure that areas covered by the flood would
get 0\% vegetation cover (i.e. 100\% bareland cover), we used our floodmap that
aligned with the creation date of the layers (30 March 2017) and set all flooded
pixels to 100\% bareland. The MODIS vegetation layers had a resolution of 250m x
250m, so we did not need to coarsen or interpolate them.

\subsubsection{Protection Status}
To assess whether dispersers preferred protected over unprotected (pastoral)
regions, we derived a binary layer separating the two categories. Corresponding
data on protection status was downloaded in shapefile format from the Peace
Parks Foundation \citep{PeaceParks.2019}. We simplified original categories to a
binary classification of protected and pastoral areas (see
\Cref{Reclassification} in \Cref{Appendix:Reclassification}). Protected areas
mainly included forest reserves, game reserves, wildlife management areas, and
NPs, whereas everything not covered by such categories was considered to be
pastoral area. Finally, we rasterized the two categories onto a binary raster (1
= protected, 0 = pastoral) with a resolution of 250m x 250m.

\subsubsection{Human Influence}
We depicted human influence in our study area using a proxy that incorporated
anthropogenic pressures arising from (1) human density, (2) farming, and (3)
roads. Previous research revealed that veterinary fences have no influence on
wild dog movement, so we did not include them in our analysis
\citep{Cozzi.2013b}.

% Human Density
(1) Spatial human density estimates were obtained through Facebook's 30m x 30m
high resolution population density dataset \citep{Facebook.2019}. The dataset
contained some human density that arised from lodges and other touristic
activities that we did not want to depict as human pressures, so we manually
removed such regions from the data. The resulting layer was coarsened to 250m x
250m by summing up human densities within 250m x 250m cells.

% Farming
(2) Information on agricultural activities was sourced from the Globeland
\citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets, from
which we only retained pixels that were classified as either \textit{cultivated
land} or \textit{croplands}. Both layers had a resolution of 30m x 30m and were
coarsened to 250m x 250m by summing up values within 250m x 250m cells. The
layers were then merged into a single binary layer depicting presence (= 1) or
absence (= 0) of farms within 250m x 250m cells.

% Roads
(3) Geo-referenced data on roads was downloaded in shapefile format from Open
Street Map \citep{OpenStreetMap.2019}. We omitted minor roads and only retained
main tar roads and trunks belonging to the biggest four categories (see
\Cref{Appendix:RoadCategories}). Finally, we rasterized roads and trunks onto a
binary raster (1 = roads, 0 = no roads) with 250m x 250m resolution.

To calculate a single proxy of human influence, we added up the layers
describing \textit{human density} (continuous), \textit{farming} (binary), and
\textit{roads} (binary). To reduce the influence of outliers, added values
were limited to a maximum of 50, which visually resulted in a good balance
between high and low anthropogenic influence and was therefore considered
appropriate for our analysis. To render the fact that humans influence their
surroundings beyond their presence, we followed \cite{Elliot.2014} and applied
to each pixel a 5km spatial buffer within which we summed up and log-transformed
human-influence values.

\subsection{Permeability Model}
\label{Modeling}
% Explanation of iSSF and Extraction
To parameterize a permeability model and thereby investigate movement and
habitat preferences of dispersing African wild dogs, we used an integrated step
selection function (iSSF, \cite{Avgar.2016}). In the iSSF framework, covariates
along realized steps are contrasted with covariates along random steps that the
animal could have taken, but decided not to. A step is here defined as the
connecting line between two consecutive GPS relocations \citep{Turchin.1998}. In
contrast to regular SSFs, iSSFs require to include movement metrics as
covariates in the corresponding conditional logistic regression model. Their
inclusion in turn allows simultaneous inference on habitat and movement
preferences, as well as to reduce potential biases in estimated habitat
preferences \citep{Forester.2009, Warton.2013, Avgar.2016}.

In order to conduct an iSSF analysis, we strictly followed published
recommendations described in \cite{Avgar.2016}. We prepared our GPS relocation
data for iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and
coerced relocations recorded during dispersal to steps that were regularly
spaced four hours apart (allowing for a minor mismatch of up to 15 minutes).
Steps that were separated by more than four hours (e.g. due to GPS failure) were
omitted from further analysis. Each remaining step was paired with 24 random
steps, generated by sampling turning angles from a uniform distribution
U(\(-\pi, \pi\)) and step lengths from a gamma distribution that was fitted
using realized step lengths (shape = 0.3677, scale = 6,302)). Together, a
realized and its 24 associated random steps formed a stratum and received a
unique identifier. Along realized and random steps we extracted environmental
covariates as listed in \Cref{ExtractedCovars}. For continuous covariates we
calculated the average value along the step, for binary covariates the
percentage cover along the step. Since our floodmaps were dynamic, we extracted
values from the floodmap that was closest in date to the actual step. We further
derived a binary variable indicating whether or not a step crossed a road and we
identified the average distance of the step to the nearest source of water or
road. The values for \textit{DistanceToRoads} and \textit{DistanceToWater} were
square-rooted to simulate a decreasing marginal impact of distance. Extracted
covariates were standardized to a mean of zero and a standard deviation of one
using a z-score transformation. We also screened covariates for correlation
using Pearson's Correlation Coefficient. None of the covariates were overly
correlated (\(|r| > 0.6\), \cite{Latham.2011}) and we retained all of them for
modeling. To run a proper iSSF analysis, we also included the movement metrics
\textit{cos(turning angle)} (\(cos(ta\_)\)) and \textit{log(step length)}
(\(log(sl\_)\)) as covariates in our models \citep{Avgar.2016}. The movement
metric \(cos(ta\_\)) served to describe the directionality of a step, as it
transformed the circular measure of (\(-\pi, +\pi\)) into a linear measure (-1,
1), where positive values indicated forward movements, whereas negative values
indicated backwards movements \citep{Turchin.1998}. The movement metric
\(log(sl\_\)), on the other hand, was used as an indicator of the preferred step
length. Since steps were regularly spaced by four hours, \(log(sl\_\)) can also
be interpreted as movement rate.

\begin{table}[h]
  \begin{center}
    \caption{Overview of environmental covariates extracted along steps (see
    \Cref{Sources} in \Cref{Appendix:Sources} for an overview of the sources and
    resampling methods). *The covariates \textit{Water} and \textit{Dryland}
    added up to 100\%, which is why only \textit{Water} was included as
    explanatory variable in our models. The same applied for the group
    \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland}, where we
    omitted \textit{Bareland} for modeling. Finally, from the group
    \textit{Protected} and \textit{Pastoral}, we only included
    \textit{Protected} in our models.}
    \label{ExtractedCovars}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{lllll}
      \hline
      \textbf{Category} &
        \textbf{Covariate} &
          \textbf{Description} &
            \textbf{Values} \\
              % \textbf{Source(s)} \\
      \midrule
      \multirow{5}{*}{Land Cover}
        & Water
          & Percentage cover of water along step
            & 0-100\%\\
              % & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
        & Dryland*
          & Percentage cover of dryland along step
            & 0-100\%\\
              % & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
        & DistanceToWater
          & Average distance of step to nearest water source
            & \(\geq\) 0m\\
              % & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
        & Shrubs/Grassland
          & Average non-tree vegetation along step
            & 0-100\%\\
              % & \cite{Dimiceli.2015} \\
        & Trees
          & Average tree-vegetation along step
            & 0-100\%\\
              % & \cite{Dimiceli.2015} \\
        & Bareland*
          & Average non-vegetated area along step
            & 0-100\%\\
              % & \cite{Dimiceli.2015} \\
      \hdashline
      \multirow{2}{*}{Protection Status}
        & Protected
          & Percentage cover of protected area along step
            & 0-100\%\\
              % & \cite{PeaceParks.2019} \\
        & Pastoral*
          & Percentage cover of unprotected area along step
            & 0-100\%\\
              % & \cite{PeaceParks.2019} \\
      \hdashline
      \multirow{3}{*}{Human Influence}
        & HumansBuffer5000
          & Average buffered (5km) human influence along step
            & \(\geq\) 0\\
              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
        & DistanceToRoads
          & Average distance of step to nearest road
            & \(\geq\) 0m\\
              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
        & RoadCrossing
          & Binary value indicating whether a step crossed a road
            & 0, 1\\
              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      \hline
      \end{tabular}
    }
  \end{center}
\end{table}

% Permeability Model
\noindent We used the iSSF framework and parameterized a permeability model that
further served to predict landscape permeability. This permeability model
operated under the assumption that dispersing wild dogs assigned a selection
score \(w(x)\) of the following exponential form to each realized and random
step:

\begin{equation}
\label{EQ2}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent That is, the selection score \(w(x)\) of a step depended on its
associated covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's
preferences for these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
probability that a step \(i\) was realized \(P(Y_{i} = 1\)) was then contingent
on the step's selection score, as well as on the selection scores of all
alternative steps in the stratum:

\begin{equation}
\label{EQ3}
  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
\end{equation}

% Conditional Logistic Regression Model
\noindent Habitat and movement preferences of interest, i.e. the \(\beta\)'s,
were estimated by comparing realized (scored 1) and random (scored 0) steps in a
conditional logistic regression model \citep{Fortin.2005}. In this model,
positive \(\beta\)-coefficients indicate \textit{selection} of a covariate,
negative \(\beta\)-coefficients \textit{avoidance} of a covariate. Selected
covariates therefore contribute positively to expected landscape permeability,
whereas avoided covariates reduce landscape permeability.

% Mixed Effects Conditional Logistic Regression Model
To deal with multiple individuals, we applied a novel modeling technique for
mixed effects conditional logistic regressions that was recently proposed by
\cite{Muff.2019}. Their approach allowed to include individuals for whom some
covariates did not vary, as well as to model random slopes for each covariate.
We implemented their method using the R-package \textit{glmmTMB}
\citep{Mollie.2017}, where we fixed the variance of the random intercept to an
artificially high value of 10\textsuperscript{12} (for an explanation, see
\cite{Muff.2019}) and used animal ID to model random intercepts and slopes.

% AIC Model Selection
We defined the movement metrics \(cos(ta\_)\) and \(log(sl\_)\) as core
covariates and ran forward model selection based on Akaike's Information
Criterion (AIC, \cite{Burnham.2002}) for environmental covariates. That is, we
started with a univariate model and iteratively increased model complexity upon
reaching a full model. We then ranked models according to AIC, assessed relative
model weights, and identified the most parsimonious model. Due to convergence
issues we were unable to model interactions between environmental covariates.

% Validation
To validate the predictive power of the most parsimonious permeability model, we
ran k-fold cross-validation for case-control studies as described in
\cite{Fortin.2009}. Using 80\% of randomly selected strata, we parameterized a
permeability model and predicted selection scores \(w(x)\) for all steps in the
remaining 20\% of strata. According to predicted selection scores we assigned
ranks 1-25 within each stratum, with rank 1 indicating the highest selection
score. We identified the realized step's rank in each stratum and tallied rank
frequencies for realized steps over all strata. Finally, we carried out a
Spearman-rank correlation analysis between ranks and associated frequencies and
we recorded the correlation coefficient (\(r_{s, realized}\)). We repeated this
procedure 100 times with replacement and computed the mean correlation
coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\% confidence
interval. For comparison, we also repeated the same procedure 100 times assuming
completely randomized preferences. We implemented randomized preferences by
omitting the realized step from each stratum and identifying the rank of a
randomly chosen random step within each stratum (now only ranks 1-24). Again, we
calculated Spearman's rank correlation coefficient (\(r_{s, random}\)), its mean
across repetitions (\(\bar{r}_{s, random}\)), and its 95\% confidence interval.
Ultimately, the validation proofed a significant prediction in case the
confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\)
did not overlap.

\subsection{Permeability Surface}
We used the most parsimonious permeability model and applied \Cref{EQ2} to
predict a permeability surface over the entire extent of the KAZA-TFCA. Stated
differently, we calculated the selection score \(w(x)\) for each 250m x 250m
pixel covering our extended study area given the pixel's environmental
covariates. Based on the definition of permeability, high selection scores
resulted in high permeability, whereas low selection scores resulted in low
permeability. To reduce the influence of outliers in predicted scores, we
followed \cite{Squires.2013} and curtailed predicted permeabilities between the
1st and 99th percentile of their original values. To predict permeability in the
first place, we needed to collapse all dynamic watermaps into a single layer,
which we achieved by identifying areas that were inundated in at least a quarter
of all our floodmaps. We therefore predicted permeability assuming a relatively
extended flood.

\subsection{Least-Cost Paths \& Least-Cost Corridors}
To identify movement corridors within the extent of KAZA-TFCA, we calculated
least-cost paths (LCPs) and least-cost corridors (LCCs) between source points
located in protectorates of our extended study area \citep{Adriaensen.2003,
Sawyer.2011}. Source points were generated by overlaying the study area with a
regular grid of points, equally spaced 100km apart and located within protected
areas that contiguously covered at least 700km\textsuperscript{2}. The threshold
of 700km\textsuperscript{2} was based on home-range requirements of African wild
dogs reported in \cite{Pomilia.2015} and served to eliminate locations that were
assumed unsuitable for supporting viable wild dog populations. Finally, we
defined centroids as source points for those protectorates with over
700km\textsuperscript{2} that did not get any source points from the regular
grid. In total, we generated 68 source points, which resulted in 2,278 unique
pairwise combinations of source points and therefore 2,278 unique LCPs and LCCs.

LCP analysis between source points was implemented using the R-package
\textit{gdistance} \citep{vanEtten.2018}. The package translated the
permeability surface into a network of nodes and used graph theory, in
particular Dijkstra's algorithm \citep{Dijkstra.1959}, to find shortest
effective distances between source points based on costs of moving from cell to
cell. The package thus required to convert the permeability surface to a
transition layer, where permeability values were translated into transition
probabilities for moving from one cell to another. In our case, the transition
probability of moving between two adjacent cells depended positively on their
averaged permeability. We allowed individuals to move from each cell to the
cell's 8 surrounding neighbors (i.e. Moores neighborhood) and applied a
geographic correction to account for the fact that diagonal neighbors were more
remote than orthogonal neighbors. Because African wild dogs have been observed
to cover massive dispersal distances (\citeauthor{DaviesMostert.2012}, 2012;
\citeauthor{Masenga.2016}, 2016; \citeauthor{Cozzi.2019}, in review), we did not
limit LCPs to a maximal effective cost. After computation, we tallied
overlapping LCPs to identify high-frequency routes.

LCPs are often criticized because they identify single-pixel-wide paths and thus
obliterate potential alternative routes with only marginally higher costs
\citep{Pinto.2009}. Moreover, they implicitly assume that dispersers have
complete knowledge of the landscape and associated movement costs and therefore
always move along the cheapest route \citep{Carroll.2012}. Yet, individuals
rarely use a single optimum path, so that LCPs are often of limited biological
relevance \citep{Pinto.2009, Pullinger.2010}. To address these shortcomings, we
also calculated LCCs \citep{Pinto.2009, Sawyer.2011}, again using the R-package
\textit{gdistance}. In contrast to LCPs, LCCs overcome the
single-pixel-width-issue and reveal additional, slightly suboptimal dispersal
routes. To identify LCCs, we first computed for each source point a cumulative
cost map (see \Cref{LCCExample}), which indicated the total minimal costs
required to get from the respective source point to any other location on the
map. A LCC between two source points was then obtained by adding up their
cumulative cost maps and masking out all cell-values exceeding the lowest
cell-value by more than 5\%. We repeated this procedure for each possible
pairwise combination of source points and thereby identified LCCs between all
source points. The resulting corridor-maps were normalized to a range from zero
to one using \(f(x) = \frac{x - min(x)}{max(x) - min(x)}\) and totaled up into a
single connectivity map.

% \begin{figure}[hbtp]
%   \begin{center}
%     \begin{tikzpicture}
%         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
%           \begin{minipage}{0.54\textwidth}
%             \begin{center}
%               \includegraphics[width = 0.5\textwidth]
%                 {99_LeastCostPaths(Example1).pdf}\\
%               \includegraphics[width = 1.0\textwidth]
%                 {99_LeastCostPaths(Example2).pdf}\\
%             \end{center}
%           \end{minipage}
%         };
%         \begin{scope}[x={(image.south east)},y={(image.north west)}]
%            %  % next four lines will help you to locate the point needed by
%            %  % forming a grid. comment these four lines in the final picture.↓
%            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
%            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
%            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
%            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
%            %  % upto here
%             % \draw[black, line width = 0.5pt] (0.389, 0.627) -- (0.080, 0.477);
%             % \draw[black, line width = 0.5pt] (0.582, 0.627) -- (0.980, 0.477);
%             % \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
%             % \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
%             % \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
%         \end{scope}
%     \end{tikzpicture}
%     \caption{Figures illustrating the process of identifying a least-cost
%     corridor between the source points A and B. (a) Example of a permeability
%     surface, which serves to determine the costs of movement. (b1) Cumulative
%     cost map for point A, depicting the total minimal costs necessary to get
%     from point A to every other location. (b2) Cumulative cost map for point B,
%     depicting the total minimal costs necessary to get from point B to every
%     other location. (c) Summed cost maps of points A and B. (d) Masked out
%     corridor containing pixels that do not exceed the cheapest pixel by more
%     than 5\%.}
%     \label{LCCExample}
%   \end{center}
% \end{figure}

\begin{figure}[hbtp]
  \begin{center}
  \begin{minipage}{.32\linewidth}
    \begin{subfigure}[t]{.9\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example1).pdf}
    \end{subfigure}
  \end{minipage}
  \begin{minipage}{.33\linewidth}
    \begin{subfigure}[t]{.9\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example2).pdf}
    \end{subfigure}
  \end{minipage}
  \begin{minipage}{.33\linewidth}
    \begin{subfigure}[t]{.9\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example3).pdf}
    \end{subfigure}
  \end{minipage}
  \caption{Figures illustrating the process of identifying a least-cost
  corridor between the source points A and B. (a) Example of a permeability
  surface, which serves to determine the costs of movement. (b1) Cumulative
  cost map for point A, depicting the total minimal costs necessary to get
  from point A to every other location. (b2) Cumulative cost map for point B,
  depicting the total minimal costs necessary to get from point B to every
  other location. (c) Summed cost maps of points A and B. (d) Masked out
  corridor containing pixels that do not exceed the cheapest pixel by more
  than 5\%.}
  \label{LCCExample}
  \end{center}
\end{figure}

\newpage
\section{Results}
\subsection{Permeability Model}
Our most parsimonious permeability model retained the covariates \textit{Water},
\textit{DistanceToWater}, \textit{Trees}, \textit{Shrubs/Grassland},
\textit{HumansBuff5000}, and, of course, the fixed covariates \(cos(ta\_)\) and
\(log(sl\_)\) (see \Cref{PermeabilityResults}a). The model strictly outperformed
all alternative models (\(\Delta AIC > 2\)) and therefore received an AIC-weight
of one (see \Cref{ModelAICs} in \Cref{AICs}). Except for the effect of
\textit{DistanceToWater}, all estimates were significant on the 95\%-confidence
interval (CI). Parameter estimates from the model showed that dispersing wild
dogs preferred low turning angles (i.e. high \(cos(ta\_)\)), indicating that
they moved in a linear manner during dispersal (\(\beta\) = 0.14; 95\% CI = 0.07
to 0.21). Similarly, they preferred a high movement rate, as evidenced by a
positive selection coefficient for \(log(sl\_)\) (\(\beta\) = 0.05, 95\% CI =
0.02 to 0.09). With respect to land cover, dispersers strictly avoided moving
through water and therefore preferably traversed dryland when available
(\(\beta\) = -0.5, 95\% CI -0.67 to -0.32). Although the effect was not
statistically clear, dispersers preferred locations in proximity to water, as
the estimated parameter for \textit{DistanceToWater} was negative (\(\beta\) =
-0.24, 95\% CI = -0.67 to 0.19). Furthermore, dispersers avoided areas that were
densely covered by trees (\(\beta\) = -0.18, CI = -0.32 to -0.04) and preferred
areas covered by shrubs/grassland (\(\beta\) = 0.40, 95\% CI = 0.24 to 0.56).
Finally, dispersers strictly avoided areas that were strongly influenced by
humans (\(\beta\) = -0.47, 95\% CI = -0.87 to -0.07).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    permeability model. Negative coefficients indicate avoidance of a covariate,
    positive coefficients selection of a covariate. Whiskers delineate the
    95\%-CIs for estimated parameters. The intercept is meaningless in the iSSF
    framework and was thus omitted. (b) Results from the k-fold cross validation
    for case control studies. The left graph shows rank frequencies of
    \textit{realized} steps according to predictions, whereas the right graph
    shows rank frequencies of \textit{randomly selected} steps according to
    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
    resulting from 100 repetitions of the k-fold cross validation. The blue
    smoothing line was fitted using a LOESS regression and solely serves to aid
    the eye in detecting the trends. Correlation coefficients suggest that our
    prediction is significant and robust, evidenced by the fact that the
    confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
    random}\) do not overlap and that there is strong and significant
    correlation between ranks and associated frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

\newpage
\noindent Results from the k-fold cross-validation suggested that our prediction
was significant and robust, as highlighted by the fact that the 95\%-CIs
intervals of \(\bar{r}_{s, realized}\) (-0.45, CI = -0.48 to -0.42) and
\(\bar{r}_{s, random}\) (0.04, CI = 0.01 to 0.08) did not overlap (see
\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations.

\subsection{Permeability Surface}
\Cref{PermeabilityMap}a illustrates the predicted permeability surface for the
extent of the KAZA-TFCA as computed from the most parsimonious permeability
model. Note that \(cos(ta\_)\) and \(log(sl\_)\) did not enter the prediction of
landscape permeability as their inclusion in our permeability model solely
served to reduce the bias in estimated habitat preferences. Our prediction
revealed substantial regional differences in predicted permeability across the
entire study area, with some regions being highly permeable whereas others are
almost insurmountable. A comparisons of median (Mdn) permeabilities across
countries (see \Cref{PermeabilityComp}) showed that in particular Angola (Mdn =
1.03) and Botswana (Mdn = 0.91) support highly permeable landscapes, whereas
Zimbabwe (Mdn = 0.19) and Zambia (Mdn = 0.15) were found to be largely
impermeable. Namibia (Mdn = 0.59), on the other hand, ranges somewhere in
between the two extremes. Throughout all countries, permeabilities inside areas
covered by the KAZA-TFCA were predicted to be vastly greater than permeabilities
outside areas covered by the KAZA-TFCA, resulting in a stark contrast of the
overall permeability inside the KAZA-TFCA (Mdn = 0.75) and outside the KAZA-TFCA
(Mdn = 0.26). Similarly, for all countries we found that protected areas provide
higher permeabilities than pastoral regions, altogether making protected areas
highly permeable (Mdn = 0.74) in comparison to pastoral regions (Mdn = 0.27). In
general, northern Botswana appeared to provide the highest density of permeable
landscapes, as judged by the bright yellow colors in \Cref{PermeabilityMap}a. A
closer look at this region (see \Cref{PermeabilityMap}b) showed that especially
close surroundings of the Okavango Delta and the Linyanti Swamp host permeable
landscapes for dispersers. Nevertheless, the waters of the Okavango Delta and
Linyanti Swamp themselves provide highly impermeable landscapes, as indicated by
the dark colors sketching the two water-bodies. Moreover, densely populated
areas such as the city of Maun or the villages Khwai, Mababe, and Sankuyo
provide low permeability, which is also true for the highly populated stripe
flanking the Okavango Delta towards its west and north-west.

\newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
            \includegraphics[width = \textwidth]{99_PermeabilityMap2.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
           %  % next four lines will help you to locate the point needed by
           %  % forming a grid. comment these four lines in the final picture.↓
           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
           %  % upto here
            \draw[black, line width = 1pt] (0.389, 0.627) -- (0.080, 0.477);
            \draw[black, line width = 1pt] (0.582, 0.627) -- (0.980, 0.477);
            \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
            \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
            \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Predicted permeability surface for the extent of the KAZA-TFCA.
    Permability was predicted by calculating selection scores \(w(x)\) for each
    pixel based on the pixel's underlying covariates. Areas that dispersers find
    easy to traverse are depicted in bright colors. Bold white lines delineate
    the borders of the KAZA-TFCA, whereas dashed white lines show country
    borders. (b) Detailed view of the Okavango Delta and Linyanti Swamp region,
    which appears to provide the most suitable landscape for dispersers. The
    white square highlights our core study area.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}
\restoregeometry

\begin{table}[h]
  \caption{Comparison of median permeability across countries, separated
  according to areas within and outside the KAZA-TFCA, as well as within and
  outside protected areas. High values indicate high permeability, whereas low
  values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{center}
    \resizebox{0.65\textwidth}{!} {
      \begin{tabular}{llllllllr}
      & & \multicolumn{2}{c}{KAZA-TFCA} && \multicolumn{2}{c}{Protection Status} \\
      \cline{3-4} \cline{6-7}
      Country & & Inside & Outside & & Protected & Pastoral & & Overall \\
      \hline
      Angola & & 1.67 & 0.59 & & 1.67 & 0.60 & & 1.03 \\
      Namibia & & 0.91 & 0.49 & & 0.98 & 0.44 & & 0.59 \\
      Botswana & & 1.17 & 0.69 & & 1.37 & 0.62 & & 0.91 \\
      Zimbabwe & & 0.25 & 0.18 & & 0.32 & 0.17 & & 0.19 \\
      Zambia & & 0.26 & 0.12 & & 0.23 & 0.13 & & 0.15 \\
      \hline
      Overall & & 0.75 & 0.26 & & 0.74 & 0.27 & & 0.38 \\
      \end{tabular}
    }
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
\Cref{LeastCost}a shows predicted LCPs and \Cref{LeastCost}b corresponding LCCs
for the extent of the KAZA-TFCA. Because LCPs identify the cheapest set of
pixels between two source points, they typically run along the center of LCCs
and traverse the same regions. The two figures thus essentially contain the same
information, albeit \Cref{LeastCost}b appears more detailed and reveals some
routes that would remain undetected in \Cref{LeastCost}a. Most notably, our
connectivity analysis revealed that several LCPs and LCCs traverse the
Chobe-Moremi region in northern Botswana (\Cref{LeastCost}a \circled{1}). In
fact, the highest frequented route in this region is passed by 637 LCPs, which
is reflected in a wide, dark red LCC in \Cref{LeastCost}b. Departing from this
central region, several high-frequency LCPs and LCCs emerge. One route
(\circled{2}) runs towards the north-western part of the KAZA-TFCA into the
Mavinga and Luengue-Luiana NPs. A further route (\circled{3}) meanders
north-east through unprotected area into the Kafue NP in Zambia. Another major
route (\circled{4}) extends east, into the Hwange NP in Zimbabwe and eventually
reaches the Matusadona NP (\circled{5}). A slightly less pronounced but still
distinguishable route (\circled{6}) originates from the Chobe-Moremi ecosystem
and runs south, through unprotected landscape, into the Central Kalahari NP in
Botswana. Alternatively, the corridor turns west (\circled{8}) and moves into
the Khaudum NP in Namibia (\circled{7}). Finally, another slightly less
accentuated path originates from southern Angola, moves through Botswana and
leads into the Khaudum NP in Namibia. Besides these medium- to high-frequency
routes, several low-frequency corridors exist and provide connectivity between
other protected areas. Strikingly, we found few direct LCPs and LCCs between
peripheral source points. That is, most paths and corridors connecting two
adjacent peripheral points detour through more central regions before they head
towards their destination at the periphery. As a consequence, there are several
peripheral regions providing little to no connectivity, as they are not crossed
by any LCPs or LCCs. Such regions are better visible in \Cref{LeastCost}b where
they appear as extended dark areas. Overall, there is a gradual decrease in the
density of LCPs and LCCs when moving from the central region towards the border
of the extended study area.

\newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \includegraphics[width = \textwidth]{99_LeastCostPaths.pdf}
            \includegraphics[width = \textwidth]{99_LeastCostCorrs.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
          % % next four lines will help you to locate the point needed by
          % % forming a grid. comment these four lines in the final picture.
          % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
          % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
          % \foreach \x in {0,1,...,9} {\node [anchor=north] at (\x/10,0) {0.\x};}
          % \foreach \y in {0,1,...,9} {\node [anchor=east] at (0,\y/10) {0.\y};}
          % % upto here
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No1) at (0.400, 0.700){1};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No2) at (0.370, 0.760){2};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No3) at (0.585, 0.780){3};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No4) at (0.647, 0.697){4};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No5) at (0.910, 0.770){5};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No6) at (0.580, 0.658){6};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No7) at (0.410, 0.640){7};
          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
            , draw=black] (No8) at (0.380, 0.665){8};
          \draw[black, line width = 1pt, densely dotted] (No1) -- (0.500, 0.710);
          \draw[black, line width = 1pt, densely dotted] (No2) -- (0.438, 0.760);
          \draw[black, line width = 1pt, densely dotted] (No3) -- (0.565, 0.805);
          \draw[black, line width = 1pt, densely dotted] (No4) -- (0.665, 0.685);
          \draw[black, line width = 1pt, densely dotted] (No5) -- (0.885, 0.765);
          \draw[black, line width = 1pt, densely dotted] (No6) -- (0.540, 0.650);
          \draw[black, line width = 1pt, densely dotted] (No7) -- (0.450, 0.645);
          \draw[black, line width = 1pt, densely dotted] (No8) -- (0.345, 0.690);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (light grey). Note that only contiguous
    protected areas covering more than 700km\textsuperscript{2} are depicted.
    Circled numbers indicate some of the highest frequented least-cost paths
    that are mentioned in the main text. Continuous black lines indicate the
    borders of the KAZA-TFCA, whereas dashed black lines delineate
    country-borders. (b) Least-cost corridors between the same source points as
    illustrated in subfigure (a). For ease of spatial reference we also labeled
    national parks that are mentioned in the main text (dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}
\restoregeometry

\newpage
\section{Discussion}
The goal of our study was to (1) assess landscape permeability for dispersing
wild dogs and to (2) subsequently use this knowledge to examine connectivity
across the greater extent of the KAZA-TFCA. To achieve these goals, we
parameterized a permeability model in which we used GPS data of 16 dispersing
wild dog coalitions to quantify their habitat preferences with respect to land
cover, protection status, and human influence. While we found significant
impacts of different land covers and human influence, the influence of
protection status was not statistically clear. We further applied the
parameterized permeability model and predicted a permeability surface for the
entire KAZA-TFCA, which depicted the ease at which dispersers crossed a specific
landscape. Finally, based on this permeability map, we computed LCPs and LCCs
between protectorates and thereby highlighted dispersal corridors of African
wild dogs.

\subsection{Permeability Model}
% Permeability Model: Land Cover, Water
Our most parsimonious permeability model revealed that, with respect to land
cover, dispersers strictly avoided areas covered by water (including swamps,
wetlands). This result corroborates previous research on wild dogs (e.g.
\cite{Abrahms.2017}) and is rather unsurprising, as the species is not well
adapted for moving through water. For residents, water is almost impermeable and
frequently serves as natural boundary to their home-ranges \citep{Cozzi.2013}.
It is conceivable that dispersers are more willing than residents to cross
waterbodies (\citeauthor{Cozzi.2019}, in review), but we did not include
residents in our analyses and could not test for such differences. Regardless of
that, water appears to pose a serious obstacle for dispersers and is therefore
likely to reduce landscape permeability for dispersing wild dogs.

% Permeability Model: Land Cover, DistanceToWater
While dispersers avoided moving \textit{through} water, they preferred moving
\textit{close to} water as indicated by avoidance of high distance to water.
While its effect was insignificant, inclusion of the covariate
\textit{DistanceToWater} strictly increased the AIC in all of our models and
suggests that distance to water influenced permeability negatively. We
hypothesize that high prey-availability (e.g. impala, kudu) in proximity to
water explains this result. It is well documented that water is a major driver
of wildlife and that it determines the distribution of many herbivores,
especially in dry biomes \citep{Western.1975}. For the Okavango Delta in
particular, it has been shown that prey abundance strongly correlates with the
extent of the flood \citep{Bonyongo.2005} and that an increasing concentration
of ungulates around remaining water-sources during the dry season might attract
dispersing wild dogs \citep{Ogutu.2014}. Following the same logic, however, apex
predators (e.g. lions, spotted hyenas) and resident wild dogs may also prefer
closeness to water \citep{Valeix.2009}, which may occasionally force dispersers
to move distant to water \citep{Ndaimani.2016} to avoid intraguild competition
\citep{Creel.1996, Mills.1997}. We could not control for the presence or absence
of apex predators or conspecifics because we lacked corresponding data, but
these considerations might explain why we did not find a statistically clear
effect of distance to water.

% Permeability Model: Land Cover, Trees & Shrubs
In line with existing literature on moving wild dogs \citep{Abrahms.2017}, we
found that dispersers avoided habitats with high percentages of tree cover,
whereas they preferred areas with high percentages of shrubs/grassland cover.
Comparable preferences were reported for residents, which typically reside in
savanna-like habitats but rarely inhabit areas with dense tree cover
\citep{Estes.2012}. This may be explained by wild dogs' hunting behavior, where
open habitats are favored because they offer increased visibility and thus
facilitate hunting \citep{Estes.1967}. Additionally, we found a strong negative
correlation between percentage cover by shrubs/grassland and percentage cover by
bareland (although we did not include bareland in our models). The observed
preference for shrubs/grassland could therefore also indicate an aversion to
bareland. Again, we hypothesize that this aversion is driven by low
prey-availability in non-vegetated habitats \citep{Pettorelli.2011}.
Unfortunately, we lacked detailed vegetation data and could not distinguish
vegetation types further.

% Permeability Model: Human Influence
Finally, considering human influence, we found that dispersers strongly avoided
human pressures arising from human density, farming, and roads. Previous
research on dispersing wild dogs was inconclusive about the impact of human
presence on permeability, as some researchers reported a high willingness of
dispersers to cross human-dominated landscapes \citep{DaviesMostert.2012},
whereas others found that humans are avoided almost entirely during dispersal
\citep{Masenga.2016}. Our results support the notion that dispersers strongly
avoid human-dominated areas, but we hypothesize that this finding depends on the
availability of alternative routes. The dipsersers monitored by
\cite{DaviesMostert.2012} originated from habitats almost exclusively surrounded
by humans, whereas our dipsersers departed from regions with various
possibilities to move through natural habitats. A lack of alternative routes
combined with strong inbreeding avoidance intrinsic in African wild dogs
\citep{McNutt.1996} might have forced the dispersers reported in
\cite{DaviesMostert.2012} to cross human-dominated landscapes despite a strong
aversion against humans. Although we found that our dispersers strictly avoided
humans, \citeauthor{Cozzi.2019} (in review) reports that human persecution
remains the main cause of mortality for dispersers, suggesting that avoidance is
not strong enough to prevent human-induced fatalities. We argue that human
influence remains a source of reduced landscape permeability and that the
proceeding transformation of natural landscapes into human altered habitats
continues to undermine wild dogs' ability to dipserse. Since our human influence
proxy did not include minor gravel roads, our finding that humans are avoided
does not contradict the observation that wild dogs preferably move on gravel
roads during activity \citep{Abrahms.2016}. With four-hourly GPS relocations we
felt that the temporal resolution was insufficient to depict such fine scale
effects \citep{Thurfjell.2014}.

\subsection{Permeability Surface}
% Permeability Surface
Our prediction of landscape permeability revealed vast regional differences.
Across the entire study area, northern Botswana stood out as it comprises the
highest density of highly permeable landscapes. The hospitality of this region
is caused by a combination of low human influence, low tree cover, high
shrubs/grassland cover and close distance to water. Thus, although swamps,
wetlands, and water (e.g. Okavango Delta, Linyanti Swamp) in northern Botswana
provide little permeability, their surroundings act as strong attractants to
dispersers. It is therefore not very surprising that the region serves as source
for the recolonization of surrounding habitats \citep{Cozzi.2013}. Moreover, we
found much higher permeability inside the KAZA-TFCA than outside its borders,
which suggests a good fit between the planned protectorates and areas that are
suitable for dispersal. We made similar observations for protection status,
where predicted permeabilities inside protectorates are much higher than in
pastoral regions. Yet, several highly permeable regions remain uncovered by the
KAZA-TFCA or other protected areas, proposing exciting opportunities for future
expansions of already existing protectorates. Across countries, we revealed that
Zambia and Zimbabwe offer the least permeable landscapes. Visual inspection of
our covariate-layers disclosed that this is mainly owed to severe anthropogenic
influence in Zambia and Zimbabwe. In contrast, Botswana and Angola host vast
uninhabited areas, resulting in more permeable landscapes for dispersers.
However, all of our predictions were based on observations from dispersing wild
dogs departing from northern Botswana, which might have introduced a bias in
estimated habitat preferences and resulting predictions.

\subsection{Least-Cost Paths \& Least-Cost Corridors}
% Least Cost Paths & Least Cost Corridors
Our LCP- and LCC-analyses disclosed several major movement corridors within the
extended study area and suggested a decent congruence between predicted
corridors and the layout of the KAZA-TFCA. That is, only few corridors run
outside the planned boundaries of the KAZA-TFCA and several currently
unprotected corridors will become protected with realization of the KAZA-TFCA.
For example, the high-frequency corridor running from southern Zambia into the
Kafue NP (\Cref{LeastCost}a \circled{3}) currently contains extended unprotected
sections that will be transformed into protectorates under the KAZA-TFCA. We
thus expect that dispersal between such regions will be vastly facilitated
thanks to the KAZA-TFCA. Our network of LCPs and LCCs furthermore highlighted
the Chobe-Moremi system in northern Botswana, which is traversed by several
LCPs and LCCs and seems to hold the potential to serve as central hub through
which dispersers gain access to more remote regions of the KAZA-TFCA. Still, our
predictions also indicated that movements between wild dog subpopulations may
often involve extended detours. This is especially the case for regions in the
northern sectors of the KAZA-TFCA. For instance, only few LCPs connect the
Zimbabwean and Zambian sectors, such that moving directly from the Matusadona NP
to the Kafue NP seems almost impossible and likely requires to make a detour
through the Chobe-Moremi region. Similarly, few LCPs directly link the Mavinga
NP in Angola and the Kafue NP in Zambia. Moreover, we discovered potential
dispersal routes between the Chobe-Moremi ecosystem and the Central Kalahari
NP that are not yet proclaimed protected areas. Nonetheless, a protected link
between the Central Kalahari NP and the Chobe-Moremi ecosystem might grant
southern populations invaluable access to the Chobe-Moremi-Hub and
subsequently to the entire KAZA-TFCA. In fact, \cite{Elliot.2014} identified a
similar corridor for dispersing lions, proposing that protecting the area would
benefit lions and wild dogs alike. The absence of direct LCPs and LCCs between
some regions of the KAZA-TFCA is likely owed to an impermeable landscape between
those regions and demonstrates that protection of such areas in their current
state would be futile. That is, impermeable landscapes first need to be
transformed into more permeable habitats before protection pays off. Although,
we noted that connectivity gradually decreases towards the borders of our study
area, this might partly be owed to boundary-effects \citep{Koen.2010}.
Interestingly, our network of LCPs and LCCs shows surprising similarities to
predicted movement corridors of dispersing lions \citep{Elliot.2014}.
Remarkably, corridors that we identified as high-frequency corridors are also
visible in the network presented by \cite{Elliot.2014}. This not only reinforces
confidence in our own predictions but it also suggests potential synergies
between conservation of the two species.

\subsection{Future Research}
% Future Research: Genetic Resistance
While we assessed landscape permeability purely based on GPS relocation
data, future studies could challenge and test our predictions using genetic data
\citep{Spear.2010}. Similar comparisons have been conducted for other species
and revealed a surprisingly high overlap between predicted landscape
permeability based on genetic and movement data \citep{Cushman.2010}. Genetic
data would also deepen our understanding of the degree to which subpopulations
admix and of the overall genetic diversity found in different subpopulations
\citep{Girman.2001}. Based on our connectivity network, we would expect to find
a low genetic diversity for populations that live in insufficiently connected
habitats, whereas populations residing at dispersal hotspots, such as the
Chobe-Linyanti ecosystem, should exhibit a relatively diverse gene-pool.

% Future Research: Better Covariates
Future research should also strive to improve representation of geo-referenced
covariates. Because we predicted connectivity for such an extensive area, we
were strongly limited in available and reliable land-cover data. As a result,
our representation of vegetation remained rather basic with only three
vegetation types (tree-cover, shrubs/grassland, bareland). Although there are a
few large-scale, high-resolution land cover datasets, such as Globeland's 30m
\citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019} land
cover datasets, we found their classifications often erroneous for our study
area. This is probably owed to the KAZA-TFCA's unique landscape and
corresponding difficulties in remote sensing satellite imagery using traditional
indices \citep{Wolski.2017}. Nevertheless, land cover is a very powerful
predictor of animal movement \citep{Thurfjell.2014}, which stresses the need for
more reliable land-cover data.

% Future Research: Competitor Presence
In a similar manner, we would desire better knowledge of the whereabouts of
other species. In particular, we would desire location data of apex predators,
potential prey, and conspecifics of wild dogs. It is well documented that
resident wild dogs frequently compete with other carnivores \citep{Creel.1998}
and we would expect to find competition between apex predators and dispersers
too. Furthermore, knowledge about the spatial distribution of prey could verify
our hypothesis that dispersers move along water due prey-availability. Lastly,
we could not control for the likelihood of encountering conspecifics during
dispersal . Yet, for dispersing meerkats, a species with surprisingly similar
social structure to wild dogs, it has been shown that the social landscape, i.e.
the presence or absence of conspecifics, strongly shapes behavior during
dispersal \citep{Cozzi.2018}. More specifically, it has been shown that
dispersers select for high density of conspecifics while still inside their
home-range, but that they avoid high densities of conspecifics once they leave
their natal home-range (we replicated this analysis for our core study area; see
\Cref{Appendix:SocialLandscape}). Spatially explicit knowledge about the
distribution of foreign wild dogs is thus desirable and would certainly deepen
our understanding of dispersal behavior. However, constructing meaningful and up
to date social landscapes is extremely challenging and requires vast amounts of
movement data, especially for such a wide-ranging species as the African wild
dog \citep{Pomilia.2015}.

% Global Resistance Mapping
Besides urging for more and better data, we also want to emphasize that our
analytical approach is neither limited to African wild dogs, nor to our study
area. All covariates used throughout this study are available on a global scale
and are likely to be important determinants of landscape permeability for other
species. Our framework could thus easily be used to assess connectivity for a
wide range of species, provided suitable relocation data of the species exists.
Expanding our analyses to other species would surely yield interesting insights
on the congruence of movement corridors of different species and thereby
highlight areas that are used by multiple species for dispersal.

% Simulating dispersal
To this end, we identified movement corridors between pre-defined start and
end-points. However, pre-defining end-points might not even be necessary, as one
could easily invert a parameterized iSSF and use it as a mechanistic movement
model to simulate dispersal. That is, one could define a set of source points at
locations where wild-dogs are assumed or known to reside and simulate dispersers
according to habitat-preferences that were derived using the iSSF framework.
End-points would not be restricted to a certain location and movement corridors
would emerge naturally as the result of a myriad of simulated dispersal events
(for an example of simulated dispersal events see
\Cref{Appendix:DispersalSimulation}).

\newpage
\subsection{Conclusion}
Our results demonstrate that indeed the majority of dispersal corridors are
covered by the planned protectorates of the KAZA-TFCA, suggesting that it will
severely contribute to the long-term viability of this species. Nevertheless,
several dispersal routes do exist outside the KAZA-TFCA and other protectorates
and suggest opportunities for further expansions in the future. Moreover, our
connectivity network of least-cost paths and least-cost corridors reveals
especially the Chobe-Linyanti region as central dispersal hub through which
dispersers gain access to more remote regions of the KAZA-TFCA. We therefore
stress the importance of northern Botswana for the conservation of this
charismatic species. Finally, our investigations showed that human pressure
remains a severe impediment to dispersal and strictly reduces landscape
connectivity. Successful conservation of the African wild dog is thus contingent
on the willingness of local authorities, policy makers, and land managers to
provide vast areas that are liberated from human strains. Ultimately, our work
contributes to the growing field of connectivity literature and has important
implications for the management and conservation of the endangered African wild
dog.

\newpage
\section{Acknowledgements}
All of this work would not have been possible without the aid of several helpful
people. I would first like to thank Prof. Arpat Ozgul, the leader and mastermind
of our research group, whose expertise in statistics, biological processes and
scientific writing was invaluable for my work. Moreover, I want to thank
Gabriele Cozzi for his excellent guidance, all the interesting discussions, his
constructive criticism, and of course his constant supply of freshly brewed
coffee. Particular gratitude goes to Dominik Behr who supervised my work on a
daily basis and greatly contributed to all aspects of the thesis. I genuinely
enjoyed the frequent meetings and brainstorms and I really feel privileged to
have had the opportunity of working with such a great scientist. I am also
greatly indebted to Prof. John Fieberg, who visited our department this summer
and invested a great amount of time to consult and improve all statistical
aspects of the thesis, as well as to deepen my personal understanding of
movement ecology. Furthermore, I thank the Okavango Research Institute, in
particular Piotr Wolski, for sharing all of their Python-scripts, floodmaps and
knowledge on developing dynamic floodmaps. Also, I want to thank Stefan Sommer
for sharing and teaching his hard earned knowledge on scientific writing.
Additionally, I want to thank all of my colleague master students who have been
wonderful companions and made the time at university a blast. Likewise, I thank
my parents for supporting all of my undertakings and for providing first class
shelter and food. Last, but most definitely not least, I also greatly thank my
girlfriend, who was always there for me and provided welcome distractions for my
mind.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\newpage
\appendix
\section{Appendix}
\subsection{Typical Flood Pulse}
\label{Appendix:FloodPulse}
\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
    \caption{Sequence of figures showing a typical flood pulse throughout the
    year. The flood arrives from the north-western corner (so called
    ``pan-handle'') of the Okavango Delta and slowly descents through the delta
    in south-eastern direction, where it nourishes several tributaries. The
    extent of the flood peaks around August or September, after which it slowly
    retracts. Between December and March the reflectance properties of water and
    dryland change, which is why often no accurate floodmaps can be obtained for
    these months using remote sensing techniques.}
    \label{FloodPulse}
  \end{center}
\end{figure}

\newpage
\subsection{Reclassification of Protected Areas}
\label{Appendix:Reclassification}
\begin{table}[h]
  \caption{Reclassification of protected areas. Data on protection status was
  obtained from the \cite{PeaceParks.2019}. We simplified original categories to
  a binary classification of protected and unprotected (pastoral) areas. Any
  area that was not represented by one of the categories listed in
  \Cref{Reclassification} was deemed to be pastoral landscape.}
  \label{Reclassification}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llp{11cm}}
      \toprule
      \textbf{Original Category} &
        \textbf{Reclassification} &
          \textbf{Comment} \\
      \midrule
      Communal Conservancy &
        Protected &
          No comment\\
      Community Forest &
        Protected &
          No comment\\
      Conservancy &
        Protected &
          No comment\\
      Forest Reserve &
        Protected &
          No comment\\
      Game Management Area &
        Protected &
          No comment\\
      Game Park &
        Protected &
          No comment\\
      Game Ranch &
        Protected &
          No comment\\
      Game Reserve &
        Protected &
          No comment\\
      Hunting Reserve &
        Protected &
          No comment\\
      National Park &
        Protected &
          No comment\\
      National Park (not yet proclaimed) &
        \textit{Removed} &
          Since not yet proclaimed, we did not include this area as protected\\
      Overlap (GMA/FR) &
        Protected &
          Mixture between game management area and forest reserve\\
      Proposed &
        Protected &
          No comment\\
      Protected Area &
        Protected &
          No comment\\
      Recreation Park &
        \textit{Removed} &
          Marine reserve\\
      Safari Area &
        Protected &
          No comment\\
      Sanctuary &
        Protected &
          No comment\\
      State Forest &
        Protected &
          No comment\\
      Wildlife Management Area &
        Protected &
          No comment\\
      World Heritage Site &
        Protected &
          No comment\\
      Freehold Conservancy &
        Protected &
          No comment\\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Road Categories}
\label{Appendix:RoadCategories}
\begin{table}[h]
  \caption{Description of road types, as sourced from Open Street Map's mapping
  guide (\url{https://wiki.openstreetmap.org/wiki/Key:highway}). Roads types
  that were considered for the purpose of this study are shaded in light gray.}
  \label{RoadsDescription}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llp{13cm}}
      \toprule
      \textbf{Group} &
        \textbf{Subgroup} &
          \textbf{Description} \\
      \midrule
      \rowcolor[gray]{0.80}
      Roads &
        motorway &
          A restricted access major divided highway, normally with 2 or more
          running lanes plus emergency hard shoulder. Equivalent to the
          Freeway, Autobahn, etc. \\
      \rowcolor[gray]{0.80}
      Roads &
        trunk &
          The most important roads in a country's system that aren't
          motorways. Need not necessarily be a divided highway. \\
      \rowcolor[gray]{0.80}
      Roads &
        primary &
          The next most important roads in a country's system. Often link
          larger towns. \\
      \rowcolor[gray]{0.80}
      Roads &
        secondary &
          The next most important roads in a country's system. Often link
          towns. \\
      Roads &
        tertiary &
          The next most important roads in a country's system. Often link
          smaller towns and villages \\
      Roads &
        unclassified &
          The least important thorough roads in a country's system, i.e. minor
          roads of a lower classification than tertiary, but which serve a
          purpose other than access to properties. Often link villages and
          hamlets. \\
      Roads &
        residential &
          Roads which serve as an access to housing, without function of
          connecting settlements. Often lined with housing. \\
      Roads &
        service &
          For access roads to, or within an industrial estate, camp site,
          business park, car park etc. \\
      \rowcolor[gray]{0.80}
      Link roads &
        motorway\_link &
          The link roads (sliproads/ramps) leading to/from a motorway from/to
          a motorway or lower class highway. Normally with the same motorway
          restrictions. \\
      \rowcolor[gray]{0.80}
      Link roads &
        trunk\_link &
          The link roads (sliproads/ramps) leading to/from a trunk road
          from/to a trunk road or lower class highway. \\
      \rowcolor[gray]{0.80}
      Link roads &
        primary\_link &
          The link roads (sliproads/ramps) leading to/from a primary road
          from/to a primary road or lower class highway. \\
      \rowcolor[gray]{0.80}
      Link roads &
        secondary\_link &
          The link roads (sliproads/ramps) leading to/from a secondary road
          from/to a secondary road or lower class highway. \\
      Link roads &
        tertiary\_link &
          The link roads (sliproads/ramps) leading to/from a tertiary road
          from/to a tertiary road or lower class highway. \\
      Special road types &
        living\_street &
          For living streets, which are residential streets where pedestrians
          have legal priority over cars, speeds are kept very low and where
          children are allowed to play on the street. \\
      Special road types &
        pedestrian &
          For roads used mainly/exclusively for pedestrians in shopping and
          some residential areas which may allow access by motorised vehicles
          only for very limited periods of the day. \\
      Special road types &
        track &
          Roads for mostly agricultural or forestry uses. \\
      Special road types &
        bus\_guideway &
          A busway where the vehicle is guided by the way (though not a railway)
          and is not suitable for other traffic. \\
      Special road types &
        escape &
          For runaway truck ramps, runaway truck lanes, emergency escape
          ramps, or truck arrester beds. It enables vehicles with braking
          failure to safely stop. \\
      Special road types &
        raceway &
          A course or track for racing \\
      Special road types &
        road &
          A road/way/street/motorway/etc. of unknown type. It can stand for
          anything ranging from a footpath to a motorway. \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Covariate Sources}
\label{Appendix:Sources}
\begin{table}[h]
  \begin{center}
    \caption{Sources from which we obtained our set of geo-referenced
    covariates, as well as an indicator of their original resolution and the
    method used for adjusting the resolution to 250m x 250m.}
    \label{Sources}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{llp{4cm}p{4cm}p{7cm}}
      \hline
      \textbf{Category} &
        \textbf{Covariate} &
          \textbf{Resolution} &
            \textbf{Resampling} &
              \textbf{Source(s)} \\
      \midrule
      Land Cover
        & Water
          & 30m x 30m (static), 500m x 500m (dynamic)
            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
      Land Cover
        & Dryland
          & 30m x 30m (static), 500m x 500m (dynamic)
            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
      Land Cover
        & DistanceToWater
          & 30m x 30m (static), 500m x 500m (dynamic)
            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
      Land Cover
        & Shrubs/Grassland
          & 250m x 250m
            & NA
              & \cite{Dimiceli.2015} \\
      Land Cover
        & Trees
          & 250m x 250m
            & NA
              & \cite{Dimiceli.2015} \\
      Land Cover
        & Bareland
          & 250m x 250m
            & NA
              & \cite{Dimiceli.2015} \\
      \hdashline
      Protection Status
        & Protected
          & NA
            & NA
              & \cite{PeaceParks.2019} \\
      Protection Status
        & Pastoral
          & NA
            & NA
              & \cite{PeaceParks.2019} \\
      \hdashline
      Human Influence
        & HumansBuffer5000
          & 30m x 30m
            & Sum of each 250m x 250m cell
              & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      Human Influence
        & DistanceToRoads
          & NA
            & NA
              & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      Human Influence
        & RoadCrossing
          & NA
            & NA
              & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      \hline
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Model Selection Results}
\label{AICs}
\begin{table}[h]
  \caption{Results from the forward model selection procedure based on AIC
  \citep{Burnham.2002} for our permeability model. Weights were calculated
  using: \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i =
  1}^n e^{-0.5} \Delta \text{AIC}_i} \). The most parsimonious model (shaded in
  light gray) outperformed all other models (\(\Delta AIC > \) 2) and received a
  weight of one. The model in the last row did not converge and no AIC value
  could be obtained. (*) represents the fixed covariates \(cos(ta\_)\) and
  \(log(sl\_)\).}
  \label{ModelAICs}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{lllll}
      \toprule
      \textbf{Covariates} &
        \textbf{AIC} &
          \textbf{\(\Delta\)AIC} &
            \textbf{AIC-Weight} &
              \textbf{LogLik} \\
      \midrule
      \rowcolor[gray]{0.80}
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees &
        89841.85 &
          0.00 &
            1.00 &
              -44905.93 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, DistanceToRoads &
        89844.46 &
          2.61 &
            0.00 &
              -44905.23 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, Protected &
        89845.78 &
          3.93 &
            0.00 &
              -44905.89 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000 &
        89847.96 &
          6.11 &
            0.00 &
              -44910.98 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees,
      RoadCrossing &
        89848.24 &
          6.38 &
            0.00 &
              -44905.12 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees,
      DistanceToRoads, Protected &
        89848.45 &
          6.60 &
            0.00 &
              -44905.22 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, Trees
        & 89850.51 &
          8.66 &
            0.00 &
              -44912.26 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000,
      DistanceToRoads &
        89850.81 &
          8.96 &
            0.00 &
              -44910.40 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, DistanceToRoads, RoadCrossing &
        89850.87 &
          9.02 &
            0.00 &
              -44904.44 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Protected &
        89851.91 &
          10.06 &
            0.00 &
              -44910.96 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, RoadCrossing &
        89854.42 &
          12.57 &
            0.00 &
              -44910.21 \\
      (*), Water, DistanceToWater, Shrubs/Grassland &
        89856.89 &
          15.03 &
            0.00 &
              -44917.44 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, DistanceToRoads &
        89858.78 &
          16.92 &
            0.00 &
              -44916.39 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, Protected &
        89860.44 &
          18.59 &
            0.00 &
              -44917.22 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, RoadCrossing &
        89862.66 &
          20.81 &
            0.00 &
              -44916.33 \\
      (*), Water, DistanceToWater, HumansBuff5000 &
        89870.14 &
          28.29 &
            0.00 &
              -44924.07 \\
      (*), Water, DistanceToWater &
        89877.72 &
          35.87 &
            0.00 &
              -44929.86 \\
      (*), Water, DistanceToWater, DistanceToRoads &
        89880.07 &
          38.22 &
            0.00 &
              -44929.04 \\
      (*), Water, DistanceToWater, Protected &
        89881.39 &
          39.54 &
            0.00 &
              -44929.70 \\
      (*), Water, DistanceToWater, Trees &
        89881.57 &
          39.72 &
            0.00 &
              -44929.79 \\
      (*), Water, DistanceToWater, RoadCrossing &
        89883.48 &
          41.63 &
            0.00 &
              -44928.74 \\
      (*), Water, HumansBuff5000 &
        89887.27 &
          45.42 &
            0.00 &
              -44934.64 \\
      (*), Water &
        89894.64 &
          52.79 &
            0.00 &
              -44940.32 \\
      (*), Water, DistanceToRoads &
        89897.24 &
          55.39 &
            0.00 &
              -44939.62 \\
      (*), Water, Protected &
        89898.47 &
          56.62 &
            0.00 &
              -44940.24 \\
      (*), Water, Trees &
        89898.53 &
          56.67 &
            0.00 &
              -44940.26 \\
      (*), Water, RoadCrossing &
        89900.73 &
          58.88 &
            0.00 &
              -44939.37 \\
      (*), Shrubs/Grassland &
        89999.47 &
          157.62 &
            0.00 &
              -44992.73 \\
      (*), DistanceToWater &
        90001.72 &
          159.87 &
            0.00 &
              -44993.86 \\
      (*), HumansBuff5000 &
        90056.07 &
          214.22 &
            0.00 &
              -45021.04 \\
      (*), Trees &
        90059.55 &
          217.69 &
            0.00 &
              -45022.77 \\
      (*), DistanceToRoads &
        90067.62 &
          225.76 &
            0.00 &
              -45026.81 \\
      (*), Protected &
        90067.85 &
          226.00 &
            0.00 &
              -45026.93 \\
      (*), RoadCrossing &
        90069.86 &
          228.00 &
            0.00 &
              -45025.93 \\
      (*), Water, Shrubs/Grassland &
        \(NA\) &
          \(NA\) &
            0.00 &
              \(NA\) \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Further Investigations}
We have been working on further investigations that did not fit the main text
and were thus moved to the Appendix of this thesis. \Cref{SchematicOverview}
gives an overview of how these additional investigations integrate into our main
work. Our main work is described in the blue boxes, which resulted in the
identification of LCPs and LCCs. However, as we have emphasized in the
discussion, the applicability of the iSSF framework expands beyond the
estimation of movement and habitat preferences and it could, in fact, be used to
simulate animal movement. We thus engaged in fitting a movement model that was
specifically intended to simulate dispersal events (red boxes). Using this
model, we simulated 1000 dispersal events across the surroundings of our core
study area and qualitatively assessed the outcome. Corresponding methods and
results are reported in \Cref{Appendix:DispersalSimulation}. Furthermore, we
motivated that data on the social landscape of dispersers would deepen our
understanding of how dispersers are influenced by the presence or absence of
conspecifics, apex predators, and prey. Although we lacked proper GPS relocation
data of apex predators and prey, we collected a considerable amount of movement
data from resident wild dogs residing in our core study area. These data cover
the periods of almost all of our dispersal events and thus hold the potential
for modeling a social landscape for our dispersers. Hence, we modeled such a
landscape for the extent of our core study area and used it as additional
covariate to parameterize a social model (yellow box). Corresponding methods and
results are reported in \Cref{Appendix:SocialLandscape}.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[node distance=5mm and 5mm]
      \node (mod2) [nonterminal
        , minimum width = 30mm
        , fill          = red!20
        ] {Movement Model};
      \node (mod1) [nonterminal
        , above         = of mod2
        , fill          = blue!20
        , minimum width = 30mm
        ] {Permeability Model};
      \node (mod3) [nonterminal
        , below         = of mod2
        , fill          = yellow!20
        , minimum width = 30mm
        ] {Social Model};
      \node (dat1) [nonterminal
        , above left    = of mod2
        , fill          = blue!20
        , yshift        = -5mm
        , minimum width = 30mm
        ] {GPS Data};
      \node (dat2) [nonterminal
        , below left    = of mod2
        , fill          = blue!20
        , yshift        = +5mm
        , minimum width = 30mm
        ] {Spatial Covariates};
      \node (for1) [nonterminal
        , right         = of mod1
        , fill          = blue!20
        , minimum width = 35mm
        ] {Permeability Map};
      \node (forfor1) [nonterminal
        , right         = of for1
        , fill          = blue!20
        , minimum width = 35mm
        , yshift        = +5mm
        ] {Least-Cost Paths};
      \node (forfor2) [nonterminal
        , right         = of for1
        , fill          = blue!20
        , minimum width = 35mm
        , yshift        = -5mm
        ] {Least-Cost Corridors};
      \node (for2) [nonterminal
        , right         = of mod2
        , fill          = red!20
        , minimum width = 35mm
        ] {Dispersal Simulation};
      \node (box) [draw
        , dotted
        , very thick
        , minimum width   = 35mm
        , minimum height  = 35mm
        ] {};
      \node [above = of box, yshift = -5mm] {iSSF};
      \path (dat1) edge[->] (box);
      \path (dat2) edge[->] (box);
      \path (mod1) edge[->] (for1);
      \path (mod2) edge[->] (for2);
      \path (for1) edge[->] (forfor1);
      \path (for1) edge[->] (forfor2);
    \end{tikzpicture}
    \caption{Schematic overview of our methodological approaches. The blue boxes
    highlight our main line of work which resulted in the identification of
    least-cost paths and least-cost corridors. The red boxes show how we
    additionally used the iSSF framework to specifically analyze movement
    behavior and subsequently to simulate dispersal events. Finally, the yellow
    box indicates our attempt to use the iSSF framework to model and analyze the
    influence of the social landscape on dispersal behavior.}
    \label{SchematicOverview}
  \end{center}
\end{figure}

\newpage
\subsection{Movement Model \& Dispersal Simulation}
\label{Appendix:DispersalSimulation}
\subsubsection{Methods}
To gain more detailed insights into the movement behavior of dispersing wild
dogs and to simulate dispersal events, we fitted a more complex movement model.
The movement model still contained the most parsimonious permeability model, but
in addition included interaction terms between movement metrics (i.e.
\(cos(ta\_)\), \(log(sl\_\))) and environmental covariates (e.g. water). For
instance, for the environmental covariate \textit{Water}, we created the
interaction terms \textit{cos(ta\_):Water} and \textit{log(sl\_):Water}. These
interactions served to learn how dispersers moved differently, depending on
environmental covariates (for comparable work see \cite{Prokopenko.2017}). We
applied the same modeling framework as explained in \Cref{Modeling} and ran
forward model selection based on AIC to identify the most parsimonious movement
model. That is, we departed from the permeability model and iteratively added
interactions with movement metrics. Note that we applied forward model selection
only on the newly identified interactions, but not for the terms that were
already in the permeability model. Since we attempted to keep our simulation
simple, we did not average models with positive AIC-weights but simply
identified the most parsimonious movement model.

Using the most parsimonious movement model, we simulated 1000 dispersal events
departing from our core study area. The simulation basically resembled an
inverted iSSF function and was set up as follows. First, we sampled a random
starting location within our core study area. Second, we assigned a random
orientation in order to be able to calculate turning angles. Third, we created
an array of 25 random steps (i.e. one stratum) following the same procedure as
described in \Cref{Modeling}. Fourth, we extracted environmental covariates
along each of the generated random steps and calculated the metrics listed in
\Cref{ExtractedCovars}. Fifth, we applied \Cref{EQ2} using parameter estimates
from the movement model and calculated the selection score \(w(x)\) of each of
the 25 random steps. Sixth, we identified and realized the step with the highest
selection score and determined the new location. We then repeated steps three to
six until 200 steps were realized. To deal with dispersers that approached a map
boundary, our simulation included the options to either stop the simulation and
skip to the next disperser, or to remove any random step ranging outside the
study area, thereby forcing the simulated disperser to stay within the
boundaries. None of the simulated dispersers approached a map boundary within
200 steps and the decision did not influence our findings. Although we ran the
simulation only for dispersers departing from our core study area, one could
easily expand the simulation to include multiple sources and thereby check how
well these sources are connected.

\newpage
\subsubsection{Results}
Along with the original covariates contained in the permeability model, our most
parsimonious movement model (see \Cref{ModelAICs2} and
\Cref{MovementModelResults}) retained three additional interactions (see red box
in \Cref{MovementModel}a). Since the sign and direction of the effects contained
in the permeability model did not markedly change, we only report on the newly
added interactions between environmental covariates and movement metrics. For an
easier interpretation of how these interactions influenced habitat selection and
movement patterns during dispersal, we prepared surface plots as depicted in
\Cref{MovementModel}b. These surface plots revealed that dispersers preferred
long steps (i.e. a high movement rate) when water cover was low, whereas they
preferred short steps (i.e. a low movement rate) when water cover was high (see
\Cref{MovementModel}b1). Furthermore, dispersers were indifferent between high
and low turning angles in proximity to water, but clearly preferred low turning
angles (i.e. a high \(cos(ta\_)\)) when distance to water was high (see
\Cref{MovementModel}b2). Finally, dispersers preferably realized steps with low
turning angles when human influence was low, whereas they preferably realized
low turning angles when human influence was high (see \Cref{MovementModel}b3).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \begin{center}
              \includegraphics[width = 0.80\textwidth]
                {99_MovementModel.pdf}\\
              \includegraphics[width = 1\textwidth]
                {99_MovementModel(Interactions).pdf}
            \end{center}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
           %  % next four lines will help you to locate the point needed by
           %  % forming a grid. comment these four lines in the final picture.↓
           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
           %  % upto here
            \draw [red, dashed] (0.13, 0.50) rectangle (0.405, 0.645);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Results from the most parsimonious movement model. Negative
    coefficients indicate avoidance of a covariate, positive coefficients
    selection of a covariate. Whiskers delineate the 95\%-CIs for estimated
    parameters. The intercept is meaningless in the iSSF framework and was thus
    omitted. (b) Corresponding surface plots for two-way interactions. Light
    areas depict a combination of covariates that were preferred by dispersers,
    whereas darker areas correspond to avoided combinations. The absolute values
    along the axes have no meaning as they were normalized to 0 (low) and 1
    (high).}
    \label{MovementModel}
  \end{center}
\end{figure}

\newpage
\noindent \Cref{Simulations}a shows the outcome of the dispersal simulations.
For better visibility of high-frequency routes, we rasterized all simulated
trajectories and computed a heatmap depicting areas that were crossed by many
simulated trajectories. Unsurprisingly, we found that the dispersal trajectories
clearly reflected preferences that we identified earlier. For instance, most
trajectories were quite directional, avoided crossing a waterbody, ran along
waterbodies and avoided human-dominated areas. More interestingly, however, the
resulting heatmap closely resembled our permeability surface (see
\Cref{Simulations}b). That is, areas providing high permeability were typically
crossed by more simulated trajectories. Yet, the heatmap also showed that some
regions are almost impossible to reach, even if they provide highly permeable
landscapes. For example, although the western section of the Okavango Delta
holds highly permeable landscapes, no simulated trajectory reached these areas.
In contrast, many trajectories ventured towards east, some of them even crossing
the border to Zimbabwe. The reason that no trajectory reached the eastern part
of the Delta is likely owed to severe human influence and an impermeable flood
that together posed an insurmountable barrier and prevented dispersers from
traversing the landscape.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = 0.83\textwidth]{99_Simulations.pdf}
    \caption{(a) Heatmap of 1000 simulated dispersal events departing from the
    core study area (white square) in northern Botswana (countries with dashed
    white lines). (b) Permeability map for the same extent, showing that
    simulated dispersal trajectories typically cover highly permeable
    landscapes.}
    \label{Simulations}
  \end{center}
\end{figure}

\newpage
\begin{table}[h]
  \caption{Results from the forward model selection procedure based on AIC
  \citep{Burnham.2002} for our movement model. Weights were calculated using:
  \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i = 1}^n
  e^{-0.5} \Delta \text{AIC}_i} \). (*) represents all covariates from the most
  parsimonious permeability model, i.e. \(cos(ta\_)\), \(log(sl\_\)),
  \textit{Water}, \textit{DistanceToWater}, \textit{Shrubs/Grassland},
  \textit{HumansBuff5000}, and \textit{Trees}. To keep the table short we only
  included models with positive AIC weights. We abstained from averaging models
  with AIC-weight \(>\) 0 because we attempted to keep our simulation simple.}
  \label{ModelAICs2}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llllll}
      \toprule
      \textbf{Covariates} &
        \textbf{AIC} &
          \textbf{\(\Delta\)AIC} &
            \textbf{AIC-Weight} &
              \textbf{LogLik} \\
      \midrule
      \rowcolor[gray]{0.80}
        (*),
        log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000 &
          89814.19 &
            0.00 &
              0.13 &
                -44889.10 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):Trees &
          89814.37 &
            0.17 & 0.12 &
              -44889.18 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater &
          89814.46 &
            0.27 &
              0.11 &
                -44890.23 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Trees &
          89814.87 &
            0.68 &
              0.09 &
                -44888.44 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Shrubs &
          89815.32 &
            1.13 &
              0.07 &
                -44888.66 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):Shrubs &
          89815.34 &
            1.15 &
              0.07 &
                -44889.67 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):Shrubs &
          89815.40 &
            1.20 &
              0.07 &
                -44888.70 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, log(sl\_):Shrubs &
          89815.67 &
            1.48 &
              0.06 &
                -44889.84 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):HumansBuff5000 &
          89815.94 &
            1.75 &
              0.05 &
                -44888.97 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Trees, log(sl\_):Shrubs &
          89816.08 &
            1.89 &
              0.05 &
                -44888.04 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):DistanceToWater &
          89816.14 &
            1.95 &
              0.05 &
                -44889.07 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Water &
          89816.16 &
            1.97 &
              0.05 &
                -44889.08 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, log(sl\_):HumansBuff5000 &
          89816.18 &
            1.99 &
              0.05 &
                -44890.09 \\
        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):Trees &
          89816.19 &
            2.00 &
              0.05 &
                -44889.10 \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\begin{table}[h]
  \caption{Selection coefficients as derived from the most parsimonious movement
  model. The movement model was a modified permeability model to which we added
  additional interaction terms between movement metrics and environmental
  covariates (e.g. \(Water:cost(ta\_)\)) in order to learn about how movement
  behavior changes depending on the environmental covariates.}
  \label{MovementModelResults}
  \begin{center}
    \resizebox{0.5\textwidth}{!}{
      \begin{tabular}{llllll}
      \toprule
      Covariate & Coefficient & SE & p-value \\
      \midrule
      cos(ta\_) & 0.149 & 0.045 & 0.001 \\
      log(sl\_) & 0.053 & 0.019 & 0.005 \\
      Water & -0.265 & 0.090 & 0.003 \\
      DistanceToWater & -0.272 & 0.226 & 0.230 \\
      Shrubs & 0.391 & 0.082 & 0.000 \\
      HumansBuff5000 & -0.466 & 0.203 & 0.022 \\
      Trees & -0.177 & 0.072 & 0.013 \\
      log(sl\_):Water & -0.040 & 0.008 & 0.000 \\
      cos(ta\_):DistanceToWater & 0.115 & 0.038 & 0.003 \\
      cos(ta\_):HumansBuff5000 & -0.047 & 0.031 & 0.134 \\
       \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Social Model}
\label{Appendix:SocialLandscape}
\subsubsection{Methods}
We used GPS relocation data of resident wild dogs and calculated a
geo-referenced stack of layers to which we refer as the \textit{social
landscape}. The social landscape was a collection of raster-layers, each
depicting the probability of encountering a resident group in a given pixel and
month. The social landscape was updated every month, using a moving window that
took into account six months of relocation data. For instance, the social
landscape dated July 1\textsuperscript{st} consisted of GPS relocation data
collected between January 1\textsuperscript{st} and June 30\textsuperscript{th}.
Because we collected relocation data of residents only inside our core study
area, we modeled the social landscape only for this region. GPS relocations were
collected at a 24-hourly interval during residence, so we kept one GPS
relocation per pack and day. To compute monthly updated social landscapes, we
calculated each resident pack's utilization distribution (UD) and corresponding
95\% home range (HR) using the kernelUD function from the R-package
\textit{adehabitatHR} \citep{Calenge.2019}. The package required to set a
smoothing parameter \(h\), which determined the sharpness of the resulting UD
and 95\%-HR. To come up with a meaningful smoothing parameter \(h\), we followed
the procedure described in \cite{Cozzi.2018}. That is, we first applied the
reference value \( h_{ref} \), which the package automatically determined. We
then obtained the corresponding UD and 95\%-HR and assessed whether the 95\%-HR
comprised of a single polygon or whether it split up into multiple polygons. In
case the 95\%-HR segregated into multiple polygons, we assumed that the
smoothing parameter was set to strictly. On the other hand, if the 95\%-HR
comprised of a single polygon, we assumed that we could further tighten the
smoothing parameter. We thus iteratively adjusted the smoothing parameter until
a value right before the 95\%-HR split up into multiple polygons. However, we
maximally varied \( h \) within 30\% and 120\% of the original \( h_{ref} \) to
avoid extreme over- or under-smoothing. Using the resulting \(h\) we computed
the final UD and corresponding 95\%-HR. Each pack's UD of a given month was
converted to a spatial probability density using \(f(x) =
\frac{x_i}{\sum_1^n{x_i}}\), so that pixel-values summed up to one. We then used
\Cref{EQ4} to merge all UDs of the same month and to calculate the probability
of not encountering any wild dogs in a given pixel and month.

\begin{equation}
\label{EQ4}
x_i = 1 - \prod_{j = 1}^{n} (1 - x_{ij})
\end{equation}

\noindent where \(x_i =\) is the probability of encountering another pack in
pixel \(i\) and \(x_{ij} =\) the probability of encountering pack \(j\) in pixel
\(i\). Finally, we inverted this probability using \((1-x_i)\) to get the
probability of encountering at least one resident pack in a given pixel and
month. Note that we kept track of each packs' 95\%-HR throughout all months,
which allowed us to identify through which HRs dispersers moved during
dispersal. An example of the final layers is given in
\Cref{SocialLandscapeExample}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = 0.95\textwidth]{99_SocialLandscapeExample.pdf}
    \caption{(a) Example of a layer depicting the probability of encountering a
    resident pack in July 2017. The layer was derived from utility distributions
    of all packs residing in the core study area. (b) Example of a layer
    depicting all 95\% home ranges of the packs residing in the core study area.
    This layer served to identify whether a disperser still moved through his
    natal pack's home range, or whether it moved through a foreign packs home
    range. The layers in (a) and (b) were updated each month to take into
    account only the latest GPS relocation data of resident individuals.}
    \label{SocialLandscapeExample}
  \end{center}
\end{figure}

\noindent To fit a social model using the newly created social landscape, we
reran our iSSF described in \Cref{Modeling}. However, this time we discarded any
stratum that did not fully lie within the core study area. The social model that
we fitted was again a modification of the most parsimonious permeability model.
Despite the covariates contained in the permeability model, we added the social
covariates listed in \Cref{SocialCovars}. Because we hypothesized that the
preference for encountering other wild dogs was different within and outside the
disperser's natal home range, we included the interaction terms
\textit{ProbEncounter:OwnHomerange} and \textit{ProbEncounter:ForeignHomeranges}
in the social model. Since we did not anticipate to use the model for
predictions, we did not engage in model selection.

\begin{table}[h]
  \begin{center}
    \caption{Overview of extracted social covariates. Values were extracted from
    the layer that was closest in date to the actual step.}
    \label{SocialCovars}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{lllll}
      \hline
      \textbf{Category} &
        \textbf{Covariate} &
          \textbf{Description} &
            \textbf{Values} \\
              % \textbf{Source(s)} \\
      \midrule
      Social Features
        & Own Homerange
          & Percentage cover of natal pack’s 95\% home-range along step
            & 0-100\% \\
      Social Features
        & ForeignHomeranges
          & Number of foreign packs’ 95\% home-ranges along step
            & \(\geq\) 0 \\
      Social Features
        & ProbEncounter
          & Average probability of encountering conspecifics along step
            & 0-100\% \\
      \hline
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsubsection{Results}
Results from our social model are depicted in \Cref{SocialModel}a (see also
\Cref{SocialModelResults}). Again, parameter estimates for the covariates
contained in the permeability model did not severely change and we will only
report on the newly added social covariates (see red box in
\Cref{SocialModel}a). Visual inspection of the corresponding surface plot (see
\Cref{SocialModel}b1) revealed that dispersers tried to maximize the probability
of encounter when moving outside their natal pack's HR. Yet, this preference was
even stronger when dispersers moved inside their natal pack's HR. A second
surface plot (see \Cref{SocialModel}b2) showed that dispersers maximized the
probability of encounter when the number of foreign HRs was low, but that they
minimized the probability of encounter when the number of foreign HRs was high.
In summary, dispersers appeared to be drawn towards other wild dogs, but only if
the number of overlapping foreign HRs was low.

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \begin{center}
              \includegraphics[width = 0.75\textwidth]
                {99_SocialModel.pdf}\\
              \includegraphics[width = 0.8\textwidth]
                {99_SocialModel(Interactions).pdf}
            \end{center}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
           %  % next four lines will help you to locate the point needed by
           %  % forming a grid. comment these four lines in the final picture.↓
           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
           %  % upto here
            \draw [red, dashed] (0.15, 0.60) rectangle (0.43, 0.75);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Estimated parameters from the social model. Negative
    coefficients indicate avoidance of a covariate, positive coefficients
    selection of a covariate. Whiskers delineate the 95\%-CIs for estimated
    parameters. The intercept is meaningless in the iSSF framework and was thus
    omitted. (b) Corresponding surface plots for two-way interactions (red box).
    Light areas in the surface plots depict preferred combinations of the
    covariates, whereas darker areas correspond to avoided combinations. The
    absolute values along the axes of the surface plots have no meaning as they
    were scaled. Rather, low values along the axis correspond to low values of
    the respective covariate. The same is true for the scalebar.}
    \label{SocialModel}
  \end{center}
\end{figure}

\newpage
\begin{table}[h]
  \caption{Selection coefficients as derived from the social model. The social
  model was a modified permeability model to which we added additional social
  covariates in order to learn about how the social landscape infleunces
  dispersal behavior.}
  \label{SocialModelResults}
  \begin{center}
    \resizebox{0.6\textwidth}{!}{
      \begin{tabular}{llllll}
        \toprule
        Covariate & Coefficient & SE & p-value \\
        \midrule
        cos(ta\_) & 0.148 & 0.066 & 0.025 \\
        log(sl\_) & 0.055 & 0.019 & 0.003 \\
        Water & -0.472 & 0.113 & 0.000 \\
        DistanceToWater & -0.193 & 0.411 & 0.639 \\
        Shrubs & 0.470 & 0.110 & 0.000 \\
        HumansBuff5000 & -1.013 & 0.295 & 0.001 \\
        Trees & -0.465 & 0.193 & 0.016 \\
        ProbEncounter & 0.724 & 0.268 & 0.007 \\
        OwnHomerange & 0.056 & 0.095 & 0.557 \\
        ForeignHomeranges & -0.099 & 0.132 & 0.452 \\
        ProbEncounter:OwnHomerange & 0.189 & 0.179 & 0.291 \\
        ProbEncounter:ForeignHomeranges & -0.504 & 0.281 & 0.072 \\
        \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Statutory Declaration}
I assure that this thesis is a result of my personal work and that no other than
the indicated aids have been used for its completion. Furthermore, I assure that
all quotations and statements that have been inferred literally or in a general
manner from published or unpublished writings are marked as such. Beyond this, I
assure that the work has not been used, neither completely nor in parts, to pass
any previous examination.\\
\vspace{15mm}\\
Zürich, \today

\vspace{5mm}
\includegraphics[width = 0.2\textwidth]{Signature.png}

\noindent David Hofmann

\end{document}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-01-12.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-01-03.tex
@@
 % Title of the document
-{\textsc{ \LARGE Running Wild: Movement Corridors of Dispersing African Wild
+{\textsc{ \LARGE Ranging Wide: Movement Corridors of Dispersing African Wild
 Dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation
@@
 KAZA-TFCA. Our results show that dispersal corridors typically run through open
-shrubs/grassland, and in the vicinity of water. Corridors rarely cross
+shrubs/grassland and in the vicinity of water. Corridors rarely cross
 waterbodies, areas with high human presence, or areas densely vegetated by
-trees. Several dispersal corridors meet in the Linyanti-Chobe ecosystem, which
-highlights the importance of northern Botswana for the conservation of this
-charismatic species.
+trees. Several dispersal corridors meet in the Okavango-Linyanty ecosystem,
+which highlights the importance of northern Botswana for the conservation of
+this charismatic species.
 \end{abstract}
@@
 % Importance of Connectivity
-Successful conservation of spatially fragmented subpopulations requires to
-assess connectivity of habitats occupied by these subpopulations. In this
-context, connectivity describes the degree to which the landscape facilitates or
-impedes movement of the focal species between habitat patches
-\citep{Taylor.1993, Clobert.2012}. Thus, connectivity is a prerequisite for
-dispersal, which in turn promotes genetic exchange, reinforcement of small
-populations, and colonization or recolonization of unoccupied habitats
+Successful conservation of fragmented and isolated subpopulations requires
+maintenance and enhancement of connectivity between habitats occupied by these
+subpopulations \citep{Hanski.1998}. In this context, connectivity describes the
+degree to which the landscape matrix facilitates or impedes movement of
+individuals and thereby genes of the focal species between habitat patches
+\citep{Taylor.1993, Clobert.2012}. Thus, sufficient connectivity facilitates
+dispersal, which in turn promotes genetic exchange, reinforcement and rescuing
+of small populations, and colonization or recolonization of unoccupied habitats
 \citep{Brown.1977, Hanski.1998, MacArthur.2001, Frankham.2002, Leigh.2012}. All
-of these processes crucially contribute to the long-term viability of numerous
-species, therefore emphasizing the importance of connectivity. However,
-increasing anthropogenic pressure and ongoing loss and fragmentation of valuable
-habitats erode connectivity on a worldwide scale \citep{Fahrig.2003,
-Barnosky.2012}. As a result, management actions for the conservation of
-endangered species increasingly aim to restore connectivity through the
-identification and preservation of key movement corridors that link isolated
-habitats \citep{Fahrig.2003, Heller.2009}. An unprecedented instance of an
-initiative to restore connectivity on a large scale is the Kavango-Zambezi
-Transfrontier Conservation Area (KAZA-TFCA), which traverses five African
-countries and spans over 520,000km\textsuperscript{2}. While the primary purpose
-of the KAZA-TFCA is to facilitate the migration of elephants across large
-landscapes, numerous other wide-ranging species will likely benefit from its
-planned protectorates too. Nevertheless, the extent to which such initiatives
-are beneficial to certain species is rarely assessed, leaving open their true
-impact on connectivity.
+these processes crucially contribute to long-term population viability, which
+emphasizes the importance of dispersal and connectivity. However, increasing
+anthropogenic pressure and ongoing habitat loss and fragmentation reduce
+connectivity on a worldwide scale \citep{Fahrig.2003, Barnosky.2012}. As a
+result, management actions for the conservation of endangered species are
+increasingly aimed towards preserving and restoring connectivity through the
+identification and preservation of key dispersal corridors that link isolated
+habitats \citep{Fahrig.2003, Heller.2009}. An unprecedented initiative to
+restore connectivity on a large scale is represented by the creation of the
+Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), a protected area
+which traverses five African countries and spans over 520,000
+km\textsuperscript{2}. While such initiatives are expected to bring benefits to
+wide-ranging species such as elephants (\textit{Loxodonta africana}), African
+wild dogs (\textit{Lycaon pictus}), and lions (\textit{Panthera leo}), the
+extent to which they are effective for targeted species and whether boundaries
+reflect natural dispersal corridors is rarely assessed.
+
+% Connectivity Assessment
+In recent years, there has been a growing body of literature that uses animal
+movement data to identify dispersal corridors and to assess connectivity on
+large scales \citep{Chetkiewicz.2006}. Identification of potential dispersal
+corridors between suitable habitat patches traditionally relies on the
+estimation of a permeability surface (elsewhere referred to as resistance
+surface, when inverted), which renders the ease or willingness at which the
+focal species traverses a specific landscape. Latest empirical methods to
+predict landscape permeability based on recorded GPS relocation data include
+\textit{resource selection functions} (RSF, \cite{Boyce.2002}), \textit{step
+selection functions} (SSF, \cite{Fortin.2005}), and \textit{path selection
+functions} (PSF, \cite{Cushman.2010}). All methods have in common that they
+infer habitat preferences of the focal species by comparing covariates (e.g.
+environmental and anthropogenic) at visited locations to the same covariates at
+randomly selected locations that are deemed available but presumably unused by
+the animal. Based on inferred habitat preferences the methods further assume
+that habitats preferred by the focal species provide high permeability, whereas
+avoided habitats provide low permeability. Regardless of the method chosen (i.e.
+RSF, SSF, or PSF), all approaches require adequate GPS relocation data that is
+representative for the process being studied. More specifically, it has been
+shown that GPS relocation data collected on dispersing individuals is better
+suited for detecting dispersal corridors than GPS relocation data collected on
+resident individuals \citep{Elliot.2014}. Few studies have, however, partitioned
+their GPS relocation data according to behavioral mode such as \textit{resident}
+vs. \textit{dispersing} or \textit{active} vs. \textit{inactive} periods of
+movement \citep{Wilson.2012, Vasudev.2015}. As a consequence, most permeability
+surfaces upon which dispersal corridors are identified are created using GPS
+relocation data that was collected on resident individuals. This introduces
+severe biases and substantially reduces the power to reveal meaningful
+corridors, for dispersing individuals have different needs and motivations
+compared to resident individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}.
+For instance, \cite{Elliot.2014} demonstrated marked differences in predicted
+permeability surfaces for dispersing lions in the KAZA-TFCA depending on whether
+permeability was assessed using data collected during residence or dispersal.
+\cite{Jackson.2016} made similar observations for African wild dogs and revealed
+higher permeability for dispersers than for residents. Moreover,
+\cite{Abrahms.2017} found a significantly greater overlap between predicted and
+observed corridors for three dispersing wild dogs when predictions were based on
+information collected only during wild dog activity, rather than across the
+entire day.
 
 % African wild dogs
-One of the species for which the degree to which it will benefit from the
-KAZA-TFCA remains unknown is the African wild dog (\textit{Lycaon pictus}), a
-keystone predator and flagship species for conservation efforts in the KAZA-TFCA
+A highly endangered flagship species for conservation efforts across the
+KAZA-TFCA, but for which systematic GPS relocation data on dispersing
+individuals is largely unavailable, is the African wild dog
 \citep{VanDerMeer.2016}. While once present across the entire Sub-Saharan
 continent, the species has been widely extirpated through human persecution,
-habitat destruction and disease outbreaks. As a result, the African wild dog has
-become Africa's most endangered large carnivore that currently only survives in
-small, spatially scattered subpopulations (see \Cref{StudyArea}a,
-\cite{Woodroffe.2012}). Within these subpopulations, wild dogs form cooperative
+habitat destruction, and disease outbreaks \citep{Woodroffe.2012}. As a result,
+the African wild dog has become one of Africa's most endangered large carnivores
+and currently only survives in small, spatially scattered subpopulations
+\citep{Woodroffe.2012}. Within these subpopulations, wild dogs form cooperative
 breeding groups of up to thirty individuals in which a single dominant pair
@@
 \citeauthor{Behr.2019}, unpublished). During such dispersal events, wild dogs
-cover vast distances, sometimes crossing perilous areas that are strongly
-influenced by humans (\citeauthor{DaviesMostert.2012}, 2012;
-\citeauthor{Masenga.2016}, 2016; \citeauthor{Cozzi.2019}, in review). Despite
-the importance of dispersal for the remaining wild dog subpopulations, little is
-known about the movement corridors of dispersing individuals and connectivity
-between subpopulations in the KAZA-TFCA in general.
-
-% Connectivity Assessment
-In recent years, there has been a growing body of literature that uses animal
-movement data to identify movement corridors and thereby assess connectivity on
-large scales \citep{Chetkiewicz.2006}. Revelation of potential movement
-corridors between selected locations traditionally relies on the estimation of a
-permeability surface (elsewhere referred to as resistance surface when
-inverted), which depicts the ease or willingness at which the focal species
-traverses a specific landscape. Latest and very powerful empirical methods to
-predict permeability based on observed movements include \textit{resource
-selection functions} (RSF, \cite{Boyce.2002}), \textit{step selection functions}
-(SSF, \cite{Fortin.2005}) and \textit{path selection functions} (PSF,
-\cite{Cushman.2010}). All methods have in common that they infer habitat
-preferences of the focal species by comparing environmental covariates at
-visited locations to environmental covariates at randomly selected locations
-that are deemed available but presumably unused by the animal. Based on inferred
-habitat preferences, the methods further assume that habitats preferred by the
-focal species provide high permeability, whereas avoided habitats provide low
-permeability. Regardless of the method chosen (i.e. RSF, SSF, or PSF), all
-approaches require adequate relocation data that is representative for the
-process being studied.
-
-% Behavioral Modes
-For connectivity studies in particular, the behavioral mode of the animal (e.g.
-resident, dispersing) during data collection should be considered when
-estimating landscape permeability and assessing movement corridors. However, few
-studies have partitioned their movement data according to behavioral mode
-\citep{Wilson.2012, Vasudev.2015}. As a consequence, most researchers estimated
-permeability for dispersing individuals as if they had the same preferences as
-resident individuals. This is a rather disputed assumption and recent research
-suggests that such an approach seriously reduces the power to reveal meaningful
-movement corridors \citep{Killeen.2014}. For instance, \cite{Elliot.2014}
-demonstrated marked differences in predicted permeability surfaces for
-dispersing lions in the KAZA-TFCA depending on whether permeability was assessed
-using data collected during residence or dispersal. \cite{Jackson.2016} made
-similar observations for African wild dogs and revealed higher permeability for
-dispersers than for residents. Moreover, \cite{Abrahms.2017} found a
-significantly greater overlap between predicted and observed corridors for three
-dispersing wild dogs when predictions were based on information collected during
-wild dog activity, rather then when ignoring the behavioral mode. Overall,
-predicted movement corridors vastly differed depending on the type of data that
-was being used \citep{Elliot.2014, Jackson.2016, Abrahms.2017}. Unfortunately,
-data of dispersing wild dogs remains scarce and is collected opportunistically
-at best \citep{DaviesMostert.2012, Masenga.2016, Jackson.2016, Abrahms.2017}.
-For instance, the dispersers reported in \cite{DaviesMostert.2012} were not
-fitted with global positioning system (GPS) collars, which is why the
-researchers only recorded straight line dispersal-distances.
-\cite{Masenga.2016}, on the other hand, were able to collect proper GPS data,
-yet only on two dispersers. Furthermore, \cite{Jackson.2016} based their work on
-occasional wild dog sightings and their distinction between residents and
-dispersers was based purely on group size, assuming that any group with fewer
-than nine members was dispersing. Finally, \cite{Abrahms.2017} used GPS
-relocation data of resident individuals during phases of activity rather than
-during true dispersal. In summary, none of these studies systematically
-monitored dispersing individuals, which is likely owed to technological and
-logistic limitations when monitoring such an elusive and wide-ranging species
-\citep{Nathan.2001, Tesson.2013}. Yet, a better understanding of how the
-landscape impedes or facilitates movement of dispersers is vital for improving
-our understanding of connectivity and for conservation of the species in
-general.
-
-% Research Goal
-Our study therefore had two aims: (1) to estimate landscape permeability for
-dispersing wild dogs, and subsequently use this information to (2) assess
-connectivity by revealing key dispersal corridors within the entire extent of
-the KAZA-TFCA. For this purpose, we collected GPS data of 16 dispersing
-coalitions originating from a free-ranging population of wild dogs in the
-eastern section of the Okavango Delta in Botswana. We highlight key dispersal
-corridors that provide connectivity between already established protected areas
-within the KAZA-TFCA and we reveal potentially important dispersal corridors
-outside the planned protectorates of the KAZA-TFCA. Our results have important
-implications for the conservation and management of the endangered African wild
-dog.
+are able to cover several hundred kilometers, sometimes crossing risky areas
+that are strongly influenced by humans \citep{DaviesMostert.2012, Masenga.2016,
+Cozzi.2020}. Despite the importance of dispersal for the remaining wild dog
+subpopulations, little is known about the corridors of dispersing individuals
+and connectivity between subpopulations in the KAZA-TFCA in general.
+
+% Covariate Motivation
+Wild dog movements and spatial behavior are known to be influenced by a suite of
+environmental and anthropogenic factors. For instance, \cite{Abrahms.2017} found
+that wild dog movements during residence mainly depended on land cover,
+protection status and anthropogenic features. Similar observations on resident
+packs were made by \cite{Pretorius.2019}, who's results indicate that pack
+presence was contingent on factors such as agricultural activity, road presence,
+and protection status of the area. With respect to dispersing individuals
+\cite{Whittington.2011} demonstrated that elevation, land cover, road density,
+and human density best explained the distribution of recorded sightings in South
+Africa. Also based on data from dispersing wild dogs, \cite{Jackson.2016} showed
+that land cover, human density and road density significantly influenced the
+probability of occurrence of dispersing wild dogs. Finally, \cite{Masenga.2016}
+reported a low willingness of dispersers to cross human-dominated landscapes.
+Large scale spatial information on the above mentioned factors can be obtained
+either through freely available online services or through remote sensed
+satellite imagery.
+
+% Our Study
+We used GPS relocation data collected on 16 wild dogs in as many dispersing
+coalitions from a free ranging population in northern Botswana to estimate the
+degree of selection or avoidance for environmental and anthropogenic landscape
+features. We then used the obtained information in combination with relevant
+spatial covariates to create a permeability surface for the entire KAZA-TFCA.
+Lastly, we used the permeability surface as input for the computation of
+least-cost paths and least-cost corridors between selected locations. This
+allowed us to highlight key dispersal corridors within and outside the
+boundaries of KAZA-TFCA and concluded with the identification of a ``dispersal
+hub'' and of dispersal corridors that run outside formally protected areas.
 
@@
 \subsection{Study Area}
+% Both Study Areas
+We differentiated between a \textit{core study area}, where all dispersers
+originated from, and an \textit{extended study area}, which was represented by a
+bounding box encompassing the KAZA-TFCA \citep{Cozzi.2020} and for which we
+calculated connectivity.
+
 % Core Study Area
-For the purpose of this study we differentiated between a \textit{core study
-area}, where all dispersers originated from, and an \textit{extended study
-area}, for which we conducted the connectivity analysis. The \textit{core study
-area} was located in the eastern section of the Okavango Delta in northern
-Botswana (\Cref{StudyArea}b, centered at -19\degree 25'30''S, 23\degree
-37'30''E; elevation ca. 950m) and comprised parts of the Moremi Game Reserve and
-adjacent wildlife management areas. The Okavango Delta is considered to be one
-of the main strongholds of African wild dogs in southern Africa and likely acts
-as a source for the recolonization of surrounding habitats
-\citep{Woodroffe.2012, Cozzi.2013}. The Okavango Delta represents the earth's
-largest inland delta and is characterized by diverse habitats, ranging from
-permanent river systems, mopane woodland and acacia forests, to open grasslands
-and seasonal swamps \citep{Broekhuis.2013}. Rainfall is periodic and lasts from
-November to March, whereas the rest of the year remains dry \citep{McNutt.1996}.
-Flooding of the delta greatly changes depending on the season and drastically
-affects surrounding landscapes (see \Cref{FloodPulse} in
-\Cref{Appendix:FloodPulse} for an illustration of a typical flood pulse). During
-maximum extent, the delta becomes a patchy conglomerate of swamps, open water,
-and islands. Because rainfalls only slowly descent from their catchment areas in
-southern Angola into the Okavango Delta, the flood usually peaks at the height
-of the dry season between August and September \citep{Wolski.2017}. Most of the
-Okavango Delta's unique basin is under protection, such that human influence in
-our core study area was mainly restricted to the surroundings of Maun and the
-smaller villages of Khwai, Mababe, and Sankuyo.
+The \textit{core study area} was located in the eastern section of the Okavango
+Delta in northern Botswana (centered at -19\degree 25'30''S, 23\degree 37'30''E;
+elevation ca. 950 m, orange rectangle in \Cref{StudyArea}b) and comprised parts
+of the Moremi Game Reserve and adjacent wildlife management areas. The Okavango
+Delta is one of the main strongholds of African wild dogs in southern Africa and
+likely acts as a source for the recolonization of surrounding habitats
+\citep{Woodroffe.2012, Cozzi.2013}. It is the earth's largest inland delta and
+is characterized by diverse habitats, ranging from permanent river systems,
+mopane woodland, and acacia forests, to open grasslands and seasonal swamps
+\citep{Broekhuis.2013}. Rainfall is seasonal and lasts from November to March,
+whereas the remainder of the year remains dry \citep{McNutt.1996}. The extent of
+the flood greatly changes within and between years depending on the amount of
+rain that descends from the catchment areas in Angola and reaches the distal
+ends of the delta between July and August (see \Cref{FloodPulse} in
+\Cref{Appendix:FloodmappingAlgorithm} for an illustration of a typical flood
+pulse). The flood thereby drastically affects surrounding landscapes, so that
+during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a
+patchy conglomerate of swamps, open water, and islands, whereas it shrinks
+substantially (ca. 5'000 km\textsuperscript{2}) during dry periods
+\citep{Gumbricht.2004, Wolski.2017}. Most of the Okavango Delta's unique basin
+is under protection, such that human influence in our core study area was mainly
+restricted to the surroundings of Maun and the smaller villages of Khwai,
+Mababe, and Sankuyo.
 
 % Large Study Area
-Our \textit{extended study area} was confined by the extent of the KAZA-TFCA
-(\Cref{StudyArea}a, centered at -17\degree 13'9''S, 23\degree 56'4''E; elevation
-ca. 1,030m), which lies in the basins of the Okavango and Zambezi rivers and
-embraces parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. With a total
-area of over 520,000km\textsuperscript{2} the KAZA-TFCA contains several already
-existing national parks (NPs) and other protected areas, and will eventually
-become earth's largest transboundary conservation area. Since we considered the
-entire rectangular extent of the KAZA-TFCA, our extended study area encompassed
-a total of over 1.3 Mio km\textsuperscript{2}. Similar to our core study area,
-the extended study area hosts a conglomerate of various landscapes, including
-savanna, grassland, and dry or moist woodland patches. Furthermore, the
-KAZA-TFCA encloses several wild dog subpopulations  that still reside in
-Namibia, Botswana, and Zimbabwe (\Cref{StudyArea}a, \cite{Woodroffe.2012}).
-Despite 36 NPs and numerous other protected areas, there is considerable human
-influence in some regions of the KAZA-TFCA originating from farms, high human
-density, and road traffic.
+The \textit{extended study area} (centered at -17\degree 13'9''S, 23\degree
+56'4''E; elevation ca. 1,030 m, red rectangle in \Cref{StudyArea}a) was confined
+by a bounding box encompassing the entire KAZA-TFCA. The KAZA-TFCA lies in the
+basins of the Okavango and Zambezi rivers and embraces parts of Angola,
+Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520,000
+km\textsuperscript{2} it contains several already-existing national parks (NPs)
+and other protected areas. It constitutes the earth's largest transboundary
+conservation area, which resulted in a rectangular study area of over 1.3 Mio
+km\textsuperscript{2}. Similar to the core study area, the extended study area
+hosts various landscapes, including savanna, grassland, and dry or moist
+woodland habitats. Despite 36 NPs and other protected areas, there is
+considerable human influence in some regions of the KAZA-TFCA originating from
+farms, high human density, and road traffic.
 
@@
 monitored and all were observed to successfully re-join their packs within one
-hour after immobilization. GPS collars were programmed to record and transmit
-GPS relocations via Iridium satellite system to a base station. The system thus
-allowed remote-tracking of collared individuals, even where field conditions
-were prohibitive or dispersal trajectories crossed international borders.
-Because we were mainly interested in recording dispersal trajectories,
-relocations were collected only every 24 hours during residence, but every 4
-hours during dispersal. Dispersal and residence were strictly distinguished
-using a combination of net squared displacement \citep{Borger.2012}, visual
-inspection of trajectories, and observations of pack constellations in the field
-(for details see \citeauthor{Cozzi.2019}, in review). Dispersal was deemed to
-have started when an individual left its natal home range, and continued until
-the individual formed a new pack or exhibited signs of resident behavior. Some
-dispersers undertook exploratory movements prior to dispersal. Such movements
-have been observed in other species as well and are reported to be very similar
-to true dispersal \citep{Killeen.2014}. We therefore classified exploratory
-movements as dispersal too. We did not attempt to distinguish sex in our
-analysis, as previous research found little differences between males and
-females during dispersal (\citeauthor{Woodroffe.2019}, 2019;
-\citeauthor{Cozzi.2019}, in review). In total, we collected 4,691 GPS
-relocations during dispersal (see \Cref{Trajectories}). On average, dispersal
-events included three individuals (min = 1, max = 7), lasted for 41 days (min =
-7, max = 133), and covered a mean cumulative distance of 370km (min = 105, max =
-610). Detailed descriptions of the dispersal events can be found in
-\citeauthor{Cozzi.2019} (in review). Although our final sample size of 16
-dispersers was relatively small, we consider this an exceptional number for such
-an endangered and elusive species.
+hour after immobilization. GPS collars were programmed to record GPS relocation
+data every 4 hours during dispersal and to regularly transmit them via Iridium
+satellite system to a base station. The system thus allowed remote-tracking of
+collared individuals, even where field conditions were prohibitive or dispersal
+trajectories crossed international borders.
+
+% Separation of Dispersers/Residents
+Because we were mainly interested in recording dispersal trajectories, we
+discarded any GPS data collected while individuals were still with their natal
+packs \citep{Cozzi.2020}. Dispersal was deemed to have started when a coalition
+had left its natal home range, and continued until it exhibited clear signs of
+resident behavior. Some coalitions undertook exploratory movements prior to
+proper dispersal. Such movements have been observed in other species and are
+reported to be very similar to proper dispersal \citep{Killeen.2014}. We
+therefore classified exploratory movements as dispersal too.
+
+% Some Summary Statistics
+In total, we collected 4'691 GPS relocations during dispersal (see
+\Cref{Trajectories}). In our analysis, we did not differentiate between male and
+female dispersing coalitions, for previous research found little differences
+between sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}. On average,
+dispersal coalitions included three individuals (min = 1, max = 7), dispersed
+for 41 days (min = 7, max = 133), and covered a mean cumulative distance of 370
+km (min = 105, max = 610).
 
@@
     from our core study area, which is indicated by the white square. The
-    coalition dispersing towards the far east of the map covered over 360km in
+    coalition dispersing towards the far east of the map covered over 360 km in
     under 10 days.}
@@
 
-\subsection{Covariates}
+\subsection{Spatial Covariates}
 % General stuff
-We used a set of geo-referenced covariates to investigate movement and habitat
-preferences of dispersing wild dogs (see \Cref{Covariates}). Covariates fell
-into one of the three categories \textit{land cover}, \textit{protection
-status}, and \textit{human influence}. For each category we obtained spatial
-raster layers from freely available online services or from remote sensed
-satellite imagery. To ensure a consistent resolution (i.e. cell-size or grain)
-across covariates, all layers were coarsened or interpolated to match a
-resolution of 250m x 250m. Processing and manipulation of data, as well as all
-spatial and statistical analyses were performed in R, version 3.6.1
-\citep{R.2019}.
+We used a set of geo-referenced covariates to investigate habitat preferences of
+dispersing wild dogs (see \Cref{Covariates}). Covariates fell into one of the
+three categories \textit{land cover}, \textit{protection status}, and
+\textit{human influence}. Note that we did not include elevation as covariate
+due to the absence of noteworthy elevational gradients in our study area. For
+each covariate we obtained spatial raster layers from freely available online
+services or from remote sensed satellite imagery. To ensure a consistent
+resolution (i.e. cell-size or grain) across covariates, all layers were
+coarsened or interpolated to match a resolution of 250m x 250m. Processing and
+manipulation of data, as well as all spatial and statistical analyses were
+performed in R, version 3.6.1 \citep{R.2019}.
 
@@
     core study area. (a) Averaged layer of all binary maps of water and dryland
-    that we acquired using our remote sensing algorithm. (b) Percentage cover of
+    that we acquired using our floodmapping algorithm. (b) Percentage cover of
     trees. (c) Percentage cover of shrubs/grassland. Note that anything not
@@
 
+\newpage
 \subsubsection{Land Cover}
 % Overview
-Land cover was represented by four layers: (1) a binary layer distinguishing
-water and dryland, and (2) three continuous layers showing the percentage cover
-by different vegetation types, together adding up to 100\%.
-
-% Dynamic Watermapping
-(1) For the purpose of this study, water included open waters, rivers, wetlands,
-and swamps, whereas anything else was presumed to be dryland. Because flooding
-of the Okavango Delta is highly variable, we decided to map the flood
-dynamically for the extent shown in \Cref{Floodmapping}. To create dynamic
-``floodmaps'', we applied an algorithm that was originally designed by the
-\textit{Okavango Research Institute} (ORI) and is explained in detail by
-\cite{Wolski.2017}. However, we rewrote and slightly amended the original Python
-code in R and thus describe the underlying principles again and especially
-indicate the discrepancies to the original algorithm.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
-    \caption{Figures describing the floodmapping algorithm. (a) The colored
-    polygons indicate permanent waters and permanent dryland. Below these
-    polygons we extracted reflectance values of MODIS Terra Band 7 and used
-    their repsective medians to calculate a classification threshold. (b)
-    Example of a classified MODIS Terra Band 7 image after application of the
-    threshold.}
-    \label{Floodmapping}
-  \end{center}
-\end{figure}
-
-% Algorithm
-To implement the algorithm, we defined two sets of polygons located in the
-region of the Okavango Delta (see \Cref{Floodmapping}). The first set consisted
-of areas known to be permanent dryland, whereas the second set consisted of
-permanent waters. Since we were unable to retrieve the original polygons used in
-\cite{Wolski.2017}, we geo-referenced one of their illustrations and traced the
-polygons there. After recreation of the polygons, we used the R-package
-\textit{getSpatialData} \citep{Schwalb.2018} to download and pre-process all
-relatively cloud-free MODIS Terra images (MCD43A4, \cite{Schaaf.2015}) available
-for the period of our dispersal events. Assessment of cloud cover was based on
-visual inspection of MODIS images on ORI's website \citep{ORI.2019}. The MCD43A4
-MODIS product is particularly useful for dynamic floodmapping because it is
-updated with new composite images every 8th day \citep{Wolski.2017}. After
-download, we classified each MODIS image into a binary map of water (flood) and
-dryland using a threshold that was identified as follows. First, we extracted
-all reflectance values of MODIS Terra Band 7 within the water- and
-dryland-polygons. Second, we computed histograms of water-reflectances and
-dryland-reflectances and empirically verified that reflectances of the two
-groups were sufficiently distinct. More specifically, we checked if
-superimposing the two histograms resulted in a bimodal plot. This was said to be
-achieved if the 99\textsuperscript{th} percentile of water-reflectances did not
-severely exceed the 1\textsuperscript{st} percentile of dryland-reflectances
-(\(p_{0.99, water} - \frac{10}{255} < p_{0.01, dryland}\)). Third, if peaks were
-sufficiently different, we calculated a threshold (\(t\)) using \Cref{EQ1}:
-
-\begin{equation}
-\label{EQ1}
-t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
-\widetilde{p}_{water})
-\end{equation}
-
-\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
-the median reflectances of water and dryland, respectively. We then classified
-all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
-all pixels with a value smaller than \(t\) as water.
-
-% Dynamic Wetmask
-Importantly, bimodality was not always achieved and in some cases no floodmap
-could be calculated. In fact, it appears that non-bimodality caused the ORI
-algorithm to fail since the end of 2018, such that no floodmaps have been
-generated since then (ORI, personal comm.). We hypothesized that this was caused
-by the application of static water-polygons that did not cover permanent waters
-correctly anymore. We therefore revised the algorithm and allowed for a more
-dynamic polygonization of water. That is, for each MODIS image we calculated new
-water-polygons comprising areas that were covered by the flood in 99\% of the
-floodmaps from the previous five years. All of the necessary floodmaps from
-previous years were kindly provided to us by ORI. Using this slightly amended
-approach, we were able to address some of the bimodality issues and to classify
-several additional floodmaps for the period of our study. Because MODIS Terra
-Band 7 had a resolution of 500m x 500m, we interpolated all maps to 250m x 250m.
-
-% Validation
-To validate and compare the performance of our own algorithm to the original
-ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
-images. To make sure that months were equally represented in the sampled dates,
-we employed stratified sampling based on months (regardless of the year) and
-randomly sampled four maps for each month. For the sampled dates we downloaded
-and classified MODIS Terra Band 7 images and compared our classified images to
-those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
-created a difference map indicating false positives and false negatives and
-computed the relative number of wrongly classified pixels. We achieved an
-overall accuracy of 97\%, which presumably is an underestimate of the true
-performance, as we introduced some errors when resampling the ORI-maps to our
-reference raster.
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
-    \caption{Validation procedure of our floodmapping algorithm. (a) Classified
-    image that was provided to us by ORI. (b) Image for the same date but now
-    classified using our own algorithm. (c) Difference image indicating false
-    positives and false negatives in our own classification.}
-    \label{FloodmappingValidation}
-  \end{center}
-\end{figure}
-
-% Static Watermap
-\noindent While water in the Okavango Delta was mapped dynamically, we used a
-static watermap for its surroundings. This static map was based on Globeland's
-land cover dataset \citep{Chen.2015}, which contains ten land cover categories.
-From the original layer we only retained the categories \textit{wetland} and
-\textit{waterbodies} and collectively reclassified them to \textit{water}.
-Globeland had an original resolution of 30m x 30m, so we coarsened the layer to
-250m x 250m using the mode of each 250m x 250m cell. Lastly, we improved river
-representation using the MERIT Hydro dataset \citep{Yamazaki.2019}, from which
-we added all rivers with a width of over 10m to our watermaps. The entire
-process resulted in several hundred watermaps, of which 72 represented the
-closest watermap to one of our GPS relocations.
-
-(2) The three vegetation layers that we used in addition to the water/dryland
-maps were derived from the MODIS Terra Vegetation Continuous Fields dataset
-(MOD44B, \cite{Dimiceli.2015}). Combined, the layers added up to 100\% and
-depicted the percentage cover of tree-vegetation (henceforth trees),
-non-tree-vegetation (henceforth shrubs/grassland), and non-vegetated (henceforth
-bareland). Since we wanted to make sure that areas covered by the flood would
-get 0\% vegetation cover (i.e. 100\% bareland cover), we used our floodmap that
-aligned with the creation date of the layers (30 March 2017) and set all flooded
-pixels to 100\% bareland. The MODIS vegetation layers had a resolution of 250m x
+Land cover was represented by four layers: a binary layer distinguishing water
+and dryland, and three continuous layers showing the percentage cover by
+different vegetation types, together adding up to 100\%.
+
+% Water Layers
+Water included open waters, rivers, wetlands, and swamps, whereas anything else
+was presumed to be dryland. Because the flood extent of the Okavango Delta is
+highly variable within and between years, we created weekly (8 days)
+``floodmaps'', following the algorithm developed in \cite{Wolski.2017}. The
+algorithm uses thresholding of SWIR reflectances of MODIS Terra satellite
+imagery (MCD43A4, \cite{Schaaf.2015}) to separate water and dryland. As we
+rewrote and slightly amended the original Python code in R, we describe the
+underlying principles and indicate the discrepancies to the original algorithm
+in \Cref{Appendix:FloodmappingAlgorithm}. While water in the Okavango Delta was
+mapped dynamically, we used a static watermap for its surroundings. This static
+map was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
+only retained the categories \textit{wetland} and \textit{waterbodies} and
+collectively reclassified them to \textit{water}. Globeland had an original
+resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
+of each 250m x 250m cell. Lastly, we improved river representation using the
+MERIT Hydro dataset \citep{Yamazaki.2019}, from which we added all rivers with a
+width of over 10m to our watermaps.
+
+% Vegetation Layers
+The three vegetation layers that we used in addition to the water/dryland maps
+were derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
+\cite{Dimiceli.2015}). Combined, the three layers added up to 100\% and depicted
+the percentage cover of tree-vegetation (henceforth trees), non-tree-vegetation
+(henceforth shrubs/grassland), and non-vegetated (henceforth bareland). To
+ensure that areas covered by the flood would get 0\% vegetation cover (i.e.
+100\% bareland cover), we used our floodmap that aligned with the creation date
+of the layers (30 March 2017) and manually defined pixels covered by the flood
+as being 100\% bareland. The MODIS vegetation layers had a resolution of 250m x
 250m, so we did not need to coarsen or interpolate them.
@@
 \subsubsection{Human Influence}
-We depicted human influence in our study area using a proxy that incorporated
-anthropogenic pressures arising from (1) human density, (2) farming, and (3)
-roads. Previous research revealed that veterinary fences have no influence on
-wild dog movement, so we did not include them in our analysis
-\citep{Cozzi.2013b}.
+We defined human influence using a proxy that incorporated anthropogenic
+pressures arising from (1) human density, (2) farming activities, and (3) roads.
+Previous research revealed that veterinary fences have no influence on wild dog
+movements \citep{Cozzi.2013b}, so we did not include them in our analysis.
 
 % Human Density
-(1) Spatial human density estimates were obtained through Facebook's 30m x 30m
+(1) We obtained spatial human density estimates through Facebook's 30m x 30m
 high resolution population density dataset \citep{Facebook.2019}. The dataset
-contained some human density that arised from lodges and other touristic
+contained some human density that arose from lodges and other touristic
 activities that we did not want to depict as human pressures, so we manually
@@
 % Farming
-(2) Information on agricultural activities was sourced from the Globeland
+(2) Information on farming activities was sourced from the Globeland
 \citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets, from
@@
 main tar roads and trunks belonging to the biggest four categories (see
-\Cref{Appendix:RoadCategories}). Finally, we rasterized roads and trunks onto a
-binary raster (1 = roads, 0 = no roads) with 250m x 250m resolution.
+\Cref{RoadsDescription} in \Cref{Appendix:RoadCategories}). Finally, we
+rasterized roads and trunks onto a binary raster (1 = roads, 0 = no roads) with
+250m x 250m resolution.
 
@@
 describing \textit{human density} (continuous), \textit{farming} (binary), and
-\textit{roads} (binary). To reduce the influence of outliers, added values
+\textit{roads} (binary). To reduce the influence of outliers, totaled values
 were limited to a maximum of 50, which visually resulted in a good balance
@@
 surroundings beyond their presence, we followed \cite{Elliot.2014} and applied
-to each pixel a 5km spatial buffer within which we summed up and log-transformed
+to each pixel a 5 km spatial buffer within which we summed up and log-transformed
 human-influence values.
 
-\subsection{Permeability Model}
+\subsection{Habitat Selection Model}
 \label{Modeling}
 % Explanation of iSSF and Extraction
-To parameterize a permeability model and thereby investigate movement and
-habitat preferences of dispersing African wild dogs, we used an integrated step
-selection function (iSSF, \cite{Avgar.2016}). In the iSSF framework, covariates
-along realized steps are contrasted with covariates along random steps that the
-animal could have taken, but decided not to. A step is here defined as the
-connecting line between two consecutive GPS relocations \citep{Turchin.1998}. In
-contrast to regular SSFs, iSSFs require to include movement metrics as
-covariates in the corresponding conditional logistic regression model. Their
-inclusion in turn allows simultaneous inference on habitat and movement
-preferences, as well as to reduce potential biases in estimated habitat
-preferences \citep{Forester.2009, Warton.2013, Avgar.2016}.
-
-In order to conduct an iSSF analysis, we strictly followed published
-recommendations described in \cite{Avgar.2016}. We prepared our GPS relocation
-data for iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and
-coerced relocations recorded during dispersal to steps that were regularly
-spaced four hours apart (allowing for a minor mismatch of up to 15 minutes).
-Steps that were separated by more than four hours (e.g. due to GPS failure) were
-omitted from further analysis. Each remaining step was paired with 24 random
-steps, generated by sampling turning angles from a uniform distribution
-U(\(-\pi, \pi\)) and step lengths from a gamma distribution that was fitted
-using realized step lengths (shape = 0.3677, scale = 6,302)). Together, a
-realized and its 24 associated random steps formed a stratum and received a
-unique identifier. Along realized and random steps we extracted environmental
-covariates as listed in \Cref{ExtractedCovars}. For continuous covariates we
-calculated the average value along the step, for binary covariates the
-percentage cover along the step. Since our floodmaps were dynamic, we extracted
-values from the floodmap that was closest in date to the actual step. We further
-derived a binary variable indicating whether or not a step crossed a road and we
-identified the average distance of the step to the nearest source of water or
-road. The values for \textit{DistanceToRoads} and \textit{DistanceToWater} were
-square-rooted to simulate a decreasing marginal impact of distance. Extracted
-covariates were standardized to a mean of zero and a standard deviation of one
-using a z-score transformation. We also screened covariates for correlation
-using Pearson's Correlation Coefficient. None of the covariates were overly
-correlated (\(|r| > 0.6\), \cite{Latham.2011}) and we retained all of them for
-modeling. To run a proper iSSF analysis, we also included the movement metrics
-\textit{cos(turning angle)} (\(cos(ta\_)\)) and \textit{log(step length)}
-(\(log(sl\_)\)) as covariates in our models \citep{Avgar.2016}. The movement
-metric \(cos(ta\_\)) served to describe the directionality of a step, as it
-transformed the circular measure of (\(-\pi, +\pi\)) into a linear measure (-1,
-1), where positive values indicated forward movements, whereas negative values
-indicated backwards movements \citep{Turchin.1998}. The movement metric
-\(log(sl\_\)), on the other hand, was used as an indicator of the preferred step
-length. Since steps were regularly spaced by four hours, \(log(sl\_\)) can also
-be interpreted as movement rate.
+We used an integrated step selection function approach (iSSF, \cite{Avgar.2016})
+to investigate dispersers' selection or avoidance for the above mentioned
+spatial covariates. In the iSSF framework, covariates along realized steps are
+contrasted with covariates along alternative random steps that the animal could
+have taken, but decided not to. A step is here defined as the connecting line
+between two consecutive GPS relocations \citep{Turchin.1998}. In contrast to
+regular SSFs, iSSFs require to include movement metrics as covariates in the
+corresponding conditional logistic regression model. Their inclusion in turn
+allows simultaneous inference on habitat and movement preferences, as well as to
+reduce potential biases in estimated habitat preferences \citep{Forester.2009,
+Warton.2013, Avgar.2016}.
+
+In order to conduct an iSSF analysis, we followed recommendations described in
+Appendix S1 in \cite{Avgar.2016}. We prepared our GPS relocation data for
+iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
+relocations recorded during dispersal to steps that were regularly spaced four
+hours apart (allowing for a minor mismatch of up to 15 minutes). Steps that were
+separated by more than four hours (e.g. due to GPS failure) were omitted from
+further analysis. Each remaining step was paired with 24 random steps, generated
+by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
+lengths from a gamma distribution that was fitted using realized step lengths
+(shape = 0.3677, scale = 6,302)). Together, a realized and its 24 associated
+random steps formed a stratum and received a unique identifier.
+
+We extracted spatial covariates along realized and random steps as listed in
+\Cref{ExtractedCovars}. For continuous covariates we calculated the average
+value along the step, for binary covariates the percentage cover along the step.
+Since our floodmaps were dynamic, we extracted values from the floodmap that was
+closest in date to the actual step. We further derived a binary variable
+indicating whether or not a step crossed a road and we identified the average
+distance of each step to the nearest source of water or to the nearest road. We
+square-rooted the values for \textit{DistanceToRoads} and
+\textit{DistanceToWater} to render a decreasing marginal impact of distance. We
+standardized extracted covariates to a mean of zero and a standard deviation of
+one using a z-score transformation. Furthermore, we screened covariates for
+correlation using Pearson's Correlation Coefficient. None of the covariates were
+overly correlated (\(|r| > 0.6\), \cite{Latham.2011}) and we retained all of
+them for modeling. To run a proper iSSF analysis, we also included two movement
+metrics, namely the cosine of the turning angle (\(cos(ta)\)) and the logarithm
+of the step length (\(log(sl)\)), as covariates in our models
+\citep{Avgar.2016}. The movement metric \(cos(ta\)) served to describe the
+directionality of a step, as it transformed the circular measure of (\(-\pi,
++\pi\)) into a linear measure (-1, 1), where positive values indicated forward
+movements, whereas negative values indicated backwards movements
+\citep{Turchin.1998}. The movement metric \(log(sl\)), on the other hand, was
+included as an indicator of the preferred step length. Since steps were
+regularly spaced by four hours, \(log(sl\)) can also be interpreted as movement
+rate.
 
@@
   \begin{center}
-    \caption{Overview of environmental covariates extracted along steps (see
+    \caption{Overview of spatial covariates extracted along steps (see
     \Cref{Sources} in \Cref{Appendix:Sources} for an overview of the sources and
@@
         & HumansBuffer5000
-          & Average buffered (5km) human influence along step
+          & Average buffered (5 km) human influence along step
             & \(\geq\) 0\\
@@
 % Permeability Model
-\noindent We used the iSSF framework and parameterized a permeability model that
-further served to predict landscape permeability. This permeability model
+\noindent We used the iSSF framework and parameterized a habitat selection model
+that further served to predict landscape permeability. This selection model
 operated under the assumption that dispersing wild dogs assigned a selection
@@
 conditional logistic regression model \citep{Fortin.2005}. In this model,
-positive \(\beta\)-coefficients indicate \textit{selection} of a covariate,
-negative \(\beta\)-coefficients \textit{avoidance} of a covariate. Selected
-covariates therefore contribute positively to expected landscape permeability,
-whereas avoided covariates reduce landscape permeability.
+positive \(\beta\)-coefficients indicate selection of a covariate, negative
+\(\beta\)-coefficients avoidance of a covariate. Selected covariates therefore
+contribute positively to expected landscape permeability, whereas avoided
+covariates contribute negatively to landscape permeability.
 
 % Mixed Effects Conditional Logistic Regression Model
-To deal with multiple individuals, we applied a novel modeling technique for
-mixed effects conditional logistic regressions that was recently proposed by
-\cite{Muff.2019}. Their approach allowed to include individuals for whom some
-covariates did not vary, as well as to model random slopes for each covariate.
-We implemented their method using the R-package \textit{glmmTMB}
-\citep{Mollie.2017}, where we fixed the variance of the random intercept to an
-artificially high value of 10\textsuperscript{12} (for an explanation, see
-\cite{Muff.2019}) and used animal ID to model random intercepts and slopes.
+To deal with multiple individuals, we applied mixed effects conditional logistic
+regression analysis following \cite{Muff.2019}. Their approach allowed to
+include individuals for whom some covariates did not vary, as well as to model
+random intercepts and slopes for each covariate. We implemented their method
+using the R-package \textit{glmmTMB} \citep{Mollie.2017}, where we fixed the
+variance of the random intercept to an artificially high value of
+10\textsuperscript{12} (for an explanation, see \cite{Muff.2019}) and used
+animal ID to model random intercepts and slopes.
 
 % AIC Model Selection
-We defined the movement metrics \(cos(ta\_)\) and \(log(sl\_)\) as core
-covariates and ran forward model selection based on Akaike's Information
-Criterion (AIC, \cite{Burnham.2002}) for environmental covariates. That is, we
-started with a univariate model and iteratively increased model complexity upon
-reaching a full model. We then ranked models according to AIC, assessed relative
-model weights, and identified the most parsimonious model. Due to convergence
-issues we were unable to model interactions between environmental covariates.
+We defined the movement metrics \(cos(ta)\) and \(log(sl)\) as core covariates
+and ran forward model selection based on Akaike's Information Criterion (AIC,
+\cite{Burnham.2002}) for all other spatial covariates. That is, we started with
+a basic model and iteratively increased model complexity upon reaching a
+full model. We then ranked models according to AIC, assessed relative model
+weights, and identified the most parsimonious model. Due to convergence issues
+we were unable to model interactions between spatial covariates.
 
 % Validation
-To validate the predictive power of the most parsimonious permeability model, we
+To validate the predictive power of the most parsimonious selection model, we
 ran k-fold cross-validation for case-control studies as described in
 \cite{Fortin.2009}. Using 80\% of randomly selected strata, we parameterized a
-permeability model and predicted selection scores \(w(x)\) for all steps in the
+selection model and predicted selection scores \(w(x)\) for all steps in the
 remaining 20\% of strata. According to predicted selection scores we assigned
@@
 \subsection{Permeability Surface}
-We used the most parsimonious permeability model and applied \Cref{EQ2} to
-predict a permeability surface over the entire extent of the KAZA-TFCA. Stated
-differently, we calculated the selection score \(w(x)\) for each 250m x 250m
-pixel covering our extended study area given the pixel's environmental
-covariates. Based on the definition of permeability, high selection scores
-resulted in high permeability, whereas low selection scores resulted in low
-permeability. To reduce the influence of outliers in predicted scores, we
-followed \cite{Squires.2013} and curtailed predicted permeabilities between the
-1st and 99th percentile of their original values. To predict permeability in the
-first place, we needed to collapse all dynamic watermaps into a single layer,
-which we achieved by identifying areas that were inundated in at least a quarter
-of all our floodmaps. We therefore predicted permeability assuming a relatively
-extended flood.
+We used the most parsimonious selection model and applied \Cref{EQ2} to predict
+a permeability surface over the entire extent of the KAZA-TFCA. In other words,
+we calculated the selection score \(w(x)\) for each 250m x 250m pixel covering
+our extended study area given the pixel's spatial covariates. Note that the
+movement metrics, i.e. \(cos(ta)\) and \(log(sl)\), did not enter the prediction
+of landscape permeability as their inclusion in our habitat selection model
+solely served to reduce the bias in estimated habitat preferences. By definition
+of permeability, high selection scores resulted in high permeability, whereas
+low selection scores resulted in low permeability. To reduce the influence of
+outliers in predicted scores, we followed \cite{Squires.2013} and curtailed
+predicted scores between the 1\textsuperscript{st} and 99\textsuperscript{th}
+percentile of their original values. To predict permeability in the first place,
+we needed to collapse all dynamic watermaps into a single layer, which we
+achieved by identifying areas that were inundated in at least a quarter of all
+our floodmaps. In order to be able to compare permeability across different
+regions, we rescaled the permeability surface to a range between 0 and 1 using
+\(f(x) = \frac{x - min(x)}{max(x) - min(x)}\). The most impermeable landscape
+therefore received a value of 0, whereas the most permeable landscape received a
+value of 1.
 
@@
 To identify movement corridors within the extent of KAZA-TFCA, we calculated
-least-cost paths (LCPs) and least-cost corridors (LCCs) between source points
-located in protectorates of our extended study area \citep{Adriaensen.2003,
-Sawyer.2011}. Source points were generated by overlaying the study area with a
-regular grid of points, equally spaced 100km apart and located within protected
-areas that contiguously covered at least 700km\textsuperscript{2}. The threshold
-of 700km\textsuperscript{2} was based on home-range requirements of African wild
-dogs reported in \cite{Pomilia.2015} and served to eliminate locations that were
-assumed unsuitable for supporting viable wild dog populations. Finally, we
-defined centroids as source points for those protectorates with over
-700km\textsuperscript{2} that did not get any source points from the regular
-grid. In total, we generated 68 source points, which resulted in 2,278 unique
-pairwise combinations of source points and therefore 2,278 unique LCPs and LCCs.
+least-cost paths (LCPs) and least-cost corridors (LCCs) between selected points
+located in protected areas of our extended study area \citep{Adriaensen.2003,
+Sawyer.2011}. Selected points were generated by overlaying the study area with a
+regular grid of points, equally spaced 100 km apart and located within protected
+areas that contiguously covered at least 700 km\textsuperscript{2}. The
+threshold of 700 km\textsuperscript{2} was based on home-range requirements of
+African wild dogs reported in \cite{Pomilia.2015} and served to eliminate
+locations that were assumed unsuitable for supporting viable wild dog
+populations. Finally, we defined centroids as source points for those protected
+areas larger than 700 km\textsuperscript{2} that did not get any source points
+from the regular grid. In total, we generated 68 points, which resulted in 2,278
+unique pairwise combinations and therefore 2,278 unique LCPs and LCCs.
 
 LCP analysis between source points was implemented using the R-package
-\textit{gdistance} \citep{vanEtten.2018}. The package translated the
+\textit{gdistance} \citep{vanEtten.2018}. The package translated the (unscaled)
 permeability surface into a network of nodes and used graph theory, in
@@
 probabilities for moving from one cell to another. In our case, the transition
-probability of moving between two adjacent cells depended positively on their
-averaged permeability. We allowed individuals to move from each cell to the
-cell's 8 surrounding neighbors (i.e. Moores neighborhood) and applied a
-geographic correction to account for the fact that diagonal neighbors were more
-remote than orthogonal neighbors. Because African wild dogs have been observed
-to cover massive dispersal distances (\citeauthor{DaviesMostert.2012}, 2012;
-\citeauthor{Masenga.2016}, 2016; \citeauthor{Cozzi.2019}, in review), we did not
-limit LCPs to a maximal effective cost. After computation, we tallied
-overlapping LCPs to identify high-frequency routes.
+probability of moving between two adjacent cells depended on their averaged
+permeability. We allowed individuals to move from each cell to the cell's 8
+surrounding neighbors (i.e. Moores neighborhood) and applied a geographic
+correction to account for the fact that diagonal neighbors were more remote than
+orthogonal neighbors. Because African wild dogs have been observed to cover
+massive dispersal distances \citep{DaviesMostert.2012, Masenga.2016,
+Cozzi.2020}, we did not limit LCPs to a maximal effective cost. After
+computation, we tallied overlapping LCPs to identify high-frequency routes.
 
@@
 complete knowledge of the landscape and associated movement costs and therefore
-always move along the cheapest route \citep{Carroll.2012}. Yet, individuals
+always move along the least-costly route \citep{Carroll.2012}. Yet, individuals
 rarely use a single optimum path, so that LCPs are often of limited biological
@@
 cost map (see \Cref{LCCExample}), which indicated the total minimal costs
-required to get from the respective source point to any other location on the
-map. A LCC between two source points was then obtained by adding up their
-cumulative cost maps and masking out all cell-values exceeding the lowest
-cell-value by more than 5\%. We repeated this procedure for each possible
-pairwise combination of source points and thereby identified LCCs between all
-source points. The resulting corridor-maps were normalized to a range from zero
-to one using \(f(x) = \frac{x - min(x)}{max(x) - min(x)}\) and totaled up into a
-single connectivity map.
+required to get from the selected point to any other location on the map. A LCC
+between two points was then obtained by adding up their cumulative cost maps and
+masking out all cell-values exceeding the lowest cell-value by more than 5\%. We
+repeated this procedure for each possible pairwise combination of source points
+and thereby identified LCCs between all 68 selected points. The resulting
+corridor-maps were normalized to a range from zero to one using \(f(x) = \frac{x
+- min(x)}{max(x) - min(x)}\) and totaled up into a single connectivity map.
 
@@
   \begin{center}
-  \begin{minipage}{.32\linewidth}
+  \begin{minipage}{.25\linewidth}
     \begin{subfigure}[t]{.9\linewidth}
@@
   \end{minipage}
-  \begin{minipage}{.33\linewidth}
+  \begin{minipage}{.25\linewidth}
     \begin{subfigure}[t]{.9\linewidth}
@@
   \end{minipage}
-  \caption{Figures illustrating the process of identifying a least-cost
-  corridor between the source points A and B. (a) Example of a permeability
-  surface, which serves to determine the costs of movement. (b1) Cumulative
-  cost map for point A, depicting the total minimal costs necessary to get
-  from point A to every other location. (b2) Cumulative cost map for point B,
-  depicting the total minimal costs necessary to get from point B to every
-  other location. (c) Summed cost maps of points A and B. (d) Masked out
-  corridor containing pixels that do not exceed the cheapest pixel by more
-  than 5\%.}
+  \caption{Figures illustrating the process of identifying a least-cost corridor
+  between the source points A and B following \citep{Pinto.2009}. (a) Example of
+  a permeability surface, which serves to determine the costs of movement. (b1)
+  Cumulative cost map for point A, depicting the total minimal costs necessary
+  to get from point A to every other location. (b2) Cumulative cost map for
+  point B, depicting the total minimal costs necessary to get from point B to
+  every other location. (c) Summed cost maps of points A and B. (d) Masked out
+  corridor containing pixels that do not exceed the cheapest pixel by more than
+  5\%.}
   \label{LCCExample}
@@
 \section{Results}
-\subsection{Permeability Model}
-Our most parsimonious permeability model retained the covariates \textit{Water},
-\textit{DistanceToWater}, \textit{Trees}, \textit{Shrubs/Grassland},
-\textit{HumansBuff5000}, and, of course, the fixed covariates \(cos(ta\_)\) and
-\(log(sl\_)\) (see \Cref{PermeabilityResults}a). The model strictly outperformed
-all alternative models (\(\Delta AIC > 2\)) and therefore received an AIC-weight
-of one (see \Cref{ModelAICs} in \Cref{AICs}). Except for the effect of
-\textit{DistanceToWater}, all estimates were significant on the 95\%-confidence
-interval (CI). Parameter estimates from the model showed that dispersing wild
-dogs preferred low turning angles (i.e. high \(cos(ta\_)\)), indicating that
-they moved in a linear manner during dispersal (\(\beta\) = 0.14; 95\% CI = 0.07
-to 0.21). Similarly, they preferred a high movement rate, as evidenced by a
-positive selection coefficient for \(log(sl\_)\) (\(\beta\) = 0.05, 95\% CI =
-0.02 to 0.09). With respect to land cover, dispersers strictly avoided moving
-through water and therefore preferably traversed dryland when available
-(\(\beta\) = -0.5, 95\% CI -0.67 to -0.32). Although the effect was not
-statistically clear, dispersers preferred locations in proximity to water, as
-the estimated parameter for \textit{DistanceToWater} was negative (\(\beta\) =
--0.24, 95\% CI = -0.67 to 0.19). Furthermore, dispersers avoided areas that were
-densely covered by trees (\(\beta\) = -0.18, CI = -0.32 to -0.04) and preferred
-areas covered by shrubs/grassland (\(\beta\) = 0.40, 95\% CI = 0.24 to 0.56).
-Finally, dispersers strictly avoided areas that were strongly influenced by
-humans (\(\beta\) = -0.47, 95\% CI = -0.87 to -0.07).
+\subsection{Habitat Selection Model}
+Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
+alternative model; see \Cref{ModelAICs} in \Cref{AICs}) retained the covariates
+\textit{Water}, \textit{DistanceToWater}, \textit{Trees},
+\textit{Shrubs/Grassland}, \textit{HumansBuff5000}, and the fixed covariates
+\(cos(ta)\) and \(log(sl)\) (\Cref{PermeabilityResults}a). Parameter estimates
+showed that dispersing wild dogs moved in a directional and fast manner, as
+indicated by a positive selection for small turning angles, i.e. high
+\(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21,
+\Cref{PermeabilityResults}a) and longer steps, i.e. high \(log(sl)\) (\(\beta\)
+= 0.05, 95\% CI = 0.02 to 0.09, \Cref{PermeabilityResults}a). Given the chance,
+dispersers avoided moving through water (\(\beta\) = -0.5, 95\% CI -0.67 to
+-0.32, \Cref{PermeabilityResults}a) but selected for locations in its vicinity,
+although the effect was not significant (\(\beta\) = -0.24, 95\% CI = -0.67 to
+0.19, \Cref{PermeabilityResults}a). Dispersers avoided areas that were densely
+covered by trees (\(\beta\) = -0.18, CI = -0.32 to -0.04,
+\Cref{PermeabilityResults}a) and preferred areas covered by shrubs/grassland
+(\(\beta\) = 0.40, 95\% CI = 0.24 to 0.56, \Cref{PermeabilityResults}a).
+Finally, dispersers avoided areas that were influenced by human presence and
+activities (\(\beta\) = -0.47, 95\% CI = -0.87 to -0.07,
+\Cref{PermeabilityResults}a).
 
@@
 intervals of \(\bar{r}_{s, realized}\) (-0.45, CI = -0.48 to -0.42) and
-\(\bar{r}_{s, random}\) (0.04, CI = 0.01 to 0.08) did not overlap (see
-\Cref{PermeabilityResults}b). Likewise, the significant correlation between
+\(\bar{r}_{s, random}\) (0.04, CI = 0.01 to 0.08) did not overlap
+(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
 ranks and corresponding frequencies for realized steps suggested a good fit
@@
 \subsection{Permeability Surface}
-\Cref{PermeabilityMap}a illustrates the predicted permeability surface for the
-extent of the KAZA-TFCA as computed from the most parsimonious permeability
-model. Note that \(cos(ta\_)\) and \(log(sl\_)\) did not enter the prediction of
-landscape permeability as their inclusion in our permeability model solely
-served to reduce the bias in estimated habitat preferences. Our prediction
-revealed substantial regional differences in predicted permeability across the
-entire study area, with some regions being highly permeable whereas others are
-almost insurmountable. A comparisons of median (Mdn) permeabilities across
-countries (see \Cref{PermeabilityComp}) showed that in particular Angola (Mdn =
-1.03) and Botswana (Mdn = 0.91) support highly permeable landscapes, whereas
-Zimbabwe (Mdn = 0.19) and Zambia (Mdn = 0.15) were found to be largely
-impermeable. Namibia (Mdn = 0.59), on the other hand, ranges somewhere in
-between the two extremes. Throughout all countries, permeabilities inside areas
-covered by the KAZA-TFCA were predicted to be vastly greater than permeabilities
-outside areas covered by the KAZA-TFCA, resulting in a stark contrast of the
-overall permeability inside the KAZA-TFCA (Mdn = 0.75) and outside the KAZA-TFCA
-(Mdn = 0.26). Similarly, for all countries we found that protected areas provide
-higher permeabilities than pastoral regions, altogether making protected areas
-highly permeable (Mdn = 0.74) in comparison to pastoral regions (Mdn = 0.27). In
-general, northern Botswana appeared to provide the highest density of permeable
-landscapes, as judged by the bright yellow colors in \Cref{PermeabilityMap}a. A
-closer look at this region (see \Cref{PermeabilityMap}b) showed that especially
-close surroundings of the Okavango Delta and the Linyanti Swamp host permeable
-landscapes for dispersers. Nevertheless, the waters of the Okavango Delta and
-Linyanti Swamp themselves provide highly impermeable landscapes, as indicated by
-the dark colors sketching the two water-bodies. Moreover, densely populated
-areas such as the city of Maun or the villages Khwai, Mababe, and Sankuyo
-provide low permeability, which is also true for the highly populated stripe
-flanking the Okavango Delta towards its west and north-west.
-
-\newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
+Our prediction revealed substantial differences in predicted permeability across
+the entire study area (\Cref{PermeabilityMap}). Comparisons of median (Mdn)
+permeability values (\Cref{PermeabilityComp}) showed that permeability inside
+the boundaries of the KAZA-TFCA (Mdn = 0.21) was three times as high as
+permeability outside it (Mdn = 0.07). A comparison across countries
+(\Cref{PermeabilityComp}) showed that Angola (Mdn = 0.29) and Botswana (Mdn =
+0.25) were characterized by comparably highly permeable landscapes, whereas
+Zimbabwe (Mdn = 0.05) and Zambia (Mdn = 0.04) were relatively impermeable.
+Namibia (Mdn = 0.16) ranged somewhere in between the two extremes. Across all
+countries, protected areas provided higher permeability (Mdn = 0.21) than
+human-dominated landscapes (Mdn = 0.07). A visual assessment of our permeability
+surface revealed that northern Botswana had the highest density of permeable
+landscapes (\Cref{PermeabilityMap}, bright yellow colors), despite the waters
+of the Okavango Delta and Linyanti Swamp themselves provided highly impermeable
+landscapes (\Cref{PermeabilityMap}, dark blue colors).
+
+% \newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
+% \begin{figure}[hbtp]
+%   \begin{center}
+%     \begin{tikzpicture}
+%         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
+%           \begin{minipage}{0.95\textwidth}
+%             \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
+%             \includegraphics[width = \textwidth]{99_PermeabilityMap2.pdf}
+%           \end{minipage}
+%         };
+%         \begin{scope}[x={(image.south east)},y={(image.north west)}]
+%            %  % next four lines will help you to locate the point needed by
+%            %  % forming a grid. comment these four lines in the final picture.↓
+%            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
+%            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
+%            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
+%            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
+%            %  % upto here
+%             \draw[black, line width = 1pt] (0.389, 0.627) -- (0.080, 0.477);
+%             \draw[black, line width = 1pt] (0.582, 0.627) -- (0.980, 0.477);
+%             \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
+%             \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
+%             \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
+%         \end{scope}
+%     \end{tikzpicture}
+%     \caption{(a) Predicted permeability surface for the extent of the KAZA-TFCA.
+%     Permability was predicted by calculating selection scores \(w(x)\) for each
+%     pixel based on the pixel's underlying covariates. Areas that dispersers find
+%     easy to traverse are depicted in bright colors. Bold white lines delineate
+%     the borders of the KAZA-TFCA, whereas dashed white lines show country
+%     borders. (b) Detailed view of the Okavango Delta and Linyanti Swamp region,
+%     which appears to provide the most suitable landscape for dispersers. The
+%     white square highlights our core study area.}
+%     \label{PermeabilityMap}
+%   \end{center}
+% \end{figure}
+% \restoregeometry
+
+% \newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
 \begin{figure}[hbtp]
@@
         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-          \begin{minipage}{0.95\textwidth}
+          \begin{minipage}{0.7\textwidth}
             \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
-            \includegraphics[width = \textwidth]{99_PermeabilityMap2.pdf}
           \end{minipage}
@@
            %  % upto here
-            \draw[black, line width = 1pt] (0.389, 0.627) -- (0.080, 0.477);
-            \draw[black, line width = 1pt] (0.582, 0.627) -- (0.980, 0.477);
-            \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
-            \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
-            \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
+          % \node[text=white, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=white] (No1) at (0.440, 0.350){1};
+          % \node[text=white, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=white] (No2) at (0.520, 0.438){2};
         \end{scope}
     \end{tikzpicture}
-    \caption{(a) Predicted permeability surface for the extent of the KAZA-TFCA.
+    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
     Permability was predicted by calculating selection scores \(w(x)\) for each
@@
     the borders of the KAZA-TFCA, whereas dashed white lines show country
-    borders. (b) Detailed view of the Okavango Delta and Linyanti Swamp region,
-    which appears to provide the most suitable landscape for dispersers. The
-    white square highlights our core study area.}
+    borders.}
     \label{PermeabilityMap}
@@
 \end{figure}
-\restoregeometry
+% \restoregeometry
+
+% \begin{table}[h]
+%   \caption{Comparison of median permeability across countries, separated
+%   according to areas within and outside the KAZA-TFCA, as well as within and
+%   outside protected areas. High values indicate high permeability, whereas low
+%   values correspond to low permeability.}
+%   \label{PermeabilityComp}
+%   \begin{center}
+%     \resizebox{\textwidth}{!} {
+%       \begin{tabular}{llllllllr}
+%       & & \multicolumn{2}{c}{KAZA-TFCA} && \multicolumn{2}{c}{Protection Status} \\
+%       \cline{3-4} \cline{6-7}
+%       Country & & Inside & Outside & & Protected & Pastoral & & Overall \\
+%       \hline
+%       Angola & & 1.67 & 0.59 & & 1.67 & 0.60 & & 1.03 \\
+%       Namibia & & 0.91 & 0.49 & & 0.98 & 0.44 & & 0.59 \\
+%       Botswana & & 1.17 & 0.69 & & 1.37 & 0.62 & & 0.91 \\
+%       Zimbabwe & & 0.25 & 0.18 & & 0.32 & 0.17 & & 0.19 \\
+%       Zambia & & 0.26 & 0.12 & & 0.23 & 0.13 & & 0.15 \\
+%       \hline
+%       Overall & & 0.75 & 0.26 & & 0.74 & 0.27 & & 0.38 \\
+%       \end{tabular}
+%     }
+%   \end{center}
+% \end{table}
 
@@
   \begin{center}
-    \resizebox{0.65\textwidth}{!} {
+    \resizebox{\textwidth}{!} {
       \begin{tabular}{llllllllr}
@@
       \hline
-      Angola & & 1.67 & 0.59 & & 1.67 & 0.60 & & 1.03 \\
-      Namibia & & 0.91 & 0.49 & & 0.98 & 0.44 & & 0.59 \\
-      Botswana & & 1.17 & 0.69 & & 1.37 & 0.62 & & 0.91 \\
-      Zimbabwe & & 0.25 & 0.18 & & 0.32 & 0.17 & & 0.19 \\
-      Zambia & & 0.26 & 0.12 & & 0.23 & 0.13 & & 0.15 \\
+      Angola & & 0.46 & 0.18 & & 0.46 & 0.18 & & 0.29 \\
+      Namibia & & 0.26 & 0.13 & & 0.27 & 0.12 & & 0.16 \\
+      Botswana & & 0.33 & 0.19 & & 0.38 & 0.17 & & 0.25 \\
+      Zimbabwe & & 0.06 & 0.05 & & 0.09 & 0.04 & & 0.05 \\
+      Zambia & & 0.07 & 0.03 & & 0.06 & 0.03 & & 0.04 \\
       \hline
-      Overall & & 0.75 & 0.26 & & 0.74 & 0.27 & & 0.38 \\
+      Overall & & 0.21 & 0.07 & & 0.21 & 0.07 & & 0.11 \\
       \end{tabular}
@@
 \subsection{Least-Cost Paths \& Least-Cost Corridors}
-\Cref{LeastCost}a shows predicted LCPs and \Cref{LeastCost}b corresponding LCCs
-for the extent of the KAZA-TFCA. Because LCPs identify the cheapest set of
-pixels between two source points, they typically run along the center of LCCs
-and traverse the same regions. The two figures thus essentially contain the same
-information, albeit \Cref{LeastCost}b appears more detailed and reveals some
-routes that would remain undetected in \Cref{LeastCost}a. Most notably, our
-connectivity analysis revealed that several LCPs and LCCs traverse the
-Chobe-Moremi region in northern Botswana (\Cref{LeastCost}a \circled{1}). In
-fact, the highest frequented route in this region is passed by 637 LCPs, which
-is reflected in a wide, dark red LCC in \Cref{LeastCost}b. Departing from this
-central region, several high-frequency LCPs and LCCs emerge. One route
-(\circled{2}) runs towards the north-western part of the KAZA-TFCA into the
-Mavinga and Luengue-Luiana NPs. A further route (\circled{3}) meanders
-north-east through unprotected area into the Kafue NP in Zambia. Another major
-route (\circled{4}) extends east, into the Hwange NP in Zimbabwe and eventually
-reaches the Matusadona NP (\circled{5}). A slightly less pronounced but still
-distinguishable route (\circled{6}) originates from the Chobe-Moremi ecosystem
-and runs south, through unprotected landscape, into the Central Kalahari NP in
-Botswana. Alternatively, the corridor turns west (\circled{8}) and moves into
-the Khaudum NP in Namibia (\circled{7}). Finally, another slightly less
-accentuated path originates from southern Angola, moves through Botswana and
-leads into the Khaudum NP in Namibia. Besides these medium- to high-frequency
-routes, several low-frequency corridors exist and provide connectivity between
-other protected areas. Strikingly, we found few direct LCPs and LCCs between
-peripheral source points. That is, most paths and corridors connecting two
-adjacent peripheral points detour through more central regions before they head
-towards their destination at the periphery. As a consequence, there are several
-peripheral regions providing little to no connectivity, as they are not crossed
-by any LCPs or LCCs. Such regions are better visible in \Cref{LeastCost}b where
-they appear as extended dark areas. Overall, there is a gradual decrease in the
-density of LCPs and LCCs when moving from the central region towards the border
-of the extended study area.
+Because LCPs identify the cheapest set of pixels between two selected points,
+they typically run along the center of LCCs and traverse the same regions. The
+two methods thus essentially return very similar information, but we show both
+for the sake of clarity and comparison and because LCPs help more precisely
+identifying discrete corridors. Our analysis revealed three major corridors
+(\Cref{LeastCost}): one runs SE-NW and connects the Okavango-Linyanti ecosystem
+in Botswana with Luengue-Luiana NP in Angola. One runs W-E between Chobe NP in
+Botswana and Zimbabwe’s Hwange NP. One runs NE-SW, completely across unprotected
+areas, and connects Kafue NP in Zambia with more central regions of the
+KAZA-TFCA. With 637 LCPs passing through it, the Okavango-Linyanti region is the
+highest frequented corridor. Several minor routes branch off these three major
+corridors; this includes a southward connection between the Okavango-Linyanti
+and the Central Kalahari Game Reserve, a southwesterly corridor connecting
+Luengue-Luiana NP with Namibia’s Kaudom NP, and a northeasterly extension of the
+Hwange corridor into Zimbabwe’s Matusadona NP. Our model did not detect any
+significant direct corridors between Zimbabwe and Zambia and only very limited
+W-E direct connection between the Okavango region and Namibia’s Khaudum NP. We
+also found little to no direct connectivity between peripheral points; that is,
+most paths and corridors connecting two adjacent peripheral points run through
+more central regions before heading towards their destination at the periphery.
+With the exception of the Central Kalahari Game Reserve corridor, our model did
+not detect any significant connectivity outside the boundaries of KAZA-TFCA.
 
@@
           % % upto here
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No1) at (0.400, 0.700){1};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No2) at (0.370, 0.760){2};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No3) at (0.585, 0.780){3};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No4) at (0.647, 0.697){4};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No5) at (0.910, 0.770){5};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No6) at (0.580, 0.658){6};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No7) at (0.410, 0.640){7};
-          \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
-            , draw=black] (No8) at (0.380, 0.665){8};
-          \draw[black, line width = 1pt, densely dotted] (No1) -- (0.500, 0.710);
-          \draw[black, line width = 1pt, densely dotted] (No2) -- (0.438, 0.760);
-          \draw[black, line width = 1pt, densely dotted] (No3) -- (0.565, 0.805);
-          \draw[black, line width = 1pt, densely dotted] (No4) -- (0.665, 0.685);
-          \draw[black, line width = 1pt, densely dotted] (No5) -- (0.885, 0.765);
-          \draw[black, line width = 1pt, densely dotted] (No6) -- (0.540, 0.650);
-          \draw[black, line width = 1pt, densely dotted] (No7) -- (0.450, 0.645);
-          \draw[black, line width = 1pt, densely dotted] (No8) -- (0.345, 0.690);
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No1) at (0.400, 0.700){1};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No2) at (0.370, 0.760){2};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No3) at (0.585, 0.780){3};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No4) at (0.647, 0.697){4};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No5) at (0.910, 0.770){5};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No6) at (0.580, 0.658){6};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No7) at (0.410, 0.640){7};
+          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
+          %   , draw=black] (No8) at (0.380, 0.665){8};
+          % \draw[black, line width = 1pt, densely dotted] (No1) -- (0.500, 0.710);
+          % \draw[black, line width = 1pt, densely dotted] (No2) -- (0.438, 0.760);
+          % \draw[black, line width = 1pt, densely dotted] (No3) -- (0.565, 0.805);
+          % \draw[black, line width = 1pt, densely dotted] (No4) -- (0.665, 0.685);
+          % \draw[black, line width = 1pt, densely dotted] (No5) -- (0.885, 0.765);
+          % \draw[black, line width = 1pt, densely dotted] (No6) -- (0.540, 0.650);
+          % \draw[black, line width = 1pt, densely dotted] (No7) -- (0.450, 0.645);
+          % \draw[black, line width = 1pt, densely dotted] (No8) -- (0.345, 0.690);
         \end{scope}
@@
     leaving from protected areas (light grey). Note that only contiguous
-    protected areas covering more than 700km\textsuperscript{2} are depicted.
-    Circled numbers indicate some of the highest frequented least-cost paths
-    that are mentioned in the main text. Continuous black lines indicate the
-    borders of the KAZA-TFCA, whereas dashed black lines delineate
-    country-borders. (b) Least-cost corridors between the same source points as
-    illustrated in subfigure (a). For ease of spatial reference we also labeled
-    national parks that are mentioned in the main text (dark-grey).}
+    protected areas covering more than 700 km\textsuperscript{2} are depicted.
+    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
+    dashed black lines delineate country-borders. (b) Least-cost corridors
+    between the same source points as illustrated in subfigure (a). For ease of
+    spatial reference we also labeled national parks that are mentioned in the
+    main text (dark-grey).}
     \label{LeastCost}
@@
 \section{Discussion}
-The goal of our study was to (1) assess landscape permeability for dispersing
-wild dogs and to (2) subsequently use this knowledge to examine connectivity
-across the greater extent of the KAZA-TFCA. To achieve these goals, we
-parameterized a permeability model in which we used GPS data of 16 dispersing
-wild dog coalitions to quantify their habitat preferences with respect to land
-cover, protection status, and human influence. While we found significant
-impacts of different land covers and human influence, the influence of
-protection status was not statistically clear. We further applied the
-parameterized permeability model and predicted a permeability surface for the
-entire KAZA-TFCA, which depicted the ease at which dispersers crossed a specific
-landscape. Finally, based on this permeability map, we computed LCPs and LCCs
-between protectorates and thereby highlighted dispersal corridors of African
-wild dogs.
-
-\subsection{Permeability Model}
-% Permeability Model: Land Cover, Water
-Our most parsimonious permeability model revealed that, with respect to land
-cover, dispersers strictly avoided areas covered by water (including swamps,
-wetlands). This result corroborates previous research on wild dogs (e.g.
-\cite{Abrahms.2017}) and is rather unsurprising, as the species is not well
-adapted for moving through water. For residents, water is almost impermeable and
-frequently serves as natural boundary to their home-ranges \citep{Cozzi.2013}.
-It is conceivable that dispersers are more willing than residents to cross
-waterbodies (\citeauthor{Cozzi.2019}, in review), but we did not include
-residents in our analyses and could not test for such differences. Regardless of
-that, water appears to pose a serious obstacle for dispersers and is therefore
-likely to reduce landscape permeability for dispersing wild dogs.
-
-% Permeability Model: Land Cover, DistanceToWater
-While dispersers avoided moving \textit{through} water, they preferred moving
-\textit{close to} water as indicated by avoidance of high distance to water.
-While its effect was insignificant, inclusion of the covariate
-\textit{DistanceToWater} strictly increased the AIC in all of our models and
-suggests that distance to water influenced permeability negatively. We
-hypothesize that high prey-availability (e.g. impala, kudu) in proximity to
-water explains this result. It is well documented that water is a major driver
-of wildlife and that it determines the distribution of many herbivores,
-especially in dry biomes \citep{Western.1975}. For the Okavango Delta in
-particular, it has been shown that prey abundance strongly correlates with the
-extent of the flood \citep{Bonyongo.2005} and that an increasing concentration
-of ungulates around remaining water-sources during the dry season might attract
-dispersing wild dogs \citep{Ogutu.2014}. Following the same logic, however, apex
-predators (e.g. lions, spotted hyenas) and resident wild dogs may also prefer
-closeness to water \citep{Valeix.2009}, which may occasionally force dispersers
-to move distant to water \citep{Ndaimani.2016} to avoid intraguild competition
+% Recap of what we did
+We combined GPS relocation data of dispersing African wild dog coalitions with
+spatial environmental and anthropogenic covariates to gain information on
+habitat selection or avoidance during dispersal. We then used this knowledge to
+create a permeability surface and assess connectivity by means of identifying
+dispersal corridors across the world’s largest transboundary conservation area,
+the Kavango-Zambezi Transfrontier Conservation Area.
+
+% Finding Regarding Human Influence
+Our results indicate that dispersers avoided anthropogenic influences,
+highlighting that human presence or activities constitute some of the main
+obstacles to connectivity. While \cite{Masenga.2016} found similar avoidance of
+human-dominated landscapes by dispersers in East Africa,
+\cite{DaviesMostert.2012} reported a high willingness of dispersers to cross
+such landscapes in South Africa. Our findings support the notion that dispersers
+forgo human-dominated landscapes and we conjecture that differences to
+observations by Davies-Mostert and colleagues are be due to the unavailability
+of alternative ``natural'' routes in South Africa. That is, dispersers in South
+Africa may be forced to cross human-dominated landscapes despite a strong
+aversion against human presence.
+
+% Finding Regarding Water
+Water was identified as being another major obstacle to dispersal. In earlier
+studies, water has been shown to be almost impermeable for resident packs and it
+is assumed that residents may in fact use it as easy-to-defend territory
+boundaries \citep{Cozzi.2013}. Although dispersers have different needs and
+motivations than residents, they have also been shown to be partially hindered
+by larger water bodies \citep{Abrahms.2017, Cozzi.2020}. Our findings thus align
+with previous research and stress the importance to accurately and dynamically
+represent water bodies, notably for such pulsing ecosystems as represented by
+the Okavango Delta.
+
+% Finding Regarding Distance To Water
+While we found that dispersers avoided moving \textit{through} water, we
+revealed a preference for moving \textit{close to} water. We hypothesize that
+high prey-availability (e.g. impala) in proximity to water explains this result.
+It is well documented that water is a major driver of wildlife and that it
+determines the distribution of many herbivores, especially in dry biomes
+\citep{Western.1975}. For the Okavango Delta in particular, it has been shown
+that prey abundance strongly correlates with the extent of the flood
+\citep{Bonyongo.2005}. An increasing concentration of ungulates around remaining
+water-sources during the dry season \citep{Ogutu.2014} therefore might attract
+dispersing wild dogs. Following the same logic, however, apex predators (e.g.
+lions, spotted hyenas) and resident wild dogs may also prefer proximity to water
+\citep{Valeix.2009} and may occasionally force dispersing wild dogs to move
+distant to water \citep{Ndaimani.2016} to avoid intraguild competition
 \citep{Creel.1996, Mills.1997}. We could not control for the presence or absence
-of apex predators or conspecifics because we lacked corresponding data, but
-these considerations might explain why we did not find a statistically clear
-effect of distance to water.
-
-% Permeability Model: Land Cover, Trees & Shrubs
-In line with existing literature on moving wild dogs \citep{Abrahms.2017}, we
-found that dispersers avoided habitats with high percentages of tree cover,
-whereas they preferred areas with high percentages of shrubs/grassland cover.
-Comparable preferences were reported for residents, which typically reside in
-savanna-like habitats but rarely inhabit areas with dense tree cover
-\citep{Estes.2012}. This may be explained by wild dogs' hunting behavior, where
-open habitats are favored because they offer increased visibility and thus
-facilitate hunting \citep{Estes.1967}. Additionally, we found a strong negative
-correlation between percentage cover by shrubs/grassland and percentage cover by
-bareland (although we did not include bareland in our models). The observed
-preference for shrubs/grassland could therefore also indicate an aversion to
-bareland. Again, we hypothesize that this aversion is driven by low
-prey-availability in non-vegetated habitats \citep{Pettorelli.2011}.
-Unfortunately, we lacked detailed vegetation data and could not distinguish
-vegetation types further.
-
-% Permeability Model: Human Influence
-Finally, considering human influence, we found that dispersers strongly avoided
-human pressures arising from human density, farming, and roads. Previous
-research on dispersing wild dogs was inconclusive about the impact of human
-presence on permeability, as some researchers reported a high willingness of
-dispersers to cross human-dominated landscapes \citep{DaviesMostert.2012},
-whereas others found that humans are avoided almost entirely during dispersal
-\citep{Masenga.2016}. Our results support the notion that dispersers strongly
-avoid human-dominated areas, but we hypothesize that this finding depends on the
-availability of alternative routes. The dipsersers monitored by
-\cite{DaviesMostert.2012} originated from habitats almost exclusively surrounded
-by humans, whereas our dipsersers departed from regions with various
-possibilities to move through natural habitats. A lack of alternative routes
-combined with strong inbreeding avoidance intrinsic in African wild dogs
-\citep{McNutt.1996} might have forced the dispersers reported in
-\cite{DaviesMostert.2012} to cross human-dominated landscapes despite a strong
-aversion against humans. Although we found that our dispersers strictly avoided
-humans, \citeauthor{Cozzi.2019} (in review) reports that human persecution
-remains the main cause of mortality for dispersers, suggesting that avoidance is
-not strong enough to prevent human-induced fatalities. We argue that human
-influence remains a source of reduced landscape permeability and that the
-proceeding transformation of natural landscapes into human altered habitats
-continues to undermine wild dogs' ability to dipserse. Since our human influence
-proxy did not include minor gravel roads, our finding that humans are avoided
-does not contradict the observation that wild dogs preferably move on gravel
-roads during activity \citep{Abrahms.2016}. With four-hourly GPS relocations we
-felt that the temporal resolution was insufficient to depict such fine scale
-effects \citep{Thurfjell.2014}.
-
-\subsection{Permeability Surface}
-% Permeability Surface
-Our prediction of landscape permeability revealed vast regional differences.
-Across the entire study area, northern Botswana stood out as it comprises the
-highest density of highly permeable landscapes. The hospitality of this region
-is caused by a combination of low human influence, low tree cover, high
-shrubs/grassland cover and close distance to water. Thus, although swamps,
-wetlands, and water (e.g. Okavango Delta, Linyanti Swamp) in northern Botswana
-provide little permeability, their surroundings act as strong attractants to
-dispersers. It is therefore not very surprising that the region serves as source
-for the recolonization of surrounding habitats \citep{Cozzi.2013}. Moreover, we
-found much higher permeability inside the KAZA-TFCA than outside its borders,
-which suggests a good fit between the planned protectorates and areas that are
-suitable for dispersal. We made similar observations for protection status,
-where predicted permeabilities inside protectorates are much higher than in
-pastoral regions. Yet, several highly permeable regions remain uncovered by the
-KAZA-TFCA or other protected areas, proposing exciting opportunities for future
-expansions of already existing protectorates. Across countries, we revealed that
-Zambia and Zimbabwe offer the least permeable landscapes. Visual inspection of
-our covariate-layers disclosed that this is mainly owed to severe anthropogenic
-influence in Zambia and Zimbabwe. In contrast, Botswana and Angola host vast
-uninhabited areas, resulting in more permeable landscapes for dispersers.
-However, all of our predictions were based on observations from dispersing wild
-dogs departing from northern Botswana, which might have introduced a bias in
-estimated habitat preferences and resulting predictions.
-
-\subsection{Least-Cost Paths \& Least-Cost Corridors}
-% Least Cost Paths & Least Cost Corridors
-Our LCP- and LCC-analyses disclosed several major movement corridors within the
-extended study area and suggested a decent congruence between predicted
-corridors and the layout of the KAZA-TFCA. That is, only few corridors run
-outside the planned boundaries of the KAZA-TFCA and several currently
-unprotected corridors will become protected with realization of the KAZA-TFCA.
-For example, the high-frequency corridor running from southern Zambia into the
-Kafue NP (\Cref{LeastCost}a \circled{3}) currently contains extended unprotected
-sections that will be transformed into protectorates under the KAZA-TFCA. We
-thus expect that dispersal between such regions will be vastly facilitated
-thanks to the KAZA-TFCA. Our network of LCPs and LCCs furthermore highlighted
-the Chobe-Moremi system in northern Botswana, which is traversed by several
-LCPs and LCCs and seems to hold the potential to serve as central hub through
-which dispersers gain access to more remote regions of the KAZA-TFCA. Still, our
-predictions also indicated that movements between wild dog subpopulations may
-often involve extended detours. This is especially the case for regions in the
-northern sectors of the KAZA-TFCA. For instance, only few LCPs connect the
-Zimbabwean and Zambian sectors, such that moving directly from the Matusadona NP
-to the Kafue NP seems almost impossible and likely requires to make a detour
-through the Chobe-Moremi region. Similarly, few LCPs directly link the Mavinga
-NP in Angola and the Kafue NP in Zambia. Moreover, we discovered potential
-dispersal routes between the Chobe-Moremi ecosystem and the Central Kalahari
-NP that are not yet proclaimed protected areas. Nonetheless, a protected link
-between the Central Kalahari NP and the Chobe-Moremi ecosystem might grant
-southern populations invaluable access to the Chobe-Moremi-Hub and
-subsequently to the entire KAZA-TFCA. In fact, \cite{Elliot.2014} identified a
-similar corridor for dispersing lions, proposing that protecting the area would
-benefit lions and wild dogs alike. The absence of direct LCPs and LCCs between
-some regions of the KAZA-TFCA is likely owed to an impermeable landscape between
-those regions and demonstrates that protection of such areas in their current
-state would be futile. That is, impermeable landscapes first need to be
-transformed into more permeable habitats before protection pays off. Although,
-we noted that connectivity gradually decreases towards the borders of our study
-area, this might partly be owed to boundary-effects \citep{Koen.2010}.
-Interestingly, our network of LCPs and LCCs shows surprising similarities to
-predicted movement corridors of dispersing lions \citep{Elliot.2014}.
-Remarkably, corridors that we identified as high-frequency corridors are also
-visible in the network presented by \cite{Elliot.2014}. This not only reinforces
-confidence in our own predictions but it also suggests potential synergies
-between conservation of the two species.
-
-\subsection{Future Research}
+of apex predators or conspecifics during dispersal as we lacked corresponding
+data, but these considerations might explain why we did not find a statistically
+clear effect of distance to water.
+
+% Finding Regarding Different Permeabilities
+According to our prediction of landscape permeability, permeability of northern
+Botswana and southern Angola are relatively high. High permeability in these
+countries is mainly caused by a combination of low human presence and
+activities, low tree cover, high shrubs/grassland cover and close distance to
+water. Although swamps, wetlands, and permanent water provide little
+permeability themselves, their surroundings act as strong attractants to
+dispersers. The low permeability that characterizes Zambia and Zimbabwe, on the
+other hand, is mainly caused by severe anthropogenic presence and other human
+activities. Despite the fact that the KAZA-TFCA covers many high permeability
+hot-spots, several permeable regions remain uncovered by the KAZA-TFCA, offering
+exciting opportunities for future expansions of already existing protected
+areas.
+
+% Finding Regarding Corridor Overlap with KAZA
+Our results moreover suggest that all major dispersal corridors of African wild
+dogs run within the boundaries of the KAZA-TFCA, demonstrating the potential
+value of such an initiative. Within the KAZA-TFCA, however, some key
+connectivity areas are not formally protected and their protection status may be
+subject to future changes under the influences of an expanding human population,
+farming, and cropping industry.
+
+% Finding Regarding Dispersal Hub
+Importantly, we identified the Okavango-Linyanti region as potential dispersal
+hub through which dispersers gain access to more peripheral regions of the
+KAZA-TFCA. While this might partly be owed to the area's central position within
+the KAZA-TFCA, it is likely also due to its environmental characteristics,
+namely due to human activities and the presence of impermeable water bodies such
+as the Okavango Delta and the Linyanti Swamp, which potentially funnel dispersal
+movements. The region is also instrumental for connecting subpopulations
+remaining in Zambia’s Kafue and Zimbabwe’s Hwange NPs. These subpopulations are
+separated by the Zambezi River, forcing dispersers to detour via northern
+Botswana to reach more peripheral areas. The key role of the Okavango-Linyanti
+region for the overall connectivity of the KATA-TFCA thus calls for actions to
+secure its protection status in the future. While the region is currently a
+Wildlife Management Area devoted to photographic tourism, it has neither the
+status of a National Park nor that of a Game Reserve and so potentially faces
+growing pressure towards exploitation of its natural resources.
+
+% Finding Regarding Non-Protected Corridor Between Zambia and Botswana
+A similar case of non-formally protected but key dispersal landscape is
+represented by the area south of Kafue NP in Zambia, for which a disruption of
+its main and narrow dispersal corridor would result in a considerable isolation
+of its subpopulations, for no other corridor links Kafue NP to the rest of the
+KAZA-TFCA.
+
+% Finding Regarding
+Our results also highlighted a potential southwards corridor between the
+Okavango-Linyanti ecosystem and the Central Kalahari Game Reserve. Some areas
+through which the corridor runs are neither part of KAZA-TFCA nor of any
+protected status. Human presence and activities along the national road that
+longitudinally traverses the corridor as well as the related increased
+mortality caused by direct persecution and collisions \citep{Cozzi.2020} may
+limit realized connectivity. In fact, despite having dispersed south of the
+Okavango Delta, none of the monitored dispersing coalitions crossed the national
+road and reached the Central Kalahari Game Reserve \citep{Cozzi.2020}.
+\citep{Elliot.2014} identified a similar corridor for dispersing lions, which
+shows that protecting the area would benefit lions and wild dogs alike.
+
+Although overall connectivity gradually decreased towards the borders of the
+study area, this might partly be a result of boundary-effects \citep{Koen.2010}.
+
+In general, our network of dispersing corridors shows high similarities to
+predicted movement corridors of dispersing lions \citep{Elliot.2014}, which not
+only reinforces confidence in our own predictions but also suggests potential
+synergies between conservation of the two species.
+
+% The goal of our study was to (1) assess landscape permeability for dispersing
+% wild dogs and to (2) subsequently use this knowledge to examine connectivity
+% across the greater extent of the KAZA-TFCA. To achieve these goals, we
+% parameterized a permeability model in which we used GPS data of 16 dispersing
+% wild dog coalitions to quantify their habitat preferences with respect to land
+% cover, protection status, and human influence. While we found significant
+% impacts of different land covers and human influence, the influence of
+% protection status was not statistically clear. We further applied the
+% parameterized permeability model and predicted a permeability surface for the
+% entire KAZA-TFCA, which depicted the ease at which dispersers crossed a specific
+% landscape. Finally, based on this permeability map, we computed LCPs and LCCs
+% between protectorates and thereby highlighted dispersal corridors of African
+% wild dogs.
+%
+% \subsection{Habitat Selection Model}
+% % Permeability Model: Land Cover, Water
+% Our most parsimonious permeability model revealed that, with respect to land
+% cover, dispersers strictly avoided areas covered by water (including swamps,
+% wetlands). This result corroborates previous research on wild dogs (e.g.
+% \cite{Abrahms.2017}) and is rather unsurprising, as the species is not well
+% adapted for moving through water. For residents, water is almost impermeable and
+% frequently serves as natural boundary to their home-ranges \citep{Cozzi.2013}.
+% It is conceivable that dispersers are more willing than residents to cross
+% waterbodies \citep{Cozzi.2020}, but we did not include residents in our analyses
+% and could not test for such differences. Regardless of that, water appears to
+% pose a serious obstacle for dispersers and is therefore likely to reduce
+% landscape permeability for dispersing wild dogs.
+%
+% % Permeability Model: Land Cover, DistanceToWater
+% While dispersers avoided moving \textit{through} water, they preferred moving
+% \textit{close to} water as indicated by avoidance of high distance to water.
+% While its effect was insignificant, inclusion of the covariate
+% \textit{DistanceToWater} strictly increased the AIC in all of our models and
+% suggests that distance to water influenced permeability negatively. We
+% hypothesize that high prey-availability (e.g. impala, kudu) in proximity to
+% water explains this result. It is well documented that water is a major driver
+% of wildlife and that it determines the distribution of many herbivores,
+% especially in dry biomes \citep{Western.1975}. For the Okavango Delta in
+% particular, it has been shown that prey abundance strongly correlates with the
+% extent of the flood \citep{Bonyongo.2005} and that an increasing concentration
+% of ungulates around remaining water-sources during the dry season might attract
+% dispersing wild dogs \citep{Ogutu.2014}. Following the same logic, however, apex
+% predators (e.g. lions, spotted hyenas) and resident wild dogs may also prefer
+% closeness to water \citep{Valeix.2009}, which may occasionally force dispersers
+% to move distant to water \citep{Ndaimani.2016} to avoid intraguild competition
+% \citep{Creel.1996, Mills.1997}. We could not control for the presence or absence
+% of apex predators or conspecifics because we lacked corresponding data, but
+% these considerations might explain why we did not find a statistically clear
+% effect of distance to water.
+%
+% % Permeability Model: Land Cover, Trees & Shrubs
+% In line with existing literature on moving wild dogs \citep{Abrahms.2017}, we
+% found that dispersers avoided habitats with high percentages of tree cover,
+% whereas they preferred areas with high percentages of shrubs/grassland cover.
+% Comparable preferences were reported for residents, which typically reside in
+% savanna-like habitats but rarely inhabit areas with dense tree cover
+% \citep{Estes.2012}. This may be explained by wild dogs' hunting behavior, where
+% open habitats are favored because they offer increased visibility and thus
+% facilitate hunting \citep{Estes.1967}. Additionally, we found a strong negative
+% correlation between percentage cover by shrubs/grassland and percentage cover by
+% bareland (although we did not include bareland in our models). The observed
+% preference for shrubs/grassland could therefore also indicate an aversion to
+% bareland. Again, we hypothesize that this aversion is driven by low
+% prey-availability in non-vegetated habitats \citep{Pettorelli.2011}.
+% Unfortunately, we lacked detailed vegetation data and could not distinguish
+% vegetation types further.
+%
+% % Permeability Model: Human Influence
+% Finally, considering human influence, we found that dispersers strongly avoided
+% human pressures arising from human density, farming, and roads. Previous
+% research on dispersing wild dogs was inconclusive about the impact of human
+% presence on permeability, as some researchers reported a high willingness of
+% dispersers to cross human-dominated landscapes \citep{DaviesMostert.2012},
+% whereas others found that humans are avoided almost entirely during dispersal
+% \citep{Masenga.2016}. Our results support the notion that dispersers strongly
+% avoid human-dominated areas, but we hypothesize that this finding depends on the
+% availability of alternative routes. The dipsersers monitored by
+% \cite{DaviesMostert.2012} originated from habitats almost exclusively surrounded
+% by humans, whereas our dipsersers departed from regions with various
+% possibilities to move through natural habitats. A lack of alternative routes
+% combined with strong inbreeding avoidance intrinsic in African wild dogs
+% \citep{McNutt.1996} might have forced the dispersers reported in
+% \cite{DaviesMostert.2012} to cross human-dominated landscapes despite a strong
+% aversion against humans. Although we found that our dispersers strictly avoided
+% humans, \citeauthor{Cozzi.2020} (in review) reports that human persecution
+% remains the main cause of mortality for dispersers, suggesting that avoidance is
+% not strong enough to prevent human-induced fatalities. We argue that human
+% influence remains a source of reduced landscape permeability and that the
+% proceeding transformation of natural landscapes into human altered habitats
+% continues to undermine wild dogs' ability to dipserse. Since our human influence
+% proxy did not include minor gravel roads, our finding that humans are avoided
+% does not contradict the observation that wild dogs preferably move on gravel
+% roads during activity \citep{Abrahms.2016}. With four-hourly GPS relocations we
+% felt that the temporal resolution was insufficient to depict such fine scale
+% effects \citep{Thurfjell.2014}.
+%
+% \subsection{Permeability Surface}
+% % Permeability Surface
+% Our prediction of landscape permeability revealed vast regional differences.
+% Across the entire study area, northern Botswana stood out as it comprises the
+% highest density of highly permeable landscapes. The hospitality of this region
+% is caused by a combination of low human influence, low tree cover, high
+% shrubs/grassland cover and close distance to water. Thus, although swamps,
+% wetlands, and water (e.g. Okavango Delta, Linyanti Swamp) in northern Botswana
+% provide little permeability, their surroundings act as strong attractants to
+% dispersers. It is therefore not very surprising that the region serves as source
+% for the recolonization of surrounding habitats \citep{Cozzi.2013}. Moreover, we
+% found much higher permeability inside the KAZA-TFCA than outside its borders,
+% which suggests a good fit between the planned protectorates and areas that are
+% suitable for dispersal. We made similar observations for protection status,
+% where predicted permeabilities inside protectorates are much higher than in
+% pastoral regions. Yet, several highly permeable regions remain uncovered by the
+% KAZA-TFCA or other protected areas, proposing exciting opportunities for future
+% expansions of already existing protectorates. Across countries, we revealed that
+% Zambia and Zimbabwe offer the least permeable landscapes. Visual inspection of
+% our covariate-layers disclosed that this is mainly owed to severe anthropogenic
+% influence in Zambia and Zimbabwe. In contrast, Botswana and Angola host vast
+% uninhabited areas, resulting in more permeable landscapes for dispersers.
+% However, all of our predictions were based on observations from dispersing wild
+% dogs departing from northern Botswana, which might have introduced a bias in
+% estimated habitat preferences and resulting predictions.
+%
+% \subsection{Least-Cost Paths \& Least-Cost Corridors}
+% % Least Cost Paths & Least Cost Corridors
+% Our LCP- and LCC-analyses disclosed several major movement corridors within the
+% extended study area and suggested a decent congruence between predicted
+% corridors and the layout of the KAZA-TFCA. That is, only few corridors run
+% outside the planned boundaries of the KAZA-TFCA and several currently
+% unprotected corridors will become protected with realization of the KAZA-TFCA.
+% For example, the high-frequency corridor running from southern Zambia into the
+% Kafue NP (\Cref{LeastCost}a \circled{3}) currently contains extended unprotected
+% sections that will be transformed into protectorates under the KAZA-TFCA. We
+% thus expect that dispersal between such regions will be vastly facilitated
+% thanks to the KAZA-TFCA. Our network of LCPs and LCCs furthermore highlighted
+% the Chobe-Moremi system in northern Botswana, which is traversed by several
+% LCPs and LCCs and seems to hold the potential to serve as central hub through
+% which dispersers gain access to more remote regions of the KAZA-TFCA. Still, our
+% predictions also indicated that movements between wild dog subpopulations may
+% often involve extended detours. This is especially the case for regions in the
+% northern sectors of the KAZA-TFCA. For instance, only few LCPs connect the
+% Zimbabwean and Zambian sectors, such that moving directly from the Matusadona NP
+% to the Kafue NP seems almost impossible and likely requires to make a detour
+% through the Chobe-Moremi region. Similarly, few LCPs directly link the Mavinga
+% NP in Angola and the Kafue NP in Zambia. Moreover, we discovered potential
+% dispersal routes between the Chobe-Moremi ecosystem and the Central Kalahari
+% NP that are not yet proclaimed protected areas. Nonetheless, a protected link
+% between the Central Kalahari NP and the Chobe-Moremi ecosystem might grant
+% southern populations invaluable access to the Chobe-Moremi-Hub and
+% subsequently to the entire KAZA-TFCA. In fact, \cite{Elliot.2014} identified a
+% similar corridor for dispersing lions, proposing that protecting the area would
+% benefit lions and wild dogs alike. The absence of direct LCPs and LCCs between
+% some regions of the KAZA-TFCA is likely owed to an impermeable landscape between
+% those regions and demonstrates that protection of such areas in their current
+% state would be futile. That is, impermeable landscapes first need to be
+% transformed into more permeable habitats before protection pays off. Although,
+% we noted that connectivity gradually decreases towards the borders of our study
+% area, this might partly be owed to boundary-effects \citep{Koen.2010}.
+% Interestingly, our network of LCPs and LCCs shows surprising similarities to
+% predicted movement corridors of dispersing lions \citep{Elliot.2014}.
+% Remarkably, corridors that we identified as high-frequency corridors are also
+% visible in the network presented by \cite{Elliot.2014}. This not only reinforces
+% confidence in our own predictions but it also suggests potential synergies
+% between conservation of the two species.
+
 % Future Research: Genetic Resistance
@@
 habitats, whereas populations residing at dispersal hotspots, such as the
-Chobe-Linyanti ecosystem, should exhibit a relatively diverse gene-pool.
+Okavango-Linyanti ecosystem, should exhibit a relatively diverse gene-pool.
 
@@
 other species. In particular, we would desire location data of apex predators,
-potential prey, and conspecifics of wild dogs. It is well documented that
-resident wild dogs frequently compete with other carnivores \citep{Creel.1998}
-and we would expect to find competition between apex predators and dispersers
-too. Furthermore, knowledge about the spatial distribution of prey could verify
-our hypothesis that dispersers move along water due prey-availability. Lastly,
-we could not control for the likelihood of encountering conspecifics during
+potential prey, and conspecifics. It is well documented that resident wild dogs
+frequently compete with other carnivores \citep{Creel.1998} and we would expect
+to find competition between dispersing wild dogs and apex predators too.
+Furthermore, knowledge about the spatial distribution of prey could verify our
+hypothesis that dispersers move along water due prey-availability. Lastly, we
+could not control for the likelihood of encountering conspecifics during
 dispersal . Yet, for dispersing meerkats, a species with surprisingly similar
@@
 
-% Global Resistance Mapping
+% Future Research: Global Resistance Mapping
 Besides urging for more and better data, we also want to emphasize that our
@@
 area. All covariates used throughout this study are available on a global scale
-and are likely to be important determinants of landscape permeability for other
-species. Our framework could thus easily be used to assess connectivity for a
-wide range of species, provided suitable relocation data of the species exists.
-Expanding our analyses to other species would surely yield interesting insights
-on the congruence of movement corridors of different species and thereby
-highlight areas that are used by multiple species for dispersal.
-
-% Simulating dispersal
+and they are likely to be important determinants of landscape permeability for
+other species. Our framework could thus easily be applied to assess connectivity
+for a wide range of species, provided suitable relocation data of the species
+exists. Expanding our analyses to other species would surely yield interesting
+insights on the congruence of movement corridors of different species and
+thereby highlight areas that are used by multiple species for dispersal.
+
+% Future Research: Simulating Dispersal
 To this end, we identified movement corridors between pre-defined start and
-end-points. However, pre-defining end-points might not even be necessary, as one
-could easily invert a parameterized iSSF and use it as a mechanistic movement
-model to simulate dispersal. That is, one could define a set of source points at
+end-points, which assumes that dispersers have a specific end-point in mind.
+However, pre-defining end-points might not even be necessary, as one could
+invert the iSSF and use it as a mechanistic movement model to simulate dispersal
+without known end-point. That is, one could define a set of source points at
 locations where wild-dogs are assumed or known to reside and simulate dispersers
@@
 
-\newpage
-\subsection{Conclusion}
-Our results demonstrate that indeed the majority of dispersal corridors are
-covered by the planned protectorates of the KAZA-TFCA, suggesting that it will
-severely contribute to the long-term viability of this species. Nevertheless,
-several dispersal routes do exist outside the KAZA-TFCA and other protectorates
-and suggest opportunities for further expansions in the future. Moreover, our
-connectivity network of least-cost paths and least-cost corridors reveals
-especially the Chobe-Linyanti region as central dispersal hub through which
-dispersers gain access to more remote regions of the KAZA-TFCA. We therefore
-stress the importance of northern Botswana for the conservation of this
-charismatic species. Finally, our investigations showed that human pressure
+% Conclusion
+Our results demonstrate that indeed the majority of dispersal corridors for
+African wild dogs are covered by the boundaries of the KAZA-TFCA, suggesting
+that the KAZA-TFCA will severely contribute to the long-term viability of this
+species. Nevertheless, some unprotected dispersal routes do exist outside the
+KAZA-TFCA and suggest opportunities for further expansions in the future.
+Moreover, our connectivity network of least-cost paths and least-cost corridors
+revealed especially the Okavango-Linyanti region as central dispersal hub
+through which dispersers gain access to more remote regions of the KAZA-TFCA. We
+therefore stress the importance of northern Botswana for the conservation of
+this charismatic species. Finally, our investigations showed that human pressure
 remains a severe impediment to dispersal and strictly reduces landscape
@@
 Gabriele Cozzi for his excellent guidance, all the interesting discussions, his
-constructive criticism, and of course his constant supply of freshly brewed
-coffee. Particular gratitude goes to Dominik Behr who supervised my work on a
-daily basis and greatly contributed to all aspects of the thesis. I genuinely
-enjoyed the frequent meetings and brainstorms and I really feel privileged to
-have had the opportunity of working with such a great scientist. I am also
-greatly indebted to Prof. John Fieberg, who visited our department this summer
-and invested a great amount of time to consult and improve all statistical
-aspects of the thesis, as well as to deepen my personal understanding of
-movement ecology. Furthermore, I thank the Okavango Research Institute, in
-particular Piotr Wolski, for sharing all of their Python-scripts, floodmaps and
-knowledge on developing dynamic floodmaps. Also, I want to thank Stefan Sommer
-for sharing and teaching his hard earned knowledge on scientific writing.
-Additionally, I want to thank all of my colleague master students who have been
-wonderful companions and made the time at university a blast. Likewise, I thank
-my parents for supporting all of my undertakings and for providing first class
-shelter and food. Last, but most definitely not least, I also greatly thank my
-girlfriend, who was always there for me and provided welcome distractions for my
-mind.
+constructive criticism, valuable revisions in the manuscript, and of course his
+constant supply of freshly brewed coffee. Particular gratitude goes to Dominik
+Behr, who supervised my work on a daily basis and greatly contributed to all
+aspects of the thesis. I genuinely enjoyed the frequent meetings and brainstorms
+and I really feel privileged to have had the opportunity of working with such a
+great scientist. I am also greatly indebted to Prof. John Fieberg, who visited
+our department this summer and invested a great amount of time to consult and
+improve all statistical aspects of the thesis, as well as to deepen my personal
+understanding of movement ecology. Furthermore, I thank the Okavango Research
+Institute, in particular Piotr Wolski, for sharing all of their Python scripts,
+floodmaps and knowledge on developing dynamic floodmaps. Also, I want to thank
+Stefan Sommer for sharing and teaching his hard earned knowledge on scientific
+writing. Additionally, I want to thank all of my colleague master students who
+have been wonderful companions and made the time at university a blast.
+Likewise, I thank my parents for supporting all of my undertakings and for
+providing first class shelter and food. Last, but most definitely not least, I
+also greatly thank my girlfriend, who was always there for me and provided
+welcome distractions for my mind.
 
@@
 \section{Appendix}
-\subsection{Typical Flood Pulse}
-\label{Appendix:FloodPulse}
+\subsection{Floodmapping Algorithm}
+\label{Appendix:FloodmappingAlgorithm}
+To implement the floodmapping algorithm, we defined two sets of polygons located
+in the region of the Okavango Delta (see \Cref{Floodmapping}). The first set
+consisted of areas known to be permanent dryland, whereas the second set
+consisted of permanent waters. Since we were unable to retrieve the original
+polygons used in \cite{Wolski.2017}, we geo-referenced one of their
+illustrations and traced the polygons there. After recreation of the polygons,
+we used the R-package \textit{getSpatialData} \citep{Schwalb.2018} to download
+and pre-process all relatively cloud-free MODIS Terra images (MCD43A4,
+\cite{Schaaf.2015}) available for the period of our dispersal events. Assessment
+of cloud cover was based on visual inspection of MODIS images on ORI's website
+\citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful for dynamic
+floodmapping because it is updated with new composite images every 8th day
+\citep{Wolski.2017}. After download, we classified each MODIS image into a
+binary map of water (flood) and dryland using a threshold that was identified as
+follows. First, we extracted all reflectance values of MODIS Terra Band 7 within
+the water- and dryland-polygons. Second, we computed histograms of
+water-reflectances and dryland-reflectances and empirically verified that
+reflectances of the two groups were sufficiently distinct. More specifically, we
+checked if superimposing the two histograms resulted in a bimodal plot. This was
+said to be achieved if the 99\textsuperscript{th} percentile of
+water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
+of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
+dryland}\)). Third, if peaks were sufficiently different, we calculated a
+threshold (\(t\)) using \Cref{EQ1}:
+
+\begin{equation}
+\label{EQ1}
+t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
+\widetilde{p}_{water})
+\end{equation}
+
+\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
+the median reflectances of water and dryland, respectively. We then classified
+all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
+all pixels with a value smaller than \(t\) as water.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
+    \caption{Figures describing the floodmapping algorithm. (a) The colored
+    polygons indicate permanent waters and permanent dryland. Below these
+    polygons we extracted reflectance values of MODIS Terra Band 7 and used
+    their repsective medians to calculate a classification threshold. (b)
+    Example of a classified MODIS Terra Band 7 image after application of the
+    threshold.}
+    \label{Floodmapping}
+  \end{center}
+\end{figure}
+
+% Dynamic Wetmask
+Importantly, bimodality was not always achieved and in some cases no floodmap
+could be calculated. In fact, it appears that non-bimodality caused the ORI
+algorithm to fail since the end of 2018, such that no floodmaps have been
+generated since then (ORI, personal comm.). We hypothesized that this was caused
+by the application of static water-polygons that did not cover permanent waters
+correctly anymore. We therefore revised the algorithm and allowed for a more
+dynamic polygonization of water. That is, for each MODIS image we calculated new
+water-polygons comprising areas that were covered by the flood in 99\% of the
+floodmaps from the previous five years. All of the necessary floodmaps from
+previous years were kindly provided to us by ORI. Using this slightly amended
+approach, we were able to address some of the bimodality issues and to classify
+several additional floodmaps for the period of our study. Because MODIS Terra
+Band 7 had a resolution of 500m x 500m, we interpolated all maps to 250m x 250m.
+
+% Validation
+To validate and compare the performance of our own algorithm to the original
+ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
+images. To make sure that months were equally represented in the sampled dates,
+we employed stratified sampling based on months (regardless of the year) and
+randomly sampled four maps for each month. For the sampled dates we downloaded
+and classified MODIS Terra Band 7 images and compared our classified images to
+those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
+created a difference map indicating false positives and false negatives and
+computed the relative number of wrongly classified pixels. We achieved an
+overall accuracy of 97\%, which presumably is an underestimate of the true
+performance, as we introduced some errors when resampling the ORI-maps to our
+reference raster.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
+    \caption{Validation procedure of our floodmapping algorithm. (a) Classified
+    image that was provided to us by ORI. (b) Image for the same date but now
+    classified using our own algorithm. (c) Difference image indicating false
+    positives and false negatives in our own classification.}
+    \label{FloodmappingValidation}
+  \end{center}
+\end{figure}
+
+% Static Watermap
+\noindent While water in the Okavango Delta was mapped dynamically, we used a
+static watermap for its surroundings. This static map was based on Globeland's
+land cover dataset \citep{Chen.2015}, which contains ten land cover categories.
+From the original layer we only retained the categories \textit{wetland} and
+\textit{waterbodies} and collectively reclassified them to \textit{water}.
+Globeland had an original resolution of 30m x 30m, so we coarsened the layer to
+250m x 250m using the mode of each 250m x 250m cell. Lastly, we improved river
+representation using the MERIT Hydro dataset \citep{Yamazaki.2019}, from which
+we added all rivers with a width of over 10m to our watermaps. The entire
+process resulted in several hundred watermaps, of which 72 represented the
+closest watermap to one of our GPS relocations.
+
 \begin{figure}[h]
@@
     \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
-    \caption{Sequence of figures showing a typical flood pulse throughout the
+    \caption{Sequence of floodmaps showing a typical flood pulse throughout the
     year. The flood arrives from the north-western corner (so called
@@
   weight of one. The model in the last row did not converge and no AIC value
-  could be obtained. (*) represents the fixed covariates \(cos(ta\_)\) and
-  \(log(sl\_)\).}
+  could be obtained. (*) represents the fixed covariates \(cos(ta)\) and
+  \(log(sl)\).}
   \label{ModelAICs}
@@
 in addition included interaction terms between movement metrics (i.e.
-\(cos(ta\_)\), \(log(sl\_\))) and environmental covariates (e.g. water). For
+\(cos(ta)\), \(log(sl\))) and environmental covariates (e.g. water). For
 instance, for the environmental covariate \textit{Water}, we created the
-interaction terms \textit{cos(ta\_):Water} and \textit{log(sl\_):Water}. These
+interaction terms \textit{cos(ta):Water} and \textit{log(sl):Water}. These
 interactions served to learn how dispersers moved differently, depending on
@@
 and low turning angles in proximity to water, but clearly preferred low turning
-angles (i.e. a high \(cos(ta\_)\)) when distance to water was high (see
+angles (i.e. a high \(cos(ta)\)) when distance to water was high (see
 \Cref{MovementModel}b2). Finally, dispersers preferably realized steps with low
@@
   e^{-0.5} \Delta \text{AIC}_i} \). (*) represents all covariates from the most
-  parsimonious permeability model, i.e. \(cos(ta\_)\), \(log(sl\_\)),
+  parsimonious permeability model, i.e. \(cos(ta)\), \(log(sl\)),
   \textit{Water}, \textit{DistanceToWater}, \textit{Shrubs/Grassland},
@@
         (*),
-        log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000 &
+        log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000 &
           89814.19 &
@@
                 -44889.10 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):Trees &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):Trees &
           89814.37 &
@@
               -44889.18 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater &
+        (*), log(sl):Water, cos(ta):DistanceToWater &
           89814.46 &
@@
                 -44890.23 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Trees &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Trees &
           89814.87 &
@@
                 -44888.44 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Shrubs &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Shrubs &
           89815.32 &
@@
                 -44888.66 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):Shrubs &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):Shrubs &
           89815.34 &
@@
                 -44889.67 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):Shrubs &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):Shrubs &
           89815.40 &
@@
                 -44888.70 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, log(sl\_):Shrubs &
+        (*), log(sl):Water, cos(ta):DistanceToWater, log(sl):Shrubs &
           89815.67 &
@@
                 -44889.84 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):HumansBuff5000 &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):HumansBuff5000 &
           89815.94 &
@@
                 -44888.97 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Trees, log(sl\_):Shrubs &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Trees, log(sl):Shrubs &
           89816.08 &
@@
                 -44888.04 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):DistanceToWater &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):DistanceToWater &
           89816.14 &
@@
                 -44889.07 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, cos(ta\_):Water &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Water &
           89816.16 &
@@
                 -44889.08 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, log(sl\_):HumansBuff5000 &
+        (*), log(sl):Water, cos(ta):DistanceToWater, log(sl):HumansBuff5000 &
           89816.18 &
@@
                 -44890.09 \\
-        (*), log(sl\_):Water, cos(ta\_):DistanceToWater, cos(ta\_):HumansBuff5000, log(sl\_):Trees &
+        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):Trees &
           89816.19 &
@@
       \midrule
-      cos(ta\_) & 0.149 & 0.045 & 0.001 \\
-      log(sl\_) & 0.053 & 0.019 & 0.005 \\
+      cos(ta) & 0.149 & 0.045 & 0.001 \\
+      log(sl) & 0.053 & 0.019 & 0.005 \\
       Water & -0.265 & 0.090 & 0.003 \\
@@
       Trees & -0.177 & 0.072 & 0.013 \\
-      log(sl\_):Water & -0.040 & 0.008 & 0.000 \\
-      cos(ta\_):DistanceToWater & 0.115 & 0.038 & 0.003 \\
-      cos(ta\_):HumansBuff5000 & -0.047 & 0.031 & 0.134 \\
+      log(sl):Water & -0.040 & 0.008 & 0.000 \\
+      cos(ta):DistanceToWater & 0.115 & 0.038 & 0.003 \\
+      cos(ta):HumansBuff5000 & -0.047 & 0.031 & 0.134 \\
        \bottomrule
@@
         \midrule
-        cos(ta\_) & 0.148 & 0.066 & 0.025 \\
-        log(sl\_) & 0.055 & 0.019 & 0.003 \\
+        cos(ta) & 0.148 & 0.066 & 0.025 \\
+        log(sl) & 0.055 & 0.019 & 0.003 \\
         Water & -0.472 & 0.113 & 0.000 \\

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-02-04.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-01-12.tex
@@
 % Title of the document
-{\textsc{ \LARGE Ranging Wide: Movement Corridors of Dispersing African Wild
-Dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation
+{\textsc{ \LARGE On the Move: Dispersal Corridors of African Wild Dogs
+(\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation
 Area}}\\[0.4cm]
@@
 currently only survives in a few scattered subpopulations in Sub-Saharan Africa.
-Long-term viability of the species thus requires the identification and
-preservation of key dispersal corridors that connect such subpopulations. We
-collected GPS data of 16 dispersing coalitions originating from a free-ranging
-population in northern Botswana and used integrated step selection analysis to
-create a permeability surface spanning over the entire Kavango-Zambezi
-Transfrontier Conservation Area (KAZA-TFCA), the largest transboundary
-conservation area on earth. Based on our permeability surface, we calculated
-least-cost paths and least-cost corridors between selected locations within the
-KAZA-TFCA. Our results show that dispersal corridors typically run through open
-shrubs/grassland and in the vicinity of water. Corridors rarely cross
-waterbodies, areas with high human presence, or areas densely vegetated by
-trees. Several dispersal corridors meet in the Okavango-Linyanty ecosystem,
-which highlights the importance of northern Botswana for the conservation of
-this charismatic species.
+Long-term viability of the species thus requires to identify and preserve key
+dispersal corridors that connect such subpopulations. We collected GPS data of
+16 dispersing coalitions originating from a free-ranging population in northern
+Botswana and used integrated step selection analysis to create a permeability
+surface spanning over the entire Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA), the largest transboundary conservation area on earth. Based on our
+permeability surface, we calculated least-cost paths and least-cost corridors
+between selected locations within the KAZA-TFCA. Our results show that dispersal
+corridors typically run through open shrubs/grassland and in the vicinity of
+water. Corridors rarely cross water bodies, areas with high human presence, or
+areas densely vegetated by trees. Several dispersal corridors meet in the
+Okavango-Linyanty ecosystem, which highlights the importance of northern
+Botswana for the conservation of this charismatic species.
 \end{abstract}
@@
 % Importance of Connectivity
-Successful conservation of fragmented and isolated subpopulations requires
-maintenance and enhancement of connectivity between habitats occupied by these
+Successful conservation of fragmented and isolated subpopulations requires to
+maintain and enhance connectivity between habitats occupied by these
 subpopulations \citep{Hanski.1998}. In this context, connectivity describes the
-degree to which the landscape matrix facilitates or impedes movement of
-individuals and thereby genes of the focal species between habitat patches
-\citep{Taylor.1993, Clobert.2012}. Thus, sufficient connectivity facilitates
-dispersal, which in turn promotes genetic exchange, reinforcement and rescuing
-of small populations, and colonization or recolonization of unoccupied habitats
-\citep{Brown.1977, Hanski.1998, MacArthur.2001, Frankham.2002, Leigh.2012}. All
-these processes crucially contribute to long-term population viability, which
-emphasizes the importance of dispersal and connectivity. However, increasing
-anthropogenic pressure and ongoing habitat loss and fragmentation reduce
-connectivity on a worldwide scale \citep{Fahrig.2003, Barnosky.2012}. As a
-result, management actions for the conservation of endangered species are
-increasingly aimed towards preserving and restoring connectivity through the
-identification and preservation of key dispersal corridors that link isolated
-habitats \citep{Fahrig.2003, Heller.2009}. An unprecedented initiative to
-restore connectivity on a large scale is represented by the creation of the
-Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), a protected area
-which traverses five African countries and spans over 520,000
-km\textsuperscript{2}. While such initiatives are expected to bring benefits to
-wide-ranging species such as elephants (\textit{Loxodonta africana}), African
-wild dogs (\textit{Lycaon pictus}), and lions (\textit{Panthera leo}), the
-extent to which they are effective for targeted species and whether boundaries
-reflect natural dispersal corridors is rarely assessed.
+degree to which the landscape matrix facilitates or impedes movements of
+individuals of the focal species between habitat patches \citep{Taylor.1993,
+Clobert.2012}. Thus, sufficient connectivity facilitates dispersal, which in
+turn promotes genetic exchange, reinforcement and rescuing of small populations,
+and colonization or recolonization of unoccupied habitats \citep{Brown.1977,
+Hanski.1998, MacArthur.2001, Frankham.2002, Leigh.2012}. These processes
+significantly contribute to long-term population viability, which emphasizes the
+importance of connectivity. However, increasing anthropogenic pressure and
+ongoing habitat loss and fragmentation reduce connectivity on a worldwide scale
+\citep{Fahrig.2003, Barnosky.2012}. As a result, management actions for the
+conservation of endangered species increasingly focus on preserving and
+restoring connectivity through the identification and preservation of key
+dispersal corridors that link isolated habitats \citep{Fahrig.2003,
+Heller.2009}. An unprecedented initiative to restore connectivity on a large
+scale is represented by the creation of the Kavango-Zambezi Transfrontier
+Conservation Area (KAZA-TFCA), a conservation area that traverses five African
+countries and spans over 520,000 km\textsuperscript{2}. While such initiatives
+are expected to bring benefits to wide-ranging species such as elephants
+(\textit{Loxodonta africana}), African wild dogs (\textit{Lycaon pictus}), and
+lions (\textit{Panthera leo}), the extent to which they are effective for
+targeted species and whether boundaries reflect natural dispersal corridors is
+rarely assessed.
 
@@
 In recent years, there has been a growing body of literature that uses animal
-movement data to identify dispersal corridors and to assess connectivity on
+relocation data to identify dispersal corridors and to assess connectivity on
 large scales \citep{Chetkiewicz.2006}. Identification of potential dispersal
@@
 functions} (PSF, \cite{Cushman.2010}). All methods have in common that they
-infer habitat preferences of the focal species by comparing covariates (e.g.
-environmental and anthropogenic) at visited locations to the same covariates at
-randomly selected locations that are deemed available but presumably unused by
-the animal. Based on inferred habitat preferences the methods further assume
-that habitats preferred by the focal species provide high permeability, whereas
-avoided habitats provide low permeability. Regardless of the method chosen (i.e.
-RSF, SSF, or PSF), all approaches require adequate GPS relocation data that is
-representative for the process being studied. More specifically, it has been
-shown that GPS relocation data collected on dispersing individuals is better
-suited for detecting dispersal corridors than GPS relocation data collected on
-resident individuals \citep{Elliot.2014}. Few studies have, however, partitioned
-their GPS relocation data according to behavioral mode such as \textit{resident}
-vs. \textit{dispersing} or \textit{active} vs. \textit{inactive} periods of
-movement \citep{Wilson.2012, Vasudev.2015}. As a consequence, most permeability
-surfaces upon which dispersal corridors are identified are created using GPS
-relocation data that was collected on resident individuals. This introduces
-severe biases and substantially reduces the power to reveal meaningful
-corridors, for dispersing individuals have different needs and motivations
-compared to resident individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}.
-For instance, \cite{Elliot.2014} demonstrated marked differences in predicted
-permeability surfaces for dispersing lions in the KAZA-TFCA depending on whether
-permeability was assessed using data collected during residence or dispersal.
-\cite{Jackson.2016} made similar observations for African wild dogs and revealed
-higher permeability for dispersers than for residents. Moreover,
-\cite{Abrahms.2017} found a significantly greater overlap between predicted and
-observed corridors for three dispersing wild dogs when predictions were based on
-information collected only during wild dog activity, rather than across the
-entire day.
+infer habitat preferences of the focal species by comparing spatial covariates
+(e.g. environmental and anthropogenic) at visited locations to the same spatial
+covariates at randomly selected locations that are deemed available but
+presumably unused by the animal. Based on inferred habitat preferences the
+methods further assume that habitats preferred by the focal species provide high
+permeability, whereas avoided habitats provide low permeability. Regardless of
+the method chosen (i.e. RSF, SSF, or PSF), all approaches require adequate GPS
+relocation data that is representative of the process being studied. More
+specifically, it has been shown that GPS relocation data collected on dispersing
+individuals are better suited for detecting dispersal corridors than GPS
+relocation data collected on resident individuals \citep{Elliot.2014}. Few
+studies have, however, partitioned their GPS relocation data according to
+behavioral modes such as \textit{resident} vs. \textit{dispersing} or
+\textit{active} vs. \textit{inactive} periods of movement \citep{Wilson.2012,
+Vasudev.2015}. As a consequence, most permeability surfaces upon which dispersal
+corridors are identified are created using GPS relocation data that was
+collected on resident individuals. This introduces severe biases and
+substantially reduces the power to reveal meaningful corridors, for dispersing
+individuals have different needs and motivations compared to resident
+individuals (\citeauthor{Killeen.2014}, 2014; \citeauthor{Elliot.2014}, 2014;
+\citeauthor{Cozzi.2020}, unpublished). For instance, \cite{Elliot.2014}
+demonstrated marked differences in predicted permeability surfaces for
+dispersing lions in the KAZA-TFCA depending on whether permeability was assessed
+using data collected during residence or dispersal. \cite{Jackson.2016} made
+similar observations for African wild dogs and revealed higher permeability for
+dispersers than for residents. Moreover, \cite{Abrahms.2017} found a
+significantly greater overlap between predicted and observed corridors for three
+dispersing wild dogs when predictions were based on information collected only
+during wild dog activity, rather than across the entire day.
 
@@
 \citep{Woodroffe.2012}. Within these subpopulations, wild dogs form cooperative
-breeding groups of up to thirty individuals in which a single dominant pair
+breeding groups of up to thirty individuals among which a single dominant pair
 monopolizes the majority of reproduction \citep{Frame.1979, Fuller.1992,
 Creel.2002}, while subordinates of both sexes help to feed newborn pups
-\citep{Malcolm.1982}. After reaching 15 months of age, male and female offspring
+\citep{Malcolm.1982}. After reaching 13 months of age, male and female offspring
 disperse from their natal pack in same-sex coalitions and search for unrelated
@@
 \citeauthor{Behr.2019}, unpublished). During such dispersal events, wild dogs
-are able to cover several hundred kilometers, sometimes crossing risky areas
-that are strongly influenced by humans \citep{DaviesMostert.2012, Masenga.2016,
-Cozzi.2020}. Despite the importance of dispersal for the remaining wild dog
-subpopulations, little is known about the corridors of dispersing individuals
-and connectivity between subpopulations in the KAZA-TFCA in general.
+can cover several hundred kilometers, sometimes crossing risky areas that are
+strongly influenced by humans (\citeauthor{DaviesMostert.2012}, 2012;
+\citeauthor{Masenga.2016}, 2016; \citeauthor{Cozzi.2020}, unpublished). Despite
+the importance of dispersal for the remaining wild dog subpopulations, little is
+known about the corridors of dispersing individuals and connectivity between
+subpopulations in the KAZA-TFCA in general.
 
@@
 that wild dog movements during residence mainly depended on land cover,
-protection status and anthropogenic features. Similar observations on resident
+protection status, and anthropogenic features. Similar observations on resident
 packs were made by \cite{Pretorius.2019}, who's results indicate that pack
 presence was contingent on factors such as agricultural activity, road presence,
-and protection status of the area. With respect to dispersing individuals
+and protection status of the area. Concerning dispersing individuals,
 \cite{Whittington.2011} demonstrated that elevation, land cover, road density,
@@
 Africa. Also based on data from dispersing wild dogs, \cite{Jackson.2016} showed
-that land cover, human density and road density significantly influenced the
+that land cover, human density, and road density significantly influenced the
 probability of occurrence of dispersing wild dogs. Finally, \cite{Masenga.2016}
 reported a low willingness of dispersers to cross human-dominated landscapes.
-Large scale spatial information on the above mentioned factors can be obtained
-either through freely available online services or through remote sensed
+Large scale spatial information on the above-mentioned factors can be obtained
+either through freely available online services or through remotely sensed
 satellite imagery.
@@
 We used GPS relocation data collected on 16 wild dogs in as many dispersing
-coalitions from a free ranging population in northern Botswana to estimate the
+coalitions from a free-ranging population in northern Botswana to estimate the
 degree of selection or avoidance for environmental and anthropogenic landscape
@@
 allowed us to highlight key dispersal corridors within and outside the
-boundaries of KAZA-TFCA and concluded with the identification of a ``dispersal
-hub'' and of dispersal corridors that run outside formally protected areas.
+boundaries of the KAZA-TFCA and to identify a potential ``dispersal hub'' as
+well as other dispersal corridors that run outside formally protected areas.
 
@@
 originated from, and an \textit{extended study area}, which was represented by a
-bounding box encompassing the KAZA-TFCA \citep{Cozzi.2020} and for which we
-calculated connectivity.
+bounding box encompassing the KAZA-TFCA (\citeauthor{Cozzi.2020}, unpublished)
+and for which we calculated connectivity.
 
@@
 \Cref{Appendix:FloodmappingAlgorithm} for an illustration of a typical flood
-pulse). The flood thereby drastically affects surrounding landscapes, so that
-during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a
-patchy conglomerate of swamps, open water, and islands, whereas it shrinks
-substantially (ca. 5'000 km\textsuperscript{2}) during dry periods
-\citep{Gumbricht.2004, Wolski.2017}. Most of the Okavango Delta's unique basin
-is under protection, such that human influence in our core study area was mainly
-restricted to the surroundings of Maun and the smaller villages of Khwai,
+pulse). The flood drastically affects surrounding landscapes, so that during
+maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a patchy
+conglomerate of swamps, open water, and islands, whereas these structures
+disappear when the flood retracts (ca. 5'000 km\textsuperscript{2}) during dry
+periods \citep{Gumbricht.2004, Wolski.2017}. Most of the Okavango Delta's unique
+basin is under protection, such that human influence in the core study area was
+mainly restricted to the surroundings of Maun and the smaller villages of Khwai,
 Mababe, and Sankuyo.
@@
 Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520,000
-km\textsuperscript{2} it contains several already-existing national parks (NPs)
+km\textsuperscript{2}, it contains several already-existing national parks (NPs)
 and other protected areas. It constitutes the earth's largest transboundary
@@
         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
+        \includegraphics[width=0.9\textwidth]{99_StudyArea2.pdf}
         };
@@
 % Collection
-We collected GPS relocation data of 16 wild dog coalitions (i.e. 7 female and 9
-male coalitions) that dispersed from the core study area between 2011 and 2019.
-Potential dispersing individuals were immobilized according to protocols
-described in \cite{Osofsky.1996} and fitted with GPS/Satellite radio collars
-(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while they were still
-with their natal pack. All required procedures were undertaken and supervised by
-a Botswana-registered wildlife veterinarian. Immobilized individuals were
-monitored and all were observed to successfully re-join their packs within one
-hour after immobilization. GPS collars were programmed to record GPS relocation
-data every 4 hours during dispersal and to regularly transmit them via Iridium
-satellite system to a base station. The system thus allowed remote-tracking of
-collared individuals, even where field conditions were prohibitive or dispersal
-trajectories crossed international borders.
+We collected GPS relocation data of 16 wild dog coalitions (7 female and 9
+male coalitions) that dispersed from the core study area between 2011 and 2019
+(see also \citeauthor{Abrahms.2017}, 2017 and \citeauthor{Cozzi.2020},
+unpublished). Potential dispersing individuals were immobilized according to
+protocols described in \cite{Osofsky.1996} and fitted with GPS/Satellite radio
+collars (\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while they were
+still with their natal pack. All required procedures were undertaken and
+supervised by a Botswana-registered wildlife veterinarian. Immobilized
+individuals were monitored and all were observed to successfully re-join their
+packs within one hour after immobilization. GPS collars were programmed to
+record GPS relocation data every 4 hours during dispersal and to regularly
+transmit them via Iridium satellite system to a base station. The system thus
+allowed remote-tracking of collared individuals, even where field conditions
+were prohibitive or dispersal trajectories crossed international borders.
 
@@
 discarded any GPS data collected while individuals were still with their natal
-packs \citep{Cozzi.2020}. Dispersal was deemed to have started when a coalition
-had left its natal home range, and continued until it exhibited clear signs of
-resident behavior. Some coalitions undertook exploratory movements prior to
-proper dispersal. Such movements have been observed in other species and are
-reported to be very similar to proper dispersal \citep{Killeen.2014}. We
-therefore classified exploratory movements as dispersal too.
+packs (\citeauthor{Cozzi.2020}, unpublished). Dispersal was deemed to have
+started when a coalition had left its natal home range, and continued until it
+exhibited clear signs of resident behavior. Some coalitions undertook
+exploratory movements prior to proper dispersal. Such movements have been
+observed in other species and are reported to be very similar to proper
+dispersal \citep{Killeen.2014}. We therefore classified exploratory movements as
+dispersal too.
 
@@
 female dispersing coalitions, for previous research found little differences
-between sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}. On average,
-dispersal coalitions included three individuals (min = 1, max = 7), dispersed
-for 41 days (min = 7, max = 133), and covered a mean cumulative distance of 370
-km (min = 105, max = 610).
+between sexes during dispersal (\citeauthor{Woodroffe.2019}, 2019;
+\citeauthor{Cozzi.2020}, unpublished). On average, dispersal coalitions included
+three individuals (min = 1, max = 7), dispersed for 41 days (min = 7, max =
+133), and covered a mean cumulative distance of 370 km (min = 105, max = 610).
 
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_Trajectories.pdf}
+    \includegraphics[width = \textwidth]{99_Trajectories2.pdf}
     \caption{Illustration of the dispersal trajectories that we recorded. Each
     color represents a different dispersal coalition. All coalitions departed
-    from our core study area, which is indicated by the white square. The
+    from the core study area, which is indicated by the white rectangle. The
     coalition dispersing towards the far east of the map covered over 360 km in
@@
 dispersing wild dogs (see \Cref{Covariates}). Covariates fell into one of the
-three categories \textit{land cover}, \textit{protection status}, and
-\textit{human influence}. Note that we did not include elevation as covariate
-due to the absence of noteworthy elevational gradients in our study area. For
-each covariate we obtained spatial raster layers from freely available online
-services or from remote sensed satellite imagery. To ensure a consistent
-resolution (i.e. cell-size or grain) across covariates, all layers were
-coarsened or interpolated to match a resolution of 250m x 250m. Processing and
-manipulation of data, as well as all spatial and statistical analyses were
-performed in R, version 3.6.1 \citep{R.2019}.
+three categories of \textit{land cover}, \textit{protection status}, and
+\textit{human influence}. We did not include elevation as a covariate due to the
+absence of noteworthy elevational gradients in our study area. For each selected
+covariate, we obtained spatial raster layers from freely available online
+services or remotely sensed satellite images. To ensure a consistent resolution
+(i.e. cell-size or grain) across covariates, all layers were coarsened or
+interpolated to match a resolution of 250m x 250m. Processing and manipulation
+of data, as well as all spatial and statistical analyses, were performed in R,
+version 3.6.1 \citep{R.2019}.
 
@@
     covariates were prepared for the extended study area, we present only an
-    extract to maximize visibility. The white square in each plot depicts our
+    extract to maximize visibility. The white rectangle in each plot depicts our
     core study area. (a) Averaged layer of all binary maps of water and dryland
@@
     trees. (c) Percentage cover of shrubs/grassland. Note that anything not
-    covered by trees or shrubs/grassland was deemed to be bareland. (d)
-    Protection status according to the Peace Parks Foundation. (e) Human
-    influence proxy, including human density, farms, and roads. (f) Distance to
-    nearest road (white lines depict actual roads).}
+    covered by trees or shrubs/grassland was deemed to be bare land. (d)
+    Protection status of the area. (e) Human influence proxy, taking into
+    account human density, farms, and roads. (f) Distance to nearest road (white
+    lines depict actual roads).}
     \label{Covariates}
@@
 and dryland, and three continuous layers showing the percentage cover by
-different vegetation types, together adding up to 100\%.
+different vegetation types, combined adding up to 100\%.
 
@@
 map was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
-only retained the categories \textit{wetland} and \textit{waterbodies} and
+only retained the categories \textit{wetland} and \textit{water bodies} and
 collectively reclassified them to \textit{water}. Globeland had an original
@@
 MERIT Hydro dataset \citep{Yamazaki.2019}, from which we added all rivers with a
-width of over 10m to our watermaps.
+width of over 10m to our floodmaps.
 
 % Vegetation Layers
-The three vegetation layers that we used in addition to the water/dryland maps
-were derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
+The three vegetation layers that we used in addition to the floodmaps were
+derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
 \cite{Dimiceli.2015}). Combined, the three layers added up to 100\% and depicted
 the percentage cover of tree-vegetation (henceforth trees), non-tree-vegetation
-(henceforth shrubs/grassland), and non-vegetated (henceforth bareland). To
+(henceforth shrubs/grassland), and non-vegetated (henceforth bare land). To
 ensure that areas covered by the flood would get 0\% vegetation cover (i.e.
-100\% bareland cover), we used our floodmap that aligned with the creation date
+100\% bare land cover), we used our floodmap that aligned with the creation date
 of the layers (30 March 2017) and manually defined pixels covered by the flood
-as being 100\% bareland. The MODIS vegetation layers had a resolution of 250m x
+as being 100\% bare land. The MODIS vegetation layers had a resolution of 250m x
 250m, so we did not need to coarsen or interpolate them.
@@
 (1) We obtained spatial human density estimates through Facebook's 30m x 30m
-high resolution population density dataset \citep{Facebook.2019}. The dataset
+high-resolution population density dataset \citep{Facebook.2019}. The dataset
 contained some human density that arose from lodges and other touristic
@@
 to investigate dispersers' selection or avoidance for the above mentioned
-spatial covariates. In the iSSF framework, covariates along realized steps are
-contrasted with covariates along alternative random steps that the animal could
-have taken, but decided not to. A step is here defined as the connecting line
-between two consecutive GPS relocations \citep{Turchin.1998}. In contrast to
-regular SSFs, iSSFs require to include movement metrics as covariates in the
-corresponding conditional logistic regression model. Their inclusion in turn
-allows simultaneous inference on habitat and movement preferences, as well as to
-reduce potential biases in estimated habitat preferences \citep{Forester.2009,
-Warton.2013, Avgar.2016}.
-
-In order to conduct an iSSF analysis, we followed recommendations described in
-Appendix S1 in \cite{Avgar.2016}. We prepared our GPS relocation data for
+spatial covariates. In the iSSF framework, covariates experienced along realized
+steps are contrasted with covariates experienced along alternative random steps
+that the animal could have taken, but decided not to. A step is here defined as
+the connecting line between two consecutive GPS relocations
+\citep{Turchin.1998}. In contrast to regular SSFs, iSSFs require to include
+movement metrics as covariates in the corresponding conditional logistic
+regression model. Their inclusion, in turn, allows simultaneous inference on
+habitat and movement preferences, as well as to reduce potential biases in
+estimated habitat preferences \citep{Forester.2009, Warton.2013, Avgar.2016}.
+
+To conduct an iSSF analysis, we followed recommendations described in
+Appendix S1 of \cite{Avgar.2016}. We prepared our GPS relocation data for
 iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
@@
 (shape = 0.3677, scale = 6,302)). Together, a realized and its 24 associated
-random steps formed a stratum and received a unique identifier.
+random steps formed a stratum of 25 steps that received a unique identifier.
 
 We extracted spatial covariates along realized and random steps as listed in
-\Cref{ExtractedCovars}. For continuous covariates we calculated the average
-value along the step, for binary covariates the percentage cover along the step.
-Since our floodmaps were dynamic, we extracted values from the floodmap that was
-closest in date to the actual step. We further derived a binary variable
-indicating whether or not a step crossed a road and we identified the average
-distance of each step to the nearest source of water or to the nearest road. We
-square-rooted the values for \textit{DistanceToRoads} and
-\textit{DistanceToWater} to render a decreasing marginal impact of distance. We
-standardized extracted covariates to a mean of zero and a standard deviation of
-one using a z-score transformation. Furthermore, we screened covariates for
-correlation using Pearson's Correlation Coefficient. None of the covariates were
-overly correlated (\(|r| > 0.6\), \cite{Latham.2011}) and we retained all of
-them for modeling. To run a proper iSSF analysis, we also included two movement
-metrics, namely the cosine of the turning angle (\(cos(ta)\)) and the logarithm
-of the step length (\(log(sl)\)), as covariates in our models
-\citep{Avgar.2016}. The movement metric \(cos(ta\)) served to describe the
-directionality of a step, as it transformed the circular measure of (\(-\pi,
-+\pi\)) into a linear measure (-1, 1), where positive values indicated forward
-movements, whereas negative values indicated backwards movements
-\citep{Turchin.1998}. The movement metric \(log(sl\)), on the other hand, was
-included as an indicator of the preferred step length. Since steps were
-regularly spaced by four hours, \(log(sl\)) can also be interpreted as movement
-rate.
+\Cref{ExtractedCovars}. For continuous covariates, we calculated the average
+value, for binary covariates the percentage cover along the step. Since our
+floodmaps were dynamic, we extracted values from the floodmap that was closest
+in date to the actual step. We further derived a binary variable indicating
+whether or not a step crossed a road and we identified the average distance of
+each step to the nearest source of water or the nearest road. We square-rooted
+the values for \textit{DistanceToRoads} and \textit{DistanceToWater} to render a
+decreasing marginal impact of distance. We standardized extracted covariates to
+a mean of zero and a standard deviation of one using a z-score transformation.
+Furthermore, we screened covariates for correlation using Pearson's Correlation
+Coefficient. None of the covariates were overly correlated (\(|r| > 0.6\),
+\cite{Latham.2011}) and we retained all of them for modeling. To run a proper
+iSSF analysis, we also included two movement metrics, namely the cosine of the
+turning angle (\(cos(ta)\)) and the logarithm of the step length (\(log(sl)\)),
+as covariates in our models \citep{Avgar.2016}. The movement metric \(cos(ta\))
+served to describe the directionality of a step, as it transformed the circular
+measure of (\(-\pi, +\pi\)) into a linear measure (-1, 1). Thus, positive values
+indicated forward movements, whereas negative values indicated backward
+movements \citep{Turchin.1998}. The movement metric \(log(sl\)), on the other
+hand, was included as an indicator of the preferred step length. Since steps
+were regularly spaced by four hours, \(log(sl\)) can also be interpreted as
+movement rate.
 
@@
 \noindent We used the iSSF framework and parameterized a habitat selection model
-that further served to predict landscape permeability. This selection model
-operated under the assumption that dispersing wild dogs assigned a selection
-score \(w(x)\) of the following exponential form to each realized and random
-step:
+that further served to predict landscape permeability. This habitat selection
+model operated under the assumption that dispersing wild dogs assigned a
+selection score \(w(x)\) of the following exponential form to each realized and
+random step:
 
@@
 positive \(\beta\)-coefficients indicate selection of a covariate, negative
-\(\beta\)-coefficients avoidance of a covariate. Selected covariates therefore
+\(\beta\)-coefficients avoidance of a covariate. Selected covariates thus
 contribute positively to expected landscape permeability, whereas avoided
@@
 full model. We then ranked models according to AIC, assessed relative model
-weights, and identified the most parsimonious model. Due to convergence issues
+weights, and identified the most parsimonious model. Due to convergence issues,
 we were unable to model interactions between spatial covariates.
@@
 score. We identified the realized step's rank in each stratum and tallied rank
-frequencies for realized steps over all strata. Finally, we carried out a
+frequencies of realized steps across all strata. Finally, we carried out a
 Spearman-rank correlation analysis between ranks and associated frequencies and
@@
 we calculated the selection score \(w(x)\) for each 250m x 250m pixel covering
-our extended study area given the pixel's spatial covariates. Note that the
+the extended study area given the pixel's spatial covariates. Note that the
 movement metrics, i.e. \(cos(ta)\) and \(log(sl)\), did not enter the prediction
@@
 \(f(x) = \frac{x - min(x)}{max(x) - min(x)}\). The most impermeable landscape
-therefore received a value of 0, whereas the most permeable landscape received a
+thus received a value of 0, whereas the most permeable landscape received a
 value of 1.
@@
 orthogonal neighbors. Because African wild dogs have been observed to cover
-massive dispersal distances \citep{DaviesMostert.2012, Masenga.2016,
-Cozzi.2020}, we did not limit LCPs to a maximal effective cost. After
-computation, we tallied overlapping LCPs to identify high-frequency routes.
+massive dispersal distances (\citeauthor{DaviesMostert.2012}, 2012;
+\citeauthor{Masenga.2016}, 2016; \citeauthor{Cozzi.2020}, unpublished), we did
+not limit LCPs to a maximal effective cost. After computation, we tallied
+overlapping LCPs to identify high-frequency routes.
 
@@
 cost map (see \Cref{LCCExample}), which indicated the total minimal costs
-required to get from the selected point to any other location on the map. A LCC
+required to get from the selected point to any other location on the map. An LCC
 between two points was then obtained by adding up their cumulative cost maps and
@@
   \caption{Figures illustrating the process of identifying a least-cost corridor
-  between the source points A and B following \citep{Pinto.2009}. (a) Example of
-  a permeability surface, which serves to determine the costs of movement. (b1)
-  Cumulative cost map for point A, depicting the total minimal costs necessary
-  to get from point A to every other location. (b2) Cumulative cost map for
-  point B, depicting the total minimal costs necessary to get from point B to
-  every other location. (c) Summed cost maps of points A and B. (d) Masked out
-  corridor containing pixels that do not exceed the cheapest pixel by more than
-  5\%.}
+  between source points A and B following \cite{Pinto.2009}. (a) Example of a
+  permeability surface, which determines the costs of movement. (b1) Cumulative
+  cost map for point A, depicting the total minimal costs necessary to get from
+  point A to every other location. (b2) Cumulative cost map for point B,
+  depicting the total minimal costs to get from point B to every other location.
+  (c) Summed cost maps of points A and B. (d) Masked out corridor containing
+  pixels that do not exceed the cheapest pixel by more than 5\%.}
   \label{LCCExample}
@@
 showed that dispersing wild dogs moved in a directional and fast manner, as
-indicated by a positive selection for small turning angles, i.e. high
-\(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21,
-\Cref{PermeabilityResults}a) and longer steps, i.e. high \(log(sl)\) (\(\beta\)
-= 0.05, 95\% CI = 0.02 to 0.09, \Cref{PermeabilityResults}a). Given the chance,
-dispersers avoided moving through water (\(\beta\) = -0.5, 95\% CI -0.67 to
--0.32, \Cref{PermeabilityResults}a) but selected for locations in its vicinity,
+indicated by selection of small turning angles, i.e. high \(cos(ta)\) (\(\beta\)
+= 0.14; 95\% CI = 0.07 to 0.21, \Cref{PermeabilityResults}a) and long steps,
+i.e. high \(log(sl)\) (\(\beta\) = 0.05, 95\% CI = 0.02 to 0.09,
+\Cref{PermeabilityResults}a). Given the chance, dispersers avoided moving
+through water (\(\beta\) = -0.5, 95\% CI -0.67 to -0.32,
+\Cref{PermeabilityResults}a) but selected for locations in its vicinity,
 although the effect was not significant (\(\beta\) = -0.24, 95\% CI = -0.67 to
@@
     \caption{(a) Estimated selection coefficients from the most parsimonious
-    permeability model. Negative coefficients indicate avoidance of a covariate,
-    positive coefficients selection of a covariate. Whiskers delineate the
-    95\%-CIs for estimated parameters. The intercept is meaningless in the iSSF
-    framework and was thus omitted. (b) Results from the k-fold cross validation
-    for case control studies. The left graph shows rank frequencies of
-    \textit{realized} steps according to predictions, whereas the right graph
-    shows rank frequencies of \textit{randomly selected} steps according to
-    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
-    resulting from 100 repetitions of the k-fold cross validation. The blue
-    smoothing line was fitted using a LOESS regression and solely serves to aid
-    the eye in detecting the trends. Correlation coefficients suggest that our
-    prediction is significant and robust, evidenced by the fact that the
-    confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
+    habitat selection model. Negative coefficients indicate avoidance of a
+    covariate, positive coefficients selection of a covariate. Whiskers
+    delineate the 95\%-CIs for estimated parameters. The intercept is
+    meaningless in the iSSF framework and was thus omitted. (b) Results from the
+    k-fold cross validation for case-control studies. The left graph shows rank
+    frequencies of \textit{realized} steps according to predictions, whereas the
+    right graph shows rank frequencies of \textit{randomly selected} steps
+    according to predictions. \(\bar{r}_s\) indicates the mean correlation
+    coefficient resulting from 100 repetitions of the k-fold cross validation.
+    The blue smoothing line was fitted using a LOESS regression and solely
+    serves to aid the eye in detecting the trends. Correlation coefficients
+    suggest that our prediction is significant and robust, evidenced by the fact
+    that the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
     random}\) do not overlap and that there is strong and significant
@@
 %     which appears to provide the most suitable landscape for dispersers. The
-%     white square highlights our core study area.}
+%     white square highlights the core study area.}
 %     \label{PermeabilityMap}
@@
         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-          \begin{minipage}{0.7\textwidth}
+          \begin{minipage}{0.65\textwidth}
             \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
@@
     \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
-    Permability was predicted by calculating selection scores \(w(x)\) for each
-    pixel based on the pixel's underlying covariates. Areas that dispersers find
-    easy to traverse are depicted in bright colors. Bold white lines delineate
-    the borders of the KAZA-TFCA, whereas dashed white lines show country
-    borders.}
+    Permability was predicted by calculating selection scores \(w(x) =
+    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each pixel based on
+    the pixel's underlying covariates (\(x_i\)) and estimated habitat
+    preferences (\(\beta_i\)). Areas that dispersers find easy to traverse are
+    depicted in bright colors. Bold white lines delineate the borders of the
+    KAZA-TFCA, whereas dashed white lines show country borders.}
     \label{PermeabilityMap}
@@
 \begin{table}[h]
-  \caption{Comparison of median permeability across countries, separated
-  according to areas within and outside the KAZA-TFCA, as well as within and
-  outside protected areas. High values indicate high permeability, whereas low
-  values correspond to low permeability.}
+  \caption{Comparison of median permeability (interquantile range in brackets)
+  across countries, separated according to areas within and outside the
+  KAZA-TFCA, as well as within and outside formally protected areas. High values
+  indicate high permeability, whereas low values correspond to low
+  permeability.}
   \label{PermeabilityComp}
   \begin{center}
-    \resizebox{\textwidth}{!} {
+    \resizebox{0.9\textwidth}{!} {
       \begin{tabular}{llllllllr}
@@
       \hline
-      Angola & & 0.46 & 0.18 & & 0.46 & 0.18 & & 0.29 \\
-      Namibia & & 0.26 & 0.13 & & 0.27 & 0.12 & & 0.16 \\
-      Botswana & & 0.33 & 0.19 & & 0.38 & 0.17 & & 0.25 \\
-      Zimbabwe & & 0.06 & 0.05 & & 0.09 & 0.04 & & 0.05 \\
-      Zambia & & 0.07 & 0.03 & & 0.06 & 0.03 & & 0.04 \\
+      Angola & & 0.46 (0.42) & 0.18 (0.40) & & 0.46 (0.42) & 0.18 (0.40) & & 0.29 (0.45) \\
+      Namibia & & 0.26 (0.39) & 0.13 (0.20) & & 0.27 (0.38) & 0.12 (0.16) & & 0.16 (0.31) \\
+      Botswana & & 0.33 (0.38) & 0.19 (0.23) & & 0.38 (0.35) & 0.17 (0.23) & & 0.25 (0.33) \\
+      Zimbabwe & & 0.06 (0.25) & 0.05 (0.04) & & 0.09 (0.26) & 0.04 (0.03) & & 0.05 (0.07) \\
+      Zambia & & 0.07 (0.11) & 0.03 (0.05) & & 0.06 (0.11) & 0.03 (0.05) & & 0.04 (0.07) \\
       \hline
-      Overall & & 0.21 & 0.07 & & 0.21 & 0.07 & & 0.11 \\
+      Overall & & 0.21 (0.39) & 0.07 (0.18) & & 0.21 (0.39) & 0.07 (0.18) & & 0.11 (0.29) \\
       \end{tabular}
@@
 Hwange corridor into Zimbabwe’s Matusadona NP. Our model did not detect any
-significant direct corridors between Zimbabwe and Zambia and only very limited
+significant direct corridors between Zimbabwe and Zambia and only a very limited
 W-E direct connection between the Okavango region and Namibia’s Khaudum NP. We
@@
 more central regions before heading towards their destination at the periphery.
-With the exception of the Central Kalahari Game Reserve corridor, our model did
-not detect any significant connectivity outside the boundaries of KAZA-TFCA.
+Except for the Central Kalahari Game Reserve corridor, our model did not detect
+any significant connectivity outside the boundaries of KAZA-TFCA.
 
@@
     between the same source points as illustrated in subfigure (a). For ease of
-    spatial reference we also labeled national parks that are mentioned in the
-    main text (dark-grey).}
+    spatial reference we also labeled some national parks (dark-grey).}
     \label{LeastCost}
@@
 \section{Discussion}
-% Recap of what we did
+\label{Discussion}
+% Recap of Study
 We combined GPS relocation data of dispersing African wild dog coalitions with
@@
 habitat selection or avoidance during dispersal. We then used this knowledge to
-create a permeability surface and assess connectivity by means of identifying
-dispersal corridors across the world’s largest transboundary conservation area,
-the Kavango-Zambezi Transfrontier Conservation Area.
-
-% Finding Regarding Human Influence
+create a permeability surface and assess connectivity by identifying dispersal
+corridors across the world’s largest transboundary conservation area, the
+Kavango-Zambezi Transfrontier Conservation Area.
+
+\subsection{Habitat Selection Model}
+% Finding: Human Influence
 Our results indicate that dispersers avoided anthropogenic influences,
@@
 forgo human-dominated landscapes and we conjecture that differences to
-observations by Davies-Mostert and colleagues are be due to the unavailability
-of alternative ``natural'' routes in South Africa. That is, dispersers in South
+observations by Davies-Mostert and colleagues are due to the unavailability of
+alternative ``natural'' routes in South Africa. That is, dispersers in South
 Africa may be forced to cross human-dominated landscapes despite a strong
@@
 
-% Finding Regarding Water
-Water was identified as being another major obstacle to dispersal. In earlier
-studies, water has been shown to be almost impermeable for resident packs and it
-is assumed that residents may in fact use it as easy-to-defend territory
-boundaries \citep{Cozzi.2013}. Although dispersers have different needs and
-motivations than residents, they have also been shown to be partially hindered
-by larger water bodies \citep{Abrahms.2017, Cozzi.2020}. Our findings thus align
-with previous research and stress the importance to accurately and dynamically
-represent water bodies, notably for such pulsing ecosystems as represented by
-the Okavango Delta.
-
-% Finding Regarding Distance To Water
+% Finding: Water
+Water was identified as being another major obstacle to dispersal. According to
+earlier studies, water is almost impermeable for resident packs and it is
+assumed that residents use water bodies as easy-to-defend territory boundaries
+\citep{Cozzi.2013}. Although dispersers have different needs and motivations
+than residents, they have also been shown to be partially hindered by larger
+water bodies (\citeauthor{Abrahms.2017}, 2017; \citeauthor{Cozzi.2020},
+unpublished). Our findings thus align with previous research and stress the
+importance to accurately and dynamically represent water bodies, notably for
+such pulsing ecosystems as represented by the Okavango Delta.
+
+% Finding: Distance To Water
 While we found that dispersers avoided moving \textit{through} water, we
@@
 determines the distribution of many herbivores, especially in dry biomes
-\citep{Western.1975}. For the Okavango Delta in particular, it has been shown
+\citep{Western.1975}. For the Okavango Delta, in particular, it has been shown
 that prey abundance strongly correlates with the extent of the flood
@@
 
-% Finding Regarding Different Permeabilities
-According to our prediction of landscape permeability, permeability of northern
-Botswana and southern Angola are relatively high. High permeability in these
-countries is mainly caused by a combination of low human presence and
+\subsection{Permeability Surface}
+% Finding: Permeabilities
+According to our prediction of landscape permeability, northern Botswana and
+southern Angola provide relatively permeable landscapes. High permeability in
+these countries is mainly caused by a combination of low human presence and
 activities, low tree cover, high shrubs/grassland cover and close distance to
@@
 other hand, is mainly caused by severe anthropogenic presence and other human
-activities. Despite the fact that the KAZA-TFCA covers many high permeability
-hot-spots, several permeable regions remain uncovered by the KAZA-TFCA, offering
-exciting opportunities for future expansions of already existing protected
-areas.
-
-% Finding Regarding Corridor Overlap with KAZA
-Our results moreover suggest that all major dispersal corridors of African wild
-dogs run within the boundaries of the KAZA-TFCA, demonstrating the potential
-value of such an initiative. Within the KAZA-TFCA, however, some key
-connectivity areas are not formally protected and their protection status may be
-subject to future changes under the influences of an expanding human population,
-farming, and cropping industry.
+activities. Even though the KAZA-TFCA covers the majority of permeability
+hot-spots, several highly permeable regions remain uncovered by the KAZA-TFCA,
+offering exciting opportunities for future expansions of already existing
+protected areas.
+
+\subsection{Least-Cost Paths \& Least-Cost Corridors}
+% Finding: Overlap Corridors and KAZA
+Our analyses of dispersal paths and corridors suggest that all major dispersal
+corridors of African wild dogs run within the boundaries of the KAZA-TFCA,
+demonstrating the potential value of such an initiative. Within the KAZA-TFCA,
+however, some key connectivity areas are not formally protected and their
+protection status may be subject to future changes under the influences of an
+expanding human population, farming, and cropping industry. Although overall
+connectivity gradually decreases towards the borders of the study area, this
+might partly be a result of boundary-effects \citep{Koen.2010}. In general, our
+network of corridors shows high similarities to predicted movement corridors of
+dispersing lions \citep{Elliot.2014}, which not only reinforces confidence in
+our predictions but also suggests potential synergies between conservation of
+the two species.
 
 % Finding Regarding Dispersal Hub
-Importantly, we identified the Okavango-Linyanti region as potential dispersal
+Importantly, we identified the Okavango-Linyanti region as a potential dispersal
 hub through which dispersers gain access to more peripheral regions of the
 KAZA-TFCA. While this might partly be owed to the area's central position within
-the KAZA-TFCA, it is likely also due to its environmental characteristics,
-namely due to human activities and the presence of impermeable water bodies such
-as the Okavango Delta and the Linyanti Swamp, which potentially funnel dispersal
-movements. The region is also instrumental for connecting subpopulations
-remaining in Zambia’s Kafue and Zimbabwe’s Hwange NPs. These subpopulations are
-separated by the Zambezi River, forcing dispersers to detour via northern
-Botswana to reach more peripheral areas. The key role of the Okavango-Linyanti
-region for the overall connectivity of the KATA-TFCA thus calls for actions to
-secure its protection status in the future. While the region is currently a
-Wildlife Management Area devoted to photographic tourism, it has neither the
-status of a National Park nor that of a Game Reserve and so potentially faces
-growing pressure towards exploitation of its natural resources.
-
-% Finding Regarding Non-Protected Corridor Between Zambia and Botswana
+the KAZA-TFCA, it is likely also due to its environmental and anthropogenic
+characteristics, namely due to the absence of human activities and the presence
+of impermeable water bodies, such as the Okavango Delta and the Linyanti Swamp,
+which potentially funnel dispersal movements. The region is also instrumental
+for connecting subpopulations remaining in Zambia’s Kafue and Zimbabwe’s Hwange
+NPs. These subpopulations are currently separated by the Zambezi River, forcing
+dispersers to detour via northern Botswana to reach more peripheral areas. The
+key role of the Okavango-Linyanti region for the overall connectivity of the
+KATA-TFCA thus calls for actions to secure its protection status in the future.
+While the region is presently a Wildlife Management Area devoted to photographic
+tourism, it has neither the status of a National Park nor that of a Game Reserve
+and so potentially faces growing pressure towards exploitation of its natural
+resources.
+
+% Finding: Corridor Between Moremi & Kafue
 A similar case of non-formally protected but key dispersal landscape is
-represented by the area south of Kafue NP in Zambia, for which a disruption of
-its main and narrow dispersal corridor would result in a considerable isolation
-of its subpopulations, for no other corridor links Kafue NP to the rest of the
+represented by the area south of Kafue NP in Zambia, for which disruption of its
+main and narrow dispersal corridor would result in considerable isolation of its
+subpopulations, for no other corridor links Kafue NP to the rest of the
 KAZA-TFCA.
 
-% Finding Regarding
-Our results also highlighted a potential southwards corridor between the
-Okavango-Linyanti ecosystem and the Central Kalahari Game Reserve. Some areas
-through which the corridor runs are neither part of KAZA-TFCA nor of any
+% Finding: Corridor Between Moremi & Kalahari
+Our results also revealed a potential southwards corridor between the
+Okavango-Linyanti ecosystem and the Central Kalahari Game Reserve.
+\cite{Elliot.2014} identified a similar corridor for dispersing lions, which
+shows that protecting the area would benefit lions and wild dogs alike. Some
+areas through which the corridor runs are neither part of KAZA-TFCA nor of any
 protected status. Human presence and activities along the national road that
-longitudinally traverses the corridor as well as the related increased
-mortality caused by direct persecution and collisions \citep{Cozzi.2020} may
-limit realized connectivity. In fact, despite having dispersed south of the
-Okavango Delta, none of the monitored dispersing coalitions crossed the national
-road and reached the Central Kalahari Game Reserve \citep{Cozzi.2020}.
-\citep{Elliot.2014} identified a similar corridor for dispersing lions, which
-shows that protecting the area would benefit lions and wild dogs alike.
-
-Although overall connectivity gradually decreased towards the borders of the
-study area, this might partly be a result of boundary-effects \citep{Koen.2010}.
-
-In general, our network of dispersing corridors shows high similarities to
-predicted movement corridors of dispersing lions \citep{Elliot.2014}, which not
-only reinforces confidence in our own predictions but also suggests potential
-synergies between conservation of the two species.
-
-% The goal of our study was to (1) assess landscape permeability for dispersing
-% wild dogs and to (2) subsequently use this knowledge to examine connectivity
-% across the greater extent of the KAZA-TFCA. To achieve these goals, we
-% parameterized a permeability model in which we used GPS data of 16 dispersing
-% wild dog coalitions to quantify their habitat preferences with respect to land
-% cover, protection status, and human influence. While we found significant
-% impacts of different land covers and human influence, the influence of
-% protection status was not statistically clear. We further applied the
-% parameterized permeability model and predicted a permeability surface for the
-% entire KAZA-TFCA, which depicted the ease at which dispersers crossed a specific
-% landscape. Finally, based on this permeability map, we computed LCPs and LCCs
-% between protectorates and thereby highlighted dispersal corridors of African
-% wild dogs.
-%
-% \subsection{Habitat Selection Model}
-% % Permeability Model: Land Cover, Water
-% Our most parsimonious permeability model revealed that, with respect to land
-% cover, dispersers strictly avoided areas covered by water (including swamps,
-% wetlands). This result corroborates previous research on wild dogs (e.g.
-% \cite{Abrahms.2017}) and is rather unsurprising, as the species is not well
-% adapted for moving through water. For residents, water is almost impermeable and
-% frequently serves as natural boundary to their home-ranges \citep{Cozzi.2013}.
-% It is conceivable that dispersers are more willing than residents to cross
-% waterbodies \citep{Cozzi.2020}, but we did not include residents in our analyses
-% and could not test for such differences. Regardless of that, water appears to
-% pose a serious obstacle for dispersers and is therefore likely to reduce
-% landscape permeability for dispersing wild dogs.
-%
-% % Permeability Model: Land Cover, DistanceToWater
-% While dispersers avoided moving \textit{through} water, they preferred moving
-% \textit{close to} water as indicated by avoidance of high distance to water.
-% While its effect was insignificant, inclusion of the covariate
-% \textit{DistanceToWater} strictly increased the AIC in all of our models and
-% suggests that distance to water influenced permeability negatively. We
-% hypothesize that high prey-availability (e.g. impala, kudu) in proximity to
-% water explains this result. It is well documented that water is a major driver
-% of wildlife and that it determines the distribution of many herbivores,
-% especially in dry biomes \citep{Western.1975}. For the Okavango Delta in
-% particular, it has been shown that prey abundance strongly correlates with the
-% extent of the flood \citep{Bonyongo.2005} and that an increasing concentration
-% of ungulates around remaining water-sources during the dry season might attract
-% dispersing wild dogs \citep{Ogutu.2014}. Following the same logic, however, apex
-% predators (e.g. lions, spotted hyenas) and resident wild dogs may also prefer
-% closeness to water \citep{Valeix.2009}, which may occasionally force dispersers
-% to move distant to water \citep{Ndaimani.2016} to avoid intraguild competition
-% \citep{Creel.1996, Mills.1997}. We could not control for the presence or absence
-% of apex predators or conspecifics because we lacked corresponding data, but
-% these considerations might explain why we did not find a statistically clear
-% effect of distance to water.
-%
-% % Permeability Model: Land Cover, Trees & Shrubs
-% In line with existing literature on moving wild dogs \citep{Abrahms.2017}, we
-% found that dispersers avoided habitats with high percentages of tree cover,
-% whereas they preferred areas with high percentages of shrubs/grassland cover.
-% Comparable preferences were reported for residents, which typically reside in
-% savanna-like habitats but rarely inhabit areas with dense tree cover
-% \citep{Estes.2012}. This may be explained by wild dogs' hunting behavior, where
-% open habitats are favored because they offer increased visibility and thus
-% facilitate hunting \citep{Estes.1967}. Additionally, we found a strong negative
-% correlation between percentage cover by shrubs/grassland and percentage cover by
-% bareland (although we did not include bareland in our models). The observed
-% preference for shrubs/grassland could therefore also indicate an aversion to
-% bareland. Again, we hypothesize that this aversion is driven by low
-% prey-availability in non-vegetated habitats \citep{Pettorelli.2011}.
-% Unfortunately, we lacked detailed vegetation data and could not distinguish
-% vegetation types further.
-%
-% % Permeability Model: Human Influence
-% Finally, considering human influence, we found that dispersers strongly avoided
-% human pressures arising from human density, farming, and roads. Previous
-% research on dispersing wild dogs was inconclusive about the impact of human
-% presence on permeability, as some researchers reported a high willingness of
-% dispersers to cross human-dominated landscapes \citep{DaviesMostert.2012},
-% whereas others found that humans are avoided almost entirely during dispersal
-% \citep{Masenga.2016}. Our results support the notion that dispersers strongly
-% avoid human-dominated areas, but we hypothesize that this finding depends on the
-% availability of alternative routes. The dipsersers monitored by
-% \cite{DaviesMostert.2012} originated from habitats almost exclusively surrounded
-% by humans, whereas our dipsersers departed from regions with various
-% possibilities to move through natural habitats. A lack of alternative routes
-% combined with strong inbreeding avoidance intrinsic in African wild dogs
-% \citep{McNutt.1996} might have forced the dispersers reported in
-% \cite{DaviesMostert.2012} to cross human-dominated landscapes despite a strong
-% aversion against humans. Although we found that our dispersers strictly avoided
-% humans, \citeauthor{Cozzi.2020} (in review) reports that human persecution
-% remains the main cause of mortality for dispersers, suggesting that avoidance is
-% not strong enough to prevent human-induced fatalities. We argue that human
-% influence remains a source of reduced landscape permeability and that the
-% proceeding transformation of natural landscapes into human altered habitats
-% continues to undermine wild dogs' ability to dipserse. Since our human influence
-% proxy did not include minor gravel roads, our finding that humans are avoided
-% does not contradict the observation that wild dogs preferably move on gravel
-% roads during activity \citep{Abrahms.2016}. With four-hourly GPS relocations we
-% felt that the temporal resolution was insufficient to depict such fine scale
-% effects \citep{Thurfjell.2014}.
-%
-% \subsection{Permeability Surface}
-% % Permeability Surface
-% Our prediction of landscape permeability revealed vast regional differences.
-% Across the entire study area, northern Botswana stood out as it comprises the
-% highest density of highly permeable landscapes. The hospitality of this region
-% is caused by a combination of low human influence, low tree cover, high
-% shrubs/grassland cover and close distance to water. Thus, although swamps,
-% wetlands, and water (e.g. Okavango Delta, Linyanti Swamp) in northern Botswana
-% provide little permeability, their surroundings act as strong attractants to
-% dispersers. It is therefore not very surprising that the region serves as source
-% for the recolonization of surrounding habitats \citep{Cozzi.2013}. Moreover, we
-% found much higher permeability inside the KAZA-TFCA than outside its borders,
-% which suggests a good fit between the planned protectorates and areas that are
-% suitable for dispersal. We made similar observations for protection status,
-% where predicted permeabilities inside protectorates are much higher than in
-% pastoral regions. Yet, several highly permeable regions remain uncovered by the
-% KAZA-TFCA or other protected areas, proposing exciting opportunities for future
-% expansions of already existing protectorates. Across countries, we revealed that
-% Zambia and Zimbabwe offer the least permeable landscapes. Visual inspection of
-% our covariate-layers disclosed that this is mainly owed to severe anthropogenic
-% influence in Zambia and Zimbabwe. In contrast, Botswana and Angola host vast
-% uninhabited areas, resulting in more permeable landscapes for dispersers.
-% However, all of our predictions were based on observations from dispersing wild
-% dogs departing from northern Botswana, which might have introduced a bias in
-% estimated habitat preferences and resulting predictions.
-%
-% \subsection{Least-Cost Paths \& Least-Cost Corridors}
-% % Least Cost Paths & Least Cost Corridors
-% Our LCP- and LCC-analyses disclosed several major movement corridors within the
-% extended study area and suggested a decent congruence between predicted
-% corridors and the layout of the KAZA-TFCA. That is, only few corridors run
-% outside the planned boundaries of the KAZA-TFCA and several currently
-% unprotected corridors will become protected with realization of the KAZA-TFCA.
-% For example, the high-frequency corridor running from southern Zambia into the
-% Kafue NP (\Cref{LeastCost}a \circled{3}) currently contains extended unprotected
-% sections that will be transformed into protectorates under the KAZA-TFCA. We
-% thus expect that dispersal between such regions will be vastly facilitated
-% thanks to the KAZA-TFCA. Our network of LCPs and LCCs furthermore highlighted
-% the Chobe-Moremi system in northern Botswana, which is traversed by several
-% LCPs and LCCs and seems to hold the potential to serve as central hub through
-% which dispersers gain access to more remote regions of the KAZA-TFCA. Still, our
-% predictions also indicated that movements between wild dog subpopulations may
-% often involve extended detours. This is especially the case for regions in the
-% northern sectors of the KAZA-TFCA. For instance, only few LCPs connect the
-% Zimbabwean and Zambian sectors, such that moving directly from the Matusadona NP
-% to the Kafue NP seems almost impossible and likely requires to make a detour
-% through the Chobe-Moremi region. Similarly, few LCPs directly link the Mavinga
-% NP in Angola and the Kafue NP in Zambia. Moreover, we discovered potential
-% dispersal routes between the Chobe-Moremi ecosystem and the Central Kalahari
-% NP that are not yet proclaimed protected areas. Nonetheless, a protected link
-% between the Central Kalahari NP and the Chobe-Moremi ecosystem might grant
-% southern populations invaluable access to the Chobe-Moremi-Hub and
-% subsequently to the entire KAZA-TFCA. In fact, \cite{Elliot.2014} identified a
-% similar corridor for dispersing lions, proposing that protecting the area would
-% benefit lions and wild dogs alike. The absence of direct LCPs and LCCs between
-% some regions of the KAZA-TFCA is likely owed to an impermeable landscape between
-% those regions and demonstrates that protection of such areas in their current
-% state would be futile. That is, impermeable landscapes first need to be
-% transformed into more permeable habitats before protection pays off. Although,
-% we noted that connectivity gradually decreases towards the borders of our study
-% area, this might partly be owed to boundary-effects \citep{Koen.2010}.
-% Interestingly, our network of LCPs and LCCs shows surprising similarities to
-% predicted movement corridors of dispersing lions \citep{Elliot.2014}.
-% Remarkably, corridors that we identified as high-frequency corridors are also
-% visible in the network presented by \cite{Elliot.2014}. This not only reinforces
-% confidence in our own predictions but it also suggests potential synergies
-% between conservation of the two species.
-
+longitudinally traverses the corridor may limit realized connectivity,
+especially since the road is reported to increase wild dog mortality through
+direct persecution and collisions (\citeauthor{Cozzi.2020}, unpublished). In
+fact, despite having dispersed south of the Okavango Delta, none of the
+monitored dispersing coalitions crossed the national road and reached the
+Central Kalahari Game Reserve (\citeauthor{Cozzi.2020}, unpublished).
+
+\subsection{Future Research}
 % Future Research: Genetic Resistance
@@
 % Future Research: Better Covariates
-Future research should also strive to improve representation of geo-referenced
-covariates. Because we predicted connectivity for such an extensive area, we
-were strongly limited in available and reliable land-cover data. As a result,
-our representation of vegetation remained rather basic with only three
-vegetation types (tree-cover, shrubs/grassland, bareland). Although there are a
-few large-scale, high-resolution land cover datasets, such as Globeland's 30m
-\citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019} land
-cover datasets, we found their classifications often erroneous for our study
-area. This is probably owed to the KAZA-TFCA's unique landscape and
-corresponding difficulties in remote sensing satellite imagery using traditional
-indices \citep{Wolski.2017}. Nevertheless, land cover is a very powerful
-predictor of animal movement \citep{Thurfjell.2014}, which stresses the need for
-more reliable land-cover data.
+Future research should also strive to improve the representation of
+geo-referenced covariates. Because we predicted connectivity for such an
+extensive area, we were strongly limited in available and reliable land cover
+data. As a result, our representation of vegetation remained rather basic with
+only three vegetation types (tree-cover, shrubs/grassland, bare land). Although
+there are a few large-scale, high-resolution land cover datasets, such as
+Globeland's 30m \citep{Chen.2015} and the European Space Agency's and 20m
+\citep{ESA.2019} land cover datasets, we found their classifications often
+erroneous for our study area. This is probably owed to the KAZA-TFCA's unique
+landscape and corresponding difficulties in remote sensing satellite imagery
+using traditional indices \citep{Wolski.2017}. Nevertheless, land cover is a
+very powerful predictor of animal movement \citep{Thurfjell.2014}, highlighting
+the need for more reliable land cover data.
 
 % Future Research: Competitor Presence
-In a similar manner, we would desire better knowledge of the whereabouts of
-other species. In particular, we would desire location data of apex predators,
-potential prey, and conspecifics. It is well documented that resident wild dogs
-frequently compete with other carnivores \citep{Creel.1998} and we would expect
-to find competition between dispersing wild dogs and apex predators too.
-Furthermore, knowledge about the spatial distribution of prey could verify our
-hypothesis that dispersers move along water due prey-availability. Lastly, we
-could not control for the likelihood of encountering conspecifics during
-dispersal . Yet, for dispersing meerkats, a species with surprisingly similar
-social structure to wild dogs, it has been shown that the social landscape, i.e.
-the presence or absence of conspecifics, strongly shapes behavior during
-dispersal \citep{Cozzi.2018}. More specifically, it has been shown that
-dispersers select for high density of conspecifics while still inside their
-home-range, but that they avoid high densities of conspecifics once they leave
-their natal home-range (we replicated this analysis for our core study area; see
+Similarly, we would desire better knowledge of the whereabouts of other species.
+In particular, we would desire location data of apex predators, prey, and
+conspecifics. It is well documented that resident wild dogs frequently compete
+with other carnivores \citep{Creel.1998} and we would expect to find competition
+between dispersing wild dogs and apex predators too. Furthermore, knowledge
+about the spatial distribution of prey could verify our hypothesis that
+dispersers move along water due to prey-availability. Lastly, we could not
+control for the likelihood of encountering conspecifics during dispersal. Yet,
+for dispersing meerkats, a species with a surprisingly similar social structure
+to wild dogs, it has been shown that the social landscape, i.e. the presence or
+absence of conspecifics, strongly shapes behavior during dispersal
+\citep{Cozzi.2018}. More specifically, it has been shown that dispersers select
+for high density of conspecifics while still inside their home-range, but that
+they avoid high densities of conspecifics once they leave their natal home-range
+(we replicated this analysis for the core study area; see
 \Cref{Appendix:SocialLandscape}). Spatially explicit knowledge about the
@@
 Besides urging for more and better data, we also want to emphasize that our
-analytical approach is neither limited to African wild dogs, nor to our study
-area. All covariates used throughout this study are available on a global scale
-and they are likely to be important determinants of landscape permeability for
-other species. Our framework could thus easily be applied to assess connectivity
-for a wide range of species, provided suitable relocation data of the species
-exists. Expanding our analyses to other species would surely yield interesting
-insights on the congruence of movement corridors of different species and
-thereby highlight areas that are used by multiple species for dispersal.
+analytical approach is neither limited to African wild dogs nor our study area.
+All covariates used throughout this study are available on a global scale and
+they are likely to be important determinants of landscape permeability for other
+species. Our framework could easily be applied to assess connectivity for a wide
+range of species, provided suitable relocation data of the species exist.
+Expanding our analyses to other species would surely yield interesting insights
+on the congruence of movement corridors and thereby highlight areas that are
+exceptionally valuable for the conservation of multiple species.
 
@@
 To this end, we identified movement corridors between pre-defined start and
-end-points, which assumes that dispersers have a specific end-point in mind.
-However, pre-defining end-points might not even be necessary, as one could
-invert the iSSF and use it as a mechanistic movement model to simulate dispersal
-without known end-point. That is, one could define a set of source points at
-locations where wild-dogs are assumed or known to reside and simulate dispersers
-according to habitat-preferences that were derived using the iSSF framework.
-End-points would not be restricted to a certain location and movement corridors
-would emerge naturally as the result of a myriad of simulated dispersal events
-(for an example of simulated dispersal events see
-\Cref{Appendix:DispersalSimulation}).
-
-% Conclusion
-Our results demonstrate that indeed the majority of dispersal corridors for
-African wild dogs are covered by the boundaries of the KAZA-TFCA, suggesting
-that the KAZA-TFCA will severely contribute to the long-term viability of this
+end-points. This assumes that dispersers have a specific end-point in mind when
+dispersing, which is likely not always the case. However, pre-defining
+end-points might not even be necessary, as one could use the iSSF framework as a
+mechanistic movement model to simulate dispersal without a known end-point. That
+is, one could define a set of source points at locations where wild-dogs are
+assumed or known to reside and simulate dispersers according to identified
+habitat-preferences. As a consequence, movement corridors would emerge naturally
+as the result of a myriad of simulated dispersal events (for an example of
+simulated dispersal events see \Cref{Appendix:DispersalSimulation}).
+
+\newpage
+\subsection{Conclusion}
+Our results demonstrate that the majority of dispersal corridors for African
+wild dogs are covered by the boundaries of the KAZA-TFCA, suggesting that the
+KAZA-TFCA will significantly contribute to the long-term viability of this
 species. Nevertheless, some unprotected dispersal routes do exist outside the
-KAZA-TFCA and suggest opportunities for further expansions in the future.
+KAZA-TFCA and indicate opportunities for further expansions in the future.
 Moreover, our connectivity network of least-cost paths and least-cost corridors
-revealed especially the Okavango-Linyanti region as central dispersal hub
-through which dispersers gain access to more remote regions of the KAZA-TFCA. We
-therefore stress the importance of northern Botswana for the conservation of
-this charismatic species. Finally, our investigations showed that human pressure
-remains a severe impediment to dispersal and strictly reduces landscape
+revealed the Okavango-Linyanti region as a central dispersal hub through which
+dispersers gain access to more remote regions of the KAZA-TFCA. Therefore, we
+stress the importance of northern Botswana for the conservation of this
+charismatic species. Finally, our investigations showed that human pressures
+remain a severe impediment to dispersal and therefore reduce landscape
 connectivity. Successful conservation of the African wild dog is thus contingent
-on the willingness of local authorities, policy makers, and land managers to
-provide vast areas that are liberated from human strains. Ultimately, our work
-contributes to the growing field of connectivity literature and has important
-implications for the management and conservation of the endangered African wild
-dog.
+on the willingness of local authorities, policymakers, and land managers to
+preserve areas that are largely liberated from human strains. Ultimately, our
+work contributes to the growing field of connectivity literature and has
+important implications for the management and conservation of the endangered
+African wild dog.
 
@@
 constant supply of freshly brewed coffee. Particular gratitude goes to Dominik
-Behr, who supervised my work on a daily basis and greatly contributed to all
-aspects of the thesis. I genuinely enjoyed the frequent meetings and brainstorms
-and I really feel privileged to have had the opportunity of working with such a
-great scientist. I am also greatly indebted to Prof. John Fieberg, who visited
-our department this summer and invested a great amount of time to consult and
-improve all statistical aspects of the thesis, as well as to deepen my personal
-understanding of movement ecology. Furthermore, I thank the Okavango Research
-Institute, in particular Piotr Wolski, for sharing all of their Python scripts,
-floodmaps and knowledge on developing dynamic floodmaps. Also, I want to thank
-Stefan Sommer for sharing and teaching his hard earned knowledge on scientific
-writing. Additionally, I want to thank all of my colleague master students who
-have been wonderful companions and made the time at university a blast.
-Likewise, I thank my parents for supporting all of my undertakings and for
-providing first class shelter and food. Last, but most definitely not least, I
-also greatly thank my girlfriend, who was always there for me and provided
-welcome distractions for my mind.
+Behr, who supervised my work daily and greatly contributed to all aspects of the
+thesis. I genuinely enjoyed the frequent meetings and brainstorms and I feel
+privileged to have had the opportunity of working with such a great scientist. I
+am also greatly indebted to Prof. John Fieberg, who visited our department this
+summer and invested a great amount of time to consult and improve all
+statistical aspects of the thesis, as well as to deepen my understanding of
+movement ecology. Furthermore, I thank the Okavango Research Institute,
+especially Piotr Wolski, for sharing all of their Python scripts, floodmaps, and
+knowledge on developing dynamic floodmaps. Also, I want to thank Stefan Sommer
+for sharing and teaching his hard-earned knowledge on scientific writing.
+Additionally, I want to thank all of my colleague master students who have been
+wonderful companions and made the time at university a blast. Likewise, I thank
+my parents for supporting all of my undertakings and for providing first-class
+shelter and food. Last, but most definitely not least, I also greatly thank my
+girlfriend, who was always there for me and provided welcome distractions for my
+mind.
 
@@
 by the application of static water-polygons that did not cover permanent waters
-correctly anymore. We therefore revised the algorithm and allowed for a more
+correctly anymore. Therefore, we revised the algorithm and allowed for a more
 dynamic polygonization of water. That is, for each MODIS image we calculated new
@@
 From the original layer we only retained the categories \textit{wetland} and
-\textit{waterbodies} and collectively reclassified them to \textit{water}.
+\textit{water bodies} and collectively reclassified them to \textit{water}.
 Globeland had an original resolution of 30m x 30m, so we coarsened the layer to
@@
     in south-eastern direction, where it nourishes several tributaries. The
-    extent of the flood peaks around August or September, after which it slowly
+    extent of the flood peaks around August or September and then slowly
     retracts. Between December and March the reflectance properties of water and
@@
           & 30m x 30m (static), 500m x 500m (dynamic)
-            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
+            & Mode of each 250m x 250m cell (static), disaggregation to 250m x 250m (dynamic)
               & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
@@
   \caption{Results from the forward model selection procedure based on AIC
-  \citep{Burnham.2002} for our permeability model. Weights were calculated
+  \citep{Burnham.2002} for our habitat selection model. Weights were calculated
   using: \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i =
@@
 gives an overview of how these additional investigations integrate into our main
-work. Our main work is described in the blue boxes, which resulted in the
-identification of LCPs and LCCs. However, as we have emphasized in the
-discussion, the applicability of the iSSF framework expands beyond the
+work. Our main work is described in the blue boxes and concluded with the
+identification of LCPs and LCCs. However, as we have emphasized in
+\Cref{Discussion}, the applicability of the iSSF framework expands beyond the
 estimation of movement and habitat preferences and it could, in fact, be used to
-simulate animal movement. We thus engaged in fitting a movement model that was
-specifically intended to simulate dispersal events (red boxes). Using this
-model, we simulated 1000 dispersal events across the surroundings of our core
-study area and qualitatively assessed the outcome. Corresponding methods and
-results are reported in \Cref{Appendix:DispersalSimulation}. Furthermore, we
-motivated that data on the social landscape of dispersers would deepen our
-understanding of how dispersers are influenced by the presence or absence of
+simulate animal movement. We thus fitted a more detailed movement model and used
+it to simulate 1000 dispersal events across the surroundings of the core study
+area (red boxes in \Cref{SchematicOverview}). Corresponding methods and results
+are reported in \Cref{Appendix:DispersalSimulation}. In \Cref{Discussion}, we
+also motivated that data on the social landscape of dispersers would deepen our
+understanding of how dispersers respond to the presence or absence of
 conspecifics, apex predators, and prey. Although we lacked proper GPS relocation
-data of apex predators and prey, we collected a considerable amount of movement
-data from resident wild dogs residing in our core study area. These data cover
-the periods of almost all of our dispersal events and thus hold the potential
-for modeling a social landscape for our dispersers. Hence, we modeled such a
-landscape for the extent of our core study area and used it as additional
-covariate to parameterize a social model (yellow box). Corresponding methods and
-results are reported in \Cref{Appendix:SocialLandscape}.
+data of apex predators and prey, we collected a considerable amount of
+relocation data on resident wild dogs residing in the core study area. These
+data cover the periods of almost all of our dispersal events and thus hold the
+potential for modeling a social landscape for our dispersers. We attempted to
+model such a landscape for the core study area and used the resulting layer as
+additional covariate to parameterize a social model (yellow box in
+\Cref{SchematicOverview}). Corresponding methods and results are reported in
+\Cref{Appendix:SocialLandscape}.
 
@@
     highlight our main line of work which resulted in the identification of
-    least-cost paths and least-cost corridors. The red boxes show how we
-    additionally used the iSSF framework to specifically analyze movement
-    behavior and subsequently to simulate dispersal events. Finally, the yellow
-    box indicates our attempt to use the iSSF framework to model and analyze the
-    influence of the social landscape on dispersal behavior.}
+    least-cost paths and least-cost corridors. The red boxes show that we
+    additionally used the iSSF framework to analyze movement behavior in more
+    detail and to simulate dispersal events. Finally, the yellow box indicates
+    our attempt to use the iSSF framework to model and analyze the influence of
+    the social landscape on dispersal behavior.}
     \label{SchematicOverview}
@@
 dogs and to simulate dispersal events, we fitted a more complex movement model.
-The movement model still contained the most parsimonious permeability model, but
-in addition included interaction terms between movement metrics (i.e.
+The movement model still contained the most parsimonious habitat selection
+model, but in addition included interaction terms between movement metrics (i.e.
 \(cos(ta)\), \(log(sl\))) and environmental covariates (e.g. water). For
@@
 forward model selection based on AIC to identify the most parsimonious movement
-model. That is, we departed from the permeability model and iteratively added
-interactions with movement metrics. Note that we applied forward model selection
-only on the newly identified interactions, but not for the terms that were
-already in the permeability model. Since we attempted to keep our simulation
-simple, we did not average models with positive AIC-weights but simply
-identified the most parsimonious movement model.
+model. That is, we departed from the habitat selection model and iteratively
+added interactions with movement metrics. Note that we applied forward model
+selection only on the newly identified interactions, but not for the terms that
+were already in the habitat selection model. Since we attempted to keep our
+simulation simple, we did not average models with positive AIC-weights but
+simply identified the most parsimonious movement model.
 
 Using the most parsimonious movement model, we simulated 1000 dispersal events
-departing from our core study area. The simulation basically resembled an
+departing from the core study area. The simulation basically resembled an
 inverted iSSF function and was set up as follows. First, we sampled a random
-starting location within our core study area. Second, we assigned a random
+starting location within the core study area. Second, we assigned a random
 orientation in order to be able to calculate turning angles. Third, we created
@@
 200 steps and the decision did not influence our findings. Although we ran the
-simulation only for dispersers departing from our core study area, one could
+simulation only for dispersers departing from the core study area, one could
 easily expand the simulation to include multiple sources and thereby check how
@@
 \subsubsection{Results}
-Along with the original covariates contained in the permeability model, our most
-parsimonious movement model (see \Cref{ModelAICs2} and
+Along with the original covariates contained in the habitat selection model, our
+most parsimonious movement model (see \Cref{ModelAICs2} and
 \Cref{MovementModelResults}) retained three additional interactions (see red box
 in \Cref{MovementModel}a). Since the sign and direction of the effects contained
-in the permeability model did not markedly change, we only report on the newly
-added interactions between environmental covariates and movement metrics. For an
-easier interpretation of how these interactions influenced habitat selection and
-movement patterns during dispersal, we prepared surface plots as depicted in
-\Cref{MovementModel}b. These surface plots revealed that dispersers preferred
-long steps (i.e. a high movement rate) when water cover was low, whereas they
-preferred short steps (i.e. a low movement rate) when water cover was high (see
-\Cref{MovementModel}b1). Furthermore, dispersers were indifferent between high
-and low turning angles in proximity to water, but clearly preferred low turning
-angles (i.e. a high \(cos(ta)\)) when distance to water was high (see
-\Cref{MovementModel}b2). Finally, dispersers preferably realized steps with low
-turning angles when human influence was low, whereas they preferably realized
-low turning angles when human influence was high (see \Cref{MovementModel}b3).
+in the habitat selection model did not markedly change, we only report on the
+newly added interactions between environmental covariates and movement metrics.
+For an easier interpretation of how these interactions influenced habitat
+selection and movement patterns during dispersal, we prepared surface plots as
+depicted in \Cref{MovementModel}b. These surface plots revealed that dispersers
+preferred long steps (i.e. a high movement rate) when water cover was low,
+whereas they preferred short steps (i.e. a low movement rate) when water cover
+was high (see \Cref{MovementModel}b1). Furthermore, dispersers were indifferent
+between high and low turning angles in proximity to water, but clearly preferred
+low turning angles (i.e. a high \(cos(ta)\)) when distance to water was high
+(see \Cref{MovementModel}b2). Finally, dispersers preferably realized steps with
+low turning angles when human influence was low, whereas they preferably
+realized low turning angles when human influence was high (see
+\Cref{MovementModel}b3).
 
@@
     whereas darker areas correspond to avoided combinations. The absolute values
-    along the axes have no meaning as they were normalized to 0 (low) and 1
-    (high).}
+    along the axes have no meaning as original values were normalized to 0 (low)
+    and 1 (high).}
     \label{MovementModel}
@@
 trajectories were quite directional, avoided crossing a waterbody, ran along
-waterbodies and avoided human-dominated areas. More interestingly, however, the
+water bodies and avoided human-dominated areas. More interestingly, however, the
 resulting heatmap closely resembled our permeability surface (see
@@
     \caption{(a) Heatmap of 1000 simulated dispersal events departing from the
-    core study area (white square) in northern Botswana (countries with dashed
-    white lines). (b) Permeability map for the same extent, showing that
-    simulated dispersal trajectories typically cover highly permeable
+    core study area (white rectangle) in northern Botswana (countries with
+    dashed white lines). (b) Permeability map for the same extent to illustrate
+    that simulated dispersal trajectories typically cover highly permeable
     landscapes.}
@@
   e^{-0.5} \Delta \text{AIC}_i} \). (*) represents all covariates from the most
-  parsimonious permeability model, i.e. \(cos(ta)\), \(log(sl\)),
+  parsimonious habitat selection model, i.e. \(cos(ta)\), \(log(sl\)),
   \textit{Water}, \textit{DistanceToWater}, \textit{Shrubs/Grassland},
@@
   \caption{Selection coefficients as derived from the most parsimonious movement
-  model. The movement model was a modified permeability model to which we added
-  additional interaction terms between movement metrics and environmental
+  model. The movement model was a modified habitat selection model to which we
+  added additional interaction terms between movement metrics and environmental
   covariates (e.g. \(Water:cost(ta\_)\)) in order to learn about how movement
@@
 collected between January 1\textsuperscript{st} and June 30\textsuperscript{th}.
-Because we collected relocation data of residents only inside our core study
+Because we collected relocation data of residents only inside the core study
 area, we modeled the social landscape only for this region. GPS relocations were
@@
 stratum that did not fully lie within the core study area. The social model that
-we fitted was again a modification of the most parsimonious permeability model.
-Despite the covariates contained in the permeability model, we added the social
-covariates listed in \Cref{SocialCovars}. Because we hypothesized that the
-preference for encountering other wild dogs was different within and outside the
-disperser's natal home range, we included the interaction terms
+we fitted was again a modification of the most parsimonious habitat selection
+model. Despite the covariates contained in the habitat selection model, we added
+the social covariates listed in \Cref{SocialCovars}. Because we hypothesized
+that the preference for encountering other wild dogs was different within and
+outside the disperser's natal home range, we included the interaction terms
 \textit{ProbEncounter:OwnHomerange} and \textit{ProbEncounter:ForeignHomeranges}
@@
       Social Features
-        & Own Homerange
+        & OwnHomerange
           & Percentage cover of natal pack’s 95\% home-range along step
@@
 \Cref{SocialModelResults}). Again, parameter estimates for the covariates
-contained in the permeability model did not severely change and we will only
-report on the newly added social covariates (see red box in
+contained in the habitat selection model did not severely change and we will
+only report on the newly added social covariates (see red box in
 \Cref{SocialModel}a). Visual inspection of the corresponding surface plot (see
@@
     covariates, whereas darker areas correspond to avoided combinations. The
-    absolute values along the axes of the surface plots have no meaning as they
-    were scaled. Rather, low values along the axis correspond to low values of
-    the respective covariate. The same is true for the scalebar.}
+    absolute values along the axes have no meaning as original values were
+    normalized to 0 (low) and 1 (high).}
     \label{SocialModel}
@@
   \caption{Selection coefficients as derived from the social model. The social
-  model was a modified permeability model to which we added additional social
-  covariates in order to learn about how the social landscape infleunces
+  model was a modified habitat selection model to which we added additional
+  social covariates in order to learn about how the social landscape infleunces
   dispersal behavior.}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-04-17.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-02-04.tex
@@
 predict landscape permeability based on recorded GPS relocation data include
-\textit{resource selection functions} (RSF, \cite{Boyce.2002}), \textit{step
-selection functions} (SSF, \cite{Fortin.2005}), and \textit{path selection
-functions} (PSF, \cite{Cushman.2010}). All methods have in common that they
+\textit{resource selection functions} (RSF, \citealp{Boyce.2002}), \textit{step
+selection functions} (SSF, \citealp{Fortin.2005}), and \textit{path selection
+functions} (PSF, \citealp{Cushman.2010}). All methods have in common that they
 infer habitat preferences of the focal species by comparing spatial covariates
@@
 The \textit{extended study area} (centered at -17\degree 13'9''S, 23\degree
-56'4''E; elevation ca. 1,030 m, red rectangle in \Cref{StudyArea}a) was confined
+56'4''E; elevation ca. 1'030 m, red rectangle in \Cref{StudyArea}a) was confined
 by a bounding box encompassing the entire KAZA-TFCA. The KAZA-TFCA lies in the
@@
 algorithm uses thresholding of SWIR reflectances of MODIS Terra satellite
-imagery (MCD43A4, \cite{Schaaf.2015}) to separate water and dryland. As we
+imagery (MCD43A4, \citealp{Schaaf.2015}) to separate water and dryland. As we
 rewrote and slightly amended the original Python code in R, we describe the
@@
 derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
-\cite{Dimiceli.2015}). Combined, the three layers added up to 100\% and depicted
-the percentage cover of tree-vegetation (henceforth trees), non-tree-vegetation
-(henceforth shrubs/grassland), and non-vegetated (henceforth bare land). To
-ensure that areas covered by the flood would get 0\% vegetation cover (i.e.
-100\% bare land cover), we used our floodmap that aligned with the creation date
-of the layers (30 March 2017) and manually defined pixels covered by the flood
-as being 100\% bare land. The MODIS vegetation layers had a resolution of 250m x
-250m, so we did not need to coarsen or interpolate them.
+\citealp{Dimiceli.2015}). Combined, the three layers added up to 100\% and
+depicted the percentage cover of tree-vegetation (henceforth trees),
+non-tree-vegetation (henceforth shrubs/grassland), and non-vegetated (henceforth
+bare land). To ensure that areas covered by the flood would get 0\% vegetation
+cover (i.e. 100\% bare land cover), we used our floodmap that aligned with the
+creation date of the layers (30 March 2017) and manually defined pixels covered
+by the flood as being 100\% bare land. The MODIS vegetation layers had a
+resolution of 250m x 250m, so we did not need to coarsen or interpolate them.
 
@@
 % Explanation of iSSF and Extraction
-We used an integrated step selection function approach (iSSF, \cite{Avgar.2016})
-to investigate dispersers' selection or avoidance for the above mentioned
-spatial covariates. In the iSSF framework, covariates experienced along realized
-steps are contrasted with covariates experienced along alternative random steps
-that the animal could have taken, but decided not to. A step is here defined as
-the connecting line between two consecutive GPS relocations
-\citep{Turchin.1998}. In contrast to regular SSFs, iSSFs require to include
-movement metrics as covariates in the corresponding conditional logistic
+We used an integrated step selection function approach (iSSF,
+\citealp{Avgar.2016}) to investigate dispersers' selection or avoidance for the
+above mentioned spatial covariates. In the iSSF framework, covariates
+experienced along realized steps are contrasted with covariates experienced
+along alternative random steps that the animal could have taken, but decided not
+to. A step is here defined as the connecting line between two consecutive GPS
+relocations \citep{Turchin.1998}. In contrast to regular SSFs, iSSFs require to
+include movement metrics as covariates in the corresponding conditional logistic
 regression model. Their inclusion, in turn, allows simultaneous inference on
@@
 lengths from a gamma distribution that was fitted using realized step lengths
-(shape = 0.3677, scale = 6,302)). Together, a realized and its 24 associated
+(shape = 0.3677, scale = 6'302). Together, a realized and its 24 associated
 random steps formed a stratum of 25 steps that received a unique identifier.
@@
 Coefficient. None of the covariates were overly correlated (\(|r| > 0.6\),
-\cite{Latham.2011}) and we retained all of them for modeling. To run a proper
+\citealp{Latham.2011}) and we retained all of them for modeling. To run a proper
 iSSF analysis, we also included two movement metrics, namely the cosine of the
@@
 variance of the random intercept to an artificially high value of
-10\textsuperscript{12} (for an explanation, see \cite{Muff.2019}) and used
+10\textsuperscript{12} (for an explanation, see \citealp{Muff.2019}) and used
 animal ID to model random intercepts and slopes.
@@
 and ran forward model selection based on Akaike's Information Criterion (AIC,
-\cite{Burnham.2002}) for all other spatial covariates. That is, we started with
-a basic model and iteratively increased model complexity upon reaching a
+\citealp{Burnham.2002}) for all other spatial covariates. That is, we started
+with a basic model and iteratively increased model complexity upon reaching a
 full model. We then ranked models according to AIC, assessed relative model
@@
 areas larger than 700 km\textsuperscript{2} that did not get any source points
-from the regular grid. In total, we generated 68 points, which resulted in 2,278
-unique pairwise combinations and therefore 2,278 unique LCPs and LCCs.
+from the regular grid. In total, we generated 68 points, which resulted in 2'278
+unique pairwise combinations and therefore 2'278 unique LCPs and LCCs.
 
@@
 All of this work would not have been possible without the aid of several helpful
-people. I would first like to thank Prof. Arpat Ozgul, the leader and mastermind
-of our research group, whose expertise in statistics, biological processes and
-scientific writing was invaluable for my work. Moreover, I want to thank
-Gabriele Cozzi for his excellent guidance, all the interesting discussions, his
-constructive criticism, valuable revisions in the manuscript, and of course his
-constant supply of freshly brewed coffee. Particular gratitude goes to Dominik
-Behr, who supervised my work daily and greatly contributed to all aspects of the
-thesis. I genuinely enjoyed the frequent meetings and brainstorms and I feel
-privileged to have had the opportunity of working with such a great scientist. I
-am also greatly indebted to Prof. John Fieberg, who visited our department this
-summer and invested a great amount of time to consult and improve all
-statistical aspects of the thesis, as well as to deepen my understanding of
-movement ecology. Furthermore, I thank the Okavango Research Institute,
-especially Piotr Wolski, for sharing all of their Python scripts, floodmaps, and
-knowledge on developing dynamic floodmaps. Also, I want to thank Stefan Sommer
-for sharing and teaching his hard-earned knowledge on scientific writing.
+people and organizations. I would first like to thank Prof. Arpat Ozgul, the
+leader and mastermind of our research group, whose expertise in statistics,
+biological processes and scientific writing was invaluable for my work.
+Moreover, I want to thank Gabriele Cozzi for his excellent guidance, all the
+interesting discussions, his constructive criticism, valuable revisions in the
+manuscript, and of course his constant supply of freshly brewed coffee.
+Particular gratitude goes to Dominik Behr, who supervised my work daily and
+greatly contributed to all aspects of the thesis. I genuinely enjoyed the
+frequent meetings and brainstorms and I feel privileged to have had the
+opportunity of working with such a great scientist. I am also greatly indebted
+to Prof. John Fieberg, who visited our department in summer and invested a great
+amount of time to consult and improve all statistical aspects of the thesis, as
+well as to deepen my understanding of movement ecology. Further, I want to thank
+the Okavango Research Institute, especially Piotr Wolski, for sharing all of
+their Python scripts, floodmaps, and knowledge on developing dynamic floodmaps.
+I also thank all organizations that enabled this project through funds, namely
+Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea Wild, Jacot
+Foundation, National Geographic Society, Parrotia Stiftung, Stiftung Temperatio,
+Wilderness Wildlife Trust Foundation, Forschungskredit der Universität Zürich,
+and a Swiss National Foundation Grant (31003A\_182286) to Prof. Arpat Ozgul. I
+thank the Ministry of Environment and Tourism of Botswana for granting
+permission to conduct this research (Research Permit EWT 8/36/4 XXXVI [33]).
 Additionally, I want to thank all of my colleague master students who have been
@@
 and pre-process all relatively cloud-free MODIS Terra images (MCD43A4,
-\cite{Schaaf.2015}) available for the period of our dispersal events. Assessment
-of cloud cover was based on visual inspection of MODIS images on ORI's website
-\citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful for dynamic
-floodmapping because it is updated with new composite images every 8th day
-\citep{Wolski.2017}. After download, we classified each MODIS image into a
-binary map of water (flood) and dryland using a threshold that was identified as
-follows. First, we extracted all reflectance values of MODIS Terra Band 7 within
-the water- and dryland-polygons. Second, we computed histograms of
-water-reflectances and dryland-reflectances and empirically verified that
-reflectances of the two groups were sufficiently distinct. More specifically, we
-checked if superimposing the two histograms resulted in a bimodal plot. This was
-said to be achieved if the 99\textsuperscript{th} percentile of
-water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
-of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
-dryland}\)). Third, if peaks were sufficiently different, we calculated a
-threshold (\(t\)) using \Cref{EQ1}:
+\citealp{Schaaf.2015}) available for the period of our dispersal events.
+Assessment of cloud cover was based on visual inspection of MODIS images on
+ORI's website \citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful
+for dynamic floodmapping because it is updated with new composite images every
+8\textsuperscript{th} day \citep{Wolski.2017}. After download, we classified
+each MODIS image into a binary map of water (flood) and dryland using a
+threshold that was identified as follows. First, we extracted all reflectance
+values of MODIS Terra Band 7 within the water- and dryland-polygons. Second, we
+computed histograms of water-reflectances and dryland-reflectances and
+empirically verified that reflectances of the two groups were sufficiently
+distinct. More specifically, we checked if superimposing the two histograms
+resulted in a bimodal plot. This was said to be achieved if the
+99\textsuperscript{th} percentile of water-reflectances did not severely exceed
+the 1\textsuperscript{st} percentile of dryland-reflectances (\(p_{0.99, water}
+- \frac{10}{255} < p_{0.01, dryland}\)). Third, if peaks were sufficiently
+different, we calculated a threshold (\(t\)) using \Cref{EQ1}:
 
@@
 interactions served to learn how dispersers moved differently, depending on
-environmental covariates (for comparable work see \cite{Prokopenko.2017}). We
+environmental covariates (for comparable work see \citealp{Prokopenko.2017}). We
 applied the same modeling framework as explained in \Cref{Modeling} and ran

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-04-27.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-04-17.tex
@@
 \usepackage{varwidth}
-% \usepackage[nomarkers,figuresonly]{endfloat}
 \usepackage{csvsimple}
@@
 \usepackage{titlesec}
+\usepackage{float}
 
@@
         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-        \includegraphics[width=0.9\textwidth]{99_StudyArea2.pdf}
+        \includegraphics[width=0.9\textwidth]{99_StudyArea.pdf}
         };
@@
     \end{tikzpicture}
-    \caption{Overview of our study area. (a) The red rectangle depicts our
-    \textit{extended study area}, which was confined by the extent of the
-    KAZA-TFCA. Purple areas indicate remaining wild dog populations according to
-    the IUCN \citep{Woodroffe.2012}. (b) The orange rectangle illustrates our
-    \textit{core study area}, from which all dispersers departed. Yellow lines
-    depict all dispersal trajectories that were recorded between the years 2011
-    and 2019. Since game reserves in Botswana virtually serve the same purpose as
-    National Parks, we will use the terms interchangeably for this region.}
+    \caption{Overview of our study area. (a) The red rectangle depicts the study
+    area, which was confined by a bounding box encompassing the entire
+    KAZA-TFCA. Pink areas indicate remaining wild dog populations according to
+    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
+    area within which dispersal coalitions were collared. Since game reserves in
+    Botswana virtually serve the same purpose as National Parks, we use the
+    terms interchangeably for this region.}
     \label{StudyArea}
@@
 In total, we collected 4'691 GPS relocations during dispersal (see
-\Cref{Trajectories}). In our analysis, we did not differentiate between male and
-female dispersing coalitions, for previous research found little differences
-between sexes during dispersal (\citeauthor{Woodroffe.2019}, 2019;
-\citeauthor{Cozzi.2020}, unpublished). On average, dispersal coalitions included
-three individuals (min = 1, max = 7), dispersed for 41 days (min = 7, max =
-133), and covered a mean cumulative distance of 370 km (min = 105, max = 610).
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_Trajectories2.pdf}
-    \caption{Illustration of the dispersal trajectories that we recorded. Each
-    color represents a different dispersal coalition. All coalitions departed
-    from the core study area, which is indicated by the white rectangle. The
-    coalition dispersing towards the far east of the map covered over 360 km in
-    under 10 days.}
-    \label{Trajectories}
-  \end{center}
-\end{figure}
+\Cref{Trajectories} in \Cref{Appendix:GPSData}). In our analysis, we did not
+differentiate between male and female dispersing coalitions, for previous
+research found little differences between sexes during dispersal
+(\citeauthor{Woodroffe.2019}, 2019; \citeauthor{Cozzi.2020}, unpublished). On
+average, dispersal coalitions included three individuals (min = 1, max = 7),
+dispersed for 41 days (min = 7, max = 133), and covered a mean cumulative
+distance of 370 km (min = 105, max = 610).
 
@@
     \includegraphics[width = \textwidth]{99_Covariates.pdf}
-    \caption{Overview of covariates that we used in our models. Although all
-    covariates were prepared for the extended study area, we present only an
-    extract to maximize visibility. The white rectangle in each plot depicts our
-    core study area. (a) Averaged layer of all binary maps of water and dryland
-    that we acquired using our floodmapping algorithm. (b) Percentage cover of
-    trees. (c) Percentage cover of shrubs/grassland. Note that anything not
-    covered by trees or shrubs/grassland was deemed to be bare land. (d)
-    Protection status of the area. (e) Human influence proxy, taking into
-    account human density, farms, and roads. (f) Distance to nearest road (white
-    lines depict actual roads).}
+    \caption{Overview of spatial covariates that we included in our models. We
+    prepared all covariates for the entire study area but for better visibility
+    we only plot them for the surroundings of the Okavango Delta. The white
+    rectangle in each plot depicts the area within which dispersal coalitions
+    were collared. (a) Averaged layer of all dynamic (binary) watermaps. (b)
+    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
+    Anything that was not covered by trees or shrubs/grassland was deemed to be
+    bare land. (d) Protection status of the area. (e) Human influence proxy
+    composed of human density, farms, and roads. (f) Distance to nearest road
+    (white lines depict actual roads).}
     \label{Covariates}
@@
 We extracted spatial covariates along realized and random steps as listed in
-\Cref{ExtractedCovars}. For continuous covariates, we calculated the average
+\Cref{Appendix:Sources}. For continuous covariates, we calculated the average
 value, for binary covariates the percentage cover along the step. Since our
@@
 movement rate.
-
-\begin{table}[h]
-  \begin{center}
-    \caption{Overview of spatial covariates extracted along steps (see
-    \Cref{Sources} in \Cref{Appendix:Sources} for an overview of the sources and
-    resampling methods). *The covariates \textit{Water} and \textit{Dryland}
-    added up to 100\%, which is why only \textit{Water} was included as
-    explanatory variable in our models. The same applied for the group
-    \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland}, where we
-    omitted \textit{Bareland} for modeling. Finally, from the group
-    \textit{Protected} and \textit{Pastoral}, we only included
-    \textit{Protected} in our models.}
-    \label{ExtractedCovars}
-    \resizebox{\textwidth}{!} {
-      \begin{tabular}{lllll}
-      \hline
-      \textbf{Category} &
-        \textbf{Covariate} &
-          \textbf{Description} &
-            \textbf{Values} \\
-              % \textbf{Source(s)} \\
-      \midrule
-      \multirow{5}{*}{Land Cover}
-        & Water
-          & Percentage cover of water along step
-            & 0-100\%\\
-              % & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
-        & Dryland*
-          & Percentage cover of dryland along step
-            & 0-100\%\\
-              % & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
-        & DistanceToWater
-          & Average distance of step to nearest water source
-            & \(\geq\) 0m\\
-              % & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
-        & Shrubs/Grassland
-          & Average non-tree vegetation along step
-            & 0-100\%\\
-              % & \cite{Dimiceli.2015} \\
-        & Trees
-          & Average tree-vegetation along step
-            & 0-100\%\\
-              % & \cite{Dimiceli.2015} \\
-        & Bareland*
-          & Average non-vegetated area along step
-            & 0-100\%\\
-              % & \cite{Dimiceli.2015} \\
-      \hdashline
-      \multirow{2}{*}{Protection Status}
-        & Protected
-          & Percentage cover of protected area along step
-            & 0-100\%\\
-              % & \cite{PeaceParks.2019} \\
-        & Pastoral*
-          & Percentage cover of unprotected area along step
-            & 0-100\%\\
-              % & \cite{PeaceParks.2019} \\
-      \hdashline
-      \multirow{3}{*}{Human Influence}
-        & HumansBuffer5000
-          & Average buffered (5 km) human influence along step
-            & \(\geq\) 0\\
-              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
-        & DistanceToRoads
-          & Average distance of step to nearest road
-            & \(\geq\) 0m\\
-              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
-        & RoadCrossing
-          & Binary value indicating whether a step crossed a road
-            & 0, 1\\
-              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
-      \hline
-      \end{tabular}
-    }
-  \end{center}
-\end{table}
 
@@
 
-% \begin{figure}[hbtp]
-%   \begin{center}
-%     \begin{tikzpicture}
-%         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-%           \begin{minipage}{0.54\textwidth}
-%             \begin{center}
-%               \includegraphics[width = 0.5\textwidth]
-%                 {99_LeastCostPaths(Example1).pdf}\\
-%               \includegraphics[width = 1.0\textwidth]
-%                 {99_LeastCostPaths(Example2).pdf}\\
-%             \end{center}
-%           \end{minipage}
-%         };
-%         \begin{scope}[x={(image.south east)},y={(image.north west)}]
-%            %  % next four lines will help you to locate the point needed by
-%            %  % forming a grid. comment these four lines in the final picture.↓
-%            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
-%            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
-%            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
-%            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
-%            %  % upto here
-%             % \draw[black, line width = 0.5pt] (0.389, 0.627) -- (0.080, 0.477);
-%             % \draw[black, line width = 0.5pt] (0.582, 0.627) -- (0.980, 0.477);
-%             % \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
-%             % \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
-%             % \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
-%         \end{scope}
-%     \end{tikzpicture}
-%     \caption{Figures illustrating the process of identifying a least-cost
-%     corridor between the source points A and B. (a) Example of a permeability
-%     surface, which serves to determine the costs of movement. (b1) Cumulative
-%     cost map for point A, depicting the total minimal costs necessary to get
-%     from point A to every other location. (b2) Cumulative cost map for point B,
-%     depicting the total minimal costs necessary to get from point B to every
-%     other location. (c) Summed cost maps of points A and B. (d) Masked out
-%     corridor containing pixels that do not exceed the cheapest pixel by more
-%     than 5\%.}
-%     \label{LCCExample}
-%   \end{center}
-% \end{figure}
-
-\begin{figure}[hbtp]
-  \begin{center}
-  \begin{minipage}{.25\linewidth}
-    \begin{subfigure}[t]{.9\linewidth}
-        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example1).pdf}
-    \end{subfigure}
-  \end{minipage}
-  \begin{minipage}{.25\linewidth}
-    \begin{subfigure}[t]{.9\linewidth}
-        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example2).pdf}
-    \end{subfigure}
-  \end{minipage}
-  \begin{minipage}{.25\linewidth}
-    \begin{subfigure}[t]{.9\linewidth}
-        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example3).pdf}
-    \end{subfigure}
-  \end{minipage}
-  \caption{Figures illustrating the process of identifying a least-cost corridor
-  between source points A and B following \cite{Pinto.2009}. (a) Example of a
-  permeability surface, which determines the costs of movement. (b1) Cumulative
-  cost map for point A, depicting the total minimal costs necessary to get from
-  point A to every other location. (b2) Cumulative cost map for point B,
-  depicting the total minimal costs to get from point B to every other location.
-  (c) Summed cost maps of points A and B. (d) Masked out corridor containing
-  pixels that do not exceed the cheapest pixel by more than 5\%.}
-  \label{LCCExample}
-  \end{center}
-\end{figure}
-
 \newpage
@@
     covariate, positive coefficients selection of a covariate. Whiskers
-    delineate the 95\%-CIs for estimated parameters. The intercept is
-    meaningless in the iSSF framework and was thus omitted. (b) Results from the
-    k-fold cross validation for case-control studies. The left graph shows rank
+    delineate the 95\%-CIs for estimated parameters. (b) Results from the k-fold
+    cross validation for case-control studies. The left graph shows rank
     frequencies of \textit{realized} steps according to predictions, whereas the
@@
     coefficient resulting from 100 repetitions of the k-fold cross validation.
-    The blue smoothing line was fitted using a LOESS regression and solely
-    serves to aid the eye in detecting the trends. Correlation coefficients
-    suggest that our prediction is significant and robust, evidenced by the fact
-    that the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-    random}\) do not overlap and that there is strong and significant
-    correlation between ranks and associated frequency for realized steps.}
+    The blue smoothing line was fitted using a locally weighted polynomial
+    regression and serves to aid the eye in detecting the trends. Correlation
+    coefficients suggest that our prediction was significant and robust,
+    evidenced by the fact that the confidence intervals of \(\bar{r}_{s,
+    realized}\) and \(\bar{r}_{s, random}\) did overlap and by the fact that
+    there was strong and significant correlation between ranks and associated
+    frequency for realized steps.}
     \label{PermeabilityResults}
@@
 
-% \newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
-% \begin{figure}[hbtp]
-%   \begin{center}
-%     \begin{tikzpicture}
-%         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-%           \begin{minipage}{0.95\textwidth}
-%             \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
-%             \includegraphics[width = \textwidth]{99_PermeabilityMap2.pdf}
-%           \end{minipage}
-%         };
-%         \begin{scope}[x={(image.south east)},y={(image.north west)}]
-%            %  % next four lines will help you to locate the point needed by
-%            %  % forming a grid. comment these four lines in the final picture.↓
-%            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
-%            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
-%            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
-%            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
-%            %  % upto here
-%             \draw[black, line width = 1pt] (0.389, 0.627) -- (0.080, 0.477);
-%             \draw[black, line width = 1pt] (0.582, 0.627) -- (0.980, 0.477);
-%             \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
-%             \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
-%             \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
-%         \end{scope}
-%     \end{tikzpicture}
-%     \caption{(a) Predicted permeability surface for the extent of the KAZA-TFCA.
-%     Permability was predicted by calculating selection scores \(w(x)\) for each
-%     pixel based on the pixel's underlying covariates. Areas that dispersers find
-%     easy to traverse are depicted in bright colors. Bold white lines delineate
-%     the borders of the KAZA-TFCA, whereas dashed white lines show country
-%     borders. (b) Detailed view of the Okavango Delta and Linyanti Swamp region,
-%     which appears to provide the most suitable landscape for dispersers. The
-%     white square highlights the core study area.}
-%     \label{PermeabilityMap}
-%   \end{center}
-% \end{figure}
-% \restoregeometry
-
-% \newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
 \begin{figure}[hbtp]
@@
 \end{figure}
-% \restoregeometry
-
-% \begin{table}[h]
-%   \caption{Comparison of median permeability across countries, separated
-%   according to areas within and outside the KAZA-TFCA, as well as within and
-%   outside protected areas. High values indicate high permeability, whereas low
-%   values correspond to low permeability.}
-%   \label{PermeabilityComp}
-%   \begin{center}
-%     \resizebox{\textwidth}{!} {
-%       \begin{tabular}{llllllllr}
-%       & & \multicolumn{2}{c}{KAZA-TFCA} && \multicolumn{2}{c}{Protection Status} \\
-%       \cline{3-4} \cline{6-7}
-%       Country & & Inside & Outside & & Protected & Pastoral & & Overall \\
-%       \hline
-%       Angola & & 1.67 & 0.59 & & 1.67 & 0.60 & & 1.03 \\
-%       Namibia & & 0.91 & 0.49 & & 0.98 & 0.44 & & 0.59 \\
-%       Botswana & & 1.17 & 0.69 & & 1.37 & 0.62 & & 0.91 \\
-%       Zimbabwe & & 0.25 & 0.18 & & 0.32 & 0.17 & & 0.19 \\
-%       Zambia & & 0.26 & 0.12 & & 0.23 & 0.13 & & 0.15 \\
-%       \hline
-%       Overall & & 0.75 & 0.26 & & 0.74 & 0.27 & & 0.38 \\
-%       \end{tabular}
-%     }
-%   \end{center}
-% \end{table}
 
@@
   \label{PermeabilityComp}
-  \begin{center}
-    \resizebox{0.9\textwidth}{!} {
-      \begin{tabular}{llllllllr}
-      & & \multicolumn{2}{c}{KAZA-TFCA} && \multicolumn{2}{c}{Protection Status} \\
-      \cline{3-4} \cline{6-7}
-      Country & & Inside & Outside & & Protected & Pastoral & & Overall \\
+  \centering
+  \begin{tabular}{llllll}
+    \toprule
+    \multicolumn{1}{c}{ } & \multicolumn{2}{c}{KAZA-TFCA} & \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{ } \\
+    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
+    Country & Inside & Outside & Protected & Pastoral & Overall\\
+    \midrule
+    Angola & 0.41 (0.42) & 0.18 (0.35) & 0.41 (0.42) & 0.19 (0.35) & 0.26 (0.40)\\
+    Botswana & 0.27 (0.30) & 0.15 (0.17) & 0.29 (0.35) & 0.16 (0.19) & 0.20 (0.26)\\
+    Namibia & 0.23 (0.30) & 0.13 (0.18) & 0.24 (0.30) & 0.11 (0.15) & 0.16 (0.24)\\
+    Zambia & 0.08 (0.11) & 0.04 (0.06) & 0.07 (0.12) & 0.04 (0.06) & 0.05 (0.08)\\
+    Zimbabwe & 0.08 (0.20) & 0.06 (0.05) & 0.10 (0.21) & 0.06 (0.04) & 0.06 (0.08)\\
     \hline
-      Angola & & 0.46 (0.42) & 0.18 (0.40) & & 0.46 (0.42) & 0.18 (0.40) & & 0.29 (0.45) \\
-      Namibia & & 0.26 (0.39) & 0.13 (0.20) & & 0.27 (0.38) & 0.12 (0.16) & & 0.16 (0.31) \\
-      Botswana & & 0.33 (0.38) & 0.19 (0.23) & & 0.38 (0.35) & 0.17 (0.23) & & 0.25 (0.33) \\
-      Zimbabwe & & 0.06 (0.25) & 0.05 (0.04) & & 0.09 (0.26) & 0.04 (0.03) & & 0.05 (0.07) \\
-      Zambia & & 0.07 (0.11) & 0.03 (0.05) & & 0.06 (0.11) & 0.03 (0.05) & & 0.04 (0.07) \\
-      \hline
-      Overall & & 0.21 (0.39) & 0.07 (0.18) & & 0.21 (0.39) & 0.07 (0.18) & & 0.11 (0.29) \\
+    Overall & 0.19 (0.32) & 0.08 (0.16) & 0.18 (0.32) & 0.08 (0.17) & 0.11 (0.23)\\
+    \bottomrule
   \end{tabular}
-    }
-  \end{center}
 \end{table}
@@
 \section{Appendix}
+\subsection{Net Squared Displacement}
+\label{Appendix:GPSData}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width = 0.8\textwidth]{99_NSD.pdf}
+    \caption{NSD displacement graph for one of our dispersers. Blue colors
+    indicate the period during which we calssified the individual as
+    dispersing.}
+    \label{NSD}
+  \end{center}
+\end{figure}
+
+\newpage
+\subsection{GPS Data}
+\label{Appendix:GPSData}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width = 0.9\textwidth]{99_Trajectories.pdf}
+    \caption{Illustration of the dispersal trajectories that we recorded. Each
+    color represents a different dispersing coalition. All coalitions departed
+    from the core study area, which is indicated by the white rectangle. The
+    coalition dispersing towards the far east of the map covered over 360 km in
+    under 10 days.}
+    \label{Trajectories}
+  \end{center}
+\end{figure}
+
+\begin{table}[h]
+  \caption{Summary statistics of all GPS relocations that have been recorded on
+  dispersing coalitions.}
+  \label{GPSData}
+  \begin{center}
+    \resizebox{\textwidth}{!}{
+      \begin{tabular}{lccccccc}
+      \toprule
+      \parbox[]{2cm}{\centering Coalition ID} &
+        \parbox[]{2cm}{\centering Sex} &
+          \parbox[]{2cm}{\centering \ Pack \\ Affiliation} &
+            \parbox[]{2cm}{\centering \# Fixes \\ Total} &
+              \parbox[]{3cm}{\centering \# Fixes \\ During Dispersal} &
+                \parbox[]{2cm}{\centering \# Days \\ Dispersing} &
+                  \parbox[]{4cm}{\centering Euclidean Dispersal\\ Distance (in km)} &
+                    \parbox[]{4cm}{\centering Cumulative Dispersal \\ Distance (in km)} \\
+      \midrule
+      Amacuro & F & MB & 954 & 583 & 137 & 23 & 1'090 \\
+      Belgium & M & ZU & 1'097 & 158 & 28 & 18 & 319 \\
+      Dalwhinnie & F & PA & 545 & 62 & 22 & 50 & 243 \\
+      Denali & F & MN & 1'096 & 173 & 33 & 11 & 528 \\
+      Everest & M & MN & 389 & 123 & 38 & 67 & 572 \\
+      Kalahari & F & HT & 1'753 & 467 & 130 & 20 & 1'963 \\
+      Karisimbi & M & MN & 438 & 141 & 34 & 45 & 251 \\
+      Liuwa & F & AP & 946 & 92 & 19 & 144 & 451 \\
+      Lupe & M & KW & 2'209 & 396 & 34 & 8 & 436 \\
+      MadameChing & F & AP & 776 & 729 & 136 & 263 & 1'560 \\
+      Mirage & M & HT & 814 & 182 & 36 & 7 & 435 \\
+      Odzala & M & AP & 1'410 & 205 & 42 & 53 & 412 \\
+      Scorpion & M & KB & 2'676 & 393 & 34 & 4 & 471 \\
+      Stetson & M & MT & 384 & 383 & 33 & 3 & 481 \\
+      Taryn & F & AP & 896 & 37 & 9 & 10 & 130 \\
+      \bottomrule
+      \end{tabular}
+    }
+  \end{center}
+\end{table}
+\newpage
+
+\newpage
 \subsection{Floodmapping Algorithm}
@@
       \toprule
-      \textbf{Original Category} &
-        \textbf{Reclassification} &
-          \textbf{Comment} \\
+      Original Category &
+        Reclassification &
+          Comment \\
       \midrule
@@
       \toprule
-      \textbf{Group} &
-        \textbf{Subgroup} &
-          \textbf{Description} \\
+      Group &
+        Subgroup &
+          Description \\
       \midrule
@@
 \newpage
-\subsection{Covariate Sources}
-\label{Appendix:Sources}
+\subsection{Human Influence Proxy}
+\label{Appendix:HumanInfluence}
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_HumanInfluence.pdf}
+    \caption{Figure that exemplifies how we combined the layers (a), (b), and
+    (c) into a single lhuman influence proxy (d). For ease of visibility we show
+    the procedure only for the extent of the Okavango Delta. Layer (a) is based
+    on Facebooks high resolution human density dataset and depicts the estimated
+    number of humans living in each 250m x 250m raster-cell. Layer (b) is a
+    binary layer and shows which raster-cells are currently used for any sort of
+    agricultural activity. Corresponding data was obtained through the Globeland
+    and Cropland land cover datasets. Layer (c) shows the presence or absence of
+    roads and is based on data from Open Street Map. We merged layers (a), (b),
+    and (c) by adding up their values, truncating the summed value to a maximum
+    of 50, log-transforming this value, and applying a focal buffer of 5km to
+    each pixel. Layer (d) depicts the final human influence layer that was used
+    as covariate in our models.}
+    \label{Trajectories}
+  \end{center}
+\end{figure}
+
+\newpage
+\subsection{Covariates \& Sources}
+
 \begin{table}[h]
   \begin{center}
-    \caption{Sources from which we obtained our set of geo-referenced
-    covariates, as well as an indicator of their original resolution and the
-    method used for adjusting the resolution to 250m x 250m.}
-    \label{Sources}
+    \caption{Overview of spatial covariates extracted along steps and their
+    respectives sources. *The covariates \textit{Water} and \textit{Dryland}
+    added up to 100\%, which is why only \textit{Water} was included as
+    explanatory variable in our models. The same applied for the group
+    \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland}, where we
+    omitted \textit{Bareland} for modeling. Finally, from the group
+    \textit{Protected} and \textit{Pastoral}, we only included
+    \textit{Protected} in our models.}
+    \label{Appendix:Sources}
     \resizebox{\textwidth}{!} {
-      \begin{tabular}{llp{4cm}p{4cm}p{7cm}}
+      \begin{tabular}{lllll}
       \hline
-      \textbf{Category} &
-        \textbf{Covariate} &
-          \textbf{Resolution} &
-            \textbf{Resampling} &
-              \textbf{Source(s)} \\
+      Category &
+        Covariate &
+          Description &
+            Values &
+              Source(s) \\
       \midrule
-      Land Cover
+      \multirow{5}{*}{Land Cover}
         & Water
-          & 30m x 30m (static), 500m x 500m (dynamic)
-            & Mode of each 250m x 250m cell (static), disaggregation to 250m x 250m (dynamic)
+          & Percentage cover of water
+            & 0-100\%
               & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
-      Land Cover
-        & Dryland
-          & 30m x 30m (static), 500m x 500m (dynamic)
-            & Mode of each 250m x 250m cell (static), disaggregation to 250m x 250m (dynamic)
-              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
-      Land Cover
+        & Dryland*
+          & Percentage cover of dryland
+            & 0-100\%
+              & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
         & DistanceToWater
-          & 30m x 30m (static), 500m x 500m (dynamic)
-            & Mode of each 250m x 250m cell (static), disaggregation to 250m x 250m (dynamic)
-              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
-      Land Cover
+          & Average distance to nearest water source
+            & \(\geq\) 0m
+              & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
         & Shrubs/Grassland
-          & 250m x 250m
-            & NA
+          & Average non-tree vegetation
+            & 0-100\%
               & \cite{Dimiceli.2015} \\
-      Land Cover
         & Trees
-          & 250m x 250m
-            & NA
+          & Average tree-vegetation
+            & 0-100\%
               & \cite{Dimiceli.2015} \\
-      Land Cover
-        & Bareland
-          & 250m x 250m
-            & NA
+        & Bareland*
+          & Average non-vegetated area
+            & 0-100\%
               & \cite{Dimiceli.2015} \\
       \hdashline
-      Protection Status
+      \multirow{2}{*}{Protection Status}
         & Protected
-          & NA
-            & NA
+          & Percentage cover of protected area
+            & 0-100\%
               & \cite{PeaceParks.2019} \\
-      Protection Status
-        & Pastoral
-          & NA
-            & NA
+        & Pastoral*
+          & Percentage cover of unprotected area
+            & 0-100\%
               & \cite{PeaceParks.2019} \\
       \hdashline
-      Human Influence
+      \multirow{3}{*}{Human Influence}
         & HumansBuffer5000
-          & 30m x 30m
-            & Sum of each 250m x 250m cell
+          & Average buffered (5 km) human influence
+            & \(\geq\) 0
               & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
-      Human Influence
         & DistanceToRoads
-          & NA
-            & NA
+          & Average distance to nearest road
+            & \(\geq\) 0m
               & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
-      Human Influence
         & RoadCrossing
-          & NA
-            & NA
+          & Binary; whether a step crossed a road
+            & 0, 1
               & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
@@
 \newpage
+\subsection{Integrated Step Selection Analysis}
+\label{Appendix:ISSF}
+
+\newpage
+\subsection{Least-Cost Paths \& Least-Cost Corridors}
+\label{Appendix:LeastCost}
+\begin{figure}[hbtp]
+  \begin{center}
+  \begin{minipage}{.25\linewidth}
+    \begin{subfigure}[t]{.9\linewidth}
+        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example1).pdf}
+    \end{subfigure}
+  \end{minipage}
+  \begin{minipage}{.25\linewidth}
+    \begin{subfigure}[t]{.9\linewidth}
+        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example2).pdf}
+    \end{subfigure}
+  \end{minipage}
+  \begin{minipage}{.25\linewidth}
+    \begin{subfigure}[t]{.9\linewidth}
+        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example3).pdf}
+    \end{subfigure}
+  \end{minipage}
+  \caption{Figures illustrating the process of identifying a least-cost corridor
+  between source points A and B following \cite{Pinto.2009}. (a) Example of a
+  permeability surface, which determines the costs of movement. (b1) Cumulative
+  cost map for point A, depicting the total minimal costs necessary to get from
+  point A to every other location. (b2) Cumulative cost map for point B,
+  depicting the total minimal costs to get from point B to every other location.
+  (c) Summed cost maps of points A and B. (d) Masked out corridor containing
+  pixels that do not exceed the cheapest pixel by more than 5\%.}
+  \label{LCCExample}
+  \end{center}
+\end{figure}
+
+\newpage
 \subsection{Model Selection Results}
@@
 \begin{table}[h]
-  \caption{Results from the forward model selection procedure based on AIC
-  \citep{Burnham.2002} for our habitat selection model. Weights were calculated
-  using: \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i =
-  1}^n e^{-0.5} \Delta \text{AIC}_i} \). The most parsimonious model (shaded in
-  light gray) outperformed all other models (\(\Delta AIC > \) 2) and received a
-  weight of one. The model in the last row did not converge and no AIC value
-  could be obtained. (*) represents the fixed covariates \(cos(ta)\) and
-  \(log(sl)\).}
+  \caption{Results from the forward model selection procedure based on Akaike's
+  Information Criterion (AIC, \citealp{Burnham.2002}) for the habitat selection
+  model. Weights were calculated using: \(\text{AIC-Weight}_i = \frac{e^{-0.5}
+  \Delta \text{AIC}_i} {\sum_{i = 1}^n e^{-0.5} \Delta \text{AIC}_i} \). The
+  most parsimonious model outperformed all other models (\(\Delta AIC > \) 2)
+  and received a weight of one. W = Water, DTW = DistanceToWater, S =
+  Shrubs/Grassland, T = Trees, P = Protected, HB5 = HumansBuff5000, RC =
+  RoadCrossing, DTR = DistanceToRoads.}
   \label{ModelAICs}
@@
       \toprule
-      \textbf{Covariates} &
-        \textbf{AIC} &
-          \textbf{\(\Delta\)AIC} &
-            \textbf{AIC-Weight} &
-              \textbf{LogLik} \\
+        Covariates & AIC & \(\Delta\)AIC & Weight & LogLik \\
         \midrule
-      \rowcolor[gray]{0.80}
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees &
-        89841.85 &
-          0.00 &
-            1.00 &
-              -44905.93 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, DistanceToRoads &
-        89844.46 &
-          2.61 &
-            0.00 &
-              -44905.23 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, Protected &
-        89845.78 &
-          3.93 &
-            0.00 &
-              -44905.89 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000 &
-        89847.96 &
-          6.11 &
-            0.00 &
-              -44910.98 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees,
-      RoadCrossing &
-        89848.24 &
-          6.38 &
-            0.00 &
-              -44905.12 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees,
-      DistanceToRoads, Protected &
-        89848.45 &
-          6.60 &
-            0.00 &
-              -44905.22 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, Trees
-        & 89850.51 &
-          8.66 &
-            0.00 &
-              -44912.26 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000,
-      DistanceToRoads &
-        89850.81 &
-          8.96 &
-            0.00 &
-              -44910.40 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, DistanceToRoads, RoadCrossing &
-        89850.87 &
-          9.02 &
-            0.00 &
-              -44904.44 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Protected &
-        89851.91 &
-          10.06 &
-            0.00 &
-              -44910.96 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, RoadCrossing &
-        89854.42 &
-          12.57 &
-            0.00 &
-              -44910.21 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland &
-        89856.89 &
-          15.03 &
-            0.00 &
-              -44917.44 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, DistanceToRoads &
-        89858.78 &
-          16.92 &
-            0.00 &
-              -44916.39 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, Protected &
-        89860.44 &
-          18.59 &
-            0.00 &
-              -44917.22 \\
-      (*), Water, DistanceToWater, Shrubs/Grassland, RoadCrossing &
-        89862.66 &
-          20.81 &
-            0.00 &
-              -44916.33 \\
-      (*), Water, DistanceToWater, HumansBuff5000 &
-        89870.14 &
-          28.29 &
-            0.00 &
-              -44924.07 \\
-      (*), Water, DistanceToWater &
-        89877.72 &
-          35.87 &
-            0.00 &
-              -44929.86 \\
-      (*), Water, DistanceToWater, DistanceToRoads &
-        89880.07 &
-          38.22 &
-            0.00 &
-              -44929.04 \\
-      (*), Water, DistanceToWater, Protected &
-        89881.39 &
-          39.54 &
-            0.00 &
-              -44929.70 \\
-      (*), Water, DistanceToWater, Trees &
-        89881.57 &
-          39.72 &
-            0.00 &
-              -44929.79 \\
-      (*), Water, DistanceToWater, RoadCrossing &
-        89883.48 &
-          41.63 &
-            0.00 &
-              -44928.74 \\
-      (*), Water, HumansBuff5000 &
-        89887.27 &
-          45.42 &
-            0.00 &
-              -44934.64 \\
-      (*), Water &
-        89894.64 &
-          52.79 &
-            0.00 &
-              -44940.32 \\
-      (*), Water, DistanceToRoads &
-        89897.24 &
-          55.39 &
-            0.00 &
-              -44939.62 \\
-      (*), Water, Protected &
-        89898.47 &
-          56.62 &
-            0.00 &
-              -44940.24 \\
-      (*), Water, Trees &
-        89898.53 &
-          56.67 &
-            0.00 &
-              -44940.26 \\
-      (*), Water, RoadCrossing &
-        89900.73 &
-          58.88 &
-            0.00 &
-              -44939.37 \\
-      (*), Shrubs/Grassland &
-        89999.47 &
-          157.62 &
-            0.00 &
-              -44992.73 \\
-      (*), DistanceToWater &
-        90001.72 &
-          159.87 &
-            0.00 &
-              -44993.86 \\
-      (*), HumansBuff5000 &
-        90056.07 &
-          214.22 &
-            0.00 &
-              -45021.04 \\
-      (*), Trees &
-        90059.55 &
-          217.69 &
-            0.00 &
-              -45022.77 \\
-      (*), DistanceToRoads &
-        90067.62 &
-          225.76 &
-            0.00 &
-              -45026.81 \\
-      (*), Protected &
-        90067.85 &
-          226.00 &
-            0.00 &
-              -45026.93 \\
-      (*), RoadCrossing &
-        90069.86 &
-          228.00 &
-            0.00 &
-              -45025.93 \\
-      (*), Water, Shrubs/Grassland &
-        \(NA\) &
-          \(NA\) &
-            0.00 &
-              \(NA\) \\
-      \bottomrule
-      \end{tabular}
-    }
-  \end{center}
-\end{table}
-
-\newpage
-\subsection{Further Investigations}
-We have been working on further investigations that did not fit the main text
-and were thus moved to the Appendix of this thesis. \Cref{SchematicOverview}
-gives an overview of how these additional investigations integrate into our main
-work. Our main work is described in the blue boxes and concluded with the
-identification of LCPs and LCCs. However, as we have emphasized in
-\Cref{Discussion}, the applicability of the iSSF framework expands beyond the
-estimation of movement and habitat preferences and it could, in fact, be used to
-simulate animal movement. We thus fitted a more detailed movement model and used
-it to simulate 1000 dispersal events across the surroundings of the core study
-area (red boxes in \Cref{SchematicOverview}). Corresponding methods and results
-are reported in \Cref{Appendix:DispersalSimulation}. In \Cref{Discussion}, we
-also motivated that data on the social landscape of dispersers would deepen our
-understanding of how dispersers respond to the presence or absence of
-conspecifics, apex predators, and prey. Although we lacked proper GPS relocation
-data of apex predators and prey, we collected a considerable amount of
-relocation data on resident wild dogs residing in the core study area. These
-data cover the periods of almost all of our dispersal events and thus hold the
-potential for modeling a social landscape for our dispersers. We attempted to
-model such a landscape for the core study area and used the resulting layer as
-additional covariate to parameterize a social model (yellow box in
-\Cref{SchematicOverview}). Corresponding methods and results are reported in
-\Cref{Appendix:SocialLandscape}.
-
-\begin{figure}[h]
-  \begin{center}
-    \begin{tikzpicture}[node distance=5mm and 5mm]
-      \node (mod2) [nonterminal
-        , minimum width = 30mm
-        , fill          = red!20
-        ] {Movement Model};
-      \node (mod1) [nonterminal
-        , above         = of mod2
-        , fill          = blue!20
-        , minimum width = 30mm
-        ] {Permeability Model};
-      \node (mod3) [nonterminal
-        , below         = of mod2
-        , fill          = yellow!20
-        , minimum width = 30mm
-        ] {Social Model};
-      \node (dat1) [nonterminal
-        , above left    = of mod2
-        , fill          = blue!20
-        , yshift        = -5mm
-        , minimum width = 30mm
-        ] {GPS Data};
-      \node (dat2) [nonterminal
-        , below left    = of mod2
-        , fill          = blue!20
-        , yshift        = +5mm
-        , minimum width = 30mm
-        ] {Spatial Covariates};
-      \node (for1) [nonterminal
-        , right         = of mod1
-        , fill          = blue!20
-        , minimum width = 35mm
-        ] {Permeability Map};
-      \node (forfor1) [nonterminal
-        , right         = of for1
-        , fill          = blue!20
-        , minimum width = 35mm
-        , yshift        = +5mm
-        ] {Least-Cost Paths};
-      \node (forfor2) [nonterminal
-        , right         = of for1
-        , fill          = blue!20
-        , minimum width = 35mm
-        , yshift        = -5mm
-        ] {Least-Cost Corridors};
-      \node (for2) [nonterminal
-        , right         = of mod2
-        , fill          = red!20
-        , minimum width = 35mm
-        ] {Dispersal Simulation};
-      \node (box) [draw
-        , dotted
-        , very thick
-        , minimum width   = 35mm
-        , minimum height  = 35mm
-        ] {};
-      \node [above = of box, yshift = -5mm] {iSSF};
-      \path (dat1) edge[->] (box);
-      \path (dat2) edge[->] (box);
-      \path (mod1) edge[->] (for1);
-      \path (mod2) edge[->] (for2);
-      \path (for1) edge[->] (forfor1);
-      \path (for1) edge[->] (forfor2);
-    \end{tikzpicture}
-    \caption{Schematic overview of our methodological approaches. The blue boxes
-    highlight our main line of work which resulted in the identification of
-    least-cost paths and least-cost corridors. The red boxes show that we
-    additionally used the iSSF framework to analyze movement behavior in more
-    detail and to simulate dispersal events. Finally, the yellow box indicates
-    our attempt to use the iSSF framework to model and analyze the influence of
-    the social landscape on dispersal behavior.}
-    \label{SchematicOverview}
-  \end{center}
-\end{figure}
-
-\newpage
-\subsection{Movement Model \& Dispersal Simulation}
-\label{Appendix:DispersalSimulation}
-\subsubsection{Methods}
-To gain more detailed insights into the movement behavior of dispersing wild
-dogs and to simulate dispersal events, we fitted a more complex movement model.
-The movement model still contained the most parsimonious habitat selection
-model, but in addition included interaction terms between movement metrics (i.e.
-\(cos(ta)\), \(log(sl\))) and environmental covariates (e.g. water). For
-instance, for the environmental covariate \textit{Water}, we created the
-interaction terms \textit{cos(ta):Water} and \textit{log(sl):Water}. These
-interactions served to learn how dispersers moved differently, depending on
-environmental covariates (for comparable work see \citealp{Prokopenko.2017}). We
-applied the same modeling framework as explained in \Cref{Modeling} and ran
-forward model selection based on AIC to identify the most parsimonious movement
-model. That is, we departed from the habitat selection model and iteratively
-added interactions with movement metrics. Note that we applied forward model
-selection only on the newly identified interactions, but not for the terms that
-were already in the habitat selection model. Since we attempted to keep our
-simulation simple, we did not average models with positive AIC-weights but
-simply identified the most parsimonious movement model.
-
-Using the most parsimonious movement model, we simulated 1000 dispersal events
-departing from the core study area. The simulation basically resembled an
-inverted iSSF function and was set up as follows. First, we sampled a random
-starting location within the core study area. Second, we assigned a random
-orientation in order to be able to calculate turning angles. Third, we created
-an array of 25 random steps (i.e. one stratum) following the same procedure as
-described in \Cref{Modeling}. Fourth, we extracted environmental covariates
-along each of the generated random steps and calculated the metrics listed in
-\Cref{ExtractedCovars}. Fifth, we applied \Cref{EQ2} using parameter estimates
-from the movement model and calculated the selection score \(w(x)\) of each of
-the 25 random steps. Sixth, we identified and realized the step with the highest
-selection score and determined the new location. We then repeated steps three to
-six until 200 steps were realized. To deal with dispersers that approached a map
-boundary, our simulation included the options to either stop the simulation and
-skip to the next disperser, or to remove any random step ranging outside the
-study area, thereby forcing the simulated disperser to stay within the
-boundaries. None of the simulated dispersers approached a map boundary within
-200 steps and the decision did not influence our findings. Although we ran the
-simulation only for dispersers departing from the core study area, one could
-easily expand the simulation to include multiple sources and thereby check how
-well these sources are connected.
-
-\newpage
-\subsubsection{Results}
-Along with the original covariates contained in the habitat selection model, our
-most parsimonious movement model (see \Cref{ModelAICs2} and
-\Cref{MovementModelResults}) retained three additional interactions (see red box
-in \Cref{MovementModel}a). Since the sign and direction of the effects contained
-in the habitat selection model did not markedly change, we only report on the
-newly added interactions between environmental covariates and movement metrics.
-For an easier interpretation of how these interactions influenced habitat
-selection and movement patterns during dispersal, we prepared surface plots as
-depicted in \Cref{MovementModel}b. These surface plots revealed that dispersers
-preferred long steps (i.e. a high movement rate) when water cover was low,
-whereas they preferred short steps (i.e. a low movement rate) when water cover
-was high (see \Cref{MovementModel}b1). Furthermore, dispersers were indifferent
-between high and low turning angles in proximity to water, but clearly preferred
-low turning angles (i.e. a high \(cos(ta)\)) when distance to water was high
-(see \Cref{MovementModel}b2). Finally, dispersers preferably realized steps with
-low turning angles when human influence was low, whereas they preferably
-realized low turning angles when human influence was high (see
-\Cref{MovementModel}b3).
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \begin{tikzpicture}
-        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-          \begin{minipage}{0.95\textwidth}
-            \begin{center}
-              \includegraphics[width = 0.80\textwidth]
-                {99_MovementModel.pdf}\\
-              \includegraphics[width = 1\textwidth]
-                {99_MovementModel(Interactions).pdf}
-            \end{center}
-          \end{minipage}
-        };
-        \begin{scope}[x={(image.south east)},y={(image.north west)}]
-           %  % next four lines will help you to locate the point needed by
-           %  % forming a grid. comment these four lines in the final picture.↓
-           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
-           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
-           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
-           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
-           %  % upto here
-            \draw [red, dashed] (0.13, 0.50) rectangle (0.405, 0.645);
-        \end{scope}
-    \end{tikzpicture}
-    \caption{(a) Results from the most parsimonious movement model. Negative
-    coefficients indicate avoidance of a covariate, positive coefficients
-    selection of a covariate. Whiskers delineate the 95\%-CIs for estimated
-    parameters. The intercept is meaningless in the iSSF framework and was thus
-    omitted. (b) Corresponding surface plots for two-way interactions. Light
-    areas depict a combination of covariates that were preferred by dispersers,
-    whereas darker areas correspond to avoided combinations. The absolute values
-    along the axes have no meaning as original values were normalized to 0 (low)
-    and 1 (high).}
-    \label{MovementModel}
-  \end{center}
-\end{figure}
-
-\newpage
-\noindent \Cref{Simulations}a shows the outcome of the dispersal simulations.
-For better visibility of high-frequency routes, we rasterized all simulated
-trajectories and computed a heatmap depicting areas that were crossed by many
-simulated trajectories. Unsurprisingly, we found that the dispersal trajectories
-clearly reflected preferences that we identified earlier. For instance, most
-trajectories were quite directional, avoided crossing a waterbody, ran along
-water bodies and avoided human-dominated areas. More interestingly, however, the
-resulting heatmap closely resembled our permeability surface (see
-\Cref{Simulations}b). That is, areas providing high permeability were typically
-crossed by more simulated trajectories. Yet, the heatmap also showed that some
-regions are almost impossible to reach, even if they provide highly permeable
-landscapes. For example, although the western section of the Okavango Delta
-holds highly permeable landscapes, no simulated trajectory reached these areas.
-In contrast, many trajectories ventured towards east, some of them even crossing
-the border to Zimbabwe. The reason that no trajectory reached the eastern part
-of the Delta is likely owed to severe human influence and an impermeable flood
-that together posed an insurmountable barrier and prevented dispersers from
-traversing the landscape.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width = 0.83\textwidth]{99_Simulations.pdf}
-    \caption{(a) Heatmap of 1000 simulated dispersal events departing from the
-    core study area (white rectangle) in northern Botswana (countries with
-    dashed white lines). (b) Permeability map for the same extent to illustrate
-    that simulated dispersal trajectories typically cover highly permeable
-    landscapes.}
-    \label{Simulations}
-  \end{center}
-\end{figure}
-
-\newpage
-\begin{table}[h]
-  \caption{Results from the forward model selection procedure based on AIC
-  \citep{Burnham.2002} for our movement model. Weights were calculated using:
-  \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i = 1}^n
-  e^{-0.5} \Delta \text{AIC}_i} \). (*) represents all covariates from the most
-  parsimonious habitat selection model, i.e. \(cos(ta)\), \(log(sl\)),
-  \textit{Water}, \textit{DistanceToWater}, \textit{Shrubs/Grassland},
-  \textit{HumansBuff5000}, and \textit{Trees}. To keep the table short we only
-  included models with positive AIC weights. We abstained from averaging models
-  with AIC-weight \(>\) 0 because we attempted to keep our simulation simple.}
-  \label{ModelAICs2}
-  \begin{center}
-    \resizebox{\textwidth}{!}{
-      \begin{tabular}{llllll}
-      \toprule
-      \textbf{Covariates} &
-        \textbf{AIC} &
-          \textbf{\(\Delta\)AIC} &
-            \textbf{AIC-Weight} &
-              \textbf{LogLik} \\
-      \midrule
-      \rowcolor[gray]{0.80}
-        (*),
-        log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000 &
-          89814.19 &
-            0.00 &
-              0.13 &
-                -44889.10 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):Trees &
-          89814.37 &
-            0.17 & 0.12 &
-              -44889.18 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater &
-          89814.46 &
-            0.27 &
-              0.11 &
-                -44890.23 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Trees &
-          89814.87 &
-            0.68 &
-              0.09 &
-                -44888.44 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Shrubs &
-          89815.32 &
-            1.13 &
-              0.07 &
-                -44888.66 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):Shrubs &
-          89815.34 &
-            1.15 &
-              0.07 &
-                -44889.67 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):Shrubs &
-          89815.40 &
-            1.20 &
-              0.07 &
-                -44888.70 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, log(sl):Shrubs &
-          89815.67 &
-            1.48 &
-              0.06 &
-                -44889.84 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):HumansBuff5000 &
-          89815.94 &
-            1.75 &
-              0.05 &
-                -44888.97 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Trees, log(sl):Shrubs &
-          89816.08 &
-            1.89 &
-              0.05 &
-                -44888.04 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):DistanceToWater &
-          89816.14 &
-            1.95 &
-              0.05 &
-                -44889.07 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Water &
-          89816.16 &
-            1.97 &
-              0.05 &
-                -44889.08 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, log(sl):HumansBuff5000 &
-          89816.18 &
-            1.99 &
-              0.05 &
-                -44890.09 \\
-        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):Trees &
-          89816.19 &
-            2.00 &
-              0.05 &
-                -44889.10 \\
-      \bottomrule
-      \end{tabular}
-    }
-  \end{center}
-\end{table}
-
-\begin{table}[h]
-  \caption{Selection coefficients as derived from the most parsimonious movement
-  model. The movement model was a modified habitat selection model to which we
-  added additional interaction terms between movement metrics and environmental
-  covariates (e.g. \(Water:cost(ta\_)\)) in order to learn about how movement
-  behavior changes depending on the environmental covariates.}
-  \label{MovementModelResults}
-  \begin{center}
-    \resizebox{0.5\textwidth}{!}{
-      \begin{tabular}{llllll}
-      \toprule
-      Covariate & Coefficient & SE & p-value \\
-      \midrule
-      cos(ta) & 0.149 & 0.045 & 0.001 \\
-      log(sl) & 0.053 & 0.019 & 0.005 \\
-      Water & -0.265 & 0.090 & 0.003 \\
-      DistanceToWater & -0.272 & 0.226 & 0.230 \\
-      Shrubs & 0.391 & 0.082 & 0.000 \\
-      HumansBuff5000 & -0.466 & 0.203 & 0.022 \\
-      Trees & -0.177 & 0.072 & 0.013 \\
-      log(sl):Water & -0.040 & 0.008 & 0.000 \\
-      cos(ta):DistanceToWater & 0.115 & 0.038 & 0.003 \\
-      cos(ta):HumansBuff5000 & -0.047 & 0.031 & 0.134 \\
-       \bottomrule
-      \end{tabular}
-    }
-  \end{center}
-\end{table}
-
-\newpage
-\subsection{Social Model}
-\label{Appendix:SocialLandscape}
-\subsubsection{Methods}
-We used GPS relocation data of resident wild dogs and calculated a
-geo-referenced stack of layers to which we refer as the \textit{social
-landscape}. The social landscape was a collection of raster-layers, each
-depicting the probability of encountering a resident group in a given pixel and
-month. The social landscape was updated every month, using a moving window that
-took into account six months of relocation data. For instance, the social
-landscape dated July 1\textsuperscript{st} consisted of GPS relocation data
-collected between January 1\textsuperscript{st} and June 30\textsuperscript{th}.
-Because we collected relocation data of residents only inside the core study
-area, we modeled the social landscape only for this region. GPS relocations were
-collected at a 24-hourly interval during residence, so we kept one GPS
-relocation per pack and day. To compute monthly updated social landscapes, we
-calculated each resident pack's utilization distribution (UD) and corresponding
-95\% home range (HR) using the kernelUD function from the R-package
-\textit{adehabitatHR} \citep{Calenge.2019}. The package required to set a
-smoothing parameter \(h\), which determined the sharpness of the resulting UD
-and 95\%-HR. To come up with a meaningful smoothing parameter \(h\), we followed
-the procedure described in \cite{Cozzi.2018}. That is, we first applied the
-reference value \( h_{ref} \), which the package automatically determined. We
-then obtained the corresponding UD and 95\%-HR and assessed whether the 95\%-HR
-comprised of a single polygon or whether it split up into multiple polygons. In
-case the 95\%-HR segregated into multiple polygons, we assumed that the
-smoothing parameter was set to strictly. On the other hand, if the 95\%-HR
-comprised of a single polygon, we assumed that we could further tighten the
-smoothing parameter. We thus iteratively adjusted the smoothing parameter until
-a value right before the 95\%-HR split up into multiple polygons. However, we
-maximally varied \( h \) within 30\% and 120\% of the original \( h_{ref} \) to
-avoid extreme over- or under-smoothing. Using the resulting \(h\) we computed
-the final UD and corresponding 95\%-HR. Each pack's UD of a given month was
-converted to a spatial probability density using \(f(x) =
-\frac{x_i}{\sum_1^n{x_i}}\), so that pixel-values summed up to one. We then used
-\Cref{EQ4} to merge all UDs of the same month and to calculate the probability
-of not encountering any wild dogs in a given pixel and month.
-
-\begin{equation}
-\label{EQ4}
-x_i = 1 - \prod_{j = 1}^{n} (1 - x_{ij})
-\end{equation}
-
-\noindent where \(x_i =\) is the probability of encountering another pack in
-pixel \(i\) and \(x_{ij} =\) the probability of encountering pack \(j\) in pixel
-\(i\). Finally, we inverted this probability using \((1-x_i)\) to get the
-probability of encountering at least one resident pack in a given pixel and
-month. Note that we kept track of each packs' 95\%-HR throughout all months,
-which allowed us to identify through which HRs dispersers moved during
-dispersal. An example of the final layers is given in
-\Cref{SocialLandscapeExample}.
-
-\begin{figure}[h]
-  \begin{center}
-    \includegraphics[width = 0.95\textwidth]{99_SocialLandscapeExample.pdf}
-    \caption{(a) Example of a layer depicting the probability of encountering a
-    resident pack in July 2017. The layer was derived from utility distributions
-    of all packs residing in the core study area. (b) Example of a layer
-    depicting all 95\% home ranges of the packs residing in the core study area.
-    This layer served to identify whether a disperser still moved through his
-    natal pack's home range, or whether it moved through a foreign packs home
-    range. The layers in (a) and (b) were updated each month to take into
-    account only the latest GPS relocation data of resident individuals.}
-    \label{SocialLandscapeExample}
-  \end{center}
-\end{figure}
-
-\noindent To fit a social model using the newly created social landscape, we
-reran our iSSF described in \Cref{Modeling}. However, this time we discarded any
-stratum that did not fully lie within the core study area. The social model that
-we fitted was again a modification of the most parsimonious habitat selection
-model. Despite the covariates contained in the habitat selection model, we added
-the social covariates listed in \Cref{SocialCovars}. Because we hypothesized
-that the preference for encountering other wild dogs was different within and
-outside the disperser's natal home range, we included the interaction terms
-\textit{ProbEncounter:OwnHomerange} and \textit{ProbEncounter:ForeignHomeranges}
-in the social model. Since we did not anticipate to use the model for
-predictions, we did not engage in model selection.
-
-\begin{table}[h]
-  \begin{center}
-    \caption{Overview of extracted social covariates. Values were extracted from
-    the layer that was closest in date to the actual step.}
-    \label{SocialCovars}
-    \resizebox{\textwidth}{!} {
-      \begin{tabular}{lllll}
-      \hline
-      \textbf{Category} &
-        \textbf{Covariate} &
-          \textbf{Description} &
-            \textbf{Values} \\
-              % \textbf{Source(s)} \\
-      \midrule
-      Social Features
-        & OwnHomerange
-          & Percentage cover of natal pack’s 95\% home-range along step
-            & 0-100\% \\
-      Social Features
-        & ForeignHomeranges
-          & Number of foreign packs’ 95\% home-ranges along step
-            & \(\geq\) 0 \\
-      Social Features
-        & ProbEncounter
-          & Average probability of encountering conspecifics along step
-            & 0-100\% \\
-      \hline
-      \end{tabular}
-    }
-  \end{center}
-\end{table}
-
-\newpage
-\subsubsection{Results}
-Results from our social model are depicted in \Cref{SocialModel}a (see also
-\Cref{SocialModelResults}). Again, parameter estimates for the covariates
-contained in the habitat selection model did not severely change and we will
-only report on the newly added social covariates (see red box in
-\Cref{SocialModel}a). Visual inspection of the corresponding surface plot (see
-\Cref{SocialModel}b1) revealed that dispersers tried to maximize the probability
-of encounter when moving outside their natal pack's HR. Yet, this preference was
-even stronger when dispersers moved inside their natal pack's HR. A second
-surface plot (see \Cref{SocialModel}b2) showed that dispersers maximized the
-probability of encounter when the number of foreign HRs was low, but that they
-minimized the probability of encounter when the number of foreign HRs was high.
-In summary, dispersers appeared to be drawn towards other wild dogs, but only if
-the number of overlapping foreign HRs was low.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \begin{tikzpicture}
-        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-          \begin{minipage}{0.95\textwidth}
-            \begin{center}
-              \includegraphics[width = 0.75\textwidth]
-                {99_SocialModel.pdf}\\
-              \includegraphics[width = 0.8\textwidth]
-                {99_SocialModel(Interactions).pdf}
-            \end{center}
-          \end{minipage}
-        };
-        \begin{scope}[x={(image.south east)},y={(image.north west)}]
-           %  % next four lines will help you to locate the point needed by
-           %  % forming a grid. comment these four lines in the final picture.↓
-           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
-           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
-           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
-           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
-           %  % upto here
-            \draw [red, dashed] (0.15, 0.60) rectangle (0.43, 0.75);
-        \end{scope}
-    \end{tikzpicture}
-    \caption{(a) Estimated parameters from the social model. Negative
-    coefficients indicate avoidance of a covariate, positive coefficients
-    selection of a covariate. Whiskers delineate the 95\%-CIs for estimated
-    parameters. The intercept is meaningless in the iSSF framework and was thus
-    omitted. (b) Corresponding surface plots for two-way interactions (red box).
-    Light areas in the surface plots depict preferred combinations of the
-    covariates, whereas darker areas correspond to avoided combinations. The
-    absolute values along the axes have no meaning as original values were
-    normalized to 0 (low) and 1 (high).}
-    \label{SocialModel}
-  \end{center}
-\end{figure}
-
-\newpage
-\begin{table}[h]
-  \caption{Selection coefficients as derived from the social model. The social
-  model was a modified habitat selection model to which we added additional
-  social covariates in order to learn about how the social landscape infleunces
-  dispersal behavior.}
-  \label{SocialModelResults}
-  \begin{center}
-    \resizebox{0.6\textwidth}{!}{
-      \begin{tabular}{llllll}
-        \toprule
-        Covariate & Coefficient & SE & p-value \\
-        \midrule
-        cos(ta) & 0.148 & 0.066 & 0.025 \\
-        log(sl) & 0.055 & 0.019 & 0.003 \\
-        Water & -0.472 & 0.113 & 0.000 \\
-        DistanceToWater & -0.193 & 0.411 & 0.639 \\
-        Shrubs & 0.470 & 0.110 & 0.000 \\
-        HumansBuff5000 & -1.013 & 0.295 & 0.001 \\
-        Trees & -0.465 & 0.193 & 0.016 \\
-        ProbEncounter & 0.724 & 0.268 & 0.007 \\
-        OwnHomerange & 0.056 & 0.095 & 0.557 \\
-        ForeignHomeranges & -0.099 & 0.132 & 0.452 \\
-        ProbEncounter:OwnHomerange & 0.189 & 0.179 & 0.291 \\
-        ProbEncounter:ForeignHomeranges & -0.504 & 0.281 & 0.072 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T & 90099.62 & 0.00 & 1.00 & -45034.81 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T + DTR & 90101.67 & 2.04 & 0.00 & -45033.83 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T + P & 90103.56 & 3.93 & 0.00 & -45034.78 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 & 90104.31 & 4.69 & 0.00 & -45039.16 \\
+        cos(ta) + log(sl) + W + S + DTW + T & 90104.98 & 5.36 & 0.00 & -45039.49 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T + DTR + P & 90105.67 & 6.04 & 0.00 & -45033.83 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T + RC & 90105.84 & 6.22 & 0.00 & -45033.92 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + DTR & 90106.27 & 6.65 & 0.00 & -45038.14 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T + DTR + RC & 90107.87 & 8.25 & 0.00 & -45032.93 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + P & 90108.23 & 8.61 & 0.00 & -45039.12 \\
+        cos(ta) + log(sl) + W + S + DTW & 90110.11 & 10.49 & 0.00 & -45044.06 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + RC & 90110.66 & 11.03 & 0.00 & -45038.33 \\
+        cos(ta) + log(sl) + W + S + DTW + DTR & 90111.70 & 12.07 & 0.00 & -45042.85 \\
+        cos(ta) + log(sl) + W + S + DTW + HB5 + T + DTR + P + RC & 90111.87 & 12.25 & 0.00 & -45032.93 \\
+        cos(ta) + log(sl) + W + S + DTW + P & 90113.60 & 13.98 & 0.00 & -45043.80 \\
+        cos(ta) + log(sl) + W + S + DTW + RC & 90115.50 & 15.88 & 0.00 & -45042.75 \\
+        cos(ta) + log(sl) + W + S + HB5 & 90132.73 & 33.11 & 0.00 & -45055.37 \\
+        cos(ta) + log(sl) + W + S + T & 90136.86 & 37.24 & 0.00 & -45057.43 \\
+        cos(ta) + log(sl) + W + S & 90137.49 & 37.87 & 0.00 & -45059.75 \\
+        cos(ta) + log(sl) + W + DTW & 90138.92 & 39.30 & 0.00 & -45060.46 \\
+        cos(ta) + log(sl) + W + S + DTR & 90139.06 & 39.44 & 0.00 & -45058.53 \\
+        cos(ta) + log(sl) + W + S + P & 90141.35 & 41.73 & 0.00 & -45059.67 \\
+        cos(ta) + log(sl) + W + S + RC & 90143.27 & 43.64 & 0.00 & -45058.63 \\
+        cos(ta) + log(sl) + W + HB5 & 90149.30 & 49.68 & 0.00 & -45065.65 \\
+        cos(ta) + log(sl) + W & 90153.70 & 54.08 & 0.00 & -45069.85 \\
+        cos(ta) + log(sl) + W + T & 90155.28 & 55.65 & 0.00 & -45068.64 \\
+        cos(ta) + log(sl) + W + DTR & 90156.32 & 56.70 & 0.00 & -45069.16 \\
+        cos(ta) + log(sl) + W + P & 90157.56 & 57.93 & 0.00 & -45069.78 \\
+        cos(ta) + log(sl) + W + RC & 90159.32 & 59.69 & 0.00 & -45068.66 \\
+        cos(ta) + log(sl) + S & 90238.43 & 138.80 & 0.00 & -45112.21 \\
+        cos(ta) + log(sl) + DTW & 90262.03 & 162.40 & 0.00 & -45124.01 \\
+        cos(ta) + log(sl) + T & 90299.89 & 200.27 & 0.00 & -45142.95 \\
+        cos(ta) + log(sl) + HB5 & 90307.55 & 207.93 & 0.00 & -45146.78 \\
+        cos(ta) + log(sl) + DTR & 90316.37 & 216.75 & 0.00 & -45151.19 \\
+        cos(ta) + log(sl) + P & 90316.87 & 217.25 & 0.00 & -45151.43 \\
+        cos(ta) + log(sl) + RC & 90318.46 & 218.84 & 0.00 & -45150.23 \\
        \bottomrule

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-05-01.tex
first commit

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[nomarkers, nolists]{endfloat}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Bound within Boundaries? How Well Do Boundaries of Protected Areas Match
Movement Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David Hofmann\textsuperscript{1\S} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Dominik Behr\textsuperscript{1,2*} \and
  Gabriele Cozzi\textsuperscript{1,2*} \and
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared senior co-authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} Lycaon pictus, KAZA-TFCA, permeability surface, least-cost
corridors, integrated step selection function, landscape connectivity

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
\noindent In response to the global deterioration and fragmentation of suitable
habitats, large portions of land have been set aside to maintain and restore
connectivity among wildlife subpopulations. While boundaries of such
conservation areas are often based on expert opinion and socio-political needs
and constraints, the extent to which their boundaries match movement corridors
of their protected species is rarely assessed. This is mainly due to lack of
adequate data. One such conservation area is the Kavango-Zambezi Transfrontier
Conservation Area (KAZA-TFCA) in Africa, which, spanning five countries and
520'000 km\textsuperscript{2}, is the largest transboundary conservation area in
the world.

\noindent Between 2011 and 2019, we collected novel high-resolution GPS
relocation data on dispersing African wild dogs (\textit{Lycaon pictus}) to
assess to which degree the KAZA-TFCA boundaries reflect natural movement
corridors of this highly mobile and critically endangered species.

\noindent We demonstrate that the major dispersal corridors were within the broad
KAZA-TFCA boundaries, yet some run outside formally-protected areas. Further, we
identified significant differences in permeability to movement across the five
KAZA-TFCA countries. Such differences were mainly due to different degrees of
human activities, which hampered dispersal movements.

\noindent Our work shows the importance of empirically assessing the adequacy of
protected areas boundaries and the need for international coordination that stem
from different contribution to connectivity
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}
As a result of the ongoing human-caused habitat deterioration and fragmentation,
successful management and conservation of isolated subpopulations requires
maintenance and enhancement of connectivity between them \citep{Hanski.1998,
Fahrig.2003, Lindenmayer.2013}. Recently, efforts have been made to facilitate
animal movements and dispersal, and to secure connectivity by protecting
contiguous natural or semi-natural areas \citep{Heller.2009, Doerr.2011,
Rudnick.2012, Cozzi.2020}. These efforts have resulted in the creation of an
ever growing number of large and often transboundary protected areas
\citep{Wolmer.2003}, whose boundaries have been drawn based largely on expert
opinions or political reasons, often revealing deficiencies
\citep{Pullinger.2010}. Due to the very nature of such vast protected areas,
which sometimes span several thousands of square kilometers and traverse
multiple countries, a thorough assessment of their effectiveness and adequacy of
their boundaries is indeed challenging \citep{Rudnick.2012}. This challenge is
complicated by a lack of movement data at the appropriate spatial scale and an
insufficient understanding of the ranging behavior, movement patterns, and
dispersal corridors of the species for which these protected areas have been
created \citep{Vasudev.2015, Tshipa.2017, Cozzi.2020}. An accurate
representation of dispersal corridors and migratory routes based on empirical
global positioning system (GPS) data is thus fundamental to assess how well
boundaries of protected areas reflect actual animal movements \citep{Beier.2008,
Doerr.2011, Rudnick.2012}.

In recent years, there has been a growing body of research that uses animal
relocation data to identify dispersal corridors and assess connectivity at large
scales \citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014,
Benz.2016, Osipova.2019}. Identification of potential corridors relies on the
estimation of permeability (respectively resistance) surfaces, which return the
ease or willingness at which the focal species traverses a specific landscape
\citep{Sawyer.2011}. Such surfaces are created on the basis of habitat
preferences, which can be estimated using a suite of selections functions,
including resource selection functions \citep{Boyce.2002}, step selection
functions \citep{Fortin.2005}, and path selection functions
\citep{Cushman.2010}. These approaches have in common that they infer habitat
preferences by comparing landscape covariates (e.g. environmental and
anthropogenic) at locations visited by the animal to the same spatial covariates
at randomly selected locations \citep{Zeller.2012}. As such, these approaches
require adequate landscape and GPS relocation data that is representative of the
process being studied \citep{Diniz.2020}. More specifically, GPS relocation
data collected on dispersing individuals has been shown to outperform GPS
relocation data collected on resident individuals when the aim is to detect
large scale dispersal corridors \citep{Elliot.2014, Diniz.2020}. Few studies
have, however, partitioned their GPS relocation data according to behavioral
modes (e.g. resident vs dispersing) of the studied species \citep{Wilson.2012,
Vasudev.2015}. As such, most permeability surfaces upon which dispersal
corridors are identified are created using GPS relocation data that was
collected on resident individuals. This introduces severe biases and
substantially reduces the power to reveal meaningful movement corridors - for
dispersing individuals have different needs and motivations compared to resident
individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020} - thus limiting our
ability to assess the adequacy of the boundaries of protected areas.

One initiative that aims at restoring and enhancing connectivity across large
scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), which
spanning over 520'000 km\textsuperscript{2} and five countries constitutes the
largest transfrontier conservation area in the world. While the KAZA-TFCA was
originally set to facilitate movements of African elephants (\textit{Loxodonta
africana}; \citealp{Tshipa.2017}), it is also key to the conservation of  other
wide-ranging species, such as African wild dogs (\textit{Lycaon pictus};
\citealp{Woodroffe.2012, Cozzi.2020}), lions (\textit{Panthera leo};
\citealp{Elliot.2014, Cushman.2018}), and cheetahs (\textit{Acinonyx jubatus};
\citealp{Weise.2017}). To date, however, few studies have attempted to assess
the appropriateness of KAZA's boundaries using relevant GPS movement  data at
the appropriate spatial scale \citep{Elliot.2014, Tshipa.2017}. Thus, how well
the KAZA-TFCA's boundaries reflect natural movement patterns and dispersal
corridors of its protected species is virtually unknown.

A highly endangered flagship species for conservation efforts across the
KAZA-TFCA, but for which systematic GPS relocation data on dispersing
individuals is unavailable, is the African wild dog (\textit{Lycaon pictus}).
While once widespread across the entire Sub-Saharan continent, the species has
been widely extirpated through human persecution, habitat destruction, and
disease outbreaks \citep{Woodroffe.2012}. As a result, the African wild dog has
become one of Africa's most endangered large carnivores, and it currently only
survives in small, spatially scattered subpopulations \citep{Woodroffe.2012}.
Within these subpopulations, wild dogs form cooperative breeding packs of up to
thirty individuals \citep{Frame.1979, Fuller.1992, Creel.2002}. Males and
females typically disperse from their natal pack, either alone or in same-sex
dispersing coalitions, and search for unrelated mates and a suitable territory
to settle \citep{McNutt.1996, Behr.2020}. During dispersal, wild dogs can cover
several hundred kilometers, sometimes crossing national boundaries and areas
strongly influenced by humans \citep{DaviesMostert.2012, Masenga.2016,
Cozzi.2020}. To date, little empirical information is available on habitat
selection and movement behavior of wild dogs during dispersal. The few studies
that have collected GPS relocation data during dispersal have demonstrated that
dispersers generally avoid human-dominated landscapes \citep{Masenga.2016,
Oneill.2020, Cozzi.2020} and areas densely forested \citep{Oneill.2020},
whereas they prefer proximity to water and dust roads \citep{Oneill.2020}.

Here, we used GPS relocation data collected on 16 wild dogs in as many
dispersing coalitions from a free-ranging population to assess the adequacy of
the boundaries of the KAZA-TFCA, the largest protected area in the world. To
this end, we estimated the degree of selection or avoidance for environmental
and anthropogenic landscape features and used the obtained information to create
a permeability surface.  This allowed us to investigate how landscape
permeability varies regionally and internationally and to identify major
dispersal corridors and dispersal hubs.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E; elevation ca.
1'030 m, red rectangle in \Cref{StudyArea} was confined by a bounding box
encompassing the entire KAZA-TFCA. was outlined by a rectangular bounding box
stretching over 1.3 Mio km\textsuperscript{2} and encompassing the entire
KAZA-TFCA. The KAZA-TFCA lies in the basins of the Okavango and Zambezi rivers
and includes parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. With a
total area of over 520'000 km\textsuperscript{2} it covers several
already-existing national parks (NPs) and other protected areas. It constitutes
the earth's largest transboundary conservation area and is characterized by
diverse landscapes, including savanna, grassland, and dry or moist woodland
habitats. Rainfall in the study area is seasonal and lasts from November to
March, whereas the rest of the year remains dry \citep{McNutt.1996,
Mendelsohn.2010}. The KAZA-TFCA also encloses the Okavango Delta, which is one
of the main strongholds of African wild dogs in southern Africa and likely acts
as a source for the recolonization of surrounding habitats
\citep{Woodroffe.2012, Cozzi.2013}. The extent of the flood in the delta greatly
changes within and between years depending on the amount of rain that descends
from the catchment areas in Angola and reaches the distal ends of the delta
between July and August (Figure XX in Appendix XX for an illustration of a
typical flood pulse). The flood drastically affects surrounding landscapes, so
that during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes
a patchy conglomerate of swamps, open water, and islands, whereas these
structures disappear when the flood retracts (ca. 5'000 km\textsuperscript{2})
during dry periods \citep{Gumbricht.2004, Wolski.2017}. Despite 36 NPs and other
protected areas, there is considerable human influence in some regions of the
KAZA-TFCA, mainly originating from human activities on farms, high human
density, and road traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red rectangle depicts the study
    area, which was confined by a bounding box encompassing the entire
    KAZA-TFCA. Pink areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
    area within which dispersal coalitions were collared. Since Game Reserves in
    Botswana virtually serve the same purpose as National Parks, we use the
    terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
We used a population of free-ranging African wild dogs inhabiting the Okavango
Delta in northern Botswana as a source population for dispersing individuals.
This population has been extensively studied since 1989 \citep{McNutt.1996,
McNutt.2008, Cozzi.2013, Cozzi.2020, Behr.2020}. We collected GPS relocation
data of 16 wild dog coalitions (7 female and 9 male coalitions) that dispersed
between 2011 and 2019 (see also \citealp{Abrahms.2017} and
\citealp{Cozzi.2020}). Candidate dispersing individuals were identified based on
procedures reported in \cite{Behr.2020}, immobilized according to protocols
described in \cite{Osofsky.1996}, and fitted with GPS/Satellite radio collars
(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while still in their
natal pack. All required procedures were undertaken and supervised by a
Botswana-registered wildlife veterinarian. Immobilized individuals were
monitored and all were observed to successfully re-join their packs within one
hour after immobilization. During dispersal, GPS collars were programmed to
record GPS relocation data every 4 hours and to regularly transmit them via
Iridium satellite system to a base station. The system thus allowed
remote-tracking of collared individuals, even where field conditions were
prohibitive or dispersal trajectories crossed international borders.

Because we were interested in recording dispersal trajectories, we discarded any
GPS data collected while individuals were still with their natal packs
\citep{Cozzi.2020}. Dispersal periods were identified based on supporting
information from the field, as well as from plotting and visually inspecting the
net squared displacement (NSD) metric for each dispersing coalition. NSD is a
metric indicating the euclidean distance of a relocation to a reference point
\citep{Borger.2012}, which in our case was the center of the dispersing
coalition's natal home range. Thus, dispersal was deemed to have started when a
coalition had left its natal home range and continued until the NSD metric
remained stationary, implying that the coalition settled (Figure XX in Appendix
XX). Some coalitions undertook exploratory movements prior to proper dispersal.
Such movements have been observed in other species and are reported to be very
similar to proper dispersal \citep{Killeen.2014}, so we classified exploratory
movements as dispersal too. In total, we collected 4'169 GPS relocations during
dispersal (Figure XX and Table XX in Appendix XX) for an average of 260
locations/coalition (range: 37 - 729). On average, dispersing coalitions
dispersed for 48 days (min = 9, max = 137), covered a mean euclidean distance of
54 km (min = 3, max = 263) and a mean cumulative distance of 596 km (min = 130,
max = 1'962). In our analysis, we did not differentiate between male and female
dispersing coalitions, for previous research found little differences between
sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.

\subsection{Spatial Covariates}
We used a set of geo-referenced covariates to investigate habitat preferences of
dispersing wild dogs (\Cref{Covariates}). Covariates fell into the categories
\textit{land cover}, \textit{protection status}, and \textit{human influence}.
Due to the absence of noteworthy elevational gradients in our study area we did
not include terrain features as covariates. For each covariate, we prepared
spatial raster layers from freely available online services or from remotely
sensed satellite imagery. To ensure a consistent resolution (i.e. cell-size or
grain) across covariates, we coarsened or interpolated all layers to match a
resolution of 250m x 250m. We performed processing and manipulation of data, as
well as all spatial and statistical analyses, using R, version 3.6.1
\citep{R.2019}).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of spatial covariates that we included in our models. We
    prepared all covariates for the entire study area but for better visibility
    we only plot them for the surroundings of the Okavango Delta. The white
    rectangle in each plot depicts the area within which dispersal coalitions
    were collared. (a) Averaged layer of all dynamic (binary) watermaps. (b)
    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
    Anything that was not covered by trees or shrubs/grassland was deemed to be
    bare land. (d) Protection status of the area. (e) Human influence proxy
    composed of human density, farms, and roads. (f) Distance to nearest road
    (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsubsection{Land Cover}
We divided land cover into water (1) and dryland (2), and we further subdivided
dryland into three categories.

\begin{enumerate}[label = (\arabic*)]

  \item Water included rivers, wetlands, and swamps. Because the
  inundation extent of the flood in the Okavango Delta is highly variable within
  and between years, we created weekly (8 daily)  ``floodmaps'' , following a
  remote sensing algorithm developed by \citep{Wolski.2017}. The algorithm uses
  thresholding of short wavelength infrared reflectances of MODIS Terra
  satellite imagery (MCD43A4, \citealp{Schaaf.2015}) to distinguish areas
  covered by water and dryland (details in Appendix XX). While we created
  dynamic floodmaps for the Okavango Delta, we assumed all other water bodies
  (e.g. Chobe river, Zambezi river) to be static within and between years. This
  static representation was based on Globeland's land cover dataset
  \citep{Chen.2015}, from which we only retained the categories \textit{wetland}
  and \textit{water bodies} and collectively reclassified them to
  \textit{water}. Globeland had an original resolution of 30m x 30m, so we
  coarsened the layer to 250m x 250m using the mode of each 250m x 250m cell. We
  further improved river representation using the rasterized MERIT Hydro dataset
  \citep{Yamazaki.2019}, from which which we added all rivers with a width of
  over 10m to our Globeland layer. Finally, we merged dynamic and static
  watermaps into a large rasterstack, covering the entire study area.

  \item We subdivided dryland into three categories as derived from the
  MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
  \citealp{Dimiceli.2015}). Combined, the three layers added up to 100\% and
  depicted the percentage cover of tree-vegetation (henceforth trees),
  non-tree-vegetation (henceforth shrubs/grassland), and non-vegetated
  (henceforth bare land).  We used our floodmap that aligned with the creation
  date of these MODIS layers (30 March 2017) and defined anything covered by
  water as 0\% vegetated. The MODIS vegetation layers had a resolution of 250m x
  250m, so we did not need to coarsen or interpolate them.

\end{enumerate}

\subsubsection{Proection Status}
We created a binary layer separating protected from unprotected land. We
downloaded corresponding data on protection status from the Peace Parks
Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}). Protected
areas included forest reserves, game reserves, wildlife management areas, and
national parks. We classified anything not covered by these categories as
unprotected (e.g. communal pastoral land, private land). We rasterized the two
categories onto a binary raster (1 = protected, 0 = pastoral) with a resolution
of 250m x 250m.

\subsubsection{Human Influence}
We created a layer representing human influence by integrating information on
(1) human density, (2) farming activities, and (3) roads.

\begin{enumerate}[label = (\arabic*)]

  \item We obtained spatial human density estimates through Facebook's 30m x 30m
  high-resolution population density dataset (\url{www.dataforgood.fb.com};
  \citealp{Facebook.2019}). The layer was coarsened to 250m x 250m by summing up
  human density values within each 250m x 250m cell.

  \item We sourced Information on farming activities from the Globeland
  \citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets, from
  which we retained areas that were classified as either \textit{cultivated
  land} or \textit{croplands}. Any other land cover classes were not pertinent
  to farming and therefore omitted. Because both layers had a resolution of 30m
  x 30m we coarsened them to 250m x 250m by assigning a value of 1 to any 250m x
  250m cell that covered farmland and a value 0 otherwise. Thus the final layer
  depicted presence (= 1) or absence (= 0) of farms within each 250m x 250m
  cell.

  \item We obtained geo-referenced data on roads from Open Street Map through
  Geofabrik (\url{www.geofabrik.de}; \citealp{Geofabrik.2019}). We only retained
  main tar roads and omitted smaller roads (see Table XX in Appendix XX) as
  these are scarcely frequented and do not represent an obstacle to dog
  movements \citep{Abrahms.2016}. Finally, we rasterized roads to a binary
  raster (1 = roads, 0 = no roads) with 250m x 250m resolution.

\end{enumerate}

\noindent To calculate a single proxy of human influence, we added up the layers
describing human density (continuous), farming (binary), and roads (binary). To
reduce the influence of outliers, totaled values were limited to a maximum of
50, which visually resulted in a good balance between high and low anthropogenic
influence and was therefore considered appropriate for our analysis. To render
the fact that humans influence their surroundings beyond their presence, we
followed \cite{Elliot.2014} and applied to each cell a 5 km spatial buffer
within which we summed up and log-transformed human-influence values (Figure XX
in Appendix XX).

\subsection{Habitat Selection Model}
We used an integrated step selection function (iSSF, \citealp{Avgar.2016}) to
investigate dispersers' selection or avoidance for the above mentioned spatial
covariates. In the iSSF framework, GPS relocations are coerced to  ``steps'',
where a step is defined as the connecting line between two consecutive GPS
relocations \citep{Turchin.1998}. Subsequently, habitat preferences are
estimated by contrasting covariates along realized steps to the same covariates
along alternative ``random''  steps \citep{Fortin.2005, Thurfjell.2014,
Avgar.2016}. Thus, we paired each realized step with 24 random steps. We
generated random steps by sampling turning angles from a uniform distribution
U(\(-\pi, \pi\)) and step lengths from a gamma distribution that was fitted
using realized steps \citep{Avgar.2016}. Together, a realized and its 24
associated random steps formed a stratum of 25 steps that received a unique
identifier. We extracted covariates along all steps (Table XX in Appendix XX),
standardized them using a z-score transformation, and checked for correlation
using Pearson's Correlation Coefficient. None of the covariates were overly
correlated (\(|r| > 0.6\), \citealp{Latham.2011}) and we retained all of them
for modeling. Our habitat selection was then based on the assumption that
dispersing wild dogs assigned a selection score \(w(x)\) of the following
exponential form to each realized and random step:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The selection score \(w(x)\) of a step depended on its associated
covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's preferences for
these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). To estimate habitat
preferences (i.e. the \(\beta\)'s) of interest, we used mixed effects
conditional logistic regression analysis as suggested by \cite{Muff.2019}. We
implemented their method using the R-package \textit{glmmTMB}
\citep{Mollie.2017} and used coalition ID to model random intercepts and slopes.
We defined two movement metrics, namely the cosine of the turning angle
(\(cos(ta)\)) and the log of the step length (\(log(sl)\)), as core covariates
and ran stepwise forward model selection based on Akaike's Information Criterion
(AIC, \citealp{Burnham.2002}) for all other covariates. The inclusion of the
movement metrics served to reduce biases in estimated habitat preferences that
may have arised due to movement preferences \citep{Avgar.2016}.

To validate the predictive power of the most parsimonious habitat selection
model, we ran k-fold cross-validation for case-control studies as described in
\cite{Fortin.2009}. This procedure compares rank frequencies of realized steps
under observed and random preferences and proofs a significant prediction in
case the confidence intervals of the spearman-rank correlation coefficients
\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap (Appendix
XX).

\subsection{Permeability Surface}
Using the most parsimonious selection model, we predicted a permeability surface
spanning the entire extent of the KAZA-TFCA. That is, we applied \Cref{EQ1} to
our spatial covariates and calculated the selection score \(w(x)\) for each
raster-cell in our study area. To reduce the influence of outliers in predicted
scores, we followed \cite{Squires.2013} and curtailed predicted scores between
the 1\textsuperscript{st} and 99\textsuperscript{th} percentile of their
original values. Because our floodmaps were dynamic, we collapsed all floodmaps
into a single static layer using areas that were covered by water in at least a
quarter of all our floodmaps. To compare permeability across different regions,
we rescaled the permeability surface to a range between 0 and 1. The most
impermeable landscape thus received a value of 0, the most permeable landscape a
value of 1.

\subsection{Least-Cost Paths \& Least-Cost Corridors}
To identify movement corridors within the extent of the KAZA-TFCA, we specified
source points and calculated factorial least-cost paths (LCPs), as well as
factorial least-cost corridors (LCCs), between them (Figure XX in Appendix XX;
\citealp{Adriaensen.2003, Sawyer.2011, Elliot.2014}. We generated source points
by overlaying the study area with a regular grid of points spaced at 100 km. We
only considered those points that fell within protected areas \(>\) 700
km\textsuperscript{2}, which conforms with home-range requirements of African
wild dogs reported in \cite{Pomilia.2015}. Finally, we defined centroids as
source points for those protected areas \(>\) 700 km\textsuperscript{2} that
were not assigned any source points from the regular grid. In total, we
generated 68 source points, which resulted in 2'278 unique pairwise combinations
and therefore 2'278 unique LCPs and LCCs.

We implemented factorial LCP analysis between source points using the R-package
\textit{gdistance} (Figure XX in Appendix XX, \citealp{vanEtten.2018}). The
package translated the (unscaled) permeability surface into a network of nodes
to find shortest effective distances between source points based on
probabilities of moving from cell to cell. In our case, the transition
probability of moving between two adjacent cells depended on their averaged
permeability. We allowed individuals to move from each cell to the cell's 8
surrounding neighbors (i.e. Moores neighborhood) and applied a geographic
correction to account for the fact that diagonal neighbors were more remote than
orthogonal neighbors. Because African wild dogs have been observed to cover
large dispersal distances \cite{DaviesMostert.2012, Masenga.2016, Cozzi.2020},
we did not limit LCPs to a maximal effective cost. After computation, we tallied
overlapping LCPs and identified high-frequency routes.

We calculated factorial LCCs \cite{Pinto.2009, Sawyer.2011, Elliot.2014}, again
using the R-package gdistance (Figure XX in Appendix XX,
\citealp{vanEtten.2018}). To identify LCCs, we first computed for each source
point a cumulative cost map, which indicated the total minimal costs required to
get from the source point to any other cell in the study area. An LCC between
two points was then obtained by adding up their cumulative cost maps and masking
out all cell-values exceeding the lowest cell-value by more than 5\%
\citep{Pinto.2009}. We repeated this procedure for each possible unique pairwise
combination of source points and thereby identified LCCs between all 68 selected
points. We normalized the resulting corridor-maps to range from zero to one and
merged them into a single connectivity map.

\section{Results}
\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table XX in Appendix XX) retained the covariates
\textit{Water}, \textit{DistanceToWater}, \textit{Trees},
\textit{Shrubs/Grassland}, and \textit{HumansBuff5000}, beside the fixed
covariates \textit{cos(ta)} and \textit{log(sl)} (\Cref{PermeabilityResults}a).
Parameter estimates showed that dispersing wild dogs moved in a directional and
fast manner, as indicated by a positive selection for small turning angles, i.e.
high \(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21,
\Cref{PermeabilityResults}a) and longer steps, i.e. high \(log(sl)\) (\(\beta\)
= 0.05, 95\% CI = 0.02 to 0.09, \Cref{PermeabilityResults}a). Dispersers avoided
moving through open water (\(\beta\) = -0.49, 95\% CI -0.67 to -0.30) but
selected for locations in its vicinity, although the latter effect was not
significant (\(\beta\) = -0.31, 95\% CI = -0.70 to 0.09,
\Cref{PermeabilityResults}a). Dispersers avoided areas that were densely covered
by trees (\(\beta\) = -0.17, CI = -0.33 to -0.01, \Cref{PermeabilityResults}a)
and preferred areas covered by shrubs/grassland (\(\beta\) = 0.44, 95\% CI =
0.30 to 0.58, \Cref{PermeabilityResults}a). Finally, dispersers avoided areas
that were influenced by human presence and activities (\(\beta\) = -0.41, 95\%
CI = -0.78 to -0.05, \Cref{PermeabilityResults}a).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. (b) Results from the k-fold
    cross validation for case-control studies. The left graph shows rank
    frequencies of \textit{realized} steps according to predictions, whereas the
    right graph shows rank frequencies of \textit{randomly selected} steps
    according to predictions. \(\bar{r}_s\) indicates the mean correlation
    coefficient resulting from 100 repetitions of the k-fold cross validation.
    The blue smoothing line was fitted using a locally weighted polynomial
    regression and serves to aid the eye in detecting the trends. Correlation
    coefficients suggest that our prediction was significant and robust,
    evidenced by the fact that the confidence intervals of \(\bar{r}_{s,
    realized}\) and \(\bar{r}_{s, random}\) did overlap and by the fact that
    there was strong and significant correlation between ranks and associated
    frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

Results from the k-fold cross-validation suggested that our prediction was
significant and robust, as highlighted by the fact that the 95\%-CIs intervals
of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) (0.04, CI = 0.01 to
0.08) did not overlap (\Cref{PermeabilityResults}b). Likewise, the significant
correlation between ranks and corresponding frequencies for realized steps
suggested a good fit between predictions and observations.

\subsection{Permeability Surface}
Our prediction of landscape permeability revealed substantial differences across
regions in the study area (\Cref{PermeabilityMap}). Comparisons of median
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
the boundaries of the KAZA-TFCA was more than two times as high as permeability
outside it. A comparison across countries showed that Angola and Botswana were
characterized by comparably highly permeable landscapes, Zimbabwe and Zambia
were relatively impermeable, and Namibia ranged in between the two extremes
(\Cref{PermeabilityComp}). Visual inspection of our covariate layers indicated
that high permeability in Botswana and Angola is mainly caused by a combination
of low human presence and activities, low tree cover, high shrubs/grassland
cover, and a close distance to water. Although swamps, wetlands, and permanent
water themselves provide little permeability, their surroundings act as strong
attractants to dispersers. The low permeability that characterizes Zambia and
Zimbabwe, on the other hand, is mainly caused by severe anthropogenic presence
and other human activities. Albeit the KAZA-TFCA covers a majority of
permeability hot-spots, several highly permeable regions remain uncovered by its
borders, offering exciting opportunities for future expansions of already
existing protected areas. Across all countries, protected areas provided higher
permeability than human-dominated landscapes. A visual assessment of our
permeability surface futhermore revealed that northern Botswana had the highest
density of permeable landscapes (\Cref{PermeabilityMap}, bright yellow colors),
although the waters of the Okavango Delta and Linyanti Swamp themselves provided
highly impermeable landscapes (\Cref{PermeabilityMap}, dark blue colors).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permability was predicted by calculating selection scores \(w(x) =
    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each pixel based on
    the pixel's underlying covariates (\(x_i\)) and estimated habitat
    preferences (\(\beta_i\)). Areas that dispersers find easy to traverse are
    depicted in bright colors. Bold white lines delineate the borders of the
    KAZA-TFCA, whereas dashed white lines show country borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Comparison of median permeability (interquantile range in brackets)
  across countries, separated according to areas within and outside the
  KAZA-TFCA, as well as within and outside formally protected areas. High values
  indicate high permeability, whereas low values correspond to low
  permeability.}
  \label{PermeabilityComp}
  \begin{tabular}{llllll}
    \toprule
    \multicolumn{1}{c}{ } & \multicolumn{2}{c}{KAZA-TFCA} & \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{ } \\
    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
    Country & Inside & Outside & Protected & Pastoral & Overall\\
    \midrule
    Angola & 0.41 (0.42) & 0.18 (0.35) & 0.41 (0.42) & 0.19 (0.35) & 0.26 (0.40) \\
    Botswana & 0.27 (0.30) & 0.15 (0.17) & 0.29 (0.35) & 0.16 (0.19) & 0.20 (0.26) \\
    Namibia & 0.23 (0.30) & 0.13 (0.18) & 0.24 (0.30) & 0.11 (0.15) & 0.16 (0.24) \\
    Zambia & 0.08 (0.11) & 0.04 (0.06) & 0.07 (0.12) & 0.04 (0.06) & 0.05 (0.08) \\
    Zimbabwe & 0.08 (0.20) & 0.06 (0.05) & 0.10 (0.21) & 0.06 (0.04) & 0.06 (0.08) \\
    \hline
    Overall & 0.19 (0.32) & 0.08 (0.16) & 0.18 (0.32) & 0.08 (0.17) & 0.11 (0.23)\\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Our least-cost analysis revealed three major movement corridors, all well-within
the KAZA-TFCA boundaries (\Cref{LeastCost}). Additionally, we found little to no
direct connectivity between peripheral points; that is, most paths and corridors
connecting two adjacent peripheral points run through more central regions
before heading towards their destination at the periphery (\Cref{LeastCost}).
One major corridor runs SE-NW and connects the Okavango-Linyanti ecosystem in
Botswana with Luengue-Luiana NP in Angola. One runs W-E between Chobe NP in
Botswana and Zimbabwe's Hwange NP. One runs NE-SW, completely across unprotected
areas, and connects Kafue NP in Zambia with more central regions of the
KAZA-TFCA. With 637 LCPs passing through it, the Okavango-Linyanti region is the
highest frequented corridor. Several minor routes branch off these three major
corridors; this includes a southward connection between the Okavango-Linyanti
and the Central Kalahari Game Reserve, a southwesterly corridor connecting
Luengue-Luiana NP with Namibia's Kaudom NP, and a northeasterly extension of the
Hwange corridor into Zimbabwe's Matusadona NP. Our model did not detect any
significant direct corridors between Zimbabwe and Zambia or Zambia and Angola,
and only a very limited W-E direct connection between the Okavango region and
Namibia's Khaudum NP. Except for the corridor into the Central Kalahari
National Park, our model did not detect any significant connectivity outside the
boundaries of KAZA-TFCA.

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (light grey). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference we also labeled some national parks (dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

\section{Discussion}
We used GPS relocation data collected on dispersing African wild dogs to
investigate whether major movement corridors are contained within the boundaries
of the world's largest transboundary conservation area. Our analysis suggests
that the KAZA-TFCA indeed encompasses all major movement corridors of African
wild dogs, demonstrating the potential value of such an initiative. However, our
approach is neither limited to the African wild dog, nor to our study area. All
covariates used throughout this study are readily available on a global scale
and many of them are likely to be important determinants of movement behavior,
landscape permeability, and connectivity for other species
\citep{Thurfjell.2014, Zeller.2012}. Interestingly, our predicted network of
least cost-paths and least-cost corridors for African wild dogs shows surprising
similarities to corridors of dispersing lions inhabiting the same ecosystem
\citep{Elliot.2014}. This not only reinforces confidence in our own predictions
but also suggests potential synergies for the conservation of these two, and
possibly more, species. Expanding this analytical framework to additional
species will yield important insights on the consistency of inter-specific
movement corridors, thus highlighting areas that are exceptionally valuable for
the conservation of several species, rather than focusing on single species.

We identified movement corridors between pre-defined start and end points
selected within already existing protected areas. This implicitly assumes that
individuals know the end point of their dispersal journey and that they have
almost complete knowledge of associated movement costs, which is clearly not the
case \citep{Abrahms.2017}. However, creation of pre-defined end points might not
be necessary, as it should be possible to simulate dispersal from known
source-points, yet without restricting the domain of potential end points
\citep{Signer.2017}. As a consequence, movement corridors would emerge naturally
as the result of a myriad of simulated dispersal events. While this approach is
conceptually straightforward, it requires a comprehensive mechanistic
understanding of dispersal movements, which is conditioned on our ability to
appropriately represent the landscape through which individuals disperse. An
approach that does not require the creation of end points will also be
instrumental in identifying corridors outside formally protected areas. In
contrast to classical corridor mapping, individual-based movement simulations
would also allow to take into account movement preferences of dispersers, and to
render the temporal dimension of dispersal. The latter consideration is
particularly important in ecosystems where connectivity changes in response to
seasonal variations in the landscape \citep{Osipova.2019}.

Our findings emphasize that human presence and activities constitute some of the
main barriers to connectivity, which is in line with findings from eastern
Africa by \cite{Masenga.2016} and \cite{Oneill.2020}. These results are,
however, in contrast with findings by \citep{DaviesMostert.2012} who reported a
high willingness of dispersers to cross human-dominated landscapes throughout
South Africa. We believe that such differences are due to the unavailability of
alternative  ``natural''  routes in South Africa. We identified water as
additional major obstacle to dispersal, which aligns with earlier studies
showing that water is almost impermeable for resident packs and only
infrequently crossed by dispersing individuals \citep{Abrahms.2017, Cozzi.2020}.
An accurate and dynamic representation of water is therefore paramount, and
particularly important for flood-pulsing ecosystems such as the Okavango Delta.
While we found that dispersers avoided moving through water, we revealed a
preference for moving close to water. We speculate that high prey-availability
in proximity of water may partly explain this result \citep{Western.1975,
Bonyongo.2005}. Following the same logic, however, apex predators (e.g. lions,
spotted hyenas) and resident wild dogs may also prefer proximity to water
\citep{Valeix.2009} forcing dispersing wild dogs away from it
\citep{Ndaimani.2016} to avoid aggressive interactions \citep{Creel.1996,
Mills.1997}. We could not control for the presence or absence of apex predators
or conspecifics during dispersal as we lacked corresponding data, but these
considerations might explain why we did not find a statistically clear effect of
distance to water. Given the influence that resident conspecific, predators,
competitors, and prey can have during dispersal \citep{Cozzi.2018,
Armansin.2019} we urge collecting and incorporating their respective spatial
information into analyses of landscape connectivity and to more precisely
reconstruct dispersal corridors.

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersers gain access to more peripheral regions of the
KAZA-TFCA. While this might partly be owed to the area's central position within
the KAZA-TFCA, it is likely also due to its unique environmental and
anthropogenic characteristics. In fact, the absence of human activities and the
presence of relatively impermeable water bodies, such as the Okavango Delta and
the Linyanti Swamp that potentially funnel dispersal movements, resulted in a
highly frequented corridor. The region is also instrumental for connecting wild
dog subpopulations in Zambia's Kafue and Zimbabwe's Hwange NPs. These
subpopulations are currently separated by the Zambezi River, forcing dispersers
to detour via northern Botswana to reach more peripheral areas. The key role of
the Okavango-Linyanti region for overall connectivity within the KATA-TFCA thus
calls for actions to secure its protection status in the future. While the
region is currently a Wildlife Management Area devoted to photographic tourism,
it has neither the status of a National Park nor that of a Game Reserve and so
potentially faces growing pressure towards exploitation of its natural resources
under the influences of an expanding human population, farming, and cropping
industry. A similar case of non-formally protected but key dispersal landscape
is represented by the area south of Kafue NP in Zambia, for which disruption of
its main and narrow dispersal corridor would result in considerable isolation of
its subpopulations.. We also revealed a potential southwards corridor between
the Okavango-Linyanti ecosystem and the Central Kalahari Game Reserve.
\cite{Elliot.2014} identified a similar corridor for dispersing lions,
suggesting that upholding and protecting a link between those ecosystems is
pivotal. Some areas through which the corridor runs are neither part of the
KAZA-TFCA nor profit from any form of protected status. In fact, human presence
and activities along the national road that longitudinally traverses this
corridor may limit realized connectivity \citep{Cozzi.2020}. In general, our
connectivity map suggests that connectivity gradually decreases towards the
borders of our study area, yet this might partly be a result of boundary effects
\citep{Koen.2010}.

Comparisons of landscape permeability across the extent of the KAZA-TFCA
revealed that northern Botswana and southern Angola provide relatively permeable
landscapes. Visual inspection of our covariate layers indicated that high
permeability in these countries is mainly caused by a combination of low human
presence and activities, low tree cover, high shrubs/grassland cover, and a
close distance to water. Although swamps, wetlands, and permanent water
themselves provide little permeability, their surroundings act as strong
attractants to dispersers. The low permeability that characterizes Zambia and
Zimbabwe, on the other hand, is mainly caused by severe anthropogenic presence
and other human activities. Albeit the KAZA-TFCA covers a majority of
permeability hot-spots, several highly permeable regions remain uncovered by its
borders, offering exciting opportunities for future expansions of already
existing protected areas.

%------------------------------------------------------------------------------
%	Gabs wanted to delete this
%------------------------------------------------------------------------------
% Because we predicted permeability and connectivity for such an extensive area,
% we were strongly limited in available and reliable land cover data. As a
% result, our representation of vegetation remained rather basic with only three
% vegetation classes (tree-cover, shrubs/grassland, bare land). Although some
% large-scale, high-resolution land cover datasets exist, such as Globeland's
% 30m \citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019}
% land cover datasets, we found their classifications often erroneous for our
% study area. This is presumably owed to the KAZA-TFCA's unique landscape and
% corresponding difficulties in remote sensing satellite imagery using
% traditional indices \citep{Wolski.2017}.

\section{Conclusion}
Our work showed that the majority of dispersal corridors for African wild dogs
are covered by the boundaries of the KAZA-TFCA, suggesting that the KAZA-TFCA
will significantly contribute to the long-term viability of this species.
Nevertheless, some unprotected dispersal routes do exist outside the KAZA-TFCA
and indicate opportunities for further expansions in the future. Moreover, our
connectivity network of least-cost paths and least-cost corridors highlights the
Okavango-Linyanti region as a central dispersal hub through which dispersers
gain access to more remote regions of the KAZA-TFCA. This stresses the
importance of northern Botswana for the conservation of this charismatic
species. Finally, our investigations showed that human pressures remain a severe
impediment to dispersal and therefore reduce landscape connectivity. Successful
conservation of the African wild dog is thus contingent on the willingness of
local authorities, policymakers, and land managers to preserve areas that are
largely liberated from human strains. Ultimately, our work contributes to the
growing field of connectivity literature and has important implications for the
management and conservation of the endangered African wild dog.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We thank Prof. J. Fieberg
for consulting on statistical aspects of this work and P. Wolski from the
Okavango Research Institute for assisting in generating dynamic flood maps. This
study was funded by Basler Stiftung für Biologische Forschung, Claraz
Foundation, Idea Wild, Jacot Foundation, National Geographic Society, Parrotia
Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-06-05.tex
first commit

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[nomarkers, nolists]{endfloat}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Bound within Boundaries: How Well Do Boundaries of Protected Areas Match
Movement Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David Hofmann\textsuperscript{1\S*} \and
  Dominik Behr\textsuperscript{1,2*} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared co-authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} Lycaon pictus, KAZA-TFCA, permeability surface, least-cost
corridors, integrated step selection function, landscape connectivity

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Protecting large portions of natural land connecting natural reserves is an
increasingly used strategy to maintain and restore connectivity among wildlife
subpopulations. One such initiative area is the Kavango-Zambezi Transfrontier
Conservation Area (KAZA-TFCA) in Africa, the world's largest transboundary
conservation area on earth, spanning five countries and 520'000
km\textsuperscript{2}. While boundaries of such conservation areas are often
determined by expert opinion as well as socio-political needs and constraints,
the extent to which these areas match species' movement corridors is rarely
examined. This is mainly due to a lack of adequate empirical data, particularly
on dispersal. Between 2011 and 2019, we collected high-resolution GPS relocation
data on 16 dispersing African wild dog coalitions (\textit{Lycaon pictus}) to
(i) measure differences in landscape permeability within and outside the
boundaries of the KAZA-TFCA and (ii) to assess to which degree the area of the
KAZA-TFCA reflects natural movement corridors of this highly mobile and
endangered species. We show that permeability within the KAZA-TFCA boundaries
was more than double compared to areas outside it. Furthermore, we identify a
five-fold permeability difference among the five KAZA-TFCA countries. These
differences were mainly caused by different degrees of human activities,
hampering dispersal movements. We further show that all major movement corridors
of wild dogs run within the KAZA-TFCA boundaries, although some minor corridors
remain outside formally protected areas. Our work highlights the importance of
empirically assessing the adequacy of protected areas and the need for
international coordination that stem from different contributions to
connectivity across regions and countries.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

\section{Introduction}
As a result of the ongoing human-caused habitat deterioration and fragmentation,
successful management and conservation of isolated subpopulations requires
maintenance and enhancement of connectivity between them \citep{Hanski.1998,
Fahrig.2003, Lindenmayer.2013}. Recently, efforts have been made to facilitate
animal movements and dispersal, and to secure connectivity by protecting
contiguous natural or semi-natural areas \citep{Heller.2009, Doerr.2011,
Rudnick.2012, Cozzi.2020}. These efforts have resulted in the creation of an
ever growing number of large and often transboundary protected areas
\citep{Wolmer.2003}. While boundaries of large scale conservation areas have
been drawn largely based on expert opinion and socio-political needs and
constraints, the usefulness of subjective boundary assessments have revealed
deficiencies in the past \citep{Pullinger.2010}. Due to the very nature of large
scale protected areas, which sometimes span several thousands of square
kilometers and traverse multiple countries, a thorough quantitative assessment
of their effectiveness and adequacy of their boundaries is challenging
\citep{Rudnick.2012}. This challenge is complicated by a lack of animal movement
data at the appropriate spatial scale and an insufficient understanding of the
ranging behavior, movement patterns, and dispersal corridors of the species for
which these protected areas have been created \citep{Vasudev.2015, Tshipa.2017,
Cozzi.2020}. Nevertheless, an accurate representation of movement corridors and
dispersal routes based on empirical global positioning system (GPS) data is
fundamental to objectively assess how well boundaries of protected areas reflect
natural animal movements \citep{Beier.2008, Doerr.2011, Rudnick.2012}.

In recent years, there has been a growing body of research that uses animal
relocation data to identify movement corridors and assess connectivity at large
scales \citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014,
Benz.2016, Osipova.2019}. Identification of potential corridors relies on the
estimation of permeability (respectively resistance) surfaces, which return the
ease or willingness at which the focal species traverses a specific landscape
\citep{Sawyer.2011}. Such surfaces are typically created on the basis of habitat
preferences, which can be estimated using a suite of selection functions
\citep{Boyce.2002, Fortin.2005, Cushman.2010}. In selection functions, habitat
preferences are inferred by comparing spatial covariates (e.g. environmental and
anthropogenic) at locations visited by the animal to the same spatial covariates
at randomly selected locations in the animal's availability domain
\citep{Zeller.2012, Thurfjell.2014}. Importantly, selection functions rely on
adequate landscape and GPS relocation data that are representative of the
process being studied \citep{Diniz.2020}. More specifically, GPS relocation data
collected on dispersing individuals has been shown to outperform GPS relocation
data collected on resident individuals when aiming at detecting large scale
dispersal corridors \citep{Elliot.2014, Diniz.2020}. However, few connectivity
studies have partitioned their GPS relocation data according to behavioral modes
(e.g. resident vs. dispersing) of the studied species \citep{Wilson.2012,
Vasudev.2015}. As such, most permeability surfaces upon which dispersal
corridors are identified are created using GPS relocation data that were
collected on resident individuals. This introduces severe biases and
substantially reduces the power to reveal meaningful movement corridors, for
dispersing individuals have different needs and drives compared to resident
individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}. Consequently, these
biases have limited our ability to assess the adequacy of the boundaries of
protected areas in securing connectivity for their protected species.

One initiative that aims at restoring and enhancing connectivity across large
scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA),
which, spanning over 520'000 km\textsuperscript{2} and five countries,
constitutes the world's largest transfrontier conservation area. While the
KAZA-TFCA was originally set to facilitate movements of African elephants
(\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key to the
conservation of other wide-ranging species such as African wild dogs
(\textit{Lycaon pictus}; \citealp{Woodroffe.2012, Cozzi.2020}), lions
(\textit{Panthera leo}; \citealp{Elliot.2014, Cushman.2018}), and cheetahs
(\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
have attempted to assess the adequacy of the KAZA-TFCA's boundaries using
relevant GPS movement data at the appropriate spatial scale \citep{Elliot.2014,
Tshipa.2017}. Thus, how well the KAZA-TFCA's boundaries reflect natural movement
patterns and dispersal corridors of its protected species is virtually unknown.

Across the KAZA-TFCA, the African wild dog (\textit{Lycaon pictus}) represents a
highly endangered flagship species for conservation efforts, but for which
systematically collected GPS relocation data on dispersing individuals are
unavailable. Once widespread across the entire Sub-Saharan continent, wild
dogs have been widely extirpated through human persecution, habitat destruction,
and disease outbreaks \citep{Woodroffe.2012}. As a result, the species has
become one of Africa's most endangered large carnivores, and currently only
survives in small, spatially scattered subpopulations \citep{Woodroffe.2012}.
Within these subpopulations, wild dogs form cooperative breeding packs of up to
thirty individuals \citep{Frame.1979, Fuller.1992, Creel.2002}, whose social
structure is strongly shaped by the process of dispersal \citep{McNutt.1996,
Woodroffe.2019, Behr.2020}. Both males and females typically disperse from their
natal pack, either alone or in same-sex dispersing coalitions, and search for
unrelated mates and a suitable territory to settle \citep{McNutt.1996,
Cozzi.2020, Behr.2020}. During dispersal, wild dogs can cover several hundred
kilometers, sometimes crossing national boundaries and areas that are strongly
influenced by humans \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019,
Cozzi.2020}. Despite the importance of dispersal for the long-term viability of
this species, little empirical information is available on habitat selection and
potential movement barriers of dispersing wild dogs. The few studies that have
collected GPS relocation data during dispersal have shown that dispersers
quickly move over large distances \citep{Woodroffe.2019}, generally avoid
human-dominated landscapes \citep{Masenga.2016, Oneill.2020, Cozzi.2020} and
areas densely forested \citep{Oneill.2020}, but prefer proximity to water and
dirt roads \citep{Oneill.2020}.

Here, we used GPS relocation data collected on 16 dispersing wild dogs in as
many dispersing coalitions from a free-ranging population in northern Botswana
to assess the adequacy of the boundaries of the KAZA-TFCA, the largest protected
area in the world. We estimated the degree of selection or avoidance for
environmental and anthropogenic landscape features. We then applied the obtained
habitat preferences to create a permeability surface spanning the entire
KAZA-TFCA and investigated how landscape permeability varies regionally and
internationally. Finally, we calculated least-cost paths and least-cost
corridors to identify major dispersal corridors, pinpoint dispersal hubs, and
verify whether the KAZA-TFCA successfully covers these features.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E; elevation ca.
1'030 m; red dotted rectangle in \Cref{StudyArea}a) was outlined by a
rectangular bounding box stretching over 1.3 Mio km\textsuperscript{2} and
encompassing the entire KAZA-TFCA (\Cref{StudyArea}b). The KAZA-TFCA lies in the
basins of the Okavango and Zambezi rivers and includes parts of Angola,
Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520'000
km\textsuperscript{2} it covers several already-existing national parks and
other protected areas. It constitutes the earth's largest transboundary
conservation area and is characterized by diverse landscapes, including savanna,
grassland, and dry or moist woodland habitats. Rainfall in the study area is
seasonal and lasts from November to March, whereas the rest of the year remains
dry \citep{Mendelsohn.2010}. The KAZA-TFCA also encloses the Okavango Delta,
which is one of the main strongholds of African wild dogs in southern Africa and
likely acts as a source for the recolonization of surrounding habitats
\citep{Woodroffe.2012, Cozzi.2013}. The Okavango Delta represents a highly
dynamic hydrological flood-pulsing system \citep{McNutt.1996, Wolski.2017}. The
extent of the flood in the delta greatly changes within and between years
depending on the amount of rain that descends from the catchment areas in Angola
and reaches the distal ends of the delta between July and August (see Figure S5
for an illustration of a typical flood pulse). The flood drastically affects
surrounding landscapes, so that during maximum extent (ca. 12'000
km\textsuperscript{2}) the delta becomes a patchy conglomerate of swamps, open
water, and islands, whereas these structures run dry when the flood retracts to
its minimum extent (ca. 5'000 km\textsuperscript{2}; \citealp{Gumbricht.2004,
Wolski.2017, McCarthy.1998}). Despite 36 NPs and other protected areas, there is
considerable human influence in some regions of the KAZA-TFCA, mainly
originating from farms, high human density, and road traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red dotted rectangle depicts
    the study area, which was confined by a bounding box encompassing the entire
    KAZA-TFCA. Gray areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
    area within which dispersal coalitions were collared. Since Game Reserves in
    Botswana virtually serve the same purpose as National Parks, we use the
    terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
We used a population of free-ranging African wild dogs inhabiting the Okavango
Delta in northern Botswana as a source population for dispersing individuals.
This population has been extensively studied since 1989 \citep{McNutt.1996,
McNutt.2008, Cozzi.2013, Cozzi.2020, Behr.2020}. Between 2011 and 2019, we
collected GPS relocation data on 16 coalitions of dispersing African wild dogs
(7 female and 9 male coalitions; see also \cite{Abrahms.2017} and
\cite{Cozzi.2020}). Candidate dispersing individuals were identified based on
criteria reported in \cite{Behr.2020}, immobilized according to protocols
described in \cite{Osofsky.1996}, and fitted with GPS/Satellite radio collars
(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while still in their
natal pack. All procedures were undertaken and supervised by a
Botswana-registered wildlife veterinarian. Immobilized individuals were
monitored and all were observed to successfully re-join their packs within one
hour after immobilization. During dispersal, GPS collars were programmed to
record GPS relocation data every 4 hours and to regularly transmit them via
Iridium satellite system to a base station. The system thus allowed
remote-tracking of collared individuals, even where field conditions were
prohibitive or dispersal trajectories crossed international borders.

Because we were interested in dispersal behavior, we discarded any GPS data
collected while individuals were still with their natal packs and also after
settlement in a new territory after dispersal \citep{Cozzi.2020}. We identified
the exact time of emigration and settlement based on direct field observations
and through visual inspection of the net squared displacement (NSD) metric. NSD
is a metric returning the euclidean distance of a relocation to a reference
point \citep{Borger.2012}, which in our case was the center of the dispersing
coalition's natal home range. Thus, dispersal was deemed to have started when a
coalition had left its natal home range and continued until the NSD metric
remained stationary, implying that the coalition had settled (Figure S1). Some
coalitions undertook exploratory movements prior to proper dispersal. Such
movements have been observed in other species and are reported to be very
similar to proper dispersal \citep{Killeen.2014}. Therefore, we also classified
exploratory movements as dispersal. In total, we collected 4'169 GPS relocations
during dispersal, i.e. between emigration and settlement (Figure S2 \& Table
S1), resulting in an average of 260 locations per dispersing coalition (range:
37 - 729). On average, coalitions dispersed for 48 days (min = 9, max = 137),
covered a mean euclidean distance of 54 km (min = 3, max = 263) and a mean
cumulative distance of 596 km (min = 130, max = 1'962). In our analysis, we did
not differentiate between male and female dispersing coalitions, for previous
research found little differences between sexes during dispersal
\citep{Woodroffe.2019, Cozzi.2020}.

\subsection{Spatial Covariates}
To investigate habitat preferences of dispersing wild dogs, we used a set of
geo-referenced covariates (\Cref{Covariates}) that we aggregated in the
categories \textit{land cover}, \textit{protection status}, and \textit{human
influence}. We did not include terrain features as covariates due to the absence
of noteworthy elevational gradients in our study area. For each covariate, we
prepared spatial raster layers from freely available online services or from
remotely sensed satellite imagery. To ensure a consistent resolution (i.e.
cell-size or grain) across covariates, we coarsened or interpolated all layers
to match a resolution of 250m x 250m. We performed processing and manipulation
of data as well as all spatial and statistical analyses using R, version 3.6.1
\citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of spatial covariates that we included in our models. We
    prepared all covariates for the entire study area but for better visibility
    we only plot them for the surroundings of the Okavango Delta. The white
    rectangle in each plot depicts the area within which dispersal coalitions
    were collared. (a) Averaged layer of all dynamic (binary) water maps. (b)
    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
    Anything that was not covered by trees or shrubs/grassland was deemed to be
    bare land. (d) Protection status of the area. (e) Human influence proxy
    composed of human density, farms, and roads. (f) Distance to nearest road
    (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsubsection{Land Cover}
We divided land cover into water and dryland. Water included rivers, wetlands,
and swamps. Because the inundation extent of the flood in the Okavango Delta is
highly variable within and between years, we created dynamic ``flood maps'' that
were updated each 8\textsuperscript{th} day following a remote sensing algorithm
developed by \citep{Wolski.2017}. The algorithm uses thresholding of short
wavelength infrared reflectances of MODIS Terra satellite imagery (MCD43A4;
\citealp{Schaaf.2015}) to distinguish areas covered by water and dryland
(details in Appendix A.3). While we created dynamic flood maps for the Okavango
Delta, we assumed the extent of all other water bodies (e.g. Chobe river,
Zambezi river) to be static within and between years. This static representation
was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
only retained the categories \textit{wetland} and \textit{water bodies} and
collectively reclassified them to \textit{water}. Globeland had an original
resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
of each 250m x 250m cell. We further improved river representation by employing
the rasterized MERIT Hydro dataset \citep{Yamazaki.2019} from which we added all
rivers with a width of over 10m to our Globeland layer. We merged dynamic and
static water maps into a large rasterstack, covering the entire study area. We
also created a rasterstack rendering the covariate \textit{distance to water} by
calculating the eculidean distance of each raster cell in the study area to the
nearest source of water.

We further subdivided dryland into three layers as derived from the MODIS Terra
Vegetation Continuous Fields dataset (MOD44B; \citealp{Dimiceli.2015}). The
three layers depicted percentage cover of tree-vegetation (henceforth
\textit{trees}), non-tree-vegetation (henceforth \textit{shrubs/grassland}), and
non-vegetated (henceforth \textit{bare land}) and added up to 100\% of dryland
coverage. We used our flood map that aligned with the creation date of these
MODIS layers and defined anything covered by water as 0\% vegetated. The MODIS
vegetation layers had a resolution of 250m x 250m and no coarsening or
interpolation was required.

\subsubsection{Protection Status}
We created a binary layer separating protected from unprotected land. We
downloaded corresponding data on protection status from the Peace Parks
Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}). Protected
areas included forest reserves, game reserves, wildlife management areas, and
national parks. We classified anything not covered by these categories as
unprotected (e.g. communal pastoral land, private land). We rasterized the two
categories to the binary raster \textit{protection status} (1 = protected, 0 =
pastoral) with a resolution of 250m x 250m.

\subsubsection{Human Influence}
We created a raster layer representing human influence by integrating
information on (1) human density, (2) farming, and (3) roads.

\begin{enumerate}[label = (\arabic*)]

  \item We obtained spatial human density estimates through a publicly available
  30m x 30m high-resolution population density dataset
  (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}). We coarsened the
  layer to 250m x 250m by summing up human density values within each 250m x
  250m cell.

  \item We sourced spatial information on farms from the Globeland
  \citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets from
  which we retained areas that were classified as either \textit{cultivated
  land} or \textit{croplands}. Any other land cover class was not pertinent to
  farming and therefore omitted. Because both layers had a resolution of 30m x
  30m we coarsened them to 250m x 250m by assigning a value of 1 to any 250m x
  250m cell that covered farmland and a value 0 otherwise. Thus, the final layer
  depicted presence (= 1) or absence (= 0) of farms within each 250m x 250m
  cell.

  \item We obtained geo-referenced data on roads from Open Street Map
  \citep{OpenStreetMap.2019}, downloaded through Geofabrik
  (\url{www.geofabrik.de}). We only retained main tarmac roads and omitted
  smaller roads (Table S2) as these are scarcely frequented and do not represent
  an obstacle to wild dog movements \citep{Abrahms.2016}. We rasterized main
  tarmac roads to the binary raster \textit{roads} (1 = roads, 0 = no roads)
  with 250m x 250m resolution. Finally, we created the covariate
  \textit{distance to roads} by calculating the eculidean distance of each
  raster cell in the study area to the nearest road.

\end{enumerate}

\noindent Because layers (1), (2), and (3) depicted features that are typically
spatially clustered and because not all dispersing coalitions moved within
meaningful distance to each of these features, we totaled the layers
\textit{human density}, \textit{farming}, and \textit{roads}. This resulted in a
single covariate called \textit{human influence}. Note that totalling the layers
implied that the three covariates received different weights (see Appendix A.6
for details). To reduce the influence of outliers, totaled values were limited
to a maximum of 50, which visually resulted in a good balance between high and
low anthropogenic influence and was therefore considered appropriate for our
analysis. To render the fact that humans influence their surroundings beyond
their presence, we followed \cite{Elliot.2014} and applied to each raster cell a
5 km focal buffer within which we summed up and log-transformed human-influence
values (Figure S6).

\subsection{Habitat Selection Model}
We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
investigate dispersers' selection or avoidance for the above-mentioned spatial
covariates. In the iSSF framework, GPS relocations are coerced to  ``steps'',
where a step is defined as the connecting line between two consecutive GPS
relocations \citep{Turchin.1998}. Subsequently, habitat preferences are
estimated by contrasting covariates along realized steps to the same covariates
along alternative ``random''  steps \citep{Fortin.2005, Thurfjell.2014,
Avgar.2016}. Thus, we paired each realized step with 24 random steps. We
generated random steps by sampling turning angles from a uniform distribution
U(\(-\pi, +\pi\)) and step lengths from a gamma distribution that was fitted
using realized steps \citep{Avgar.2016}. Together, a realized step and its 24
associated random steps formed a stratum of 25 steps that received a unique
identifier. Along all steps we extracted the above mentioned covariates (see
also Table S3). We also derived a binary covariate indicating whether or not a
step crossed a road and we identified the average distance of each step to the
nearest source of water and the nearest road. We square-rooted the values for
\textit{distance to roads} and \textit{distance to water} to render a decreasing
marginal impact of distance. Additionally, we derived a binary indicator of
whether a step crossed a road or not. We standardized extracted values using a
z-score transformation, and checked for correlation using Pearson's Correlation
Coefficient \(r\). None of the covariates were overly correlated (\(|r| > 0.6\);
\citealp{Latham.2011}) and we retained all of them for modeling. Our habitat
selection was then based on the assumption that dispersing wild dogs assigned a
selection score \(w(x)\) of the following exponential form to each realized and
random step \citep{Fortin.2005, Thurfjell.2014, Avgar.2016}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The selection score \(w(x)\) of a step depended on its associated
covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's preferences for
these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). To estimate habitat
preferences (i.e. the \(\beta\)'s) of interest, we used mixed effects
conditional logistic regression analysis as suggested by \cite{Muff.2020}. We
implemented their method using the R-package \textit{glmmTMB}
\citep{Mollie.2017} and used dispersing coalition ID to model random intercepts
and slopes. We defined two movement metrics, namely the cosine of the turning
angle (\(cos(ta)\)) and the log of the step length (\(log(sl)\)), as core
covariates and ran stepwise forward model selection based on Akaike's
Information Criterion (AIC; \citealp{Burnham.2002}) for all other covariates.
The inclusion of movement metrics served to reduce biases in estimated habitat
preferences that may have arised due to movement preferences \citep{Avgar.2016}.

To validate the predictive power of the most parsimonious habitat selection
model, we ran k-fold cross-validation for case-control studies as described in
\cite{Fortin.2009}. This validation procedure compares rank frequencies of
realized steps under observed and random preferences and proves a significant
prediction in case the confidence intervals of the spearman-rank correlation
coefficients \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not
overlap (details in Appendix A.8).

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\Cref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell in our study area. Because our flood maps were dynamic, we
collapsed all flood maps into a single static layer using areas that were
covered by water in at least a quarter of all our flood maps. We also calculated
a corresponding layer indicating the euclidean distance of each raster cell to
the nearest source of water on the static water map. To reduce the influence of
outliers in predicted permeability scores we followed \cite{Squires.2013} and
curtailed predicted scores between the 1\textsuperscript{st} and
99\textsuperscript{th} percentile of their original values. To compare
permeability across different regions, we rescaled the permeability surface to a
range between 0 and 1. The most impermeable landscape thus received a value of
0, the most permeable landscape a value of 1. Based on the permeability surface
we calculated median permeability values within and outside the KAZA-TFCA,
within and outside formally protected areas, and within each of the five
KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify movement corridors of dispersing wild dogs, we specified source
points and calculated factorial least-cost paths (LCPs) as well as factorial
least-cost corridors (LCCs) among all source points (Figure S.7;
\citealp{Adriaensen.2003, Sawyer.2011, Elliot.2014}. We generated source points
by overlaying the study area with a regular grid of points spaced at 100 km. We
only considered those points that fell within protected areas \(>\) 700
km\textsuperscript{2} as source points, which conforms with home-range
requirements of African wild dogs reported in \cite{Pomilia.2015}. Finally, we
defined centroids as source points for those protected areas \(>\) 700
km\textsuperscript{2} that were not assigned any source points from the regular
grid. In total, we generated 68 source points, which resulted in 2'278 unique
pairwise combinations and therefore 2'278 unique LCPs and LCCs.

We implemented factorial LCP analysis between source points using the R-package
\textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). The package translated
the (unscaled) permeability surface into a network of nodes to find shortest
effective distances between source points based on probabilities of moving from
cell to cell. In our case, the transition probability of moving between two
adjacent cells depended on their averaged permeability. We allowed individuals
to move from each cell to the cell's eight surrounding neighbors (i.e. Moores
neighborhood) and applied a geographic correction to account for the fact that
diagonal neighbors were more remote than orthogonal neighbors. Because African
wild dogs have been observed to cover large dispersal distances
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}, we did not limit LCPs to a
maximal effective cost. After computation, we tallied overlapping LCPs and
identified high-frequency routes.

We also calculated factorial LCCs \citep{Pinto.2009, Sawyer.2011, Elliot.2014},
again using the R-package \textit{gdistance} (Figure S.7;
\citealp{vanEtten.2017}). To identify LCCs, we first computed for each source
point a cumulative cost map, which indicated the total minimal costs required to
get from the source point to any other location in the study area. We then
obtained an LCC between two source points by adding up their cumulative cost
maps and masking out all cell-values exceeding the lowest cell-value by more
than 5\% \citep{Pinto.2009}. We repeated this procedure for each possible unique
pairwise combination of source points and thereby identified LCCs between all 68
selected source points. We normalized the resulting corridor-maps to range from
zero to one and tallied them into a single connectivity map.

\section{Results}
\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)} and
\textit{log(sl)} (\Cref{PermeabilityResults}a). Parameter estimates showed that
dispersing wild dogs moved in a directional and fast manner, as indicated by a
positive selection for small turning angles, i.e. high \(cos(ta)\) (\(\beta\) =
0.14; 95\% CI = 0.07 to 0.21, \Cref{PermeabilityResults}a) and longer steps,
i.e. high \(log(sl)\) (\(\beta\) = 0.06, 95\% CI = 0.02 to 0.09,
\Cref{PermeabilityResults}a). Dispersers avoided moving through water (\(\beta\)
= -0.52, 95\% CI -0.77 to -0.26, \Cref{PermeabilityResults}a) but selected for
locations in its vicinity, although the latter effect was not significant
(\(\beta\) = -0.32, 95\% CI = -0.72 to 0.08, \Cref{PermeabilityResults}a).
Dispersers avoided areas that were densely covered by trees (\(\beta\) = -0.31,
CI = -0.46 to -0.15, \Cref{PermeabilityResults}a) and preferred areas covered by
shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42,
\Cref{PermeabilityResults}a). Finally, dispersers avoided areas that were
influenced by human presence and activities (\(\beta\) = -0.41, 95\% CI = -0.78
to -0.05, \Cref{PermeabilityResults}a).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. (b) Results from the k-fold
    cross validation for case-control studies. The left graph shows rank
    frequencies of \textit{realized} steps according to predictions, whereas the
    right graph shows rank frequencies of \textit{randomly selected} steps
    according to predictions. \(\bar{r}_s\) indicates the mean correlation
    coefficient resulting from 100 repetitions of the k-fold cross validation.
    The blue smoothing line was fitted using a locally weighted polynomial
    regression and serves to aid the eye in detecting the trends. Correlation
    coefficients suggest that our prediction was significant and robust,
    evidenced by the fact that the confidence intervals of \(\bar{r}_{s,
    realized}\) and \(\bar{r}_{s, random}\) did overlap and by the fact that
    there was strong and significant correlation between ranks and associated
    frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

Results from the k-fold cross-validation suggested that our prediction was
significant and robust, as highlighted by the fact that the 95\%-CIs intervals
of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap
(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations (\Cref{PermeabilityResults}b).

\subsection{Permeability Surface}
Our prediction of landscape permeability revealed substantial differences across
regions in the study area (\Cref{PermeabilityMap}). Comparisons of median
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
KAZA-TFCA is more than two times as high as permeability outside it.
Permeability varies by country and reveals a five-fold permeability difference
among them. Angola and Botswana are characterized by comparably highly permeable
landscapes, Zimbabwe and Zambia are relatively impermeable, and Namibia ranges
in between the two extremes (\Cref{PermeabilityComp}). Visual inspection of our
covariate layers indicated that high permeability in Botswana and Angola is
mainly caused by a combination of low human presence and activities, low tree
cover, high shrubs/grassland cover, and a close distance to water. Although
swamps, wetlands, and permanent water themselves provide little permeability,
their surroundings act as strong attractants to dispersers. The low permeability
that characterizes Zambia and Zimbabwe, on the other hand, is mainly caused by
substantial human influences. Albeit the KAZA-TFCA covers a majority of
permeability hot-spots, several highly permeable regions remain uncovered by its
borders. Across all countries, protected areas provide roughly double the
permeability of human-dominated landscapes (\Cref{PermeabilityComp}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permeability was predicted by calculating selection scores \(w(x) =
    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each raster cell
    based on the raster cell's underlying covariates (\(x_i\)) and estimated
    habitat preferences (\(\beta_i\)). Areas that dispersers find easy to
    traverse are depicted in bright colors. Bold white lines delineate the
    borders of the KAZA-TFCA, whereas dashed white lines show country borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Comparison of median permeability (interquantile range in brackets)
  across countries, separated into areas within and outside the KAZA-TFCA, as
  well as within and outside formally protected areas. High values indicate high
  permeability, whereas low values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{tabular}{llllll}
    \toprule
    \multicolumn{1}{c}{} & \multicolumn{2}{c}{KAZA-TFCA} &
    \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{} \\
    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
    Country & Inside & Outside & Protected & Pastoral & Overall\\
    \midrule
    Angola & 0.36 (0.41) & 0.12 (0.32) & 0.36 (0.41) & 0.12 (0.33) & 0.20 (0.39)\\
    Botswana & 0.25 (0.30) & 0.15 (0.16) & 0.28 (0.35) & 0.15 (0.18) & 0.19 (0.25)\\
    Namibia & 0.22 (0.30) & 0.13 (0.18) & 0.24 (0.30) & 0.11 (0.15) & 0.16 (0.24)\\
    Zambia & 0.05 (0.09) & 0.03 (0.06) & 0.04 (0.10) & 0.03 (0.05) & 0.03 (0.07)\\
    Zimbabwe & 0.07 (0.16) & 0.06 (0.05) & 0.08 (0.17) & 0.05 (0.05) & 0.06 (0.07)\\
    \hline
    Overall & 0.16 (0.30) & 0.07 (0.15) & 0.15 (0.30) & 0.07 (0.15) & 0.10 (0.22)\\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Our least-cost analysis revealed three major movement corridors of which all
were well-within the KAZA-TFCA boundaries (\Cref{LeastCost}). Furthermore, we
found little to no direct connectivity between peripheral points; that is, most
paths and corridors connecting two adjacent peripheral points run through more
central regions before heading towards their destination at the periphery
(\Cref{LeastCost}). One major corridor runs SE-NW and connects the
Okavango-Linyanti ecosystem in Botswana with Luengue-Luiana NP in Angola. A
further corridor runs W-E between Chobe NP in Botswana and Zimbabwe's Hwange NP.
A third major corridor runs NE-SW, completely across unprotected areas, and
connects Kafue NP in Zambia with more central regions of the KAZA-TFCA. Several
minor corridors branch off from these three major corridors; these include a
southward connection between the Okavango-Linyanti and the Central Kalahari Game
Reserve, a southwesterly corridor connecting Luengue-Luiana NP with Namibia's
Kaudom NP, and a northeasterly extension of the Hwange corridor into Zimbabwe's
Matusadona NP. According to our predictions, the corridor towards Kafue NP and
the corridor passing the Okavango-Linyanti region are the highest frequented
routes within KAZA (maybe numbers?). Our model did not detect any significant
direct corridors between Zimbabwe and Zambia or Zambia and Angola, and only a
very limited W-E direct connection between the Okavango region and Namibia's
Khaudum NP. Except for the corridor into the Central Kalahari National Park, our
model did not detect any significant connectivity outside the boundaries of
KAZA-TFCA.

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (light grey). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference, we also labeled some national parks (dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

\section{Discussion}
We used GPS relocation data collected on dispersing African wild dogs to
investigate whether their main movement corridors are contained within the
boundaries of the world's largest transboundary conservation area, namely the
KAZA-TFCA. Our analysis suggests that the KAZA-TFCA indeed encompasses all major
movement corridors of African wild dogs, demonstrating the potential value of
such an initiative. However, our approach is neither limited to the African wild
dog, nor to our study area. All covariates used throughout this study are
readily available on a global scale and many of them are likely to be important
determinants of movement behavior, landscape permeability, and connectivity for
other species \citep{Thurfjell.2014, Zeller.2012}. Interestingly, our predicted
network of least cost-paths and corridors for African wild dogs shows surprising
similarities to corridors of dispersing lions inhabiting the same ecosystem
\citep{Elliot.2014}. This not only reinforces confidence in our own predictions
but also suggests potential synergies for the conservation of these two, and
possibly more, species. Expanding this analytical framework to additional
species will yield important insights on the consistency of inter-specific
movement corridors, thus highlighting areas that are exceptionally valuable for
the conservation of several species.

Our approach of identifying movement corridors based on pre-defined start and
end points implicitly assumes that individuals know the end point of their
dispersal journey and that they have almost complete knowledge of associated
movement costs. Considering the fact that dispersers often move into unknown
territory \citep{Woodroffe.2019, Cozzi.2020}, this may not necessarily be the
case \citep{Abrahms.2017, Woodroffe.2019}. However, specification of pre-defined
end points might not be necessary, as it should be possible to simulate
dispersal from known source-points, yet without restricting the domain of
potential end points \citep{Signer.2017}. As a consequence, movement corridors
would emerge naturally as the result of a myriad of simulated dispersal events.
While this approach is conceptually straightforward, it requires a comprehensive
mechanistic understanding of dispersal movements, which is conditional on our
ability to collect additional dispersal data and adequately represent the
landscape through which individuals disperse. An approach that does not require
the creation of end points will also be instrumental in identifying corridors
outside formally protected areas. In contrast to classical corridor mapping,
individual-based movement simulations would further allow taking into account
movement preferences of dispersers as well as rendering the temporal aspect of
dispersal (for instance to examine how quickly dispersers are able to move
between locations). The latter consideration is particularly important in
ecosystems where connectivity may change in response to seasonal variations in
the landscape \citep{Osipova.2019}.

Our results emphasize that human presence and activities constitute some of the
main barriers to connectivity. Our result that human influence is detrimental to
dispersal conforms to findings in wild dog populations from eastern Africa
\citep{Masenga.2016, Oneill.2020} but conflicts with findings by
\cite{DaviesMostert.2012}, who reported a high willingness of dispersers to
cross human-dominated landscapes throughout South Africa. We believe that such
differences are due to the unavailability of alternative ``natural'' routes,
which may have forced dispersers to cross human dominated landscapes despite a
strong aversion to do so. This may hint at a considerable ability of dispersing
wild dogs to adapt to prevailing habitat condition and could imply a limited
model transferability between dispersersers from our study system to dispersers
found in other ecosystems, such as in South Africa \citep{Whittington.2011,
DaviesMostert.2012} or Kenya \citep{Woodroffe.2019, Oneill.2020}. However, we
believe that our study population adequately represents wild dog populations
residing within KAZA-TFCA where more natural dispersal routes are available.
Successful conservation of this species is thus contingent on policymakers' and
local authorities' ability and willingness to provide and conserve areas that
remain free from anthropogenic pressures. This is not only paramount in light of
providing connectivity and facilitating dispersal, but also in terms of reducing
human-caused mortality during dispersal. In fact, previous studies have shown
that human-caused mortality during dispersal, especially in proximity to
settlements and roads, represents a major threat to the dispersal ability of
this species \citep{Woodroffe.2019, Cozzi.2020}.

Despite anthropogenic pressures, we identified water as another major obstacle
to dispersal. This is consistent with earlier studies showing that water bodies
are almost impenetrable by resident packs \citep{Abrahms.2017, Cozzi.2020} and
only infrequently crossed by dispersing individuals \citep{Cozzi.2020}. An
accurate and dynamic representation of water is thus imperative and particularly
relevant in flood-pulsing ecosystems such as the Okavango Delta. Although
dispersers avoided moving through water, we revealed a preference for moving in
proximity to water. This preference is likely due to high prey-availability in
vicinity of water \citep{Western.1975, Bonyongo.2005}. However, by the same
reasoning apex predators (e.g. lions, spotted hyenas) and resident wild dogs may
prefer proximity to water \citep{Valeix.2010} and occasionally force dispersing
wild dogs to move into prey-poor areas that are more distant to water
\citep{Ndaimani.2016} to avoid aggressive intraguild interactions
\citep{Creel.1996, Mills.1997, Webster.2012}. We could not control for the
presence or absence of apex predators or conspecifics during dispersal as we
lacked corresponding data, but these considerations may explain why we did not
find a significant effect of \textit{distance to water}. Given the influence
that resident conspecifics, predators, competitors, and prey can have during
dispersal \citep{Cozzi.2018, Armansin.2019} future studies should strive to
collect and incorporate corresponding spatial information into analyses of
landscape connectivity to more precisely reconstruct dispersal corridors.

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersersing wild dogs gain access to more peripheral regions of
the KAZA-TFCA. While this might partly be owed to the area's central position
within the KAZA-TFCA, it is likely also due to its unique environmental and
anthropogenic characteristics. It appears that the absence of human activities
and the presence of relatively impermeable water bodies, such as the Okavango
Delta and the Linyanti Swamp funnel dispersal movements, resulting in a highly
frequented corridor. The region is also instrumental for connecting wild dog
subpopulations in Zambia's Kafue and Zimbabwe's Hwange NPs. These subpopulations
are currently separated by the Zambezi River, forcing dispersers to detour via
northern Botswana to reach more peripheral areas. The key role of the
Okavango-Linyanti region for overall connectivity within the KATA-TFCA thus
calls for actions to secure its protection status in the future. While the
region is currently a Wildlife Management Area devoted to photographic tourism,
it has neither the status of a National Park nor that of a Game Reserve and so
potentially faces growing pressure towards exploitation of its natural resources
under the influences of an expanding human population, farming, and cropping
industry. A similar case of non-formally protected but key dispersal landscape
is represented by the area south of Kafue NP in Zambia, for which a disruption
of its main and narrow dispersal corridor would result in considerable isolation
of its subpopulations. We also revealed a potential southwards corridor between
the Okavango-Linyanti ecosystem and the Central Kalahari National Park.
\cite{Elliot.2014} identified a similar corridor for dispersing lions,
suggesting that upholding and protecting a link between those ecosystems is
pivotal. Some areas through which the corridor runs are neither part of the
KAZA-TFCA nor profit from any form of protected status. In fact, human presence
and activities along the national road that longitudinally traverses this
corridor may limit realized connectivity \citep{Cozzi.2020}. Finally, our
connectivity map suggested that connectivity within the extent of the KAZA-TFCA
gradually decreases towards the borders of our study area, yet this might partly
be a result of boundary effects \citep{Koen.2010}.

%------------------------------------------------------------------------------
%	Gabs wanted to delete this
%------------------------------------------------------------------------------
% Because we predicted permeability and connectivity for such an extensive area,
% we were strongly limited in available and reliable land cover data. As a
% result, our representation of vegetation remained rather basic with only three
% vegetation classes (tree-cover, shrubs/grassland, bare land). Although some
% large-scale, high-resolution land cover datasets exist, such as Globeland's
% 30m \citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019}
% land cover datasets, we found their classifications often erroneous for our
% study area. This is presumably owed to the KAZA-TFCA's unique landscape and
% corresponding difficulties in remote sensing satellite imagery using
% traditional indices \citep{Wolski.2017}.

\section{Conclusion}
Our work showed that the majority of dispersal corridors for African wild dogs
are covered by the boundaries of the KAZA-TFCA, suggesting that the KAZA-TFCA
will significantly contribute to the long-term viability of this species.
Moreover, our connectivity network highlights the Okavango-Linyanti region as a
central dispersal hub through which dispersers gain access to more remote
regions of the KAZA-TFCA. This fidning emphasizes the importance of northern
Botswana for the conservation of this charismatic species. Finally, our
investigations showed that human influence constitutes one of the main barriers
to dispersal and thus reduces landscape connectivity. Successful conservation of
the wild dogs is therefore contingent on the willingness of local authorities,
policymakers, and land managers to preserve areas that are largely liberated
from human strains. Ultimately, our work contributes to the growing field of
connectivity literature and has important implications for the management and
conservation of the highly endangered African wild dog.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Prof. J. Fieberg, who consulted all statistical aspects of this work and P.
Wolski from the Okavango Research Institute who assisted us in generating
dynamic flood maps. This study was funded by Basler Stiftung für Biologische
Forschung, Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic
Society, Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust
Foundation, Forschungkredit der Universität Zürich, and a Swiss National Science
Foundation Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-06-08.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-06-05.tex
@@
 \usepackage[nameinlink]{cleveref}
-% \usepackage{lineno}
+\usepackage{lineno}
 \usepackage{multirow}
@@
 
+% Change how equations are referenced
+\renewcommand{\theequation}{Equation \arabic{equation}}%
+
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-\title{Bound within Boundaries: How Well Do Boundaries of Protected Areas Match
-Movement Corridors of Their Most Mobile Protected Species?}
+\title{Bound within Boundaries: How Well Do Protected Areas Match Movement
+Corridors of Their Most Mobile Protected Species?}
 
@@
 \author{
-  David Hofmann\textsuperscript{1\S*} \and
-  Dominik Behr\textsuperscript{1,2*} \and
+  David D. Hofmann\textsuperscript{1,\S,*} \and
+  Dominik M. Behr\textsuperscript{1,2,*} \and
   John W. McNutt\textsuperscript{2} \and
@@
 
-\textsuperscript{*} Shared co-authorship
+\textsuperscript{*} Shared first authorship
 
@@
 \begin{abstract}
-Protecting large portions of natural land connecting natural reserves is an
-increasingly used strategy to maintain and restore connectivity among wildlife
-subpopulations. One such initiative area is the Kavango-Zambezi Transfrontier
-Conservation Area (KAZA-TFCA) in Africa, the world's largest transboundary
-conservation area on earth, spanning five countries and 520'000
-km\textsuperscript{2}. While boundaries of such conservation areas are often
-determined by expert opinion as well as socio-political needs and constraints,
-the extent to which these areas match species' movement corridors is rarely
-examined. This is mainly due to a lack of adequate empirical data, particularly
-on dispersal. Between 2011 and 2019, we collected high-resolution GPS relocation
-data on 16 dispersing African wild dog coalitions (\textit{Lycaon pictus}) to
-(i) measure differences in landscape permeability within and outside the
-boundaries of the KAZA-TFCA and (ii) to assess to which degree the area of the
-KAZA-TFCA reflects natural movement corridors of this highly mobile and
-endangered species. We show that permeability within the KAZA-TFCA boundaries
-was more than double compared to areas outside it. Furthermore, we identify a
+\noindent \textbf{1. Background}: Protecting large portions of natural land
+connecting natural reserves is an increasingly used strategy to maintain and
+restore connectivity among wildlife subpopulations. One such initiative is the
+Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA) in Africa, the
+world's largest transboundary conservation area, spanning five countries and
+520'000 km\textsuperscript{2}. While boundaries of such conservation areas are
+often determined by expert opinion as well as socio-political needs and
+constraints, the extent to which they match species' movement corridors is
+rarely examined. This is mainly due to a lack of adequate empirical data,
+particularly on movement behavior of wide-ranging species during dispersal.
+
+\noindent \textbf{2. Methods}: Between 2011 and 2019, we collected
+high-resolution GPS relocation data on 16 coalitions of dispersing African wild
+dogs (\textit{Lycaon pictus}) from a free ranging population in northern
+Botswana. We combined the collected data with relevant spatial covariates and
+used integrated step selection analysis to estimate habitat preferences during
+dispersal and create a permeability surface spanning the entire KAZA-TFCA. To
+examine whether the KAZA-TFCA covers major movement corridors of this highly
+mobile species, we compared landscape permeability across different regions of
+our study area. We also calculated least-cost paths and corridors between
+selected locations and verified that corridors are encompassed within the
+boundaries of the KAZA-TFCA.
+
+\noindent \textbf{3. Results}: We show that permeability within the KAZA-TFCA is
+more than double compared to areas outside it. Furthermore, we report a
 five-fold permeability difference among the five KAZA-TFCA countries. These
-differences were mainly caused by different degrees of human activities,
-hampering dispersal movements. We further show that all major movement corridors
-of wild dogs run within the KAZA-TFCA boundaries, although some minor corridors
-remain outside formally protected areas. Our work highlights the importance of
-empirically assessing the adequacy of protected areas and the need for
-international coordination that stem from different contributions to
-connectivity across regions and countries.
+differences are mainly caused by different degrees of human activities across
+regions, hampering dispersal movements. We further present a connectivity
+network, consisting of least-cost paths and corridors. We thereby show that all
+major movement corridors of wild dogs run within the KAZA-TFCA boundaries,
+although some minor corridors remain outside formally protected areas.
+
+\noindent \textbf{4. Synthesis and Applications}: Our work highlights the
+importance of empirically assessing the adequacy of protected areas and the need
+for international coordination that stem from different contributions to
+connectivity across regions and countries. Moreover, the presented framework is
+readily expandable to include additional species and regions in order to
+highlight consistencies among potential dispersal corridors as well as dispersal
+corridors outside our ecosystem.
+
 \end{abstract}
@@
 % Create linenumbers
-% \linenumbers
+\linenumbers
 
 \section{Introduction}
-As a result of the ongoing human-caused habitat deterioration and fragmentation,
-successful management and conservation of isolated subpopulations requires
-maintenance and enhancement of connectivity between them \citep{Hanski.1998,
-Fahrig.2003, Lindenmayer.2013}. Recently, efforts have been made to facilitate
-animal movements and dispersal, and to secure connectivity by protecting
-contiguous natural or semi-natural areas \citep{Heller.2009, Doerr.2011,
-Rudnick.2012, Cozzi.2020}. These efforts have resulted in the creation of an
-ever growing number of large and often transboundary protected areas
-\citep{Wolmer.2003}. While boundaries of large scale conservation areas have
-been drawn largely based on expert opinion and socio-political needs and
-constraints, the usefulness of subjective boundary assessments have revealed
-deficiencies in the past \citep{Pullinger.2010}. Due to the very nature of large
-scale protected areas, which sometimes span several thousands of square
-kilometers and traverse multiple countries, a thorough quantitative assessment
+Connectivity among subpopulations is a crucial pre-requisite for many species to
+thrive and persist \citep{Fahrig.2003}. A sufficient degree of connectivity
+enables dispersal, which in turn contributes to population viability through
+gene-flow, the ability to reinforce and rescue small populations, to colonize or
+recolonize unoccupied habitats, as well as to shift range in response to climate
+change \citep{Brown.1977, Hanski.1998, MacArthur.2001, Frankham.2002,
+Heller.2009, Leigh.2012}. Particularly in light of ongoing habitat fragmentation
+and destruction worldwide \citep{Fahrig.2003, Lindenmayer.2013}, the protection
+of natural or semi natural areas that serve as movement corridors and provide
+connectivity has become an utmost task for conservation management
+\citep{Heller.2009, Doerr.2011, Rudnick.2012, Cozzi.2020}. This has resulted in
+an ever-growing number of large and often transboundary protected areas that are
+set to facilitate dispersal between habitats \citep{Wolmer.2003}. Due to the
+very nature of large scale protected areas, a thorough quantitative assessment
 of their effectiveness and adequacy of their boundaries is challenging
-\citep{Rudnick.2012}. This challenge is complicated by a lack of animal movement
-data at the appropriate spatial scale and an insufficient understanding of the
-ranging behavior, movement patterns, and dispersal corridors of the species for
-which these protected areas have been created \citep{Vasudev.2015, Tshipa.2017,
-Cozzi.2020}. Nevertheless, an accurate representation of movement corridors and
-dispersal routes based on empirical global positioning system (GPS) data is
-fundamental to objectively assess how well boundaries of protected areas reflect
-natural animal movements \citep{Beier.2008, Doerr.2011, Rudnick.2012}.
+\citep{Rudnick.2012}. In result, boundaries are typically drawn according to
+expert opinion and socio-political needs rather than empirical information. In
+comparison to qantitative assessments, however, such qualitative approaches have
+revealed deficiencies in the past and potentially hold the risk of misdirecting
+valuable and scarce conservation funds \citep{Pullinger.2010, Sawyer.2011}. An
+accurate representation of movement corridors and dispersal routes based on
+empirical data is thus fundamental to objectively assess how well boundaries of
+existing or planned protected areas reflect natural movements of their protected
+species \citep{Beier.2008, Doerr.2011, Rudnick.2012}.
 
-In recent years, there has been a growing body of research that uses animal
-relocation data to identify movement corridors and assess connectivity at large
-scales \citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014,
-Benz.2016, Osipova.2019}. Identification of potential corridors relies on the
-estimation of permeability (respectively resistance) surfaces, which return the
-ease or willingness at which the focal species traverses a specific landscape
-\citep{Sawyer.2011}. Such surfaces are typically created on the basis of habitat
-preferences, which can be estimated using a suite of selection functions
-\citep{Boyce.2002, Fortin.2005, Cushman.2010}. In selection functions, habitat
-preferences are inferred by comparing spatial covariates (e.g. environmental and
-anthropogenic) at locations visited by the animal to the same spatial covariates
-at randomly selected locations in the animal's availability domain
+In recent years, a growing body of research has used animal relocation data to
+identify movement corridors and assess connectivity at large scales
+\citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014, Benz.2016,
+Osipova.2019}. Identification of potential corridors typically relies on the
+estimation of permeability surfaces, which return the ease or willingness at
+which the focal species traverses a specific landscape \citep{Sawyer.2011}. Such
+surfaces are typically created on the basis of a species' habitat preferences,
+which can be quantified using a suite of selection functions \citep{Boyce.2002,
+Fortin.2005, Cushman.2010, Zeller.2012}. In selection functions, habitat
+preferences are estimated by comparing spatial covariates (e.g. environmental
+and anthropogenic) at locations visited by the animal to the same spatial
+covariates at randomly selected locations in the animal's availability domain
 \citep{Zeller.2012, Thurfjell.2014}. Importantly, selection functions rely on
-adequate landscape and GPS relocation data that are representative of the
-process being studied \citep{Diniz.2020}. More specifically, GPS relocation data
-collected on dispersing individuals has been shown to outperform GPS relocation
-data collected on resident individuals when aiming at detecting large scale
-dispersal corridors \citep{Elliot.2014, Diniz.2020}. However, few connectivity
-studies have partitioned their GPS relocation data according to behavioral modes
-(e.g. resident vs. dispersing) of the studied species \citep{Wilson.2012,
-Vasudev.2015}. As such, most permeability surfaces upon which dispersal
-corridors are identified are created using GPS relocation data that were
-collected on resident individuals. This introduces severe biases and
-substantially reduces the power to reveal meaningful movement corridors, for
-dispersing individuals have different needs and drives compared to resident
-individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}. Consequently, these
-biases have limited our ability to assess the adequacy of the boundaries of
-protected areas in securing connectivity for their protected species.
+adequate landscape and relocation data that are representative of the process
+being studied \citep{Diniz.2020}. More specifically, relocation data collected
+on dispersing individuals has been shown to outperform relocation data collected
+on resident individuals, especially when the aim is to detect large scale
+dispersal corridors \citep{Elliot.2014, Diniz.2020}. However, dispersal data is
+inherently difficult to collect and few connectivity studies have partitioned
+their data according to behavioral modes (e.g. resident vs. dispersing) of the
+studied species \citep{Wilson.2012, Vasudev.2015}. As such, most permeability
+surfaces upon which dispersal corridors are identified are created using
+relocation data that were collected on resident individuals. This introduces
+severe biases and substantially reduces the power to reveal meaningful movement
+corridors, for dispersing individuals have different needs and drives compared
+to resident individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}.
+Consequently, these biases have limited our ability to assess the effectiveness
+of protected areas in securing connectivity for their protected species.
 
 One initiative that aims at restoring and enhancing connectivity across large
-scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA),
-which, spanning over 520'000 km\textsuperscript{2} and five countries,
-constitutes the world's largest transfrontier conservation area. While the
-KAZA-TFCA was originally set to facilitate movements of African elephants
-(\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key to the
-conservation of other wide-ranging species such as African wild dogs
+scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), which
+constitutes the world's largest transfrontier conservation area, spanning over
+520'000 km\textsuperscript{2} and five countries (\url{www.kavangozambezi.org}).
+While the KAZA-TFCA was originally set to facilitate movements of African
+elephants (\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key
+to the conservation of other wide-ranging species such as African wild dogs
 (\textit{Lycaon pictus}; \citealp{Woodroffe.2012, Cozzi.2020}), lions
@@
 (\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
-have attempted to assess the adequacy of the KAZA-TFCA's boundaries using
-relevant GPS movement data at the appropriate spatial scale \citep{Elliot.2014,
-Tshipa.2017}. Thus, how well the KAZA-TFCA's boundaries reflect natural movement
-patterns and dispersal corridors of its protected species is virtually unknown.
+have attempted to assess the adequacy of the KAZA-TFCA using relevant global
+positioning system (GPS) relocation data of its protected species at the
+appropriate spatial scale \citep{Elliot.2014, Tshipa.2017}. Thus, how well the
+boundaries of the KAZA-TFCA reflect natural movement patterns and dispersal
+corridors of its most mobile protected species is virtually unknown.
 
 Across the KAZA-TFCA, the African wild dog (\textit{Lycaon pictus}) represents a
-highly endangered flagship species for conservation efforts, but for which
-systematically collected GPS relocation data on dispersing individuals are
-unavailable. Once widespread across the entire Sub-Saharan continent, wild
-dogs have been widely extirpated through human persecution, habitat destruction,
-and disease outbreaks \citep{Woodroffe.2012}. As a result, the species has
-become one of Africa's most endangered large carnivores, and currently only
-survives in small, spatially scattered subpopulations \citep{Woodroffe.2012}.
-Within these subpopulations, wild dogs form cooperative breeding packs of up to
-thirty individuals \citep{Frame.1979, Fuller.1992, Creel.2002}, whose social
-structure is strongly shaped by the process of dispersal \citep{McNutt.1996,
-Woodroffe.2019, Behr.2020}. Both males and females typically disperse from their
-natal pack, either alone or in same-sex dispersing coalitions, and search for
+highly mobile and endangered flagship species for conservation efforts, but for
+which systematically collected GPS relocation data on dispersing individuals are
+unavailable. Once widespread across the entire Sub-Saharan continent, wild dogs
+have been widely extirpated through human persecution, habitat destruction, and
+disease outbreaks \citep{Woodroffe.2012}. As a result, the species has become
+one of Africa's most endangered large carnivores, and currently only survives in
+small, spatially scattered subpopulations \citep{Woodroffe.2012}. Within these
+subpopulations, wild dogs form cooperative breeding packs of up to thirty
+individuals \citep{Frame.1979, Fuller.1992, Creel.2002}, whose social structure
+is strongly governed the process of dispersal \citep{McNutt.1996,
+Woodroffe.2019, Behr.2020}. Both males and females disperse from their natal
+pack, either alone or in same-sex dispersing coalitions, and search for
 unrelated mates and a suitable territory to settle \citep{McNutt.1996,
@@
 this species, little empirical information is available on habitat selection and
-potential movement barriers of dispersing wild dogs. The few studies that have
+potential movement barriers during dispersal. The few studies that have
 collected GPS relocation data during dispersal have shown that dispersers
@@
 
-Here, we used GPS relocation data collected on 16 dispersing wild dogs in as
-many dispersing coalitions from a free-ranging population in northern Botswana
-to assess the adequacy of the boundaries of the KAZA-TFCA, the largest protected
-area in the world. We estimated the degree of selection or avoidance for
-environmental and anthropogenic landscape features. We then applied the obtained
-habitat preferences to create a permeability surface spanning the entire
-KAZA-TFCA and investigated how landscape permeability varies regionally and
-internationally. Finally, we calculated least-cost paths and least-cost
-corridors to identify major dispersal corridors, pinpoint dispersal hubs, and
-verify whether the KAZA-TFCA successfully covers these features.
+Here, we collected GPS relocation data on 16 dispersing wild dogs in as many
+dispersing coalitions from a free-ranging population in northern Botswana and
+used it to assess the adequacy of the KAZA-TFCA in securing connectivity for
+this species. To achieve this, we estimated the degree of selection or avoidance
+for environmental and anthropogenic landscape features during dispersal. We then
+applied the obtained habitat preferences and predicted a permeability surface
+spanning the entire KAZA-TFCA and we investigated how landscape permeability
+varies regionally and internationally. Finally, we calculated least-cost paths
+and corridors to identify major movement corridors, pinpoint dispersal hubs, and
+verify that the KAZA-TFCA successfully covers these features.
 
@@
 McNutt.2008, Cozzi.2013, Cozzi.2020, Behr.2020}. Between 2011 and 2019, we
-collected GPS relocation data on 16 coalitions of dispersing African wild dogs
-(7 female and 9 male coalitions; see also \cite{Abrahms.2017} and
-\cite{Cozzi.2020}). Candidate dispersing individuals were identified based on
-criteria reported in \cite{Behr.2020}, immobilized according to protocols
-described in \cite{Osofsky.1996}, and fitted with GPS/Satellite radio collars
-(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while still in their
-natal pack. All procedures were undertaken and supervised by a
+systematically collected GPS relocation data on 16 coalitions of dispersing
+African wild dogs (7 female and 9 male coalitions; see also \cite{Abrahms.2017},
+\cite{Cozzi.2020}, and \cite{Behr.2020}). Candidate dispersing individuals were
+identified based on criteria reported in \cite{Behr.2020}, immobilized according
+to protocols described in \cite{Osofsky.1996}, and fitted with GPS/Satellite
+radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while
+still in their natal pack. All procedures were undertaken and supervised by a
 Botswana-registered wildlife veterinarian. Immobilized individuals were
@@
 
-Because we were interested in dispersal behavior, we discarded any GPS data
+Because we were interested in dispersal behavior only, we discarded any GPS data
 collected while individuals were still with their natal packs and also after
@@
 and through visual inspection of the net squared displacement (NSD) metric. NSD
-is a metric returning the euclidean distance of a relocation to a reference
-point \citep{Borger.2012}, which in our case was the center of the dispersing
+quantifies the euclidean distance of a relocation to a reference point
+\citep{Borger.2012}, which in our case was the center of the dispersing
 coalition's natal home range. Thus, dispersal was deemed to have started when a
 coalition had left its natal home range and continued until the NSD metric
-remained stationary, implying that the coalition had settled (Figure S1). Some
-coalitions undertook exploratory movements prior to proper dispersal. Such
-movements have been observed in other species and are reported to be very
-similar to proper dispersal \citep{Killeen.2014}. Therefore, we also classified
-exploratory movements as dispersal. In total, we collected 4'169 GPS relocations
-during dispersal, i.e. between emigration and settlement (Figure S2 \& Table
-S1), resulting in an average of 260 locations per dispersing coalition (range:
-37 - 729). On average, coalitions dispersed for 48 days (min = 9, max = 137),
-covered a mean euclidean distance of 54 km (min = 3, max = 263) and a mean
-cumulative distance of 596 km (min = 130, max = 1'962). In our analysis, we did
-not differentiate between male and female dispersing coalitions, for previous
-research found little differences between sexes during dispersal
+remained stationary, implying that the coalition had successfully settled
+(Figure S1). Some coalitions undertook exploratory movements prior to proper
+dispersal. Such movements have been observed in other species and are reported
+to be very similar to proper dispersal \citep{Killeen.2014}. Therefore, we also
+classified exploratory movements as dispersal. In total, we collected 4'169 GPS
+relocations during dispersal, i.e. between emigration and settlement (Figure S2
+\& Table S1), resulting in an average of 260 locations per dispersing coalition
+(min = 37, max = 729). On average, coalitions dispersed for 48 days (min = 9,
+max = 137), covered a mean euclidean distance of 54 km (min = 3, max = 263) and
+a mean cumulative distance of 596 km (min = 130, max = 1'962). In our analysis,
+we did not differentiate between male and female dispersing coalitions, for
+previous research found little differences between sexes during dispersal
 \citep{Woodroffe.2019, Cozzi.2020}.
@@
 highly variable within and between years, we created dynamic ``flood maps'' that
-were updated each 8\textsuperscript{th} day following a remote sensing algorithm
-developed by \citep{Wolski.2017}. The algorithm uses thresholding of short
-wavelength infrared reflectances of MODIS Terra satellite imagery (MCD43A4;
-\citealp{Schaaf.2015}) to distinguish areas covered by water and dryland
-(details in Appendix A.3). While we created dynamic flood maps for the Okavango
-Delta, we assumed the extent of all other water bodies (e.g. Chobe river,
-Zambezi river) to be static within and between years. This static representation
-was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
-only retained the categories \textit{wetland} and \textit{water bodies} and
-collectively reclassified them to \textit{water}. Globeland had an original
-resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
-of each 250m x 250m cell. We further improved river representation by employing
-the rasterized MERIT Hydro dataset \citep{Yamazaki.2019} from which we added all
-rivers with a width of over 10m to our Globeland layer. We merged dynamic and
-static water maps into a large rasterstack, covering the entire study area. We
-also created a rasterstack rendering the covariate \textit{distance to water} by
-calculating the eculidean distance of each raster cell in the study area to the
-nearest source of water.
+were updated every 8\textsuperscript{th} day following a remote sensing
+algorithm developed by \cite{Wolski.2017}. The algorithm uses thresholding of
+short wavelength infrared reflectances of MODIS Terra satellite imagery
+(MCD43A4; \citealp{Schaaf.2015}) to distinguish areas covered by water and
+dryland (details in Appendix A.3). While we created dynamic flood maps for the
+Okavango Delta, we assumed the extent of all other water bodies (e.g. Chobe
+river, Zambezi river) to be static within and between years. This static
+representation was based on Globeland's land cover dataset \citep{Chen.2015},
+from which we only retained the categories \textit{wetland} and \textit{water
+bodies} and collectively reclassified them to \textit{water}. Globeland had an
+original resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using
+the mode of each 250m x 250m cell. We further improved river representation by
+employing the rasterized MERIT Hydro dataset \citep{Yamazaki.2019} from which we
+added all rivers with a width of over 10m to our Globeland layer. We merged
+dynamic and static water maps into a large rasterstack, covering the entire
+study area. We also created a rasterstack rendering the covariate
+\textit{distance to water} by calculating the eculidean distance of each raster
+cell in the study area to the nearest source of water.
 
@@
 We created a binary layer separating protected from unprotected land. We
-downloaded corresponding data on protection status from the Peace Parks
-Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}). Protected
-areas included forest reserves, game reserves, wildlife management areas, and
-national parks. We classified anything not covered by these categories as
-unprotected (e.g. communal pastoral land, private land). We rasterized the two
-categories to the binary raster \textit{protection status} (1 = protected, 0 =
-pastoral) with a resolution of 250m x 250m.
+downloaded corresponding data on protection status in shapefile format from the
+Peace Parks Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}).
+Protected areas included forest reserves, game reserves, wildlife management
+areas, and national parks. We classified anything not covered by these
+categories as unprotected (e.g. communal pastoral land, private land). We
+rasterized the two categories to the binary raster \textit{protection status} (1
+= protected, 0 = pastoral) with a resolution of 250m x 250m.
 
@@
 \textit{human density}, \textit{farming}, and \textit{roads}. This resulted in a
-single covariate called \textit{human influence}. Note that totalling the layers
+single covariate called \textit{human influence}. Note that totaling the layers
 implied that the three covariates received different weights (see Appendix A.6
@@
 step crossed a road and we identified the average distance of each step to the
-nearest source of water and the nearest road. We square-rooted the values for
-\textit{distance to roads} and \textit{distance to water} to render a decreasing
-marginal impact of distance. Additionally, we derived a binary indicator of
-whether a step crossed a road or not. We standardized extracted values using a
+nearest nearest road and the nearest source of water. We square-rooted the
+values for \textit{distance to roads} and \textit{distance to water} to render a
+decreasing marginal impact of distance. We standardized extracted values using a
 z-score transformation, and checked for correlation using Pearson's Correlation
@@
 \citealp{Latham.2011}) and we retained all of them for modeling. Our habitat
-selection was then based on the assumption that dispersing wild dogs assigned a
-selection score \(w(x)\) of the following exponential form to each realized and
-random step \citep{Fortin.2005, Thurfjell.2014, Avgar.2016}:
+selection model was then based on the assumption that dispersing wild dogs
+assigned a selection score \(w(x)\) of the following exponential form to each
+realized and random step \citep{Fortin.2005, Thurfjell.2014, Avgar.2016}:
 
@@
 surface spanning the entire extent of the KAZA-TFCA. That is, we applied
-\Cref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
+\ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
 for each raster cell in our study area. Because our flood maps were dynamic, we
 collapsed all flood maps into a single static layer using areas that were
-covered by water in at least a quarter of all our flood maps. We also calculated
-a corresponding layer indicating the euclidean distance of each raster cell to
-the nearest source of water on the static water map. To reduce the influence of
-outliers in predicted permeability scores we followed \cite{Squires.2013} and
-curtailed predicted scores between the 1\textsuperscript{st} and
-99\textsuperscript{th} percentile of their original values. To compare
-permeability across different regions, we rescaled the permeability surface to a
-range between 0 and 1. The most impermeable landscape thus received a value of
-0, the most permeable landscape a value of 1. Based on the permeability surface
-we calculated median permeability values within and outside the KAZA-TFCA,
-within and outside formally protected areas, and within each of the five
-KAZA-TFCA countries.
+covered by water in at least 10\% of all our flood maps. We also calculated a
+corresponding layer indicating the square-rooted euclidean distance of each
+raster cell to the nearest source of water. To reduce the influence of outliers
+in predicted permeability scores we followed \cite{Squires.2013} and curtailed
+predicted scores between the 1\textsuperscript{st} and 99\textsuperscript{th}
+percentile of their original values. To compare permeability across different
+regions, we normalized the permeability surface to a range between 0 and 1. The
+most impermeable landscape thus received a value of 0, the most permeable
+landscape a value of 1. Based on the permeability surface we calculated median
+permeability values within and outside the KAZA-TFCA, within and outside
+formally protected areas, and within each of the five KAZA-TFCA countries.
 
@@
 least-cost corridors (LCCs) among all source points (Figure S.7;
-\citealp{Adriaensen.2003, Sawyer.2011, Elliot.2014}. We generated source points
+\citealp{Adriaensen.2003, Sawyer.2011, Elliot.2014}). We generated source points
 by overlaying the study area with a regular grid of points spaced at 100 km. We
 only considered those points that fell within protected areas \(>\) 700
-km\textsuperscript{2} as source points, which conforms with home-range
-requirements of African wild dogs reported in \cite{Pomilia.2015}. Finally, we
-defined centroids as source points for those protected areas \(>\) 700
-km\textsuperscript{2} that were not assigned any source points from the regular
-grid. In total, we generated 68 source points, which resulted in 2'278 unique
-pairwise combinations and therefore 2'278 unique LCPs and LCCs.
+km\textsuperscript{2}, which conforms with home-range requirements of African
+wild dogs reported in \cite{Pomilia.2015}. Finally, we defined centroids as
+source points for those protected areas \(>\) 700 km\textsuperscript{2} that
+were not assigned any source points from the regular grid. In total, we
+generated 68 source points, which resulted in 2'278 unique pairwise combinations
+and therefore 2'278 unique LCPs and LCCs.
 
@@
 Okavango-Linyanti ecosystem in Botswana with Luengue-Luiana NP in Angola. A
-further corridor runs W-E between Chobe NP in Botswana and Zimbabwe's Hwange NP.
+second corridor runs W-E between Chobe NP in Botswana and Zimbabwe's Hwange NP.
 A third major corridor runs NE-SW, completely across unprotected areas, and
@@
 the corridor passing the Okavango-Linyanti region are the highest frequented
-routes within KAZA (maybe numbers?). Our model did not detect any significant
-direct corridors between Zimbabwe and Zambia or Zambia and Angola, and only a
-very limited W-E direct connection between the Okavango region and Namibia's
-Khaudum NP. Except for the corridor into the Central Kalahari National Park, our
-model did not detect any significant connectivity outside the boundaries of
-KAZA-TFCA.
+routes within the KAZA-TFCA (\Cref{LeastCost}, dark red colors). Our model did
+not detect any significant direct corridors between Zimbabwe and Zambia or
+Zambia and Angola, and only a very limited W-E direct connection between the
+Okavango region and Namibia's Khaudum NP. Except for the corridor into the
+Central Kalahari National Park, our model did not detect any significant
+connectivity outside the boundaries of the KAZA-TFCA.
 
@@
 movement corridors of African wild dogs, demonstrating the potential value of
-such an initiative. However, our approach is neither limited to the African wild
+such an initiative. We thus expemplified how pertinent dispersal data of a
+highly mobile species can be used to assess the adequacy of already existing or
+planned protected areas. Our approach is neither limited to the African wild
 dog, nor to our study area. All covariates used throughout this study are
@@
 similarities to corridors of dispersing lions inhabiting the same ecosystem
-\citep{Elliot.2014}. This not only reinforces confidence in our own predictions
-but also suggests potential synergies for the conservation of these two, and
-possibly more, species. Expanding this analytical framework to additional
-species will yield important insights on the consistency of inter-specific
-movement corridors, thus highlighting areas that are exceptionally valuable for
-the conservation of several species.
-
-Our approach of identifying movement corridors based on pre-defined start and
-end points implicitly assumes that individuals know the end point of their
-dispersal journey and that they have almost complete knowledge of associated
-movement costs. Considering the fact that dispersers often move into unknown
-territory \citep{Woodroffe.2019, Cozzi.2020}, this may not necessarily be the
-case \citep{Abrahms.2017, Woodroffe.2019}. However, specification of pre-defined
-end points might not be necessary, as it should be possible to simulate
-dispersal from known source-points, yet without restricting the domain of
-potential end points \citep{Signer.2017}. As a consequence, movement corridors
-would emerge naturally as the result of a myriad of simulated dispersal events.
-While this approach is conceptually straightforward, it requires a comprehensive
-mechanistic understanding of dispersal movements, which is conditional on our
-ability to collect additional dispersal data and adequately represent the
-landscape through which individuals disperse. An approach that does not require
-the creation of end points will also be instrumental in identifying corridors
-outside formally protected areas. In contrast to classical corridor mapping,
-individual-based movement simulations would further allow taking into account
-movement preferences of dispersers as well as rendering the temporal aspect of
-dispersal (for instance to examine how quickly dispersers are able to move
-between locations). The latter consideration is particularly important in
-ecosystems where connectivity may change in response to seasonal variations in
-the landscape \citep{Osipova.2019}.
-
-Our results emphasize that human presence and activities constitute some of the
-main barriers to connectivity. Our result that human influence is detrimental to
-dispersal conforms to findings in wild dog populations from eastern Africa
-\citep{Masenga.2016, Oneill.2020} but conflicts with findings by
-\cite{DaviesMostert.2012}, who reported a high willingness of dispersers to
-cross human-dominated landscapes throughout South Africa. We believe that such
-differences are due to the unavailability of alternative ``natural'' routes,
-which may have forced dispersers to cross human dominated landscapes despite a
-strong aversion to do so. This may hint at a considerable ability of dispersing
-wild dogs to adapt to prevailing habitat condition and could imply a limited
-model transferability between dispersersers from our study system to dispersers
-found in other ecosystems, such as in South Africa \citep{Whittington.2011,
-DaviesMostert.2012} or Kenya \citep{Woodroffe.2019, Oneill.2020}. However, we
-believe that our study population adequately represents wild dog populations
-residing within KAZA-TFCA where more natural dispersal routes are available.
-Successful conservation of this species is thus contingent on policymakers' and
-local authorities' ability and willingness to provide and conserve areas that
-remain free from anthropogenic pressures. This is not only paramount in light of
-providing connectivity and facilitating dispersal, but also in terms of reducing
-human-caused mortality during dispersal. In fact, previous studies have shown
-that human-caused mortality during dispersal, especially in proximity to
-settlements and roads, represents a major threat to the dispersal ability of
-this species \citep{Woodroffe.2019, Cozzi.2020}.
-
-Despite anthropogenic pressures, we identified water as another major obstacle
-to dispersal. This is consistent with earlier studies showing that water bodies
-are almost impenetrable by resident packs \citep{Abrahms.2017, Cozzi.2020} and
-only infrequently crossed by dispersing individuals \citep{Cozzi.2020}. An
-accurate and dynamic representation of water is thus imperative and particularly
-relevant in flood-pulsing ecosystems such as the Okavango Delta. Although
-dispersers avoided moving through water, we revealed a preference for moving in
-proximity to water. This preference is likely due to high prey-availability in
-vicinity of water \citep{Western.1975, Bonyongo.2005}. However, by the same
-reasoning apex predators (e.g. lions, spotted hyenas) and resident wild dogs may
-prefer proximity to water \citep{Valeix.2010} and occasionally force dispersing
-wild dogs to move into prey-poor areas that are more distant to water
-\citep{Ndaimani.2016} to avoid aggressive intraguild interactions
-\citep{Creel.1996, Mills.1997, Webster.2012}. We could not control for the
-presence or absence of apex predators or conspecifics during dispersal as we
-lacked corresponding data, but these considerations may explain why we did not
-find a significant effect of \textit{distance to water}. Given the influence
-that resident conspecifics, predators, competitors, and prey can have during
-dispersal \citep{Cozzi.2018, Armansin.2019} future studies should strive to
-collect and incorporate corresponding spatial information into analyses of
-landscape connectivity to more precisely reconstruct dispersal corridors.
+\citep{Elliot.2014, Cushman.2018}. This not only reinforces confidence in our
+own predictions but also suggests potential synergies for the conservation of
+these two, and possibly more, species. Expanding this analytical framework to
+additional species will yield important insights on the consistency of
+inter-specific movement corridors, thus highlighting areas that are
+exceptionally valuable for the conservation of several species.
 
@@
 the KAZA-TFCA. While this might partly be owed to the area's central position
-within the KAZA-TFCA, it is likely also due to its unique environmental and
+within our study area, it is likely also due to its unique environmental and
 anthropogenic characteristics. It appears that the absence of human activities
@@
 frequented corridor. The region is also instrumental for connecting wild dog
-subpopulations in Zambia's Kafue and Zimbabwe's Hwange NPs. These subpopulations
-are currently separated by the Zambezi River, forcing dispersers to detour via
-northern Botswana to reach more peripheral areas. The key role of the
-Okavango-Linyanti region for overall connectivity within the KATA-TFCA thus
-calls for actions to secure its protection status in the future. While the
-region is currently a Wildlife Management Area devoted to photographic tourism,
-it has neither the status of a National Park nor that of a Game Reserve and so
-potentially faces growing pressure towards exploitation of its natural resources
-under the influences of an expanding human population, farming, and cropping
-industry. A similar case of non-formally protected but key dispersal landscape
-is represented by the area south of Kafue NP in Zambia, for which a disruption
-of its main and narrow dispersal corridor would result in considerable isolation
-of its subpopulations. We also revealed a potential southwards corridor between
-the Okavango-Linyanti ecosystem and the Central Kalahari National Park.
-\cite{Elliot.2014} identified a similar corridor for dispersing lions,
+subpopulations in Zambia's Kafue and Zimbabwe's Hwange NPs
+\citep{Woodroffe.2012}. These subpopulations are currently separated by the
+Zambezi River, forcing dispersers to detour via northern Botswana to reach more
+peripheral areas. The key role of the Okavango-Linyanti region for overall
+connectivity within the KATA-TFCA thus calls for actions to secure its
+protection status in the future. While the region is currently a Wildlife
+Management Area devoted to photographic tourism, it has neither the status of a
+National Park nor that of a Game Reserve and so potentially faces growing
+pressure towards exploitation of its natural resources under the influences of
+an expanding human population, farming, and cropping industry. A similar case of
+non-formally protected but key dispersal landscape is represented by the area
+south of Kafue NP in Zambia, for which a disruption of its main and narrow
+dispersal corridor would result in considerable isolation of its subpopulations.
+We also revealed a potential southwards corridor between the Okavango-Linyanti
+ecosystem and the Central Kalahari National Park. \cite{Elliot.2014} and
+\citep{Cushman.2018} identified a similar corridor for dispersing lions,
 suggesting that upholding and protecting a link between those ecosystems is
@@
 
-%------------------------------------------------------------------------------
-%	Gabs wanted to delete this
-%------------------------------------------------------------------------------
-% Because we predicted permeability and connectivity for such an extensive area,
-% we were strongly limited in available and reliable land cover data. As a
-% result, our representation of vegetation remained rather basic with only three
-% vegetation classes (tree-cover, shrubs/grassland, bare land). Although some
-% large-scale, high-resolution land cover datasets exist, such as Globeland's
-% 30m \citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019}
-% land cover datasets, we found their classifications often erroneous for our
-% study area. This is presumably owed to the KAZA-TFCA's unique landscape and
-% corresponding difficulties in remote sensing satellite imagery using
-% traditional indices \citep{Wolski.2017}.
+Our approach of identifying movement corridors based on pre-defined start and
+end points implicitly assumes that individuals know the end point of their
+dispersal journey and that they have almost complete knowledge of associated
+movement costs \citep{Panzacchi.2016}. Since dispersers often move into unknown
+territory, this may not necessarily be the case \citep{Abrahms.2017,
+Woodroffe.2019, Cozzi.2020}. However, specification of pre-defined end points
+might not be necessary, as the parametrized iSSF model can be applied as a
+mechanistic movement model to simulate dispersal events from known source
+points, yet without restricting the domain of potential end points
+\citep{Signer.2017}. Consequently, movement corridors would emerge more
+naturally as the result of a myriad of simulated dispersal events. An approach
+that does not require the creation of end points will also be instrumental in
+identifying corridors outside formally protected areas. In contrast to classical
+corridor mapping, individual-based movement simulations would further allow
+considering movement preferences of dispersers as well as rendering the temporal
+aspect of dispersal (for instance to examine how quickly dispersers are able to
+move between locations). The latter consideration is particularly important in
+ecosystems where connectivity may change in response to seasonal variations in
+the landscape \citep{Osipova.2019}. While a simulation based approach is
+conceptually straightforward, it requires a comprehensive mechanistic
+understanding of dispersal movements, which is conditional on our ability to
+collect additional dispersal data and adequately represent the landscape through
+which individuals disperse.
+
+Our results further emphasize that human presence and activities constitute some
+of the main barriers to connectivity. Our result that human influence is
+detrimental to dispersal conforms to findings on dispersing wild dogs from
+eastern Africa \citep{Masenga.2016, Oneill.2020} but conflicts with findings by
+\cite{DaviesMostert.2012}, who reported a comparably high willingness of
+dispersers to cross human-dominated landscapes in South Africa. We believe that
+such differences are due to the unavailability of alternative ``natural''
+routes, which may have forced dispersers in South Africa to cross human
+dominated landscapes despite a strong aversion to do so. These contrasting
+findings may hint at a considerable ability of dispersing wild dogs to adapt to
+prevailing habitat conditions (see e.g. \cite{Woodroffe.2011}) and could imply a
+limited model transferability between dispersersers from our study system to
+dispersers found in other ecosystems, such as in South Africa
+\citep{Whittington.2011, DaviesMostert.2012} or Kenya \citep{Woodroffe.2019,
+Oneill.2020}. However, we believe that our study population adequately
+represents wild dog populations residing within the greater extent of the
+KAZA-TFCA, where more natural dispersal routes are readily available. Successful
+conservation of this species is thus strictly contingent on policymakers' and
+local authorities' willingness and ability to provide and conserve areas that
+remain free from anthropogenic pressures. This is not only paramount in light of
+providing connectivity and facilitating dispersal, but also in terms of reducing
+human-caused mortality during dispersal. In fact, previous studies have shown
+that human-caused mortality during dispersal, especially in proximity to
+settlements and roads, represents a major threat to the dispersal ability of
+this species \citep{Woodroffe.2019, Cozzi.2020}.
+
+Besides anthropogenic pressures, we identified water as another major obstacle
+to dispersal. This corroborates earlier studies which showed that water bodies
+(including swamps) are almost impenetrable by resident packs
+\citep{Abrahms.2017, Cozzi.2020} and only infrequently crossed by dispersing
+individuals \citep{Cozzi.2020}. An accurate and dynamic representation of water
+is thus imperative and particularly relevant in seasonal or flood-pulsing
+ecosystems such as the Okavango Delta. In fact, future studies could investigate
+how connectivity across such systems changes in result to seasonal changes in
+the environment (see also \cite{Osipova.2019}).
+
+Although dispersers avoided moving through water, we revealed a preference for
+moving in proximity to water. This preference is likely explained by high
+prey-availability in vicinity of water \citep{Western.1975, Bonyongo.2005}.
+However, following the same reasoning, apex predators (e.g. lions, spotted
+hyenas) and resident wild dogs may prefer proximity to water
+\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs into
+prey-poorer areas that are more distant to water \citep{Ndaimani.2016} and pose
+a lower risk of aggressive intraguild interactions \citep{Creel.1996,
+Mills.1997, Webster.2012}. We could not control for the presence or absence of
+apex predators or conspecifics during dispersal as we lacked corresponding data,
+but these considerations may explain why we did not find a significant effect of
+\textit{distance to water}. Given the influence that resident conspecifics,
+predators, competitors, and prey can have on dispersers \citep{Cozzi.2018,
+Armansin.2019} future studies should strive to collect and incorporate
+corresponding ``social'' information into analyses of landscape connectivity to
+more precisely reconstruct dispersal corridors.
 
 \section{Conclusion}
-Our work showed that the majority of dispersal corridors for African wild dogs
-are covered by the boundaries of the KAZA-TFCA, suggesting that the KAZA-TFCA
-will significantly contribute to the long-term viability of this species.
-Moreover, our connectivity network highlights the Okavango-Linyanti region as a
-central dispersal hub through which dispersers gain access to more remote
-regions of the KAZA-TFCA. This fidning emphasizes the importance of northern
-Botswana for the conservation of this charismatic species. Finally, our
-investigations showed that human influence constitutes one of the main barriers
-to dispersal and thus reduces landscape connectivity. Successful conservation of
-the wild dogs is therefore contingent on the willingness of local authorities,
-policymakers, and land managers to preserve areas that are largely liberated
-from human strains. Ultimately, our work contributes to the growing field of
-connectivity literature and has important implications for the management and
-conservation of the highly endangered African wild dog.
+Our work shows how dispersal data of a highly mobile species can be used
+identify movement corridors and assess the adequacy of protected areas. In our
+case, the predicted dispersal corridors of African wild dogs were well contained
+within the boundaries of the world's largest transboundary conservation area,
+namely the KAZA-TFCA, suggesting that it will significantly contribute to the
+long-term viability of this species. Moreover, our connectivity network allowed
+to reveal potential dispersal hubs thorugh which dispersers gain access to more
+remote regions in the study area. Finally, our investigations showed that human
+influence constitutes one of the main barriers to dispersal and substantially
+reduces landscape connectivity accross large scales. Successful conservation of
+wide-ranging species, such as exemplified by the African wild dog, will
+therefore be contingent on the willingness of local authorities, policymakers,
+and land managers to preserve areas that remain free from human strains.
+Ultimately, our work contributes to the growing field of connectivity literature
+and provides and easily expandible framework for the assessement of existing or
+planned protected areas.
+
+\section{Author's Contributions}
+D.D.H., D.M.B., and G.C. conceived the study and designed methodology; D.M.B.,
+G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the data; G.C.
+and A.O. assisted with modelling; D.D.H., D.M.B., and G.C. wrote the first draft
+of the manuscript and all authors contributed to the drafts at several stages
+and gave final approval for publication.
+
+\section{Data Availability}
+GPS movement data of dispersing coalitions will be made available on dryad at
+the time of publication.
 
@@
 to Prof. J. Fieberg, who consulted all statistical aspects of this work and P.
-Wolski from the Okavango Research Institute who assisted us in generating
+Wolski, from the Okavango Research Institute, who assisted us in generating
 dynamic flood maps. This study was funded by Basler Stiftung für Biologische

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-06-12.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-06-08.tex
@@
 \begin{abstract}
-\noindent \textbf{1. Background}: Protecting large portions of natural land
-connecting natural reserves is an increasingly used strategy to maintain and
-restore connectivity among wildlife subpopulations. One such initiative is the
-Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA) in Africa, the
-world's largest transboundary conservation area, spanning five countries and
-520'000 km\textsuperscript{2}. While boundaries of such conservation areas are
-often determined by expert opinion as well as socio-political needs and
-constraints, the extent to which they match species' movement corridors is
-rarely examined. This is mainly due to a lack of adequate empirical data,
-particularly on movement behavior of wide-ranging species during dispersal.
+\noindent 1. Conserving and managing large portions of land to connect wildlife
+reserves is increasingly used to maintain and restore connectivity among
+wildlife populations. Boundaries of such conservation areas are, however, often
+determined based on expert opinion and socio-political constraints, but the
+extent to which they match species' movement corridors is rarely examined. This
+is mainly due to a lack of data, particularly on wide-ranging movement behavior
+such as dispersal. Yet, empirically assessing the adequacy of protected areas is
+key for the implementation of targeted management actions and use of limited
+conservation funds.
 
-\noindent \textbf{2. Methods}: Between 2011 and 2019, we collected
-high-resolution GPS relocation data on 16 coalitions of dispersing African wild
-dogs (\textit{Lycaon pictus}) from a free ranging population in northern
-Botswana. We combined the collected data with relevant spatial covariates and
-used integrated step selection analysis to estimate habitat preferences during
-dispersal and create a permeability surface spanning the entire KAZA-TFCA. To
-examine whether the KAZA-TFCA covers major movement corridors of this highly
-mobile species, we compared landscape permeability across different regions of
-our study area. We also calculated least-cost paths and corridors between
-selected locations and verified that corridors are encompassed within the
-boundaries of the KAZA-TFCA.
+\noindent 2. Between 2011 and 2019, we collected high-resolution GPS movement
+data on 16 dispersing African wild dog (\textit{Lycaon pictus}) coalitions from
+a free-ranging population in the Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA). Spanning five countries and 520'000 km\textsuperscript{2} the
+KAZA-TFCA is the world's largest transboundary conservation area and a prime
+example of international conservation efforts. We used an integrated step
+selection analysis to create a permeability surface for the KAZA-TFCA. We then
+compared landscape permeability across different regions within the KAZA-TFCA as
+well as outside its boundaries. Lastly, we calculated least-cost paths and
+corridors to verify that major movement corridors are adequately encompassed
+within the KAZA-TFCA.
 
-\noindent \textbf{3. Results}: We show that permeability within the KAZA-TFCA is
-more than double compared to areas outside it. Furthermore, we report a
-five-fold permeability difference among the five KAZA-TFCA countries. These
-differences are mainly caused by different degrees of human activities across
-regions, hampering dispersal movements. We further present a connectivity
-network, consisting of least-cost paths and corridors. We thereby show that all
-major movement corridors of wild dogs run within the KAZA-TFCA boundaries,
-although some minor corridors remain outside formally protected areas.
+\noindent 3. Permeability within the boundaries of the KAZA-TFCA was more than
+double compared to areas outside it. Furthermore, we observed a five-fold
+permeability difference among the five countries. We further showed that major
+movement corridors of wild dogs run within the KAZA-TFCA boundaries, although
+some minor corridors remained outside formally protected areas.
 
-\noindent \textbf{4. Synthesis and Applications}: Our work highlights the
-importance of empirically assessing the adequacy of protected areas and the need
-for international coordination that stem from different contributions to
-connectivity across regions and countries. Moreover, the presented framework is
-readily expandable to include additional species and regions in order to
-highlight consistencies among potential dispersal corridors as well as dispersal
-corridors outside our ecosystem.
+\noindent 4. Differences in permeability were mainly caused by different degrees
+of human activities across regions, which hampered dispersal. Rivers and swamps
+also limited dispersal, while other landscape features had a limited effect.
 
+\noindent 5. \textit{Synthesis and Applications}: In this study, we showed how
+pertinent dispersal data of a highly mobile species can be used to empirically
+evaluate the adequacy of already-existing protected areas or could be used for
+the planning of future such initiatives. Furthermore, the ability to reveal
+regional differences in landscape permeability is fundamental for a coordinated
+effort towards maintaining or restoring connectivity, especially where
+international effort is required.
 \end{abstract}
@@
 Connectivity among subpopulations is a crucial pre-requisite for many species to
-thrive and persist \citep{Fahrig.2003}. A sufficient degree of connectivity
-enables dispersal, which in turn contributes to population viability through
-gene-flow, the ability to reinforce and rescue small populations, to colonize or
-recolonize unoccupied habitats, as well as to shift range in response to climate
-change \citep{Brown.1977, Hanski.1998, MacArthur.2001, Frankham.2002,
-Heller.2009, Leigh.2012}. Particularly in light of ongoing habitat fragmentation
-and destruction worldwide \citep{Fahrig.2003, Lindenmayer.2013}, the protection
-of natural or semi natural areas that serve as movement corridors and provide
-connectivity has become an utmost task for conservation management
-\citep{Heller.2009, Doerr.2011, Rudnick.2012, Cozzi.2020}. This has resulted in
-an ever-growing number of large and often transboundary protected areas that are
-set to facilitate dispersal between habitats \citep{Wolmer.2003}. Due to the
-very nature of large scale protected areas, a thorough quantitative assessment
-of their effectiveness and adequacy of their boundaries is challenging
-\citep{Rudnick.2012}. In result, boundaries are typically drawn according to
-expert opinion and socio-political needs rather than empirical information. In
-comparison to qantitative assessments, however, such qualitative approaches have
-revealed deficiencies in the past and potentially hold the risk of misdirecting
-valuable and scarce conservation funds \citep{Pullinger.2010, Sawyer.2011}. An
-accurate representation of movement corridors and dispersal routes based on
-empirical data is thus fundamental to objectively assess how well boundaries of
-existing or planned protected areas reflect natural movements of their protected
-species \citep{Beier.2008, Doerr.2011, Rudnick.2012}.
+thrive and persist \citep{Fahrig.2003}. Accordingly, preserving and protecting
+movement corridors between wildlife reserves has become an utmost task for
+conservation management \citep{Heller.2009, Doerr.2011, Rudnick.2012,
+Cozzi.2020} and has resulted in an ever-growing number of large and often
+transboundary protected areas \citep{Wolmer.2003}. Due to the very nature of
+such large protected areas, a thorough quantitative assessment of their
+effectiveness and adequacy of their boundaries is challenging
+\citep{Rudnick.2012}. This challenge is complicated by a lack of animal movement
+data at the appropriate spatial scale and an insufficient understanding of the
+movement patterns and dispersal corridors of the species for which these
+protected areas have been created \citep{Vasudev.2015, Tshipa.2017, Cozzi.2020}.
+The use of subjective boundary assessments and the application of inappropriate
+movement data, however, have revealed deficiencies in the past
+\citep{Clevenger.2002, Pullinger.2010, Sawyer.2011, Elliot.2014, Abrahms.2016}.
+Thus, an empirical assessment of the adequacy of already-existing or planned
+protected areas using pertinent movement data of dispersing animals is paramount
+for targeted use of valuable and scarce conservation funds
+\citep{Pullinger.2010, Sawyer.2011}.
 
@@
 individuals \citep{Frame.1979, Fuller.1992, Creel.2002}, whose social structure
-is strongly governed the process of dispersal \citep{McNutt.1996,
+is strongly governed by the process of dispersal \citep{McNutt.1996,
 Woodroffe.2019, Behr.2020}. Both males and females disperse from their natal
@@
 this species. To achieve this, we estimated the degree of selection or avoidance
-for environmental and anthropogenic landscape features during dispersal. We then
-applied the obtained habitat preferences and predicted a permeability surface
-spanning the entire KAZA-TFCA and we investigated how landscape permeability
-varies regionally and internationally. Finally, we calculated least-cost paths
-and corridors to identify major movement corridors, pinpoint dispersal hubs, and
-verify that the KAZA-TFCA successfully covers these features.
+for environmental and anthropogenic landscape features. We then applied the
+obtained habitat preferences and predicted a permeability surface spanning the
+entire KAZA-TFCA and we investigated how landscape permeability varies
+regionally and internationally and compared permeability within and outside the
+KAZA-TFCA boundaries. Finally, we calculated least-cost paths and corridors to
+identify major movement corridors, pinpoint dispersal hubs, and verify that the
+KAZA-TFCA successfully covers these features.
 
@@
 likely acts as a source for the recolonization of surrounding habitats
-\citep{Woodroffe.2012, Cozzi.2013}. The Okavango Delta represents a highly
-dynamic hydrological flood-pulsing system \citep{McNutt.1996, Wolski.2017}. The
-extent of the flood in the delta greatly changes within and between years
-depending on the amount of rain that descends from the catchment areas in Angola
-and reaches the distal ends of the delta between July and August (see Figure S5
-for an illustration of a typical flood pulse). The flood drastically affects
+\citep{Woodroffe.2012}. The Okavango Delta represents a highly dynamic
+hydrological flood-pulsing system \citep{McNutt.1996, Wolski.2017}. The extent
+of the flood in the delta greatly changes within and between years depending on
+the amount of rain that descends from the catchment areas in Angola and reaches
+the distal ends of the delta between July and August (see Figure S5 for an
+illustration of a typical flood pulse). The flood drastically affects
 surrounding landscapes, so that during maximum extent (ca. 12'000
@@
 decreasing marginal impact of distance. We standardized extracted values using a
-z-score transformation, and checked for correlation using Pearson's Correlation
+z-score transformation and checked for correlation using Pearson's Correlation
 Coefficient \(r\). None of the covariates were overly correlated (\(|r| > 0.6\);
@@
 Kaudom NP, and a northeasterly extension of the Hwange corridor into Zimbabwe's
-Matusadona NP. According to our predictions, the corridor towards Kafue NP and
-the corridor passing the Okavango-Linyanti region are the highest frequented
-routes within the KAZA-TFCA (\Cref{LeastCost}, dark red colors). Our model did
-not detect any significant direct corridors between Zimbabwe and Zambia or
-Zambia and Angola, and only a very limited W-E direct connection between the
-Okavango region and Namibia's Khaudum NP. Except for the corridor into the
-Central Kalahari National Park, our model did not detect any significant
-connectivity outside the boundaries of the KAZA-TFCA.
+Matusadona NP. According to our predictions, the landscapes in the
+Okavango-Linyanti region are the highest frequented dispersal routes within the
+KAZA-TFCA (\Cref{LeastCost}b). Our model did not detect any significant direct
+corridors between Zimbabwe and Zambia or Zambia and Angola, and only a very
+limited W-E direct connection between the Okavango region and Namibia's Khaudum
+NP. Except for the corridor into the Central Kalahari National Park, our model
+did not detect any significant connectivity outside the boundaries of the
+KAZA-TFCA.
 
@@
 planned protected areas. Our approach is neither limited to the African wild
-dog, nor to our study area. All covariates used throughout this study are
-readily available on a global scale and many of them are likely to be important
-determinants of movement behavior, landscape permeability, and connectivity for
-other species \citep{Thurfjell.2014, Zeller.2012}. Interestingly, our predicted
-network of least cost-paths and corridors for African wild dogs shows surprising
+dog, nor to our study area and thus applicable to any study system. All
+covariates used throughout this study are readily available on a global scale
+and many of them are likely to be important determinants of movement behavior,
+landscape permeability, and connectivity for other species
+\citep{Thurfjell.2014, Zeller.2012}. Interestingly, our predicted network of
+least cost-paths and corridors for African wild dogs shows surprising
 similarities to corridors of dispersing lions inhabiting the same ecosystem
@@
 
+Our results emphasize that human presence and activities constitute some of the
+main barriers to connectivity. Our result that human influence is detrimental to
+dispersal conforms to findings on dispersing wild dogs from eastern Africa
+\citep{Masenga.2016, Oneill.2020} but conflicts with findings from South Africa
+by \cite{DaviesMostert.2012}, who reported a comparably high willingness of
+dispersers to cross human-dominated landscapes. We believe that such differences
+are due to the unavailability of alternative routes through natural landscapes,
+which may have forced dispersers in South Africa to cross human dominated
+landscapes despite a strong aversion to do so. In this regard, our
+representation of dispersal corridors and the resulting connectivity appear
+conservative, as dispersers may be able to make the best out of a bad situation
+and cross landscapes characterized by considerably unfavorable conditions
+\citep{Palomares.2000, Elliot.2014}. Successful conservation of this species is
+thus strictly contingent on policymakers' and local authorities' willingness and
+ability to provide and conserve areas that remain free from anthropogenic
+pressures. This is not only paramount in light of providing connectivity and
+facilitating dispersal, but also in terms of reducing human-caused mortality
+during dispersal. In fact, previous studies have shown that human-caused
+mortality during dispersal, especially in proximity to settlements and roads,
+represents a major threat to the dispersal ability of this species
+\citep{Woodroffe.2019, Cozzi.2020}.
+
+Besides anthropogenic pressures, we identified water as another major obstacle
+to dispersal. This corroborates earlier studies which showed that water bodies
+(including swamps) are almost impenetrable by resident packs
+\citep{Abrahms.2017} and only infrequently crossed by dispersing individuals
+\citep{Cozzi.2020}. An accurate and dynamic representation of water is thus
+imperative and particularly relevant in seasonal or flood-pulsing ecosystems
+such as the Okavango Delta. In fact, future studies could investigate how
+connectivity across such systems changes in result to seasonal changes in the
+environment (see also \cite{Osipova.2019}). Although dispersers avoided moving
+through water, they selected locations in its vicinity. This preference may be
+caused by the presence of prey close to water \citep{Western.1975,
+Bonyongo.2005}. For the same reason, however, competitors such as lions, spotted
+hyenas, and resident wild dogs may also use areas close to water sources
+\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to move
+into prey-poorer areas away from water \citep{Ndaimani.2016}. We could not
+control for the presence or absence of apex predators or conspecifics during
+dispersal as we lacked corresponding data, but these considerations may explain
+why we did not find a significant effect of \textit{distance to water}. Given
+the influence that resident conspecifics, predators, competitors, and prey can
+have on dispersers \citep{Cozzi.2018, Armansin.2019} future studies should
+strive to collect and incorporate corresponding ``social'' information into
+analyses of landscape connectivity to more precisely reconstruct dispersal
+corridors.
+
 Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
-through which dispersersing wild dogs gain access to more peripheral regions of
-the KAZA-TFCA. While this might partly be owed to the area's central position
-within our study area, it is likely also due to its unique environmental and
+through which dispersing wild dogs gain access to more peripheral regions of the
+KAZA-TFCA. While this might partly be owed to the area's central position within
+our study area, it is likely also due to its unique environmental and
 anthropogenic characteristics. It appears that the absence of human activities
@@
 peripheral areas. The key role of the Okavango-Linyanti region for overall
-connectivity within the KATA-TFCA thus calls for actions to secure its
+connectivity within the KAZA-TFCA thus calls for actions to secure its
 protection status in the future. While the region is currently a Wildlife
@@
 
-Our results further emphasize that human presence and activities constitute some
-of the main barriers to connectivity. Our result that human influence is
-detrimental to dispersal conforms to findings on dispersing wild dogs from
-eastern Africa \citep{Masenga.2016, Oneill.2020} but conflicts with findings by
-\cite{DaviesMostert.2012}, who reported a comparably high willingness of
-dispersers to cross human-dominated landscapes in South Africa. We believe that
-such differences are due to the unavailability of alternative ``natural''
-routes, which may have forced dispersers in South Africa to cross human
-dominated landscapes despite a strong aversion to do so. These contrasting
-findings may hint at a considerable ability of dispersing wild dogs to adapt to
-prevailing habitat conditions (see e.g. \cite{Woodroffe.2011}) and could imply a
-limited model transferability between dispersersers from our study system to
-dispersers found in other ecosystems, such as in South Africa
-\citep{Whittington.2011, DaviesMostert.2012} or Kenya \citep{Woodroffe.2019,
-Oneill.2020}. However, we believe that our study population adequately
-represents wild dog populations residing within the greater extent of the
-KAZA-TFCA, where more natural dispersal routes are readily available. Successful
-conservation of this species is thus strictly contingent on policymakers' and
-local authorities' willingness and ability to provide and conserve areas that
-remain free from anthropogenic pressures. This is not only paramount in light of
-providing connectivity and facilitating dispersal, but also in terms of reducing
-human-caused mortality during dispersal. In fact, previous studies have shown
-that human-caused mortality during dispersal, especially in proximity to
-settlements and roads, represents a major threat to the dispersal ability of
-this species \citep{Woodroffe.2019, Cozzi.2020}.
-
-Besides anthropogenic pressures, we identified water as another major obstacle
-to dispersal. This corroborates earlier studies which showed that water bodies
-(including swamps) are almost impenetrable by resident packs
-\citep{Abrahms.2017, Cozzi.2020} and only infrequently crossed by dispersing
-individuals \citep{Cozzi.2020}. An accurate and dynamic representation of water
-is thus imperative and particularly relevant in seasonal or flood-pulsing
-ecosystems such as the Okavango Delta. In fact, future studies could investigate
-how connectivity across such systems changes in result to seasonal changes in
-the environment (see also \cite{Osipova.2019}).
-
-Although dispersers avoided moving through water, we revealed a preference for
-moving in proximity to water. This preference is likely explained by high
-prey-availability in vicinity of water \citep{Western.1975, Bonyongo.2005}.
-However, following the same reasoning, apex predators (e.g. lions, spotted
-hyenas) and resident wild dogs may prefer proximity to water
-\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs into
-prey-poorer areas that are more distant to water \citep{Ndaimani.2016} and pose
-a lower risk of aggressive intraguild interactions \citep{Creel.1996,
-Mills.1997, Webster.2012}. We could not control for the presence or absence of
-apex predators or conspecifics during dispersal as we lacked corresponding data,
-but these considerations may explain why we did not find a significant effect of
-\textit{distance to water}. Given the influence that resident conspecifics,
-predators, competitors, and prey can have on dispersers \citep{Cozzi.2018,
-Armansin.2019} future studies should strive to collect and incorporate
-corresponding ``social'' information into analyses of landscape connectivity to
-more precisely reconstruct dispersal corridors.
-
-\section{Conclusion}
-Our work shows how dispersal data of a highly mobile species can be used
+Our work shows how dispersal data of a highly mobile species can be used to
 identify movement corridors and assess the adequacy of protected areas. In our
@@
 long-term viability of this species. Moreover, our connectivity network allowed
-to reveal potential dispersal hubs thorugh which dispersers gain access to more
+revealing potential dispersal hubs through which dispersers gain access to more
 remote regions in the study area. Finally, our investigations showed that human
@@
 Ultimately, our work contributes to the growing field of connectivity literature
-and provides and easily expandible framework for the assessement of existing or
+and provides and easily expandable framework for the assessement of existing or
 planned protected areas.

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Support.tex
first commit

\documentclass[abstract=off,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% For supplementary material
\newcommand{\beginappendix}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
  \setcounter{equation}{0}
  \renewcommand{\theequation}{Equation S\arabic{equation}}%
  \setcounter{section}{0}
  \renewcommand{\thesection}{A.\arabic{section}}%
}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{\textbf{Appendix}\\ Bound within Boundaries: How Well Do Protected Areas
Match Movement Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S,*} \and
  Dominik M. Behr\textsuperscript{1,2,*} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared first authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} dispersal, habitat selection, integrated step selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, least-cost corridors, Lycaon pictus, permeability surface,
protected areas, wildlife management

\end{flushleft}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

% Change to appendix counters
\appendix
\beginappendix

%------------------------------------------------------------------------------
%	Appendix S1: Net Squared Displacement
%------------------------------------------------------------------------------
\section{Net Squared Displacement}
\begin{figure}[hbpt]
  \begin{center}
    \includegraphics[width = \textwidth]{99_NSD.pdf}
    \caption{NSD displacement through time for one of our dispersers. The blue
    line indicates the period during which we calssified the individual as
    dispersing.}
    \label{NSD}
  \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S2: GPS Data
%------------------------------------------------------------------------------
\newpage
\section{GPS Data}
\begin{figure}[hbtp]
  \begin{center}
    \includegraphics[width = 0.86\textwidth]{99_Trajectories.pdf}
    \caption{Illustration of all trajectories that we recorded. Each color
    represents a different dispersing coalition. All coalitions departed from
    the area which is indicated by the white dashed rectangle. The coalition
    dispersing towards the far east of the map covered over 360 km in under 10
    days. Satellite background imagery was provided by Microsoft Bing.}
    \label{Trajectories}
  \end{center}
\end{figure}

\begin{table}[hbtp]
  \caption{Summary statistics of all GPS relocations that have been recorded on
  dispersing coalitions.}
  \label{GPSData}
  \begin{center}
    \resizebox{0.8\textwidth}{!}{
      \begin{tabular}{lccccccc}
      \toprule
      \parbox[]{2cm}{\centering Coalition ID} &
        \parbox[]{1cm}{\centering Sex} &
          \parbox[]{2cm}{\centering \ Pack \\ Affiliation} &
            \parbox[]{1.5cm}{\centering \# Fixes \\ Total} &
              \parbox[]{1.5cm}{\centering \# Fixes \\ During \\ Dispersal} &
                \parbox[]{1.5cm}{\centering \# Days \\ Dispersing} &
                  \parbox[]{2cm}{\centering Euclidean \\ Dispersal\\ Distance \\ (in km)} &
                    \parbox[]{2cm}{\centering Cumulative \\ Dispersal \\ Distance \\ (in km)} \\
      \midrule
      Abel & M & MU & 894 & 45 & 9 & 131 & 205 \\
      Amacuro & F & MB & 954 & 583 & 137 & 23 & 1'090 \\
      Belgium & M & ZU & 1'097 & 158 & 28 & 18 & 319 \\
      Dalwhinnie & F & PA & 545 & 62 & 22 & 50 & 243 \\
      Denali & F & MN & 1'096 & 173 & 33 & 11 & 528 \\
      Everest & M & MN & 389 & 123 & 38 & 67 & 572 \\
      Kalahari & F & HT & 1'753 & 467 & 130 & 20 & 1'963 \\
      Karisimbi & M & MN & 438 & 141 & 34 & 45 & 251 \\
      Liuwa & F & AP & 946 & 92 & 19 & 144 & 451 \\
      Lupe & M & KW & 2'209 & 396 & 34 & 8 & 436 \\
      MadameChing & F & AP & 776 & 729 & 136 & 263 & 1'560 \\
      Mirage & M & HT & 814 & 182 & 36 & 7 & 435 \\
      Odzala & M & AP & 1'410 & 205 & 42 & 53 & 412 \\
      Scorpion & M & KB & 2'676 & 393 & 34 & 4 & 471 \\
      Stetson & M & MT & 384 & 383 & 33 & 3 & 481 \\
      Taryn & F & AP & 896 & 37 & 9 & 10 & 130 \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

%------------------------------------------------------------------------------
%	Appendix S3: Spatial Covariates
%------------------------------------------------------------------------------
\newpage
\section{Spatial Covariates}
To investigate habitat preferences of dispersing wild dogs, we used a set of
geo-referenced covariates that we aggregated in the categories \textit{land
cover}, \textit{protection status}, and \textit{human influence}. We did not
include terrain features as covariates due to the absence of noteworthy
elevational gradients in our study area. For each covariate, we prepared spatial
raster layers from freely available online services or from remotely sensed
satellite imagery. To ensure a consistent resolution (i.e. cell-size or grain)
across covariates, we coarsened or interpolated all layers to match a resolution
of 250m x 250m. We performed processing and manipulation of data as well as all
spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.

\subsection{Land Cover}
\subsubsection{Water}
The covariate water included rivers, wetlands, and swamps. Because the
inundation extent of the flood in the Okavango Delta is highly variable within
and between years, we created dynamic ``flood maps'' that were updated every
8\textsuperscript{th} day following a remote sensing algorithm developed by the
Okavango Research Institute (ORI; \citealp{Wolski.2017}). To implement the
algorithm, we defined two sets of polygons located in the region of the Okavango
Delta (\Cref{Floodmapping}). The first set consisted of areas known to be
permanent dryland, whereas the second set consisted of permanent waters. Since
we were unable to retrieve the original polygons used in \cite{Wolski.2017}, we
geo-referenced and digitized the polygons reported in their publication. After
recreating the polygons, we used the R-package \textit{getSpatialData}
\citep{Schwalb.2018} to download and pre-process all relatively cloud-free MODIS
Terra images (MCD43A4; \citealp{Schaaf.2015}) available for the period of our
dispersal events. Assessment of cloud cover was based on visual inspection of
MODIS images on ORI's website
(\url{www.okavangodata.ub.bw/ori/monitoring/flood_maps}). After download, we
classified each MODIS image into a binary map of water (flood) and dryland using
a threshold that was identified as follows. First, we extracted all reflectance
values of MODIS Terra Band 7 within the water- and dryland-polygons. Second, we
computed histograms of water-reflectances and dryland-reflectances and
empirically verified that reflectances of the two groups were sufficiently
distinct. More specifically, we checked if superimposing the histograms of
water-reflectances and dryland-reflectances resulted in a bimodal histogram.
This was said to be achieved if the 99\textsuperscript{th} percentile of
water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
dryland}\)). Third, if bimodality was achieved, we calculated a threshold
(\(t\)) using \ref{EQ1}:

\begin{equation}
\label{EQ1}
t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
\widetilde{p}_{water})
\end{equation}

\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
the median reflectances of water and dryland, respectively. We then classified
all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
all pixels with a value smaller than \(t\) as water.

\begin{figure}[hbtp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
    \caption{Images describing the flood mapping algorithm. (a) The colored
    polygons indicate permanent waters (blue) and permanent dryland (green).
    Below these polygons we extracted reflectance values of MODIS Terra Band 7
    and used their repsective medians to calculate a classification threshold
    \(t\). (b) Example of a classified MODIS Terra Band 7 image after
    application of the threshold. The satellite image in the background was
    provided by Microsoft Bing.}
    \label{Floodmapping}
  \end{center}
\end{figure}

% Dynamic Wetmask
\noindent Importantly, bimodality was not always achieved and in some cases no
flood map could be calculated. In fact, it appears that non-bimodality caused
the ORI algorithm to fail since the end of 2018, which is why no flood maps have
been generated since then (ORI, personal comm.). We hypothesized that this was
caused by the application of static water-polygons that did not cover permanent
waters correctly anymore. Therefore, we revised the algorithm and allowed for a
more dynamic polygonization of water. That is, for each MODIS image we
calculated new water-polygons comprising areas that were covered by the flood in
99\% of the flood maps from the previous five years. All of the necessary flood
maps from previous years were kindly provided to us by ORI. Using this slightly
amended approach, we were able to address some of the bimodality issues and to
classify several additional flood maps for the period of our study. Because
MODIS Terra Band 7 had a resolution of 500m x 500m, we interpolated all maps to
250m x 250m.

% Validation
To validate and compare the performance of our own algorithm to the original
ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
images. To make sure that months were equally represented in the sampled dates,
we employed stratified sampling based on months (regardless of the year) and
randomly sampled four maps for each month. For the sampled dates we downloaded
and classified MODIS Terra Band 7 images and compared our classified images to
those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
created a difference map indicating false positives and false negatives and
computed the relative number of wrongly classified pixels. We achieved an
overall accuracy of 97\%, which presumably is an underestimate of the true
performance, as we introduced some errors when resampling the ORI-maps to our
reference grid.

\begin{figure}[hbtp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
    \caption{Validation procedure of our flood mapping algorithm. (a) Classified
    image that was provided to us by ORI. (b) Image for the same date but now
    classified using our own algorithm. (c) Difference image indicating false
    positives and false negatives in our own classification.}
    \label{FloodmappingValidation}
  \end{center}
\end{figure}

% Static Watermap
\noindent While we created dynamic flood maps for the Okavango Delta, we assumed
the extent of all other water bodies (e.g. Chobe river, Zambezi river) to be
static within and between years. This static representation was based on
Globeland's land cover dataset \citep{Chen.2015}, from which we only retained
the categories \textit{wetland} and \textit{water bodies} and collectively
reclassified them to \textit{water}. Globeland had an original resolution of 30m
x 30m, so we coarsened the layer to 250m x 250m using the mode of each 250m x
250m cell. We further improved river representation by employing the rasterized
MERIT Hydro dataset \citep{Yamazaki.2019} from which we added all rivers with a
width of over 10m to our Globeland layer. We merged dynamic and static water
maps into a large rasterstack, covering the entire study area. We also created a
rasterstack rendering the covariate \textit{distance to water} by calculating
the Euclidean distance of each raster cell in the study area to the nearest
source of water.

\subsubsection{Dryland}
We subdivided dryland into three layers as derived from the MODIS Terra
Vegetation Continuous Fields dataset (MOD44B; \citealp{Dimiceli.2015}). The
three layers depicted percentage cover of tree-vegetation (henceforth
\textit{trees}), non-tree-vegetation (henceforth \textit{shrubs/grassland}), and
non-vegetated (henceforth \textit{bare land}) and added up to 100\% of dryland
coverage. We used our flood map that aligned with the creation date of these
MODIS layers and defined anything covered by water as 0\% vegetated. The MODIS
vegetation layers had a resolution of 250m x 250m and no coarsening or
interpolation was required.

\subsection{Protection Status}
We created a binary layer separating protected from unprotected land. We
downloaded corresponding data on protection status in shapefile format from the
Peace Parks Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}).
Protected areas included forest reserves, game reserves, wildlife management
areas, and national parks. We classified anything not covered by these
categories as unprotected (e.g. communal pastoral land, private land). We
rasterized the two categories to the binary raster \textit{protection status} (1
= protected, 0 = unprotected) with a resolution of 250m x 250m.

\subsection{Human Influence}
We created a raster layer representing human influence by integrating
information on (1) human density, (2) farming, and (3) roads.

\begin{enumerate}[label = (\arabic*)]

  \item We obtained spatial human density estimates through a publicly available
  30m x 30m high-resolution population density dataset
  (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}). We coarsened the
  layer to 250m x 250m by summing up human density values within each 250m x
  250m cell.

  \item We sourced spatial information on farms from the Globeland
  \citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets from
  which we retained areas that were classified as either \textit{cultivated
  land} or \textit{croplands}. Any other land cover class was not pertinent to
  farming and therefore omitted. Because both layers had a resolution of 30m x
  30m we coarsened them to 250m x 250m by assigning a value of 1 to any 250m x
  250m cell that covered farmland and a value 0 otherwise. Thus, the final layer
  depicted presence (= 1) or absence (= 0) of farms within each 250m x 250m
  cell.

  \item We obtained geo-referenced data on roads from Open Street Map
  \citep{OpenStreetMap.2019}, downloaded through Geofabrik
  (\url{www.geofabrik.de}). We only retained main tarmac roads and omitted
  smaller roads (Table S2) as these are scarcely frequented and do not represent
  an obstacle to wild dog movements \citep{Abrahms.2016}. We rasterized main
  tarmac roads to the binary raster \textit{roads} (1 = roads, 0 = no roads)
  with 250m x 250m resolution. Finally, we created the covariate
  \textit{distance to roads} by calculating the Euclidean distance of each
  raster cell in the study area to the nearest road.

\end{enumerate}

\begin{table}[hbtp]
  \caption{Description of road types, as sourced from Open Street Map's mapping
  guide (\url{https://wiki.openstreetmap.org/wiki/Key:highway}). Roads types
  that were considered for the purpose of this study are shaded in light gray.}
  \label{RoadsDescription}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llp{12cm}}
      \toprule
      Group &
        Subgroup &
          Description \\
      \midrule
      \rowcolor[gray]{0.90}
      Roads &
        motorway &
          A restricted access major divided highway, normally with 2 or more
          running lanes plus emergency hard shoulder. Equivalent to the
          Freeway, Autobahn, etc. \\
      \rowcolor[gray]{0.90}
      Roads &
        trunk &
          The most important roads in a country's system that aren't
          motorways. Need not necessarily be a divided highway. \\
      \rowcolor[gray]{0.90}
      Roads &
        primary &
          The next most important roads in a country's system. Often link
          larger towns. \\
      \rowcolor[gray]{0.90}
      Roads &
        secondary &
          The next most important roads in a country's system. Often link
          towns. \\
      Roads &
        tertiary &
          The next most important roads in a country's system. Often link
          smaller towns and villages \\
      Roads &
        unclassified &
          The least important thorough roads in a country's system, i.e. minor
          roads of a lower classification than tertiary, but which serve a
          purpose other than access to properties. Often link villages and
          hamlets. \\
      Roads &
        residential &
          Roads which serve as an access to housing, without function of
          connecting settlements. Often lined with housing. \\
      Roads &
        service &
          For access roads to, or within an industrial estate, camp site,
          business park, car park etc. \\
      \rowcolor[gray]{0.90}
      Link roads &
        motorway\_link &
          The link roads (sliproads/ramps) leading to/from a motorway from/to
          a motorway or lower class highway. Normally with the same motorway
          restrictions. \\
      \rowcolor[gray]{0.90}
      Link roads &
        trunk\_link &
          The link roads (sliproads/ramps) leading to/from a trunk road
          from/to a trunk road or lower class highway. \\
      \rowcolor[gray]{0.90}
      Link roads &
        primary\_link &
          The link roads (sliproads/ramps) leading to/from a primary road
          from/to a primary road or lower class highway. \\
      \rowcolor[gray]{0.90}
      Link roads &
        secondary\_link &
          The link roads (sliproads/ramps) leading to/from a secondary road
          from/to a secondary road or lower class highway. \\
      Link roads &
        tertiary\_link &
          The link roads (sliproads/ramps) leading to/from a tertiary road
          from/to a tertiary road or lower class highway. \\
      Special road types &
        living\_street &
          For living streets, which are residential streets where pedestrians
          have legal priority over cars, speeds are kept very low and where
          children are allowed to play on the street. \\
      Special road types &
        pedestrian &
          For roads used mainly/exclusively for pedestrians in shopping and
          some residential areas which may allow access by motorised vehicles
          only for very limited periods of the day. \\
      Special road types &
        track &
          Roads for mostly agricultural or forestry uses. \\
      Special road types &
        bus\_guideway &
          A busway where the vehicle is guided by the way (though not a railway)
          and is not suitable for other traffic. \\
      Special road types &
        escape &
          For runaway truck ramps, runaway truck lanes, emergency escape
          ramps, or truck arrester beds. It enables vehicles with braking
          failure to safely stop. \\
      Special road types &
        raceway &
          A course or track for racing \\
      Special road types &
        road &
          A road/way/street/motorway/etc. of unknown type. It can stand for
          anything ranging from a footpath to a motorway. \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\noindent Because layers (1), (2), and (3) depicted features that are typically
spatially clustered and because not all dispersing coalitions moved within
meaningful distance to each of these features, we totaled values from the layers
describing \textit{human density} (continuous), \textit{farming} (binary), and
\textit{roads} (binary). This approach implied that \textit{roads} and
\textit{farms} entered the final layer with a value of 1, whereas human density
entered the final layer with a value \(\geq 0\) and potentially unbound. To
reduce the influence of outliers in human density estimates, totaled values were
limited to a maximum of 50, which visually resulted in a good balance between
high and low anthropogenic influence and was therefore considered appropriate
for our analysis. To render the fact that humans influence their surroundings
beyond their presence, we followed \cite{Elliot.2014} and applied to each
raster-cell a 5km focal buffer within which we summed up and log-transformed
human-influence values.

\begin{figure}[hbtp]
  \begin{center}
    \includegraphics[width = 0.95\textwidth]{99_HumanInfluence.pdf}
    \caption{Sequence of figures that exemplifies how we combined the layers
    (a), (b), and (c) into a single layer for human influence (d). For better
    visibility we show the procedure only for the extent of the Okavango Delta.
    The layer in (a) is based on Facebook's high resolution human density
    dataset (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}) and depicts
    the estimated number of humans living in each 250m x 250m raster-cell
    (coarsened from 30m x 30m). The layer in (b) is a binary layer and shows
    whether raster-cells are cover any sort of agricultural fields.
    Corresponding data was obtained through the Globeland and Cropland land
    cover datasets \citep{Chen.2015, Xiong.2017}. The layer in (c) shows the
    presence or absence of roads and is based on data from Open Street Map
    \citep{OpenStreetMap.2019}. We merged the layers in (a), (b), and (c) by
    summing up their values, truncating the summed values to a maximum of 50. We
    then log-transforming the values and applied to each raster cell a focal
    buffer of 5km within which we totaled human influence values. The layer in
    (d) depicts the final human influence layer that entered our habitat
    selection model.}
    \label{HumanInfluence}
  \end{center}
\end{figure}

%------------------------------------------------------------------------------
%  Appendix S4: Flood Pulse
%------------------------------------------------------------------------------
\newpage
\section{Typical Flood Pulse}
\begin{figure}[hbtp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
    \caption{Sequence of flood maps showing a typical flood pulse throughout the
    year. The flood arrives from the north-western corner (so called
    ``pan-handle'') of the Okavango Delta and slowly descends through the delta
    in south-eastern direction, where it nourishes several tributaries. The
    extent of the flood peaks around August or September and then slowly
    retracts. Between December and March the reflectance properties of water and
    dryland change, which is why often no accurate flood maps can be obtained
    for these months using remote sensing techniques \citep{Wolski.2017}.}
    \label{FloodPulse}
  \end{center}
\end{figure}

%------------------------------------------------------------------------------
%  Appendix S5: Step Selection Analysis
%------------------------------------------------------------------------------
\newpage
\section{Integrated Step Selection Function}
We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
investigate dispersers' selection or avoidance of spatial covariates. In the
iSSF framework, covariates experienced along realized steps are contrasted with
covariates experienced along alternative random steps that the animal could have
taken but decided not to. A step in this framework is defined as the connecting
line between two consecutive GPS relocations \citep{Turchin.1998}. In contrast
to regular SSFs, iSSFs require to include movement metrics as covariates in the
corresponding conditional logistic regression model. Their inclusion, in turn,
allows simultaneous inference on habitat and movement preferences, as well as to
reduce potential biases in estimated habitat preferences \citep{Forester.2009,
Warton.2013, Avgar.2016}.

To conduct iSSF analysis, we followed the recommendations described in Appendix
S1 of the publication by \cite{Avgar.2016}. We prepared our GPS relocation data
for iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
relocations recorded during dispersal to steps that were regularly spaced four
hours apart. Steps that were separated by more than four hours (e.g. due to GPS
failure) were omitted from further analysis (allowing for a minor mismatch of up
to 15 minutes). Each remaining step was paired with 24 random steps, generated
by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
lengths from a gamma distribution that was fitted using realized step lengths
(shape = 0.3677, scale = 6'302). Together, a realized and its 24 associated
random steps formed a stratum of 25 steps that received a unique identifier.

We extracted spatial covariates along realized and random steps
(\Cref{Appendix:Sources}). For continuous covariates, we calculated the average
value, for categorical covariates the percentage cover along the step. We
further derived a binary variable indicating whether a step crossed a road. We
square-rooted extracted values to render a decreasing marginal impact of
distance. We scaled covariates using a z-score transformation and screened for
correlation using Pearson's Correlation Coefficient. None of the covariates were
overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we retained all of
them for modeling. Despite the covariates mentioned in \Cref{Appendix:Sources},
we included two movement metrics, namely the cosine of the turning angle
(\(cos(ta)\)) and the logarithm of the step length (\(log(sl)\)) in our
regression model \citep{Avgar.2016}. The movement metric \(cos(ta)\) serves to
describe the directionality of a step, as it transforms the circular measure of
(\(-\pi\) to \(\pi\)) into a linear measure (-1, 1). Thus, positive values
indicate forward movements, whereas negative values indicated backward movements
\citep{Turchin.1998}. The movement metric \(log(sl)\), on the other hand, is as
an indicator of the preferred step length. Since in our case steps were spaced
by four hours, \(log(sl)\) can also be interpreted as movement rate.

\begin{table}[hbtp]
  \begin{center}
    \caption{Overview of spatial covariates and their sources. We extracted
    covariates along realized and random steps. For continuous covariates we
    calculated average values along steps, for categorical covariates the
    percentage coverage along steps. We also prepared two covariates indicating
    the distance to water and distance to roads, respectively. We square-rooted
    the values for these two covariates to render a decreasing marginal impact
    of the effect of distance. Finally, we derived a binary indicator of whether
    a step crossed a road or not.}
    \label{Appendix:Sources}
    \resizebox{\textwidth}{!} {
      \begin{threeparttable}
        \begin{tabular}{lllll}
        \hline
        Category &
          Covariate &
            Description &
              Values &
                Source \\
        \midrule
        \multirow{5}{*}{Land Cover}
          & Water
            & Percentage cover of water
              & 0-100\%
                & (1) (2) (3) \\
          & Dryland*
            & Percentage cover of dryland
              & 0-100\%
                & (1) (2) (3) \\
          & DistanceToWater
            & Average distance to nearest water source
              & \(\geq\) 0m
                & (1) (2) (3) \\
          & Shrubs/Grassland
            & Average non-tree vegetation
              & 0-100\%
                & (4) \\
          & Trees
            & Average tree-vegetation
              & 0-100\%
                & (4) \\
          & Bareland*
            & Average non-vegetated area
              & 0-100\%
                & (4) \\
        \hdashline
        \multirow{2}{*}{Protection Status}
          & Protected
            & Percentage cover of protected area
              & 0-100\%
                & (5) \\
          & Unprotected*
            & Percentage cover of unprotected area
              & 0-100\%
                & (5) \\
        \hdashline
        \multirow{3}{*}{Human Influence}
          & Human Influence
            & Average human influence
              & \(\geq\) 0
                & (1) (6) (7) (8)\\
          & DistanceToRoads
            & Average distance to nearest road
              & \(\geq\) 0m
                & (1) (6) (7) (8)\\
          & RoadCrossing
            & Binary; whether a step crossed a road
              & 0, 1
                & (1) (6) (7) (8)\\
        \hline
        \end{tabular}
        \begin{tablenotes}
          \item \textit{Sources:} (1) \cite{Chen.2015} (2) \cite{Schaaf.2015} (3)
          \cite{Yamazaki.2019} (4) \cite{Dimiceli.2015} (5) \cite{PeaceParks.2019}
          (6) \cite{Facebook.2019} (7) \cite{OpenStreetMap.2019} (8)
          \cite{Xiong.2017}
          \item
          \item \textit{* Note:} The covariates \textit{Water} and
          \textit{Dryland} added up to 100\%, which is why only \textit{Water} was
          included as explanatory variable in our models. The same applied for the
          group \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland},
          where we omitted \textit{Bareland} for modeling. Finally, from the group
          \textit{Protected} and \textit{Unprotected}, we only included
          \textit{Protected} in our models.
        \end{tablenotes}
      \end{threeparttable}
    }
  \end{center}
\end{table}

\noindent We then used the iSSF framework to parameterize a habitat selection
model that further served to predict landscape permeability. This habitat
selection model operated under the assumption that dispersing wild dogs assigned
a selection score \(w(x)\) of the following exponential form to each realized
and random step:

\begin{equation}
\label{EQ2}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent That is, the selection score \(w(x)\) of a step depended on its
associated covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's
preferences for these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
probability that a step \(i\) was realized \(P(Y_{i} = 1\)) was then contingent
on the step's selection score, as well as on the selection scores of all
alternative steps in the stratum:

\begin{equation}
\label{EQ3}
  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
\end{equation}

\noindent Habitat and movement preferences of interest, i.e. the \(\beta\)'s,
were then estimated by comparing realized (scored 1) and random (scored 0) steps
in a conditional logistic regression model \citep{Fortin.2005}. In this model,
positive \(\beta\)-coefficients indicate selection of a covariate, negative
\(\beta\)-coefficients avoidance of a covariate. To deal with multiple
individuals, we applied mixed effects conditional logistic regression analysis
following \cite{Muff.2020}. We implemented their method using the R-package
\textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID to model
random intercepts and slopes.

We defined the movement metrics \(cos(ta)\) and \(log(sl)\) as core covariates
and ran forward model selection based on Akaike's Information Criterion (AIC;
\citealp{Burnham.2002}) for all other covariates. We ranked models according to
AIC, assessed relative model weights, and identified the most parsimonious
model. Due to convergence issues, we were unable to model interactions between
covariates.

To validate the predictive power of the most parsimonious habitat selection
model, we ran k-fold cross-validation for case-control studies as described in
\cite{Fortin.2009}. Using 80\% of randomly selected strata, we parameterized a
habitat selection model and predicted selection scores \(w(x)\) for all steps in
the remaining 20\% of strata. According to predicted selection scores we
assigned ranks 1-25 within each stratum, with rank 1 indicating the highest
selection score. We identified the realized step's rank in each stratum and
tallied rank frequencies of realized steps across all strata. Finally, we
carried out a Spearman-rank correlation analysis between ranks and associated
frequencies and we recorded the correlation coefficient (\(r_{s, realized}\)).
We repeated this procedure 100 times with replacement and computed the mean
correlation coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\%
confidence interval. For comparison, we also repeated the same procedure 100
times assuming completely randomized preferences. We implemented randomized
preferences by omitting the realized step from each stratum and identifying the
rank of a randomly chosen random step within each stratum (now only ranks 1-24).
Again, we calculated Spearman's rank correlation coefficient (\(r_{s,
random}\)), its mean across repetitions (\(\bar{r}_{s, random}\)), and its 95\%
confidence interval. Ultimately, the validation proved a significant prediction
in case the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
random}\) did not overlap.

%------------------------------------------------------------------------------
%  Appendix S6: Least Cost Paths and Least Cost Corridors
%------------------------------------------------------------------------------
\newpage
\section{Identification of Least-Cost Paths \& Corridors}
\subsection{Least-Cost Paths}
We implemented factorial LCP analysis between source points using the R-package
\textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). The package translated
the (unscaled) permeability surface into a network of nodes to find shortest
effective distances between source points based on probabilities of moving from
cell to cell. In our case, the transition probability of moving between two
adjacent cells depended on their averaged permeability. We allowed individuals
to move from each cell to the cell's eight surrounding neighbors (i.e. Moores
neighborhood) and applied a geographic correction to account for the fact that
diagonal neighbors were more remote than orthogonal neighbors. Because African
wild dogs have been observed to cover large dispersal distances
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}, we did not limit LCPs to a
maximal effective cost. After computation, we tallied overlapping LCPs and
identified high-frequency routes.

\subsection{Least-Cost Corridors}
We calculated factorial LCCs \citep{Pinto.2009, Sawyer.2011, Elliot.2014}, again
using the R-package \textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). To
identify LCCs, we first computed for each source point a cumulative cost map,
which indicated the total minimal costs required to get from the source point to
any other location in the study area. We then obtained an LCC between two source
points by adding up their cumulative cost maps and masking out all cell-values
exceeding the lowest cell-value by more than 5\% \citep{Pinto.2009}. We repeated
this procedure for each possible unique pairwise combination of source points
and thereby identified LCCs between all 68 selected source points. We normalized
the resulting corridor-maps to range from zero to one and tallied them into a
single connectivity map.

\begin{figure}[hbtp]
  \begin{center}
  \begin{minipage}{.32\linewidth}
    \begin{subfigure}[t]{\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example1).pdf}
    \end{subfigure}
  \end{minipage}
  \begin{minipage}{.32\linewidth}
    \begin{subfigure}[t]{\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example2).pdf}
    \end{subfigure}
  \end{minipage}
  \begin{minipage}{.32\linewidth}
    \begin{subfigure}[t]{\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example3).pdf}
    \end{subfigure}
  \end{minipage}
  \caption{Images illustrating the process of identifying a least-cost corridor
  between source points A and B following \cite{Pinto.2009}. (a) Example of a
  permeability surface, which determines the costs of movement. (b1) Cumulative
  cost map for point A, depicting the total minimal costs necessary to get from
  point A to every other location. (b2) Cumulative cost map for point B,
  depicting the total minimal costs to get from point B to every other location.
  (c) Summed cost maps of points A and B. (d) Masked out corridor containing
  pixels that do not exceed the cheapest pixel by more than 5\%.}
  \label{LCCExample}
  \end{center}
\end{figure}

%------------------------------------------------------------------------------
%  Appendix S7: AIC Model Selection
%------------------------------------------------------------------------------
\newpage
\section{Model Selection Results}
\begin{table}[hbpt]
  \caption{Results from the forward model selection procedure based on Akaike's
  Information Criterion (AIC; \citealp{Burnham.2002}) for the habitat selection
  model. The most parsimonious model outperformed all other models (\(\Delta AIC
  > \) 2) and received a weight of one.}
  \label{ModelAICs}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{threeparttable}
        \begin{tabular}{lllll}
        \toprule
          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik \\
          \midrule
          cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\
          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\
          cos(ta) + log(sl) + W + T + DTW + HI & 90094.79 & 3.38 & 0.00 & -45034.39 \\
          cos(ta) + log(sl) + W + T + DTW + HI + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\
          cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\
          cos(ta) + log(sl) + W + T + DTW + HI + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\
          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\
          cos(ta) + log(sl) + W + T + DTW + HI + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\
          cos(ta) + log(sl) + W + T + DTW + HI + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\
          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\
          cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\
          cos(ta) + log(sl) + W + T + DTW + HI + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\
          cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\
          cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\
          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\
          cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\
          cos(ta) + log(sl) + W + T + HI & 90123.71 & 32.31 & 0.00 & -45050.85 \\
          cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\
          cos(ta) + log(sl) + W + T + S & 90129.36 & 37.96 & 0.00 & -45053.68 \\
          cos(ta) + log(sl) + W + T + DTR & 90130.11 & 38.71 & 0.00 & -45054.05 \\
          cos(ta) + log(sl) + W + T + P & 90131.75 & 40.35 & 0.00 & -45054.88 \\
          cos(ta) + log(sl) + W + T + RC & 90133.53 & 42.13 & 0.00 & -45053.77 \\
          cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\
          cos(ta) + log(sl) + W + HI & 90149.30 & 57.90 & 0.00 & -45065.65 \\
          cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\
          cos(ta) + log(sl) + W & 90153.70 & 62.30 & 0.00 & -45069.85 \\
          cos(ta) + log(sl) + W + DTR & 90156.32 & 64.92 & 0.00 & -45069.16 \\
          cos(ta) + log(sl) + W + P & 90157.23 & 65.83 & 0.00 & -45069.62 \\
          cos(ta) + log(sl) + W + RC & 90159.32 & 67.91 & 0.00 & -45068.66 \\
          cos(ta) + log(sl) + S & 90161.07 & 69.67 & 0.00 & -45073.54 \\
          cos(ta) + log(sl) + DTW & 90262.03 & 170.63 & 0.00 & -45124.01 \\
          cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\
          cos(ta) + log(sl) + HI & 90307.55 & 216.15 & 0.00 & -45146.78 \\
          cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\
          cos(ta) + log(sl) + P & 90316.62 & 225.22 & 0.00 & -45151.31 \\
          cos(ta) + log(sl) + RC & 90318.46 & 227.06 & 0.00 & -45150.23 \\
         \bottomrule
       \end{tabular}
       \begin{tablenotes}
         \item \textit{Note:} W = Water, DTW = Distance To Water, S =
         Shrubs/Grassland, T = Trees, P = Protected, HI = Human Influence, RC =
         Road Crossing, DTR = Distance To Roads.
       \end{tablenotes}
     \end{threeparttable}
    }
  \end{center}
\end{table}

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}

### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Support_Backup_2020-06-12.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Support.tex
@@
 \textbf{Keywords:} dispersal, habitat selection, integrated step selection
-function, Kavango-Zambezi Transfrontier Conservation Area, landscape
+function,Kavango-Zambezi Transfrontier Conservation Area, landscape
 connectivity, least-cost corridors, Lycaon pictus, permeability surface,
@@
     represents a different dispersing coalition. All coalitions departed from
-    the area which is indicated by the white dashed rectangle. The coalition
-    dispersing towards the far east of the map covered over 360 km in under 10
-    days. Satellite background imagery was provided by Microsoft Bing.}
+    the core study area, which is indicated by the white rectangle. The
+    coalition dispersing towards the far east of the map covered over 360 km in
+    under 10 days. Satellite background imagery was provided by Microsoft Bing.}
     \label{Trajectories}
@@
   \begin{center}
-    \resizebox{0.8\textwidth}{!}{
+    \resizebox{0.9\textwidth}{!}{
       \begin{tabular}{lccccccc}
@@
       \midrule
-      Abel & M & MU & 894 & 45 & 9 & 131 & 205 \\
       Amacuro & F & MB & 954 & 583 & 137 & 23 & 1'090 \\
@@
 \section{Spatial Covariates}
-To investigate habitat preferences of dispersing wild dogs, we used a set of
-geo-referenced covariates that we aggregated in the categories \textit{land
-cover}, \textit{protection status}, and \textit{human influence}. We did not
-include terrain features as covariates due to the absence of noteworthy
-elevational gradients in our study area. For each covariate, we prepared spatial
-raster layers from freely available online services or from remotely sensed
-satellite imagery. To ensure a consistent resolution (i.e. cell-size or grain)
-across covariates, we coarsened or interpolated all layers to match a resolution
-of 250m x 250m. We performed processing and manipulation of data as well as all
-spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.
+\subsection{Land Cover}
+We divided land cover into water and dryland. Water included rivers, wetlands,
+and swamps. Because the inundation extent of the flood in the Okavango Delta is
+highly variable within and between years, we created dynamic ``flood maps'' that
+were updated every 8\textsuperscript{th} day following a remote sensing
+algorithm developed by \cite{Wolski.2017}. The algorithm uses thresholding of
+short wavelength infrared reflectances of MODIS Terra satellite imagery
+(MCD43A4; \citealp{Schaaf.2015}) to distinguish areas covered by water and
+dryland (details in Appendix A.3). While we created dynamic flood maps for the
+Okavango Delta, we assumed the extent of all other water bodies (e.g. Chobe
+river, Zambezi river) to be static within and between years. This static
+representation was based on Globeland's land cover dataset \citep{Chen.2015},
+from which we only retained the categories \textit{wetland} and \textit{water
+bodies} and collectively reclassified them to \textit{water}. Globeland had an
+original resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using
+the mode of each 250m x 250m cell. We further improved river representation by
+employing the rasterized MERIT Hydro dataset \citep{Yamazaki.2019} from which we
+added all rivers with a width of over 10m to our Globeland layer. We merged
+dynamic and static water maps into a large rasterstack, covering the entire
+study area. We also created a rasterstack rendering the covariate
+\textit{distance to water} by calculating the eculidean distance of each raster
+cell in the study area to the nearest source of water.
 
-\subsection{Land Cover}
-\subsubsection{Water}
-The covariate water included rivers, wetlands, and swamps. Because the
-inundation extent of the flood in the Okavango Delta is highly variable within
-and between years, we created dynamic ``flood maps'' that were updated every
-8\textsuperscript{th} day following a remote sensing algorithm developed by the
-Okavango Research Institute (ORI; \citealp{Wolski.2017}). To implement the
-algorithm, we defined two sets of polygons located in the region of the Okavango
-Delta (\Cref{Floodmapping}). The first set consisted of areas known to be
-permanent dryland, whereas the second set consisted of permanent waters. Since
-we were unable to retrieve the original polygons used in \cite{Wolski.2017}, we
-geo-referenced and digitized the polygons reported in their publication. After
-recreating the polygons, we used the R-package \textit{getSpatialData}
-\citep{Schwalb.2018} to download and pre-process all relatively cloud-free MODIS
-Terra images (MCD43A4; \citealp{Schaaf.2015}) available for the period of our
-dispersal events. Assessment of cloud cover was based on visual inspection of
-MODIS images on ORI's website
-(\url{www.okavangodata.ub.bw/ori/monitoring/flood_maps}). After download, we
-classified each MODIS image into a binary map of water (flood) and dryland using
-a threshold that was identified as follows. First, we extracted all reflectance
-values of MODIS Terra Band 7 within the water- and dryland-polygons. Second, we
-computed histograms of water-reflectances and dryland-reflectances and
-empirically verified that reflectances of the two groups were sufficiently
-distinct. More specifically, we checked if superimposing the histograms of
-water-reflectances and dryland-reflectances resulted in a bimodal histogram.
-This was said to be achieved if the 99\textsuperscript{th} percentile of
-water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
-of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
-dryland}\)). Third, if bimodality was achieved, we calculated a threshold
-(\(t\)) using \ref{EQ1}:
-
-\begin{equation}
-\label{EQ1}
-t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
-\widetilde{p}_{water})
-\end{equation}
-
-\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
-the median reflectances of water and dryland, respectively. We then classified
-all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
-all pixels with a value smaller than \(t\) as water.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
-    \caption{Images describing the flood mapping algorithm. (a) The colored
-    polygons indicate permanent waters (blue) and permanent dryland (green).
-    Below these polygons we extracted reflectance values of MODIS Terra Band 7
-    and used their repsective medians to calculate a classification threshold
-    \(t\). (b) Example of a classified MODIS Terra Band 7 image after
-    application of the threshold. The satellite image in the background was
-    provided by Microsoft Bing.}
-    \label{Floodmapping}
-  \end{center}
-\end{figure}
-
-% Dynamic Wetmask
-\noindent Importantly, bimodality was not always achieved and in some cases no
-flood map could be calculated. In fact, it appears that non-bimodality caused
-the ORI algorithm to fail since the end of 2018, which is why no flood maps have
-been generated since then (ORI, personal comm.). We hypothesized that this was
-caused by the application of static water-polygons that did not cover permanent
-waters correctly anymore. Therefore, we revised the algorithm and allowed for a
-more dynamic polygonization of water. That is, for each MODIS image we
-calculated new water-polygons comprising areas that were covered by the flood in
-99\% of the flood maps from the previous five years. All of the necessary flood
-maps from previous years were kindly provided to us by ORI. Using this slightly
-amended approach, we were able to address some of the bimodality issues and to
-classify several additional flood maps for the period of our study. Because
-MODIS Terra Band 7 had a resolution of 500m x 500m, we interpolated all maps to
-250m x 250m.
-
-% Validation
-To validate and compare the performance of our own algorithm to the original
-ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
-images. To make sure that months were equally represented in the sampled dates,
-we employed stratified sampling based on months (regardless of the year) and
-randomly sampled four maps for each month. For the sampled dates we downloaded
-and classified MODIS Terra Band 7 images and compared our classified images to
-those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
-created a difference map indicating false positives and false negatives and
-computed the relative number of wrongly classified pixels. We achieved an
-overall accuracy of 97\%, which presumably is an underestimate of the true
-performance, as we introduced some errors when resampling the ORI-maps to our
-reference grid.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
-    \caption{Validation procedure of our flood mapping algorithm. (a) Classified
-    image that was provided to us by ORI. (b) Image for the same date but now
-    classified using our own algorithm. (c) Difference image indicating false
-    positives and false negatives in our own classification.}
-    \label{FloodmappingValidation}
-  \end{center}
-\end{figure}
-
-% Static Watermap
-\noindent While we created dynamic flood maps for the Okavango Delta, we assumed
-the extent of all other water bodies (e.g. Chobe river, Zambezi river) to be
-static within and between years. This static representation was based on
-Globeland's land cover dataset \citep{Chen.2015}, from which we only retained
-the categories \textit{wetland} and \textit{water bodies} and collectively
-reclassified them to \textit{water}. Globeland had an original resolution of 30m
-x 30m, so we coarsened the layer to 250m x 250m using the mode of each 250m x
-250m cell. We further improved river representation by employing the rasterized
-MERIT Hydro dataset \citep{Yamazaki.2019} from which we added all rivers with a
-width of over 10m to our Globeland layer. We merged dynamic and static water
-maps into a large rasterstack, covering the entire study area. We also created a
-rasterstack rendering the covariate \textit{distance to water} by calculating
-the Euclidean distance of each raster cell in the study area to the nearest
-source of water.
-
-\subsubsection{Dryland}
-We subdivided dryland into three layers as derived from the MODIS Terra
+We further subdivided dryland into three layers as derived from the MODIS Terra
 Vegetation Continuous Fields dataset (MOD44B; \citealp{Dimiceli.2015}). The
@@
 rasterized the two categories to the binary raster \textit{protection status} (1
-= protected, 0 = unprotected) with a resolution of 250m x 250m.
+= protected, 0 = pastoral) with a resolution of 250m x 250m.
 
@@
   with 250m x 250m resolution. Finally, we created the covariate
-  \textit{distance to roads} by calculating the Euclidean distance of each
+  \textit{distance to roads} by calculating the eculidean distance of each
   raster cell in the study area to the nearest road.
@@
 \end{enumerate}
+
+\noindent Because layers (1), (2), and (3) depicted features that are typically
+spatially clustered and because not all dispersing coalitions moved within
+meaningful distance to each of these features, we totaled the layers
+\textit{human density}, \textit{farming}, and \textit{roads}. This resulted in a
+single covariate called \textit{human influence}. Note that totaling the layers
+implied that the three covariates received different weights (see Appendix A.6
+for details). To reduce the influence of outliers, totaled values were limited
+to a maximum of 50, which visually resulted in a good balance between high and
+low anthropogenic influence and was therefore considered appropriate for our
+analysis. To render the fact that humans influence their surroundings beyond
+their presence, we followed \cite{Elliot.2014} and applied to each raster cell a
+5 km focal buffer within which we summed up and log-transformed human-influence
+values (Figure S6).
+
+%------------------------------------------------------------------------------
+%	Appendix S3: Floodmapping Algorithm
+%------------------------------------------------------------------------------
+\newpage
+\section{Flood-Mapping Algorithm}
+To implement the flood mapping algorithm, we defined two sets of polygons
+located in the region of the Okavango Delta (\Cref{Floodmapping}). The first set
+consisted of areas known to be permanent dryland, whereas the second set
+consisted of permanent waters. Since we were unable to retrieve the original
+polygons used in \cite{Wolski.2017}, we geo-referenced and digitized the
+polygons reported in their publication. After recreating the polygons, we used
+the R-package \textit{getSpatialData} \citep{Schwalb.2018} to download and
+pre-process all relatively cloud-free MODIS Terra images (MCD43A4;
+\citealp{Schaaf.2015}) available for the period of our dispersal events.
+Assessment of cloud cover was based on visual inspection of MODIS images on
+ORI's website \citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful
+for dynamic flood mapping because it is updated with new composite images every
+8\textsuperscript{th} day \citep{Wolski.2017}. After download, we classified
+each MODIS image into a binary map of water (flood) and dryland using a
+threshold that was identified as follows. First, we extracted all reflectance
+values of MODIS Terra Band 7 within the water- and dryland-polygons. Second, we
+computed histograms of water-reflectances and dryland-reflectances and
+empirically verified that reflectances of the two groups were sufficiently
+distinct. More specifically, we checked if superimposing the two histograms
+resulted in a bimodal plot. This was said to be achieved if the
+99\textsuperscript{th} percentile of water-reflectances did not severely exceed
+the 1\textsuperscript{st} percentile of dryland-reflectances (\(p_{0.99, water}
+- \frac{10}{255} < p_{0.01, dryland}\)). Third, if peaks were sufficiently
+different, we calculated a threshold (\(t\)) using \ref{EQ1}:
+
+\begin{equation}
+\label{EQ1}
+t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
+\widetilde{p}_{water})
+\end{equation}
+
+\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
+the median reflectances of water and dryland, respectively. We then classified
+all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
+all pixels with a value smaller than \(t\) as water.
+
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
+    \caption{Images describing the flood mapping algorithm. (a) The colored
+    polygons indicate permanent waters and permanent dryland. Below these
+    polygons we extracted reflectance values of MODIS Terra Band 7 and used
+    their repsective medians to calculate a classification threshold. (b)
+    Example of a classified MODIS Terra Band 7 image after application of the
+    threshold.}
+    \label{Floodmapping}
+  \end{center}
+\end{figure}
+
+% Dynamic Wetmask
+\noindent Importantly, bimodality was not always achieved and in some cases no
+flood map could be calculated. In fact, it appears that non-bimodality caused
+the ORI algorithm to fail since the end of 2018, such that no flood maps have
+been generated since then (ORI, personal comm.). We hypothesized that this was
+caused by the application of static water-polygons that did not cover permanent
+waters correctly anymore. Therefore, we revised the algorithm and allowed for a
+more dynamic polygonization of water. That is, for each MODIS image we
+calculated new water-polygons comprising areas that were covered by the flood in
+99\% of the flood maps from the previous five years. All of the necessary
+flood maps from previous years were kindly provided to us by ORI. Using this
+slightly amended approach, we were able to address some of the bimodality issues
+and to classify several additional flood maps for the period of our study.
+Because MODIS Terra Band 7 had a resolution of 500m x 500m, we interpolated all
+maps to 250m x 250m.
+
+% Validation
+To validate and compare the performance of our own algorithm to the original
+ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
+images. To make sure that months were equally represented in the sampled dates,
+we employed stratified sampling based on months (regardless of the year) and
+randomly sampled four maps for each month. For the sampled dates we downloaded
+and classified MODIS Terra Band 7 images and compared our classified images to
+those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
+created a difference map indicating false positives and false negatives and
+computed the relative number of wrongly classified pixels. We achieved an
+overall accuracy of 97\%, which presumably is an underestimate of the true
+performance, as we introduced some errors when resampling the ORI-maps to our
+reference raster.
+
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
+    \caption{Validation procedure of our flood mapping algorithm. (a) Classified
+    image that was provided to us by ORI. (b) Image for the same date but now
+    classified using our own algorithm. (c) Difference image indicating false
+    positives and false negatives in our own classification.}
+    \label{FloodmappingValidation}
+  \end{center}
+\end{figure}
+
+% Static Watermap
+\noindent While water in the Okavango Delta was mapped dynamically, we used a
+static watermap for its surroundings. This static map was based on Globeland's
+land cover dataset \citep{Chen.2015}, which contains ten land cover categories.
+From the original layer we only retained the categories \textit{wetland} and
+\textit{water bodies} and collectively reclassified them to \textit{water}.
+Globeland had an original resolution of 30m x 30m, so we coarsened the layer to
+250m x 250m using the mode of each 250m x 250m cell. Lastly, we improved river
+representation using the MERIT Hydro dataset \citep{Yamazaki.2019}, from which
+we added all rivers with a width of over 10m to our watermaps. The entire
+process resulted in several hundred watermaps, of which 90 represented the
+closest watermap to one of our GPS relocations.
+
+%------------------------------------------------------------------------------
+%	Appendix S4: Flood Pulse
+%------------------------------------------------------------------------------
+\newpage
+\section{Typical Flood Pulse}
+
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
+    \caption{Sequence of flood maps showing a typical flood pulse throughout the
+    year. The flood arrives from the north-western corner (so called
+    ``pan-handle'') of the Okavango Delta and slowly descents through the delta
+    in south-eastern direction, where it nourishes several tributaries. The
+    extent of the flood peaks around August or September and then slowly
+    retracts. Between December and March the reflectance properties of water and
+    dryland change, which is why often no accurate flood maps can be obtained
+    for these months using remote sensing techniques \citep{Wolski.2017}.}
+    \label{FloodPulse}
+  \end{center}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%  Appendix S5: Road Categories
+%------------------------------------------------------------------------------
+\newpage
+\section{Road Categories}
 
@@
 
-\noindent Because layers (1), (2), and (3) depicted features that are typically
-spatially clustered and because not all dispersing coalitions moved within
-meaningful distance to each of these features, we totaled values from the layers
-describing \textit{human density} (continuous), \textit{farming} (binary), and
+%------------------------------------------------------------------------------
+%	Appendix S6: Human Influence
+%------------------------------------------------------------------------------
+\newpage
+\section{Human Influence Proxy}
+Because the layers \textit{human density}, \textit{farming}, and \textit{roads}
+depicted features that were clustered in space, and because not all dispersing
+coalitions moved within meaningful distance of all these features, we merged
+them into a single general proxy rendering human influence
+\Cref{HumanInfluence}. That is, we totaled values from the layers describing
+\textit{human density} (continuous), \textit{farming} (binary), and
 \textit{roads} (binary). This approach implied that \textit{roads} and
@@
   \begin{center}
-    \includegraphics[width = 0.95\textwidth]{99_HumanInfluence.pdf}
-    \caption{Sequence of figures that exemplifies how we combined the layers
-    (a), (b), and (c) into a single layer for human influence (d). For better
-    visibility we show the procedure only for the extent of the Okavango Delta.
-    The layer in (a) is based on Facebook's high resolution human density
-    dataset (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}) and depicts
-    the estimated number of humans living in each 250m x 250m raster-cell
-    (coarsened from 30m x 30m). The layer in (b) is a binary layer and shows
-    whether raster-cells are cover any sort of agricultural fields.
-    Corresponding data was obtained through the Globeland and Cropland land
-    cover datasets \citep{Chen.2015, Xiong.2017}. The layer in (c) shows the
-    presence or absence of roads and is based on data from Open Street Map
-    \citep{OpenStreetMap.2019}. We merged the layers in (a), (b), and (c) by
-    summing up their values, truncating the summed values to a maximum of 50. We
-    then log-transforming the values and applied to each raster cell a focal
-    buffer of 5km within which we totaled human influence values. The layer in
-    (d) depicts the final human influence layer that entered our habitat
-    selection model.}
+    \includegraphics[width = \textwidth]{99_HumanInfluence.pdf}
+    \caption{Sequence of figures that exemplify how we combined the layers
+    depicted in (a), (b), and (c) into a single layer for human influence (d).
+    For better visibility we show the procedure only for the extent of the
+    Okavango Delta. The layer in (a) is based on Facebook's high resolution
+    human density dataset (\url{www.dataforgood.fb.com};
+    \citealp{Facebook.2019}) and depicts the estimated number of humans living
+    in each 250m x 250m raster-cell (coarsened from 30m x 30m). The layer in (b)
+    is a binary layer and shows whether raster-cells are cover any sort of
+    agricultural fields. Corresponding data was obtained through the Globeland
+    and Cropland land cover datasets \citep{Chen.2015, Xiong.2017}. The layer in
+    (c) shows the presence or absence of roads and is based on data from Open
+    Street Map \citep{OpenStreetMap.2019}. We merged the layers in (a), (b), and
+    (c) by summing up their values, truncating the summed values to a maximum of
+    50. We then log-transforming the values and applied to each raster cell a
+    focal buffer of 5km within which we totaled human influence values. The
+    layer in (d) depicts the final human influence layer that entered our
+    habitat selection model.}
     \label{HumanInfluence}
@@
 %------------------------------------------------------------------------------
-%  Appendix S4: Flood Pulse
+%	Appendix S7: Covariates
 %------------------------------------------------------------------------------
 \newpage
-\section{Typical Flood Pulse}
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
-    \caption{Sequence of flood maps showing a typical flood pulse throughout the
-    year. The flood arrives from the north-western corner (so called
-    ``pan-handle'') of the Okavango Delta and slowly descends through the delta
-    in south-eastern direction, where it nourishes several tributaries. The
-    extent of the flood peaks around August or September and then slowly
-    retracts. Between December and March the reflectance properties of water and
-    dryland change, which is why often no accurate flood maps can be obtained
-    for these months using remote sensing techniques \citep{Wolski.2017}.}
-    \label{FloodPulse}
-  \end{center}
-\end{figure}
-
-%------------------------------------------------------------------------------
-%  Appendix S5: Step Selection Analysis
-%------------------------------------------------------------------------------
-\newpage
-\section{Integrated Step Selection Function}
-We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
-investigate dispersers' selection or avoidance of spatial covariates. In the
-iSSF framework, covariates experienced along realized steps are contrasted with
-covariates experienced along alternative random steps that the animal could have
-taken but decided not to. A step in this framework is defined as the connecting
-line between two consecutive GPS relocations \citep{Turchin.1998}. In contrast
-to regular SSFs, iSSFs require to include movement metrics as covariates in the
-corresponding conditional logistic regression model. Their inclusion, in turn,
-allows simultaneous inference on habitat and movement preferences, as well as to
-reduce potential biases in estimated habitat preferences \citep{Forester.2009,
-Warton.2013, Avgar.2016}.
-
-To conduct iSSF analysis, we followed the recommendations described in Appendix
-S1 of the publication by \cite{Avgar.2016}. We prepared our GPS relocation data
-for iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
-relocations recorded during dispersal to steps that were regularly spaced four
-hours apart. Steps that were separated by more than four hours (e.g. due to GPS
-failure) were omitted from further analysis (allowing for a minor mismatch of up
-to 15 minutes). Each remaining step was paired with 24 random steps, generated
-by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
-lengths from a gamma distribution that was fitted using realized step lengths
-(shape = 0.3677, scale = 6'302). Together, a realized and its 24 associated
-random steps formed a stratum of 25 steps that received a unique identifier.
-
-We extracted spatial covariates along realized and random steps
-(\Cref{Appendix:Sources}). For continuous covariates, we calculated the average
-value, for categorical covariates the percentage cover along the step. We
-further derived a binary variable indicating whether a step crossed a road. We
-square-rooted extracted values to render a decreasing marginal impact of
-distance. We scaled covariates using a z-score transformation and screened for
-correlation using Pearson's Correlation Coefficient. None of the covariates were
-overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we retained all of
-them for modeling. Despite the covariates mentioned in \Cref{Appendix:Sources},
-we included two movement metrics, namely the cosine of the turning angle
-(\(cos(ta)\)) and the logarithm of the step length (\(log(sl)\)) in our
-regression model \citep{Avgar.2016}. The movement metric \(cos(ta)\) serves to
-describe the directionality of a step, as it transforms the circular measure of
-(\(-\pi\) to \(\pi\)) into a linear measure (-1, 1). Thus, positive values
-indicate forward movements, whereas negative values indicated backward movements
-\citep{Turchin.1998}. The movement metric \(log(sl)\), on the other hand, is as
-an indicator of the preferred step length. Since in our case steps were spaced
-by four hours, \(log(sl)\) can also be interpreted as movement rate.
+\section{Covariates \& Sources}
 
@@
                 & (5) \\
-          & Unprotected*
+          & Pastoral*
             & Percentage cover of unprotected area
@@
           where we omitted \textit{Bareland} for modeling. Finally, from the group
-          \textit{Protected} and \textit{Unprotected}, we only included
+          \textit{Protected} and \textit{Pastoral}, we only included
           \textit{Protected} in our models.
@@
 
-\noindent We then used the iSSF framework to parameterize a habitat selection
-model that further served to predict landscape permeability. This habitat
-selection model operated under the assumption that dispersing wild dogs assigned
-a selection score \(w(x)\) of the following exponential form to each realized
-and random step:
+\newpage
+\section{Integrated Step Selection Function}
+We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
+investigate dispersers' selection or avoidance of spatial covariates. In the
+iSSF framework, covariates experienced along realized steps are contrasted with
+covariates experienced along alternative random steps that the animal could have
+taken but decided not to. A step in this framework is defined as the connecting
+line between two consecutive GPS relocations \citep{Turchin.1998}. In contrast
+to regular SSFs, iSSFs require to include movement metrics as covariates in the
+corresponding conditional logistic regression model. Their inclusion, in turn,
+allows simultaneous inference on habitat and movement preferences, as well as to
+reduce potential biases in estimated habitat preferences \citep{Forester.2009,
+Warton.2013, Avgar.2016}.
+
+To conduct iSSF analysis, we followed the recommendations described in Appendix
+S1 of the publication by \cite{Avgar.2016}. We prepared our GPS relocation data
+for iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
+relocations recorded during dispersal to steps that were regularly spaced four
+hours apart. Steps that were separated by more than four hours (e.g. due to GPS
+failure) were omitted from further analysis (allowing for a minor mismatch of up
+to 15 minutes). Each remaining step was paired with 24 random steps, generated
+by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
+lengths from a gamma distribution that was fitted using realized step lengths
+(shape = 0.3677, scale = 6'302). Together, a realized and its 24 associated
+random steps formed a stratum of 25 steps that received a unique identifier.
+
+We extracted spatial covariates along realized and random steps. For continuous
+covariates, we calculated the average value, for categorical covariates the
+percentage cover along the step. We further derived a binary variable indicating
+whether or not a step crossed a road and we identified the average distance of
+each step to the nearest source of water or the nearest road. We square-rooted
+these values to render a decreasing marginal impact of distance. We standardized
+continuous covariates to a mean of zero and a standard deviation of one using a
+z-score transformation. Furthermore, we screened covariates for correlation
+using Pearson's Correlation Coefficient. None of the covariates were correlated
+(\(|r| > 0.6\); \citealp{Latham.2011}) and we retained all of them for modeling.
+In our regression model we included two additional covariates, namely the cosine
+of the turning angle (\(cos(ta)\)) and the logarithm of the step length
+(\(log(sl)\)) \citep{Avgar.2016}. The movement metric \(cos(ta)\) serves to
+describe the directionality of a step, as it transforms the circular measure of
+(\(-\pi\) to \(\pi\)) into a linear measure (-1, 1). Thus, positive values
+indicate forward movements, whereas negative values indicated backward movements
+\citep{Turchin.1998}. The movement metric \(log(sl)\), on the other hand, is as
+an indicator of the preferred step length. Since in our case steps were spaced
+by four hours, \(log(sl)\) can also be interpreted as movement rate.
+
+We then used the iSSF framework and parameterized a habitat selection model that
+further served to predict landscape permeability. This habitat selection model
+operated under the assumption that dispersing wild dogs assigned a selection
+score \(w(x)\) of the following exponential form to each realized and random
+step:
 
@@
 
-\noindent That is, the selection score \(w(x)\) of a step depended on its
-associated covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's
-preferences for these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
-probability that a step \(i\) was realized \(P(Y_{i} = 1\)) was then contingent
-on the step's selection score, as well as on the selection scores of all
-alternative steps in the stratum:
+The selection score \(w(x)\) of a step depended on its associated covariates
+(\(x_1, x_2, ..., x_n\)), as well as on the animal's preferences for these
+covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The probability that a step
+\(i\) was realized \(P(Y_{i} = 1\)) was then contingent on the step's selection
+score, as well as on the selection scores of all alternative steps in the
+stratum:
 
@@
 \noindent Habitat and movement preferences of interest, i.e. the \(\beta\)'s,
-were then estimated by comparing realized (scored 1) and random (scored 0) steps
-in a conditional logistic regression model \citep{Fortin.2005}. In this model,
+were estimated by comparing realized (scored 1) and random (scored 0) steps in a
+conditional logistic regression model \citep{Fortin.2005}. In this model,
 positive \(\beta\)-coefficients indicate selection of a covariate, negative
@@
 following \cite{Muff.2020}. We implemented their method using the R-package
-\textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID to model
+\textit{glmmTMB} \citep{Mollie.2017} and used dispersal coalition ID to model
 random intercepts and slopes.
@@
 and ran forward model selection based on Akaike's Information Criterion (AIC;
-\citealp{Burnham.2002}) for all other covariates. We ranked models according to
+\citealp{Burnham.2002} for all other covariates. We ranked models according to
 AIC, assessed relative model weights, and identified the most parsimonious
@@
 
-To validate the predictive power of the most parsimonious habitat selection
-model, we ran k-fold cross-validation for case-control studies as described in
+To validate the predictive power of the most parsimonious selection model, we
+ran k-fold cross-validation for case-control studies as described in
 \cite{Fortin.2009}. Using 80\% of randomly selected strata, we parameterized a
-habitat selection model and predicted selection scores \(w(x)\) for all steps in
-the remaining 20\% of strata. According to predicted selection scores we
-assigned ranks 1-25 within each stratum, with rank 1 indicating the highest
-selection score. We identified the realized step's rank in each stratum and
-tallied rank frequencies of realized steps across all strata. Finally, we
-carried out a Spearman-rank correlation analysis between ranks and associated
-frequencies and we recorded the correlation coefficient (\(r_{s, realized}\)).
-We repeated this procedure 100 times with replacement and computed the mean
-correlation coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\%
-confidence interval. For comparison, we also repeated the same procedure 100
-times assuming completely randomized preferences. We implemented randomized
-preferences by omitting the realized step from each stratum and identifying the
-rank of a randomly chosen random step within each stratum (now only ranks 1-24).
-Again, we calculated Spearman's rank correlation coefficient (\(r_{s,
-random}\)), its mean across repetitions (\(\bar{r}_{s, random}\)), and its 95\%
-confidence interval. Ultimately, the validation proved a significant prediction
-in case the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-random}\) did not overlap.
+selection model and predicted selection scores \(w(x)\) for all steps in the
+remaining 20\% of strata. According to predicted selection scores we assigned
+ranks 1-25 within each stratum, with rank 1 indicating the highest selection
+score. We identified the realized step's rank in each stratum and tallied rank
+frequencies of realized steps across all strata. Finally, we carried out a
+Spearman-rank correlation analysis between ranks and associated frequencies and
+we recorded the correlation coefficient (\(r_{s, realized}\)). We repeated this
+procedure 100 times with replacement and computed the mean correlation
+coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\% confidence
+interval. For comparison, we also repeated the same procedure 100 times assuming
+completely randomized preferences. We implemented randomized preferences by
+omitting the realized step from each stratum and identifying the rank of a
+randomly chosen random step within each stratum (now only ranks 1-24). Again, we
+calculated Spearman's rank correlation coefficient (\(r_{s, random}\)), its mean
+across repetitions (\(\bar{r}_{s, random}\)), and its 95\% confidence interval.
+Ultimately, the validation proofed a significant prediction in case the
+confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\)
+did not overlap.
 
-%------------------------------------------------------------------------------
-%  Appendix S6: Least Cost Paths and Least Cost Corridors
-%------------------------------------------------------------------------------
 \newpage
 \section{Identification of Least-Cost Paths \& Corridors}
-\subsection{Least-Cost Paths}
 We implemented factorial LCP analysis between source points using the R-package
@@
 
-\subsection{Least-Cost Corridors}
-We calculated factorial LCCs \citep{Pinto.2009, Sawyer.2011, Elliot.2014}, again
-using the R-package \textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). To
-identify LCCs, we first computed for each source point a cumulative cost map,
-which indicated the total minimal costs required to get from the source point to
-any other location in the study area. We then obtained an LCC between two source
-points by adding up their cumulative cost maps and masking out all cell-values
-exceeding the lowest cell-value by more than 5\% \citep{Pinto.2009}. We repeated
-this procedure for each possible unique pairwise combination of source points
-and thereby identified LCCs between all 68 selected source points. We normalized
-the resulting corridor-maps to range from zero to one and tallied them into a
-single connectivity map.
+We also calculated factorial LCCs \citep{Pinto.2009, Sawyer.2011, Elliot.2014},
+again using the R-package \textit{gdistance} (Figure S.7;
+\citealp{vanEtten.2017}). To identify LCCs, we first computed for each source
+point a cumulative cost map, which indicated the total minimal costs required to
+get from the source point to any other location in the study area. We then
+obtained an LCC between two source points by adding up their cumulative cost
+maps and masking out all cell-values exceeding the lowest cell-value by more
+than 5\% \citep{Pinto.2009}. We repeated this procedure for each possible unique
+pairwise combination of source points and thereby identified LCCs between all 68
+selected source points. We normalized the resulting corridor-maps to range from
+zero to one and tallied them into a single connectivity map.
 
@@
 
-%------------------------------------------------------------------------------
-%  Appendix S7: AIC Model Selection
-%------------------------------------------------------------------------------
 \newpage
 \section{Model Selection Results}
+
 \begin{table}[hbpt]
@@
           \midrule
-          cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\
-          cos(ta) + log(sl) + W + T + DTW + HI & 90094.79 & 3.38 & 0.00 & -45034.39 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 & 90094.79 & 3.38 & 0.00 & -45034.39 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\
           cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\
           cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\
           cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\
           cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\
+          cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\
           cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\
-          cos(ta) + log(sl) + W + T + HI & 90123.71 & 32.31 & 0.00 & -45050.85 \\
+          cos(ta) + log(sl) + W + T + HB5 & 90123.71 & 32.31 & 0.00 & -45050.85 \\
           cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\
@@
           cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\
-          cos(ta) + log(sl) + W + HI & 90149.30 & 57.90 & 0.00 & -45065.65 \\
+          cos(ta) + log(sl) + W + HB5 & 90149.30 & 57.90 & 0.00 & -45065.65 \\
           cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\
@@
           cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\
-          cos(ta) + log(sl) + HI & 90307.55 & 216.15 & 0.00 & -45146.78 \\
+          cos(ta) + log(sl) + HB5 & 90307.55 & 216.15 & 0.00 & -45146.78 \\
           cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\

### 2020-08-21 Chapter_1/04_Manuscript/99_MovementModelAICs.tex
first commit

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Thu Jul  9 09:17:26 2020
\begin{tabular}{lrrrr}
  \toprule
Covariates & AIC & DeltaAIC & Weight & LogLik \\ 
  \midrule
cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
   \bottomrule
\end{tabular}

### 2020-08-21 Chapter_1/04_Manuscript/Manuscript.tex
first commit

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
% \usepackage[nomarkers, nolists]{endfloat}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Step by Step: Using Integrated Step Selection Analysis to Simulate Wild
Dog Dispersal and Assess Landscape Connectivity}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2} \and
  Dominik M. Behr\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\vspace{4cm}

\textbf{Running Title:} Simulating Wild Dog Dispersal.

\vspace{0.5cm}

\textbf{Keywords:} dispersal, habitat selection, integrated step selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, least-cost corridors, Lycaon pictus, permeability surface,
protected areas, wildlife management

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
This is the abstract.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% % Create linenumbers
% \linenumbers

\section{Introduction}
An animals movement trajectory can be seen as the result of an interplay between
habitat and movement preferences.

In recent decades, least-cost analysis has become the workhorse to study
landscape connectivty at large scales. Originally introduced by XX, the concept
and relating methods have been refined to more realistically render animal's
movement capabilities and to determine valuable movement corridors that require
protection. (talk a bit about different methods of LCPs, see methods in
gdistance package). More recently, more attention has been assigned to the
collection of movement data is better tailored towards assessing landscape
connectivity. In particular, it has been proposed and verified that relocation
data collected dispersing individuals leads to more accurate and reliable
estimates of landscape connectivity. This superiority of dispersal data in
comparison to data that stem from residents is mainly owed to the fact that
animals behave vastly different during residency. Despite substantial advances
in ecologists ability to track animals in space and time, dispersal remains one
of the most difficult behavioral modes to observe in wild animals. Especially
for wide ranging and long-lived species, dispersal is difficult to predict and
observe, such that data remains scarce.

Many connectivity modelling techniques, especially least-cost analysis,
implicitly assume that the studied animal has partial or even complete knowledge
of the landscape and associated movement costs. While this assumption may be
reasonable for migrating animals that move between a limited number of habitats,
yet it is unlikely to hold for dispersers. In contrast to migrants, dispersers
move into unknown territory and are therefore confronted with novel landscapes.
Consequently, dispersers are more likely to adjust their movement behavior
\textit{ad hoc} instead of preplanning an entire trajectory. As a result,
methods that assume complete knowledge of the landscape and subsequent optimal
movement behavior likely misrepresent true dispersal behavior. As such, methods
that allow a more random approach to a dispersers movement behavior may more
realistically render the movement corridors of dispersing individuals.

In addition, individual based simulations will allow to explicitly model
dispersal between subpopulations in models of population dynamics.

Reliable identification of dispersal corridors will become increasingly
important with the uprise of ever-growing and often transboundary conservation
areas. One such instance is the KAZA-TFCA, a massive conservation area spanning
five countries and over 520'000 km\textsuperscript{2}. The KAZA holds the
potential of re-establishing dispersal routes for many of its protected species,
including the african wild dog \textit{Lycaon pictus}. This species has
experienced severe population declines caused by human induced mortality and
deadly diseases. In result, the species currently marks the KAZA's most
endangered large carnivore and has been a assigned a very high conservation
priority. Importantly, due to their inherent mobility and intrinsic need for
vast undisturbed landscapes, AWDs have been proposed as surrogate species for
landscape connectivity (see recent paper on multispecies connectivity).
Nevertheless, the species has received little attention in the connectivity
literature, mainly due to the difficulty in observing wild dog dispersal. In a
previous paper, we addressed this issue and developed a habitat selection model
based on which we predicted landscape connectivity using least-cost corridor
analysis. We now expand on this knowledge and develop a more detailed movement
model of dispersing wild dogs. We then use this model to simulate thousands of
dispersers moving throughout the KAZA. Based on said simulations, we compute
heatmaps and identify potential dispersal hotspots and compare them to the
dispersal routes identified in (Hofmann ....). We also apply the data to feed a
network-analysis, based on which we determine network-metrics pertinent to
landscape connectivity.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E;
\Cref{StudyArea}a) stretched over 1.3 Mio km\textsuperscript{2} and ecompassed
the entire KAZA-TFCA (\Cref{StudyArea}b). The KAZA-TFCA is the world's largest
transboundary conservation area and comprises parts of Angola, Botswana,
Namibia, Zimbabwe, and Zambia, covering a total area of over 520'000
km\textsuperscript{2}. Its landscape varies regionally and ranges from savanna,
to grassland, and from dry to moist woodland habitats. A dominant
hydrogeographical feature in its center is the Okavango Delta, the earth's
largest inland delta and home to a vast diversity of mammal species. The delta
and its surroundings are considered a stronghold for African wild dogs which may
act as a source for the recolonization of surrounding habitats. The wet season
in the region lasts from November to March, albeit the main floodwaters reach
the Delta between July and August after descending through the Angolian
highlands \citep{McNutt.1996, Wolski.2017}.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_StudyArea}
    \caption{Study area}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
Between 2011 and 2020, we collected GPS relocation data on free-ranging wild
dogs inhabiting the Okavango Delta in northern Botswana. We identified potential
dispersers based on criteria reported in \cite{Behr.2020}, immobilized them
according to protocols described in \cite{Osofsky.1996}, and fitted them with
GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
Berlin}). Handling and collaring of all individuals was carried out and
supervised by a Botswana-registered wildlife veterinarian. Of all collared
individuals, a total of 16 individuals dispersed, each in a same-sex coalition
(7 female and 9 male coalitions). When we observed dispersal we remotely
increased the GPS fixrate from 24 to 4 hours. Collected relocations were then
regularly transmitted via Iridium satellite system to a base station. This
allowed remote tracking of dispersers even if they left the main study area or
crossed international borders. To effectively distinguish between resident and
dispersersing individuals, we applied the net-squared displacement metric. This
metric measures the Euclidean distance of a collared individual to a reference
point \citep{Borger.2012}. In our case, these reference points resembled the
center of the dispersing coalition's natal home range. Hence, dispersal was
deemed to have started when an individual left its natal home range and ended
when the individual became stationary again. Previous research found no
differences between males and females during dispersal \citep{Woodroffe.2019,
Cozzi.2020}, so we did not distinguish sexes in our analyses. GPS relocations
were then converted to steps, where a step represented the straight-line
distance travelled between two consequtive GPS relocations
\citep{Turching.1998}.

\subsection{Covariates}
To represent environmental covariates spatially, we prepared a set of raster
layers depicting water-cover (dynamically updated), square rooted distance to
water (dynamically updated), tree-cover, and shrub/grassland-cover. We also
created a proxy for human influence, rendering anthropogenic pressures stemming
from human-density, agricultural sites, and roads. We prepared all layers at a
resolution of 250m by 250m for the extent of the entire KAZA-TFCA. A more
detailed description of the derivation and preparation of each environmental
covariate is given in \cite{Hofmann.2020}. Because wild dogs follow a diurnal
activity pattern, we coded a binary variable indicating whether an observed step
started during phases of main wild dog activity (07:00 a.m. and 07:00 p.m.) or
low wild dog activity (anything else).

\subsection{Movement Model}
We used integrated step selection functions (iSSF) to parametrize a mechanistic
movement model of dispersing wild dogs. In the iSSF framework, observed steps
receive a selection score \(w(x)\) that depends on the covariates \(X\)
experienced along a step, as well as the animals preferences towards these
covariates \(\beta\). In contrast to regular step selection analysis,
\textit{integrated} step selection analysis allows simulatneous inference on
movement and habitat preferences of the studied animal. Moreover, potential
interactions between habitat and movement preferences can be modelled. Thus, the
method produces more accurate selection estimates and allows to use the
resulting models as proper mechanistic movement models from which movement can
be generated. To conduct iSSf analysis, we paired each observed step with 24
random steps. Random steps basically resembled potential alternatives that the
animal could have realized but decided not to. We generated random steps by
sampling random turning angles from a uniform distribution (\(-\pi, +\pi\)) and
step lengths from a gamma distirbution that was fitted to observed steps (scale
= 6308, shape = 0.37). While the number of random steps is inversely
proportional to the sampling error \citep{Avgar.2016}, we found only minor
changes in inference when sampling further random steps. Along each step we
extracted spatial covariates using the \textit{velox} package. We also
calculated step metrics, namely the natural logarithm of the step length
\(log(sl\_\)) and the cosine of the turning angle \(cos(ta\_)\). We scaled all
continuous covariates to a mean of zero and a standard deviation of one. We then
used the r-package \textit{glmmTMB} to fit mixed effects conditional logistic
regression models as proposed by \citep{Muff.2020}. Our movement model was built
around a previously published habitat selection model of dispersing wild dogs.
Because the earlier model was used to predict landscape permeability,
interactions between environmental and spatial covariates were impossible to
model. Our current movement model, in contrast, allowed the inclusion of
additional interactions that describe differences in movement behavior induced
through different environments. Hence, we started with the habitat model and
iteratively increased model complexity by including additional interactions
between movement metrics and environmental covariates. That is, we proposed all
two-way interactions between spatial covariates and movement metrics. For
instance, for the covariate \textit{water} we proposed the interactions
Water:cos(ta\_) and Water:log(sl\_). Despite interactions between environmental
factors and movement metrics, we also proposed the interaction
\(log(sl\_):MainActivity\) to account for the diurnal activity pattern of wild
dogs. A preliminary revealed little differences in turning angles during main
and low activity, hence we did not include the interaction
\(cos(ta\_):MainActivity\). We then ran stepwise modal forward selection absed
on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}) values and
identified the most parsimonious movement model.

\subsection{Dispersal Simulation}
We used the most parsimonious movement model to simulate dispersing wild dogs
departing from predefined source points. The simulation basically resembled an
inverted iSSF function and was set up as follows. First, we determined a source
point at which a disperser was initiated. To allow calculation of turning
angles, we assumed a random initial orientation of the animal. Second, we
proposed 25 random steps originating at the predefined source point. We
generated random steps by sampling turning angles from a uniform distribution
(\(-\pi, +\pi\)) and step lengths from a gamma distribution fitted to observed
step lengths. To prevent unrealistically large steps, we capped the gamma
distribution 35km, the farthest distance travelled in four hours according to
our data. Third, along each random step we extracted environmental covariates
calculated step metrics. Fourth, we applied the parametrized dispersal model to
predict selection scores \(w(x)\) and to determine the probability of a step
being realized. Fifth, we sampled one of the proposed based on their
probabilities and calculated the animal's new position. We then repeated steps
two to five until 2000 steps were realized. In case a proposed random step left
our study extent, we removed the step from the set of random steps, thereby
forcing the disperser to stay within the boundaries of the study area.

\subsection{Source Points}
Simulations from step selection functions are known to be sensitive to the
location of source points, which is why we followed a twofold approach to sample
release points across the KAZA-TFCA. In a first approach, we used the same 68
source points between which we already computed least-cost paths and least-cost
corridors (see xx). These source points were regularly spaced 100 km apart,
located within protected areas (\(>\) 700 km\textsuperscript{2}). Because of
the regulary spacing, not all protected areas received source points. Hence, we
placed additional source points at the center of such protected areas. At each
of the generated source point we released 1000 dispersers, implying a total of
68'000 simulated dispersers. In a second approach, we used the same 68 source
points to define catchment areas inside protected areas. That is, for each of
the 68 source points we defined its voronoi polygon within the point's protected
area. Within these catchment areas we then randomly sampled 1000 new source
points from which we released another 68'000 dispersers. Overall, we simulated
134'000 dispersers for a total of 268 Mio. steps.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_SourcePoints}
    \caption{Illustration of source points from which dispersal was simulated.
    Although in reality we simulated 1'000 per source point, we illustrate an
    example assuming 10 dispersers from each source point. (a) Static source
    points similar to the source points reported in ... (b) Source points that
    were randomized within the catchment areas (dark gray, delinated by solid
    black lines).}
    \label{SourcePoints}
  \end{center}
\end{figure}

\subsection{Heatmaps}
Using the simulated trajectories we created heatmaps to examine through which
areas most simulated individuals dispersed. We rasterized all trajectories and
calculated how often each raster-cell was traversed by simulated dispersers. If
the same trajectory crossed a pixel twice, it was only counted once. We achieved
high performance rasterization of spatial lines using the recently developed
R-package \textit{terra} \citep{Hijmans.2020}. To examine if and how ``heat''
changes in response to changes in the location of source points and the number
of simulated steps, we followed a 2 x 6 design and created heatmaps for both
point sampling regines, as well as for 68, 125, 250, 500, 1000, and 2000
dispersal steps. We quantified the similarity of the resulting 12 heatmaps to
the permeability and least-cost corridor maps presented in (Hofmann ...) we used
Bhattacharyya's affinity. Bhattacharyya's affinity ranges from zero (complete
separation) to one (perfect match) and has earlier been proposed to compare the
overlap of utilisation distributions (Fieberg).

\subsection{Network Analysis I: Areas Reached}
Based on the simulated trajectories we identified to which other areas each
source area is connected. For instance, if a trajectory originated at source
point one and intersected with source areas two and three, we assumed that
source areas two and three were within reach of source are one. To get a sense
of the strength of connections, we also calculated how often each of these
connections was realized. This procedure resulted in 6 x 2 edge lists which we
further used to generate weighted networks and to calculate network metrics.

\subsection{Network Analysis II: Betweenness}
We coerced all simulated trajectories into a network consisting of vertices
(relocations) and edges (connections between relocations). To do so, we created
raster layers at multiple resolutions and identified each trajectory's
transition matrix on these rasters see figure xx \citep{Bastille.2018}. We then
merged the transition matrices of all trajectories and calculated cumulative
transitions between all raster-cells. This resulted in an edge-list, containing
all observed from-to connections as well as their frequency. Using this
edge-list, we generated a weighted graph using the r-package \textit{igraph}.
Based on this graph we calculated betweenness scores as well as the degree of
each raster cell. Betweenness indicates how often a specific raster-cell lies on
a shortest path between two other raster-cells and is a useful metric to detect
movement corridors. Degree, on the other hand, indicates how many connections a
raster cell has and therefore serves to illustrate the xxx of specific nodes.
When calculating betweenness, we used the transition frequency as weighting
factor. That is, a higher transition frequency contributed to a higher
betweenness score.

\subsection{Network Analysis III: Transitions}

\section{Results}
Compared to the base model, the most parsimonious movement model included
several additional interactions (\Cref{MovementModel} and Table S1). The model
indicates that dispersers move directional, particularly when distant to water,
yet less so in human dominated landscapes. Furthermore, dispersers prefer large
steps, especially during main activity. In contrast, step lengths tend to be
shorter when water- or tree-cover is high. In general, dispersers avoid water,
prefer proximity to water, avoid dense tree-cover, prefer shrubs/grassland, and,
finally, avoid human dominated landscapes.

\begin{figure}
  \begin{center}
    \begin{subfigure}[c]{0.8\textwidth}
      \includegraphics[width=\textwidth]{99_MovementModel}
      % \subcaption{Subfigure Bild Nr. 1}
    \end{subfigure}
    % \begin{subfigure}[c]{\textwidth}
    %   \includegraphics[width=\textwidth]{99_MovementModel(Interactions)}
    %   % \subcaption{Subfigure Bild Nr. 2}
    % \end{subfigure}
    \caption{Most parsimonious movement model. Whiskers delineate the 95\%
    Confidence-Intervals. Significance codes: * \(p < 0.10\), ** \(p < 0.05\),
    *** \(p < 0.01\).}
    \label{MovementModel}
  \end{center}
\end{figure}

\subsection{Heatmaps}
Six of the twelve rasterized dispersal trajectories are presented as heatmaps in
\Cref{Heatmaps}. As can be seen, differences that stem from the method of point
sampling disappear as more steps are simulated. This is to be expected as the
influence of the origin becomes smaller and smaller as the animal moves through
the landscape. However, there are striking differences when simualtions are only
run for few iterations.

\begin{figure}
  \includegraphics[width=\textwidth]{99_Heatmaps}
  \caption{}
  \label{Heatmaps}
\end{figure}

Bhattacharyya's affinity index supports the notion that heatmaps become more
similar to the earlier developed permeability and corridor maps as the number of
simulated steps increases. Furthermore, it appears that randomly sampled source
points contribute to a higher similarity too, albeit differences due to the
sampling regime vanish as the number of simulated steps increases. In fact, both
maps are almost identical to each other after 2000 steps. Furthermore, the
connectivity networks become increasingly similar to the previously published
permeability and corridor maps. Still, some differences remain even after 2000
simulated steps, highlighting that some severe impediments in the landscape
exist.

\begin{table}[hbtp]
  \caption{Summary statistics of all GPS relocations that have been recorded on
  dispersing coalitions.}
  \label{GPSData}
  \begin{center}
  \resizebox{\textwidth}{!}{
    \begin{tabular}{lllllllll}
    \toprule
    sampling & Metric & Map & 68 & 125 & 250 & 500 & 1000 & 2000 \\
    \midrule
    Static & Affinity & Corr & 0.64 & 0.70 & 0.76 & 0.80 & 0.82 & 0.84 \\
    Static & Affinity & Perm & 0.67 & 0.74 & 0.80 & 0.85 & 0.89 & 0.91 \\
    Static & Correlation & Corr & 0.32 & 0.38 & 0.45 & 0.51 & 0.56 & 0.60 \\
    Static & Correlation & Perm & 0.47 & 0.57 & 0.66 & 0.72 & 0.78 & 0.82 \\
    \hdashline
    Random & Affinity & Corr & 0.70 & 0.75 & 0.79 & 0.82 & 0.84 & 0.85 \\
    Random & Affinity & Perm & 0.74 & 0.79 & 0.83 & 0.87 & 0.90 & 0.92 \\
    Random & Correlation & Corr & 0.38 & 0.43 & 0.48 & 0.53 & 0.57 & 0.61 \\
    Random & Correlation & Perm & 0.58 & 0.64 & 0.69 & 0.75 & 0.79 & 0.83 \\
    \hdashline
    n.a. & Affinity & Heat & 0.76 & 0.86 & 0.94 & 0.97 & 0.99 & 0.99 \\
    n.a. & Correlation & Heat & 0.90 & 0.94 & 0.97 & 0.98 & 0.99 & 1.00 \\
    \bottomrule
    \end{tabular}
  }
  \end{center}
\end{table}

\section{Discussion}
Our connectivity network further suggests that dispersers from the Okavango
Delta more likely disperse towards east than west. Indeed, only x out of our y
observed dispersers ever reached the western part of the delta. Only when the
flood retracts a small pathway between the city of Maun and the floodwaters of
the delta emerges and enables dispersers to move towards the detal's western
part.

Our work suggests that the selection of source points significantly impacts
resulting connectivity networks. Especially when dispersal durations are short,
wrongly placed source points lead to vastly different results. Signer et al.
used estimated utilisation distributions by means of simulated movements. They
used a rather long burn in period prior to alleviate the problem of selecting
meaningful source points. However, this approach only works when individuals
move around a point of attraciton. This is typically not the case when
simulating dispersers, introducing an important trade-off. The researcher can
decide to increase the number of simulated steps, hence reducing the influence
of starting locations, yet this also inevitably increases estimated
connectivity.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modelling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing coalitions will be made available on dryad at
the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Johannes Signer for assisting with the simulation algorithm. This study was
funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung
Temperatio, Wilderness Wildlife Trust Foundation, Forschungkredit der
Universität Zürich, and a Swiss National Science Foundation Grant
(31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}

### 2020-08-21 Chapter_1/04_Manuscript/Manuscript_Support.tex
first commit

--- 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Support.tex
@@
 \usepackage[nameinlink]{cleveref}
-\usepackage{lineno}
+% \usepackage{lineno}
 \usepackage{multirow}
 \usepackage{arydshln}
+% \usepackage[nomarkers, nolists]{endfloat}
 \usepackage[flushleft]{threeparttable}
@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Bound within Boundaries: How Well Do Protected Areas
-Match Movement Corridors of Their Most Mobile Protected Species?}
+\title{\textbf{Appendix}\\ Step by Step: Using Integrated Step Selection
+Analysis to Simulate Wild Dog Dispersal and Assess Landscape Connectivity}
 
@@
 \author{
-  David D. Hofmann\textsuperscript{1,\S,*} \and
-  Dominik M. Behr\textsuperscript{1,2,*} \and
+  David D. Hofmann\textsuperscript{1,\S} \and
   John W. McNutt\textsuperscript{2} \and
   Arpat Ozgul\textsuperscript{1} \and
-  Gabriele Cozzi\textsuperscript{1,2}
+  Gabriele Cozzi\textsuperscript{1,2} \and
+  Dominik M. Behr\textsuperscript{1,2}
 }
@@
 
-\textsuperscript{*} Shared first authorship
-
 \vspace{4cm}
 
-\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.
+\textbf{Running Title:} Simulating Wild Dog Dispersal.
 
@@
 
-% Create linenumbers
-\linenumbers
+% % Create linenumbers
+% \linenumbers
 
@@
 %------------------------------------------------------------------------------
-%	Appendix S1: Net Squared Displacement
+%	Appendix S1: All Movement Models
 %------------------------------------------------------------------------------
-\section{Net Squared Displacement}
-\begin{figure}[hbpt]
+\section{Movement Models}
+% latex table generated in R 3.6.3 by xtable 1.8-4 package
+% Thu Jul  9 09:17:26 2020
+\begin{table}[hbpt]
+  \caption{Results from the forward model selection procedure based on Akaike's
+  Information Criterion (AIC; \citealp{Burnham.2002}) for the movement model.
+  The base model upon which we based our movement model is depicted in the last
+  row. We omitted any other model with an AIC weight of zero.}
+  \label{ModelAICs}
   \begin{center}
-    \includegraphics[width = \textwidth]{99_NSD.pdf}
-    \caption{NSD displacement through time for one of our dispersers. The blue
-    line indicates the period during which we calssified the individual as
-    dispersing.}
-    \label{NSD}
-  \end{center}
-\end{figure}
-
-%------------------------------------------------------------------------------
-%	Appendix S2: GPS Data
-%------------------------------------------------------------------------------
-\newpage
-\section{GPS Data}
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = 0.86\textwidth]{99_Trajectories.pdf}
-    \caption{Illustration of all trajectories that we recorded. Each color
-    represents a different dispersing coalition. All coalitions departed from
-    the area which is indicated by the white dashed rectangle. The coalition
-    dispersing towards the far east of the map covered over 360 km in under 10
-    days. Satellite background imagery was provided by Microsoft Bing.}
-    \label{Trajectories}
-  \end{center}
-\end{figure}
-
-\begin{table}[hbtp]
-  \caption{Summary statistics of all GPS relocations that have been recorded on
-  dispersing coalitions.}
-  \label{GPSData}
-  \begin{center}
-    \resizebox{0.8\textwidth}{!}{
-      \begin{tabular}{lccccccc}
+    \resizebox{\textwidth}{!}{
+      \begin{threeparttable}
+        \begin{tabular}{llllll}
         \toprule
-      \parbox[]{2cm}{\centering Coalition ID} &
-        \parbox[]{1cm}{\centering Sex} &
-          \parbox[]{2cm}{\centering \ Pack \\ Affiliation} &
-            \parbox[]{1.5cm}{\centering \# Fixes \\ Total} &
-              \parbox[]{1.5cm}{\centering \# Fixes \\ During \\ Dispersal} &
-                \parbox[]{1.5cm}{\centering \# Days \\ Dispersing} &
-                  \parbox[]{2cm}{\centering Euclidean \\ Dispersal\\ Distance \\ (in km)} &
-                    \parbox[]{2cm}{\centering Cumulative \\ Dispersal \\ Distance \\ (in km)} \\
+          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik & Model ID \\
           \midrule
-      Abel & M & MU & 894 & 45 & 9 & 131 & 205 \\
-      Amacuro & F & MB & 954 & 583 & 137 & 23 & 1'090 \\
-      Belgium & M & ZU & 1'097 & 158 & 28 & 18 & 319 \\
-      Dalwhinnie & F & PA & 545 & 62 & 22 & 50 & 243 \\
-      Denali & F & MN & 1'096 & 173 & 33 & 11 & 528 \\
-      Everest & M & MN & 389 & 123 & 38 & 67 & 572 \\
-      Kalahari & F & HT & 1'753 & 467 & 130 & 20 & 1'963 \\
-      Karisimbi & M & MN & 438 & 141 & 34 & 45 & 251 \\
-      Liuwa & F & AP & 946 & 92 & 19 & 144 & 451 \\
-      Lupe & M & KW & 2'209 & 396 & 34 & 8 & 436 \\
-      MadameChing & F & AP & 776 & 729 & 136 & 263 & 1'560 \\
-      Mirage & M & HT & 814 & 182 & 36 & 7 & 435 \\
-      Odzala & M & AP & 1'410 & 205 & 42 & 53 & 412 \\
-      Scorpion & M & KB & 2'676 & 393 & 34 & 4 & 471 \\
-      Stetson & M & MT & 384 & 383 & 33 & 3 & 481 \\
-      Taryn & F & AP & 896 & 37 & 9 & 10 & 130 \\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 & 1\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 & 2\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 & 3\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 & 4\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 & 5\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 & 6\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 & 7\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 & 8\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 & 9\\
+          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 & 10\\
+          \hdashline
+          ...\\
+          \hdashline
+          Base Model: cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 787.67 & 0.00 & -45030.70 & NA\\
          \bottomrule
        \end{tabular}
-    }
-  \end{center}
-\end{table}
-
-%------------------------------------------------------------------------------
-%	Appendix S3: Spatial Covariates
-%------------------------------------------------------------------------------
-\newpage
-\section{Spatial Covariates}
-To investigate habitat preferences of dispersing wild dogs, we used a set of
-geo-referenced covariates that we aggregated in the categories \textit{land
-cover}, \textit{protection status}, and \textit{human influence}. We did not
-include terrain features as covariates due to the absence of noteworthy
-elevational gradients in our study area. For each covariate, we prepared spatial
-raster layers from freely available online services or from remotely sensed
-satellite imagery. To ensure a consistent resolution (i.e. cell-size or grain)
-across covariates, we coarsened or interpolated all layers to match a resolution
-of 250m x 250m. We performed processing and manipulation of data as well as all
-spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.
-
-\subsection{Land Cover}
-\subsubsection{Water}
-The covariate water included rivers, wetlands, and swamps. Because the
-inundation extent of the flood in the Okavango Delta is highly variable within
-and between years, we created dynamic ``flood maps'' that were updated every
-8\textsuperscript{th} day following a remote sensing algorithm developed by the
-Okavango Research Institute (ORI; \citealp{Wolski.2017}). To implement the
-algorithm, we defined two sets of polygons located in the region of the Okavango
-Delta (\Cref{Floodmapping}). The first set consisted of areas known to be
-permanent dryland, whereas the second set consisted of permanent waters. Since
-we were unable to retrieve the original polygons used in \cite{Wolski.2017}, we
-geo-referenced and digitized the polygons reported in their publication. After
-recreating the polygons, we used the R-package \textit{getSpatialData}
-\citep{Schwalb.2018} to download and pre-process all relatively cloud-free MODIS
-Terra images (MCD43A4; \citealp{Schaaf.2015}) available for the period of our
-dispersal events. Assessment of cloud cover was based on visual inspection of
-MODIS images on ORI's website
-(\url{www.okavangodata.ub.bw/ori/monitoring/flood_maps}). After download, we
-classified each MODIS image into a binary map of water (flood) and dryland using
-a threshold that was identified as follows. First, we extracted all reflectance
-values of MODIS Terra Band 7 within the water- and dryland-polygons. Second, we
-computed histograms of water-reflectances and dryland-reflectances and
-empirically verified that reflectances of the two groups were sufficiently
-distinct. More specifically, we checked if superimposing the histograms of
-water-reflectances and dryland-reflectances resulted in a bimodal histogram.
-This was said to be achieved if the 99\textsuperscript{th} percentile of
-water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
-of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
-dryland}\)). Third, if bimodality was achieved, we calculated a threshold
-(\(t\)) using \ref{EQ1}:
-
-\begin{equation}
-\label{EQ1}
-t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
-\widetilde{p}_{water})
-\end{equation}
-
-\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
-the median reflectances of water and dryland, respectively. We then classified
-all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
-all pixels with a value smaller than \(t\) as water.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
-    \caption{Images describing the flood mapping algorithm. (a) The colored
-    polygons indicate permanent waters (blue) and permanent dryland (green).
-    Below these polygons we extracted reflectance values of MODIS Terra Band 7
-    and used their repsective medians to calculate a classification threshold
-    \(t\). (b) Example of a classified MODIS Terra Band 7 image after
-    application of the threshold. The satellite image in the background was
-    provided by Microsoft Bing.}
-    \label{Floodmapping}
-  \end{center}
-\end{figure}
-
-% Dynamic Wetmask
-\noindent Importantly, bimodality was not always achieved and in some cases no
-flood map could be calculated. In fact, it appears that non-bimodality caused
-the ORI algorithm to fail since the end of 2018, which is why no flood maps have
-been generated since then (ORI, personal comm.). We hypothesized that this was
-caused by the application of static water-polygons that did not cover permanent
-waters correctly anymore. Therefore, we revised the algorithm and allowed for a
-more dynamic polygonization of water. That is, for each MODIS image we
-calculated new water-polygons comprising areas that were covered by the flood in
-99\% of the flood maps from the previous five years. All of the necessary flood
-maps from previous years were kindly provided to us by ORI. Using this slightly
-amended approach, we were able to address some of the bimodality issues and to
-classify several additional flood maps for the period of our study. Because
-MODIS Terra Band 7 had a resolution of 500m x 500m, we interpolated all maps to
-250m x 250m.
-
-% Validation
-To validate and compare the performance of our own algorithm to the original
-ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
-images. To make sure that months were equally represented in the sampled dates,
-we employed stratified sampling based on months (regardless of the year) and
-randomly sampled four maps for each month. For the sampled dates we downloaded
-and classified MODIS Terra Band 7 images and compared our classified images to
-those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
-created a difference map indicating false positives and false negatives and
-computed the relative number of wrongly classified pixels. We achieved an
-overall accuracy of 97\%, which presumably is an underestimate of the true
-performance, as we introduced some errors when resampling the ORI-maps to our
-reference grid.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
-    \caption{Validation procedure of our flood mapping algorithm. (a) Classified
-    image that was provided to us by ORI. (b) Image for the same date but now
-    classified using our own algorithm. (c) Difference image indicating false
-    positives and false negatives in our own classification.}
-    \label{FloodmappingValidation}
-  \end{center}
-\end{figure}
-
-% Static Watermap
-\noindent While we created dynamic flood maps for the Okavango Delta, we assumed
-the extent of all other water bodies (e.g. Chobe river, Zambezi river) to be
-static within and between years. This static representation was based on
-Globeland's land cover dataset \citep{Chen.2015}, from which we only retained
-the categories \textit{wetland} and \textit{water bodies} and collectively
-reclassified them to \textit{water}. Globeland had an original resolution of 30m
-x 30m, so we coarsened the layer to 250m x 250m using the mode of each 250m x
-250m cell. We further improved river representation by employing the rasterized
-MERIT Hydro dataset \citep{Yamazaki.2019} from which we added all rivers with a
-width of over 10m to our Globeland layer. We merged dynamic and static water
-maps into a large rasterstack, covering the entire study area. We also created a
-rasterstack rendering the covariate \textit{distance to water} by calculating
-the Euclidean distance of each raster cell in the study area to the nearest
-source of water.
-
-\subsubsection{Dryland}
-We subdivided dryland into three layers as derived from the MODIS Terra
-Vegetation Continuous Fields dataset (MOD44B; \citealp{Dimiceli.2015}). The
-three layers depicted percentage cover of tree-vegetation (henceforth
-\textit{trees}), non-tree-vegetation (henceforth \textit{shrubs/grassland}), and
-non-vegetated (henceforth \textit{bare land}) and added up to 100\% of dryland
-coverage. We used our flood map that aligned with the creation date of these
-MODIS layers and defined anything covered by water as 0\% vegetated. The MODIS
-vegetation layers had a resolution of 250m x 250m and no coarsening or
-interpolation was required.
-
-\subsection{Protection Status}
-We created a binary layer separating protected from unprotected land. We
-downloaded corresponding data on protection status in shapefile format from the
-Peace Parks Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}).
-Protected areas included forest reserves, game reserves, wildlife management
-areas, and national parks. We classified anything not covered by these
-categories as unprotected (e.g. communal pastoral land, private land). We
-rasterized the two categories to the binary raster \textit{protection status} (1
-= protected, 0 = unprotected) with a resolution of 250m x 250m.
-
-\subsection{Human Influence}
-We created a raster layer representing human influence by integrating
-information on (1) human density, (2) farming, and (3) roads.
-
-\begin{enumerate}[label = (\arabic*)]
-
-  \item We obtained spatial human density estimates through a publicly available
-  30m x 30m high-resolution population density dataset
-  (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}). We coarsened the
-  layer to 250m x 250m by summing up human density values within each 250m x
-  250m cell.
-
-  \item We sourced spatial information on farms from the Globeland
-  \citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets from
-  which we retained areas that were classified as either \textit{cultivated
-  land} or \textit{croplands}. Any other land cover class was not pertinent to
-  farming and therefore omitted. Because both layers had a resolution of 30m x
-  30m we coarsened them to 250m x 250m by assigning a value of 1 to any 250m x
-  250m cell that covered farmland and a value 0 otherwise. Thus, the final layer
-  depicted presence (= 1) or absence (= 0) of farms within each 250m x 250m
-  cell.
-
-  \item We obtained geo-referenced data on roads from Open Street Map
-  \citep{OpenStreetMap.2019}, downloaded through Geofabrik
-  (\url{www.geofabrik.de}). We only retained main tarmac roads and omitted
-  smaller roads (Table S2) as these are scarcely frequented and do not represent
-  an obstacle to wild dog movements \citep{Abrahms.2016}. We rasterized main
-  tarmac roads to the binary raster \textit{roads} (1 = roads, 0 = no roads)
-  with 250m x 250m resolution. Finally, we created the covariate
-  \textit{distance to roads} by calculating the Euclidean distance of each
-  raster cell in the study area to the nearest road.
-
-\end{enumerate}
-
-\begin{table}[hbtp]
-  \caption{Description of road types, as sourced from Open Street Map's mapping
-  guide (\url{https://wiki.openstreetmap.org/wiki/Key:highway}). Roads types
-  that were considered for the purpose of this study are shaded in light gray.}
-  \label{RoadsDescription}
-  \begin{center}
-    \resizebox{\textwidth}{!}{
-      \begin{tabular}{llp{12cm}}
-      \toprule
-      Group &
-        Subgroup &
-          Description \\
-      \midrule
-      \rowcolor[gray]{0.90}
-      Roads &
-        motorway &
-          A restricted access major divided highway, normally with 2 or more
-          running lanes plus emergency hard shoulder. Equivalent to the
-          Freeway, Autobahn, etc. \\
-      \rowcolor[gray]{0.90}
-      Roads &
-        trunk &
-          The most important roads in a country's system that aren't
-          motorways. Need not necessarily be a divided highway. \\
-      \rowcolor[gray]{0.90}
-      Roads &
-        primary &
-          The next most important roads in a country's system. Often link
-          larger towns. \\
-      \rowcolor[gray]{0.90}
-      Roads &
-        secondary &
-          The next most important roads in a country's system. Often link
-          towns. \\
-      Roads &
-        tertiary &
-          The next most important roads in a country's system. Often link
-          smaller towns and villages \\
-      Roads &
-        unclassified &
-          The least important thorough roads in a country's system, i.e. minor
-          roads of a lower classification than tertiary, but which serve a
-          purpose other than access to properties. Often link villages and
-          hamlets. \\
-      Roads &
-        residential &
-          Roads which serve as an access to housing, without function of
-          connecting settlements. Often lined with housing. \\
-      Roads &
-        service &
-          For access roads to, or within an industrial estate, camp site,
-          business park, car park etc. \\
-      \rowcolor[gray]{0.90}
-      Link roads &
-        motorway\_link &
-          The link roads (sliproads/ramps) leading to/from a motorway from/to
-          a motorway or lower class highway. Normally with the same motorway
-          restrictions. \\
-      \rowcolor[gray]{0.90}
-      Link roads &
-        trunk\_link &
-          The link roads (sliproads/ramps) leading to/from a trunk road
-          from/to a trunk road or lower class highway. \\
-      \rowcolor[gray]{0.90}
-      Link roads &
-        primary\_link &
-          The link roads (sliproads/ramps) leading to/from a primary road
-          from/to a primary road or lower class highway. \\
-      \rowcolor[gray]{0.90}
-      Link roads &
-        secondary\_link &
-          The link roads (sliproads/ramps) leading to/from a secondary road
-          from/to a secondary road or lower class highway. \\
-      Link roads &
-        tertiary\_link &
-          The link roads (sliproads/ramps) leading to/from a tertiary road
-          from/to a tertiary road or lower class highway. \\
-      Special road types &
-        living\_street &
-          For living streets, which are residential streets where pedestrians
-          have legal priority over cars, speeds are kept very low and where
-          children are allowed to play on the street. \\
-      Special road types &
-        pedestrian &
-          For roads used mainly/exclusively for pedestrians in shopping and
-          some residential areas which may allow access by motorised vehicles
-          only for very limited periods of the day. \\
-      Special road types &
-        track &
-          Roads for mostly agricultural or forestry uses. \\
-      Special road types &
-        bus\_guideway &
-          A busway where the vehicle is guided by the way (though not a railway)
-          and is not suitable for other traffic. \\
-      Special road types &
-        escape &
-          For runaway truck ramps, runaway truck lanes, emergency escape
-          ramps, or truck arrester beds. It enables vehicles with braking
-          failure to safely stop. \\
-      Special road types &
-        raceway &
-          A course or track for racing \\
-      Special road types &
-        road &
-          A road/way/street/motorway/etc. of unknown type. It can stand for
-          anything ranging from a footpath to a motorway. \\
-      \bottomrule
-      \end{tabular}
-    }
-  \end{center}
-\end{table}
-
-\noindent Because layers (1), (2), and (3) depicted features that are typically
-spatially clustered and because not all dispersing coalitions moved within
-meaningful distance to each of these features, we totaled values from the layers
-describing \textit{human density} (continuous), \textit{farming} (binary), and
-\textit{roads} (binary). This approach implied that \textit{roads} and
-\textit{farms} entered the final layer with a value of 1, whereas human density
-entered the final layer with a value \(\geq 0\) and potentially unbound. To
-reduce the influence of outliers in human density estimates, totaled values were
-limited to a maximum of 50, which visually resulted in a good balance between
-high and low anthropogenic influence and was therefore considered appropriate
-for our analysis. To render the fact that humans influence their surroundings
-beyond their presence, we followed \cite{Elliot.2014} and applied to each
-raster-cell a 5km focal buffer within which we summed up and log-transformed
-human-influence values.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = 0.95\textwidth]{99_HumanInfluence.pdf}
-    \caption{Sequence of figures that exemplifies how we combined the layers
-    (a), (b), and (c) into a single layer for human influence (d). For better
-    visibility we show the procedure only for the extent of the Okavango Delta.
-    The layer in (a) is based on Facebook's high resolution human density
-    dataset (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}) and depicts
-    the estimated number of humans living in each 250m x 250m raster-cell
-    (coarsened from 30m x 30m). The layer in (b) is a binary layer and shows
-    whether raster-cells are cover any sort of agricultural fields.
-    Corresponding data was obtained through the Globeland and Cropland land
-    cover datasets \citep{Chen.2015, Xiong.2017}. The layer in (c) shows the
-    presence or absence of roads and is based on data from Open Street Map
-    \citep{OpenStreetMap.2019}. We merged the layers in (a), (b), and (c) by
-    summing up their values, truncating the summed values to a maximum of 50. We
-    then log-transforming the values and applied to each raster cell a focal
-    buffer of 5km within which we totaled human influence values. The layer in
-    (d) depicts the final human influence layer that entered our habitat
-    selection model.}
-    \label{HumanInfluence}
-  \end{center}
-\end{figure}
-
-%------------------------------------------------------------------------------
-%  Appendix S4: Flood Pulse
-%------------------------------------------------------------------------------
-\newpage
-\section{Typical Flood Pulse}
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
-    \caption{Sequence of flood maps showing a typical flood pulse throughout the
-    year. The flood arrives from the north-western corner (so called
-    ``pan-handle'') of the Okavango Delta and slowly descends through the delta
-    in south-eastern direction, where it nourishes several tributaries. The
-    extent of the flood peaks around August or September and then slowly
-    retracts. Between December and March the reflectance properties of water and
-    dryland change, which is why often no accurate flood maps can be obtained
-    for these months using remote sensing techniques \citep{Wolski.2017}.}
-    \label{FloodPulse}
-  \end{center}
-\end{figure}
-
-%------------------------------------------------------------------------------
-%  Appendix S5: Step Selection Analysis
-%------------------------------------------------------------------------------
-\newpage
-\section{Integrated Step Selection Function}
-We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
-investigate dispersers' selection or avoidance of spatial covariates. In the
-iSSF framework, covariates experienced along realized steps are contrasted with
-covariates experienced along alternative random steps that the animal could have
-taken but decided not to. A step in this framework is defined as the connecting
-line between two consecutive GPS relocations \citep{Turchin.1998}. In contrast
-to regular SSFs, iSSFs require to include movement metrics as covariates in the
-corresponding conditional logistic regression model. Their inclusion, in turn,
-allows simultaneous inference on habitat and movement preferences, as well as to
-reduce potential biases in estimated habitat preferences \citep{Forester.2009,
-Warton.2013, Avgar.2016}.
-
-To conduct iSSF analysis, we followed the recommendations described in Appendix
-S1 of the publication by \cite{Avgar.2016}. We prepared our GPS relocation data
-for iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
-relocations recorded during dispersal to steps that were regularly spaced four
-hours apart. Steps that were separated by more than four hours (e.g. due to GPS
-failure) were omitted from further analysis (allowing for a minor mismatch of up
-to 15 minutes). Each remaining step was paired with 24 random steps, generated
-by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
-lengths from a gamma distribution that was fitted using realized step lengths
-(shape = 0.3677, scale = 6'302). Together, a realized and its 24 associated
-random steps formed a stratum of 25 steps that received a unique identifier.
-
-We extracted spatial covariates along realized and random steps
-(\Cref{Appendix:Sources}). For continuous covariates, we calculated the average
-value, for categorical covariates the percentage cover along the step. We
-further derived a binary variable indicating whether a step crossed a road. We
-square-rooted extracted values to render a decreasing marginal impact of
-distance. We scaled covariates using a z-score transformation and screened for
-correlation using Pearson's Correlation Coefficient. None of the covariates were
-overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we retained all of
-them for modeling. Despite the covariates mentioned in \Cref{Appendix:Sources},
-we included two movement metrics, namely the cosine of the turning angle
-(\(cos(ta)\)) and the logarithm of the step length (\(log(sl)\)) in our
-regression model \citep{Avgar.2016}. The movement metric \(cos(ta)\) serves to
-describe the directionality of a step, as it transforms the circular measure of
-(\(-\pi\) to \(\pi\)) into a linear measure (-1, 1). Thus, positive values
-indicate forward movements, whereas negative values indicated backward movements
-\citep{Turchin.1998}. The movement metric \(log(sl)\), on the other hand, is as
-an indicator of the preferred step length. Since in our case steps were spaced
-by four hours, \(log(sl)\) can also be interpreted as movement rate.
-
-\begin{table}[hbtp]
-  \begin{center}
-    \caption{Overview of spatial covariates and their sources. We extracted
-    covariates along realized and random steps. For continuous covariates we
-    calculated average values along steps, for categorical covariates the
-    percentage coverage along steps. We also prepared two covariates indicating
-    the distance to water and distance to roads, respectively. We square-rooted
-    the values for these two covariates to render a decreasing marginal impact
-    of the effect of distance. Finally, we derived a binary indicator of whether
-    a step crossed a road or not.}
-    \label{Appendix:Sources}
-    \resizebox{\textwidth}{!} {
-      \begin{threeparttable}
-        \begin{tabular}{lllll}
-        \hline
-        Category &
-          Covariate &
-            Description &
-              Values &
-                Source \\
-        \midrule
-        \multirow{5}{*}{Land Cover}
-          & Water
-            & Percentage cover of water
-              & 0-100\%
-                & (1) (2) (3) \\
-          & Dryland*
-            & Percentage cover of dryland
-              & 0-100\%
-                & (1) (2) (3) \\
-          & DistanceToWater
-            & Average distance to nearest water source
-              & \(\geq\) 0m
-                & (1) (2) (3) \\
-          & Shrubs/Grassland
-            & Average non-tree vegetation
-              & 0-100\%
-                & (4) \\
-          & Trees
-            & Average tree-vegetation
-              & 0-100\%
-                & (4) \\
-          & Bareland*
-            & Average non-vegetated area
-              & 0-100\%
-                & (4) \\
-        \hdashline
-        \multirow{2}{*}{Protection Status}
-          & Protected
-            & Percentage cover of protected area
-              & 0-100\%
-                & (5) \\
-          & Unprotected*
-            & Percentage cover of unprotected area
-              & 0-100\%
-                & (5) \\
-        \hdashline
-        \multirow{3}{*}{Human Influence}
-          & Human Influence
-            & Average human influence
-              & \(\geq\) 0
-                & (1) (6) (7) (8)\\
-          & DistanceToRoads
-            & Average distance to nearest road
-              & \(\geq\) 0m
-                & (1) (6) (7) (8)\\
-          & RoadCrossing
-            & Binary; whether a step crossed a road
-              & 0, 1
-                & (1) (6) (7) (8)\\
-        \hline
-        \end{tabular}
        \begin{tablenotes}
-          \item \textit{Sources:} (1) \cite{Chen.2015} (2) \cite{Schaaf.2015} (3)
-          \cite{Yamazaki.2019} (4) \cite{Dimiceli.2015} (5) \cite{PeaceParks.2019}
-          (6) \cite{Facebook.2019} (7) \cite{OpenStreetMap.2019} (8)
-          \cite{Xiong.2017}
-          \item
-          \item \textit{* Note:} The covariates \textit{Water} and
-          \textit{Dryland} added up to 100\%, which is why only \textit{Water} was
-          included as explanatory variable in our models. The same applied for the
-          group \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland},
-          where we omitted \textit{Bareland} for modeling. Finally, from the group
-          \textit{Protected} and \textit{Unprotected}, we only included
-          \textit{Protected} in our models.
+         \item \textit{Note:} W = Water, DTW = Distance To Water, S =
+         Shrubs/Grassland, T = Trees, P = Protected, HI = Human Influence, RC =
+         Road Crossing, DTR = Distance To Roads, MA = Main Activity.
        \end{tablenotes}
@@
 
-\noindent We then used the iSSF framework to parameterize a habitat selection
-model that further served to predict landscape permeability. This habitat
-selection model operated under the assumption that dispersing wild dogs assigned
-a selection score \(w(x)\) of the following exponential form to each realized
-and random step:
-
-\begin{equation}
-\label{EQ2}
-  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
-\end{equation}
-
-\noindent That is, the selection score \(w(x)\) of a step depended on its
-associated covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's
-preferences for these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
-probability that a step \(i\) was realized \(P(Y_{i} = 1\)) was then contingent
-on the step's selection score, as well as on the selection scores of all
-alternative steps in the stratum:
-
-\begin{equation}
-\label{EQ3}
-  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
-  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
-\end{equation}
-
-\noindent Habitat and movement preferences of interest, i.e. the \(\beta\)'s,
-were then estimated by comparing realized (scored 1) and random (scored 0) steps
-in a conditional logistic regression model \citep{Fortin.2005}. In this model,
-positive \(\beta\)-coefficients indicate selection of a covariate, negative
-\(\beta\)-coefficients avoidance of a covariate. To deal with multiple
-individuals, we applied mixed effects conditional logistic regression analysis
-following \cite{Muff.2020}. We implemented their method using the R-package
-\textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID to model
-random intercepts and slopes.
-
-We defined the movement metrics \(cos(ta)\) and \(log(sl)\) as core covariates
-and ran forward model selection based on Akaike's Information Criterion (AIC;
-\citealp{Burnham.2002}) for all other covariates. We ranked models according to
-AIC, assessed relative model weights, and identified the most parsimonious
-model. Due to convergence issues, we were unable to model interactions between
-covariates.
-
-To validate the predictive power of the most parsimonious habitat selection
-model, we ran k-fold cross-validation for case-control studies as described in
-\cite{Fortin.2009}. Using 80\% of randomly selected strata, we parameterized a
-habitat selection model and predicted selection scores \(w(x)\) for all steps in
-the remaining 20\% of strata. According to predicted selection scores we
-assigned ranks 1-25 within each stratum, with rank 1 indicating the highest
-selection score. We identified the realized step's rank in each stratum and
-tallied rank frequencies of realized steps across all strata. Finally, we
-carried out a Spearman-rank correlation analysis between ranks and associated
-frequencies and we recorded the correlation coefficient (\(r_{s, realized}\)).
-We repeated this procedure 100 times with replacement and computed the mean
-correlation coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\%
-confidence interval. For comparison, we also repeated the same procedure 100
-times assuming completely randomized preferences. We implemented randomized
-preferences by omitting the realized step from each stratum and identifying the
-rank of a randomly chosen random step within each stratum (now only ranks 1-24).
-Again, we calculated Spearman's rank correlation coefficient (\(r_{s,
-random}\)), its mean across repetitions (\(\bar{r}_{s, random}\)), and its 95\%
-confidence interval. Ultimately, the validation proved a significant prediction
-in case the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-random}\) did not overlap.
-
-%------------------------------------------------------------------------------
-%  Appendix S6: Least Cost Paths and Least Cost Corridors
-%------------------------------------------------------------------------------
-\newpage
-\section{Identification of Least-Cost Paths \& Corridors}
-\subsection{Least-Cost Paths}
-We implemented factorial LCP analysis between source points using the R-package
-\textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). The package translated
-the (unscaled) permeability surface into a network of nodes to find shortest
-effective distances between source points based on probabilities of moving from
-cell to cell. In our case, the transition probability of moving between two
-adjacent cells depended on their averaged permeability. We allowed individuals
-to move from each cell to the cell's eight surrounding neighbors (i.e. Moores
-neighborhood) and applied a geographic correction to account for the fact that
-diagonal neighbors were more remote than orthogonal neighbors. Because African
-wild dogs have been observed to cover large dispersal distances
-\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}, we did not limit LCPs to a
-maximal effective cost. After computation, we tallied overlapping LCPs and
-identified high-frequency routes.
-
-\subsection{Least-Cost Corridors}
-We calculated factorial LCCs \citep{Pinto.2009, Sawyer.2011, Elliot.2014}, again
-using the R-package \textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). To
-identify LCCs, we first computed for each source point a cumulative cost map,
-which indicated the total minimal costs required to get from the source point to
-any other location in the study area. We then obtained an LCC between two source
-points by adding up their cumulative cost maps and masking out all cell-values
-exceeding the lowest cell-value by more than 5\% \citep{Pinto.2009}. We repeated
-this procedure for each possible unique pairwise combination of source points
-and thereby identified LCCs between all 68 selected source points. We normalized
-the resulting corridor-maps to range from zero to one and tallied them into a
-single connectivity map.
-
-\begin{figure}[hbtp]
+\begin{figure}[htpb]
   \begin{center}
-  \begin{minipage}{.32\linewidth}
-    \begin{subfigure}[t]{\linewidth}
-        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example1).pdf}
-    \end{subfigure}
-  \end{minipage}
-  \begin{minipage}{.32\linewidth}
-    \begin{subfigure}[t]{\linewidth}
-        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example2).pdf}
-    \end{subfigure}
-  \end{minipage}
-  \begin{minipage}{.32\linewidth}
-    \begin{subfigure}[t]{\linewidth}
-        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example3).pdf}
-    \end{subfigure}
-  \end{minipage}
-  \caption{Images illustrating the process of identifying a least-cost corridor
-  between source points A and B following \cite{Pinto.2009}. (a) Example of a
-  permeability surface, which determines the costs of movement. (b1) Cumulative
-  cost map for point A, depicting the total minimal costs necessary to get from
-  point A to every other location. (b2) Cumulative cost map for point B,
-  depicting the total minimal costs to get from point B to every other location.
-  (c) Summed cost maps of points A and B. (d) Masked out corridor containing
-  pixels that do not exceed the cheapest pixel by more than 5\%.}
-  \label{LCCExample}
+    \includegraphics[width = \textwidth]{99_AllMovementModels}
+    \caption{Coefficients of all movement models with an AIC weight above one.}
+    \label{AllMovementModels}
   \end{center}
 \end{figure}
-
-%------------------------------------------------------------------------------
-%  Appendix S7: AIC Model Selection
-%------------------------------------------------------------------------------
-\newpage
-\section{Model Selection Results}
-\begin{table}[hbpt]
-  \caption{Results from the forward model selection procedure based on Akaike's
-  Information Criterion (AIC; \citealp{Burnham.2002}) for the habitat selection
-  model. The most parsimonious model outperformed all other models (\(\Delta AIC
-  > \) 2) and received a weight of one.}
-  \label{ModelAICs}
-  \begin{center}
-    \resizebox{\textwidth}{!}{
-      \begin{threeparttable}
-        \begin{tabular}{lllll}
-        \toprule
-          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik \\
-          \midrule
-          cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\
-          cos(ta) + log(sl) + W + T + DTW + HI & 90094.79 & 3.38 & 0.00 & -45034.39 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\
-          cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\
-          cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\
-          cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\
-          cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\
-          cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\
-          cos(ta) + log(sl) + W + T + HI & 90123.71 & 32.31 & 0.00 & -45050.85 \\
-          cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\
-          cos(ta) + log(sl) + W + T + S & 90129.36 & 37.96 & 0.00 & -45053.68 \\
-          cos(ta) + log(sl) + W + T + DTR & 90130.11 & 38.71 & 0.00 & -45054.05 \\
-          cos(ta) + log(sl) + W + T + P & 90131.75 & 40.35 & 0.00 & -45054.88 \\
-          cos(ta) + log(sl) + W + T + RC & 90133.53 & 42.13 & 0.00 & -45053.77 \\
-          cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\
-          cos(ta) + log(sl) + W + HI & 90149.30 & 57.90 & 0.00 & -45065.65 \\
-          cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\
-          cos(ta) + log(sl) + W & 90153.70 & 62.30 & 0.00 & -45069.85 \\
-          cos(ta) + log(sl) + W + DTR & 90156.32 & 64.92 & 0.00 & -45069.16 \\
-          cos(ta) + log(sl) + W + P & 90157.23 & 65.83 & 0.00 & -45069.62 \\
-          cos(ta) + log(sl) + W + RC & 90159.32 & 67.91 & 0.00 & -45068.66 \\
-          cos(ta) + log(sl) + S & 90161.07 & 69.67 & 0.00 & -45073.54 \\
-          cos(ta) + log(sl) + DTW & 90262.03 & 170.63 & 0.00 & -45124.01 \\
-          cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\
-          cos(ta) + log(sl) + HI & 90307.55 & 216.15 & 0.00 & -45146.78 \\
-          cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\
-          cos(ta) + log(sl) + P & 90316.62 & 225.22 & 0.00 & -45151.31 \\
-          cos(ta) + log(sl) + RC & 90318.46 & 227.06 & 0.00 & -45150.23 \\
-         \bottomrule
-       \end{tabular}
-       \begin{tablenotes}
-         \item \textit{Note:} W = Water, DTW = Distance To Water, S =
-         Shrubs/Grassland, T = Trees, P = Protected, HI = Human Influence, RC =
-         Road Crossing, DTR = Distance To Roads.
-       \end{tablenotes}
-     \end{threeparttable}
-    }
-  \end{center}
-\end{table}
 

### 2020-08-24 Chapter_0/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript.r
Various updates

@@ -5,7 +5,7 @@
 rm(list = ls())
 
 # Set the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/00_WildDogs"
+wd <- "/home/david/Schreibtisch/15. PhD/Chapter_0"
 setwd(wd)
 
 # Set a seed
@@ -39,20 +39,7 @@ library(ggmap)
 library(mapview)
 library(knitr)
 library(kableExtra)
-
-# Load custom functions
-source("Functions.r")
-
-############################################################
-#### Useful Functions
-############################################################
-# Function to darken a color
-darken <- function(color, factor = 1.4){
-    col <- col2rgb(color)
-    col <- col / factor
-    col <- rgb(t(col), maxColorValue = 255)
-    col
-}
+library(davidoff)
 
 ############################################################
 #### General Statistics

### 2020-08-24 Chapter_1/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation.r

@@ -5,7 +5,7 @@
 rm(list = ls())
 
 # Set the working directory
-setwd("C:/Users/david/switchdrive/University/15. PhD/00_WildDogs")
+setwd("/home/david/Schreibtisch/15. PhD/Chapter_1")
 
 # Load required packages
 library(tidyverse)
@@ -132,7 +132,7 @@ p2 <- tm_shape(africa) +
 
 # Store the plot
 CairoPDF("Test", width = 5.25, height = 6)
-p1
+p2
 dev.off()
 
 ############################################################
@@ -141,7 +141,7 @@ dev.off()
 # Load required data
 africa2 <- "03_Data/02_CleanData/00_General_Africa.shp" %>% readOGR()
 kaza <- "03_Data/02_CleanData/00_General_KAZA_KAZA.shp" %>% readOGR()
-prot <- "03_Data/02_CleanData/02_LandUseTypes_Protected_PeaceParks(1Class).shp" %>% readOGR()
+prot <- "03_Data/02_CleanData/02_LandUseTypes_Protected_PeaceParks.shp" %>% readOGR()
 water <- "03_Data/02_CleanData/03_LandscapeFeatures_MajorWaters_GEOFABRIK.shp" %>% readOGR(.)
 
 # Clip africa layer
@@ -229,24 +229,16 @@ dev.off()
 # Prepare a map of Kaza
 p2 <- tm_shape(prot) +
     tm_polygons(
-      , col           = "gray30"
-      , border.col    = "black"
-      , alpha         = 0.5
-      , lwd           = 0.2
+      , col           = "gray40"
+      , border.col    = NA
+      , lwd           = 0
       , legend.show   = F
     ) +
-  tm_shape(water) +
-    tm_polygons(
-        col           = "cornflowerblue"
-      , border.col    = "cornflowerblue"
-      , border.alpha  = 0.6
-      , lwd           = 0.2
-    ) +
   tm_shape(kaza, is.master = T) +
-    tm_borders(
+    tm_polygons(
         col = "orange"
-      , lty = 1
-      , lwd = 2
+      , alpha = 0.5
+      , border.col = "orange"
     ) +
   tm_shape(africa) +
     tm_borders(

### 2020-08-31 Chapter_0/04_Manuscript/Manuscript.tex
Working on paper revision

@@
 \usepackage[nomarkers, nolists]{endfloat}
+\usepackage{todonotes}
 
@@
 example for international conservation efforts. We used integrated step
-selection analysis to estimate habitat preferences of dispersers and to create a
-permeability surface for the entire KAZA-TFCA. We compared landscape
-permeability across different regions within the KAZA-TFCA as well as outside
-its boundaries. Lastly, we calculated least-cost paths and corridors to verify
-that major movement routes were adequately encompassed within the KAZA-TFCA.
+selection analysis to estimate relative selection strenghts\todo{as per reviewer
+1} of dispersers and to create a permeability surface for the entire KAZA-TFCA.
+We compared landscape permeability across different regions within the KAZA-TFCA
+as well as outside its boundaries. Lastly, we calculated least-cost paths and
+corridors to verify that major movement routes were adequately encompassed
+within the KAZA-TFCA.
 
@@
 the need for a coordinated effort towards maintaining or restoring connectivity,
-especially where international effort is required.
+especially where transboundary dispersal occurs.\todo{According to suggestions
+by reviewer 2}
 \end{abstract}
@@
 thrive and persist \citep{Fahrig.2003}. Accordingly, preserving and protecting
-movement corridors between wildlife reserves has become an utmost task for
-conservation management \citep{Doerr.2011, Rudnick.2012}, resulting in an
-ever-growing number of large and often transboundary protected areas. While
-boundaries of such areas are often drawn according to expert opinion and
-socio-political needs, subjective assessments have revealed deficiencies in the
-past \citep{Clevenger.2002, Pullinger.2010}. Thus, an empirical assessment of
-the adequacy of already-existing or planned protected areas using pertinent
-animal movement data is paramount for targeted use of valuable and scarce
-conservation funds \citep{Pullinger.2010}.
+movement corridors between wildlife reserves has become a task of utmost
+\todo{Reviewer one pointed out an awkward use of the word utmost. I have changed
+it according to his suggestion.} importance for conservation management
+\citep{Doerr.2011, Rudnick.2012}, resulting in an ever-growing number of large
+and often transboundary protected areas. While boundaries of such areas are
+often drawn according to expert opinion and socio-political needs, subjective
+assessments have revealed deficiencies in the past \citep{Clevenger.2002,
+Pullinger.2010}. Thus, an empirical assessment of the adequacy of
+already-existing or planned protected areas using pertinent animal movement data
+is paramount for targeted use of valuable and scarce conservation funds
+\citep{Pullinger.2010}.
 
@@
 specific landscape \citep{Sawyer.2011}. Such surfaces are created based on a
-species' habitat preferences, which can be quantified using a suite of selection
-functions \citep{Zeller.2012}. Specifically, habitat preferences are estimated
-by comparing spatial covariates (e.g. environmental and anthropogenic) at
-locations visited by the animal to the same spatial covariates at randomly
-selected locations \citep{Zeller.2012}. Importantly, selection functions rely on
-adequate landscape and relocation data that are representative of the process
-being studied \citep{Diniz.2020}. For instance, relocation data collected on
-dispersing individuals has been shown to outperform data collected on resident
-individuals in the detection of large-scale movement corridors
-\citep{Elliot.2014, Diniz.2020}. Nevertheless, dispersal data is inherently
-difficult to collect and remains scarce in the connectivity literature
-\citep{Vasudev.2015}. As such, most permeability surfaces upon which movement
-corridors are identified are created using relocation data collected on resident
-individuals. This introduces severe biases and substantially reduces the power
-to reveal meaningful movement corridors, for dispersing individuals have
-different needs and drives compared to resident individuals \citep{Elliot.2014,
-Cozzi.2020}. Such biases have limited our ability to meaningfully assess the
-effectiveness of protected areas in securing connectivity for their protected
-species.
+species' relative selection strengths\todo{reviewer 1}, which can be quantified
+using a suite of selection functions \citep{Zeller.2012}. Specifically, relative
+selection strenghts \todo{reviewer 1} are estimated by comparing spatial
+covariates (e.g. environmental and anthropogenic) at locations visited by the
+animal to the same spatial covariates at locations avilable to, yet unused by
+\todo{Reviewer 1 pointed out correctly that we should talk about available
+locations, rather than randomly selected locations} the animal
+\citep{Zeller.2012}. Importantly, selection functions rely on adequate landscape
+and relocation data that are representative of the process being studied
+\citep{Diniz.2020}. For instance, relocation data collected on dispersing
+individuals has been shown to outperform data collected on resident individuals
+in the detection of large-scale movement corridors \citep{Elliot.2014,
+Diniz.2020}. Nevertheless, dispersal data is inherently difficult to collect and
+remains scarce in the connectivity literature \citep{Vasudev.2015}. As such,
+most permeability surfaces upon which movement corridors are identified are
+created using relocation data collected on resident individuals. While some
+species use the same habitats during residence and dispersal
+\citep{Fattebert.2015}, others depict vastly different habitat selection
+depending on the movement mode \citep{Elliot.2014, Abrahms.2017}. Thus, the use
+of data collected during residency can introduce severe biases in estimated
+landscape permeability and has likely limited our ability to meaningfully assess
+the effectiveness of protected areas in securing connectivity for their
+protected species.\todo{Reviewer 2 desired to change from: this introduces - to:
+this can introduce}
 
@@
 have attempted to assess the adequacy of the KAZA-TFCA using relevant global
-positioning system (GPS) relocation data of its protected species at the
-appropriate spatial scale \citep{Elliot.2014, Tshipa.2017}. Thus, how well the
-boundaries of the KAZA-TFCA reflect natural movement patterns and dispersal
-corridors of its most mobile protected species is virtually unknown.
+positioning system (GPS) relocation data of its protected species at large
+spatial scales \citep{Elliot.2014, Tshipa.2017}. Thus, how well the boundaries
+of the KAZA-TFCA reflect natural movement patterns and dispersal corridors of
+its most mobile protected species is virtually unknown.
 
@@
 analyzed it to assess the adequacy of the KAZA-TFCA in securing connectivity. We
-estimated the degree of selection or avoidance for environmental and
-anthropogenic landscape features and used the obtained habitat preferences to
-predict a permeability surface spanning the entire KAZA-TFCA. We then
-investigated how landscape permeability varies regionally and internationally
-and compared permeability within and outside the KAZA-TFCA boundaries. Finally,
-we calculated least-cost paths and corridors to identify major movement routes
-and to verify that these are successfully covered by the KAZA-TFCA.
+estimated the relative selection strengths\todo{reviewer 1} towards or agains
+environmental and anthropogenic landscape features and used the obtained habitat
+coefficients to predict a permeability surface spanning the entire KAZA-TFCA. We
+then investigated how landscape permeability varies regionally and
+internationally and compared permeability within and outside the KAZA-TFCA
+boundaries. Finally, we calculated least-cost paths and corridors to identify
+major movement routes and to verify that these are successfully covered by the
+KAZA-TFCA.
 
@@
 (7 female and 9 male coalitions). Candidate dispersing individuals were
-identified based on criteria reported in \cite{Behr.2020}, immobilized according
-to protocols described in \cite{Osofsky.1996}, and fitted with GPS/Satellite
-radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while
-still with their natal pack. All procedures were undertaken and supervised by a
-Botswana-registered wildlife veterinarian. During dispersal, GPS collars were
+identified based on age, number of same‐sex siblings, pack size, and presence of
+unrelated individuals of the opposite sex in ther pack citep \citep{McNutt.1996,
+Behr.2020}\todo{We originally only pointed to Behr and did not include the
+criteria. The editor desired to include those}. Individuals were immobilized
+according to protocols described in \cite{Osofsky.1996}, and fitted with
+GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
+Berlin, Germany}) while still with their natal pack. All procedures were
+undertaken and supervised by a Botswana-registered wildlife veterinarian. Fully
+assembled collars produced a tag weight of approximately xx kg, which conforms
+to the proposed threshold of 3\% of an animal's body weight
+\citep{Kenward.2000}. Collars were mounted using leather belts and included a
+drop off mechanism triggered by a slowly decomposing piece of cloth.\todo{Editor
+desired more information on GPS collar} During dispersal, GPS collars were
 programmed to record GPS relocation data every 4 hours and to regularly transmit
@@
 visual inspection of the net squared displacement (NSD) metric. NSD quantifies
-the Euclidean distance of a relocation to a reference point \citep{Borger.2012},
-which in our case was the center of the dispersing coalition's natal home range.
-Thus, dispersal was deemed to have started when a coalition had left its natal
-home range and continued until the NSD metric remained stationary, implying that
-the coalition had successfully settled (Figure S1). In total, we collected 4'169
-GPS relocations during dispersal (Figure S2 \& Table S1), resulting in an
-average of 260 locations per dispersing coalition (min = 37, max = 729). In our
-analysis, we did not differentiate between male and female dispersing
-coalitions, for previous research found little differences between sexes during
-dispersal \citep{Woodroffe.2019, Cozzi.2020}.
+the squared \todo{Reviewer 1 pointed out correctly that we need to talk about
+the \textit{squared} euclidean distance} Euclidean distance of a relocation to a
+reference point \citep{Borger.2012}, which in our case was the center of the
+dispersing coalition's natal home range. Thus, dispersal was deemed to have
+started when a coalition had left its natal home range and continued until the
+NSD metric remained stationary, implying that the coalition had successfully
+settled (Figure S1). In our analysis, we did not differentiate between male and
+female dispersing coalitions, for previous research found little differences
+between sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.
 
 \subsection{Spatial Covariates}
-To investigate habitat preferences of dispersing wild dogs, we used a set of
-geo-referenced covariates (\Cref{Covariates}) that we aggregated in the
-categories \textit{land cover} (which included water cover, distance to water,
-shrubs/grassland cover, and tree cover), \textit{protection status} (protected
-vs. unprotected), and \textit{human pressure} (which included human influence,
-presence of roads, and distance to roads). For each of these covariates we
-prepared spatial raster layers from freely available online services or from
-remotely sensed satellite imagery. To ensure a consistent resolution (i.e.
-cell-size or grain) across covariates, we coarsened or interpolated all layers
-to match a resolution of 250m x 250m. For further details on the preparation of
-each covariate, see Appendix A.3. We performed processing and manipulation of
-data as well as all spatial and statistical analyses using R, version 3.6.1
-\citep{R.2019}.
+To investigate habitat relative selection strength\todo{As per reviewer 1} of
+dispersing wild dogs, we used a set of geo-referenced covariates
+(\Cref{Covariates}) that we aggregated in the categories \textit{land cover}
+(which included water cover, distance to water, shrubs/grassland cover, and tree
+cover), \textit{protection status} (protected vs. unprotected), and
+\textit{human pressure} (which included human density, presence of roads, and
+distance to roads). For each of these covariates we prepared spatial raster
+layers from freely available online services or from remotely sensed satellite
+imagery. To ensure a consistent resolution (i.e. cell-size or grain) across
+covariates, we coarsened or interpolated all layers to match a resolution of
+250m x 250m. For further details on the preparation of each covariate, see
+Appendix A.3. We performed processing and manipulation of data as well as all
+spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.
 
@@
 We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
-investigate dispersers' selection or avoidance for the above-mentioned spatial
-covariates. That is, we paired each realized step (i.e. the connecting line
-between two consecutive GPS relocations; \citealp{Turchin.1998}) with 24 random
-steps. We generated random steps by sampling turning angles from a uniform
-distribution U(\(-\pi, +\pi\)) and step lengths from a gamma distribution that
-was fitted using realized steps \citep{Avgar.2016}. A realized step and its 24
-associated random steps formed a stratum that received a unique identifier.
-Along each step, we extracted the above-mentioned covariates (Table S3),
-standardized extracted values using a z-score transformation, and checked for
-correlation using Pearson's Correlation Coefficient \(r\). None of the
-covariates were overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we
-retained all of them for modeling. Our habitat selection model then assumed that
-dispersing wild dogs assigned a selection score \(w(x)\) of the following
-exponential form to each realized and random step \citep{Fortin.2005}:
+investigate dispersers' relative selection strentgh \todo{Reviewer 1 suggested
+to use relative selection strength, rather than preference} for the
+above-mentioned spatial covariates. That is, we paired each realized step (i.e.
+the connecting line between two consecutive GPS relocations;
+\citealp{Turchin.1998}) with 24 random steps. We generated random steps by
+sampling turning angles from a uniform distribution U(\(-\pi, +\pi\)) and step
+lengths from a gamma distribution that was fitted using realized steps
+\citep{Avgar.2016}. A realized step and its 24 associated random steps formed a
+stratum that received a unique identifier. Along each step, we extracted the
+above-mentioned covariates (Table S3), standardized extracted values using a
+z-score transformation, and checked for correlation using Pearson's Correlation
+Coefficient \(r\). None of the covariates were overly correlated (\(|r| > 0.6\);
+\citealp{Latham.2011}) and we retained all of them for modeling. Our habitat
+selection model then assumed that dispersing wild dogs assigned a selection
+score \(w(x)\) of the following exponential form to each realized and random
+step \citep{Fortin.2005}:
 
@@
 \noindent The selection score \(w(x)\) of a step depended on its associated
-covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's preferences for
-these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). To estimate habitat
-preferences (i.e. the \(\beta\)'s), we used mixed effects conditional logistic
+covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's relative
+selection strength \todo{As per reviewer 1} for these covariates (\(\beta_1,
+\beta_2, ..., \beta_n\)). To estimate relative selection strengths \todo{As per
+reviewer 1} (i.e. the \(\beta\)'s), we used mixed effects conditional logistic
 regression analysis as suggested by \cite{Muff.2020}. We implemented their
 method using the R-package \textit{glmmTMB} \citep{Mollie.2017} and used
-dispersing coalition ID to model random intercepts and slopes. We defined two
-movement metrics, namely the cosine of the turning angle (\(cos(ta)\)) and the
-log of the step length (\(log(sl)\)), as core covariates and ran stepwise
-forward model selection based on Akaike's Information Criterion (AIC;
-\citealp{Burnham.2002}) for all other covariates. The inclusion of movement
-metrics served to reduce biases in estimated habitat preferences that may have
-arisen due to movement preferences \citep{Avgar.2016}. To validate the
-predictive power of the most parsimonious habitat selection model, we ran k-fold
-cross-validation for case-control studies as described in \cite{Fortin.2009}
-(details in Appendix A.5).
+dispersing coalition ID to model random slopes. We also modeled random
+intercepts with an arbitrary high variance of 1e6 to make use of the
+\textit{poisson trick} (see \cite{Muff.2020}). We defined three movement
+metrics, namely the cosine of the turning angle (\(cos(ta)\)), the step length
+\(sl_\) and the natural logarithm of the step length (\(log(sl)\)), as core
+covariates and ran stepwise forward model selection based on Akaike's
+Information Criterion (AIC; \citealp{Burnham.2002}) for all other
+covariates.\todo{Reviewer 1 asked for more details in the random effects. I
+therefore decided to update this section too} The inclusion of movement metrics
+served to reduce biases in estimated habitat selection coefficients \todo{as per
+revierwer 1} that may have arisen due to movement behavior \citep{Avgar.2016}.
+To validate the predictive power of the most parsimonious habitat selection
+model, we ran k-fold cross-validation for case-control studies as described in
+\cite{Fortin.2009} (details in Appendix A.5).
 
@@
 \section{Results}
+\subsection{Dispersal Events}
+In total, we collected 4'169 GPS relocations during dispersal (Figure S2 \&
+Table S1), resulting in an average of 261 (\(SD\) = 207) locations per
+dispersing coalition. Dispersers on dispersed on average for 48 days (\(SD\) =
+44), covered a mean euclidean distance of 54 km (\(SD\) = 71) and a cumulative
+distance of 597 km (\(SD\) = 508). One female coalition dispersed far east into
+the Hwange Hwange National Park and covered a cumulative distance of over 360 km
+in under 9 days.\todo{Editor desired a quick summary of the dispersal data.}
+
 \subsection{Habitat Selection Model}
@@
 \textit{human influence}, beside the fixed covariates \textit{cos(ta)} and
-\textit{log(sl)} (\Cref{PermeabilityResults}a). Parameter estimates showed that
-dispersing wild dogs moved in a directional and fast manner, as indicated by a
-positive selection for small turning angles, i.e. high \(cos(ta)\) (\(\beta\) =
-0.14; 95\% CI = 0.07 to 0.21) and longer steps, i.e. high \(log(sl)\) (\(\beta\)
-= 0.06, 95\% CI = 0.02 to 0.09). Dispersers avoided moving through water
-(\(\beta\) = -0.52, 95\% CI -0.77 to -0.26) but selected for locations in its
-vicinity, although the latter effect was not significant (\(\beta\) = -0.32,
-95\% CI = -0.72 to 0.08). Dispersers avoided areas that were densely covered by
-trees (\(\beta\) = -0.31, CI = -0.46 to -0.15) and preferred areas covered by
-shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42). Finally, dispersers
-avoided areas that were influenced by humans (\(\beta\) = -0.41, 95\% CI = -0.78
-to -0.05).
+\textit{log(sl)} (\Cref{PermeabilityResults}a). The positive coefficient for
+\(log(sl)\) (\(\beta\) = 0.06, 95\% CI = 0.02 to 0.09) showed that our fitted
+(yet selection-biased) gamma distribution produced step-lengths slightly shorter
+than under a selection-free gamma distribution. Similarly, the positive
+coefficient for \(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21) indicates
+that the uniform distribution for turning angles produced turning angles with
+too little directionality.\todo{Reviewer 1 nicely explained that the
+coefficients for log(sl) and cos(ta) imply that the initial distributions were
+selection-biased} With respect to environmental covariates, dispersers avoided
+moving through water (\(\beta\) = -0.52, 95\% CI -0.77 to -0.26) but selected
+for locations in its vicinity, although the latter effect was not significant
+(\(\beta\) = -0.32, 95\% CI = -0.72 to 0.08). Dispersers avoided areas that were
+densely covered by trees (\(\beta\) = -0.31, CI = -0.46 to -0.15) and preferred
+areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42).
+Finally, dispersers avoided areas that were influenced by humans (\(\beta\) =
+-0.41, 95\% CI = -0.78 to -0.05).
 
@@
     based on the raster cell's underlying covariates (\(x_i\)) and estimated
-    habitat preferences (\(\beta_i\)). Areas that dispersers find easy to
+    selection strenghts (\(\beta_i\)). Areas that dispersers find easy to
     traverse are depicted in bright colors. Bold white lines delineate the
@@
 Our results emphasize that human influences constitute some of the main barriers
-to connectivity among wild dog populations. This conforms to findings on
-dispersing wild dogs from eastern Africa \citep{Masenga.2016, Oneill.2020} but
-conflicts with findings from South Africa by \cite{DaviesMostert.2012}, who
-reported a high willingness of dispersers to cross human-dominated landscapes.
-We believe that such differences are due to the unavailability of alternative
-routes through natural landscapes, which may have forced dispersers in South
-Africa to cross human dominated landscapes despite a strong aversion to do so.
-In this regard, our representation of dispersal corridors and the resulting
-connectivity appear conservative, as dispersers may be able to make the best out
-of a bad situation and cross landscapes characterized by considerably
-unfavorable conditions \citep{Palomares.2000, Elliot.2014}. Nevertheless,
-successful conservation of this species relies on policymakers' and local
-authorities' willingness and ability to provide and conserve natural areas that
-remain free from anthropogenic pressures. This is not only paramount in light of
-increasing connectivity and facilitating dispersal, but also in terms of
-reducing human-caused mortality during dispersal. In fact, previous studies have
-shown that human-caused mortality represents a major threat to wild dogs'
-ability to disperse \citep{Woodroffe.2019, Cozzi.2020}.
+to connectivity among wild dog populations. According to our model, dispersers
+avoided corssing human dominated landscapes whenever given the
+choice.\todo{According to reviewer 1 we should point out that our model compares
+available vs. unavailable} This conforms to findings on dispersing wild dogs
+from eastern Africa \citep{Masenga.2016, Oneill.2020} but conflicts with
+findings from South Africa by \cite{DaviesMostert.2012}, who reported a high
+willingness of dispersers to cross human-dominated landscapes. We believe that
+such differences are due to the unavailability of alternative routes through
+natural landscapes, which may have forced dispersers in South Africa to cross
+human dominated landscapes despite a strong aversion to do so. In this regard,
+our representation of dispersal corridors and the resulting connectivity appear
+conservative, as dispersers may be able to make the best out of a bad situation
+and cross landscapes characterized by considerably unfavorable conditions
+\citep{Palomares.2000, Elliot.2014}. Nevertheless, successful conservation of
+this species relies on policymakers' and local authorities' willingness and
+ability to provide and conserve natural areas that remain free from
+anthropogenic pressures. This is not only paramount in light of increasing
+connectivity and facilitating dispersal, but also in terms of reducing
+human-caused mortality during dispersal. In fact, previous studies have shown
+that human-caused mortality represents a major threat to wild dogs' ability to
+disperse \citep{Woodroffe.2019, Cozzi.2020}.
 
@@
 Although dispersers avoided moving through water, they selected locations in its
-vicinity. This preference may be caused by the occurence of prey close to water
+vicinity. This behavior may be caused by the occurence of prey close to water
 \citep{Bonyongo.2005}. For the same reason, however, competitors such as lions,
@@
 \citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to move
-into prey-poorer areas away from water. Given the influence that resident
-conspecifics, competitors, and prey can have on dispersers \citep{Cozzi.2018,
-Armansin.2019} future studies should strive to collect and incorporate intra-
-and interspecific relationships into analyses of landscape connectivity.
+into prey-poorer areas away from water. This could also explain the effect of
+distane to water was insignificant.\todo{Reviewer 1 suggested to add this} Given
+the influence that resident conspecifics, competitors, and prey can have on
+dispersers \citep{Cozzi.2018, Armansin.2019} future studies should strive to
+collect and incorporate intra- and interspecific relationships into analyses of
+landscape connectivity.
+
+Overall, our findings on habitat selection of dispersing wild dogs coincide with
+findings from Kenya by \cite{Oneill.2020}, suggesting that there are strong
+commonalities across ecosystems. Hence, despite wild dogs' ability to cope with
+and adapt to a wide range of habitats, we feel that that the fundamental factors
+included in our study and their influences can be generalized to other
+populations.\todo{Editor desired that we discuss the applicability of our
+ecosystem to other ecosystems.}
 
@@
 dispersal journey and that they have almost complete knowledge of associated
-movement costs \citep{Panzacchi.2016}. Since dispersers often move into unknown
-territory, this may not necessarily be the case \citep{Abrahms.2017,
-Cozzi.2020}. However, specification of pre-defined end points might not be
-necessary, as the parametrized iSSF model can be used as mechanistic movement
-model to simulate dispersal events from known source points, yet without
-restricting the domain of potential end points \citep{Signer.2017}.
-Consequently, movement corridors would emerge more naturally as the result of a
-myriad of simulated dispersal events. While a simulation-based approach is
-conceptually straightforward, it requires a comprehensive mechanistic
-understanding of dispersal movements, which is conditional on our ability to
-collect additional dispersal data and adequately represent the landscape through
-which individuals move.
+movement costs \citep{Koen.2014, Panzacchi.2016}. Since dispersers often move
+into unknown territory, this may not necessarily be the case
+\citep{Abrahms.2017, Cozzi.2020}. However, specification of pre-defined end
+points might not be necessary, as the parametrized iSSF model can be used as
+mechanistic movement model to simulate dispersal events from known source
+points, yet without restricting the domain of potential end points
+\citep{Signer.2017}. Consequently, movement corridors (i.e. transient
+utilisation distribution) would emerge more naturally as the result of a myriad
+of simulated dispersal events. A very similar approach could be applied to
+produce permeability surfaces from steady state utilistation distributions
+\citep{Avgar.2016, Signer.2017}. \citep{Signer.2017} have shown that such
+permeability surfaces are less prone to overestimateing permeability in areas
+that lie far from suitable habitats. While a simulation-based approach is
+conceptually straightforward, computational requirements are tremendous,
+especially for such a large extent as constituted by the KAZA-TFCA.
+\todo{Reviewer 1 pointed out that the permeability surface likely overestimates
+permeability in areas far from suitable habitat. Additionally, reviewer 2
+mentioned that our dataset by far suffices for simulations and we should not
+argue with sample size.}
 
-Our work shows how dispersal data of a highly mobile species can be used t
-identify movement corridors and to assess the adequacy of protected areas. In
-our case, the predicted movement corridors of African wild dogs were well
-contained within the boundaries of the world's largest transboundary
-conservation area, namely the KAZA-TFCA, suggesting that it will significantly
-contribute to the long-term viability of this species. Moreover, our
-connectivity network allowed revealing potential dispersal hubs through which
-dispersers gain access to more remote regions of the study area. Finally, our
-investigations showed that human influence constitutes one of the main barriers
-to dispersal and substantially reduces landscape connectivity. Successful
-conservation of wide-ranging species, such as exemplified by the African wild
-dog, will therefore be contingent on the willingness of local authorities,
-policymakers, and land managers to preserve areas that remain free from human
-strains. Ultimately, our work contributes to the growing field of connectivity
-studies and provides and easily expandable framework for assessing the adequacy
-of already-existing or planned protected areas.
+Our work shows how dispersal data of a highly mobile species can be used to
+\todo{to instead of t, pointed out by the editor and reviewer 1} identify
+movement corridors and to assess the adequacy of protected areas. In our case,
+the predicted movement corridors of African wild dogs were well contained within
+the boundaries of the world's largest transboundary conservation area, namely
+the KAZA-TFCA, suggesting that it will significantly contribute to the long-term
+viability of this species. Moreover, our connectivity network allowed revealing
+potential dispersal hubs through which dispersers gain access to more remote
+regions of the study area. Finally, our investigations showed that human
+influence constitutes one of the main barriers to dispersal and substantially
+reduces landscape connectivity. Successful conservation of wide-ranging species,
+such as exemplified by the African wild dog, will therefore be contingent on the
+willingness of local authorities, policymakers, and land managers to preserve
+areas that remain free from human strains. Ultimately, our work contributes to
+the growing field of connectivity studies and provides and easily expandable
+framework for assessing the adequacy of already-existing or planned protected
+areas.
 
@@
 Wolski, from the Okavango Research Institute, who assisted us in generating
-dynamic water maps. This study was funded by Basler Stiftung für Biologische
-Forschung, Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic
-Society, Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust
-Foundation, Forschungkredit der Universität Zürich, and a Swiss National Science
-Foundation Grant (31003A\_182286) to A. Ozgul.
+dynamic water maps. We also thank the anonymous reviewers for their valuable
+comments and suggestions. This study was funded by Basler Stiftung für
+Biologische Forschung, Claraz Foundation, Idea Wild, Jacot Foundation, National
+Geographic Society, Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife
+Trust Foundation, Forschungkredit der Universität Zürich, and a Swiss National
+Science Foundation Grant (31003A\_182286) to A. Ozgul.
 

### 2020-09-02 Chapter_0/04_Manuscript/99_PermeabilityModelAICs.tex
Worked on the revision for BES paper

@@
 % latex table generated in R 3.6.3 by xtable 1.8-4 package
-% Thu Jun  4 15:54:16 2020
+% Wed Sep  2 09:04:38 2020
 \begin{tabular}{lrrrr}
@@
   \midrule
-cos(ta) + log(sl) + W + T + DTW + HB5 + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 & 90094.79 & 3.38 & 0.00 & -45034.39 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\ 
-  cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\ 
-  cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\ 
-  cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\ 
-  cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\ 
-  cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\ 
-  cos(ta) + log(sl) + W + T + HB5 & 90123.71 & 32.31 & 0.00 & -45050.85 \\ 
-  cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\ 
-  cos(ta) + log(sl) + W + T + S & 90129.36 & 37.96 & 0.00 & -45053.68 \\ 
-  cos(ta) + log(sl) + W + T + DTR & 90130.11 & 38.71 & 0.00 & -45054.05 \\ 
-  cos(ta) + log(sl) + W + T + P & 90131.75 & 40.35 & 0.00 & -45054.88 \\ 
-  cos(ta) + log(sl) + W + T + RC & 90133.53 & 42.13 & 0.00 & -45053.77 \\ 
-  cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\ 
-  cos(ta) + log(sl) + W + HB5 & 90149.30 & 57.90 & 0.00 & -45065.65 \\ 
-  cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\ 
-  cos(ta) + log(sl) + W & 90153.70 & 62.30 & 0.00 & -45069.85 \\ 
-  cos(ta) + log(sl) + W + DTR & 90156.32 & 64.92 & 0.00 & -45069.16 \\ 
-  cos(ta) + log(sl) + W + P & 90157.23 & 65.83 & 0.00 & -45069.62 \\ 
-  cos(ta) + log(sl) + W + RC & 90159.32 & 67.91 & 0.00 & -45068.66 \\ 
-  cos(ta) + log(sl) + S & 90161.07 & 69.67 & 0.00 & -45073.54 \\ 
-  cos(ta) + log(sl) + DTW & 90262.03 & 170.63 & 0.00 & -45124.01 \\ 
-  cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\ 
-  cos(ta) + log(sl) + HB5 & 90307.55 & 216.15 & 0.00 & -45146.78 \\ 
-  cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\ 
-  cos(ta) + log(sl) + P & 90316.62 & 225.22 & 0.00 & -45151.31 \\ 
-  cos(ta) + log(sl) + RC & 90318.46 & 227.06 & 0.00 & -45150.23 \\ 
+cos(ta) + sl + log(sl) + W + T + DTW + HI + S & 90068.15 & 0.00 & 1.00 & -45017.08 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI & 90071.84 & 3.69 & 0.00 & -45020.92 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR & 90071.94 & 3.79 & 0.00 & -45016.97 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + P & 90071.94 & 3.79 & 0.00 & -45016.97 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + RC & 90073.46 & 5.30 & 0.00 & -45015.73 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + DTR & 90075.66 & 7.50 & 0.00 & -45020.83 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + P & 90075.66 & 7.51 & 0.00 & -45020.83 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + P & 90075.79 & 7.64 & 0.00 & -45016.89 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + S & 90076.71 & 8.56 & 0.00 & -45023.36 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + RC & 90076.84 & 8.69 & 0.00 & -45019.42 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + RC & 90077.20 & 9.05 & 0.00 & -45015.60 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW & 90080.08 & 11.92 & 0.00 & -45027.04 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90080.96 & 12.81 & 0.00 & -45015.48 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + DTR & 90082.95 & 14.79 & 0.00 & -45026.47 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + P & 90083.31 & 15.16 & 0.00 & -45026.66 \\ 
+  cos(ta) + sl + log(sl) + W + T + HI & 90103.28 & 35.13 & 0.00 & -45038.64 \\ 
+  cos(ta) + sl + log(sl) + W + T & 90109.40 & 41.25 & 0.00 & -45043.70 \\ 
+  cos(ta) + sl + log(sl) + W + T + S & 90110.35 & 42.20 & 0.00 & -45042.17 \\ 
+  cos(ta) + sl + log(sl) + W + T + DTR & 90112.55 & 44.40 & 0.00 & -45043.27 \\ 
+  cos(ta) + sl + log(sl) + W + T + P & 90113.11 & 44.96 & 0.00 & -45043.56 \\ 
+  cos(ta) + sl + log(sl) + W + T + RC & 90113.60 & 45.45 & 0.00 & -45041.80 \\ 
+  cos(ta) + sl + log(sl) + W + DTW & 90118.55 & 50.40 & 0.00 & -45048.28 \\ 
+  cos(ta) + sl + log(sl) + W + HI & 90128.70 & 60.54 & 0.00 & -45053.35 \\ 
+  cos(ta) + sl + log(sl) + W + S & 90132.22 & 64.06 & 0.00 & -45055.11 \\ 
+  cos(ta) + sl + log(sl) + W & 90134.85 & 66.69 & 0.00 & -45058.42 \\ 
+  cos(ta) + sl + log(sl) + W + DTR & 90138.31 & 70.16 & 0.00 & -45058.16 \\ 
+  cos(ta) + sl + log(sl) + W + P & 90138.50 & 70.35 & 0.00 & -45058.25 \\ 
+  cos(ta) + sl + log(sl) + W + RC & 90139.30 & 71.15 & 0.00 & -45056.65 \\ 
+  cos(ta) + sl + log(sl) + S & 90141.98 & 73.83 & 0.00 & -45061.99 \\ 
+  cos(ta) + sl + log(sl) + DTW & 90225.64 & 157.49 & 0.00 & -45103.82 \\ 
+  cos(ta) + sl + log(sl) + T & 90271.73 & 203.58 & 0.00 & -45126.86 \\ 
+  cos(ta) + sl + log(sl) + HI & 90273.18 & 205.02 & 0.00 & -45127.59 \\ 
+  cos(ta) + sl + log(sl) + P & 90285.24 & 217.08 & 0.00 & -45133.62 \\ 
+  cos(ta) + sl + log(sl) + DTR & 90285.33 & 217.18 & 0.00 & -45133.67 \\ 
+  cos(ta) + sl + log(sl) + RC &  &  & 0.00 &  \\ 
+  cos(ta) + sl + log(sl) + W + T + DTW + RC &  &  & 0.00 &  \\ 
    \bottomrule

### 2020-09-02 Chapter_0/04_Manuscript/Manuscript.tex
Worked on the revision for BES paper

@@
 undertaken and supervised by a Botswana-registered wildlife veterinarian. Fully
-assembled collars produced a tag weight of approximately xx kg, which conforms
-to the proposed threshold of 3\% of an animal's body weight
-\citep{Kenward.2000}. Collars were mounted using leather belts and included a
-drop off mechanism triggered by a slowly decomposing piece of cloth.\todo{Editor
-desired more information on GPS collar} During dispersal, GPS collars were
-programmed to record GPS relocation data every 4 hours and to regularly transmit
-them via Iridium satellite system to a base station.
+assembled collars produced a tag weight of approximately 330 g, accounting for
+approximately 1.2\% of a wild dog's average body weight. Collars were mounted
+using leather belts and included a drop off mechanism, triggered by a slowly
+decomposing piece of cloth.\todo{Editor desired more information on GPS collar}
+During dispersal, GPS collars were programmed to record GPS relocation data
+every 4 hours and to regularly transmit them via Iridium satellite system to a
+base station.
 
@@
 \subsection{Spatial Covariates}
-To investigate habitat relative selection strength\todo{As per reviewer 1} of
+To investigate relative selection strengths \todo{As per reviewer 1} of
 dispersing wild dogs, we used a set of geo-referenced covariates
-(\Cref{Covariates}) that we aggregated in the categories \textit{land cover}
-(which included water cover, distance to water, shrubs/grassland cover, and tree
-cover), \textit{protection status} (protected vs. unprotected), and
-\textit{human pressure} (which included human density, presence of roads, and
-distance to roads). For each of these covariates we prepared spatial raster
-layers from freely available online services or from remotely sensed satellite
-imagery. To ensure a consistent resolution (i.e. cell-size or grain) across
-covariates, we coarsened or interpolated all layers to match a resolution of
-250m x 250m. For further details on the preparation of each covariate, see
-Appendix A.3. We performed processing and manipulation of data as well as all
-spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.
+(\Cref{Covariates}) that we aggregated in the categories \textit{land cover},
+\textit{protection status}, and \textit{human influence}. Land cover comprised
+of the covariates water cover, distance to water, percentage cover by
+shrubs/grassland, and percentage cover by trees. To capture the pulsing behavior
+of the Okavango Delta we frequently updated water cover layers and corresponding
+distance to water layers. Protection status contained a binary indicator of
+whether an area was protected or not. Human influence included covariates
+rendering human density, the presence of roads, and the distance to
+roads.\todo{Reviewer 2 found the description of the covariates unclear} For each
+of these covariates we prepared spatial raster layers from freely available
+online services or from remotely sensed satellite imagery. To ensure a
+consistent resolution (i.e. cell-size or grain) across covariates, we coarsened
+or interpolated all layers to match a resolution of 250m x 250m. For further
+details on the preparation of each covariate, see Appendix A.3. We performed
+processing and manipulation of data as well as all spatial and statistical
+analyses using R, version 3.6.1 \citep{R.2019}.
 
@@
 method using the R-package \textit{glmmTMB} \citep{Mollie.2017} and used
-dispersing coalition ID to model random slopes. We also modeled random
-intercepts with an arbitrary high variance of 1e6 to make use of the
-\textit{poisson trick} (see \cite{Muff.2020}). We defined three movement
+dispersing coalition ID to model random slopes. We also modelled random
+intercepts with an arbitrary high variance of 10\textsuperscript{6} to make use
+of the poisson trick (see \citealp{Muff.2020}). We defined three movement
 metrics, namely the cosine of the turning angle (\(cos(ta)\)), the step length
-\(sl_\) and the natural logarithm of the step length (\(log(sl)\)), as core
+(\(sl\)) and the natural logarithm of the step length (\(log(sl)\)), as core
 covariates and ran stepwise forward model selection based on Akaike's
@@
 between source points using the R-package \textit{gdistance} (for further
-details see Appendix A.6). After computation, we tallied overlapping LCPs and
+details see Appendix A.7). After computation, we tallied overlapping LCPs and
 LCCs, respectively, into single connectivity maps.
@@
 Table S1), resulting in an average of 261 (\(SD\) = 207) locations per
-dispersing coalition. Dispersers on dispersed on average for 48 days (\(SD\) =
+dispersing coalition. Coalitions on average dispersed for 48 days (\(SD\) =
 44), covered a mean euclidean distance of 54 km (\(SD\) = 71) and a cumulative
@@
 \textit{log(sl)} (\Cref{PermeabilityResults}a). The positive coefficient for
-\(log(sl)\) (\(\beta\) = 0.06, 95\% CI = 0.02 to 0.09) showed that our fitted
-(yet selection-biased) gamma distribution produced step-lengths slightly shorter
-than under a selection-free gamma distribution. Similarly, the positive
+\(log(sl)\) (\(\beta\) = 0.06, 95\% CI = 0.02 to 0.09) showed that our initial
+(selection-biased) gamma distribution produced step-lengths slightly shorter
+than a selection-free gamma distribution would. Similarly, the positive
 coefficient for \(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21) indicates
-that the uniform distribution for turning angles produced turning angles with
-too little directionality.\todo{Reviewer 1 nicely explained that the
+that our uniform distribution for turning angles produced turning angles with
+too little directionality. \todo{Reviewer 1 nicely explained that the
 coefficients for log(sl) and cos(ta) imply that the initial distributions were
@@
 Finally, dispersers avoided areas that were influenced by humans (\(\beta\) =
--0.41, 95\% CI = -0.78 to -0.05).
+-0.41, 95\% CI = -0.78 to -0.05). With the exepction of \textit{DistanceToWater}
+(\(SD_{RE}\) = 0.57), random effects revealed little variability in different
+dispersal coalitions' selection strengths (i.e. \(SD_{RE} \leq 0.22\), Figure
+S7).\todo{Reviewer 1 desired more information on the random effects}
 
@@
     \midrule
-    Angola & 0.36 (0.41) & 0.12 (0.32) & 0.36 (0.41) & 0.12 (0.33) & 0.20 (0.39)\\
-    Botswana & 0.25 (0.30) & 0.15 (0.16) & 0.28 (0.35) & 0.15 (0.18) & 0.19 (0.25)\\
-    Namibia & 0.22 (0.30) & 0.13 (0.18) & 0.24 (0.30) & 0.11 (0.15) & 0.16 (0.24)\\
-    Zambia & 0.05 (0.09) & 0.03 (0.06) & 0.04 (0.10) & 0.03 (0.05) & 0.03 (0.07)\\
-    Zimbabwe & 0.07 (0.16) & 0.06 (0.05) & 0.08 (0.17) & 0.05 (0.05) & 0.06 (0.07)\\
+    Angola & 0.35 (0.41) & 0.12 (0.32) & 0.35 (0.41) & 0.12 (0.32) & 0.19 (0.38)\\
+    Botswana & 0.24 (0.30) & 0.14 (0.16) & 0.27 (0.35) & 0.14 (0.18) & 0.18 (0.25)\\
+    Namibia & 0.20 (0.30) & 0.12 (0.17) & 0.22 (0.30) & 0.10 (0.14) & 0.14 (0.24)\\
+    Zambia & 0.05 (0.09) & 0.02 (0.05) & 0.04 (0.09) & 0.03 (0.05) & 0.03 (0.06)\\
+    Zimbabwe & 0.06 (0.16) & 0.05 (0.04) & 0.07 (0.17) & 0.04 (0.04) & 0.05 (0.06)\\
     \hline
-    Overall & 0.16 (0.30) & 0.07 (0.15) & 0.15 (0.30) & 0.07 (0.15) & 0.10 (0.22)\\
+    Overall & 0.15 (0.29) & 0.06 (0.14) & 0.14 (0.30) & 0.06 (0.14) & 0.08 (0.21)\\
     \bottomrule

### 2020-09-02 Chapter_0/04_Manuscript/Manuscript_Support.tex
Working on paper revision
Worked on the revision for BES paper

@@
 \usepackage[flushleft]{threeparttable}
+\usepackage{todonotes}
 
@@
 %------------------------------------------------------------------------------
-\section{Net Squared Displacement}
+\section{Net Squared Displacement}\todo{Reviewer 1 pointed out that we need to
+replace calssified with classified}
 \begin{figure}[hbpt]
@@
     \caption{NSD displacement through time for one of our dispersers. The blue
-    line indicates the period during which we calssified the individual as
+    line indicates the period during which we classified the individual as
     dispersing.}
@@
       Taryn & F & AP & 896 & 37 & 9 & 10 & 130 \\
+      \toprule
+      \toprule
+      \textit{Mean} & - & - & 1080 & 261 & 48 & 54 & 597 \\
+      \textit{(SD)} & - & - & (649) & (207) & (44) & (71) & (508) \\
       \bottomrule
@@
 cover}, \textit{protection status}, and \textit{human influence}. We did not
-include terrain features as covariates due to the absence of noteworthy
-elevational gradients in our study area. For each covariate, we prepared spatial
-raster layers from freely available online services or from remotely sensed
-satellite imagery. To ensure a consistent resolution (i.e. cell-size or grain)
-across covariates, we coarsened or interpolated all layers to match a resolution
-of 250m x 250m. We performed processing and manipulation of data as well as all
+include any terrain features due to the absence of noteworthy elevational
+gradients in our study area. For each covariate, we prepared spatial raster
+layers from freely available online services or from remotely sensed satellite
+imagery. To ensure a consistent resolution (i.e. cell-size or grain) across
+covariates, we coarsened or interpolated all layers to match a resolution of
+250m x 250m. We performed processing and manipulation of data as well as all
 spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.
@@
 %------------------------------------------------------------------------------
-%  Appendix S6: Least Cost Paths and Least Cost Corridors
+%  Appendix S6: Random Effects Variability
+%------------------------------------------------------------------------------
+\newpage
+\section{Habitat Selection Model: Random Effects}
+\begin{figure}[hbtp]
+  \includegraphics[width = \textwidth]{99_RandomEffects}
+  \caption{Plot of random effects showing variability across dispersal
+  coalitions.}
+  \label{RandomEffects}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%  Appendix S7: Least Cost Paths and Least Cost Corridors
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%  Appendix S7: AIC Model Selection
+%  Appendix S8: AIC Model Selection
 %------------------------------------------------------------------------------
@@
           \midrule
-          cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\
-          cos(ta) + log(sl) + W + T + DTW + HI & 90094.79 & 3.38 & 0.00 & -45034.39 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\
-          cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\
-          cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\
-          cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\
-          cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\
-          cos(ta) + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\
-          cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\
-          cos(ta) + log(sl) + W + T + HI & 90123.71 & 32.31 & 0.00 & -45050.85 \\
-          cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\
-          cos(ta) + log(sl) + W + T + S & 90129.36 & 37.96 & 0.00 & -45053.68 \\
-          cos(ta) + log(sl) + W + T + DTR & 90130.11 & 38.71 & 0.00 & -45054.05 \\
-          cos(ta) + log(sl) + W + T + P & 90131.75 & 40.35 & 0.00 & -45054.88 \\
-          cos(ta) + log(sl) + W + T + RC & 90133.53 & 42.13 & 0.00 & -45053.77 \\
-          cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\
-          cos(ta) + log(sl) + W + HI & 90149.30 & 57.90 & 0.00 & -45065.65 \\
-          cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\
-          cos(ta) + log(sl) + W & 90153.70 & 62.30 & 0.00 & -45069.85 \\
-          cos(ta) + log(sl) + W + DTR & 90156.32 & 64.92 & 0.00 & -45069.16 \\
-          cos(ta) + log(sl) + W + P & 90157.23 & 65.83 & 0.00 & -45069.62 \\
-          cos(ta) + log(sl) + W + RC & 90159.32 & 67.91 & 0.00 & -45068.66 \\
-          cos(ta) + log(sl) + S & 90161.07 & 69.67 & 0.00 & -45073.54 \\
-          cos(ta) + log(sl) + DTW & 90262.03 & 170.63 & 0.00 & -45124.01 \\
-          cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\
-          cos(ta) + log(sl) + HI & 90307.55 & 216.15 & 0.00 & -45146.78 \\
-          cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\
-          cos(ta) + log(sl) + P & 90316.62 & 225.22 & 0.00 & -45151.31 \\
-          cos(ta) + log(sl) + RC & 90318.46 & 227.06 & 0.00 & -45150.23 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S & 90068.15 & 0.00 & 1.00 & -45017.08 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI & 90071.84 & 3.69 & 0.00 & -45020.92 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR & 90071.94 & 3.79 & 0.00 & -45016.97 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S + P & 90071.94 & 3.79 & 0.00 & -45016.97 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S + RC & 90073.46 & 5.30 & 0.00 & -45015.73 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + DTR & 90075.66 & 7.50 & 0.00 & -45020.83 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + P & 90075.66 & 7.51 & 0.00 & -45020.83 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + P & 90075.79 & 7.64 & 0.00 & -45016.89 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + S & 90076.71 & 8.56 & 0.00 & -45023.36 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + RC & 90076.84 & 8.69 & 0.00 & -45019.42 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + RC & 90077.20 & 9.05 & 0.00 & -45015.60 \\
+          cos(ta) + sl + log(sl) + W + T + DTW & 90080.08 & 11.92 & 0.00 & -45027.04 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90080.96 & 12.81 & 0.00 & -45015.48 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + DTR & 90082.95 & 14.79 & 0.00 & -45026.47 \\
+          cos(ta) + sl + log(sl) + W + T + DTW + P & 90083.31 & 15.16 & 0.00 & -45026.66 \\
+          cos(ta) + sl + log(sl) + W + T + HI & 90103.28 & 35.13 & 0.00 & -45038.64 \\
+          cos(ta) + sl + log(sl) + W + T & 90109.40 & 41.25 & 0.00 & -45043.70 \\
+          cos(ta) + sl + log(sl) + W + T + S & 90110.35 & 42.20 & 0.00 & -45042.17 \\
+          cos(ta) + sl + log(sl) + W + T + DTR & 90112.55 & 44.40 & 0.00 & -45043.27 \\
+          cos(ta) + sl + log(sl) + W + T + P & 90113.11 & 44.96 & 0.00 & -45043.56 \\
+          cos(ta) + sl + log(sl) + W + T + RC & 90113.60 & 45.45 & 0.00 & -45041.80 \\
+          cos(ta) + sl + log(sl) + W + DTW & 90118.55 & 50.40 & 0.00 & -45048.28 \\
+          cos(ta) + sl + log(sl) + W + HI & 90128.70 & 60.54 & 0.00 & -45053.35 \\
+          cos(ta) + sl + log(sl) + W + S & 90132.22 & 64.06 & 0.00 & -45055.11 \\
+          cos(ta) + sl + log(sl) + W & 90134.85 & 66.69 & 0.00 & -45058.42 \\
+          cos(ta) + sl + log(sl) + W + DTR & 90138.31 & 70.16 & 0.00 & -45058.16 \\
+          cos(ta) + sl + log(sl) + W + P & 90138.50 & 70.35 & 0.00 & -45058.25 \\
+          cos(ta) + sl + log(sl) + W + RC & 90139.30 & 71.15 & 0.00 & -45056.65 \\
+          cos(ta) + sl + log(sl) + S & 90141.98 & 73.83 & 0.00 & -45061.99 \\
+          cos(ta) + sl + log(sl) + DTW & 90225.64 & 157.49 & 0.00 & -45103.82 \\
+          cos(ta) + sl + log(sl) + T & 90271.73 & 203.58 & 0.00 & -45126.86 \\
+          cos(ta) + sl + log(sl) + HI & 90273.18 & 205.02 & 0.00 & -45127.59 \\
+          cos(ta) + sl + log(sl) + P & 90285.24 & 217.08 & 0.00 & -45133.62 \\
+          cos(ta) + sl + log(sl) + DTR & 90285.33 & 217.18 & 0.00 & -45133.67 \\
+          cos(ta) + sl + log(sl) + RC & - & - & - & - \\
+          cos(ta) + sl + log(sl) + W + T + DTW + RC & - & - & - & - \\
          \bottomrule
@@
          Shrubs/Grassland, T = Trees, P = Protected, HI = Human Influence, RC =
-         Road Crossing, DTR = Distance To Roads.
+         Road Crossing, DTR = Distance To Roads. The two models at the bottom
+         failed to converge, which is why no AIC value could be obtained.
        \end{tablenotes}

### 2020-09-02 Chapter_0/02_R-Scripts/02_Analysis/02_SSF(HabitatSelectionModel).R

@@ -496,6 +496,84 @@ models <- subset(models, Weight > 0)
 # Store the result for later
 write_rds(models, "03_Data/03_Results/99_ModelSelection.rds")
 
+##############################################################
+#### Examine Random Effects
+##############################################################
+# Reload model
+mod <- read_rds("03_Data/03_Results/99_ModelSelection.rds")
+mod <- models$Model[[1]]
+
+# Check out coefficients per individual
+coeffs <- coef(mod)
+ranefs <- ranef(mod, condVar = T)
+coeffs$cond$id
+ranefs$cond$id
+
+# Note: ranef yields the difference between the individual specific effect and
+# the mean level effect. coef, on the other hand, yields the individual specific
+# effect. Thus, the followin two lines yield (approximately) the same
+mean(coeffs$cond$id$cos_ta_) + ranefs$cond$id$cos_ta_[3]
+
+# We now want to visualize the individual variation. There are two possibilities
+# for this: lme4::dotplot() or a ggplot. The dotplot is easier, yet not
+# customizable. Let's first do the dotplot, then recreate it in ggplot.
+lme4:::dotplot.ranef.mer(ranef(mod)$cond)
+
+# Maybe scalefree?
+lme4:::dotplot.ranef.mer(ranef(mod)$cond, scales = list(x = list(relation = "free")))
+
+# Prepare dataframe that we need to plot the same in ggplot
+rfs <- ranefs$cond$id %>%
+  rownames_to_column() %>%
+  gather(key = Covariate, value = Mean, 2:9)
+names(rfs)[1] <- "id"
+
+# We need to add the conditional variance
+condVar <- attributes(ranefs$cond$id)$condVar
+names(condVar) <- attributes(ranefs$cond$id)$names
+condVar <- as.data.frame(do.call(rbind, condVar))
+names(condVar) <- attributes(ranefs$cond$id)$row.names
+condVar <- rownames_to_column(condVar)
+names(condVar)[1] <- "Covariate"
+condVar <- gather(condVar, key = id, value = Variance, 2:17)
+
+# Join data to rfs dataframe
+rfs <- left_join(rfs, condVar)
+
+# Rename stuff nicely
+rfs$Covariate <- gsub(rfs$Covariate, pattern = "cos_ta_", replacement = "cos(ta)")
+rfs$Covariate <- gsub(rfs$Covariate, pattern = "log_sl_", replacement = "log(sl)")
[…]

### 2020-09-02 Chapter_0/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript.r

@@ -1408,7 +1408,7 @@ aics$Covariates <- aics$Covariates %>%
   gsub(pattern = "\\bDistanceToRoads\\b", replacement = "DTR") %>%
   gsub(pattern = "\\bRoadCrossing\\b", replacement = "RC") %>%
   gsub(pattern = ",", replacement = " +") %>%
-  paste("cos(ta) + log(sl) +", .)
+  paste("cos(ta) + sl + log(sl) +", .)
 
 # We will also add cos(ta) and log(sl) as covariates
 aics
@@ -1446,12 +1446,13 @@ best_table <- getCoeffs(best, pvalue = TRUE) %>%
 
 # Rename Covariates
 best_table$Covariate[best_table$Covariate == "Shrubs"] <- "Shrubs/Grassland"
-best_table$Covariate[best_table$Covariate == "cos(ta_)"] <- "cos(ta)"
-best_table$Covariate[best_table$Covariate == "log(sl_)"] <- "log(sl)"
+best_table$Covariate[best_table$Covariate == "cos_ta_"] <- "cos(ta)"
+best_table$Covariate[best_table$Covariate == "sl_"] <- "sl"
+best_table$Covariate[best_table$Covariate == "log_sl_"] <- "log(sl)"
 best_table$Covariate[best_table$Covariate == "HumansBuff5000"] <- "HumanInfluence"
 
 # Print the result as a table to a tex file
-options(scipen=999)
+options(scipen = 999)
 print(xtable(best_table)
   , floating            = FALSE
   , latex.environments  = NULL
@@ -1464,8 +1465,9 @@ print(xtable(best_table)
 # Prepare and store a plot of the coefficients
 coeffs <- getCoeffs(best)[-1, ]
 coeffs$Covariate[coeffs$Covariate == "Shrubs"] <- "Shrubs/Grassland"
-coeffs$Covariate[coeffs$Covariate == "cos(ta_)"] <- "cos(ta)"
-coeffs$Covariate[coeffs$Covariate == "log(sl_)"] <- "log(sl)"
+coeffs$Covariate[coeffs$Covariate == "cos_ta_"] <- "cos(ta)"
+coeffs$Covariate[coeffs$Covariate == "sl_"] <- "sl"
+coeffs$Covariate[coeffs$Covariate == "log_sl_"] <- "log(sl)"
[…]

### 2020-09-06 Chapter_0/04_Manuscript/Manuscript.tex
Worked on revision

@@
 
-\noindent 4. Differences in permeability were mainly caused by different degrees
-of human activities across regions, which hampered dispersal. Rivers, swamps  or
-open water also limited dispersal, while other landscape features had a limited
-effect.
+\noindent 4. Differences in permeability were mainly related to different
+degrees of human activities across regions, as well as to the presence of
+rivers, swamps or open water. The relation between other landscape features and
+landscape permeability was less evident.\todo{Editor indicated that the
+relationships are correlative and not necessarily causal.}
 
@@
 Berlin, Germany}) while still with their natal pack. All procedures were
-undertaken and supervised by a Botswana-registered wildlife veterinarian. Fully
-assembled collars produced a tag weight of approximately 330 g, accounting for
-approximately 1.2\% of a wild dog's average body weight. Collars were mounted
-using leather belts and included a drop off mechanism, triggered by a slowly
-decomposing piece of cloth.\todo{Editor desired more information on GPS collar}
-During dispersal, GPS collars were programmed to record GPS relocation data
-every 4 hours and to regularly transmit them via Iridium satellite system to a
-base station.
+undertaken and supervised by a Botswana-registered wildlife veterinarian.
+Collars produced a tag weight of 1.2\% of a wild dog's average body weight,
+which is well within published recommendations \citep{Kenward.2000}. Collars
+were attached through a leather belt that included a drop off mechanism,
+triggered by a slowly decomposing piece of cloth.\todo{Editor desired more
+information on GPS collar} During dispersal, GPS collars were programmed to
+record GPS relocation data every 4 hours and to regularly transmit them via
+Iridium satellite system to a base station.
 
@@
 selection strength \todo{As per reviewer 1} for these covariates (\(\beta_1,
-\beta_2, ..., \beta_n\)). To estimate relative selection strengths \todo{As per
-reviewer 1} (i.e. the \(\beta\)'s), we used mixed effects conditional logistic
-regression analysis as suggested by \cite{Muff.2020}. We implemented their
-method using the R-package \textit{glmmTMB} \citep{Mollie.2017} and used
-dispersing coalition ID to model random slopes. We also modelled random
-intercepts with an arbitrary high variance of 10\textsuperscript{6} to make use
-of the poisson trick (see \citealp{Muff.2020}). We defined three movement
-metrics, namely the cosine of the turning angle (\(cos(ta)\)), the step length
-(\(sl\)) and the natural logarithm of the step length (\(log(sl)\)), as core
-covariates and ran stepwise forward model selection based on Akaike's
-Information Criterion (AIC; \citealp{Burnham.2002}) for all other
-covariates.\todo{Reviewer 1 asked for more details in the random effects. I
-therefore decided to update this section too} The inclusion of movement metrics
-served to reduce biases in estimated habitat selection coefficients \todo{as per
-revierwer 1} that may have arisen due to movement behavior \citep{Avgar.2016}.
-To validate the predictive power of the most parsimonious habitat selection
-model, we ran k-fold cross-validation for case-control studies as described in
-\cite{Fortin.2009} (details in Appendix A.5).
+\beta_2, ..., \beta_n\)). To estimate relative selection strength \todo{As per
+reviewer 1} (i.e. the \(\beta\)'s) for each covariate, we used mixed effects
+conditional logistic regression analysis as suggested by \cite{Muff.2020}. We
+implemented their method using the R-package \textit{glmmTMB}
+\citep{Mollie.2017} and used dispersing coalition ID to model random slopes. We
+also modelled random intercepts with an arbitrary high variance of
+10\textsuperscript{6} to make use of the poisson trick (see
+\citealp{Muff.2020}). We defined three movement metrics, namely the cosine of
+the turning angle (\(cos(ta)\)), the step length (\(sl\)) and the natural
+logarithm of the step length (\(log(sl)\)), as core covariates and ran stepwise
+forward model selection based on Akaike's Information Criterion (AIC;
+\citealp{Burnham.2002}) for all other covariates.\todo{Reviewer 1 asked for more
+details in the random effects. I therefore decided to update this section too}
+The inclusion of movement metrics served to reduce biases in estimated habitat
+selection coefficients \todo{as per revierwer 1} that may have arisen due to
+movement behavior \citep{Avgar.2016}. To validate the predictive power of the
+most parsimonious habitat selection model, we ran k-fold cross-validation for
+case-control studies as described in \cite{Fortin.2009} (details in Appendix
+A.5).
 

### 2020-09-10 Chapter_0/04_Manuscript/Manuscript.tex
Worked on revisions

@@
 \noindent 1. Conserving and managing large portions of land to connect wildlife
-reserves is increasingly used to maintain and restore connectivity among
-wildlife populations. Boundaries of such conservation areas are often determined
-based on expert opinion and socio-political constraints, yet the extent to which
-they match species' movement corridors is rarely examined. This is mainly due to
-a lack of data, particularly on wide-ranging movement behavior such as
-dispersal. Nevertheless, empirically assessing the adequacy of protected areas
-is key for the implementation of targeted management actions and efficient use
-of limited conservation funds.
+reserves is an increasingly used strategy to maintain and restore connectivity
+among wildlife populations. Boundaries of such conservation areas are often
+determined based on expert opinion and socio-political constraints, yet the
+extent to which they match species' movement corridors is rarely examined. This
+is mainly due to a lack of data, particularly on wide-ranging movement behavior
+such as dispersal. Nevertheless, empirically assessing the adequacy of protected
+areas is key for the implementation of targeted management actions and efficient
+use of limited conservation funds.
 
@@
 example for international conservation efforts. We used integrated step
-selection analysis to estimate relative selection strenghts\todo{as per reviewer
-1} of dispersers and to create a permeability surface for the entire KAZA-TFCA.
-We compared landscape permeability across different regions within the KAZA-TFCA
-as well as outside its boundaries. Lastly, we calculated least-cost paths and
-corridors to verify that major movement routes were adequately encompassed
-within the KAZA-TFCA.
+selection analysis to estimate habitat selection of dispersers and to create a
+permeability surface for the entire KAZA-TFCA. We compared landscape
+permeability across different regions within the KAZA-TFCA as well as outside
+its boundaries. Lastly, we calculated least-cost paths and corridors to verify
+that major movement routes were adequately encompassed within the KAZA-TFCA.
 
@@
 double compared to areas outside it. Furthermore, we observed a five-fold
-permeability difference among the five KAZA-TFCA countries. We further showed
+permeability difference among the five KAZA-TFCA countries. We also showed
 that major movement corridors of wild dogs run within the KAZA-TFCA, although
@@
 \noindent 4. Differences in permeability were mainly related to different
-degrees of human activities across regions, as well as to the presence of
-rivers, swamps or open water. The relation between other landscape features and
-landscape permeability was less evident.\todo{Editor indicated that the
+degrees of human activities across regions, and to the presence or absence of
+rivers, swamps or open water. The relation between permeability and other
+landscape features was less evident.\todo{Editor indicated that the
 relationships are correlative and not necessarily causal.}
@@
 evaluate the adequacy of already-existing or planned protected areas.
-Furthermore, observed regional differences in landscape permeability highlight
-the need for a coordinated effort towards maintaining or restoring connectivity,
-especially where transboundary dispersal occurs.\todo{According to suggestions
-by reviewer 2}
+Furthermore, we observed regional differences in landscape permeability that
+highlight the need for a coordinated effort towards maintaining or restoring
+connectivity, especially where transboundary dispersal occurs.\todo{According to
+suggestions of reviewer 2}
 \end{abstract}
@@
 \todo{Reviewer one pointed out an awkward use of the word utmost. I have changed
-it according to his suggestion.} importance for conservation management
-\citep{Doerr.2011, Rudnick.2012}, resulting in an ever-growing number of large
-and often transboundary protected areas. While boundaries of such areas are
-often drawn according to expert opinion and socio-political needs, subjective
-assessments have revealed deficiencies in the past \citep{Clevenger.2002,
-Pullinger.2010}. Thus, an empirical assessment of the adequacy of
-already-existing or planned protected areas using pertinent animal movement data
-is paramount for targeted use of valuable and scarce conservation funds
-\citep{Pullinger.2010}.
+it according to his suggestion.} importance \citep{Doerr.2011, Rudnick.2012},
+resulting in an ever-growing number of large and often transboundary protected
+areas. While boundaries of such areas are often drawn according to expert
+opinion and socio-political needs, subjective assessments have revealed
+deficiencies in the past \citep{Clevenger.2002, Pullinger.2010}. Thus, an
+empirical evaluation of the adequacy of already-existing or planned protected
+areas using pertinent animal movement data is paramount for targeted use of
+valuable and scarce conservation funds \citep{Pullinger.2010}.
 
@@
 identify movement corridors and assess connectivity at large scales (e.g.
-\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). Identification of
-potential corridors typically relies on the estimation of permeability surfaces,
-which return the ease or willingness at which the focal species traverses a
-specific landscape \citep{Sawyer.2011}. Such surfaces are created based on a
-species' relative selection strengths\todo{reviewer 1}, which can be quantified
-using a suite of selection functions \citep{Zeller.2012}. Specifically, relative
-selection strenghts \todo{reviewer 1} are estimated by comparing spatial
-covariates (e.g. environmental and anthropogenic) at locations visited by the
-animal to the same spatial covariates at locations avilable to, yet unused by
-\todo{Reviewer 1 pointed out correctly that we should talk about available
-locations, rather than randomly selected locations} the animal
+\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). The use of relocation
+data has shifted the focus from a structural to a functional view on
+connectivity, taking into account how species respond to environmental features
+\citep{Taylor.2006, Fattebert.2015}.\todo{As per reviewer 2} Identification of
+potential movement corridors typically relies on the estimation of permeability
+surfaces, which return the ease or willingness at which the focal species
+traverses a specific landscape \citep{Sawyer.2011}. Such surfaces are created
+based on species' relative selection strengths \citep{Avgar.2017}\todo{reviewer
+1}, which can be quantified using a suite of selection functions
+\citep{Zeller.2012}. Specifically, selection strengths \todo{reviewer 1} are
+estimated by comparing spatial covariates (e.g. environmental and anthropogenic)
+at locations visited by the animal to the same spatial covariates at locations
+avilable to the animal \todo{Reviewer 1 pointed out correctly that we should
+talk about available locations, rather than randomly selected locations}
 \citep{Zeller.2012}. Importantly, selection functions rely on adequate landscape
 and relocation data that are representative of the process being studied
-\citep{Diniz.2020}. For instance, relocation data collected on dispersing
-individuals has been shown to outperform data collected on resident individuals
-in the detection of large-scale movement corridors \citep{Elliot.2014,
-Diniz.2020}. Nevertheless, dispersal data is inherently difficult to collect and
-remains scarce in the connectivity literature \citep{Vasudev.2015}. As such,
-most permeability surfaces upon which movement corridors are identified are
-created using relocation data collected on resident individuals. While some
-species use the same habitats during residence and dispersal
-\citep{Fattebert.2015}, others depict vastly different habitat selection
-depending on the movement mode \citep{Elliot.2014, Abrahms.2017}. Thus, the use
-of data collected during residency can introduce severe biases in estimated
-landscape permeability and has likely limited our ability to meaningfully assess
-the effectiveness of protected areas in securing connectivity for their
-protected species.\todo{Reviewer 2 desired to change from: this introduces - to:
-this can introduce}
+\citep{Diniz.2020}. Altough selection during residence and dispersal may
+coincide \citep{Fattebert.2015},\todo{second reviewer pointed out that space use
+can be alike} it appears that relocation data collected on dispersing
+individuals outperforms data collected on resident individuals in the detection
+of large-scale movement corridors \citep{Elliot.2014, Diniz.2020}. Nevertheless,
+dispersal data is inherently difficult to collect and remains scarce in the
+connectivity literature \citep{Vasudev.2015}. As such, most permeability
+surfaces upon which movement corridors are identified are created using
+relocation data collected on resident individuals. This has likely limited our
+ability to meaningfully assess the effectiveness of protected areas in securing
+functional connectivity for their protected species.\todo{Reviewer 2 desired to
+change from: this introduces - to: this can introduce}
 
@@
 positioning system (GPS) relocation data of its protected species at large
-spatial scales \citep{Elliot.2014, Tshipa.2017}. Thus, how well the boundaries
-of the KAZA-TFCA reflect natural movement patterns and dispersal corridors of
-its most mobile protected species is virtually unknown.
+spatial scales \citep{Elliot.2014, Tshipa.2017}.\todo{as per editor} Thus, how
+well the boundaries of the KAZA-TFCA reflect natural movement patterns and
+dispersal corridors of its most mobile protected species is virtually unknown.
 
@@
 \citep{McNutt.1996, Cozzi.2020, Behr.2020}. During dispersal, wild dogs can
-cover several hundred kilometers \citep{Masenga.2016, Woodroffe.2019,
-Cozzi.2020}. Despite the importance of dispersal for the long-term viability of
-this species, little empirical information is available on habitat selection and
-potential movement barriers during dispersal. The few studies that have
-collected dispersal data have shown that dispersers quickly move over large
-distances, avoid human-dominated landscapes and areas densely covered by trees,
-but prefer proximity to water \citep{Masenga.2016, Woodroffe.2019, Oneill.2020,
-Cozzi.2020}.
+cover several hundred kilometers and cross international borders
+\citep{Masenga.2016, Woodroffe.2019, Cozzi.2020}. Despite the importance of
+dispersal for the long-term viability of this species, little empirical
+information is available on habitat selection and potential movement barriers
+during dispersal. The few studies that have collected dispersal data have shown
+that dispersers quickly move over large distances, avoid human-dominated
+landscapes and areas densely covered by trees, but prefer proximity to water
+\citep{Masenga.2016, Woodroffe.2019, Oneill.2020, Cozzi.2020}.
 
-Here, we collected GPS relocation data on 16 dispersing wild dogs in as many
-dispersing coalitions from a free-ranging population in northern Botswana and
-analyzed it to assess the adequacy of the KAZA-TFCA in securing connectivity. We
-estimated the relative selection strengths\todo{reviewer 1} towards or agains
-environmental and anthropogenic landscape features and used the obtained habitat
-coefficients to predict a permeability surface spanning the entire KAZA-TFCA. We
-then investigated how landscape permeability varies regionally and
-internationally and compared permeability within and outside the KAZA-TFCA
-boundaries. Finally, we calculated least-cost paths and corridors to identify
-major movement routes and to verify that these are successfully covered by the
-KAZA-TFCA.
+Here, we collected and analyzed GPS relocation data on 16 dispersing wild dogs
+in as many dispersing coalitions from a free-ranging population in northern
+Botswana to assess the adequacy of the KAZA-TFCA in securing connectivity. We
+estimated relative selection strengths\todo{reviewer 1} towards environmental
+and anthropogenic landscape features, and used the obtained coefficients to
+predict a permeability surface spanning the entire KAZA-TFCA. We then
+investigated how landscape permeability varied regionally and internationally,
+and we compared permeability within and outside the KAZA-TFCA boundaries.
+Finally, we calculated least-cost paths and corridors to identify major movement
+routes and to verify that these were successfully covered by the KAZA-TFCA.
 
@@
 Wolski.2017}. The extent of the flood in the delta greatly changes within and
-between years depending on the amount of rain that descends from the catchment
+between years, depending on the amount of rain that descends from the catchment
 areas in Angola and reaches the distal ends of the delta between July and August
@@
 and other protected areas, there is considerable human influence in some regions
-of the KAZA-TFCA, mainly originating from farms, high human density, and road
+of the KAZA-TFCA, mainly originating from farms, human density, and road
 traffic.
@@
 identified based on age, number of same‐sex siblings, pack size, and presence of
-unrelated individuals of the opposite sex in ther pack citep \citep{McNutt.1996,
-Behr.2020}\todo{We originally only pointed to Behr and did not include the
-criteria. The editor desired to include those}. Individuals were immobilized
-according to protocols described in \cite{Osofsky.1996}, and fitted with
-GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
-Berlin, Germany}) while still with their natal pack. All procedures were
-undertaken and supervised by a Botswana-registered wildlife veterinarian.
-Collars produced a tag weight of 1.2\% of a wild dog's average body weight,
-which is well within published recommendations \citep{Kenward.2000}. Collars
-were attached through a leather belt that included a drop off mechanism,
-triggered by a slowly decomposing piece of cloth.\todo{Editor desired more
-information on GPS collar} During dispersal, GPS collars were programmed to
-record GPS relocation data every 4 hours and to regularly transmit them via
-Iridium satellite system to a base station.
+unrelated individuals of the opposite sex in their pack \citep{McNutt.1996,
+Behr.2020}\todo{The editor desired to include criteria}. Selected individuals
+were immobilized according to protocols described in \cite{Osofsky.1996}, and
+fitted with GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic
+Aerospace GmbH, Berlin, Germany}) while still with their natal pack. All
+required procedures were undertaken and supervised by a Botswana-registered
+wildlife veterinarian. Collars produced a tag weight of 300 gramms, accounting
+for 1.2\% of a wild dog's average body weight. They were attached using leather
+belts that included a drop off mechanism, triggered by a decomposing piece of
+cloth.\todo{Editor desired more information on GPS collar} During dispersal, GPS
+collars were programmed to record a GPS relocation every 4 hours and to
+regularly transmit data via Iridium satellite system to a base station.
 
@@
 started when a coalition had left its natal home range and continued until the
-NSD metric remained stationary, implying that the coalition had successfully
+NSD metric remained stationary, indicating that the coalition had successfully
 settled (Figure S1). In our analysis, we did not differentiate between male and
@@
 \textit{protection status}, and \textit{human influence}. Land cover comprised
-of the covariates water cover, distance to water, percentage cover by
+the covariates water cover, distance to water, percentage cover by
 shrubs/grassland, and percentage cover by trees. To capture the pulsing behavior
-of the Okavango Delta we frequently updated water cover layers and corresponding
-distance to water layers. Protection status contained a binary indicator of
-whether an area was protected or not. Human influence included covariates
-rendering human density, the presence of roads, and the distance to
-roads.\todo{Reviewer 2 found the description of the covariates unclear} For each
-of these covariates we prepared spatial raster layers from freely available
-online services or from remotely sensed satellite imagery. To ensure a
-consistent resolution (i.e. cell-size or grain) across covariates, we coarsened
-or interpolated all layers to match a resolution of 250m x 250m. For further
-details on the preparation of each covariate, see Appendix A.3. We performed
-processing and manipulation of data as well as all spatial and statistical
-analyses using R, version 3.6.1 \citep{R.2019}.
+of the Okavango Delta, we frequently updated layers for water cover and
+corresponding layers depicting distance to water. Protection status contained a
+binary indicator of whether an area was protected or not. Human influence
+included covariates rendering human density, the presence of roads, and the
+distance to roads.\todo{Reviewer 2 found the description of the covariates
+unclear} We prepared all covariates as spatial raster layers from freely
+available online services and from remotely sensed satellite imagery. To ensure
+a consistent resolution (i.e. cell-size or grain) across covariates, we
+coarsened or interpolated all layers to a resolution of 250m x 250m. For further
+details on the preparation of each covariate and their sources, see Appendix
+A.3. We performed processing and manipulation of data as well as all spatial and
+statistical analyses using R, version 3.6.1 \citep{R.2019}.
 
@@
 We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
-investigate dispersers' relative selection strentgh \todo{Reviewer 1 suggested
-to use relative selection strength, rather than preference} for the
+investigate dispersers' relative selection strentghs towards \todo{Reviewer 1
+suggested to use relative selection strength, rather than preference} the
 above-mentioned spatial covariates. That is, we paired each realized step (i.e.
 the connecting line between two consecutive GPS relocations;
-\citealp{Turchin.1998}) with 24 random steps. We generated random steps by
-sampling turning angles from a uniform distribution U(\(-\pi, +\pi\)) and step
-lengths from a gamma distribution that was fitted using realized steps
-\citep{Avgar.2016}. A realized step and its 24 associated random steps formed a
-stratum that received a unique identifier. Along each step, we extracted the
-above-mentioned covariates (Table S3), standardized extracted values using a
-z-score transformation, and checked for correlation using Pearson's Correlation
-Coefficient \(r\). None of the covariates were overly correlated (\(|r| > 0.6\);
-\citealp{Latham.2011}) and we retained all of them for modeling. Our habitat
-selection model then assumed that dispersing wild dogs assigned a selection
-score \(w(x)\) of the following exponential form to each realized and random
-step \citep{Fortin.2005}:
+\citealp{Turchin.1998}) with 24 random steps that were generated by sampling
+turning angles from a uniform distribution U(\(-\pi, +\pi\)) and step lengths
+from a gamma distribution fitted to realized steps \citep{Avgar.2016}. A
+realized step and its 24 associated random steps formed a stratum and received a
+unique identifier. Along each step, we extracted the above-mentioned covariates
+(Table S3), standardized extracted values using a z-score transformation, and
+checked for correlation using Pearson's Correlation Coefficient \(r\). None of
+the covariates were overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and
+we retained all of them for modeling. Our habitat selection model then assumed
+that dispersing wild dogs assigned a selection score \(w(x)\) of the following
+exponential form to each realized and random step \citep{Fortin.2005}:
 
@@
 covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's relative
-selection strength \todo{As per reviewer 1} for these covariates (\(\beta_1,
-\beta_2, ..., \beta_n\)). To estimate relative selection strength \todo{As per
-reviewer 1} (i.e. the \(\beta\)'s) for each covariate, we used mixed effects
-conditional logistic regression analysis as suggested by \cite{Muff.2020}. We
-implemented their method using the R-package \textit{glmmTMB}
-\citep{Mollie.2017} and used dispersing coalition ID to model random slopes. We
-also modelled random intercepts with an arbitrary high variance of
-10\textsuperscript{6} to make use of the poisson trick (see
+selection strengths \todo{As per reviewer 1} towards these covariates
+(\(\beta_1, \beta_2, ..., \beta_n\)). To estimate relative selection strengths
+\todo{As per reviewer 1} (i.e. the \(\beta\)'s) for each covariate, we used
+mixed effects conditional logistic regression analysis as suggested by
+\cite{Muff.2020}. We implemented their method using the R-package
+\textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID to model
+random slopes. We also modelled random intercepts with an arbitrary high
+variance of 10\textsuperscript{6} to make use of the poisson trick (see
 \citealp{Muff.2020}). We defined three movement metrics, namely the cosine of
@@
 water in at least 10\% of the cases. Using the resulting static map we also
-calculated a layer returning the distance to water. To reduce the influence of
-outliers in predicted permeability scores we followed \cite{Squires.2013} and
+calculated a layer returning the distance to water. Because the extent of the
+delta covers merely 6\% of the entire KAZA-TFCA, the use of static maps should
+introduce little bias in predicted permeability. To reduce the influence of
+outliers in predicted permeability scores, we followed \cite{Squires.2013} and
 curtailed predicted scores between the 1\textsuperscript{st} and
@@
 permeability across different regions, we normalized the permeability surface to
-a range between 0 (most impermeable) and 1 (most permeable). We then determined
+a range between 0 (most impermeable) and 1 (most permeable), and we determined
 median permeability within and outside the KAZA-TFCA, within and outside
@@
 km\textsuperscript{2}, which conforms with home-range requirements of African
-wild dogs reported in \cite{Pomilia.2015}. Finally, we defined centroids as
-source points for those protected areas \(>\) 700 km\textsuperscript{2} that
-were not assigned any source points from the regular grid. In total, we
-generated 68 source points, which resulted in 2'278 unique pairwise combinations
-and therefore 2'278 unique LCPs and LCCs. We computed factorial LCPs and LCCs
-between source points using the R-package \textit{gdistance} (for further
-details see Appendix A.7). After computation, we tallied overlapping LCPs and
-LCCs, respectively, into single connectivity maps.
+wild dogs \citep{Pomilia.2015}. Finally, we defined centroids as source points
+for those protected areas \(>\) 700 km\textsuperscript{2} that were not assigned
+any source points from the regular grid. Because wild dogs residing outside of
+protected areas are rare and likely unviable \citep{VanDerMeer.2014}, we
+consider our selection of source points within protected areas to be
+appropriate. In total, we generated 68 source points, which resulted in 2'278
+unique pairwise combinations and therefore 2'278 unique LCPs and LCCs. We
+computed factorial LCPs and LCCs between source points using the R-package
+\textit{gdistance} (further details in Appendix A.7). After computation, we
+tallied overlapping LCPs and LCCs, respectively, into single connectivity maps.
 
@@
 Table S1), resulting in an average of 261 (\(SD\) = 207) locations per
-dispersing coalition. Coalitions on average dispersed for 48 days (\(SD\) =
-44), covered a mean euclidean distance of 54 km (\(SD\) = 71) and a cumulative
+dispersing coalition. Coalitions on average dispersed for 48 days (\(SD\) = 44),
+covered a mean Euclidean distance of 54 km (\(SD\) = 71) and a cumulative
 distance of 597 km (\(SD\) = 508). One female coalition dispersed far east into
@@
 \textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
-\textit{human influence}, beside the fixed covariates \textit{cos(ta)} and
-\textit{log(sl)} (\Cref{PermeabilityResults}a). The positive coefficient for
-\(log(sl)\) (\(\beta\) = 0.06, 95\% CI = 0.02 to 0.09) showed that our initial
-(selection-biased) gamma distribution produced step-lengths slightly shorter
-than a selection-free gamma distribution would. Similarly, the positive
-coefficient for \(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21) indicates
+\textit{human influence}, beside the fixed covariates \textit{cos(ta)},
+\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). The positive
+coefficients for \(sl\) (\(\beta\) = 0.03, 95\% CI = -0.09 to 0.15) and
+\(log(sl)\) (\(\beta\) = 0.12, 95\% CI = 0.02 to 0.22) indicated that our
+initial (selection-biased) gamma distribution produced step-lengths slightly
+shorter than a selection-free gamma distribution would. Similarly, the positive
+coefficient for \(cos(ta)\) (\(\beta\) = 0.10; 95\% CI = 0.05 to 0.15) suggested
 that our uniform distribution for turning angles produced turning angles with
@@
 selection-biased} With respect to environmental covariates, dispersers avoided
-moving through water (\(\beta\) = -0.52, 95\% CI -0.77 to -0.26) but selected
+moving through water (\(\beta\) = -0.53, 95\% CI -0.79 to -0.27) but selected
 for locations in its vicinity, although the latter effect was not significant
-(\(\beta\) = -0.32, 95\% CI = -0.72 to 0.08). Dispersers avoided areas that were
-densely covered by trees (\(\beta\) = -0.31, CI = -0.46 to -0.15) and preferred
+(\(\beta\) = -0.33, 95\% CI = -0.73 to 0.08). Dispersers avoided areas that were
+densely covered by trees (\(\beta\) = -0.31, CI = -0.47 to -0.15) and preferred
 areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42).
 Finally, dispersers avoided areas that were influenced by humans (\(\beta\) =
--0.41, 95\% CI = -0.78 to -0.05). With the exepction of \textit{DistanceToWater}
-(\(SD_{RE}\) = 0.57), random effects revealed little variability in different
-dispersal coalitions' selection strengths (i.e. \(SD_{RE} \leq 0.22\), Figure
-S7).\todo{Reviewer 1 desired more information on the random effects}
+-0.45, 95\% CI = -0.82 to -0.08). With the exepction of \textit{distance to
+water} (\(SD_{RE}\) = 0.57), random effects revealed little variability between
+different dispersal coalitions' selection strengths (i.e. all with \(SD_{RE}
+\leq 0.22\), Figure S7).\todo{Reviewer 1 desired more information on the random
+effects}
 
@@
     covariate, positive coefficients selection of a covariate. Whiskers
-    delineate the 95\%-CIs for estimated parameters. (b) Results from the k-fold
-    cross validation for case-control studies. The left graph shows rank
-    frequencies of \textit{realized} steps according to predictions, whereas the
-    right graph shows rank frequencies of \textit{randomly selected} steps
-    according to predictions. \(\bar{r}_s\) indicates the mean correlation
-    coefficient resulting from 100 repetitions of the k-fold cross validation.
-    The blue smoothing line was fitted using a locally weighted polynomial
-    regression and serves to aid the eye in detecting the trends. Correlation
-    coefficients suggest that our prediction was significant and robust,
-    evidenced by the fact that the confidence intervals of \(\bar{r}_{s,
-    realized}\) and \(\bar{r}_{s, random}\) did overlap and by the fact that
-    there was strong and significant correlation between ranks and associated
-    frequency for realized steps.}
+    delineate the 95\%-CIs for estimated parameters. Significance codes: ** \(p
+    < 0.05\), *** \(p < 0.01\).(b) Results from the k-fold cross validation for
+    case-control studies. The left graph shows rank frequencies of
+    \textit{realized} steps according to predictions, whereas the right graph
+    shows rank frequencies of \textit{randomly selected} steps according to
+    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
+    resulting from 100 repetitions of the k-fold cross validation. The blue
+    smoothing line was fitted using a locally weighted polynomial regression and
+    serves to aid the eye in detecting the trends. Correlation coefficients
+    suggest that our prediction was significant and robust, evidenced by the
+    fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
+    \(\bar{r}_{s, random}\) did overlap and by the fact that there was strong
+    and significant correlation between ranks and associated frequency for
+    realized steps.}
     \label{PermeabilityResults}
@@
 permeability values (\Cref{PermeabilityComp}) showed that permeability inside
-the KAZA-TFCA is more than two times as high as permeability outside it.
-Permeability varies by country, with a five-fold permeability difference among
-them. Angola and Botswana are characterized by comparably highly permeable
-landscapes, Zimbabwe and Zambia are relatively impermeable, and Namibia ranges
+the KAZA-TFCA was more than two times as high as permeability outside it.
+Permeability varied by country, with a five-fold permeability difference among
+them. Angola and Botswana were characterized by comparably highly permeable
+landscapes, Zimbabwe and Zambia were relatively impermeable, and Namibia ranged
 in between the two extremes (\Cref{PermeabilityComp}). Visual inspection of our
-covariate layers indicated that high permeability in Angola and Botswana is
-mainly caused by a combination of low human influences, low tree cover, high
+covariate layers indicated that high permeability in Angola and Botswana was
+mainly related to a combination of low human influence, low tree cover, high
 shrubs/grassland cover, and a close distance to water. Although swamps,
-wetlands, and permanent water themselves provide little permeability, their
-surroundings act as strong attractants to dispersers. The low permeability that
-characterizes Zambia and Zimbabwe, on the other hand, is mainly caused by
-substantial human influences. Albeit the KAZA-TFCA covers most permeability
-hot-spots, several highly permeable regions remain uncovered by its borders.
-Across all countries, protected areas provide roughly double the permeability of
-unprotected landscapes (\Cref{PermeabilityComp}).
+wetlands, and permanent water themselves provided little permeability, their
+surroundings acted as strong attractants to dispersers. The low permeability
+that characterized Zambia and Zimbabwe, on the other hand, was mainly caused by
+substantial human influence. Albeit the KAZA-TFCA covered most permeability
+hot-spots, several highly permeable regions remained uncovered by its borders.
+Across all countries, protected areas provided roughly double the permeability
+of unprotected landscapes (\Cref{PermeabilityComp}).
 
@@
     based on the raster cell's underlying covariates (\(x_i\)) and estimated
-    selection strenghts (\(\beta_i\)). Areas that dispersers find easy to
+    selection strength (\(\beta_i\)). Areas that dispersers find easy to
     traverse are depicted in bright colors. Bold white lines delineate the
@@
 were well-contained within the KAZA-TFCA boundaries (\Cref{LeastCost}). One
-major corridor runs SE-NW and connects the Okavango-Linyanti ecosystem in
-Botswana with Luengue-Luiana NP in Angola. A second corridor runs W-E between
-Chobe NP in Botswana and Zimbabwe's Hwange NP. A third major corridor runs
-NE-SW, completely across unprotected areas, and connects Kafue NP in Zambia with
-more central regions of the KAZA-TFCA. Several minor corridors branch off from
-these three major corridors; these include a southward connection between the
+major corridor ran SE-NW and connected the Okavango-Linyanti ecosystem in
+Botswana with Luengue-Luiana NP in Angola. A second corridor ran W-E between
+Chobe NP in Botswana and Zimbabwe's Hwange NP. A third major corridor ran NE-SW,
+completely across unprotected areas, and connected Kafue NP in Zambia with more
+central regions of the KAZA-TFCA. Several minor corridors branched off from
+these three major corridors; these included a southward connection between the
 Okavango-Linyanti and the Central Kalahari Game Reserve, a southwesterly
@@
 northeasterly extension of the Hwange corridor into Zimbabwe's Matusadona NP.
-According to our predictions, the landscapes in the Okavango-Linyanti region are
-the highest frequented dispersal routes within the KAZA-TFCA
+According to our predictions, the landscapes in the Okavango-Linyanti region
+were the highest frequented dispersal routes within the KAZA-TFCA
 (\Cref{LeastCost}b). Our model did not detect any significant direct corridors
@@
 points; that is, most paths and corridors connecting two adjacent peripheral
-points run through more central regions before heading towards their destination
+points ran through more central regions before heading towards their destination
 at the periphery (\Cref{LeastCost}).
@@
     \caption{(a) Source points (black dots) and corresponding least-cost paths
-    leaving from protected areas (light grey). Note that only contiguous
+    leaving from protected areas (dark grey borders). Note that only contiguous
     protected areas covering more than 700 km\textsuperscript{2} are depicted.
@@
 are likely to be important determinants of movement behavior, landscape
-permeability, and connectivity for other species \citep{Thurfjell.2014,
-Zeller.2012}. Interestingly, our predicted network of least cost-paths and
+permeability, and connectivity for other species \citep{Zeller.2012,
+Thurfjell.2014}. Interestingly, our predicted network of least cost-paths and
 corridors for African wild dogs shows surprising similarities to corridors of
@@
 movement corridors, thus highlighting areas that are exceptionally valuable for
-the conservation of several species.
+the conservation of several species (e.g. \citealp{Brennan.2020}).
 
@@
 and cross landscapes characterized by considerably unfavorable conditions
-\citep{Palomares.2000, Elliot.2014}. Nevertheless, successful conservation of
-this species relies on policymakers' and local authorities' willingness and
-ability to provide and conserve natural areas that remain free from
-anthropogenic pressures. This is not only paramount in light of increasing
-connectivity and facilitating dispersal, but also in terms of reducing
-human-caused mortality during dispersal. In fact, previous studies have shown
-that human-caused mortality represents a major threat to wild dogs' ability to
-disperse \citep{Woodroffe.2019, Cozzi.2020}.
+\citep{Palomares.2000, Elliot.2014}. Unfortunately, \cite{DaviesMostert.2012}
+only recorded net dispersal distances, thereby precluding a more indepth
+understanding of the impact of humans on dispersal behavior.\todo{As per
+reviewer 2} Nevertheless, successful conservation of this species relies on
+policymakers' and local authorities' willingness and ability to provide and
+conserve natural areas that remain free from anthropogenic pressures. This is
+not only paramount in light of increasing connectivity and facilitating
+dispersal, but also in terms of reducing human-caused mortality during
+dispersal. In fact, previous studies have shown that human-caused mortality
+represents a major threat to wild dogs' ability to disperse
+\citep{Woodroffe.2019, Cozzi.2020}.
 
-Besides human influences, we identified water as additional obstacle to
+Besides human influence, we identified water as additional obstacle to
 dispersal. This corroborates earlier studies showing that water bodies are
@@
 \citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to move
-into prey-poorer areas away from water. This could also explain the effect of
-distane to water was insignificant.\todo{Reviewer 1 suggested to add this} Given
-the influence that resident conspecifics, competitors, and prey can have on
-dispersers \citep{Cozzi.2018, Armansin.2019} future studies should strive to
-collect and incorporate intra- and interspecific relationships into analyses of
-landscape connectivity.
+into prey-poorer areas away from water \citep{Creel.2002, Mills.1997}. This
+reasoning may also explain why the effect of distance to water was
+insignificant.\todo{Reviewer 1 suggested to add this} Given the influence that
+resident conspecifics, competitors, and prey can have on dispersers
+\citep{Cozzi.2018, Armansin.2019} future studies should strive to collect and
+incorporate intra- and interspecific relationships into analyses of landscape
+connectivity.
 
-Overall, our findings on habitat selection of dispersing wild dogs coincide with
-findings from Kenya by \cite{Oneill.2020}, suggesting that there are strong
-commonalities across ecosystems. Hence, despite wild dogs' ability to cope with
-and adapt to a wide range of habitats, we feel that that the fundamental factors
-included in our study and their influences can be generalized to other
-populations.\todo{Editor desired that we discuss the applicability of our
-ecosystem to other ecosystems.}
+Overall, our findings on habitat selection during dispersal coincide with
+findings from dipsersing wild dogs in Kenya \citep{Oneill.2020}, suggesting that
+there are strong commonalities dispersers from these very different ecosystems.
+Thus, despite wild dogs' ability to cope with diverse habitats and adapt to
+changing conditions \citep{Woodroffe.2011}, the fundamental factors included in
+our study appear to influence wild dogs from other ecosystems alike.
+Nevertheless, expanding our analysis to dispersers emigrating from other source
+populations would invaluably contribute to our understanding of
+dispersal.\todo{Editor desired info on applicability}
 
@@
 (e.g. Okavango Delta, Linyanti Swamp) funnel dispersal movements, resulting in a
-highly frequented corridor. The key role of the Okavango-Linyanti region for
-overall connectivity within the KAZA-TFCA thus calls for actions to secure its
-protection status in the future. While the region is currently a Wildlife
-Management Area, it has neither the status of a National Park nor that of a Game
-Reserve. A similar case of non-formally protected but key dispersal landscape is
-represented by the area south of Kafue NP in Zambia, for which a disruption of
-its main and narrow dispersal corridor would result in considerable isolation of
-its subpopulations. We also revealed a potential southwards corridor between the
-Okavango-Linyanti ecosystem and the Central Kalahari National Park.
-\cite{Elliot.2014} and \cite{Cushman.2018} identified a similar corridor for
-dispersing lions, suggesting that upholding and protecting a link between those
-ecosystems is pivotal. Some areas through which the corridor runs are neither
-part of the KAZA-TFCA nor profit from any form of protection status. In fact,
-human presence and activities along the national road that longitudinally
-traverses this corridor may limit realized connectivity \citep{Cozzi.2020}.
+highly frequented corridor. Furthermore, the lack of permeable areas between
+peripheral source points often resulted in corridors stretching longer Euclidean
+distances, in an arc-shaped route via a stretch of suitable habitat through the
+Okavango-Linyanty ecosystem. This is an expected outcome in case structural and
+functional connectivity coincide, i.e. when dispersers follow suitable habitats
+to disperse \citep{Fattebert.2015, Hauenstein.2019}.\todo{As per reviewer 2} The
+key role of the Okavango-Linyanti region for overall connectivity within the
+KAZA-TFCA thus calls for actions to secure its protection status in the future.
+While the region is currently a Wildlife Management Area, it has neither the
+status of a National Park nor that of a Game Reserve. A similar case of
+non-formally protected but key dispersal landscape is represented by the area
+south of Kafue NP in Zambia, for which a disruption of its main and narrow
+dispersal corridor would result in considerable isolation of its subpopulations.
+We also revealed a potential southwards corridor between the Okavango-Linyanti
+ecosystem and the Central Kalahari National Park. \cite{Elliot.2014} and
+\cite{Cushman.2018} identified a similar corridor for dispersing lions,
+suggesting that upholding and protecting a link between those ecosystems is
+pivotal. Some areas through which the corridor runs are neither part of the
+KAZA-TFCA nor profit from any form of protection status. Human presence and
+activities along the national road that longitudinally traverses this corridor
+may limit functional connectivity \citep{Cozzi.2020}.
 
@@
 dispersal journey and that they have almost complete knowledge of associated
-movement costs \citep{Koen.2014, Panzacchi.2016}. Since dispersers often move
-into unknown territory, this may not necessarily be the case
-\citep{Abrahms.2017, Cozzi.2020}. However, specification of pre-defined end
-points might not be necessary, as the parametrized iSSF model can be used as
-mechanistic movement model to simulate dispersal events from known source
-points, yet without restricting the domain of potential end points
-\citep{Signer.2017}. Consequently, movement corridors (i.e. transient
-utilisation distribution) would emerge more naturally as the result of a myriad
-of simulated dispersal events. A very similar approach could be applied to
-produce permeability surfaces from steady state utilistation distributions
-\citep{Avgar.2016, Signer.2017}. \citep{Signer.2017} have shown that such
-permeability surfaces are less prone to overestimateing permeability in areas
-that lie far from suitable habitats. While a simulation-based approach is
-conceptually straightforward, computational requirements are tremendous,
-especially for such a large extent as constituted by the KAZA-TFCA.
-\todo{Reviewer 1 pointed out that the permeability surface likely overestimates
-permeability in areas far from suitable habitat. Additionally, reviewer 2
-mentioned that our dataset by far suffices for simulations and we should not
-argue with sample size.}
+movement costs \citep{Panzacchi.2016}. Since dispersers often move into unknown
+territory, this may not necessarily be the case \citep{Abrahms.2017,
+Cozzi.2020}. However, specification of pre-defined end points might not be
+necessary, as the parametrized iSSF model can be used as mechanistic movement
+model to simulate dispersal events from known source points, yet without
+restricting the domain of potential end points \citep{Signer.2017}.
+Consequently, movement corridors would emerge more naturally as the result of a
+myriad of simulated dispersal events \citep{Zeller.2020}. Individual based
+simulations could also serve to generate permeability surfaces that are based on
+steady state utilistation distributions \citep{Avgar.2016, Signer.2017}. Such
+surfaces have been shown to reduce the risk of overstimating permeability,
+particularly in areas that lie far from suitable habitats \cite{Signer.2017}.
+While a simulation-based approach is conceptually straightforward, computational
+requirements are tremendous, especially for such a large extent as constituted
+by the KAZA-TFCA. \todo{Reviewer 1 pointed out that the permeability surface
+likely overestimates permeability in areas far from suitable habitat.
+Additionally, reviewer 2 mentioned that our dataset by far suffices for
+simulations and we should not argue with sample size.}
 
@@
 GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication.
+the time of publication. We will also provide example R-code on Github to
+illustrate our least-cost modelling approach
+(\url{https://github.com/DavidDHofmann}).\todo{As per editor}
 
@@
 dynamic water maps. We also thank the anonymous reviewers for their valuable
-comments and suggestions. This study was funded by Basler Stiftung für
-Biologische Forschung, Claraz Foundation, Idea Wild, Jacot Foundation, National
-Geographic Society, Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife
-Trust Foundation, Forschungkredit der Universität Zürich, and a Swiss National
-Science Foundation Grant (31003A\_182286) to A. Ozgul.
+comments and suggestions. This study was funded by Albert-Heim Stiftung, Basler
+Stiftung für Biologische Forschung, Claraz Foundation, Idea Wild, Jacot
+Foundation, National Geographic Society, Parrotia Stiftung, Stiftung Temperatio,
+Wilderness Wildlife Trust Foundation, Forschungkredit der Universität Zürich,
+and a Swiss National Science Foundation Grant (31003A\_182286) to A. Ozgul.
 

### 2020-09-21 Chapter_0/04_Manuscript/Manuscript.tex
Finalized revisions

@@
 permeability difference among the five KAZA-TFCA countries. We also showed
-that major movement corridors of wild dogs run within the KAZA-TFCA, although
+that major movement corridors of wild dogs ran within the KAZA-TFCA, although
 some minor routes remained outside formally protected areas.
@@
 rivers, swamps or open water. The relation between permeability and other
-landscape features was less evident.\todo{Editor indicated that the
-relationships are correlative and not necessarily causal.}
+landscape features was less evident.
 
@@
 highlight the need for a coordinated effort towards maintaining or restoring
-connectivity, especially where transboundary dispersal occurs.\todo{According to
-suggestions of reviewer 2}
+connectivity, especially where transboundary dispersal occurs.
 \end{abstract}
@@
 movement corridors between wildlife reserves has become a task of utmost
-\todo{Reviewer one pointed out an awkward use of the word utmost. I have changed
-it according to his suggestion.} importance \citep{Doerr.2011, Rudnick.2012},
-resulting in an ever-growing number of large and often transboundary protected
-areas. While boundaries of such areas are often drawn according to expert
-opinion and socio-political needs, subjective assessments have revealed
-deficiencies in the past \citep{Clevenger.2002, Pullinger.2010}. Thus, an
-empirical evaluation of the adequacy of already-existing or planned protected
-areas using pertinent animal movement data is paramount for targeted use of
-valuable and scarce conservation funds \citep{Pullinger.2010}.
+importance \citep{Doerr.2011, Rudnick.2012}, resulting in an ever-growing number
+of large and often transboundary protected areas. While boundaries of such areas
+are often drawn according to expert opinion and socio-political needs,
+subjective assessments have revealed deficiencies in the past
+\citep{Clevenger.2002, Pullinger.2010}. Thus, an empirical evaluation of the
+adequacy of already-existing or planned protected areas using pertinent animal
+movement data is paramount for targeted use of valuable and scarce conservation
+funds \citep{Pullinger.2010}.
 
@@
 identify movement corridors and assess connectivity at large scales (e.g.
-\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). The use of relocation
-data has shifted the focus from a structural to a functional view on
-connectivity, taking into account how species respond to environmental features
-\citep{Taylor.2006, Fattebert.2015}.\todo{As per reviewer 2} Identification of
+\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). Identification of
 potential movement corridors typically relies on the estimation of permeability
@@
 traverses a specific landscape \citep{Sawyer.2011}. Such surfaces are created
-based on species' relative selection strengths \citep{Avgar.2017}\todo{reviewer
-1}, which can be quantified using a suite of selection functions
-\citep{Zeller.2012}. Specifically, selection strengths \todo{reviewer 1} are
-estimated by comparing spatial covariates (e.g. environmental and anthropogenic)
-at locations visited by the animal to the same spatial covariates at locations
-avilable to the animal \todo{Reviewer 1 pointed out correctly that we should
-talk about available locations, rather than randomly selected locations}
-\citep{Zeller.2012}. Importantly, selection functions rely on adequate landscape
-and relocation data that are representative of the process being studied
-\citep{Diniz.2020}. Altough selection during residence and dispersal may
-coincide \citep{Fattebert.2015},\todo{second reviewer pointed out that space use
-can be alike} it appears that relocation data collected on dispersing
-individuals outperforms data collected on resident individuals in the detection
-of large-scale movement corridors \citep{Elliot.2014, Diniz.2020}. Nevertheless,
-dispersal data is inherently difficult to collect and remains scarce in the
-connectivity literature \citep{Vasudev.2015}. As such, most permeability
-surfaces upon which movement corridors are identified are created using
-relocation data collected on resident individuals. This has likely limited our
-ability to meaningfully assess the effectiveness of protected areas in securing
-functional connectivity for their protected species.\todo{Reviewer 2 desired to
-change from: this introduces - to: this can introduce}
+based on species' relative selection strengths \citep{Avgar.2017}, which can be
+quantified using a suite of selection functions \citep{Zeller.2012}.
+Specifically, selection strengths are estimated by comparing spatial covariates
+(e.g. environmental and anthropogenic) at locations visited by the animal to the
+same spatial covariates at locations avilable to the animal \citep{Zeller.2012}.
+Importantly, selection functions require adequate landscape and relocation data
+that are representative of the process being studied \citep{Diniz.2020}. Altough
+selection during residence and dispersal may coincide \citep{Fattebert.2015}, it
+appears that relocation data collected on dispersing individuals outperforms
+data collected on resident individuals in the detection of large-scale movement
+corridors \citep{Elliot.2014, Abrahms.2017, Diniz.2020}. Nevertheless, dispersal
+data is inherently difficult to collect and remains scarce in the connectivity
+literature \citep{Vasudev.2015}. As such, most permeability surfaces upon which
+movement corridors are identified are created using relocation data collected on
+resident individuals. This has likely limited our ability to meaningfully assess
+the effectiveness of protected areas in securing functional connectivity for
+their protected species.
 
@@
 (\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
-have attempted to assess the adequacy of the KAZA-TFCA using relevant global
-positioning system (GPS) relocation data of its protected species at large
-spatial scales \citep{Elliot.2014, Tshipa.2017}.\todo{as per editor} Thus, how
-well the boundaries of the KAZA-TFCA reflect natural movement patterns and
-dispersal corridors of its most mobile protected species is virtually unknown.
+have attempted to assess the adequacy of the KAZA-TFCA using global positioning
+system (GPS) relocation data of its protected species at large spatial scales
+\citep{Elliot.2014, Tshipa.2017, Brennan.2020}. Thus, how well the boundaries of
+the KAZA-TFCA reflect natural movement patterns and dispersal corridors of its
+most mobile protected species is virtually unknown.
 
@@
 Botswana to assess the adequacy of the KAZA-TFCA in securing connectivity. We
-estimated relative selection strengths\todo{reviewer 1} towards environmental
-and anthropogenic landscape features, and used the obtained coefficients to
-predict a permeability surface spanning the entire KAZA-TFCA. We then
-investigated how landscape permeability varied regionally and internationally,
-and we compared permeability within and outside the KAZA-TFCA boundaries.
-Finally, we calculated least-cost paths and corridors to identify major movement
-routes and to verify that these were successfully covered by the KAZA-TFCA.
+estimated relative selection strengths towards environmental and anthropogenic
+landscape features, and used the obtained coefficients to predict a permeability
+surface spanning the entire KAZA-TFCA. We then investigated how landscape
+permeability varied regionally and internationally, and we compared permeability
+within and outside the KAZA-TFCA boundaries. Finally, we calculated least-cost
+paths and corridors to identify major movement routes and to verify that these
+were successfully covered by the KAZA-TFCA.
 
@@
 unrelated individuals of the opposite sex in their pack \citep{McNutt.1996,
-Behr.2020}\todo{The editor desired to include criteria}. Selected individuals
-were immobilized according to protocols described in \cite{Osofsky.1996}, and
-fitted with GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic
-Aerospace GmbH, Berlin, Germany}) while still with their natal pack. All
-required procedures were undertaken and supervised by a Botswana-registered
-wildlife veterinarian. Collars produced a tag weight of 300 gramms, accounting
-for 1.2\% of a wild dog's average body weight. They were attached using leather
-belts that included a drop off mechanism, triggered by a decomposing piece of
-cloth.\todo{Editor desired more information on GPS collar} During dispersal, GPS
-collars were programmed to record a GPS relocation every 4 hours and to
-regularly transmit data via Iridium satellite system to a base station.
+Behr.2020}. Selected individuals were immobilized according to protocols
+described in \cite{Osofsky.1996}, and fitted with GPS/Satellite radio collars
+(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin, Germany}) while still
+with their natal pack. Collars weighed 330g, accounting for about 1.5\% of a
+wild dog's body weight. A 5cm long decomposable cotton piece was added to the
+collar belt to guarantee collar drop-off after about 12-18 months. All required
+procedures were undertaken and supervised by a Botswana-registered wildlife
+veterinarian. During dispersal, GPS collars were programmed to record a GPS
+relocation every 4 hours and to regularly transmit data via iridium satellite
+system to a base station.
 
@@
 visual inspection of the net squared displacement (NSD) metric. NSD quantifies
-the squared \todo{Reviewer 1 pointed out correctly that we need to talk about
-the \textit{squared} euclidean distance} Euclidean distance of a relocation to a
-reference point \citep{Borger.2012}, which in our case was the center of the
-dispersing coalition's natal home range. Thus, dispersal was deemed to have
-started when a coalition had left its natal home range and continued until the
-NSD metric remained stationary, indicating that the coalition had successfully
-settled (Figure S1). In our analysis, we did not differentiate between male and
-female dispersing coalitions, for previous research found little differences
-between sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.
+the squared Euclidean distance of a relocation to a reference point
+\citep{Borger.2012}, which in our case was the center of the dispersing
+coalition's natal home range. Thus, dispersal was deemed to have started when a
+coalition had left its natal home range and continued until the NSD metric
+remained stationary, indicating that the coalition had successfully settled
+(Figure S1). In our analysis, we did not differentiate between male and female
+dispersing coalitions, for previous research found little differences between
+sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.
 
 \subsection{Spatial Covariates}
-To investigate relative selection strengths \todo{As per reviewer 1} of
-dispersing wild dogs, we used a set of geo-referenced covariates
-(\Cref{Covariates}) that we aggregated in the categories \textit{land cover},
-\textit{protection status}, and \textit{human influence}. Land cover comprised
-the covariates water cover, distance to water, percentage cover by
-shrubs/grassland, and percentage cover by trees. To capture the pulsing behavior
-of the Okavango Delta, we frequently updated layers for water cover and
-corresponding layers depicting distance to water. Protection status contained a
-binary indicator of whether an area was protected or not. Human influence
-included covariates rendering human density, the presence of roads, and the
-distance to roads.\todo{Reviewer 2 found the description of the covariates
-unclear} We prepared all covariates as spatial raster layers from freely
-available online services and from remotely sensed satellite imagery. To ensure
-a consistent resolution (i.e. cell-size or grain) across covariates, we
-coarsened or interpolated all layers to a resolution of 250m x 250m. For further
-details on the preparation of each covariate and their sources, see Appendix
-A.3. We performed processing and manipulation of data as well as all spatial and
-statistical analyses using R, version 3.6.1 \citep{R.2019}.
+To investigate relative selection strengths of dispersing wild dogs, we used a
+set of geo-referenced covariates (\Cref{Covariates}) that we aggregated in the
+categories \textit{land cover}, \textit{protection status}, and
+\textit{anthropogenic}. \textit{Land cover} comprised the covariates water cover
+(binary), distance to water (continuous), percentage cover by shrubs/grassland
+(continuous), and percentage cover by trees (continuous). To capture the pulsing
+behavior of the Okavango Delta, we classified satellite imagery and frequently
+updated layers for water cover and corresponding layers depicting distance to
+water. \textit{Protection status} contained a binary covariate, indicating
+whether an area was protected or not. \textit{Anthropogenic} included covariates
+rendering the presence of roads (binary), the distance to roads (continuous),
+and a proxy for human influence (continuous) that took into account human
+density, farming, and roads. We prepared all covariates as spatial raster layers
+from freely available online services and from remotely sensed satellite
+imagery. To ensure a consistent resolution (i.e. cell-size or grain) across
+covariates, we coarsened or interpolated all layers to a resolution of 250m x
+250m. For further details on the preparation and source of each covariate, see
+Appendix A.3. We performed processing and manipulation of data as well as all
+spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.
 
@@
 We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
-investigate dispersers' relative selection strentghs towards \todo{Reviewer 1
-suggested to use relative selection strength, rather than preference} the
-above-mentioned spatial covariates. That is, we paired each realized step (i.e.
-the connecting line between two consecutive GPS relocations;
-\citealp{Turchin.1998}) with 24 random steps that were generated by sampling
-turning angles from a uniform distribution U(\(-\pi, +\pi\)) and step lengths
-from a gamma distribution fitted to realized steps \citep{Avgar.2016}. A
-realized step and its 24 associated random steps formed a stratum and received a
-unique identifier. Along each step, we extracted the above-mentioned covariates
-(Table S3), standardized extracted values using a z-score transformation, and
-checked for correlation using Pearson's Correlation Coefficient \(r\). None of
-the covariates were overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and
-we retained all of them for modeling. Our habitat selection model then assumed
-that dispersing wild dogs assigned a selection score \(w(x)\) of the following
-exponential form to each realized and random step \citep{Fortin.2005}:
+investigate dispersers' relative selection strentghs towards the above-mentioned
+spatial covariates. That is, we paired each realized step (i.e. the connecting
+line between two consecutive GPS relocations; \citealp{Turchin.1998}) with 24
+random steps that were generated by sampling turning angles from a uniform
+distribution U(\(-\pi, +\pi\)) and step lengths from a gamma distribution fitted
+to realized steps \citep{Avgar.2016}. A realized step and its 24 associated
+random steps formed a stratum and received a unique identifier. Along each step,
+we extracted the above-mentioned covariates (Table S3), standardized extracted
+values using a z-score transformation, and checked for correlation using
+Pearson's Correlation Coefficient \(r\). None of the covariates were overly
+correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we retained all of them
+for modeling. Our habitat selection model then assumed that dispersing wild dogs
+assigned a selection score \(w(x)\) of the following exponential form to each
+realized and random step \citep{Fortin.2005}:
 
@@
 covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's relative
-selection strengths \todo{As per reviewer 1} towards these covariates
-(\(\beta_1, \beta_2, ..., \beta_n\)). To estimate relative selection strengths
-\todo{As per reviewer 1} (i.e. the \(\beta\)'s) for each covariate, we used
-mixed effects conditional logistic regression analysis as suggested by
-\cite{Muff.2020}. We implemented their method using the R-package
-\textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID to model
-random slopes. We also modelled random intercepts with an arbitrary high
-variance of 10\textsuperscript{6} to make use of the poisson trick (see
+selection strengths towards these covariates (\(\beta_1, \beta_2, ...,
+\beta_n\)). To estimate relative selection strengths (i.e. the \(\beta\)'s) for
+each covariate, we used mixed effects conditional logistic regression analysis
+as suggested by \cite{Muff.2020}. We implemented their method using the
+R-package \textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID
+to model random slopes. We also modelled random intercepts with an arbitrary
+high variance of 10\textsuperscript{6} to make use of the poisson trick (see
 \citealp{Muff.2020}). We defined three movement metrics, namely the cosine of
-the turning angle (\(cos(ta)\)), the step length (\(sl\)) and the natural
-logarithm of the step length (\(log(sl)\)), as core covariates and ran stepwise
-forward model selection based on Akaike's Information Criterion (AIC;
-\citealp{Burnham.2002}) for all other covariates.\todo{Reviewer 1 asked for more
-details in the random effects. I therefore decided to update this section too}
-The inclusion of movement metrics served to reduce biases in estimated habitat
-selection coefficients \todo{as per revierwer 1} that may have arisen due to
+the turning angle (\(cos(ta)\)), the step length (\(sl\)) and the logarithm of
+the step length (\(log(sl)\)), as core covariates and ran stepwise forward model
+selection based on Akaike's Information Criterion (AIC; \citealp{Burnham.2002})
+for all other covariates. The inclusion of movement metrics served to reduce
+biases in estimated habitat selection coefficients that may have arisen due to
 movement behavior \citep{Avgar.2016}. To validate the predictive power of the
@@
 \ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
-for each raster cell. Because our representation of water was dynamic (to
-properly render the pulsing behavior of the Okavango Delta) we collapsed all
-dynamic water maps into a single static map using areas that were covered by
-water in at least 10\% of the cases. Using the resulting static map we also
-calculated a layer returning the distance to water. Because the extent of the
-delta covers merely 6\% of the entire KAZA-TFCA, the use of static maps should
-introduce little bias in predicted permeability. To reduce the influence of
+for each raster cell. Because our representation of water was dynamic we
+collapsed all dynamic water maps into a single map using areas that were covered
+by water in at least 10\% of the cases. We used the resulting map to also
+calculate a layer returning the distance to water. To reduce the influence of
 outliers in predicted permeability scores, we followed \cite{Squires.2013} and
@@
 any source points from the regular grid. Because wild dogs residing outside of
-protected areas are rare and likely unviable \citep{VanDerMeer.2014}, we
-consider our selection of source points within protected areas to be
-appropriate. In total, we generated 68 source points, which resulted in 2'278
-unique pairwise combinations and therefore 2'278 unique LCPs and LCCs. We
-computed factorial LCPs and LCCs between source points using the R-package
-\textit{gdistance} (further details in Appendix A.7). After computation, we
-tallied overlapping LCPs and LCCs, respectively, into single connectivity maps.
+protected areas are rare and unviable \citep{VanDerMeer.2014}, we consider our
+selection of source points within protected areas to be appropriate. In total,
+we generated 68 source points, which resulted in 2'278 unique pairwise
+combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
+LCPs and LCCs between source points using the R-package \textit{gdistance}
+(further details in Appendix A.7). After computation, we tallied overlapping
+LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
+discrete paths, whereas LCCs return continuous corridors, we present both
+methods.
 
@@
 covered a mean Euclidean distance of 54 km (\(SD\) = 71) and a cumulative
-distance of 597 km (\(SD\) = 508). One female coalition dispersed far east into
-the Hwange Hwange National Park and covered a cumulative distance of over 360 km
-in under 9 days.\todo{Editor desired a quick summary of the dispersal data.}
+distance of 597 km (\(SD\) = 508).
 
@@
 \textit{human influence}, beside the fixed covariates \textit{cos(ta)},
-\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). The positive
-coefficients for \(sl\) (\(\beta\) = 0.03, 95\% CI = -0.09 to 0.15) and
-\(log(sl)\) (\(\beta\) = 0.12, 95\% CI = 0.02 to 0.22) indicated that our
-initial (selection-biased) gamma distribution produced step-lengths slightly
-shorter than a selection-free gamma distribution would. Similarly, the positive
-coefficient for \(cos(ta)\) (\(\beta\) = 0.10; 95\% CI = 0.05 to 0.15) suggested
-that our uniform distribution for turning angles produced turning angles with
-too little directionality. \todo{Reviewer 1 nicely explained that the
-coefficients for log(sl) and cos(ta) imply that the initial distributions were
-selection-biased} With respect to environmental covariates, dispersers avoided
-moving through water (\(\beta\) = -0.53, 95\% CI -0.79 to -0.27) but selected
-for locations in its vicinity, although the latter effect was not significant
-(\(\beta\) = -0.33, 95\% CI = -0.73 to 0.08). Dispersers avoided areas that were
-densely covered by trees (\(\beta\) = -0.31, CI = -0.47 to -0.15) and preferred
-areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42).
-Finally, dispersers avoided areas that were influenced by humans (\(\beta\) =
--0.45, 95\% CI = -0.82 to -0.08). With the exepction of \textit{distance to
-water} (\(SD_{RE}\) = 0.57), random effects revealed little variability between
-different dispersal coalitions' selection strengths (i.e. all with \(SD_{RE}
-\leq 0.22\), Figure S7).\todo{Reviewer 1 desired more information on the random
-effects}
+\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). With respect to
+environmental covariates, dispersers avoided moving through water (\(\beta\) =
+-0.53, 95\% CI -0.79 to -0.27) but selected for locations in its vicinity,
+although the latter effect was not significant (\(\beta\) = -0.33, 95\% CI =
+-0.73 to 0.08). Dispersers avoided areas that were densely covered by trees
+(\(\beta\) = -0.31, CI = -0.47 to -0.15) and preferred areas covered by
+shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42). Finally, dispersers
+avoided areas that were influenced by humans (\(\beta\) = -0.45, 95\% CI = -0.82
+to -0.08). With the exepction of \textit{distance to water}, random effects
+revealed little variability between dispersal coalitions (Appendix A6).
+
+% The positive coefficient for \(cos(ta)\) (\(\beta\) = 0.10; 95\% CI = 0.05 to
+% 0.15) suggested that the uniform distribution for turning angles produced too
+% little directionality. Similarly, the positive coefficient for \(log(sl)\)
+% (\(\beta\) = 0.12, 95\% CI = 0.02 to 0.22) indicated that our fitted gamma
+% distribution produced step-lengths slightly shorter than the true
+% selection-free gamma distribution would have. The estimate for \(sl\), on the
+% other hand, was insignificant (\(\beta\) = 0.03, 95\% CI = -0.09 to 0.15).
 
@@
 Our results emphasize that human influences constitute some of the main barriers
-to connectivity among wild dog populations. According to our model, dispersers
-avoided corssing human dominated landscapes whenever given the
-choice.\todo{According to reviewer 1 we should point out that our model compares
-available vs. unavailable} This conforms to findings on dispersing wild dogs
-from eastern Africa \citep{Masenga.2016, Oneill.2020} but conflicts with
-findings from South Africa by \cite{DaviesMostert.2012}, who reported a high
-willingness of dispersers to cross human-dominated landscapes. We believe that
-such differences are due to the unavailability of alternative routes through
-natural landscapes, which may have forced dispersers in South Africa to cross
-human dominated landscapes despite a strong aversion to do so. In this regard,
-our representation of dispersal corridors and the resulting connectivity appear
-conservative, as dispersers may be able to make the best out of a bad situation
-and cross landscapes characterized by considerably unfavorable conditions
-\citep{Palomares.2000, Elliot.2014}. Unfortunately, \cite{DaviesMostert.2012}
-only recorded net dispersal distances, thereby precluding a more indepth
-understanding of the impact of humans on dispersal behavior.\todo{As per
-reviewer 2} Nevertheless, successful conservation of this species relies on
-policymakers' and local authorities' willingness and ability to provide and
-conserve natural areas that remain free from anthropogenic pressures. This is
-not only paramount in light of increasing connectivity and facilitating
-dispersal, but also in terms of reducing human-caused mortality during
-dispersal. In fact, previous studies have shown that human-caused mortality
-represents a major threat to wild dogs' ability to disperse
-\citep{Woodroffe.2019, Cozzi.2020}.
+to connectivity among wild dog populations. This conforms to findings on
+dispersing wild dogs from eastern Africa \citep{Masenga.2016, Oneill.2020} but
+conflicts with findings from South Africa by \cite{DaviesMostert.2012}, who
+reported a high willingness of dispersers to cross human-dominated landscapes.
+Such differences may arise from the fact that our model infers preferences by
+comparing \textit{used} and \textit{available} habitats, whereas
+\cite{DaviesMostert.2012} only recorded net dispersal distances, thereby
+precluding such an analysis. Thus, we believe that differences to
+\cite{DaviesMostert.2012} may be explained by the unavailability of alternative
+routes through natural landscapes, which may have forced dispersers in South
+Africa to cross human dominated landscapes despite a strong aversion to do so.
+In this regard, our representation of dispersal corridors and the resulting
+connectivity appear conservative, as dispersers may be able to make the best out
+of a bad situation and cross landscapes characterized by considerably
+unfavorable conditions \citep{Palomares.2000, Elliot.2014}. Nevertheless,
+successful conservation of this species relies on policymakers' and local
+authorities' willingness and ability to provide and conserve natural areas that
+remain free from anthropogenic pressures. This is not only paramount in light of
+increasing connectivity and facilitating dispersal, but also in terms of
+reducing human-caused mortality during dispersal. In fact, previous studies have
+shown that human-caused mortality represents a major threat to wild dogs'
+ability to disperse \citep{Woodroffe.2019, Cozzi.2020}.
 
@@
 spotted hyenas, and resident wild dogs may also use areas close to water
-\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to move
-into prey-poorer areas away from water \citep{Creel.2002, Mills.1997}. This
-reasoning may also explain why the effect of distance to water was
-insignificant.\todo{Reviewer 1 suggested to add this} Given the influence that
+\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to switch
+behavior and move into prey-poorer areas away from water \citep{Creel.2002,
+Mills.1997}. This reasoning may also explain the large confidence intervals
+surrounding the corresponding \(\beta\)-estimate. Given the influence that
 resident conspecifics, competitors, and prey can have on dispersers
@@
 findings from dipsersing wild dogs in Kenya \citep{Oneill.2020}, suggesting that
-there are strong commonalities dispersers from these very different ecosystems.
-Thus, despite wild dogs' ability to cope with diverse habitats and adapt to
-changing conditions \citep{Woodroffe.2011}, the fundamental factors included in
-our study appear to influence wild dogs from other ecosystems alike.
+there are strong commonalities between dispersers from these very different
+ecosystems. Thus, despite wild dogs' ability to cope with diverse habitats and
+adapt to changing conditions \citep{Woodroffe.2011}, the fundamental factors
+included in our study appear to influence wild dogs from other ecosystems alike.
 Nevertheless, expanding our analysis to dispersers emigrating from other source
-populations would invaluably contribute to our understanding of
-dispersal.\todo{Editor desired info on applicability}
+populations would invaluably contribute to our understanding of dispersal.
 
@@
 functional connectivity coincide, i.e. when dispersers follow suitable habitats
-to disperse \citep{Fattebert.2015, Hauenstein.2019}.\todo{As per reviewer 2} The
-key role of the Okavango-Linyanti region for overall connectivity within the
-KAZA-TFCA thus calls for actions to secure its protection status in the future.
-While the region is currently a Wildlife Management Area, it has neither the
-status of a National Park nor that of a Game Reserve. A similar case of
-non-formally protected but key dispersal landscape is represented by the area
-south of Kafue NP in Zambia, for which a disruption of its main and narrow
-dispersal corridor would result in considerable isolation of its subpopulations.
-We also revealed a potential southwards corridor between the Okavango-Linyanti
-ecosystem and the Central Kalahari National Park. \cite{Elliot.2014} and
-\cite{Cushman.2018} identified a similar corridor for dispersing lions,
-suggesting that upholding and protecting a link between those ecosystems is
-pivotal. Some areas through which the corridor runs are neither part of the
-KAZA-TFCA nor profit from any form of protection status. Human presence and
-activities along the national road that longitudinally traverses this corridor
-may limit functional connectivity \citep{Cozzi.2020}.
+to disperse \citep{Fattebert.2015, Hauenstein.2019}. The key role of the
+Okavango-Linyanti region for overall connectivity within the KAZA-TFCA thus
+calls for actions to secure its protection status in the future. While the
+region is currently a Wildlife Management Area, it has neither the status of a
+National Park nor that of a Game Reserve. A similar case of non-formally
+protected but key dispersal landscape is represented by the area south of Kafue
+NP in Zambia, for which a disruption of its main and narrow dispersal corridor
+would result in considerable isolation of its subpopulations. We also revealed a
+potential southwards corridor between the Okavango-Linyanti ecosystem and the
+Central Kalahari National Park. \cite{Elliot.2014} and \cite{Cushman.2018}
+identified a similar corridor for dispersing lions, suggesting that upholding
+and protecting a link between those ecosystems is pivotal. Some areas through
+which the corridor runs are neither part of the KAZA-TFCA nor profit from any
+form of protection status. Human presence and activities along the national road
+that longitudinally traverses this corridor may limit functional connectivity
+\citep{Cozzi.2020}.
 
@@
 Consequently, movement corridors would emerge more naturally as the result of a
-myriad of simulated dispersal events \citep{Zeller.2020}. Individual based
-simulations could also serve to generate permeability surfaces that are based on
-steady state utilistation distributions \citep{Avgar.2016, Signer.2017}. Such
-surfaces have been shown to reduce the risk of overstimating permeability,
-particularly in areas that lie far from suitable habitats \cite{Signer.2017}.
-While a simulation-based approach is conceptually straightforward, computational
-requirements are tremendous, especially for such a large extent as constituted
-by the KAZA-TFCA. \todo{Reviewer 1 pointed out that the permeability surface
-likely overestimates permeability in areas far from suitable habitat.
-Additionally, reviewer 2 mentioned that our dataset by far suffices for
-simulations and we should not argue with sample size.}
+myriad of simulated dispersal events \citep{Zeller.2020}.
+
+Besides estimating corridors, individual based simulations may be used to
+generate permeability surfaces \citep{Avgar.2016, Signer.2017}. Such surfaces
+have been shown to reduce the risk of overestimating \(w(x)\) and conesquently
+connectivity, particularly in areas that lie far from suitable habitats
+\citep{Signer.2017}. While a simulation-based approach is conceptually
+straightforward, computational requirements are high, especially for such a
+large extent as constituted by the KAZA-TFCA.
 
 Our work shows how dispersal data of a highly mobile species can be used to
-\todo{to instead of t, pointed out by the editor and reviewer 1} identify
-movement corridors and to assess the adequacy of protected areas. In our case,
-the predicted movement corridors of African wild dogs were well contained within
-the boundaries of the world's largest transboundary conservation area, namely
-the KAZA-TFCA, suggesting that it will significantly contribute to the long-term
-viability of this species. Moreover, our connectivity network allowed revealing
-potential dispersal hubs through which dispersers gain access to more remote
-regions of the study area. Finally, our investigations showed that human
-influence constitutes one of the main barriers to dispersal and substantially
-reduces landscape connectivity. Successful conservation of wide-ranging species,
-such as exemplified by the African wild dog, will therefore be contingent on the
-willingness of local authorities, policymakers, and land managers to preserve
-areas that remain free from human strains. Ultimately, our work contributes to
-the growing field of connectivity studies and provides and easily expandable
-framework for assessing the adequacy of already-existing or planned protected
-areas.
+identify movement corridors and to assess the adequacy of protected areas. In
+our case, the predicted movement corridors of African wild dogs were well
+contained within the boundaries of the world's largest transboundary
+conservation area, namely the KAZA-TFCA, suggesting that it will significantly
+contribute to the long-term viability of this species. Moreover, our
+connectivity network allowed revealing potential dispersal hubs through which
+dispersers gain access to more remote regions of the study area. Finally, our
+investigations showed that human influence constitutes one of the main barriers
+to dispersal and substantially reduces landscape connectivity. Successful
+conservation of wide-ranging species, such as exemplified by the African wild
+dog, will therefore be contingent on the willingness of local authorities,
+policymakers, and land managers to preserve areas that remain free from human
+strains. Ultimately, our work contributes to the growing field of connectivity
+studies and provides and easily expandable framework for assessing the adequacy
+of already-existing or planned protected areas.
 
@@
 the time of publication. We will also provide example R-code on Github to
-illustrate our least-cost modelling approach
-(\url{https://github.com/DavidDHofmann}).\todo{As per editor}
+demonstrate our least-cost modelling approach
+(\url{https://github.com/DavidDHofmann}).
 

### 2020-09-21 Chapter_0/04_Manuscript/Manuscript_Support.tex
Worked on revisions
Finalized revisions

@@
 %------------------------------------------------------------------------------
-\section{Net Squared Displacement}\todo{Reviewer 1 pointed out that we need to
-replace calssified with classified}
+\section{Net Squared Displacement}
+
 \begin{figure}[hbpt]
@@
   \begin{center}
-    \resizebox{0.8\textwidth}{!}{
+    \resizebox{\textwidth}{!}{
       \begin{tabular}{lccccccc}
@@
       \toprule
-      \textit{Mean} & - & - & 1080 & 261 & 48 & 54 & 597 \\
+      \textit{Mean} & - & - & 1'080 & 261 & 48 & 54 & 597 \\
       \textit{(SD)} & - & - & (649) & (207) & (44) & (71) & (508) \\
@@
 
-\subsection{Human Influence}
+\subsection{Anthropogenic}
 We created a raster layer representing human influence by integrating
@@
         \hdashline
-        \multirow{3}{*}{Human Influence}
+        \multirow{3}{*}{Anthropogenic}
           & Human Influence
@@
         \begin{tablenotes}
-          \item \textit{Sources:} (1) \cite{Chen.2015} (2) \cite{Schaaf.2015} (3)
-          \cite{Yamazaki.2019} (4) \cite{Dimiceli.2015} (5) \cite{PeaceParks.2019}
-          (6) \cite{Facebook.2019} (7) \cite{OpenStreetMap.2019} (8)
-          \cite{Xiong.2017}
+          \item \textit{Sources:} (1) \cite{Chen.2015} (2) \cite{Schaaf.2015}
+          (3) \cite{Yamazaki.2019} (4) \cite{Dimiceli.2015} (5)
+          \cite{PeaceParks.2019} (6) \cite{Facebook.2019} (7)
+          \cite{OpenStreetMap.2019} (8) \cite{Xiong.2017}
           \item
           \item \textit{* Note:} The covariates \textit{Water} and
-          \textit{Dryland} added up to 100\%, which is why only \textit{Water} was
-          included as explanatory variable in our models. The same applied for the
-          group \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland},
-          where we omitted \textit{Bareland} for modeling. Finally, from the group
-          \textit{Protected} and \textit{Unprotected}, we only included
-          \textit{Protected} in our models.
+          \textit{Dryland} added up to 100\%, which is why only \textit{Water}
+          was included as explanatory variable in our models. The same applied
+          for the group \textit{Shrubs/Grassland}, \textit{Trees}, and
+          \textit{Bareland}, where we omitted \textit{Bareland} for modeling.
+          Finally, from the group \textit{Protected} and \textit{Unprotected},
+          we only included \textit{Protected} in our models.
         \end{tablenotes}

### 2020-09-23 Chapter_0/04_Manuscript/Manuscript.tex
Finalized Revision

@@
 degrees of human activities across regions, and to the presence or absence of
-rivers, swamps or open water. The relation between permeability and other
-landscape features was less evident.
+rivers, swamps or open water. The relationship between permeability and other
+landscape features was less pronounced.
 
@@
 discrete paths, whereas LCCs return continuous corridors, we present both
-methods.
+methods. R-codes showcasing the main steps for sections 2.5 to 2.6 can be
+downloaded from GitHub
+(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).
 
@@
 behavior and move into prey-poorer areas away from water \citep{Creel.2002,
-Mills.1997}. This reasoning may also explain the large confidence intervals
-surrounding the corresponding \(\beta\)-estimate. Given the influence that
-resident conspecifics, competitors, and prey can have on dispersers
-\citep{Cozzi.2018, Armansin.2019} future studies should strive to collect and
-incorporate intra- and interspecific relationships into analyses of landscape
-connectivity.
+Mills.1997}. This may explain the large confidence intervals for the
+corresponding \(\beta\)-estimate of \textit{distance to water}. Given the
+influence that resident conspecifics, competitors, and prey can have on
+dispersers \citep{Cozzi.2018, Armansin.2019} future studies should strive to
+collect and incorporate intra- and interspecific relationships into analyses of
+landscape connectivity.
 
 Overall, our findings on habitat selection during dispersal coincide with
-findings from dipsersing wild dogs in Kenya \citep{Oneill.2020}, suggesting that
-there are strong commonalities between dispersers from these very different
-ecosystems. Thus, despite wild dogs' ability to cope with diverse habitats and
-adapt to changing conditions \citep{Woodroffe.2011}, the fundamental factors
-included in our study appear to influence wild dogs from other ecosystems alike.
-Nevertheless, expanding our analysis to dispersers emigrating from other source
-populations would invaluably contribute to our understanding of dispersal.
+findings from dipsersing wild dogs in Kenya \citep{Oneill.2020} and Tanzania
+\citep{Masenga.2016}, suggesting that there are strong commonalities between
+dispersers from these very different ecosystems. Thus, despite wild dogs'
+ability to cope with diverse habitats and adapt to changing conditions
+\citep{Woodroffe.2011}, the fundamental factors included in our study appear to
+influence wild dogs from other ecosystems alike. Nevertheless, expanding our
+analysis to dispersers emigrating from other source populations would invaluably
+contribute to our understanding of dispersal.
 
@@
 \citep{Signer.2017}. While a simulation-based approach is conceptually
-straightforward, computational requirements are high, especially for such a
-large extent as constituted by the KAZA-TFCA.
+straightforward, computational requirements for such a large spatial extent as
+the KAZA-TFCA are very high, making the use of this approach challenging. We
+urge future studies to optimize the simulation of corridors from iSSFs that
+capture a more mechanistic model of dispersal.
 
@@
 GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication. We will also provide example R-code on Github to
-demonstrate our least-cost modelling approach
-(\url{https://github.com/DavidDHofmann}).
+the time of publication. R-code showcasing the main steps for sections 2.5 to
+2.6 can be downloaded from GitHub
+(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).
 

### 2020-09-23 Chapter_0/04_Manuscript/Manuscript_Support.tex
Finalized Revision

@@
 \newpage
+\newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
 \section{GPS Data}
@@
   \begin{center}
-    \includegraphics[width = 0.86\textwidth]{99_Trajectories.pdf}
+    \includegraphics[width = 0.84\textwidth]{99_Trajectories.pdf}
     \caption{Illustration of all trajectories that we recorded. Each color
@@
   \caption{Summary statistics of all GPS relocations that have been recorded on
-  dispersing coalitions.}
+  dispersing coalitions. The coalition ID refers to the individual in the
+  dispersal coalition that was equipped with a GPS collar. All coalitions
+  consisted of same-sex individuals.}
   \label{GPSData}
   \begin{center}
-    \resizebox{\textwidth}{!}{
+    \resizebox{0.95\textwidth}{!}{
       \begin{tabular}{lccccccc}
@@
 \end{table}
+\restoregeometry
 
@@
   \caption{Plot of random effects showing variability across dispersal
-  coalitions.}
+  coalitions. The coalition ID refers to the individual in the dispersal
+  coalition that was equipped with a GPS collar.}
   \label{RandomEffects}

### 2020-10-02 Chapter_0/04_Manuscript/Manuscript.tex
Worked on revision

@@
 by water in at least 10\% of the cases. We used the resulting map to also
-calculate a layer returning the distance to water. To reduce the influence of
-outliers in predicted permeability scores, we followed \cite{Squires.2013} and
-curtailed predicted scores between the 1\textsuperscript{st} and
-99\textsuperscript{th} percentile of their original values. To compare
-permeability across different regions, we normalized the permeability surface to
-a range between 0 (most impermeable) and 1 (most permeable), and we determined
-median permeability within and outside the KAZA-TFCA, within and outside
-formally protected areas, and within each of the five KAZA-TFCA countries.
+calculate a layer returning the distance to water. Given that the delta only
+covers 6\% of the study extent, ignoring seasonality should introduce little
+bias in predicted permeability. To reduce the influence of outliers in predicted
+permeability scores, we followed \cite{Squires.2013} and curtailed predicted
+scores between the 1\textsuperscript{st} and 99\textsuperscript{th} percentile
+of their original values. To compare permeability across different regions, we
+normalized the permeability surface to a range between 0 (most impermeable) and
+1 (most permeable), and we determined median permeability within and outside the
+KAZA-TFCA, within and outside formally protected areas, and within each of the
+five KAZA-TFCA countries.
 
@@
 environmental covariates, dispersers avoided moving through water (\(\beta\) =
--0.53, 95\% CI -0.79 to -0.27) but selected for locations in its vicinity,
-although the latter effect was not significant (\(\beta\) = -0.33, 95\% CI =
--0.73 to 0.08). Dispersers avoided areas that were densely covered by trees
-(\(\beta\) = -0.31, CI = -0.47 to -0.15) and preferred areas covered by
-shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42). Finally, dispersers
-avoided areas that were influenced by humans (\(\beta\) = -0.45, 95\% CI = -0.82
-to -0.08). With the exepction of \textit{distance to water}, random effects
-revealed little variability between dispersal coalitions (Appendix A6).
+-0.53, 95\% CI -0.79 to -0.27, SD\textsubscript{RE} = 0.17) but selected for
+locations in its vicinity, although the latter effect was not significant
+(\(\beta\) = -0.33, 95\% CI = -0.73 to 0.08, SD\textsubscript{RE} = 0.57).
+Dispersers avoided areas that were densely covered by trees (\(\beta\) = -0.31,
+CI = -0.47 to -0.15, SD\textsubscript{RE} = 0.14) and preferred areas covered by
+shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42, SD\textsubscript{RE}
+\(<\) 0.01). Finally, dispersers avoided areas that were influenced by humans
+(\(\beta\) = -0.45, 95\% CI = -0.82 to -0.08, SD\textsubscript{RE} = 0.22). With
+the exepction of \textit{distance to water}, random effects revealed little
+variability between dispersal coalitions (Appendix A6).
 

### 2020-10-02 Chapter_0/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript.r

@@ -1598,6 +1598,10 @@ ranefs <- ranef(mod, condVar = T)
 coeffs$cond$id
 ranefs$cond$id
 
+# Calculate the SD of the REs
+apply(coeffs$cond$id, 2, sd)
+apply(ranefs$cond$id, 2, sd)
+
 # Note: ranef yields the difference between the individual specific effect and
 # the mean level effect. coef, on the other hand, yields the individual specific
 # effect. Thus, the followin two lines yield (approximately) the same

### 2020-10-03 Chapter_0/04_Manuscript/Manuscript.tex
Finalized revision

--- 2020-09-23 Chapter_0/04_Manuscript/Manuscript.tex
@@
 
-\noindent 2. Between 2011 and 2019, we collected high-resolution GPS movement
-data on 16 dispersing African wild dog (\textit{Lycaon pictus}) coalitions from
-a free-ranging population in the Kavango-Zambezi Transfrontier Conservation Area
+\noindent 2. Between 2011 and 2019, we collected high-resolution GPS data on 16
+dispersing African wild dog (\textit{Lycaon pictus}) coalitions from a
+free-ranging population in the Kavango-Zambezi Transfrontier Conservation Area
 (KAZA-TFCA). Spanning five countries and 520'000 km\textsuperscript{2} the
@@
 selection analysis to estimate habitat selection of dispersers and to create a
-permeability surface for the entire KAZA-TFCA. We compared landscape
-permeability across different regions within the KAZA-TFCA as well as outside
-its boundaries. Lastly, we calculated least-cost paths and corridors to verify
-that major movement routes were adequately encompassed within the KAZA-TFCA.
+permeability surface for the KAZA-TFCA. We compared landscape permeability
+across different regions within the KAZA-TFCA as well as outside its boundaries.
+Lastly, we calculated least-cost paths and corridors to verify that major
+movement routes were adequately encompassed within the KAZA-TFCA.
 
@@
 double compared to areas outside it. Furthermore, we observed a five-fold
-permeability difference among the five KAZA-TFCA countries. We also showed
-that major movement corridors of wild dogs ran within the KAZA-TFCA, although
-some minor routes remained outside formally protected areas.
+permeability difference among the five KAZA-TFCA countries. We also showed that
+major movement corridors of wild dogs ran within the KAZA-TFCA, although some
+minor routes remained formally unprotected.
 
@@
 degrees of human activities across regions, and to the presence or absence of
-rivers, swamps or open water. The relationship between permeability and other
+rivers, swamps and open water. The relationship between permeability and other
 landscape features was less pronounced.
@@
 (e.g. environmental and anthropogenic) at locations visited by the animal to the
-same spatial covariates at locations avilable to the animal \citep{Zeller.2012}.
-Importantly, selection functions require adequate landscape and relocation data
-that are representative of the process being studied \citep{Diniz.2020}. Altough
-selection during residence and dispersal may coincide \citep{Fattebert.2015}, it
-appears that relocation data collected on dispersing individuals outperforms
-data collected on resident individuals in the detection of large-scale movement
-corridors \citep{Elliot.2014, Abrahms.2017, Diniz.2020}. Nevertheless, dispersal
-data is inherently difficult to collect and remains scarce in the connectivity
-literature \citep{Vasudev.2015}. As such, most permeability surfaces upon which
-movement corridors are identified are created using relocation data collected on
-resident individuals. This has likely limited our ability to meaningfully assess
-the effectiveness of protected areas in securing functional connectivity for
-their protected species.
+same spatial covariates at locations available to the animal
+\citep{Zeller.2012}. Importantly, selection functions require adequate landscape
+and relocation data that are representative of the process being studied
+\citep{Diniz.2020}. Altough selection during residence and dispersal may
+coincide \citep{Fattebert.2015}, it appears that relocation data collected on
+dispersing individuals outperforms data collected on resident individuals in the
+detection of large-scale movement corridors \citep{Elliot.2014, Abrahms.2017,
+Diniz.2020}. Nevertheless, dispersal data is inherently difficult to collect and
+remains scarce in the connectivity literature \citep{Vasudev.2015}. As such,
+most permeability surfaces upon which movement corridors are identified are
+created using relocation data collected on resident individuals. This has likely
+limited our ability to meaningfully assess the effectiveness of protected areas
+in securing connectivity for their protected species.
 
@@
 extirpated through human persecution, habitat destruction, and disease outbreaks
-\citep{Woodroffe.2012}. As a result, the species has become one of Africa's most
-endangered large carnivores, and currently only survives in small, spatially
-scattered subpopulations \citep{Woodroffe.2012}. Within these subpopulations,
-wild dogs form cooperative breeding packs of up to thirty individuals
-\citep{Creel.2002}, whose social structure is strongly governed by the process
-of dispersal \citep{McNutt.1996, Behr.2020}. Both males and females disperse
-from their natal pack, either alone or in same-sex dispersing coalitions, and
-search for unrelated mates and a suitable territory to settle
-\citep{McNutt.1996, Cozzi.2020, Behr.2020}. During dispersal, wild dogs can
-cover several hundred kilometers and cross international borders
-\citep{Masenga.2016, Woodroffe.2019, Cozzi.2020}. Despite the importance of
-dispersal for the long-term viability of this species, little empirical
-information is available on habitat selection and potential movement barriers
-during dispersal. The few studies that have collected dispersal data have shown
-that dispersers quickly move over large distances, avoid human-dominated
-landscapes and areas densely covered by trees, but prefer proximity to water
-\citep{Masenga.2016, Woodroffe.2019, Oneill.2020, Cozzi.2020}.
+\citep{Woodroffe.2012}. For these reasons, viable populations mainly occur in
+spatially scattered subpopulations within protected areas \citep{Woodroffe.1999,
+Woodroffe.2012, VanDerMeer.2014}. Within these subpopulations, wild dogs form
+cooperative breeding packs of up to thirty individuals \citep{Creel.2002}, whose
+social structure is strongly governed by the process of dispersal
+\citep{McNutt.1996, Behr.2020}. Both males and females disperse from their natal
+pack, either alone or in same-sex dispersing coalitions, and search for
+unrelated mates and a suitable territory to settle \citep{McNutt.1996,
+Cozzi.2020, Behr.2020}. During dispersal, wild dogs can cover several hundred
+kilometers and cross international borders \citep{Masenga.2016, Woodroffe.2019,
+Cozzi.2020}. Despite the importance of dispersal for the long-term viability of
+this species, little empirical information is available on habitat selection and
+potential movement barriers during dispersal. The few studies that have
+collected dispersal data have shown that dispersers quickly move over large
+distances, avoid human-dominated landscapes and areas densely covered by trees,
+but prefer proximity to water \citep{Masenga.2016, Woodroffe.2019, Oneill.2020,
+Cozzi.2020}.
 
@@
 by water in at least 10\% of the cases. We used the resulting map to also
-calculate a layer returning the distance to water. To reduce the influence of
-outliers in predicted permeability scores, we followed \cite{Squires.2013} and
-curtailed predicted scores between the 1\textsuperscript{st} and
-99\textsuperscript{th} percentile of their original values. To compare
-permeability across different regions, we normalized the permeability surface to
-a range between 0 (most impermeable) and 1 (most permeable), and we determined
-median permeability within and outside the KAZA-TFCA, within and outside
-formally protected areas, and within each of the five KAZA-TFCA countries.
+calculate a layer returning the distance to water. Because the delta only covers
+5\% of the KAZA-TFCA, we considered the use of a single water map to be
+appropriate. To reduce the influence of outliers in predicted permeability
+scores, we followed \cite{Squires.2013} and curtailed predicted scores between
+the 1\textsuperscript{st} and 99\textsuperscript{th} percentile of their
+original values. To compare permeability across different regions, we normalized
+the permeability surface to a range between 0 (most impermeable) and 1 (most
+permeable), and we determined median permeability within and outside the
+KAZA-TFCA, within and outside formally protected areas, and within each of the
+five KAZA-TFCA countries.
 
@@
 \textit{human influence}, beside the fixed covariates \textit{cos(ta)},
-\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). With respect to
-environmental covariates, dispersers avoided moving through water (\(\beta\) =
--0.53, 95\% CI -0.79 to -0.27) but selected for locations in its vicinity,
-although the latter effect was not significant (\(\beta\) = -0.33, 95\% CI =
--0.73 to 0.08). Dispersers avoided areas that were densely covered by trees
-(\(\beta\) = -0.31, CI = -0.47 to -0.15) and preferred areas covered by
-shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42). Finally, dispersers
-avoided areas that were influenced by humans (\(\beta\) = -0.45, 95\% CI = -0.82
-to -0.08). With the exepction of \textit{distance to water}, random effects
-revealed little variability between dispersal coalitions (Appendix A6).
-
-% The positive coefficient for \(cos(ta)\) (\(\beta\) = 0.10; 95\% CI = 0.05 to
-% 0.15) suggested that the uniform distribution for turning angles produced too
-% little directionality. Similarly, the positive coefficient for \(log(sl)\)
-% (\(\beta\) = 0.12, 95\% CI = 0.02 to 0.22) indicated that our fitted gamma
-% distribution produced step-lengths slightly shorter than the true
-% selection-free gamma distribution would have. The estimate for \(sl\), on the
-% other hand, was insignificant (\(\beta\) = 0.03, 95\% CI = -0.09 to 0.15).
+\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). Dispersers
+avoided moving through water (\(\beta\) = -0.53, 95\% \(CI\) -0.79 to -0.27) but
+selected for locations in its vicinity, although the latter effect was not
+significant (\(\beta\) = -0.33, 95\% \(CI\) = -0.73 to 0.08). Dispersers avoided
+areas that were densely covered by trees (\(\beta\) = -0.31, \(CI\) = -0.47 to
+-0.15) and preferred areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\%
+\(CI\) = 0.07 to 0.42). Finally, dispersers avoided areas that were influenced
+by humans (\(\beta\) = -0.45, 95\% \(CI\) = -0.82 to -0.08). Except for
+\textit{distance to water} (\(SD\)\textsubscript{RandomEffect} = 0.57), we
+observed little variation between dispersal coalitions' relative selection
+strengths (\(SD\)\textsubscript{RandomEffect} \(<\) 0.22 for all other
+covariates, see also Figure S7).
 
@@
 necessary, as the parametrized iSSF model can be used as mechanistic movement
-model to simulate dispersal events from known source points, yet without
-restricting the domain of potential end points \citep{Signer.2017}.
-Consequently, movement corridors would emerge more naturally as the result of a
-myriad of simulated dispersal events \citep{Zeller.2020}.
+model to simulate dispersal from known source points, yet without restricting
+the domain of potential end points \citep{Signer.2017}. Consequently, movement
+corridors would emerge more naturally as the result of a myriad of simulated
+dispersal events \citep{Allen.2016, Zeller.2020}.
 
 Besides estimating corridors, individual based simulations may be used to
-generate permeability surfaces \citep{Avgar.2016, Signer.2017}. Such surfaces
-have been shown to reduce the risk of overestimating \(w(x)\) and conesquently
-connectivity, particularly in areas that lie far from suitable habitats
-\citep{Signer.2017}. While a simulation-based approach is conceptually
-straightforward, computational requirements for such a large spatial extent as
-the KAZA-TFCA are very high, making the use of this approach challenging. We
-urge future studies to optimize the simulation of corridors from iSSFs that
-capture a more mechanistic model of dispersal.
+generate permeability surfaces \citep{Avgar.2016, Signer.2017}. Such simulation
+based surfaces have been shown to reduce the risk of overestimating permeability
+(\(w(x)\)) and conesquently connectivity, particularly in areas that lie far
+from suitable habitats \citep{Signer.2017}. While a simulation-based approach is
+conceptually straightforward, computational requirements for such a large
+spatial extent as the KAZA-TFCA are very high, making the use of this approach
+challenging. We therefore urge future studies to optimize the simulation of
+movement from iSSFs to capture a more mechanistic model of dispersal.
 

### 2021-01-04 Chapter_0/04_Manuscript/Manuscript.tex
Implemented minor revisions

@@
 \ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
-for each raster cell. Because our representation of water was dynamic we
+for each raster cell. Because our representation of water was dynamic, we
 collapsed all dynamic water maps into a single map using areas that were covered
@@
 \subsection{Least-Cost Paths and Corridors}
-To identify movement corridors of dispersing wild dogs, we specified source
-points and calculated factorial least-cost paths (LCPs) as well as factorial
-least-cost corridors (LCCs) among them \citep{Elliot.2014}. We generated source
-points by overlaying the study area with a regular grid of points spaced at 100
-km. We only considered those points that fell within protected areas \(>\) 700
-km\textsuperscript{2}, which conforms with home-range requirements of African
-wild dogs \citep{Pomilia.2015}. Finally, we defined centroids as source points
-for those protected areas \(>\) 700 km\textsuperscript{2} that were not assigned
-any source points from the regular grid. Because wild dogs residing outside of
-protected areas are rare and unviable \citep{VanDerMeer.2014}, we consider our
-selection of source points within protected areas to be appropriate. In total,
-we generated 68 source points, which resulted in 2'278 unique pairwise
+To identify potential movement corridors of dispersing wild dogs, we specified
+source points and calculated factorial least-cost paths (LCPs) as well as
+factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To specify
+source points, we applied the omnidirectional \textit{go-through} approach
+proposed by \citep{Koen.2014} and placed 68 regularly spaced source points along
+the map border. This approach has been shown to reduce potential biases arising
+from the selection of inappropriate source locations \citep{Koen.2014,
+Pitman.2017}. To assess the sensitivity of our results with respect to the
+location of source points, we also applied an alternative method and placed 68
+source points within protected areas that are large enough to sustain viable
+wild dog populations (further details and corresponding results in Appendix
+A.9). In either case, the 68 source points resulted in 2'278 unique pairwise
 combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
 LCPs and LCCs between source points using the R-package \textit{gdistance}
-(further details in Appendix A.7). After computation, we tallied overlapping
+(details in Appendix A.6). After computation, we tallied overlapping
 LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
@@
 strengths (\(SD\)\textsubscript{RandomEffect} \(<\) 0.22 for all other
-covariates, see also Figure S7).
+covariates, see also Figure S8).
 
@@
     \end{minipage}
-    \caption{(a) Source points (black dots) and corresponding least-cost paths
-    leaving from protected areas (dark grey borders). Note that only contiguous
-    protected areas covering more than 700 km\textsuperscript{2} are depicted.
-    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
-    dashed black lines delineate country-borders. (b) Least-cost corridors
-    between the same source points as illustrated in subfigure (a). For ease of
-    spatial reference, we also labeled some national parks (NPs, in dark-grey).}
+    \caption{(a) Source points (black semicircles along the map border) and
+    corresponding least-cost paths between them. Continuous thin black lines
+    indicate the borders of the KAZA-TFCA, whereas dashed black lines delineate
+    country-borders. (b) Least-cost corridors between the same source points as
+    illustrated in subfigure (a). For ease of spatial reference, we also labeled
+    some national parks (NPs, in dark-grey).}
     \label{LeastCost}

### 2021-01-04 Chapter_0/04_Manuscript/Manuscript_Support.tex
Implemented minor revisions

@@
 %------------------------------------------------------------------------------
-%  Appendix S6: Random Effects Variability
-%------------------------------------------------------------------------------
-\newpage
-\section{Habitat Selection Model: Random Effects}
-\begin{figure}[hbtp]
-  \includegraphics[width = \textwidth]{99_RandomEffects}
-  \caption{Plot of random effects showing variability across dispersal
-  coalitions. The coalition ID refers to the individual in the dispersal
-  coalition that was equipped with a GPS collar.}
-  \label{RandomEffects}
-\end{figure}
-
-%------------------------------------------------------------------------------
-%  Appendix S7: Least Cost Paths and Least Cost Corridors
+%  Appendix S6: Least Cost Paths and Least Cost Corridors
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%  Appendix S8: AIC Model Selection
+%  Appendix S7: AIC Model Selection
 %------------------------------------------------------------------------------
@@
 
+%------------------------------------------------------------------------------
+%  Appendix S8: Random Effects Variability
+%------------------------------------------------------------------------------
+\newpage
+\section{Habitat Selection Model: Random Effects}
+\begin{figure}[hbtp]
+  \includegraphics[width = \textwidth]{99_RandomEffects}
+  \caption{Plot of random effects showing variability across dispersal
+  coalitions. The coalition ID refers to the individual in the dispersal
+  coalition that was equipped with a GPS collar.}
+  \label{RandomEffects}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%  Appendix S9: Alternative Source Points
+%------------------------------------------------------------------------------
+\newpage
+\section{Alternative Source Points}
+Because we were interested in assessing the sensitivity of our results with
+respect to the location of source points, we implemented two competing
+approaches to select source points. In a first run, we applied an
+omnidirectional \textit{go-through} approach as proposed by \cite{Koen.2014} and
+applied by \cite{Pitman.2017}. That is, we distributed 68 equally spaced source
+points along the map border and we calculated least-cost paths and least-cost
+corridors for all pairwise combinations of them. In contrast to other methods,
+this approach has been shown to be less subjective and less sensitive to the
+selection of unreasonable source points. Furthermore, such an approach enables
+the identification of dispersal routes that do not necessarily connect patches
+inhabited by the species. In a second run, we followed the approach described in
+\cite{Elliot.2014} and overlayed the study area with a regular grid of points
+spaced 100 km apart. From this regular grid we only considered those points that
+fell within protected areas \(>\) 700 km\textsuperscript{2}, which conforms with
+home-range requirements of African wild dogs \citep{Pomilia.2015}. Finally, we
+defined centroids as source points for those protected areas \(>\) 700
+km\textsuperscript{2} that were not assigned any source points from the regular
+grid. Because wild dogs residing outside of protected areas are rare and
+unviable \citep{VanDerMeer.2014}, we consider such a selection of source points
+within protected areas to be reasonable. The second procedure also resulted in
+68 source points, implying that we calculated 2'278 unique pairwise combinations
+and therefore 2'278 unique LCPs and LCCs per approach.
+\Cref{AlternativeSourcePoints} illustrates the resulting least-cost paths and
+least-cost corridors and highlights that both methods produce qualitatively
+similar results. In fact, map values of \Cref{AlternativeSourcePoints} (b1) and
+(b2) correlate with \(r\) = 0.86. Nevertheless, one can see that connectivity
+following the first approach gets somewhat  ``stretched'' towards the map
+borders, which is expected given that the source points are all located at the
+borders. Regardless of the method, we find that the major dispersal corridors
+run within the borders of the KAZA-TFCA and that mainly impenetrable landscape
+remains beyond its borders.
+
+\begin{figure}[hbtp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_LeastCost_Suppl}
+  \caption{Comparison of least-cost paths (a) and least-cost corridors (b) using
+  different source points. In subfigures (a1) and (b1), 68 source points were
+  placed along the map border (black and white semicircles). In subfigures (a2)
+  and (b2), on the other hand, source points (black and white circles) were
+  placed in protected areas (gray) that were large enough to sustain viable wild
+  dog populations (i.e. larger than 700 km\textsuperscript{2},
+  \citep{Pomilia.2015}). Continuous thin black lines indicate the borders of the
+  KAZA-TFCA, whereas dashed black lines delineate country-borders. For ease of
+  spatial reference, we also labeled some national parks (NPs, in dark-grey).}
+  \label{AlternativeSourcePoints}
+  \end{center}
+\end{figure}
+
 \newpage

### 2021-01-18 Chapter_0/04_Manuscript/Manuscript.tex
Revision No. 2 implemented

@@
 source points and calculated factorial least-cost paths (LCPs) as well as
-factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To specify
-source points, we applied the omnidirectional \textit{go-through} approach
-proposed by \citep{Koen.2014} and placed 68 regularly spaced source points along
-the map border. This approach has been shown to reduce potential biases arising
-from the selection of inappropriate source locations \citep{Koen.2014,
-Pitman.2017}. To assess the sensitivity of our results with respect to the
-location of source points, we also applied an alternative method and placed 68
-source points within protected areas that are large enough to sustain viable
-wild dog populations (further details and corresponding results in Appendix
-A.9). In either case, the 68 source points resulted in 2'278 unique pairwise
-combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
-LCPs and LCCs between source points using the R-package \textit{gdistance}
-(details in Appendix A.6). After computation, we tallied overlapping
-LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
-discrete paths, whereas LCCs return continuous corridors, we present both
-methods. R-codes showcasing the main steps for sections 2.5 to 2.6 can be
-downloaded from GitHub
+factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To select
+source points, we followed the omnidirectional \textit{go-through} approach
+proposed by \cite{Koen.2014} and placed 68 regularly spaced source points along
+the map border \citep{Koen.2014, Pitman.2017}. While this approach tends to
+identify high connectivity towards the map boundaries, it reduces potential
+biases caused by the selection of unreasonable source points \citep{Koen.2014}.
+To assess the sensitivity of our results with respect to the location of source
+points, we reran the same analysis using 68 source points located within
+protected areas that are large enough to sustain viable wild dog populations
+(further details and corresponding results in Appendix A.9). In either case, the
+68 source points resulted in 2'278 unique pairwise combinations and therefore
+2'278 unique LCPs and LCCs. We computed factorial LCPs and LCCs between source
+points using the R-package \textit{gdistance} (details in Appendix A.6). After
+computation, we tallied overlapping LCPs and LCCs, respectively, into single
+connectivity maps. Because LCPs return discrete paths, whereas LCCs return
+continuous corridors, we present both methods. R-codes showcasing the main steps
+for sections 2.5 to 2.6 can be downloaded from GitHub
 (\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).
@@
     habitat selection model. Negative coefficients indicate avoidance of a
-    covariate, positive coefficients selection of a covariate. Whiskers
-    delineate the 95\%-CIs for estimated parameters. Significance codes: ** \(p
-    < 0.05\), *** \(p < 0.01\).(b) Results from the k-fold cross validation for
-    case-control studies. The left graph shows rank frequencies of
-    \textit{realized} steps according to predictions, whereas the right graph
-    shows rank frequencies of \textit{randomly selected} steps according to
-    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
-    resulting from 100 repetitions of the k-fold cross validation. The blue
-    smoothing line was fitted using a locally weighted polynomial regression and
-    serves to aid the eye in detecting the trends. Correlation coefficients
-    suggest that our prediction was significant and robust, evidenced by the
-    fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
-    \(\bar{r}_{s, random}\) did overlap and by the fact that there was strong
-    and significant correlation between ranks and associated frequency for
-    realized steps.}
+    covariate, positive coefficients selection of a covariate. \(ta\) = turning
+    angle, \(sl\) = step length. Whiskers delineate the 95\%-CIs for estimated
+    parameters. Significance codes: ** \(p < 0.05\), *** \(p < 0.01\). (b)
+    Results from the k-fold cross validation for case-control studies. The left
+    graph shows rank frequencies of \textit{realized} steps according to
+    predictions, whereas the right graph shows rank frequencies of
+    \textit{randomly selected} steps according to predictions. \(\bar{r}_s\)
+    indicates the mean correlation coefficient resulting from 100 repetitions of
+    the k-fold cross validation. The blue smoothing line was fitted using a
+    locally weighted polynomial regression and serves to aid the eye in
+    detecting the trends. Correlation coefficients suggest that our prediction
+    was significant and robust, evidenced by the fact that the confidence
+    intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did
+    overlap and by the fact that there was strong and significant correlation
+    between ranks and associated frequency for realized steps.}
     \label{PermeabilityResults}
@@
     \caption{(a) Source points (black semicircles along the map border) and
-    corresponding least-cost paths between them. Continuous thin black lines
-    indicate the borders of the KAZA-TFCA, whereas dashed black lines delineate
+    corresponding least-cost paths between them. Continuous black lines indicate
+    the borders of the KAZA-TFCA, whereas dashed black lines delineate
     country-borders. (b) Least-cost corridors between the same source points as
@@
 NP in Zambia, for which a disruption of its main and narrow dispersal corridor
-would result in considerable isolation of its subpopulations. We also revealed a
-potential southwards corridor between the Okavango-Linyanti ecosystem and the
-Central Kalahari National Park. \cite{Elliot.2014} and \cite{Cushman.2018}
-identified a similar corridor for dispersing lions, suggesting that upholding
-and protecting a link between those ecosystems is pivotal. Some areas through
-which the corridor runs are neither part of the KAZA-TFCA nor profit from any
-form of protection status. Human presence and activities along the national road
-that longitudinally traverses this corridor may limit functional connectivity
-\citep{Cozzi.2020}.
+or high disperser mortality due for example to human persecution or vehicle
+collision would result in considerable isolation of its subpopulations. We also
+revealed a potential southwards corridor between the Okavango-Linyanti ecosystem
+and the Central Kalahari National Park. \cite{Elliot.2014} and
+\cite{Cushman.2018} identified a similar corridor for dispersing lions,
+suggesting that upholding and protecting a link between those ecosystems is
+pivotal. Some areas through which the corridor runs are neither part of the
+KAZA-TFCA nor profit from any form of protection status. Human presence and
+activities along the national road that longitudinally traverses this corridor
+may limit functional connectivity \citep{Cozzi.2020}.
 

### 2021-01-18 Chapter_0/04_Manuscript/Manuscript_Support.tex
Revision No. 2 implemented

--- 2020-09-23 Chapter_0/04_Manuscript/Manuscript_Support.tex
@@
 
+%------------------------------------------------------------------------------
+%  Appendix S8: Random Effects Variability
+%------------------------------------------------------------------------------
+\newpage
+\section{Habitat Selection Model: Random Effects}
+\begin{figure}[hbtp]
+  \includegraphics[width = \textwidth]{99_RandomEffects}
+  \caption{Plot of random effects showing variability across dispersal
+  coalitions. The coalition ID refers to the individual in the dispersal
+  coalition that was equipped with a GPS collar.}
+  \label{RandomEffects}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%  Appendix S9: Alternative Source Points
+%------------------------------------------------------------------------------
+\newpage
+\section{Alternative Source Points}
+Because we were interested in assessing the sensitivity of our connectivity maps
+with respect to the location of source points, we implemented two competing
+approaches to select source points. In a first run, we applied an
+omnidirectional \textit{go-through} approach as proposed by \cite{Koen.2014} and
+applied by \cite{Pitman.2017}. That is, we distributed 68 equally spaced source
+points along the map border and we calculated least-cost paths and least-cost
+corridors for all pairwise combinations of them. In contrast to other methods,
+this approach has been shown to reduce unwanted biases arising from the
+selection of unreasonable source points. Furthermore, such an approach enables
+the identification of dispersal routes that do not necessarily run between
+occupied patches of suitable habitat. In a second run, we followed the approach
+described in \cite{Elliot.2014} and overlayed the study area with a regular grid
+of points spaced 100 km apart. From this regular grid we only considered those
+points that fell within protected areas \(>\) 700 km\textsuperscript{2}, which
+conforms with home-range requirements of African wild dogs \citep{Pomilia.2015}.
+Finally, we defined centroids as source points for those protected areas \(>\)
+700 km\textsuperscript{2} that were not assigned any source points from the
+regular grid. Because wild dogs residing outside of protected areas are rare and
+unviable \citep{VanDerMeer.2014}, we consider such a selection of source points
+within protected areas to be reasonable. The second approach also resulted in 68
+source points, implying 2'278 unique pairwise combinations and therefore 2'278
+unique LCPs and LCCs per approach. \Cref{AlternativeSourcePoints} illustrates
+the resulting least-cost paths and least-cost corridors and suggests that both
+methods produce qualitatively similar results. In fact, map values of
+\Cref{AlternativeSourcePoints} (b1) and (b2) correlate with \(r\) = 0.86.
+Nevertheless, one can see that connectivity towards the map borders is more
+pronounced using the first approach and that the final connectivity map in (b1)
+appears to be a  ``stretched'' version of the map in (b2). This is an expected
+result given that all source points were located at the map borders. Regardless
+of the method, we find that the major dispersal corridors run within the borders
+of the KAZA-TFCA and that mainly impenetrable landscape remains beyond its
+borders.
+
+\begin{figure}[hbtp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_LeastCost_Suppl}
+  \caption{Comparison of least-cost paths (a1 and a2) and least-cost corridors
+  (b1 and b2) using different source points. In subfigures (a1) and (b1) 68
+  source points were placed along the map border (black and white semicircles).
+  In subfigures (a2) and (b2), on the other hand, source points (black and white
+  circles) were placed in protected areas (gray) that were larger than 700
+  km\textsuperscript{2}, which conforms to home range requirements for viable
+  wild dog populations \citep{Pomilia.2015}. Continuous black lines indicate the
+  borders of the KAZA-TFCA, whereas dashed black lines delineate
+  country-borders. For ease of spatial reference, we also labeled some national
+  parks (NPs, in dark-grey).}
+  \label{AlternativeSourcePoints}
+  \end{center}
+\end{figure}
+
 \newpage

### 2021-02-04 Chapter_1/02_R-Scripts/01_DataCleaning/09_LandCover(Copernicus).r
back to old data

@@ -1,142 +1,0 @@
-################################################################################
-#### Cleaning Copernicus Land Cover Data
-################################################################################
-# Description: In this script I'll clean the land cover data for KAZA downloaded
-# from here: https://lcviewer.vito.be/download
-
-# Clear R's brain
-rm(list = ls())
-
-# Set working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load packages
-library(terra)      # To handle rasters
-library(raster)     # To handle rasters
-library(tidyverse)  # For data wrangling
-
-# Make use of multiple cores
-beginCluster()
-
-################################################################################
-#### Stitching the Tiles
-################################################################################
-# Load the data
-dat <- dir(
-    path       = "03_Data/01_RawData/COPERNICUS"
-  , full.names = T
-  , pattern    = ".tif$"
-)
-dat <- lapply(dat, raster)
-
-# Merge the tiles
-merged <- do.call(merge, dat)
-
-################################################################################
-#### Cropping, Aggregating, and Simplifying the Stitched Raster
-################################################################################
-# Load the reference shapefile and raster
-s <- shapefile("03_Data/02_CleanData/00_General_Shapefile.shp")
-r <- raster("03_Data/02_CleanData/00_General_Raster.tif")
-
-# Crop the merged tiles to our reference shapefile
-merged <- crop(merged, s)
-
-# Store the merged object to file
-writeRaster(
-    x         = merged
-  , filename  = "03_Data/01_RawData/COPERNICUS/Copernicus.tif"
-  , overwrite = T
-)
-
-# Aggregate to coarser resolution
-fact <- res(r)[1] / res(merged)[1]
-coarse <- aggregate(merged, fact = round(fact), fun = modal)
-
-# Check out the distribution of values
-freq(coarse, useNA = "ifany")
-
-# Load land cover classes
-info <- read_csv("03_Data/01_RawData/COPERNICUS/LandCoverClasses.csv")
-info <- arrange(info, Code)
-
-# Not all classes are represented in the cropped study area
[…]

### 2021-02-04 Chapter_1/02_R-Scripts/01_DataCleaning/17_Floodmap(MODIS).r

@@ -1,0 +1,332 @@
+################################################################################
+#### Download and Classify MODIS MCD43A4 Imagery
+################################################################################
+# Description: This script allows to download, process and classify MODIS
+# imagery automatically.
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(raster)           # To handle raster data
+library(terra)            # To handle raster data quickly
+library(tidyverse)        # For data wrangling
+library(lubridate)        # To handle dates
+library(floodmapr)        # To floodmap
+library(pbmcapply)        # To run stuff in parallel with progressbar
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Are you running the script for validation purposes?
+validation <- F
+
+# Do you want to use dynamic floodmapping?
+dynamic <- T
+
+################################################################################
+#### Specify Dates
+################################################################################
+# Specify dates for which you want inundation maps
+dates <- c(
+    "2012-01-09"
+  , "2012-01-17"
+  , "2012-01-25"
+  , "2012-02-02"
+  , "2012-02-10"
+  , "2015-11-17"
+  , "2015-11-25"
+  , "2015-12-03"
+  , "2015-12-11"
+  , "2015-12-19"
+  , "2015-12-27"
+  , "2016-01-01"
+  , "2016-01-09"
+  , "2016-01-17"
+  , "2016-02-02"
+  , "2016-02-10"
+  , "2016-02-18"
+  , "2016-03-21"
+  , "2016-03-29"
+  , "2016-11-08"
+  , "2016-11-16"
+  , "2016-11-24"
+  , "2016-12-02"
+  , "2017-03-14"
+  , "2017-03-22"
+  , "2017-11-01"
+  , "2017-11-09"
+  , "2017-11-17"
+  , "2017-11-25"
+  , "2017-12-03"
+  , "2017-12-11"
+  , "2017-12-19"
+  , "2017-12-27"
+  , "2018-01-01"
+  , "2018-01-09"
+  , "2018-01-17"
+  , "2018-03-06"
+  , "2018-03-14"
+  , "2018-03-22"
+  , "2018-03-30"
+  , "2018-04-07"
+  , "2018-04-15"
+  , "2018-04-23"
+  , "2018-05-01"
+  , "2018-05-09"
+  , "2018-05-17"
+  , "2018-05-25"
[…]

### 2021-02-04 Chapter_1/02_R-Scripts/01_DataCleaning/23_Floodmap(ORI).r

@@ -1,121 +1,0 @@
-################################################################################
-#### Preparation of the Ori Floodmaps
-################################################################################
-# Description: Resampling and preparation of the ORI floodmaps. We will also
-# check for which dates we need to download additional floodmaps.
-
-# Clean environment
-rm(list = ls())
-
-# Change the working directory.
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# load packages
-library(tidyverse)    # For data wrangling
-library(raster)       # To handle raster data
-library(lubridate)    # To handle dates
-
-# Allow for multicore use
-beginCluster()
-
-################################################################################
-#### Compare Dispersal and Ori Dates
-################################################################################
-# Identify the ori files in the working directory
-files <- dir(
-    path        = "03_Data/01_RawData/ORI/01_Maps"
-  , pattern     = "*.tif$"
-  , full.names  = T
-)
-
-# Create a vector that identifies the dates for which we have ORI data
-ori_dates <- files %>%
-  basename() %>%
-  substr(start = 1, stop = 10) %>%
-  as.Date()
-
-# Put filenames of their respective date into a dataframe
-files <- data.frame(Filename = files, Date = ori_dates)
-
-# Load the disperser's tracks
-dispersal <- "03_Data/02_CleanData/00_General_Dispersers_POPECOL.csv" %>%
-  read_csv() %>%
-  subset(State == "Disperser")
-
-# Extract unique dates
-dis_dates <- unique(dispersal$Timestamp) %>% as.Date()
-
-# For all these dispersal dates we want to know the closest ORI dates
-closest1 <- as.Date(NA)
-closest2 <- as.Date(NA)
-for (i in 1:length(dis_dates)){
-  closest1[i] <- ori_dates[which(abs(dis_dates[i] - ori_dates) ==
-    min(abs(dis_dates[i] - ori_dates)))][1]
-  closest2[i] <- ori_dates[which(abs(dis_dates[i] - ori_dates) ==
-    min(abs(dis_dates[i] - ori_dates)))][2]
-}
-
-# Put the dates together
-dates <- data.frame(
-    Dispersal   = dis_dates
-  , ClosestORI1 = closest1
-  , ClosestORI2 = closest2
-  , Difference  = abs(dis_dates - closest1)
-)
-
-# Look at the result
-dates
-
-# Let's see how close we get with the ori data
[…]

### 2021-02-04 Chapter_1/02_R-Scripts/01_DataCleaning/24_ModisVegetationCover.r

@@ -1,0 +1,280 @@
+################################################################################
+#### Preparation of the Modis Vegetation Data
+################################################################################
+# Description: Stitching, resampling and preparation of the modis continuous
+# vegetation maps. We will again create "pseudodynamic" layers that take the
+# dynamic watermaps into account.
+
+# Clean environment
+rm(list = ls())
+
+# Change the working directory.
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# load packages
+library(tidyverse)    # For data wrangling
+library(raster)       # For handling spatial data
+library(terra)        # For handling spatial data
+library(gdalUtils)    # To stitch files
+library(pbmcapply)    # To run stuff in parallel
+
+################################################################################
+#### Extracting TIFF files
+################################################################################
+# Identify files as they were downloaded
+files <- dir(
+    path        = "03_Data/01_RawData/MODIS/MOD44B"
+  , pattern     = "hdf"
+  , full.names  = T
+)
+
+# Run a loop that takes each hdf dataset and extracts the layers of interest
+# (Percent Tree Cover, Percent Non-Tree Cover, Percent-Non-Vegetated)
+for (i in 1:length(files)){
+  subs  <- get_subdatasets(files[i])
+  for (j in 1:3){
+    part1 <- substr(subs, start = 67, stop = 72)
+    part2 <- substr(subs, start = 113, stop = nchar(subs))
+    part3 <- ".tif"
+    names <- paste0(part1, part2, part3)
+    names <- paste0(wd, "/03_Data/01_RawData/MODIS/MOD44B/", names)
+    gdal_translate(subs[j], dst_dataset = names[j])
+  }
+}
+
+################################################################################
+#### Stitching TIFFs together
+################################################################################
+# Load the reference raster
+r <- rast("03_Data/02_CleanData/00_General_Raster.tif")
+
+# Specify layers that need to be stitched
+patterns <- c(
+    "Tree_Cover"
+  , "NonVegetated"
+  , "NonTree_Vegetation"
+)
+
+# We also prepare the names of the merged layers
[…]

### 2021-02-04 Chapter_1/02_R-Scripts/02_Analysis/00_SSF(SSF).R

@@ -22,298 +22,244 @@ library(pbmcapply)    # For multicore abilities
 
 # Set seed for reproducability (need to change random sampler for parallel)
 set.seed(1234)
-RNGkind("L'Ecuyer-CMRG")
-mc.reset.stream()
-
-# Big loop over the two methods
-methods <- c("iSSF", "TiSSF")
-for (i in 1:length(methods)){
-
-  # Select method
-  method <- methods[i]
 
 ################################################################################
 #### Data Cleaning
 ################################################################################
+# Load the gps data
+data <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL.csv")
 
-  # Load the gps data
-  data <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL.csv")
+# We only need dispersers' data
+data <- subset(data, State == "Disperser")
 
-  # We only need dispersers' data
-  data <- subset(data, State == "Disperser")
+# Let's create a timestamp that is rounded to the nearest hour
+data$TimestampRounded <- round_date(data$Timestamp, "1 hour")
 
-  # Let's create a timestamp that is rounded to the nearest hour
-  data$TimestampRounded <- round_date(data$Timestamp, "1 hour")
+# Make sure there are no duplicates!
+table(duplicated(data[, c("DogName", "TimestampRounded")]))
 
-  # Make sure there are no duplicates!
-  table(duplicated(data[, c("DogName", "TimestampRounded")]))
+# Count number of dogs
+length(unique(data$DogName))
 
-  # Count number of dogs
-  length(unique(data$DogName))
+# Nest data by dog
+data <- data %>% group_by(DogName) %>% nest()
 
-  # Nest data by dog
-  data <- data %>% group_by(DogName) %>% nest()
-
-  # Resample to two (minimally) hours. Can't use 1 hourly fixes because step
-  # speeds are unproportionally large with 1 hour.
-  data$data <- suppressMessages(
-    pbmclapply(1:nrow(data)
-      , ignore.interactive = T
-      , mc.cores           = detectCores() - 1
-      , FUN                = function(x){
-        resFix2(data$data[[x]], hours = 2, start = 1, tol = 0.5)
-      }
-    )
+# Resample to two (minimally) 2 hours
+data$data <- suppressMessages(
+  pbmclapply(1:nrow(data)
[…]

### 2021-02-04 Chapter_1/02_R-Scripts/02_Analysis/03_MovementModel.R

@@ -1,0 +1,883 @@
+################################################################################
+#### Step Selection Function - Model Selection
+################################################################################
+# Description: Here, we'll refit the base model described in Hofmann (2021) to
+# verify that its still the "best" model for the new data. Afterwards, we'll
+# expand it to a proper movement model by allowing interactions between
+# environmental covariates and movement metrics.
+
+# Clear R's brain
+rm(list = ls())
+
+# Surpress scientific notation
+options(scipen = 999)
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(davidoff)     # Custom functions
+library(parallel)     # To run stuff in parallel
+library(corrplot)     # To plot correlations
+library(cowplot)      # For nice plots
+library(ggpubr)       # For nice plots
+library(glmmTMB)      # For modelling
+
+# Identify number of cores for parallel computing
+cores <- detectCores()
+
+################################################################################
+#### Loading Data
+################################################################################
+# Load data
+dat <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF_Extracted).csv")
+
+# Keep only desired columns
+dat <- dat %>% select(c(
+      id
+    , step_id_
+    , case_
+    , sl_
+    , ta_
+    , inactive
+    , Water
+    , DistanceToWater
+    , Trees
+    , Shrubs
+    , HumansBuff5000 = HumanInfluenceBuffer_5000
+))
+
+# We want to add the log of the step speed and the cosine of the turning angle
+# and calculate the sqrt of the DistanceToWater
+dat <- dat %>% mutate(
+    log_sl_             = log(sl_)
+  , cos_ta_             = cos(ta_)
+  , SqrtDistanceToWater = sqrt(DistanceToWater)
+)
+
+# Let's also move all movement metrics to the front
+dat <- dat %>% select(c(
+  id, step_id_, case_, sl_, log_sl_, ta_, cos_ta_, inactive, everything()
+))
+
+################################################################################
+#### Scaling Data
+################################################################################
[…]

### 2021-02-04 Chapter_2/02_R-Scripts/01_DataCleaning/11_Agriculture(Copernicus).r

@@ -1,0 +1,49 @@
+################################################################################
+#### Resampling and Cropping of the Copernicus Agriculture Data
+################################################################################
+# Description: We can use copernicus's agricultural fields to augment our
+# croplands layer. In this script I therefore extract agricultural fields so we
+# can merge them.
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(raster)   # To handle raster data
+
+# Make use of multiple cores
+beginCluster()
+
+# Import the Copernicus dataset
+crops <- raster("03_Data/01_RawData/COPERNICUS/Copernicus.tif")
+
+# Load the reference raster
+r <- raster("03_Data/02_CleanData/00_General_Raster.tif")
+
+# Keep only crops
+crops <- crops == 40
+
+# Aggregate the copernicus dataset to match the resolution of the reference
+# raster
+fact <- res(r)[1] / res(crops)[1]
+crops <- aggregate(crops, fact = round(fact), fun = max)
+
+# Resample the layer to match the reference raster
+crops_res <- resample(crops, r, "ngb")
+
+# Check NAs
+sum(is.na(values(crops_res)))
+
+# Save the result to file
+writeRaster(
+    x         = crops_res
+  , filename  = "03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_COPERNICUS.tif"
+  , overwrite = TRUE
+)
+
+# End cluster
+endCluster()

### 2021-02-04 Chapter_2/02_R-Scripts/02_Analysis/01_CovariateExtraction(Old).R

@@ -1,0 +1,658 @@
+############################################################
+#### Step Selection Function - Extraction of Covariates
+############################################################
+# Description: In this script we create regular tracks (either 4 or 24 hours)
+# and create random steps for the step selection function. Finally, we also
+# extract the desired covariates along each of the steps. Note that the script
+# needs to be run twice to get the 4 hourly and 24 hourly extractions. Also be
+# warned that the extraction takes a good amount of time to complete.
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load packages
+library(rgdal)
+library(raster)
+library(data.table)
+library(lubridate)
+library(amt)
+library(tidyverse)
+library(spatstat)
+library(rgeos)
+library(maptools)
+library(velox)
+library(parallel)
+library(davidoff)
+
+# Set seed
+set.seed(1234)
+
+# Load the merged dataset containing all GPS fixes
+data <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL.csv")
+
+# Do you want to extract for the 4 hourly or the 24 hourly fixes?
+# selection <- 24
+selection <- 4
+
+############################################################
+#### Data Cleaning (For 4 Hourly Fixes)
+############################################################
+# For this part we will use the GPS fixes from dispersers only
+tracks <- data %>%
+
+  # Subset to the dispersers only (4 hourly fixes are actually only available
+  # during dispersal)
+  filter(., State == "Disperser") %>%
+
+  # Ultimately, we will only be using fixes that were collected during the
+  # following hours: 3, 7, 15, 19, 23 (rounded to 15 minutes)
+  filter(.
+    , strftime(round_date(Timestamp, "30 minutes")
+      , tz      = "UTC"
+      , format  = "%H:%M:%S") %in%
+      c("03:00:00", "07:00:00", "15:00:00", "19:00:00", "23:00:00")) %>%
+
+  # We also remove popecols fixes before october 2016 because the fixrate scheme
+  # was slightly different before that. We still want to conserve the fixes from
+  # Abrahms' dispersers
+  filter(.,
+    Timestamp >= as.POSIXct('2016-10-01 10:00:00') |
+    DogName %in% c("Lupe", "Scorpion", "Stetson")
+  ) %>%
+
[…]

### 2021-02-04 Chapter_2/02_R-Scripts/02_Analysis/08_RasterizeSimulations.R

@@ -1,0 +1,487 @@
+################################################################################
+#### Rasterization of Simulated Dispersal Trajectories
+################################################################################
+# Description: In this script, we rasterize the simulated dispersal
+# trajectories. To achieve this, we'll first rasterize trajectories by their
+# source points. Afterwards, we can combine them them into a single "heatmap".
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/media/david/My Passport/Backups/WildDogs/15. PhD/00_WildDogs"
+setwd(wd)
+
+# Load required packages
+library(terra)            # For quick raster manipulation
+library(raster)           # For general raster manipulation
+library(tidyverse)        # For data wrangling
+library(rgeos)            # For manipulating vector data
+library(lubridate)        # For working with dates
+library(viridis)          # For nicer colors
+library(tictoc)           # To keep track of processing time
+library(pbmcapply)        # To show progress bar in mclapply calls
+library(tmap)             # For nice spatial plots
+library(Cairo)            # To store plots
+library(davidoff)         # Custom functions
+
+################################################################################
+#### Function to Rasterize Simulated Trajectory
+################################################################################
+# Function to rasterize trajectories of a desired source point after a desired
+# amount of time (i.e. after a desired number of steps)
+rasterizeSims <- function(
+      simulations = NULL      # Simulated trajectories
+    , steps       = 400       # How many steps should be considered
+    , sampling    = "Static"  # For which sampling method we want to rasterize
+  ){
+
+  # Create tracks that fulfill the above requirements
+  sub_traj <- sims2tracks(
+      simulations = simulations
+    , steps       = steps
+    , sampling    = sampling
+  )
+
+  # Remove data by creating spatial lines
+  sub_traj <- as(sub_traj, "SpatialLines")
+
+  # Assign correct CRS
+  crs(sub_traj) <- CRS("+init=epsg:4326")
+
[…]

### 2021-02-04 Chapter_2/02_R-Scripts/02_Analysis/09_CompareHeatmaps.R

@@ -1,0 +1,119 @@
+################################################################################
+#### Comparison of Heatmap of Simulated Trajectories and Least Cost Corridors
+################################################################################
+# Description: In this script, we are going to compare the rasterized simulate
+# trajectories (heatmaps) to the least-cost corridors that we predicted in an
+# earlier paper. Basically, we want to know how similar the heatmaps are to the
+# permeability and least-cost maps.
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/media/david/My Passport/Backups/WildDogs/15. PhD/00_WildDogs"
+setwd(wd)
+
+# Load required packages
+library(raster)     # For general raster manipulation
+library(tidyverse)  # For data wrangling
+library(davidoff)   # Custom functions
+
+################################################################################
+#### Prepare Data
+################################################################################
+# Reload rasterized maps. We have to load both, the tibble (containing some
+# useful information), as well as the rasters themselves. The reason the two
+# files are seperate is that in the tibble the heatmaps are not linked correctly
+# anymore.
+rasterized <- read_rds("03_Data/03_Results/99_RasterizedSimulationsBootstrap.rds")
+heatmaps <- stack("03_Data/03_Results/99_RasterizedSimulationsBootstrap.tif")
+
+# Remove unneccessary columns
+rasterized$filename <- NULL
+rasterized$heatmap2 <- NULL
+
+# Load least-cost corridors and the permeability map that we want to compare our
+# heatmaps to.
+perm <- raster("03_Data/03_Results/99_PermeabilityMap.tif")
+corr <- raster("03_Data/03_Results/99_LeastCostCorridors.tif")
+
+# Aggregate corridor- and permeability-map to same resolution as heatmaps
+perm <- aggregate(perm, fact = 10, fun = "mean")
+corr <- aggregate(corr, fact = 10, fun = "mean")
+
+# In the heatmaps there is one row missing. Let's extend them to match the
+# permeability and corridor maps. Note that we'll replace the missing row simply
+# with the last row containing values. Not the most beautiful solution but it
+# will do the job.
[…]

### 2021-02-04 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/SocialLandscape.R

@@ -1,0 +1,101 @@
+################################################################################
+#### Visualization of the Social Landscape
+################################################################################
+# Clear R's brain
+rm(list = ls())
+
+# Set a seed
+set.seed(123)
+
+# Load required packages
+library(raster)
+library(tidyverse)
+library(adehabitatHR)
+library(viridis)
+
+# Create an empty raster
+r <- raster(ncol = 30, nrow = 30, xmn = 0, xmx = 900, ymn = 0, ymx = 900)
+values(r) <- rnorm(mean = 0, sd = 1, n = ncell(r))
+
+# Simulate 3 points of attraction
+points <- sampleRandom(r, size = 3, sp = T)
+
+# Calculate distance
+distance <- distanceFromPoints(r, points)
+
+# Create likelihood surface (with a bit of randomness)
+like <- - distance + 75 * r
+like <- (like - minValue(like)) / (maxValue(like) - minValue(like))
+
+# Visualize it
+plot(like)
+
+# Function to simulate movement
+move <- function(cell = NULL, like = NULL, steps = NULL, directions = 8){
+  path <- c()
+  for (i in 1:steps){
+    path[i] <- cell
+    neighbors <- adjacent(like, cell, directions = directions, pairs = F)
+    values <- values(like)[neighbors]
+    probs <- values / sum(values)
+    cell <- sample(neighbors, prob = probs, size = 1)
+  }
+  return(path)
+}
+
+# Use it to simulate a couple of tracks
+tracks <- tibble(TrackID = 1:5)
+tracks$Coords <- lapply(1:5, function(x){
+  track <- move(cell = 500, like = like, steps = 500, directions = 16)
+  track <- SpatialPoints(xyFromCell(r, cell = track))
+  return(track)
+})
+
+# Create utilization distribution
+r <- extend(r, c(-200, 1100, -200, 1100))
+r <- disaggregate(r, fact = 5)
+tracks$UD <- lapply(tracks$Coords, function(x){
+  kernelUD(x, h = "href", grid = as(r, "SpatialPixels"))
+})
+
+# Get 95% Home range
+tracks$HR <- lapply(tracks$UD, function(x){
+  getverticeshr(x, percent = 95)
+})
+
+# Convert UDs to rasters
+tracks$UD <- lapply(tracks$UD, raster)
+
+# Plot 1
[…]

### 2021-02-17 Chapter_0/04_Manuscript/Manuscript.tex
Various updates

@@
 function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, least-cost corridors, Lycaon pictus, permeability surface,
-protected areas, wildlife management
+connectivity, African wild dog, movement
 
@@
 remained stationary, indicating that the coalition had successfully settled
-(Figure S1). In our analysis, we did not differentiate between male and female
-dispersing coalitions, for previous research found little differences between
-sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.
+(Figure S1 in Supporting Information). In our analysis, we did not differentiate
+between male and female dispersing coalitions, for previous research found
+little differences between sexes during dispersal \citep{Woodroffe.2019,
+Cozzi.2020}.
 
@@
 \section{Data Availability}
-GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication. R-code showcasing the main steps for sections 2.5 to
-2.6 can be downloaded from GitHub
+GPS movement data of dispersing coalitions is available in the Dryad Digital
+Repository doi:10.5061/dryad.dncjsxkzn \citep{Hofmann.2021}. R-code showcasing
+the main steps for sections 2.5 to 2.6 can be obtained through GitHub
 (\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

### 2021-02-17 Chapter_0/04_Manuscript/Manuscript_Support.tex
Various updates

@@
 % Create linenumbers
-\linenumbers
+% \linenumbers
 

### 2021-02-25 Chapter_1/04_Manuscript/Manuscript.tex
Worked on revision
back to old data
Various updates

@@
 %------------------------------------------------------------------------------
-\title{Step by Step: Using Integrated Step Selection Analysis to Simulate Wild
-Dog Dispersal and Assess Landscape Connectivity}
+\title{Step by Step: Using Step Selection Analysis to Simulate Dispersal and
+Assess Landscape Connectivity}
 
@@
 distance travelled between two consequtive GPS relocations
-\citep{Turching.1998}.
+\citep{Turchin.1998}.
 
@@
 and low activity, hence we did not include the interaction
-\(cos(ta\_):MainActivity\). We then ran stepwise modal forward selection absed
+\(cos(ta\_):MainActivity\). We then ran stepwise model forward selection based
 on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}) values and
-identified the most parsimonious movement model.
+identified the most parsimonious movement model. Although several models
+received an AIC weight above one, we decided to only consider the  ``best''
+model for simplicity. In either way, the models with positive weights contained
+almost identical covariates and an averaged model would have only given weak
+support to additional covariates.
 

### 2021-02-25 Chapter_1/02_R-Scripts/02_Analysis/04_BacktransformModel.R

@@ -1,0 +1,438 @@
+################################################################################
+#### Backtransform Scaled Model
+################################################################################
+# Description: Here, I want to backtransform the scaled model to retrieve
+# coefficients for the movement metrics on the untransformed scale. However,
+# this script is only supposed to make sure the function that I use to
+# backtansform the data actually works. I can verify this by fitting an unscaled
+# model and then compare the coefficients to a model with scaled covariates that
+# I backtransform.
+
+# Clear R's brain
+rm(list = ls())
+
+# Surpress scientific notation
+options(scipen = 999)
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(davidoff)     # Custom functions
+library(glmmTMB)      # For modelling
+
+################################################################################
+#### Loading Data
+################################################################################
+# Load step selection data
+dat <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF_Extracted).csv")
+
+# Keep only desired columns
+dat <- dat %>% select(c(
+      id
+    , step_id_
+    , case_
+    , sl_
+    , ta_
+    , inactive
+    , Water
+    , DistanceToWater
+    , Trees
+    , Shrubs
+    , HumansBuff5000 = HumanInfluenceBuffer_5000
+))
+
+# This is a crucial step. In the original model, step lengths were included in
+# meters. This resulted in convergence issues, which I overcame by standardizing
+# also the movement metrics. Alternatively, I could have also converted step
+# lengths to kilometers, yet I realized that the model fit was better using
+# standardized step lengths. Here, I will convert the step lengths to kilometers
+# which will allow me to refit the full model and see if my
[…]

### 2021-02-25 Chapter_1/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/MovementModel_conflict-20210222-111048.R

@@ -1,0 +1,278 @@
+################################################################################
+#### Plot Movement Model Results
+################################################################################
+# Description: A simple plot of the Dispersal Durations
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)  # For data wrangling
+library(lubridate)  # To handle dates nicely
+library(ggpubr)     # For nice plots
+library(davidoff)   # Custom functions
+library(glmmTMB)    # To handle model results
+library(cowplot)    # To add image to ggplot
+library(lemon)      # For nice axis labels
+library(imager)     # To import images
+library(viridis)    # For nice colors
+library(raster)     # To create rasters (visInt)
+library(ggdark)     # Dark ggplot theme
+
+################################################################################
+#### Prepare Data
+################################################################################
+# Old directory
+# wd <- "/media/david/My Passport/Backups/WildDogs/15. PhD/00_WildDogs"
+# setwd(wd)
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load the movement model
+move.mod <- readRDS("03_Data/03_Results/99_MovementModel.rds")
+
+# Select best model
+best <- move.mod$Model[[1]]
+
+# Summary of the best model
+summary(best)
+
+# Calculate Confidence intervals
+coeffs <- getCoeffs(best, pvalue = TRUE)[-1, ] %>%
+  mutate(
+      LCI_90 = Coefficient - 1.645 * SE
+    , UCI_90 = Coefficient + 1.645 * SE
+    , LCI_95 = Coefficient - 1.96 * SE
+    , UCI_95 = Coefficient + 1.96 * SE
+    , LCI_99 = Coefficient - 2.575 * SE
+    , UCI_99 = Coefficient + 2.575 * SE
+  )
+
+# Add stars indicating the significance
+coeffs$Significance <- sapply(1:nrow(coeffs), function(x){
+  if (coeffs$pvalue[x] <= 0.01){
+    return("***")
[…]

### 2021-03-08 Chapter_0/04_Manuscript/Manuscript.tex
Minor modifications

@@
 
-\section{Methods}
+\section{Materials and Methods}
 \subsection{Study Area}
@@
 
-\section{Data Availability}
-GPS movement data of dispersing coalitions is available in the Dryad Digital
+\section{Data Availability Statement}
+GPS movement data of dispersing coalitions available via the Dryad Digital
 Repository doi:10.5061/dryad.dncjsxkzn \citep{Hofmann.2021}. R-code showcasing

### 2021-03-22 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

@@
 
-\textbf{Running Title:} Simulating Wild Dog Dispersal.
+\textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
+Assess Landscape Connectivity
 
@@
 
-\textbf{Keywords:} dispersal, habitat selection, integrated step selection
-function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, least-cost corridors, Lycaon pictus, permeability surface,
-protected areas, wildlife management
+\textbf{Keywords:} dispersal, simulation, integrated step selection,
+Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
+pictus
 
@@
 \begin{abstract}
-This is the abstract.
+
+During the past two decades, least-cost analysis and circuit theory have been
+the two workhorses in connectivity analyses. This is largely thanks to their
+ease of use and intuitive nature which has facilitated the application of the
+methods across a broad range of animals. However, recent innovations in movement
+ecology have brought forward new methods that are similarly intuitive but
+provide several advantages over traditional connectivity modelling techniques.
+
+Here, we propose the simulation of dispersal trajectories as a much more generic
+way to identify dispersal barriers and potential movement corridors. Moreover,
+we show how such simulations can be used to identify zones of high
+human-wildlife conflicht, as well as to pinpoint potential "dispersal traps"
+where dispersing individuals get trapped due to unavailability of suitable
+alternatives. Finally, we discuss the benefits and pitfalls of dispersal
+simualtions.
 \end{abstract}
@@
 \section{Introduction}
+% Dispersal & Connectivity
+Dispersal of individuals is an important process governing the dynamics wild
+animal populations that are distributed in space \citep{Clobert.2012}. It is
+defined as the movement of individuals from their natal location to the site of
+first reproduction \cite{Howard.1960} and allows to avoid inbreeding
+\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012}, to rescue small and
+unviable populations \citep{Brown.1977}, and to promote the colonization of
+unoccupied habitats \citep{Hanski.1998, MacArthur.2001}. Information on movement
+behavior during dispersal and knowledge about the factors that limit dispersal
+is therefore critical for a comprehensive understanding of landscape
+connectivity and population viability \citep{Baguette.2013, Vasudev.2015}.
+
+% Technological Advancements
+Thanks to novel technologies developed over the past decades, particularly of
+GPS/Satellite radio-collars, the study of dispersal has accelerated
+\citep{Jonsson.2016, Williams.2019}. Additionally, the advent of publicly
+accessible satellite imagery and sophisticated remote sensing techniques to
+represent the physical landscape through which individuals disperse has pushed
+the field towards a  ``golden age of animal tracking'' \citep{Kays.2015}.
+Concurrently, the increased availability of large amounts of empirical data and
+an increased computational power have led to the development of numerous
+analytical methods to study movement behavior during dispersal
+\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012, Elliot.2014,
+Diniz.2019}.
+
+% Measuring Connectivity
+Resource selection functions \citep{Boyce.2002} and derived methods such as step
+selection functions \citep{Fortin.2005} and path selection functions
+\citep{Cushman.2010} have proven particularly useful for studying animal
+movement and landscape connectivity. These methods allow estimating habitat
+preferences of the focal species by comparing covariates at locations visited by
+the animal to the scame covariates at locations available to, but not visited by
+the animal. The so estimated preferences can then be used to predict a
+permeability surface, indicating the expected ease at which an animal can
+traverse a given area \citep{Zeller.2012}. Utimately, the permeability surface
+serves as foundation to a connectivity model and to reveal movement corridors.
+In this regard, the literature has long relied on least-cost methods
+\citep{Adriaensen.2003} and circuit theory \citep{McRae.2006, McRae.2008}.
+Unfortunately, both of these approaches make several limiting assumptions that
+are hardly ever met in reality, particularly by dispersing animals. For
+instance, to calculate a least-cost path one has to pre-define start- and
+end-points. This implicitly assumes that animals have a preconceived end-point
+in mind and choose a cost-minimizing route accordingly. To calculate a
+least-costly route, one also needs to assume that animals have an infinite
+perceptual range and that a route can always be realized, regardless of the
+imposed costs. In contrast to least-cost methods, circuit theory assumes a
+perceptual range of a single pixel, which is equally unrealistic and rarely
+captures the scale of selection (cite someone). Moreover, directional biases
+cannot be captured by circuit theory as a random walk is implicitly assumed. In
+reality, however, dispersers often depict directional persistence and move in an
+almost straight line to cover much ground very quickly. A mix of the two methods
+can be seein in least-cost corridor approaches or randomized shortest paths have
+been proposed to mitigate these issues by allowing animals to deviate from the
+``optimal'' least-costly route \citep{Pinto.2009, Panzacchi.2016,
+VanMoorter.2021}, they still fail to render a more realistic perceptual range of
+the animal. In addition, neither of the two methods render the temporal
+dimension of movement. Thusfar, only few have explored the usefulness of
+individual based movement models when estimating landscape connectivity
+\citep{Kanangaraj.2013, Hauenstein.2019, Zeller.2020}.
+
+Nevertheless, collecting data on dispersers remains logistically challenging due
+to dispersers' inherent mobility and due to the fact that dispersers only make
+up a small fraction of the entire population, thereby making it difficult for
+scientists to target apporpriate individuals for their study.
+
 An animals movement trajectory can be seen as the result of an interplay between
@@
 
+However, these models can not account for a correlation between turning angles
+and step lengths, unless the two are sampled jointly from a copula distribution
+(Hodel et al.).
+
+A relatively recent addition to the step selection family has been integrated
+step selection analysis.... Thanks to recently published papers and packages,
+the suite of step selection functions has become accessible to most researchers
+working with relocation data. Despite this, the use of such methods rarely
+extends beyond assessing relative habitat preferences. However, once
+parametrized, integrated step selection models could potentially serve as
+powerful framework to simulate animal movement and assess landscape
+connectivity. Here, we exemplify such analyses and show how ... coud be applied
+to investigate ...
+
+In contrast to Signer et et al, our focus lied on the transient utilisation of
+the landscape.
+
 In addition, individual based simulations will allow to explicitly model
@@
 potential of re-establishing dispersal routes for many of its protected species,
-including the african wild dog \textit{Lycaon pictus}. This species has
-experienced severe population declines caused by human induced mortality and
-deadly diseases. In result, the species currently marks the KAZA's most
-endangered large carnivore and has been a assigned a very high conservation
-priority. Importantly, due to their inherent mobility and intrinsic need for
-vast undisturbed landscapes, AWDs have been proposed as surrogate species for
-landscape connectivity (see recent paper on multispecies connectivity).
-Nevertheless, the species has received little attention in the connectivity
-literature, mainly due to the difficulty in observing wild dog dispersal. In a
-previous paper, we addressed this issue and developed a habitat selection model
-based on which we predicted landscape connectivity using least-cost corridor
-analysis. We now expand on this knowledge and develop a more detailed movement
-model of dispersing wild dogs. We then use this model to simulate thousands of
-dispersers moving throughout the KAZA. Based on said simulations, we compute
-heatmaps and identify potential dispersal hotspots and compare them to the
-dispersal routes identified in (Hofmann ....). We also apply the data to feed a
-network-analysis, based on which we determine network-metrics pertinent to
-landscape connectivity.
+including the african wild dog \textit{Lycaon pictus}.  Persecution by humans,
+habitat loss, and reduced connectivity are major causes of the decline of the
+species \citep{Woodroffe.2012}. In result, the species currently marks the
+KAZA's most endangered large carnivore and has been a assigned a very high
+conservation priority. Importantly, due to their inherent mobility and intrinsic
+need for vast undisturbed landscapes, AWDs have been proposed as surrogate
+species for landscape connectivity (see recent paper on multispecies
+connectivity). Nevertheless, the species has received little attention in the
+connectivity literature, mainly due to the difficulty in observing wild dog
+dispersal. In a previous paper, we addressed this issue and developed a habitat
+selection model based on which we predicted landscape connectivity using
+least-cost corridor analysis. We now expand on this knowledge and develop a more
+detailed movement model of dispersing wild dogs. We then use this model to
+simulate thousands of dispersers moving throughout the KAZA. Based on said
+simulations, we compute heatmaps and identify potential dispersal hotspots and
+compare them to the dispersal routes identified in (Hofmann ....). We also apply
+the data to feed a network-analysis, based on which we determine network-metrics
+pertinent to landscape connectivity.
 
@@
 hydrogeographical feature in its center is the Okavango Delta, the earth's
-largest inland delta and home to a vast diversity of mammal species. The delta
-and its surroundings are considered a stronghold for African wild dogs which may
-act as a source for the recolonization of surrounding habitats. The wet season
-in the region lasts from November to March, albeit the main floodwaters reach
-the Delta between July and August after descending through the Angolian
-highlands \citep{McNutt.1996, Wolski.2017}.
+largest inland delta. The delta and its surroundings are considered a stronghold
+for African wild dogs and may act as a source for the recolonization of
+surrounding habitats. The wet season within our study area lasts from November
+to March and is out of phase with the main flooding of the Okavango Delta which
+peaks between July and August \citep{McNutt.1996, Wolski.2017}.
 
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_StudyArea}
+    \includegraphics[width = \textwidth]{99_StudyArea.png}
     \caption{Study area}
@@
 \subsection{GPS Relocation Data}
-Between 2011 and 2020, we collected GPS relocation data on free-ranging wild
-dogs inhabiting the Okavango Delta in northern Botswana. We identified potential
-dispersers based on criteria reported in \cite{Behr.2020}, immobilized them
-according to protocols described in \cite{Osofsky.1996}, and fitted them with
-GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
-Berlin}). Handling and collaring of all individuals was carried out and
-supervised by a Botswana-registered wildlife veterinarian. Of all collared
-individuals, a total of 16 individuals dispersed, each in a same-sex coalition
-(7 female and 9 male coalitions). When we observed dispersal we remotely
-increased the GPS fixrate from 24 to 4 hours. Collected relocations were then
-regularly transmitted via Iridium satellite system to a base station. This
-allowed remote tracking of dispersers even if they left the main study area or
-crossed international borders. To effectively distinguish between resident and
-dispersersing individuals, we applied the net-squared displacement metric. This
-metric measures the Euclidean distance of a collared individual to a reference
-point \citep{Borger.2012}. In our case, these reference points resembled the
-center of the dispersing coalition's natal home range. Hence, dispersal was
-deemed to have started when an individual left its natal home range and ended
-when the individual became stationary again. Previous research found no
-differences between males and females during dispersal \citep{Woodroffe.2019,
-Cozzi.2020}, so we did not distinguish sexes in our analyses. GPS relocations
-were then converted to steps, where a step represented the straight-line
-distance travelled between two consequtive GPS relocations
-\citep{Turchin.1998}.
+Between 2011 and 2020, we collected GPS relocation data on dispersing wild dogs
+from a free-ranging wild dog population inhabiting the Okavango Delta in
+northern Botswana. We identified potential dispersers based on age, number of
+same‐sex siblings, pack size, and presence of unrelated individuals of the
+opposite sex in their pack \citep{McNutt.1996, Behr.2020}. We immobilized
+selected individuals using a cocktail of Ketamine/Xylazine/Atropine
+\citep{Osofsky.1996, Cozzi.2020} which we injected with a dart, fired from a
+CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). The
+corresponding protocols have been approved by the Ministry of Environment,
+Natural Resources Conservation and Tourism of Botswana (permit EWT 8/36/4
+XXXVI). After immobilization, individuals were fitted with GPS/Satellite radio
+collars (\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) that included
+an automated drop-off mechanism. Handling and collaring of all individuals was
+carried out and supervised by a Botswana-registered wildlife veterinarian. All
+of the immobilized individuals quickly rejoined their pack after the procedure.
+From all collared individuals, 16 individuals eventually dispersed in separate
+same-sex coalitions and their trajectories were successfully recorded (7 female
+and 9 male coalitions).
+
+During dispersal, collars were programmed to record a GPS fix every 4 hours.
+Recorded relocations were regularly transmitted over the Iridium satellite
+system, which allowed remote tracking of individuals, even if they left the main
+study area and crossed international borders. To effectively distinguish between
+periods of residency and dispersal, we applied the net-squared displacement
+metric to the observed movement data. This metric measures the squared Euclidean
+distance of a collared individual to a reference point \citep{Borger.2012},
+which we set at the center of each disperser's natal home range. Hence,
+dispersal was deemed to have started when an individual left its natal home
+range and ended when the individual became stationary again. For the purpose of
+this study, we discarded andy data that was collected during residency. Because
+previous research could not find any differences between males and females
+during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish
+sexes for the sake of our analyses. Ultimately, we converted the collected GPS
+relocations to steps, where a step represented the straight-line distance
+travelled between two consequtive GPS relocations \citep{Turchin.1998}.
 
 \subsection{Covariates}
-To represent environmental covariates spatially, we prepared a set of raster
-layers depicting water-cover (dynamically updated), square rooted distance to
-water (dynamically updated), tree-cover, and shrub/grassland-cover. We also
-created a proxy for human influence, rendering anthropogenic pressures stemming
-from human-density, agricultural sites, and roads. We prepared all layers at a
-resolution of 250m by 250m for the extent of the entire KAZA-TFCA. A more
-detailed description of the derivation and preparation of each environmental
-covariate is given in \cite{Hofmann.2020}. Because wild dogs follow a diurnal
-activity pattern, we coded a binary variable indicating whether an observed step
-started during phases of main wild dog activity (07:00 a.m. and 07:00 p.m.) or
-low wild dog activity (anything else).
+To represent habitat covariates, we prepared a set of spatial raster layers
+depicting water-cover (dynamically updated), distance to water (dynamically
+updated), tree-cover, and shrub/grassland-cover. To render a decreasing impact
+with increasing distance, we square rooted the distance to water covariate
+layer. We also created a proxy for human influence, rendering anthropogenic
+pressures stemming from human-density, agricultural sites, and roads. We
+rendered all layers at a resolution of 250m by 250m for the entire extent of the
+KAZA-TFCA. Further details on the derivation and preparation of each
+environmental covariate is given in \cite{Hofmann.2020}.
+
+Besides habitat covariates, we also computed movement metrics that we used as
+movement covariates in our models. These covariates included the step length
+(\(sl\)), its natural logarithm (\(log(sl)\)), as well as the cosine of the
+relative turning angle (\(cos(ta)\)). Because wild dogs follow a diurnal
+activity pattern, we furthermore coded a binary variable indicating whether an
+observed step was realized during main activity (xx to xx) or low activity (xx
+to xx).
 
@@
 We used integrated step selection functions (iSSF) to parametrize a mechanistic
-movement model of dispersing wild dogs. In the iSSF framework, observed steps
-receive a selection score \(w(x)\) that depends on the covariates \(X\)
+movement model of dispersing wild dogs. In this framework, observed steps are
+contrasted to random steps that the animal could have realized, but decided not
+to. To compare steps, it is assumed that animals assign a relative selection
+score \(w(x)\) to each step. The score depends on the covariates \(X\)
 experienced along a step, as well as the animals preferences towards these
-covariates \(\beta\). In contrast to regular step selection analysis,
+covariates \(\beta\). The probability of a step being realized is then
+contingent on the step's selection score, as well as the selection scores of all
+other step in the same stratum. In contrast to regular step selection analysis,
 \textit{integrated} step selection analysis allows simulatneous inference on
 movement and habitat preferences of the studied animal. Moreover, potential
-interactions between habitat and movement preferences can be modelled. Thus, the
-method produces more accurate selection estimates and allows to use the
-resulting models as proper mechanistic movement models from which movement can
-be generated. To conduct iSSf analysis, we paired each observed step with 24
+interactions between habitat and movement preferences can be included to
+investigate how movement behavior depends on habitat characteristics. In result,
+the method produces less biased selection estimates and allows to apply the
+resulting model as a fully mechanistic movement model based on which movement
+can be simulated. To conduct iSSF analysis, we paired each observed step with 24
 random steps. Random steps basically resembled potential alternatives that the
@@
 step lengths from a gamma distirbution that was fitted to observed steps (scale
-= 6308, shape = 0.37). While the number of random steps is inversely
+= 6'308, shape = 0.37). While the number of random steps is inversely
 proportional to the sampling error \citep{Avgar.2016}, we found only minor
-changes in inference when sampling further random steps. Along each step we
-extracted spatial covariates using the \textit{velox} package. We also
-calculated step metrics, namely the natural logarithm of the step length
-\(log(sl\_\)) and the cosine of the turning angle \(cos(ta\_)\). We scaled all
-continuous covariates to a mean of zero and a standard deviation of one. We then
-used the r-package \textit{glmmTMB} to fit mixed effects conditional logistic
-regression models as proposed by \citep{Muff.2020}. Our movement model was built
-around a previously published habitat selection model of dispersing wild dogs.
-Because the earlier model was used to predict landscape permeability,
-interactions between environmental and spatial covariates were impossible to
-model. Our current movement model, in contrast, allowed the inclusion of
-additional interactions that describe differences in movement behavior induced
-through different environments. Hence, we started with the habitat model and
-iteratively increased model complexity by including additional interactions
-between movement metrics and environmental covariates. That is, we proposed all
-two-way interactions between spatial covariates and movement metrics. For
-instance, for the covariate \textit{water} we proposed the interactions
-Water:cos(ta\_) and Water:log(sl\_). Despite interactions between environmental
-factors and movement metrics, we also proposed the interaction
-\(log(sl\_):MainActivity\) to account for the diurnal activity pattern of wild
-dogs. A preliminary revealed little differences in turning angles during main
-and low activity, hence we did not include the interaction
-\(cos(ta\_):MainActivity\). We then ran stepwise model forward selection based
-on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}) values and
-identified the most parsimonious movement model. Although several models
-received an AIC weight above one, we decided to only consider the  ``best''
-model for simplicity. In either way, the models with positive weights contained
-almost identical covariates and an averaged model would have only given weak
-support to additional covariates.
+differences in model performance when sampling additional random steps and
+deemed 24 random steps sufficient. Along each step, we extracted spatial
+covariates using the \textit{velox} package and calculated the movement metrics
+\(sl\), \(log(sl\)), and \(cos(ta)\). To facilitate model convergence, we
+standardized all continuous covariates to a mean of zero and a standard
+deviation of one. We then used the r-package \textit{glmmTMB} to fit mixed
+effects conditional logistic regression models as proposed by \citep{Muff.2020}.
 
-\subsection{Dispersal Simulation}
-We used the most parsimonious movement model to simulate dispersing wild dogs
-departing from predefined source points. The simulation basically resembled an
-inverted iSSF function and was set up as follows. First, we determined a source
-point at which a disperser was initiated. To allow calculation of turning
-angles, we assumed a random initial orientation of the animal. Second, we
-proposed 25 random steps originating at the predefined source point. We
-generated random steps by sampling turning angles from a uniform distribution
-(\(-\pi, +\pi\)) and step lengths from a gamma distribution fitted to observed
-step lengths. To prevent unrealistically large steps, we capped the gamma
-distribution 35km, the farthest distance travelled in four hours according to
-our data. Third, along each random step we extracted environmental covariates
-calculated step metrics. Fourth, we applied the parametrized dispersal model to
-predict selection scores \(w(x)\) and to determine the probability of a step
-being realized. Fifth, we sampled one of the proposed based on their
-probabilities and calculated the animal's new position. We then repeated steps
-two to five until 2000 steps were realized. In case a proposed random step left
-our study extent, we removed the step from the set of random steps, thereby
-forcing the disperser to stay within the boundaries of the study area.
+Our movement model was and extension of the habitat model presented in Hofmann
+et al 2021. We will refer to it as the base model. Because the base model was
+used to predict landscape permeability, no interactions between the habitat and
+movement metrics were modeled. Thus, we expanded the base model and by proposing
+additional interactions between movement metrics and environmental covariates.
+For this, we started with the base model and iteratively increased model
+complexity by adding all proposed two-way interactions between habitat
+covariates and movement covariates. For instance, for the covariate
+\textit{water} we proposed the interactions \(Water:log(sl)\),
+\(Water:log(sl)\), and \(Water:cos(ta)\). Furthermore, we proposed the
+interactions \(sl:cos(ta)\) and \(log(sl):cos(ta)\) to account for a correlation
+between turning angles and step lengths, as well as the interactions
+\(sl:MainActivity\) and \(log(sl):MainActivity\) to account for the fact that
+step lengths diffuer due to the diurnal activity pattern. We then ran stepwise
+model forward selection based on Akaike's Information Criterion (AIC,
+\citealp{Burnham.2002}) values and identified the most parsimonious movement
+model. Although several models received an AIC weight above one, we decided to
+only consider the  ``best'' model for simplicity. In either way, the models with
+positive weights contained almost identical covariates and an averaged model
+would have only given weak support to additional covariates.
 
 \subsection{Source Points}
-Simulations from step selection functions are known to be sensitive to the
-location of source points, which is why we followed a twofold approach to sample
-release points across the KAZA-TFCA. In a first approach, we used the same 68
-source points between which we already computed least-cost paths and least-cost
-corridors (see xx). These source points were regularly spaced 100 km apart,
-located within protected areas (\(>\) 700 km\textsuperscript{2}). Because of
-the regulary spacing, not all protected areas received source points. Hence, we
-placed additional source points at the center of such protected areas. At each
-of the generated source point we released 1000 dispersers, implying a total of
-68'000 simulated dispersers. In a second approach, we used the same 68 source
-points to define catchment areas inside protected areas. That is, for each of
-the 68 source points we defined its voronoi polygon within the point's protected
-area. Within these catchment areas we then randomly sampled 1000 new source
-points from which we released another 68'000 dispersers. Overall, we simulated
-134'000 dispersers for a total of 268 Mio. steps.
+We randomly placed source points within protected areas larger (\(>\) 700
+km\textsuperscript{2}), which conforms to the average home range requirement of
+resident wild dogs \citep{Pomilia.2015}. By distributing source points randomly,
+the number of source points per km\textsuperscript{2} within protected areas was
+approximately equal across our study area. In total, we distributed 50'000
+source points across all protected areas, each representing the starting point
+of a dispersal trajectory. To render potential immigrants into the study system,
+we randomly placed 10'000 additional source points inside a xx km buffer around
+the study area. Hence, a total of 60'000 source points were generated, thereby
+resulting in 60'000 simulated dispersers. It is worth pointing out that our
+choice of  placing source points completely random is arbitrary, as one as well
+adjust the sampling frequency based on the density of dispersers in the
+respective source area.
+
+% Simulations from step selection functions are known to be sensitive to the
+% location of source points, which is why we followed a twofold approach to sample
+% release points across the KAZA-TFCA. In a first approach, we used the same 68
+% source points between which we already computed least-cost paths and least-cost
+% corridors (see xx). These source points were regularly spaced 100 km apart,
+% located within protected areas (\(>\) 700 km\textsuperscript{2}). Because of
+% the regulary spacing, not all protected areas received source points. Hence, we
+% placed additional source points at the center of such protected areas. At each
+% of the generated source point we released 1000 dispersers, implying a total of
+% 68'000 simulated dispersers. In a second approach, we used the same 68 source
+% points to define catchment areas inside protected areas. That is, for each of
+% the 68 source points we defined its voronoi polygon within the point's protected
+% area. Within these catchment areas we then randomly sampled 1000 new source
+% points from which we released another 68'000 dispersers. Overall, we simulated
+% 134'000 dispersers for a total of 268 Mio. steps.
 
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_SourcePoints}
+    \includegraphics[width = \textwidth]{99_SourcePoints.png}
     \caption{Illustration of source points from which dispersal was simulated.
@@
 
+\subsection{Dispersal Simulation}
+We used the most parsimonious movement model to simulate dispersing wild dogs
+departing from the above described source points. The simulation resembled an
+inverted iSSF and was set up as follows. First, we determined a source point at
+which a disperser was initiated. To allow calculation of turning angles, we
+assumed a random orientation of the animal. Second, we proposed a set of 25
+random steps leaving from the source point. We generated random steps by
+sampling turning angles from a uniform distribution (\(-\pi, +\pi\)) and step
+lengths from our fitted gamma distribution. To prevent unrealistically large
+steps, we capped sampled step lengths at 35km, which is the farthest distance
+travelled by our dispersers within 4 hours. Third, along each random step we
+extracted habitat covariates and calculated movement covariates. Fourth, we
+applied the parametrized movement model to predict the selection scores \(w(x)\)
+for all steps. Fifth, we translated selection scores into probabilitities by
+applying equation xy. Sixth, we sampled one of the proposed based on assigned
+probabilities and calculated the animal's new position. We then repeated steps
+two to six until a total of 2000 steps were realized.
+
+To mitigate the influence of edge effects, we added a buffer zone around each
+covariate layer. The values within the buffer zone where resampled from the true
+data. As such, simulated dispersers were allowed to leave and reenter the study
+area through this buffer zone. In case one of the proposed random steps left the
+outer boundary of the buffer, we replaced the step with an alternative step that
+lied fully within the buffer zone. Thus, we force the disperser to bounce of the
+borders of the study area.
+
 \subsection{Heatmaps}
-Using the simulated trajectories we created heatmaps to examine through which
-areas most simulated individuals dispersed. We rasterized all trajectories and
-calculated how often each raster-cell was traversed by simulated dispersers. If
-the same trajectory crossed a pixel twice, it was only counted once. We achieved
-high performance rasterization of spatial lines using the recently developed
-R-package \textit{terra} \citep{Hijmans.2020}. To examine if and how ``heat''
-changes in response to changes in the location of source points and the number
-of simulated steps, we followed a 2 x 6 design and created heatmaps for both
-point sampling regines, as well as for 68, 125, 250, 500, 1000, and 2000
-dispersal steps. We quantified the similarity of the resulting 12 heatmaps to
-the permeability and least-cost corridor maps presented in (Hofmann ...) we used
-Bhattacharyya's affinity. Bhattacharyya's affinity ranges from zero (complete
-separation) to one (perfect match) and has earlier been proposed to compare the
-overlap of utilisation distributions (Fieberg).
+Using the simulated trajectories, we created heatmaps that indicate the
+frequency at which each raster-cell in the study area was visited. If the same
+trajectory crossed a pixel twice, it was only counted once, i.e. we did not
+consider revisits. We achieved high performance rasterization of all simulated
+trajectories using the recently developed R-package \textit{terra}
+\citep{Hijmans.2020}.
+
+To examine if and how ``heat'' changes in response to changes in the location of
+source points and the number of simulated steps, we followed a 2 x 6 design and
+created heatmaps for both point sampling regines, as well as for 68, 125, 250,
+500, 1000, and 2000 dispersal steps. We quantified the similarity of the
+resulting 12 heatmaps to the permeability and least-cost corridor maps presented
+in (Hofmann ...) we used Bhattacharyya's affinity. Bhattacharyya's affinity
+ranges from zero (complete separation) to one (perfect match) and has earlier
+been proposed to compare the overlap of utilisation distributions (Fieberg).
+
+\subsection{Betweenness}
 
@@
 raster layers at multiple resolutions and identified each trajectory's
-transition matrix on these rasters see figure xx \citep{Bastille.2018}. We then
-merged the transition matrices of all trajectories and calculated cumulative
-transitions between all raster-cells. This resulted in an edge-list, containing
-all observed from-to connections as well as their frequency. Using this
-edge-list, we generated a weighted graph using the r-package \textit{igraph}.
-Based on this graph we calculated betweenness scores as well as the degree of
-each raster cell. Betweenness indicates how often a specific raster-cell lies on
-a shortest path between two other raster-cells and is a useful metric to detect
-movement corridors. Degree, on the other hand, indicates how many connections a
-raster cell has and therefore serves to illustrate the xxx of specific nodes.
-When calculating betweenness, we used the transition frequency as weighting
-factor. That is, a higher transition frequency contributed to a higher
-betweenness score.
+transition matrix on these rasters see figure xx \citep{BastilleRousseau.2018}.
+We then merged the transition matrices of all trajectories and calculated
+cumulative transitions between all raster-cells. This resulted in an edge-list,
+containing all observed from-to connections as well as their frequency. Using
+this edge-list, we generated a weighted graph using the r-package
+\textit{igraph}. Based on this graph we calculated betweenness scores as well as
+the degree of each raster cell. Betweenness indicates how often a specific
+raster-cell lies on a shortest path between two other raster-cells and is a
+useful metric to detect movement corridors. Degree, on the other hand, indicates
+how many connections a raster cell has and therefore serves to illustrate the
+xxx of specific nodes. When calculating betweenness, we used the transition
+frequency as weighting factor. That is, a higher transition frequency
+contributed to a higher betweenness score.
 
@@
   \begin{center}
-    \begin{subfigure}[c]{0.8\textwidth}
     \includegraphics[width=\textwidth]{99_MovementModel}
-      % \subcaption{Subfigure Bild Nr. 1}
-    \end{subfigure}
-    % \begin{subfigure}[c]{\textwidth}
-    %   \includegraphics[width=\textwidth]{99_MovementModel(Interactions)}
-    %   % \subcaption{Subfigure Bild Nr. 2}
-    % \end{subfigure}
-    \caption{Most parsimonious movement model. Whiskers delineate the 95\%
-    Confidence-Intervals. Significance codes: * \(p < 0.10\), ** \(p < 0.05\),
-    *** \(p < 0.01\).}
+    \caption{(a) Most parsimonious movement model for dispersing wild dogs. The
+    model includes a habitat kernel, a movement kernel, as well as their
+    interactions. The orange and blue line segments delineate the 90\%, 95\%,
+    and 99\% Confidence-Intervals around the respective \(\beta\) coefficients.
+    Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
+    Results from the k-fold cross validation procedure. The upper plot shows
+    rank frequencies of predicted realized scroes according to model predictions
+    with known preferences, whereas the lower plot shows rank frequencies when
+    assuming random preferences. The blue ribbon shows the prediction interval
+    around a loess smoothing regression that we fitted to ease the
+    interpretation of the plots.}
     \label{MovementModel}
@@
 
+In this regard, our approach is rather similar to dispersal kernels, yet it
+enables to render directional biases, which are currently difficult to implement
+using such methods.
+
+We completely randomized the location of source points within protected areas.
+However, in some cases prior knowledge about the density of potential dispersers
+is available and can be used to adjust the number of simulated individuals
+accordingly. Alternatively, instead of tweaking the number of simulated
+individuals, one could assign a weight to each trajectory that depends on the
+density of potential dispersers in the source areas. As such, trajectories from
+areas with high density would enter the heatmap with above average weigth.
+
+The parametrized movement model could also be manipulated to investigate how
+different habitat preferences influence landscape connectivity and to test the
+sensitivity of results with respect to the exact preferences of individuals.
+
+One of the major benefits of individual-based simulations is the ability to make
+the temporal dimension of movement explicit. This allows to investigate how
+connectivity depends on the dispersal duration, something that is not possible
+with traditional least-cost or circuit theory methods.
+
+Optimally, one should simulate additional dispersers until the amount of newly
+gained information lies beyond a certain threshold, i.e. until some sort of
+convergence is achieved. However, due to the myriad of outcomes in movement
+behavior, this would often imply countless repetitions for each source points at
+each location where dispersal is possibly initiated. This is computationally
+infeasible.
+
 Our work suggests that the selection of source points significantly impacts
@@
 connectivity.
+
+Comparable simulations could also serve to predict the likely whereabouts of
+dispersers as they move. This information can then be utilized as an  ``early
+warning'' system, such that potential encounters can be gauged ad hoc.
 

### 2021-03-22 Chapter_1/02_R-Scripts/02_Analysis/08_RasterizeSimulations.R

@@ -22,7 +22,6 @@ library(viridis)          # For nicer colors
 library(tictoc)           # To keep track of processing time
 library(pbmcapply)        # To show progress bar in mclapply calls
 library(tmap)             # For nice spatial plots
-library(Cairo)            # To store plots
 library(davidoff)         # Custom functions
 
 ################################################################################
@@ -33,29 +32,25 @@ library(davidoff)         # Custom functions
 rasterizeSims <- function(
       simulations = NULL      # Simulated trajectories
     , steps       = 400       # How many steps should be considered
-    , sampling    = "Static"  # For which sampling method we want to rasterize
   ){
 
   # Create tracks that fulfill the above requirements
+  cat("Creating spatial lines...\n")
   sub_traj <- sims2tracks(
       simulations = simulations
     , steps       = steps
-    , sampling    = sampling
+    , id          = "TrackID"
+    , keep.data   = F
   )
-
-  # Remove data by creating spatial lines
   sub_traj <- as(sub_traj, "SpatialLines")
-
-  # Assign correct CRS
   crs(sub_traj) <- CRS("+init=epsg:4326")
 
-  # Coerce to "vect" for faster rasterization
+  # Coerce lines to "vect" and crop raster
   sub_traj <- vect(sub_traj)
-
-  # Speed up rasterization by cropping the raster
   r_crop <- crop(r, ext(sub_traj))
 
   # Rasterize lines onto the cropped raster
+  cat("Rasterizing spatial lines...\n")
   heatmap <- rasterizeTerra(sub_traj, r_crop)
 
   # Return the resulting heatmap
@@ -72,19 +67,18 @@ r <- rast("03_Data/02_CleanData/00_General_Raster.tif")
 r <- aggregate(r, fact = 10)
 
 # Load the simulated dispersal trajectories
-sims <- read_rds("03_Data/03_Results/99_DispersalSimulationSub.rds")
+sims <- read_rds("03_Data/03_Results/99_DispersalSimulation.rds")
 
 # Check out the number of rows
 nrow(sims) / 1e6
 
-water <- raster("03_Data/02_CleanData/01_LandCover_WaterCoverAveraged_MERGED.tif")
-sub <- subset(sims, SourceArea %in% c(34, 35))
-sub <- subset(sims, TrackID %in% sample(unique(sub$TrackID), 1000))
-sub$ID <- sub$TrackID
-sub$PointSampling <- "Random"
[…]

### 2021-03-25 Chapter_1/04_Manuscript/Manuscript.tex
Improved rasterization speed drastically

@@
 \begin{abstract}
+For many species, a sufficient degree of landscape connectivity is a crucial
+prerequisite for long-term species persistence. This is especially true for
+animals that exhibit long-distance dispersal, as dispersers typically cross a
+variety of habitats and potentially move into unfamiliar territory. Estimating
+habitat permeability and quantifying landscape connectivity have therefore
+become two major tasks in conservation ecology. For this, least-cost analysis
+and circuit theory have been the two workhorses during the past two decades.
+This is largely owed to their ease of use and intuitive nature which has
+facilitated the application of the methods across a broad range of the animal
+kingdom. However, both methods make several restricting assumptions that limit
+their usefulness in reality.
 
-During the past two decades, least-cost analysis and circuit theory have been
-the two workhorses in connectivity analyses. This is largely thanks to their
-ease of use and intuitive nature which has facilitated the application of the
-methods across a broad range of animals. However, recent innovations in movement
-ecology have brought forward new methods that are similarly intuitive but
-provide several advantages over traditional connectivity modelling techniques.
+Recent innovations in movement ecology, particularly on step selection analysis,
+have brought forward novel ways to look at connectivity. For instance,
+integrated step selection functions provide a means to parametrize a fully
+mechanistic movement model based on which virtual dispersers could be simulated.
+While this approach has been used to infer habitat utilization, it may also
+serve to highlight landscape connectivity and pinpoint dispersal corridors.
 
 Here, we propose the simulation of dispersal trajectories as a much more generic
-way to identify dispersal barriers and potential movement corridors. Moreover,
-we show how such simulations can be used to identify zones of high
-human-wildlife conflicht, as well as to pinpoint potential "dispersal traps"
-where dispersing individuals get trapped due to unavailability of suitable
-alternatives. Finally, we discuss the benefits and pitfalls of dispersal
-simualtions.
+way to identify dispersal barriers and potential movement corridors. To achieve
+this, we applied integrated step selection analysis and uses data from 16
+dispersing wild dogs to parametrize a fully mechanistic movement model for
+dispersing African wild dogs. The model rendered dispersers' habitat and
+movement preferences, as well as potential interactions between them. Based on
+the model, we simulated myriads of dispersal trajectories and inferred dispersal
+corridors across the earth's largest transboundary conservation area, the
+Kavango-Zambezi Transfrontier Conservation Area. We also exemplify how such
+simulations could be analysed using network theory. Finally, we discuss the
+benefits and pitfalls of dispersal simualtions and highlight potential
+improvements to be made in the future.
 \end{abstract}
@@
 \section{Introduction}
-% Dispersal & Connectivity
+% Importance of Dispersal & Connectivity
 Dispersal of individuals is an important process governing the dynamics wild
@@
 defined as the movement of individuals from their natal location to the site of
-first reproduction \cite{Howard.1960} and allows to avoid inbreeding
+first reproduction \cite{Howard.1960} and allows species to avoid inbreeding
 \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012}, to rescue small and
 unviable populations \citep{Brown.1977}, and to promote the colonization of
-unoccupied habitats \citep{Hanski.1998, MacArthur.2001}. Information on movement
-behavior during dispersal and knowledge about the factors that limit dispersal
-is therefore critical for a comprehensive understanding of landscape
-connectivity and population viability \citep{Baguette.2013, Vasudev.2015}.
+unoccupied habitats \citep{Hanski.1998, MacArthur.2001}. Successfull dispersal
+requires a sufficient degree of landscape connectivity, which is why the
+protection of dispersal corridors has become a major goal in conservation
+science. Information on movement behavior during dispersal and knowledge about
+the factors that limit dispersal is therefore critical for a comprehensive
+understanding of landscape connectivity and population viability
+\citep{Baguette.2013, Vasudev.2015}. In addition, reliable modelling techniques
+to identify dispersal corridors based on empirical data are necessary
+\citep{Diniz.2019}.
 
-% Technological Advancements
+% Advancements in GPS Technology and Movement Ecology
 Thanks to novel technologies developed over the past decades, particularly of
-GPS/Satellite radio-collars, the study of dispersal has accelerated
-\citep{Jonsson.2016, Williams.2019}. Additionally, the advent of publicly
-accessible satellite imagery and sophisticated remote sensing techniques to
-represent the physical landscape through which individuals disperse has pushed
-the field towards a  ``golden age of animal tracking'' \citep{Kays.2015}.
-Concurrently, the increased availability of large amounts of empirical data and
-an increased computational power have led to the development of numerous
-analytical methods to study movement behavior during dispersal
-\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012, Elliot.2014,
-Diniz.2019}.
+GPS/Satellite radio-collars, the study of dispersal and connectivity using
+telemetry data has accelerated \citep{Jonsson.2016, Williams.2019}.
+Additionally, the advent of publicly accessible satellite imagery and
+sophisticated remote sensing techniques to represent the physical landscape
+through which individuals disperse have heralded the ``golden age of animal
+tracking'' \citep{Kays.2015}. Concurrently, the increased availability of large
+amounts of empirical data and an increased computational power have led to the
+development of several modelling techniques that allow studying movement
+behavior during dispersal and highlighting major movement corridors
+\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012, Diniz.2019}.
 
-% Measuring Connectivity
-Resource selection functions \citep{Boyce.2002} and derived methods such as step
-selection functions \citep{Fortin.2005} and path selection functions
-\citep{Cushman.2010} have proven particularly useful for studying animal
-movement and landscape connectivity. These methods allow estimating habitat
-preferences of the focal species by comparing covariates at locations visited by
-the animal to the scame covariates at locations available to, but not visited by
-the animal. The so estimated preferences can then be used to predict a
-permeability surface, indicating the expected ease at which an animal can
+% Resource Selection & Connectivity
+\textit{Resource selection functions} \citep{Boyce.2002} and derived methods
+such as \textit{step selection functions} \citep{Fortin.2005} and \textit{path
+selection functions} \citep{Cushman.2010} have proven particularly useful for
+studying animal movement \citep{Fieberg.2020}. These methods allow estimating
+habitat preferences of the focal species by comparing covariates at locations
+visited by the animal to the scame covariates at locations available to, but not
+visited by the animal. The so estimated preferences can then be used to predict
+a permeability surface, indicating the expected ease at which an animal can
 traverse a given area \citep{Zeller.2012}. Utimately, the permeability surface
-serves as foundation to a connectivity model and to reveal movement corridors.
-In this regard, the literature has long relied on least-cost methods
-\citep{Adriaensen.2003} and circuit theory \citep{McRae.2006, McRae.2008}.
-Unfortunately, both of these approaches make several limiting assumptions that
-are hardly ever met in reality, particularly by dispersing animals. For
-instance, to calculate a least-cost path one has to pre-define start- and
-end-points. This implicitly assumes that animals have a preconceived end-point
-in mind and choose a cost-minimizing route accordingly. To calculate a
-least-costly route, one also needs to assume that animals have an infinite
-perceptual range and that a route can always be realized, regardless of the
-imposed costs. In contrast to least-cost methods, circuit theory assumes a
-perceptual range of a single pixel, which is equally unrealistic and rarely
-captures the scale of selection (cite someone). Moreover, directional biases
-cannot be captured by circuit theory as a random walk is implicitly assumed. In
-reality, however, dispersers often depict directional persistence and move in an
-almost straight line to cover much ground very quickly. A mix of the two methods
-can be seein in least-cost corridor approaches or randomized shortest paths have
-been proposed to mitigate these issues by allowing animals to deviate from the
-``optimal'' least-costly route \citep{Pinto.2009, Panzacchi.2016,
-VanMoorter.2021}, they still fail to render a more realistic perceptual range of
-the animal. In addition, neither of the two methods render the temporal
-dimension of movement. Thusfar, only few have explored the usefulness of
-individual based movement models when estimating landscape connectivity
-\citep{Kanangaraj.2013, Hauenstein.2019, Zeller.2020}.
+serves as input to a connectivity model that is used to reveal movement
+corridors. In this regard, two of the most prominent connectivity models in the
+literature are least-cost path analysis \citep{Adriaensen.2003} and circuit
+theory \citep{McRae.2006, McRae.2008}.
 
-Nevertheless, collecting data on dispersers remains logistically challenging due
-to dispersers' inherent mobility and due to the fact that dispersers only make
-up a small fraction of the entire population, thereby making it difficult for
-scientists to target apporpriate individuals for their study.
+% Least-cost path
+In least-cost path analysis, connectivity is estimated by identifying
+least-costly routes between pre-defined start- and endpoints
+\citep{Adriaensen.2003}. For this, the permeability surface is converted into a
+network graph and permeability scores are translated into probabilities for
+moving from one cell to another. Least-costly routes between start- and
+endpoints can then be computed using Dijkstra’s algorithm \citep{Dijkstra.1959}.
+Although intuitively appealing, the method suffers from several weaknesses.
+First of all, a least-cost path always exist, even if the associated costs are
+unreasonably high. Second, it assumes that animals have a preconceived end-point
+in mind and choose a cost-minimizing route accordingly. While this assumption
+may be reasonable for migrating animals that move between a limited number of
+habitats, it is unlikely to hold for dispersers that move over long distances
+into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Finally, the
+method also requires that animals have an infinite perceptual range, otherwise
+they could not compute an optimal path. Some of these issues have been addressed
+using alternative least-cost approaches. For instance, instead of computing a
+least-cost path, a least cost-corridor that also considers slightly suboptimal
+routes \citep{Pinto.2009} can be calculated. The corridor relaxes the assumption
+of perfect knowledge and overcomes the single pixel width issue. Alternatively,
+the randomized least-cost path has been proposed, which allows animals to
+deviate from the least-cost route by a pre-defined factor \(\theta\)
+\citep{Panzacchi.2016, VanMoorter.2021}.
 
+% Circuit theory
+In contrast to least-cost approaches, circuit theory makes use of electrical
+circuit theory to quantify landscape connectivity \citep{McRae.2006,
+McRae.2008}. Here, the permeability surface is converted into an electrical
+circuit containing nodes and resistor, where resistance values depend on habitat
+permeability. A source point is connected to a current source, whereas an
+endpoint is grounded. Hence, the current flowing through each node in the
+circuit can be computed and is supposed to resemble the relative frequency at
+which each location in the study area is being visited. By revealing not only a
+single path or corridor, circuit theory overcomes the single-pixel width issue
+inherent to least-cost approaches. Intrinsically, circuit theory is based on the
+assumption that animals follow a random walk and can therefore not render
+directional biases. In reality, however, dispersers often move in a very
+directional fashion as they want to cover as much ground in as little time as
+possible. Furthermore, it is assumed that individuals have a perceptual range of
+a single pixel, which rarely captures the true spatial scale of selection (cite
+someone).
+
+% Temporal scale
+Neither least-cost analyes or circuit theory are capable of rendering the
+temporal dimension of movement. Such influences could, however, be studied using
+individual based movement models. Still, only few have explored these
+possibilities \citep{Kanangaraj.2013, Hauenstein.2019, Zeller.2020}.
+
+% Step selection analysis
+While step selection functions were initially intended to learn about habitat
+preferences of the focal species \citep{Fortin.2005}, recent improvements may
+elevate its applicability into other areas. Originally, \cite{Fortin.2005}
+proposed step selection functions to account for serial autocorrelation inherent
+to GPS relocation data. In this method, covariates along realized steps (the
+connecting lines between two consecutive relocations; ; \citealp{Turchin.1998})
+are contrasted to covariates along  ``alternative''  or  ``random steps'' that
+are generated by randomly sampling turning angles and step lengths based on
+observed movement characteristics. It is then assumed that animals assign a
+selection score \(w(x)\) of the following form to each step:
+
+\begin{equation}
+\label{EQ1}
+  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
+\end{equation}
+
+The selection score \(w(x)\) of each step depends on its associated covariates
+(\(x_1, x_2, ..., x_n\)) and on the animal's relative selection strengths (i.e.
+preferences) towards these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
+probability that a step \(i\) is realized \(P(Y_{i} = 1\)) then depends on the
+step's selection score, as well as on the selection scores of all alternative
+steps:
+
+\begin{equation}
+\label{EQ2}
+  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
+  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
+\end{equation}
+
+\noindent Preferences of interest, i.e. the \(\beta\)'s, can be estimated by
+comparing realized (scored 1) and random (scored 0) steps in a conditional
+logistic regression model \citep{Fortin.2005}. In this model, positive
+\(\beta\)-coefficients indicate selection of a covariate, whereas negative
+\(\beta\)-coefficients indicate avoidance of a covariate. To deal with multiple
+individuals, one can either average estimates of individual models
+\citep{Murtaugh.2007, Fieberg.2010} or apply mixed effects conditional logistic
+regression analysis as recently proposed by \cite{Muff.2020}.
+
+% Integrated step selection analysis
 An animals movement trajectory can be seen as the result of an interplay between
-habitat and movement preferences.
+habitat and movement preferences. Traditional step selection functions cannot
+account for a correlation between turning angles and step lengths, unless the
+two are sampled jointly from a copula distribution \citep{Hodel.2021}. Recently,
+however, SSFs have been generalized to \textit{integrated} SSFs (iSSFs), which
+tear apart movement and habitat preferences of the studied animals
+\citep{Avgar.2016}. The method thus allows to render an animals preferences with
+respect to prevailing habitat conditions, as well as with respect to movement
+characteristics. This reduces potential biases in estimated preferences and
+enables parametrization of a fully mechanistic movement model from which
+movement and space use can be simulated \citep{Avgar.2016, Signer.2017}.
+\cite{Signer.2017} used integrated step selection analysis to simulate steady
+state utilization distributions of resident animals. However, the degree to
+which such simulations could be used to simulate dispersers and infer movement
+corridors is unknown.
 
-In recent decades, least-cost analysis has become the workhorse to study
-landscape connectivty at large scales. Originally introduced by XX, the concept
-and relating methods have been refined to more realistically render animal's
-movement capabilities and to determine valuable movement corridors that require
-protection. (talk a bit about different methods of LCPs, see methods in
-gdistance package). More recently, more attention has been assigned to the
-collection of movement data is better tailored towards assessing landscape
-connectivity. In particular, it has been proposed and verified that relocation
-data collected dispersing individuals leads to more accurate and reliable
-estimates of landscape connectivity. This superiority of dispersal data in
-comparison to data that stem from residents is mainly owed to the fact that
-animals behave vastly different during residency. Despite substantial advances
-in ecologists ability to track animals in space and time, dispersal remains one
-of the most difficult behavioral modes to observe in wild animals. Especially
-for wide ranging and long-lived species, dispersal is difficult to predict and
-observe, such that data remains scarce.
+% Study Species & Study Area
+One of the species for which long-term species persistence relies on sufficient
+landscape connectivity is the endangered African wild dog \textit{Lycon pictus}.
+This species has once been wide-spread across sub-Saharan Africa but disappeared
+from a vast majority of its historic range due to ongoing persecution by humans,
+habitat destruction, and deadly diseases. As of today, only 6'000 free-ranging
+individuals remain in small and spatially scattered subpopulations. Within those
+subpopulations, wild dogs form cohesive packs comprising 8 to 12 adults and
+their offspring \cite{McNutt.1995}. After reaching sexual maturity, male and
+female offspring form same-sex coalitions and disperse from their natal pack
+\citep{McNutt.1996, Behr.2020}. New packs are formed when dispersing coalitions
+join unrelated opposite-sex dispersing coalitions \citep{McNutt.1996}.
+Dispersing wild dogs can cover several hundred kilometers across a variety of
+landscapes \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
+One of the few strongholds for this species lies near the Moremi Game Reserve in
+northern Botswana, which is part of the world's largest transboundary protected
+area, namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
+This area has originally been intended to facilitate migration of elephants, but
+is expected to benefit a multitude of other species \citep{Elliot.2014,
+Brennan.2020, Hofmann.2021}.
 
-Many connectivity modelling techniques, especially least-cost analysis,
-implicitly assume that the studied animal has partial or even complete knowledge
-of the landscape and associated movement costs. While this assumption may be
-reasonable for migrating animals that move between a limited number of habitats,
-yet it is unlikely to hold for dispersers. In contrast to migrants, dispersers
-move into unknown territory and are therefore confronted with novel landscapes.
-Consequently, dispersers are more likely to adjust their movement behavior
-\textit{ad hoc} instead of preplanning an entire trajectory. As a result,
-methods that assume complete knowledge of the landscape and subsequent optimal
-movement behavior likely misrepresent true dispersal behavior. As such, methods
-that allow a more random approach to a dispersers movement behavior may more
-realistically render the movement corridors of dispersing individuals.
-
-However, these models can not account for a correlation between turning angles
-and step lengths, unless the two are sampled jointly from a copula distribution
-(Hodel et al.).
-
-A relatively recent addition to the step selection family has been integrated
-step selection analysis.... Thanks to recently published papers and packages,
-the suite of step selection functions has become accessible to most researchers
-working with relocation data. Despite this, the use of such methods rarely
-extends beyond assessing relative habitat preferences. However, once
-parametrized, integrated step selection models could potentially serve as
-powerful framework to simulate animal movement and assess landscape
-connectivity. Here, we exemplify such analyses and show how ... coud be applied
-to investigate ...
-
-In contrast to Signer et et al, our focus lied on the transient utilisation of
-the landscape.
-
-In addition, individual based simulations will allow to explicitly model
-dispersal between subpopulations in models of population dynamics.
+In a previous paper, we assessed landscape connectivity within the KAZA-TFCA for
+dispersing African wild dogs using a least-cost corridor approach. For this, we
+fitted a basic habitat selection model based on which we predicted landscape
+permeability. We now expand on this knowledge and develop a more detailed
+movement model of dispersing wild dogs. We then use this model to simulate
+thousands of dispersers moving throughout the KAZA. Based on said simulations,
+we compute heatmaps and identify potential dispersal hotspots and compare them
+to the dispersal routes identified in \citep{Hofmann.2021}. We also showcase how
+simulated dispersal data can be analysed using network-analysis and how network
+metrics relevant to landscape connectivity can be computed. Our results suggest
+that a simulation based approach yields several major benefits over traditional
+connectivity modelling techniques. Most importantly, simulations provide a more
+generic view on how connectivity emerges and to which degree connectivity
+depends on the dispersal duration. In addtion, by generating proper dispersal
+trajectories, network theory can be applied to calculate network metrics that
+are pertinent to connectivity analysis. Finally, we put forward additional
+opportunities using simulations that go beyond the scope of this paper.
 
@@
 connectivity literature, mainly due to the difficulty in observing wild dog
-dispersal. In a previous paper, we addressed this issue and developed a habitat
-selection model based on which we predicted landscape connectivity using
-least-cost corridor analysis. We now expand on this knowledge and develop a more
-detailed movement model of dispersing wild dogs. We then use this model to
-simulate thousands of dispersers moving throughout the KAZA. Based on said
-simulations, we compute heatmaps and identify potential dispersal hotspots and
-compare them to the dispersal routes identified in (Hofmann ....). We also apply
-the data to feed a network-analysis, based on which we determine network-metrics
-pertinent to landscape connectivity.
+dispersal.
 
@@
 northern Botswana. We identified potential dispersers based on age, number of
-same‐sex siblings, pack size, and presence of unrelated individuals of the
-opposite sex in their pack \citep{McNutt.1996, Behr.2020}. We immobilized
-selected individuals using a cocktail of Ketamine/Xylazine/Atropine
-\citep{Osofsky.1996, Cozzi.2020} which we injected with a dart, fired from a
-CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). The
-corresponding protocols have been approved by the Ministry of Environment,
-Natural Resources Conservation and Tourism of Botswana (permit EWT 8/36/4
-XXXVI). After immobilization, individuals were fitted with GPS/Satellite radio
-collars (\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) that included
-an automated drop-off mechanism. Handling and collaring of all individuals was
-carried out and supervised by a Botswana-registered wildlife veterinarian. All
-of the immobilized individuals quickly rejoined their pack after the procedure.
-From all collared individuals, 16 individuals eventually dispersed in separate
-same-sex coalitions and their trajectories were successfully recorded (7 female
-and 9 male coalitions).
+same‐sex siblings, pack size, and presence of unrelated opposite-sex individuals
+in the pack \citep{McNutt.1996, Behr.2020}. We immobilized individuals using a
+cocktail of Ketamine/Xylazine/Atropine \citep{Osofsky.1996, Cozzi.2020},
+injected with a dart that was fired from a CO\textsubscript{2}-pressurized gun
+(\textit{DAN-Inject, Denmark}). Protocols for the anesthization protocols were
+approved by the Ministry of Environment, Natural Resources Conservation and
+Tourism of Botswana (permit EWT 8/36/4 XXXVI). After immobilization, individuals
+were fitted with GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic
+Aerospace GmbH, Berlin}) that included an automated drop-off mechanism. Handling
+and collaring of all individuals was carried out and supervised by a
+Botswana-registered wildlife veterinarian. All of the immobilized individuals
+quickly rejoined their pack within one hour after the procedure. Out of all
+collared individuals, 16 eventually dispersed in separate same-sex coalitions
+and their trajectories were successfully recorded (7 female and 9 male
+coalitions).
 
@@
 sexes for the sake of our analyses. Ultimately, we converted the collected GPS
-relocations to steps, where a step represented the straight-line distance
+relocations to steps, where each step represented the straight-line distance
 travelled between two consequtive GPS relocations \citep{Turchin.1998}.
@@
 score \(w(x)\) to each step. The score depends on the covariates \(X\)
-experienced along a step, as well as the animals preferences towards these
-covariates \(\beta\). The probability of a step being realized is then
-contingent on the step's selection score, as well as the selection scores of all
-other step in the same stratum. In contrast to regular step selection analysis,
+experienced along a step, as well as the animals preferences \(\beta\) towards
+these covariates. The probability of a step being realized is then contingent on
+the step's selection score, as well as the selection scores of all other step in
+the same stratum. In contrast to regular step selection analysis,
 \textit{integrated} step selection analysis allows simulatneous inference on
@@
 \subsection{Dispersal Simulation}
-We used the most parsimonious movement model to simulate dispersing wild dogs
-departing from the above described source points. The simulation resembled an
-inverted iSSF and was set up as follows. First, we determined a source point at
-which a disperser was initiated. To allow calculation of turning angles, we
-assumed a random orientation of the animal. Second, we proposed a set of 25
-random steps leaving from the source point. We generated random steps by
-sampling turning angles from a uniform distribution (\(-\pi, +\pi\)) and step
-lengths from our fitted gamma distribution. To prevent unrealistically large
-steps, we capped sampled step lengths at 35km, which is the farthest distance
-travelled by our dispersers within 4 hours. Third, along each random step we
-extracted habitat covariates and calculated movement covariates. Fourth, we
+We used the most parsimonious movement model to simulate a total 80'000
+dispersing wild dogs across the KAZA-TFCA. The simulation basically resembled an
+inverted iSSF and was set up as follows. First, we sampled a source point at
+which a disperser was initiated and assumed a random initial orientation of the
+animal. The orientation was essential to permit calculation of relative turning
+angles. Second, we generated a set of 25 random steps by sampling turning angles
+from a uniform distribution (\(-\pi, +\pi\)) and step lengths from the fitted
+gamma distribution. Similarly to our input data, each step corresponded to a
+temporal difference of 4 hours. To prevent unrealistically large steps, we
+capped sampled step lengths at 35km, which corresponds to the farthest distance
+travelled by one of our dispersers within 4 hours. Third, along each random step
+we extracted habitat covariates and calculated movement covariates. Fourth, we
 applied the parametrized movement model to predict the selection scores \(w(x)\)
-for all steps. Fifth, we translated selection scores into probabilitities by
-applying equation xy. Sixth, we sampled one of the proposed based on assigned
-probabilities and calculated the animal's new position. We then repeated steps
-two to six until a total of 2000 steps were realized.
+for all steps. Fifth, we translated selection scores into probabilities by
+applying \Cref{EQ2}. Sixth, we sampled one of the random steps based on assigned
+probabilities and determined the animal's new position. We then repeated steps
+two to six until a total of 2'000 steps were realized.
 
-To mitigate the influence of edge effects, we added a buffer zone around each
-covariate layer. The values within the buffer zone where resampled from the true
-data. As such, simulated dispersers were allowed to leave and reenter the study
-area through this buffer zone. In case one of the proposed random steps left the
-outer boundary of the buffer, we replaced the step with an alternative step that
-lied fully within the buffer zone. Thus, we force the disperser to bounce of the
-borders of the study area.
+Inevitably, some simulated dispersers would approach a map boundary and one of
+the generated random steps would leave the study area. However, beyond the study
+area no covariates could be extracted and no selection scores can be calculated.
+To mitigate the influence of such edge effects, we buffered all covariate layers
+by xx percent and filled the buffer zone with covariate values resampled from
+the observed data. As such, simulated dispersers were allowed to leave and
+reenter the study area through this buffer zone. In case one of the proposed
+random steps left the buffer, we resampled the corresponding step until it lied
+fully within the buffer zone. This implies that we forced the disperser to
+bounce off the borders of the buffer zone.
+
+To meaningfully select source points, we randomly sampled coordinates within
+protected areas larger (\(>\) 700 km\textsuperscript{2}), which conforms to the
+average home range requirement of resident wild dogs \citep{Pomilia.2015}. By
+distributing source points randomly, the number of source points per
+km\textsuperscript{2} within protected areas was approximately equal. In total,
+we distributed 50'000 source points across protected areas. To render potential
+immigrants into the study area, we placed 10'000 additional random points inside
+a xx km buffer around the study area. In total, we generated 80'000 source
+points, thereby resulting in 80'000 simulated dispersers, each of which moved
+2'000 steps. Overall, this implied 160 Mio. simulated steps.
 
@@
 
-\subsection{Betweenness}
-
-\subsection{Network Analysis I: Areas Reached}
-Based on the simulated trajectories we identified to which other areas each
-source area is connected. For instance, if a trajectory originated at source
-point one and intersected with source areas two and three, we assumed that
-source areas two and three were within reach of source are one. To get a sense
-of the strength of connections, we also calculated how often each of these
-connections was realized. This procedure resulted in 6 x 2 edge lists which we
-further used to generate weighted networks and to calculate network metrics.
-
-\subsection{Network Analysis II: Betweenness}
+\subsection{Network Analysis}
 We coerced all simulated trajectories into a network consisting of vertices
@@
 
-\subsection{Network Analysis III: Transitions}
+% Based on the simulated trajectories we identified to which other areas each
+% source area is connected. For instance, if a trajectory originated at source
+% point one and intersected with source areas two and three, we assumed that
+% source areas two and three were within reach of source are one. To get a sense
+% of the strength of connections, we also calculated how often each of these
+% connections was realized. This procedure resulted in 6 x 2 edge lists which we
+% further used to generate weighted networks and to calculate network metrics.
 
 \section{Results}
+\subsection{Movement Model}
 Compared to the base model, the most parsimonious movement model included
@@
 
+\subsection{Dispersal Simulation}
+
 \subsection{Heatmaps}
@@
 
-\begin{table}[hbtp]
-  \caption{Summary statistics of all GPS relocations that have been recorded on
-  dispersing coalitions.}
-  \label{GPSData}
-  \begin{center}
-  \resizebox{\textwidth}{!}{
-    \begin{tabular}{lllllllll}
-    \toprule
-    sampling & Metric & Map & 68 & 125 & 250 & 500 & 1000 & 2000 \\
-    \midrule
-    Static & Affinity & Corr & 0.64 & 0.70 & 0.76 & 0.80 & 0.82 & 0.84 \\
-    Static & Affinity & Perm & 0.67 & 0.74 & 0.80 & 0.85 & 0.89 & 0.91 \\
-    Static & Correlation & Corr & 0.32 & 0.38 & 0.45 & 0.51 & 0.56 & 0.60 \\
-    Static & Correlation & Perm & 0.47 & 0.57 & 0.66 & 0.72 & 0.78 & 0.82 \\
-    \hdashline
-    Random & Affinity & Corr & 0.70 & 0.75 & 0.79 & 0.82 & 0.84 & 0.85 \\
-    Random & Affinity & Perm & 0.74 & 0.79 & 0.83 & 0.87 & 0.90 & 0.92 \\
-    Random & Correlation & Corr & 0.38 & 0.43 & 0.48 & 0.53 & 0.57 & 0.61 \\
-    Random & Correlation & Perm & 0.58 & 0.64 & 0.69 & 0.75 & 0.79 & 0.83 \\
-    \hdashline
-    n.a. & Affinity & Heat & 0.76 & 0.86 & 0.94 & 0.97 & 0.99 & 0.99 \\
-    n.a. & Correlation & Heat & 0.90 & 0.94 & 0.97 & 0.98 & 0.99 & 1.00 \\
-    \bottomrule
-    \end{tabular}
-  }
-  \end{center}
-\end{table}
+% \begin{table}[hbtp]
+%   \caption{Summary statistics of all GPS relocations that have been recorded on
+%   dispersing coalitions.}
+%   \label{GPSData}
+%   \begin{center}
+%   \resizebox{\textwidth}{!}{
+%     \begin{tabular}{lllllllll}
+%     \toprule
+%     sampling & Metric & Map & 68 & 125 & 250 & 500 & 1000 & 2000 \\
+%     \midrule
+%     Static & Affinity & Corr & 0.64 & 0.70 & 0.76 & 0.80 & 0.82 & 0.84 \\
+%     Static & Affinity & Perm & 0.67 & 0.74 & 0.80 & 0.85 & 0.89 & 0.91 \\
+%     Static & Correlation & Corr & 0.32 & 0.38 & 0.45 & 0.51 & 0.56 & 0.60 \\
+%     Static & Correlation & Perm & 0.47 & 0.57 & 0.66 & 0.72 & 0.78 & 0.82 \\
+%     \hdashline
+%     Random & Affinity & Corr & 0.70 & 0.75 & 0.79 & 0.82 & 0.84 & 0.85 \\
+%     Random & Affinity & Perm & 0.74 & 0.79 & 0.83 & 0.87 & 0.90 & 0.92 \\
+%     Random & Correlation & Corr & 0.38 & 0.43 & 0.48 & 0.53 & 0.57 & 0.61 \\
+%     Random & Correlation & Perm & 0.58 & 0.64 & 0.69 & 0.75 & 0.79 & 0.83 \\
+%     \hdashline
+%     n.a. & Affinity & Heat & 0.76 & 0.86 & 0.94 & 0.97 & 0.99 & 0.99 \\
+%     n.a. & Correlation & Heat & 0.90 & 0.94 & 0.97 & 0.98 & 0.99 & 1.00 \\
+%     \bottomrule
+%     \end{tabular}
+%   }
+%   \end{center}
+% \end{table}
+
+\subsection{Network Analysis}
 
@@
 part.
+
+Empircally informed dispersal models will also be invaluable in spatially
+explicit population models in which dispersal is actually rendered
+mechanistically.
+
+In addition, individual based simulations will allow to explicitly model
+dispersal between subpopulations in models of population dynamics.
 

### 2021-03-25 Chapter_1/02_R-Scripts/02_Analysis/07_CombineSimulations.R

@@ -82,25 +82,10 @@ format(object.size(sims), units = "Gb")
 write_rds(sims, "03_Data/03_Results/99_DispersalSimulation.rds")
 sims <- read_rds("03_Data/03_Results/99_DispersalSimulation.rds")
 
-# # Function to subset simulations
-# subsims <- function(simulations, area = NULL, nid = NULL){
-#
-#   # Sample ids
-#   ids <- unique(simulations$TrackID[simulations$Area == area])
-#   ids <- sample(ids, size = nid, replace = F)
-#
-#   # Subset to respective simulations
-#   sub <- simulations[simulations$TrackID %in% ids, ]
-#
-#   # Return the subset
-#   return(sub)
-# }
-#
-# # Subset simulations
-# sub <- subsims(sims, nid = 100, area = "Main")
-# sub <- sims2tracks(sub, id = "TrackID", steps = 500)
-# plot(sub)
-#
+# Remove raw data of simulations to free some space
+file.remove(files_main)
+file.remove(files_buffer)
+
 # ################################################################################
 # #### Verify Turning Angles and Step Lengths
 # ################################################################################

### 2021-03-25 Chapter_1/02_R-Scripts/02_Analysis/08_RasterizeSimulations.R

@@ -19,39 +19,111 @@ library(tidyverse)        # For data wrangling
 library(rgeos)            # For manipulating vector data
 library(lubridate)        # For working with dates
 library(viridis)          # For nicer colors
-library(tictoc)           # To keep track of processing time
 library(pbmcapply)        # To show progress bar in mclapply calls
 library(tmap)             # For nice spatial plots
 library(davidoff)         # Custom functions
+library(spatstat)         # For quick rasterization
+library(maptools)         # For quick rasterization
+
+# sim2tracks2 <- function(simulation = NULL){
+#
+#   # Create spatial line
+#   lines <- vect(
+#       x    = as.matrix(simulation[, c("x", "y")])
+#     , type = "lines"
+#     , crs  = CRS("+init=epsg:4326")
+#   )
+#
+#   # Return it
+#   return(lines)
+#
+# }
+#
+# sims2tracks2 <- function(
+#       simulations = NULL
+#     , id          = "TrackID"
+#     , messages    = T
+#   ){
+#
+#   # Nest by id
+#   nested <- nest(simulations, data = -all_of(id))
+#
+#   # Prepare progress bar
+#   if (messages) {
+#     pb <- txtProgressBar(
+#         min     = 0
+#       , max     = nrow(nested)
+#       , initial = 0
+#       , style   = 3
+#       , width   = 55
+#     )
+#   }
+#
+#   # Create lines
+#   lines <- list()
+#   for (i in 1:nrow(nested)){
+#     lines[[i]] <- sim2tracks2(simulation = nested$data[[i]])
+#     if (messages) {
+#       setTxtProgressBar(pb, i)
+#     }
+#   }
+#
+#   # Put them together
+#   lines <- do.call(c, lines)
+#
+#   # Return lines
+#   return(lines)
+# }
 
 ################################################################################
-#### Function to Rasterize Simulated Trajectory
+#### Function to Rasterize Tracks
 ################################################################################
-# Function to rasterize trajectories of a desired source point after a desired
 # amount of time (i.e. after a desired number of steps)
 rasterizeSims <- function(
       simulations = NULL      # Simulated trajectories
+    , raster      = NULL      # Raster onto which we rasterize
[…]

### 2021-03-25 Chapter_1/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/MovementModel_conflict-20210222-111048.R

@@ -1,278 +1,0 @@
-################################################################################
-#### Plot Movement Model Results
-################################################################################
-# Description: A simple plot of the Dispersal Durations
-
-# Clear R's brain
-rm(list = ls())
-
-# Set the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)  # For data wrangling
-library(lubridate)  # To handle dates nicely
-library(ggpubr)     # For nice plots
-library(davidoff)   # Custom functions
-library(glmmTMB)    # To handle model results
-library(cowplot)    # To add image to ggplot
-library(lemon)      # For nice axis labels
-library(imager)     # To import images
-library(viridis)    # For nice colors
-library(raster)     # To create rasters (visInt)
-library(ggdark)     # Dark ggplot theme
-
-################################################################################
-#### Prepare Data
-################################################################################
-# Old directory
-# wd <- "/media/david/My Passport/Backups/WildDogs/15. PhD/00_WildDogs"
-# setwd(wd)
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load the movement model
-move.mod <- readRDS("03_Data/03_Results/99_MovementModel.rds")
-
-# Select best model
-best <- move.mod$Model[[1]]
-
-# Summary of the best model
-summary(best)
-
-# Calculate Confidence intervals
-coeffs <- getCoeffs(best, pvalue = TRUE)[-1, ] %>%
-  mutate(
-      LCI_90 = Coefficient - 1.645 * SE
-    , UCI_90 = Coefficient + 1.645 * SE
-    , LCI_95 = Coefficient - 1.96 * SE
-    , UCI_95 = Coefficient + 1.96 * SE
-    , LCI_99 = Coefficient - 2.575 * SE
-    , UCI_99 = Coefficient + 2.575 * SE
-  )
-
-# Add stars indicating the significance
-coeffs$Significance <- sapply(1:nrow(coeffs), function(x){
-  if (coeffs$pvalue[x] <= 0.01){
-    return("***")
[…]

### 2021-04-22 Chapter_1/04_Manuscript/99_MovementModelAICs.tex
Worked on manuscript for 1 chap

@@
 % latex table generated in R 3.6.3 by xtable 1.8-4 package
-% Thu Jul  9 09:17:26 2020
-\begin{tabular}{lrrrr}
+% Thu Apr 22 09:09:15 2021
+\begin{tabular}{lrrrrr}
   \toprule
-Covariates & AIC & DeltaAIC & Weight & LogLik \\ 
+Covariates & AIC & DeltaAIC & Weight & LogLik & ID \\ 
   \midrule
-cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
+cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) & 89392.88 & 0.00 & 0.15 & -44670.44 &   1 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:log\_sl & 89393.92 & 1.04 & 0.09 & -44669.96 &   2 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + DTW:log\_sl & 89394.13 & 1.25 & 0.08 & -44670.06 &   3 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:log\_sl & 89394.25 & 1.37 & 0.08 & -44670.13 &   4 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl & 89394.36 & 1.48 & 0.07 & -44672.18 &   5 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + log\_sl:LA & 89394.44 & 1.56 & 0.07 & -44670.22 &   6 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:sl & 89394.56 & 1.68 & 0.07 & -44670.28 &   7 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:log\_sl & 89394.57 & 1.69 & 0.07 & -44670.29 &   8 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:cos(ta) & 89394.59 & 1.71 & 0.07 & -44670.30 &   9 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:cos(ta) & 89394.63 & 1.75 & 0.06 & -44670.31 &  10 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + sl:cos(ta) & 89394.68 & 1.80 & 0.06 & -44672.34 &  11 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:log\_sl & 89394.69 & 1.81 & 0.06 & -44670.35 &  12 \\ 
+  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:cos(ta) & 89394.84 & 1.96 & 0.06 & -44670.42 &  13 \\ 
    \bottomrule

### 2021-04-22 Chapter_1/04_Manuscript/Manuscript.tex
Worked on manuscript for 1 chap

@@
 \usepackage{arydshln}
-% \usepackage[nomarkers, nolists]{endfloat}
+\usepackage[flushleft]{threeparttable}
+\usepackage[nomarkers, nolists]{endfloat}
 
@@
 movement model of dispersing wild dogs. We then use this model to simulate
-thousands of dispersers moving throughout the KAZA. Based on said simulations,
+thousands of dispersers moving across the KAZA-TFCA. Based on said simulations,
 we compute heatmaps and identify potential dispersal hotspots and compare them
@@
 simulated dispersal data can be analysed using network-analysis and how network
-metrics relevant to landscape connectivity can be computed. Our results suggest
+metrics relevant to landscape connectivity can be computed. Our results show
 that a simulation based approach yields several major benefits over traditional
@@
 \subsection{Study Area}
-The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E;
-\Cref{StudyArea}a) stretched over 1.3 Mio km\textsuperscript{2} and ecompassed
-the entire KAZA-TFCA (\Cref{StudyArea}b). The KAZA-TFCA is the world's largest
-transboundary conservation area and comprises parts of Angola, Botswana,
-Namibia, Zimbabwe, and Zambia, covering a total area of over 520'000
-km\textsuperscript{2}. Its landscape varies regionally and ranges from savanna,
-to grassland, and from dry to moist woodland habitats. A dominant
-hydrogeographical feature in its center is the Okavango Delta, the earth's
-largest inland delta. The delta and its surroundings are considered a stronghold
-for African wild dogs and may act as a source for the recolonization of
-surrounding habitats. The wet season within our study area lasts from November
-to March and is out of phase with the main flooding of the Okavango Delta which
-peaks between July and August \citep{McNutt.1996, Wolski.2017}.
+The study area was centered at -17\degree 13'9''S, 23\degree 56'4''E
+(\Cref{StudyArea}a) and was represented by a rectangular bounding box that
+stretched over 1.3 Mio. km\textsuperscript{2}. Its extent encompassed the entire
+KAZA-TFCA (\Cref{StudyArea}b), which is the world's largest transboundary
+conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
+and Zambia, covering a total of 520'000 km\textsuperscript{2}. Its landscape
+varies regionally and ranges from savanna, to grassland, and from dry to moist
+woodland habitats. A dominant hydrogeographical feature in its center is the
+flood-pulsing Okavango Delta, the earth's largest inland delta. The delta and
+its surroundings are considered a stronghold for African wild dogs and may act
+as a source for the recolonization of surrounding habitats \citep{Cozzi.2013}.
+The wet season within the study area lasts from November to March and is out of
+phase with the main flooding of the Okavango Delta which peaks between July and
+August \citep{McNutt.1996, Wolski.2017}. While large portions of land within the
+KAZA-TFCA are designated protected areas or national parks that remain largely
+free from humans, substantial human influence originates from roads,
+agricultural sites and several small to large settlements.
 
@@
     \includegraphics[width = \textwidth]{99_StudyArea.png}
-    \caption{Study area}
+    \caption{(a) The study area of our research was confined by a bounding box
+    encompassing the entire KAZA-TFCA, which comprises parts of Angola, Namibia,
+    Botswana, Zimbabwe, and Zambia. (b) The KAZA-TFCA represents the world's
+    largest terrestrial conservation area and covers a total of 520'000
+    km\textsuperscript{2}. It will connect or reconnect multiple already
+    existing national parks and protected areas (green polygons) that are
+    distributed across its extent. The African wild dog dispersal data
+    considered in this study was collected on free-ranging individuals departing
+    from the Moremi National Park in northern Botswana (since game reserves in
+    Botswana serve the same purpose as national parks, we refer to them as
+    national parks for simplicity).}
     \label{StudyArea}
@@
 \subsection{GPS Relocation Data}
-Between 2011 and 2020, we collected GPS relocation data on dispersing wild dogs
-from a free-ranging wild dog population inhabiting the Okavango Delta in
-northern Botswana. We identified potential dispersers based on age, number of
+Between 2011 and 2019, we collected GPS relocation data from dispersing wild
+dogs in a free-ranging wild dog population inhabiting the Moremi National Park
+in northern Botswana. We identified potential dispersers based on age, number of
 same‐sex siblings, pack size, and presence of unrelated opposite-sex individuals
@@
 cocktail of Ketamine/Xylazine/Atropine \citep{Osofsky.1996, Cozzi.2020},
-injected with a dart that was fired from a CO\textsubscript{2}-pressurized gun
-(\textit{DAN-Inject, Denmark}). Protocols for the anesthization protocols were
-approved by the Ministry of Environment, Natural Resources Conservation and
-Tourism of Botswana (permit EWT 8/36/4 XXXVI). After immobilization, individuals
-were fitted with GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic
-Aerospace GmbH, Berlin}) that included an automated drop-off mechanism. Handling
-and collaring of all individuals was carried out and supervised by a
-Botswana-registered wildlife veterinarian. All of the immobilized individuals
-quickly rejoined their pack within one hour after the procedure. Out of all
-collared individuals, 16 eventually dispersed in separate same-sex coalitions
-and their trajectories were successfully recorded (7 female and 9 male
-coalitions).
+injected with a dart, fired from a CO\textsubscript{2}-pressurized gun
+(\textit{DAN-Inject, Denmark}). Anesthesia protocols were approved by the
+Ministry of Environment, Natural Resources Conservation and Tourism of Botswana
+(permit EWT 8/36/4 XXXVI). After immobilization, individuals were fitted with
+GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
+Berlin}) that included an automated drop-off mechanism. Handling and collaring
+of all individuals was carried out and supervised by a Botswana-registered
+wildlife veterinarian. All of the immobilized individuals usually rejoined their
+pack within one hour after the procedure. Out of all collared individuals, 16
+individuals eventually dispersed in separate same-sex coalitions and their
+trajectories were successfully recorded (7 female and 9 male coalitions).
 
 During dispersal, collars were programmed to record a GPS fix every 4 hours.
-Recorded relocations were regularly transmitted over the Iridium satellite
+Collected relocations were regularly transmitted over the Iridium satellite
 system, which allowed remote tracking of individuals, even if they left the main
-study area and crossed international borders. To effectively distinguish between
-periods of residency and dispersal, we applied the net-squared displacement
-metric to the observed movement data. This metric measures the squared Euclidean
-distance of a collared individual to a reference point \citep{Borger.2012},
-which we set at the center of each disperser's natal home range. Hence,
+study area and crossed international borders. To distinguish between periods of
+residency and dispersal, we applied the net-squared displacement metric to the
+observed movement data. This metric measures the squared Euclidean distance of a
+collared individual to a reference point \citep{Borger.2012}. As a reference
+point, we used the center of each disperser's natal home range, such that
 dispersal was deemed to have started when an individual left its natal home
-range and ended when the individual became stationary again. For the purpose of
-this study, we discarded andy data that was collected during residency. Because
-previous research could not find any differences between males and females
-during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish
-sexes for the sake of our analyses. Ultimately, we converted the collected GPS
-relocations to steps, where each step represented the straight-line distance
-travelled between two consequtive GPS relocations \citep{Turchin.1998}.
+range and ended once individuals became sedentary again. For the purpose of this
+study, we discarded all data that was collected during residency and only
+retained movement data during dispersal. Previous research suggests that females
+and males behave similarly during dispersal \citep{Woodroffe.2019, Cozzi.2020},
+so we did not distinguish between sexes in our analyses. Ultimately, we
+converted the collected GPS coordinates to steps, where each step represented
+the straight-line distance traveled by and individual between two consecutive
+GPS relocations \citep{Turchin.1998}.
 
 \subsection{Covariates}
-To represent habitat covariates, we prepared a set of spatial raster layers
-depicting water-cover (dynamically updated), distance to water (dynamically
-updated), tree-cover, and shrub/grassland-cover. To render a decreasing impact
-with increasing distance, we square rooted the distance to water covariate
-layer. We also created a proxy for human influence, rendering anthropogenic
+We represented the physical landscape in the study area using a set of habitat
+covariates rendering water-cover, distance to water, tree-cover, and
+shrub/grassland-cover. Because water cover greatly changes within and between
+years around the Okavango Delta, we applied a remote sensing technique that
+allowed us to generate frequently updated water cover layers and corresponding
+distance to water layers (see \citealp{Wolski.2017} and \citealp{Hofmann.2021}).
+Hence, the obtained layers temporally aligned with our dispersal events.
+Furthermore, we computed a proxy for human influence, depicting anthropogenic
 pressures stemming from human-density, agricultural sites, and roads. We
-rendered all layers at a resolution of 250m by 250m for the entire extent of the
-KAZA-TFCA. Further details on the derivation and preparation of each
-environmental covariate is given in \cite{Hofmann.2020}.
+coarsened or interpolated all layers to a target resolution of 250 m by 250 m
+and cropped them to the extent of the KAZA-TFCA. Further details on the sources
+and preparation of each habitat covariate are given in \cite{Hofmann.2020}.
 
 Besides habitat covariates, we also computed movement metrics that we used as
-movement covariates in our models. These covariates included the step length
-(\(sl\)), its natural logarithm (\(log(sl)\)), as well as the cosine of the
-relative turning angle (\(cos(ta)\)). Because wild dogs follow a diurnal
-activity pattern, we furthermore coded a binary variable indicating whether an
-observed step was realized during main activity (xx to xx) or low activity (xx
-to xx).
+movement covariates in our models. Movement metrics were calculated for each
+step and included the step length (\textsf{sl}), its natural logarithm
+(\textsf{log(sl)}), and the cosine of the relative turning angle
+(\textsf{cos(ta)}). Because wild dogs follow a diurnal activity pattern, we also
+coded a binary variable (\textsf{LowActivity}) indicating whether a step was
+realized during low activity (17:00 to 07:00 local time) or high activity (09:00
+to 17:00 local time). We conducted all cleaning and analysis, as well as all
+simulations using the statistical software {\tt R}, version 3.6.6
+\citep{R.2019}. The corresponding {\tt R}-scripts are provided on Github.
 
@@
 We used integrated step selection functions (iSSF) to parametrize a mechanistic
-movement model of dispersing wild dogs. In this framework, observed steps are
-contrasted to random steps that the animal could have realized, but decided not
-to. To compare steps, it is assumed that animals assign a relative selection
-score \(w(x)\) to each step. The score depends on the covariates \(X\)
-experienced along a step, as well as the animals preferences \(\beta\) towards
-these covariates. The probability of a step being realized is then contingent on
-the step's selection score, as well as the selection scores of all other step in
-the same stratum. In contrast to regular step selection analysis,
-\textit{integrated} step selection analysis allows simulatneous inference on
-movement and habitat preferences of the studied animal. Moreover, potential
-interactions between habitat and movement preferences can be included to
-investigate how movement behavior depends on habitat characteristics. In result,
-the method produces less biased selection estimates and allows to apply the
+movement model of dispersing wild dogs. In contrast to regular step selection
+functions, \textit{integrated} step selection functions allow simultaneous
+inference on movement and habitat preferences of the studied animal, and on
+potential interactions between habitat and movement preferences. In result, the
+method produces less biased selection estimates and allows to treat the
 resulting model as a fully mechanistic movement model based on which movement
 can be simulated. To conduct iSSF analysis, we paired each observed step with 24
-random steps. Random steps basically resembled potential alternatives that the
-animal could have realized but decided not to. We generated random steps by
-sampling random turning angles from a uniform distribution (\(-\pi, +\pi\)) and
-step lengths from a gamma distirbution that was fitted to observed steps (scale
-= 6'308, shape = 0.37). While the number of random steps is inversely
-proportional to the sampling error \citep{Avgar.2016}, we found only minor
-differences in model performance when sampling additional random steps and
-deemed 24 random steps sufficient. Along each step, we extracted spatial
-covariates using the \textit{velox} package and calculated the movement metrics
-\(sl\), \(log(sl\)), and \(cos(ta)\). To facilitate model convergence, we
+random steps, alltogether forming a stratum that received a unique identifier.
+Each of the 24 random steps served as pseudeo-absence and resembled a potential
+alternative that the animal could have realized but decided not to. We generated
+random steps by sampling random turning angles from a uniform distribution
+(\(-\pi, +\pi\)) and step lengths from a gamma distirbution that was fitted to
+observed steps (scale = 6'308, shape = 0.37). While the number of random steps
+is inversely proportional to the sampling error \citep{Avgar.2016}, a
+preliminary analysis revealed only minor differences in model performance when
+sampling additional random steps, hence we deemed 24 random steps to be
+sufficient. Along each step, we extracted and averaged spatial covariates using
+the {\tt velox} package and we calculated the movement metrics \textsf{sl},
+\textsf{log(sl)}, and \textsf{cos(ta)}. To facilitate model convergence, we
 standardized all continuous covariates to a mean of zero and a standard
-deviation of one. We then used the r-package \textit{glmmTMB} to fit mixed
-effects conditional logistic regression models as proposed by \citep{Muff.2020}.
+deviation of one. Since correlation among covariates was low (\(|r| > 0.6\);
+\citealp{Latham.2011}), we retained all of them for modeling.
 
-Our movement model was and extension of the habitat model presented in Hofmann
-et al 2021. We will refer to it as the base model. Because the base model was
-used to predict landscape permeability, no interactions between the habitat and
-movement metrics were modeled. Thus, we expanded the base model and by proposing
-additional interactions between movement metrics and environmental covariates.
-For this, we started with the base model and iteratively increased model
-complexity by adding all proposed two-way interactions between habitat
-covariates and movement covariates. For instance, for the covariate
-\textit{water} we proposed the interactions \(Water:log(sl)\),
-\(Water:log(sl)\), and \(Water:cos(ta)\). Furthermore, we proposed the
-interactions \(sl:cos(ta)\) and \(log(sl):cos(ta)\) to account for a correlation
+To compare realized steps (scored 1) to random steps (scored 2), we assumed that
+animals assigned a selection score \(w(x)\) to each step. The score depended on
+the step's covariates \(X\), as well as the animals preferences \(\beta\)
+towards these covariates. The probability of a step being realized was then
+contingent on the step's selection score, as well as on the selection scores of
+all other step in the same stratum. To estimate the preferences of interest, we
+fitted conditional logistic regression models in the r-package {\tt glmmTMB}
+following the method proposed by \citep{Muff.2020}. This approach enables to fit
+random slopes in addition to random intercepts. Here, we used animal IDs as
+random effect and modelled random slopes for each covariate. To apply the
+``poisson trick'', we fixed the random intercept variance at an arbitrary high
+value of 10\textsuperscript{6} \citep{Muff.2020}.
+
+The structure of our movement model was based on a habitat selection model for
+dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model, no
+interactions between the habitat and movement covariates were considered, so we
+slightly expanded this base model by proposing interactions between movement and
+habitat covariates. More specifically, we started with the base model and
+incrementally increased model complexity by adding all possible two-way
+interactions between habitat covariates and movement covariates. For instance,
+for the covariate \textsf{water}, we proposed the interactions
+\textsf{Water:log(sl)}, \textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}.
+Besides those combinations, we also proposed the interactions
+\textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)} to account for a correlation
 between turning angles and step lengths, as well as the interactions
-\(sl:MainActivity\) and \(log(sl):MainActivity\) to account for the fact that
-step lengths diffuer due to the diurnal activity pattern. We then ran stepwise
-model forward selection based on Akaike's Information Criterion (AIC,
-\citealp{Burnham.2002}) values and identified the most parsimonious movement
-model. Although several models received an AIC weight above one, we decided to
-only consider the  ``best'' model for simplicity. In either way, the models with
-positive weights contained almost identical covariates and an averaged model
-would have only given weak support to additional covariates.
+\textsf{sl:LowActivity} and \textsf{log(sl):LowActivity} to account for the fact
+that step lengths may differ due to wild dogs' diurnal activity pattern. To
+compare competing models and assess the most parsimonious movement model, we ran
+stepwise forward model selection based on Akaike's Information Criterion (AIC,
+\citealp{Burnham.2002}) and computed model weights.
+
+We validated the predictive power of the selected movement model using a k-fold
+cross-validation procedure for case-control studies presented in
+\cite{Fortin.2009}. We randomly split the input data into training and testing
+data with an 80:20 ratio. Using the training data we parametrized a movement
+model based on which we predicted selection scores \(w(x)\) for all steps in the
+test data. Within each stratum we then assigned ranks 1-25 to each step based on
+predicted selection scores such that rank 1 was given to the step with the
+highest score \(w(x)\). Across all strata we determined the realized step's rank
+and calculated rank frequencies for realized steps. Finally, we computed
+Spearman's rank correlation between ranks and associated frequencies \(r_{s,
+realized}\). We replicated the entire procedure 100 times and computed the mean
+correlation coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\%
+confidence interval across all replicates. For comparison, we repeated the same
+procedure 100 times assuming random preferences. Random preferences were
+implemented by discarding the realized step from all strata and identifying the
+rank of a random step in each stratum. Again, we calculated Spearman's rank
+correlation coefficient (\(r_{s, random}\)), its mean across repetitions
+(\(\bar{r}_{s, random}\)), and its 95\% confidence interval. This validation
+proofs a significant prediction in case the confidence intervals of
+\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap.
+
+\subsection{Dispersal Simulation}
+We used the most parsimonious movement model to simulate a total of 80'000
+dispersing wild dogs moving across the KAZA-TFCA. The simulation resembled an
+inverted iSSF and was set up as follows. (1) We sampled a source point at which
+a disperser was initiated. We assumed a random initial orientation of the animal
+in order to permit calculation of relative turning angles. (2) We generated a
+set of 25 random steps by sampling turning angles from a uniform distribution
+(\(-\pi, +\pi\)) and step lengths from the fitted gamma distribution. In line
+with our training data, each step resembled the straight line movement realized
+within 4 hours. To prevent the sampling of unreasonably large steps, we capped
+sampled step lengths to a maximum of 35 km, which corresponds to the farthest
+distance ever traveled within 4 hours by our dispersers. (3) Along each random
+step, we extracted habitat covariates and calculated movement covariates. (4) We
+applied the parameterized movement model to predict the selection score \(w(x)\)
+for each of the proposed steps and translated selection scores into
+probabilities by applying \Cref{EQ2}. (5) We sampled one of the random steps
+based on predicted probabilities and used it to determine the animal's new
+position. We then repeated steps (2) to (5) until 2'000 steps were realized.
+
+Because some simulated dispersers would inevitably approach a map boundary, some
+of the proposed random steps may leave the study area, such that covariates
+could be extracted and no selection score could be calculated. To mitigate the
+influence of such edge effects, we followed \citep{Koen.2010} and artificially
+expanded all covariate layers by adding a 100 km buffer zone within which we
+randomized covariate values by resampling values from the observed covariate
+layers. This allowed simulated dispersers to leave and reenter the study area
+through this buffer zone. Nevertheless, if a disperser approached this outer
+boundary and one of the proposed random steps left the buffer, we resampled
+transgressing steps until they fully lied within the buffer zone, thereby
+forcing dispersers to bounce off the outer boundary. The simulation was mainly
+coded in the programming language {\tt R}, yet a few helper functions were
+written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
+\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
 
 \subsection{Source Points}
-We randomly placed source points within protected areas larger (\(>\) 700
-km\textsuperscript{2}), which conforms to the average home range requirement of
-resident wild dogs \citep{Pomilia.2015}. By distributing source points randomly,
-the number of source points per km\textsuperscript{2} within protected areas was
-approximately equal across our study area. In total, we distributed 50'000
-source points across all protected areas, each representing the starting point
-of a dispersal trajectory. To render potential immigrants into the study system,
-we randomly placed 10'000 additional source points inside a xx km buffer around
-the study area. Hence, a total of 60'000 source points were generated, thereby
-resulting in 60'000 simulated dispersers. It is worth pointing out that our
-choice of  placing source points completely random is arbitrary, as one as well
-adjust the sampling frequency based on the density of dispersers in the
-respective source area.
-
-% Simulations from step selection functions are known to be sensitive to the
-% location of source points, which is why we followed a twofold approach to sample
-% release points across the KAZA-TFCA. In a first approach, we used the same 68
-% source points between which we already computed least-cost paths and least-cost
-% corridors (see xx). These source points were regularly spaced 100 km apart,
-% located within protected areas (\(>\) 700 km\textsuperscript{2}). Because of
-% the regulary spacing, not all protected areas received source points. Hence, we
-% placed additional source points at the center of such protected areas. At each
-% of the generated source point we released 1000 dispersers, implying a total of
-% 68'000 simulated dispersers. In a second approach, we used the same 68 source
-% points to define catchment areas inside protected areas. That is, for each of
-% the 68 source points we defined its voronoi polygon within the point's protected
-% area. Within these catchment areas we then randomly sampled 1000 new source
-% points from which we released another 68'000 dispersers. Overall, we simulated
-% 134'000 dispersers for a total of 268 Mio. steps.
+To initiate virtual dispersers at biologically meaningful source points, we
+randomly placed source points within contiguous protected areas larger \(>\) 700
+km\textsuperscript{2} (\Cref{SourcePoints} (a)). This conforms to the average
+home range requirement of resident wild dogs \citep{Pomilia.2015} and allowed us
+to remove areas too small to host viable wild dog populations. By distributing
+source points randomly, the number of source points per km\textsuperscript{2}
+within protected areas was approximately equal across our study area. Across all
+protected areas, we distributed 50'000 source points, each representing the
+starting point of a single dispersal trajectory. To render potential immigrants
+into the study system, we placed additional 30'000 source points inside the
+buffer zone around the main study area (\Cref{SourcePoints} (b)).
 
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_SourcePoints.png}
-    \caption{Illustration of source points from which dispersal was simulated.
-    Although in reality we simulated 1'000 per source point, we illustrate an
-    example assuming 10 dispersers from each source point. (a) Static source
-    points similar to the source points reported in ... (b) Source points that
-    were randomized within the catchment areas (dark gray, delinated by solid
-    black lines).}
+    \includegraphics[width = \textwidth]{99_SourceAreas.png}
+    \caption{(a) Illustration of an artificially expanded covariate layer (water
+    cover). We expanded the layer by adding a 100 km buffer (light blue) which
+    we filled with values sampled from the original layer. (b) Different source
+    areas from which we released virtual dispersers. We initiated 50'000
+    dispersers within the main study area (green) and another 30'000 dispersers
+    within a virtual buffer (blue).}
     \label{SourcePoints}
@@
 
-\subsection{Dispersal Simulation}
-We used the most parsimonious movement model to simulate a total 80'000
-dispersing wild dogs across the KAZA-TFCA. The simulation basically resembled an
-inverted iSSF and was set up as follows. First, we sampled a source point at
-which a disperser was initiated and assumed a random initial orientation of the
-animal. The orientation was essential to permit calculation of relative turning
-angles. Second, we generated a set of 25 random steps by sampling turning angles
-from a uniform distribution (\(-\pi, +\pi\)) and step lengths from the fitted
-gamma distribution. Similarly to our input data, each step corresponded to a
-temporal difference of 4 hours. To prevent unrealistically large steps, we
-capped sampled step lengths at 35km, which corresponds to the farthest distance
-travelled by one of our dispersers within 4 hours. Third, along each random step
-we extracted habitat covariates and calculated movement covariates. Fourth, we
-applied the parametrized movement model to predict the selection scores \(w(x)\)
-for all steps. Fifth, we translated selection scores into probabilities by
-applying \Cref{EQ2}. Sixth, we sampled one of the random steps based on assigned
-probabilities and determined the animal's new position. We then repeated steps
-two to six until a total of 2'000 steps were realized.
+\subsection{Heatmaps}
+To identify dispersal hotspots across our study area, we rasterized all
+simulated trajectories and created a heatmap indicating the absolute frequency
+at which each raster-cell in the study area was visited by our virtual
+dispersers. If the same trajectory crossed a raster-cell twice, it was only
+counted once. That is, we did not consider revisits, thereby mitigating biases
+arising from individuals that were trapped in a certain locations and moved
+around in circles. We achieved high performance rasterization of all simulated
+trajectories using the R-package {\tt terra} \citep{Hijmans.2020}. While one
+could easily subset the data to visualize heatmaps assuming different dispersal
+durations, we retained all 2'000 simulated steps to create a single heatmap.
 
-Inevitably, some simulated dispersers would approach a map boundary and one of
-the generated random steps would leave the study area. However, beyond the study
-area no covariates could be extracted and no selection scores can be calculated.
-To mitigate the influence of such edge effects, we buffered all covariate layers
-by xx percent and filled the buffer zone with covariate values resampled from
-the observed data. As such, simulated dispersers were allowed to leave and
-reenter the study area through this buffer zone. In case one of the proposed
-random steps left the buffer, we resampled the corresponding step until it lied
-fully within the buffer zone. This implies that we forced the disperser to
-bounce off the borders of the buffer zone.
+% To examine if and how ``heat'' changes in response to changes in the location of
+% source points and the number of simulated steps, we followed a 2 x 6 design and
+% created heatmaps for both point sampling regines, as well as for 68, 125, 250,
+% 500, 1000, and 2000 dispersal steps. We quantified the similarity of the
+% resulting 12 heatmaps to the permeability and least-cost corridor maps presented
+% in (Hofmann ...) we used Bhattacharyya's affinity. Bhattacharyya's affinity
+% ranges from zero (complete separation) to one (perfect match) and has earlier
+% been proposed to compare the overlap of utilisation distributions (Fieberg).
 
-To meaningfully select source points, we randomly sampled coordinates within
-protected areas larger (\(>\) 700 km\textsuperscript{2}), which conforms to the
-average home range requirement of resident wild dogs \citep{Pomilia.2015}. By
-distributing source points randomly, the number of source points per
-km\textsuperscript{2} within protected areas was approximately equal. In total,
-we distributed 50'000 source points across protected areas. To render potential
-immigrants into the study area, we placed 10'000 additional random points inside
-a xx km buffer around the study area. In total, we generated 80'000 source
-points, thereby resulting in 80'000 simulated dispersers, each of which moved
-2'000 steps. Overall, this implied 160 Mio. simulated steps.
+\subsection{Network Analysis I}
+To pinpoint areas that are of high relevance for connecting remote regions in
+our study area, we adopted a network view on the simulated trajecotries and
+computed the betweenness metric. More specifically, we overlaid the study area
+(including the buffer) with a with a raster that had a cell size of 5 x 5 km,
+where each raster-cell represented a node in the final network. We then used the
+simulated trajectories to determine all transitions occurring from one cell to
+another, as well as the frequency at which these transitions occurred. This
+resulted in an edge-list that we translated into a weighted network using the
+r-package {\tt igraph}. Because {\tt igraph} handles the weights \(\omega\) as
+costs, we inverted the traversal frequency at each cell by applying \(\omega =
+\frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal Frequency_i}\). Finally, we
+calculated the betweenness metric for each node in the final network. This
+metric indicates how often a specific raster-cell lies on a shortest path
+between two other raster-cells and is therefore a useful metric to detect
+movement corridors \citep{BastilleRousseau.2018}.
 
-\subsection{Heatmaps}
-Using the simulated trajectories, we created heatmaps that indicate the
-frequency at which each raster-cell in the study area was visited. If the same
-trajectory crossed a pixel twice, it was only counted once, i.e. we did not
-consider revisits. We achieved high performance rasterization of all simulated
-trajectories using the recently developed R-package \textit{terra}
-\citep{Hijmans.2020}.
-
-To examine if and how ``heat'' changes in response to changes in the location of
-source points and the number of simulated steps, we followed a 2 x 6 design and
-created heatmaps for both point sampling regines, as well as for 68, 125, 250,
-500, 1000, and 2000 dispersal steps. We quantified the similarity of the
-resulting 12 heatmaps to the permeability and least-cost corridor maps presented
-in (Hofmann ...) we used Bhattacharyya's affinity. Bhattacharyya's affinity
-ranges from zero (complete separation) to one (perfect match) and has earlier
-been proposed to compare the overlap of utilisation distributions (Fieberg).
-
-\subsection{Network Analysis}
-We coerced all simulated trajectories into a network consisting of vertices
-(relocations) and edges (connections between relocations). To do so, we created
-raster layers at multiple resolutions and identified each trajectory's
-transition matrix on these rasters see figure xx \citep{BastilleRousseau.2018}.
-We then merged the transition matrices of all trajectories and calculated
-cumulative transitions between all raster-cells. This resulted in an edge-list,
-containing all observed from-to connections as well as their frequency. Using
-this edge-list, we generated a weighted graph using the r-package
-\textit{igraph}. Based on this graph we calculated betweenness scores as well as
-the degree of each raster cell. Betweenness indicates how often a specific
-raster-cell lies on a shortest path between two other raster-cells and is a
-useful metric to detect movement corridors. Degree, on the other hand, indicates
-how many connections a raster cell has and therefore serves to illustrate the
-xxx of specific nodes. When calculating betweenness, we used the transition
-frequency as weighting factor. That is, a higher transition frequency
-contributed to a higher betweenness score.
-
-% Based on the simulated trajectories we identified to which other areas each
-% source area is connected. For instance, if a trajectory originated at source
-% point one and intersected with source areas two and three, we assumed that
-% source areas two and three were within reach of source are one. To get a sense
-% of the strength of connections, we also calculated how often each of these
-% connections was realized. This procedure resulted in 6 x 2 edge lists which we
-% further used to generate weighted networks and to calculate network metrics.
+\subsection{Network Analysis II}
+Besides this, we also adopted an alternative network view and identified the
+frequency at which dispersers originating from one national park successfully
+moved into an other national park. We achieved this by looking at each pairwise
+combination of national parks and determining the relative number of
+trajectories that originated in one national park and had at least one
+subsequent coordinate laying in the other national park. This allowed us to
+determine the average duration it took a simulated disperser to move from one
+national park to another, and to point out combinations of national parks
+between no links were observed.
 
@@
 \subsection{Movement Model}
-Compared to the base model, the most parsimonious movement model included
-several additional interactions (\Cref{MovementModel} and Table S1). The model
-indicates that dispersers move directional, particularly when distant to water,
-yet less so in human dominated landscapes. Furthermore, dispersers prefer large
-steps, especially during main activity. In contrast, step lengths tend to be
-shorter when water- or tree-cover is high. In general, dispersers avoid water,
-prefer proximity to water, avoid dense tree-cover, prefer shrubs/grassland, and,
-finally, avoid human dominated landscapes.
+Compared to the base model reported in \citep{Hofmann.2021}, our most
+parsimonious movement model included several additional interactions between
+habitat and movement covariates (\Cref{MovementModel} and
+\Cref{MovementModelNumbers}). Although multiple models received an AIC weight
+above zero (T1 in Appendix S1), we only consiered results from the  ``best''
+model for further analyses. Since all models with positive AIC weight contained
+similar covariates, this decision only marginally influenced subsequent results.
+Results from the selected model are given in \Cref{MovementModelNumbers} and
+illustrated in \Cref{MovementModel} (a). Additional plots to ease with the
+interpretation of the model are provided in Appendix XX.
+
+When holding constant for movement behavior, we find that dispersing wild dogs
+avoid water, but prefer its proximity. Dispersers also avoid densely forested
+woodlands, yet prefer open shrublands or grasslands. Finally, dispersers avoid
+moving through landscapes that are influenced by humans.
+
+When looking at the movement kernel, we observe several significant
+interactions. However, except for the interaction \textsf{sl:LowActivity},
+estimated slopes are relatively flat, suggesting that our initial distributions
+for step lengths and turning angles were only marginally biased. For instance,
+the positive effect for \textsf{cos(ta)} indicates that observed turning angles
+are slightly more directional compared to the turning angles proposed by our
+uniform distribution. On the other hand, the strongly significant and negative
+interaction \textsf{sl:LowActivity} reveals that our fitted gamma distribution
+produced steps lengths that were substantially larger that those realized by the
+dispersers during low wild dog activity.
+
+However, we also find a strongly significant and negative interaction between
+the step length and main-activity indicator, showing that realized step lengths
+tended to be much shorter during periods of activity in comparison to the steps
+proposed by our fitted gamma distribution.
+
+Finally, the significant interactions \textsf{cos(ta):HumanInfluence} and
+\textsf{cos(ta):DistanceToWater\textsuperscript{2}} indicate that dispersers
+move much more tortuous in areas influenced by humans but more directed when
+distant to water.
+
+Results from the k-fold cross-validation procedure show that our model
+substantially outperforms a random guess, as the confidence intervals of r.. and
+r.. do not overlap. Additionally, we find that the rank correlation slightly
+improved in comparison to the base model reported in \citep{Hofmann.2021}.
 
@@
     Results from the k-fold cross validation procedure. The upper plot shows
-    rank frequencies of predicted realized scroes according to model predictions
+    rank frequencies of predicted realized scores according to model predictions
     with known preferences, whereas the lower plot shows rank frequencies when
@@
 
+\begin{table}
+  \begin{center}
+  \caption{Most parsimonious movement model for dispersing wild dogs. The model
+  comprises of a movement and habitat kernel, where the movement kernels
+  describes preferences with regards to movement behavior, whereas the habitat
+  kernel describes preferences with respect to the habitat. Finally, the two
+  kernels can interact, such that movement preferences are contingent on habitat
+  conditions.}
+  \label{MovementModelNumbers}
+  \resizebox{\textwidth}{!} {
+    \begin{threeparttable}
+      \begin{tabular}{llrrrc}
+        \toprule
+        Kernel & Covariate & Coefficient & SE & pvalue & Significance \\
+        \midrule
+        \multirow{5}{*}{Habitat Kernel}
+         & Water & -0.546 & 0.112 & \(<\) 0.001 & *** \\
+         & DistanceToWater \textsuperscript{0.5} & -0.390 & 0.231 & 0.092 & * \\
+         & Woodland & -0.364 & 0.086 & \(<\) 0.001 & *** \\
+         & Shrubs/Grassland & 0.288 & 0.092 & 0.002 & *** \\
+         & HumanInfluence & -0.535 & 0.229 & 0.019 & ** \\
+        \hdashline
+        \multirow{6}{*}{Movement Kernel}
+         & sl & 0.075 & 0.037 & 0.042 & ** \\
+         & cos(ta) & 0.105 & 0.031 & 0.001 & *** \\
+         & log(sl) & 0.146 & 0.051 & 0.004 & *** \\
+         & cos(ta) : sl & 0.049 & 0.026 & 0.064 & * \\
+         & cos(ta) : log(sl) & 0.076 & 0.026 & 0.003 & *** \\
+         & sl : LowActivity & -0.917 & 0.113 & \(<\) 0.001 & *** \\
+        \hdashline
+        \multirow{5}{*}{Interaction}
+         & sl : Water & -0.305 & 0.076 & \(<\) 0.001 & *** \\
+         & sl : Woodland & -0.089 & 0.039 & 0.023 & ** \\
+         & sl : Shrubs/Grassland & 0.124 & 0.058 & 0.032 & ** \\
+         & sl : DistanceToWater \textsuperscript{0.5} & -0.058 & 0.031 & 0.056 & * \\
+         & cos(ta) : HumanInfluence & -0.040 & 0.022 & 0.070 & * \\
+         & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063 & 0.026 & 0.017 & ** \\
+         \bottomrule
+      \end{tabular}
+       \begin{tablenotes}
+         \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
+         \quad *** \(p < 0.01\)}
+       \end{tablenotes}
+    \end{threeparttable}
+    }
+  \end{center}
+\end{table}
+
 \subsection{Dispersal Simulation}
+On a machine with an AMD Ryzen 7 2700X processor with 8 x 3.6 GHz and 64 GB of
+RAM, a single batch of 1'000 simulated dispersers took roughly 90 minutes to
+compute (\(\mu = 88.90\), \(\sigma = 1.87\)). As such, the simulation of all
+80'000 dispersers terminated after 120 hours, i.e. five days. However,
+comparable computations will be substantially quicker for smaller study areas,
+as the covariate extraction from large rasters was computationally the most
+expensive task.
+
+On average, step lengths realized by the simulated dispersers (\(\mu = 2'093\)
+m, \(\sigma = 3'067\)) were slightly shorter than those by observed dispersers
+(\(\mu = 2'326\) m, \(\sigma = 3'323\)). Simultaneously, a slightly smaller
+\(cos(ta)\) indicated that simulated dispersers moved with marginally lower
+directionality (\(\mu = 0.057\), \(\sigma = 0.071\)) compared to observed
+dispersers (\(\mu = 0.078\), \(\sigma = 0.072\)). These differences in step
+lengths and turning angles can be attributed to minor disparities between
+habitat conditions at the area within which we collected training data and
+habitat conditions within the entire study area.
+
+Out of the 50'000 dispersers initiated in protected areas, only 4.5\% eventually
+hit a map boundary, highlighting that boundary effects should be neglectable. In
+contrast, 78\% of the 30'000 dispersers originating from the buffer zone hit a
+map boundary at least once.
 
 \subsection{Heatmaps}
-Six of the twelve rasterized dispersal trajectories are presented as heatmaps in
-\Cref{Heatmaps}. As can be seen, differences that stem from the method of point
-sampling disappear as more steps are simulated. This is to be expected as the
-influence of the origin becomes smaller and smaller as the animal moves through
-the landscape. However, there are striking differences when simualtions are only
-run for few iterations.
+\Cref{Heatmap} depicts the heatmap rendering the traversal frequency at each
+pixel of the study area across all simulated individuals and steps. The map
+illustrates that large portions of land beyond the borders of the KAZA-TFCA are
+only infrequently visited by dispersers (dark blue areas). On the contrary, we
+observe that within the KAZA-TFCA extensive regions are regularly visited by
+dispersers (bright yellow and red areas). Most notably, the region in northern
+Botswana south of the Linyanti swamp glows in rich red and stands out as highly
+frequented dispersal hub. Nevertheless, massive water bodies such as the
+Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, pose considerable
+dispersal barriers within and therefore limit realized connectivity the
+KAZA-TFCA.
 
 \begin{figure}
-  \includegraphics[width=\textwidth]{99_Heatmaps}
+  \includegraphics[width=\textwidth]{99_Heatmap.png}
   \caption{}
-  \label{Heatmaps}
+  \label{Heatmap}
 \end{figure}
 
-Bhattacharyya's affinity index supports the notion that heatmaps become more
-similar to the earlier developed permeability and corridor maps as the number of
-simulated steps increases. Furthermore, it appears that randomly sampled source
-points contribute to a higher similarity too, albeit differences due to the
-sampling regime vanish as the number of simulated steps increases. In fact, both
-maps are almost identical to each other after 2000 steps. Furthermore, the
-connectivity networks become increasingly similar to the previously published
-permeability and corridor maps. Still, some differences remain even after 2000
-simulated steps, highlighting that some severe impediments in the landscape
-exist.
+\subsection{Network Analysis I}
+The results of our first network analyses are presented in \Cref{Betweenness}.
+In contrast to the heatmap, \Cref{Betweenness} puts much more emphasis on
+discrete dispersal corridors. Nonetheless, the dispersal hub in northern
+Botswana stands out again and is traversed by a corridor that receives a
+comparably high betweenness score. This implies that the region is particularly
+crucial for connecting other pixels in the study system and hence represents a
+proper dispersal hub. Towards east, the corridor runs through the Chobe National
+Park into the Hwange national park, where it branches out and further extends
+into the distant Matusadona National Park in Zimbabwe. Northwest of the Linyanty
+ecosystem, the same corridor expands into Angola, where it splits and finally
+traverses over a long stretch of unprotected area into the Kafue National Park
+in Zambia. Several additional corridors with slightly lower betweenness scores
+exist, yet most of them run within the boundaries of the KAZA-TFCA. In general,
+only few corridors appear to directly link the peripheral regions of the
+KAZA-TFCA. For instance, there are no viable dispesal corridors between the
+Matusadona National Park in Zimbabwe and the Kafue National Park in Zimbabwe.
+Similarly, there are no direct links between the Zimbabwean and Angolan
+``spikes'' of the KAZA-TFCA.
 
-% \begin{table}[hbtp]
-%   \caption{Summary statistics of all GPS relocations that have been recorded on
-%   dispersing coalitions.}
-%   \label{GPSData}
-%   \begin{center}
-%   \resizebox{\textwidth}{!}{
-%     \begin{tabular}{lllllllll}
-%     \toprule
-%     sampling & Metric & Map & 68 & 125 & 250 & 500 & 1000 & 2000 \\
-%     \midrule
-%     Static & Affinity & Corr & 0.64 & 0.70 & 0.76 & 0.80 & 0.82 & 0.84 \\
-%     Static & Affinity & Perm & 0.67 & 0.74 & 0.80 & 0.85 & 0.89 & 0.91 \\
-%     Static & Correlation & Corr & 0.32 & 0.38 & 0.45 & 0.51 & 0.56 & 0.60 \\
-%     Static & Correlation & Perm & 0.47 & 0.57 & 0.66 & 0.72 & 0.78 & 0.82 \\
-%     \hdashline
-%     Random & Affinity & Corr & 0.70 & 0.75 & 0.79 & 0.82 & 0.84 & 0.85 \\
-%     Random & Affinity & Perm & 0.74 & 0.79 & 0.83 & 0.87 & 0.90 & 0.92 \\
-%     Random & Correlation & Corr & 0.38 & 0.43 & 0.48 & 0.53 & 0.57 & 0.61 \\
-%     Random & Correlation & Perm & 0.58 & 0.64 & 0.69 & 0.75 & 0.79 & 0.83 \\
-%     \hdashline
-%     n.a. & Affinity & Heat & 0.76 & 0.86 & 0.94 & 0.97 & 0.99 & 0.99 \\
-%     n.a. & Correlation & Heat & 0.90 & 0.94 & 0.97 & 0.98 & 0.99 & 1.00 \\
-%     \bottomrule
-%     \end{tabular}
-%   }
-%   \end{center}
-% \end{table}
+\begin{figure}
+  \includegraphics[width=\textwidth]{99_Betweenness.png}
+  \caption{Betweenness scores of each raster cell in a raster with 5 x 5 km
+  resolution. Betweenness scores were determined based on simulated dispersal
+  events. A high betweenness score highlights cells that are exceptionally
+  relevant in connecting different regions in the study area. That is, the
+  higher the betweenness score, the more often a pixel lies on a shortest path
+  between adjacent areas. In this sense the metric can be used to pinpoint
+  discrete movement corridors. Note that we square-rooted betweenness scores to
+  improve visibility of corrdiors with low scores.}
+  \label{Betweenness}
+\end{figure}
 
-\subsection{Network Analysis}
+\subsection{Network Analysis II}
+The results from the second network analysis that was focused on connectivity
+between national parks are given in \Cref{AreasReached}. The map shows between
+which national parks direct links exist and how frequent they are, as well as
+the average duration a disperser had to move to realize those links. For
+instance, 6.8 \% of the simulated dispersers originating from the Moremi
+National Park successfully reached the Chobe National Park and 4.2 \% reached
+the Hwange National Park in Zimbabwe. On average, dispersers moved for 623 steps
+before arriving at Chobe (SD = 520) and for 1'413 steps before arriving at
+Hwange (SD = 371).
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{99_AreasReached.png}
+  \caption{Network on simulated dispersal trajectories highlighting the
+  connections between national parks (dark green). Yellow bubbles represent the
+  center of the different national parks and are sized in relation to the number
+  of simulated dispersers originating from each park. Colored arrows between
+  national parks illustrate between which national parks the simulated
+  dispersers successfully moved and the color of each arrow shows the average
+  number of steps that were necessary to realize those connections.
+  Additionally, the line thickness indicates the relative number of dispersers
+  originating from a national park that realized those connections.}
+  \label{AreasReached}
+\end{figure}
 
@@
 
-Empircally informed dispersal models will also be invaluable in spatially
-explicit population models in which dispersal is actually rendered
-mechanistically.
+All of our findings are well in line with our previous work, where we have
+highlighted dispersal corridors for wild dogs using least-cost approaches. This
+suggests that qualitative results are quite insensitive to the exact
+methodological approach. Still, we believe that a simulation based approach
+offers possibilities for much richer inferences compared to traditional
+approaches. This is largely due to the fact that proper movement trajectories
+are generated that can be analysed \textit{as if} they were generated by real
+dispersers. This is of course contingent on the assumption that underlying
+models are adequately representing movement behavior of the focal species and
+calls for further methods to validate the predictive power of such simulations.
 
-In addition, individual based simulations will allow to explicitly model
-dispersal between subpopulations in models of population dynamics.
+While the segment running into Kafue receives a high betweenness score, it was
+actually only rarely traversed by our simulated dispersers, as can be seen from
+the dark colors in this region in \Cref{Heatmap}. It is therefore worth noting
+that the betweenness metric highlights crucial bottlenecks that are relevant for
+connecting remote regions, yet does not directly yield information about the
+absolute frequency at which these bottlenecks are used.
+
+We have previously attributed the weak significance of distance to water to the
+fact that we did not control for the presence or absence of conspecifics. We
+stick to this reasoning as our expanded model still shows a rather large
+uncertainty around the respective beta coefficients. To better gauge the
+imporance and influence of this covariate, future studies will need to control
+for inter- and intra-speficic interactions that may explain why and when
+dispersers are attracted to or afraid of waterbodies.
+
+For our simulations, we represented the Okavango Delta statically and assumed a
+relatively extended flood. This resulted in a quasi-barrier, formed by the
+Okavango-Delta and the adjacent city of Maun, which was rarely traversed by
+simulated individuals. Out of the 499 dispersers initiated inside the Moremi
+National Park, only 101 managed to reach the south-western section of the Delta,
+whereas 284 eventually reached the equally distant Linyanti swamp. In this
+regard, the heatmap presented in \Cref{Heatmap} may be most representative of
+the period shortly after the wet-season, when floodlevels in the Delta are at
+their maximum. During the dry season, however, the flood considerably retracts
+and potentially clears the way for wild dogs dispersing from the Moremi-Game
+reserve into the south-western section of the Delta. Future studies could relax
+the assumption of a static flood and attempt to update floodlevels as the
+dispersers move. This would allow studying how connectivity within the ecosystem
+evolves over time as the flood climaxes and retracts again. In fact, one of the
+major advantages of such simulation based approaches is that a dynamic
+environment can be rendered, as time is explicit in these models
+\citep{Zeller.2020}. This contrasts with traditional modelling approaches such
+least-cost analysis or circuit theory, where the temporal dimension cannot be
+made explicit. An explicit view on time also directly translates in insights on
+the duration required by dispersers to move between distinct patches such as
+national parks or spatially segregated subpopulations.
+
+Comparable simulations that are based on empirical data are also a fundamental
+component for spatially realistic population models in which dispersal is
+rendered more realistically and does not merely depend on the distance between
+habitat patches.
+
+We did not model mortality during dispersal in our simulations. This was a
+simplifying assumption and only inaccurately reflects dispersal in reality. It
+is well known that wild dogs often die during dispersal in result of deadly
+encounters with other predators, but also due to road kills and illegal
+shootings by humans. In result, mortality during dispersal may limit realized
+connectivity, especially in areas that wild dogs are unfamiliar with or in areas
+with high potential for human-wildlife conflict \citep{Cozzi.2020}.
+Interestingly though, \cite{Behr.2021} have recently demonstrated that mortality
+during dispersal is lower compared to residency, suggesting the presence some
+fitness benefits to dispering individuals.
 
@@
 
-Comparable simulations could also serve to predict the likely whereabouts of
-dispersers as they move. This information can then be utilized as an  ``early
-warning'' system, such that potential encounters can be gauged ad hoc.
+In some European countries, the comeback of large predators, such as bears,
+lynx, and the wolf, has triggered emotional discussions and raised public
+concern \citep{Behr.2017}, particularly in areas with free-roaming livestock
+that may be prayed upon by the returned species. In cases were recent locations
+of such predators are known, an early warning system based on simulations could
+serve to forewarn about potential encounters and and thereby increase public
+acceptance of large predators.
 
@@
 D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modelling; D.D.H., D.M.B., and G.C. wrote the
+data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
 first draft of the manuscript and all authors contributed to the drafts at

### 2021-04-22 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/GeneralMetrics.R

@@ -1,0 +1,151 @@
+################################################################################
+#### Some Descriptive Metrics
+################################################################################
+# Description: A simple plot of the Dispersal Durations
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(raster)       # For spatial stuff
+library(rgdal)        # For spatial stuff
+library(rgeos)        # For spatial stuff
+library(davidoff)     # Custom functions
+library(pbmcapply)    # For multicoring
+
+################################################################################
+#### Dispersal Simulation
+################################################################################
+# Get an idea of the dispersal simulation duration
+info <- read_csv("03_Data/03_Results/99_Simulations/Report.csv")
+
+# Average and sd of simulation duration
+hist(info$duration, breaks = 20)
+mean(info$duration)
+sd(info$duration)
+
+# Load simulated and observed trajectories
+sims  <- read_rds("03_Data/03_Results/99_DispersalSimulation.rds")
+obs   <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF).csv")
+
+# Check how many simulated trajectories eventually hit a map boundary
+sims %>%
+  select(TrackID, BoundaryHit, Area) %>%
+  group_by(TrackID) %>%
+  summarize(BoundaryHit = max(BoundaryHit, na.rm = T), Area = unique(Area)) %>%
+  group_by(Area) %>%
+  summarize(BoundaryHit = mean(BoundaryHit))
+
+# Get an idea of the step lengths and turning angles of simulated and observed dispersers
+obs   <- subset(obs, case_)
+sims <- select(sims, ta_, sl_)
+obs <- select(obs, ta_, sl_)
+
+# Put them together
+sims$Type <- "Simulated"
+obs$Type <- "Observed"
+dat <- rbind(sims, obs)
+
+# Average and sd of simulated step lengths
+dat %>%
+  group_by(Type) %>%
+  summarize(
+      mean_cos_ta = mean(cos(ta_), na.rm = T)
+    , sd_cos_ta   = sd(cos(ta_), na.rm   = T)
+    , mean_sl     = mean(sl_, na.rm      = T)
[…]

### 2021-04-22 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/MovementModelAICs.R

@@ -1,0 +1,348 @@
+################################################################################
+#### Plot of Model AICs
+################################################################################
+# Description: Plot of model AICs
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(lubridate)    # To handle dates nicely
+library(ggpubr)       # For nice plots
+library(davidoff)     # Custom functions
+library(glmmTMB)      # To handle model results
+library(lemon)        # For nice axis labels
+library(viridis)      # For nice colors
+library(ggdark)       # Dark ggplot theme
+library(latex2exp)    # For latex expressions
+library(msir)         # For prediction interval around loess
+library(RColorBrewer) # For nice colors
+library(xtable)       # To write model results to a .tex file
+
+# Suppress the scientific notion
+options(scipen = 999)
+
+################################################################################
+#### Prepare Data
+################################################################################
+# Load the model aics
+aics <- readRDS("03_Data/03_Results/99_MovementModelAICs.rds")
+
+# Also remove the ModelID column
+aics$ModelID <- NULL
+
+# Remove models with 0 weight
+aics <- subset(aics, Weight > 0)
+
+# Replace covariates with short code and add cos(ta) and log(sl) as covariatess
+aics$Covariates <- aics$Covariates %>%
+  gsub(pattern = "\\bShrubs, SqrtDistanceToWater, Trees, Water, HumansBuff5000\\b", replacement = "Base Model") %>%
+  gsub(pattern = "\\bWater\\b", replacement = "WA") %>%
+  gsub(pattern = "\\bShrubs\\b", replacement = "SH") %>%
+  gsub(pattern = "\\bTrees\\b", replacement = "WO") %>%
[…]

### 2021-04-30 Chapter_1/04_Manuscript/Manuscript.tex
Worked on manuscript

@@
 \begin{abstract}
-For many species, a sufficient degree of landscape connectivity is a crucial
-prerequisite for long-term species persistence. This is especially true for
-animals that exhibit long-distance dispersal, as dispersers typically cross a
-variety of habitats and potentially move into unfamiliar territory. Estimating
-habitat permeability and quantifying landscape connectivity have therefore
-become two major tasks in conservation ecology. For this, least-cost analysis
-and circuit theory have been the two workhorses during the past two decades.
-This is largely owed to their ease of use and intuitive nature which has
-facilitated the application of the methods across a broad range of the animal
-kingdom. However, both methods make several restricting assumptions that limit
-their usefulness in reality.
-
-Recent innovations in movement ecology, particularly on step selection analysis,
-have brought forward novel ways to look at connectivity. For instance,
-integrated step selection functions provide a means to parametrize a fully
-mechanistic movement model based on which virtual dispersers could be simulated.
-While this approach has been used to infer habitat utilization, it may also
-serve to highlight landscape connectivity and pinpoint dispersal corridors.
-
-Here, we propose the simulation of dispersal trajectories as a much more generic
-way to identify dispersal barriers and potential movement corridors. To achieve
-this, we applied integrated step selection analysis and uses data from 16
-dispersing wild dogs to parametrize a fully mechanistic movement model for
-dispersing African wild dogs. The model rendered dispersers' habitat and
-movement preferences, as well as potential interactions between them. Based on
-the model, we simulated myriads of dispersal trajectories and inferred dispersal
-corridors across the earth's largest transboundary conservation area, the
-Kavango-Zambezi Transfrontier Conservation Area. We also exemplify how such
-simulations could be analysed using network theory. Finally, we discuss the
-benefits and pitfalls of dispersal simualtions and highlight potential
-improvements to be made in the future.
+For many species, dispersal represents a crucial process that allows to avoid
+inbreeding, to colonize new habitats and to reinforce small and unviable
+subpopulations. Successful dispersal is therefore a crucial pre-requisite for
+long-term species persistence in wild animal populations. However, the ability
+to disperse is contingent a sufficient degree of landscape connectivity, which
+is why the estimation of connectivity and identification of dispersal corridors
+has become a task of extraordinary importance in ecological studies and for the
+planning of conservation initiatives.
+
+Over the past two decades, scientists have mainly relied on analytical tools
+such as least-cost methods and circuit theory to investigate landscape
+connectivity. This is largely owed to their ease of use and intuitive nature,
+which have facilitated their application in the past. Despite their usefulness
+for various ecological applications, both methods make several restricting
+assumptions that can limit their applicability in reality. Least-cost analysis,
+for instance, suffers from the fact that animals are assumed to have perfect
+knowledge about the entire landscape and choose cost-minimizing routes
+accordingly. On the contrary, circuit theory assumes that animals walk according
+to a complete random walk, which is equally unlikely to hold in reality.
+
+Fortunately, recent innovations in movement ecology have brought forward novel
+opportunities to study dispersal and landscape connectivity. In particular, the
+rich suite of resource selection functions has undergone substantial
+improvements in recent years. Most notably, step selection functions have been
+generalized to \textit{integrated} step selection functions, which essentially
+represent fully mechanistic movement models based on which individual animal
+movement could be predicted and simulated. While such models have been applied
+to simualte the steady state utilization distribution, similar approaches could
+be useful for investigating transient movement behavior and study landscape
+connectivity in a much more natural way.
+
+Here, we showcase the use of integrated step selection analysis to simulate
+dispersal of a top-predator across a large study area. For this, we collected
+data on 16 dispersing wild dogs and analysed it in an integrated step selection
+analysis to produce a fully mechanistic movement model rendering wild dog
+dispersal. Based on this model, we simulated 80'000 dispersers moving across the
+Kavango-Zambezi Transfrontier Conservation Area, the world's largest
+transboundary conservation area. We summarize the simulation output in a heatmap
+indicating the intensity of use across the KAZA-TFCA, and we apply network
+theory to produce a set of complementary maps that further reveal major
+dispersal hotspots and crucial bottlenecks. Finally, we discuss the benefits and
+pitfalls of dispersal simualtions and we highlight improvements to be made in
+the future.
 \end{abstract}
@@
 
+\onehalfspacing
+\tableofcontents
+\doublespacing
+
 % Change page numbering
+\newpage
 \pagenumbering{arabic}
@@
 individual based movement models. Still, only few have explored these
-possibilities \citep{Kanangaraj.2013, Hauenstein.2019, Zeller.2020}.
+possibilities \citep{Kanagaraj.2013, Hauenstein.2019, Zeller.2020}.
 
@@
 % Integrated step selection analysis
-An animals movement trajectory can be seen as the result of an interplay between
-habitat and movement preferences. Traditional step selection functions cannot
-account for a correlation between turning angles and step lengths, unless the
-two are sampled jointly from a copula distribution \citep{Hodel.2021}. Recently,
-however, SSFs have been generalized to \textit{integrated} SSFs (iSSFs), which
-tear apart movement and habitat preferences of the studied animals
-\citep{Avgar.2016}. The method thus allows to render an animals preferences with
-respect to prevailing habitat conditions, as well as with respect to movement
-characteristics. This reduces potential biases in estimated preferences and
-enables parametrization of a fully mechanistic movement model from which
-movement and space use can be simulated \citep{Avgar.2016, Signer.2017}.
-\cite{Signer.2017} used integrated step selection analysis to simulate steady
-state utilization distributions of resident animals. However, the degree to
-which such simulations could be used to simulate dispersers and infer movement
-corridors is unknown.
+An animals movement trajectory can be viewed as the result of an interplay
+between habitat and movement preferences. Traditional step selection functions
+cannot account for a correlation between turning angles and step lengths, unless
+the two are sampled jointly from a copula distribution \citep{Hodel.2021}.
+Recently, however, SSFs have been generalized to \textit{integrated} SSFs
+(iSSFs), which tear apart movement and habitat preferences of the studied
+animals \citep{Avgar.2016}. The method thus allows to render an animals
+preferences with respect to prevailing habitat conditions, as well as with
+respect to movement characteristics. This reduces potential biases in estimated
+preferences and enables parametrization of a fully mechanistic movement model
+from which movement and space use can be simulated \citep{Avgar.2016,
+Signer.2017}. \cite{Signer.2017} used integrated step selection analysis to
+simulate steady state utilization distributions of resident animals. However,
+the degree to which such simulations could be used to simulate dispersers and
+infer movement corridors is unknown.
 
@@
 (\Cref{StudyArea}a) and was represented by a rectangular bounding box that
-stretched over 1.3 Mio. km\textsuperscript{2}. Its extent encompassed the entire
-KAZA-TFCA (\Cref{StudyArea}b), which is the world's largest transboundary
+stretched over 1.3 Mio. km\textsuperscript{2}, ecompassing the entire KAZA-TFCA
+(\Cref{StudyArea}b). The KAZA-TFCA is the world's largest transboundary
 conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
 and Zambia, covering a total of 520'000 km\textsuperscript{2}. Its landscape
-varies regionally and ranges from savanna, to grassland, and from dry to moist
+varies regionally and ranges from savanna, to grassland and from dry to moist
 woodland habitats. A dominant hydrogeographical feature in its center is the
-flood-pulsing Okavango Delta, the earth's largest inland delta. The delta and
-its surroundings are considered a stronghold for African wild dogs and may act
-as a source for the recolonization of surrounding habitats \citep{Cozzi.2013}.
-The wet season within the study area lasts from November to March and is out of
-phase with the main flooding of the Okavango Delta which peaks between July and
-August \citep{McNutt.1996, Wolski.2017}. While large portions of land within the
-KAZA-TFCA are designated protected areas or national parks that remain largely
-free from humans, substantial human influence originates from roads,
-agricultural sites and several small to large settlements.
+flood-pulsing Okavango Delta. The delta and its surroundings are considered a
+stronghold for African wild dogs that may act as a source for the recolonization
+of surrounding habitats \citep{Cozzi.2013}. The wet season within the KAZA-TFCA
+lasts from November to March and is out of phase with the main flooding of the
+Okavango Delta which peaks between July and August \citep{McNutt.1996,
+Wolski.2017}. While large portions of land within the KAZA-TFCA are designated
+national parks or other protected areas, substantial human influence remains due
+to roads, agricultural sites and settlements and villages.
 
@@
     largest terrestrial conservation area and covers a total of 520'000
-    km\textsuperscript{2}. It will connect or reconnect multiple already
-    existing national parks and protected areas (green polygons) that are
-    distributed across its extent. The African wild dog dispersal data
-    considered in this study was collected on free-ranging individuals departing
-    from the Moremi National Park in northern Botswana (since game reserves in
-    Botswana serve the same purpose as national parks, we refer to them as
-    national parks for simplicity).}
+    km\textsuperscript{2}. Its purpose is to re-establish connectivity between
+    already-existing national parks and other protected areas (green polygons).
+    The dispersal data used in this study was collected on a free-ranging
+    African wild dog population inhabiting the Moremi National Park in northern
+    Botswana.}
     \label{StudyArea}
@@
 \subsection{GPS Relocation Data}
-Between 2011 and 2019, we collected GPS relocation data from dispersing wild
-dogs in a free-ranging wild dog population inhabiting the Moremi National Park
-in northern Botswana. We identified potential dispersers based on age, number of
-same‐sex siblings, pack size, and presence of unrelated opposite-sex individuals
-in the pack \citep{McNutt.1996, Behr.2020}. We immobilized individuals using a
-cocktail of Ketamine/Xylazine/Atropine \citep{Osofsky.1996, Cozzi.2020},
-injected with a dart, fired from a CO\textsubscript{2}-pressurized gun
-(\textit{DAN-Inject, Denmark}). Anesthesia protocols were approved by the
-Ministry of Environment, Natural Resources Conservation and Tourism of Botswana
-(permit EWT 8/36/4 XXXVI). After immobilization, individuals were fitted with
-GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
-Berlin}) that included an automated drop-off mechanism. Handling and collaring
-of all individuals was carried out and supervised by a Botswana-registered
-wildlife veterinarian. All of the immobilized individuals usually rejoined their
-pack within one hour after the procedure. Out of all collared individuals, 16
-individuals eventually dispersed in separate same-sex coalitions and their
-trajectories were successfully recorded (7 female and 9 male coalitions).
+Between 2011 and 2019, we collected GPS relocation data on dispersing wild dogs
+from a free-ranging wild dog population inhabiting the Moremi National Park in
+northern Botswana. We selected potential dispersers based on age, pack size,
+number of same‐sex siblings within the pack, and presence of unrelated
+opposite-sex individuals in the pack \citep{McNutt.1996, Behr.2020}. We
+immobilized selected individuals using a cocktail of Ketamine/Xylazine/Atropine
+\citep{Osofsky.1996, Cozzi.2020} that was injected by dart, fired from a
+CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). Anesthesia
+protocols were approved by the Ministry of Environment, Natural Resources
+Conservation and Tourism of Botswana (permit EWT 8/36/4 XXXVI). Immobilized
+individuals were fitted with GPS/Satellite radio collars (\textit{Vertex Lite;
+Vectronic Aerospace GmbH, Berlin}) that guaranteed automated drop-off through a
+decomposable piece of cotton. Handling and collaring of all individuals was
+supervised by a Botswana-registered wildlife veterinarian and all individuals
+rejoined their pack within one hour after immobilization. 16 collared
+individuals eventually dispersed, each in a separate same-sex coalition (7
+female and 9 male coalitions).
 
@@
 system, which allowed remote tracking of individuals, even if they left the main
-study area and crossed international borders. To distinguish between periods of
-residency and dispersal, we applied the net-squared displacement metric to the
-observed movement data. This metric measures the squared Euclidean distance of a
-collared individual to a reference point \citep{Borger.2012}. As a reference
-point, we used the center of each disperser's natal home range, such that
-dispersal was deemed to have started when an individual left its natal home
-range and ended once individuals became sedentary again. For the purpose of this
-study, we discarded all data that was collected during residency and only
-retained movement data during dispersal. Previous research suggests that females
-and males behave similarly during dispersal \citep{Woodroffe.2019, Cozzi.2020},
-so we did not distinguish between sexes in our analyses. Ultimately, we
-converted the collected GPS coordinates to steps, where each step represented
-the straight-line distance traveled by and individual between two consecutive
-GPS relocations \citep{Turchin.1998}.
+study area and crossed international borders. In some instances exact dispersal
+dates were known from field observations. In other cases, we distinguished
+between residency and dispersal using the net-squared displacement metric. This
+metric measures the squared Euclidean distance a GPS relocation to a reference
+point \citep{Borger.2012}, which in our case was set to the center of each
+individual's natal home range. As such, dispersal was deemed to have started
+when an individual left its natal home range and ended once individuals became
+sedentary again. Because behavior during dispersal is more pertinent for
+assessing connectivity, we discarded all data that was collected during
+residency and only retained GPS data recorded during dispersal. As previous
+research revealed similar behavior of females and males during dispersal
+\citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish between sexes. After
+collection, we converted collected GPS coordinates (n = 4'169) to steps, where
+each step represented the straight-line distance traveled by and individual
+between two consecutive GPS relocations \citep{Turchin.1998}. To ensure a
+regular sampling interval, we removed fixes that were not collected on the
+4-hourly schedule.
 
 \subsection{Covariates}
-We represented the physical landscape in the study area using a set of habitat
-covariates rendering water-cover, distance to water, tree-cover, and
+We represented the physical landscape across the study area using a set of
+habitat covariates including water-cover, distance to water, tree-cover, and
 shrub/grassland-cover. Because water cover greatly changes within and between
@@
 allowed us to generate frequently updated water cover layers and corresponding
-distance to water layers (see \citealp{Wolski.2017} and \citealp{Hofmann.2021}).
-Hence, the obtained layers temporally aligned with our dispersal events.
-Furthermore, we computed a proxy for human influence, depicting anthropogenic
-pressures stemming from human-density, agricultural sites, and roads. We
-coarsened or interpolated all layers to a target resolution of 250 m by 250 m
-and cropped them to the extent of the KAZA-TFCA. Further details on the sources
-and preparation of each habitat covariate are given in \cite{Hofmann.2020}.
+distance to water layers for the close surroundings of the Delta (see
+\citealp{Wolski.2017} and \citealp{Hofmann.2021}). Thus, those layers temporally
+aligned with our dispersal events. We also computed a proxy for human influence,
+depicting anthropogenic pressures stemming from human-density, agricultural
+sites, and roads. All obtained layers were coarsened or interpolated to a target
+resolution of 250 m by 250 m and cropped slightly beyond the extent of the
+KAZA-TFCA. Further details on the sources and preparation of each habitat
+covariate are given in \cite{Hofmann.2021}.
 
@@
 coded a binary variable (\textsf{LowActivity}) indicating whether a step was
-realized during low activity (17:00 to 07:00 local time) or high activity (09:00
-to 17:00 local time). We conducted all cleaning and analysis, as well as all
-simulations using the statistical software {\tt R}, version 3.6.6
-\citep{R.2019}. The corresponding {\tt R}-scripts are provided on Github.
+realized during periods of low wild dog activity (17:00 to 07:00 local time) or
+high wild dogactivity (09:00 to 17:00 local time). Handling and manipulation of
+all data, as well as all modelling and simulating was achieved with statistical
+software {\tt R}, version 3.6.6 \citep{R.2019}. {\tt R}-scripts to reproduce the
+analyses are provided on Github.
 
@@
 functions, \textit{integrated} step selection functions allow simultaneous
-inference on movement and habitat preferences of the studied animal, and on
-potential interactions between habitat and movement preferences. In result, the
-method produces less biased selection estimates and allows to treat the
-resulting model as a fully mechanistic movement model based on which movement
-can be simulated. To conduct iSSF analysis, we paired each observed step with 24
-random steps, alltogether forming a stratum that received a unique identifier.
-Each of the 24 random steps served as pseudeo-absence and resembled a potential
-alternative that the animal could have realized but decided not to. We generated
-random steps by sampling random turning angles from a uniform distribution
-(\(-\pi, +\pi\)) and step lengths from a gamma distirbution that was fitted to
-observed steps (scale = 6'308, shape = 0.37). While the number of random steps
-is inversely proportional to the sampling error \citep{Avgar.2016}, a
-preliminary analysis revealed only minor differences in model performance when
-sampling additional random steps, hence we deemed 24 random steps to be
-sufficient. Along each step, we extracted and averaged spatial covariates using
-the {\tt velox} package and we calculated the movement metrics \textsf{sl},
-\textsf{log(sl)}, and \textsf{cos(ta)}. To facilitate model convergence, we
-standardized all continuous covariates to a mean of zero and a standard
-deviation of one. Since correlation among covariates was low (\(|r| > 0.6\);
-\citealp{Latham.2011}), we retained all of them for modeling.
-
-To compare realized steps (scored 1) to random steps (scored 2), we assumed that
-animals assigned a selection score \(w(x)\) to each step. The score depended on
-the step's covariates \(X\), as well as the animals preferences \(\beta\)
-towards these covariates. The probability of a step being realized was then
-contingent on the step's selection score, as well as on the selection scores of
-all other step in the same stratum. To estimate the preferences of interest, we
-fitted conditional logistic regression models in the r-package {\tt glmmTMB}
-following the method proposed by \citep{Muff.2020}. This approach enables to fit
-random slopes in addition to random intercepts. Here, we used animal IDs as
-random effect and modelled random slopes for each covariate. To apply the
-``poisson trick'', we fixed the random intercept variance at an arbitrary high
-value of 10\textsuperscript{6} \citep{Muff.2020}.
-
-The structure of our movement model was based on a habitat selection model for
-dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model, no
-interactions between the habitat and movement covariates were considered, so we
-slightly expanded this base model by proposing interactions between movement and
-habitat covariates. More specifically, we started with the base model and
-incrementally increased model complexity by adding all possible two-way
-interactions between habitat covariates and movement covariates. For instance,
-for the covariate \textsf{water}, we proposed the interactions
-\textsf{Water:log(sl)}, \textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}.
-Besides those combinations, we also proposed the interactions
-\textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)} to account for a correlation
-between turning angles and step lengths, as well as the interactions
-\textsf{sl:LowActivity} and \textsf{log(sl):LowActivity} to account for the fact
-that step lengths may differ due to wild dogs' diurnal activity pattern. To
-compare competing models and assess the most parsimonious movement model, we ran
-stepwise forward model selection based on Akaike's Information Criterion (AIC,
-\citealp{Burnham.2002}) and computed model weights.
-
-We validated the predictive power of the selected movement model using a k-fold
-cross-validation procedure for case-control studies presented in
+inference on movement and habitat preferences of the studied animal, as well as
+on potential interactions between the two. This results in less biased selection
+estimates and allows to treat the resulting model as a fully mechanistic
+movement model based on which movement can be simulated \citep{Avgar.2016}. To
+conduct iSSF analysis, we paired each observed step with 24 random steps.
+Together, an observed and its random steps formed a stratum that received a
+unique identifier. The 24 random steps served as pseudeo-absence and represented
+a potential alternative that the animal could have realized but decided not to.
+We generated random steps by sampling random turning angles from a uniform
+distribution (\(-\pi, +\pi\)) and step lengths from a gamma distirbution that
+was fitted to observed steps (scale = 6'308, shape = 0.37). Although a larger
+number of random steps reduces the sampling error \citep{Avgar.2016},
+preliminary analyses revealed that 24 steps were sufficient in our case. Along
+each step, we extracted and averaged spatial covariates using the {\tt velox}
+package and we calculated the movement metrics \textsf{sl}, \textsf{log(sl)},
+and \textsf{cos(ta)}. To facilitate model convergence, we standardized all
+continuous covariates to a mean of zero and a standard deviation of one. Since
+correlation among covariates was low (\(|r| > 0.6\); \citealp{Latham.2011}), we
+retained all of them for modeling.
+
+To contrast realized steps (scored 1) and random steps (scored 2) and fit a
+movement model, we assumed that animals assigned a selection score \(w(x)\) to
+each step. The score depended on the step's covariates \(X\), as well as the
+animals preferences \(\beta\) towards these covariates. The probability of a
+step being realized was then contingent on the step's selection score, as well
+as on the selection scores of all other step in the same stratum. To estimate
+the preferences of interest, we ran conditional logistic regression models in
+the r-package {\tt glmmTMB}. More specifically, we applied the method developed
+by \citep{Muff.2020}, which allows to fit random slopes. Here, we treated animal
+IDs as random effect and modelled random slopes for each covariate and we fixed
+the random intercept variance at an arbitrary high value of
+10\textsuperscript{6} to make use of the  ``poission''-trick  \citep{Muff.2020}.
+
+The movement model was based on a habitat selection model for dispersing wild
+dogs presented in \cite{Hofmann.2021}. In the original model, no interactions
+between the habitat and movement covariates were considered, so we slightly
+expanded this base model by proposing interactions between movement and habitat
+covariates. More specifically, we started with the base model and incrementally
+increased model complexity by adding all possible two-way interactions between
+habitat covariates and movement covariates. For instance, for the covariate
+\textsf{water}, we proposed the interactions \textsf{Water:log(sl)},
+\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides those combinations,
+we also proposed the interactions \textsf{sl:cos(ta)} and
+\textsf{log(sl):cos(ta)} to account for a correlation between turning angles and
+step lengths, as well as the interactions \textsf{sl:LowActivity} and
+\textsf{log(sl):LowActivity} to account for the fact that step lengths may
+differ due to wild dogs' diurnal activity pattern. To compare competing models
+and assess the most parsimonious movement model, we ran stepwise forward model
+selection based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
+
+We validated the predictive power of the most parsimonious movement model using
+k-fold cross-validation for case-control studies as presented
 \cite{Fortin.2009}. We randomly split the input data into training and testing
@@
 \subsection{Dispersal Simulation}
-We used the most parsimonious movement model to simulate a total of 80'000
-dispersing wild dogs moving across the KAZA-TFCA. The simulation resembled an
-inverted iSSF and was set up as follows. (1) We sampled a source point at which
-a disperser was initiated. We assumed a random initial orientation of the animal
-in order to permit calculation of relative turning angles. (2) We generated a
-set of 25 random steps by sampling turning angles from a uniform distribution
-(\(-\pi, +\pi\)) and step lengths from the fitted gamma distribution. In line
-with our training data, each step resembled the straight line movement realized
-within 4 hours. To prevent the sampling of unreasonably large steps, we capped
-sampled step lengths to a maximum of 35 km, which corresponds to the farthest
-distance ever traveled within 4 hours by our dispersers. (3) Along each random
-step, we extracted habitat covariates and calculated movement covariates. (4) We
-applied the parameterized movement model to predict the selection score \(w(x)\)
-for each of the proposed steps and translated selection scores into
-probabilities by applying \Cref{EQ2}. (5) We sampled one of the random steps
-based on predicted probabilities and used it to determine the animal's new
-position. We then repeated steps (2) to (5) until 2'000 steps were realized.
-
-Because some simulated dispersers would inevitably approach a map boundary, some
-of the proposed random steps may leave the study area, such that covariates
-could be extracted and no selection score could be calculated. To mitigate the
-influence of such edge effects, we followed \citep{Koen.2010} and artificially
-expanded all covariate layers by adding a 100 km buffer zone within which we
-randomized covariate values by resampling values from the observed covariate
-layers. This allowed simulated dispersers to leave and reenter the study area
-through this buffer zone. Nevertheless, if a disperser approached this outer
-boundary and one of the proposed random steps left the buffer, we resampled
-transgressing steps until they fully lied within the buffer zone, thereby
-forcing dispersers to bounce off the outer boundary. The simulation was mainly
-coded in the programming language {\tt R}, yet a few helper functions were
-written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
-\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
+We used the most parsimonious movement model to simulate 80'000 dispersing wild
+dogs moving across the KAZA-TFCA. The simulation resembled an inverted iSSF and
+worked as follows. (1) We defined a random source point and assumed a random
+initial orientation of the animal. (2) Departing from the source point, we
+generated 25 random steps by sampling relative turning angles from a uniform
+distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
+distribution. As such, each step corresponded to the straight line movement
+conducted within 4 hours. To prevent unreasonably large steps, we capped sampled
+values to a maximum of 35 km, which corresponds to the farthest distance ever
+traveled within 4 hours by on of our dispersers. (3) Along each random step, we
+extracted and averaged habitat covariates and calculated movement covariates.
+Extracted values were again standardized using the same transformations that
+were used on the input data. (4) We applied the parameterized movement model and
+predicted the selection score \(w(x)\) for each of the proposed steps and
+translated selection scores into probabilities by applying \Cref{EQ2}. (5) We
+randomly sampled one of the random steps based on predicted probabilities and
+used it to determine the animal's new position. We repeated steps (2) to (5)
+until a total of 2'000 steps were realized.
+
+To minimize the influence of edge effects and to deal with random steps leaving
+the study area, we followed \citep{Koen.2010} and artificially expanded all
+covariate layers by adding a 100 km buffer zone. Within the buffer zone, we
+randomized covariate values by resampling values from the original covariate
+layers. In result, simulated dispersers were able to leave and reenter the main
+study area through this buffer zone. In cases where a disperser approached the
+outer map boundary such that one of the proposed random steps transgressed its
+border, we resampled transgressing steps until they fully lied within the buffer
+zone. Although the algorithm for the simulation was mainly coded in {\tt R}, we
+wrote a few helper functions in {\tt C++} and imported them into {\tt R} using
+the {\tt Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}.
 
 \subsection{Source Points}
-To initiate virtual dispersers at biologically meaningful source points, we
-randomly placed source points within contiguous protected areas larger \(>\) 700
-km\textsuperscript{2} (\Cref{SourcePoints} (a)). This conforms to the average
-home range requirement of resident wild dogs \citep{Pomilia.2015} and allowed us
-to remove areas too small to host viable wild dog populations. By distributing
-source points randomly, the number of source points per km\textsuperscript{2}
-within protected areas was approximately equal across our study area. Across all
-protected areas, we distributed 50'000 source points, each representing the
-starting point of a single dispersal trajectory. To render potential immigrants
-into the study system, we placed additional 30'000 source points inside the
-buffer zone around the main study area (\Cref{SourcePoints} (b)).
+We initiated virtual dispersers at 50'000 randomly selected source points within
+contiguous protected areas larger \(>\) 700 km\textsuperscript{2}
+(\Cref{SourcePoints} (a)). This conforms to the average home range requirement
+of resident wild dogs \citep{Pomilia.2015} and allowed us to remove areas too
+small to host viable wild dog populations. By distributing source points
+randomly, the number of source points per km\textsuperscript{2} within protected
+areas was approximately equal across the study area. To render potential
+immigrants into the study system, we placed additional 30'000 source points
+inside the buffer zone around the main study area (\Cref{SourcePoints} (b)).
+This resulted in a total of 80'000 source points, each representing the start
+point of an individual disperser.
 
@@
     \includegraphics[width = \textwidth]{99_SourceAreas.png}
-    \caption{(a) Illustration of an artificially expanded covariate layer (water
-    cover). We expanded the layer by adding a 100 km buffer (light blue) which
-    we filled with values sampled from the original layer. (b) Different source
-    areas from which we released virtual dispersers. We initiated 50'000
-    dispersers within the main study area (green) and another 30'000 dispersers
-    within a virtual buffer (blue).}
+    \caption{(a) Different source areas from which we released virtual
+    dispersers. We only considered contiguous protected areas (national parks,
+    game reserves, and forest reserves) that were larger than 700
+    km\textsuperscript{2} (green). This area corresponds to the average home
+    range requirement for viable wild dog populations \citep{Pomilia.2015}. To
+    render potential immigrants into the study system, we also initiated
+    dispersers within a buffer zone (blue) surrounding the main study area. (b)
+    Source points from which dipsersers were released. 50'000 dispersers were
+    released from the main study area (green dots) and another 30'000 dispersers
+    within the virtual buffer (blue dots).}
     \label{SourcePoints}
@@
 
-\subsection{Heatmaps}
-To identify dispersal hotspots across our study area, we rasterized all
-simulated trajectories and created a heatmap indicating the absolute frequency
-at which each raster-cell in the study area was visited by our virtual
-dispersers. If the same trajectory crossed a raster-cell twice, it was only
-counted once. That is, we did not consider revisits, thereby mitigating biases
-arising from individuals that were trapped in a certain locations and moved
-around in circles. We achieved high performance rasterization of all simulated
-trajectories using the R-package {\tt terra} \citep{Hijmans.2020}. While one
-could easily subset the data to visualize heatmaps assuming different dispersal
-durations, we retained all 2'000 simulated steps to create a single heatmap.
-
-% To examine if and how ``heat'' changes in response to changes in the location of
-% source points and the number of simulated steps, we followed a 2 x 6 design and
-% created heatmaps for both point sampling regines, as well as for 68, 125, 250,
-% 500, 1000, and 2000 dispersal steps. We quantified the similarity of the
-% resulting 12 heatmaps to the permeability and least-cost corridor maps presented
-% in (Hofmann ...) we used Bhattacharyya's affinity. Bhattacharyya's affinity
-% ranges from zero (complete separation) to one (perfect match) and has earlier
-% been proposed to compare the overlap of utilisation distributions (Fieberg).
-
-\subsection{Network Analysis I}
-To pinpoint areas that are of high relevance for connecting remote regions in
-our study area, we adopted a network view on the simulated trajecotries and
-computed the betweenness metric. More specifically, we overlaid the study area
-(including the buffer) with a with a raster that had a cell size of 5 x 5 km,
-where each raster-cell represented a node in the final network. We then used the
-simulated trajectories to determine all transitions occurring from one cell to
-another, as well as the frequency at which these transitions occurred. This
-resulted in an edge-list that we translated into a weighted network using the
-r-package {\tt igraph}. Because {\tt igraph} handles the weights \(\omega\) as
-costs, we inverted the traversal frequency at each cell by applying \(\omega =
-\frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal Frequency_i}\). Finally, we
-calculated the betweenness metric for each node in the final network. This
-metric indicates how often a specific raster-cell lies on a shortest path
-between two other raster-cells and is therefore a useful metric to detect
-movement corridors \citep{BastilleRousseau.2018}.
-
-\subsection{Network Analysis II}
-Besides this, we also adopted an alternative network view and identified the
-frequency at which dispersers originating from one national park successfully
-moved into an other national park. We achieved this by looking at each pairwise
-combination of national parks and determining the relative number of
-trajectories that originated in one national park and had at least one
-subsequent coordinate laying in the other national park. This allowed us to
-determine the average duration it took a simulated disperser to move from one
-national park to another, and to point out combinations of national parks
-between no links were observed.
+\subsection{Heatmap}
+To identify dispersal hotspots across our study area using the simulated
+dispersal events, we rasterized all simulated trajectories and created a heatmap
+indicating the absolute frequency at which each raster-cell in the study area
+was visited by our virtual dispersers. If the same trajectory crossed a
+raster-cell twice, it was only counted once, meaning that we did not consider
+revisits. This mitigated biases arising due to trapped individuals that were
+moving in circles. To achieve high performance rasterization of all simulated
+trajectories, we used the R-package {\tt terra} \citep{Hijmans.2020}.
+
+\subsection{Network Analysis I: Betweenness}
+To pinpoint areas of exceptional relevance for connecting remote regions in our
+study area, we converted simulated trajectories into a network and calculated
+betweenness scores. More specifically, we overlaid the study area (including the
+buffer) with a regular raster that was resolved at 5 x 5 km and where each
+raster-cell represented a node in the final network. We then used the simulated
+trajectories to determine all transitions occurring from one cell to another, as
+well as their frequency. This resulted in an edge-list that we translated into a
+weighted network using the r-package {\tt igraph}. Because {\tt igraph} handles
+the weights \(\omega\) as costs, we inverted the traversal frequency at each
+cell by applying \(\omega = \frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal
+Frequency_i}\). Finally, we calculated the betweenness metric for each node in
+the final network. This metric indicates how often a specific raster-cell lies
+on a shortest path between two other raster-cells and is therefore a useful
+metric to detect movement corridors \citep{BastilleRousseau.2018}.
+
+\subsection{Network Analysis II: Interpatch-Connectivity}
+We also adopted an alternative network view and assessed interpatch connectivity
+between national parks. The decision to focus on national parks was purely out
+of simplicity and could easily be expanded to any other patch of interest. To
+determine interpatch-connectivity, we identified the relative frequency at which
+dispersers originating from one national park successfully moved into an other
+national park. This allowed us to determine \textit{if} and \textit{how often}
+dispersers moved between certain national parks, as well as to estimate the
+average number of steps that was necessary for those movements.
 

### 2021-05-05 Chapter_1/04_Manuscript/99_MovementModelAICs.tex
Various updates

--- 2020-08-21 Chapter_1/04_Manuscript/99_MovementModelAICs.tex
@@
 % latex table generated in R 3.6.3 by xtable 1.8-4 package
-% Thu Jul  9 09:17:26 2020
+% Tue May  4 15:07:44 2021
 \begin{tabular}{lrrrr}
@@
   \midrule
-cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
-  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
+Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) & 89392.88 & 0.00 & 0.15 & -44670.44 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:log(sl) & 89393.92 & 1.04 & 0.09 & -44669.96 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + DTW:log(sl) & 89394.13 & 1.25 & 0.08 & -44670.06 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:log(sl) & 89394.25 & 1.37 & 0.08 & -44670.13 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl & 89394.36 & 1.48 & 0.07 & -44672.18 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + log(sl):LA & 89394.44 & 1.56 & 0.07 & -44670.22 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:sl & 89394.56 & 1.68 & 0.07 & -44670.28 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:log(sl) & 89394.57 & 1.69 & 0.07 & -44670.29 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:cos(ta) & 89394.59 & 1.71 & 0.07 & -44670.30 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:cos(ta) & 89394.63 & 1.75 & 0.06 & -44670.31 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + sl:cos(ta) & 89394.68 & 1.80 & 0.06 & -44672.34 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:log(sl) & 89394.69 & 1.81 & 0.06 & -44670.35 \\ 
+  Base Model + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:cos(ta) & 89394.84 & 1.96 & 0.06 & -44670.42 \\ 
    \bottomrule

### 2021-05-05 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

@@
 
-\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
+\textsuperscript{2} Botswana Predator Conservation, Private Bag 13, Maun,
 Botswana.
@@
 
-\textbf{Keywords:} dispersal, simulation, integrated step selection,
+\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
 Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
@@
 \begin{abstract}
-For many species, dispersal represents a crucial process that allows to avoid
-inbreeding, to colonize new habitats and to reinforce small and unviable
-subpopulations. Successful dispersal is therefore a crucial pre-requisite for
-long-term species persistence in wild animal populations. However, the ability
-to disperse is contingent a sufficient degree of landscape connectivity, which
-is why the estimation of connectivity and identification of dispersal corridors
-has become a task of extraordinary importance in ecological studies and for the
-planning of conservation initiatives.
-
-Over the past two decades, scientists have mainly relied on analytical tools
+Dispersal is an important process that allows species to avoid inbreeding, to
+colonize new habitats and to reinforce non-viable subpopulations. Successful
+dispersal thus represents a crucial pre-requisite for long-term species
+persistence in wild animal populations. However, the ability to disperse is
+contingent a sufficient degree of landscape connectivity, which is why the
+estimation of connectivity and identification of dispersal corridors has become
+a task of extraordinary importance in ecological studies.
+
+Over the past two decades, ecologists have primarily relied on analytical tools
 such as least-cost methods and circuit theory to investigate landscape
-connectivity. This is largely owed to their ease of use and intuitive nature,
-which have facilitated their application in the past. Despite their usefulness
-for various ecological applications, both methods make several restricting
-assumptions that can limit their applicability in reality. Least-cost analysis,
-for instance, suffers from the fact that animals are assumed to have perfect
-knowledge about the entire landscape and choose cost-minimizing routes
-accordingly. On the contrary, circuit theory assumes that animals walk according
-to a complete random walk, which is equally unlikely to hold in reality.
-
-Fortunately, recent innovations in movement ecology have brought forward novel
-opportunities to study dispersal and landscape connectivity. In particular, the
-rich suite of resource selection functions has undergone substantial
-improvements in recent years. Most notably, step selection functions have been
-generalized to \textit{integrated} step selection functions, which essentially
-represent fully mechanistic movement models based on which individual animal
-movement could be predicted and simulated. While such models have been applied
-to simualte the steady state utilization distribution, similar approaches could
-be useful for investigating transient movement behavior and study landscape
-connectivity in a much more natural way.
+connectivity. Despite their usefulness for various ecological applications, both
+methods make several restricting assumptions that limit their suitability in
+reality. To address their shortcomings, simulations from individual-based
+movement models have been proposed and applied. To overcome their shortcomings,
+simulations from individual-based movement models have been proposed. Yet, due
+to the infinite number of possibilities of coming up with an individual-based
+model, a unified and objective framework is missing.
+
+Recent innovations in movement ecology have brought forward novel opportunities
+to study animal dispersal and landscape connectivity. In particular, the rich
+suite of resource selection functions, including point-, step-, and
+path-selection functions have undergone substantial improvements over the past
+years. Most notably, step-selection functions have been generalized to
+\textit{integrated} step selection functions, which essentially represent fully
+mechanistic movement models based on which individual animal movement could be
+simulated. While such models have been applied to study \textit{steady-state}
+utilization distribution, a similar approach may be useful for investigating
+\textit{transient} movement behavior and thereby study landscape connectivity by
+means of simulated dispersal events.
 
 Here, we showcase the use of integrated step selection analysis to simulate
-dispersal of a top-predator across a large study area. For this, we collected
-data on 16 dispersing wild dogs and analysed it in an integrated step selection
-analysis to produce a fully mechanistic movement model rendering wild dog
-dispersal. Based on this model, we simulated 80'000 dispersers moving across the
-Kavango-Zambezi Transfrontier Conservation Area, the world's largest
-transboundary conservation area. We summarize the simulation output in a heatmap
-indicating the intensity of use across the KAZA-TFCA, and we apply network
-theory to produce a set of complementary maps that further reveal major
-dispersal hotspots and crucial bottlenecks. Finally, we discuss the benefits and
-pitfalls of dispersal simualtions and we highlight improvements to be made in
-the future.
+dispersal of the highly endangered African wild dog across the world's largest
+transboundary conservation area, the Kavango-Zambezi Transfrontier Conservation
+Area (KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
+coalitions departing from northern Botswana. We combine the movement data with
+relevant spatial covariate layers and used integrated step selection functions
+to parametrize a fully mechanistic movement model rendering wild dog dispersal.
+Based on this model, we simulate 80'000 dispersers, originating from protected
+areas and moving across the extent of the KAZA-TFCA. We then generate a heatmap
+and use network theory to reveal dispersal hotspots and crucial bottlenecks
+across the study area. Finally, we discuss the benefits and pitfalls of such
+dispersal simualtions and highlight potential improvements to be made in the
+future.
 \end{abstract}
@@
 
-% % Create linenumbers
-% \linenumbers
-
-\section{Introduction}
+% Create linenumbers
+\linenumbers
+
+\section{Introduction (90\%)}
+
 % Importance of Dispersal & Connectivity
-Dispersal of individuals is an important process governing the dynamics wild
-animal populations that are distributed in space \citep{Clobert.2012}. It is
-defined as the movement of individuals from their natal location to the site of
-first reproduction \cite{Howard.1960} and allows species to avoid inbreeding
-\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012}, to rescue small and
-unviable populations \citep{Brown.1977}, and to promote the colonization of
-unoccupied habitats \citep{Hanski.1998, MacArthur.2001}. Successfull dispersal
-requires a sufficient degree of landscape connectivity, which is why the
-protection of dispersal corridors has become a major goal in conservation
-science. Information on movement behavior during dispersal and knowledge about
-the factors that limit dispersal is therefore critical for a comprehensive
-understanding of landscape connectivity and population viability
-\citep{Baguette.2013, Vasudev.2015}. In addition, reliable modelling techniques
-to identify dispersal corridors based on empirical data are necessary
-\citep{Diniz.2019}.
-
-% Advancements in GPS Technology and Movement Ecology
+\subsection{Importance of Dispersal \& Connectivity (90\%)}
+Dispersal is defined as the movement of individuals from their natal location to
+the site of first reproduction \cite{Howard.1960}. It is a vital process
+governing the dynamics wild animal populations that are distributed in space
+\citep{Hanski.1998, Clobert.2012} and may strongly affect population dynamics at
+different spatial and social scales \citep{Hanski.1999a, Clobert.2012}.
+Dispersal allows species to avoid inbreeding and maintain genetic diversity
+\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013}, to
+rescue small and non-viable populations \citep{Brown.1977}, and to promote the
+colonization of unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}.
+However, successful dispersal requires a sufficient degree of functional
+connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the identification
+and protection of major dispersal corridors has become an important task in
+conservation science \citep{Doerr.2011, Rudnick.2012}. To achieve this, reliable
+information on movement behavior during dispersal and knowledge about factors
+that limit dispersal and connectivity is paramount \citep{Baguette.2013,
+Vasudev.2015}.
+
+% Advancements in GPS Technology & Movement Ecology
+\subsection{Advancements in GPS Technology \& Movement Ecology (90\%)}
 Thanks to novel technologies developed over the past decades, particularly of
-GPS/Satellite radio-collars, the study of dispersal and connectivity using
-telemetry data has accelerated \citep{Jonsson.2016, Williams.2019}.
+GPS/Satellite radio-collars, the use of GPS data to study animal dispersal and
+connectivity has accelerated \citep{Elliot.2014, Jonsson.2016, Williams.2019}.
 Additionally, the advent of publicly accessible satellite imagery and
 sophisticated remote sensing techniques to represent the physical landscape
-through which individuals disperse have heralded the ``golden age of animal
-tracking'' \citep{Kays.2015}. Concurrently, the increased availability of large
-amounts of empirical data and an increased computational power have led to the
-development of several modelling techniques that allow studying movement
-behavior during dispersal and highlighting major movement corridors
-\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012, Diniz.2019}.
+through which individuals disperse have heralded a ``golden age of animal
+tracking'' \citep{Kays.2015}. Concurrently, the availability of large amounts of
+empirical data and an increased computational power have led to the development
+of numerous techniques to study dispersal movements and highlight critical
+corridors between subpopulations \citep{Boyce.2002, Fortin.2005, Cushman.2010,
+Zeller.2012, Diniz.2020}.
 
 % Resource Selection & Connectivity
+\subsection{Resource Selection \& Connectivity (90\%)}
 \textit{Resource selection functions} \citep{Boyce.2002} and derived methods
@@
 selection functions} \citep{Cushman.2010} have proven particularly useful for
-studying animal movement \citep{Fieberg.2020}. These methods allow estimating
-habitat preferences of the focal species by comparing covariates at locations
-visited by the animal to the scame covariates at locations available to, but not
-visited by the animal. The so estimated preferences can then be used to predict
-a permeability surface, indicating the expected ease at which an animal can
-traverse a given area \citep{Zeller.2012}. Utimately, the permeability surface
-serves as input to a connectivity model that is used to reveal movement
-corridors. In this regard, two of the most prominent connectivity models in the
-literature are least-cost path analysis \citep{Adriaensen.2003} and circuit
-theory \citep{McRae.2006, McRae.2008}.
-
-% Least-cost path
-In least-cost path analysis, connectivity is estimated by identifying
-least-costly routes between pre-defined start- and endpoints
-\citep{Adriaensen.2003}. For this, the permeability surface is converted into a
-network graph and permeability scores are translated into probabilities for
-moving from one cell to another. Least-costly routes between start- and
-endpoints can then be computed using Dijkstra’s algorithm \citep{Dijkstra.1959}.
-Although intuitively appealing, the method suffers from several weaknesses.
-First of all, a least-cost path always exist, even if the associated costs are
-unreasonably high. Second, it assumes that animals have a preconceived end-point
-in mind and choose a cost-minimizing route accordingly. While this assumption
-may be reasonable for migrating animals that move between a limited number of
-habitats, it is unlikely to hold for dispersers that move over long distances
-into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Finally, the
-method also requires that animals have an infinite perceptual range, otherwise
-they could not compute an optimal path. Some of these issues have been addressed
-using alternative least-cost approaches. For instance, instead of computing a
-least-cost path, a least cost-corridor that also considers slightly suboptimal
-routes \citep{Pinto.2009} can be calculated. The corridor relaxes the assumption
-of perfect knowledge and overcomes the single pixel width issue. Alternatively,
-the randomized least-cost path has been proposed, which allows animals to
-deviate from the least-cost route by a pre-defined factor \(\theta\)
-\citep{Panzacchi.2016, VanMoorter.2021}.
-
-% Circuit theory
-In contrast to least-cost approaches, circuit theory makes use of electrical
-circuit theory to quantify landscape connectivity \citep{McRae.2006,
-McRae.2008}. Here, the permeability surface is converted into an electrical
-circuit containing nodes and resistor, where resistance values depend on habitat
-permeability. A source point is connected to a current source, whereas an
-endpoint is grounded. Hence, the current flowing through each node in the
-circuit can be computed and is supposed to resemble the relative frequency at
-which each location in the study area is being visited. By revealing not only a
-single path or corridor, circuit theory overcomes the single-pixel width issue
-inherent to least-cost approaches. Intrinsically, circuit theory is based on the
-assumption that animals follow a random walk and can therefore not render
-directional biases. In reality, however, dispersers often move in a very
-directional fashion as they want to cover as much ground in as little time as
-possible. Furthermore, it is assumed that individuals have a perceptual range of
-a single pixel, which rarely captures the true spatial scale of selection (cite
-someone).
-
-% Temporal scale
-Neither least-cost analyes or circuit theory are capable of rendering the
-temporal dimension of movement. Such influences could, however, be studied using
-individual based movement models. Still, only few have explored these
-possibilities \citep{Kanagaraj.2013, Hauenstein.2019, Zeller.2020}.
-
-% Step selection analysis
-While step selection functions were initially intended to learn about habitat
-preferences of the focal species \citep{Fortin.2005}, recent improvements may
-elevate its applicability into other areas. Originally, \cite{Fortin.2005}
-proposed step selection functions to account for serial autocorrelation inherent
-to GPS relocation data. In this method, covariates along realized steps (the
-connecting lines between two consecutive relocations; ; \citealp{Turchin.1998})
-are contrasted to covariates along  ``alternative''  or  ``random steps'' that
-are generated by randomly sampling turning angles and step lengths based on
-observed movement characteristics. It is then assumed that animals assign a
-selection score \(w(x)\) of the following form to each step:
-
-\begin{equation}
-\label{EQ1}
-  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
-\end{equation}
-
-The selection score \(w(x)\) of each step depends on its associated covariates
-(\(x_1, x_2, ..., x_n\)) and on the animal's relative selection strengths (i.e.
-preferences) towards these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
-probability that a step \(i\) is realized \(P(Y_{i} = 1\)) then depends on the
-step's selection score, as well as on the selection scores of all alternative
-steps:
-
-\begin{equation}
-\label{EQ2}
-  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
-  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
-\end{equation}
-
-\noindent Preferences of interest, i.e. the \(\beta\)'s, can be estimated by
-comparing realized (scored 1) and random (scored 0) steps in a conditional
-logistic regression model \citep{Fortin.2005}. In this model, positive
-\(\beta\)-coefficients indicate selection of a covariate, whereas negative
-\(\beta\)-coefficients indicate avoidance of a covariate. To deal with multiple
-individuals, one can either average estimates of individual models
-\citep{Murtaugh.2007, Fieberg.2010} or apply mixed effects conditional logistic
-regression analysis as recently proposed by \cite{Muff.2020}.
-
-% Integrated step selection analysis
-An animals movement trajectory can be viewed as the result of an interplay
-between habitat and movement preferences. Traditional step selection functions
-cannot account for a correlation between turning angles and step lengths, unless
-the two are sampled jointly from a copula distribution \citep{Hodel.2021}.
-Recently, however, SSFs have been generalized to \textit{integrated} SSFs
-(iSSFs), which tear apart movement and habitat preferences of the studied
-animals \citep{Avgar.2016}. The method thus allows to render an animals
-preferences with respect to prevailing habitat conditions, as well as with
-respect to movement characteristics. This reduces potential biases in estimated
-preferences and enables parametrization of a fully mechanistic movement model
-from which movement and space use can be simulated \citep{Avgar.2016,
-Signer.2017}. \cite{Signer.2017} used integrated step selection analysis to
+studying animal movement \citep{Fieberg.2020} and modelling connectivity
+\citep{Diniz.2020}. These methods allow estimating habitat preferences of the
+focal species by comparing covariates at locations visited by the animal to the
+same covariates at locations available to, but not visited by the animal
+\citep{Boyce.2002, Fortin.2005, Cushman.2010, Thurfjell.2014}. The so estimated
+preferences can then be used to predict a permeability surface, indicating the
+expected ease at which an animal can traverse a given area \citep{Spear.2010,
+Zeller.2012, Etherington.2016}. Ultimately, the permeability surface serves as
+input to a connectivity model that is used to reveal movement corridors
+\citep{Diniz.2020}. Two of the most prominent connectivity models are least-cost
+path analysis (LCP analysis; \citealp{Adriaensen.2003}) and circuit theory (CT
+\citealp{McRae.2006, McRae.2008}), both graph-based methods that estimate
+conductance of the landscape. Despite their intuitive nature and ease of use,
+both methods make rigorous assumptions about animal movement that may or may not
+be fulfilled in reality \citep{Diniz.2020}.
+
+% Issues with Least-Cost Paths & Circuit Theory
+\subsection{Issues with Least-Cost Paths \& Circuit Theory (90\%)}
+In LCP analysis, for instance, a least costly path always exists, even if
+associated movement costs are unreasonably high and will never be incurred by a
+dispersing individual. The method also presumes that animals have an infinite
+perceptual range, a preconceived end-point in mind, and choose a cost-minimizing
+route accordingly. These assumptions may be reasonable for migrating animals,
+yet they are unlikely to be true for dispersers that move over long distances
+into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Finally,
+LCPs are only one pixel wide, meaning that their absolute size depends on the
+resolution of chosen covariate layers \citep{Diniz.2020}. Although some of these
+issues can be alleviated using less stringent versions of the LCP algorithm
+(e.g. least-cost \textit{corridors} \citep{Pinto.2009}, \textit{thresholded}
+least-cost paths \citep{Landguth.2012}, and \textit{randomized} least-cost paths
+\citep{Panzacchi.2016, VanMoorter.2021}), a certain degree of arbitrariness in
+the assumptions remains. CT entails similar restrictions. Because CT only
+considers movements from the source cell to its 4 or 8 adjacent cells, it
+implicitly posits a perceptual range of a single pixel. It is also built on the
+assumption of a complete random walk \citep{Diniz.2020}, implying that
+directional biases cannot be rendered, albeit being very common in dispersal
+movements \citep{Cozzi.2020, Hofmann.2021}. Ultimately, both LCP analysis and CT
+omit the temporal dimension of dispersal. In result, statements about the
+expected duration required to traverse a certain corridor are impossible.
+Because movement is not modelled explicitely, neither of the methods can render
+possible interactions between movement behavior and landscape characteristics
+such that connectivity mainly arises as a result of the landscape structure.
+This is referred to as structural connectivity and stands in contrast to
+functional connectivity, which also renders the behavioral response of the
+animal with respect to prevailing habitat conditions \citep{Tischendorf.2000}.
+Even though functional connectivity is more difficult to estimate, a functional
+view is the ultimate goal in conservation science because it has direct
+consequences for gene flow \citep{Baguette.2013}.
+
+% What about IBMMs?
+\subsection{What about IBMMs? (90\%)}
+To address the issues inherent to LCPs and CT, individual-based movement models
+(IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
+dispersal is simulated explicitly, based on movement rules that determine how
+individuals move over and interact with the prevailing landscape
+\citep{Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020,
+Vasudev.2021}. Using the simulated trajectories, one can calculate a set of
+connectivity metrics, such as interpatch-connectivity and traversal frequency
+across the landscape to reveal major dispersal hotspots \citep{Kanagaraj.2013,
+BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}. Even though IBMMS can be
+employed to overcome any of the shortcomings intrinsic to LCPs and CT, as well
+as to provide a more functional view on connectivity, they can be challenging to
+fit and require vast amounts of data collected during dispersal
+\citep{Diniz.2020}.
+
+% Step Selection Analysis
+\subsection{Step Selection Analysis (90\%)}
+Here, we investigate the usefulness of integrated step selection functions
+(ISSFs, \citealp{Avgar.2016}), as a relatively simple but powerful IBMM based on
+which dispersal can be simulated. While regular SSFs were intended to learn
+about relative habitat preferences of the focal species \citep{Fortin.2005,
+Thurfjell.2014, Avgar.2017}, the method has been generalized and now enables to
+jointly study habitat and movement preferences, as well as potential
+interactions between movement and habitat preferences \citep{Avgar.2016,
+Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple method
+to model complex movement behavior where movement results from two intertwined
+behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}). Importantly, a
+parametrized ISSF model can be viewed as a fully mechanistic movement model
+based on which individual movement trajectories can be simulated
+\citep{Avgar.2016, Signer.2017}. In fact, \cite{Signer.2017} used ISSF to
 simulate steady state utilization distributions of resident animals. However,
-the degree to which such simulations could be used to simulate dispersers and
-infer movement corridors is unknown.
+the degree to which such simulations are helpful in detecting movement corridors
+and modeling landscape connectivity is unknown.
 
 % Study Species & Study Area
-One of the species for which long-term species persistence relies on sufficient
+\subsection{Study Species \& Study Area (90\%)}
+One of the species for which long-term viability relies on sufficient
 landscape connectivity is the endangered African wild dog \textit{Lycon pictus}.
-This species has once been wide-spread across sub-Saharan Africa but disappeared
-from a vast majority of its historic range due to ongoing persecution by humans,
+While once present across entire sub-Saharan Africa, wild dogs have disappeared
+from a vast majority of their historic range due to persecution by humans,
 habitat destruction, and deadly diseases. As of today, only 6'000 free-ranging
-individuals remain in small and spatially scattered subpopulations. Within those
-subpopulations, wild dogs form cohesive packs comprising 8 to 12 adults and
-their offspring \cite{McNutt.1995}. After reaching sexual maturity, male and
-female offspring form same-sex coalitions and disperse from their natal pack
-\citep{McNutt.1996, Behr.2020}. New packs are formed when dispersing coalitions
-join unrelated opposite-sex dispersing coalitions \citep{McNutt.1996}.
-Dispersing wild dogs can cover several hundred kilometers across a variety of
-landscapes \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
-One of the few strongholds for this species lies near the Moremi Game Reserve in
-northern Botswana, which is part of the world's largest transboundary protected
-area, namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
-This area has originally been intended to facilitate migration of elephants, but
-is expected to benefit a multitude of other species \citep{Elliot.2014,
-Brennan.2020, Hofmann.2021}.
-
-In a previous paper, we assessed landscape connectivity within the KAZA-TFCA for
-dispersing African wild dogs using a least-cost corridor approach. For this, we
-fitted a basic habitat selection model based on which we predicted landscape
-permeability. We now expand on this knowledge and develop a more detailed
-movement model of dispersing wild dogs. We then use this model to simulate
-thousands of dispersers moving across the KAZA-TFCA. Based on said simulations,
-we compute heatmaps and identify potential dispersal hotspots and compare them
-to the dispersal routes identified in \citep{Hofmann.2021}. We also showcase how
-simulated dispersal data can be analysed using network-analysis and how network
-metrics relevant to landscape connectivity can be computed. Our results show
-that a simulation based approach yields several major benefits over traditional
-connectivity modelling techniques. Most importantly, simulations provide a more
-generic view on how connectivity emerges and to which degree connectivity
-depends on the dispersal duration. In addtion, by generating proper dispersal
-trajectories, network theory can be applied to calculate network metrics that
-are pertinent to connectivity analysis. Finally, we put forward additional
-opportunities using simulations that go beyond the scope of this paper.
-
-Reliable identification of dispersal corridors will become increasingly
-important with the uprise of ever-growing and often transboundary conservation
-areas. One such instance is the KAZA-TFCA, a massive conservation area spanning
-five countries and over 520'000 km\textsuperscript{2}. The KAZA holds the
-potential of re-establishing dispersal routes for many of its protected species,
-including the african wild dog \textit{Lycaon pictus}.  Persecution by humans,
-habitat loss, and reduced connectivity are major causes of the decline of the
-species \citep{Woodroffe.2012}. In result, the species currently marks the
-KAZA's most endangered large carnivore and has been a assigned a very high
-conservation priority. Importantly, due to their inherent mobility and intrinsic
-need for vast undisturbed landscapes, AWDs have been proposed as surrogate
-species for landscape connectivity (see recent paper on multispecies
-connectivity). Nevertheless, the species has received little attention in the
-connectivity literature, mainly due to the difficulty in observing wild dog
-dispersal.
+individuals remain in small and spatially scattered subpopulations
+\citep{Woodroffe.2012}. Within those subpopulations, wild dogs form cohesive
+packs comprising 8 to 12 adults and their offspring \cite{McNutt.1995}. After
+reaching sexual maturity, male and female offspring form same-sex coalitions and
+disperse from their natal pack \citep{McNutt.1996, Behr.2020}. New packs are
+formed when dispersing coalitions join unrelated opposite-sex dispersing
+coalitions \citep{McNutt.1996}. Dispersing wild dogs can cover several hundred
+kilometers across a variety of landscapes \citep{DaviesMostert.2012,
+Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the few strongholds for this
+species lies near the Moremi Game Reserve in northern Botswana, which is part of
+the world's largest transboundary protected area, namely the Kavango-Zambezi
+Transfrontier Conservation Area (KAZA-TFCA). This area has originally been
+intended to facilitate migration of elephants, but is expected to benefit a
+multitude of other species \citep{Elliot.2014, Brennan.2020, Hofmann.2021}.
+
+% Previous Paper
+\subsection{Previous Paper (90\%)}
+In a previous study, we assessed landscape connectivity for dispersing African
+wild dogs within the KAZA-TFCA using a least-cost corridor approach
+\citep{Hofmann.2021}. For this, we fitted a basic habitat selection model based
+on which we predicted landscape permeability. We now expand on this knowledge
+and use ISSF to develop a more detailed movement model of dispersing wild dogs.
+We then use this model to simulate dispersers moving across the KAZA-TFCA. Based
+on simulations, we compute heatmaps and identify potential dispersal hotspots.
+We also showcase how network metrics relevant to landscape connectivity can be
+computed. Our results show that a simulation-based approach yields several major
+benefits over traditional connectivity modeling techniques. Most importantly,
+simulations provide a more generic view on how connectivity emerges and to which
+degree connectivity depends on the dispersal duration. In addtion, by generating
+proper dispersal trajectories, network theory can be applied to calculate
+network metrics that are pertinent to connectivity analysis. Finally, we put
+forward additional opportunities using simulations that go beyond the scope of
+this paper.
 
 \section{Methods}
-\subsection{Study Area}
+\subsection{Study Area (90\%)}
 The study area was centered at -17\degree 13'9''S, 23\degree 56'4''E
@@
 woodland habitats. A dominant hydrogeographical feature in its center is the
-flood-pulsing Okavango Delta. The delta and its surroundings are considered a
-stronghold for African wild dogs that may act as a source for the recolonization
-of surrounding habitats \citep{Cozzi.2013}. The wet season within the KAZA-TFCA
-lasts from November to March and is out of phase with the main flooding of the
-Okavango Delta which peaks between July and August \citep{McNutt.1996,
-Wolski.2017}. While large portions of land within the KAZA-TFCA are designated
-national parks or other protected areas, substantial human influence remains due
-to roads, agricultural sites and settlements and villages.
+flood-pulsing Okavango Delta. The wet season within the KAZA-TFCA lasts from
+November to March and is out of phase with the main flooding of the Okavango
+Delta which peaks between July and August \citep{McNutt.1996, Wolski.2017}.
+While large portions within the KAZA-TFCA are designated national parks or other
+protected areas, substantial human influence remains due to roads, agricultural
+sites and settlements and villages.
 
@@
     \includegraphics[width = \textwidth]{99_StudyArea.png}
-    \caption{(a) The study area of our research was confined by a bounding box
-    encompassing the entire KAZA-TFCA, which comprises parts of Angola, Namibia,
-    Botswana, Zimbabwe, and Zambia. (b) The KAZA-TFCA represents the world's
-    largest terrestrial conservation area and covers a total of 520'000
-    km\textsuperscript{2}. Its purpose is to re-establish connectivity between
-    already-existing national parks and other protected areas (green polygons).
-    The dispersal data used in this study was collected on a free-ranging
-    African wild dog population inhabiting the Moremi National Park in northern
-    Botswana.}
+    \caption{(a) The study area was confined by a bounding box encompassing the
+    entire KAZA-TFCA and spanned parts of Angola, Namibia, Botswana, Zimbabwe,
+    and Zambia. (b) The KAZA-TFCA represents the world's largest terrestrial
+    conservation area and covers a total of 520'000 km\textsuperscript{2}. Its
+    purpose is to re-establish connectivity between already-existing national
+    parks (dark green) and other protected areas (light green). The dispersal
+    data used in this study was collected on a free-ranging African wild dog
+    population inhabiting the Moremi National Park in northern Botswana.}
     \label{StudyArea}
@@
 
-\subsection{GPS Relocation Data}
+\subsection{GPS Relocation Data (90\%)}
 Between 2011 and 2019, we collected GPS relocation data on dispersing wild dogs
@@
 \citep{Osofsky.1996, Cozzi.2020} that was injected by dart, fired from a
-CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). Anesthesia
-protocols were approved by the Ministry of Environment, Natural Resources
-Conservation and Tourism of Botswana (permit EWT 8/36/4 XXXVI). Immobilized
+CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). Immobilized
 individuals were fitted with GPS/Satellite radio collars (\textit{Vertex Lite;
@@
 decomposable piece of cotton. Handling and collaring of all individuals was
-supervised by a Botswana-registered wildlife veterinarian and all individuals
+supervised by a Botswana-registered wildlife veterinarian. All individuals
 rejoined their pack within one hour after immobilization. 16 collared
-individuals eventually dispersed, each in a separate same-sex coalition (7
-female and 9 male coalitions).
-
+individuals eventually dispersed, each in a separate same-sex dispersal
+coalition (7 female and 9 male coalitions).
+
+Because behavior during dispersal is more pertinent for assessing landscape
+connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded all data that was
+collected during residency and only retained GPS data recorded during dispersal.
 During dispersal, collars were programmed to record a GPS fix every 4 hours.
@@
 system, which allowed remote tracking of individuals, even if they left the main
-study area and crossed international borders. In some instances exact dispersal
-dates were known from field observations. In other cases, we distinguished
-between residency and dispersal using the net-squared displacement metric. This
-metric measures the squared Euclidean distance a GPS relocation to a reference
-point \citep{Borger.2012}, which in our case was set to the center of each
+study area and crossed international borders. In some instances, exact dispersal
+dates were known from field observations. Otherwise, we distinguished between
+residency and dispersal using the net-squared displacement metric. This metric
+measures the squared Euclidean distance of a GPS relocation to a reference point
+\citep{Borger.2012}, which in our case was set to the center of each
 individual's natal home range. As such, dispersal was deemed to have started
 when an individual left its natal home range and ended once individuals became
-sedentary again. Because behavior during dispersal is more pertinent for
-assessing connectivity, we discarded all data that was collected during
-residency and only retained GPS data recorded during dispersal. As previous
-research revealed similar behavior of females and males during dispersal
-\citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish between sexes. After
-collection, we converted collected GPS coordinates (n = 4'169) to steps, where
-each step represented the straight-line distance traveled by and individual
-between two consecutive GPS relocations \citep{Turchin.1998}. To ensure a
-regular sampling interval, we removed fixes that were not collected on the
-4-hourly schedule.
-
-\subsection{Covariates}
+sedentary again. As previous research revealed similar behavior of females and
+males during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did not
+distinguish between sexes. After collection, we converted collected GPS
+coordinates (n = 4'169) to steps, where each step represented the straight-line
+distance traveled by and individual between two consecutive GPS relocations
+\citep{Turchin.1998}. To ensure a regular sampling interval, we removed fixes
+that were not successfully collected on the 4-hourly schedule.
+
+\subsection{Covariates (90\%)}
 We represented the physical landscape across the study area using a set of
-habitat covariates including water-cover, distance to water, tree-cover, and
+habitat covariates including water-cover, distance to water, woodland-cover, and
 shrub/grassland-cover. Because water cover greatly changes within and between
-years around the Okavango Delta, we applied a remote sensing technique that
-allowed us to generate frequently updated water cover layers and corresponding
-distance to water layers for the close surroundings of the Delta (see
-\citealp{Wolski.2017} and \citealp{Hofmann.2021}). Thus, those layers temporally
-aligned with our dispersal events. We also computed a proxy for human influence,
-depicting anthropogenic pressures stemming from human-density, agricultural
-sites, and roads. All obtained layers were coarsened or interpolated to a target
-resolution of 250 m by 250 m and cropped slightly beyond the extent of the
-KAZA-TFCA. Further details on the sources and preparation of each habitat
-covariate are given in \cite{Hofmann.2021}.
+years in the Okavango Delta, we applied a remote sensing algorithm and generated
+frequently updated water cover layers and corresponding distance to water layers
+(see \citealp{Wolski.2017} and \citealp{Hofmann.2021}). Resulting layers thus
+temporally aligned with our dispersal events. We furthermore computed a proxy
+for human influence, depicting anthropogenic pressures stemming from
+human-density, agricultural sites, and roads. All spatial layers were coarsened
+or interpolated to a target resolution of 250 m by 250 m. Further details on the
+sources and preparation of each habitat covariate are given in
+\cite{Hofmann.2021}.
 
@@
 realized during periods of low wild dog activity (17:00 to 07:00 local time) or
-high wild dogactivity (09:00 to 17:00 local time). Handling and manipulation of
-all data, as well as all modelling and simulating was achieved with statistical
-software {\tt R}, version 3.6.6 \citep{R.2019}. {\tt R}-scripts to reproduce the
-analyses are provided on Github.
-
-\subsection{Movement Model}
-We used integrated step selection functions (iSSF) to parametrize a mechanistic
-movement model of dispersing wild dogs. In contrast to regular step selection
-functions, \textit{integrated} step selection functions allow simultaneous
-inference on movement and habitat preferences of the studied animal, as well as
-on potential interactions between the two. This results in less biased selection
-estimates and allows to treat the resulting model as a fully mechanistic
-movement model based on which movement can be simulated \citep{Avgar.2016}. To
-conduct iSSF analysis, we paired each observed step with 24 random steps.
-Together, an observed and its random steps formed a stratum that received a
-unique identifier. The 24 random steps served as pseudeo-absence and represented
-a potential alternative that the animal could have realized but decided not to.
-We generated random steps by sampling random turning angles from a uniform
-distribution (\(-\pi, +\pi\)) and step lengths from a gamma distirbution that
-was fitted to observed steps (scale = 6'308, shape = 0.37). Although a larger
-number of random steps reduces the sampling error \citep{Avgar.2016},
-preliminary analyses revealed that 24 steps were sufficient in our case. Along
-each step, we extracted and averaged spatial covariates using the {\tt velox}
-package and we calculated the movement metrics \textsf{sl}, \textsf{log(sl)},
-and \textsf{cos(ta)}. To facilitate model convergence, we standardized all
-continuous covariates to a mean of zero and a standard deviation of one. Since
-correlation among covariates was low (\(|r| > 0.6\); \citealp{Latham.2011}), we
-retained all of them for modeling.
-
-To contrast realized steps (scored 1) and random steps (scored 2) and fit a
-movement model, we assumed that animals assigned a selection score \(w(x)\) to
-each step. The score depended on the step's covariates \(X\), as well as the
-animals preferences \(\beta\) towards these covariates. The probability of a
-step being realized was then contingent on the step's selection score, as well
-as on the selection scores of all other step in the same stratum. To estimate
-the preferences of interest, we ran conditional logistic regression models in
-the r-package {\tt glmmTMB}. More specifically, we applied the method developed
-by \citep{Muff.2020}, which allows to fit random slopes. Here, we treated animal
-IDs as random effect and modelled random slopes for each covariate and we fixed
-the random intercept variance at an arbitrary high value of
-10\textsuperscript{6} to make use of the  ``poission''-trick  \citep{Muff.2020}.
-
-The movement model was based on a habitat selection model for dispersing wild
+high wild dog activity (09:00 to 17:00 local time). Handling and manipulation of
+all data, as well as all models and simulations were implemented with the
+statistical software {\tt R}, version 3.6.6 \citep{R.2019}. Several helper
+functions were written in {\tt C++} and imported into {\tt R} using the {\tt
+Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}
+
+\subsection{Movement Model (80\%)}
+We used ISSFs to parametrize a mechanistic movement model for dispersing wild
+dogs \citep{Avgar.2016}. To conduct iSSF analysis, we paired each observed step
+with 24 random steps. An observed and its 24 random steps thus formed a stratum
+and received a unique identifier. We generated random steps by sampling random
+turning angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths
+from a gamma distirbution that was fitted to observed steps (scale = 6'308,
+shape = 0.37). Along each step, we extracted and averaged spatial covariates
+using the {\tt velox}. We also calculated the movement metrics \textsf{sl},
+\textsf{log(sl)}, and \textsf{cos(ta)}. To facilitate model convergence, we
+standardized all continuous covariates to a mean of zero and a standard
+deviation of one. Since correlation among covariates was low (\(|r| > 0.6\);
+\citealp{Latham.2011}), we retained all of them for modeling.
+
+To contrast realized steps (scored 1) and random steps (scored 0), we assumed
+that animals assigned a selection score \(w(x)\) of the exponential form to each
+step. The selection score \(w(x)\) of each step depended on its associated
+covariates (\(x_1, x_2, ..., x_n\)) and on the animal's relative selection
+strengths (i.e. preferences) towards these covariates (\(\beta_1, \beta_2, ...,
+\beta_n\)):
+
+\begin{equation}
+\label{EQ1}
+  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
+\end{equation}
+
+The probability of a step being realized was then contingent on the step's
+selection score, as well as on the selection scores of all other step in the
+same stratum:
+
+\begin{equation}
+\label{EQ2}
+  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
+  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
+\end{equation}
+
+To estimate the preferences of interest, we ran conditional logistic regression
+models in the r-package {\tt glmmTMB}. To handle multiple individuals, we
+applied the mixed effects method developed by \citep{Muff.2020}, which allows to
+fit random slopes in addition to random intercepts. Thus, we treated animal IDs
+as random effect and modeled random slopes for each covariate. We fixed the
+random intercept variance at an arbitrary high value of 10\textsuperscript{6} to
+make use of the  ``poission''-trick \citep{Muff.2020}.
+
+The movement model was based on the habitat selection model for dispersing wild
 dogs presented in \cite{Hofmann.2021}. In the original model, no interactions
@@
 
-\subsection{Dispersal Simulation}
+\subsection{Dispersal Simulation (80\%)}
 We used the most parsimonious movement model to simulate 80'000 dispersing wild
 dogs moving across the KAZA-TFCA. The simulation resembled an inverted iSSF and
-worked as follows. (1) We defined a random source point and assumed a random
+was set up as follows. (1) We defined a random source point and assumed a random
 initial orientation of the animal. (2) Departing from the source point, we
@@
 distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
-distribution. As such, each step corresponded to the straight line movement
-conducted within 4 hours. To prevent unreasonably large steps, we capped sampled
-values to a maximum of 35 km, which corresponds to the farthest distance ever
-traveled within 4 hours by on of our dispersers. (3) Along each random step, we
-extracted and averaged habitat covariates and calculated movement covariates.
-Extracted values were again standardized using the same transformations that
-were used on the input data. (4) We applied the parameterized movement model and
-predicted the selection score \(w(x)\) for each of the proposed steps and
-translated selection scores into probabilities by applying \Cref{EQ2}. (5) We
-randomly sampled one of the random steps based on predicted probabilities and
-used it to determine the animal's new position. We repeated steps (2) to (5)
-until a total of 2'000 steps were realized.
+distribution. Each step corresponded to the straight line movement conducted
+within 4 hours. To prevent unreasonably large steps, we capped sampled values to
+a maximum of 35 km, which corresponds to the farthest distance ever traveled
+within 4 hours by on of our dispersers. (3) Along each random step, we extracted
+and averaged habitat covariates and calculated movement covariates. Extracted
+values were again standardized using the same transformations that were used on
+the input data. (4) We applied the parameterized movement model and predicted
+the selection score \(w(x)\) for each random step and translated selection
+scores into probabilities using \Cref{EQ2}. (5) We randomly sampled one of the
+steps based on predicted probabilities and determined the animal's new position.
+We repeated steps (2) to (5) until a total of 2'000 steps were realized.
 
@@
 layers. In result, simulated dispersers were able to leave and reenter the main
-study area through this buffer zone. In cases where a disperser approached the
-outer map boundary such that one of the proposed random steps transgressed its
-border, we resampled transgressing steps until they fully lied within the buffer
-zone. Although the algorithm for the simulation was mainly coded in {\tt R}, we
-wrote a few helper functions in {\tt C++} and imported them into {\tt R} using
-the {\tt Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}.
-
-\subsection{Source Points}
+study area through this buffer zone. In cases where proposed random steps
+transgressed the outer border of this buffer zone, we resampled transgressing
+steps until they fully lied within the buffer.
+
+\subsection{Source Points (90\%)}
 We initiated virtual dispersers at 50'000 randomly selected source points within
 contiguous protected areas larger \(>\) 700 km\textsuperscript{2}
-(\Cref{SourcePoints} (a)). This conforms to the average home range requirement
-of resident wild dogs \citep{Pomilia.2015} and allowed us to remove areas too
-small to host viable wild dog populations. By distributing source points
-randomly, the number of source points per km\textsuperscript{2} within protected
-areas was approximately equal across the study area. To render potential
-immigrants into the study system, we placed additional 30'000 source points
-inside the buffer zone around the main study area (\Cref{SourcePoints} (b)).
-This resulted in a total of 80'000 source points, each representing the start
-point of an individual disperser.
+(\Cref{SourcePoints}a). This conforms to the average home range requirement of
+resident wild dogs \citep{Pomilia.2015} and allowed us to remove areas too small
+to host viable wild dog populations. By distributing source points randomly, the
+number of source points per km\textsuperscript{2} within protected areas was
+approximately equal across the study area. To render potential immigrants into
+the study system, we placed additional 30'000 source points inside the buffer
+zone around the main study area (\Cref{SourcePoints}b). This resulted in a total
+of 80'000 source points, each representing the start point of an individual
+disperser.
 
@@
 
-\subsection{Heatmap}
-To identify dispersal hotspots across our study area using the simulated
-dispersal events, we rasterized all simulated trajectories and created a heatmap
+\subsection{Heatmap (100\%)}
+To identify dispersal hotspots across our study area, we created a heatmap
 indicating the absolute frequency at which each raster-cell in the study area
-was visited by our virtual dispersers. If the same trajectory crossed a
-raster-cell twice, it was only counted once, meaning that we did not consider
-revisits. This mitigated biases arising due to trapped individuals that were
-moving in circles. To achieve high performance rasterization of all simulated
-trajectories, we used the R-package {\tt terra} \citep{Hijmans.2020}.
-
-\subsection{Network Analysis I: Betweenness}
+was visited by virtual dispersers. For this, we rasterized all simulated
+trajectories and tallied them into a single map. If the same trajectory crossed
+a raster-cell twice, it was only counted once. This way, we did not consider
+revisits and mitigated biases arising from trapped individuals that were moving
+in circles. To achieve high performance rasterization, we used the R-package
+{\tt terra} \citep{Hijmans.2020}.
+
+\subsection{Betweenness (80\%)}
 To pinpoint areas of exceptional relevance for connecting remote regions in our
 study area, we converted simulated trajectories into a network and calculated
-betweenness scores. More specifically, we overlaid the study area (including the
-buffer) with a regular raster that was resolved at 5 x 5 km and where each
-raster-cell represented a node in the final network. We then used the simulated
-trajectories to determine all transitions occurring from one cell to another, as
-well as their frequency. This resulted in an edge-list that we translated into a
-weighted network using the r-package {\tt igraph}. Because {\tt igraph} handles
-the weights \(\omega\) as costs, we inverted the traversal frequency at each
-cell by applying \(\omega = \frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal
-Frequency_i}\). Finally, we calculated the betweenness metric for each node in
-the final network. This metric indicates how often a specific raster-cell lies
-on a shortest path between two other raster-cells and is therefore a useful
-metric to detect movement corridors \citep{BastilleRousseau.2018}.
-
-\subsection{Network Analysis II: Interpatch-Connectivity}
-We also adopted an alternative network view and assessed interpatch connectivity
-between national parks. The decision to focus on national parks was purely out
-of simplicity and could easily be expanded to any other patch of interest. To
-determine interpatch-connectivity, we identified the relative frequency at which
-dispersers originating from one national park successfully moved into an other
-national park. This allowed us to determine \textit{if} and \textit{how often}
-dispersers moved between certain national parks, as well as to estimate the
-average number of steps that was necessary for those movements.
+betweenness scores. For this, we overlaid the study area (including the buffer)
+with a regular raster resolved at 5 x 5 km. We then used the simulated
+trajectories to determine all transitions occurring from one raster-cell to
+another, as well as the frequency at which those transitions occurred. This
+resulted in an edge-list that we translated into a weighted network using the
+r-package {\tt igraph} \citep{Gabor.2006}. Because {\tt igraph} handles edge
+weights (\(\omega\)) as costs, we inverted the traversal frequency in each cell
+by applying \(\omega = \frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal
+Frequency_i}\). Finally, we used the weighted network to calculate betweenness
+scores for each raster-cell in the overlaid raster. The betweenness metric
+indicates how often a specific raster-cell lies on a shortest path between two
+other raster-cells and is therefore a useful metric to detect movement corridors
+\citep{BastilleRousseau.2018}.
+
+\subsection{Interpatch-Connectivity (80\%)}
+We also assessed inter-patch connectivity between national parks in our study
+area. The decision to focus on national parks was purely out of simplicity and
+the same logic could easily be expanded to include other protected areas as
+well. To quantify inter-patch connectivity, we computed the relative frequency
+at which dispersers originating from one national park successfully moved into
+another national park. This allowed us to determine \textit{if} and \textit{how
+often} dispersers moved between certain national parks. Moreover, because time
+was explicit in our IBMM, we were able to estimate \textit{how long} dispersers
+had to move to realize those connections.
 
 \section{Results}
-\subsection{Movement Model}
+\subsection{Movement Model (20\%)}
 Compared to the base model reported in \citep{Hofmann.2021}, our most
 parsimonious movement model included several additional interactions between
-habitat and movement covariates (\Cref{MovementModel} and
+habitat covariates and movement covariates (\Cref{MovementModel} and
 \Cref{MovementModelNumbers}). Although multiple models received an AIC weight
-above zero (T1 in Appendix S1), we only consiered results from the  ``best''
-model for further analyses. Since all models with positive AIC weight contained
-similar covariates, this decision only marginally influenced subsequent results.
-Results from the selected model are given in \Cref{MovementModelNumbers} and
-illustrated in \Cref{MovementModel} (a). Additional plots to ease with the
-interpretation of the model are provided in Appendix XX.
-
-When holding constant for movement behavior, we find that dispersing wild dogs
-avoid water, but prefer its proximity. Dispersers also avoid densely forested
-woodlands, yet prefer open shrublands or grasslands. Finally, dispersers avoid
-moving through landscapes that are influenced by humans.
-
-When looking at the movement kernel, we observe several significant
-interactions. However, except for the interaction \textsf{sl:LowActivity},
-estimated slopes are relatively flat, suggesting that our initial distributions
-for step lengths and turning angles were only marginally biased. For instance,
-the positive effect for \textsf{cos(ta)} indicates that observed turning angles
-are slightly more directional compared to the turning angles proposed by our
-uniform distribution. On the other hand, the strongly significant and negative
-interaction \textsf{sl:LowActivity} reveals that our fitted gamma distribution
-produced steps lengths that were substantially larger that those realized by the
-dispersers during low wild dog activity.
-
-However, we also find a strongly significant and negative interaction between
-the step length and main-activity indicator, showing that realized step lengths
-tended to be much shorter during periods of activity in comparison to the steps
-proposed by our fitted gamma distribution.
-
-Finally, the significant interactions \textsf{cos(ta):HumanInfluence} and
-\textsf{cos(ta):DistanceToWater\textsuperscript{2}} indicate that dispersers
-move much more tortuous in areas influenced by humans but more directed when
-distant to water.
-
-Results from the k-fold cross-validation procedure show that our model
-substantially outperforms a random guess, as the confidence intervals of r.. and
-r.. do not overlap. Additionally, we find that the rank correlation slightly
-improved in comparison to the base model reported in \citep{Hofmann.2021}.
+above zero (Table 1 in Appendix S1), we only considered results from the most
+parsimonious model. Since all models with positive AIC weight contained similar
+covariates, this decision only marginally influenced subsequent results. Results
+from the selected model are given in \Cref{MovementModelNumbers} and illustrated
+in \Cref{MovementModel} (a). Additional plots that facilitate the interpretation
+of the model are provided in Appendix S2.
+
+When looking at the habitat kernel and holding constant the movement kernel, we
+find that dispersing wild dogs avoid water but prefer its proximity. Dispersers
+also avoid densely forested woodlands, yet prefer open shrublands or grasslands.
+Finally, dispersers avoid moving through landscapes that are influenced by
+humans. These results align with our previous findings reported in
+\cite{Hofmann.2021}.
+
+When looking at the movement kernel, we observe several significant estimates.
+However, except for the interaction \textsf{sl:LowActivity}, effect sizes are
+relatively small, suggesting that our proposal distributions for step lengths
+and turning angles were only marginally biased. For instance, the positive and
+significant effect for \textsf{cos(ta)} indicates that realized turning angles
+were slightly more directional than the turning angles proposed by our uniform
+distribution, implying that realized steps in fact followed a von Mises
+Distribution with positive concentration parameter. On the other hand, the
+significant and negative interaction for \textsf{sl:LowActivity} reveals that
+wild dogs moved over shorter distances during low wild dog activity compared to
+the steps suggested by our gamma distribution.
+
+Finally, we turn to the interactions between the two kernels. Here, we observe
+that movement behavior significantly differs depending on habitat covariates. In
+general, steps tend to be shorter in areas with high water cover and high
+woodland cover, yet larger in areas with open shrubs and grasslands. Similarly,
+dispersers realized shorter steps in the vicinity to water, even though the size
+of this effect is negligible. Finally, it appears that dispersers move more
+tortuous in areas influenced by humans, and more directional when far from
+water.
+
+As can be seen in \Cref{MovementModel}b, the k-fold cross-validation procedure
+reveals that our model substantially outperforms a random guess, as indicated by
+the fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
+\(\bar{r}_{s, random}\) do not overlap. Furthermore, the significant correlation
+between ranks and corresponding frequencies for realized steps indicates a good
+fit between predictions and observations (\Cref{MovementModel}b). Finally, we
+find that the rank correlation slightly improved in comparison to the base model
+(\(\bar{r}_{s, realized} = -0.55\); \citealp{Hofmann.2021}).
 
@@
         \toprule
-        Kernel & Covariate & Coefficient & SE & pvalue & Significance \\
+        Kernel & Covariate & Coefficient & SE & p-value & Sign. \\
         \midrule
@@
 
-\subsection{Dispersal Simulation}
+\subsection{Dispersal Simulation (60\%)}
 On a machine with an AMD Ryzen 7 2700X processor with 8 x 3.6 GHz and 64 GB of
@@
 compute (\(\mu = 88.90\), \(\sigma = 1.87\)). As such, the simulation of all
-80'000 dispersers terminated after 120 hours, i.e. five days. However,
-comparable computations will be substantially quicker for smaller study areas,
-as the covariate extraction from large rasters was computationally the most
-expensive task.
-
-On average, step lengths realized by the simulated dispersers (\(\mu = 2'093\)
-m, \(\sigma = 3'067\)) were slightly shorter than those by observed dispersers
-(\(\mu = 2'326\) m, \(\sigma = 3'323\)). Simultaneously, a slightly smaller
-\(cos(ta)\) indicated that simulated dispersers moved with marginally lower
-directionality (\(\mu = 0.057\), \(\sigma = 0.071\)) compared to observed
-dispersers (\(\mu = 0.078\), \(\sigma = 0.072\)). These differences in step
-lengths and turning angles can be attributed to minor disparities between
+80'000 dispersers terminated after 120 hours, i.e. five days. Comparable
+computations will be substantially quicker for smaller study areas or lower
+resolution covariates, as the covariate extraction from large rasters was
+computationally the most expensive task.
+
+On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
+2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
+dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)). Simultaneously,
+simulated dispersers moved marginally less directional (\(\mu_{cos(ta)} =
+0.057\), \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers
+(\(\mu_{cos(ta)} = 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in
+step lengths and turning angles can be attributed to minor disparities between
 habitat conditions at the area within which we collected training data and
@@
 Out of the 50'000 dispersers initiated in protected areas, only 4.5\% eventually
-hit a map boundary, highlighting that boundary effects should be neglectable. In
-contrast, 78\% of the 30'000 dispersers originating from the buffer zone hit a
-map boundary at least once.
-
-\subsection{Heatmaps}
-\Cref{Heatmap} depicts the heatmap rendering the traversal frequency at each
-pixel of the study area across all simulated individuals and steps. The map
-illustrates that large portions of land beyond the borders of the KAZA-TFCA are
-only infrequently visited by dispersers (dark blue areas). On the contrary, we
-observe that within the KAZA-TFCA extensive regions are regularly visited by
-dispersers (bright yellow and red areas). Most notably, the region in northern
-Botswana south of the Linyanti swamp glows in rich red and stands out as highly
-frequented dispersal hub. Nevertheless, massive water bodies such as the
-Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, pose considerable
-dispersal barriers within and therefore limit realized connectivity the
-KAZA-TFCA.
+hit a map boundary, suggesting that biases due to boundary effects should be
+minimal. In contrast, 78\% of the 30'000 dispersers originating from the buffer
+zone eventually hit a map boundary.
+
+\subsection{Heatmap (50\%)}
+\Cref{Heatmap} depicts the heatmap of all 80'000 simulated trajectories rendered
+after 2'000 steps. The map highlights that large portions of land beyond the
+borders of the KAZA-TFCA are only infrequently visited by dispersers (dark blue
+areas), whereas within the KAZA-TFCA we observe several extensive regions that
+are regularly visited (bright yellow and red areas). Most notably, the region in
+northern Botswana south of the Linyanti swamp stands out as highly frequented
+dispersal hub. Still, the presence of several massive water bodies, such as the
+Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in
+considerable dispersal barriers that limit realized connectivity within the
+KAZA-TFCA. Similarly, dispersal appears to be very limited in Zambia's and
+Zimbabwe's part of the KAZA-TFCA, where only few areas are traversed by
+dispersers, which can largely be attributed to a high human influence in these
+regions. Outside the KAZA-TFCA, the most frequently traversed regions include
+the areas of the Central Kalahari National Park in Botswana, the area south-west
+of the Khaudum National Park in Namibia, and the area around the Liuwa Plains
+National Park in Zambia.
 
@@
   \includegraphics[width=\textwidth]{99_Heatmap.png}
-  \caption{}
+  \caption{Heatmap of all 80'000 simulated trajectories after 2'000 steps.
+  Simulations were based on an integrated step selection model, fit to the
+  movement data of dispersing African wild dogs. To generate the heatmap, we
+  rasterized and summed all simulated trajectories. Additional heatmaps
+  showcasing the traversal frequency after different numbers of steps are
+  provided in xx in Appendix S3.}
   \label{Heatmap}
@@
 
-\subsection{Network Analysis I}
-The results of our first network analyses are presented in \Cref{Betweenness}.
-In contrast to the heatmap, \Cref{Betweenness} puts much more emphasis on
-discrete dispersal corridors. Nonetheless, the dispersal hub in northern
-Botswana stands out again and is traversed by a corridor that receives a
-comparably high betweenness score. This implies that the region is particularly
-crucial for connecting other pixels in the study system and hence represents a
-proper dispersal hub. Towards east, the corridor runs through the Chobe National
-Park into the Hwange national park, where it branches out and further extends
-into the distant Matusadona National Park in Zimbabwe. Northwest of the Linyanty
-ecosystem, the same corridor expands into Angola, where it splits and finally
-traverses over a long stretch of unprotected area into the Kafue National Park
-in Zambia. Several additional corridors with slightly lower betweenness scores
-exist, yet most of them run within the boundaries of the KAZA-TFCA. In general,
-only few corridors appear to directly link the peripheral regions of the
-KAZA-TFCA. For instance, there are no viable dispesal corridors between the
+\subsection{Betweenness (50\%)}
+Betweenness scores of each raster-cell the study area are presented in
+\Cref{Betweenness}. While the heatmap emphasized frequently visited areas, the
+map of betweenness reveals rather discrete dispersal corridors. As can be seen,
+the dispersal hotspot in northern Botswana is traversed by a corridor that
+receives a comparably high betweenness score. This implies that the region is
+particularly crucial for connecting other regions in the study system and hence
+represents a proper network hub. Towards east, the corridor runs through the
+Chobe National Park into the Hwange national park, where it branches out and
+further extends into the distant Matusadona National Park in Zimbabwe. Northwest
+of the Linyanty ecosystem, the same corridor expands into Angola, where it
+splits and finally traverses over a long stretch of unprotected area into the
+Kafue National Park in Zambia. Several additional corridors with slightly lower
+betweenness scores exist, yet most of them run within the boundaries of the
+KAZA-TFCA. In general, only few corridors directly link the peripheral regions
+of the KAZA-TFCA. For instance, there are only few corridors between the
 Matusadona National Park in Zimbabwe and the Kafue National Park in Zimbabwe.
@@
 
-\subsection{Network Analysis II}
-The results from the second network analysis that was focused on connectivity
-between national parks are given in \Cref{AreasReached}. The map shows between
-which national parks direct links exist and how frequent they are, as well as
-the average duration a disperser had to move to realize those links. For
-instance, 6.8 \% of the simulated dispersers originating from the Moremi
-National Park successfully reached the Chobe National Park and 4.2 \% reached
-the Hwange National Park in Zimbabwe. On average, dispersers moved for 623 steps
-before arriving at Chobe (SD = 520) and for 1'413 steps before arriving at
-Hwange (SD = 371).
+\subsection{Interpatch Connectivity (0\%)}
+The results of the analysis of interpatch connectivity is given in
+\Cref{AreasReached}. The map shows between which national parks direct links
+exist and how frequent they are, as well as the average duration a disperser had
+to move to realize those links. For instance, 6.8 \% of the simulated dispersers
+originating from the Moremi National Park successfully reached the Chobe
+National Park and 4.2 \% reached the Hwange National Park in Zimbabwe. On
+average, dispersers moved for 623 steps before arriving at Chobe (SD = 520) and
+for 1'413 steps before arriving at Hwange (SD = 371).
 
@@
 
-\section{Discussion}
+\section{Discussion (10\%)}
 Our connectivity network further suggests that dispersers from the Okavango
@@
 suggests that qualitative results are quite insensitive to the exact
-methodological approach. Still, we believe that a simulation based approach
+methodological approach. Still, we believe that a simulation-based approach
 offers possibilities for much richer inferences compared to traditional
@@
 evolves over time as the flood climaxes and retracts again. In fact, one of the
-major advantages of such simulation based approaches is that a dynamic
+major advantages of such simulation-based approaches is that a dynamic
 environment can be rendered, as time is explicit in these models
-\citep{Zeller.2020}. This contrasts with traditional modelling approaches such
+\citep{Zeller.2020}. This contrasts with traditional modeling approaches such
 least-cost analysis or circuit theory, where the temporal dimension cannot be
@@
 during dispersal is lower compared to residency, suggesting the presence some
-fitness benefits to dispering individuals.
+fitness benefits to dispersing individuals.
 
@@
 density of potential dispersers in the source areas. As such, trajectories from
-areas with high density would enter the heatmap with above average weigth.
+areas with high density would enter the heatmap with above average weight.
 
@@
 
+Future studies could also investigate the sensitivity of such simulations with
+respect to estimated habitat preferences. Here, we used point estimates and did
+not vary degree by which individuals preferred or avoided certain covariates.
+Nevertheless, the degree to which our results depend on these estimates is
+unknown. Because data are usually scarce for dispersal studies on endangered
+species, point estimates may be inaccurate \citep{Wiegand.2003,
+KramerSchadt.2007}.
+
 \section{Authors' Contributions}
@@
 GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication.
+the time of publication. Access to all R-scripts for our analyses is provided
+through Github.
 

### 2021-05-05 Chapter_1/04_Manuscript/Manuscript_Support.tex
Worked on manuscript for 1 chap
Various updates

@@
   The base model upon which we based our movement model is depicted in the last
-  row. We omitted any other model with an AIC weight of zero.}
+  row. We omitted all models with an AIC weight of zero.}
   \label{ModelAICs}
@@
       \begin{threeparttable}
-        \begin{tabular}{llllll}
+        \begin{tabular}{lllll}
         \toprule
-          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik & Model ID \\
+          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik \\
           \midrule
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 & 1\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 & 2\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 & 3\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 & 4\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 & 5\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 & 6\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 & 7\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 & 8\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 & 9\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 & 10\\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) & 89392.88 & 0.00 & 0.15 & -44670.44 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:log(sl) & 89393.92 & 1.04 & 0.09 & -44669.96 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + DTW:log(sl) & 89394.13 & 1.25 & 0.08 & -44670.06 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:log(sl) & 89394.25 & 1.37 & 0.08 & -44670.13 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl & 89394.36 & 1.48 & 0.07 & -44672.18 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + log(sl):LA & 89394.44 & 1.56 & 0.07 & -44670.22 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:sl & 89394.56 & 1.68 & 0.07 & -44670.28 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:log(sl) & 89394.57 & 1.69 & 0.07 & -44670.29 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:cos(ta) & 89394.59 & 1.71 & 0.07 & -44670.30 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:cos(ta) & 89394.63 & 1.75 & 0.06 & -44670.31 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + sl:cos(ta) & 89394.68 & 1.80 & 0.06 & -44672.34 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:log(sl) & 89394.69 & 1.81 & 0.06 & -44670.35 \\
+          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:cos(ta) & 89394.84 & 1.96 & 0.06 & -44670.42 \\
           \hdashline
-          ...\\
+          \\
+          \vdots & \vdots & \vdots & \vdots & \vdots \\
+          \\
           \hdashline
-          Base Model: cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 787.67 & 0.00 & -45030.70 & NA\\
+          \\
+          \textbf{Base Model}: cos(ta) + sl + log(sl) + WA + WO + DTW + HI + SH & 90091.40 & 787.67 & 0.00 & -45030.70\\
+          \\
          \bottomrule
@@
        \begin{tablenotes}
-         \item \textit{Note:} W = Water, DTW = Distance To Water, S =
-         Shrubs/Grassland, T = Trees, P = Protected, HI = Human Influence, RC =
-         Road Crossing, DTR = Distance To Roads, MA = Main Activity.
+         \item \textit{Note:} ta = Turning Angle, sl = Step Length, LA = Low Activity, WA = Water,
+         DTW = Distance To Water, SH = Shrubs/Grassland, WO = Woodland, HI =
+         Human Influence.
        \end{tablenotes}
@@
 
-\begin{figure}[htpb]
+% \begin{figure}[htpb]
+%   \begin{center}
+%     \includegraphics[width = \textwidth]{99_MovementModelAICs.png}
+%     \caption{Coefficients of all movement models with an AIC weight above zero.}
+%     \label{AllMovementModels}
+%   \end{center}
+% \end{figure}
+
+%------------------------------------------------------------------------------
+%	Appendix S2: Evolution of Heatmaps
+%------------------------------------------------------------------------------
+\newpage
+\newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
+\section{Evolution of Heatmaps}
+\begin{figure}[hbtp]
   \begin{center}
-    \includegraphics[width = \textwidth]{99_AllMovementModels}
-    \caption{Coefficients of all movement models with an AIC weight above one.}
-    \label{AllMovementModels}
+    \includegraphics[width = 0.88\textwidth]{99_HeatmapsIndividual.png}
+    \caption{Heatmaps produced after 125, 500, and 2000 simulated steps,
+    respectively. The left panel (a1, a2, a3) was generated based on simulations
+    initiated within the main study area, the right panel (b1, b2, b3) was
+    generated based on simulations initiated within the buffer area. To produce
+    the heatmap presented in the main text, we summed up values from maps a3 and
+    b3.}
+    \label{Trajectories}
   \end{center}
 \end{figure}
+\restoregeometry
+
+%------------------------------------------------------------------------------
+%	Appendix S3: Evolution of Betweenness Maps
+%------------------------------------------------------------------------------
+\newpage
+\newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
+\section{Evolution of Betweenness}
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width = 0.88\textwidth]{99_BetweennessIndividual.png}
+    \caption{Maps of betweenness scores produced after 125 (a), 500 (b), and
+    2000 (c) simulated steps, respectively. A high betweenness score indicates
+    that the respective area is paramount for linking other regions in the study
+    area. It therefore highlights movement corridors that provide access to
+    various destinations.}
+    \label{Trajectories}
+  \end{center}
+\end{figure}
+\restoregeometry
 

### 2021-05-05 Chapter_1/02_R-Scripts/02_Analysis/11_AreasReached.R

@@ -20,6 +20,7 @@ library(rgeos)          # For spatial ananylsis
 library(igraph)         # For network analysis
 library(davidoff)       # Custom functions
 library(rgdal)          # To load spatial data
+library(ggnetwork)      # Package to plot networks with ggplot
 
 ################################################################################
 #### Rasterize National Parks

### 2021-05-05 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/AreasReached.R

@@ -33,7 +33,7 @@ library(gridExtra)      # To combine ggplots
 visits <- read_rds("03_Data/03_Results/99_AreasReached.rds")
 
 # Load protected areas
-prot    <- readOGR("03_Data/02_CleanData/02_LandUse_Protected_PEACEPARKS.shp")
+prot <- readOGR("03_Data/02_CleanData/02_LandUse_Protected_PEACEPARKS.shp")
 
 # Keep only national parks
 prot <- subset(prot, Desig == "National Park")
@@ -86,9 +86,9 @@ crs(kaza_ext) <- CRS("+init=epsg:4326")
 kaza$Name <- "KAZA-TFCA Borders"
 
 # Convert objects to sf
-kaza                  <- st_as_sf(kaza)
-africa                <- st_as_sf(africa)
-prot                  <- st_as_sf(prot)
+kaza   <- st_as_sf(kaza)
+africa <- st_as_sf(africa)
+prot   <- st_as_sf(prot)
 
 # Convert heatmap to dataframe
 r <- as.data.frame(r, xy = T)

### 2021-05-14 Chapter_1/04_Manuscript/Manuscript.tex
worked on manuscript

@@
 %------------------------------------------------------------------------------
-\title{Step by Step: Using Step Selection Analysis to Simulate Dispersal and
+\title{Step by Step: Using Step Selection Functions to Simulate Dispersal and
 Assess Landscape Connectivity}
+
+% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
+% Connectivity}
 
@@
 estimation of connectivity and identification of dispersal corridors has become
-a task of extraordinary importance in ecological studies.
+a task of extraordinary importance for conservation authorities worldwide.
 
 Over the past two decades, ecologists have primarily relied on analytical tools
-such as least-cost methods and circuit theory to investigate landscape
-connectivity. Despite their usefulness for various ecological applications, both
-methods make several restricting assumptions that limit their suitability in
-reality. To address their shortcomings, simulations from individual-based
-movement models have been proposed and applied. To overcome their shortcomings,
-simulations from individual-based movement models have been proposed. Yet, due
-to the infinite number of possibilities of coming up with an individual-based
-model, a unified and objective framework is missing.
+such as least-cost analysis and circuit theory to model and investigate
+landscape connectivity. Despite their usefulness for a diverse suite of
+ecological applications, both methods make several restricting assumptions that
+limit their suitability in reality. To address these shortcomings, dispersal
+simulations from individual-based movement models have been proposed and
+applied. Yet, due to the almost infinite amount of non-trivial decisions a
+modeler faces when parametrizing such models, a unified and objective framework
+is missing.
 
 Recent innovations in movement ecology have brought forward novel opportunities
-to study animal dispersal and landscape connectivity. In particular, the rich
-suite of resource selection functions, including point-, step-, and
-path-selection functions have undergone substantial improvements over the past
+to study animal dispersal and estimate landscape connectivity. In particular,
+the rich suite of resource selection functions, namely point-, step-, and
+path-selection functions, have undergone substantial improvements over the past
 years. Most notably, step-selection functions have been generalized to
 \textit{integrated} step selection functions, which essentially represent fully
-mechanistic movement models based on which individual animal movement could be
+mechanistic movement models based on which an individual's movement could be
 simulated. While such models have been applied to study \textit{steady-state}
-utilization distribution, a similar approach may be useful for investigating
-\textit{transient} movement behavior and thereby study landscape connectivity by
-means of simulated dispersal events.
-
-Here, we showcase the use of integrated step selection analysis to simulate
-dispersal of the highly endangered African wild dog across the world's largest
+utilization distribution resident animals, a similar approach may be useful for
+investigating \textit{transient} movement behavior and study landscape
+connectivity.
+
+Here, we showcase the use of integrated step selection functions to simulate
+dispersal of the endangered African wild dog across the world's largest
 transboundary conservation area, the Kavango-Zambezi Transfrontier Conservation
 Area (KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
-coalitions departing from northern Botswana. We combine the movement data with
-relevant spatial covariate layers and used integrated step selection functions
-to parametrize a fully mechanistic movement model rendering wild dog dispersal.
-Based on this model, we simulate 80'000 dispersers, originating from protected
-areas and moving across the extent of the KAZA-TFCA. We then generate a heatmap
-and use network theory to reveal dispersal hotspots and crucial bottlenecks
-across the study area. Finally, we discuss the benefits and pitfalls of such
-dispersal simualtions and highlight potential improvements to be made in the
-future.
+coalitions in combination with relevant habitat covariates. We analyse the data
+using integrated step selection functions and parametrize a fully mechanistic
+movement model rendering wild dog dispersal. Based on this model, we simulate
+80'000 dispersers, originating from protected areas and moving across the extent
+of the KAZA-TFCA. We then generate a heatmap indicating regions frequently
+visited by dispersers, use network theory to reveal dispersal hotspots and
+crucial bottlenecks across the study area. Finally, we discuss the benefits and
+pitfalls of such dispersal simualtions and highlight potential improvements to
+be made in the future.
+
 \end{abstract}
@@
 
-\section{Introduction (90\%)}
+\section{Introduction}
 
@@
 \subsection{Importance of Dispersal \& Connectivity (90\%)}
-Dispersal is defined as the movement of individuals from their natal location to
-the site of first reproduction \cite{Howard.1960}. It is a vital process
-governing the dynamics wild animal populations that are distributed in space
-\citep{Hanski.1998, Clobert.2012} and may strongly affect population dynamics at
-different spatial and social scales \citep{Hanski.1999a, Clobert.2012}.
-Dispersal allows species to avoid inbreeding and maintain genetic diversity
-\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013}, to
-rescue small and non-viable populations \citep{Brown.1977}, and to promote the
-colonization of unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}.
-However, successful dispersal requires a sufficient degree of functional
-connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the identification
-and protection of major dispersal corridors has become an important task in
-conservation science \citep{Doerr.2011, Rudnick.2012}. To achieve this, reliable
-information on movement behavior during dispersal and knowledge about factors
-that limit dispersal and connectivity is paramount \citep{Baguette.2013,
-Vasudev.2015}.
+Dispersal is defined as the movement of individuals away from their natal
+location to the site of first reproduction \cite{Howard.1960}. It is a vital
+process governing the dynamics wild animal populations that are distributed in
+space \citep{Hanski.1998, Clobert.2012} and may strongly affect population
+dynamics at different spatial and social scales \citep{Hanski.1999a,
+Clobert.2012}. Dispersal allows species to avoid inbreeding and maintain genetic
+diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
+Baguette.2013}, to rescue small and non-viable populations \citep{Brown.1977},
+and to promote the colonization of unoccupied habitats \citep{Hanski.1999b,
+MacArthur.2001}. However, successful dispersal requires a sufficient degree of
+functional connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the
+identification and protection of major dispersal corridors has become an
+important task in conservation science \citep{Doerr.2011, Rudnick.2012}. In
+order to pinpoint relevant dispersal hotspots, reliable information on movement
+behavior during dispersal and knowledge about factors that limit dispersal and
+connectivity is paramount \citep{Baguette.2013, Vasudev.2015}.
 
@@
 input to a connectivity model that is used to reveal movement corridors
-\citep{Diniz.2020}. Two of the most prominent connectivity models are least-cost
-path analysis (LCP analysis; \citealp{Adriaensen.2003}) and circuit theory (CT
-\citealp{McRae.2006, McRae.2008}), both graph-based methods that estimate
-conductance of the landscape. Despite their intuitive nature and ease of use,
-both methods make rigorous assumptions about animal movement that may or may not
-be fulfilled in reality \citep{Diniz.2020}.
+\citep{Diniz.2020}. In this regard, two of the most prominent connectivity
+models are least-cost path analysis (LCP analysis; \citealp{Adriaensen.2003})
+and circuit theory (CT \citealp{McRae.2006, McRae.2008}), both graph-based
+methods that estimate conductance of the landscape. Despite their intuitive
+nature and ease of use, both methods make rigorous assumptions about animal
+movement that are often not fulfilled in reality \citep{Diniz.2020}.
 
@@
 route accordingly. These assumptions may be reasonable for migrating animals,
-yet they are unlikely to be true for dispersers that move over long distances
-into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Finally,
-LCPs are only one pixel wide, meaning that their absolute size depends on the
-resolution of chosen covariate layers \citep{Diniz.2020}. Although some of these
-issues can be alleviated using less stringent versions of the LCP algorithm
-(e.g. least-cost \textit{corridors} \citep{Pinto.2009}, \textit{thresholded}
-least-cost paths \citep{Landguth.2012}, and \textit{randomized} least-cost paths
-\citep{Panzacchi.2016, VanMoorter.2021}), a certain degree of arbitrariness in
-the assumptions remains. CT entails similar restrictions. Because CT only
-considers movements from the source cell to its 4 or 8 adjacent cells, it
-implicitly posits a perceptual range of a single pixel. It is also built on the
-assumption of a complete random walk \citep{Diniz.2020}, implying that
-directional biases cannot be rendered, albeit being very common in dispersal
-movements \citep{Cozzi.2020, Hofmann.2021}. Ultimately, both LCP analysis and CT
-omit the temporal dimension of dispersal. In result, statements about the
-expected duration required to traverse a certain corridor are impossible.
-Because movement is not modelled explicitely, neither of the methods can render
-possible interactions between movement behavior and landscape characteristics
-such that connectivity mainly arises as a result of the landscape structure.
-This is referred to as structural connectivity and stands in contrast to
-functional connectivity, which also renders the behavioral response of the
-animal with respect to prevailing habitat conditions \citep{Tischendorf.2000}.
-Even though functional connectivity is more difficult to estimate, a functional
-view is the ultimate goal in conservation science because it has direct
-consequences for gene flow \citep{Baguette.2013}.
+yet they are unlikely to hold for dispersers, which typically move over long
+distances into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}.
+Finally, LCPs are only one pixel wide, meaning that their absolute size depends
+on the resolution of chosen covariate layers \citep{Diniz.2020}. Although some
+of these deficiencies can be addressed using less stringent versions of the LCP
+algorithm (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
+\textit{thresholded} least-cost paths \citep{Landguth.2012}, and
+\textit{randomized} least-cost paths \citep{Panzacchi.2016, VanMoorter.2021}), a
+certain degree of arbitrariness in the assumptions remains. CT entails similarly
+unreasonable restrictions. Because CT only considers movements from the source
+cell to its 4 or 8 adjacent cells, it implicitly posits a perceptual range of a
+single pixel. As such, the perceptual range cannot be defined based on
+biological observations, but is imposed by the resolution of the reference grid.
+CT is also built on the assumption of a complete random walk \citep{Diniz.2020},
+implying that directional biases cannot be rendered, albeit being very common in
+dispersal movements \citep{Cozzi.2020, Hofmann.2021}. Ultimately, neither LCP
+analysis nor CT are capable of rendering the temporal dimension of dispersal.
+Statements about the expected duration required to traverse a certain corridor
+are therefore impossible, albeit being important factors. Likewise, because
+movement is not modelled explicitely, neither of the methods allows to render
+interactions between movement and habitat preferences of the focal species. This
+implies that connectivity mainly arises as a result of the landscape structure
+and is referred to as structural connectivity. Structural connectivity stands in
+contrast to functional connectivity, which also renders the behavioral response
+of the animal with respect to prevailing habitat conditions
+\citep{Tischendorf.2000}. Even though functional connectivity is more difficult
+to estimate, a functional view is the ultimate goal in conservation science
+because it has direct consequences for gene flow \citep{Baguette.2013}.
 
@@
 \subsection{Study Species \& Study Area (90\%)}
-One of the species for which long-term viability relies on sufficient
-landscape connectivity is the endangered African wild dog \textit{Lycon pictus}.
-While once present across entire sub-Saharan Africa, wild dogs have disappeared
-from a vast majority of their historic range due to persecution by humans,
-habitat destruction, and deadly diseases. As of today, only 6'000 free-ranging
-individuals remain in small and spatially scattered subpopulations
-\citep{Woodroffe.2012}. Within those subpopulations, wild dogs form cohesive
-packs comprising 8 to 12 adults and their offspring \cite{McNutt.1995}. After
-reaching sexual maturity, male and female offspring form same-sex coalitions and
-disperse from their natal pack \citep{McNutt.1996, Behr.2020}. New packs are
-formed when dispersing coalitions join unrelated opposite-sex dispersing
-coalitions \citep{McNutt.1996}. Dispersing wild dogs can cover several hundred
-kilometers across a variety of landscapes \citep{DaviesMostert.2012,
-Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the few strongholds for this
-species lies near the Moremi Game Reserve in northern Botswana, which is part of
-the world's largest transboundary protected area, namely the Kavango-Zambezi
-Transfrontier Conservation Area (KAZA-TFCA). This area has originally been
-intended to facilitate migration of elephants, but is expected to benefit a
-multitude of other species \citep{Elliot.2014, Brennan.2020, Hofmann.2021}.
+One of the species for which long-term viability relies on sufficient landscape
+connectivity is the endangered African wild dog \textit{Lycon pictus}. While
+once present across entire sub-Saharan Africa, wild dogs have disappeared from a
+vast majority of their historic range due to persecution by humans, habitat
+fragmentation and destruction, and deadly diseases \citep{Woodroffe.2012,
+Woodroffe.2020}. As of today, only 6'000 free-ranging individuals remain in
+small and spatially scattered subpopulations \citep{Woodroffe.2012}. Within
+those subpopulations, wild dogs form cohesive packs comprising 8 to 12 adults
+and their offspring \cite{McNutt.1995}. After reaching sexual maturity, male and
+female offspring form same-sex coalitions and disperse from their natal pack
+\citep{McNutt.1996, Behr.2020}. New packs are formed when dispersing coalitions
+join unrelated opposite-sex dispersing coalitions \citep{McNutt.1996}.
+Dispersing wild dogs can cover several hundred kilometers across a variety of
+landscapes \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
+One of the few strongholds for this species lies near the Moremi Game Reserve in
+northern Botswana, which is part of the world's largest transboundary protected
+area, namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
+This area has originally been intended to facilitate migration of elephants, but
+is expected to benefit a multitude of other species \citep{Elliot.2014,
+Brennan.2020, Hofmann.2021}.
 
@@
 on which we predicted landscape permeability. We now expand on this knowledge
-and use ISSF to develop a more detailed movement model of dispersing wild dogs.
-We then use this model to simulate dispersers moving across the KAZA-TFCA. Based
-on simulations, we compute heatmaps and identify potential dispersal hotspots.
-We also showcase how network metrics relevant to landscape connectivity can be
-computed. Our results show that a simulation-based approach yields several major
-benefits over traditional connectivity modeling techniques. Most importantly,
-simulations provide a more generic view on how connectivity emerges and to which
-degree connectivity depends on the dispersal duration. In addtion, by generating
-proper dispersal trajectories, network theory can be applied to calculate
-network metrics that are pertinent to connectivity analysis. Finally, we put
-forward additional opportunities using simulations that go beyond the scope of
-this paper.
+and use ISSF to develop a more detailed movement model of dispersing wild dogs
+(\Cref{GraphicalAbstract}). We then use this model to simulate dispersers moving
+across the KAZA-TFCA. Based on simulations, we compute heatmaps and identify
+potential dispersal hotspots. We also showcase how network metrics relevant to
+landscape connectivity can be computed. Our results show that a simulation-based
+approach yields several major benefits over traditional connectivity modeling
+techniques. Most importantly, simulations provide a more generic view on how
+connectivity emerges and to which degree connectivity depends on the dispersal
+duration. In addtion, by generating proper dispersal trajectories, network
+theory can be applied to calculate network metrics that are pertinent to
+connectivity analysis.
+
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
+    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
+    data and habitat covariates have to be collected. The combined data is then
+    analyzed in an integrated step selection model, which enables the
+    parametrization of the focal species' habitat and movement kernels. The
+    parametrized model is then treated as an individual-based movement model and
+    used to simulate dispersal trajectories. Ultimately, simulated trajectories
+    serve to produce a set of maps that are pertinent to landscape connectivity.
+    This includes a heatmap, indicating the relative traversal frequency across
+    the study area, a betweenness map, highlighting movement corridors and
+    bottlenecks, and, finally, an inter-patch connectivity map, where the
+    frequency of connections and their average duration can be depicted. Photos:
+    Whom to cite? Vectronics or Photographers?}
+    \label{GraphicalAbstract}
+  \end{center}
+\end{figure}
 
@@
 \subsection{Study Area (90\%)}
-The study area was centered at -17\degree 13'9''S, 23\degree 56'4''E
-(\Cref{StudyArea}a) and was represented by a rectangular bounding box that
-stretched over 1.3 Mio. km\textsuperscript{2}, ecompassing the entire KAZA-TFCA
-(\Cref{StudyArea}b). The KAZA-TFCA is the world's largest transboundary
-conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
-and Zambia, covering a total of 520'000 km\textsuperscript{2}. Its landscape
-varies regionally and ranges from savanna, to grassland and from dry to moist
-woodland habitats. A dominant hydrogeographical feature in its center is the
-flood-pulsing Okavango Delta. The wet season within the KAZA-TFCA lasts from
-November to March and is out of phase with the main flooding of the Okavango
-Delta which peaks between July and August \citep{McNutt.1996, Wolski.2017}.
-While large portions within the KAZA-TFCA are designated national parks or other
-protected areas, substantial human influence remains due to roads, agricultural
-sites and settlements and villages.
+The study area was defined by a bounding box centered at -17\degree 13'9''S,
+23\degree 56'4''E (\Cref{StudyArea}a) stretching over 1.3 Mio.
+km\textsuperscript{2} and ecompassing the entire KAZA-TFCA (\Cref{StudyArea}b).
+The KAZA-TFCA represents the world's largest transboundary conservation area and
+comprises parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. It covers a
+total of 520'000 km\textsuperscript{2} and hosts a diverse landscape, ranging
+from savanna to grassland and from dry to moist woodland habitats. In its center
+lies the Okavango Delta, a dominant hydrogeographical feature and the world's
+largest flood-pulsing inland delta. The wet season within the KAZA-TFCA lasts
+from November to March and is out of phase with the flood in the Okavango Delta,
+which peaks between July and August \citep{McNutt.1996, Wolski.2017}. Although
+large portions within the KAZA-TFCA are designated national parks or other
+protected areas, considerable human influence remains due to roads, agricultural
+sites and settlements and villages that are distributed across the KAZA-TFCA's
+landscape.
 
@@
     \includegraphics[width = \textwidth]{99_StudyArea.png}
-    \caption{(a) The study area was confined by a bounding box encompassing the
-    entire KAZA-TFCA and spanned parts of Angola, Namibia, Botswana, Zimbabwe,
-    and Zambia. (b) The KAZA-TFCA represents the world's largest terrestrial
-    conservation area and covers a total of 520'000 km\textsuperscript{2}. Its
-    purpose is to re-establish connectivity between already-existing national
-    parks (dark green) and other protected areas (light green). The dispersal
-    data used in this study was collected on a free-ranging African wild dog
-    population inhabiting the Moremi National Park in northern Botswana.}
+    \caption{Illustration of the study area located in southern Africa. (a) The
+    study area was confined by a bounding box spanning the entire KAZA-TFCA and
+    encompassing parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
+    The KAZA-TFCA represents the world's largest terrestrial conservation area
+    and covers a total of 520'000 km\textsuperscript{2}. Its purpose is to
+    re-establish connectivity between already-existing national parks (dark
+    green) and other protected areas (light green). The dispersal data used in
+    this study was collected on a free-ranging African wild dog population
+    inhabiting the Moremi National Park in northern Botswana.}
     \label{StudyArea}
@@
 from a free-ranging wild dog population inhabiting the Moremi National Park in
-northern Botswana. We selected potential dispersers based on age, pack size,
-number of same‐sex siblings within the pack, and presence of unrelated
-opposite-sex individuals in the pack \citep{McNutt.1996, Behr.2020}. We
-immobilized selected individuals using a cocktail of Ketamine/Xylazine/Atropine
-\citep{Osofsky.1996, Cozzi.2020} that was injected by dart, fired from a
-CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). Immobilized
-individuals were fitted with GPS/Satellite radio collars (\textit{Vertex Lite;
-Vectronic Aerospace GmbH, Berlin}) that guaranteed automated drop-off through a
-decomposable piece of cotton. Handling and collaring of all individuals was
-supervised by a Botswana-registered wildlife veterinarian. All individuals
-rejoined their pack within one hour after immobilization. 16 collared
-individuals eventually dispersed, each in a separate same-sex dispersal
-coalition (7 female and 9 male coalitions).
-
-Because behavior during dispersal is more pertinent for assessing landscape
-connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded all data that was
-collected during residency and only retained GPS data recorded during dispersal.
-During dispersal, collars were programmed to record a GPS fix every 4 hours.
-Collected relocations were regularly transmitted over the Iridium satellite
-system, which allowed remote tracking of individuals, even if they left the main
-study area and crossed international borders. In some instances, exact dispersal
-dates were known from field observations. Otherwise, we distinguished between
-residency and dispersal using the net-squared displacement metric. This metric
-measures the squared Euclidean distance of a GPS relocation to a reference point
-\citep{Borger.2012}, which in our case was set to the center of each
-individual's natal home range. As such, dispersal was deemed to have started
-when an individual left its natal home range and ended once individuals became
-sedentary again. As previous research revealed similar behavior of females and
-males during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did not
-distinguish between sexes. After collection, we converted collected GPS
-coordinates (n = 4'169) to steps, where each step represented the straight-line
-distance traveled by and individual between two consecutive GPS relocations
-\citep{Turchin.1998}. To ensure a regular sampling interval, we removed fixes
-that were not successfully collected on the 4-hourly schedule.
+northern Botswana \citep{Cozzi.2020, Hofmann.2021}. We selected potential
+dispersers based on age, pack size, number of same‐sex siblings within the pack,
+and presence of unrelated opposite-sex individuals in the pack
+\citep{McNutt.1996, Behr.2020}. We immobilized selected individuals using a
+cocktail of Ketamine/Xylazine/Atropine \citep{Osofsky.1996, Cozzi.2020} that was
+injected by dart, fired from a CO\textsubscript{2}-pressurized gun
+(\textit{DAN-Inject, Denmark}). Immobilized individuals were fitted with
+GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
+Berlin}) that guaranteed automated drop-off through a decomposable piece of
+cotton. Handling and collaring of all individuals was supervised by a
+Botswana-registered wildlife veterinarian and all individuals quickly rejoined
+their pack after immobilization.
+
+16 collared individuals eventually dispersed, each in a separate same-sex
+dispersal coalition (7 female and 9 male coalitions). During dispersal, collars
+were programmed to record a GPS fix every 4 hours. Collected relocations were
+regularly transmitted over the Iridium satellite system, which allowed remote
+tracking of individuals, even if they left the main study area and crossed
+international borders. Because behavior during dispersal is more pertinent for
+assessing landscape connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded
+all data that was collected during residency and only retained GPS data recorded
+during dispersal. In some instances, exact dispersal dates were known from field
+observations. Where this was not the case, determined dispersal phases using the
+net-squared displacement metric. Net squared displacement measures the squared
+Euclidean distance of a GPS relocation to a reference point \citep{Borger.2012},
+which in our case was set to the center of each individual's natal home range.
+As such, dispersal was deemed to have started when an individual left its natal
+home range and ended once individuals became sedentary again. As previous
+research revealed similar behavior of females and males during dispersal
+\citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish between sexes. After
+collection, we converted collected GPS coordinates (n = 4'169) to steps, where
+each step represented the straight-line distance traveled by and individual
+between two consecutive GPS relocations \citep{Turchin.1998}. To ensure a
+regular sampling interval, we removed fixes that were not successfully collected
+on the 4-hourly schedule (\( \pm \) 15 minutes).
 
@@
 We represented the physical landscape across the study area using a set of
-habitat covariates including water-cover, distance to water, woodland-cover, and
-shrub/grassland-cover. Because water cover greatly changes within and between
-years in the Okavango Delta, we applied a remote sensing algorithm and generated
-frequently updated water cover layers and corresponding distance to water layers
-(see \citealp{Wolski.2017} and \citealp{Hofmann.2021}). Resulting layers thus
-temporally aligned with our dispersal events. We furthermore computed a proxy
-for human influence, depicting anthropogenic pressures stemming from
-human-density, agricultural sites, and roads. All spatial layers were coarsened
-or interpolated to a target resolution of 250 m by 250 m. Further details on the
-sources and preparation of each habitat covariate are given in
-\cite{Hofmann.2021}.
-
-Besides habitat covariates, we also computed movement metrics that we used as
+habitat covariates that included water-cover, distance to water, woodland-cover,
+and shrub/grassland-cover. Because water cover greatly changes within and
+between years in the Okavango Delta, we applied a remote sensing algorithm and
+generated frequently updated water cover layers and corresponding distance to
+water layers (see \citealp{Wolski.2017} and Appendix A3 in
+\citealp{Hofmann.2021}). Resulting water layers thus temporally aligned with our
+dispersal events. We furthermore computed a proxy for human influence, rendering
+anthropogenic pressures stemming from human-density, agricultural sites, and
+roads. All spatial layers were coarsened or interpolated to a target resolution
+of 250 m by 250 m. Further details on the sources and preparation of each
+habitat covariate are given in \cite{Hofmann.2021}.
+
+Besides habitat covariates, we computed movement metrics that we used as
 movement covariates in our models. Movement metrics were calculated for each
@@
 (\textsf{log(sl)}), and the cosine of the relative turning angle
-(\textsf{cos(ta)}). Because wild dogs follow a diurnal activity pattern, we also
-coded a binary variable (\textsf{LowActivity}) indicating whether a step was
-realized during periods of low wild dog activity (17:00 to 07:00 local time) or
-high wild dog activity (09:00 to 17:00 local time). Handling and manipulation of
-all data, as well as all models and simulations were implemented with the
-statistical software {\tt R}, version 3.6.6 \citep{R.2019}. Several helper
-functions were written in {\tt C++} and imported into {\tt R} using the {\tt
-Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}
+(\textsf{cos(ta)}) (for details see \citep{Avgar.2016, Fieberg.2020}). Because
+wild dogs follow a diurnal activity pattern, we also coded a binary variable
+(\textsf{LowActivity}) indicating whether a step was realized during periods of
+low wild dog activity (17:00 to 09:00 local time) or high wild dog activity
+(09:00 to 17:00 local time). Handling and manipulation of all data, as well as
+all models and simulations were implemented with the statistical software {\tt
+R}, version 3.6.6 \citep{R.2019}. Several helper functions were written in {\tt
+C++} and imported into {\tt R} using the {\tt Rcpp} package
+\citep{Eddelbuettel.2011, Eddelbuettel.2013}
 
@@
 We used ISSFs to parametrize a mechanistic movement model for dispersing wild
-dogs \citep{Avgar.2016}. To conduct iSSF analysis, we paired each observed step
-with 24 random steps. An observed and its 24 random steps thus formed a stratum
-and received a unique identifier. We generated random steps by sampling random
+dogs \citep{Avgar.2016}. To conduct ISSF analysis, we paired each realized step
+with 24 random steps. An observed and its 24 random steps formed a stratum and
+received a unique identifier. We generated random steps by sampling random
 turning angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths
-from a gamma distirbution that was fitted to observed steps (scale = 6'308,
+from a gamma distirbution that was fitted to realized steps (scale = 6'308,
 shape = 0.37). Along each step, we extracted and averaged spatial covariates
-using the {\tt velox}. We also calculated the movement metrics \textsf{sl},
-\textsf{log(sl)}, and \textsf{cos(ta)}. To facilitate model convergence, we
-standardized all continuous covariates to a mean of zero and a standard
-deviation of one. Since correlation among covariates was low (\(|r| > 0.6\);
-\citealp{Latham.2011}), we retained all of them for modeling.
+using the {\tt velox} package \citep{Hunziker.2017}. We also calculated the
+movement metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)} for each
+observed and random step. To facilitate model convergence, we standardized all
+continuous covariates to a mean of zero and a standard deviation of one. Since
+correlation among covariates was low (\(|r| > 0.6\); \citealp{Latham.2011}), we
+retained all of them for modeling.
 
@@
 that animals assigned a selection score \(w(x)\) of the exponential form to each
-step. The selection score \(w(x)\) of each step depended on its associated
-covariates (\(x_1, x_2, ..., x_n\)) and on the animal's relative selection
-strengths (i.e. preferences) towards these covariates (\(\beta_1, \beta_2, ...,
-\beta_n\)):
+step \citep{Fortin.2005}. The selection score \(w(x)\) of each step depended on
+its associated covariates (\(x_1, x_2, ..., x_n\)) and on the animal's
+preferences (i.e. relative selection strengths; \citealp{Avgar.2017}) towards
+these covariates (\(\beta_1, \beta_2, ..., \beta_n\)):
 
@@
 To estimate the preferences of interest, we ran conditional logistic regression
-models in the r-package {\tt glmmTMB}. To handle multiple individuals, we
-applied the mixed effects method developed by \citep{Muff.2020}, which allows to
-fit random slopes in addition to random intercepts. Thus, we treated animal IDs
-as random effect and modeled random slopes for each covariate. We fixed the
-random intercept variance at an arbitrary high value of 10\textsuperscript{6} to
-make use of the  ``poission''-trick \citep{Muff.2020}.
-
-The movement model was based on the habitat selection model for dispersing wild
-dogs presented in \cite{Hofmann.2021}. In the original model, no interactions
-between the habitat and movement covariates were considered, so we slightly
-expanded this base model by proposing interactions between movement and habitat
-covariates. More specifically, we started with the base model and incrementally
-increased model complexity by adding all possible two-way interactions between
-habitat covariates and movement covariates. For instance, for the covariate
-\textsf{water}, we proposed the interactions \textsf{Water:log(sl)},
-\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides those combinations,
-we also proposed the interactions \textsf{sl:cos(ta)} and
-\textsf{log(sl):cos(ta)} to account for a correlation between turning angles and
-step lengths, as well as the interactions \textsf{sl:LowActivity} and
-\textsf{log(sl):LowActivity} to account for the fact that step lengths may
-differ due to wild dogs' diurnal activity pattern. To compare competing models
-and assess the most parsimonious movement model, we ran stepwise forward model
-selection based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
+analysis in the r-package {\tt glmmTMB}. To handle multiple individuals, we
+applied the mixed effects technnique developed by \citep{Muff.2020}, which
+allows to effectively model random slopes. Thus, we treated animal IDs as random
+effect and modeled random slopes for each covariate. We fixed the random
+intercept variance at an arbitrary high value of 10\textsuperscript{6} to make
+use of the  ``poission''-trick \citep{Muff.2020}.
+
+The formula for the movement model was based on the habitat selection model for
+dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model, no
+interactions between habitat covariates (\textsf{Water,
+DistanceToWater\textsuperscript{0.5}, Woodland, Shrubs/Grassland, Human
+Influence}) and movement covariates (\textsf{sl, log(sl), cos(ta)}) were
+considered. Here, we slightly expanded this base model and proposed interactions
+between all movement and habitat covariates. More specifically, we started with
+the base model and incrementally increased model complexity by adding all
+possible two-way interactions between habitat covariates and movement
+covariates. For instance, for the covariate \textsf{Water}, we proposed the
+interactions \textsf{Water:log(sl)}, \textsf{Water:log(sl)}, and
+\textsf{Water:cos(ta)}. Besides those combinations, we also proposed the
+interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)} to account for a
+correlation between turning angles and step lengths, as well as the interactions
+\textsf{sl:LowActivity} and \textsf{log(sl):LowActivity} to account for the fact
+that step lengths may differ due to wild dogs' diurnal activity pattern. To
+compare competing models and assess the most parsimonious movement model, we ran
+stepwise forward model selection based on Akaike's Information Criterion (AIC,
+\citealp{Burnham.2002}).
 
 We validated the predictive power of the most parsimonious movement model using
-k-fold cross-validation for case-control studies as presented
-\cite{Fortin.2009}. We randomly split the input data into training and testing
-data with an 80:20 ratio. Using the training data we parametrized a movement
-model based on which we predicted selection scores \(w(x)\) for all steps in the
-test data. Within each stratum we then assigned ranks 1-25 to each step based on
-predicted selection scores such that rank 1 was given to the step with the
-highest score \(w(x)\). Across all strata we determined the realized step's rank
-and calculated rank frequencies for realized steps. Finally, we computed
-Spearman's rank correlation between ranks and associated frequencies \(r_{s,
-realized}\). We replicated the entire procedure 100 times and computed the mean
-correlation coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\%
-confidence interval across all replicates. For comparison, we repeated the same
-procedure 100 times assuming random preferences. Random preferences were
-implemented by discarding the realized step from all strata and identifying the
-rank of a random step in each stratum. Again, we calculated Spearman's rank
-correlation coefficient (\(r_{s, random}\)), its mean across repetitions
-(\(\bar{r}_{s, random}\)), and its 95\% confidence interval. This validation
-proofs a significant prediction in case the confidence intervals of
-\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap.
+k-fold cross-validation for case-control studies as suggested by
+\cite{Fortin.2009}. For this, we randomly assigned 80\% of the strata to a
+training set and the remaining 20\% to a testing set. Using the training data we
+parametrized a movement model based on which we predicted selection scores
+\(w(x)\) for all steps in the test data. Within each stratum, we then assigned
+ranks 1-25 to each step based on predicted selection scores, where rank 1 was
+given to the step with the highest score \(w(x)\). Across all strata we
+determined the realized step's rank and we calculated rank frequencies of
+realized steps across all strata. Finally, we computed Spearman's rank
+correlation between ranks and associated frequencies \(r_{s, realized}\). We
+replicated the entire procedure 100 times and computed the mean correlation
+coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\% confidence interval
+across all replicates. For comparison, we repeated the same procedure 100 times
+assuming random preferences, which we implemented by discarding the realized
+step from all strata and identifying the rank of a random step in each stratum.
+Again, we calculated Spearman's rank correlation coefficient (\(r_{s,
+random}\)), its mean across repetitions (\(\bar{r}_{s, random}\)), and its 95\%
+confidence interval. This validation ultimately proofs a significant prediction
+in case the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
+random}\) do not overlap.
 
 \subsection{Dispersal Simulation (80\%)}
-We used the most parsimonious movement model to simulate 80'000 dispersing wild
-dogs moving across the KAZA-TFCA. The simulation resembled an inverted iSSF and
-was set up as follows. (1) We defined a random source point and assumed a random
-initial orientation of the animal. (2) Departing from the source point, we
-generated 25 random steps by sampling relative turning angles from a uniform
+We used the most parsimonious movement model to simulate 80'000 virtual
+dispersers moving across the KAZA-TFCA. The simulation resembled an inverted
+ISSF and was set up as follows. (1) We defined a random source point and assumed
+a random initial orientation of the animal. (2) Departing from the source point,
+we generated 25 random steps by sampling turning angles from a uniform
 distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
-distribution. Each step corresponded to the straight line movement conducted
-within 4 hours. To prevent unreasonably large steps, we capped sampled values to
-a maximum of 35 km, which corresponds to the farthest distance ever traveled
-within 4 hours by on of our dispersers. (3) Along each random step, we extracted
-and averaged habitat covariates and calculated movement covariates. Extracted
-values were again standardized using the same transformations that were used on
-the input data. (4) We applied the parameterized movement model and predicted
-the selection score \(w(x)\) for each random step and translated selection
-scores into probabilities using \Cref{EQ2}. (5) We randomly sampled one of the
-steps based on predicted probabilities and determined the animal's new position.
-We repeated steps (2) to (5) until a total of 2'000 steps were realized.
+distribution. Similar to the input data, each random step represented the
+straight line movement within 4 hours. To prevent unreasonably large steps, we
+capped sampled step lengths to a maximum of 35 km, which corresponded to the
+farthest distance ever traveled within 4 hours by one of our dispersers. (3)
+Along each random step, we extracted and averaged habitat covariates and we
+calculated movement covariates. To ensure compatible scales, we standardized
+extracted values using the same parameters applied to our input data. (4) We
+applied the parameterized movement model to predict the selection score
+\(w(x)\), which we translated into probabilities using \Cref{EQ2}. (5) We
+sampled one of the random steps based on predicted probabilities and determined
+the animal's new position. We repeated steps (2) to (5) until 2'000 steps were
+realized, implying a total 160 Mio. simulated steps.
 
@@
 the study area, we followed \citep{Koen.2010} and artificially expanded all
-covariate layers by adding a 100 km buffer zone. Within the buffer zone, we
+covariate layers by adding a 100 km wide buffer zone. Within the buffer zone, we
 randomized covariate values by resampling values from the original covariate
-layers. In result, simulated dispersers were able to leave and reenter the main
-study area through this buffer zone. In cases where proposed random steps
-transgressed the outer border of this buffer zone, we resampled transgressing
-steps until they fully lied within the buffer.
+layers. Through this buffer zone, simulated dispersers were able to leave and
+re-enter the main study area. In cases where proposed random steps transgressed
+the border of this buffer zone, we resampled transgressing steps until they
+fully lied within the buffer, thereby forcing simulated individuals to ``bounce
+off'' from such invisible borders.
 
 \subsection{Source Points (90\%)}
-We initiated virtual dispersers at 50'000 randomly selected source points within
-contiguous protected areas larger \(>\) 700 km\textsuperscript{2}
-(\Cref{SourcePoints}a). This conforms to the average home range requirement of
-resident wild dogs \citep{Pomilia.2015} and allowed us to remove areas too small
-to host viable wild dog populations. By distributing source points randomly, the
-number of source points per km\textsuperscript{2} within protected areas was
-approximately equal across the study area. To render potential immigrants into
-the study system, we placed additional 30'000 source points inside the buffer
-zone around the main study area (\Cref{SourcePoints}b). This resulted in a total
-of 80'000 source points, each representing the start point of an individual
-disperser.
+We released 80'000 virtual dispersers from 80'000 unique source points
+distributed across the study area. 50'000 virtual dispersers were released from
+randomly selected source points within contiguous protected areas larger \(>\)
+700 km\textsuperscript{2} (\Cref{SourcePoints}a), which conforms to average home
+range requirements of resident wild dogs \citep{Pomilia.2015} and allowed us to
+remove patches too small to host viable populations. By distributing source
+points randomly, the number of source points per km\textsuperscript{2} was
+approximately equal within protected areas. To render potential immigrants into
+the study system, we released another 30'000 dispersers at random locations
+inside the 100 km wide buffer zone surrounding the main study area
+(\Cref{SourcePoints}b).
 
@@
 trajectories and tallied them into a single map. If the same trajectory crossed
-a raster-cell twice, it was only counted once. This way, we did not consider
-revisits and mitigated biases arising from trapped individuals that were moving
-in circles. To achieve high performance rasterization, we used the R-package
-{\tt terra} \citep{Hijmans.2020}.
+a raster-cell twice, it was only counted once. This way, we mitigated potential
+biases resulting from trapped individuals that were moving in circles. To
+achieve high performance rasterization, we used the R-package {\tt terra}
+\citep{Hijmans.2020}.
 
@@
 betweenness scores. For this, we overlaid the study area (including the buffer)
-with a regular raster resolved at 5 x 5 km. We then used the simulated
-trajectories to determine all transitions occurring from one raster-cell to
-another, as well as the frequency at which those transitions occurred. This
-resulted in an edge-list that we translated into a weighted network using the
-r-package {\tt igraph} \citep{Gabor.2006}. Because {\tt igraph} handles edge
-weights (\(\omega\)) as costs, we inverted the traversal frequency in each cell
-by applying \(\omega = \frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal
-Frequency_i}\). Finally, we used the weighted network to calculate betweenness
-scores for each raster-cell in the overlaid raster. The betweenness metric
-indicates how often a specific raster-cell lies on a shortest path between two
-other raster-cells and is therefore a useful metric to detect movement corridors
-\citep{BastilleRousseau.2018}.
-
-\subsection{Interpatch-Connectivity (80\%)}
-We also assessed inter-patch connectivity between national parks in our study
-area. The decision to focus on national parks was purely out of simplicity and
-the same logic could easily be expanded to include other protected areas as
-well. To quantify inter-patch connectivity, we computed the relative frequency
-at which dispersers originating from one national park successfully moved into
-another national park. This allowed us to determine \textit{if} and \textit{how
-often} dispersers moved between certain national parks. Moreover, because time
-was explicit in our IBMM, we were able to estimate \textit{how long} dispersers
-had to move to realize those connections.
+with a regular raster resolved at 5 x 5 km. The centerpoint of each raster-cell
+acted as node in the final network and we used the simulated trajectories to
+determine all transitions occurring from one raster-cell to another, as well as
+the frequency at which those transitions occurred. This resulted in an edge-list
+that we translated into a weighted network using the r-package {\tt igraph}
+\citep{Gabor.2006}. Because {\tt igraph} handles edge weights (\(\omega\)) as
+costs, we inverted the traversal frequency in each cell by applying \(\omega =
+\frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal Frequency_i}\). Finally, we
+used the weighted network to calculate the betweenness score of each
+raster-cell. Betweenness measures how often a specific raster-cell lies on a
+shortest path between two other raster-cells and is therefore a useful metric to
+detect movement corridors \citep{BastilleRousseau.2018}.
+
+\subsection{Inter-Patch Connectivity (80\%)}
+To exemplify how simulations can be used to provide a network view on
+connectivity, we assessed inter-patch connectivity between national parks
+located in our study area. The decision to focus on national parks was purely
+out of simplicity and does not imply that connections between other regions are
+impossible. In fact, the same logic could easily be expanded to include other
+protected areas as well. To quantify inter-patch connectivity, we computed the
+relative frequency at which dispersers originating from one national park
+successfully moved into another national park. Successful movement was said to
+be achieved if the individuals' trajectory ever intersected with the
+corresponding national park. We also recorded the step-number of the first step
+that ever intersected with the national park's polygon. This allowed us to
+determine \textit{if} and \textit{how often} dispersers moved between certain
+national parks, as well as to estimate \textit{how long} dispersers had to move
+to realize those connections.
 
 \section{Results}
-\subsection{Movement Model (20\%)}
+\subsection{Movement Model (60\%)}
 Compared to the base model reported in \citep{Hofmann.2021}, our most
-parsimonious movement model included several additional interactions between
+parsimonious movement model contained several additional interactions between
 habitat covariates and movement covariates (\Cref{MovementModel} and
@@
 parsimonious model. Since all models with positive AIC weight contained similar
-covariates, this decision only marginally influenced subsequent results. Results
-from the selected model are given in \Cref{MovementModelNumbers} and illustrated
-in \Cref{MovementModel} (a). Additional plots that facilitate the interpretation
-of the model are provided in Appendix S2.
-
-When looking at the habitat kernel and holding constant the movement kernel, we
-find that dispersing wild dogs avoid water but prefer its proximity. Dispersers
-also avoid densely forested woodlands, yet prefer open shrublands or grasslands.
-Finally, dispersers avoid moving through landscapes that are influenced by
-humans. These results align with our previous findings reported in
-\cite{Hofmann.2021}.
-
-When looking at the movement kernel, we observe several significant estimates.
+covariates (Table S1), this decision only marginally influenced subsequent
+analyses. Plots that aid with the interpretation of the final model are provided
+in Appendix S2.
+
+When looking at the habitat kernel (and holding constant the movement kernel),
+we find that dispersing wild dogs avoid water but prefer its proximity.
+Dispersers also avoid woodlands, yet prefer shrublands or grasslands. Finally,
+dispersers avoid moving through landscapes that are influenced by humans. These
+results align with our previous findings reported in \cite{Hofmann.2021}.
+
+Results from the movement kernel indicate several significant estimates.
 However, except for the interaction \textsf{sl:LowActivity}, effect sizes are
-relatively small, suggesting that our proposal distributions for step lengths
-and turning angles were only marginally biased. For instance, the positive and
-significant effect for \textsf{cos(ta)} indicates that realized turning angles
-were slightly more directional than the turning angles proposed by our uniform
-distribution, implying that realized steps in fact followed a von Mises
-Distribution with positive concentration parameter. On the other hand, the
-significant and negative interaction for \textsf{sl:LowActivity} reveals that
-wild dogs moved over shorter distances during low wild dog activity compared to
-the steps suggested by our gamma distribution.
-
-Finally, we turn to the interactions between the two kernels. Here, we observe
-that movement behavior significantly differs depending on habitat covariates. In
-general, steps tend to be shorter in areas with high water cover and high
-woodland cover, yet larger in areas with open shrubs and grasslands. Similarly,
-dispersers realized shorter steps in the vicinity to water, even though the size
-of this effect is negligible. Finally, it appears that dispersers move more
-tortuous in areas influenced by humans, and more directional when far from
-water.
-
-As can be seen in \Cref{MovementModel}b, the k-fold cross-validation procedure
-reveals that our model substantially outperforms a random guess, as indicated by
-the fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
-\(\bar{r}_{s, random}\) do not overlap. Furthermore, the significant correlation
+relatively small, indicating that our proposal distributions for step lengths
+and turning angles were only marginally biased by habitat selection. For
+instance, the positive and significant estimate for \textsf{cos(ta)} indicates
+that realized turning angles were slightly more directional than the turning
+angles proposed by our uniform distribution. This shows that realized steps in
+reality followed a von Mises Distribution with positive concentration parameter.
+On the other hand, the significant and negative interaction for
+\textsf{sl:LowActivity} reveals that wild dogs realized shorter steps during low
+wild dog activity compared to the steps suggested by our gamma distribution.
+
+When looking at the interactions between the movement and habitat kernel, we
+observe several significant interactions, highlighting that movement behavior
+differs depending on habitat conditions. For example, there's ample evidence
+that the propensity to realize a log step in areas covered by water is
+substantially smaller compared to the propensity of realizing a small step.
+Similarly, long steps are less likely to be realized in areas covered by trees,
+compared to more open regions. Conversely, it appears that steps tend to longer
+be in the vicinity to water, even though the size of this effect is negligible.
+Finally, it seems that dispersers move more tortuous in areas influenced by
+humans, but more directional when far from water.
+
+The k-fold cross-validation procedure reveals that our model substantially
+outperforms a random guess (\Cref{MovementModel}b). Confidence intervals of
+\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap and
+therefore proof a reliable prediction. Furthermore, the significant correlation
 between ranks and corresponding frequencies for realized steps indicates a good
-fit between predictions and observations (\Cref{MovementModel}b). Finally, we
-find that the rank correlation slightly improved in comparison to the base model
-(\(\bar{r}_{s, realized} = -0.55\); \citealp{Hofmann.2021}).
+fit between predictions and observations (\Cref{MovementModel}b). In comparison
+to the base model (\(\bar{r}_{s, realized} = -0.55\); \citealp{Hofmann.2021}),
+the inclusion of interactions between movement and habitat covariates slightly
+improved model performance.
 
@@
   \caption{Most parsimonious movement model for dispersing wild dogs. The model
-  comprises of a movement and habitat kernel, where the movement kernels
-  describes preferences with regards to movement behavior, whereas the habitat
-  kernel describes preferences with respect to the habitat. Finally, the two
-  kernels can interact, such that movement preferences are contingent on habitat
-  conditions.}
+  consists of a movement kernel, a habitat kernel, and their interactions. The
+  movement kernel describes preferences with regards to movement behavior,
+  whereas the habitat kernel describes preferences with respect to habitat
+  conditions. Interactions between the two kernels indicate that movement
+  preferences are contingent on habitat conditions. Note that all covariates
+  were standardized to a mean of zero and standard deviation of 1. Plots to aid
+  with the interpretation of this model are given in Appendix S2.}
   \label{MovementModelNumbers}
@@
 
-\subsection{Dispersal Simulation (60\%)}
-On a machine with an AMD Ryzen 7 2700X processor with 8 x 3.6 GHz and 64 GB of
-RAM, a single batch of 1'000 simulated dispersers took roughly 90 minutes to
-compute (\(\mu = 88.90\), \(\sigma = 1.87\)). As such, the simulation of all
-80'000 dispersers terminated after 120 hours, i.e. five days. Comparable
-computations will be substantially quicker for smaller study areas or lower
-resolution covariates, as the covariate extraction from large rasters was
-computationally the most expensive task.
+\subsection{Dispersal Simulation (80\%)}
+On a machine with an octacore AMD Ryzen 7 2700X processor (8 x 3.6 GHz) and 64
+GB of RAM, a single batch of 1'000 simulated dispersers took roughly 90 minutes
+to compute (\(\mu = 88.90\), \(\sigma = 1.87\)) and the simulation of all 80'000
+dispersers terminated after 120 hours, i.e. five days. Comparable computations
+will be substantially faster for smaller study areas or lower resolution
+covariates, as the covariate extraction from large rasters was computationally
+the most expensive task.
 
@@
 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
-dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)). Simultaneously,
-simulated dispersers moved marginally less directional (\(\mu_{cos(ta)} =
-0.057\), \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers
-(\(\mu_{cos(ta)} = 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in
-step lengths and turning angles can be attributed to minor disparities between
-habitat conditions at the area within which we collected training data and
-habitat conditions within the entire study area.
-
-Out of the 50'000 dispersers initiated in protected areas, only 4.5\% eventually
-hit a map boundary, suggesting that biases due to boundary effects should be
-minimal. In contrast, 78\% of the 30'000 dispersers originating from the buffer
-zone eventually hit a map boundary.
-
-\subsection{Heatmap (50\%)}
+dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
+dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
+\(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
+0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
+turning angles can be attributed to minor disparities between habitat conditions
+at the area within which we collected training data and habitat conditions
+within the entire study area. Out of the 50'000 dispersers initiated in
+protected areas, only 4.5\% eventually hit a map boundary, suggesting that
+biases due to boundary effects should be limited. In contrast, 78\% of the
+30'000 dispersers originating from the buffer zone eventually hit a map
+boundary.
+
+\subsection{Heatmap (80\%)}
 \Cref{Heatmap} depicts the heatmap of all 80'000 simulated trajectories rendered
@@
 borders of the KAZA-TFCA are only infrequently visited by dispersers (dark blue
-areas), whereas within the KAZA-TFCA we observe several extensive regions that
-are regularly visited (bright yellow and red areas). Most notably, the region in
-northern Botswana south of the Linyanti swamp stands out as highly frequented
-dispersal hub. Still, the presence of several massive water bodies, such as the
-Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in
-considerable dispersal barriers that limit realized connectivity within the
-KAZA-TFCA. Similarly, dispersal appears to be very limited in Zambia's and
-Zimbabwe's part of the KAZA-TFCA, where only few areas are traversed by
-dispersers, which can largely be attributed to a high human influence in these
-regions. Outside the KAZA-TFCA, the most frequently traversed regions include
-the areas of the Central Kalahari National Park in Botswana, the area south-west
-of the Khaudum National Park in Namibia, and the area around the Liuwa Plains
-National Park in Zambia.
+areas), whereas within the KAZA-TFCA several extensive regions are regularly
+visited (bright yellow and red areas). Most notably, the region in northern
+Botswana south of the Linyanti swamp stands out as highly frequented dispersal
+hub. Still, the presence of several massive water bodies, such as the Okavango
+Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in considerable
+dispersal barriers that limit realized connectivity within the KAZA-TFCA.
+Similarly, dispersal appears to be very limited in Zambia's and Zimbabwe's part
+of the KAZA-TFCA, where only few areas are regularly traversed by dispersers.
+This which can largely be attributed to the strong high human influences in
+these regions. Outside the KAZA-TFCA, the most heavily trafficked regions
+include the areas of the Central Kalahari National Park in Botswana, the area
+south-west of the Khaudum National Park in Namibia, and the area around the
+Liuwa Plains National Park in Zambia.
 
@@
   \includegraphics[width=\textwidth]{99_Heatmap.png}
-  \caption{Heatmap of all 80'000 simulated trajectories after 2'000 steps.
-  Simulations were based on an integrated step selection model, fit to the
-  movement data of dispersing African wild dogs. To generate the heatmap, we
-  rasterized and summed all simulated trajectories. Additional heatmaps
-  showcasing the traversal frequency after different numbers of steps are
-  provided in xx in Appendix S3.}
+  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
+  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
+  integrated step selection model that we fit to the movement data of dispersing
+  African wild dogs. To generate the heatmap, we rasterized and tallied all
+  simulated trajectories. Consequently, the map highlights areas that are
+  frequently traversed by virtual dispersers. Further heatmaps showing the
+  traversal frequency after different numbers of steps are provided in Appendix
+  S3.}
   \label{Heatmap}
@@
 
-\subsection{Betweenness (50\%)}
+\subsection{Betweenness (80\%)}
 Betweenness scores of each raster-cell the study area are presented in
-\Cref{Betweenness}. While the heatmap emphasized frequently visited areas, the
-map of betweenness reveals rather discrete dispersal corridors. As can be seen,
-the dispersal hotspot in northern Botswana is traversed by a corridor that
-receives a comparably high betweenness score. This implies that the region is
-particularly crucial for connecting other regions in the study system and hence
-represents a proper network hub. Towards east, the corridor runs through the
-Chobe National Park into the Hwange national park, where it branches out and
-further extends into the distant Matusadona National Park in Zimbabwe. Northwest
-of the Linyanty ecosystem, the same corridor expands into Angola, where it
-splits and finally traverses over a long stretch of unprotected area into the
-Kafue National Park in Zambia. Several additional corridors with slightly lower
-betweenness scores exist, yet most of them run within the boundaries of the
-KAZA-TFCA. In general, only few corridors directly link the peripheral regions
-of the KAZA-TFCA. For instance, there are only few corridors between the
-Matusadona National Park in Zimbabwe and the Kafue National Park in Zimbabwe.
-Similarly, there are no direct links between the Zimbabwean and Angolan
-``spikes'' of the KAZA-TFCA.
+\Cref{Betweenness} and reveals a set of discrete dispersal corridors. As can be
+seen, the dispersal hotspot in northern Botswana reported above is crossed by
+a corridor that receives a relativley high betweenness score. This implies that
+the region is particularly crucial for connecting other regions in the study
+system and hence represents a proper hub in the  ``network'' . Towards east, the
+corridor runs through the Chobe National Park into the Hwange national park,
+where it branches out and further extends into the distant Matusadona National
+Park in Zimbabwe. Northwest of the Linyanty ecosystem, the same corridor expands
+into Angola, where it splits and finally traverses over a long stretch of
+unprotected area into the Kafue National Park in Zambia. Several additional
+corridors with slightly lower betweenness scores exist, yet most of them run
+within the boundaries of the KAZA-TFCA. In general, only few corridors directly
+link the peripheral regions of the KAZA-TFCA. For instance, there are only few
+corridors between the Matusadona National Park in Zimbabwe and the Kafue
+National Park in Zimbabwe. Similarly, there are no direct links between the
+Zimbabwean and Angolan ``spikes'' of the KAZA-TFCA.
 
@@
 
-\subsection{Interpatch Connectivity (0\%)}
-The results of the analysis of interpatch connectivity is given in
-\Cref{AreasReached}. The map shows between which national parks direct links
-exist and how frequent they are, as well as the average duration a disperser had
-to move to realize those links. For instance, 6.8 \% of the simulated dispersers
-originating from the Moremi National Park successfully reached the Chobe
-National Park and 4.2 \% reached the Hwange National Park in Zimbabwe. On
-average, dispersers moved for 623 steps before arriving at Chobe (SD = 520) and
-for 1'413 steps before arriving at Hwange (SD = 371).
+\subsection{Interpatch Connectivity (50\%)}
+Results from the analysis of interpatch connectivity are given in
+\Cref{AreasReached}. It is worth pointing out that the figure is only intended
+as an example; for clarity we limited the network on national parks, albeit
+plenty of links to other protected areas exist. The map shows all realized links
+by simulated dispersers between national parks and indicates the average
+duration a disperser had to move to realize those links. For instance, 6.8 \% of
+the simulated dispersers originating from the Moremi National Park successfully
+reached the Chobe National Park and 4.2 \% the Hwange National Park in Zimbabwe.
+On average, dispersers moved for 623 steps before arriving at Chobe (SD = 520)
+and for 1'413 steps before arriving at Hwange (SD = 371). While the map only
+depicts connections between national parks, this does not imply that no
+connections to other protected areas exist and is the result of our simplifying
+assumptions above. Interestingly, while we identified a potential dispersal
+corridor between Angola's NPs and the Kafue NP in Zambia, \Cref{AreasReached}
+suggests that this link is only rarely realized and requires very long dispersal
+events. In contrast, we find that dispersal between the Moremi NP and Chobe NP
+are relatively frequent and require fewer steps, which can be expected given
+that the areas are located close to each other.
 
@@
   \includegraphics[width=\textwidth]{99_AreasReached.png}
-  \caption{Network on simulated dispersal trajectories highlighting the
-  connections between national parks (dark green). Yellow bubbles represent the
-  center of the different national parks and are sized in relation to the number
-  of simulated dispersers originating from each park. Colored arrows between
-  national parks illustrate between which national parks the simulated
-  dispersers successfully moved and the color of each arrow shows the average
-  number of steps that were necessary to realize those connections.
-  Additionally, the line thickness indicates the relative number of dispersers
-  originating from a national park that realized those connections.}
+  \caption{Network on simulated dispersal trajectories highlighting connections
+  between national parks (dark green). Yellow bubbles represent the center of
+  the different national parks and are sized in relation to the number of
+  simulated dispersers originating from each park. Black dots represent national
+  parks that were smaller than 700 km\textsuperscript{2} and therefore did not
+  serve as source areas. Arrows between national parks illustrate between which
+  national parks the simulated dispersers successfully moved and the color of
+  each arrow shows the average number of steps (4-hourly movements) that were
+  necessary to realize those connections. Additionally, the line thickness
+  indicates the relative number of dispersers originating from a national park
+  that realized those connections. Note that a similar network view could be
+  adopted to investigate connectivity between other protected areas need not to
+  be restricted to national parks.}
   \label{AreasReached}
@@
 
-\section{Discussion (10\%)}
-Our connectivity network further suggests that dispersers from the Okavango
-Delta more likely disperse towards east than west. Indeed, only x out of our y
-observed dispersers ever reached the western part of the delta. Only when the
-flood retracts a small pathway between the city of Maun and the floodwaters of
-the delta emerges and enables dispersers to move towards the detal's western
-part.
-
-All of our findings are well in line with our previous work, where we have
-highlighted dispersal corridors for wild dogs using least-cost approaches. This
-suggests that qualitative results are quite insensitive to the exact
-methodological approach. Still, we believe that a simulation-based approach
-offers possibilities for much richer inferences compared to traditional
-approaches. This is largely due to the fact that proper movement trajectories
-are generated that can be analysed \textit{as if} they were generated by real
-dispersers. This is of course contingent on the assumption that underlying
-models are adequately representing movement behavior of the focal species and
-calls for further methods to validate the predictive power of such simulations.
+\section{Discussion}
+\subsection{Short Summary (90\%)}
+To this end, we used integrated step selection functions to analyse data of
+dispersing wild dogs and parametrize a fully mechanistic movement model
+describing how dispersers move through the available landscape. We employed the
+parametrized model to simulate 80'000 dispersing wild dogs, moving across the
+extent of the KAZA-TFCA, the world's largest transboundary conservation area.
+Based on simulated dispersal trajectories, we prepared a set of complementary
+maps, all geared towards a better understanding of dispersal and landscape
+connectivity. The set of maps included a heatmap, revealing frequently traversed
+areas, a betweenness-map, delineating critical dispersal corridors connecting
+different regions, and a map of inter-patch connectivity, indicating presence or
+absence of functional links between national parks, as well as the average
+dispersal duration required realize those links. With this, we showcase that
+integrated step selection analysis holds the potential to serve as simple, yet
+powerful mechanistic movement model to simulate animal movement and assess
+landscape connectivity. Importantly, such an individual-based approach overcomes
+several shortcomings of traditional connectivity models, such as least-cost
+analyses and circuit theory. Nevertheless, we support the idea that simulations
+only serve as a complement, and not as substitute to more traditional
+connectivity modeling techniques.
+
+\subsection{Movement Model (50 \%)}
+Our most parsimonious movement model comprised of a habitat kernel and a
+movement kernel that describe in detail how dispersers move through the
+available landscape. Parameter estimates revealed that dispersing wild dogs
+avoid water, prefer proximity to water, avoid woodland, prefer shrubs/grassland,
+and avoid areas dominated by humans. This is in line with an earlier dispersal
+model, were we primarily investigated the habitat kernel \cite{Hofmann.2021},
+and suggests that the introduction of additional covariates only marginally
+impacted qualitative results. Even though corresponding effect sizes were
+moderate, the model suggested that movement behavior of dispersers differs
+depending on habitat characteristics. For instance, step lengths appear to be
+shorter (i.e. animals move slower) in areas covered by water in comparison to
+steps realized on dryland. We found similar interactions between movement
+covariates and other habitat covariates, highlighting how a simple ISSF can be
+employed to model relatively complex dispersal patterns. Correspondingly, the
+ISSF framework allowed us to generically model resting behavior by including a
+binary variable, signalling that dispersers moved were substantially slower
+between 17:00 and 09:00 o'clock compared to the rest of the day.
+
+While this (shorter steps on water) could be caused by a bias in our data,
+where the straight line distance between two GPS relocations erroneously
+suggests movement through water, where in reality the animal circumvented
+water-bodies, we hypothesize that this slow-down is caused by dispersers that
+move through the KAZA-TFCA's floodlands and get trapped by the flood. In fact,
+we repeatedly observed individuals venturing across the Okavango Delta,
+suddenly getting surrounded by the advancing flood. Until the flood retracts,
+these individuals often remain on distinct ``islands'' on which movement is
+restricted to rather small steps. This also aligns with the finding that
+turning angles tend to be larger when individuals are close to water (or
+conversely, steps are larger when far from water). It seems that trapped
+individuals wander around, searching for a passage to escape through the
+indundated floodlands, unable to move in a straight-line manner. Quite
+similarly, our model revealed that steps are marginally shorter on woodland,
+but larger on shrubs/grassland, suggesting that wild dogs slow down when
+moving through areas densely covered by trees, but speed up when moving over
+shrubs/grassland. This can probably be attributed to resting periods, during
+which dispersing individuals seek protection in the shade. To effectively
+model this behavior, however, one would need to include a three-way
+interaction between the movement kernel, habitat kernel and a temporal
+measure. Instead, we generically modeled resting behavior using a binary
+variable, which, signaled that steps were substantially shorter between 17:00
+and 07:00 o'clock compared to the rest of the day. This result is rather
+unsurprising as it is well known that wild dogs' follow a diurnal activity
+pattern.
+
+\subsection{Simulation (10\%)}
+By treating the parametrized ISSF model as a fully mechanistic movement model,
+we were able to simulate 80'000 dispersers originating from large protected
+areas. Each individual was simulated over 2'000 steps, which resulted in a total
+of 160 Mio. simulated steps. Although the simulation only terminated after 6
+days, this was largely owed to the massive extent considered (ca. 1.4 Mio. km
+\textsuperscript{2}). Moreover, a smaller number simulated dispersers will often
+suffice, as the relative traversal frequency across the study area in our case
+converged rather quickly (Appendices).
+
+\subsection{Maps (10\%)}
+Our heatmap suggests that a large number of dispersers traverses the regions of
+the Moremi NP and the Chobe NP. from the Okavango Delta more likely disperse
+towards east than west. Indeed, only x out of our y observed dispersers ever
+reached the western part of the delta. Only when the flood retracts a small
+pathway between the city of Maun and the floodwaters of the delta emerges and
+enables dispersers to move towards the detal's western part.
 
@@
 that the betweenness metric highlights crucial bottlenecks that are relevant for
-connecting remote regions, yet does not directly yield information about the
-absolute frequency at which these bottlenecks are used.
+connecting remote regions, it does not directly yield information about the
+frequency at which these bottlenecks are used.
+
+\subsection{General (20\%)}
+A major benefit of the simulation-based approached is that the domain of
+endpoints does not need to be determined \textit{a priori}. Instead, each
+endpoint emerges naturally as the result of a simulated dispersal trajectory.
+This is particularly important when modeling dispersers, mainly because
+dispersing individuals venture into unknown territory and do not necessarily
+move towards a predetermined endpoint. Nevertheless, in cases were individuals
+do exhibit such preferences, the incorporation of attraction points in the
+corresponding ISSF movement model allows to render them.
+
+While endpoints emerge naturally from the dispersal process, startpoints still
+need to be specified \textit{a priori} by the modeler. Here, we placed source
+points within protected areas large enough to sustain viable wild dog
+populations. This was done under the simplifying assumption that wild dogs only
+survive in formally protected areas, which is in line with scientific findings
+\citep{Woodroffe.1999, Woodroffe.2012, VanDerMeer.2014, Loveridge.2019}. In some
+cases, exact locations of source populations are known and source points can
+easily be placed accordingly \citep{Kanagaraj.2013}. In other cases, comparable
+knowledge may be lacking and it may be more appropriate to delineate likely
+source patches using habitat suitability models \citep{Squires.2013}. Because we
+did not investigate the sensitivity of our results with respect to the exact
+location of source points, this is something that needs further investigation in
+the future.
+
+Another advantage of simulation-based approaches is that they render time
+explicitly. This enables to answer questions such as: ``\textit{How long will it
+take a disperser to move from A to B?}'' or \textit{Is it possible for a
+disperser to move from A to B within X days?}. These are interesting questions
+and they shift the focus from a structural to a more functional point of view.
+However, an explicit representation of time requires that speed (step length)
+and directionality during motion is approapriately modeled
+\citep{Kanagaraj.2013}. Because ISSFs enable to model these two components
+adequately, the method offers a powerful framwork for simulations. Besides this,
+one also needs to decide on a meaningful dispersal duration when simulating
+movement. We decided to simulate individuals for 2'000 steps, which is at the
+upper end of observed dispersal durations and likely resulted in overestimated
+landscape connectivity. Nevertheless, it requires little tweaking to subset from
+the generated simulations to any dispersal duration desired. In fact, one could
+randomly sample dispersal durations based on observed dispersal events. In most
+cases, however, it will be more convenient and insightful to simulate relatively
+extensive dispersal events and only subsample afterwards.
+
+While we have assumed a set of static covariates when simulating dispersal, an
+explicit representation theoretically allows to render seasonality in covariate
+layers. This is an important aspect in ecosystems where seasonality
+substantially influences landscape connectivity. With least-cost analysis and
+circuit theory, seasonality can merely be incorporated by producing a multitude
+of permeability surfaces, each depicting landscape permeability in a different
+season, and then applying the connectivity models to those surfaces
+\citep{Benz.2016, Osipova.2019}. With individual-based simulations, on the other
+hand, seasonal covariates can be updated as the simulated dispersers move. As
+such, seasonality would directly influence movement, i.e. the process that
+ultimately leads to connectivity. For instance, in our simulation we represented
+the Okavango Delta statically and assumed a relatively extended flood. In this
+regard, the maps presented in the results section may be most representative of
+the period shortly after the wet-season, when floodlevels in the Delta are at
+their maximum. During the dry season, however, the flood considerably retracts
+and potentially clears the way for wild dogs dispersing from the Moremi-Game
+reserve into the south-western section of the Delta. Consequently, instead of
+using a static floodmap, one could render the flood dynamically. Hence, the
+floodlevels would be updated as the dispersers move, which would allow studying
+how connectivity evolves as the flood climaxes and retracts again.
+
+We simulated dispersal using point estimates from our most parsimonious movement
+model, yet the degree to which our results depend on those estimates is unknown.
+Given that data from dispersal studies on endangered species is scarce, point
+estimates may be quite inaccurate, therefore leading to erronous inference
+\citep{Wiegand.2003, KramerSchadt.2007}. Rather than using point estimates, an
+alternative may be to simulate dispersers using a set of randomized preferences
+imposed by the uncertainty reported in the model output. We urge future studies
+to further investigate investigate the sensitivity of ISSF simulations with
+respect to estimated habitat preferences.
 
@@
 uncertainty around the respective beta coefficients. To better gauge the
-imporance and influence of this covariate, future studies will need to control
+importance and influence of this covariate, future studies will need to control
 for inter- and intra-speficic interactions that may explain why and when
 dispersers are attracted to or afraid of waterbodies.
-
-For our simulations, we represented the Okavango Delta statically and assumed a
-relatively extended flood. This resulted in a quasi-barrier, formed by the
-Okavango-Delta and the adjacent city of Maun, which was rarely traversed by
-simulated individuals. Out of the 499 dispersers initiated inside the Moremi
-National Park, only 101 managed to reach the south-western section of the Delta,
-whereas 284 eventually reached the equally distant Linyanti swamp. In this
-regard, the heatmap presented in \Cref{Heatmap} may be most representative of
-the period shortly after the wet-season, when floodlevels in the Delta are at
-their maximum. During the dry season, however, the flood considerably retracts
-and potentially clears the way for wild dogs dispersing from the Moremi-Game
-reserve into the south-western section of the Delta. Future studies could relax
-the assumption of a static flood and attempt to update floodlevels as the
-dispersers move. This would allow studying how connectivity within the ecosystem
-evolves over time as the flood climaxes and retracts again. In fact, one of the
-major advantages of such simulation-based approaches is that a dynamic
-environment can be rendered, as time is explicit in these models
-\citep{Zeller.2020}. This contrasts with traditional modeling approaches such
-least-cost analysis or circuit theory, where the temporal dimension cannot be
-made explicit. An explicit view on time also directly translates in insights on
-the duration required by dispersers to move between distinct patches such as
-national parks or spatially segregated subpopulations.
 
@@
 
+An important benefit of ISSF simulations is that the framework always considers
+availability. That is, the propensity of a simulated disperser to realize a
+certain step is always contingent on the set of alternative steps. As such, a
+disperser surrounded by relatively unsuitable habitat will still move and
+disperse insteado f getting stuck.
+
+Even though connectivity is generally believed to promote population viability,
+it has also been pointed out that improved connectivity may cause ecological
+traps, especially when connectivity into or through human-dominated landscapes
+is promoted. In such instances, connectivity increases the risk of encountering
+humans and facilitates persecution by humans. By overlapping simulated
+trajectories with a map of human influence, such ecological traps could be
+pinpointed.
+
 We did not model mortality during dispersal in our simulations. This was a
-simplifying assumption and only inaccurately reflects dispersal in reality. It
-is well known that wild dogs often die during dispersal in result of deadly
-encounters with other predators, but also due to road kills and illegal
-shootings by humans. In result, mortality during dispersal may limit realized
-connectivity, especially in areas that wild dogs are unfamiliar with or in areas
-with high potential for human-wildlife conflict \citep{Cozzi.2020}.
-Interestingly though, \cite{Behr.2021} have recently demonstrated that mortality
-during dispersal is lower compared to residency, suggesting the presence some
-fitness benefits to dispersing individuals.
-
-In this regard, our approach is rather similar to dispersal kernels, yet it
-enables to render directional biases, which are currently difficult to implement
-using such methods.
-
-We completely randomized the location of source points within protected areas.
-However, in some cases prior knowledge about the density of potential dispersers
-is available and can be used to adjust the number of simulated individuals
-accordingly. Alternatively, instead of tweaking the number of simulated
-individuals, one could assign a weight to each trajectory that depends on the
-density of potential dispersers in the source areas. As such, trajectories from
-areas with high density would enter the heatmap with above average weight.
-
-The parametrized movement model could also be manipulated to investigate how
-different habitat preferences influence landscape connectivity and to test the
-sensitivity of results with respect to the exact preferences of individuals.
+simplifying assumption that only inaccurately reflects true dispersal. It is
+well known that wild dogs often die during dispersal in result of deadly
+encounters with competing predators, but also due to road kills or human
+persecution \citep{Woodroffe.2012, Behr.2021b}. In result, mortality during
+dispersal may limit realized connectivity, especially in areas that wild dogs
+are unfamiliar with and in areas with a high potential for human-wildlife
+conflict \citep{Cozzi.2020}.
+
+Due to a lack of information on wild dog densities across the extent of the
+KAZA-TFCA, we assumed and simulated an almost uniform distribution of dispersers
+across protected areas. In some cases, prior knowledge about the density of
+potential dispersers is available and can be used to adjust the number of
+simulated individuals accordingly. Instead of tweaking the number of simulated
+individuals, one could also weight each trajectory relative to the density of
+potential dispersers in the corresponding source areas, such that trajectories
+from areas with high density would enter the heatmap with a higher weight.
 
@@
 infeasible.
-
-Our work suggests that the selection of source points significantly impacts
-resulting connectivity networks. Especially when dispersal durations are short,
-wrongly placed source points lead to vastly different results. Signer et al.
-used estimated utilisation distributions by means of simulated movements. They
-used a rather long burn in period prior to alleviate the problem of selecting
-meaningful source points. However, this approach only works when individuals
-move around a point of attraciton. This is typically not the case when
-simulating dispersers, introducing an important trade-off. The researcher can
-decide to increase the number of simulated steps, hence reducing the influence
-of starting locations, yet this also inevitably increases estimated
-connectivity.
 
@@
 
-Future studies could also investigate the sensitivity of such simulations with
-respect to estimated habitat preferences. Here, we used point estimates and did
-not vary degree by which individuals preferred or avoided certain covariates.
-Nevertheless, the degree to which our results depend on these estimates is
-unknown. Because data are usually scarce for dispersal studies on endangered
-species, point estimates may be inaccurate \citep{Wiegand.2003,
-KramerSchadt.2007}.
+We do not dismiss the application of least-cost analysis or circuit theory in
+any way. Rather, we propose to use simulations from IBMMs complementary to
+least-cost analysis and circuit theory in order to provide a more comprehensive
+analysis of landscape connectivity. In fact, simulations from IBMMs could be
+employed to identify likely connections between habitat patches in a first place
+and only afterwards least-cost paths and corridors can be drawn between such
+likely connections. Thus far, dispersal kernels or thresholded least cost paths
+have been used for this.
+
+Each map that we produced from simulated trajectories accentuates a different
+aspect of connectivity. The heatmap, for example, puts emphasis on areas that
+are frequently traversed, even if such areas represent ``dead ends'' where
+dispersers get trapped and bounce off from obstacles. Hence, the map is useful
+to determine locations where a disperser is likely to show up, yet not
+necessarily to determine corridors that are worthwile to be protected. The
+betweenness map, on the other hand, brings out those areas that are relevant in
+connecting different regions in the landscape. In this regard, the betweenness
+map is most pertinent to the delineation of dispersal corridors and may serve as
+an alternative to least-cost algorithms. Finally, the inter-patch connectivity
+map illustrates the frequency at which dispersal between distinct habitat
+patches occurs, as well as the average dispersal duration that passes when
+individuals move between those patches. This knowledge serves to gauge the
+amount of geneflow between habitat patches and can be employed to determine
+whether two habitats are connected or not.
 

### 2021-06-24 Chapter_1/04_Manuscript/Manuscript.tex
various updates

@@
 contingent a sufficient degree of landscape connectivity, which is why the
-estimation of connectivity and identification of dispersal corridors has become
-a task of extraordinary importance for conservation authorities worldwide.
+estimation of connectivity and identification and preservation of dispersal
+corridors has become a task of extraordinary importance for conservation
+authorities worldwide.
 
@@
 simulations from individual-based movement models have been proposed and
-applied. Yet, due to the almost infinite amount of non-trivial decisions a
+applied. Yet, due to the almost infinite number of non-trivial decisions a
 modeler faces when parametrizing such models, a unified and objective framework
@@
 
-Here, we showcase the use of integrated step selection functions to simulate
-dispersal of the endangered African wild dog across the world's largest
-transboundary conservation area, the Kavango-Zambezi Transfrontier Conservation
-Area (KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
+Here, we showcase the use of integrated step selection functions as a simple
+individual-based and spatially explicit model to simulate dispersal of the
+endangered African wild dog across the world's largest transboundary
+conservation area, the Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
 coalitions in combination with relevant habitat covariates. We analyse the data
 using integrated step selection functions and parametrize a fully mechanistic
-movement model rendering wild dog dispersal. Based on this model, we simulate
-80'000 dispersers, originating from protected areas and moving across the extent
-of the KAZA-TFCA. We then generate a heatmap indicating regions frequently
-visited by dispersers, use network theory to reveal dispersal hotspots and
-crucial bottlenecks across the study area. Finally, we discuss the benefits and
-pitfalls of such dispersal simualtions and highlight potential improvements to
-be made in the future.
+model describing how dispersing wild dogs move through the landscape. Based on
+this model, we simulate 80'000 dispersers moving across the extent of the
+KAZA-TFCA. We then generate a set of maps, each focused on a different aspect of
+connectivity within the KAZA-TFCA. Finally, we discuss the benefits and pitfalls
+of such a simulation-based approach and highlight potential improvements to be
+made in the future.
 
@@
 location to the site of first reproduction \cite{Howard.1960}. It is a vital
-process governing the dynamics wild animal populations that are distributed in
-space \citep{Hanski.1998, Clobert.2012} and may strongly affect population
-dynamics at different spatial and social scales \citep{Hanski.1999a,
+process governing the social structure of wild animal populations that are
+distributed in space \citep{Hanski.1998, Clobert.2012} and may strongly affect
+population dynamics at different spatial and social scales \citep{Hanski.1999a,
 Clobert.2012}. Dispersal allows species to avoid inbreeding and maintain genetic
 diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
-Baguette.2013}, to rescue small and non-viable populations \citep{Brown.1977},
-and to promote the colonization of unoccupied habitats \citep{Hanski.1999b,
+Baguette.2013}, to rescue small, non-viable populations \citep{Brown.1977}, and
+to promote the colonization of unoccupied habitats \citep{Hanski.1999b,
 MacArthur.2001}. However, successful dispersal requires a sufficient degree of
-functional connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the
-identification and protection of major dispersal corridors has become an
-important task in conservation science \citep{Doerr.2011, Rudnick.2012}. In
-order to pinpoint relevant dispersal hotspots, reliable information on movement
-behavior during dispersal and knowledge about factors that limit dispersal and
-connectivity is paramount \citep{Baguette.2013, Vasudev.2015}.
+landscape connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the
+identification and protection of major dispersal corridors has become a
+fundamental task in conservation science \citep{Nathan.2008, Doerr.2011,
+Rudnick.2012}. Importantly, the ability to pinpoint relevant dispersal hotspots
+requires information on movement behavior during dispersal and knowledge about
+factors that limit dispersal and therefore connectivity \citep{Baguette.2013,
+Vasudev.2015}.
 
@@
 empirical data and an increased computational power have led to the development
-of numerous techniques to study dispersal movements and highlight critical
-corridors between subpopulations \citep{Boyce.2002, Fortin.2005, Cushman.2010,
+of numerous techniques to study dispersal and highlight critical corridors
+between subpopulations \citep{Boyce.2002, Fortin.2005, Cushman.2010,
 Zeller.2012, Diniz.2020}.
@@
 selection functions} \citep{Cushman.2010} have proven particularly useful for
-studying animal movement \citep{Fieberg.2020} and modelling connectivity
+studying animal movement \citep{Fieberg.2020} and modeling connectivity
 \citep{Diniz.2020}. These methods allow estimating habitat preferences of the
@@
 % Issues with Least-Cost Paths & Circuit Theory
-\subsection{Issues with Least-Cost Paths \& Circuit Theory (90\%)}
+\subsection{Issues with Least-Cost Paths (90\%)}
 In LCP analysis, for instance, a least costly path always exists, even if
@@
 dispersing individual. The method also presumes that animals have an infinite
-perceptual range, a preconceived end-point in mind, and choose a cost-minimizing
-route accordingly. These assumptions may be reasonable for migrating animals,
-yet they are unlikely to hold for dispersers, which typically move over long
-distances into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}.
-Finally, LCPs are only one pixel wide, meaning that their absolute size depends
-on the resolution of chosen covariate layers \citep{Diniz.2020}. Although some
-of these deficiencies can be addressed using less stringent versions of the LCP
-algorithm (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
-\textit{thresholded} least-cost paths \citep{Landguth.2012}, and
-\textit{randomized} least-cost paths \citep{Panzacchi.2016, VanMoorter.2021}), a
-certain degree of arbitrariness in the assumptions remains. CT entails similarly
-unreasonable restrictions. Because CT only considers movements from the source
-cell to its 4 or 8 adjacent cells, it implicitly posits a perceptual range of a
-single pixel. As such, the perceptual range cannot be defined based on
-biological observations, but is imposed by the resolution of the reference grid.
-CT is also built on the assumption of a complete random walk \citep{Diniz.2020},
-implying that directional biases cannot be rendered, albeit being very common in
-dispersal movements \citep{Cozzi.2020, Hofmann.2021}. Ultimately, neither LCP
-analysis nor CT are capable of rendering the temporal dimension of dispersal.
-Statements about the expected duration required to traverse a certain corridor
-are therefore impossible, albeit being important factors. Likewise, because
-movement is not modelled explicitely, neither of the methods allows to render
-interactions between movement and habitat preferences of the focal species. This
-implies that connectivity mainly arises as a result of the landscape structure
-and is referred to as structural connectivity. Structural connectivity stands in
-contrast to functional connectivity, which also renders the behavioral response
-of the animal with respect to prevailing habitat conditions
-\citep{Tischendorf.2000}. Even though functional connectivity is more difficult
-to estimate, a functional view is the ultimate goal in conservation science
-because it has direct consequences for gene flow \citep{Baguette.2013}.
+perceptual range and a preconceived end-point in mind, such that they choose a
+cost-minimizing route accordingly. These assumptions may be reasonable for
+migrating animals, yet they are unlikely to hold true for dispersers, which
+typically move over long distances into unknown territory \citep{Koen.2014,
+Abrahms.2017, Cozzi.2020}. Another issue of LCPs analysis concerns the fact that
+least-costly routes, by their very nature, are only one pixel wide. This implies
+that the width of inferred movement routes depends on the resolution of chosen
+covariate layers and may not be biologically meaningful \citep{Diniz.2020}.
+Although some of these deficiencies can be addressed using less stringent
+versions of the LCP algorithm (e.g. least-cost \textit{corridors}
+\citep{Pinto.2009}, \textit{thresholded} least-cost paths \citep{Landguth.2012},
+and \textit{randomized} least-cost paths \citep{Panzacchi.2016,
+VanMoorter.2021}), a certain degree of arbitrariness remains.
+
+\subsection{Issues with Circuit Theory (90\%)}
+CT entails similarly unreasonable restrictions that are hardly ever met. Because
+CT only allows movements from a source cell to its 4 or 8 adjacent cells, it
+implicitly posits a perceptual range of a single pixel. Assuming that covariate
+layers are usually resolve between 30 m and 1 km, this hardly ever renders the
+true capability of animals to perceive the environment. Moreover, CT builds on
+the assumption of a complete random walk \citep{Diniz.2020}, implying that
+directional biases cannot be rendered. Nevertheless, directionality is a
+commonly observed characteristic in animal movement \citep{Bovet.1991,
+Schultz.2001}, especially in dispersing individuals \citep{Cozzi.2020,
+Hofmann.2021}.
+
+\subsection{Issues of Both Methods}
+Finally, neither LCP analysis nor CT are capable of rendering the temporal
+dimension of dispersal. Statements about the expected duration required to
+traverse a certain corridor are therefore impossible. Likewise, because movement
+is not modeled explicitly, neither of the methods allows to render interactions
+between movement and habitat preferences of the focal species. Connectivity
+therefore merely arises in result to the landscape structure, which is usually
+referred to as structural connectivity. While structural connectivity yields
+insights in the \textit{potential} of the landscape to be traversed, it does not
+enable to quantify the \textit{actual} gene flow through the area. Consequently,
+a functional on connectivity, which also renders the behavioral response of the
+animal with respect to prevailing habitat conditions, is often more desirable
+\citep{Tischendorf.2000, Baguette.2013}.
 
@@
 (IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
-dispersal is simulated explicitly, based on movement rules that determine how
-individuals move over and interact with the prevailing landscape
-\citep{Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020,
-Vasudev.2021}. Using the simulated trajectories, one can calculate a set of
-connectivity metrics, such as interpatch-connectivity and traversal frequency
-across the landscape to reveal major dispersal hotspots \citep{Kanagaraj.2013,
-BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}. Even though IBMMS can be
-employed to overcome any of the shortcomings intrinsic to LCPs and CT, as well
-as to provide a more functional view on connectivity, they can be challenging to
-fit and require vast amounts of data collected during dispersal
-\citep{Diniz.2020}.
+dispersal is simulated spatially explicitly, based on movement rules that
+determine how individuals move over and interact with the prevailing landscape
+\citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004,
+Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019,
+Zeller.2020, Vasudev.2021}. Using the simulated trajectories, one can calculate
+a set of connectivity metrics, such as inter-patch connectivity and traversal
+frequency to reveal major dispersal corridors \citep{Kanagaraj.2013,
+BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}. However, while IBMMS can
+be employed to overcome any of the shortcomings intrinsic to LCPs and CT and
+provide a more functional view on connectivity, they can be challenging to fit
+and require vast amounts of data collected during dispersal \citep{Diniz.2020}.
+Moreover, most individual-based models are heavily influenced by subjective
+decisions, rendering replication and validation of results difficult.
 
@@
 vast majority of their historic range due to persecution by humans, habitat
-fragmentation and destruction, and deadly diseases \citep{Woodroffe.2012,
-Woodroffe.2020}. As of today, only 6'000 free-ranging individuals remain in
-small and spatially scattered subpopulations \citep{Woodroffe.2012}. Within
-those subpopulations, wild dogs form cohesive packs comprising 8 to 12 adults
-and their offspring \cite{McNutt.1995}. After reaching sexual maturity, male and
-female offspring form same-sex coalitions and disperse from their natal pack
-\citep{McNutt.1996, Behr.2020}. New packs are formed when dispersing coalitions
-join unrelated opposite-sex dispersing coalitions \citep{McNutt.1996}.
-Dispersing wild dogs can cover several hundred kilometers across a variety of
-landscapes \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
-One of the few strongholds for this species lies near the Moremi Game Reserve in
-northern Botswana, which is part of the world's largest transboundary protected
-area, namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
-This area has originally been intended to facilitate migration of elephants, but
-is expected to benefit a multitude of other species \citep{Elliot.2014,
+fragmentation and destruction, and deadly diseases \citep{Woodroffe.2012}. As of
+today, only 6'000 free-ranging individuals remain in small and spatially
+scattered subpopulations \citep{Woodroffe.2012}. Within those subpopulations,
+wild dogs form cohesive packs comprising 8 to 12 adults and their offspring
+\cite{McNutt.1995}. After reaching sexual maturity, male and female offspring
+form same-sex coalitions and disperse from their natal pack \citep{McNutt.1996,
+Behr.2020}. New packs are formed when dispersing coalitions join unrelated
+opposite-sex dispersing coalitions \citep{McNutt.1996}. Dispersing wild dogs can
+cover several hundred kilometers across a variety of landscapes
+\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the
+few strongholds for this species lies near the Moremi Game Reserve in northern
+Botswana, which is part of the world's largest transboundary protected area,
+namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA). This
+area has originally been intended to facilitate migration of elephants, but is
+expected to benefit a multitude of other species \citep{Elliot.2014,
 Brennan.2020, Hofmann.2021}.
@@
 In a previous study, we assessed landscape connectivity for dispersing African
-wild dogs within the KAZA-TFCA using a least-cost corridor approach
-\citep{Hofmann.2021}. For this, we fitted a basic habitat selection model based
-on which we predicted landscape permeability. We now expand on this knowledge
-and use ISSF to develop a more detailed movement model of dispersing wild dogs
-(\Cref{GraphicalAbstract}). We then use this model to simulate dispersers moving
-across the KAZA-TFCA. Based on simulations, we compute heatmaps and identify
-potential dispersal hotspots. We also showcase how network metrics relevant to
-landscape connectivity can be computed. Our results show that a simulation-based
-approach yields several major benefits over traditional connectivity modeling
-techniques. Most importantly, simulations provide a more generic view on how
-connectivity emerges and to which degree connectivity depends on the dispersal
-duration. In addtion, by generating proper dispersal trajectories, network
-theory can be applied to calculate network metrics that are pertinent to
-connectivity analysis.
+wild dogs within the KAZA-TFCA using a least-cost methods \citep{Hofmann.2021}.
+Specifically, we fitted a basic habitat selection model using which we predicted
+a permeability surface and computed least-cost paths and corridors. We now
+expand this model and use ISSFs to develop a more detailed movement model of
+dispersing wild dogs (\Cref{GraphicalAbstract}). We then employ the model to
+simulate dispersers moving across the KAZA-TFCA. Based on simulations, we
+generate three distinc connectivity maps, all sheding light onto a different
+aspect of connectivity. With this work we hope to show that a simulation-based
+approach yields several benefits over traditional connectivity modeling
+techniques such as least-cost analyses and circuit theory. Most importantly,
+simulations provide a more generic view on how connectivity emerges and to which
+degree connectivity depends on the dispersal duration. In addtion, by generating
+proper dispersal trajectories, network theory can be applied to calculate
+network metrics that are pertinent to connectivity analysis.
 
@@
 shape = 0.37). Along each step, we extracted and averaged spatial covariates
-using the {\tt velox} package \citep{Hunziker.2017}. We also calculated the
+using the {\tt velox} package \citep{Hunziker.2021}. We also calculated the
 movement metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)} for each
@@
 
+\subsection{Convegence (80\%)}
+To verify that the number of simulated individuals sufficed to ensure reliable
+estimates of connectivity, we evaluated how the relative traversal frequency
+across the landscape depended on the number of simulated trajectories.
+Specifically, we placed 1'000 ``checkpoints'' each with an extent of 5 km x 5 km
+at random locations inside the main study area. We then determined the relative
+traversal frequency by simulated trajectories through each checkpoint for
+different numbers of simulations (1 to 50'000). We replicated each random sample
+100 times and computed the mean across replicates, as well as the 95\%
+xx-interval. Note that we only considered the 50'000 trajectories originating
+from the main study area.
+
 \subsection{Heatmap (100\%)}
@@
 indicating the absolute frequency at which each raster-cell in the study area
-was visited by virtual dispersers. For this, we rasterized all simulated
-trajectories and tallied them into a single map. If the same trajectory crossed
-a raster-cell twice, it was only counted once. This way, we mitigated potential
-biases resulting from trapped individuals that were moving in circles. To
-achieve high performance rasterization, we used the R-package {\tt terra}
-\citep{Hijmans.2020}.
+was visited by virtual dispersers \citep{Hauenstein.2019, Peer.2008}. For this,
+we rasterized all simulated trajectories and tallied them into a single map. If
+the same trajectory crossed a raster-cell twice, it was only counted once,
+thereby mitigating potential biases resulting from individuals that were trapped
+and moved in circles. To achieve high performance rasterization, we used the
+R-package {\tt terra} \citep{Hijmans.2020}.
 
 \subsection{Betweenness (80\%)}
-To pinpoint areas of exceptional relevance for connecting remote regions in our
-study area, we converted simulated trajectories into a network and calculated
-betweenness scores. For this, we overlaid the study area (including the buffer)
-with a regular raster resolved at 5 x 5 km. The centerpoint of each raster-cell
-acted as node in the final network and we used the simulated trajectories to
-determine all transitions occurring from one raster-cell to another, as well as
-the frequency at which those transitions occurred. This resulted in an edge-list
-that we translated into a weighted network using the r-package {\tt igraph}
-\citep{Gabor.2006}. Because {\tt igraph} handles edge weights (\(\omega\)) as
-costs, we inverted the traversal frequency in each cell by applying \(\omega =
-\frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal Frequency_i}\). Finally, we
-used the weighted network to calculate the betweenness score of each
-raster-cell. Betweenness measures how often a specific raster-cell lies on a
-shortest path between two other raster-cells and is therefore a useful metric to
-detect movement corridors \citep{BastilleRousseau.2018}.
+To pinpoint areas of exceptional relevance for connecting remote regions inside
+our study area, we converted simulated trajectories into a network and
+calculated betweenness scores \citep{BastilleRousseau.2018}. For this, we
+overlaid the study area (including the buffer) with a regular raster resolved at
+5 x 5 km. The centerpoint of each raster-cell acted as node in the final network
+and we used the simulated trajectories to determine all transitions occurring
+from one raster-cell to another, as well as the frequency at which those
+transitions occurred. This resulted in an edge-list that we translated into a
+weighted network using the r-package {\tt igraph} \citep{Gabor.2006}. Because
+{\tt igraph} handles edge weights (\(\omega\)) as costs, we inverted the
+traversal frequency in each cell by applying \(\omega = \frac{\sum_i^n{Traversal
+Frequency_i}/n}{Traversal Frequency_i}\). Finally, we used the weighted network
+to calculate the betweenness score of each raster-cell. Betweenness measures how
+often a specific raster-cell lies on a shortest path between two other
+raster-cells and is therefore a useful metric to detect movement corridors
+\citep{BastilleRousseau.2018}.
 
@@
 \section{Results}
-\subsection{Movement Model (60\%)}
+\subsection{Movement Model (80\%)}
 Compared to the base model reported in \citep{Hofmann.2021}, our most
-parsimonious movement model contained several additional interactions between
+parsimonious movement model retained several additional interactions between
 habitat covariates and movement covariates (\Cref{MovementModel} and
-\Cref{MovementModelNumbers}). Although multiple models received an AIC weight
+\Cref{MovementModelNumbers}). Although several models received an AIC weight
 above zero (Table 1 in Appendix S1), we only considered results from the most
-parsimonious model. Since all models with positive AIC weight contained similar
-covariates (Table S1), this decision only marginally influenced subsequent
-analyses. Plots that aid with the interpretation of the final model are provided
-in Appendix S2.
-
-When looking at the habitat kernel (and holding constant the movement kernel),
-we find that dispersing wild dogs avoid water but prefer its proximity.
-Dispersers also avoid woodlands, yet prefer shrublands or grasslands. Finally,
-dispersers avoid moving through landscapes that are influenced by humans. These
-results align with our previous findings reported in \cite{Hofmann.2021}.
-
-Results from the movement kernel indicate several significant estimates.
-However, except for the interaction \textsf{sl:LowActivity}, effect sizes are
-relatively small, indicating that our proposal distributions for step lengths
-and turning angles were only marginally biased by habitat selection. For
-instance, the positive and significant estimate for \textsf{cos(ta)} indicates
-that realized turning angles were slightly more directional than the turning
-angles proposed by our uniform distribution. This shows that realized steps in
-reality followed a von Mises Distribution with positive concentration parameter.
-On the other hand, the significant and negative interaction for
-\textsf{sl:LowActivity} reveals that wild dogs realized shorter steps during low
-wild dog activity compared to the steps suggested by our gamma distribution.
-
-When looking at the interactions between the movement and habitat kernel, we
-observe several significant interactions, highlighting that movement behavior
-differs depending on habitat conditions. For example, there's ample evidence
-that the propensity to realize a log step in areas covered by water is
-substantially smaller compared to the propensity of realizing a small step.
-Similarly, long steps are less likely to be realized in areas covered by trees,
-compared to more open regions. Conversely, it appears that steps tend to longer
-be in the vicinity to water, even though the size of this effect is negligible.
-Finally, it seems that dispersers move more tortuous in areas influenced by
-humans, but more directional when far from water.
+parsimonious model for simplicity. All models with positive AIC weight included
+similar covariates (Table S1), so this decision only marginally influenced
+subsequent analyses. Plots that aid with the interpretation of the final model
+are provided in Appendix S2.
+
+Assuming that all other covariates are held constant, the habitat kernel reveals
+that dispersing wild dogs avoid water but prefer its proximity. Similarly,
+dispersers avoid areas that are covered by woodlands, yet prefer regions covered
+by shrublands or grasslands. Finally, dispersers avoid movement through
+landscapes that are dominated by humans. Effect sizes are strong and, except for
+effect for \textsf{distance to water}, statistically clear on the 5\%
+significance level.
+
+With regards to the movement kernel, the positive estimate for \textsf{cos(ta)}
+indicates that dispersers move with directional persistence, unlike what was
+proposed by the uniform turning angle distribution. Moreover, directionality is
+particularly pronounced when dispersers realize large steps (move quickly), as
+indicated by the positive estimates for \textsf{cos(ta):sl} and
+\textsf{cos(ta):log(sl)}. Finally, the negative estimate for the interaction
+\textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (move
+slowly) outside the main activity periods during sunrise and sunset. Aside from
+the interaction \textsf{sl:LowActivity}, which appears to strongly influence
+movmement behavior, effect sizes are moderate, but mostly significant on the 5\%
+significance level.
+
+When looking at the interactions between the movement and habitat kernels, we
+observe that movement behavior is contingent on habitat conditions. For example,
+there's strong evidence that dispersers realize smaller steps in areas covered
+by water or areas covered by wooldand. Conversely, it appears that steps tend to
+larger in regions dominated by shrubs/grassland, and shorter when the distance
+to water is high. Correspondingly, the model suggests that directionality is
+lower in areas dominated by humans, yet more pronounced when dispersers are far
+from water. However, except for the effect of \textsf{sl:Water}, effect sizes
+are small.
 
@@
     model includes a habitat kernel, a movement kernel, as well as their
-    interactions. The orange and blue line segments delineate the 90\%, 95\%,
-    and 99\% Confidence-Intervals around the respective \(\beta\) coefficients.
+    interactions. The horizontal line segments delineate the 90\%, 95\%, and
+    99\% Confidence-Intervals for the respective \(\beta\)-coefficients.
     Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
     Results from the k-fold cross validation procedure. The upper plot shows
-    rank frequencies of predicted realized scores according to model predictions
-    with known preferences, whereas the lower plot shows rank frequencies when
-    assuming random preferences. The blue ribbon shows the prediction interval
-    around a loess smoothing regression that we fitted to ease the
-    interpretation of the plots.}
+    rank frequencies of realized steps according to model predictions with known
+    preferences, whereas the lower plot shows rank frequencies of realized steps
+    when assuming random preferences. The blue ribbon shows the prediction
+    interval around a loess smoothing regression that we fitted to ease the
+    interpretation of the plots. The significant correlation between rank and
+    associated frequency in (b1) highlights that the most parsimonious model
+    successfully outperforms a random guess (b2) and assigns comparably high
+    selection scores to realized steps.}
     \label{MovementModel}
@@
         \hdashline
-        \multirow{5}{*}{Interaction}
+        \multirow{5}{*}{Interactions}
          & sl : Water & -0.305 & 0.076 & \(<\) 0.001 & *** \\
@@
 On a machine with an octacore AMD Ryzen 7 2700X processor (8 x 3.6 GHz) and 64
-GB of RAM, a single batch of 1'000 simulated dispersers took roughly 90 minutes
-to compute (\(\mu = 88.90\), \(\sigma = 1.87\)) and the simulation of all 80'000
-dispersers terminated after 120 hours, i.e. five days. Comparable computations
-will be substantially faster for smaller study areas or lower resolution
-covariates, as the covariate extraction from large rasters was computationally
-the most expensive task.
-
-On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
-2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
-dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
-dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
-\(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
-0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
-turning angles can be attributed to minor disparities between habitat conditions
-at the area within which we collected training data and habitat conditions
-within the entire study area. Out of the 50'000 dispersers initiated in
-protected areas, only 4.5\% eventually hit a map boundary, suggesting that
-biases due to boundary effects should be limited. In contrast, 78\% of the
-30'000 dispersers originating from the buffer zone eventually hit a map
+GB of RAM, a batch of 1'000 simulated dispersers moving over 2'000 steps
+required 90 minutes to compute (\(\mu = 88.90\), \(\sigma = 1.87\)).
+Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
+terminated after 120 hours or five days. Comparable simulations will be
+substantially faster for smaller study areas and lower resolution covariates, as
+the covariate extraction from large and high-resolution rasters was
+computationally the most demanding task. Out of the 50'000 dispersers initiated
+inside the main source area \Cref{SourcePoints}(a), only 4.5\% eventually hit a
+map boundary, suggesting that we succesfully limited biases due to boundary
+effects. In contrast, 78\% of the 30'000 dispersers originating from the buffer
+zone eventually hit a map boundary, yet this was to be expected since many of
+those dispersers originated from source points located close to the map
 boundary.
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{99_Convergence}
+    \caption{Convergence}
+    \label{Convergence}
+  \end{center}
+\end{figure}
+
+% On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
+% 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
+% dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
+% dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
+% \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
+% 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
+% turning angles can be attributed to minor disparities between habitat conditions
+% at the area within which we collected training data and habitat conditions
+% within the entire study area.
+
 \subsection{Heatmap (80\%)}
-\Cref{Heatmap} depicts the heatmap of all 80'000 simulated trajectories rendered
-after 2'000 steps. The map highlights that large portions of land beyond the
-borders of the KAZA-TFCA are only infrequently visited by dispersers (dark blue
-areas), whereas within the KAZA-TFCA several extensive regions are regularly
-visited (bright yellow and red areas). Most notably, the region in northern
-Botswana south of the Linyanti swamp stands out as highly frequented dispersal
-hub. Still, the presence of several massive water bodies, such as the Okavango
-Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in considerable
-dispersal barriers that limit realized connectivity within the KAZA-TFCA.
-Similarly, dispersal appears to be very limited in Zambia's and Zimbabwe's part
-of the KAZA-TFCA, where only few areas are regularly traversed by dispersers.
-This which can largely be attributed to the strong high human influences in
-these regions. Outside the KAZA-TFCA, the most heavily trafficked regions
-include the areas of the Central Kalahari National Park in Botswana, the area
-south-west of the Khaudum National Park in Namibia, and the area around the
-Liuwa Plains National Park in Zambia.
+\Cref{Heatmap} depicts the heatmap of all 80'000 simulated trajectories
+resulting after 2'000 steps. The map shows that large portions of land beyond
+the borders of the KAZA-TFCA are only infrequently visited by dispersers (dark
+blue areas), whereas within the KAZA-TFCA several extensive regions are
+regularly traversed (bright yellow and red areas). Most notably, the region in
+northern Botswana south of the Linyanti swamp stands out as highly frequented
+dispersal hotspot. Still, the presence of several massive water bodies, such as
+the Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in
+considerable dispersal barriers that limit realized connectivity within the
+KAZA-TFCA. Similarly, dispersal appears to be very limited in Zambia's and
+Zimbabwe's part of the KAZA-TFCA, where only few areas are successfully
+traversed by dispersers. This can largely be attributed to substantial human
+influences resulting from high human density, roads, and agricultural
+activities. Outside the KAZA-TFCA, the most heavily used regions include the
+areas inside the Central Kalahari National Park in Botswana, the area south-west
+of the Khaudum National Park in Namibia, and the area around the Liuwa Plains
+National Park in Zambia.
 
@@
   simulated trajectories. Consequently, the map highlights areas that are
-  frequently traversed by virtual dispersers. Further heatmaps showing the
-  traversal frequency after different numbers of steps are provided in Appendix
-  S3.}
+  frequently traversed by virtual dispersers. Additional heatmaps showing the
+  traversal frequency for different numbers of simulated steps are provided in
+  Appendix S3.}
   \label{Heatmap}
@@
 \subsection{Betweenness (80\%)}
-Betweenness scores of each raster-cell the study area are presented in
-\Cref{Betweenness} and reveals a set of discrete dispersal corridors. As can be
-seen, the dispersal hotspot in northern Botswana reported above is crossed by
-a corridor that receives a relativley high betweenness score. This implies that
-the region is particularly crucial for connecting other regions in the study
-system and hence represents a proper hub in the  ``network'' . Towards east, the
-corridor runs through the Chobe National Park into the Hwange national park,
-where it branches out and further extends into the distant Matusadona National
-Park in Zimbabwe. Northwest of the Linyanty ecosystem, the same corridor expands
-into Angola, where it splits and finally traverses over a long stretch of
-unprotected area into the Kafue National Park in Zambia. Several additional
-corridors with slightly lower betweenness scores exist, yet most of them run
-within the boundaries of the KAZA-TFCA. In general, only few corridors directly
-link the peripheral regions of the KAZA-TFCA. For instance, there are only few
-corridors between the Matusadona National Park in Zimbabwe and the Kafue
-National Park in Zimbabwe. Similarly, there are no direct links between the
-Zimbabwean and Angolan ``spikes'' of the KAZA-TFCA.
+Betweenness scores after 2'000 simulated steps are presented in
+\Cref{Betweenness} and reveal a set of discrete dispersal corridors. Again, the
+region in northern Botswana stands out as crucial dispersal corridor that
+receives the highest betweenness score across the study area. Consequently, the
+region serves as crucial dispersal hub that connects more remote regions in the
+study system. Towards east, the extension of this corridor runs through the
+Chobe National Park into the Hwange national park. From there, a further
+extension connects to the distant Matusadona National Park in Zimbabwe.
+Northwest of the Linyanty ecosystem, a major corridor expands into Angola, where
+it splits and finally traverses over a long stretch of unprotected area into the
+Kafue National Park in Zambia. Several additional corridors with slightly lower
+betweenness scores exist, yet most of them run within the boundaries of the
+KAZA-TFCA. In general, only few corridors directly link the peripheral regions
+of the KAZA-TFCA. For instance, there are only few corridors between the
+Matusadona National Park in Zimbabwe and the Kafue National Park in Zimbabwe.
+Similarly, there are no direct links between the Zimbabwean and Angolan
+``spikes'' of the KAZA-TFCA.
 
@@
   \includegraphics[width=\textwidth]{99_Betweenness.png}
-  \caption{Betweenness scores of each raster cell in a raster with 5 x 5 km
-  resolution. Betweenness scores were determined based on simulated dispersal
-  events. A high betweenness score highlights cells that are exceptionally
-  relevant in connecting different regions in the study area. That is, the
-  higher the betweenness score, the more often a pixel lies on a shortest path
-  between adjacent areas. In this sense the metric can be used to pinpoint
-  discrete movement corridors. Note that we square-rooted betweenness scores to
-  improve visibility of corrdiors with low scores.}
+  \caption{Map of betweenness scores, highlighting distinct dispersal corridors
+  and potential bottlenecks. A high betweenness score indicates that the
+  respective cells are exceptionally relevant in connecting different regions in
+  the study area. That is, the higher the betweenness score, the more often a
+  pixel lies on a shortest path between adjacent areas. In this sense the metric
+  can be used to pinpoint discrete movement corridors. Note that we
+  square-rooted betweenness scores to improve visibility of corrdiors with low
+  scores. Betweenness scores were determined by converting simulated dispersal
+  trajectories into a large network. Note that we square-rooted betweenness
+  scores to improve the visibility of corridors with low betweenness scores.}
   \label{Betweenness}
@@
 
-\subsection{Interpatch Connectivity (50\%)}
+\subsection{Interpatch Connectivity (80\%)}
 Results from the analysis of interpatch connectivity are given in
-\Cref{AreasReached}. It is worth pointing out that the figure is only intended
-as an example; for clarity we limited the network on national parks, albeit
-plenty of links to other protected areas exist. The map shows all realized links
-by simulated dispersers between national parks and indicates the average
-duration a disperser had to move to realize those links. For instance, 6.8 \% of
-the simulated dispersers originating from the Moremi National Park successfully
-reached the Chobe National Park and 4.2 \% the Hwange National Park in Zimbabwe.
-On average, dispersers moved for 623 steps before arriving at Chobe (SD = 520)
-and for 1'413 steps before arriving at Hwange (SD = 371). While the map only
-depicts connections between national parks, this does not imply that no
-connections to other protected areas exist and is the result of our simplifying
-assumptions above. Interestingly, while we identified a potential dispersal
-corridor between Angola's NPs and the Kafue NP in Zambia, \Cref{AreasReached}
-suggests that this link is only rarely realized and requires very long dispersal
-events. In contrast, we find that dispersal between the Moremi NP and Chobe NP
-are relatively frequent and require fewer steps, which can be expected given
-that the areas are located close to each other.
+\Cref{AreasReached}, but it is worth pointing out that the figure is only
+intended as an example, as for clarity we limited the network on national parks
+(NPs), albeit plenty of links between other protected areas exist. As can be
+seen, interpatch connectivity between NPs in Angola, Namibia, and Botswana is
+comparably high and dispersal durations short. In contrast, it appears that
+connectivity towards the Kafue NP in Zambia and is low and requires long
+dispersal events to realize a link. Similarly, there is a lack of connections to
+Zimbabwe's Chizarira and Matusadona NP and the more distant Lower Zambezi and
+Mana Pools NPs.
+
+% The map shows all realized links by simulated dispersers between national parks
+% and indicates the average duration a disperser had to move to realize those
+% links. For instance, 6.8 \% of the simulated dispersers originating from the
+% Moremi National Park successfully reached the Chobe National Park and 4.2 \% the
+% Hwange National Park in Zimbabwe. On average, dispersers moved for 623 steps
+% before arriving at Chobe (SD = 520) and for 1'413 steps before arriving at
+% Hwange (SD = 371). While the map only depicts connections between national
+% parks, this does not imply that no connections to other protected areas exist
+% and is the result of our simplifying assumptions above. Interestingly, while we
+% identified a potential dispersal corridor between Angola's NPs and the Kafue NP
+% in Zambia, \Cref{AreasReached} suggests that this link is only rarely realized
+% and requires very long dispersal events. In contrast, we find that dispersal
+% between the Moremi NP and Chobe NP are relatively frequent and require fewer
+% steps, which can be expected given that the areas are located close to each
+% other.
 
@@
 
-\section{Discussion}
+\section{Discussion} % NEEDS TO BE ENRICHED WITH MORE CITATIONS
 \subsection{Short Summary (90\%)}
-To this end, we used integrated step selection functions to analyse data of
-dispersing wild dogs and parametrize a fully mechanistic movement model
-describing how dispersers move through the available landscape. We employed the
-parametrized model to simulate 80'000 dispersing wild dogs, moving across the
-extent of the KAZA-TFCA, the world's largest transboundary conservation area.
-Based on simulated dispersal trajectories, we prepared a set of complementary
-maps, all geared towards a better understanding of dispersal and landscape
-connectivity. The set of maps included a heatmap, revealing frequently traversed
-areas, a betweenness-map, delineating critical dispersal corridors connecting
-different regions, and a map of inter-patch connectivity, indicating presence or
-absence of functional links between national parks, as well as the average
-dispersal duration required realize those links. With this, we showcase that
-integrated step selection analysis holds the potential to serve as simple, yet
-powerful mechanistic movement model to simulate animal movement and assess
-landscape connectivity. Importantly, such an individual-based approach overcomes
-several shortcomings of traditional connectivity models, such as least-cost
-analyses and circuit theory. Nevertheless, we support the idea that simulations
-only serve as a complement, and not as substitute to more traditional
-connectivity modeling techniques.
-
-\subsection{Movement Model (50 \%)}
-Our most parsimonious movement model comprised of a habitat kernel and a
-movement kernel that describe in detail how dispersers move through the
-available landscape. Parameter estimates revealed that dispersing wild dogs
-avoid water, prefer proximity to water, avoid woodland, prefer shrubs/grassland,
-and avoid areas dominated by humans. This is in line with an earlier dispersal
-model, were we primarily investigated the habitat kernel \cite{Hofmann.2021},
-and suggests that the introduction of additional covariates only marginally
-impacted qualitative results. Even though corresponding effect sizes were
-moderate, the model suggested that movement behavior of dispersers differs
-depending on habitat characteristics. For instance, step lengths appear to be
-shorter (i.e. animals move slower) in areas covered by water in comparison to
-steps realized on dryland. We found similar interactions between movement
-covariates and other habitat covariates, highlighting how a simple ISSF can be
-employed to model relatively complex dispersal patterns. Correspondingly, the
-ISSF framework allowed us to generically model resting behavior by including a
-binary variable, signalling that dispersers moved were substantially slower
-between 17:00 and 09:00 o'clock compared to the rest of the day.
-
-While this (shorter steps on water) could be caused by a bias in our data,
-where the straight line distance between two GPS relocations erroneously
-suggests movement through water, where in reality the animal circumvented
-water-bodies, we hypothesize that this slow-down is caused by dispersers that
-move through the KAZA-TFCA's floodlands and get trapped by the flood. In fact,
-we repeatedly observed individuals venturing across the Okavango Delta,
-suddenly getting surrounded by the advancing flood. Until the flood retracts,
-these individuals often remain on distinct ``islands'' on which movement is
-restricted to rather small steps. This also aligns with the finding that
-turning angles tend to be larger when individuals are close to water (or
-conversely, steps are larger when far from water). It seems that trapped
-individuals wander around, searching for a passage to escape through the
-indundated floodlands, unable to move in a straight-line manner. Quite
-similarly, our model revealed that steps are marginally shorter on woodland,
-but larger on shrubs/grassland, suggesting that wild dogs slow down when
-moving through areas densely covered by trees, but speed up when moving over
-shrubs/grassland. This can probably be attributed to resting periods, during
-which dispersing individuals seek protection in the shade. To effectively
-model this behavior, however, one would need to include a three-way
-interaction between the movement kernel, habitat kernel and a temporal
-measure. Instead, we generically modeled resting behavior using a binary
-variable, which, signaled that steps were substantially shorter between 17:00
-and 07:00 o'clock compared to the rest of the day. This result is rather
-unsurprising as it is well known that wild dogs' follow a diurnal activity
-pattern.
-
-\subsection{Simulation (10\%)}
-By treating the parametrized ISSF model as a fully mechanistic movement model,
-we were able to simulate 80'000 dispersers originating from large protected
-areas. Each individual was simulated over 2'000 steps, which resulted in a total
-of 160 Mio. simulated steps. Although the simulation only terminated after 6
-days, this was largely owed to the massive extent considered (ca. 1.4 Mio. km
-\textsuperscript{2}). Moreover, a smaller number simulated dispersers will often
-suffice, as the relative traversal frequency across the study area in our case
-converged rather quickly (Appendices).
+% Maybe put this at the end?
+We used integrated step selection analysis to analyse data of dispersing wild
+dogs and to parametrize a fully mechanistic movement model describing how
+dispersers move through the available landscape. We employed the parametrized
+model as an individual-based movement model and simulated 80'000 dispersing wild
+dogs moving 2'000 steps across the extent of the KAZA-TFCA, the world's largest
+transboundary conservation area. Based on simulated dispersal trajectories, we
+derived a set of three complementary maps, all geared towards a better
+understanding of dispersal and landscape connectivity. The set of maps included
+a heatmap, revealing frequently traversed areas, a betweenness-map, delineating
+critical dispersal corridors and bottlenecks, and a map of inter-patch
+connectivity, indicating the presence or absence of functional links between
+national parks. We thereby showcase how integrated step selection functions can
+be utilized as simple, yet powerful framework to parametrize movement models and
+simulate animal movement with the aim of assessing landscape connectivity. Such
+an individual-based simulation overcomes several conceptual shortcomings
+inherent to more traditional connectivity modeling techniques, such as
+least-cost analyses and circuit theory. Nevertheless, we suggest to view such
+individual-based simulations as complement, and not as substitute, to
+traditional connectivity models.
+
+\subsection{Movement Model (80 \%)}
+Our movement model of dispersing wild dogs comprised a habitat kernel, a
+movement kernel, and their interactions. Together, the kernels described habitat
+and movement preferences of dispersing individuals, as well as how movement
+behavior was affected by habitat conditions. Parameter estimates from the
+habitat kernel revealed that dispersing wild dogs avoid water, prefer its
+proximity, avoid woodland, prefer shrubs/grassland, and avoid areas dominated by
+humans. These findings are consistent with the findings of previous studies on
+dispersing wild dogs \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019,
+Oneill.2020}, as well as with an earlier dispersal model that we developed
+primarily to investigate the habitat kernel of dispersers \citep{Hofmann.2021}.
+
+By expanding our initial model to include a proper movement kernel, we were able
+to model several additional complexities inherent to dispersal. It is, for
+instance, well known that turning angles and step lengths tend to be correlated
+in dispersing animals due to a preference for directional persistence. That is,
+larger steps usually coincide with smaller turning angles and vice versa (cite
+someone). While this correlation structure could be effectively rendered by
+sampling turning angles and step lengths from copula probability distributions
+\citep{Hodel.2021}, the ISSF framework allowed us to model similar behavior
+directly in the movement model. Besides accounting for directional persistence,
+we further exploited the ISSF framework and included interactions between
+movement covariates to represent the fact that wild dogs mainly move during the
+more temperate morning and evening hours. The final model thus rendered a
+pulse-like movement behavior, where dispersers realize larger steps during
+morning and evening hours, but shorter steps during the rest of the day.
+
+By including interactions between habitat and movement covariates, we also
+acknowledged that movement behavior of dispersers may depend on habitat
+conditions. For example, the final model contained an interaction between water
+cover and step length, highlighting that dispersers are more likely to realize
+short steps (i.e. move slower) in areas covered by water. Likewise, the
+interaction between water cover and turning angles revealed that dispersers move
+less directional when across water bodies. We believe that this behavior is owed
+to the fact that wild dogs wade or swim to traverse waterbodies, thus resulting
+in slower, more tortuous movements. Besides this, our model also suggested that
+dipsersers preferably realize shorter steps when moving through woodland, but
+larger steps when moving across shrubs/grassland. This can likely be attributed
+to resting periods during which dispersers seek shade and protection below the
+woodland canopy. Since we did not include three way interactions, we could not
+test this hypothesis.
+
+\subsection{Simulation (80\%)}
+% I need to better distinguish our approach from other IBMMs, maybe use the term
+% iSSF-IBMM?
+Based on the above outlined model results, we simulated 80'000 dispersers moving
+2'000 steps across the landscapes of the KAZA-TFCA. Our dispersal simulation
+thus comprised 160 Mio. simulated steps, which required six days of computation
+on a modern desktop machine. This rather long simulation duration was primarily
+owed to the massive extent considered (ca. 1.8 Mio. km \textsuperscript{2}) and
+the large number of dispersers simulated. Most connectivity studies consider a
+much smaller extent (e.g. \citealp{Kanagaraj.2013, Abrahms.2017, Zeller.2020})
+and should be able to attain faster simulation speeds. We also believe that
+fewer simulated dispersers will often suffice, as the relative traversal
+frequency of simulated individuals through randomly placed checkpoints in our
+study area converged rather quickly (Appendices), yet this will vary depending
+on the structure of the landscape and dispersal ability of the focal species.
+
+A simulation-based approached as proposed in this article offers several
+advantages over more traditional connectivity modeling techniques such as
+least-cost analyses or circuit theory. In contrast to least-cost analyses, for
+instance, an individual-based simulation does not require predetermined
+endpoints. Instead, endpoints emerge naturally as the result of simulated
+dispersal trajectories. Not having to assume known endpoints is particularly
+important for dispersal studies because dispersers often venture into unfamiliar
+territory and are therefore unlikely to know the endpoint of their journey. In
+addition, without known endpoints, movement corridors are no longer enforced
+between start- and endpoint, permitting to detect potential dead ends where
+dispersers get trapped due to insurmountable obstacles.
+
+IBMMs furthermore yield the advantage of an explicit representation of time.
+This enables to answer questions such as: ``\textit{How long will it take a
+disperser to move from A to B?}'' or \textit{Is it possible for a disperser to
+move from A to B within X days?}, all undoubtly interesting questions that shift
+the focus from a structural to a more functional view on connectivity. Similar
+questions cannot be answered using least-cost methods or circuit theory because
+the time component is omitted from those models. An explicit representation of
+time in IBMMs also yields exciting opportunities for studying seasonal changes
+to connectivity \cite{Zeller.2020}. With least-cost methods or circuit theory a
+dynamic analysis of connectivity is impractical because the same analyses needs
+to be repeated using different permeability surfaces (e.g. \citealp{Benz.2016,
+Osipova.2019}). With IBMMs, on the other hand, simulated individuals can
+directly respond to a dynamic environment within the model and the environment
+is allowed to change ``as the dispersers move''.
+
+While an explicit representation of time yields several advantages, it also
+requires that step lengths and turning angles are modeled properly
+\citep{Kanagaraj.2013}, so that dispersal durations between habitat patches can
+be reliably approximated. Accounting for and correctly rendering this movement
+kernel is the main strength of ISSFs, which is why we believe that the framework
+is particularly well suitded for simulating animal dispersal and assessing
+landscape connectivity. In principle, the ISSF framework allows to define an
+infinite amount of movement kernels, each representing movement behavior under
+different conditions \citep{Fieberg.2020}. It can therefore accomodate highly
+diverse movement modes, including fast paced directional movements and slow
+paced tortuous movements. Compared to models that disregard an animal's movement
+kernel, this allows to model directional peristence, a movement characteristic
+regularly observed in dispersing animals. To date, however, we only accounted
+for 1. order autocorrelation. That is, we only considered correlations between
+two subsequent steps. Modeling higher-order autocorrelation may be desirable,
+yet will require more extensive data.
+
+Overall, IBMMs enable to generate a rich insight into the emergence of
+connectivty when compared to the outputs of least-cost analysis or circuit
+theory. By generating proper movement trajectories, a broad variety of summary
+statistics can be derived. Specifically, by turning simulated trajectories into
+a network, network theoretical metrics relevant to connectivity can be applied
+\citep{BastilleRousseau.2018}. In contrast to least-cost path analysis and
+circuit theory, a simulation-based approach permits studying inter-patch
+connectivity. This facilitates the identification of asymmetrical links and
+statements about source-sink dynamics \citep{Ferreras.2001, Revilla.2004,
+Kanagaraj.2013}. Our results are also in concert with previous studies on
+functional connectivity based on individual-based models that showed that
+including the behavioral ecology of the target species and the landscape
+structure are imperative when assessing connectivity \citep{Gustafson.1996,
+Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004, Revilla.2008,
+Kanagaraj.2013} rather than considering it purely a function of distance.
+
+Besides the benefits of IBMMs, we also want to point out some caveats and not so
+trivial modeling decisions. In particular, we will discuss five modeling
+challenges: (1) number of simulated individuals, (2) location of source points,
+(3) dispersal duration, (4) boundary behavior, and (5) individual variability.
+
+(1) When simulating dispersal, the modeler needs to decide on the number of
+simulated individuals. This decision involves the total (absolute) number of
+simulated individuals across the study area, as well as the relative number of
+simulated individuals per spatial entity (e.g. protected area, habitat patch,
+source point). With regards to the total number of simulated individuals, a
+larger number is usually desirable as it yields a more balanced view on
+connectivity. However, this comes at the cost of computational efficiency,
+implying that a trade-off needs to be managed. One simple solution to handle the
+trade-off would be to simulate only as many individuals as are needed to achieve
+convergence in the target metric. In our case, examination of the target metric
+\textsf{traversal frequency} revealed that convergence was achieved with
+relatively few individuals, such that additional dispersers contributed only
+little additional information to our results. With regards to the relative
+number of simulated individuals, we see two feasible approaches. First, one
+could initiate dispersers in relation to population density at the respective
+area. This would directly reflect the fact that population densities are not
+homogeneous across space. (Say more) Alternatively, however, one could also
+distribute dispersers homogeneously, but then weigh simulated trajectories by
+the population density at the source patch (Say more). Because we lacked
+reliable information on density estimates for the extent of the KAZA-TFCA, we
+opted for a solution where dispersers were distributed homogeneously.
+
+(2) While endpoints do not need to be defined in IBMMs, source points will still
+need to be provided by the modeler. Here, we randomly placed source points
+within protected areas large enough to sustain viable wild dog populations,
+knowing that the species primarily survives in these formally protected areas
+\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. In
+some cases, exact locations of potential source populations are known and source
+points can be placed accordingly \citep{Kanagaraj.2013}. Moreover, if abundance
+estimates are available, these can be used to inform the relative number of
+dispersers initiated at each location. In other cases, comparable knowledge may
+be lacking and it could be more beneficial to delineate likely source patches
+based on habitat suitability models (e.g. \citealp{Squires.2013}). Either way, a
+meaningful selection of source points is not a problem unique to IBMMs and
+applies to any connectivity modelling technique. As highlighted by
+\cite{Signer.2017} the influence of the exact location of source points
+decreases as the number of simulated steps is increased, yet their goal was
+assessing steady state utilization distributions and not to assess landscape
+connectivity.
+
+(3) When employing IBMMs, one also needs to decide on meaningful dispersal
+durations (number of simulated steps). We decided to simulate individuals for
+2'000 steps, which is at the upper end of observed dispersal durations and may
+have resulted in an overestimated representation of landscape connectivity.
+Instead of enforcing homogeneous dispersal durations, one could also sample
+dispersal durations from on observed dispersal events. Alternatively, after each
+step, individuals could be faced with a probability to settle, where the
+probability is contingent on the surrounding landscape and cumulative dispersal
+distance travelled. Because we only observed few dispersal events and due to the
+wild dog's ability to cover vast dispersal distance \citep{DaviesMostert.2012,
+Masenga.2016, Cozzi.2020} we decided not to apply this approach. Regardless of
+this, we think that in most cases it will be more insightful to simulate
+relatively extensive dispersal events and only subsample afterwards.
+
+(4) Unless simulated individuals are strongly drawn towards a point of
+attraction, some simulated individuals will eventually approach a map boundary
+and the modeller needs to define a rule to handle such situations. One feasible
+option would be to terminate the trajectory as soon as it hits a map boundary,
+implicitly assuming that the simulated animal left the study area and will not
+return. The downside to this approach is that source points located near map
+boundaries will inevitably produce many such dead ends. Alternatively, one could
+remove transgressing random steps from the set of proposed random steps, thereby
+forcing virtual dispersers to bounce off virtual boundaries and remain within
+the main study area. This may seem like an elegant solution, yet again, it
+potentially results in substantial edge effects when source points are located
+near map boundaries. Finally, one could extend the study area by an artificial
+buffer zone through which dispersers are allowed to leave and re-enter the main
+study area. Of course this does not fully mitigate the issue, since some
+dispersers may still approach the outer map boundary, however, as
+\cite{Koen.2010} have shown, such an artificial buffer can often help to
+mitigate edge effects. In theoretical applications, this issue can be
+circumvented by simulating movement on a torus \citep{Hodel.2021b}, yet this is
+unfeasible in most real world applications.
+
+(5) We simulated dispersal using point estimates from our most parsimonious
+movement model but did not investigate the sensitivity of our results with
+respect to those estimates. Uncertainty is rather common in dispersal studies on
+endangered species, as data tends to be scarce, resulting in model estimates
+large confidence intervals \citep{Wiegand.2003, KramerSchadt.2007}. To address
+this, one may explore a broader range of preferences instead of using point
+estimates initiate dispersers with randomized preferences with variability
+imposed by the uncertainty in the movement model. We therefore urge future
+studies to investigate the sensitivity of ISSF simulations with respect to
+estimated preferences.
 
 \subsection{Maps (10\%)}
-Our heatmap suggests that a large number of dispersers traverses the regions of
-the Moremi NP and the Chobe NP. from the Okavango Delta more likely disperse
-towards east than west. Indeed, only x out of our y observed dispersers ever
-reached the western part of the delta. Only when the flood retracts a small
-pathway between the city of Maun and the floodwaters of the delta emerges and
-enables dispersers to move towards the detal's western part.
-
-While the segment running into Kafue receives a high betweenness score, it was
-actually only rarely traversed by our simulated dispersers, as can be seen from
-the dark colors in this region in \Cref{Heatmap}. It is therefore worth noting
-that the betweenness metric highlights crucial bottlenecks that are relevant for
-connecting remote regions, it does not directly yield information about the
-frequency at which these bottlenecks are used.
+The heatmap resulting from our dispersal simulation suggests that a large
+portion of dispersers traverses the Moremi NP and the Chobe NP in northern
+Botswana. Since we already highlighted the same area in our previous analysis
+\citep{Hofmann.2021}, this result was to be expected. Nevertheless, we believe
+it reinforces our notion that the area acts as crucial hub inside the KAZA-TFCA.
+The very same area stands out on the betweenness map, showing that the region is
+not only frequently traversed, but actually serves as stepping stone into more
+remote regions of the KAZA-TFCA. As such, the area exemplifies a region were
+both traversal frequency and betweenness scores are high. In contrast, the
+betweenness map suggests the presence of a corridor moving from Angola into
+Zambia's Kafue NP, yet when looking at the heatmap, one realizes that the area
+is not used very often. Consequently, despite the corridor's importance for
+linking Angola's NPs to Zambia's Np, only very few simulated dispersers
+successfully traversed it. This is also reflected in the interpatch connectivity
+map, where the Kafue NP only receives very few links from the central region of
+the KAZA-TFCA.
+
+Each map that we produced from simulated trajectories accentuates a different
+aspect of connectivity. The heatmap, for example, puts emphasis on areas where
+movement is concentrated, regardless whether such areas represent ``dead ends''
+that do not necessarily connect distinct patches. Therefore, the heatmap is
+useful to determine locations where a disperser are likely to move to, yet not
+necessarily to uncover corridors that link valuable habitats. The betweenness
+map, on the other hand, brings out those areas that are relevant in connecting
+different regions in the landscape. In this regard, the betweenness map is most
+pertinent to the delineation of dispersal corridors and may serve as an
+alternative to least-cost algorithms. Finally, the inter-patch connectivity map
+illustrates the frequency at which dispersal between distinct habitat patches
+occurs, as well as the average dispersal duration that passes when individuals
+move between those patches. This knowledge serves to gauge the amount of
+geneflow between habitat patches and can be employed to determine whether two
+habitats are connected or not.
 
 \subsection{General (20\%)}
-A major benefit of the simulation-based approached is that the domain of
-endpoints does not need to be determined \textit{a priori}. Instead, each
-endpoint emerges naturally as the result of a simulated dispersal trajectory.
-This is particularly important when modeling dispersers, mainly because
-dispersing individuals venture into unknown territory and do not necessarily
-move towards a predetermined endpoint. Nevertheless, in cases were individuals
-do exhibit such preferences, the incorporation of attraction points in the
-corresponding ISSF movement model allows to render them.
-
-While endpoints emerge naturally from the dispersal process, startpoints still
-need to be specified \textit{a priori} by the modeler. Here, we placed source
-points within protected areas large enough to sustain viable wild dog
-populations. This was done under the simplifying assumption that wild dogs only
-survive in formally protected areas, which is in line with scientific findings
-\citep{Woodroffe.1999, Woodroffe.2012, VanDerMeer.2014, Loveridge.2019}. In some
-cases, exact locations of source populations are known and source points can
-easily be placed accordingly \citep{Kanagaraj.2013}. In other cases, comparable
-knowledge may be lacking and it may be more appropriate to delineate likely
-source patches using habitat suitability models \citep{Squires.2013}. Because we
-did not investigate the sensitivity of our results with respect to the exact
-location of source points, this is something that needs further investigation in
-the future.
-
-Another advantage of simulation-based approaches is that they render time
-explicitly. This enables to answer questions such as: ``\textit{How long will it
-take a disperser to move from A to B?}'' or \textit{Is it possible for a
-disperser to move from A to B within X days?}. These are interesting questions
-and they shift the focus from a structural to a more functional point of view.
-However, an explicit representation of time requires that speed (step length)
-and directionality during motion is approapriately modeled
-\citep{Kanagaraj.2013}. Because ISSFs enable to model these two components
-adequately, the method offers a powerful framwork for simulations. Besides this,
-one also needs to decide on a meaningful dispersal duration when simulating
-movement. We decided to simulate individuals for 2'000 steps, which is at the
-upper end of observed dispersal durations and likely resulted in overestimated
-landscape connectivity. Nevertheless, it requires little tweaking to subset from
-the generated simulations to any dispersal duration desired. In fact, one could
-randomly sample dispersal durations based on observed dispersal events. In most
-cases, however, it will be more convenient and insightful to simulate relatively
-extensive dispersal events and only subsample afterwards.
-
-While we have assumed a set of static covariates when simulating dispersal, an
-explicit representation theoretically allows to render seasonality in covariate
-layers. This is an important aspect in ecosystems where seasonality
-substantially influences landscape connectivity. With least-cost analysis and
-circuit theory, seasonality can merely be incorporated by producing a multitude
-of permeability surfaces, each depicting landscape permeability in a different
-season, and then applying the connectivity models to those surfaces
-\citep{Benz.2016, Osipova.2019}. With individual-based simulations, on the other
-hand, seasonal covariates can be updated as the simulated dispersers move. As
-such, seasonality would directly influence movement, i.e. the process that
-ultimately leads to connectivity. For instance, in our simulation we represented
-the Okavango Delta statically and assumed a relatively extended flood. In this
-regard, the maps presented in the results section may be most representative of
-the period shortly after the wet-season, when floodlevels in the Delta are at
-their maximum. During the dry season, however, the flood considerably retracts
-and potentially clears the way for wild dogs dispersing from the Moremi-Game
-reserve into the south-western section of the Delta. Consequently, instead of
-using a static floodmap, one could render the flood dynamically. Hence, the
-floodlevels would be updated as the dispersers move, which would allow studying
-how connectivity evolves as the flood climaxes and retracts again.
-
-We simulated dispersal using point estimates from our most parsimonious movement
-model, yet the degree to which our results depend on those estimates is unknown.
-Given that data from dispersal studies on endangered species is scarce, point
-estimates may be quite inaccurate, therefore leading to erronous inference
-\citep{Wiegand.2003, KramerSchadt.2007}. Rather than using point estimates, an
-alternative may be to simulate dispersers using a set of randomized preferences
-imposed by the uncertainty reported in the model output. We urge future studies
-to further investigate investigate the sensitivity of ISSF simulations with
-respect to estimated habitat preferences.
-
+% Mortality during dispersal
+While we did not attempt to model mortality during dispersal, it is
+well known that dispersers regularly die, mainly due to deadly encounters with
+predators, road kills, or persecution by humans \citep{Woodroffe.2012,
+Behr.2021b}. Mortality during dispersal could therefore substantially limit
+functional connectivity, especially in areas where the likelihood of
+encountering competitors and humans is high \citep{Cozzi.2020}. If corresponding
+information is available, IBMMs would readily allow to model mortality. Each
+timestep, a binary draw determines whether the animal survives or not and this
+probability can depend on landscape characteristics as well as the dispersal
+duration.
+
+% Other studies that use similar simulations
+Our approach of simulating movement to assess connectivity is most closely
+related to the works of \cite{Clark.2015} and \cite{Zeller.2020}, who used
+regular step selection functions to model habitat preferences by black bears and
+simulated movement to assess connectivity. However, neither of these studies
+attempted to jointly model habitat and movement kernels, thereby precluding a
+more mechanistic understanding of movement. Moreover, both studies parametrized
+models using data of resident black bearsk, which may lead to a biased view on
+connectivity, as dispersers may move substantially further, thereby improving
+connectivity among habitat patches \citep{Elliot.2014}.
+
+% Spatially Realistic Population Models
+Besides enabling a more mechanistic understanding of connectivity, the ability
+to realistically render movement during dispersal also forms the foundation for
+spatially realistic population models. In these models, dispersal is not merely
+represented by a dispersal kernel (e.g. ...) but mechanistically rendered
+\citep{Revilla.2008}. Such population models can ultimately be employed to
+determine the required level of dispersal to achieve metapopulation viability
+\citep{DaviesMostert.2012}.
+
+% Simulations for an early warning system?
+Simulations resulting from IBMMs could furthermore be utilized as simple tool to
+predict the likely whereabouts of GPS collared animals into the near future. In
+some European countries, the comeback of large predators, such as bears, lynx,
+and the wolf, has triggered emotional discussions and raised public concern
+\citep{Behr.2017}, particularly in areas with free-roaming livestock that may be
+prayed upon by the returned species. An early warning system based on
+simulations could thus serve to forewarn about potential encounters and and
+thereby increase public acceptance of large predators.
+
+% Social landscape
 We have previously attributed the weak significance of distance to water to the
@@
 for inter- and intra-speficic interactions that may explain why and when
-dispersers are attracted to or afraid of waterbodies.
-
-Comparable simulations that are based on empirical data are also a fundamental
-component for spatially realistic population models in which dispersal is
-rendered more realistically and does not merely depend on the distance between
-habitat patches.
-
-An important benefit of ISSF simulations is that the framework always considers
-availability. That is, the propensity of a simulated disperser to realize a
-certain step is always contingent on the set of alternative steps. As such, a
-disperser surrounded by relatively unsuitable habitat will still move and
-disperse insteado f getting stuck.
-
-Even though connectivity is generally believed to promote population viability,
-it has also been pointed out that improved connectivity may cause ecological
-traps, especially when connectivity into or through human-dominated landscapes
-is promoted. In such instances, connectivity increases the risk of encountering
-humans and facilitates persecution by humans. By overlapping simulated
-trajectories with a map of human influence, such ecological traps could be
-pinpointed.
-
-We did not model mortality during dispersal in our simulations. This was a
-simplifying assumption that only inaccurately reflects true dispersal. It is
-well known that wild dogs often die during dispersal in result of deadly
-encounters with competing predators, but also due to road kills or human
-persecution \citep{Woodroffe.2012, Behr.2021b}. In result, mortality during
-dispersal may limit realized connectivity, especially in areas that wild dogs
-are unfamiliar with and in areas with a high potential for human-wildlife
-conflict \citep{Cozzi.2020}.
-
-Due to a lack of information on wild dog densities across the extent of the
-KAZA-TFCA, we assumed and simulated an almost uniform distribution of dispersers
-across protected areas. In some cases, prior knowledge about the density of
-potential dispersers is available and can be used to adjust the number of
-simulated individuals accordingly. Instead of tweaking the number of simulated
-individuals, one could also weight each trajectory relative to the density of
-potential dispersers in the corresponding source areas, such that trajectories
-from areas with high density would enter the heatmap with a higher weight.
-
-One of the major benefits of individual-based simulations is the ability to make
-the temporal dimension of movement explicit. This allows to investigate how
-connectivity depends on the dispersal duration, something that is not possible
-with traditional least-cost or circuit theory methods.
-
-Optimally, one should simulate additional dispersers until the amount of newly
-gained information lies beyond a certain threshold, i.e. until some sort of
-convergence is achieved. However, due to the myriad of outcomes in movement
-behavior, this would often imply countless repetitions for each source points at
-each location where dispersal is possibly initiated. This is computationally
-infeasible.
-
-In some European countries, the comeback of large predators, such as bears,
-lynx, and the wolf, has triggered emotional discussions and raised public
-concern \citep{Behr.2017}, particularly in areas with free-roaming livestock
-that may be prayed upon by the returned species. In cases were recent locations
-of such predators are known, an early warning system based on simulations could
-serve to forewarn about potential encounters and and thereby increase public
-acceptance of large predators.
-
-We do not dismiss the application of least-cost analysis or circuit theory in
-any way. Rather, we propose to use simulations from IBMMs complementary to
-least-cost analysis and circuit theory in order to provide a more comprehensive
-analysis of landscape connectivity. In fact, simulations from IBMMs could be
-employed to identify likely connections between habitat patches in a first place
-and only afterwards least-cost paths and corridors can be drawn between such
-likely connections. Thus far, dispersal kernels or thresholded least cost paths
-have been used for this.
-
-Each map that we produced from simulated trajectories accentuates a different
-aspect of connectivity. The heatmap, for example, puts emphasis on areas that
-are frequently traversed, even if such areas represent ``dead ends'' where
-dispersers get trapped and bounce off from obstacles. Hence, the map is useful
-to determine locations where a disperser is likely to show up, yet not
-necessarily to determine corridors that are worthwile to be protected. The
-betweenness map, on the other hand, brings out those areas that are relevant in
-connecting different regions in the landscape. In this regard, the betweenness
-map is most pertinent to the delineation of dispersal corridors and may serve as
-an alternative to least-cost algorithms. Finally, the inter-patch connectivity
-map illustrates the frequency at which dispersal between distinct habitat
-patches occurs, as well as the average dispersal duration that passes when
-individuals move between those patches. This knowledge serves to gauge the
-amount of geneflow between habitat patches and can be employed to determine
-whether two habitats are connected or not.
+dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
+instance, found that elk movement was significantly impacted by the density of
+wolf in the area, such that habitat preferences strictly differed depending on
+the presence or absence of wolves. The decision to settle is likely related to
+the presence or absence of conspecifics. Hence, the exact dispersal duration and
+distance will not be independent of current wild dog densities. In dispersing
+wolves, for instance, the longest dispersal distances have been observed in
+low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
+may thus be determined by the by the amount of islotation between subpopulations
+and population densities \citep{DaviesMostert.2012}.
+
+% Downsides of connectivity?
+Even though connectivity is generally thought to promote population viability,
+it is also related to various aspects that may cause ecological damage, such as
+increased human-wildlife conflicts or facilitated spread of deadly diseases.
+
+% Conclusion
+To this end, we have exemplified the use of ISSFs as a simple framework based on
+which researchers can parametrize dispersers' habitat and movement preferences
+and at the same time use the parametrized model to as an individual-based
+movement model to simulate dispersal and examine landscape connectivity.
+Furthermore, we suggest to analyse simulated trajectories using three
+complementary maps, each focused on a different aspect of connectivity.
+Ultimately, we hope to have sparked interest in the uprising framework of step
+selection functions for investigating dispersal behavior and landscape
+connectivity. Nevertheless, we do not attempt to dismiss the application of
+traditional connectivity models such as least-cost methods or circuit theory by
+any means. Rather, we propose to use simulations from ISSF-IBMMs as a simple but
+powerful tool that complemets earlier methods in an attempt provide a more
+comprehensive understanding of landscape connectivity.
 

### 2021-06-24 Chapter_2/04_Manuscript/Manuscript.tex
various updates

--- 2021-05-05 Chapter_1/04_Manuscript/Manuscript.tex
@@
 %------------------------------------------------------------------------------
-\title{Step by Step: Using Step Selection Analysis to Simulate Dispersal and
-Assess Landscape Connectivity}
+\title{Dispersal Through Seasonal Landscapes}
+
+% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
+% Connectivity}
 
@@
 \begin{abstract}
-Dispersal is an important process that allows species to avoid inbreeding, to
-colonize new habitats and to reinforce non-viable subpopulations. Successful
-dispersal thus represents a crucial pre-requisite for long-term species
-persistence in wild animal populations. However, the ability to disperse is
-contingent a sufficient degree of landscape connectivity, which is why the
-estimation of connectivity and identification of dispersal corridors has become
-a task of extraordinary importance in ecological studies.
-
-Over the past two decades, ecologists have primarily relied on analytical tools
-such as least-cost methods and circuit theory to investigate landscape
-connectivity. Despite their usefulness for various ecological applications, both
-methods make several restricting assumptions that limit their suitability in
-reality. To address their shortcomings, simulations from individual-based
-movement models have been proposed and applied. To overcome their shortcomings,
-simulations from individual-based movement models have been proposed. Yet, due
-to the infinite number of possibilities of coming up with an individual-based
-model, a unified and objective framework is missing.
-
-Recent innovations in movement ecology have brought forward novel opportunities
-to study animal dispersal and landscape connectivity. In particular, the rich
-suite of resource selection functions, including point-, step-, and
-path-selection functions have undergone substantial improvements over the past
-years. Most notably, step-selection functions have been generalized to
-\textit{integrated} step selection functions, which essentially represent fully
-mechanistic movement models based on which individual animal movement could be
-simulated. While such models have been applied to study \textit{steady-state}
-utilization distribution, a similar approach may be useful for investigating
-\textit{transient} movement behavior and thereby study landscape connectivity by
-means of simulated dispersal events.
-
-Here, we showcase the use of integrated step selection analysis to simulate
-dispersal of the highly endangered African wild dog across the world's largest
-transboundary conservation area, the Kavango-Zambezi Transfrontier Conservation
-Area (KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
-coalitions departing from northern Botswana. We combine the movement data with
-relevant spatial covariate layers and used integrated step selection functions
-to parametrize a fully mechanistic movement model rendering wild dog dispersal.
-Based on this model, we simulate 80'000 dispersers, originating from protected
-areas and moving across the extent of the KAZA-TFCA. We then generate a heatmap
-and use network theory to reveal dispersal hotspots and crucial bottlenecks
-across the study area. Finally, we discuss the benefits and pitfalls of such
-dispersal simualtions and highlight potential improvements to be made in the
-future.
+This is the abstract.
 \end{abstract}
@@
 
-\section{Introduction (90\%)}
+\section{Introduction}
 
-% Importance of Dispersal & Connectivity
-\subsection{Importance of Dispersal \& Connectivity (90\%)}
-Dispersal is defined as the movement of individuals from their natal location to
-the site of first reproduction \cite{Howard.1960}. It is a vital process
-governing the dynamics wild animal populations that are distributed in space
-\citep{Hanski.1998, Clobert.2012} and may strongly affect population dynamics at
-different spatial and social scales \citep{Hanski.1999a, Clobert.2012}.
-Dispersal allows species to avoid inbreeding and maintain genetic diversity
-\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013}, to
-rescue small and non-viable populations \citep{Brown.1977}, and to promote the
-colonization of unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}.
-However, successful dispersal requires a sufficient degree of functional
-connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the identification
-and protection of major dispersal corridors has become an important task in
-conservation science \citep{Doerr.2011, Rudnick.2012}. To achieve this, reliable
-information on movement behavior during dispersal and knowledge about factors
-that limit dispersal and connectivity is paramount \citep{Baguette.2013,
-Vasudev.2015}.
+% Taken from Zeller.2020
+Connectivity for wildlife is typically modeled in static environments and
+estimated for a single snapshot in time (e.g., [6,7]). However, wildlife
+responses to landscapes are inherently dynamic, as are the landscapes themselves
+(e.g., [8])
 
-% Advancements in GPS Technology & Movement Ecology
-\subsection{Advancements in GPS Technology \& Movement Ecology (90\%)}
-Thanks to novel technologies developed over the past decades, particularly of
-GPS/Satellite radio-collars, the use of GPS data to study animal dispersal and
-connectivity has accelerated \citep{Elliot.2014, Jonsson.2016, Williams.2019}.
-Additionally, the advent of publicly accessible satellite imagery and
-sophisticated remote sensing techniques to represent the physical landscape
-through which individuals disperse have heralded a ``golden age of animal
-tracking'' \citep{Kays.2015}. Concurrently, the availability of large amounts of
-empirical data and an increased computational power have led to the development
-of numerous techniques to study dispersal movements and highlight critical
-corridors between subpopulations \citep{Boyce.2002, Fortin.2005, Cushman.2010,
-Zeller.2012, Diniz.2020}.
-
-% Resource Selection & Connectivity
-\subsection{Resource Selection \& Connectivity (90\%)}
-\textit{Resource selection functions} \citep{Boyce.2002} and derived methods
-such as \textit{step selection functions} \citep{Fortin.2005} and \textit{path
-selection functions} \citep{Cushman.2010} have proven particularly useful for
-studying animal movement \citep{Fieberg.2020} and modelling connectivity
-\citep{Diniz.2020}. These methods allow estimating habitat preferences of the
-focal species by comparing covariates at locations visited by the animal to the
-same covariates at locations available to, but not visited by the animal
-\citep{Boyce.2002, Fortin.2005, Cushman.2010, Thurfjell.2014}. The so estimated
-preferences can then be used to predict a permeability surface, indicating the
-expected ease at which an animal can traverse a given area \citep{Spear.2010,
-Zeller.2012, Etherington.2016}. Ultimately, the permeability surface serves as
-input to a connectivity model that is used to reveal movement corridors
-\citep{Diniz.2020}. Two of the most prominent connectivity models are least-cost
-path analysis (LCP analysis; \citealp{Adriaensen.2003}) and circuit theory (CT
-\citealp{McRae.2006, McRae.2008}), both graph-based methods that estimate
-conductance of the landscape. Despite their intuitive nature and ease of use,
-both methods make rigorous assumptions about animal movement that may or may not
-be fulfilled in reality \citep{Diniz.2020}.
-
-% Issues with Least-Cost Paths & Circuit Theory
-\subsection{Issues with Least-Cost Paths \& Circuit Theory (90\%)}
-In LCP analysis, for instance, a least costly path always exists, even if
-associated movement costs are unreasonably high and will never be incurred by a
-dispersing individual. The method also presumes that animals have an infinite
-perceptual range, a preconceived end-point in mind, and choose a cost-minimizing
-route accordingly. These assumptions may be reasonable for migrating animals,
-yet they are unlikely to be true for dispersers that move over long distances
-into unknown territory \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Finally,
-LCPs are only one pixel wide, meaning that their absolute size depends on the
-resolution of chosen covariate layers \citep{Diniz.2020}. Although some of these
-issues can be alleviated using less stringent versions of the LCP algorithm
-(e.g. least-cost \textit{corridors} \citep{Pinto.2009}, \textit{thresholded}
-least-cost paths \citep{Landguth.2012}, and \textit{randomized} least-cost paths
-\citep{Panzacchi.2016, VanMoorter.2021}), a certain degree of arbitrariness in
-the assumptions remains. CT entails similar restrictions. Because CT only
-considers movements from the source cell to its 4 or 8 adjacent cells, it
-implicitly posits a perceptual range of a single pixel. It is also built on the
-assumption of a complete random walk \citep{Diniz.2020}, implying that
-directional biases cannot be rendered, albeit being very common in dispersal
-movements \citep{Cozzi.2020, Hofmann.2021}. Ultimately, both LCP analysis and CT
-omit the temporal dimension of dispersal. In result, statements about the
-expected duration required to traverse a certain corridor are impossible.
-Because movement is not modelled explicitely, neither of the methods can render
-possible interactions between movement behavior and landscape characteristics
-such that connectivity mainly arises as a result of the landscape structure.
-This is referred to as structural connectivity and stands in contrast to
-functional connectivity, which also renders the behavioral response of the
-animal with respect to prevailing habitat conditions \citep{Tischendorf.2000}.
-Even though functional connectivity is more difficult to estimate, a functional
-view is the ultimate goal in conservation science because it has direct
-consequences for gene flow \citep{Baguette.2013}.
-
-% What about IBMMs?
-\subsection{What about IBMMs? (90\%)}
-To address the issues inherent to LCPs and CT, individual-based movement models
-(IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
-dispersal is simulated explicitly, based on movement rules that determine how
-individuals move over and interact with the prevailing landscape
-\citep{Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020,
-Vasudev.2021}. Using the simulated trajectories, one can calculate a set of
-connectivity metrics, such as interpatch-connectivity and traversal frequency
-across the landscape to reveal major dispersal hotspots \citep{Kanagaraj.2013,
-BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}. Even though IBMMS can be
-employed to overcome any of the shortcomings intrinsic to LCPs and CT, as well
-as to provide a more functional view on connectivity, they can be challenging to
-fit and require vast amounts of data collected during dispersal
-\citep{Diniz.2020}.
-
-% Step Selection Analysis
-\subsection{Step Selection Analysis (90\%)}
-Here, we investigate the usefulness of integrated step selection functions
-(ISSFs, \citealp{Avgar.2016}), as a relatively simple but powerful IBMM based on
-which dispersal can be simulated. While regular SSFs were intended to learn
-about relative habitat preferences of the focal species \citep{Fortin.2005,
-Thurfjell.2014, Avgar.2017}, the method has been generalized and now enables to
-jointly study habitat and movement preferences, as well as potential
-interactions between movement and habitat preferences \citep{Avgar.2016,
-Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple method
-to model complex movement behavior where movement results from two intertwined
-behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}). Importantly, a
-parametrized ISSF model can be viewed as a fully mechanistic movement model
-based on which individual movement trajectories can be simulated
-\citep{Avgar.2016, Signer.2017}. In fact, \cite{Signer.2017} used ISSF to
-simulate steady state utilization distributions of resident animals. However,
-the degree to which such simulations are helpful in detecting movement corridors
-and modeling landscape connectivity is unknown.
-
-% Study Species & Study Area
-\subsection{Study Species \& Study Area (90\%)}
-One of the species for which long-term viability relies on sufficient
-landscape connectivity is the endangered African wild dog \textit{Lycon pictus}.
-While once present across entire sub-Saharan Africa, wild dogs have disappeared
-from a vast majority of their historic range due to persecution by humans,
-habitat destruction, and deadly diseases. As of today, only 6'000 free-ranging
-individuals remain in small and spatially scattered subpopulations
-\citep{Woodroffe.2012}. Within those subpopulations, wild dogs form cohesive
-packs comprising 8 to 12 adults and their offspring \cite{McNutt.1995}. After
-reaching sexual maturity, male and female offspring form same-sex coalitions and
-disperse from their natal pack \citep{McNutt.1996, Behr.2020}. New packs are
-formed when dispersing coalitions join unrelated opposite-sex dispersing
-coalitions \citep{McNutt.1996}. Dispersing wild dogs can cover several hundred
-kilometers across a variety of landscapes \citep{DaviesMostert.2012,
-Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the few strongholds for this
-species lies near the Moremi Game Reserve in northern Botswana, which is part of
-the world's largest transboundary protected area, namely the Kavango-Zambezi
-Transfrontier Conservation Area (KAZA-TFCA). This area has originally been
-intended to facilitate migration of elephants, but is expected to benefit a
-multitude of other species \citep{Elliot.2014, Brennan.2020, Hofmann.2021}.
-
-% Previous Paper
-\subsection{Previous Paper (90\%)}
-In a previous study, we assessed landscape connectivity for dispersing African
-wild dogs within the KAZA-TFCA using a least-cost corridor approach
-\citep{Hofmann.2021}. For this, we fitted a basic habitat selection model based
-on which we predicted landscape permeability. We now expand on this knowledge
-and use ISSF to develop a more detailed movement model of dispersing wild dogs.
-We then use this model to simulate dispersers moving across the KAZA-TFCA. Based
-on simulations, we compute heatmaps and identify potential dispersal hotspots.
-We also showcase how network metrics relevant to landscape connectivity can be
-computed. Our results show that a simulation-based approach yields several major
-benefits over traditional connectivity modeling techniques. Most importantly,
-simulations provide a more generic view on how connectivity emerges and to which
-degree connectivity depends on the dispersal duration. In addtion, by generating
-proper dispersal trajectories, network theory can be applied to calculate
-network metrics that are pertinent to connectivity analysis. Finally, we put
-forward additional opportunities using simulations that go beyond the scope of
-this paper.
+Modeling connectivity in a static environment may mask the importance of connectivity for
+maintaining wildlife population dynamics and may result in identifying corridors that are only
+periodically functional [15,16,18]
 
 \section{Methods}
-\subsection{Study Area (90\%)}
-The study area was centered at -17\degree 13'9''S, 23\degree 56'4''E
-(\Cref{StudyArea}a) and was represented by a rectangular bounding box that
-stretched over 1.3 Mio. km\textsuperscript{2}, ecompassing the entire KAZA-TFCA
-(\Cref{StudyArea}b). The KAZA-TFCA is the world's largest transboundary
-conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
-and Zambia, covering a total of 520'000 km\textsuperscript{2}. Its landscape
-varies regionally and ranges from savanna, to grassland and from dry to moist
-woodland habitats. A dominant hydrogeographical feature in its center is the
-flood-pulsing Okavango Delta. The wet season within the KAZA-TFCA lasts from
-November to March and is out of phase with the main flooding of the Okavango
-Delta which peaks between July and August \citep{McNutt.1996, Wolski.2017}.
-While large portions within the KAZA-TFCA are designated national parks or other
-protected areas, substantial human influence remains due to roads, agricultural
-sites and settlements and villages.
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_StudyArea.png}
-    \caption{(a) The study area was confined by a bounding box encompassing the
-    entire KAZA-TFCA and spanned parts of Angola, Namibia, Botswana, Zimbabwe,
-    and Zambia. (b) The KAZA-TFCA represents the world's largest terrestrial
-    conservation area and covers a total of 520'000 km\textsuperscript{2}. Its
-    purpose is to re-establish connectivity between already-existing national
-    parks (dark green) and other protected areas (light green). The dispersal
-    data used in this study was collected on a free-ranging African wild dog
-    population inhabiting the Moremi National Park in northern Botswana.}
-    \label{StudyArea}
-  \end{center}
-\end{figure}
-
-\subsection{GPS Relocation Data (90\%)}
-Between 2011 and 2019, we collected GPS relocation data on dispersing wild dogs
-from a free-ranging wild dog population inhabiting the Moremi National Park in
-northern Botswana. We selected potential dispersers based on age, pack size,
-number of same‐sex siblings within the pack, and presence of unrelated
-opposite-sex individuals in the pack \citep{McNutt.1996, Behr.2020}. We
-immobilized selected individuals using a cocktail of Ketamine/Xylazine/Atropine
-\citep{Osofsky.1996, Cozzi.2020} that was injected by dart, fired from a
-CO\textsubscript{2}-pressurized gun (\textit{DAN-Inject, Denmark}). Immobilized
-individuals were fitted with GPS/Satellite radio collars (\textit{Vertex Lite;
-Vectronic Aerospace GmbH, Berlin}) that guaranteed automated drop-off through a
-decomposable piece of cotton. Handling and collaring of all individuals was
-supervised by a Botswana-registered wildlife veterinarian. All individuals
-rejoined their pack within one hour after immobilization. 16 collared
-individuals eventually dispersed, each in a separate same-sex dispersal
-coalition (7 female and 9 male coalitions).
-
-Because behavior during dispersal is more pertinent for assessing landscape
-connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded all data that was
-collected during residency and only retained GPS data recorded during dispersal.
-During dispersal, collars were programmed to record a GPS fix every 4 hours.
-Collected relocations were regularly transmitted over the Iridium satellite
-system, which allowed remote tracking of individuals, even if they left the main
-study area and crossed international borders. In some instances, exact dispersal
-dates were known from field observations. Otherwise, we distinguished between
-residency and dispersal using the net-squared displacement metric. This metric
-measures the squared Euclidean distance of a GPS relocation to a reference point
-\citep{Borger.2012}, which in our case was set to the center of each
-individual's natal home range. As such, dispersal was deemed to have started
-when an individual left its natal home range and ended once individuals became
-sedentary again. As previous research revealed similar behavior of females and
-males during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did not
-distinguish between sexes. After collection, we converted collected GPS
-coordinates (n = 4'169) to steps, where each step represented the straight-line
-distance traveled by and individual between two consecutive GPS relocations
-\citep{Turchin.1998}. To ensure a regular sampling interval, we removed fixes
-that were not successfully collected on the 4-hourly schedule.
-
-\subsection{Covariates (90\%)}
-We represented the physical landscape across the study area using a set of
-habitat covariates including water-cover, distance to water, woodland-cover, and
-shrub/grassland-cover. Because water cover greatly changes within and between
-years in the Okavango Delta, we applied a remote sensing algorithm and generated
-frequently updated water cover layers and corresponding distance to water layers
-(see \citealp{Wolski.2017} and \citealp{Hofmann.2021}). Resulting layers thus
-temporally aligned with our dispersal events. We furthermore computed a proxy
-for human influence, depicting anthropogenic pressures stemming from
-human-density, agricultural sites, and roads. All spatial layers were coarsened
-or interpolated to a target resolution of 250 m by 250 m. Further details on the
-sources and preparation of each habitat covariate are given in
-\cite{Hofmann.2021}.
-
-Besides habitat covariates, we also computed movement metrics that we used as
-movement covariates in our models. Movement metrics were calculated for each
-step and included the step length (\textsf{sl}), its natural logarithm
-(\textsf{log(sl)}), and the cosine of the relative turning angle
-(\textsf{cos(ta)}). Because wild dogs follow a diurnal activity pattern, we also
-coded a binary variable (\textsf{LowActivity}) indicating whether a step was
-realized during periods of low wild dog activity (17:00 to 07:00 local time) or
-high wild dog activity (09:00 to 17:00 local time). Handling and manipulation of
-all data, as well as all models and simulations were implemented with the
-statistical software {\tt R}, version 3.6.6 \citep{R.2019}. Several helper
-functions were written in {\tt C++} and imported into {\tt R} using the {\tt
-Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}
-
-\subsection{Movement Model (80\%)}
-We used ISSFs to parametrize a mechanistic movement model for dispersing wild
-dogs \citep{Avgar.2016}. To conduct iSSF analysis, we paired each observed step
-with 24 random steps. An observed and its 24 random steps thus formed a stratum
-and received a unique identifier. We generated random steps by sampling random
-turning angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths
-from a gamma distirbution that was fitted to observed steps (scale = 6'308,
-shape = 0.37). Along each step, we extracted and averaged spatial covariates
-using the {\tt velox}. We also calculated the movement metrics \textsf{sl},
-\textsf{log(sl)}, and \textsf{cos(ta)}. To facilitate model convergence, we
-standardized all continuous covariates to a mean of zero and a standard
-deviation of one. Since correlation among covariates was low (\(|r| > 0.6\);
-\citealp{Latham.2011}), we retained all of them for modeling.
-
-To contrast realized steps (scored 1) and random steps (scored 0), we assumed
-that animals assigned a selection score \(w(x)\) of the exponential form to each
-step. The selection score \(w(x)\) of each step depended on its associated
-covariates (\(x_1, x_2, ..., x_n\)) and on the animal's relative selection
-strengths (i.e. preferences) towards these covariates (\(\beta_1, \beta_2, ...,
-\beta_n\)):
-
-\begin{equation}
-\label{EQ1}
-  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
-\end{equation}
-
-The probability of a step being realized was then contingent on the step's
-selection score, as well as on the selection scores of all other step in the
-same stratum:
-
-\begin{equation}
-\label{EQ2}
-  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
-  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
-\end{equation}
-
-To estimate the preferences of interest, we ran conditional logistic regression
-models in the r-package {\tt glmmTMB}. To handle multiple individuals, we
-applied the mixed effects method developed by \citep{Muff.2020}, which allows to
-fit random slopes in addition to random intercepts. Thus, we treated animal IDs
-as random effect and modeled random slopes for each covariate. We fixed the
-random intercept variance at an arbitrary high value of 10\textsuperscript{6} to
-make use of the  ``poission''-trick \citep{Muff.2020}.
-
-The movement model was based on the habitat selection model for dispersing wild
-dogs presented in \cite{Hofmann.2021}. In the original model, no interactions
-between the habitat and movement covariates were considered, so we slightly
-expanded this base model by proposing interactions between movement and habitat
-covariates. More specifically, we started with the base model and incrementally
-increased model complexity by adding all possible two-way interactions between
-habitat covariates and movement covariates. For instance, for the covariate
-\textsf{water}, we proposed the interactions \textsf{Water:log(sl)},
-\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides those combinations,
-we also proposed the interactions \textsf{sl:cos(ta)} and
-\textsf{log(sl):cos(ta)} to account for a correlation between turning angles and
-step lengths, as well as the interactions \textsf{sl:LowActivity} and
-\textsf{log(sl):LowActivity} to account for the fact that step lengths may
-differ due to wild dogs' diurnal activity pattern. To compare competing models
-and assess the most parsimonious movement model, we ran stepwise forward model
-selection based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
-
-We validated the predictive power of the most parsimonious movement model using
-k-fold cross-validation for case-control studies as presented
-\cite{Fortin.2009}. We randomly split the input data into training and testing
-data with an 80:20 ratio. Using the training data we parametrized a movement
-model based on which we predicted selection scores \(w(x)\) for all steps in the
-test data. Within each stratum we then assigned ranks 1-25 to each step based on
-predicted selection scores such that rank 1 was given to the step with the
-highest score \(w(x)\). Across all strata we determined the realized step's rank
-and calculated rank frequencies for realized steps. Finally, we computed
-Spearman's rank correlation between ranks and associated frequencies \(r_{s,
-realized}\). We replicated the entire procedure 100 times and computed the mean
-correlation coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\%
-confidence interval across all replicates. For comparison, we repeated the same
-procedure 100 times assuming random preferences. Random preferences were
-implemented by discarding the realized step from all strata and identifying the
-rank of a random step in each stratum. Again, we calculated Spearman's rank
-correlation coefficient (\(r_{s, random}\)), its mean across repetitions
-(\(\bar{r}_{s, random}\)), and its 95\% confidence interval. This validation
-proofs a significant prediction in case the confidence intervals of
-\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap.
-
-\subsection{Dispersal Simulation (80\%)}
-We used the most parsimonious movement model to simulate 80'000 dispersing wild
-dogs moving across the KAZA-TFCA. The simulation resembled an inverted iSSF and
-was set up as follows. (1) We defined a random source point and assumed a random
-initial orientation of the animal. (2) Departing from the source point, we
-generated 25 random steps by sampling relative turning angles from a uniform
-distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
-distribution. Each step corresponded to the straight line movement conducted
-within 4 hours. To prevent unreasonably large steps, we capped sampled values to
-a maximum of 35 km, which corresponds to the farthest distance ever traveled
-within 4 hours by on of our dispersers. (3) Along each random step, we extracted
-and averaged habitat covariates and calculated movement covariates. Extracted
-values were again standardized using the same transformations that were used on
-the input data. (4) We applied the parameterized movement model and predicted
-the selection score \(w(x)\) for each random step and translated selection
-scores into probabilities using \Cref{EQ2}. (5) We randomly sampled one of the
-steps based on predicted probabilities and determined the animal's new position.
-We repeated steps (2) to (5) until a total of 2'000 steps were realized.
-
-To minimize the influence of edge effects and to deal with random steps leaving
-the study area, we followed \citep{Koen.2010} and artificially expanded all
-covariate layers by adding a 100 km buffer zone. Within the buffer zone, we
-randomized covariate values by resampling values from the original covariate
-layers. In result, simulated dispersers were able to leave and reenter the main
-study area through this buffer zone. In cases where proposed random steps
-transgressed the outer border of this buffer zone, we resampled transgressing
-steps until they fully lied within the buffer.
-
-\subsection{Source Points (90\%)}
-We initiated virtual dispersers at 50'000 randomly selected source points within
-contiguous protected areas larger \(>\) 700 km\textsuperscript{2}
-(\Cref{SourcePoints}a). This conforms to the average home range requirement of
-resident wild dogs \citep{Pomilia.2015} and allowed us to remove areas too small
-to host viable wild dog populations. By distributing source points randomly, the
-number of source points per km\textsuperscript{2} within protected areas was
-approximately equal across the study area. To render potential immigrants into
-the study system, we placed additional 30'000 source points inside the buffer
-zone around the main study area (\Cref{SourcePoints}b). This resulted in a total
-of 80'000 source points, each representing the start point of an individual
-disperser.
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_SourceAreas.png}
-    \caption{(a) Different source areas from which we released virtual
-    dispersers. We only considered contiguous protected areas (national parks,
-    game reserves, and forest reserves) that were larger than 700
-    km\textsuperscript{2} (green). This area corresponds to the average home
-    range requirement for viable wild dog populations \citep{Pomilia.2015}. To
-    render potential immigrants into the study system, we also initiated
-    dispersers within a buffer zone (blue) surrounding the main study area. (b)
-    Source points from which dipsersers were released. 50'000 dispersers were
-    released from the main study area (green dots) and another 30'000 dispersers
-    within the virtual buffer (blue dots).}
-    \label{SourcePoints}
-  \end{center}
-\end{figure}
-
-\subsection{Heatmap (100\%)}
-To identify dispersal hotspots across our study area, we created a heatmap
-indicating the absolute frequency at which each raster-cell in the study area
-was visited by virtual dispersers. For this, we rasterized all simulated
-trajectories and tallied them into a single map. If the same trajectory crossed
-a raster-cell twice, it was only counted once. This way, we did not consider
-revisits and mitigated biases arising from trapped individuals that were moving
-in circles. To achieve high performance rasterization, we used the R-package
-{\tt terra} \citep{Hijmans.2020}.
-
-\subsection{Betweenness (80\%)}
-To pinpoint areas of exceptional relevance for connecting remote regions in our
-study area, we converted simulated trajectories into a network and calculated
-betweenness scores. For this, we overlaid the study area (including the buffer)
-with a regular raster resolved at 5 x 5 km. We then used the simulated
-trajectories to determine all transitions occurring from one raster-cell to
-another, as well as the frequency at which those transitions occurred. This
-resulted in an edge-list that we translated into a weighted network using the
-r-package {\tt igraph} \citep{Gabor.2006}. Because {\tt igraph} handles edge
-weights (\(\omega\)) as costs, we inverted the traversal frequency in each cell
-by applying \(\omega = \frac{\sum_i^n{Traversal Frequency_i}/n}{Traversal
-Frequency_i}\). Finally, we used the weighted network to calculate betweenness
-scores for each raster-cell in the overlaid raster. The betweenness metric
-indicates how often a specific raster-cell lies on a shortest path between two
-other raster-cells and is therefore a useful metric to detect movement corridors
-\citep{BastilleRousseau.2018}.
-
-\subsection{Interpatch-Connectivity (80\%)}
-We also assessed inter-patch connectivity between national parks in our study
-area. The decision to focus on national parks was purely out of simplicity and
-the same logic could easily be expanded to include other protected areas as
-well. To quantify inter-patch connectivity, we computed the relative frequency
-at which dispersers originating from one national park successfully moved into
-another national park. This allowed us to determine \textit{if} and \textit{how
-often} dispersers moved between certain national parks. Moreover, because time
-was explicit in our IBMM, we were able to estimate \textit{how long} dispersers
-had to move to realize those connections.
 
 \section{Results}
-\subsection{Movement Model (20\%)}
-Compared to the base model reported in \citep{Hofmann.2021}, our most
-parsimonious movement model included several additional interactions between
-habitat covariates and movement covariates (\Cref{MovementModel} and
-\Cref{MovementModelNumbers}). Although multiple models received an AIC weight
-above zero (Table 1 in Appendix S1), we only considered results from the most
-parsimonious model. Since all models with positive AIC weight contained similar
-covariates, this decision only marginally influenced subsequent results. Results
-from the selected model are given in \Cref{MovementModelNumbers} and illustrated
-in \Cref{MovementModel} (a). Additional plots that facilitate the interpretation
-of the model are provided in Appendix S2.
 
-When looking at the habitat kernel and holding constant the movement kernel, we
-find that dispersing wild dogs avoid water but prefer its proximity. Dispersers
-also avoid densely forested woodlands, yet prefer open shrublands or grasslands.
-Finally, dispersers avoid moving through landscapes that are influenced by
-humans. These results align with our previous findings reported in
-\cite{Hofmann.2021}.
-
-When looking at the movement kernel, we observe several significant estimates.
-However, except for the interaction \textsf{sl:LowActivity}, effect sizes are
-relatively small, suggesting that our proposal distributions for step lengths
-and turning angles were only marginally biased. For instance, the positive and
-significant effect for \textsf{cos(ta)} indicates that realized turning angles
-were slightly more directional than the turning angles proposed by our uniform
-distribution, implying that realized steps in fact followed a von Mises
-Distribution with positive concentration parameter. On the other hand, the
-significant and negative interaction for \textsf{sl:LowActivity} reveals that
-wild dogs moved over shorter distances during low wild dog activity compared to
-the steps suggested by our gamma distribution.
-
-Finally, we turn to the interactions between the two kernels. Here, we observe
-that movement behavior significantly differs depending on habitat covariates. In
-general, steps tend to be shorter in areas with high water cover and high
-woodland cover, yet larger in areas with open shrubs and grasslands. Similarly,
-dispersers realized shorter steps in the vicinity to water, even though the size
-of this effect is negligible. Finally, it appears that dispersers move more
-tortuous in areas influenced by humans, and more directional when far from
-water.
-
-As can be seen in \Cref{MovementModel}b, the k-fold cross-validation procedure
-reveals that our model substantially outperforms a random guess, as indicated by
-the fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
-\(\bar{r}_{s, random}\) do not overlap. Furthermore, the significant correlation
-between ranks and corresponding frequencies for realized steps indicates a good
-fit between predictions and observations (\Cref{MovementModel}b). Finally, we
-find that the rank correlation slightly improved in comparison to the base model
-(\(\bar{r}_{s, realized} = -0.55\); \citealp{Hofmann.2021}).
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\textwidth]{99_MovementModel}
-    \caption{(a) Most parsimonious movement model for dispersing wild dogs. The
-    model includes a habitat kernel, a movement kernel, as well as their
-    interactions. The orange and blue line segments delineate the 90\%, 95\%,
-    and 99\% Confidence-Intervals around the respective \(\beta\) coefficients.
-    Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
-    Results from the k-fold cross validation procedure. The upper plot shows
-    rank frequencies of predicted realized scores according to model predictions
-    with known preferences, whereas the lower plot shows rank frequencies when
-    assuming random preferences. The blue ribbon shows the prediction interval
-    around a loess smoothing regression that we fitted to ease the
-    interpretation of the plots.}
-    \label{MovementModel}
-  \end{center}
-\end{figure}
-
-\begin{table}
-  \begin{center}
-  \caption{Most parsimonious movement model for dispersing wild dogs. The model
-  comprises of a movement and habitat kernel, where the movement kernels
-  describes preferences with regards to movement behavior, whereas the habitat
-  kernel describes preferences with respect to the habitat. Finally, the two
-  kernels can interact, such that movement preferences are contingent on habitat
-  conditions.}
-  \label{MovementModelNumbers}
-  \resizebox{\textwidth}{!} {
-    \begin{threeparttable}
-      \begin{tabular}{llrrrc}
-        \toprule
-        Kernel & Covariate & Coefficient & SE & p-value & Sign. \\
-        \midrule
-        \multirow{5}{*}{Habitat Kernel}
-         & Water & -0.546 & 0.112 & \(<\) 0.001 & *** \\
-         & DistanceToWater \textsuperscript{0.5} & -0.390 & 0.231 & 0.092 & * \\
-         & Woodland & -0.364 & 0.086 & \(<\) 0.001 & *** \\
-         & Shrubs/Grassland & 0.288 & 0.092 & 0.002 & *** \\
-         & HumanInfluence & -0.535 & 0.229 & 0.019 & ** \\
-        \hdashline
-        \multirow{6}{*}{Movement Kernel}
-         & sl & 0.075 & 0.037 & 0.042 & ** \\
-         & cos(ta) & 0.105 & 0.031 & 0.001 & *** \\
-         & log(sl) & 0.146 & 0.051 & 0.004 & *** \\
-         & cos(ta) : sl & 0.049 & 0.026 & 0.064 & * \\
-         & cos(ta) : log(sl) & 0.076 & 0.026 & 0.003 & *** \\
-         & sl : LowActivity & -0.917 & 0.113 & \(<\) 0.001 & *** \\
-        \hdashline
-        \multirow{5}{*}{Interaction}
-         & sl : Water & -0.305 & 0.076 & \(<\) 0.001 & *** \\
-         & sl : Woodland & -0.089 & 0.039 & 0.023 & ** \\
-         & sl : Shrubs/Grassland & 0.124 & 0.058 & 0.032 & ** \\
-         & sl : DistanceToWater \textsuperscript{0.5} & -0.058 & 0.031 & 0.056 & * \\
-         & cos(ta) : HumanInfluence & -0.040 & 0.022 & 0.070 & * \\
-         & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063 & 0.026 & 0.017 & ** \\
-         \bottomrule
-      \end{tabular}
-       \begin{tablenotes}
-         \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
-         \quad *** \(p < 0.01\)}
-       \end{tablenotes}
-    \end{threeparttable}
-    }
-  \end{center}
-\end{table}
-
-\subsection{Dispersal Simulation (60\%)}
-On a machine with an AMD Ryzen 7 2700X processor with 8 x 3.6 GHz and 64 GB of
-RAM, a single batch of 1'000 simulated dispersers took roughly 90 minutes to
-compute (\(\mu = 88.90\), \(\sigma = 1.87\)). As such, the simulation of all
-80'000 dispersers terminated after 120 hours, i.e. five days. Comparable
-computations will be substantially quicker for smaller study areas or lower
-resolution covariates, as the covariate extraction from large rasters was
-computationally the most expensive task.
-
-On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
-2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
-dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)). Simultaneously,
-simulated dispersers moved marginally less directional (\(\mu_{cos(ta)} =
-0.057\), \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers
-(\(\mu_{cos(ta)} = 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in
-step lengths and turning angles can be attributed to minor disparities between
-habitat conditions at the area within which we collected training data and
-habitat conditions within the entire study area.
-
-Out of the 50'000 dispersers initiated in protected areas, only 4.5\% eventually
-hit a map boundary, suggesting that biases due to boundary effects should be
-minimal. In contrast, 78\% of the 30'000 dispersers originating from the buffer
-zone eventually hit a map boundary.
-
-\subsection{Heatmap (50\%)}
-\Cref{Heatmap} depicts the heatmap of all 80'000 simulated trajectories rendered
-after 2'000 steps. The map highlights that large portions of land beyond the
-borders of the KAZA-TFCA are only infrequently visited by dispersers (dark blue
-areas), whereas within the KAZA-TFCA we observe several extensive regions that
-are regularly visited (bright yellow and red areas). Most notably, the region in
-northern Botswana south of the Linyanti swamp stands out as highly frequented
-dispersal hub. Still, the presence of several massive water bodies, such as the
-Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in
-considerable dispersal barriers that limit realized connectivity within the
-KAZA-TFCA. Similarly, dispersal appears to be very limited in Zambia's and
-Zimbabwe's part of the KAZA-TFCA, where only few areas are traversed by
-dispersers, which can largely be attributed to a high human influence in these
-regions. Outside the KAZA-TFCA, the most frequently traversed regions include
-the areas of the Central Kalahari National Park in Botswana, the area south-west
-of the Khaudum National Park in Namibia, and the area around the Liuwa Plains
-National Park in Zambia.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_Heatmap.png}
-  \caption{Heatmap of all 80'000 simulated trajectories after 2'000 steps.
-  Simulations were based on an integrated step selection model, fit to the
-  movement data of dispersing African wild dogs. To generate the heatmap, we
-  rasterized and summed all simulated trajectories. Additional heatmaps
-  showcasing the traversal frequency after different numbers of steps are
-  provided in xx in Appendix S3.}
-  \label{Heatmap}
-\end{figure}
-
-\subsection{Betweenness (50\%)}
-Betweenness scores of each raster-cell the study area are presented in
-\Cref{Betweenness}. While the heatmap emphasized frequently visited areas, the
-map of betweenness reveals rather discrete dispersal corridors. As can be seen,
-the dispersal hotspot in northern Botswana is traversed by a corridor that
-receives a comparably high betweenness score. This implies that the region is
-particularly crucial for connecting other regions in the study system and hence
-represents a proper network hub. Towards east, the corridor runs through the
-Chobe National Park into the Hwange national park, where it branches out and
-further extends into the distant Matusadona National Park in Zimbabwe. Northwest
-of the Linyanty ecosystem, the same corridor expands into Angola, where it
-splits and finally traverses over a long stretch of unprotected area into the
-Kafue National Park in Zambia. Several additional corridors with slightly lower
-betweenness scores exist, yet most of them run within the boundaries of the
-KAZA-TFCA. In general, only few corridors directly link the peripheral regions
-of the KAZA-TFCA. For instance, there are only few corridors between the
-Matusadona National Park in Zimbabwe and the Kafue National Park in Zimbabwe.
-Similarly, there are no direct links between the Zimbabwean and Angolan
-``spikes'' of the KAZA-TFCA.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_Betweenness.png}
-  \caption{Betweenness scores of each raster cell in a raster with 5 x 5 km
-  resolution. Betweenness scores were determined based on simulated dispersal
-  events. A high betweenness score highlights cells that are exceptionally
-  relevant in connecting different regions in the study area. That is, the
-  higher the betweenness score, the more often a pixel lies on a shortest path
-  between adjacent areas. In this sense the metric can be used to pinpoint
-  discrete movement corridors. Note that we square-rooted betweenness scores to
-  improve visibility of corrdiors with low scores.}
-  \label{Betweenness}
-\end{figure}
-
-\subsection{Interpatch Connectivity (0\%)}
-The results of the analysis of interpatch connectivity is given in
-\Cref{AreasReached}. The map shows between which national parks direct links
-exist and how frequent they are, as well as the average duration a disperser had
-to move to realize those links. For instance, 6.8 \% of the simulated dispersers
-originating from the Moremi National Park successfully reached the Chobe
-National Park and 4.2 \% reached the Hwange National Park in Zimbabwe. On
-average, dispersers moved for 623 steps before arriving at Chobe (SD = 520) and
-for 1'413 steps before arriving at Hwange (SD = 371).
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_AreasReached.png}
-  \caption{Network on simulated dispersal trajectories highlighting the
-  connections between national parks (dark green). Yellow bubbles represent the
-  center of the different national parks and are sized in relation to the number
-  of simulated dispersers originating from each park. Colored arrows between
-  national parks illustrate between which national parks the simulated
-  dispersers successfully moved and the color of each arrow shows the average
-  number of steps that were necessary to realize those connections.
-  Additionally, the line thickness indicates the relative number of dispersers
-  originating from a national park that realized those connections.}
-  \label{AreasReached}
-\end{figure}
-
-\section{Discussion (10\%)}
-Our connectivity network further suggests that dispersers from the Okavango
-Delta more likely disperse towards east than west. Indeed, only x out of our y
-observed dispersers ever reached the western part of the delta. Only when the
-flood retracts a small pathway between the city of Maun and the floodwaters of
-the delta emerges and enables dispersers to move towards the detal's western
-part.
-
-All of our findings are well in line with our previous work, where we have
-highlighted dispersal corridors for wild dogs using least-cost approaches. This
-suggests that qualitative results are quite insensitive to the exact
-methodological approach. Still, we believe that a simulation-based approach
-offers possibilities for much richer inferences compared to traditional
-approaches. This is largely due to the fact that proper movement trajectories
-are generated that can be analysed \textit{as if} they were generated by real
-dispersers. This is of course contingent on the assumption that underlying
-models are adequately representing movement behavior of the focal species and
-calls for further methods to validate the predictive power of such simulations.
-
-While the segment running into Kafue receives a high betweenness score, it was
-actually only rarely traversed by our simulated dispersers, as can be seen from
-the dark colors in this region in \Cref{Heatmap}. It is therefore worth noting
-that the betweenness metric highlights crucial bottlenecks that are relevant for
-connecting remote regions, yet does not directly yield information about the
-absolute frequency at which these bottlenecks are used.
-
-We have previously attributed the weak significance of distance to water to the
-fact that we did not control for the presence or absence of conspecifics. We
-stick to this reasoning as our expanded model still shows a rather large
-uncertainty around the respective beta coefficients. To better gauge the
-imporance and influence of this covariate, future studies will need to control
-for inter- and intra-speficic interactions that may explain why and when
-dispersers are attracted to or afraid of waterbodies.
-
-For our simulations, we represented the Okavango Delta statically and assumed a
-relatively extended flood. This resulted in a quasi-barrier, formed by the
-Okavango-Delta and the adjacent city of Maun, which was rarely traversed by
-simulated individuals. Out of the 499 dispersers initiated inside the Moremi
-National Park, only 101 managed to reach the south-western section of the Delta,
-whereas 284 eventually reached the equally distant Linyanti swamp. In this
-regard, the heatmap presented in \Cref{Heatmap} may be most representative of
-the period shortly after the wet-season, when floodlevels in the Delta are at
-their maximum. During the dry season, however, the flood considerably retracts
-and potentially clears the way for wild dogs dispersing from the Moremi-Game
-reserve into the south-western section of the Delta. Future studies could relax
-the assumption of a static flood and attempt to update floodlevels as the
-dispersers move. This would allow studying how connectivity within the ecosystem
-evolves over time as the flood climaxes and retracts again. In fact, one of the
-major advantages of such simulation-based approaches is that a dynamic
-environment can be rendered, as time is explicit in these models
-\citep{Zeller.2020}. This contrasts with traditional modeling approaches such
-least-cost analysis or circuit theory, where the temporal dimension cannot be
-made explicit. An explicit view on time also directly translates in insights on
-the duration required by dispersers to move between distinct patches such as
-national parks or spatially segregated subpopulations.
-
-Comparable simulations that are based on empirical data are also a fundamental
-component for spatially realistic population models in which dispersal is
-rendered more realistically and does not merely depend on the distance between
-habitat patches.
-
-We did not model mortality during dispersal in our simulations. This was a
-simplifying assumption and only inaccurately reflects dispersal in reality. It
-is well known that wild dogs often die during dispersal in result of deadly
-encounters with other predators, but also due to road kills and illegal
-shootings by humans. In result, mortality during dispersal may limit realized
-connectivity, especially in areas that wild dogs are unfamiliar with or in areas
-with high potential for human-wildlife conflict \citep{Cozzi.2020}.
-Interestingly though, \cite{Behr.2021} have recently demonstrated that mortality
-during dispersal is lower compared to residency, suggesting the presence some
-fitness benefits to dispersing individuals.
-
-In this regard, our approach is rather similar to dispersal kernels, yet it
-enables to render directional biases, which are currently difficult to implement
-using such methods.
-
-We completely randomized the location of source points within protected areas.
-However, in some cases prior knowledge about the density of potential dispersers
-is available and can be used to adjust the number of simulated individuals
-accordingly. Alternatively, instead of tweaking the number of simulated
-individuals, one could assign a weight to each trajectory that depends on the
-density of potential dispersers in the source areas. As such, trajectories from
-areas with high density would enter the heatmap with above average weight.
-
-The parametrized movement model could also be manipulated to investigate how
-different habitat preferences influence landscape connectivity and to test the
-sensitivity of results with respect to the exact preferences of individuals.
-
-One of the major benefits of individual-based simulations is the ability to make
-the temporal dimension of movement explicit. This allows to investigate how
-connectivity depends on the dispersal duration, something that is not possible
-with traditional least-cost or circuit theory methods.
-
-Optimally, one should simulate additional dispersers until the amount of newly
-gained information lies beyond a certain threshold, i.e. until some sort of
-convergence is achieved. However, due to the myriad of outcomes in movement
-behavior, this would often imply countless repetitions for each source points at
-each location where dispersal is possibly initiated. This is computationally
-infeasible.
-
-Our work suggests that the selection of source points significantly impacts
-resulting connectivity networks. Especially when dispersal durations are short,
-wrongly placed source points lead to vastly different results. Signer et al.
-used estimated utilisation distributions by means of simulated movements. They
-used a rather long burn in period prior to alleviate the problem of selecting
-meaningful source points. However, this approach only works when individuals
-move around a point of attraciton. This is typically not the case when
-simulating dispersers, introducing an important trade-off. The researcher can
-decide to increase the number of simulated steps, hence reducing the influence
-of starting locations, yet this also inevitably increases estimated
-connectivity.
-
-In some European countries, the comeback of large predators, such as bears,
-lynx, and the wolf, has triggered emotional discussions and raised public
-concern \citep{Behr.2017}, particularly in areas with free-roaming livestock
-that may be prayed upon by the returned species. In cases were recent locations
-of such predators are known, an early warning system based on simulations could
-serve to forewarn about potential encounters and and thereby increase public
-acceptance of large predators.
-
-Future studies could also investigate the sensitivity of such simulations with
-respect to estimated habitat preferences. Here, we used point estimates and did
-not vary degree by which individuals preferred or avoided certain covariates.
-Nevertheless, the degree to which our results depend on these estimates is
-unknown. Because data are usually scarce for dispersal studies on endangered
-species, point estimates may be inaccurate \citep{Wiegand.2003,
-KramerSchadt.2007}.
+\section{Discussion}
 

### 2021-06-24 Chapter_2/04_Manuscript/Manuscript_Support.tex
various updates

--- 2021-05-05 Chapter_1/04_Manuscript/Manuscript_Support.tex
@@
 %------------------------------------------------------------------------------
-%	Appendix S2: Evolution of Heatmaps
+%	Appendix S2: Convergence
+%------------------------------------------------------------------------------
+\newpage
+\section{Convergence}
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_Convergence.png}
+    \caption{}
+    \label{Trajectories}
+  \end{center}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%	Appendix S3: Evolution of Heatmaps
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S3: Evolution of Betweenness Maps
+%	Appendix S4: Evolution of Betweenness Maps
 %------------------------------------------------------------------------------
@@
 
+\section{Step Selection Summary}
+Resource Selection Functions (RSFs) and Resource Selection Probability Functions
+(RSPFs) are routinely used to model habitat selection by animals using data from
+Very High Frequency (VHF) and Global Positioning System (GPS) locations. A
+RS(P)F is defined as any statistical model deployed to estimate the relative
+probability of selecting a resource unit versus alternative possible resource
+units \citep{Manly.2002}.
+
+Compared to a RSF, a RSPF yields the actual probability that an available
+resource unit is selected and can be estimated using weighted distribution
+theory \citep{Lele.2006}.
+
+Compared to RSFs, the key feature of SSFs is link- ing consecutive animal
+locations (most commonly taken at regular time intervals) that can be defined as
+steps \citep{Turchin.1998} (Figure 1).
+
+With step selection functions, GPS trajectories are broken down into steps, each
+representing the straight line segment between two consecutive GPS relocations.
+Each observed step is then paired with a set of random steps, having the same
+starting point but differing in length and orientation \citep{Fortin.2005}. To
+sample step lengths and orientations (i.e. turning angles), \cite{Fortin.2005}
+sampled values from observed steps. Later studies, on the other hand, rely on
+parametrized distributions, fitted to observed steps. By comparing
+characteristics at observed and random steps using conditional logistic
+regression, features influencing elk movements can be determined. The step
+selection function takes the form \citep{Fortin.2005}:
+
+\begin{equation}
+\label{EQ1}
+\hat{w}(x) = exp(\beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3 + ... + \beta_n x_n)
+\end{equation}
+
+Where coefficients \(\beta_1\) to \(\beta_n\) are estimated using conditional
+logistic regression and are associated with environmental covariates \(x_1\) to
+\(x_n\), respectively. Ultimately, steps with higher \(\hat{w}(x)\) have higher
+\textbf{odds} of being chosen.
+
+As the time interval between relocations increases, it can be expected that
+autocorrelation in the data decreases.
+
+How to state results:
+\begin{itemize}
+  \item  Assuming that other environmental factors remain constant, animals were
+  less likely to ...
+  \item Animals generally avoided steps in areas made up of a large proportion
+  of ...
+  \item  An increase in ... did not generally decrease the odds of animals
+  choosing these steps, as indicated by the positive association between ... and
+  the probability of animals making steps covering the two most widely available
+  vegetation cover types
+\end{itemize}
+
 \newpage

### 2021-06-24 Chapter_1/04_Manuscript/Manuscript.tex
various updates

@@
 \begin{abstract}
-Dispersal is an important process that allows species to avoid inbreeding, to
-colonize new habitats and to reinforce non-viable subpopulations. Successful
+Dispersal is an important process that allows species to avoid inbreeding,
+colonize new habitats and reinforce non-viable subpopulations. Successful
 dispersal thus represents a crucial pre-requisite for long-term species
@@
 contingent a sufficient degree of landscape connectivity, which is why the
-estimation of connectivity and identification and preservation of dispersal
-corridors has become a task of extraordinary importance for conservation
-authorities worldwide.
+estimation of connectivity and preservation of dispersal corridors has become a
+task of extraordinary importance for conservation authorities worldwide.
 
@@
 ecological applications, both methods make several restricting assumptions that
-limit their suitability in reality. To address these shortcomings, dispersal
-simulations from individual-based movement models have been proposed and
-applied. Yet, due to the almost infinite number of non-trivial decisions a
-modeler faces when parametrizing such models, a unified and objective framework
-is missing.
+limit their suitability in reality. Individual-based dispersal simulations have
+been proposed to address these shortcomings, yet due to the sheer amount of
+non-trivial modeling decisions decisions required, a unified and objective
+framework to simulate dispersal is missing.
 
@@
 simulated. While such models have been applied to study \textit{steady-state}
-utilization distribution resident animals, a similar approach may be useful for
-investigating \textit{transient} movement behavior and study landscape
-connectivity.
+utilization distribution resident animals, a similar approach may be useful to
+investigate \textit{transient} movement behavior and landscape connectivity.
 
@@
 coalitions in combination with relevant habitat covariates. We analyse the data
-using integrated step selection functions and parametrize a fully mechanistic
-model describing how dispersing wild dogs move through the landscape. Based on
-this model, we simulate 80'000 dispersers moving across the extent of the
-KAZA-TFCA. We then generate a set of maps, each focused on a different aspect of
-connectivity within the KAZA-TFCA. Finally, we discuss the benefits and pitfalls
-of such a simulation-based approach and highlight potential improvements to be
-made in the future.
+using integrated step selection functions, thereby parametrizing a fully
+mechanistic movement model describing how dispersing wild dogs move through the
+landscape. Based on this model, we simulate 80'000 dispersers moving across the
+extent of the KAZA-TFCA. and generate a set of maps, each focused on a different
+aspect of connectivity. Finally, we discuss the benefits and pitfalls of such a
+simulation-based approach and highlight potential improvements to be made in the
+future.
 
@@
 population dynamics at different spatial and social scales \citep{Hanski.1999a,
-Clobert.2012}. Dispersal allows species to avoid inbreeding and maintain genetic
-diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
-Baguette.2013}, to rescue small, non-viable populations \citep{Brown.1977}, and
-to promote the colonization of unoccupied habitats \citep{Hanski.1999b,
+Clobert.2012}. Dispersal allows species to maintain genetic diversity
+\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013}, to
+rescue small, non-viable populations \citep{Brown.1977}, and to promote the
+colonization or recolonization of unoccupied habitats \citep{Hanski.1999b,
 MacArthur.2001}. However, successful dispersal requires a sufficient degree of
@@
 fundamental task in conservation science \citep{Nathan.2008, Doerr.2011,
-Rudnick.2012}. Importantly, the ability to pinpoint relevant dispersal hotspots
-requires information on movement behavior during dispersal and knowledge about
-factors that limit dispersal and therefore connectivity \citep{Baguette.2013,
+Rudnick.2012}. The ability to pinpoint relevant dispersal hotspots requires
+information on movement behavior during dispersal and knowledge about factors
+that limit dispersal and therefore connectivity \citep{Baguette.2013,
 Vasudev.2015}.
@@
 input to a connectivity model that is used to reveal movement corridors
-\citep{Diniz.2020}. In this regard, two of the most prominent connectivity
-models are least-cost path analysis (LCP analysis; \citealp{Adriaensen.2003})
-and circuit theory (CT \citealp{McRae.2006, McRae.2008}), both graph-based
-methods that estimate conductance of the landscape. Despite their intuitive
-nature and ease of use, both methods make rigorous assumptions about animal
-movement that are often not fulfilled in reality \citep{Diniz.2020}.
+\citep{Diniz.2020}. Two of the most prominent connectivity models are least-cost
+path analysis (LCP analysis; \citealp{Adriaensen.2003}) and circuit theory (CT;
+\citealp{McRae.2006, McRae.2008}), both graph-based methods that estimate
+conductance of the landscape to infer likely movement corridors. Despite their
+intuitive nature and ease of use, both methods make rigorous assumptions about
+animal movement that are often not fulfilled in reality \citep{Diniz.2020}.
 
@@
 perceptual range and a preconceived end-point in mind, such that they choose a
-cost-minimizing route accordingly. These assumptions may be reasonable for
-migrating animals, yet they are unlikely to hold true for dispersers, which
-typically move over long distances into unknown territory \citep{Koen.2014,
-Abrahms.2017, Cozzi.2020}. Another issue of LCPs analysis concerns the fact that
-least-costly routes, by their very nature, are only one pixel wide. This implies
-that the width of inferred movement routes depends on the resolution of chosen
-covariate layers and may not be biologically meaningful \citep{Diniz.2020}.
-Although some of these deficiencies can be addressed using less stringent
-versions of the LCP algorithm (e.g. least-cost \textit{corridors}
-\citep{Pinto.2009}, \textit{thresholded} least-cost paths \citep{Landguth.2012},
-and \textit{randomized} least-cost paths \citep{Panzacchi.2016,
-VanMoorter.2021}), a certain degree of arbitrariness remains.
+cost-minimizing route accordingly. These assumptions may be fulfilled by
+migrating animals that typically move between a discrete set of habitats through
+familiar landscapes. Dispersers, on the other hand, usually move over long
+distances into unknown territory and are therefore less likely to be aware of
+associated movement costs \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Another
+issue of LCPs analysis concerns the fact that least-costly routes, by their very
+nature, are only one pixel wide \citep{Pinto.2009}. This neglects the fact that
+alternative routes with similar costs may exist and implies that the width of
+inferred movement routes depends on the resolution of chosen covariate layers
+and may not be biologically meaningful \citep{Diniz.2020}. Although some of
+these deficiencies can be addressed using less stringent versions of the LCP
+algorithm (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
+\textit{thresholded} least-cost paths \citep{Landguth.2012}, and
+\textit{randomized} least-cost paths \citep{Panzacchi.2016, VanMoorter.2021}), a
+certain degree of arbitrariness remains.
 
 \subsection{Issues with Circuit Theory (90\%)}
-CT entails similarly unreasonable restrictions that are hardly ever met. Because
-CT only allows movements from a source cell to its 4 or 8 adjacent cells, it
-implicitly posits a perceptual range of a single pixel. Assuming that covariate
-layers are usually resolve between 30 m and 1 km, this hardly ever renders the
-true capability of animals to perceive the environment. Moreover, CT builds on
-the assumption of a complete random walk \citep{Diniz.2020}, implying that
-directional biases cannot be rendered. Nevertheless, directionality is a
-commonly observed characteristic in animal movement \citep{Bovet.1991,
-Schultz.2001}, especially in dispersing individuals \citep{Cozzi.2020,
-Hofmann.2021}.
+CT entails similarly unreasonable restrictions that are hardly ever met. For
+example, because CT only allows movements from a source cell to its 4 or 8
+adjacent cells, it implicitly posits that animals exhibit a perceptual range of
+a single pixel. Given that covariate layers are usually resolved with a pixel
+size between 30 m x 30 m and 1 km x 1 km, this hardly ever renders the true
+capability of animals to perceive the environment. Moreover, CT is built around
+the assumption of a complete random walk \citep{Diniz.2020}, entailing that
+directional biases cannot be rendered. Nevertheless, directionality is a common
+characteristic in animal movement \citep{Bovet.1991, Schultz.2001}, especially
+in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}.
 
@@
 Finally, neither LCP analysis nor CT are capable of rendering the temporal
-dimension of dispersal. Statements about the expected duration required to
-traverse a certain corridor are therefore impossible. Likewise, because movement
-is not modeled explicitly, neither of the methods allows to render interactions
-between movement and habitat preferences of the focal species. Connectivity
-therefore merely arises in result to the landscape structure, which is usually
-referred to as structural connectivity. While structural connectivity yields
-insights in the \textit{potential} of the landscape to be traversed, it does not
-enable to quantify the \textit{actual} gene flow through the area. Consequently,
-a functional on connectivity, which also renders the behavioral response of the
-animal with respect to prevailing habitat conditions, is often more desirable
-\citep{Tischendorf.2000, Baguette.2013}.
+dimension of dispersal \citep{Diniz.2020}. Statements about the expected
+duration required to traverse a certain corridor are therefore impossible.
+Likewise, because movement is not modeled explicitly, interactions between
+movement and habitat preferences of the focal species cannot be rendered.
+Connectivity therefore merely arises in result to the landscape structure, which
+is usually referred to as structural connectivity. While structural connectivity
+yields insights in the \textit{potential} of the landscape to be traversed, it
+does not enable to quantify the \textit{actual} gene flow through the area.
+Consequently, a functional view on connectivity, which also renders the
+behavioral response of the animal with respect to prevailing habitat conditions,
+is often more desirable \citep{Tischendorf.2000, Baguette.2013}.
 
@@
 (IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
-dispersal is simulated spatially explicitly, based on movement rules that
-determine how individuals move over and interact with the prevailing landscape
-\citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004,
-Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019,
-Zeller.2020, Vasudev.2021}. Using the simulated trajectories, one can calculate
-a set of connectivity metrics, such as inter-patch connectivity and traversal
-frequency to reveal major dispersal corridors \citep{Kanagaraj.2013,
-BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}. However, while IBMMS can
-be employed to overcome any of the shortcomings intrinsic to LCPs and CT and
-provide a more functional view on connectivity, they can be challenging to fit
-and require vast amounts of data collected during dispersal \citep{Diniz.2020}.
-Moreover, most individual-based models are heavily influenced by subjective
-decisions, rendering replication and validation of results difficult.
+dispersal trajectories are simulated spatially explicitly, based on movement
+rules that determine how individuals move over and interact with the prevailing
+landscape \citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004,
+Revilla.2004, Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016,
+Hauenstein.2019, Zeller.2020, Vasudev.2021}. Using the simulated trajectories,
+one can calculate a set of connectivity metrics, such as inter-patch
+connectivity and traversal frequency, to reveal major dispersal corridors
+\citep{Kanagaraj.2013, BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}.
+However, while IBMMS can be employed to overcome any of the shortcomings
+intrinsic to LCPs and CT, they are subject to a vast amount of subjective,
+non-trivial modeling decisions. Moreover, they can be challenging to fit and
+require vast amounts of movement data, ideally collected during dispersal
+\citep{Diniz.2020}. Consequently, alternative methods that require fewer
+modeling decisions and are straight forward to apply are desirable.
 
@@
 about relative habitat preferences of the focal species \citep{Fortin.2005,
-Thurfjell.2014, Avgar.2017}, the method has been generalized and now enables to
-jointly study habitat and movement preferences, as well as potential
-interactions between movement and habitat preferences \citep{Avgar.2016,
-Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple method
-to model complex movement behavior where movement results from two intertwined
-behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}). Importantly, a
-parametrized ISSF model can be viewed as a fully mechanistic movement model
-based on which individual movement trajectories can be simulated
+Thurfjell.2014, Avgar.2017}, the method has recently been generalized to
+\textit{integrated} SSFs and now enables to jointly study habitat and movement
+preferences, as well as potential interactions between them \citep{Avgar.2016,
+Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple means to
+model complex movement behavior, where movement is viewed as the result of two
+intertwined behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}).
+Importantly, a parametrized ISSF model can be employed as a fully mechanistic
+movement model based on which individual movement trajectories can be simulated
 \citep{Avgar.2016, Signer.2017}. In fact, \cite{Signer.2017} used ISSF to
-simulate steady state utilization distributions of resident animals. However,
-the degree to which such simulations are helpful in detecting movement corridors
-and modeling landscape connectivity is unknown.
+simulate steady state utilization distributions of resident animals that were
+moving around a point of attraction. However, the degree to which such
+simulations are helpful in detecting movement corridors and modeling landscape
+connectivity remains to be investigated.
 
@@
 \cite{McNutt.1995}. After reaching sexual maturity, male and female offspring
-form same-sex coalitions and disperse from their natal pack \citep{McNutt.1996,
+form same-sex coalitions and disperse from their natal pack in search for
+potential mating partners and a suitable territory to settle \citep{McNutt.1996,
 Behr.2020}. New packs are formed when dispersing coalitions join unrelated
@@
 area has originally been intended to facilitate migration of elephants, but is
-expected to benefit a multitude of other species \citep{Elliot.2014,
+expected to provide benefits to a multitude of other species \citep{Elliot.2014,
 Brennan.2020, Hofmann.2021}.
@@
 In a previous study, we assessed landscape connectivity for dispersing African
-wild dogs within the KAZA-TFCA using a least-cost methods \citep{Hofmann.2021}.
-Specifically, we fitted a basic habitat selection model using which we predicted
-a permeability surface and computed least-cost paths and corridors. We now
-expand this model and use ISSFs to develop a more detailed movement model of
-dispersing wild dogs (\Cref{GraphicalAbstract}). We then employ the model to
-simulate dispersers moving across the KAZA-TFCA. Based on simulations, we
-generate three distinc connectivity maps, all sheding light onto a different
-aspect of connectivity. With this work we hope to show that a simulation-based
-approach yields several benefits over traditional connectivity modeling
-techniques such as least-cost analyses and circuit theory. Most importantly,
-simulations provide a more generic view on how connectivity emerges and to which
-degree connectivity depends on the dispersal duration. In addtion, by generating
-proper dispersal trajectories, network theory can be applied to calculate
-network metrics that are pertinent to connectivity analysis.
+wild dogs within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}.
+Specifically, we fitted a basic habitat selection model and predicted a
+permeability surface that we used to compute least-cost paths and corridors. We
+now expand on this knowledge and use ISSFs to develop a more mechanistic
+movement model of dispersing wild dogs (\Cref{GraphicalAbstract}). We employ the
+model to simulate dispersers moving across the KAZA-TFCA and generate three
+distinct connectivity maps, each shedding light onto a different aspect of
+connectivity. With this work, we exemplify how ISSFs can be utilized for
+dispersal simulations and we discuss several benefits of this approach over
+traditional connectivity modeling techniques such as least-cost analyses and
+circuit theory. Most importantly, simulations based on ISSFs provide a more
+generic view on how connectivity emerges and to which degree connectivity
+depends on the dispersal duration. In addition, by generating proper dispersal
+trajectories, network theory can be applied to calculate network metrics that
+are pertinent to connectivity analysis.
 
@@
 23\degree 56'4''E (\Cref{StudyArea}a) stretching over 1.3 Mio.
-km\textsuperscript{2} and ecompassing the entire KAZA-TFCA (\Cref{StudyArea}b).
+km\textsuperscript{2} and encompassing the entire KAZA-TFCA (\Cref{StudyArea}b).
 The KAZA-TFCA represents the world's largest transboundary conservation area and
 comprises parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. It covers a
-total of 520'000 km\textsuperscript{2} and hosts a diverse landscape, ranging
+total of 520'000 km\textsuperscript{2} and hosts diverse landscapes, ranging
 from savanna to grassland and from dry to moist woodland habitats. In its center
@@
 dispersal coalition (7 female and 9 male coalitions). During dispersal, collars
-were programmed to record a GPS fix every 4 hours. Collected relocations were
-regularly transmitted over the Iridium satellite system, which allowed remote
-tracking of individuals, even if they left the main study area and crossed
+were programmed to record a GPS fix every 4 hours, all of which were regularly
+transmitted over the Iridium satellite system, thereby allowing to remotely
+track individuals, even if they left the main study area and crossed
 international borders. Because behavior during dispersal is more pertinent for
@@
 during dispersal. In some instances, exact dispersal dates were known from field
-observations. Where this was not the case, determined dispersal phases using the
+observations, whereas in other cases we determined dispersal phases using the
 net-squared displacement metric. Net squared displacement measures the squared
@@
 which in our case was set to the center of each individual's natal home range.
-As such, dispersal was deemed to have started when an individual left its natal
+Thus, dispersal was deemed to have started when an individual left its natal
 home range and ended once individuals became sedentary again. As previous
-research revealed similar behavior of females and males during dispersal
+research found no differences in behaviors of females and males during dispersal
 \citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish between sexes. After
 collection, we converted collected GPS coordinates (n = 4'169) to steps, where
-each step represented the straight-line distance traveled by and individual
+each step represented the straight-line distance traveled by an individual
 between two consecutive GPS relocations \citep{Turchin.1998}. To ensure a
@@
 dogs \citep{Avgar.2016}. To conduct ISSF analysis, we paired each realized step
-with 24 random steps. An observed and its 24 random steps formed a stratum and
-received a unique identifier. We generated random steps by sampling random
-turning angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths
-from a gamma distirbution that was fitted to realized steps (scale = 6'308,
-shape = 0.37). Along each step, we extracted and averaged spatial covariates
-using the {\tt velox} package \citep{Hunziker.2021}. We also calculated the
-movement metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)} for each
-observed and random step. To facilitate model convergence, we standardized all
-continuous covariates to a mean of zero and a standard deviation of one. Since
-correlation among covariates was low (\(|r| > 0.6\); \citealp{Latham.2011}), we
-retained all of them for modeling.
+with 24 random steps. An observed step plus its 24 random steps formed a stratum
+and received a unique identifier. As suggested by \cite{Avgar.2016}, we
+generated random steps by sampling random turning angles from a uniform
+distribution (\(-\pi, +\pi\)) and step lengths from a gamma distribution that
+was fitted to realized steps (scale = 6'308, shape = 0.37). Along each step, we
+extracted and averaged spatial covariates using the {\tt velox} package
+\citep{Hunziker.2021}. We also calculated the movement metrics \textsf{sl},
+\textsf{log(sl)}, and \textsf{cos(ta)} for each observed and random step. To
+facilitate model convergence, we standardized all continuous covariates to a
+mean of zero and a standard deviation of one. Since correlation among covariates
+was low (\(|r| > 0.6\); \citealp{Latham.2011}), we retained all of them for
+modeling.
 
@@
 that animals assigned a selection score \(w(x)\) of the exponential form to each
-step \citep{Fortin.2005}. The selection score \(w(x)\) of each step depended on
-its associated covariates (\(x_1, x_2, ..., x_n\)) and on the animal's
-preferences (i.e. relative selection strengths; \citealp{Avgar.2017}) towards
-these covariates (\(\beta_1, \beta_2, ..., \beta_n\)):
+step \citep{Fortin.2005}. The selection score \(w(x)\) of each step thus
+depended on its associated covariates (\(x_1, x_2, ..., x_n\)) and on the
+animal's preferences (i.e. relative selection strengths; \citealp{Avgar.2017})
+towards these covariates (\(\beta_1, \beta_2, ..., \beta_n\)):
 
@@
 
-To estimate the preferences of interest, we ran conditional logistic regression
-analysis in the r-package {\tt glmmTMB}. To handle multiple individuals, we
-applied the mixed effects technnique developed by \citep{Muff.2020}, which
-allows to effectively model random slopes. Thus, we treated animal IDs as random
-effect and modeled random slopes for each covariate. We fixed the random
-intercept variance at an arbitrary high value of 10\textsuperscript{6} to make
-use of the  ``poission''-trick \citep{Muff.2020}.
-
-The formula for the movement model was based on the habitat selection model for
-dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model, no
-interactions between habitat covariates (\textsf{Water,
-DistanceToWater\textsuperscript{0.5}, Woodland, Shrubs/Grassland, Human
-Influence}) and movement covariates (\textsf{sl, log(sl), cos(ta)}) were
-considered. Here, we slightly expanded this base model and proposed interactions
-between all movement and habitat covariates. More specifically, we started with
-the base model and incrementally increased model complexity by adding all
-possible two-way interactions between habitat covariates and movement
-covariates. For instance, for the covariate \textsf{Water}, we proposed the
-interactions \textsf{Water:log(sl)}, \textsf{Water:log(sl)}, and
-\textsf{Water:cos(ta)}. Besides those combinations, we also proposed the
-interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)} to account for a
-correlation between turning angles and step lengths, as well as the interactions
-\textsf{sl:LowActivity} and \textsf{log(sl):LowActivity} to account for the fact
-that step lengths may differ due to wild dogs' diurnal activity pattern. To
-compare competing models and assess the most parsimonious movement model, we ran
-stepwise forward model selection based on Akaike's Information Criterion (AIC,
-\citealp{Burnham.2002}).
+We ran conditional logistic regression analysis in the r-package {\tt glmmTMB}
+to estimate preferences of interest. To handle multiple individuals, we applied
+the mixed effects technique developed by \citep{Muff.2020}, which allows to
+effectively model random slopes. Thus, we treated animal IDs as random effect
+and modeled random slopes for each covariate. We fixed the random intercept
+variance at an arbitrary high value of 10\textsuperscript{6} to make use of the
+``poission''-trick \citep{Muff.2020}.
+
+The structure of the movement model was based on the habitat selection model for
+dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model
+(referred to as base model hereafter), no interactions between habitat
+covariates (\textsf{Water, DistanceToWater\textsuperscript{0.5}, Woodland,
+Shrubs/Grassland, Human Influence}) and movement covariates (\textsf{sl,
+log(sl), cos(ta)}) were considered. Hence, we slightly expanded this base model
+and proposed interactions between all movement and habitat covariates. More
+specifically, we started with the base model and incrementally increased model
+complexity by adding all possible two-way interactions between habitat
+covariates and movement covariates. For instance, for the covariate
+\textsf{Water}, we proposed the interactions \textsf{Water:log(sl)},
+\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides those combinations,
+we also proposed the interactions \textsf{sl:cos(ta)} and
+\textsf{log(sl):cos(ta)} to account for a correlation between turning angles and
+step lengths, as well as the interactions \textsf{sl:LowActivity} and
+\textsf{log(sl):LowActivity} to account for the fact that step lengths may
+differ due to wild dogs' diurnal activity pattern. To compare competing models
+and assess the most parsimonious movement model, we ran stepwise forward model
+selection based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
 
@@
 extracted values using the same parameters applied to our input data. (4) We
-applied the parameterized movement model to predict the selection score
-\(w(x)\), which we translated into probabilities using \Cref{EQ2}. (5) We
-sampled one of the random steps based on predicted probabilities and determined
-the animal's new position. We repeated steps (2) to (5) until 2'000 steps were
-realized, implying a total 160 Mio. simulated steps.
+applied the parametrized movement model to predict the selection score \(w(x)\)
+for each step and we translated predicted scores into probabilities using
+\Cref{EQ2}. (5) We sampled one of the random steps based on predicted
+probabilities and determined the animal's new position. We repeated steps (2) to
+(5) until 2'000 steps were realized, implying a total 160 Mio. simulated steps.
 
@@
 fully lied within the buffer, thereby forcing simulated individuals to ``bounce
-off'' from such invisible borders.
+off'' such invisible borders.
 
@@
 
-\subsection{Convegence (80\%)}
+\subsection{Convergence (80\%)}
 To verify that the number of simulated individuals sufficed to ensure reliable
@@
 across the landscape depended on the number of simulated trajectories.
-Specifically, we placed 1'000 ``checkpoints'' each with an extent of 5 km x 5 km
-at random locations inside the main study area. We then determined the relative
-traversal frequency by simulated trajectories through each checkpoint for
-different numbers of simulations (1 to 50'000). We replicated each random sample
-100 times and computed the mean across replicates, as well as the 95\%
-xx-interval. Note that we only considered the 50'000 trajectories originating
-from the main study area.
+Specifically, we distributed 1'000 squared ``checkpoints'', each with an extent
+of 5 km x 5 km at random locations inside the main study area. We then
+determined the relative traversal frequency by simulated trajectories through
+each checkpoint for different numbers of simulations (1 to 50'000). We
+repeatedly sampled trajectories 100 times and computed the mean traversal
+frequency across replicates, as well as the 95\% prediction-interval. We
+expected that the mean traversal frequency converges towards a steady state with
+increasing simulations.
 
@@
 we rasterized all simulated trajectories and tallied them into a single map. If
-the same trajectory crossed a raster-cell twice, it was only counted once,
-thereby mitigating potential biases resulting from individuals that were trapped
-and moved in circles. To achieve high performance rasterization, we used the
+the same trajectory crossed a raster-cell twice, we only counted it once,
+thereby mitigating potential biases caused by individuals that were trapped and
+moved in circles. To achieve high performance rasterization, we used the
 R-package {\tt terra} \citep{Hijmans.2020}.
@@
 overlaid the study area (including the buffer) with a regular raster resolved at
-5 x 5 km. The centerpoint of each raster-cell acted as node in the final network
-and we used the simulated trajectories to determine all transitions occurring
-from one raster-cell to another, as well as the frequency at which those
+5 x 5 km. The centerpoint of each raster-cell served as node in the final
+network and we used the simulated trajectories to determine all transitions
+occurring from one node to another, as well as the frequency at which those
 transitions occurred. This resulted in an edge-list that we translated into a
@@
 traversal frequency in each cell by applying \(\omega = \frac{\sum_i^n{Traversal
-Frequency_i}/n}{Traversal Frequency_i}\). Finally, we used the weighted network
-to calculate the betweenness score of each raster-cell. Betweenness measures how
-often a specific raster-cell lies on a shortest path between two other
-raster-cells and is therefore a useful metric to detect movement corridors
+Frequency_i}/n}{Traversal Frequency_i}\). Consequently, edges that were
+traversed frequently were assigned low costs. Finally, we used the weighted
+network to calculate the betweenness score of each raster-cell. Betweenness
+measures how often a specific raster-cell lies on a shortest path between two
+other raster-cells and is therefore a useful metric to detect movement corridors
 \citep{BastilleRousseau.2018}.
@@
 \subsection{Inter-Patch Connectivity (80\%)}
-To exemplify how simulations can be used to provide a network view on
-connectivity, we assessed inter-patch connectivity between national parks
-located in our study area. The decision to focus on national parks was purely
-out of simplicity and does not imply that connections between other regions are
-impossible. In fact, the same logic could easily be expanded to include other
-protected areas as well. To quantify inter-patch connectivity, we computed the
-relative frequency at which dispersers originating from one national park
-successfully moved into another national park. Successful movement was said to
-be achieved if the individuals' trajectory ever intersected with the
-corresponding national park. We also recorded the step-number of the first step
-that ever intersected with the national park's polygon. This allowed us to
-determine \textit{if} and \textit{how often} dispersers moved between certain
-national parks, as well as to estimate \textit{how long} dispersers had to move
-to realize those connections.
+We assessed inter-patch connectivity between national parks located in our study
+area to examine functional links between distinct patches in the KAZA-TFCA. The
+decision to focus on national parks was purely out of simplicity and does not
+imply that connections between other regions are impossible. In fact, the same
+logic could easily be expanded to include other protected areas. To quantify
+inter-patch connectivity, we computed the relative frequency at which dispersers
+originating from one national park successfully moved into another national
+park. Successful movement was said to be achieved if the individuals' trajectory
+intersected with the corresponding national park at least once. We also recorded
+the number of steps required until the first intersection with the polygon of
+the respective national park. This allowed us to determine \textit{if} and
+\textit{how often} dispersers moved between certain national parks, as well as
+\textit{how long} dispersers had to move to realize those connections.
 
@@
 
-Assuming that all other covariates are held constant, the habitat kernel reveals
-that dispersing wild dogs avoid water but prefer its proximity. Similarly,
-dispersers avoid areas that are covered by woodlands, yet prefer regions covered
-by shrublands or grasslands. Finally, dispersers avoid movement through
-landscapes that are dominated by humans. Effect sizes are strong and, except for
-effect for \textsf{distance to water}, statistically clear on the 5\%
+Assuming that all other covariates are held constant at their means, the habitat
+kernel reveals that dispersing wild dogs avoid water but prefer its proximity.
+Similarly, dispersers avoid areas that are covered by woodlands, yet prefer
+regions covered by shrublands or grasslands. Finally, dispersers avoid movement
+through landscapes that are dominated by humans. Effect sizes are strong and,
+except for effect of \textsf{distance to water}, statistically clear on the 5\%
 significance level.
@@
 \textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (move
-slowly) outside the main activity periods during sunrise and sunset. Aside from
-the interaction \textsf{sl:LowActivity}, which appears to strongly influence
-movmement behavior, effect sizes are moderate, but mostly significant on the 5\%
-significance level.
-
-When looking at the interactions between the movement and habitat kernels, we
+slower) outside the main activity periods (during sunrise and sunset). Aside
+from the interaction \textsf{sl:LowActivity}, which appears to strongly
+influence movmement behavior, effect sizes are moderate, but mostly significant
+on the 5\% significance level.
+
+When looking at the interactions between movement and habitat kernels, we
 observe that movement behavior is contingent on habitat conditions. For example,
 there's strong evidence that dispersers realize smaller steps in areas covered
-by water or areas covered by wooldand. Conversely, it appears that steps tend to
-larger in regions dominated by shrubs/grassland, and shorter when the distance
-to water is high. Correspondingly, the model suggests that directionality is
-lower in areas dominated by humans, yet more pronounced when dispersers are far
-from water. However, except for the effect of \textsf{sl:Water}, effect sizes
-are small.
+by water or areas covered by wooldand, yet it appears that steps are larger in
+regions dominated by shrubs/grassland, and shorter when the distance to water is
+high. Correspondingly, the model suggests that directionality is lower in areas
+dominated by humans but more pronounced when dispersers are far from water.
+However, except for the effect of \textsf{sl:Water}, effect sizes and
+statistical significance are moderate.
 
 The k-fold cross-validation procedure reveals that our model substantially
-outperforms a random guess (\Cref{MovementModel}b). Confidence intervals of
+outperforms a random guess (\Cref{MovementModel}b) and therefore correctly
+assigns a high selection score to realized steps. Confidence intervals of
 \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap and
@@
 Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
-terminated after 120 hours or five days. Comparable simulations will be
+terminated after 120 hours (i.e. five days). Comparable simulations will be
 substantially faster for smaller study areas and lower resolution covariates, as
@@
 inside the main source area \Cref{SourcePoints}(a), only 4.5\% eventually hit a
-map boundary, suggesting that we succesfully limited biases due to boundary
+map boundary, suggesting that we successfully prevented biases due to boundary
 effects. In contrast, 78\% of the 30'000 dispersers originating from the buffer
@@
 
+\subsection{Convergence (50\%)}
+Our examination of the traversal frequency as a function of the number of
+simulated dispersers shows that already after a few simulations the mean
+traversal frequency remains stable and on average changes only little when
+adding further dispersers (\Cref{Convergence} (a) and (b)). While variability
+keeps decreasing with additional dispersers, the marginal benefit of having
+further dispersers decreases following a negative-exponential trend
+(\Cref{Convergence} (c)).
+
 \begin{figure}
@@
     \includegraphics[width=\textwidth]{99_Convergence}
-    \caption{Convergence}
+    \caption{Relative traversal frequency across 1'000 checkpoints (5 km x 5)
+    distributed in the study area. The relative traversal frequency is plotted
+    against the number of simulated individuals to visualize how quickly the
+    metric converges to a steady state. (a) Replicated (100 times) relative
+    traversal frequencies across three randomly chosen checkpoints as well as
+    the corresponding 95\% prediction interval (PI). (b) Averaged relative
+    traversal frequency across all checkpoints and replicates including a 95\%
+    PI. (c) Width of the PI in relation to the number of simulated dispersers.}
     \label{Convergence}
@@
 dispersal hotspot. Still, the presence of several massive water bodies, such as
-the Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, results in
+the Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, poses
 considerable dispersal barriers that limit realized connectivity within the
-KAZA-TFCA. Similarly, dispersal appears to be very limited in Zambia's and
-Zimbabwe's part of the KAZA-TFCA, where only few areas are successfully
-traversed by dispersers. This can largely be attributed to substantial human
-influences resulting from high human density, roads, and agricultural
-activities. Outside the KAZA-TFCA, the most heavily used regions include the
-areas inside the Central Kalahari National Park in Botswana, the area south-west
-of the Khaudum National Park in Namibia, and the area around the Liuwa Plains
+KAZA-TFCA. Similarly, dispersal across Zambia's and Zimbabwe's part of the
+KAZA-TFCA appears to be limited, as only few areas are successfully traversed by
+dispersers. This can largely be attributed to substantial human influences
+resulting from high human density, roads, and agricultural activities in these
+areas. Outside the KAZA-TFCA, the most heavily used regions include the areas
+inside the Central Kalahari National Park in Botswana, the area south-west of
+the Khaudum National Park in Namibia, and the area around the Liuwa Plains
 National Park in Zambia.
@@
 \Cref{Betweenness} and reveal a set of discrete dispersal corridors. Again, the
-region in northern Botswana stands out as crucial dispersal corridor that
-receives the highest betweenness score across the study area. Consequently, the
-region serves as crucial dispersal hub that connects more remote regions in the
-study system. Towards east, the extension of this corridor runs through the
-Chobe National Park into the Hwange national park. From there, a further
-extension connects to the distant Matusadona National Park in Zimbabwe.
-Northwest of the Linyanty ecosystem, a major corridor expands into Angola, where
-it splits and finally traverses over a long stretch of unprotected area into the
-Kafue National Park in Zambia. Several additional corridors with slightly lower
-betweenness scores exist, yet most of them run within the boundaries of the
-KAZA-TFCA. In general, only few corridors directly link the peripheral regions
-of the KAZA-TFCA. For instance, there are only few corridors between the
-Matusadona National Park in Zimbabwe and the Kafue National Park in Zimbabwe.
-Similarly, there are no direct links between the Zimbabwean and Angolan
-``spikes'' of the KAZA-TFCA.
+region in northern Botswana stands out as crucial dispersal hub that connects
+more remote regions in the study system. Towards east, the extension of this
+corridor runs through the Chobe National Park into the Hwange national park.
+From there, a further extension connects to the distant Matusadona National Park
+in Zimbabwe. Northwest of the Linyanty ecosystem, a major corridor expands into
+Angola, where it splits and finally traverses over a long stretch of unprotected
+area into the Kafue National Park in Zambia. Several additional corridors with
+slightly lower betweenness scores exist, yet most of them run within the
+boundaries of the KAZA-TFCA. In general, only few corridors directly link the
+peripheral regions of the KAZA-TFCA. For instance, there are only few corridors
+between the Matusadona National Park in Zimbabwe and the Kafue National Park in
+Zimbabwe. Similarly, there are no direct links between the Zimbabwean and
+Angolan ``spikes'' of the KAZA-TFCA.
 
@@
 
-\subsection{Interpatch Connectivity (80\%)}
-Results from the analysis of interpatch connectivity are given in
-\Cref{AreasReached}, but it is worth pointing out that the figure is only
-intended as an example, as for clarity we limited the network on national parks
-(NPs), albeit plenty of links between other protected areas exist. As can be
-seen, interpatch connectivity between NPs in Angola, Namibia, and Botswana is
-comparably high and dispersal durations short. In contrast, it appears that
-connectivity towards the Kafue NP in Zambia and is low and requires long
-dispersal events to realize a link. Similarly, there is a lack of connections to
-Zimbabwe's Chizarira and Matusadona NP and the more distant Lower Zambezi and
-Mana Pools NPs.
+\subsection{Inter-Patch Connectivity (80\%)}
+Results from the analysis of inter-patch connectivity are given in
+\Cref{AreasReached}. The figure shows all realized links by simulated dispersers
+between national parks and indicates the average duration a disperser had to
+move to realize those links. It is again worth pointing out that the figure is
+only intended as an example and for clarity we only consider connectivity
+between national parks (NPs), albeit plenty of links between other protected
+areas exist. As can be seen from the number and thicknesses of arrows,
+inter-patch connectivity between NPs in Angola, Namibia, and Botswana is
+comparably high. In addition, the bright colors highlight that most dispersal
+events between those areas are short. In contrast, we see that connections into
+the Kafue NP in Zambia require more steps and are fewer in general. Similarly,
+there is a lack of connections into Zimbabwe's Chizarira and Matusadona NP and
+the more distant Lower Zambezi and Mana Pools NPs. In some cases, one can also
+detect imbalances between ingoing and outgoing links, hinting at potential
+source-sink dynamics that occur due to asymmetries in landscape permeability
+depending on the origin. For instance, while a large portion of dispersers from
+the Chizaria NP in Zimbabwe manage to move into the Hwange NP, there are
+comparably few dispersers that succeed in the opposite direction.
 
@@
 
-\section{Discussion} % NEEDS TO BE ENRICHED WITH MORE CITATIONS
+\section{Discussion}
 \subsection{Short Summary (90\%)}
 % Maybe put this at the end?
-We used integrated step selection analysis to analyse data of dispersing wild
-dogs and to parametrize a fully mechanistic movement model describing how
-dispersers move through the available landscape. We employed the parametrized
-model as an individual-based movement model and simulated 80'000 dispersing wild
-dogs moving 2'000 steps across the extent of the KAZA-TFCA, the world's largest
-transboundary conservation area. Based on simulated dispersal trajectories, we
-derived a set of three complementary maps, all geared towards a better
-understanding of dispersal and landscape connectivity. The set of maps included
-a heatmap, revealing frequently traversed areas, a betweenness-map, delineating
-critical dispersal corridors and bottlenecks, and a map of inter-patch
-connectivity, indicating the presence or absence of functional links between
-national parks. We thereby showcase how integrated step selection functions can
-be utilized as simple, yet powerful framework to parametrize movement models and
-simulate animal movement with the aim of assessing landscape connectivity. Such
-an individual-based simulation overcomes several conceptual shortcomings
-inherent to more traditional connectivity modeling techniques, such as
-least-cost analyses and circuit theory. Nevertheless, we suggest to view such
-individual-based simulations as complement, and not as substitute, to
+We used ISSFs to analyse data of dispersing wild dogs and parametrize a fully
+mechanistic movement model describing how dispersers move through the available
+landscape. We employed the parametrized model as an individual-based movement
+model and simulated 80'000 dispersing wild dogs moving 2'000 steps across the
+extent of the KAZA-TFCA, the world's largest transboundary conservation area.
+Based on simulated dispersal trajectories, we derived a set of three
+complementary maps, all geared towards a better understanding of dispersal and
+landscape connectivity. The set of maps included a heatmap, revealing frequently
+traversed areas, a betweenness-map, delineating critical dispersal corridors and
+bottlenecks, and a map of inter-patch connectivity, indicating the presence or
+absence of functional links between national parks. We thereby showcase how
+ISSFs can be utilized as simple, yet powerful framework to parametrize movement
+models and simulate animal movement with the aim of assessing landscape
+connectivity. Such an individual-based simulation overcomes several conceptual
+shortcomings inherent to more traditional connectivity modeling techniques, such
+as least-cost analyses and circuit theory. Nevertheless, we suggest to view
+similar individual-based simulations as complement, and not as substitute, to
 traditional connectivity models.
@@
 Oneill.2020}, as well as with an earlier dispersal model that we developed
-primarily to investigate the habitat kernel of dispersers \citep{Hofmann.2021}.
+to investigate the habitat kernel of dispersers \citep{Hofmann.2021}.
 
@@
 to model several additional complexities inherent to dispersal. It is, for
-instance, well known that turning angles and step lengths tend to be correlated
-in dispersing animals due to a preference for directional persistence. That is,
-larger steps usually coincide with smaller turning angles and vice versa (cite
-someone). While this correlation structure could be effectively rendered by
-sampling turning angles and step lengths from copula probability distributions
-\citep{Hodel.2021}, the ISSF framework allowed us to model similar behavior
-directly in the movement model. Besides accounting for directional persistence,
-we further exploited the ISSF framework and included interactions between
-movement covariates to represent the fact that wild dogs mainly move during the
-more temperate morning and evening hours. The final model thus rendered a
-pulse-like movement behavior, where dispersers realize larger steps during
-morning and evening hours, but shorter steps during the rest of the day.
+instance, well known that dispersers typically move with directional persistence
+and that step lengths are correlated with turning angles. That is, larger steps
+usually coincide with smaller turning angles and vice versa (cite someone).
+While this correlation structure could be rendered by sampling turning angles
+and step lengths from copula probability distributions \citep{Hodel.2021}, the
+ISSF framework allowed us to model similar behavior directly in the movement
+model. Besides accounting for directional persistence, we further exploited the
+ISSF framework and included interactions between movement covariates to
+represent the fact that wild dogs mainly move during the more temperate morning
+and evening hours but tend to rest during the remainder of the day. The final
+model thus rendered a pulse-like movement behavior, where dispersers realize
+larger steps during morning and evening hours, but shorter steps during the rest
+of the day.
 
 By including interactions between habitat and movement covariates, we also
-acknowledged that movement behavior of dispersers may depend on habitat
-conditions. For example, the final model contained an interaction between water
-cover and step length, highlighting that dispersers are more likely to realize
-short steps (i.e. move slower) in areas covered by water. Likewise, the
-interaction between water cover and turning angles revealed that dispersers move
-less directional when across water bodies. We believe that this behavior is owed
-to the fact that wild dogs wade or swim to traverse waterbodies, thus resulting
-in slower, more tortuous movements. Besides this, our model also suggested that
+acknowledged that movement behavior of dispersers depends on habitat conditions.
+For example, the final model contained an interaction between water cover and
+step length, highlighting that dispersers are more likely to realize short steps
+(i.e. move slower) in areas covered by water. Likewise, the interaction between
+water cover and turning angles revealed that dispersers move less directional
+when moving across water bodies. We believe that this behavior is owed to the
+fact that wild dogs wade or swim to traverse waterbodies, thus resulting in
+slower, more tortuous movements. Besides this, our model also suggested that
 dipsersers preferably realize shorter steps when moving through woodland, but
@@
 2'000 steps across the landscapes of the KAZA-TFCA. Our dispersal simulation
-thus comprised 160 Mio. simulated steps, which required six days of computation
+thus comprised 160 Mio. simulated steps, which required five days of computation
 on a modern desktop machine. This rather long simulation duration was primarily
-owed to the massive extent considered (ca. 1.8 Mio. km \textsuperscript{2}) and
-the large number of dispersers simulated. Most connectivity studies consider a
-much smaller extent (e.g. \citealp{Kanagaraj.2013, Abrahms.2017, Zeller.2020})
-and should be able to attain faster simulation speeds. We also believe that
-fewer simulated dispersers will often suffice, as the relative traversal
+owed to the massive extent considered (including the buffer ca. 1.8 Mio. km
+\textsuperscript{2}) and the large number of dispersers simulated. Most
+connectivity studies consider much smaller extents (e.g.
+\citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020})
+and will therefore be able to attain faster simulation speeds. We also believe
+that fewer simulated dispersers will often suffice, as the relative traversal
 frequency of simulated individuals through randomly placed checkpoints in our
@@
 
-A simulation-based approached as proposed in this article offers several
-advantages over more traditional connectivity modeling techniques such as
-least-cost analyses or circuit theory. In contrast to least-cost analyses, for
-instance, an individual-based simulation does not require predetermined
-endpoints. Instead, endpoints emerge naturally as the result of simulated
-dispersal trajectories. Not having to assume known endpoints is particularly
-important for dispersal studies because dispersers often venture into unfamiliar
-territory and are therefore unlikely to know the endpoint of their journey. In
-addition, without known endpoints, movement corridors are no longer enforced
-between start- and endpoint, permitting to detect potential dead ends where
-dispersers get trapped due to insurmountable obstacles.
+A simulation-based approach as proposed in this article offers several striking
+advantages over more traditional connectivity modeling techniques such as LCP
+analysis or CT. In contrast to LCP analyses, for instance, an individual-based
+simulation does not require to assume known endpoints, as endpoints emerge
+naturally as the result of simulated dispersal trajectories. Liberating modelers
+from the assumption of known endpoints is particularly important in dispersal
+studies, as dispersers often venture into unfamiliar territory and are therefore
+unlikely to know the destination of their journey. Additionally, without
+pre-defined endpoints, movement corridors are no longer limited to certain
+combinations of start- and endpoints, thereby permitting to detect movement
+corridors that lead towards dead ends and insurmountable obstacles (i.e.
+ecological traps).
 
@@
 disperser to move from A to B?}'' or \textit{Is it possible for a disperser to
-move from A to B within X days?}, all undoubtly interesting questions that shift
-the focus from a structural to a more functional view on connectivity. Similar
-questions cannot be answered using least-cost methods or circuit theory because
-the time component is omitted from those models. An explicit representation of
-time in IBMMs also yields exciting opportunities for studying seasonal changes
-to connectivity \cite{Zeller.2020}. With least-cost methods or circuit theory a
-dynamic analysis of connectivity is impractical because the same analyses needs
-to be repeated using different permeability surfaces (e.g. \citealp{Benz.2016,
-Osipova.2019}). With IBMMs, on the other hand, simulated individuals can
-directly respond to a dynamic environment within the model and the environment
-is allowed to change ``as the dispersers move''.
+move from A to B within X days?}, all undoubtedly interesting questions that
+shift the focus from a structural to a more functional view on connectivity.
+Similar questions cannot be answered using LCP methods or CT because the time
+component is omitted from those models. An explicit representation of time in
+IBMMs also yields opportunities for studying seasonal changes to connectivity
+\cite{Zeller.2020}. With least-cost methods or circuit theory a dynamic analysis
+of connectivity is impractical because the same analyses needs to be repeated
+using different permeability surfaces (e.g. \citealp{Benz.2016, Osipova.2019}).
+With IBMMs, on the other hand, simulated individuals can directly respond to a
+dynamic environment within the model and the environment is allowed to change
+``as the dispersers move''.
 
@@
 
-Overall, IBMMs enable to generate a rich insight into the emergence of
-connectivty when compared to the outputs of least-cost analysis or circuit
-theory. By generating proper movement trajectories, a broad variety of summary
-statistics can be derived. Specifically, by turning simulated trajectories into
-a network, network theoretical metrics relevant to connectivity can be applied
+Overall, IBMMs enable a rich examination of the emergence of connectivity when
+compared to the outputs of least-cost analysis or circuit theory. By generating
+proper movement trajectories, a broad variety of summary statistics can be
+derived. Specifically, by turning simulated trajectories into a network, network
+theoretical metrics relevant to connectivity can be applied
 \citep{BastilleRousseau.2018}. In contrast to least-cost path analysis and
@@
 low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
-may thus be determined by the by the amount of islotation between subpopulations
+may thus be determined by the by the amount of isolation between subpopulations
 and population densities \citep{DaviesMostert.2012}.
+
+% We did not validate all of our predictions
+Even though our simulations generated several interesting results, most of them
+need to be validated using independent data. We learned that between some
+national parks asymmetries between the number of ingoing and outgoing dispersers
+exist, suggesting that some patches act as sinks, whereas others serve as
+sources for dispersers.
 

### 2021-06-28 Chapter_1/04_Manuscript/Manuscript.tex
some typos and grammatical checking

@@
 kernel is the main strength of ISSFs, which is why we believe that the framework
-is particularly well suitded for simulating animal dispersal and assessing
+is particularly well suited for simulating animal dispersal and assessing
 landscape connectivity. In principle, the ISSF framework allows to define an
 infinite amount of movement kernels, each representing movement behavior under
-different conditions \citep{Fieberg.2020}. It can therefore accomodate highly
+different conditions \citep{Fieberg.2020}. It can therefore accommodate highly
 diverse movement modes, including fast paced directional movements and slow
 paced tortuous movements. Compared to models that disregard an animal's movement
-kernel, this allows to model directional peristence, a movement characteristic
+kernel, this allows to model directional persistence, a movement characteristic
 regularly observed in dispersing animals. To date, however, we only accounted
@@
 meaningful selection of source points is not a problem unique to IBMMs and
-applies to any connectivity modelling technique. As highlighted by
+applies to any connectivity modeling technique. As highlighted by
 \cite{Signer.2017} the influence of the exact location of source points
@@
 probability is contingent on the surrounding landscape and cumulative dispersal
-distance travelled. Because we only observed few dispersal events and due to the
+distance traveled. Because we only observed few dispersal events and due to the
 wild dog's ability to cover vast dispersal distance \citep{DaviesMostert.2012,
@@
 attraction, some simulated individuals will eventually approach a map boundary
-and the modeller needs to define a rule to handle such situations. One feasible
+and the modeler needs to define a rule to handle such situations. One feasible
 option would be to terminate the trajectory as soon as it hits a map boundary,
@@
 linking Angola's NPs to Zambia's Np, only very few simulated dispersers
-successfully traversed it. This is also reflected in the interpatch connectivity
-map, where the Kafue NP only receives very few links from the central region of
-the KAZA-TFCA.
+successfully traversed it. This is also reflected in the inter-patch
+connectivity map, where the Kafue NP only receives very few links from the
+central region of the KAZA-TFCA.
 
@@
 move between those patches. This knowledge serves to gauge the amount of
-geneflow between habitat patches and can be employed to determine whether two
+gene flow between habitat patches and can be employed to determine whether two
 habitats are connected or not.
@@
 more mechanistic understanding of movement. Moreover, both studies parametrized
-models using data of resident black bearsk, which may lead to a biased view on
+models using data of resident black bears, which may lead to a biased view on
 connectivity, as dispersers may move substantially further, thereby improving
@@
 importance and influence of this covariate, future studies will need to control
-for inter- and intra-speficic interactions that may explain why and when
+for inter- and intra-specific interactions that may explain why and when
 dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
@@
 any means. Rather, we propose to use simulations from ISSF-IBMMs as a simple but
-powerful tool that complemets earlier methods in an attempt provide a more
+powerful tool that complements earlier methods in an attempt provide a more
 comprehensive understanding of landscape connectivity.

### 2021-07-01 Chapter_1/04_Manuscript/Manuscript.tex
Updated graphics and discussion

@@
 dispersal simulations and we discuss several benefits of this approach over
-traditional connectivity modeling techniques such as least-cost analyses and
+traditional connectivity modeling techniques such as least-cost analysis and
 circuit theory. Most importantly, simulations based on ISSFs provide a more
@@
 (\textsf{cos(ta)}) (for details see \citep{Avgar.2016, Fieberg.2020}). Because
-wild dogs follow a diurnal activity pattern, we also coded a binary variable
-(\textsf{LowActivity}) indicating whether a step was realized during periods of
-low wild dog activity (17:00 to 09:00 local time) or high wild dog activity
-(09:00 to 17:00 local time). Handling and manipulation of all data, as well as
-all models and simulations were implemented with the statistical software {\tt
-R}, version 3.6.6 \citep{R.2019}. Several helper functions were written in {\tt
-C++} and imported into {\tt R} using the {\tt Rcpp} package
+wild dogs follow a diurnal activity pattern \citep{Castello.2018}, we also coded
+a binary variable (\textsf{LowActivity}) indicating whether a step was realized
+during periods of low wild dog activity (17:00 to 09:00 local time) or high wild
+dog activity (09:00 to 17:00 local time). Handling and manipulation of all data,
+as well as all models and simulations were implemented with the statistical
+software {\tt R}, version 3.6.6 \citep{R.2019}. Several helper functions were
+written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
 \citep{Eddelbuettel.2011, Eddelbuettel.2013}
@@
 
-\subsection{Convergence (50\%)}
+\subsection{Convergence (80\%)}
 Our examination of the traversal frequency as a function of the number of
-simulated dispersers shows that already after a few simulations the mean
-traversal frequency remains stable and on average changes only little when
-adding further dispersers (\Cref{Convergence} (a) and (b)). While variability
-keeps decreasing with additional dispersers, the marginal benefit of having
-further dispersers decreases following a negative-exponential trend
-(\Cref{Convergence} (c)).
+simulated dispersers shows that the mean traversal frequency stabilizes already
+after very few simulations and changes only little when adding further
+dispersers (\Cref{Convergence} (a) and (b)). While variability keeps decreasing
+with additional dispersers, the marginal benefit of adding further dispersers
+steeply decreases with a negative-exponential trend (\Cref{Convergence} (c)).
 
@@
     \includegraphics[width=\textwidth]{99_Convergence}
-    \caption{Relative traversal frequency across 1'000 checkpoints (5 km x 5)
-    distributed in the study area. The relative traversal frequency is plotted
-    against the number of simulated individuals to visualize how quickly the
-    metric converges to a steady state. (a) Replicated (100 times) relative
-    traversal frequencies across three randomly chosen checkpoints as well as
-    the corresponding 95\% prediction interval (PI). (b) Averaged relative
-    traversal frequency across all checkpoints and replicates including a 95\%
-    PI. (c) Width of the PI in relation to the number of simulated dispersers.}
+    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5)
+    distributed randomly across the study area. The relative traversal frequency
+    is plotted against the number of simulated individuals to visualize how
+    quickly the metric converges to a steady state. (a) Replicated (100 times)
+    relative traversal frequencies across three randomly chosen checkpoints as
+    well as the corresponding 95\% prediction interval (PI). (b) Averaged
+    relative traversal frequency across all checkpoints and replicates including
+    a 95\% PI. (c) Width of the PI in relation to the number of simulated
+    dispersers.}
     \label{Convergence}
@@
 \subsection{Short Summary (90\%)}
-% Maybe put this at the end?
-We used ISSFs to analyse data of dispersing wild dogs and parametrize a fully
+We used ISSFs to analyse data of dispersing wild dogs and to parametrize a fully
 mechanistic movement model describing how dispersers move through the available
 landscape. We employed the parametrized model as an individual-based movement
-model and simulated 80'000 dispersing wild dogs moving 2'000 steps across the
-extent of the KAZA-TFCA, the world's largest transboundary conservation area.
-Based on simulated dispersal trajectories, we derived a set of three
-complementary maps, all geared towards a better understanding of dispersal and
-landscape connectivity. The set of maps included a heatmap, revealing frequently
-traversed areas, a betweenness-map, delineating critical dispersal corridors and
-bottlenecks, and a map of inter-patch connectivity, indicating the presence or
-absence of functional links between national parks. We thereby showcase how
-ISSFs can be utilized as simple, yet powerful framework to parametrize movement
-models and simulate animal movement with the aim of assessing landscape
-connectivity. Such an individual-based simulation overcomes several conceptual
-shortcomings inherent to more traditional connectivity modeling techniques, such
-as least-cost analyses and circuit theory. Nevertheless, we suggest to view
-similar individual-based simulations as complement, and not as substitute, to
-traditional connectivity models.
+model to simulate 80'000 dispersing wild dogs moving 2'000 steps across the
+extent of the KAZA-TFCA, the world's largest transboundary conservation area,
+located in southern Africa. Based on simulated dispersal trajectories, we
+derived three complementary maps, all geared towards a better understanding of
+dispersal and landscape connectivity. These maps included a heatmap, revealing
+frequently traversed areas, a betweenness-map, delineating critical dispersal
+corridors, and a map of inter-patch connectivity, indicating the presence or
+absence of functional links between national parks. We thereby showcase that
+ISSFs offer a simple, yet powerful framework to parametrize movement models and
+simulate dispersal to assess landscape connectivity. Importantly,
+individual-based simulations from ISSFs overcome several conceptual shortcomings
+inherent to more traditional connectivity modeling techniques, such as
+least-cost path analysis and circuit theory.
 
 \subsection{Movement Model (80 \%)}
-Our movement model of dispersing wild dogs comprised a habitat kernel, a
-movement kernel, and their interactions. Together, the kernels described habitat
-and movement preferences of dispersing individuals, as well as how movement
-behavior was affected by habitat conditions. Parameter estimates from the
-habitat kernel revealed that dispersing wild dogs avoid water, prefer its
-proximity, avoid woodland, prefer shrubs/grassland, and avoid areas dominated by
-humans. These findings are consistent with the findings of previous studies on
-dispersing wild dogs \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019,
-Oneill.2020}, as well as with an earlier dispersal model that we developed
-to investigate the habitat kernel of dispersers \citep{Hofmann.2021}.
-
-By expanding our initial model to include a proper movement kernel, we were able
-to model several additional complexities inherent to dispersal. It is, for
-instance, well known that dispersers typically move with directional persistence
-and that step lengths are correlated with turning angles. That is, larger steps
-usually coincide with smaller turning angles and vice versa (cite someone).
-While this correlation structure could be rendered by sampling turning angles
-and step lengths from copula probability distributions \citep{Hodel.2021}, the
-ISSF framework allowed us to model similar behavior directly in the movement
-model. Besides accounting for directional persistence, we further exploited the
-ISSF framework and included interactions between movement covariates to
-represent the fact that wild dogs mainly move during the more temperate morning
-and evening hours but tend to rest during the remainder of the day. The final
-model thus rendered a pulse-like movement behavior, where dispersers realize
-larger steps during morning and evening hours, but shorter steps during the rest
-of the day.
-
-By including interactions between habitat and movement covariates, we also
-acknowledged that movement behavior of dispersers depends on habitat conditions.
-For example, the final model contained an interaction between water cover and
-step length, highlighting that dispersers are more likely to realize short steps
-(i.e. move slower) in areas covered by water. Likewise, the interaction between
-water cover and turning angles revealed that dispersers move less directional
-when moving across water bodies. We believe that this behavior is owed to the
-fact that wild dogs wade or swim to traverse waterbodies, thus resulting in
-slower, more tortuous movements. Besides this, our model also suggested that
-dipsersers preferably realize shorter steps when moving through woodland, but
-larger steps when moving across shrubs/grassland. This can likely be attributed
-to resting periods during which dispersers seek shade and protection below the
-woodland canopy. Since we did not include three way interactions, we could not
-test this hypothesis.
+Our movement model of dispersing wild dogs consisted of a habitat kernel, a
+movement kernel, and their interactions. Together, the kernels in detail
+described habitat and movement preferences of dispersing wild dogs, as well as
+how movement behavior was affected by habitat conditions. Parameter estimates
+from the habitat kernel revealed that dispersing wild dogs avoid water, prefer
+its proximity, avoid woodland, prefer shrubs/grassland, and avoid areas
+dominated by humans. These findings are consistent with findings from previous
+studies on dispersing wild dogs \citep{DaviesMostert.2012, Masenga.2016,
+Woodroffe.2019, Oneill.2020}, as well as with a base model that we developed
+with a main focus on dispersers' habitat kernel \citep{Hofmann.2021}.
+
+By expanding our base model by a proper movement kernel, we were able to model
+several additional complexities inherent to dispersal. For instance, it is well
+known that dispersers move with directional persistence and that their step
+lengths are correlated with turning angles \citep{Morales.2004, Borger.2012}.
+That is, larger steps usually imply smaller turning angles and vice versa. While
+this behavior could be captured by sampling turning angles and step lengths from
+copula probability distributions \citep{Hodel.2021}, the ISSF framework allowed
+us to model similar behavior directly in the movement model. Besides accounting
+for directional persistence and correlations between step lengths and turning
+angles, we also allowed for interactions that rendered the fact that wild dogs
+mainly move during the morning and evening hours, whereas they tend to rest
+during the remainder of the day. The final model thus rendered a burst-like
+movement behavior, where dispersers realize larger steps during the darker
+morning and evening hours, but shorter steps during the rest of the day. While
+we accounted for 1. order autocorrelation (i.e. correlation between two
+subsequent steps), higher order autocorrelation is conceivable and might be
+desirable to model. While ISSFs are conceptually capable of rendering such
+behavior, it unfortunately requires vast amounts of GPS data that is not
+intercepted by missing fixes, making it impractical to model higher-order
+autocorrelation using ISSFs in reality.
+
+By allowing interactions between habitat covariates and movement covariates, we
+furthermore accounted for the fact that movement and habitat preferences are
+interdependent. For example, the final model retained an interaction between
+water cover and step length, showing that dispersers are more likely to realize
+short steps (i.e. move slower) in areas covered by water. Likewise, the
+parameter estimate for the interaction between water cover and turning angles
+revealed that dispersers move less directional when moving across water bodies.
+We believe that this behavior is owed to the fact that wild dogs wade or swim
+when traversing waterbodies, thus resulting in slower, more tortuous movements.
+Besides this, our model also suggested that dipsersers preferably realize
+shorter steps when moving through woodland, but larger steps when moving across
+shrubs/grassland. This can likely be linked to wild dogs' resting behavior, as
+wild dogs usually use open areas to quickly move over long distances
+\citep{Abrahms.2017} but seek shade and protection below the woodland canopy
+when resting \citep{Creel.2002}. However, we did not consider three way
+interactions and could therefore not test this hypothesis.
 
 \subsection{Simulation (80\%)}
-% I need to better distinguish our approach from other IBMMs, maybe use the term
-% iSSF-IBMM?
-Based on the above outlined model results, we simulated 80'000 dispersers moving
-2'000 steps across the landscapes of the KAZA-TFCA. Our dispersal simulation
-thus comprised 160 Mio. simulated steps, which required five days of computation
-on a modern desktop machine. This rather long simulation duration was primarily
-owed to the massive extent considered (including the buffer ca. 1.8 Mio. km
-\textsuperscript{2}) and the large number of dispersers simulated. Most
-connectivity studies consider much smaller extents (e.g.
-\citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020})
-and will therefore be able to attain faster simulation speeds. We also believe
-that fewer simulated dispersers will often suffice, as the relative traversal
-frequency of simulated individuals through randomly placed checkpoints in our
-study area converged rather quickly (Appendices), yet this will vary depending
-on the structure of the landscape and dispersal ability of the focal species.
-
-A simulation-based approach as proposed in this article offers several striking
-advantages over more traditional connectivity modeling techniques such as LCP
-analysis or CT. In contrast to LCP analyses, for instance, an individual-based
-simulation does not require to assume known endpoints, as endpoints emerge
-naturally as the result of simulated dispersal trajectories. Liberating modelers
-from the assumption of known endpoints is particularly important in dispersal
-studies, as dispersers often venture into unfamiliar territory and are therefore
-unlikely to know the destination of their journey. Additionally, without
-pre-defined endpoints, movement corridors are no longer limited to certain
-combinations of start- and endpoints, thereby permitting to detect movement
-corridors that lead towards dead ends and insurmountable obstacles (i.e.
-ecological traps).
-
-IBMMs furthermore yield the advantage of an explicit representation of time.
-This enables to answer questions such as: ``\textit{How long will it take a
-disperser to move from A to B?}'' or \textit{Is it possible for a disperser to
-move from A to B within X days?}, all undoubtedly interesting questions that
-shift the focus from a structural to a more functional view on connectivity.
-Similar questions cannot be answered using LCP methods or CT because the time
-component is omitted from those models. An explicit representation of time in
-IBMMs also yields opportunities for studying seasonal changes to connectivity
-\cite{Zeller.2020}. With least-cost methods or circuit theory a dynamic analysis
-of connectivity is impractical because the same analyses needs to be repeated
-using different permeability surfaces (e.g. \citealp{Benz.2016, Osipova.2019}).
-With IBMMs, on the other hand, simulated individuals can directly respond to a
-dynamic environment within the model and the environment is allowed to change
-``as the dispersers move''.
-
-While an explicit representation of time yields several advantages, it also
-requires that step lengths and turning angles are modeled properly
-\citep{Kanagaraj.2013}, so that dispersal durations between habitat patches can
-be reliably approximated. Accounting for and correctly rendering this movement
-kernel is the main strength of ISSFs, which is why we believe that the framework
-is particularly well suited for simulating animal dispersal and assessing
-landscape connectivity. In principle, the ISSF framework allows to define an
-infinite amount of movement kernels, each representing movement behavior under
-different conditions \citep{Fieberg.2020}. It can therefore accommodate highly
-diverse movement modes, including fast paced directional movements and slow
-paced tortuous movements. Compared to models that disregard an animal's movement
-kernel, this allows to model directional persistence, a movement characteristic
-regularly observed in dispersing animals. To date, however, we only accounted
-for 1. order autocorrelation. That is, we only considered correlations between
-two subsequent steps. Modeling higher-order autocorrelation may be desirable,
-yet will require more extensive data.
-
-Overall, IBMMs enable a rich examination of the emergence of connectivity when
-compared to the outputs of least-cost analysis or circuit theory. By generating
-proper movement trajectories, a broad variety of summary statistics can be
-derived. Specifically, by turning simulated trajectories into a network, network
-theoretical metrics relevant to connectivity can be applied
-\citep{BastilleRousseau.2018}. In contrast to least-cost path analysis and
-circuit theory, a simulation-based approach permits studying inter-patch
-connectivity. This facilitates the identification of asymmetrical links and
-statements about source-sink dynamics \citep{Ferreras.2001, Revilla.2004,
-Kanagaraj.2013}. Our results are also in concert with previous studies on
-functional connectivity based on individual-based models that showed that
-including the behavioral ecology of the target species and the landscape
-structure are imperative when assessing connectivity \citep{Gustafson.1996,
+Based on the above described movement model, we simulated 80'000 dispersers
+moving 2'000 steps across the landscapes of the KAZA-TFCA. This simulation of
+160 Mio. steps required five days of computation on a modern desktop machine.
+The long simulation time was primarily caused by the massive extent considered
+(ca. 1.8 Mio. km \textsuperscript{2} when including the buffer) and the large
+number of dispersers simulated. Most connectivity studies are limited to much
+smaller extents (e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016,
+Abrahms.2017, Zeller.2020}) and will therefore require shorter simulation times.
+We also believe that fewer simulated dispersers will often suffice, as the
+relative traversal frequency by simulated individuals through randomly placed
+checkpoints converged rather quickly. In any case, the exact number of required
+simulations will vary depending on the structure of the landscape and the
+dispersal ability of the focal species.
+
+A simulation-based approach as proposed in this article offers several
+advantages over traditional connectivity modeling techniques such as LCP
+analysis or CT. In contrast to LCP analysis, for instance, an individual-based
+simulation does not require to assume known endpoints. Instead, each endpoint
+emerges naturally from a simulated dispersal trajectory. This is particularly
+valuable for dispersal studies, because dispersers often venture into unfamiliar
+territory and are unlikely to know the destination of their journey
+\citep{Elliot.2014, Abrahms.2017, Cozzi.2020}. A connectivity model that does
+not require pre-defined endpoints also ensures that movement corridors are not
+enforced between certain start- and endpoints and permits to reveal frequently
+used routes that potentially lead into ecological traps \citep{Dwernychuk.1972,
+VanDerMeer.2014}.
+
+In comparison to LCP and CT approaches, IBMMs furthermore yield the advantage of
+an explicit representation of time. The ability to render time enables to answer
+questions such as: ``\textit{How long will it take a disperser to move from A to
+B?}'' or \textit{Is it possible for a disperser to move from A to B within X
+days?}. These are important questions that shift the focus from a structural to
+a more functional view on connectivity. An explicit representation of time also
+yields opportunities to study how connectivity depends on seasonal changes in
+the environment (\textit{dynamic connectivity}; \citealp{Zeller.2020}). With LCP
+or CT, a dynamic connectivity analysis is currently impractical because the same
+analysis needs to be repeated for different permeability surfaces, each
+resembling a snapshot in time for a different seasonal state of the environment
+(e.g. \citealp{Benz.2016, Osipova.2019}). With IBMMs, on the other hand, the
+environment is allowed to change dynamically ``as the dispersers move'' such
+that simulated individuals can respond to seasonal factors directly within the
+simulation.
+
+While an explicit representation of time appears beneficial, it also requires
+that step lengths and turning angles are modeled correctly
+\citep{Kanagaraj.2013}, so that the time required to move between areas can be
+reliably estimated. Correctly rendering movement behavior under varying
+environmental conditions is one of the main strength of ISSFs \citep{Avgar.2016,
+Prokopenko.2017, Fieberg.2020}, which is why we believe that the framework is
+exceptionally well suited for simulating animal dispersal and assessing
+landscape connectivity. In fact, the ISSF framework allows to define an infinite
+amount of movement kernels, each representing movement preferences under
+different environmental conditions \citep{Fieberg.2020}. ISSFs can therefore
+accommodate highly diverse movement modes, including fast paced directional
+movements and slow paced tortuous movements.
+
+Whereas the results of LCP analysis and CT are usually restricted to single
+``conductance'' maps, individual-based simulations can be processed using a
+diverse suite of different approaches, each enabling to focus on a different
+aspect of connectivity. By turning simulated trajectories into a network, for
+instance, researchers can capitalize on concepts from network theory and
+calculate network metrics that are pertinent to landscape connectivity
+\citep{BastilleRousseau.2018}. Here, we have exemplified this by calculating the
+betweenness metric, highlighting major movement corridors. Alternatively, can
+learn about inter-patch connectivity and potential source-sink dynamics between
+patches \citep{Ferreras.2001, Revilla.2004, Kanagaraj.2013} based on simulated
+dispersal trajectories. Ultimately, by accounting for habitat- and movement
+preferences of the focal species, simulations from ISSFs are capable of
+rendering the behavioral ecology of the focal species in detail, thereby
+permitting a more functional view on connectivity that is not solely driven by
+the landscape structure and distances between patches \citep{Gustafson.1996,
 Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004, Revilla.2008,
-Kanagaraj.2013} rather than considering it purely a function of distance.
-
-Besides the benefits of IBMMs, we also want to point out some caveats and not so
-trivial modeling decisions. In particular, we will discuss five modeling
-challenges: (1) number of simulated individuals, (2) location of source points,
-(3) dispersal duration, (4) boundary behavior, and (5) individual variability.
+Kanagaraj.2013}.
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{99_ModelingDecisions.png}
+  \caption{Five major modeling decisions that a researcher needs to consider
+  when simulating dispersers to assess landscape connectivity. (1) Number of
+  simulated trajectories. (2) Handling individual variation: we used point
+  estimates when simulating dispersers, assuming that there was no individual
+  variation. Alternatively, however, one could draw preferences for each
+  simulated individually based on model uncertainty. (3) Dispersal duration.
+  While one could draw the number of simulated steps based on observed dispersal
+  events, we opted for an alternative and simulated each individual for 2'000
+  steps, which was at the upper end of observed dispersal durations. This allows
+  to easily shorten the generated trajectories afterwards and to investigate the
+  sensitivity of results with regards to the dispersal duration. (4) Boundary
+  behavior. We allowed dispersers to leave and re-enter the main study area
+  through a buffer zone with randomized covariate values. Alternatively, one
+  could discard transgressing random steps, thereby forcing dispersers to remain
+  within the study area or simply terminate the simulation, assuming the
+  individual has disappeared. (5) Source point location. The location of source
+  points should optimally be biologically informed. For our simulation, we
+  placed source points within protected areas large enough to sustain viable
+  wild dog populations. Conceivable alternatives include placing source points
+  according to a habitat suitability model or based on abundance information.
+  Importantly, one must also consider potential immigrants from outside the main
+  study area.}
+  \label{ModelingDecisions}
+\end{figure}
+
+Despite the benefits that ISSF simulations offer, we also want to confer some of
+the non-trivial modeling decisions involved. In particular, we will discuss five
+modeling decisions (\Cref{ModelingDecisions}): (1) number of simulated
+individuals, (2) location of source points, (3) dispersal duration, (4) boundary
+behavior, and (5) how to handle individual variability.
 
 (1) When simulating dispersal, the modeler needs to decide on the number of
-simulated individuals. This decision involves the total (absolute) number of
-simulated individuals across the study area, as well as the relative number of
-simulated individuals per spatial entity (e.g. protected area, habitat patch,
-source point). With regards to the total number of simulated individuals, a
-larger number is usually desirable as it yields a more balanced view on
-connectivity. However, this comes at the cost of computational efficiency,
-implying that a trade-off needs to be managed. One simple solution to handle the
-trade-off would be to simulate only as many individuals as are needed to achieve
-convergence in the target metric. In our case, examination of the target metric
-\textsf{traversal frequency} revealed that convergence was achieved with
-relatively few individuals, such that additional dispersers contributed only
-little additional information to our results. With regards to the relative
-number of simulated individuals, we see two feasible approaches. First, one
-could initiate dispersers in relation to population density at the respective
-area. This would directly reflect the fact that population densities are not
-homogeneous across space. (Say more) Alternatively, however, one could also
-distribute dispersers homogeneously, but then weigh simulated trajectories by
-the population density at the source patch (Say more). Because we lacked
-reliable information on density estimates for the extent of the KAZA-TFCA, we
-opted for a solution where dispersers were distributed homogeneously.
+simulated individuals. This decision includes the \textit{absolute} number of
+simulated individuals across the study area, as well as the \textit{relative}
+number of simulated individuals per spatial entity (e.g. protected area, habitat
+patch, source point). Because each additionally simulated individual adds
+information about landscape connectivity and potential movement routes, a larger
+absolute number of simulated individuals is always desirable, yet this comes at
+the cost of computational efficiency. Consequently, a trade-off needs to be
+managed. One solution to handle the trade-off would be to define a target metric
+and to only simulate so many individuals until convergence in the target metric
+is achieved. Here, we used the \textit{relative traversal frequency} across
+checkpoints as target metric and revealed that convergence was achieved already
+after few simulated individuals. With regards to the relative number of
+simulated individuals, we see two feasible approaches. First, one could initiate
+dispersers in relation to population density at the respective area. This would
+directly reflect the fact that population densities are not homogeneous across
+space. (Say more) Alternatively, however, one could also distribute dispersers
+homogeneously, but then weigh simulated trajectories by the population density
+at the source patch (Say more). Because we lacked reliable information on
+density estimates for the extent of the KAZA-TFCA, we opted for a solution where
+dispersers were distributed homogeneously.
 
@@
 Instead of enforcing homogeneous dispersal durations, one could also sample
-dispersal durations from on observed dispersal events. Alternatively, after each
+dispersal durations from observed dispersal events. Alternatively, after each
 step, individuals could be faced with a probability to settle, where the
@@
 
+Approach also similar to \cite{Potts.2013}, yet they used regular SSFs and
+examined steady sate UDs.
+
 % Spatially Realistic Population Models
@@
 
-% Downsides of connectivity?
-Even though connectivity is generally thought to promote population viability,
-it is also related to various aspects that may cause ecological damage, such as
-increased human-wildlife conflicts or facilitated spread of deadly diseases.
-
 % Conclusion
-To this end, we have exemplified the use of ISSFs as a simple framework based on
-which researchers can parametrize dispersers' habitat and movement preferences
-and at the same time use the parametrized model to as an individual-based
-movement model to simulate dispersal and examine landscape connectivity.
-Furthermore, we suggest to analyse simulated trajectories using three
-complementary maps, each focused on a different aspect of connectivity.
-Ultimately, we hope to have sparked interest in the uprising framework of step
-selection functions for investigating dispersal behavior and landscape
-connectivity. Nevertheless, we do not attempt to dismiss the application of
-traditional connectivity models such as least-cost methods or circuit theory by
-any means. Rather, we propose to use simulations from ISSF-IBMMs as a simple but
-powerful tool that complements earlier methods in an attempt provide a more
-comprehensive understanding of landscape connectivity.
+To this end, we have used data on dispersing wild dogs to exemplify how ISSFs
+can be used to parametrize an individual-based movement model that is further
+employed to simulate dispersal and examine landscape connectivity. We also
+presented three complementary connectivity maps derived from simulated
+trajectories, each focused on a different aspect of connectivity. Furthermore,
+we discussed the potential advantages and disadvantages of the proposed
+framework compared to traditional connectivity modeling techniques such as LCPs
+and CT. With this article, we hope to have sparked interest in the uprising
+framework of step selection functions for investigating dispersal behavior and
+landscape connectivity. Nevertheless, we do not attempt to dismiss the
+application of traditional connectivity models by any means. Rather, we propose
+to use simulations from ISSF-models as a simple but powerful tool to provide a
+more comprehensive understanding of dispersal and landscape connectivity.
+
+%------------------------------------------------------------------------------
+%  Further topics one could discuss
+%------------------------------------------------------------------------------
+% % Downsides of connectivity?
+% Even though connectivity is in general thought to promote population viability,
+% it is also related to various aspects that may cause ecological damage, such as
+% increased connectivity into human-dominated areas or facilitated spread of
+% deadly diseases \citep{Brearly.2013}. Nevertheless, information on the
+% relationship between connectivity and the prevalence of deadly diseases in
+% African wild dogs is currently lacking.
 

### 2021-07-01 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/Convergence.R

@@ -93,7 +93,15 @@ p2 <- convergence %>%
     labels = function(x){format(x, big.mark = "'")}
   ) +
   xlab("# Simulated Trajectories") +
-  ylab("Mean Relative Traversal Frequency")
+  ylab("Mean Relative Traversal Frequency") +
+  geom_segment(aes(x = 31000, y = 0.008, xend = 23000, yend = 0.00791)
+    , arrow = arrow(length = unit(0.02, "npc")
+    , type  = "closed"
+  )
+    , colour  = "black"
+    , size    = 0.2
+  ) +
+  annotate("text", x = 31000, y = 0.00802, label = "95% Prediction-Interval (95%-PI)", fontface = 3)
 
 # Check the width of the confidence interval over time
 p3 <- convergence %>%

### 2021-07-14 Chapter_1/04_Manuscript/Manuscript.tex
Updated manuscript

@@
 \usepackage[nameinlink]{cleveref}
-\usepackage{lineno}
+% \usepackage{lineno}
 \usepackage{multirow}
@@
 \begin{abstract}
+% Dispersal & Connectivity
 Dispersal is an important process that allows species to avoid inbreeding,
@@
 estimation of connectivity and preservation of dispersal corridors has become a
-task of extraordinary importance for conservation authorities worldwide.
-
+task of extraordinary importance for conservation authorities.
+
+% Quantifying Connectivity
 Over the past two decades, ecologists have primarily relied on analytical tools
@@
 ecological applications, both methods make several restricting assumptions that
-limit their suitability in reality. Individual-based dispersal simulations have
+limit their usefulness in reality. Individual-based dispersal simulations have
 been proposed to address these shortcomings, yet due to the sheer amount of
-non-trivial modeling decisions decisions required, a unified and objective
-framework to simulate dispersal is missing.
-
+non-trivial modeling decisions decisions, a unified and objective framework to
+simulate dispersal is missing.
+
+% Step Selection Functions
 Recent innovations in movement ecology have brought forward novel opportunities
@@
 utilization distribution resident animals, a similar approach may be useful to
-investigate \textit{transient} movement behavior and landscape connectivity.
-
-Here, we showcase the use of integrated step selection functions as a simple
-individual-based and spatially explicit model to simulate dispersal of the
-endangered African wild dog across the world's largest transboundary
+investigate \textit{transient} movement behavior during dispersal and landscape
+connectivity.
+
+% Current Study
+Here, we showcase the use of integrated step selection functions as simple,
+individual-based, and spatially explicit movement model to simulate dispersal of
+the endangered African wild dog across the world's largest transboundary
 conservation area, the Kavango-Zambezi Transfrontier Conservation Area
@@
 mechanistic movement model describing how dispersing wild dogs move through the
-landscape. Based on this model, we simulate 80'000 dispersers moving across the
-extent of the KAZA-TFCA. and generate a set of maps, each focused on a different
-aspect of connectivity. Finally, we discuss the benefits and pitfalls of such a
-simulation-based approach and highlight potential improvements to be made in the
-future.
+available landscape. Based on this model, we simulate 80'000 dispersers moving
+across the extent of the KAZA-TFCA and generate a set of three connectivity
+maps, each focused on a different aspect of connectivity. Finally, we discuss
+the benefits and pitfalls of such a simulation-based approach and highlight
+potential improvements to be made in the future.
 
@@
 \citep{Diniz.2020}. Two of the most prominent connectivity models are least-cost
-path analysis (LCP analysis; \citealp{Adriaensen.2003}) and circuit theory (CT;
+analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory (CT;
 \citealp{McRae.2006, McRae.2008}), both graph-based methods that estimate
@@
 
-% Issues with Least-Cost Paths & Circuit Theory
+% Issues with Least-Cost Paths
 \subsection{Issues with Least-Cost Paths (90\%)}
@@
 and may not be biologically meaningful \citep{Diniz.2020}. Although some of
-these deficiencies can be addressed using less stringent versions of the LCP
-algorithm (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
+these deficiencies can be addressed using less stringent versions of the least
+cost algorithm (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
 \textit{thresholded} least-cost paths \citep{Landguth.2012}, and
@@
 
+% Issues with Circuit Theory
 \subsection{Issues with Circuit Theory (90\%)}
-CT entails similarly unreasonable restrictions that are hardly ever met. For
-example, because CT only allows movements from a source cell to its 4 or 8
-adjacent cells, it implicitly posits that animals exhibit a perceptual range of
-a single pixel. Given that covariate layers are usually resolved with a pixel
-size between 30 m x 30 m and 1 km x 1 km, this hardly ever renders the true
-capability of animals to perceive the environment. Moreover, CT is built around
-the assumption of a complete random walk \citep{Diniz.2020}, entailing that
-directional biases cannot be rendered. Nevertheless, directionality is a common
-characteristic in animal movement \citep{Bovet.1991, Schultz.2001}, especially
-in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}.
-
+CT entails similarly unreasonable restrictions that are rarely ever met. For
+example, because CT only allows movements from a source cell to its four or
+eight adjacent cells, it implicitly posits that animals exhibit a perceptual
+range of a single pixel. Given that covariate layers are usually resolved with a
+pixel size between 30 m x 30 m and 1 km x 1 km, this hardly ever renders the
+true capability of animals to perceive the environment. Moreover, CT is built
+around the assumption of a complete random walk \citep{Diniz.2020}, entailing
+that directional biases cannot be rendered. Nevertheless, directionality is a
+common characteristic in animal movement \citep{Bovet.1991, Schultz.2001},
+especially in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}.
+
+% Issues with Both Methods
 \subsection{Issues of Both Methods}
@@
     \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
-    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
-    data and habitat covariates have to be collected. The combined data is then
-    analyzed in an integrated step selection model, which enables the
-    parametrization of the focal species' habitat and movement kernels. The
-    parametrized model is then treated as an individual-based movement model and
-    used to simulate dispersal trajectories. Ultimately, simulated trajectories
-    serve to produce a set of maps that are pertinent to landscape connectivity.
-    This includes a heatmap, indicating the relative traversal frequency across
-    the study area, a betweenness map, highlighting movement corridors and
-    bottlenecks, and, finally, an inter-patch connectivity map, where the
-    frequency of connections and their average duration can be depicted. Photos:
-    Whom to cite? Vectronics or Photographers?}
+    \caption{Flowchart of the simulation-based connectivity analysis as proposed
+    in this article. First, GPS data and habitat covariates have to be
+    collected. The combined data is then analyzed in an integrated step
+    selection model, which enables the parametrization of the focal species'
+    habitat and movement kernels. The parametrized model is then treated as an
+    individual-based movement model and used to simulate dispersal trajectories.
+    Ultimately, simulated trajectories serve to produce a set of maps that are
+    pertinent to landscape connectivity. This includes a heatmap, indicating the
+    relative traversal frequency across the study area, a betweenness map,
+    highlighting movement corridors and bottlenecks, and, finally, an
+    inter-patch connectivity map, where the frequency of connections and their
+    average duration can be depicted. \textcolor{red}{Photos: Whom to cite?
+    Vectronics or Photographers?}}
     \label{GraphicalAbstract}
@@
 23\degree 56'4''E (\Cref{StudyArea}a) stretching over 1.3 Mio.
-km\textsuperscript{2} and encompassing the entire KAZA-TFCA (\Cref{StudyArea}b).
+km\textsuperscript{2} and encompassed the entire KAZA-TFCA (\Cref{StudyArea}b).
 The KAZA-TFCA represents the world's largest transboundary conservation area and
@@
 
-\begin{figure}[htbp]
+\begin{figure}[h]
   \begin{center}
-    \includegraphics[width = \textwidth]{99_StudyArea.png}
+    \begin{tikzpicture}
+        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
+        \includegraphics[width=\textwidth]{99_StudyArea.png}
+        };
+        \begin{scope}[x={(image.south east)},y={(image.north west)}]
+            % % next four lines will help you to locate the point needed by forming a grid.
+            % % comment these four lines in the final picture.
+            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
+            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
+            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
+            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
+            % % upto here
+            \draw[densely dotted, black] (0.169, 0.222) -- (0.364, 0.955);
+            \draw[densely dotted, black] (0.169, 0.157) -- (0.364, 0.074);
+        \end{scope}
+    \end{tikzpicture}
     \caption{Illustration of the study area located in southern Africa. (a) The
     study area was confined by a bounding box spanning the entire KAZA-TFCA and
-    encompassing parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
+    encompassed parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
     The KAZA-TFCA represents the world's largest terrestrial conservation area
@@
     re-establish connectivity between already-existing national parks (dark
-    green) and other protected areas (light green). The dispersal data used in
-    this study was collected on a free-ranging African wild dog population
-    inhabiting the Moremi National Park in northern Botswana.}
+    green) and other protected areas (light green).}
     \label{StudyArea}
@@
 effectively model random slopes. Thus, we treated animal IDs as random effect
-and modeled random slopes for each covariate. We fixed the random intercept
-variance at an arbitrary high value of 10\textsuperscript{6} to make use of the
-``poission''-trick \citep{Muff.2020}.
+and modeled random slopes for each covariate.
 
@@
 estimates of connectivity, we evaluated how the relative traversal frequency
-across the landscape depended on the number of simulated trajectories.
-Specifically, we distributed 1'000 squared ``checkpoints'', each with an extent
-of 5 km x 5 km at random locations inside the main study area. We then
-determined the relative traversal frequency by simulated trajectories through
-each checkpoint for different numbers of simulations (1 to 50'000). We
-repeatedly sampled trajectories 100 times and computed the mean traversal
-frequency across replicates, as well as the 95\% prediction-interval. We
-expected that the mean traversal frequency converges towards a steady state with
-increasing simulations.
+across the landscape changed as we increased the number of simulated
+individuals. For this, we distributed 1'000 rectangular ``checkpoints'', each
+with an extent of 5 km x 5 km at random locations inside the main study area. We
+then determined the relative traversal frequency by simulated trajectories
+through each checkpoint for different numbers of simulations (1 to 50'000). To
+assess variability in the relative traversal frequency, we repeatedly sampled
+trajectories 100 times and computed the mean traversal frequency across
+replicates, as well as the 95\% prediction-interval. We deemed that a checkpoint
+converged as soon as the width of the prediction interval for the traversal
+frequency across replicates dropped below a value of 0.01.
 
@@
 
-% On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
-% 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
-% dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
-% dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
-% \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
-% 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
-% turning angles can be attributed to minor disparities between habitat conditions
-% at the area within which we collected training data and habitat conditions
-% within the entire study area.
-
 \subsection{Heatmap (80\%)}
@@
 move to realize those links. It is again worth pointing out that the figure is
-only intended as an example and for clarity we only consider connectivity
+only intended as an example, as for clarity we only considered connectivity
 between national parks (NPs), albeit plenty of links between other protected
-areas exist. As can be seen from the number and thicknesses of arrows,
+areas exist. As can be seen from the number, thickness, and color of arrows,
 inter-patch connectivity between NPs in Angola, Namibia, and Botswana is
-comparably high. In addition, the bright colors highlight that most dispersal
-events between those areas are short. In contrast, we see that connections into
-the Kafue NP in Zambia require more steps and are fewer in general. Similarly,
-there is a lack of connections into Zimbabwe's Chizarira and Matusadona NP and
-the more distant Lower Zambezi and Mana Pools NPs. In some cases, one can also
-detect imbalances between ingoing and outgoing links, hinting at potential
-source-sink dynamics that occur due to asymmetries in landscape permeability
-depending on the origin. For instance, while a large portion of dispersers from
-the Chizaria NP in Zimbabwe manage to move into the Hwange NP, there are
-comparably few dispersers that succeed in the opposite direction.
+comparably high and dispersal events between those areas short. In contrast, we
+see that connections into the Kafue NP in Zambia require more steps and are
+fewer in general. Similarly, there is a lack of connections into Zimbabwe's
+Chizarira and Matusadona NP and the more distant Lower Zambezi and Mana Pools
+NPs. In some cases, one can also detect imbalances between ingoing and outgoing
+links, hinting at potential source-sink dynamics that occur due to asymmetries
+in landscape permeability depending on the origin. For instance, while a large
+portion of dispersers from the Chizaria NP in Zimbabwe manage to move into the
+Hwange NP, there are comparably few dispersers that succeed in the opposite
+direction.
 
@@
 \section{Discussion}
+
+% Short Summary
 \subsection{Short Summary (90\%)}
@@
 model to simulate 80'000 dispersing wild dogs moving 2'000 steps across the
-extent of the KAZA-TFCA, the world's largest transboundary conservation area,
-located in southern Africa. Based on simulated dispersal trajectories, we
-derived three complementary maps, all geared towards a better understanding of
-dispersal and landscape connectivity. These maps included a heatmap, revealing
-frequently traversed areas, a betweenness-map, delineating critical dispersal
-corridors, and a map of inter-patch connectivity, indicating the presence or
-absence of functional links between national parks. We thereby showcase that
-ISSFs offer a simple, yet powerful framework to parametrize movement models and
-simulate dispersal to assess landscape connectivity. Importantly,
-individual-based simulations from ISSFs overcome several conceptual shortcomings
-inherent to more traditional connectivity modeling techniques, such as
-least-cost path analysis and circuit theory.
-
+extent of the KAZA-TFCA, the world's largest transboundary conservation area.
+Based on simulated dispersal trajectories, we derived three complementary maps,
+each geared towards a better understanding of dispersal and landscape
+connectivity. The set of maps included a heatmap, revealing frequently traversed
+areas, a betweenness-map, delineating critical dispersal corridors, and a map of
+inter-patch connectivity, indicating the presence or absence of functional links
+between national parks as well as the average dispersal duration required to
+realize those links. We thereby showcase that ISSFs offer a simple, yet powerful
+framework to parametrize movement models and simulate dispersal to assess
+landscape connectivity. Importantly, individual-based simulations from ISSFs
+overcome several conceptual shortcomings inherent to more traditional
+connectivity modeling techniques, such as least-cost path analysis and circuit
+theory.
+
+% Movement Model
 \subsection{Movement Model (80 \%)}
-Our movement model of dispersing wild dogs consisted of a habitat kernel, a
-movement kernel, and their interactions. Together, the kernels in detail
-described habitat and movement preferences of dispersing wild dogs, as well as
-how movement behavior was affected by habitat conditions. Parameter estimates
-from the habitat kernel revealed that dispersing wild dogs avoid water, prefer
-its proximity, avoid woodland, prefer shrubs/grassland, and avoid areas
-dominated by humans. These findings are consistent with findings from previous
-studies on dispersing wild dogs \citep{DaviesMostert.2012, Masenga.2016,
-Woodroffe.2019, Oneill.2020}, as well as with a base model that we developed
-with a main focus on dispersers' habitat kernel \citep{Hofmann.2021}.
-
-By expanding our base model by a proper movement kernel, we were able to model
+Our movement model of dispersing wild dogs comprised a habitat kernel, a
+movement kernel, and their interactions. Thus, the model rendered habitat and
+movement preferences of dispersers, as well as how their movement preferences
+were affected by habitat conditions. Parameter estimates from the habitat kernel
+revealed that dispersers avoid water, prefer its proximity, avoid woodland,
+prefer shrubs/grassland, and avoid areas dominated by humans. These results are
+consistent with findings from previous studies on dispersing wild dogs
+\citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020}, as well
+as with a base model that we developed focussing on dispersing wild dogs'
+habitat kernel \citep{Hofmann.2021}.
+
+By expanding the base model by a proper movement kernel, we were able to model
 several additional complexities inherent to dispersal. For instance, it is well
-known that dispersers move with directional persistence and that their step
-lengths are correlated with turning angles \citep{Morales.2004, Borger.2012}.
-That is, larger steps usually imply smaller turning angles and vice versa. While
-this behavior could be captured by sampling turning angles and step lengths from
+known that dispersers usually move with directional persistence
+\citep{Cozzi.2018, Hofmann.2021} and that step lengths are typically correlated
+with turning angles \citep{Morales.2004, Borger.2012}. That is, larger steps
+usually coincide with smaller turning angles and vice versa. While such behavior
+could be captured by jointly sampling turning angles and step lengths from
 copula probability distributions \citep{Hodel.2021}, the ISSF framework allowed
-us to model similar behavior directly in the movement model. Besides accounting
-for directional persistence and correlations between step lengths and turning
-angles, we also allowed for interactions that rendered the fact that wild dogs
-mainly move during the morning and evening hours, whereas they tend to rest
-during the remainder of the day. The final model thus rendered a burst-like
-movement behavior, where dispersers realize larger steps during the darker
-morning and evening hours, but shorter steps during the rest of the day. While
-we accounted for 1. order autocorrelation (i.e. correlation between two
-subsequent steps), higher order autocorrelation is conceivable and might be
-desirable to model. While ISSFs are conceptually capable of rendering such
-behavior, it unfortunately requires vast amounts of GPS data that is not
-intercepted by missing fixes, making it impractical to model higher-order
-autocorrelation using ISSFs in reality.
+us to model similar behavior directly using the movement model. Besides
+accounting for directional persistence and correlations between step lengths and
+turning angles, we also allowed for interactions that rendered the fact that
+wild dogs mainly move during the darker morning and evening hours, whereas they
+tend to rest during the remainder of the day.
 
@@
 water cover and step length, showing that dispersers are more likely to realize
-short steps (i.e. move slower) in areas covered by water. Likewise, the
-parameter estimate for the interaction between water cover and turning angles
-revealed that dispersers move less directional when moving across water bodies.
-We believe that this behavior is owed to the fact that wild dogs wade or swim
-when traversing waterbodies, thus resulting in slower, more tortuous movements.
-Besides this, our model also suggested that dipsersers preferably realize
-shorter steps when moving through woodland, but larger steps when moving across
-shrubs/grassland. This can likely be linked to wild dogs' resting behavior, as
-wild dogs usually use open areas to quickly move over long distances
-\citep{Abrahms.2017} but seek shade and protection below the woodland canopy
-when resting \citep{Creel.2002}. However, we did not consider three way
-interactions and could therefore not test this hypothesis.
+short steps (i.e. move slower) in areas covered by water and large steps in
+areas located on dryland. Likewise, the parameter estimate for the interaction
+between water cover and turning angles revealed that dispersers move less
+directional across water bodies than across dryland. We believe that this is
+owed to the fact that wild dogs wade or swim when traversing waterbodies, thus
+resulting in slower, more tortuous movements. Besides this, our model also
+suggested that dipsersers preferably realize shorter steps when moving through
+woodland, but larger steps when moving across shrubs/grassland. This can likely
+be linked to wild dogs' resting behavior, as wild dogs usually use open areas to
+quickly move over long distances \citep{Abrahms.2017} but seek shade and
+protection below the woodland canopy when resting \citep{Creel.2002}.
 
@@
 Based on the above described movement model, we simulated 80'000 dispersers
-moving 2'000 steps across the landscapes of the KAZA-TFCA. This simulation of
-160 Mio. steps required five days of computation on a modern desktop machine.
-The long simulation time was primarily caused by the massive extent considered
-(ca. 1.8 Mio. km \textsuperscript{2} when including the buffer) and the large
-number of dispersers simulated. Most connectivity studies are limited to much
-smaller extents (e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016,
-Abrahms.2017, Zeller.2020}) and will therefore require shorter simulation times.
-We also believe that fewer simulated dispersers will often suffice, as the
-relative traversal frequency by simulated individuals through randomly placed
-checkpoints converged rather quickly. In any case, the exact number of required
-simulations will vary depending on the structure of the landscape and the
-dispersal ability of the focal species.
-
+moving 2'000 steps across the landscapes of the KAZA-TFCA. On a modern desktop
+machine, this simulation required five days of computation. The long simulation
+time was primarily caused by the massive extent considered (ca. 1.8 Mio. km
+\textsuperscript{2} when including the buffer) and the large number of
+dispersers simulated. Most connectivity studies are limited to much smaller
+extents (e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017,
+Zeller.2020}) and will therefore achieve faster simulation times. We also
+believe that fewer simulated dispersers will often suffice, as the relative
+traversal frequency by simulated individuals through randomly placed checkpoints
+in the study area converged already after 10'500 simulated individuals in our
+case. The required number of simulated individuals will, however, vary depending
+on the structure of the landscape and the dispersal ability of the focal
+species.
+
+\subsection{Maps (70\%)}
+The heatmap resulting from our dispersal simulation highlighted that a large
+portion of simulated dispersers traversed the Moremi NP and the Chobe NP in
+northern Botswana. We already recognized the same area as dispersal hotspot
+using least-cost path and least-cost corridor analysis \citep{Hofmann.2021}, yet
+some researchers questioned whether this was just the logical consequence of the
+region being in the center of the study area and least-costly routes being
+enforced between pre-defined start and endpoints. Using least-cost approaches,
+this claim is difficult to disprove, as all identified routes have to completely
+run within the study area and will always be enforced between a start and
+endpoint. With our simulation, on the other hand, dispersers were able to leave
+the study area and were not enforced to move towards a known endpoint. Despite
+this, a majority of simulated individuals traversed the central region in
+northern Botsana, so we conclude that this dispersal hotspot is not caused by
+geometric properties but results from landscape characteristics and the location
+of source areas.
+
+Overall, the heatmap gives a good overview of the intensity of use in different
+areas, yet it is not well suited for pinpointing discrete movement corridors,
+which is why we also computed a betweenness-map. In contrast to the heatmap, the
+betweenness-map puts stronger emphasis on areas that are used as stepping stones
+into other regions of the study area and thereby highlights discrete dispersal
+corridors or bottlenecks \citep{BastilleRousseau.2018}. Interestingly, the
+central region in northern Botswana again stands out, implying that the region
+is not only frequently visited, but also promotes the relocation of individuals
+into more remote regions of the KAZA-TFCA. While this is an example of an area
+were both the traversal frequency \textit{and} the betweeness score is high,
+there are other instances where only one of the metrics is pronounced. For
+example, while the area between the Lengue-Luiana NP in Angola and the Kafue NP
+in Zambia receives a high betweenness-score, we find that the same area is only
+rarely traversed by dispersers according to the heatmap. Consequently, despite
+the region's importance for linking Angola's NPs to Zambia's NP, only few
+simulated dispersers actually manage to successfully traversed it. Conversely,
+while the area inside the Central Kalahri NP is traversed by many dispersers,
+the betweenness map indicates that the same region does not serve as major
+stepping stone into other regions of the study area.
+
+To complete the picture, we also computed inter-patch connectivity between NPs,
+highlighting functional links and expected dispersal durations between each
+national park in the study area. The map showed that movements from Angola into
+Zambia's Kafue NP are not only rare, but they also require many steps until they
+are realized. Conversely, we find that dispersal between the Moremi NP and Chobe
+NP are relatively frequent and require fewer steps, which can be expected given
+that the areas are located close to each other.
+
+Together, these examples nicely illustrate how powerful a combination of
+different connectivity metrics can be in deepening our understanding of
+landscape connectivity. Each map that we produced from simulated trajectories
+accentuated a different aspect of connectivity, together providing a
+comprehensive view on dispersal and landscape connectivity. The heatmap, for
+example, put emphasis on areas where movement is concentrated, regardless
+whether such areas truly contribute to geneflow or whether they represent ``dead
+ends'' that do not connect distinct patches. The betweenness map, on the other
+hand, pronounced those areas that are relevant in connecting different regions
+in the landscape and highlights potential bottlenecks. Finally, the map of
+inter-patch connectivity illustrated the frequency at which dispersal between
+distinct patches occurs, as well as the average dispersal duration required for
+individuals to move between them.
+
+% Other studies that use similar simulations
+\subsection{Related Literature (80\%)}
+Our approach of simulating movement to assess connectivity is closely related to
+a series of previously published papers. \cite{Clark.2015}, for instance, fitted
+a regular SSF to American black bears (\textit{Ursus americanus}) and employed
+the estimated model parameters to simulate movement and identify the most likely
+movement corridors between four habitat patches. For the same species,
+\cite{Zeller.2020} used regular SSFs and forecasted seasonal habitat
+connectivity under changing land-use. As both of these studies relied on
+\textit{regular} SSFs, rather than \textit{integrated} SSFs, neither of them was
+able to account the interdependence between habitat and movement preferences. As
+such, movement behavior was assumed to be independent of habitat conditions. In
+addition, both studies lacked data collected during dispersal and instead
+employed data on residents to estimate connectivity. Although preferences during
+residence and dispersal may coincide for some species \citep{Fattebert.2015},
+there is compelling evidence suggesting that dispersers more readily cross areas
+avoided by residents \citep{Elliot.2014, Gaston.2016, Abrahms.2017,
+Keeley.2017}. The use of data collected during residence may therefore result in
+biased model estimates that distort our view on connectivity, causing a
+misallocation of scarce conservation funds \citep{Elliot.2014}. Another set of
+related studies that uses simulations from (regular and integrated) SSFs has
+been conducted by \cite{Potts.2013} and \cite{Signer.2017}, yet the primary
+focus of these papers lied on the estimation of steady-state utilization
+distributions and not the investigation of connectivity between habitat patches.
+
+\subsection{Benefits \& Modeling Decisions with ISSF Simulations (70\%)}
+% Does not enforce route towards known endpoint
 A simulation-based approach as proposed in this article offers several
-advantages over traditional connectivity modeling techniques such as LCP
-analysis or CT. In contrast to LCP analysis, for instance, an individual-based
-simulation does not require to assume known endpoints. Instead, each endpoint
-emerges naturally from a simulated dispersal trajectory. This is particularly
-valuable for dispersal studies, because dispersers often venture into unfamiliar
-territory and are unlikely to know the destination of their journey
-\citep{Elliot.2014, Abrahms.2017, Cozzi.2020}. A connectivity model that does
-not require pre-defined endpoints also ensures that movement corridors are not
-enforced between certain start- and endpoints and permits to reveal frequently
-used routes that potentially lead into ecological traps \citep{Dwernychuk.1972,
-VanDerMeer.2014}.
-
-In comparison to LCP and CT approaches, IBMMs furthermore yield the advantage of
-an explicit representation of time. The ability to render time enables to answer
+advantages over traditional connectivity modeling techniques such as LCPA or CT.
+In contrast to LCPA, for instance, an individual-based simulation does not
+require to assume known endpoints. Instead, each endpoint emerges naturally from
+a simulated dispersal trajectory. The ability of not having to provide
+pre-determined is particularly valuable for dispersal studies, because
+dispersers often venture into unfamiliar territory and are therefore unlikely to
+know the destination of their journey \citep{Elliot.2014, Abrahms.2017,
+Cozzi.2020}. Moreover, LCPA always enforces a connection towards the predefined
+endpoints, even if associated movement costs are unreasonably high. With
+simulations from ISSFs this is no longer the case. A connectivity model that
+does not require pre-defined endpoints also ensures that movement corridors are
+not enforced between certain start- and endpoints, which permits to detect
+potential routes that do not lead into suitable habitats but into ecological
+traps \citep{Dwernychuk.1972, VanDerMeer.2014} or areas with a high
+susceptibility for human wildlife conflicts \citep{Cushman.2018}.
+
+% Explicit Representation of Time
+In contrast to LCPA and CT, simulations from ISSFs furthermore yield the
+advantage of an explicit representation of time. This enables to answer
 questions such as: ``\textit{How long will it take a disperser to move from A to
-B?}'' or \textit{Is it possible for a disperser to move from A to B within X
-days?}. These are important questions that shift the focus from a structural to
-a more functional view on connectivity. An explicit representation of time also
-yields opportunities to study how connectivity depends on seasonal changes in
-the environment (\textit{dynamic connectivity}; \citealp{Zeller.2020}). With LCP
-or CT, a dynamic connectivity analysis is currently impractical because the same
-analysis needs to be repeated for different permeability surfaces, each
-resembling a snapshot in time for a different seasonal state of the environment
-(e.g. \citealp{Benz.2016, Osipova.2019}). With IBMMs, on the other hand, the
-environment is allowed to change dynamically ``as the dispersers move'' such
-that simulated individuals can respond to seasonal factors directly within the
-simulation.
-
-While an explicit representation of time appears beneficial, it also requires
-that step lengths and turning angles are modeled correctly
-\citep{Kanagaraj.2013}, so that the time required to move between areas can be
-reliably estimated. Correctly rendering movement behavior under varying
-environmental conditions is one of the main strength of ISSFs \citep{Avgar.2016,
-Prokopenko.2017, Fieberg.2020}, which is why we believe that the framework is
-exceptionally well suited for simulating animal dispersal and assessing
-landscape connectivity. In fact, the ISSF framework allows to define an infinite
-amount of movement kernels, each representing movement preferences under
-different environmental conditions \citep{Fieberg.2020}. ISSFs can therefore
-accommodate highly diverse movement modes, including fast paced directional
-movements and slow paced tortuous movements.
-
-Whereas the results of LCP analysis and CT are usually restricted to single
-``conductance'' maps, individual-based simulations can be processed using a
-diverse suite of different approaches, each enabling to focus on a different
-aspect of connectivity. By turning simulated trajectories into a network, for
-instance, researchers can capitalize on concepts from network theory and
-calculate network metrics that are pertinent to landscape connectivity
-\citep{BastilleRousseau.2018}. Here, we have exemplified this by calculating the
-betweenness metric, highlighting major movement corridors. Alternatively, can
-learn about inter-patch connectivity and potential source-sink dynamics between
-patches \citep{Ferreras.2001, Revilla.2004, Kanagaraj.2013} based on simulated
-dispersal trajectories. Ultimately, by accounting for habitat- and movement
-preferences of the focal species, simulations from ISSFs are capable of
-rendering the behavioral ecology of the focal species in detail, thereby
-permitting a more functional view on connectivity that is not solely driven by
-the landscape structure and distances between patches \citep{Gustafson.1996,
-Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004, Revilla.2008,
-Kanagaraj.2013}.
+B?}'' or ``\textit{Is it possible for a disperser to move from A to B within X
+days?}'' These are important questions that shift the focus from a structural to
+a more functional view on connectivity, which is usually desirable because
+functional connectivity it is directly related to geneflow \citep{Taylor.1993,
+Tischendorf.2000}. An explicit representation of time also yields opportunities
+for studying how seasonality affects connectivity and to investigate whether
+some dispersal corridors are only available temporarily (\textit{dynamic
+connectivity}; \citealp{Zeller.2020}). With LCPA or CT, incorporating
+seasonality is currently impractical, as both methods require a static
+permeability surface as inputs. Hence, the only possibility to study seasonality
+effects is to repeat the same analysis using different permeability surfaces,
+each rendering the environment at a different point in time (e.g.
+\citealp{Benz.2016, Osipova.2019}). With simulations from ISSFs, on the other
+hand, the environment can be rendered dynamically ``as the dispersers move'',
+such that simulated individuals can respond to seasonal factors directly within
+the simulation. Hence, rather than employing a set of static habitat layers,
+each layer would be updated as the dispersers move, thus correctly rendering
+seasonal changes in the environment.
+
+% Requires Detailed Knowledge on Step Lengths and Turning Angles
+While an explicit representation of time offers multiple benefits, it requires
+that step lengths and turning angles are modeled properly
+\citep{Kanagaraj.2013}, so that dispersal durations between areas can be
+estimated reliably. Correctly rendering step lengths and turning angles under
+varying environmental conditions is one of the key strengths of ISSFs
+\citep{Avgar.2016, Prokopenko.2017, Fieberg.2020}, which is why we believe that
+the framework is exceptionally well suited for simulating dispersal and
+assessing landscape connectivity. In addition, the framework enables to model
+autocorrelation between step lengths and turning angles, thereby incorporating
+directional persistence. Here, we only considered first order autocorrelation,
+i.e. correlation between two consecutive steps. Although higher order
+autocorrelation is conceivable and might be desirable to model, this requires
+vast amounts of GPS data that is not intercepted by missing fixes and is
+therefore often impractical to model in reality.
+
+\subsection{Further Considerations (70\%)}
+% Modelling Mortality
+Although we did not render mortality, animals regularly die during dispersal,
+mainly due to deadly encounters with predators, road kills, and persecution by
+humans \citep{Bonnet.1999, Woodroffe.2012, Behr.2021b}. Mortality during
+dispersal could therefore substantially limit functional connectivity
+\citep{Bowler.2009}, especially in areas where the likelihood of encountering
+competitors and humans is high \citep{Cozzi.2020}. If corresponding information
+is available, mortality can and should be included in ISSF simulations.
+
+% Spatially Realistic Population Models
+The ability to realistically render movement during dispersal not only serves to
+investigate landscape connectivity, but also forms the foundation for more
+realistic spatially explicit population models in which dispersal is not merely
+rendered through dispersal kernels or cellular automata movements
+\citep{Visintin.2020}, but mechanistically based on observed movement and
+habitat preferences (e.g. \citealp{Revilla.2008}, \citealp{Kleinmann.2017}).
+Such models can ultimately be employed to conduct population viability analyses
+\citep{Boyce.1992} in which species' dispersal abilities are taken into account.
 
@@
 
-Despite the benefits that ISSF simulations offer, we also want to confer some of
-the non-trivial modeling decisions involved. In particular, we will discuss five
-modeling decisions (\Cref{ModelingDecisions}): (1) number of simulated
-individuals, (2) location of source points, (3) dispersal duration, (4) boundary
-behavior, and (5) how to handle individual variability.
-
-(1) When simulating dispersal, the modeler needs to decide on the number of
-simulated individuals. This decision includes the \textit{absolute} number of
-simulated individuals across the study area, as well as the \textit{relative}
-number of simulated individuals per spatial entity (e.g. protected area, habitat
-patch, source point). Because each additionally simulated individual adds
-information about landscape connectivity and potential movement routes, a larger
-absolute number of simulated individuals is always desirable, yet this comes at
-the cost of computational efficiency. Consequently, a trade-off needs to be
-managed. One solution to handle the trade-off would be to define a target metric
-and to only simulate so many individuals until convergence in the target metric
-is achieved. Here, we used the \textit{relative traversal frequency} across
-checkpoints as target metric and revealed that convergence was achieved already
-after few simulated individuals. With regards to the relative number of
-simulated individuals, we see two feasible approaches. First, one could initiate
-dispersers in relation to population density at the respective area. This would
-directly reflect the fact that population densities are not homogeneous across
-space. (Say more) Alternatively, however, one could also distribute dispersers
-homogeneously, but then weigh simulated trajectories by the population density
-at the source patch (Say more). Because we lacked reliable information on
-density estimates for the extent of the KAZA-TFCA, we opted for a solution where
-dispersers were distributed homogeneously.
-
-(2) While endpoints do not need to be defined in IBMMs, source points will still
-need to be provided by the modeler. Here, we randomly placed source points
-within protected areas large enough to sustain viable wild dog populations,
-knowing that the species primarily survives in these formally protected areas
-\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. In
-some cases, exact locations of potential source populations are known and source
-points can be placed accordingly \citep{Kanagaraj.2013}. Moreover, if abundance
-estimates are available, these can be used to inform the relative number of
-dispersers initiated at each location. In other cases, comparable knowledge may
-be lacking and it could be more beneficial to delineate likely source patches
-based on habitat suitability models (e.g. \citealp{Squires.2013}). Either way, a
-meaningful selection of source points is not a problem unique to IBMMs and
-applies to any connectivity modeling technique. As highlighted by
-\cite{Signer.2017} the influence of the exact location of source points
-decreases as the number of simulated steps is increased, yet their goal was
-assessing steady state utilization distributions and not to assess landscape
-connectivity.
-
-(3) When employing IBMMs, one also needs to decide on meaningful dispersal
-durations (number of simulated steps). We decided to simulate individuals for
-2'000 steps, which is at the upper end of observed dispersal durations and may
-have resulted in an overestimated representation of landscape connectivity.
-Instead of enforcing homogeneous dispersal durations, one could also sample
-dispersal durations from observed dispersal events. Alternatively, after each
-step, individuals could be faced with a probability to settle, where the
-probability is contingent on the surrounding landscape and cumulative dispersal
-distance traveled. Because we only observed few dispersal events and due to the
-wild dog's ability to cover vast dispersal distance \citep{DaviesMostert.2012,
-Masenga.2016, Cozzi.2020} we decided not to apply this approach. Regardless of
-this, we think that in most cases it will be more insightful to simulate
-relatively extensive dispersal events and only subsample afterwards.
-
-(4) Unless simulated individuals are strongly drawn towards a point of
-attraction, some simulated individuals will eventually approach a map boundary
-and the modeler needs to define a rule to handle such situations. One feasible
-option would be to terminate the trajectory as soon as it hits a map boundary,
-implicitly assuming that the simulated animal left the study area and will not
-return. The downside to this approach is that source points located near map
-boundaries will inevitably produce many such dead ends. Alternatively, one could
-remove transgressing random steps from the set of proposed random steps, thereby
-forcing virtual dispersers to bounce off virtual boundaries and remain within
-the main study area. This may seem like an elegant solution, yet again, it
-potentially results in substantial edge effects when source points are located
-near map boundaries. Finally, one could extend the study area by an artificial
-buffer zone through which dispersers are allowed to leave and re-enter the main
-study area. Of course this does not fully mitigate the issue, since some
-dispersers may still approach the outer map boundary, however, as
-\cite{Koen.2010} have shown, such an artificial buffer can often help to
-mitigate edge effects. In theoretical applications, this issue can be
-circumvented by simulating movement on a torus \citep{Hodel.2021b}, yet this is
-unfeasible in most real world applications.
-
-(5) We simulated dispersal using point estimates from our most parsimonious
-movement model but did not investigate the sensitivity of our results with
-respect to those estimates. Uncertainty is rather common in dispersal studies on
-endangered species, as data tends to be scarce, resulting in model estimates
-large confidence intervals \citep{Wiegand.2003, KramerSchadt.2007}. To address
-this, one may explore a broader range of preferences instead of using point
-estimates initiate dispersers with randomized preferences with variability
-imposed by the uncertainty in the movement model. We therefore urge future
-studies to investigate the sensitivity of ISSF simulations with respect to
-estimated preferences.
-
-\subsection{Maps (10\%)}
-The heatmap resulting from our dispersal simulation suggests that a large
-portion of dispersers traverses the Moremi NP and the Chobe NP in northern
-Botswana. Since we already highlighted the same area in our previous analysis
-\citep{Hofmann.2021}, this result was to be expected. Nevertheless, we believe
-it reinforces our notion that the area acts as crucial hub inside the KAZA-TFCA.
-The very same area stands out on the betweenness map, showing that the region is
-not only frequently traversed, but actually serves as stepping stone into more
-remote regions of the KAZA-TFCA. As such, the area exemplifies a region were
-both traversal frequency and betweenness scores are high. In contrast, the
-betweenness map suggests the presence of a corridor moving from Angola into
-Zambia's Kafue NP, yet when looking at the heatmap, one realizes that the area
-is not used very often. Consequently, despite the corridor's importance for
-linking Angola's NPs to Zambia's Np, only very few simulated dispersers
-successfully traversed it. This is also reflected in the inter-patch
-connectivity map, where the Kafue NP only receives very few links from the
-central region of the KAZA-TFCA.
-
-Each map that we produced from simulated trajectories accentuates a different
-aspect of connectivity. The heatmap, for example, puts emphasis on areas where
-movement is concentrated, regardless whether such areas represent ``dead ends''
-that do not necessarily connect distinct patches. Therefore, the heatmap is
-useful to determine locations where a disperser are likely to move to, yet not
-necessarily to uncover corridors that link valuable habitats. The betweenness
-map, on the other hand, brings out those areas that are relevant in connecting
-different regions in the landscape. In this regard, the betweenness map is most
-pertinent to the delineation of dispersal corridors and may serve as an
-alternative to least-cost algorithms. Finally, the inter-patch connectivity map
-illustrates the frequency at which dispersal between distinct habitat patches
-occurs, as well as the average dispersal duration that passes when individuals
-move between those patches. This knowledge serves to gauge the amount of
-gene flow between habitat patches and can be employed to determine whether two
-habitats are connected or not.
-
-\subsection{General (20\%)}
-% Mortality during dispersal
-While we did not attempt to model mortality during dispersal, it is
-well known that dispersers regularly die, mainly due to deadly encounters with
-predators, road kills, or persecution by humans \citep{Woodroffe.2012,
-Behr.2021b}. Mortality during dispersal could therefore substantially limit
-functional connectivity, especially in areas where the likelihood of
-encountering competitors and humans is high \citep{Cozzi.2020}. If corresponding
-information is available, IBMMs would readily allow to model mortality. Each
-timestep, a binary draw determines whether the animal survives or not and this
-probability can depend on landscape characteristics as well as the dispersal
-duration.
-
-% Other studies that use similar simulations
-Our approach of simulating movement to assess connectivity is most closely
-related to the works of \cite{Clark.2015} and \cite{Zeller.2020}, who used
-regular step selection functions to model habitat preferences by black bears and
-simulated movement to assess connectivity. However, neither of these studies
-attempted to jointly model habitat and movement kernels, thereby precluding a
-more mechanistic understanding of movement. Moreover, both studies parametrized
-models using data of resident black bears, which may lead to a biased view on
-connectivity, as dispersers may move substantially further, thereby improving
-connectivity among habitat patches \citep{Elliot.2014}.
-
-Approach also similar to \cite{Potts.2013}, yet they used regular SSFs and
-examined steady sate UDs.
-
-% Spatially Realistic Population Models
-Besides enabling a more mechanistic understanding of connectivity, the ability
-to realistically render movement during dispersal also forms the foundation for
-spatially realistic population models. In these models, dispersal is not merely
-represented by a dispersal kernel (e.g. ...) but mechanistically rendered
-\citep{Revilla.2008}. Such population models can ultimately be employed to
-determine the required level of dispersal to achieve metapopulation viability
-\citep{DaviesMostert.2012}.
-
-% Simulations for an early warning system?
-Simulations resulting from IBMMs could furthermore be utilized as simple tool to
-predict the likely whereabouts of GPS collared animals into the near future. In
-some European countries, the comeback of large predators, such as bears, lynx,
-and the wolf, has triggered emotional discussions and raised public concern
-\citep{Behr.2017}, particularly in areas with free-roaming livestock that may be
-prayed upon by the returned species. An early warning system based on
-simulations could thus serve to forewarn about potential encounters and and
-thereby increase public acceptance of large predators.
-
-% Social landscape
-We have previously attributed the weak significance of distance to water to the
-fact that we did not control for the presence or absence of conspecifics. We
-stick to this reasoning as our expanded model still shows a rather large
-uncertainty around the respective beta coefficients. To better gauge the
-importance and influence of this covariate, future studies will need to control
-for inter- and intra-specific interactions that may explain why and when
-dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
-instance, found that elk movement was significantly impacted by the density of
-wolf in the area, such that habitat preferences strictly differed depending on
-the presence or absence of wolves. The decision to settle is likely related to
-the presence or absence of conspecifics. Hence, the exact dispersal duration and
-distance will not be independent of current wild dog densities. In dispersing
-wolves, for instance, the longest dispersal distances have been observed in
-low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
-may thus be determined by the by the amount of isolation between subpopulations
-and population densities \citep{DaviesMostert.2012}.
-
-% We did not validate all of our predictions
-Even though our simulations generated several interesting results, most of them
-need to be validated using independent data. We learned that between some
-national parks asymmetries between the number of ingoing and outgoing dispersers
-exist, suggesting that some patches act as sinks, whereas others serve as
-sources for dispersers.
-
-% Conclusion
+Despite the benefits that simulations from ISSFs offer, we also want to confer
+some of the non-trivial modeling decisions involved. In particular, we will
+discuss five modeling decisions (\Cref{ModelingDecisions}): (1) number of
+simulated individuals, (2) location of source points, (3) dispersal duration,
+(4) boundary behavior, and (5) how to handle individual variability.
+
+(1) When simulating dispersal using ISSFs, the modeler needs to decide on the
+number of simulated individuals. This decision includes the \textit{absolute}
+number of simulated individuals across the entire study area, as well as the
+\textit{relative} number of simulated individuals per spatial entity (e.g.
+protected area, habitat patch, source point). With respect to the
+\textit{absolute} number of simulated individuals, a higher number is always
+desirable, as each additional disperser provides novel information about
+landscape connectivity. Of course this comes at the cost of computational
+efficiency, such that a trade-off needs to be managed. We propose to handle this
+trade-off by defining a target metric and only simulating additional until
+convergence in the target metric is observed. Here, we employed the
+\textit{relative traversal frequency} across checkpoints as target metric and
+found that convergence across all checkpoints was achieved already after 10'500
+simulated individuals. With regards to the \textit{relative} number of simulated
+individuals, we see several feasible approaches. If corresponding data is
+available, one could distribute dispersers in relation to known abundances,
+reflecting that population densities are not necessarily homogeneous across
+space. Alternatively, one could also distribute dispersers homogeneously, yet
+after the simulation weigh each simulated trajectory according to population
+densities at the respective source patch. Again this requires information on the
+spatial abundance of the focal speciesl. Finally, if such information is
+missing, one can distribute dispersers homogeneously across space. This is the
+approach that we employed and resulted in larger source areas generating a
+larger number of dispersers.
+
+(2) While we simulated dispersal using point estimates from our most
+parsimonious movement model but did not investigate the sensitivity of our
+results with respect to those estimates. Uncertainty is rather common in
+dispersal studies on endangered species, as data tends to be scarce, resulting
+in model estimates large confidence intervals \citep{Wiegand.2003,
+KramerSchadt.2007}. To address this, one may explore a broader range of
+preferences instead of using point estimates initiate dispersers with randomized
+preferences with variability imposed by the uncertainty in the movement model.
+We therefore urge future studies to investigate the sensitivity of ISSF
+simulations with respect to estimated preferences.
+
+(3) When employing ISSFs to simulate dispersers, one also needs to decide on
+meaningful dispersal durations (i.e. number of simulated steps). If
+corresponding data is available, dispersal durations could be sampled from
+observed events, such that each individual would only be simulated until its
+assigned dispersal duration has been achieved. Due to the low number of observed
+dispersal events and due to the great variability in wild dogs' dispersal
+distances \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020} we opted against
+this approach. Instead, we simulated individuals for 2'000 steps, which is at
+the upper end of observed dispersal durations and may have resulted in an overly
+optimistic representation of landscape connectivity. Nevertheless, it is
+relatively straight forward to shorten simulated trajectories in order to
+investigate the sensitivity of results with regards to the dispersal duration.
+Alternatively, if detailed information on settlement behavior is available, the
+dispersal simulation could include a settlement submodel, where after each
+simulated step the simulated individual decides whether or not to settle, based
+the number of realized steps, environmental conditions in the landscape,
+abundance of conspecifics or comptetitors etc.
+
+(4) Unless simulated individuals are drawn towards a point of attraction, some
+individuals will inevitably approach a map boundary such that some of the
+proposed random steps will leave the study area such that no selection score can
+be computed. One option to handle this situation would be to simply terminate
+the simulation as soon as one of the random steps leaves the study area,
+assuming that the simulated animal left the study area and will not return. This
+can be problematic when many individuals are initiated close to map boundaries,
+especially since a single random step leaving the study area forces termination
+of the simulation. As an alternative, one could resample transgressing random
+steps until all proposed random steps lie fully within the study area. This will
+force simulated dispersers to bounce off those boundaries and remain within the
+main study area. Finally, one could also extend the study area by an artificial
+buffer zone with randomized covariate values through which dispersers are
+allowed to leave and re-enter the main study area. Although dispersers might
+still approach the boundary of the buffer, it has been shown that adding an
+artifical buffer helps to mitigate edge effects \cite{Koen.2010}. A last
+solution that only applies in theoretical applications is to simulate movement
+on a torus \citep{Hodel.2021b}.
+
+(5) To initiate the simulation of a disperser, the modeler needs to define a
+source point. In some cases, exact locations of source populations are known and
+source points can be placed accordingly \citep{Kanagaraj.2013}. Moreover, if
+abundance estimates are available, these can be used to inform the relative
+number of dispersers initiated at each location. The selection of source points
+is thus directly related to the relative number of simulated individuals. Here,
+we randomly placed source points within protected areas large enough to sustain
+viable wild dog populations. Given that the species primarily survives in these
+formally protected areas \citep{Woodroffe.1999, DaviesMostert.2012,
+Woodroffe.2012, VanDerMeer.2014} we consider this decision to be appropriate. In
+other cases, comparable knowledge may be lacking and it could be more beneficial
+to delineate likely source patches based on habitat suitability models (e.g.
+\citealp{Squires.2013}). After all, the challenge of selecting meaningful source
+points is not unique individual-based simulations also applies to LCPA and
+CT. However, as highlighted by \cite{Signer.2017}, the influence of the exact
+location of source points decreases as the number of simulated steps is
+increased.
+
+\subsection{Conclusion (80\%)}
 To this end, we have used data on dispersing wild dogs to exemplify how ISSFs
@@
 we discussed the potential advantages and disadvantages of the proposed
-framework compared to traditional connectivity modeling techniques such as LCPs
+framework compared to traditional connectivity modeling techniques such as LCPA
 and CT. With this article, we hope to have sparked interest in the uprising
@@
 %------------------------------------------------------------------------------
-%  Further topics one could discuss
+%  Further Results to Report
 %------------------------------------------------------------------------------
+% % Some step length, turning angle comparisons
+% On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
+% 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
+% dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
+% dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
+% \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
+% 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
+% turning angles can be attributed to minor disparities between habitat conditions
+% at the area within which we collected training data and habitat conditions
+% within the entire study area.
+
+%------------------------------------------------------------------------------
+%  Further Topics to Discuss
+%------------------------------------------------------------------------------
+
+% % Simulations for an early warning system?
+% Finally, simulations from ISSFs could be utilized as forecasting tool to predict
+% the likely whereabouts of GPS collared animals into the near future. In some
+% European countries, the reintroduction of large predators, such as bears, lynx,
+% and wolves, has triggered emotional discussions and raised public concern
+% \citep{Behr.2017}, particularly in areas where locals depend on income from
+% free-roaming livestock. An early warning system based on simulations could thus
+% serve to inform about potential encounters with large carnivores, such that
+% farmers can secure their livestock accordingly. This would hopefully increase
+% public acceptance of large predators.
+
 % % Downsides of connectivity?
@@
 
+% % Social landscape
+% We have previously attributed the weak significance of distance to water to the
+% fact that we did not control for the presence or absence of conspecifics. We
+% stick to this reasoning as our expanded model still shows a rather large
+% uncertainty around the respective beta coefficients. To better gauge the
+% importance and influence of this covariate, future studies will need to control
+% for inter- and intra-specific interactions that may explain why and when
+% dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
+% instance, found that elk movement was significantly impacted by the density of
+% wolf in the area, such that habitat preferences strictly differed depending on
+% the presence or absence of wolves. The decision to settle is likely related to
+% the presence or absence of conspecifics. Hence, the exact dispersal duration and
+% distance will not be independent of current wild dog densities. In dispersing
+% wolves, for instance, the longest dispersal distances have been observed in
+% low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
+% may thus be determined by the by the amount of isolation between subpopulations
+% and population densities \citep{DaviesMostert.2012}
+
+% % Can generate multiple metrics
+% Another benefit of simulations from ISSFs is that their output can be processed
+% using a diverse suite of approaches, each enabling to focus on a different
+% aspect of connectivity. the results from LCPA and CT, on the other hand, are
+% usually restricted to single ``conductance'' maps. By turning simulated
+% trajectories into a network, for instance, researchers can capitalize on
+% concepts from network theory and calculate network metrics that are pertinent to
+% landscape connectivity \citep{BastilleRousseau.2018}. Here, we exemplified
+% this by calculating the betweenness metric, highlighting major movement
+% corridors. Alternatively, can learn about inter-patch connectivity and potential
+% source-sink dynamics between patches \citep{Ferreras.2001, Revilla.2004,
+% Kanagaraj.2013} based on simulated dispersal trajectories. Ultimately, by
+% accounting for habitat- and movement preferences of the focal species,
+% simulations from ISSFs are capable of rendering the behavioral ecology of the
+% focal species in detail, thereby permitting a more functional view on
+% connectivity that is not solely driven by the landscape structure and distances
+% between patches \citep{Gustafson.1996, Gardner.2004, Graf.2007,
+% KramerSchadt.2004, Revilla.2004, Revilla.2008, Kanagaraj.2013}.
+
+% % We did not validate all of our predictions
+% Even though our simulations generated several interesting results, most of them
+% need to be validated using independent data. We learned that between some
+% national parks asymmetries between the number of ingoing and outgoing dispersers
+% exist, suggesting that some patches act as sinks, whereas others serve as
+% sources for dispersers.
+
 \section{Authors' Contributions}
@@
 GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication. Access to all R-scripts for our analyses is provided
-through Github.
+the time of publication. Access to all R-scripts is provided through Github.
 

### 2021-07-14 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/Convergence.R

@@ -16,17 +16,68 @@ library(ggpubr)         # To arrange multiple plots
 # Load data
 convergence <- read_rds("03_Data/03_Results/99_Convergence.rds")
 
+# Arrange the data a bit nicer
+convergence$CheckID <- as.numeric(convergence$CheckID)
+convergence <- arrange(convergence, NTracks, Replicate, CheckID)
+
 # Give the checkpoint ID a nicer name
 convergence$CheckID <- paste0("Checkpoint ", convergence$CheckID)
 
-# Summarize
+# Summarize across all checkpoints
 summarized <- convergence %>%
   group_by(NTracks, Replicate) %>%
-  summarize(
-      RelativeTraversals = mean(RelativeTraversals)
-    , .groups            = "drop"
-  ) %>%
+  summarize(RelativeTraversals = mean(RelativeTraversals), .groups = "drop") %>%
   subset(NTracks > 0)
+
+# Lets check the number of "converged" checkpoints over time
+converged <- convergence %>%
+  group_by(NTracks, CheckID) %>%
+  summarize(
+      MeanRelativeTraversals = mean(RelativeTraversals)
+    , Upper                  = quantile(RelativeTraversals, 0.975)
+    , Lower                  = quantile(RelativeTraversals, 0.025)
+    , .groups                = "drop"
+  ) %>%
+  subset(NTracks > 0) %>%
+  mutate(Width = Upper - Lower) %>%
+  mutate(Converged = Width < 0.01) %>%
+  group_by(NTracks) %>%
+  summarize(Converged = sum(Converged) / n())
+
+# After how many tracks do all checkpoints remain converged?
+first <- which(sapply(1:nrow(converged), function(x){
+  sum(converged$Converged[x:nrow(converged)]) == (nrow(converged) - x + 1)
+}))[1]
+first <- converged$NTracks[first]
+
+# Plot
+p0 <- ggplot(converged, aes(x = NTracks, y = Converged)) +
+  geom_line() +
+  theme_classic() +
+  coord_capped_cart(
+      left   = "both"
+    , bottom = "both"
+  ) +
+  scale_x_continuous(
+    labels = function(x){format(x, big.mark = "'")}
+  ) +
+  xlab("# Simulated Trajectories") +
+  ylab("Converged Checkpoints (in %)") +
[…]

### 2021-07-14 Chapter_4/02_R-Scripts/02_AnalyseData.r

@@ -17,6 +17,9 @@ library(survival)      # To run conditional logistic regression
 # Set working directory
 setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
 
+# Define number of cores allowed to use
+mcores <- detectCores() - 1
+
 ################################################################################
 #### Helpful Functions
 ################################################################################
@@ -47,41 +50,61 @@ absAngle <- function(x, y, rad = T){
 }
 
 ################################################################################
-#### Prepare Data
+#### Substample Observed Data
 ################################################################################
 # Load observed movement data
 obs <- read_csv("03_Data/01_RawData/ObservedMovements.csv")
 
-# Specify the different design combinations
+# Specify the different design combinations. Note that a forgiveness of 1 refers
+# to a regular step selection function
 dat <- expand_grid(
-    Missingness = seq(0, 0.5, by = 0.1)
-  , Futility    = 1:5
-  , Replicate   = 1:10
+    Missingness = seq(0, 0.8, by = 0.1)  # Fraction of the fixes that is removed
+  , Forgiveness = 1:5                    # Allowed lag of steps (in steps)
+  , Replicate   = 1:10                   # Number of replicates for each design
 )
 
 # Create datasets with rarified observations. That is, randomly remove fixes,
-# from 0% to 50%.
+# from 0% to 80%.
 dat <- mutate(dat, Observations = map(Missingness, function(x){
   obs[sort(sample(1:nrow(obs), size = nrow(obs) * (1 - x))), ]
 }))
 
-# Prepare data for regular step selection analysis
-dat <- mutate(dat, SSF = map2(Observations, Futility, function(x, y){
+# Unnest and nest again (take out the animal ID)
+dat <- dat %>%
+  unnest(Observations) %>%
+  group_by(Missingness, Forgiveness, Replicate, ID) %>%
+  nest() %>%
+  rename(Observations = data)
+
+################################################################################
+#### Convert to Steps and Calculate Step Metrics
+################################################################################
+# Prepare data for step selection analysis
[…]

### 2021-07-16 Chapter_1/04_Manuscript/Manuscript.tex
Rewriting introduction

@@
 
-\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
-Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
-pictus
+\textbf{Keywords:} dispersal, simulation, movement, integrated step selection
+function, Kavango-Zambezi Transfrontier Conservation Area, landscape
+connectivity, Lycaon pictus
 
@@
 \begin{abstract}
-% Dispersal & Connectivity
-Dispersal is an important process that allows species to avoid inbreeding,
-colonize new habitats and reinforce non-viable subpopulations. Successful
-dispersal thus represents a crucial pre-requisite for long-term species
-persistence in wild animal populations. However, the ability to disperse is
-contingent a sufficient degree of landscape connectivity, which is why the
-estimation of connectivity and preservation of dispersal corridors has become a
-task of extraordinary importance for conservation authorities.
-
-% Quantifying Connectivity
-Over the past two decades, ecologists have primarily relied on analytical tools
-such as least-cost analysis and circuit theory to model and investigate
-landscape connectivity. Despite their usefulness for a diverse suite of
-ecological applications, both methods make several restricting assumptions that
-limit their usefulness in reality. Individual-based dispersal simulations have
-been proposed to address these shortcomings, yet due to the sheer amount of
-non-trivial modeling decisions decisions, a unified and objective framework to
-simulate dispersal is missing.
-
-% Step Selection Functions
-Recent innovations in movement ecology have brought forward novel opportunities
-to study animal dispersal and estimate landscape connectivity. In particular,
-the rich suite of resource selection functions, namely point-, step-, and
-path-selection functions, have undergone substantial improvements over the past
-years. Most notably, step-selection functions have been generalized to
-\textit{integrated} step selection functions, which essentially represent fully
-mechanistic movement models based on which an individual's movement could be
-simulated. While such models have been applied to study \textit{steady-state}
-utilization distribution resident animals, a similar approach may be useful to
-investigate \textit{transient} movement behavior during dispersal and landscape
-connectivity.
-
-% Current Study
-Here, we showcase the use of integrated step selection functions as simple,
-individual-based, and spatially explicit movement model to simulate dispersal of
-the endangered African wild dog across the world's largest transboundary
-conservation area, the Kavango-Zambezi Transfrontier Conservation Area
-(KAZA-TFCA). For this, we utilize data collected on 16 dispersing wild dog
-coalitions in combination with relevant habitat covariates. We analyse the data
-using integrated step selection functions, thereby parametrizing a fully
-mechanistic movement model describing how dispersing wild dogs move through the
-available landscape. Based on this model, we simulate 80'000 dispersers moving
-across the extent of the KAZA-TFCA and generate a set of three connectivity
-maps, each focused on a different aspect of connectivity. Finally, we discuss
-the benefits and pitfalls of such a simulation-based approach and highlight
-potential improvements to be made in the future.
+
+%------------------------------------------------------------------------------
+% Short Version
+%------------------------------------------------------------------------------
+% The ability to disperse is contingent a sufficient degree of landscape
+% connectivity, which is why the identification and preservation of movement
+% corridors that promote connectivity has become a task of extraordinary
+% importance. Currently, ecologists mainly rely on least-cost analysis or circuit
+% theory to investigate connectivity, albeit both methods make several assumptions
+% are hardly met in reality. To address these issues, simulations from
+% individual-based movement models have been proposed, yet a unified framework to
+% simulate dispersal and quantify connectivity is lacking.
+%
+% Here, we employ a simple three-step workflow that combines several
+% already-existing methods to assess connectivity using simulated dispersal
+% trajectories. In the first step, we use integrated step selection functions to
+% parametrize a mechanistic movement model rendering dispersal behavior. In a
+% second step, we apply the parametrized model as an individual-based movement
+% model to simulate dispersal trajectories. Finally, in the third step, we
+% consolidate simulated trajectories into three complementary connectivity maps,
+% each focusing on a different aspect of landscape connectivity.
+%
+% We apply these steps to assess landscape connectivity for dispersing African
+% wild dogs (\textit{Lycaon pictus}) southern Africa. We thereby demonstrate that
+% simulations from integrated step selection functions offer a simple, yet
+% powerful alternative to traditional connectivity modeling techniques.
+
+%------------------------------------------------------------------------------
+%  LONG VERSION
+%------------------------------------------------------------------------------
+% Importance of Dispersal & Connectivity
+Dispersal of individuals is a crucial pre-requisite for long-term species
+persistence. However, the ability to disperse is contingent a sufficient degree
+of landscape connectivity, which is why the identification and preservation of
+movement corridors that promote connectivity has become a task of exceptional
+importance.
+
+% Quantifying connectivity
+Over the past two decades, ecologists have primarily relied on least-cost
+analysis and circuit theory to investigate connectivity. Unfortunately, both
+methods make several restrictive assumptions that limit their usefulness in
+reality. Least-cost analysis, for instance, assumes that animals move towards a
+pre-conceived endpoint and choose a cost-minimizing route accordingly. Circuit
+theory, in contrast, presumes a random walk without any directional persistence.
+To address these shortcomings, simulations from individual-based movement models
+have been proposed and applied, yet a unified framework to simulate dispersal
+and to derive connectivity maps based on simulated dispersal events is lacking.
+
+% What we do
+% and synthesize a simple
+% three-step workflow that uses
+Here, we combine several already-existing methods to simulate dispersal and
+investigate landscape connectivity in three distinct steps. In the first step,
+integrated step selection functions are used to parametrize a mechanistic
+movement model. In the second step, the parametrized model is applied as an
+individual-based movement model to simulate dispersal trajectories. In the third
+step, simulated trajectories are consolidated and synthesized into three
+complementary connectivity maps, each focusing on a different aspect of
+landscape connectivity. The set of maps includes a heatmap, revealing frequently
+traversed areas, a betweenness-map, delineating critical dispersal corridors,
+and a map of inter-patch connectivity, depicting the presence or absence of
+functional links between distinct habitat patches.
+
+% Case study
+As a case study, we apply the proposed workflow to assess landscape connectivity
+for dispersing African wild dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi
+Transfrontier Conservation Area, the world's largest transboundary conservation
+area. With this, we demonstrate that the outlined workflow offers a simple, yet
+powerful alternative to traditional connectivity modeling techniques. Although
+such individual-based simulations necessitate informed decisions about the
+number of simulated individuals and the duration of simulated dispersal events,
+they enable to relax several of the unrealistic assumptions inherent to
+least-cost analysis or circuit theory.
 
@@
 % Create linenumbers
-% \linenumbers
+\linenumbers
 
@@
 % Importance of Dispersal & Connectivity
-\subsection{Importance of Dispersal \& Connectivity (90\%)}
-Dispersal is defined as the movement of individuals away from their natal
-location to the site of first reproduction \cite{Howard.1960}. It is a vital
-process governing the social structure of wild animal populations that are
-distributed in space \citep{Hanski.1998, Clobert.2012} and may strongly affect
-population dynamics at different spatial and social scales \citep{Hanski.1999a,
-Clobert.2012}. Dispersal allows species to maintain genetic diversity
-\citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013}, to
-rescue small, non-viable populations \citep{Brown.1977}, and to promote the
-colonization or recolonization of unoccupied habitats \citep{Hanski.1999b,
-MacArthur.2001}. However, successful dispersal requires a sufficient degree of
-landscape connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the
-identification and protection of major dispersal corridors has become a
-fundamental task in conservation science \citep{Nathan.2008, Doerr.2011,
-Rudnick.2012}. The ability to pinpoint relevant dispersal hotspots requires
-information on movement behavior during dispersal and knowledge about factors
-that limit dispersal and therefore connectivity \citep{Baguette.2013,
-Vasudev.2015}.
-
-% Advancements in GPS Technology & Movement Ecology
-\subsection{Advancements in GPS Technology \& Movement Ecology (90\%)}
-Thanks to novel technologies developed over the past decades, particularly of
-GPS/Satellite radio-collars, the use of GPS data to study animal dispersal and
-connectivity has accelerated \citep{Elliot.2014, Jonsson.2016, Williams.2019}.
-Additionally, the advent of publicly accessible satellite imagery and
-sophisticated remote sensing techniques to represent the physical landscape
-through which individuals disperse have heralded a ``golden age of animal
-tracking'' \citep{Kays.2015}. Concurrently, the availability of large amounts of
-empirical data and an increased computational power have led to the development
-of numerous techniques to study dispersal and highlight critical corridors
-between subpopulations \citep{Boyce.2002, Fortin.2005, Cushman.2010,
-Zeller.2012, Diniz.2020}.
-
-% Resource Selection & Connectivity
-\subsection{Resource Selection \& Connectivity (90\%)}
-\textit{Resource selection functions} \citep{Boyce.2002} and derived methods
-such as \textit{step selection functions} \citep{Fortin.2005} and \textit{path
-selection functions} \citep{Cushman.2010} have proven particularly useful for
-studying animal movement \citep{Fieberg.2020} and modeling connectivity
-\citep{Diniz.2020}. These methods allow estimating habitat preferences of the
-focal species by comparing covariates at locations visited by the animal to the
-same covariates at locations available to, but not visited by the animal
-\citep{Boyce.2002, Fortin.2005, Cushman.2010, Thurfjell.2014}. The so estimated
-preferences can then be used to predict a permeability surface, indicating the
-expected ease at which an animal can traverse a given area \citep{Spear.2010,
-Zeller.2012, Etherington.2016}. Ultimately, the permeability surface serves as
-input to a connectivity model that is used to reveal movement corridors
-\citep{Diniz.2020}. Two of the most prominent connectivity models are least-cost
-analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory (CT;
-\citealp{McRae.2006, McRae.2008}), both graph-based methods that estimate
-conductance of the landscape to infer likely movement corridors. Despite their
-intuitive nature and ease of use, both methods make rigorous assumptions about
-animal movement that are often not fulfilled in reality \citep{Diniz.2020}.
-
-% Issues with Least-Cost Paths
-\subsection{Issues with Least-Cost Paths (90\%)}
-In LCP analysis, for instance, a least costly path always exists, even if
-associated movement costs are unreasonably high and will never be incurred by a
-dispersing individual. The method also presumes that animals have an infinite
-perceptual range and a preconceived end-point in mind, such that they choose a
-cost-minimizing route accordingly. These assumptions may be fulfilled by
-migrating animals that typically move between a discrete set of habitats through
-familiar landscapes. Dispersers, on the other hand, usually move over long
-distances into unknown territory and are therefore less likely to be aware of
-associated movement costs \citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Another
-issue of LCPs analysis concerns the fact that least-costly routes, by their very
-nature, are only one pixel wide \citep{Pinto.2009}. This neglects the fact that
-alternative routes with similar costs may exist and implies that the width of
-inferred movement routes depends on the resolution of chosen covariate layers
-and may not be biologically meaningful \citep{Diniz.2020}. Although some of
-these deficiencies can be addressed using less stringent versions of the least
-cost algorithm (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
-\textit{thresholded} least-cost paths \citep{Landguth.2012}, and
-\textit{randomized} least-cost paths \citep{Panzacchi.2016, VanMoorter.2021}), a
-certain degree of arbitrariness remains.
-
-% Issues with Circuit Theory
-\subsection{Issues with Circuit Theory (90\%)}
-CT entails similarly unreasonable restrictions that are rarely ever met. For
-example, because CT only allows movements from a source cell to its four or
-eight adjacent cells, it implicitly posits that animals exhibit a perceptual
-range of a single pixel. Given that covariate layers are usually resolved with a
-pixel size between 30 m x 30 m and 1 km x 1 km, this hardly ever renders the
-true capability of animals to perceive the environment. Moreover, CT is built
-around the assumption of a complete random walk \citep{Diniz.2020}, entailing
-that directional biases cannot be rendered. Nevertheless, directionality is a
-common characteristic in animal movement \citep{Bovet.1991, Schultz.2001},
-especially in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}.
+\subsection{Dispersal \& Connectivity Models (90\%)}
+Successful dispersal of individuals requires a sufficient degree of landscape
+connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the identification
+and protection of major dispersal corridors has become a fundamental task in
+conservation science \citep{Nathan.2008, Doerr.2011, Rudnick.2012}. To pinpoint
+relevant dispersal corridors, a comprehensive understanding of the factors that
+limit dispersal of the focal species and an adequate connectivity model are
+required \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the two
+most prominent connectivity models are least-cost path analysis (LCPA;
+\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
+McRae.2008}), both graph-based approaches that estimate conductance of the
+landscape based on predicted habitat permeability \citep{Zeller.2012}.
 
 % Issues with Both Methods
-\subsection{Issues of Both Methods}
-Finally, neither LCP analysis nor CT are capable of rendering the temporal
-dimension of dispersal \citep{Diniz.2020}. Statements about the expected
-duration required to traverse a certain corridor are therefore impossible.
-Likewise, because movement is not modeled explicitly, interactions between
-movement and habitat preferences of the focal species cannot be rendered.
-Connectivity therefore merely arises in result to the landscape structure, which
-is usually referred to as structural connectivity. While structural connectivity
-yields insights in the \textit{potential} of the landscape to be traversed, it
-does not enable to quantify the \textit{actual} gene flow through the area.
-Consequently, a functional view on connectivity, which also renders the
-behavioral response of the animal with respect to prevailing habitat conditions,
-is often more desirable \citep{Tischendorf.2000, Baguette.2013}.
+\subsection{Issues with Traditional Methods (90\%)}
+Despite their intuitive nature and ease of use, both methods make several
+assumptions that are hardly ever met in reality \citep{Diniz.2020}. With LCPA,
+for instance, a least costly route always exists, even if associated movement
+costs are unreasonably high and will never be incurred by a dispersing
+individual. The method also presumes that animals move towards a pre-conceived
+endpoint and choose a cost-minimizing route accordingly. While these assumptions
+may by fulfilled by migrating animals, dispersers usually move into unfamiliar
+territory and are therefore less likely to be aware of associated movement costs
+\citep{Koen.2014, Abrahms.2017, Cozzi.2020}. CT relaxes these assumptions, yet
+posits that animals move according to a random walk, entailing that directional
+biases cannot be rendered \citep{Diniz.2020}. In reality, however, directional
+persistence is regularly observed \citep{Bovet.1991, Schultz.2001}, especially
+in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}. Some of these
+deficiencies can be addressed using alternative graph-based approaches, such as
+least-cost \textit{corridors} \citep{Pinto.2009}, \textit{thresholded}
+least-cost paths \citep{Landguth.2012}, or \textit{randomized} least-cost paths
+\citep{Panzacchi.2016}, yet a certain degree of arbitrariness remains. Finally,
+none of the methods is capable of rendering the temporal dimension of dispersal
+\citep{Diniz.2020}, meaning that statements about the expected duration required
+to traverse a certain corridor are impossible.
 
@@
 \subsection{What about IBMMs? (90\%)}
-To address the issues inherent to LCPs and CT, individual-based movement models
-(IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
-dispersal trajectories are simulated spatially explicitly, based on movement
-rules that determine how individuals move over and interact with the prevailing
-landscape \citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004,
-Revilla.2004, Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016,
-Hauenstein.2019, Zeller.2020, Vasudev.2021}. Using the simulated trajectories,
-one can calculate a set of connectivity metrics, such as inter-patch
-connectivity and traversal frequency, to reveal major dispersal corridors
-\citep{Kanagaraj.2013, BastilleRousseau.2018, Hauenstein.2019, Zeller.2020}.
-However, while IBMMS can be employed to overcome any of the shortcomings
-intrinsic to LCPs and CT, they are subject to a vast amount of subjective,
-non-trivial modeling decisions. Moreover, they can be challenging to fit and
-require vast amounts of movement data, ideally collected during dispersal
-\citep{Diniz.2020}. Consequently, alternative methods that require fewer
-modeling decisions and are straight forward to apply are desirable.
-
-% Step Selection Analysis
-\subsection{Step Selection Analysis (90\%)}
-Here, we investigate the usefulness of integrated step selection functions
-(ISSFs, \citealp{Avgar.2016}), as a relatively simple but powerful IBMM based on
-which dispersal can be simulated. While regular SSFs were intended to learn
-about relative habitat preferences of the focal species \citep{Fortin.2005,
-Thurfjell.2014, Avgar.2017}, the method has recently been generalized to
-\textit{integrated} SSFs and now enables to jointly study habitat and movement
-preferences, as well as potential interactions between them \citep{Avgar.2016,
-Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple means to
-model complex movement behavior, where movement is viewed as the result of two
-intertwined behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}).
-Importantly, a parametrized ISSF model can be employed as a fully mechanistic
-movement model based on which individual movement trajectories can be simulated
-\citep{Avgar.2016, Signer.2017}. In fact, \cite{Signer.2017} used ISSF to
-simulate steady state utilization distributions of resident animals that were
-moving around a point of attraction. However, the degree to which such
-simulations are helpful in detecting movement corridors and modeling landscape
-connectivity remains to be investigated.
+To overcome the unrealistic assumptions inherent to traditional connectivity
+modeling techniques, simulations from individual-based movement models (IBMMs)
+have been proposed and applied \citep{Diniz.2020}. In these models, dispersal
+trajectories are simulated based on movement rules that determine how
+individuals move across or interact with the prevailing landscape
+\citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004,
+Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019,
+Zeller.2020, Vasudev.2021}. Instead of enforcing connections towards a known
+endpoint, IBMMs allow individuals to choose their route as they move and enable
+to render potential interactions between movement behavior and environmental
+conditions. IBMMs furthermore enable to model directional persistence and to
+study the temporal dimension of dispersal \citep{Diniz.2020}. Despite these
+advantages, however, a unified framework to simulate dispersal and to derive
+connectivity maps based on simulated dispersal events is lacking.
+
+% The Power of Step Selection Functions
+\subsection{Step Selection Functions (90\%)}
+A promising framework that has been developed more recently...
+Over the past decade, the increased availability of large amounts of data and an
+increased computational power have led to the development of numerous techniques
+to study animal movement and resource selection during dispersal
+\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012}. Especially
+\textit{resource selection functions} \citep{Boyce.2002}, \textit{step selection
+functions} \citep{Fortin.2005}, and \textit{path selection functions}
+\citep{Cushman.2010} have pushed the field of movement ecology forward. These
+methods allow estimating habitat preferences of the focal species by comparing
+covariates at locations visited by the animal to the same covariates at
+locations available to, but not visited by the animal \citep{Boyce.2002,
+Fortin.2005, Cushman.2010, Thurfjell.2014}. While regular SSFs were intended to
+learn about relative habitat preferences of the focal species
+\citep{Fortin.2005, Thurfjell.2014, Avgar.2017}, the method has recently been
+generalized to \textit{integrated} SSFs and now enables to jointly study habitat
+and movement preferences, as well as potential interactions between them
+\citep{Avgar.2016, Signer.2017, Fieberg.2020}. ISSFs therefore provide a
+relatively simple means to model complex movement behavior, where movement is
+viewed as the result of two intertwined behavioral kernels (e.g.
+\citealp{Prokopenko.2017, Munden.2020}). Importantly, a parametrized ISSF model
+represents a fully mechanistic movement model based on which individual movement
+trajectories can be simulated \citep{Avgar.2016, Signer.2017}. In fact,
+\cite{Signer.2017} used ISSF to simulate steady state utilization distributions
+of resident animals that were moving around a point of attraction. However, the
+degree to which such simulations are helpful in detecting movement corridors and
+modeling landscape connectivity remains to be investigated.
+
+% What we do
+\subsection{Proposed Framework (90\%)}
+Here, we therefore propose to combine several already-existing methods into a
+simple three-step workflow to simulate dispersal and investigate landscape
+connectivity. In the first step, integrated step selection functions are used to
+parametrize a mechanistic movement model rendering dispersal behavior. In the
+second step, the parametrized model is applied as an individual-based movement
+model to simulate dispersal trajectories. Finally, in the third step, simulated
+trajectories are consolidated into three complementary connectivity maps, each
+focusing on a different aspect of landscape connectivity. The set of maps
+includes a heatmap, revealing frequently traversed areas, a betweenness-map,
+delineating critical dispersal corridors, and a map of inter-patch connectivity,
+depicting the presence or absence of functional links between distinct habitat
+patches.
 
@@
 \subsection{Study Species \& Study Area (90\%)}
-One of the species for which long-term viability relies on sufficient landscape
-connectivity is the endangered African wild dog \textit{Lycon pictus}. While
+We exemplify application of the proposed workflow by investigating landscape
+connectivity for the endangered African wild dog (\textit{Lycon pictus}). While
 once present across entire sub-Saharan Africa, wild dogs have disappeared from a
-vast majority of their historic range due to persecution by humans, habitat
-fragmentation and destruction, and deadly diseases \citep{Woodroffe.2012}. As of
-today, only 6'000 free-ranging individuals remain in small and spatially
-scattered subpopulations \citep{Woodroffe.2012}. Within those subpopulations,
-wild dogs form cohesive packs comprising 8 to 12 adults and their offspring
-\cite{McNutt.1995}. After reaching sexual maturity, male and female offspring
-form same-sex coalitions and disperse from their natal pack in search for
-potential mating partners and a suitable territory to settle \citep{McNutt.1996,
-Behr.2020}. New packs are formed when dispersing coalitions join unrelated
-opposite-sex dispersing coalitions \citep{McNutt.1996}. Dispersing wild dogs can
-cover several hundred kilometers across a variety of landscapes
+vast majority of their historic range, mainly due to persecution by humans,
+habitat fragmentation and destruction, and deadly diseases
+\citep{Woodroffe.2012}. Within the few remaining subpopulations, wild dogs form
+cohesive packs comprising 8 to 12 adults and their offspring \cite{McNutt.1995}.
+After reaching sexual maturity, male and female offspring form same-sex
+coalitions and disperse from their natal pack in search for potential mating
+partners and a suitable territory to settle \citep{McNutt.1996, Behr.2020}. New
+packs are formed when dispersing coalitions join unrelated opposite-sex
+dispersing coalitions \citep{McNutt.1996}. During dispersal, wild dogs can cover
+several hundred kilometers across a variety of landscapes
 \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the
-few strongholds for this species lies near the Moremi Game Reserve in northern
+main strongholds for this species lies near the Moremi Game Reserve in northern
 Botswana, which is part of the world's largest transboundary protected area,
 namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA). This
-area has originally been intended to facilitate migration of elephants, but is
-expected to provide benefits to a multitude of other species \citep{Elliot.2014,
-Brennan.2020, Hofmann.2021}.
+area was originally devoted towards facilitating the migration of elephants, yet
+it can be expected to yield benefits to a great multitude of other species
+\citep{Elliot.2014, Brennan.2020, Hofmann.2021}.
 
@@
 \subsection{Previous Paper (90\%)}
-In a previous study, we assessed landscape connectivity for dispersing African
-wild dogs within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}.
-Specifically, we fitted a basic habitat selection model and predicted a
-permeability surface that we used to compute least-cost paths and corridors. We
-now expand on this knowledge and use ISSFs to develop a more mechanistic
-movement model of dispersing wild dogs (\Cref{GraphicalAbstract}). We employ the
-model to simulate dispersers moving across the KAZA-TFCA and generate three
-distinct connectivity maps, each shedding light onto a different aspect of
-connectivity. With this work, we exemplify how ISSFs can be utilized for
-dispersal simulations and we discuss several benefits of this approach over
-traditional connectivity modeling techniques such as least-cost analysis and
-circuit theory. Most importantly, simulations based on ISSFs provide a more
-generic view on how connectivity emerges and to which degree connectivity
-depends on the dispersal duration. In addition, by generating proper dispersal
-trajectories, network theory can be applied to calculate network metrics that
-are pertinent to connectivity analysis.
+We previously assessed landscape connectivity for dispersing African wild dogs
+within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}. More
+specifically, we fitted a habitat selection model and predicted a permeability
+surface based on which we computed least-cost paths and corridors between
+protected areas. We now expand on this knowledge and use ISSFs to develop a more
+detailed, mechanistic movement model of dispersing wild dogs
+(\Cref{GraphicalAbstract}). We then employ the model to simulate 80'000
+dispersers moving 2'000 steps across the KAZA-TFCA. Finally, we generate a
+heatmap, betweenness map and a map of interpatch connectivity. We thereby
+exemplify how ISSFs can be utilized for dispersal simulations and we discuss
+several benefits of this approach over traditional connectivity modeling
+techniques such as least-cost analysis and circuit theory. Most importantly,
+simulations based on ISSFs provide a more generic view on how connectivity
+emerges and to which degree connectivity depends on the dispersal duration. In
+addition, by generating proper dispersal trajectories, network theory can be
+applied to calculate network metrics that are pertinent to connectivity
+analysis.
 
@@
 \subsection{GPS Relocation Data (90\%)}
-Between 2011 and 2019, we collected GPS relocation data on dispersing wild dogs
-from a free-ranging wild dog population inhabiting the Moremi National Park in
-northern Botswana \citep{Cozzi.2020, Hofmann.2021}. We selected potential
-dispersers based on age, pack size, number of same‐sex siblings within the pack,
-and presence of unrelated opposite-sex individuals in the pack
-\citep{McNutt.1996, Behr.2020}. We immobilized selected individuals using a
-cocktail of Ketamine/Xylazine/Atropine \citep{Osofsky.1996, Cozzi.2020} that was
-injected by dart, fired from a CO\textsubscript{2}-pressurized gun
-(\textit{DAN-Inject, Denmark}). Immobilized individuals were fitted with
-GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic Aerospace GmbH,
-Berlin}) that guaranteed automated drop-off through a decomposable piece of
-cotton. Handling and collaring of all individuals was supervised by a
-Botswana-registered wildlife veterinarian and all individuals quickly rejoined
-their pack after immobilization.
-
-16 collared individuals eventually dispersed, each in a separate same-sex
-dispersal coalition (7 female and 9 male coalitions). During dispersal, collars
-were programmed to record a GPS fix every 4 hours, all of which were regularly
-transmitted over the Iridium satellite system, thereby allowing to remotely
-track individuals, even if they left the main study area and crossed
-international borders. Because behavior during dispersal is more pertinent for
-assessing landscape connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded
-all data that was collected during residency and only retained GPS data recorded
-during dispersal. In some instances, exact dispersal dates were known from field
-observations, whereas in other cases we determined dispersal phases using the
-net-squared displacement metric. Net squared displacement measures the squared
-Euclidean distance of a GPS relocation to a reference point \citep{Borger.2012},
-which in our case was set to the center of each individual's natal home range.
-Thus, dispersal was deemed to have started when an individual left its natal
-home range and ended once individuals became sedentary again. As previous
-research found no differences in behaviors of females and males during dispersal
-\citep{Woodroffe.2019, Cozzi.2020}, we did not distinguish between sexes. After
-collection, we converted collected GPS coordinates (n = 4'169) to steps, where
-each step represented the straight-line distance traveled by an individual
-between two consecutive GPS relocations \citep{Turchin.1998}. To ensure a
-regular sampling interval, we removed fixes that were not successfully collected
-on the 4-hourly schedule (\( \pm \) 15 minutes).
+GPS data of 16 dispersing wild dogs was obtained as described in
+\citep{Cozzi.2020, Hofmann.2021}. We collected data on 7 female and 9 male
+dispersing coalitions between 2011 and 2019 from a free-ranging population
+inhabiting the Moremi National Park in northern Botswana. In some instances,
+exact dispersal dates were known from field observations, in other cases we
+determined dispersal phases using the net-squared displacement metric. Net
+squared displacement measures the squared Euclidean distance of a GPS relocation
+to a reference point \citep{Borger.2012}, which in our case was set to the
+center of each individual's natal home range. Thus, dispersal was deemed to have
+started when an individual left its natal home range and ended once individuals
+became sedentary again. Because behavior during dispersal is more pertinent to
+landscape connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded all data
+that was collected during residency. During dispersal, collars were programmed
+to record a GPS fix every four hours and regularly transmitted over the Iridium
+satellite system. As previous research found no differences in behaviors of
+females and males during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did
+not distinguish between sexes. After collection, we converted GPS coordinates (n
+= 4'169) to steps, where each step represented the straight-line distance
+traveled by an individual between two consecutive GPS relocations
+\citep{Turchin.1998}. To ensure a regular sampling interval, we removed fixes
+that were not successfully collected on the 4-hourly schedule (\( \pm \) 15
+minutes).
 
@@
 water layers (see \citealp{Wolski.2017} and Appendix A3 in
-\citealp{Hofmann.2021}). Resulting water layers thus temporally aligned with our
-dispersal events. We furthermore computed a proxy for human influence, rendering
-anthropogenic pressures stemming from human-density, agricultural sites, and
-roads. All spatial layers were coarsened or interpolated to a target resolution
-of 250 m by 250 m. Further details on the sources and preparation of each
-habitat covariate are given in \cite{Hofmann.2021}.
+\citealp{Hofmann.2021}). Resulting layers thus temporally aligned with each
+dispersal event, ensuring minimal temporal lag between observed movements and
+corresponding water layer. We furthermore computed a proxy for human influence,
+rendering anthropogenic pressures stemming from human-density, agricultural
+sites, and roads. All spatial layers were coarsened or interpolated to a target
+resolution of 250 m by 250 m. A detailed description of the preparation of each
+habitat covariate is given in \cite{Hofmann.2021}.
 
@@
 written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
-\citep{Eddelbuettel.2011, Eddelbuettel.2013}
+\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
 
@@
 dogs \citep{Avgar.2016}. To conduct ISSF analysis, we paired each realized step
-with 24 random steps. An observed step plus its 24 random steps formed a stratum
-and received a unique identifier. As suggested by \cite{Avgar.2016}, we
-generated random steps by sampling random turning angles from a uniform
+with 24 random steps, such that a realized step plus its 24 random steps formed
+a stratum that received a unique identifier. As suggested by \cite{Avgar.2016},
+we generated random steps by sampling random turning angles from a uniform
 distribution (\(-\pi, +\pi\)) and step lengths from a gamma distribution that
@@
 dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model
-(referred to as base model hereafter), no interactions between habitat
+(hereafter referred to as base model), no interactions between habitat
 covariates (\textsf{Water, DistanceToWater\textsuperscript{0.5}, Woodland,
@@
 distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
-distribution. Similar to the input data, each random step represented the
-straight line movement within 4 hours. To prevent unreasonably large steps, we
+distribution (similar to the input data, each random step represented the
+straight line movement within 4 hours). To prevent unreasonably large steps, we
 capped sampled step lengths to a maximum of 35 km, which corresponded to the
@@
 distributed across the study area. 50'000 virtual dispersers were released from
-randomly selected source points within contiguous protected areas larger \(>\)
-700 km\textsuperscript{2} (\Cref{SourcePoints}a), which conforms to average home
-range requirements of resident wild dogs \citep{Pomilia.2015} and allowed us to
-remove patches too small to host viable populations. By distributing source
-points randomly, the number of source points per km\textsuperscript{2} was
-approximately equal within protected areas. To render potential immigrants into
-the study system, we released another 30'000 dispersers at random locations
-inside the 100 km wide buffer zone surrounding the main study area
-(\Cref{SourcePoints}b).
-
-\begin{figure}[htbp]
-  \begin{center}
-    \includegraphics[width = \textwidth]{99_SourceAreas.png}
-    \caption{(a) Different source areas from which we released virtual
-    dispersers. We only considered contiguous protected areas (national parks,
-    game reserves, and forest reserves) that were larger than 700
-    km\textsuperscript{2} (green). This area corresponds to the average home
-    range requirement for viable wild dog populations \citep{Pomilia.2015}. To
-    render potential immigrants into the study system, we also initiated
-    dispersers within a buffer zone (blue) surrounding the main study area. (b)
-    Source points from which dipsersers were released. 50'000 dispersers were
-    released from the main study area (green dots) and another 30'000 dispersers
-    within the virtual buffer (blue dots).}
-    \label{SourcePoints}
-  \end{center}
-\end{figure}
+randomly selected source points within contiguous protected areas that were
+large enough to sustain viable wild dog populations (\(>\) 700
+km\textsuperscript{2}; \citealp{Pomilia.2015}) and another 30'000 were released
+at random locations inside the 100 km wide buffer to render potential immigrants
+into the study system. By distributing source points randomly, the number of
+source points per km\textsuperscript{2} was approximately equal within protected
+areas and the buffer zone.
+
+% \begin{figure}[htbp]
+%   \begin{center}
+%     \includegraphics[width = \textwidth]{99_SourceAreas.png}
+%     \caption{(a) Different source areas from which we released virtual
+%     dispersers. We only considered contiguous protected areas (national parks,
+%     game reserves, and forest reserves) that were larger than 700
+%     km\textsuperscript{2} (green). This area corresponds to the average home
+%     range requirement for viable wild dog populations \citep{Pomilia.2015}. To
+%     render potential immigrants into the study system, we also initiated
+%     dispersers within a buffer zone (blue) surrounding the main study area. (b)
+%     Source points from which dipsersers were released. 50'000 dispersers were
+%     released from the main study area (green dots) and another 30'000 dispersers
+%     within the virtual buffer (blue dots).}
+%     \label{SourcePoints}
+%   \end{center}
+% \end{figure}
 
@@
 then determined the relative traversal frequency by simulated trajectories
-through each checkpoint for different numbers of simulations (1 to 50'000). To
-assess variability in the relative traversal frequency, we repeatedly sampled
-trajectories 100 times and computed the mean traversal frequency across
-replicates, as well as the 95\% prediction-interval. We deemed that a checkpoint
-converged as soon as the width of the prediction interval for the traversal
-frequency across replicates dropped below a value of 0.01.
+through each checkpoint as we increased the number of simulated individuals (1
+to 50'000). To assess variability in the relative traversal frequency, we
+repeatedly sampled trajectories 100 times and computed the mean traversal
+frequency across replicates, as well as the 95\% prediction-interval. We deemed
+that a checkpoint converged as soon as the width of the prediction interval for
+the traversal frequency across replicates dropped below a value of 0.01.
 
@@
 the same trajectory crossed a raster-cell twice, we only counted it once,
-thereby mitigating potential biases caused by individuals that were trapped and
-moved in circles. To achieve high performance rasterization, we used the
-R-package {\tt terra} \citep{Hijmans.2020}.
+thereby mitigating potential biases caused by individuals that moved in circles
+because they were trapped by dispersal barriers. To achieve high performance
+rasterization, we used the R-package {\tt terra} \citep{Hijmans.2020}.
 
@@
 imply that connections between other regions are impossible. In fact, the same
-logic could easily be expanded to include other protected areas. To quantify
+process could easily be expanded to include other protected areas. To quantify
 inter-patch connectivity, we computed the relative frequency at which dispersers
@@
 intersected with the corresponding national park at least once. We also recorded
-the number of steps required until the first intersection with the polygon of
-the respective national park. This allowed us to determine \textit{if} and
-\textit{how often} dispersers moved between certain national parks, as well as
-\textit{how long} dispersers had to move to realize those connections.
+the number of steps required until the first intersection with the respective
+national park. This allowed us to determine \textit{if} and \textit{how often}
+dispersers moved between certain national parks, as well as \textit{how long}
+dispersers had to move to realize those connections.
 
@@
 
-Assuming that all other covariates are held constant at their means, the habitat
-kernel reveals that dispersing wild dogs avoid water but prefer its proximity.
+Assuming that all other covariates are held at their means, the habitat kernel
+reveals that dispersing wild dogs avoid water but prefer its proximity.
 Similarly, dispersers avoid areas that are covered by woodlands, yet prefer
@@
 inside the main source area \Cref{SourcePoints}(a), only 4.5\% eventually hit a
-map boundary, suggesting that we successfully prevented biases due to boundary
-effects. In contrast, 78\% of the 30'000 dispersers originating from the buffer
-zone eventually hit a map boundary, yet this was to be expected since many of
-those dispersers originated from source points located close to the map
-boundary.
+map boundary, suggesting that biases due to boundary effects are moderate. In
+contrast, 78\% of the 30'000 dispersers originating from the buffer zone
+eventually hit a map boundary, yet this was to be expected since many of those
+dispersers originated from source points located close to the map boundary.
 
@@
 \subsection{Further Considerations (70\%)}
-% Modelling Mortality
+% Modeling Mortality
 Although we did not render mortality, animals regularly die during dispersal,
@@
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_ModelingDecisions.png}
-  \caption{Five major modeling decisions that a researcher needs to consider
-  when simulating dispersers to assess landscape connectivity. (1) Number of
-  simulated trajectories. (2) Handling individual variation: we used point
-  estimates when simulating dispersers, assuming that there was no individual
-  variation. Alternatively, however, one could draw preferences for each
-  simulated individually based on model uncertainty. (3) Dispersal duration.
-  While one could draw the number of simulated steps based on observed dispersal
-  events, we opted for an alternative and simulated each individual for 2'000
-  steps, which was at the upper end of observed dispersal durations. This allows
-  to easily shorten the generated trajectories afterwards and to investigate the
-  sensitivity of results with regards to the dispersal duration. (4) Boundary
-  behavior. We allowed dispersers to leave and re-enter the main study area
-  through a buffer zone with randomized covariate values. Alternatively, one
-  could discard transgressing random steps, thereby forcing dispersers to remain
-  within the study area or simply terminate the simulation, assuming the
-  individual has disappeared. (5) Source point location. The location of source
-  points should optimally be biologically informed. For our simulation, we
-  placed source points within protected areas large enough to sustain viable
-  wild dog populations. Conceivable alternatives include placing source points
-  according to a habitat suitability model or based on abundance information.
-  Importantly, one must also consider potential immigrants from outside the main
-  study area.}
-  \label{ModelingDecisions}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\textwidth]{99_ModelingDecisions.png}
+%   \caption{Five major modeling decisions that a researcher needs to consider
+%   when simulating dispersers to assess landscape connectivity. (1) Number of
+%   simulated trajectories. (2) Handling individual variation: we used point
+%   estimates when simulating dispersers, assuming that there was no individual
+%   variation. Alternatively, however, one could draw preferences for each
+%   simulated individually based on model uncertainty. (3) Dispersal duration.
+%   While one could draw the number of simulated steps based on observed dispersal
+%   events, we opted for an alternative and simulated each individual for 2'000
+%   steps, which was at the upper end of observed dispersal durations. This allows
+%   to easily shorten the generated trajectories afterwards and to investigate the
+%   sensitivity of results with regards to the dispersal duration. (4) Boundary
+%   behavior. We allowed dispersers to leave and re-enter the main study area
+%   through a buffer zone with randomized covariate values. Alternatively, one
+%   could discard transgressing random steps, thereby forcing dispersers to remain
+%   within the study area or simply terminate the simulation, assuming the
+%   individual has disappeared. (5) Source point location. The location of source
+%   points should optimally be biologically informed. For our simulation, we
+%   placed source points within protected areas large enough to sustain viable
+%   wild dog populations. Conceivable alternatives include placing source points
+%   according to a habitat suitability model or based on abundance information.
+%   Importantly, one must also consider potential immigrants from outside the main
+%   study area.}
+%   \label{ModelingDecisions}
+% \end{figure}
 
@@
 some of the non-trivial modeling decisions involved. In particular, we will
-discuss five modeling decisions (\Cref{ModelingDecisions}): (1) number of
-simulated individuals, (2) location of source points, (3) dispersal duration,
-(4) boundary behavior, and (5) how to handle individual variability.
+discuss five modeling decisions: (1) number of simulated individuals, (2)
+location of source points, (3) dispersal duration, (4) boundary behavior, and
+(5) how to handle individual variability.
 

### 2021-07-16 Chapter_1/04_Manuscript/Manuscript__2021_07_15.tex
Rewriting introduction

--- 2021-07-16 Chapter_1/04_Manuscript/Manuscript.tex
@@
 
-\textbf{Keywords:} dispersal, simulation, movement, integrated step selection
-function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, Lycaon pictus
+\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
+Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
+pictus
 
@@
 \begin{abstract}
-
-%------------------------------------------------------------------------------
-% Short Version
-%------------------------------------------------------------------------------
-% The ability to disperse is contingent a sufficient degree of landscape
-% connectivity, which is why the identification and preservation of movement
-% corridors that promote connectivity has become a task of extraordinary
-% importance. Currently, ecologists mainly rely on least-cost analysis or circuit
-% theory to investigate connectivity, albeit both methods make several assumptions
-% are hardly met in reality. To address these issues, simulations from
-% individual-based movement models have been proposed, yet a unified framework to
-% simulate dispersal and quantify connectivity is lacking.
-%
-% Here, we employ a simple three-step workflow that combines several
-% already-existing methods to assess connectivity using simulated dispersal
-% trajectories. In the first step, we use integrated step selection functions to
-% parametrize a mechanistic movement model rendering dispersal behavior. In a
-% second step, we apply the parametrized model as an individual-based movement
-% model to simulate dispersal trajectories. Finally, in the third step, we
-% consolidate simulated trajectories into three complementary connectivity maps,
-% each focusing on a different aspect of landscape connectivity.
-%
-% We apply these steps to assess landscape connectivity for dispersing African
-% wild dogs (\textit{Lycaon pictus}) southern Africa. We thereby demonstrate that
-% simulations from integrated step selection functions offer a simple, yet
-% powerful alternative to traditional connectivity modeling techniques.
-
-%------------------------------------------------------------------------------
-%  LONG VERSION
-%------------------------------------------------------------------------------
 % Importance of Dispersal & Connectivity
@@
 of landscape connectivity, which is why the identification and preservation of
-movement corridors that promote connectivity has become a task of exceptional
-importance.
+movement corridors that promote connectivity has become a task of extraordinary
+importance for conservation authorities.
 
 % Quantifying connectivity
-Over the past two decades, ecologists have primarily relied on least-cost
-analysis and circuit theory to investigate connectivity. Unfortunately, both
-methods make several restrictive assumptions that limit their usefulness in
-reality. Least-cost analysis, for instance, assumes that animals move towards a
-pre-conceived endpoint and choose a cost-minimizing route accordingly. Circuit
-theory, in contrast, presumes a random walk without any directional persistence.
-To address these shortcomings, simulations from individual-based movement models
-have been proposed and applied, yet a unified framework to simulate dispersal
-and to derive connectivity maps based on simulated dispersal events is lacking.
+During the past two decades, ecologists have largely relied on graph-based
+methods, such as least-cost analysis or circuit theory, to investigate
+connectivity. Despite their popularity, both methods make several restrictive
+assumptions that limit their usefulness in reality. Least-cost analysis, for
+instance, assumes that animals move towards a pre-conceived endpoint and choose
+a cost-minimizing route accordingly. Circuit theory, in contrast, presumes that
+animals move according to a random walk without any directional persistence. To
+overcome these shortcomings, dispersal simulations from individual-based
+movement models have been proposed, yet a unified framework to simulate
+dispersal and assess connectivity is lacking.
 
 % What we do
-% and synthesize a simple
-% three-step workflow that uses
-Here, we combine several already-existing methods to simulate dispersal and
-investigate landscape connectivity in three distinct steps. In the first step,
-integrated step selection functions are used to parametrize a mechanistic
-movement model. In the second step, the parametrized model is applied as an
-individual-based movement model to simulate dispersal trajectories. In the third
-step, simulated trajectories are consolidated and synthesized into three
-complementary connectivity maps, each focusing on a different aspect of
-landscape connectivity. The set of maps includes a heatmap, revealing frequently
+Here, we examine the use of integrated step selection analysis as a simple but
+powerful framework for fitting individual-based movement models and simulating
+dispersal. For this, we analyse data of 16 dispersing wild dogs (\textit{Lycaon
+pictus}) and parametrize a fully mechanistic movement model describing how
+dispersers move through the available landscape. We then employ the model to
+simulate 80'000 dispersing wild dogs moving across the world's largest
+transboundary conservation area. Finally, we use the simulated trajectories to
+derive three complementary connectivity maps, each focusing on a different
+aspect of connectivity. The set of maps includes a heatmap, revealing frequently
 traversed areas, a betweenness-map, delineating critical dispersal corridors,
@@
 
-% Case study
-As a case study, we apply the proposed workflow to assess landscape connectivity
-for dispersing African wild dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi
-Transfrontier Conservation Area, the world's largest transboundary conservation
-area. With this, we demonstrate that the outlined workflow offers a simple, yet
-powerful alternative to traditional connectivity modeling techniques. Although
-such individual-based simulations necessitate informed decisions about the
-number of simulated individuals and the duration of simulated dispersal events,
-they enable to relax several of the unrealistic assumptions inherent to
-least-cost analysis or circuit theory.
+% Relevance
+We conclude that simulations from integrated step selection functions indeed
+offer a powerful alternative to traditional connectivity modeling techniques and
+requires fewer unrealistic assumptions about dispersal. Nevertheless, some
+non-trivial modeling decisions such as the number and duration of simulated
+dispersal events need to be considered.
 
@@
 % Create linenumbers
-\linenumbers
+% \linenumbers
 
 \section{Introduction}
-
 % Importance of Dispersal & Connectivity
-\subsection{Dispersal \& Connectivity Models (90\%)}
+\subsection{Dispersal \& Connectivity (90\%)}
 Successful dispersal of individuals requires a sufficient degree of landscape
@@
 and protection of major dispersal corridors has become a fundamental task in
-conservation science \citep{Nathan.2008, Doerr.2011, Rudnick.2012}. To pinpoint
-relevant dispersal corridors, a comprehensive understanding of the factors that
-limit dispersal of the focal species and an adequate connectivity model are
-required \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the two
-most prominent connectivity models are least-cost path analysis (LCPA;
+conservation science \citep{Nathan.2008, Doerr.2011, Rudnick.2012}. Pinpointing
+relevant dispersal hotspots not only necessitates information on movement
+behavior and knowledge about factors that limit dispersal \citep{Baguette.2013,
+Vasudev.2015, Hofmann.2021}, but also an adequate connectivity model
+\citep{Diniz.2020}. To date, two of the most prominent connectivity models used
+to infer movement corridors are least-cost path analysis (LCPA;
 \citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
 McRae.2008}), both graph-based approaches that estimate conductance of the
-landscape based on predicted habitat permeability \citep{Zeller.2012}.
-
-% Issues with Both Methods
-\subsection{Issues with Traditional Methods (90\%)}
-Despite their intuitive nature and ease of use, both methods make several
-assumptions that are hardly ever met in reality \citep{Diniz.2020}. With LCPA,
-for instance, a least costly route always exists, even if associated movement
-costs are unreasonably high and will never be incurred by a dispersing
-individual. The method also presumes that animals move towards a pre-conceived
-endpoint and choose a cost-minimizing route accordingly. While these assumptions
-may by fulfilled by migrating animals, dispersers usually move into unfamiliar
-territory and are therefore less likely to be aware of associated movement costs
-\citep{Koen.2014, Abrahms.2017, Cozzi.2020}. CT relaxes these assumptions, yet
-posits that animals move according to a random walk, entailing that directional
-biases cannot be rendered \citep{Diniz.2020}. In reality, however, directional
-persistence is regularly observed \citep{Bovet.1991, Schultz.2001}, especially
-in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}. Some of these
-deficiencies can be addressed using alternative graph-based approaches, such as
-least-cost \textit{corridors} \citep{Pinto.2009}, \textit{thresholded}
-least-cost paths \citep{Landguth.2012}, or \textit{randomized} least-cost paths
-\citep{Panzacchi.2016}, yet a certain degree of arbitrariness remains. Finally,
-none of the methods is capable of rendering the temporal dimension of dispersal
-\citep{Diniz.2020}, meaning that statements about the expected duration required
-to traverse a certain corridor are impossible.
+landscape based on estimated habitat permeability for the focal species
+\citep{Zeller.2012}.
+
+% Issues with Least-Cost Paths
+\subsection{Issues with LCPs and CT (90\%)}
+Despite their intuitive nature and ease of use, least-cost analysis and circuit
+theory make several rigorous assumptions that are hardly ever met in reality
+\citep{Diniz.2020}. With LCP analysis, for instance, a least costly route always
+exists, even if associated movement costs are unreasonably high and will never
+be incurred by a dispersing individual. The method also presumes that animals
+move towards a pre-conceived endpoint and choose a cost-minimizing route
+accordingly, which further requires perfect knowledge about landscape and
+associated movement costs. While this may be a reasonable assumption for
+migrating animals, dispersers usually move into unfamiliar territory and are
+therefore unaware of associated movement costs and potential endpoints
+\citep{Koen.2014, Abrahms.2017, Cozzi.2020}. Circuit theory relaxes these
+assumptions, yet requires that animals move according to a random walk,
+entailing that directional biases cannot be rendered \citep{Diniz.2020}.
+Nevertheless, directional movements are regularly observed \citep{Bovet.1991,
+Schultz.2001}, particularly in dispersing individuals \citep{Cozzi.2020,
+Hofmann.2021}. Some of these deficiencies can be addressed using intermediate
+methods, such as (e.g. least-cost \textit{corridors} \citep{Pinto.2009},
+\textit{thresholded} least-cost paths \citep{Landguth.2012}, and
+\textit{randomized} least-cost paths \citep{Panzacchi.2016}, yet a certain
+degree of arbitrariness remains. In addition, neither of these methods is
+capable of rendering the temporal dimension of dispersal \citep{Diniz.2020},
+such that statements about the expected duration required to traverse a certain
+corridor are impossible.
 
@@
 \subsection{What about IBMMs? (90\%)}
-To overcome the unrealistic assumptions inherent to traditional connectivity
-modeling techniques, simulations from individual-based movement models (IBMMs)
-have been proposed and applied \citep{Diniz.2020}. In these models, dispersal
-trajectories are simulated based on movement rules that determine how
-individuals move across or interact with the prevailing landscape
-\citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004,
-Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019,
-Zeller.2020, Vasudev.2021}. Instead of enforcing connections towards a known
-endpoint, IBMMs allow individuals to choose their route as they move and enable
-to render potential interactions between movement behavior and environmental
-conditions. IBMMs furthermore enable to model directional persistence and to
-study the temporal dimension of dispersal \citep{Diniz.2020}. Despite these
-advantages, however, a unified framework to simulate dispersal and to derive
-connectivity maps based on simulated dispersal events is lacking.
-
-% The Power of Step Selection Functions
-\subsection{Step Selection Functions (90\%)}
-A promising framework that has been developed more recently...
-Over the past decade, the increased availability of large amounts of data and an
-increased computational power have led to the development of numerous techniques
-to study animal movement and resource selection during dispersal
-\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012}. Especially
-\textit{resource selection functions} \citep{Boyce.2002}, \textit{step selection
-functions} \citep{Fortin.2005}, and \textit{path selection functions}
-\citep{Cushman.2010} have pushed the field of movement ecology forward. These
-methods allow estimating habitat preferences of the focal species by comparing
-covariates at locations visited by the animal to the same covariates at
-locations available to, but not visited by the animal \citep{Boyce.2002,
-Fortin.2005, Cushman.2010, Thurfjell.2014}. While regular SSFs were intended to
-learn about relative habitat preferences of the focal species
-\citep{Fortin.2005, Thurfjell.2014, Avgar.2017}, the method has recently been
-generalized to \textit{integrated} SSFs and now enables to jointly study habitat
-and movement preferences, as well as potential interactions between them
-\citep{Avgar.2016, Signer.2017, Fieberg.2020}. ISSFs therefore provide a
-relatively simple means to model complex movement behavior, where movement is
-viewed as the result of two intertwined behavioral kernels (e.g.
-\citealp{Prokopenko.2017, Munden.2020}). Importantly, a parametrized ISSF model
-represents a fully mechanistic movement model based on which individual movement
-trajectories can be simulated \citep{Avgar.2016, Signer.2017}. In fact,
-\cite{Signer.2017} used ISSF to simulate steady state utilization distributions
-of resident animals that were moving around a point of attraction. However, the
-degree to which such simulations are helpful in detecting movement corridors and
-modeling landscape connectivity remains to be investigated.
-
-% What we do
-\subsection{Proposed Framework (90\%)}
-Here, we therefore propose to combine several already-existing methods into a
-simple three-step workflow to simulate dispersal and investigate landscape
-connectivity. In the first step, integrated step selection functions are used to
-parametrize a mechanistic movement model rendering dispersal behavior. In the
-second step, the parametrized model is applied as an individual-based movement
-model to simulate dispersal trajectories. Finally, in the third step, simulated
-trajectories are consolidated into three complementary connectivity maps, each
-focusing on a different aspect of landscape connectivity. The set of maps
-includes a heatmap, revealing frequently traversed areas, a betweenness-map,
-delineating critical dispersal corridors, and a map of inter-patch connectivity,
-depicting the presence or absence of functional links between distinct habitat
-patches.
+To overcome the issues inherent to LCPs and CT, individual-based movement models
+(IBMMs) have been proposed and applied \citep{Diniz.2020}. In these models,
+individual dispersal trajectories are simulated based according to movement
+rules that determine how individuals move across or interact with the prevailing
+landscape \citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004,
+Revilla.2004, Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016,
+Hauenstein.2019, Zeller.2020, Vasudev.2021}. To quantify landscape connectivity
+from simulated trajectories, as set of connectivity metrics, such as inter-patch
+connectivity and traversal frequency, to reveal major dispersal corridors
+\citep{Kanagaraj.2013, BastilleRousseau.2018, Hauenstein.2019, Zeller.2020} can
+be calculated. However, while IBMMS can be employed to overcome any of the
+shortcomings intrinsic to LCPs and CT, they are subject to a vast amount of
+subjective, non-trivial modeling decisions. Moreover, they can be challenging to
+fit and require vast amounts of movement data, ideally collected during
+dispersal \citep{Diniz.2020}. Consequently, alternative methods that require
+fewer modeling decisions and are straight forward to apply are desirable.
+
+% Step Selection Analysis
+\subsection{Step Selection Analysis (90\%)}
+Here, we investigate the usefulness of integrated step selection functions
+(ISSFs, \citealp{Avgar.2016}), as a relatively simple but powerful IBMM based on
+which dispersal can be simulated. While regular SSFs were intended to learn
+about relative habitat preferences of the focal species \citep{Fortin.2005,
+Thurfjell.2014, Avgar.2017}, the method has recently been generalized to
+\textit{integrated} SSFs and now enables to jointly study habitat and movement
+preferences, as well as potential interactions between them \citep{Avgar.2016,
+Signer.2017, Fieberg.2020}. ISSFs therefore provide a relatively simple means to
+model complex movement behavior, where movement is viewed as the result of two
+intertwined behavioral kernels (e.g. \citealp{Prokopenko.2017, Munden.2020}).
+Importantly, a parametrized ISSF model can be employed as a fully mechanistic
+movement model based on which individual movement trajectories can be simulated
+\citep{Avgar.2016, Signer.2017}. In fact, \cite{Signer.2017} used ISSF to
+simulate steady state utilization distributions of resident animals that were
+moving around a point of attraction. However, the degree to which such
+simulations are helpful in detecting movement corridors and modeling landscape
+connectivity remains to be investigated.
 
@@
 \subsection{Study Species \& Study Area (90\%)}
-We exemplify application of the proposed workflow by investigating landscape
-connectivity for the endangered African wild dog (\textit{Lycon pictus}). While
+One of the species for which long-term viability relies on sufficient landscape
+connectivity is the endangered African wild dog \textit{Lycon pictus}. While
 once present across entire sub-Saharan Africa, wild dogs have disappeared from a
-vast majority of their historic range, mainly due to persecution by humans,
-habitat fragmentation and destruction, and deadly diseases
-\citep{Woodroffe.2012}. Within the few remaining subpopulations, wild dogs form
-cohesive packs comprising 8 to 12 adults and their offspring \cite{McNutt.1995}.
-After reaching sexual maturity, male and female offspring form same-sex
-coalitions and disperse from their natal pack in search for potential mating
-partners and a suitable territory to settle \citep{McNutt.1996, Behr.2020}. New
-packs are formed when dispersing coalitions join unrelated opposite-sex
-dispersing coalitions \citep{McNutt.1996}. During dispersal, wild dogs can cover
-several hundred kilometers across a variety of landscapes
+vast majority of their historic range due to persecution by humans, habitat
+fragmentation and destruction, and deadly diseases \citep{Woodroffe.2012}. As of
+today, only 6'000 free-ranging individuals remain in small and spatially
+scattered subpopulations \citep{Woodroffe.2012}. Within those subpopulations,
+wild dogs form cohesive packs comprising 8 to 12 adults and their offspring
+\cite{McNutt.1995}. After reaching sexual maturity, male and female offspring
+form same-sex coalitions and disperse from their natal pack in search for
+potential mating partners and a suitable territory to settle \citep{McNutt.1996,
+Behr.2020}. New packs are formed when dispersing coalitions join unrelated
+opposite-sex dispersing coalitions \citep{McNutt.1996}. Dispersing wild dogs can
+cover several hundred kilometers across a variety of landscapes
 \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the
-main strongholds for this species lies near the Moremi Game Reserve in northern
+few strongholds for this species lies near the Moremi Game Reserve in northern
 Botswana, which is part of the world's largest transboundary protected area,
 namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA). This
-area was originally devoted towards facilitating the migration of elephants, yet
-it can be expected to yield benefits to a great multitude of other species
-\citep{Elliot.2014, Brennan.2020, Hofmann.2021}.
+area has originally been intended to facilitate migration of elephants, but is
+expected to provide benefits to a multitude of other species \citep{Elliot.2014,
+Brennan.2020, Hofmann.2021}.
 
@@
 We previously assessed landscape connectivity for dispersing African wild dogs
-within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}. More
-specifically, we fitted a habitat selection model and predicted a permeability
-surface based on which we computed least-cost paths and corridors between
-protected areas. We now expand on this knowledge and use ISSFs to develop a more
-detailed, mechanistic movement model of dispersing wild dogs
-(\Cref{GraphicalAbstract}). We then employ the model to simulate 80'000
-dispersers moving 2'000 steps across the KAZA-TFCA. Finally, we generate a
-heatmap, betweenness map and a map of interpatch connectivity. We thereby
-exemplify how ISSFs can be utilized for dispersal simulations and we discuss
-several benefits of this approach over traditional connectivity modeling
-techniques such as least-cost analysis and circuit theory. Most importantly,
-simulations based on ISSFs provide a more generic view on how connectivity
-emerges and to which degree connectivity depends on the dispersal duration. In
-addition, by generating proper dispersal trajectories, network theory can be
-applied to calculate network metrics that are pertinent to connectivity
-analysis.
+within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}.
+Specifically, we fitted a habitat selection model and predicted a permeability
+surface that we used to compute least-cost paths and corridors between protected
+areas. We now expand on this knowledge and use ISSFs to develop a more detailed,
+mechanistic movement model of dispersing wild dogs (\Cref{GraphicalAbstract}).
+We employ the model to simulate dispersers moving across the KAZA-TFCA and
+generate three distinct connectivity maps, each shedding light onto a different
+aspect of connectivity. The set of maps includes a heatmap, revealing frequently
+traversed areas, a betweenness-map, delineating critical dispersal corridors and
+bottlenecks, and a map of inter-patch connectivity, depicting the presence or
+absence of functional links between national parks as well as the average
+dispersal duration required to realize those links. With this work, we exemplify
+how ISSFs can be utilized for dispersal simulations and we discuss several
+benefits of this approach over traditional connectivity modeling techniques such
+as least-cost analysis and circuit theory. Most importantly, simulations based
+on ISSFs provide a more generic view on how connectivity emerges and to which
+degree connectivity depends on the dispersal duration. In addition, by
+generating proper dispersal trajectories, network theory can be applied to
+calculate network metrics that are pertinent to connectivity analysis.
 
@@
 \subsection{GPS Relocation Data (90\%)}
-GPS data of 16 dispersing wild dogs was obtained as described in
-\citep{Cozzi.2020, Hofmann.2021}. We collected data on 7 female and 9 male
-dispersing coalitions between 2011 and 2019 from a free-ranging population
-inhabiting the Moremi National Park in northern Botswana. In some instances,
-exact dispersal dates were known from field observations, in other cases we
-determined dispersal phases using the net-squared displacement metric. Net
-squared displacement measures the squared Euclidean distance of a GPS relocation
-to a reference point \citep{Borger.2012}, which in our case was set to the
-center of each individual's natal home range. Thus, dispersal was deemed to have
-started when an individual left its natal home range and ended once individuals
-became sedentary again. Because behavior during dispersal is more pertinent to
+GPS data of dispersing wild dogs was obtained as described in \citep{Cozzi.2020,
+Hofmann.2021}. We collected data on 7 female and 9 male dispersing coalitions
+between 2011 and 2019 from a free-ranging population inhabiting the Moremi
+National Park in northern Botswana. In some instances, exact dispersal dates
+were known from field observations, in other cases we determined dispersal
+phases using the net-squared displacement metric. Net squared displacement
+measures the squared Euclidean distance of a GPS relocation to a reference point
+\citep{Borger.2012}, which in our case was set to the center of each
+individual's natal home range. Thus, dispersal was deemed to have started when
+an individual left its natal home range and ended once individuals became
+sedentary again. Because behavior during dispersal is more pertinent to
 landscape connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded all data
@@
 water layers (see \citealp{Wolski.2017} and Appendix A3 in
-\citealp{Hofmann.2021}). Resulting layers thus temporally aligned with each
-dispersal event, ensuring minimal temporal lag between observed movements and
-corresponding water layer. We furthermore computed a proxy for human influence,
-rendering anthropogenic pressures stemming from human-density, agricultural
-sites, and roads. All spatial layers were coarsened or interpolated to a target
-resolution of 250 m by 250 m. A detailed description of the preparation of each
-habitat covariate is given in \cite{Hofmann.2021}.
+\citealp{Hofmann.2021}). Resulting water layers thus temporally aligned with
+each dispersal event, ensuring minimal temporal lag between observed movements
+and corresponding water layer. We furthermore computed a proxy for human
+influence, rendering anthropogenic pressures stemming from human-density,
+agricultural sites, and roads. All spatial layers were coarsened or interpolated
+to a target resolution of 250 m by 250 m. A detailed description of the
+preparation of each habitat covariate is given in \cite{Hofmann.2021}.
 
@@
 written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
-\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
+\citep{Eddelbuettel.2011, Eddelbuettel.2013}
 
@@
 dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model
-(hereafter referred to as base model), no interactions between habitat
+(referred to as base model hereafter), no interactions between habitat
 covariates (\textsf{Water, DistanceToWater\textsuperscript{0.5}, Woodland,
@@
 then determined the relative traversal frequency by simulated trajectories
-through each checkpoint as we increased the number of simulated individuals (1
-to 50'000). To assess variability in the relative traversal frequency, we
-repeatedly sampled trajectories 100 times and computed the mean traversal
-frequency across replicates, as well as the 95\% prediction-interval. We deemed
-that a checkpoint converged as soon as the width of the prediction interval for
-the traversal frequency across replicates dropped below a value of 0.01.
+through each checkpoint for different numbers of simulations (1 to 50'000). To
+assess variability in the relative traversal frequency, we repeatedly sampled
+trajectories 100 times and computed the mean traversal frequency across
+replicates, as well as the 95\% prediction-interval. We deemed that a checkpoint
+converged as soon as the width of the prediction interval for the traversal
+frequency across replicates dropped below a value of 0.01.
 
@@
 the same trajectory crossed a raster-cell twice, we only counted it once,
-thereby mitigating potential biases caused by individuals that moved in circles
-because they were trapped by dispersal barriers. To achieve high performance
-rasterization, we used the R-package {\tt terra} \citep{Hijmans.2020}.
+thereby mitigating potential biases caused by individuals that were trapped and
+moved in circles. To achieve high performance rasterization, we used the
+R-package {\tt terra} \citep{Hijmans.2020}.
 

### 2021-09-08 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

@@
 \author{
-  David D. Hofmann\textsuperscript{1,\S} \and
+  David D. Hofmann\textsuperscript{1,2,\S} \and
+  Gabriele Cozzi\textsuperscript{1,2} \and
   John W. McNutt\textsuperscript{2} \and
   Arpat Ozgul\textsuperscript{1} \and
-  Gabriele Cozzi\textsuperscript{1,2} \and
   Dominik M. Behr\textsuperscript{1,2}
@@
 %------------------------------------------------------------------------------
-% Importance of Dispersal & Connectivity
-Dispersal of individuals is a crucial pre-requisite for long-term species
-persistence. However, the ability to disperse is contingent a sufficient degree
-of landscape connectivity, which is why the identification and preservation of
-movement corridors that promote connectivity has become a task of exceptional
-importance.
-
-% Quantifying connectivity
-Over the past two decades, ecologists have primarily relied on least-cost
-analysis and circuit theory to investigate connectivity. Unfortunately, both
-methods make several restrictive assumptions that limit their usefulness in
-reality. Least-cost analysis, for instance, assumes that animals move towards a
-pre-conceived endpoint and choose a cost-minimizing route accordingly. Circuit
-theory, in contrast, presumes a random walk without any directional persistence.
-To address these shortcomings, simulations from individual-based movement models
-have been proposed and applied, yet a unified framework to simulate dispersal
-and to derive connectivity maps based on simulated dispersal events is lacking.
+% Importance of dispersal & connectivity and mow its measured
+Dispersal of individuals is a crucial prerequisite for long-term species
+persistence but depends on a sufficient degree of landscape connectivity. To
+date, connectivity has primarily been quantified using least-cost analysis and
+circuit theory, albeit both methods make assumptions that are rarely met in
+reality. Least-cost analysis assumes that animals move towards a preconceived
+destination and choose a cost-minimizing route accordingly, whereas circuit
+theory presumes a complete random walk without directional persistence. While
+these shortcomings can be overcome by spatio-temporally explicitly simulating
+dispersal trajectories from individual-based movement models, a unified
+approach for such simulations is lacking.
 
 % What we do
-% and synthesize a simple
-% three-step workflow that uses
-Here, we combine several already-existing methods to simulate dispersal and
-investigate landscape connectivity in three distinct steps. In the first step,
-integrated step selection functions are used to parametrize a mechanistic
-movement model. In the second step, the parametrized model is applied as an
-individual-based movement model to simulate dispersal trajectories. In the third
-step, simulated trajectories are consolidated and synthesized into three
-complementary connectivity maps, each focusing on a different aspect of
-landscape connectivity. The set of maps includes a heatmap, revealing frequently
-traversed areas, a betweenness-map, delineating critical dispersal corridors,
-and a map of inter-patch connectivity, depicting the presence or absence of
-functional links between distinct habitat patches.
+Here, we present a simple three-step workflow to simulate dispersal and asses
+landscape connectivity starting from empirical GPS movement data. We exemplify
+the workflow using dispersal data collected on the endangered African wild dog
+(\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation Area,
+the world's largest transboundary conservation area. In step one, we use
+integrated step selection functions to parametrize a mechanistic movement model
+describing habitat and movement preferences of dispersers. In step two, we
+employ the parametrized model to simulate individual dispersal trajectories
+across the landscape. In step three, we interpret simulated trajectories by
+converting them into three complementary connectivity maps: a heatmap revealing
+frequently traversed areas, a betweenness-map delineating dispersal corridors,
+and a map of inter-patch connectivity depicting the presence and intensity of
+functional links between habitat patches.
+
+% Some results
+We find that simulating dispersal from indvidual-based movement models fitted
+via step selection functions are readily interpretable and offer great
+flexibility to realistically render dispersal movements. Moreover, by not
+enforcing connections towards known endpoints but instead letting dispersers
+react to the environment as they go, such simulations overcome shortcomings
+inherent least-cost path analysis, while at the same time allowing for
+autocorrelation in simulated steps, which can not be incorporated using circuit
+theory. Inspite of these advantages, simulating dispersal entails numerous
+non-trivial modeling decisions, such as the number of simulated individuals and
+the duration of simulated dispersal events.
 
 % Case study
-As a case study, we apply the proposed workflow to assess landscape connectivity
-for dispersing African wild dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi
-Transfrontier Conservation Area, the world's largest transboundary conservation
-area. With this, we demonstrate that the outlined workflow offers a simple, yet
-powerful alternative to traditional connectivity modeling techniques. Although
-such individual-based simulations necessitate informed decisions about the
-number of simulated individuals and the duration of simulated dispersal events,
-they enable to relax several of the unrealistic assumptions inherent to
-least-cost analysis or circuit theory.
+We  show that simulations from step-selection functions offer a simple yet
+powerful alternative to traditional connectivity modeling techniques, although
+necessitating informed decisions about the number of simulated individuals and
+the duration of simulated dispersal events. Ultimately, our workflow permits a
+more mechanistic understanding of dispersal behaviour and landscape connectivity
+making it useful for a variety of applications in ecological, evolutionary, and
+conservation research.
 
@@
 % Importance of Dispersal & Connectivity
-\subsection{Dispersal \& Connectivity Models (90\%)}
-Successful dispersal of individuals requires a sufficient degree of landscape
+\subsection{Importance of Connectivity \& Connectivity Models}
+Dispersal of individuals is a vital process that allows species to maintain
+genetic diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
+Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and colonize
+or recolonize unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However,
+the ability to disperse is contingent on a sufficient degree of landscape
 connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the identification
-and protection of major dispersal corridors has become a fundamental task in
-conservation science \citep{Nathan.2008, Doerr.2011, Rudnick.2012}. To pinpoint
-relevant dispersal corridors, a comprehensive understanding of the factors that
-limit dispersal of the focal species and an adequate connectivity model are
-required \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the two
-most prominent connectivity models are least-cost path analysis (LCPA;
-\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
-McRae.2008}), both graph-based approaches that estimate conductance of the
-landscape based on predicted habitat permeability \citep{Zeller.2012}.
+and protection of dispersal corridors that promote connectivity has become a
+task of fundamental importance \citep{Nathan.2008, Doerr.2011, Rudnick.2012}.
+The identification of dispersal corridors not only necessitates a comprehensive
+understanding of the factors that limit dispersal, but also an appropriate model
+to estimate connectivity \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To
+date, the two most prominent connectivity models in the literature are
+least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory
+(CT; \citealp{McRae.2006, McRae.2008}), both graph-based approaches that
+quantify conductance of the landscape based on habitat permeability
+\citep{Zeller.2012}.
 
 % Issues with Both Methods
-\subsection{Issues with Traditional Methods (90\%)}
-Despite their intuitive nature and ease of use, both methods make several
-assumptions that are hardly ever met in reality \citep{Diniz.2020}. With LCPA,
-for instance, a least costly route always exists, even if associated movement
-costs are unreasonably high and will never be incurred by a dispersing
-individual. The method also presumes that animals move towards a pre-conceived
-endpoint and choose a cost-minimizing route accordingly. While these assumptions
-may by fulfilled by migrating animals, dispersers usually move into unfamiliar
-territory and are therefore less likely to be aware of associated movement costs
-\citep{Koen.2014, Abrahms.2017, Cozzi.2020}. CT relaxes these assumptions, yet
-posits that animals move according to a random walk, entailing that directional
-biases cannot be rendered \citep{Diniz.2020}. In reality, however, directional
-persistence is regularly observed \citep{Bovet.1991, Schultz.2001}, especially
-in dispersing individuals \citep{Cozzi.2020, Hofmann.2021}. Some of these
-deficiencies can be addressed using alternative graph-based approaches, such as
-least-cost \textit{corridors} \citep{Pinto.2009}, \textit{thresholded}
-least-cost paths \citep{Landguth.2012}, or \textit{randomized} least-cost paths
-\citep{Panzacchi.2016}, yet a certain degree of arbitrariness remains. Finally,
-none of the methods is capable of rendering the temporal dimension of dispersal
-\citep{Diniz.2020}, meaning that statements about the expected duration required
-to traverse a certain corridor are impossible.
+\subsection{Issues with Traditional Connectivity Models}
+Despite their intuitive nature and ease of use, both methods make rigorous
+assumptions that are hardly ever met in reality \citep{Diniz.2019}. With LCPA,
+for instance, a least-costly route always exists, even if associated movement
+costs are unreasonably high and will never be incurred by dispersing individuals
+\citep{Sawyer.2011}. The method also presumes that animals move towards a
+preconceived endpoint and choose a cost-minimizing route accordingly
+\citep{Abrahms.2017}. While these assumptions may be fulfilled by migrating
+animals, they are unlikely to hold for dispersers, as dispersers typically move
+across unfamiliar territory and are therefore less aware of associated movement
+costs \citep{Koen.2014, Cozzi.2020}. CT relaxes these assumptions, but posits
+that animals move according to a random walk, entailing that autocorrelation
+between subsequent movements cannot be rendered \citep{Diniz.2019}. In reality,
+however, autocorrelation in animal movements is regularly observed
+\citep{Bovet.1991, Schultz.2001}, especially during dispersal \citep{Cozzi.2020,
+Hofmann.2021}. Some of these deficiencies can be alleviated using less stringent
+graph-based methods, such as least-cost \textit{corridors} \citep{Pinto.2009},
+\textit{thresholded} least-cost paths \citep{Landguth.2012}, or
+\textit{randomized} least-cost paths \citep{Panzacchi.2016}, yet a certain
+degree of arbitrariness remains. Besides this, graph-based methods are incapable
+of rendering the temporal dimension of dispersal, meaning that statements about
+the approximate duration required to move between habitats are impossible
+\citep{Martensen.2017, Diniz.2019}.
 
 % What about IBMMs?
-\subsection{What about IBMMs? (90\%)}
-To overcome the unrealistic assumptions inherent to traditional connectivity
-modeling techniques, simulations from individual-based movement models (IBMMs)
-have been proposed and applied \citep{Diniz.2020}. In these models, dispersal
-trajectories are simulated based on movement rules that determine how
-individuals move across or interact with the prevailing landscape
-\citep{Gustafson.1996, Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004,
-Revilla.2008, Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019,
-Zeller.2020, Vasudev.2021}. Instead of enforcing connections towards a known
-endpoint, IBMMs allow individuals to choose their route as they move and enable
-to render potential interactions between movement behavior and environmental
-conditions. IBMMs furthermore enable to model directional persistence and to
-study the temporal dimension of dispersal \citep{Diniz.2020}. Despite these
-advantages, however, a unified framework to simulate dispersal and to derive
-connectivity maps based on simulated dispersal events is lacking.
-
-% The Power of Step Selection Functions
-\subsection{Step Selection Functions (90\%)}
-A promising framework that has been developed more recently...
-Over the past decade, the increased availability of large amounts of data and an
-increased computational power have led to the development of numerous techniques
-to study animal movement and resource selection during dispersal
-\citep{Boyce.2002, Fortin.2005, Cushman.2010, Zeller.2012}. Especially
-\textit{resource selection functions} \citep{Boyce.2002}, \textit{step selection
-functions} \citep{Fortin.2005}, and \textit{path selection functions}
-\citep{Cushman.2010} have pushed the field of movement ecology forward. These
-methods allow estimating habitat preferences of the focal species by comparing
-covariates at locations visited by the animal to the same covariates at
-locations available to, but not visited by the animal \citep{Boyce.2002,
-Fortin.2005, Cushman.2010, Thurfjell.2014}. While regular SSFs were intended to
-learn about relative habitat preferences of the focal species
-\citep{Fortin.2005, Thurfjell.2014, Avgar.2017}, the method has recently been
-generalized to \textit{integrated} SSFs and now enables to jointly study habitat
-and movement preferences, as well as potential interactions between them
-\citep{Avgar.2016, Signer.2017, Fieberg.2020}. ISSFs therefore provide a
-relatively simple means to model complex movement behavior, where movement is
-viewed as the result of two intertwined behavioral kernels (e.g.
-\citealp{Prokopenko.2017, Munden.2020}). Importantly, a parametrized ISSF model
-represents a fully mechanistic movement model based on which individual movement
-trajectories can be simulated \citep{Avgar.2016, Signer.2017}. In fact,
-\cite{Signer.2017} used ISSF to simulate steady state utilization distributions
-of resident animals that were moving around a point of attraction. However, the
-degree to which such simulations are helpful in detecting movement corridors and
-modeling landscape connectivity remains to be investigated.
-
-% What we do
-\subsection{Proposed Framework (90\%)}
-Here, we therefore propose to combine several already-existing methods into a
-simple three-step workflow to simulate dispersal and investigate landscape
-connectivity. In the first step, integrated step selection functions are used to
-parametrize a mechanistic movement model rendering dispersal behavior. In the
-second step, the parametrized model is applied as an individual-based movement
-model to simulate dispersal trajectories. Finally, in the third step, simulated
-trajectories are consolidated into three complementary connectivity maps, each
-focusing on a different aspect of landscape connectivity. The set of maps
-includes a heatmap, revealing frequently traversed areas, a betweenness-map,
-delineating critical dispersal corridors, and a map of inter-patch connectivity,
-depicting the presence or absence of functional links between distinct habitat
-patches.
-
-% Study Species & Study Area
-\subsection{Study Species \& Study Area (90\%)}
-We exemplify application of the proposed workflow by investigating landscape
-connectivity for the endangered African wild dog (\textit{Lycon pictus}). While
-once present across entire sub-Saharan Africa, wild dogs have disappeared from a
-vast majority of their historic range, mainly due to persecution by humans,
-habitat fragmentation and destruction, and deadly diseases
-\citep{Woodroffe.2012}. Within the few remaining subpopulations, wild dogs form
-cohesive packs comprising 8 to 12 adults and their offspring \cite{McNutt.1995}.
-After reaching sexual maturity, male and female offspring form same-sex
-coalitions and disperse from their natal pack in search for potential mating
-partners and a suitable territory to settle \citep{McNutt.1996, Behr.2020}. New
-packs are formed when dispersing coalitions join unrelated opposite-sex
-dispersing coalitions \citep{McNutt.1996}. During dispersal, wild dogs can cover
-several hundred kilometers across a variety of landscapes
-\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. One of the
-main strongholds for this species lies near the Moremi Game Reserve in northern
-Botswana, which is part of the world's largest transboundary protected area,
-namely the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA). This
-area was originally devoted towards facilitating the migration of elephants, yet
-it can be expected to yield benefits to a great multitude of other species
-\citep{Elliot.2014, Brennan.2020, Hofmann.2021}.
-
-% Previous Paper
-\subsection{Previous Paper (90\%)}
-We previously assessed landscape connectivity for dispersing African wild dogs
-within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}. More
-specifically, we fitted a habitat selection model and predicted a permeability
-surface based on which we computed least-cost paths and corridors between
-protected areas. We now expand on this knowledge and use ISSFs to develop a more
-detailed, mechanistic movement model of dispersing wild dogs
-(\Cref{GraphicalAbstract}). We then employ the model to simulate 80'000
-dispersers moving 2'000 steps across the KAZA-TFCA. Finally, we generate a
-heatmap, betweenness map and a map of interpatch connectivity. We thereby
-exemplify how ISSFs can be utilized for dispersal simulations and we discuss
-several benefits of this approach over traditional connectivity modeling
-techniques such as least-cost analysis and circuit theory. Most importantly,
-simulations based on ISSFs provide a more generic view on how connectivity
-emerges and to which degree connectivity depends on the dispersal duration. In
-addition, by generating proper dispersal trajectories, network theory can be
-applied to calculate network metrics that are pertinent to connectivity
-analysis.
+\subsection{What about IBMMs?}
+To overcome the unrealistic assumptions of LCPA and CT, dispersal simulations
+from individual-based movement models (IBMMs) have been proposed and applied
+\citep{Diniz.2019}. In these models, dispersal trajectories are simulated based
+on movement rules that govern how individuals move across and interact with the
+prevailing landscape (e.g. \citealp{Gustafson.1996, Gardner.2004, Graf.2007,
+KramerSchadt.2004, Revilla.2004, Revilla.2008, Kanagaraj.2013, Clark.2015,
+Allen.2016, Hauenstein.2019, Zeller.2020, Vasudev.2021}). Once the trajectories
+are simulated, they are further processed and converted into measures of
+connectivity \citep{Kanagaraj.2013, Zeller.2020}. Because IBMMs do not enforce
+connections towards preconceived endpoints, simulated individuals are allowed to
+dynamically adjust their behavior as they disperse. This enables to render
+potential interactions between movement behavior and habitat conditions,
+shifting the focus from a structural to a more functional view on connectivity
+\citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}. Furthermore, IBMMs
+render movement sequentially, meaning that the temporal dimension of dispersal
+becomes explicit and that autocorrelation between movements can be incorporated
+\citep{Diniz.2019}. In spite of these advantages, a unified framework to
+simulate dispersal and assess connectivity using IBMMs is lacking.
+
+% Proposed Solution: Three-Step Workflow
+\subsection{Proposed Solution: Three-Step Workflow}
+In this paper, we draw on recent innovations in the field of movement ecology
+and network theory to propose a simple three-step workflow for the simulation of
+dispersal and assessment of connectivity (\Cref{GraphicalAbstract}). In the
+first step, GPS data of the focal species is combined with relevant habitat
+covariates and analyzed in an integrated step selection framework to fit a
+movement model. In the second step, the parametrized model is employed as an
+IBMM to simulate dispersal trajectories across the study area. Finally, in the
+third step, simulated trajectories are converted into a set of three
+complementary connectivity maps, each focusing on a different aspect of
+connectivity. The set of maps includes a heatmap, revealing frequently traversed
+areas, a betweenness-map, delineating dispersal corridors and bottlenecks, and a
+map of inter-patch connectivity, depicting the presence or absence, intensity,
+and dispersal duration of functional links between habitat patches.
+
+% Some Background on SSFs
+\subsection{Some Background on SSFs}
+To fit a movement model, we propose to employ the framework of step selection
+functions (SSFs; \citealp{Fortin.2005, Thurfjell.2014, Avgar.2016}). Originally,
+SSFs were developed to learn about habitat preferences of the focal species
+using GPS data \citep{Fortin.2005}. The method works by comparing covariates
+along realized steps (the straight-line connecting two consecutive GPS
+relocations; \citealp{Turchin.1998}) to the same covariates along randomly
+generated alternative steps \citep{Fortin.2005, Thurfjell.2014} using
+conditional logistic regression \citep{Fortin.2005, Muff.2020}. Recently,
+\cite{Avgar.2016} have introduced \textit{integrated} SSFs (ISSFs), which is a
+generalization of regular SSFs and requires that movement metrics are included
+in the corresponding regression model. This not only reduces biases in estimated
+preferences, but also permits simultaneous inference on a species' habitat
+kernel (i.e. habitat preferences), its movement kernel (i.e. movement
+preferences/capabilities), and potential interactions between the two kernels
+\citep{Avgar.2016, Fieberg.2020}. Importantly, a model that was fitted using
+ISSFs resembles a fully mechanistic IBMM, based on which movement can be
+simulated \citep{Avgar.2016, Signer.2017}. Previously, ISSFs have been applied
+to simulate utilization distributions of resident animals \citep{Avgar.2016,
+Signer.2017}, yet the method remains underutilized for the simulation of
+dispersal and detection of movement corridors.
+
+% Some Background on Connectivity "Metrics"
+\subsection{Some Background on  Connectivity ``Metrics'' }
+Irrespective of the chosen modeling framework, simulated dispersal trajectories
+need to be processed into measures of connectivity. To date, the most common
+practice to translate simulations into connectivity measures is to compute
+inter-patch connectivity between distinct habitats \citep{Diniz.2019}.
+Inter-patch connectivity is a simple ratio between the number of dispersers
+successfully moving between habitat patches and the total number of simulated
+individuals from each patch (e.g. \citealp{Kanagaraj.2013}). As an alternative,
+some researchers have tallied overlapping trajectories into what might be called
+``heatmaps'', i.e. maps that depict the frequency at which each spatial unit in
+the study area was traversed by simulated dispersers (e.g. \citealp{Peer.2008,
+Hauenstein.2019, Zeller.2020}). More recently, \cite{BastilleRousseau.2018}
+proposed a method to convert trajectories into a network, so that metrics from
+network theory can be computed. With regards to connectivity, the betweenness
+metric seems promising, as it indicates the number of shortest paths going
+through each node in the network and therefore highlights areas of exceptional
+importance for connecting different regions in the study area. Because each of
+the three metrics metrics focuses on a very different aspect of landscape
+connectivity, we propose to view them as complementary and to use them in
+concert when assessing connectivity.
 
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
+    \includegraphics[width = \textwidth]{99_GraphicalAbstract2.pdf}
     \caption{Flowchart of the simulation-based connectivity analysis as proposed
@@
     selection model, which enables the parametrization of the focal species'
-    habitat and movement kernels. The parametrized model is then treated as an
-    individual-based movement model and used to simulate dispersal trajectories.
-    Ultimately, simulated trajectories serve to produce a set of maps that are
-    pertinent to landscape connectivity. This includes a heatmap, indicating the
-    relative traversal frequency across the study area, a betweenness map,
-    highlighting movement corridors and bottlenecks, and, finally, an
-    inter-patch connectivity map, where the frequency of connections and their
-    average duration can be depicted. \textcolor{red}{Photos: Whom to cite?
-    Vectronics or Photographers?}}
+    habitat and movement kernels and results in a mechanistic movement model.
+    The parametrized model is then treated as an individual-based movement model
+    and used to simulate dispersal trajectories. Ultimately, simulated
+    trajectories serve to produce a set of maps that are pertinent to landscape
+    connectivity. This includes a heatmap, indicating the traversal frequency
+    across each spatial unit of the study area, a betweenness map, highlighting
+    movement corridors and bottlenecks, and, finally, an inter-patch
+    connectivity map, where the frequency of connections and their average
+    duration can be depicted.}
     \label{GraphicalAbstract}
@@
 
+% Case Study
+\subsection{Case Study}
+To exemplify the application of the proposed workflow in
+(\Cref{GraphicalAbstract}), we investigate landscape connectivity for the
+endangered African wild dog (\textit{Lycon pictus}) as a case study. While once
+present across entire sub-Saharan Africa, this species has disappeared from a
+vast majority of its historic range, mainly due to persecution by humans,
+habitat fragmentation and destruction, and deadly diseases
+\citep{Woodroffe.2012}. Some of the biggest remaining populations reside near
+the Moremi Game Reserve in northern Botswana \citep{McNutt.1996, Cozzi.2013,
+Behr.2020}, which is part of the Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA), the world's largest transboundary protected area. We previously
+assessed landscape connectivity for dispersing African wild dogs within this
+ecosystem using least-cost methods \citep{Hofmann.2021}. More specifically, we
+used ISSFs to fit a habitat selection model and predicted a permeability surface
+based on which we computed least-cost paths and corridors between protected
+areas. We now expand on this knowledge and use ISSFs to develop a more detailed,
+mechanistic movement model of dispersing wild dogs (\Cref{GraphicalAbstract}).
+We then employ the model to simulate 80'000 dispersers moving 2'000 steps across
+the landscape of the KAZA-TFCA. Finally, we condense the simulated trajectories
+into a heatmap, betweenness-map, and inter-patch connectivity map. We thereby
+demonstrate that simulations from ISSFs provide a viable alternative to LCPA or
+CT and deliver a more comprehensive understanding of the emergence of
+connectivity. Nevertheless, we recognize that simulations from ISSFs entail
+several non-trivial modeling-decisions, such as the number of simulated
+dispersers and the duration of simulated dispersal events.
+
 \section{Methods}
-\subsection{Study Area (90\%)}
-The study area was defined by a bounding box centered at -17\degree 13'9''S,
-23\degree 56'4''E (\Cref{StudyArea}a) stretching over 1.3 Mio.
-km\textsuperscript{2} and encompassed the entire KAZA-TFCA (\Cref{StudyArea}b).
-The KAZA-TFCA represents the world's largest transboundary conservation area and
-comprises parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia. It covers a
-total of 520'000 km\textsuperscript{2} and hosts diverse landscapes, ranging
-from savanna to grassland and from dry to moist woodland habitats. In its center
-lies the Okavango Delta, a dominant hydrogeographical feature and the world's
-largest flood-pulsing inland delta. The wet season within the KAZA-TFCA lasts
-from November to March and is out of phase with the flood in the Okavango Delta,
-which peaks between July and August \citep{McNutt.1996, Wolski.2017}. Although
-large portions within the KAZA-TFCA are designated national parks or other
-protected areas, considerable human influence remains due to roads, agricultural
-sites and settlements and villages that are distributed across the KAZA-TFCA's
-landscape.
+\subsection{Study Area}
+The study area of our wild dog case study was defined by a bounding box centered
+at -17\degree 13'9''S, 23\degree 56'4''E (\Cref{StudyArea}a) and stretched over
+1.3 Mio. km\textsuperscript{2}, encompassing the entire KAZA-TFCA
+(\Cref{StudyArea}b). The KAZA-TFCA comprises parts of Angola, Botswana, Namibia,
+Zimbabwe, and Zambia and hosts diverse landscapes, ranging from savannah to
+grassland and from dry to moist woodland habitats. In its center lies the
+Okavango Delta, a dominant hydrogeographical feature and the world's largest
+flood-pulsing inland delta. Although large portions of the KAZA-TFCA are
+designated national parks and other protected areas, considerable human
+influence remains from roads, agricultural sites and settlements.
 
@@
     study area was confined by a bounding box spanning the entire KAZA-TFCA and
-    encompassed parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
+    comprised parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
     The KAZA-TFCA represents the world's largest terrestrial conservation area
-    and covers a total of 520'000 km\textsuperscript{2}. Its purpose is to
+    and covers a total of 520'000 km\textsuperscript{2}. Its main purpose is to
     re-establish connectivity between already-existing national parks (dark
@@
 
-\subsection{GPS Relocation Data (90\%)}
+\subsection{GPS Data}
 GPS data of 16 dispersing wild dogs was obtained as described in
-\citep{Cozzi.2020, Hofmann.2021}. We collected data on 7 female and 9 male
-dispersing coalitions between 2011 and 2019 from a free-ranging population
-inhabiting the Moremi National Park in northern Botswana. In some instances,
-exact dispersal dates were known from field observations, in other cases we
-determined dispersal phases using the net-squared displacement metric. Net
-squared displacement measures the squared Euclidean distance of a GPS relocation
-to a reference point \citep{Borger.2012}, which in our case was set to the
-center of each individual's natal home range. Thus, dispersal was deemed to have
-started when an individual left its natal home range and ended once individuals
-became sedentary again. Because behavior during dispersal is more pertinent to
-landscape connectivity \citep{Elliot.2014, Abrahms.2017}, we discarded all data
-that was collected during residency. During dispersal, collars were programmed
-to record a GPS fix every four hours and regularly transmitted over the Iridium
-satellite system. As previous research found no differences in behaviors of
-females and males during dispersal \citep{Woodroffe.2019, Cozzi.2020}, we did
-not distinguish between sexes. After collection, we converted GPS coordinates (n
-= 4'169) to steps, where each step represented the straight-line distance
-traveled by an individual between two consecutive GPS relocations
-\citep{Turchin.1998}. To ensure a regular sampling interval, we removed fixes
-that were not successfully collected on the 4-hourly schedule (\( \pm \) 15
-minutes).
-
-\subsection{Covariates (90\%)}
-We represented the physical landscape across the study area using a set of
-habitat covariates that included water-cover, distance to water, woodland-cover,
-and shrub/grassland-cover. Because water cover greatly changes within and
-between years in the Okavango Delta, we applied a remote sensing algorithm and
-generated frequently updated water cover layers and corresponding distance to
-water layers (see \citealp{Wolski.2017} and Appendix A3 in
+\cite{Cozzi.2020} and \cite{Hofmann.2021}. We collected data on 7 female and 9
+male dispersers between 2011 and 2019 from a free-ranging population inhabiting
+the Moremi National Park in northern Botswana. Because behavior during dispersal
+is more pertinent to landscape connectivity than behavior during residence
+\citep{Elliot.2014, Abrahms.2017}, we discarded all data that was collected
+during residency. For some individuals, exact dispersal dates were known from
+field observations, whereas in other cases, we determined dispersal phases using
+net-squared displacement. Net-squared displacement is a metric that measures the
+squared Euclidean distance of a GPS relocation to a reference point
+\citep{Borger.2012}, which in our case was set to the center of each
+individual's natal home range. Thus, dispersal was deemed to have started when
+an individual left its natal home range and ended once individuals became
+sedentary again. During dispersal, GPS collars recorded a fix every 4 hours and
+they regularly transmitted data over the Iridium satellite system. After
+collection, we converted GPS coordinates of dispersers (n = 4'169) to steps,
+where each step represented the straight-line connecting two consecutive GPS
+relocations \citep{Turchin.1998}. To ensure a regular sampling interval, we
+removed fixes that were not successfully collected on the 4-hourly schedule (\(
+\pm \) 15 minutes) and identified bursts of regularly sampled GPS fixes.
+
+\subsection{Covariates}
+We represented the physical landscape in our study area using a set of habitat
+covariates, including water-cover, distance-to-water, woodland-cover, and
+shrub/grassland-cover. Because water-cover greatly changes within and between
+years in the Okavango Delta, we applied a remote sensing algorithm and obtained
+weekly updated water-cover layers and corresponding distance-to-water layers
+from MODIS satellite imagery (see \citealp{Wolski.2017} and Appendix A3 in
 \citealp{Hofmann.2021}). Resulting layers thus temporally aligned with each
 dispersal event, ensuring minimal temporal lag between observed movements and
-corresponding water layer. We furthermore computed a proxy for human influence,
+corresponding water layers. We furthermore included a proxy for human influence,
 rendering anthropogenic pressures stemming from human-density, agricultural
@@
 resolution of 250 m by 250 m. A detailed description of the preparation of each
-habitat covariate is given in \cite{Hofmann.2021}.
+habitat covariate is provided in \cite{Hofmann.2021}.
 
 Besides habitat covariates, we computed movement metrics that we used as
-movement covariates in our models. Movement metrics were calculated for each
-step and included the step length (\textsf{sl}), its natural logarithm
-(\textsf{log(sl)}), and the cosine of the relative turning angle
-(\textsf{cos(ta)}) (for details see \citep{Avgar.2016, Fieberg.2020}). Because
-wild dogs follow a diurnal activity pattern \citep{Castello.2018}, we also coded
-a binary variable (\textsf{LowActivity}) indicating whether a step was realized
-during periods of low wild dog activity (17:00 to 09:00 local time) or high wild
-dog activity (09:00 to 17:00 local time). Handling and manipulation of all data,
-as well as all models and simulations were implemented with the statistical
-software {\tt R}, version 3.6.6 \citep{R.2019}. Several helper functions were
-written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
+movement covariates in our ISSF models \citep{Avgar.2016, Fieberg.2020}.
+Movement metrics were calculated for each step and included the step length
+(\textsf{sl}), its natural logarithm (\textsf{log(sl)}), and the cosine of the
+relative turning angle (\textsf{cos(ta)}). Because wild dogs follow a distinct
+activity pattern \citep{Castello.2018}, we also coded a binary variable
+(\textsf{LowActivity}) indicating whether a step was realized during periods of
+low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
+(17:00 to 09:00 local time). Handling and manipulation of all data, as well as
+all models and simulations were implemented with the statistical software {\tt
+R}, version 3.6.6 \citep{R.2020}. Some helper functions were written in {\tt
+C++} and imported into {\tt R} using the {\tt Rcpp} package
 \citep{Eddelbuettel.2011, Eddelbuettel.2013}.
 
-\subsection{Movement Model (80\%)}
+\subsection{Movement Model}
 We used ISSFs to parametrize a mechanistic movement model for dispersing wild
 dogs \citep{Avgar.2016}. To conduct ISSF analysis, we paired each realized step
-with 24 random steps, such that a realized step plus its 24 random steps formed
-a stratum that received a unique identifier. As suggested by \cite{Avgar.2016},
-we generated random steps by sampling random turning angles from a uniform
-distribution (\(-\pi, +\pi\)) and step lengths from a gamma distribution that
-was fitted to realized steps (scale = 6'308, shape = 0.37). Along each step, we
-extracted and averaged spatial covariates using the {\tt velox} package
-\citep{Hunziker.2021}. We also calculated the movement metrics \textsf{sl},
-\textsf{log(sl)}, and \textsf{cos(ta)} for each observed and random step. To
-facilitate model convergence, we standardized all continuous covariates to a
-mean of zero and a standard deviation of one. Since correlation among covariates
-was low (\(|r| > 0.6\); \citealp{Latham.2011}), we retained all of them for
-modeling.
+with 24 random steps, so that a realized step plus its 24 random steps formed a
+stratum that received a unique identifier. As suggested by \cite{Avgar.2016}, we
+generated random steps by sampling random turning angles from a von Mises
+distribution with concentration parameter \(\kappa = 0\) (i.e. a uniform
+distribution on (\(-\pi, +\pi\))) and step lengths from a gamma distribution
+that was fitted to realized steps (scale \(\theta\) = 6'308 and shape \(k\) =
+0.37)). Along each step, we extracted and averaged values from the habitat
+covariate layers using the {\tt velox} package \citep{Hunziker.2021}. We also
+calculated the movement metrics \textsf{sl}, \textsf{log(sl)}, and
+\textsf{cos(ta)} for each realized and random step. To facilitate model
+convergence, we standardized all continuous covariates to a mean of zero and a
+standard deviation of one. Since correlation among covariates was low (\(|r| <
+0.6\); \citealp{Latham.2011}), we retained all of them for modeling.
 
@@
 step \citep{Fortin.2005}. The selection score \(w(x)\) of each step thus
-depended on its associated covariates (\(x_1, x_2, ..., x_n\)) and on the
+depended on the step's associated covariates (\(x_1, x_2, ..., x_n\)) and on the
 animal's preferences (i.e. relative selection strengths; \citealp{Avgar.2017})
@@
 
-We ran conditional logistic regression analysis in the r-package {\tt glmmTMB}
-to estimate preferences of interest. To handle multiple individuals, we applied
-the mixed effects technique developed by \citep{Muff.2020}, which allows to
-effectively model random slopes. Thus, we treated animal IDs as random effect
-and modeled random slopes for each covariate.
-
-The structure of the movement model was based on the habitat selection model for
-dispersing wild dogs presented in \cite{Hofmann.2021}. In the original model
-(hereafter referred to as base model), no interactions between habitat
-covariates (\textsf{Water, DistanceToWater\textsuperscript{0.5}, Woodland,
-Shrubs/Grassland, Human Influence}) and movement covariates (\textsf{sl,
-log(sl), cos(ta)}) were considered. Hence, we slightly expanded this base model
-and proposed interactions between all movement and habitat covariates. More
+\noindent To estimate preferences (i.e. the \(\beta\)'s), we used mixed effects
+conditional logistic regression analysis \citep{Muff.2020} that we implemented
+using the r-package {\tt glmmTMB} \citep{Brooks.2017}. We treated animal IDs as
+random effects and modeled random slopes for each covariate.
+
+The structure of the movement model was based on a habitat selection model that
+we previously developed \citep{Hofmann.2021}. In the original model (hereafter
+referred to as \textit{base model}), no interactions between habitat covariates
+(i.e. \textsf{Water, DistanceToWater\textsuperscript{0.5}, Woodland,
+Shrubs/Grassland, Human Influence}) and movement covariates (i.e. \textsf{sl,
+log(sl), cos(ta)}) were considered. Hence, we slightly expanded the base model
+and proposed interactions between all movement covariates and habitat
+covariates, assuming that movement behavior depends on habitat conditions. More
 specifically, we started with the base model and incrementally increased model
@@
 covariates and movement covariates. For instance, for the covariate
-\textsf{Water}, we proposed the interactions \textsf{Water:log(sl)},
-\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides those combinations,
-we also proposed the interactions \textsf{sl:cos(ta)} and
-\textsf{log(sl):cos(ta)} to account for a correlation between turning angles and
-step lengths, as well as the interactions \textsf{sl:LowActivity} and
-\textsf{log(sl):LowActivity} to account for the fact that step lengths may
-differ due to wild dogs' diurnal activity pattern. To compare competing models
-and assess the most parsimonious movement model, we ran stepwise forward model
-selection based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
+\textsf{Water}, we proposed the interactions \textsf{Water:sl},
+\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides these interactions,
+we also allowed for correlations between turning angles and step lengths by
+proposing the interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)}.
+Furthermore, we formed the interactions \textsf{sl:LowActivity} and
+\textsf{log(sl):LowActivity} to render that step lengths are likely to be
+shorter during periods of inactivity. To assess the most parsimonious movement
+model among model candidates, we ran stepwise forward model selection based on
+Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
 
 We validated the predictive power of the most parsimonious movement model using
-k-fold cross-validation for case-control studies as suggested by
-\cite{Fortin.2009}. For this, we randomly assigned 80\% of the strata to a
-training set and the remaining 20\% to a testing set. Using the training data we
-parametrized a movement model based on which we predicted selection scores
-\(w(x)\) for all steps in the test data. Within each stratum, we then assigned
-ranks 1-25 to each step based on predicted selection scores, where rank 1 was
-given to the step with the highest score \(w(x)\). Across all strata we
-determined the realized step's rank and we calculated rank frequencies of
-realized steps across all strata. Finally, we computed Spearman's rank
-correlation between ranks and associated frequencies \(r_{s, realized}\). We
-replicated the entire procedure 100 times and computed the mean correlation
-coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\% confidence interval
-across all replicates. For comparison, we repeated the same procedure 100 times
-assuming random preferences, which we implemented by discarding the realized
-step from all strata and identifying the rank of a random step in each stratum.
-Again, we calculated Spearman's rank correlation coefficient (\(r_{s,
-random}\)), its mean across repetitions (\(\bar{r}_{s, random}\)), and its 95\%
-confidence interval. This validation ultimately proofs a significant prediction
-in case the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-random}\) do not overlap.
-
-\subsection{Dispersal Simulation (80\%)}
+k-fold cross-validation for case-control studies \cite{Fortin.2009}.
+Specifically, we randomly assigned 80\% of the strata to a training set and the
+remaining 20\% to a testing set. With the training data we parametrized a
+movement model and predicted selection scores \(w(x)\) for all steps in the test
+data. Within each stratum, we then assigned ranks 1-25 to each step based on
+predicted selection scores, so that rank 1 was given to the step with the
+highest score \(w(x)\). Within each strata, we determined the realized step's
+rank and calculated rank frequencies of realized steps across all strata.
+Finally, we computed Spearman's rank correlation between ranks and associated
+frequencies \(r_{s, realized}\). We replicated this procedure 100 times and
+computed the mean correlation coefficient (\(\bar{r}_{s, realized}\)), as well
+as its 95\% confidence interval across all replicates. For comparison, we
+repeated the same procedure 100 times assuming random preferences. Random
+preferences were implemented by discarding the realized step from all strata and
+identifying the rank of a random step in each stratum. Again, we calculated
+Spearman's rank correlation coefficient (\(r_{s, random}\)), its mean across
+repetitions (\(\bar{r}_{s, random}\)), and its 95\% confidence interval.
+Ultimately, this validation proves a significant prediction in case the
+confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do
+not overlap \citep{Fortin.2009}.
+
+\subsection{Dispersal Simulation}
 We used the most parsimonious movement model to simulate 80'000 virtual
-dispersers moving across the KAZA-TFCA. The simulation resembled an inverted
-ISSF and was set up as follows. (1) We defined a random source point and assumed
-a random initial orientation of the animal. (2) Departing from the source point,
-we generated 25 random steps by sampling turning angles from a uniform
-distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
-distribution (similar to the input data, each random step represented the
-straight line movement within 4 hours). To prevent unreasonably large steps, we
-capped sampled step lengths to a maximum of 35 km, which corresponded to the
-farthest distance ever traveled within 4 hours by one of our dispersers. (3)
-Along each random step, we extracted and averaged habitat covariates and we
-calculated movement covariates. To ensure compatible scales, we standardized
-extracted values using the same parameters applied to our input data. (4) We
-applied the parametrized movement model to predict the selection score \(w(x)\)
-for each step and we translated predicted scores into probabilities using
-\Cref{EQ2}. (5) We sampled one of the random steps based on predicted
-probabilities and determined the animal's new position. We repeated steps (2) to
-(5) until 2'000 steps were realized, implying a total 160 Mio. simulated steps.
-
-To minimize the influence of edge effects and to deal with random steps leaving
-the study area, we followed \citep{Koen.2010} and artificially expanded all
-covariate layers by adding a 100 km wide buffer zone. Within the buffer zone, we
-randomized covariate values by resampling values from the original covariate
-layers. Through this buffer zone, simulated dispersers were able to leave and
-re-enter the main study area. In cases where proposed random steps transgressed
-the border of this buffer zone, we resampled transgressing steps until they
-fully lied within the buffer, thereby forcing simulated individuals to ``bounce
-off'' such invisible borders.
-
-\subsection{Source Points (90\%)}
+dispersers moving across the KAZA-TFCA. The simulation resembled an
+``inverted''  ISSF and was set up as follows. (1) We defined a random source
+point and assumed a random initial orientation of the animal. (2) Departing from
+the source point, we generated 25 random steps by sampling turning angles from a
+uniform distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
+distribution. Similar to the input data, each random step represented the
+straight line movement possible within 4 hours. To prevent unreasonably large
+steps, we restricted sampled step lengths to a maximum of 35 km, which
+corresponded to the farthest distance ever traveled within 4 hours by one of our
+dispersers. (3) Along each random step, we extracted and averaged values from
+the habitat covariate layers and we calculated movement covariates. To ensure
+compatible scales with the fitted movement model, we standardized extracted
+values using the same scaling parameters that we applied to standardize our
+input data. (4) We applied the parametrized movement model to predict the
+selection score \(w(x)\) for each step and we translated predicted scores into
+probabilities using \ref{EQ2}. (5) We sampled one of the random steps based on
+predicted probabilities and determined the animal's new position. We then
+repeated steps (2) to (5) until 2'000 steps were realized.
+
+To mitigate edge effects and to deal with random steps leaving the study area,
+we followed \cite{Koen.2010} and artificially expanded all covariate layers by
+a 100 km wide buffer zone. Within the buffer zone, we randomized covariate
+values by resampling values from the original covariate layers. Through this
+buffer zone, simulated dispersers were able to leave and re-enter the main study
+area. In cases where proposed random steps transgressed the outer border of this
+buffer zone, we resampled transgressing steps until they fully lied within the
+buffer zone, forcing individuals to be repelled from these map borders.
+
+\subsection{Source Points}
 We released 80'000 virtual dispersers from 80'000 unique source points
@@
 large enough to sustain viable wild dog populations (\(>\) 700
-km\textsuperscript{2}; \citealp{Pomilia.2015}) and another 30'000 were released
-at random locations inside the 100 km wide buffer to render potential immigrants
-into the study system. By distributing source points randomly, the number of
-source points per km\textsuperscript{2} was approximately equal within protected
-areas and the buffer zone.
-
-% \begin{figure}[htbp]
-%   \begin{center}
-%     \includegraphics[width = \textwidth]{99_SourceAreas.png}
-%     \caption{(a) Different source areas from which we released virtual
-%     dispersers. We only considered contiguous protected areas (national parks,
-%     game reserves, and forest reserves) that were larger than 700
-%     km\textsuperscript{2} (green). This area corresponds to the average home
-%     range requirement for viable wild dog populations \citep{Pomilia.2015}. To
-%     render potential immigrants into the study system, we also initiated
-%     dispersers within a buffer zone (blue) surrounding the main study area. (b)
-%     Source points from which dipsersers were released. 50'000 dispersers were
-%     released from the main study area (green dots) and another 30'000 dispersers
-%     within the virtual buffer (blue dots).}
-%     \label{SourcePoints}
-%   \end{center}
-% \end{figure}
-
-\subsection{Convergence (80\%)}
-To verify that the number of simulated individuals sufficed to ensure reliable
-estimates of connectivity, we evaluated how the relative traversal frequency
-across the landscape changed as we increased the number of simulated
-individuals. For this, we distributed 1'000 rectangular ``checkpoints'', each
-with an extent of 5 km x 5 km at random locations inside the main study area. We
-then determined the relative traversal frequency by simulated trajectories
-through each checkpoint as we increased the number of simulated individuals (1
-to 50'000). To assess variability in the relative traversal frequency, we
-repeatedly sampled trajectories 100 times and computed the mean traversal
-frequency across replicates, as well as the 95\% prediction-interval. We deemed
-that a checkpoint converged as soon as the width of the prediction interval for
-the traversal frequency across replicates dropped below a value of 0.01.
-
-\subsection{Heatmap (100\%)}
+km\textsuperscript{2}; \citealp{Pomilia.2015}), whereas another 30'000
+dispersers were released at random locations inside the 100 km wide buffer to
+render potential immigrants into the study system. By distributing source points
+randomly, the number of source points per km\textsuperscript{2} was
+approximately equal.
+
+\subsection{Convergence}
+\textcolor{red}{Maybe this paragraph should go into another paper?} To verify
+that the number of simulated individuals sufficed to ensure reliable estimates
+of connectivity, we evaluated how the relative traversal frequency across the
+landscape changed as we increased the number of simulated individuals. For this,
+we distributed 1'000 rectangular ``checkpoints'', each with an extent of 5 km x
+5 km at random locations inside the main study area (excluding the buffer). At
+each checkpoint, we then determined the relative traversal frequency by
+simulated dispersers as we increased the number of simulated individuals from 1
+to 50'000. To assess variability in the relative traversal frequency, we
+repeatedly subsampled 100 times from all 50'000 and computed the mean traversal
+frequency across replicates, as well as its 95\% confidence-interval. We deemed
+that a checkpoint converged as soon the width of the confidence-interval dropped
+below a value of 0.01.
+
+\subsection{Heatmap}
 To identify dispersal hotspots across our study area, we created a heatmap
-indicating the absolute frequency at which each raster-cell in the study area
-was visited by virtual dispersers \citep{Hauenstein.2019, Peer.2008}. For this,
-we rasterized all simulated trajectories and tallied them into a single map. If
-the same trajectory crossed a raster-cell twice, we only counted it once,
-thereby mitigating potential biases caused by individuals that moved in circles
-because they were trapped by dispersal barriers. To achieve high performance
-rasterization, we used the R-package {\tt terra} \citep{Hijmans.2020}.
-
-\subsection{Betweenness (80\%)}
-To pinpoint areas of exceptional relevance for connecting remote regions inside
-our study area, we converted simulated trajectories into a network and
-calculated betweenness scores \citep{BastilleRousseau.2018}. For this, we
-overlaid the study area (including the buffer) with a regular raster resolved at
-5 x 5 km. The centerpoint of each raster-cell served as node in the final
-network and we used the simulated trajectories to determine all transitions
-occurring from one node to another, as well as the frequency at which those
-transitions occurred. This resulted in an edge-list that we translated into a
-weighted network using the r-package {\tt igraph} \citep{Gabor.2006}. Because
-{\tt igraph} handles edge weights (\(\omega\)) as costs, we inverted the
-traversal frequency in each cell by applying \(\omega = \frac{\sum_i^n{Traversal
-Frequency_i}/n}{Traversal Frequency_i}\). Consequently, edges that were
-traversed frequently were assigned low costs. Finally, we used the weighted
-network to calculate the betweenness score of each raster-cell. Betweenness
-measures how often a specific raster-cell lies on a shortest path between two
-other raster-cells and is therefore a useful metric to detect movement corridors
-\citep{BastilleRousseau.2018}.
-
-\subsection{Inter-Patch Connectivity (80\%)}
-We assessed inter-patch connectivity between national parks located in our study
-area to examine functional links between distinct patches in the KAZA-TFCA. The
-decision to focus on national parks was purely out of simplicity and does not
-imply that connections between other regions are impossible. In fact, the same
-process could easily be expanded to include other protected areas. To quantify
+indicating the absolute frequency at which each raster-cell in our study area
+was visited by virtual dispersers (e.g. \citealp{Peer.2008, Hauenstein.2019,
+Zeller.2020}). For this, we rasterized all simulated trajectories and tallied
+them into a single map. If the same trajectory crossed a raster-cell twice, we
+only counted it once, thereby mitigating biases from individuals that were
+trapped and moved in circles. To achieve high performance rasterization, we used
+the R-package {\tt terra} \citep{Hijmans.2021b}.
+
+\subsection{Betweenness}
+To pinpoint discrete movement corridors and bottlenecks, we converted simulated
+trajectories into a network and calculated betweenness scores for each
+raster-cell in the study area \citep{BastilleRousseau.2018}. Betweenness
+measures how often a specific network-node (i.e. raster-cell) lies on a shortest
+path between any other pair of nodes and is therefore pertinent to connectivity
+\citep{BastilleRousseau.2018}. To transform simulated trajectories into a
+network, we overlaid the study area (including the buffer) with a raster
+resolved at 5 km x 5 km, where the center of each raster-cell served as node in
+the final network. To identify edges (connections) between the nodes, we used
+the simulated trajectories and determined all transitions occurring from one
+node to another, as well as the frequency at which those transitions occurred.
+This resulted in an edge-list that we translated into a weighted network using
+the r-package {\tt igraph} \citep{Gabor.2006}. The weight of each edge was
+determined by the frequency of transitions, yet because {\tt igraph} handles
+edge weights (\(\omega\)) as costs, we inverted the traversal frequency in each
+cell by applying \(\omega = \frac{mean(Traversal Frequency)}{Traversal
+Frequency_i}\). Consequently, edges that were traversed frequently were assigned
+small weights and vice versa. Finally, we used the weighted network to calculate
+betweenness scores for all network nodes.
+
+\subsection{Inter-Patch Connectivity}
+To examine functional links between distinct patches in the KAZA-TFCA, we
+calculated inter-patch connectivity between national parks. The decision to
+focus on national parks was purely out of simplicity and does not imply that
+connections between other protected areas are impossible. To quantify
 inter-patch connectivity, we computed the relative frequency at which dispersers
@@
 the number of steps required until the first intersection with the respective
-national park. This allowed us to determine \textit{if} and \textit{how often}
+national park, allowing us to compute average dispersal durations from one park
+to another. In summary, we determined \textit{if} and \textit{how often}
 dispersers moved between certain national parks, as well as \textit{how long}
@@
 \section{Results}
-\subsection{Movement Model (80\%)}
-Compared to the base model reported in \citep{Hofmann.2021}, our most
-parsimonious movement model retained several additional interactions between
-habitat covariates and movement covariates (\Cref{MovementModel} and
-\Cref{MovementModelNumbers}). Although several models received an AIC weight
-above zero (Table 1 in Appendix S1), we only considered results from the most
-parsimonious model for simplicity. All models with positive AIC weight included
-similar covariates (Table S1), so this decision only marginally influenced
-subsequent analyses. Plots that aid with the interpretation of the final model
-are provided in Appendix S2.
-
-Assuming that all other covariates are held at their means, the habitat kernel
-reveals that dispersing wild dogs avoid water but prefer its proximity.
-Similarly, dispersers avoid areas that are covered by woodlands, yet prefer
-regions covered by shrublands or grasslands. Finally, dispersers avoid movement
-through landscapes that are dominated by humans. Effect sizes are strong and,
-except for effect of \textsf{distance to water}, statistically clear on the 5\%
-significance level.
+\subsection{Movement Model}
+Our most parsimonious movement model resulting from the ISSF analysis consisted
+of a movement kernel, a habitat kernel, and their interactions. Compared to the
+base model reported in \citep{Hofmann.2021}, the movement model retained several
+additional interactions between habitat covariates and movement covariates
+(\Cref{MovementModel} and \Cref{MovementModelNumbers}). Although multiple models
+received an AIC weight above zero (Table 1 in Appendix S1), we only considered
+results from the most parsimonious model for simplicity. Since all models with
+positive AIC weight contained similar covariates (Table S1), this decision only
+marginally influenced subsequent steps. To aid with the interpretation of the
+final model, we followed \cite{Fieberg.2020} and produced a collection of
+auxiliary plots that are provided in Appendix S2.
+
+When all other covariates held constant at their means, the habitat kernel
+reveals that dispersing wild dogs avoid water but prefer its proximity, that
+dispersers avoid areas that are covered by woodlands, yet prefer regions covered
+by shrublands or grasslands, and that dispersers avoid movement through
+landscapes that are dominated by humans. All of these effects are strong and
+significant, except for the effect of \textsf{distance to water}, which is
+insignificant on the 5\% significance level.
 
@@
 indicates that dispersers move with directional persistence, unlike what was
-proposed by the uniform turning angle distribution. Moreover, directionality is
-particularly pronounced when dispersers realize large steps (move quickly), as
-indicated by the positive estimates for \textsf{cos(ta):sl} and
+proposed by the uniform turning angle distribution. This preference is
+particularly pronounced when dispersers realize large steps (i.e. move quickly),
+as indicated by the positive estimates for \textsf{cos(ta):sl} and
 \textsf{cos(ta):log(sl)}. Finally, the negative estimate for the interaction
-\textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (move
-slower) outside the main activity periods (during sunrise and sunset). Aside
-from the interaction \textsf{sl:LowActivity}, which appears to strongly
-influence movmement behavior, effect sizes are moderate, but mostly significant
-on the 5\% significance level.
-
-When looking at the interactions between movement and habitat kernels, we
-observe that movement behavior is contingent on habitat conditions. For example,
-there's strong evidence that dispersers realize smaller steps in areas covered
-by water or areas covered by wooldand, yet it appears that steps are larger in
-regions dominated by shrubs/grassland, and shorter when the distance to water is
-high. Correspondingly, the model suggests that directionality is lower in areas
-dominated by humans but more pronounced when dispersers are far from water.
-However, except for the effect of \textsf{sl:Water}, effect sizes and
-statistical significance are moderate.
-
-The k-fold cross-validation procedure reveals that our model substantially
-outperforms a random guess (\Cref{MovementModel}b) and therefore correctly
-assigns a high selection score to realized steps. Confidence intervals of
+\textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (i.e. move
+slower) between 09:00 and 17:00 o'clock than during the rest of the day. Aside
+from the interaction \textsf{sl:LowActivity}, which has a very strong effect,
+effects are moderate strength yet significant on the 5\% significance level;
+only the interaction \textsf{cos(ta):sl} appears as insignificant on the 5\%
+significance level.
+
+Finally, several significant interactions between movement and habitat
+covariates suggest that movement behavior is contingent on habitat conditions.
+For example, there's strong evidence that dispersers realize shorter steps in
+areas covered by water or in forested areas. Furthermore, it appears that
+dispersers realize larger steps in areas that are dominated by shrubs/grassland,
+but shorter steps in areas that are distant to water. Finally, it seems that the
+preference for directional persistence is less pronounced when dispersers cross
+human-dominated landscapes, yet more pronounced at great distance from water.
+Nevertheless, some of these effects are only weakly significant and exhibit
+small effect sizes.
+
+The k-fold cross-validation of the movement model shows that the model
+substantially outperforms a random guess (\Cref{MovementModel}b) and correctly
+assigns high selection scores to realized steps. Confidence intervals of
 \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap and
-therefore proof a reliable prediction. Furthermore, the significant correlation
+therefore prove a reliable prediction. Furthermore, the significant correlation
 between ranks and corresponding frequencies for realized steps indicates a good
@@
 the inclusion of interactions between movement and habitat covariates slightly
-improved model performance.
+improved model performance (\(\bar{r}_{s, realized} = -0.65\).
 
@@
     \caption{(a) Most parsimonious movement model for dispersing wild dogs. The
-    model includes a habitat kernel, a movement kernel, as well as their
+    model comprises a habitat kernel, a movement kernel, as well as their
     interactions. The horizontal line segments delineate the 90\%, 95\%, and
-    99\% Confidence-Intervals for the respective \(\beta\)-coefficients.
+    99\% confidence-intervals for the respective \(\beta\)-coefficients.
     Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
@@
 
-\begin{table}
-  \begin{center}
-  \caption{Most parsimonious movement model for dispersing wild dogs. The model
-  consists of a movement kernel, a habitat kernel, and their interactions. The
-  movement kernel describes preferences with regards to movement behavior,
-  whereas the habitat kernel describes preferences with respect to habitat
-  conditions. Interactions between the two kernels indicate that movement
-  preferences are contingent on habitat conditions. Note that all covariates
-  were standardized to a mean of zero and standard deviation of 1. Plots to aid
-  with the interpretation of this model are given in Appendix S2.}
-  \label{MovementModelNumbers}
-  \resizebox{\textwidth}{!} {
-    \begin{threeparttable}
-      \begin{tabular}{llrrrc}
-        \toprule
-        Kernel & Covariate & Coefficient & SE & p-value & Sign. \\
-        \midrule
-        \multirow{5}{*}{Habitat Kernel}
-         & Water & -0.546 & 0.112 & \(<\) 0.001 & *** \\
-         & DistanceToWater \textsuperscript{0.5} & -0.390 & 0.231 & 0.092 & * \\
-         & Woodland & -0.364 & 0.086 & \(<\) 0.001 & *** \\
-         & Shrubs/Grassland & 0.288 & 0.092 & 0.002 & *** \\
-         & HumanInfluence & -0.535 & 0.229 & 0.019 & ** \\
-        \hdashline
-        \multirow{6}{*}{Movement Kernel}
-         & sl & 0.075 & 0.037 & 0.042 & ** \\
-         & cos(ta) & 0.105 & 0.031 & 0.001 & *** \\
-         & log(sl) & 0.146 & 0.051 & 0.004 & *** \\
-         & cos(ta) : sl & 0.049 & 0.026 & 0.064 & * \\
-         & cos(ta) : log(sl) & 0.076 & 0.026 & 0.003 & *** \\
-         & sl : LowActivity & -0.917 & 0.113 & \(<\) 0.001 & *** \\
-        \hdashline
-        \multirow{5}{*}{Interactions}
-         & sl : Water & -0.305 & 0.076 & \(<\) 0.001 & *** \\
-         & sl : Woodland & -0.089 & 0.039 & 0.023 & ** \\
-         & sl : Shrubs/Grassland & 0.124 & 0.058 & 0.032 & ** \\
-         & sl : DistanceToWater \textsuperscript{0.5} & -0.058 & 0.031 & 0.056 & * \\
-         & cos(ta) : HumanInfluence & -0.040 & 0.022 & 0.070 & * \\
-         & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063 & 0.026 & 0.017 & ** \\
-         \bottomrule
-      \end{tabular}
-       \begin{tablenotes}
-         \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
-         \quad *** \(p < 0.01\)}
-       \end{tablenotes}
-    \end{threeparttable}
-    }
-  \end{center}
-\end{table}
-
-\subsection{Dispersal Simulation (80\%)}
-On a machine with an octacore AMD Ryzen 7 2700X processor (8 x 3.6 GHz) and 64
-GB of RAM, a batch of 1'000 simulated dispersers moving over 2'000 steps
-required 90 minutes to compute (\(\mu = 88.90\), \(\sigma = 1.87\)).
-Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
+\subsection{Dispersal Simulation}
+On a machine with an octa-core AMD Ryzen 7 2700X processor (8 x 3.6 GHz, 16
+logical cores) and 64 GB of RAM, a batch of 1'000 simulated dispersers moving
+over 2'000 steps required 90 minutes to compute (\(\mu = 88.90\), \(\sigma =
+1.87\)). Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
 terminated after 120 hours (i.e. five days). Comparable simulations will be
@@
 computationally the most demanding task. Out of the 50'000 dispersers initiated
-inside the main source area \Cref{SourcePoints}(a), only 4.5\% eventually hit a
-map boundary, suggesting that biases due to boundary effects are moderate. In
-contrast, 78\% of the 30'000 dispersers originating from the buffer zone
-eventually hit a map boundary, yet this was to be expected since many of those
-dispersers originated from source points located close to the map boundary.
-
-\subsection{Convergence (80\%)}
-Our examination of the traversal frequency as a function of the number of
-simulated dispersers shows that the mean traversal frequency stabilizes already
-after very few simulations and changes only little when adding further
-dispersers (\Cref{Convergence} (a) and (b)). While variability keeps decreasing
-with additional dispersers, the marginal benefit of adding further dispersers
-steeply decreases with a negative-exponential trend (\Cref{Convergence} (c)).
+inside the main source area, only 4.5\% were eventually repelled by a map
+boundary, suggesting that biases due to boundary effects should be minimal.
+
+\subsection{Convergence}
+Our examination of the traversal frequencies across checkpoints shows that we
+reached convergence across all checkpoints after 10'500 simulated individuals
+(\Cref{Convergence}a and b). Even though variability kept decreasing as we
+increased the number of simulated individuals, the marginal benefit of each
+additional disperser diminished quickly (\Cref{Convergence}c).
 
@@
     \includegraphics[width=\textwidth]{99_Convergence}
-    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5)
-    distributed randomly across the study area. The relative traversal frequency
-    is plotted against the number of simulated individuals to visualize how
-    quickly the metric converges to a steady state. (a) Replicated (100 times)
-    relative traversal frequencies across three randomly chosen checkpoints as
-    well as the corresponding 95\% prediction interval (PI). (b) Averaged
-    relative traversal frequency across all checkpoints and replicates including
-    a 95\% PI. (c) Width of the PI in relation to the number of simulated
-    dispersers.}
+    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5
+    km) distributed randomly across the study area. The relative traversal
+    frequency is plotted against the number of simulated individuals to
+    visualize how quickly the metric converges to a steady state. (a) Replicated
+    (100 times) relative traversal frequencies across three randomly chosen
+    checkpoints as well as the corresponding 95\% prediction interval (PI). (b)
+    Averaged relative traversal frequency across all checkpoints and replicates
+    including a 95\% PI. (c) Width of the PI in relation to the number of
+    simulated dispersers.}
     \label{Convergence}
@@
 
-\subsection{Heatmap (80\%)}
-\Cref{Heatmap} depicts the heatmap of all 80'000 simulated trajectories
-resulting after 2'000 steps. The map shows that large portions of land beyond
-the borders of the KAZA-TFCA are only infrequently visited by dispersers (dark
-blue areas), whereas within the KAZA-TFCA several extensive regions are
-regularly traversed (bright yellow and red areas). Most notably, the region in
-northern Botswana south of the Linyanti swamp stands out as highly frequented
-dispersal hotspot. Still, the presence of several massive water bodies, such as
-the Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, poses
-considerable dispersal barriers that limit realized connectivity within the
-KAZA-TFCA. Similarly, dispersal across Zambia's and Zimbabwe's part of the
-KAZA-TFCA appears to be limited, as only few areas are successfully traversed by
-dispersers. This can largely be attributed to substantial human influences
-resulting from high human density, roads, and agricultural activities in these
-areas. Outside the KAZA-TFCA, the most heavily used regions include the areas
-inside the Central Kalahari National Park in Botswana, the area south-west of
-the Khaudum National Park in Namibia, and the area around the Liuwa Plains
-National Park in Zambia.
+\subsection{Heatmap}
+\Cref{Heatmap} depicts the heatmap from all simulated individuals after 2'000
+steps (additional maps in Appendix SX). The map shows that large portions of
+land beyond the borders of the KAZA-TFCA are only infrequently visited by
+dispersers (dark blue areas), whereas within the KAZA-TFCA several extensive
+regions are regularly traversed (bright yellow and red areas). Most notably, the
+region in northern Botswana south of the Linyanti swamp stands out as a highly
+frequented dispersal hotspot. Still, the presence of several massive water
+bodies, such as the Okavango Delta, the Makgadikgadi Pan, and the Linyanti
+swamp, restricts dispersal movements and limits realized connectivity within the
+KAZA-TFCA. Similarly, the heatmap shows that dispersal across Zambia's and
+Zimbabwe's part of the KAZA-TFCA is limited, which can largely be attributed to
+substantial human influences resulting from high human density, roads, and
+agricultural activities in these areas. Outside the KAZA-TFCA, the most heavily
+used regions include the areas inside the Central Kalahari National Park in
+Botswana, the area south-west of the Khaudum National Park in Namibia, and the
+area around the Liuwa Plains National Park in Zambia. Even though the heatmap
+facilitates the identification of areas that are intensely used by virtual
+dispersers, it is not straightforward to spot dispersal corridors.
+Consequently, we turn to the betweenness map.
 
@@
   moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
-  integrated step selection model that we fit to the movement data of dispersing
-  African wild dogs. To generate the heatmap, we rasterized and tallied all
-  simulated trajectories. Consequently, the map highlights areas that are
-  frequently traversed by virtual dispersers. Additional heatmaps showing the
-  traversal frequency for different numbers of simulated steps are provided in
-  Appendix S3.}
+  integrated step selection model that we fitted to the movement data of
+  dispersing African wild dogs. To generate the heatmap, we rasterized and
+  tallied all simulated trajectories. Consequently, the map highlights areas
+  that are frequently traversed by virtual dispersers. Additional heatmaps
+  showing the traversal frequency when individuals move fewer than 2'000 steps
+  are provided in Appendix S3. For spatial reference we plotted a few selected
+  national parks (dark gray).}
   \label{Heatmap}
@@
 
-\subsection{Betweenness (80\%)}
-Betweenness scores after 2'000 simulated steps are presented in
-\Cref{Betweenness} and reveal a set of discrete dispersal corridors. Again, the
-region in northern Botswana stands out as crucial dispersal hub that connects
-more remote regions in the study system. Towards east, the extension of this
-corridor runs through the Chobe National Park into the Hwange national park.
-From there, a further extension connects to the distant Matusadona National Park
-in Zimbabwe. Northwest of the Linyanty ecosystem, a major corridor expands into
-Angola, where it splits and finally traverses over a long stretch of unprotected
-area into the Kafue National Park in Zambia. Several additional corridors with
-slightly lower betweenness scores exist, yet most of them run within the
-boundaries of the KAZA-TFCA. In general, only few corridors directly link the
-peripheral regions of the KAZA-TFCA. For instance, there are only few corridors
-between the Matusadona National Park in Zimbabwe and the Kafue National Park in
-Zimbabwe. Similarly, there are no direct links between the Zimbabwean and
-Angolan ``spikes'' of the KAZA-TFCA.
+\subsection{Betweenness}
+Betweenness scores emerging after 2'000 simulated steps are presented in
+\Cref{Betweenness} and reveal a set of discrete dispersal corridors (additional
+maps in Appendix SX). Again, the region in northern Botswana stands out as a
+central dispersal hub that connects more remote regions in the study system.
+Towards east, the extension of this corridor runs through the Chobe National
+Park into the Hwange National Park. From there, a further extension connects to
+the distant Matusadona National Park in Zimbabwe. Northwest of the Linyanty
+ecosystem, a major corridor expands into Angola, where it splits and finally
+traverses over a long stretch of unprotected area into the Kafue National Park
+in Zambia. Several additional corridors with slightly lower betweenness scores
+exist, yet most of them run within the boundaries of the KAZA-TFCA. In general,
+only a few corridors directly link the peripheral regions of the KAZA-TFCA. For
+instance, there are few corridors between the Matusadona National Park in
+Zimbabwe and the Kafue National Park in Zimbabwe. Similarly, there are no direct
+links between the Zimbabwean and Angolan ``spikes'' of the KAZA-TFCA. In
+comparison to the heatmap, the betweenness map facilitates the identification of
+linear corridors and bottlenecks between habitat patches.
 
@@
   \caption{Map of betweenness scores, highlighting distinct dispersal corridors
-  and potential bottlenecks. A high betweenness score indicates that the
-  respective cells are exceptionally relevant in connecting different regions in
-  the study area. That is, the higher the betweenness score, the more often a
-  pixel lies on a shortest path between adjacent areas. In this sense the metric
-  can be used to pinpoint discrete movement corridors. Note that we
-  square-rooted betweenness scores to improve visibility of corrdiors with low
-  scores. Betweenness scores were determined by converting simulated dispersal
-  trajectories into a large network. Note that we square-rooted betweenness
-  scores to improve the visibility of corridors with low betweenness scores.}
+  and potential bottlenecks across the extent of the KAZA-TFCA. A high
+  betweenness score indicates that the respective area is exceptionally
+  important for connecting different regions in the study area. In this sense
+  the metric can be used to pinpoint discrete movement corridors
+  \citep{BastilleRousseau.2018}. Note that we square-rooted betweenness scores
+  to improve visibility of corridors with low scores.}
   \label{Betweenness}
@@
 
-\subsection{Inter-Patch Connectivity (80\%)}
+\subsection{Inter-Patch Connectivity}
 Results from the analysis of inter-patch connectivity are given in
-\Cref{AreasReached}. The figure shows all realized links by simulated dispersers
-between national parks and indicates the average duration a disperser had to
-move to realize those links. It is again worth pointing out that the figure is
-only intended as an example, as for clarity we only considered connectivity
+\Cref{AreasReached}, which shows all realized links by simulated dispersers
+between national parks. The map furthermore indicates the relative frequency at
+which dispersers originating from one national park successfully reached another
+national park, as well as the average duration dispersers had to move to realize
+those links. It is again worth pointing out that the figure is only intended as
+an illustration, as for clarity we only considered inter-patch connectivity
 between national parks (NPs), albeit plenty of links between other protected
-areas exist. As can be seen from the number, thickness, and color of arrows,
-inter-patch connectivity between NPs in Angola, Namibia, and Botswana is
-comparably high and dispersal events between those areas short. In contrast, we
-see that connections into the Kafue NP in Zambia require more steps and are
-fewer in general. Similarly, there is a lack of connections into Zimbabwe's
-Chizarira and Matusadona NP and the more distant Lower Zambezi and Mana Pools
-NPs. In some cases, one can also detect imbalances between ingoing and outgoing
-links, hinting at potential source-sink dynamics that occur due to asymmetries
-in landscape permeability depending on the origin. For instance, while a large
-portion of dispersers from the Chizaria NP in Zimbabwe manage to move into the
-Hwange NP, there are comparably few dispersers that succeed in the opposite
-direction.
+areas exist. Overall, inter-patch connectivity between NPs in Angola, Namibia,
+and Botswana appears to be high, with relatively short dispersal durations
+between national parks. In contrast, we observe that connections from the
+central region into the Kafue NP in Zambia require rather long dispersal events
+and are rather infrequent. Similarly, relatively few connections lead into
+Zimbabwe's Chizarira and Matusadona NP. In some cases, we also find imbalances
+between ingoing and outgoing links, hinting at potential source-sink dynamics
+that occur due to asymmetries in disperser's willingness to cross an area,
+depending on the direction in which the area is traversed. For instance, while a
+fair share of dispersers originating from the Chizaria NP in Zimbabwe manages to
+move into the Hwange NP, there are comparably few dispersers that succeed in the
+opposite direction.
 
@@
   \includegraphics[width=\textwidth]{99_AreasReached.png}
-  \caption{Network on simulated dispersal trajectories highlighting connections
-  between national parks (dark green). Yellow bubbles represent the center of
-  the different national parks and are sized in relation to the number of
-  simulated dispersers originating from each park. Black dots represent national
-  parks that were smaller than 700 km\textsuperscript{2} and therefore did not
-  serve as source areas. Arrows between national parks illustrate between which
+  \caption{Map of inter-patch connectivity, highlighting connections between
+  national parks (dark green). Yellow bubbles represent the center of the
+  different national parks and are sized in relation to the number of simulated
+  dispersers originating from each park. Black dots represent national parks
+  that were smaller than 700 km\textsuperscript{2} and therefore did not serve
+  as source areas. Arrows between national parks illustrate between which
   national parks the simulated dispersers successfully moved and the color of
@@
 % Short Summary
-\subsection{Short Summary (90\%)}
+\subsection{Short Summary}
 We used ISSFs to analyse data of dispersing wild dogs and to parametrize a fully
@@
 overcome several conceptual shortcomings inherent to more traditional
-connectivity modeling techniques, such as least-cost path analysis and circuit
-theory.
+connectivity modeling techniques, such as LCPA and CT.
 
 % Movement Model
-\subsection{Movement Model (80 \%)}
-Our movement model of dispersing wild dogs comprised a habitat kernel, a
-movement kernel, and their interactions. Thus, the model rendered habitat and
-movement preferences of dispersers, as well as how their movement preferences
-were affected by habitat conditions. Parameter estimates from the habitat kernel
-revealed that dispersers avoid water, prefer its proximity, avoid woodland,
-prefer shrubs/grassland, and avoid areas dominated by humans. These results are
-consistent with findings from previous studies on dispersing wild dogs
-\citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020}, as well
-as with a base model that we developed focussing on dispersing wild dogs'
-habitat kernel \citep{Hofmann.2021}.
-
-By expanding the base model by a proper movement kernel, we were able to model
-several additional complexities inherent to dispersal. For instance, it is well
-known that dispersers usually move with directional persistence
-\citep{Cozzi.2018, Hofmann.2021} and that step lengths are typically correlated
-with turning angles \citep{Morales.2004, Borger.2012}. That is, larger steps
-usually coincide with smaller turning angles and vice versa. While such behavior
-could be captured by jointly sampling turning angles and step lengths from
-copula probability distributions \citep{Hodel.2021}, the ISSF framework allowed
-us to model similar behavior directly using the movement model. Besides
-accounting for directional persistence and correlations between step lengths and
-turning angles, we also allowed for interactions that rendered the fact that
-wild dogs mainly move during the darker morning and evening hours, whereas they
-tend to rest during the remainder of the day.
-
-By allowing interactions between habitat covariates and movement covariates, we
-furthermore accounted for the fact that movement and habitat preferences are
-interdependent. For example, the final model retained an interaction between
-water cover and step length, showing that dispersers are more likely to realize
-short steps (i.e. move slower) in areas covered by water and large steps in
-areas located on dryland. Likewise, the parameter estimate for the interaction
-between water cover and turning angles revealed that dispersers move less
-directional across water bodies than across dryland. We believe that this is
-owed to the fact that wild dogs wade or swim when traversing waterbodies, thus
-resulting in slower, more tortuous movements. Besides this, our model also
-suggested that dipsersers preferably realize shorter steps when moving through
-woodland, but larger steps when moving across shrubs/grassland. This can likely
-be linked to wild dogs' resting behavior, as wild dogs usually use open areas to
-quickly move over long distances \citep{Abrahms.2017} but seek shade and
-protection below the woodland canopy when resting \citep{Creel.2002}.
-
-\subsection{Simulation (80\%)}
-Based on the above described movement model, we simulated 80'000 dispersers
-moving 2'000 steps across the landscapes of the KAZA-TFCA. On a modern desktop
-machine, this simulation required five days of computation. The long simulation
-time was primarily caused by the massive extent considered (ca. 1.8 Mio. km
-\textsuperscript{2} when including the buffer) and the large number of
-dispersers simulated. Most connectivity studies are limited to much smaller
+\subsection{Movement Model}
+Because our movement model of dispersing wild dogs comprised two interacting
+kernels, it effectively rendered habitat and movement preferences of dispersers,
+as well as how preferences depended on habitat conditions. Results from the
+habitat kernel were largely in concert with previous studies that investigated
+habitat selection by dispersing wild dogs \citep{DaviesMostert.2012,
+Masenga.2016, Woodroffe.2019, Oneill.2020, Hofmann.2021}. However, by also
+incorporating a movement kernel, we were able to model several additional
+complexities inherent to dispersal. For instance, we could accommodate that
+dispersers move with directional persistence \citep{Cozzi.2018, Hofmann.2021}
+and exhibit step lengths that are correlated with turning angles
+\citep{Morales.2004, Borger.2012} by including the appropriate interactions in
+the movement model. While correlations between step lengths and turning angles
+could also be rendered by sampling them from copula probability distributions
+\citep{Hodel.2021a}, the ISSF framework allowed us to directly incorporate them
+in the movement model. In addition, by forming interactions between habitat
+covariates and movement covariates, we could render potential dependencies
+between movement and habitat preferences. For example, our final model contained
+an interaction between water-cover and step length, showing that dispersers
+realize shorter steps in areas covered by water. Likewise, we found that
+dispersers move more tortuously across water bodies than over dryland, which is
+clearly to be expected, given that wild dogs have to wade or swim when
+traversing waterbodies. The ability of accompanying such effects in a single
+model is one of the great strengths of ISSFs \citep{Avgar.2016, Fieberg.2020},
+which is why we believe the method offers a suitable framework for simulating
+movement and assessing connectivity.
+
+\subsection{Simulation}
+Our simulation of 80'000 dispersers moving 2'000 steps across the landscapes of
+the KAZA-TFCA required five days of computation on a modern desktop machine. The
+long simulation time was primarily caused by the massive extent considered (ca.
+1.8 Mio. km\textsuperscript{2} when including the buffer) and the large number
+of dispersers simulated. Most connectivity studies are limited to much smaller
 extents (e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017,
@@
 traversal frequency by simulated individuals through randomly placed checkpoints
-in the study area converged already after 10'500 simulated individuals in our
-case. The required number of simulated individuals will, however, vary depending
-on the structure of the landscape and the dispersal ability of the focal
-species.
-
-\subsection{Maps (70\%)}
-The heatmap resulting from our dispersal simulation highlighted that a large
-portion of simulated dispersers traversed the Moremi NP and the Chobe NP in
-northern Botswana. We already recognized the same area as dispersal hotspot
-using least-cost path and least-cost corridor analysis \citep{Hofmann.2021}, yet
-some researchers questioned whether this was just the logical consequence of the
-region being in the center of the study area and least-costly routes being
-enforced between pre-defined start and endpoints. Using least-cost approaches,
-this claim is difficult to disprove, as all identified routes have to completely
-run within the study area and will always be enforced between a start and
-endpoint. With our simulation, on the other hand, dispersers were able to leave
-the study area and were not enforced to move towards a known endpoint. Despite
-this, a majority of simulated individuals traversed the central region in
-northern Botsana, so we conclude that this dispersal hotspot is not caused by
-geometric properties but results from landscape characteristics and the location
-of source areas.
-
-Overall, the heatmap gives a good overview of the intensity of use in different
-areas, yet it is not well suited for pinpointing discrete movement corridors,
-which is why we also computed a betweenness-map. In contrast to the heatmap, the
-betweenness-map puts stronger emphasis on areas that are used as stepping stones
-into other regions of the study area and thereby highlights discrete dispersal
-corridors or bottlenecks \citep{BastilleRousseau.2018}. Interestingly, the
-central region in northern Botswana again stands out, implying that the region
-is not only frequently visited, but also promotes the relocation of individuals
-into more remote regions of the KAZA-TFCA. While this is an example of an area
-were both the traversal frequency \textit{and} the betweeness score is high,
-there are other instances where only one of the metrics is pronounced. For
-example, while the area between the Lengue-Luiana NP in Angola and the Kafue NP
-in Zambia receives a high betweenness-score, we find that the same area is only
-rarely traversed by dispersers according to the heatmap. Consequently, despite
-the region's importance for linking Angola's NPs to Zambia's NP, only few
-simulated dispersers actually manage to successfully traversed it. Conversely,
-while the area inside the Central Kalahri NP is traversed by many dispersers,
-the betweenness map indicates that the same region does not serve as major
-stepping stone into other regions of the study area.
-
-To complete the picture, we also computed inter-patch connectivity between NPs,
-highlighting functional links and expected dispersal durations between each
-national park in the study area. The map showed that movements from Angola into
-Zambia's Kafue NP are not only rare, but they also require many steps until they
-are realized. Conversely, we find that dispersal between the Moremi NP and Chobe
-NP are relatively frequent and require fewer steps, which can be expected given
-that the areas are located close to each other.
-
-Together, these examples nicely illustrate how powerful a combination of
-different connectivity metrics can be in deepening our understanding of
-landscape connectivity. Each map that we produced from simulated trajectories
-accentuated a different aspect of connectivity, together providing a
-comprehensive view on dispersal and landscape connectivity. The heatmap, for
-example, put emphasis on areas where movement is concentrated, regardless
-whether such areas truly contribute to geneflow or whether they represent ``dead
-ends'' that do not connect distinct patches. The betweenness map, on the other
-hand, pronounced those areas that are relevant in connecting different regions
-in the landscape and highlights potential bottlenecks. Finally, the map of
-inter-patch connectivity illustrated the frequency at which dispersal between
-distinct patches occurs, as well as the average dispersal duration required for
-individuals to move between them.
+in the study area converged already after 10'500 simulated individuals. However,
+the required number of simulated individuals to achieve reliable estimates of
+connectivity will vary depending on the structure of the landscape and the
+dispersal ability of the focal species.
+
+\subsection{Maps}
+The heatmap generated from simulated dispersal trajectories highlighted that
+numerous individuals traversed the Moremi NP and the Chobe NP in northern
+Botswana. We previously uncovered the same area as potential dispersal hotspot
+using least-cost methods \citep{Hofmann.2021}, yet it has been questioned
+whether this was a consequence of the region being located in the center of the
+study area and connections being enforced between predefined start and
+endpoints. With the current simulations, connections were no longer enforced and
+simulated individuals were able to leave the main study area. Nonetheless, a
+majority of simulated individuals traversed the area in northern Botswana. This
+suggests that the dispersal hotspot is not merely an artifact of the applied
+method but truly results from landscape characteristics. The same region is also
+pronounced on the betweenness map, implying that it facilitates the relocation
+of individuals into more remote regions of the KAZA-TFCA. While this is an
+example of an area where both the heatmap and the betweenness map attest great
+importance, there are other instances where this is not true. For example, while
+the area between the Lengue-Luiana NP in Angola and the Kafue NP in Zambia
+receives a high betweenness-score, the heatmap shows that the area is only
+rarely traversed by dispersers. Hence, despite the region's importance for
+linking Angola's NPs to Zambia's NP, only few simulated dispersers actually used
+the corridor. Conversely, we find that the Central Kalahari NP receives a low
+betweenness score, despite being highly frequented by simulated dispersers.
+Besides highlighting frequently traversed areas and dispersal corridors, we also
+also consulted inter-patch connectivity between NPs, depicting the presence or
+absence of functional links between national parks. Because dispersal movements
+were rendered sequentially, we could also derive the average dispersal duration
+that was required to realize those links. This enabled us to show that movements
+from Angola into Zambia's Kafue NP are rare and require a large number of steps
+, whereas dispersal between the Moremi NP and Chobe NP occurs frequently and
+requires relatively few steps. All in all, the rich inference that can be drawn
+from the ensemble of proposed connectivity maps.
 
 % Other studies that use similar simulations
-\subsection{Related Literature (80\%)}
-Our approach of simulating movement to assess connectivity is closely related to
-a series of previously published papers. \cite{Clark.2015}, for instance, fitted
-a regular SSF to American black bears (\textit{Ursus americanus}) and employed
-the estimated model parameters to simulate movement and identify the most likely
-movement corridors between four habitat patches. For the same species,
-\cite{Zeller.2020} used regular SSFs and forecasted seasonal habitat
-connectivity under changing land-use. As both of these studies relied on
-\textit{regular} SSFs, rather than \textit{integrated} SSFs, neither of them was
-able to account the interdependence between habitat and movement preferences. As
-such, movement behavior was assumed to be independent of habitat conditions. In
-addition, both studies lacked data collected during dispersal and instead
-employed data on residents to estimate connectivity. Although preferences during
-residence and dispersal may coincide for some species \citep{Fattebert.2015},
-there is compelling evidence suggesting that dispersers more readily cross areas
-avoided by residents \citep{Elliot.2014, Gaston.2016, Abrahms.2017,
-Keeley.2017}. The use of data collected during residence may therefore result in
-biased model estimates that distort our view on connectivity, causing a
-misallocation of scarce conservation funds \citep{Elliot.2014}. Another set of
-related studies that uses simulations from (regular and integrated) SSFs has
-been conducted by \cite{Potts.2013} and \cite{Signer.2017}, yet the primary
-focus of these papers lied on the estimation of steady-state utilization
-distributions and not the investigation of connectivity between habitat patches.
-
-\subsection{Benefits \& Modeling Decisions with ISSF Simulations (70\%)}
+\subsection{Related Literature}
+Our approach of simulating movement to assess connectivity is related to a
+series of previously published papers. \cite{Clark.2015}, for instance,
+collected GPS data on American black bears (\textit{Ursus americanus}) and used
+SSFs to fit a habitat selection model. They then used the fitted model to
+simulate movement and identify likely movement corridors between four distinct
+habitat patches. For the same species, \cite{Zeller.2020} used SSFs and
+forecasted seasonal habitat connectivity under changing land-use. Because both
+of these studies relied on \textit{regular} SSFs, rather than
+\textit{integrated} SSFs, they could not account for interdependencies among
+habitat and movement preferences \citep{Avgar.2016}. In addition, both studies
+employed data collected on resident individuals instead of dispersers, although
+evidence suggests that residents are more reluctant to cross areas that are
+readily traversed by dispersers \citep{Elliot.2014, Gaston.2016, Abrahms.2017,
+Keeley.2017}. The application of data collected on resident animals may
+therefore result in an underestimate of connectivity \citep{Elliot.2014}. Two
+further studies that used SSFs to simulate animal movement have been conducted
+by \cite{Potts.2013} and \cite{Signer.2017}, yet the primary purpose here was to
+estimate steady-state utilization distributions of resident animals and not on
+the analysis of dispersal and connectivity.
+
+\subsection{Advantages of ISSF Simulations}
 % Does not enforce route towards known endpoint
 A simulation-based approach as proposed in this article offers several
-advantages over traditional connectivity modeling techniques such as LCPA or CT.
-In contrast to LCPA, for instance, an individual-based simulation does not
-require to assume known endpoints. Instead, each endpoint emerges naturally from
-a simulated dispersal trajectory. The ability of not having to provide
-pre-determined is particularly valuable for dispersal studies, because
-dispersers often venture into unfamiliar territory and are therefore unlikely to
-know the destination of their journey \citep{Elliot.2014, Abrahms.2017,
-Cozzi.2020}. Moreover, LCPA always enforces a connection towards the predefined
-endpoints, even if associated movement costs are unreasonably high. With
-simulations from ISSFs this is no longer the case. A connectivity model that
-does not require pre-defined endpoints also ensures that movement corridors are
-not enforced between certain start- and endpoints, which permits to detect
-potential routes that do not lead into suitable habitats but into ecological
-traps \citep{Dwernychuk.1972, VanDerMeer.2014} or areas with a high
-susceptibility for human wildlife conflicts \citep{Cushman.2018}.
+advantages over LCPA and CT. In contrast to LCPA, for instance, an
+individual-based simulation does not require to assume known endpoints. Instead,
+each endpoint emerges naturally from a simulated dispersal trajectory. The
+ability of not needing to provide endpoints is particularly valuable for
+dispersal studies, because dispersers often venture into unfamiliar territory
+and are therefore unlikely to know the destination of their journey
+\citep{Elliot.2014, Abrahms.2017, Cozzi.2020}. Moreover, LCPA always enforces a
+connection towards the pre-defined endpoints, even if associated movement costs
+are unreasonably high. With simulations from ISSFs this is no longer the case. A
+connectivity model that does not require pre-defined endpoints also ensures that
+movement corridors are not enforced between certain start- and endpoints, which
+permits to detect potential routes that do not lead into suitable habitats but
+into ecological traps \citep{Dwernychuk.1972, VanDerMeer.2014} or areas with a
+high susceptibility for human wildlife conflicts \citep{Cushman.2018}.
 
@@
 In contrast to LCPA and CT, simulations from ISSFs furthermore yield the
-advantage of an explicit representation of time. This enables to answer
+advantage of an explicit representation of time, which enables to answer
 questions such as: ``\textit{How long will it take a disperser to move from A to
 B?}'' or ``\textit{Is it possible for a disperser to move from A to B within X
-days?}'' These are important questions that shift the focus from a structural to
-a more functional view on connectivity, which is usually desirable because
-functional connectivity it is directly related to geneflow \citep{Taylor.1993,
-Tischendorf.2000}. An explicit representation of time also yields opportunities
+days?}''. An explicit representation of time also yields opportunities
 for studying how seasonality affects connectivity and to investigate whether
@@
 % Requires Detailed Knowledge on Step Lengths and Turning Angles
-While an explicit representation of time offers multiple benefits, it requires
+While an explicit representation of time provides several benefits, it requires
 that step lengths and turning angles are modeled properly
@@
 
-\subsection{Further Considerations (70\%)}
-% Modeling Mortality
-Although we did not render mortality, animals regularly die during dispersal,
-mainly due to deadly encounters with predators, road kills, and persecution by
-humans \citep{Bonnet.1999, Woodroffe.2012, Behr.2021b}. Mortality during
-dispersal could therefore substantially limit functional connectivity
-\citep{Bowler.2009}, especially in areas where the likelihood of encountering
-competitors and humans is high \citep{Cozzi.2020}. If corresponding information
-is available, mortality can and should be included in ISSF simulations.
-
-% Spatially Realistic Population Models
-The ability to realistically render movement during dispersal not only serves to
-investigate landscape connectivity, but also forms the foundation for more
-realistic spatially explicit population models in which dispersal is not merely
-rendered through dispersal kernels or cellular automata movements
-\citep{Visintin.2020}, but mechanistically based on observed movement and
-habitat preferences (e.g. \citealp{Revilla.2008}, \citealp{Kleinmann.2017}).
-Such models can ultimately be employed to conduct population viability analyses
-\citep{Boyce.1992} in which species' dispersal abilities are taken into account.
-
+\subsection{Disadvantages of ISSF Simulations}
+Despite the benefits that simulations from ISSFs offer, we also want to confer
+some of the non-trivial modeling decisions involved.
+
+In particular, it is worth pointing out five modeling decisions: (1) number of
+simulated individuals, (2) location of source points, (3) dispersal duration,
+(4) boundary behavior, and (5) how to handle uncertainty and individual
+variability.
+
+(1) When simulating dispersal, the modeler needs to decide on the number of
+simulated individuals. A higher number is always desirable, as each additional
+disperser provides novel information about landscape connectivity. However, this
+comes at the cost of computational efficiency, implying that a trade-off needs
+to be managed. As noted by \cite{Signer.2017}, the trade-off can be handled by
+simulating additional individuals only until estimated metrics converge. Here,
+we employed the relative traversal frequency across checkpoints as target metric
+and found that convergence across all checkpoints was already achieved after
+10'500 simulated individuals, yet we recognize that this strongly depends on the
+focal species dispersal ability and landscape characteristics.
+
+(2) Aside from specifying the absolute number of simulated individuals, one also
+needs to define a source point within a suitable source area for each
+individual. Here, we placed source points within protected areas large enough to
+sustain viable wild dog populations. Given that wild dogs primarily survive in
+formally protected areas \citep{Woodroffe.1999, DaviesMostert.2012,
+Woodroffe.2012, VanDerMeer.2014} we considered this decision to be appropriate.
+Due to a lack of precise knowledge about wild dog abundances in the different
+protected areas, we distributed source points randomly within them. If, however,
+corresponding data is available, source points can be distributed accordingly,
+reflecting the fact that source areas do not necessarily produce an identical
+number of dispersers. Alternatively, source points can be distributed
+homogeneously, but be weighted afterwards according to estimated densities in
+the respective source area. In cases where knowledge about suitable source areas
+is lacking, these could also be delineated using habitat suitability models
+(e.g. \citealp{Squires.2013}). After all, the challenge of selecting meaningful
+source areas and source points is not unique to individual-based simulations and
+also applies to LCPA or CT.
+
+(3) When employing ISSFs to simulate dispersers, it is also required to decide
+on meaningful dispersal durations (i.e. number of simulated steps). When
+sufficient dispersal data of the focal species is available, dispersal durations
+can be sampled from observed dispersal events. Due to the low number of observed
+dispersal events and due to the great variability in wild dogs' dispersal
+distances \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020} we opted against
+this approach. Instead, we simulated individuals for 2'000 steps, which is at
+the upper end of observed dispersal durations. Once the trajectories have been
+simulated, it is straightforward to subsample them and investigate the
+sensitivity of derived results with regards to different dispersal durations.
+
+(4) Unless simulated individuals are strongly drawn towards a point of
+attraction (e.g. \cite{Signer.2017}), dispersers eventually approach a map
+boundary, so that one or several of the proposed random steps leave the study
+area. In theoretical applications, this issue can be circumvented by simulating
+movement on a torus \citep{Hodel.2021b}. For real data, however, alternative
+solutions are needed. One approach would be to terminate the simulation,
+assuming that the simulated animal left the study area forever. This can be
+problematic for individuals that are initiated in areas that are located close
+to map boundaries, especially since already a single random step leaving the
+study area breaks the simulation. Here, instead of breaking the simulation loop,
+we simply resampled transgressing random steps until they fully lied within the
+study area. This enforced simulated dispersers to be repelled by map boundaries
+and to remain within the main study area. Additionally, we artificially
+increased the study area using a buffer zone with randomized covariate values.
+This enabled virtual dispersers to leave and re-enter the main study area. The
+same method has been shown to effectively mitigate edge effects for graph-based
+connectivity models \citep{Koen.2010}.
+
+(5) We simulated dispersal using point estimates from our most parsimonious
+movement model. Depending on the amount of data and individual variability,
+these estimates can be subject to substantial uncertainty. For dispersal studies
+in particular, the low amount of data typically results in model coefficients
+with large confidence intervals \citep{Wiegand.2003, KramerSchadt.2007}. In
+these cases, point estimates may lead to biased connectivity estimates, which is
+why we urge future studies to investigate the sensitivity of ISSF simulations
+with respect to employed model parameters and to simulate dispersal based on
+parameters that encapsulate model uncertainty.
+
+\subsection{Conclusion}
+To this end, we proposed and applied a simple three-step workflow that uses
+ISSFs to parametrize an individual-based movement model from which dispersal can
+be simulated with the purpose of informing about landscape connectivity. By
+explicitly simulating dispersal trajectories, simulations enable a more
+mechanistic understanding of connectivity, and overcome several of the
+unrealistic assumptions inherent to graph-based connectivity models, such as
+leas-cost analysis or circuit theory. We exemplified the application of the
+proposed workflow using data of dispersing wild dogs to assess landscape
+connectivity for the species within the KAZA-TFCA ecosystem. With this, we hope
+to have sparked interest in the powerful framework of step selection functions
+for investigating dispersal behavior and landscape connectivity. Nevertheless,
+we propose to view simulations from ISSF-models as complementary and not as
+substitutes to traditional connectivity modeling techniques.
+
+%------------------------------------------------------------------------------
+%  Further Results to Report
+%------------------------------------------------------------------------------
+% % Some step length, turning angle comparisons
+% On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
+% 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
+% dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
+% dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
+% \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
+% 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
+% turning angles can be attributed to minor disparities between habitat conditions
+% at the area within which we collected training data and habitat conditions
+% within the entire study area.
+
+%------------------------------------------------------------------------------
+%  Further Topics to Discuss
+%------------------------------------------------------------------------------
+
+% % Simulations for an early warning system?
+% Finally, simulations from ISSFs could be utilized as forecasting tool to predict
+% the likely whereabouts of GPS collared animals into the near future. In some
+% European countries, the reintroduction of large predators, such as bears, lynx,
+% and wolves, has triggered emotional discussions and raised public concern
+% \citep{Behr.2017}, particularly in areas where locals depend on income from
+% free-roaming livestock. An early warning system based on simulations could thus
+% serve to inform about potential encounters with large carnivores, such that
+% farmers can secure their livestock accordingly. This would hopefully increase
+% public acceptance of large predators.
+
+% % Downsides of connectivity?
+% Even though connectivity is in general thought to promote population viability,
+% it is also related to various aspects that may cause ecological damage, such as
+% increased connectivity into human-dominated areas or facilitated spread of
+% deadly diseases \citep{Brearly.2013}. Nevertheless, information on the
+% relationship between connectivity and the prevalence of deadly diseases in
+% African wild dogs is currently lacking.
+
+% % Social landscape
+% We have previously attributed the weak significance of distance to water to the
+% fact that we did not control for the presence or absence of conspecifics. We
+% stick to this reasoning as our expanded model still shows a rather large
+% uncertainty around the respective beta coefficients. To better gauge the
+% importance and influence of this covariate, future studies will need to control
+% for inter- and intra-specific interactions that may explain why and when
+% dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
+% instance, found that elk movement was significantly impacted by the density of
+% wolf in the area, such that habitat preferences strictly differed depending on
+% the presence or absence of wolves. The decision to settle is likely related to
+% the presence or absence of conspecifics. Hence, the exact dispersal duration and
+% distance will not be independent of current wild dog densities. In dispersing
+% wolves, for instance, the longest dispersal distances have been observed in
+% low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
+% may thus be determined by the by the amount of isolation between subpopulations
+% and population densities \citep{DaviesMostert.2012}
+
+% % Can generate multiple metrics
+% Another benefit of simulations from ISSFs is that their output can be processed
+% using a diverse suite of approaches, each enabling to focus on a different
+% aspect of connectivity. the results from LCPA and CT, on the other hand, are
+% usually restricted to single ``conductance'' maps. By turning simulated
+% trajectories into a network, for instance, researchers can capitalize on
+% concepts from network theory and calculate network metrics that are pertinent to
+% landscape connectivity \citep{BastilleRousseau.2018}. Here, we exemplified
+% this by calculating the betweenness metric, highlighting major movement
+% corridors. Alternatively, can learn about inter-patch connectivity and potential
+% source-sink dynamics between patches \citep{Ferreras.2001, Revilla.2004,
+% Kanagaraj.2013} based on simulated dispersal trajectories. Ultimately, by
+% accounting for habitat- and movement preferences of the focal species,
+% simulations from ISSFs are capable of rendering the behavioral ecology of the
+% focal species in detail, thereby permitting a more functional view on
+% connectivity that is not solely driven by the landscape structure and distances
+% between patches \citep{Gustafson.1996, Gardner.2004, Graf.2007,
+% KramerSchadt.2004, Revilla.2004, Revilla.2008, Kanagaraj.2013}.
+
+% % We did not validate all of our predictions
+% Even though our simulations generated several interesting results, most of them
+% need to be validated using independent data. We learned that between some
+% national parks asymmetries between the number of ingoing and outgoing dispersers
+% exist, suggesting that some patches act as sinks, whereas others serve as
+% sources for dispersers.
+
+% % Modeling Mortality
+% Although we did not render mortality, animals regularly die during dispersal,
+% mainly due to deadly encounters with predators, road kills, and persecution by
+% humans \citep{Bonnet.1999, Woodroffe.2012, Behr.2021b}. Mortality during
+% dispersal could therefore substantially limit functional connectivity
+% \citep{Bowler.2009}, especially in areas where the likelihood of encountering
+% competitors and humans is high \citep{Cozzi.2020}. If corresponding information
+% is available, mortality can and should be included in ISSF simulations.
+
+% % Spatially Realistic Population Models
+% The ability to realistically render movement during dispersal not only serves to
+% investigate landscape connectivity, but also forms the foundation for more
+% realistic spatially explicit population models in which dispersal is not merely
+% rendered through dispersal kernels or cellular automata movements
+% \citep{Visintin.2020}, but mechanistically based on observed movement and
+% habitat preferences (e.g. \citealp{Revilla.2008}, \citealp{Kleinmann.2017}).
+% Such models can ultimately be employed to conduct population viability analyses
+% \citep{Boyce.1992} in which species' dispersal abilities are taken into account.
+
+% % Modeling Decisions
 % \begin{figure}
@@
 
-Despite the benefits that simulations from ISSFs offer, we also want to confer
-some of the non-trivial modeling decisions involved. In particular, we will
-discuss five modeling decisions: (1) number of simulated individuals, (2)
-location of source points, (3) dispersal duration, (4) boundary behavior, and
-(5) how to handle individual variability.
-
-(1) When simulating dispersal using ISSFs, the modeler needs to decide on the
-number of simulated individuals. This decision includes the \textit{absolute}
-number of simulated individuals across the entire study area, as well as the
-\textit{relative} number of simulated individuals per spatial entity (e.g.
-protected area, habitat patch, source point). With respect to the
-\textit{absolute} number of simulated individuals, a higher number is always
-desirable, as each additional disperser provides novel information about
-landscape connectivity. Of course this comes at the cost of computational
-efficiency, such that a trade-off needs to be managed. We propose to handle this
-trade-off by defining a target metric and only simulating additional until
-convergence in the target metric is observed. Here, we employed the
-\textit{relative traversal frequency} across checkpoints as target metric and
-found that convergence across all checkpoints was achieved already after 10'500
-simulated individuals. With regards to the \textit{relative} number of simulated
-individuals, we see several feasible approaches. If corresponding data is
-available, one could distribute dispersers in relation to known abundances,
-reflecting that population densities are not necessarily homogeneous across
-space. Alternatively, one could also distribute dispersers homogeneously, yet
-after the simulation weigh each simulated trajectory according to population
-densities at the respective source patch. Again this requires information on the
-spatial abundance of the focal speciesl. Finally, if such information is
-missing, one can distribute dispersers homogeneously across space. This is the
-approach that we employed and resulted in larger source areas generating a
-larger number of dispersers.
-
-(2) While we simulated dispersal using point estimates from our most
-parsimonious movement model but did not investigate the sensitivity of our
-results with respect to those estimates. Uncertainty is rather common in
-dispersal studies on endangered species, as data tends to be scarce, resulting
-in model estimates large confidence intervals \citep{Wiegand.2003,
-KramerSchadt.2007}. To address this, one may explore a broader range of
-preferences instead of using point estimates initiate dispersers with randomized
-preferences with variability imposed by the uncertainty in the movement model.
-We therefore urge future studies to investigate the sensitivity of ISSF
-simulations with respect to estimated preferences.
-
-(3) When employing ISSFs to simulate dispersers, one also needs to decide on
-meaningful dispersal durations (i.e. number of simulated steps). If
-corresponding data is available, dispersal durations could be sampled from
-observed events, such that each individual would only be simulated until its
-assigned dispersal duration has been achieved. Due to the low number of observed
-dispersal events and due to the great variability in wild dogs' dispersal
-distances \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020} we opted against
-this approach. Instead, we simulated individuals for 2'000 steps, which is at
-the upper end of observed dispersal durations and may have resulted in an overly
-optimistic representation of landscape connectivity. Nevertheless, it is
-relatively straight forward to shorten simulated trajectories in order to
-investigate the sensitivity of results with regards to the dispersal duration.
-Alternatively, if detailed information on settlement behavior is available, the
-dispersal simulation could include a settlement submodel, where after each
-simulated step the simulated individual decides whether or not to settle, based
-the number of realized steps, environmental conditions in the landscape,
-abundance of conspecifics or comptetitors etc.
-
-(4) Unless simulated individuals are drawn towards a point of attraction, some
-individuals will inevitably approach a map boundary such that some of the
-proposed random steps will leave the study area such that no selection score can
-be computed. One option to handle this situation would be to simply terminate
-the simulation as soon as one of the random steps leaves the study area,
-assuming that the simulated animal left the study area and will not return. This
-can be problematic when many individuals are initiated close to map boundaries,
-especially since a single random step leaving the study area forces termination
-of the simulation. As an alternative, one could resample transgressing random
-steps until all proposed random steps lie fully within the study area. This will
-force simulated dispersers to bounce off those boundaries and remain within the
-main study area. Finally, one could also extend the study area by an artificial
-buffer zone with randomized covariate values through which dispersers are
-allowed to leave and re-enter the main study area. Although dispersers might
-still approach the boundary of the buffer, it has been shown that adding an
-artifical buffer helps to mitigate edge effects \cite{Koen.2010}. A last
-solution that only applies in theoretical applications is to simulate movement
-on a torus \citep{Hodel.2021b}.
-
-(5) To initiate the simulation of a disperser, the modeler needs to define a
-source point. In some cases, exact locations of source populations are known and
-source points can be placed accordingly \citep{Kanagaraj.2013}. Moreover, if
-abundance estimates are available, these can be used to inform the relative
-number of dispersers initiated at each location. The selection of source points
-is thus directly related to the relative number of simulated individuals. Here,
-we randomly placed source points within protected areas large enough to sustain
-viable wild dog populations. Given that the species primarily survives in these
-formally protected areas \citep{Woodroffe.1999, DaviesMostert.2012,
-Woodroffe.2012, VanDerMeer.2014} we consider this decision to be appropriate. In
-other cases, comparable knowledge may be lacking and it could be more beneficial
-to delineate likely source patches based on habitat suitability models (e.g.
-\citealp{Squires.2013}). After all, the challenge of selecting meaningful source
-points is not unique individual-based simulations also applies to LCPA and
-CT. However, as highlighted by \cite{Signer.2017}, the influence of the exact
-location of source points decreases as the number of simulated steps is
-increased.
-
-\subsection{Conclusion (80\%)}
-To this end, we have used data on dispersing wild dogs to exemplify how ISSFs
-can be used to parametrize an individual-based movement model that is further
-employed to simulate dispersal and examine landscape connectivity. We also
-presented three complementary connectivity maps derived from simulated
-trajectories, each focused on a different aspect of connectivity. Furthermore,
-we discussed the potential advantages and disadvantages of the proposed
-framework compared to traditional connectivity modeling techniques such as LCPA
-and CT. With this article, we hope to have sparked interest in the uprising
-framework of step selection functions for investigating dispersal behavior and
-landscape connectivity. Nevertheless, we do not attempt to dismiss the
-application of traditional connectivity models by any means. Rather, we propose
-to use simulations from ISSF-models as a simple but powerful tool to provide a
-more comprehensive understanding of dispersal and landscape connectivity.
-
-%------------------------------------------------------------------------------
-%  Further Results to Report
-%------------------------------------------------------------------------------
-% % Some step length, turning angle comparisons
-% On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
-% 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
-% dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
-% dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
-% \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
-% 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
-% turning angles can be attributed to minor disparities between habitat conditions
-% at the area within which we collected training data and habitat conditions
-% within the entire study area.
-
-%------------------------------------------------------------------------------
-%  Further Topics to Discuss
-%------------------------------------------------------------------------------
-
-% % Simulations for an early warning system?
-% Finally, simulations from ISSFs could be utilized as forecasting tool to predict
-% the likely whereabouts of GPS collared animals into the near future. In some
-% European countries, the reintroduction of large predators, such as bears, lynx,
-% and wolves, has triggered emotional discussions and raised public concern
-% \citep{Behr.2017}, particularly in areas where locals depend on income from
-% free-roaming livestock. An early warning system based on simulations could thus
-% serve to inform about potential encounters with large carnivores, such that
-% farmers can secure their livestock accordingly. This would hopefully increase
-% public acceptance of large predators.
-
-% % Downsides of connectivity?
-% Even though connectivity is in general thought to promote population viability,
-% it is also related to various aspects that may cause ecological damage, such as
-% increased connectivity into human-dominated areas or facilitated spread of
-% deadly diseases \citep{Brearly.2013}. Nevertheless, information on the
-% relationship between connectivity and the prevalence of deadly diseases in
-% African wild dogs is currently lacking.
-
-% % Social landscape
-% We have previously attributed the weak significance of distance to water to the
-% fact that we did not control for the presence or absence of conspecifics. We
-% stick to this reasoning as our expanded model still shows a rather large
-% uncertainty around the respective beta coefficients. To better gauge the
-% importance and influence of this covariate, future studies will need to control
-% for inter- and intra-specific interactions that may explain why and when
-% dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
-% instance, found that elk movement was significantly impacted by the density of
-% wolf in the area, such that habitat preferences strictly differed depending on
-% the presence or absence of wolves. The decision to settle is likely related to
-% the presence or absence of conspecifics. Hence, the exact dispersal duration and
-% distance will not be independent of current wild dog densities. In dispersing
-% wolves, for instance, the longest dispersal distances have been observed in
-% low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
-% may thus be determined by the by the amount of isolation between subpopulations
-% and population densities \citep{DaviesMostert.2012}
-
-% % Can generate multiple metrics
-% Another benefit of simulations from ISSFs is that their output can be processed
-% using a diverse suite of approaches, each enabling to focus on a different
-% aspect of connectivity. the results from LCPA and CT, on the other hand, are
-% usually restricted to single ``conductance'' maps. By turning simulated
-% trajectories into a network, for instance, researchers can capitalize on
-% concepts from network theory and calculate network metrics that are pertinent to
-% landscape connectivity \citep{BastilleRousseau.2018}. Here, we exemplified
-% this by calculating the betweenness metric, highlighting major movement
-% corridors. Alternatively, can learn about inter-patch connectivity and potential
-% source-sink dynamics between patches \citep{Ferreras.2001, Revilla.2004,
-% Kanagaraj.2013} based on simulated dispersal trajectories. Ultimately, by
-% accounting for habitat- and movement preferences of the focal species,
-% simulations from ISSFs are capable of rendering the behavioral ecology of the
-% focal species in detail, thereby permitting a more functional view on
-% connectivity that is not solely driven by the landscape structure and distances
-% between patches \citep{Gustafson.1996, Gardner.2004, Graf.2007,
-% KramerSchadt.2004, Revilla.2004, Revilla.2008, Kanagaraj.2013}.
-
-% % We did not validate all of our predictions
-% Even though our simulations generated several interesting results, most of them
-% need to be validated using independent data. We learned that between some
-% national parks asymmetries between the number of ingoing and outgoing dispersers
-% exist, suggesting that some patches act as sinks, whereas others serve as
-% sources for dispersers.
-
 \section{Authors' Contributions}
@@
 \section{Data Availability}
-GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication. Access to all R-scripts is provided through Github.
+GPS movement data of dispersing wild dogs will be made available on dryad at the
+time of publication. Access to R-scripts that exemplify the application of the
+proposed framework to simulated data are provided through Github.
 

### 2021-09-08 Chapter_1/04_Manuscript/Manuscript_2021-08-30.tex
Various updates

--- 2021-09-08 Chapter_1/04_Manuscript/Manuscript.tex
@@
 %------------------------------------------------------------------------------
-% Importance of dispersal & connectivity and mow its measured
-Dispersal of individuals is a crucial prerequisite for long-term species
-persistence but depends on a sufficient degree of landscape connectivity. To
-date, connectivity has primarily been quantified using least-cost analysis and
-circuit theory, albeit both methods make assumptions that are rarely met in
-reality. Least-cost analysis assumes that animals move towards a preconceived
-destination and choose a cost-minimizing route accordingly, whereas circuit
-theory presumes a complete random walk without directional persistence. While
-these shortcomings can be overcome by spatio-temporally explicitly simulating
-dispersal trajectories from individual-based movement models, a unified
-approach for such simulations is lacking.
+% Importance of Dispersal & Connectivity
+Dispersal of individuals is crucial prerequisite for long-term species
+persistence. However, the ability to disperse is contingent on a sufficient
+degree of landscape connectivity, which is why the identification of movement
+corridors that contribute to connectivity has become a task of extraordinary
+importance.
+
+% Quantifying connectivity
+Over the past two decades, ecologists have investigated connectivity using
+graph-based methods, such as least-cost analysis and circuit theory.
+Unfortunately, these methods make several assumptions that are rarely met in
+reality. Least-cost analysis, for instance, assumes that animals move towards a
+preconceived endpoint and choose a cost-minimizing route accordingly. Circuit
+theory, in contrast, presumes a random walk without any directional persistence.
+To overcome these unrealistic assumptions, dispersal simulations from
+individual-based movement models have been proposed and applied, yet a unified
+framework to simulate dispersal and assess connectivity is lacking.
 
 % What we do
-Here, we present a simple three-step workflow to simulate dispersal and asses
-landscape connectivity starting from empirical GPS movement data. We exemplify
-the workflow using dispersal data collected on the endangered African wild dog
-(\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation Area,
-the world's largest transboundary conservation area. In step one, we use
-integrated step selection functions to parametrize a mechanistic movement model
-describing habitat and movement preferences of dispersers. In step two, we
-employ the parametrized model to simulate individual dispersal trajectories
-across the landscape. In step three, we interpret simulated trajectories by
-converting them into three complementary connectivity maps: a heatmap revealing
-frequently traversed areas, a betweenness-map delineating dispersal corridors,
-and a map of inter-patch connectivity depicting the presence and intensity of
-functional links between habitat patches.
-
-% Some results
-We find that simulating dispersal from indvidual-based movement models fitted
-via step selection functions are readily interpretable and offer great
-flexibility to realistically render dispersal movements. Moreover, by not
-enforcing connections towards known endpoints but instead letting dispersers
-react to the environment as they go, such simulations overcome shortcomings
-inherent least-cost path analysis, while at the same time allowing for
-autocorrelation in simulated steps, which can not be incorporated using circuit
-theory. Inspite of these advantages, simulating dispersal entails numerous
-non-trivial modeling decisions, such as the number of simulated individuals and
-the duration of simulated dispersal events.
+Here, we combine several already-existing methods to synthesize a simple
+three-step workflow for simulating dispersal and assessing connectivity. In step
+one, integrated step selection functions are used to parametrize a mechanistic
+movement model. In step two, the parametrized model is applied as an
+individual-based movement model to simulate dispersal trajectories. In step
+three, simulated trajectories are consolidated into three complementary
+connectivity maps, each focusing on a different aspect of connectivity. The set
+of maps includes a heatmap, revealing frequently traversed areas, a
+betweenness-map, delineating dispersal corridors and bottlenecks, and a map of
+inter-patch connectivity, depicting the presence or absence of functional links
+between habitat patches.
 
 % Case study
-We  show that simulations from step-selection functions offer a simple yet
-powerful alternative to traditional connectivity modeling techniques, although
-necessitating informed decisions about the number of simulated individuals and
-the duration of simulated dispersal events. Ultimately, our workflow permits a
-more mechanistic understanding of dispersal behaviour and landscape connectivity
-making it useful for a variety of applications in ecological, evolutionary, and
-conservation research.
+We exemplify the application of the proposed workflow to assess landscape
+connectivity for dispersing African wild dogs (\textit{Lycaon pictus}) in the
+Kavango-Zambezi Transfrontier Conservation Area, the world's largest
+transboundary conservation area. We thereby demonstrate that simulations from
+integrated step selection functions offer a simple but powerful alternative to
+traditional connectivity modeling techniques, although necessitating informed
+decisions about the number of simulated individuals and the duration of
+simulated dispersal events. After all, the proposed workflow enables to relax
+several unrealistic assumptions inherent to graph-based methods and permits a
+more mechanistic understanding of landscape connectivity.
 
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_GraphicalAbstract2.pdf}
+    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
     \caption{Flowchart of the simulation-based connectivity analysis as proposed
@@
 \end{figure}
+
+\begin{table}
+  \begin{center}
+  \caption{Most parsimonious movement model for dispersing wild dogs. The
+  movement kernel describes preferences of dispersers with regards to movement
+  behavior (may also be viewed as movement capability), whereas the habitat
+  kernel describes preferences with respect to habitat conditions. Interactions
+  between the two kernels indicate that the two kernels are not independent,
+  meaning that movement preferences are contingent on habitat conditions. Note
+  that all covariates were standardized to a mean of zero and standard deviation
+  of 1. Plots to aid with the interpretation of this model are given in Appendix
+  S2.}
+  \label{MovementModelNumbers}
+  \resizebox{\textwidth}{!} {
+    \begin{threeparttable}
+      \begin{tabular}{llrrrc}
+        \toprule
+        Kernel & Covariate & Coefficient & SE & p-value & Sign. \\
+        \midrule
+        \multirow{5}{*}{Habitat Kernel}
+         & Water & -0.546 & 0.112 & \(<\) 0.001 & *** \\
+         & DistanceToWater \textsuperscript{0.5} & -0.390 & 0.231 & 0.092 & * \\
+         & Woodland & -0.364 & 0.086 & \(<\) 0.001 & *** \\
+         & Shrubs/Grassland & 0.288 & 0.092 & 0.002 & *** \\
+         & HumanInfluence & -0.535 & 0.229 & 0.019 & ** \\
+        \hdashline
+        \multirow{6}{*}{Movement Kernel}
+         & sl & 0.075 & 0.037 & 0.042 & ** \\
+         & cos(ta) & 0.105 & 0.031 & 0.001 & *** \\
+         & log(sl) & 0.146 & 0.051 & 0.004 & *** \\
+         & cos(ta) : sl & 0.049 & 0.026 & 0.064 & * \\
+         & cos(ta) : log(sl) & 0.076 & 0.026 & 0.003 & *** \\
+         & sl : LowActivity & -0.917 & 0.113 & \(<\) 0.001 & *** \\
+        \hdashline
+        \multirow{5}{*}{Interactions}
+         & sl : Water & -0.305 & 0.076 & \(<\) 0.001 & *** \\
+         & sl : Woodland & -0.089 & 0.039 & 0.023 & ** \\
+         & sl : Shrubs/Grassland & 0.124 & 0.058 & 0.032 & ** \\
+         & sl : DistanceToWater \textsuperscript{0.5} & -0.058 & 0.031 & 0.056 & * \\
+         & cos(ta) : HumanInfluence & -0.040 & 0.022 & 0.070 & * \\
+         & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063 & 0.026 & 0.017 & ** \\
+         \bottomrule
+      \end{tabular}
+       \begin{tablenotes}
+         \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
+         \quad *** \(p < 0.01\)}
+       \end{tablenotes}
+    \end{threeparttable}
+    }
+  \end{center}
+\end{table}
 

### 2021-09-08 Chapter_1/04_Manuscript/Manuscript_Support.tex
Various updates

@@
 \usepackage[flushleft]{threeparttable}
+\usepackage{pdflscape}
 
@@
 
-\textbf{Keywords:} dispersal, habitat selection, integrated step selection
+\textbf{Keywords:} dispersal, simulation, movement, integrated step selection
 function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, least-cost corridors, Lycaon pictus, permeability surface,
-protected areas, wildlife management
+connectivity, Lycaon pictus
 
@@
 %------------------------------------------------------------------------------
-%	Appendix S1: All Movement Models
+%	Appendix S1: Source Points
 %------------------------------------------------------------------------------
-\section{Movement Models}
+\newpage
+\section{Source Areas \& Points}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_SourceAreas.png}
+    \caption{(a) Different source areas from which we released virtual
+    dispersers. We only considered contiguous protected areas (national parks,
+    game reserves, and forest reserves) that were larger than 700
+    km\textsuperscript{2}. This size corresponds to the average home range
+    requirement for viable wild dog populations \citep{Pomilia.2015}. To render
+    potential immigrants into the study system, we also initiated dispersers
+    within a buffer zone (blue) surrounding the main study area. (b) Source
+    points from which dipsersers were released. 50'000 dispersers were released
+    within the main study area (green dots) and another 30'000 dispersers within
+    the virtual buffer (blue dots).}
+    \label{SourcePoints}
+  \end{center}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%	Appendix S2: All Movement Models
+%------------------------------------------------------------------------------
+\newpage
+\begin{landscape}
+\thispagestyle{empty}
+\section{Model Selection Results}
 % latex table generated in R 3.6.3 by xtable 1.8-4 package
@@
   \caption{Results from the forward model selection procedure based on Akaike's
-  Information Criterion (AIC; \citealp{Burnham.2002}) for the movement model.
-  The base model upon which we based our movement model is depicted in the last
-  row. We omitted all models with an AIC weight of zero.}
+  Information Criterion (AIC; \citealp{Burnham.2002}). The model in the top row
+  was the model that we used to simulate movement of dispersers. The base model
+  upon which we based our movement model is depicted in the last row. We omitted
+  all models with an AIC weight of zero from the table.}
   \label{ModelAICs}
   \begin{center}
-    \resizebox{\textwidth}{!}{
+    \resizebox{1.75\textwidth}{!}{
       \begin{threeparttable}
@@
 \end{table}
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \includegraphics[width = \textwidth]{99_MovementModelAICs.png}
-%     \caption{Coefficients of all movement models with an AIC weight above zero.}
-%     \label{AllMovementModels}
-%   \end{center}
-% \end{figure}
+\vfill
+\raisebox{-1cm}{\makebox[\linewidth]{\thepage}}
+\end{landscape}
 
 %------------------------------------------------------------------------------
-%	Appendix S2: Evolution of Heatmaps
+%	Appendix S3: Movement Model interpretation
+%------------------------------------------------------------------------------
+\newpage
+% \newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
+\section{Movement Model Interpretation}
+To ease with the interpretation of the most parsimonious movement model, we
+followed recommendations published in \cite{Fieberg.2020} and produced a series
+of plots highlighting how the habitat and movement kernel depended on covariate
+values (\Cref{Interpretation}). To visualize the movement kernel and its
+interactions with other covariates, we used model estimates and updated our
+tentative distribution parameters for turning angles (von Mises distribution
+with concentration \(\kappa = 0\)) and step lengths (gamma distribution with
+scale \(\theta\) = 6'308 and shape \(k\) = 0.37) by applying the function
+\textsf{update\_vonmises()} from the R-package \textsf{amt} \citep{Amt.2019}.
+This allowed us to compute probability densities of turning angles and step
+lengths under varying values of the associated covariates, while holding all
+other covariates constant (\Cref{Interpretation}, a1-a8). Moreover, we
+investigated the habitat kernel by computing relative selection strengths (RSS)
+between a set of steps where values of the covariate of interest was varied to a
+reference step where the covariate value was fixed to its centered value. To
+illustrate model uncertainty, we also generated large-sample confidence
+intervals using standard errors associated with each model estimate
+(\Cref{Interpretation}, b1-b5).
+
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width = \textwidth]{99_MovementModelInterpretation.png}
+    \caption{Auxiliary plots that help with the interpretation of the most
+    parsimonious movement model from Table S1. The plots were generated
+    following recommendations reported in \cite{Fieberg.2020}. Subplots
+    \textbf{a1} to \textbf{a8} highlight dispersing wild dogs' movement kernel
+    and indicate how the kernel is influenced by interactions with other
+    covariates. Subplots \textbf{b1} to \textbf{b5} depict results from
+    dispersing wild dogs' habitat kernel and highlight differences in predicted
+    relative selection scores (RSS) when varying values of the covariate of
+    interest. For each covariate, predictions were made on the range of values
+    that was observed in the real data, assuming that all other covariates were
+    centered and that steps were realized during periods of ``high'' wild dog
+    activity. Plot a1, for example, can be interpreted as follows: the
+    probability of realizing a step with a low turning angle is much higher when
+    the corresponding step is large. Moreover, b1 can be interpreted as follows:
+    relative probability of using a step decreases as the amount of water cover
+    along the step increases.}
+    \label{Interpretation}
+  \end{center}
+\end{figure}
+% \restoregeometry
+
+%------------------------------------------------------------------------------
+%	Appendix S4: Heatmaps
 %------------------------------------------------------------------------------
@@
 \newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
-\section{Evolution of Heatmaps}
+\section{Heatmaps in Relation to Number of Simulated Steps}
 \begin{figure}[hbtp]
   \begin{center}
-    \includegraphics[width = 0.88\textwidth]{99_HeatmapsIndividual.png}
-    \caption{Heatmaps produced after 125, 500, and 2000 simulated steps,
-    respectively. The left panel (a1, a2, a3) was generated based on simulations
-    initiated within the main study area, the right panel (b1, b2, b3) was
-    generated based on simulations initiated within the buffer area. To produce
-    the heatmap presented in the main text, we summed up values from maps a3 and
-    b3.}
-    \label{Trajectories}
+    \includegraphics[width = 0.86\textwidth]{99_HeatmapsIndividual.png}
+    \caption{Heatmaps produced when considering 125, 500, and 2000 simulated
+    steps, respectively. The left panel (a1, a2, a3) was generated based on
+    simulations initiated within the main study area, the right panel (b1, b2,
+    b3) was generated based on simulations initiated within the buffer area. To
+    produce the heatmap presented in the main manuscript (Figure 5), we tallied
+    the values from maps a3 and b3.}
+    \label{Heatman}
   \end{center}
@@
 %------------------------------------------------------------------------------
-%	Appendix S3: Evolution of Betweenness Maps
+%	Appendix S5: Evolution of Betweenness Maps
 %------------------------------------------------------------------------------
@@
 \newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
-\section{Evolution of Betweenness}
+\section{Betweenness Maps in Relation to Number of Simulated Steps}
 \begin{figure}[hbtp]
   \begin{center}
-    \includegraphics[width = 0.88\textwidth]{99_BetweennessIndividual.png}
-    \caption{Maps of betweenness scores produced after 125 (a), 500 (b), and
-    2000 (c) simulated steps, respectively. A high betweenness score indicates
-    that the respective area is paramount for linking other regions in the study
-    area. It therefore highlights movement corridors that provide access to
-    various destinations.}
-    \label{Trajectories}
+    \includegraphics[width = 0.86\textwidth]{99_BetweennessIndividual.png}
+    \caption{Maps of betweenness scores produced when considering (a) 125, (b)
+    500, (c) and 2000 simulated steps, respectively. A high betweenness score
+    indicates that the respective area has a high importance for linking other
+    regions in the study area.}
+    \label{Betweenness}
   \end{center}

### 2021-09-08 Chapter_1/04_Manuscript/Manuscript_Support_2021-08-30.tex
Various updates

--- 2021-09-08 Chapter_1/04_Manuscript/Manuscript_Support.tex
@@
 %------------------------------------------------------------------------------
-%	Appendix S3: Movement Model interpretation
+%	Appendix S3: Most Parsimonious Model
+%------------------------------------------------------------------------------
+\newpage
+\begin{table}
+  \begin{center}
+  \caption{Most parsimonious movement model for dispersing wild dogs. The
+  movement kernel describes preferences of dispersers with regards to movement
+  behavior (may also be viewed as movement capability), whereas the habitat
+  kernel describes preferences with respect to habitat conditions. Interactions
+  between the two kernels indicate that the two kernels are not independent,
+  meaning that movement preferences are contingent on habitat conditions. Note
+  that all covariates were standardized to a mean of zero and standard deviation
+  of 1. Plots to aid with the interpretation of this model are given in Appendix
+  S2.}
+  \label{MovementModelNumbers}
+  \resizebox{\textwidth}{!} {
+    \begin{threeparttable}
+      \begin{tabular}{llrrrc}
+        \toprule
+        Kernel & Covariate & Coefficient & SE & p-value & Sign. \\
+        \midrule
+        \multirow{5}{*}{Habitat Kernel}
+         & Water & -0.546 & 0.112 & \(<\) 0.001 & *** \\
+         & DistanceToWater \textsuperscript{0.5} & -0.390 & 0.231 & 0.092 & * \\
+         & Woodland & -0.364 & 0.086 & \(<\) 0.001 & *** \\
+         & Shrubs/Grassland & 0.288 & 0.092 & 0.002 & *** \\
+         & HumanInfluence & -0.535 & 0.229 & 0.019 & ** \\
+        \hdashline
+        \multirow{6}{*}{Movement Kernel}
+         & sl & 0.075 & 0.037 & 0.042 & ** \\
+         & cos(ta) & 0.105 & 0.031 & 0.001 & *** \\
+         & log(sl) & 0.146 & 0.051 & 0.004 & *** \\
+         & cos(ta) : sl & 0.049 & 0.026 & 0.064 & * \\
+         & cos(ta) : log(sl) & 0.076 & 0.026 & 0.003 & *** \\
+         & sl : LowActivity & -0.917 & 0.113 & \(<\) 0.001 & *** \\
+        \hdashline
+        \multirow{5}{*}{Interactions}
+         & sl : Water & -0.305 & 0.076 & \(<\) 0.001 & *** \\
+         & sl : Woodland & -0.089 & 0.039 & 0.023 & ** \\
+         & sl : Shrubs/Grassland & 0.124 & 0.058 & 0.032 & ** \\
+         & sl : DistanceToWater \textsuperscript{0.5} & -0.058 & 0.031 & 0.056 & * \\
+         & cos(ta) : HumanInfluence & -0.040 & 0.022 & 0.070 & * \\
+         & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063 & 0.026 & 0.017 & ** \\
+         \bottomrule
+      \end{tabular}
+       \begin{tablenotes}
+         \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
+         \quad *** \(p < 0.01\)}
+       \end{tablenotes}
+    \end{threeparttable}
+    }
+  \end{center}
+\end{table}
+
+%------------------------------------------------------------------------------
+%	Appendix S4: Movement Model interpretation
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S4: Heatmaps
+%	Appendix S5: Heatmaps
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S5: Evolution of Betweenness Maps
+%	Appendix S6: Evolution of Betweenness Maps
 %------------------------------------------------------------------------------

### 2021-09-08 Backups/02_R-Scripts/01_DataCleaning/21_HumanInfluence(Worldpop).r

@@ -1,0 +1,103 @@
+################################################################################
+#### Human Influence Layer
+################################################################################
+# Description: Here I combine the layers from Worldpop, Geofabrik, Croplands,
+# Gabriele and Globelands to create two layers. First, a binary human influence
+# layer that depicts in each cell whether there are humans or not. Second, I
+# create a human density map, indicating whether human influence is strong or
+# weak.
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(raster)       # To manipulate spatial rasters
+library(terra)        # To manipulate spatial rasters
+library(rgeos)        # To manipulate vector data
+library(rgdal)        # To handle vector data
+library(viridis)      # For nice colors
+library(pbmcapply)    # For multicore with progress bar
+library(tidyverse)    # For data wrangling
+
+################################################################################
+#### Human Base and Density Layers
+################################################################################
+# Import data representing human influence
+population  <- rast("03_Data/02_CleanData/04_AnthropogenicFeatures_HumanDensity_WORLDPOP.tif")
+farms_gabs  <- rast("03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_GABRIELE.tif")
+farms_crops <- rast("03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_CROPLANDS.tif")
+farms_glob  <- rast("03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_GLOBELAND.tif")
+roads       <- rast("03_Data/02_CleanData/04_AnthropogenicFeatures_Roads_GEOFABRIK.tif")
+
+# Combine all the layers (make sure that the farms are not double counted)
+merged <- population + roads + max(farms_crops, farms_gabs, farms_glob)
+
+# Let's check how the resulting values are distributed (we only care about
+# values larger than 0)
+summary(values(merged)[values(merged) > 0])
+hist(values(merged)[values(merged) > 0])
+
+# Apparently there are some heavy outliers. Let's remove values beyond 50 (set
+# them to 50)
+values(merged)[values(merged) > 50] <- 50
+
+# Look at the histogram again
+hist(values(merged)[values(merged) > 0])
+
[…]

### 2021-09-08 Backups/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/KAZA-TFCA.R

@@ -1,0 +1,293 @@
+################################################################################
+#### Plot of KAZA-TFCA
+################################################################################
+# Description: Multiple plots depicting the KAZA-TFCA
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+wd <- "C:/Users/david/switchdrive/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)  # For data wrangling
+library(raster)     # To handle spatial data
+library(rgeos)      # To manipulate spatial data
+library(rgdal)      # To read and write spatial data
+library(tmap)       # To plot nice maps
+library(smoothr)    # To smooth spatial objects
+library(davidoff)   # Custom functions
+
+################################################################################
+#### Data Preparation
+################################################################################
+# Load required data
+africa <- "03_Data/02_CleanData/00_General_Africa.shp" %>%
+  readOGR()
+africa2 <- "03_Data/02_CleanData/00_General_Africa.shp" %>%
+  readOGR()
+kaza <- "03_Data/02_CleanData/00_General_KAZA_KAZA.shp" %>%
+  readOGR()
+prot <- "03_Data/02_CleanData/02_LandUseTypes_Protected_PeaceParks.shp" %>%
+  readOGR()
+water <- "03_Data/02_CleanData/03_LandscapeFeatures_MajorWaters_GEOFABRIK.shp" %>%
+  readOGR(.)
+
+# Buffer slightly
+africa <- gBuffer(africa, width = 0.01)
+
+# Disaggregate
+africa <- disaggregate(africa)
+
+# Identify sizes of areas
+africa$Size <- gArea(africa, byid = T)
+
+# Keep only the largest two
+africa <- subset(africa, Size %in% sort(africa$Size, decreasing = T)[1:2])
+
+# Simplify and smoothen africa shape
+africa <- gSimplify(africa, tol = 0.5)
+africa <- smooth(africa, method = "ksmooth")
+
+# Clip africa layer
+africa <- gIntersection(africa2, africa, byid = T)
+
[…]

### 2021-09-08 Backups/AreasReached2.R

@@ -1,396 +1,0 @@
-################################################################################
-#### Visualize Connections Between National Parks
-################################################################################
-# Description: In this script, we visualize the interpatch connectivity
-
-# Clear R's brain
-rm(list = ls())
-
-# Change the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)      # For data wrangling
-library(raster)         # To handle spatial data
-library(igraph)         # For network analysis
-library(rgdal)          # To load spatial data
-library(sf)             # To plot spatial stuff with ggplot
-library(ggspatial)      # To add scale bars etc to plots
-library(rgeos)          # To calculate areas
-library(ggnetwork)      # To plot network using ggplot
-library(viridis)        # For nice colors
-library(davidoff)       # Custom functions
-library(cowplot)        # Grab ggplot legend
-library(gtable)         # To combine ggplots
-library(grid)           # To combine ggplots
-library(gridExtra)      # To combine ggplots
-
-################################################################################
-#### Prepare Network
-################################################################################
-# Reload data on areas reached
-visits <- read_rds("03_Data/03_Results/99_AreasReached2.rds")
-
-# Load network areas and points
-areas <- readOGR("03_Data/03_Results/99_NetworkAreas.shp")
-points <- readOGR("03_Data/03_Results/99_NetworkPoints.shp")
-
-# Prepare layout
-lay <- coordinates(points)
-# lay <- coordinates(gCentroid(areas, byid = T))
-
-# Prepare network
-net <- graph_from_data_frame(
-    d        = visits
-  , vertices = unique(areas$ID)
-  , directed = T
-)
-
-# Check attributes
-vertex_attr(net)
-edge_attr(net)
-
-# First visualization
-plot(areas
-  , col    = "gray20"
-  , border = "black"
-  , bg     = "black"
-  , lwd    = 0.4
-)
-plot(net
-  , layout             = lay
-  , vertex.label       = NA
-  , vertex.size        = 1
-  , edge.arrow.size    = 0
[…]

### 2021-09-24 Chapter_1/04_Manuscript/Manuscript.tex
Worked on Dom's comments

@@
 \usepackage[nomarkers, nolists]{endfloat}
+\usepackage[colorinlistoftodos]{todonotes}
 
@@
 % Importance of dispersal & connectivity and mow its measured
-Dispersal of individuals is a crucial prerequisite for long-term species
-persistence but depends on a sufficient degree of landscape connectivity. To
-date, connectivity has primarily been quantified using least-cost analysis and
-circuit theory, albeit both methods make assumptions that are rarely met in
-reality. Least-cost analysis assumes that animals move towards a preconceived
-destination and choose a cost-minimizing route accordingly, whereas circuit
-theory presumes a complete random walk without directional persistence. While
-these shortcomings can be overcome by spatio-temporally explicitly simulating
-dispersal trajectories from individual-based movement models, a unified
-approach for such simulations is lacking.
+Dispersal of individuals is crucial for long-term species persistence and
+depends on a sufficient degree of landscape connectivity. Thusfar, researchers
+have investigated connectivity primarily using least-cost analysis and circuit
+theory, although both methods make assumptions that are rarely met in reality.
+Least-cost analysis assumes that animals move towards a known endpoint and are
+knowledgeable about the least costly route to reach it. Circuit theory relies on
+the assumption of a complete random walk without any directional persistence
+along the dispersing route. Albeit these shortcomings can be overcome by
+spatio-temporally explicitly simulating dispersing trajectories across the
+landscape, a unified approach for such a simulation is lacking.
 
 % What we do
-Here, we present a simple three-step workflow to simulate dispersal and asses
-landscape connectivity starting from empirical GPS movement data. We exemplify
-the workflow using dispersal data collected on the endangered African wild dog
-(\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation Area,
-the world's largest transboundary conservation area. In step one, we use
-integrated step selection functions to parametrize a mechanistic movement model
-describing habitat and movement preferences of dispersers. In step two, we
-employ the parametrized model to simulate individual dispersal trajectories
-across the landscape. In step three, we interpret simulated trajectories by
-converting them into three complementary connectivity maps: a heatmap revealing
-frequently traversed areas, a betweenness-map delineating dispersal corridors,
-and a map of inter-patch connectivity depicting the presence and intensity of
-functional links between habitat patches.
+Here, we present a simple three-step workflow to simulate dispersal and assess
+connectivity starting from empirical GPS movement data. To exemplify its
+application, we employ the workflow to dispersal data collected on the
+endangered African wild dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi
+Transfrontier Conservation Area, the world’s largest transboundary conservation
+area. In step one, we use integrated step selection functions to fit a
+mechanistic movement model describing habitat and movement preferences of
+dispersers. In step two, we apply the parameterized model to simulate individual
+dispersal trajectories. In step three, we compress the simulated trajectories
+into three complementary connectivity maps: a heatmap that highlights frequently
+traversed areas, a betweenness map that pinpoints dispersal corridors, and a map
+of interpatch connectivity, indicating the presence or absence of functional
+links between habitat patches.
 
 % Some results
-We find that simulating dispersal from indvidual-based movement models fitted
-via step selection functions are readily interpretable and offer great
-flexibility to realistically render dispersal movements. Moreover, by not
-enforcing connections towards known endpoints but instead letting dispersers
-react to the environment as they go, such simulations overcome shortcomings
-inherent least-cost path analysis, while at the same time allowing for
-autocorrelation in simulated steps, which can not be incorporated using circuit
-theory. Inspite of these advantages, simulating dispersal entails numerous
-non-trivial modeling decisions, such as the number of simulated individuals and
-the duration of simulated dispersal events.
+Our results show that dispersing African wild dogs move with directional
+persistence in the vicinity to waterbodies and preferrably through areas with
+little human influence. This is reflected in the simulation with resulted in
+several dispersal hotspots and corridors along waterbodies through areas with
+little or no anthropogenic pressure. Connectivity between national parks inside
+the KAZA-TFCA appears to be good, albeit the parks in Zambia and Zimbabwe are
+comparably less well connected to the rest of KAZA-TFCA, suggesting the
+necessity for additional protected areas that serve as ``stepping stones'' into
+these regions.
 
@@
 Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and colonize
-or recolonize unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However,
-the ability to disperse is contingent on a sufficient degree of landscape
-connectivity \citep{Fahrig.2003, Clobert.2012}, which is why the identification
-and protection of dispersal corridors that promote connectivity has become a
-task of fundamental importance \citep{Nathan.2008, Doerr.2011, Rudnick.2012}.
-The identification of dispersal corridors not only necessitates a comprehensive
-understanding of the factors that limit dispersal, but also an appropriate model
-to estimate connectivity \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To
-date, the two most prominent connectivity models in the literature are
-least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory
-(CT; \citealp{McRae.2006, McRae.2008}), both graph-based approaches that
-quantify conductance of the landscape based on habitat permeability
-\citep{Zeller.2012}.
+unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
+to disperse depends on a sufficient degree of landscape connectivity
+\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
+dispersal corridors that promote connectivity a task of fundamental importance
+\citep{Nathan.2008, Doerr.2011, Rudnick.2012}. The identification of dispersal
+corridors not only necessitates a comprehensive understanding of the factors
+that limit dispersal, but also an appropriate model to estimate connectivity
+\citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the two most
+commonly used connectivity models are least-cost path analysis (LCPA;
+\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
+McRae.2008}), both graph-based methods that quantify conductance of the
+landscape based on habitat permeability \citep{Zeller.2012}. However, both
+approaches rest on assumptions that appear unsuitable for dispersers, which is
+why  simulating dispersal from individual-based movement models provides a
+promising alternative approach to modeling connectivity \citep{Diniz.2019}.
 
@@
 \subsection{Issues with Traditional Connectivity Models}
-Despite their intuitive nature and ease of use, both methods make rigorous
-assumptions that are hardly ever met in reality \citep{Diniz.2019}. With LCPA,
-for instance, a least-costly route always exists, even if associated movement
-costs are unreasonably high and will never be incurred by dispersing individuals
-\citep{Sawyer.2011}. The method also presumes that animals move towards a
+Traditional connectivity models make assumptions that are rarely met for
+dispersers. LCPA, for instance, assumes that individuals move towards a
 preconceived endpoint and choose a cost-minimizing route accordingly
-\citep{Abrahms.2017}. While these assumptions may be fulfilled by migrating
-animals, they are unlikely to hold for dispersers, as dispersers typically move
-across unfamiliar territory and are therefore less aware of associated movement
-costs \citep{Koen.2014, Cozzi.2020}. CT relaxes these assumptions, but posits
-that animals move according to a random walk, entailing that autocorrelation
-between subsequent movements cannot be rendered \citep{Diniz.2019}. In reality,
-however, autocorrelation in animal movements is regularly observed
-\citep{Bovet.1991, Schultz.2001}, especially during dispersal \citep{Cozzi.2020,
-Hofmann.2021}. Some of these deficiencies can be alleviated using less stringent
-graph-based methods, such as least-cost \textit{corridors} \citep{Pinto.2009},
-\textit{thresholded} least-cost paths \citep{Landguth.2012}, or
-\textit{randomized} least-cost paths \citep{Panzacchi.2016}, yet a certain
-degree of arbitrariness remains. Besides this, graph-based methods are incapable
-of rendering the temporal dimension of dispersal, meaning that statements about
-the approximate duration required to move between habitats are impossible
+\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be fulfilled by
+migrating animals, it is unlikely to hold for dispersers, as dispersers
+typically move across unfamiliar territory and are therefore less aware of
+associated movement costs and potential endpoints \citep{Koen.2014, Cozzi.2020}.
+CT, on the contrary, posits that animals move according to a random walk,
+entailing that autocorrelation between subsequent movements cannot be rendered
+\citep{Diniz.2019}. For dispersers, however, autocorrelated movements are
+regularly observed \citep{Cozzi.2020, Hofmann.2021}, implying that subsequent
+relocations tend to fall onto a straight line. Moreover, graph-based methods are
+unable to reflect the temporal dimension of dispersal, meaning that statements
+about the approximate duration required to move between habitats impossible
 \citep{Martensen.2017, Diniz.2019}.
@@
 \subsection{What about IBMMs?}
-To overcome the unrealistic assumptions of LCPA and CT, dispersal simulations
-from individual-based movement models (IBMMs) have been proposed and applied
-\citep{Diniz.2019}. In these models, dispersal trajectories are simulated based
-on movement rules that govern how individuals move across and interact with the
-prevailing landscape (e.g. \citealp{Gustafson.1996, Gardner.2004, Graf.2007,
-KramerSchadt.2004, Revilla.2004, Revilla.2008, Kanagaraj.2013, Clark.2015,
-Allen.2016, Hauenstein.2019, Zeller.2020, Vasudev.2021}). Once the trajectories
-are simulated, they are further processed and converted into measures of
-connectivity \citep{Kanagaraj.2013, Zeller.2020}. Because IBMMs do not enforce
-connections towards preconceived endpoints, simulated individuals are allowed to
-dynamically adjust their behavior as they disperse. This enables to render
-potential interactions between movement behavior and habitat conditions,
-shifting the focus from a structural to a more functional view on connectivity
-\citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}. Furthermore, IBMMs
-render movement sequentially, meaning that the temporal dimension of dispersal
-becomes explicit and that autocorrelation between movements can be incorporated
-\citep{Diniz.2019}. In spite of these advantages, a unified framework to
-simulate dispersal and assess connectivity using IBMMs is lacking.
+The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
+trajectories from individual-based movement models (IBMMs) and by converting
+simulated trajectories into meaningful measures of connectivity
+\citep{Diniz.2019}). IBMMs allow to explicitly simulate how individuals move
+across and interact with the encountered landscape \citep{Kanagaraj.2013,
+Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020} and to render potential
+interactions between movement behavior and habitat conditions. This shifts the
+focus from a structural to a more functional view on connectivity
+\citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}. Furthermore,
+simulations from IBMMs generate movement sequentially, making the temporal
+dimension of dispersal explicit and allowing to model autocorrelation between
+subsequent movements \citep{Diniz.2019}. Finally, simulations from IBMMs do not
+enforce connections towards preconceived endpoints, thereby mitigating biases
+arising from misplaced endpoints. Despite these advantages, a unified framework
+to simulate dispersal and assess connectivity using IBMMs is lacking.
 
@@
 \subsection{Proposed Solution: Three-Step Workflow}
-In this paper, we draw on recent innovations in the field of movement ecology
-and network theory to propose a simple three-step workflow for the simulation of
-dispersal and assessment of connectivity (\Cref{GraphicalAbstract}). In the
-first step, GPS data of the focal species is combined with relevant habitat
-covariates and analyzed in an integrated step selection framework to fit a
-movement model. In the second step, the parametrized model is employed as an
-IBMM to simulate dispersal trajectories across the study area. Finally, in the
-third step, simulated trajectories are converted into a set of three
-complementary connectivity maps, each focusing on a different aspect of
-connectivity. The set of maps includes a heatmap, revealing frequently traversed
-areas, a betweenness-map, delineating dispersal corridors and bottlenecks, and a
-map of inter-patch connectivity, depicting the presence or absence, intensity,
-and dispersal duration of functional links between habitat patches.
-
-% Some Background on SSFs
-\subsection{Some Background on SSFs}
-To fit a movement model, we propose to employ the framework of step selection
-functions (SSFs; \citealp{Fortin.2005, Thurfjell.2014, Avgar.2016}). Originally,
-SSFs were developed to learn about habitat preferences of the focal species
-using GPS data \citep{Fortin.2005}. The method works by comparing covariates
-along realized steps (the straight-line connecting two consecutive GPS
-relocations; \citealp{Turchin.1998}) to the same covariates along randomly
-generated alternative steps \citep{Fortin.2005, Thurfjell.2014} using
-conditional logistic regression \citep{Fortin.2005, Muff.2020}. Recently,
-\cite{Avgar.2016} have introduced \textit{integrated} SSFs (ISSFs), which is a
-generalization of regular SSFs and requires that movement metrics are included
-in the corresponding regression model. This not only reduces biases in estimated
-preferences, but also permits simultaneous inference on a species' habitat
-kernel (i.e. habitat preferences), its movement kernel (i.e. movement
-preferences/capabilities), and potential interactions between the two kernels
-\citep{Avgar.2016, Fieberg.2020}. Importantly, a model that was fitted using
-ISSFs resembles a fully mechanistic IBMM, based on which movement can be
-simulated \citep{Avgar.2016, Signer.2017}. Previously, ISSFs have been applied
-to simulate utilization distributions of resident animals \citep{Avgar.2016,
-Signer.2017}, yet the method remains underutilized for the simulation of
-dispersal and detection of movement corridors.
-
-% Some Background on Connectivity "Metrics"
-\subsection{Some Background on  Connectivity ``Metrics'' }
-Irrespective of the chosen modeling framework, simulated dispersal trajectories
-need to be processed into measures of connectivity. To date, the most common
-practice to translate simulations into connectivity measures is to compute
-inter-patch connectivity between distinct habitats \citep{Diniz.2019}.
-Inter-patch connectivity is a simple ratio between the number of dispersers
-successfully moving between habitat patches and the total number of simulated
-individuals from each patch (e.g. \citealp{Kanagaraj.2013}). As an alternative,
-some researchers have tallied overlapping trajectories into what might be called
-``heatmaps'', i.e. maps that depict the frequency at which each spatial unit in
-the study area was traversed by simulated dispersers (e.g. \citealp{Peer.2008,
-Hauenstein.2019, Zeller.2020}). More recently, \cite{BastilleRousseau.2018}
-proposed a method to convert trajectories into a network, so that metrics from
-network theory can be computed. With regards to connectivity, the betweenness
-metric seems promising, as it indicates the number of shortest paths going
-through each node in the network and therefore highlights areas of exceptional
-importance for connecting different regions in the study area. Because each of
-the three metrics metrics focuses on a very different aspect of landscape
-connectivity, we propose to view them as complementary and to use them in
-concert when assessing connectivity.
+Here, we draw on recent advancements in the fields of movement ecology and
+network theory to propose a simple three-step workflow for simulating dispersal
+movements and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
+step one, we combine GPS movement data of dispersing individuals with relevant
+habitat covariates to fit a mechanistic movement model using integrated step
+selection functions (ISSFs, \citealp{Avgar.2016}). ISSFs permit
+simultaneous inference on the focal species' habitat kernel (i.e. habitat
+preferences), its movement kernel (i.e. movement preferences/capabilities), and
+potential interactions between the two \citep{Avgar.2016, Fieberg.2021}. In step
+two,  we simulate individual dispersal trajectories using the movement model
+parametrized in step one. Finally, in step three, we convert simulated
+trajectories into three complementary connectivity maps, each highlighting a
+different aspect of connectivity. The proposed set of maps includes (i) a
+heatmap revealing areas that are frequently traversed by simulated dispersers
+(e.g. \citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map
+delineating dispersal corridors and bottlenecks (e.g.
+\citealp{BastilleRousseau.2018}), (iii) and a map of inter-patch connectivity,
+depicting the presence or absence and intensity of use of specific connections,
+and the average dispersal duration between corresponding habitat patches (e.g.
+\citealp{Kanagaraj.2013}).
+
+% Introduction of the study species
+\subsection{Introduction of the Study Species}
+The endangered African wild dog (\textit{Lycaon pictus}) represents a highly
+mobile species whose persistence hinges on a  sufficient degree of landscape
+connectivity. Once common throughout sub-Saharan Africa, this species has
+disappeared from much of its historic range, largely due to human persecution,
+habitat fragmentation, and disease outbreaks \citep{Woodroffe.2012}. Wild dogs
+typically emigrate in single-sex coalitions \citep{McNutt.1996, Behr.2020}
+capable of dispersing several hundred kilometers \citep{DaviesMostert.2012,
+Masenga.2016, Cozzi.2020}. Although previous studies have investigated
+connectivity using LCPA \citep{Hofmann.2021} or CT \citep{Brennan.2020}, a more
+comprehensive and mechanistic understanding of connectivity is missing for this
+species (but see \citealp{Creel.2020}). With fewer than 6,000 free-ranging wild
+dogs remaining in fragmented populations \citep{Woodroffe.2012}, reliable
+information on landscape connectivity is essential for the conservation of this
+endangered carnivore.
+
+% % Some Background on SSFs
+% \subsection{Some Background on SSFs}
+% To fit a movement model, we propose to employ the framework of step selection
+% functions (SSFs; \citealp{Fortin.2005, Thurfjell.2014, Avgar.2016}). Originally,
+% SSFs were developed to learn about habitat preferences of the focal species
+% using GPS data \citep{Fortin.2005}. The method works by comparing covariates
+% along realized steps (the straight-line connecting two consecutive GPS
+% relocations; \citealp{Turchin.1998}) to the same covariates along randomly
+% generated alternative steps \citep{Fortin.2005, Thurfjell.2014} using
+% conditional logistic regression \citep{Fortin.2005, Muff.2020}. Recently,
+% \cite{Avgar.2016} have introduced \textit{integrated} SSFs (ISSFs), which is a
+% generalization of regular SSFs and requires that movement metrics are included
+% in the corresponding regression model. This not only reduces biases in estimated
+% preferences, but also permits simultaneous inference on a species' habitat
+% kernel (i.e. habitat preferences), its movement kernel (i.e. movement
+% preferences/capabilities), and potential interactions between the two kernels
+% \citep{Avgar.2016, Fieberg.2021}. Importantly, a model that was fitted using
+% ISSFs resembles a fully mechanistic IBMM, based on which movement can be
+% simulated \citep{Avgar.2016, Signer.2017}. Previously, ISSFs have been applied
+% to simulate utilization distributions of resident animals \citep{Avgar.2016,
+% Signer.2017}, yet the method remains underutilized for the simulation of
+% dispersal and detection of movement corridors.
+
+% % Some Background on Connectivity "Metrics"
+% \subsection{Some Background on  Connectivity ``Metrics'' }
+% Irrespective of the chosen modeling framework, simulated dispersal trajectories
+% need to be processed into measures of connectivity. To date, the most common
+% practice to translate simulations into connectivity measures is to compute
+% inter-patch connectivity between distinct habitats \citep{Diniz.2019}.
+% Inter-patch connectivity is a simple ratio between the number of dispersers
+% successfully moving between habitat patches and the total number of simulated
+% individuals from each patch (e.g. \citealp{Kanagaraj.2013}). As an alternative,
+% some researchers have tallied overlapping trajectories into what might be called
+% ``heatmaps'', i.e. maps that depict the frequency at which each spatial unit in
+% the study area was traversed by simulated dispersers (e.g. \citealp{Peer.2008,
+% Hauenstein.2019, Zeller.2020}). More recently, \cite{BastilleRousseau.2018}
+% proposed a method to convert trajectories into a network, so that metrics from
+% network theory can be computed. With regards to connectivity, the betweenness
+% metric seems promising, as it indicates the number of shortest paths going
+% through each node in the network and therefore highlights areas of exceptional
+% importance for connecting different regions in the study area. Because each of
+% the three metrics metrics focuses on a very different aspect of landscape
+% connectivity, we propose to view them as complementary and to use them in
+% concert when assessing connectivity.
 
@@
 \subsection{Case Study}
-To exemplify the application of the proposed workflow in
-(\Cref{GraphicalAbstract}), we investigate landscape connectivity for the
-endangered African wild dog (\textit{Lycon pictus}) as a case study. While once
-present across entire sub-Saharan Africa, this species has disappeared from a
-vast majority of its historic range, mainly due to persecution by humans,
-habitat fragmentation and destruction, and deadly diseases
-\citep{Woodroffe.2012}. Some of the biggest remaining populations reside near
-the Moremi Game Reserve in northern Botswana \citep{McNutt.1996, Cozzi.2013,
-Behr.2020}, which is part of the Kavango-Zambezi Transfrontier Conservation Area
-(KAZA-TFCA), the world's largest transboundary protected area. We previously
-assessed landscape connectivity for dispersing African wild dogs within this
-ecosystem using least-cost methods \citep{Hofmann.2021}. More specifically, we
-used ISSFs to fit a habitat selection model and predicted a permeability surface
-based on which we computed least-cost paths and corridors between protected
-areas. We now expand on this knowledge and use ISSFs to develop a more detailed,
-mechanistic movement model of dispersing wild dogs (\Cref{GraphicalAbstract}).
-We then employ the model to simulate 80'000 dispersers moving 2'000 steps across
-the landscape of the KAZA-TFCA. Finally, we condense the simulated trajectories
-into a heatmap, betweenness-map, and inter-patch connectivity map. We thereby
-demonstrate that simulations from ISSFs provide a viable alternative to LCPA or
-CT and deliver a more comprehensive understanding of the emergence of
-connectivity. Nevertheless, we recognize that simulations from ISSFs entail
-several non-trivial modeling-decisions, such as the number of simulated
-dispersers and the duration of simulated dispersal events.
+In this study, we illustrate the proposed three-step workflow
+(\Cref{GraphicalAbstract}) to assess landscape connectivity for African wild
+dogs within the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), the
+world's largest transboundary conservation area. To this end, we use GPS
+movement data of 16 dispersing wild dogs originating from a free-ranging
+population in northern Botswana to parametrize a mechanstic movement model,
+which we then employed to simulate 80,000 dispersal trajectories across the
+landscape of the KAZA-TFCA. We anticipated that simulations based on our
+three-step workflow would overcome several of the above highlighted conceptual
+shortcomings of traditional connectivity models and provide a more comprehensive
+view of landscape connectivity.
 
@@
 \subsection{Study Area}
-The study area of our wild dog case study was defined by a bounding box centered
-at -17\degree 13'9''S, 23\degree 56'4''E (\Cref{StudyArea}a) and stretched over
-1.3 Mio. km\textsuperscript{2}, encompassing the entire KAZA-TFCA
+The study area centered at -17\degree 13'9''S, 23\degree 56'4''E
+(\Cref{StudyArea}a) is located in southern Africa (\Cref{StudyArea}a) and spans
+more than 1.3 Mio. km\textsuperscript{2}, encompassing the entire KAZA-TFCA
 (\Cref{StudyArea}b). The KAZA-TFCA comprises parts of Angola, Botswana, Namibia,
-Zimbabwe, and Zambia and hosts diverse landscapes, ranging from savannah to
+Zimbabwe, and Zambia and hosts diverse landscapes; ranging from savannah to
 grassland and from dry to moist woodland habitats. In its center lies the
 Okavango Delta, a dominant hydrogeographical feature and the world's largest
-flood-pulsing inland delta. Although large portions of the KAZA-TFCA are
+flood-pulsing inland delta. Although large portions of the KAZA-TFCA are part of
 designated national parks and other protected areas, considerable human
-influence remains from roads, agricultural sites and settlements.
+influence from roads, agricultural sites, and settlements remains.
 
@@
             % % upto here
-            \draw[densely dotted, black] (0.169, 0.222) -- (0.364, 0.955);
-            \draw[densely dotted, black] (0.169, 0.157) -- (0.364, 0.074);
+            \draw[densely dotted, blue] (0.169, 0.222) -- (0.364, 0.955);
+            \draw[densely dotted, blue] (0.169, 0.157) -- (0.364, 0.074);
         \end{scope}
@@
 
-\subsection{GPS Data}
-GPS data of 16 dispersing wild dogs was obtained as described in
-\cite{Cozzi.2020} and \cite{Hofmann.2021}. We collected data on 7 female and 9
-male dispersers between 2011 and 2019 from a free-ranging population inhabiting
-the Moremi National Park in northern Botswana. Because behavior during dispersal
-is more pertinent to landscape connectivity than behavior during residence
-\citep{Elliot.2014, Abrahms.2017}, we discarded all data that was collected
-during residency. For some individuals, exact dispersal dates were known from
-field observations, whereas in other cases, we determined dispersal phases using
-net-squared displacement. Net-squared displacement is a metric that measures the
-squared Euclidean distance of a GPS relocation to a reference point
-\citep{Borger.2012}, which in our case was set to the center of each
-individual's natal home range. Thus, dispersal was deemed to have started when
-an individual left its natal home range and ended once individuals became
-sedentary again. During dispersal, GPS collars recorded a fix every 4 hours and
-they regularly transmitted data over the Iridium satellite system. After
-collection, we converted GPS coordinates of dispersers (n = 4'169) to steps,
-where each step represented the straight-line connecting two consecutive GPS
-relocations \citep{Turchin.1998}. To ensure a regular sampling interval, we
-removed fixes that were not successfully collected on the 4-hourly schedule (\(
-\pm \) 15 minutes) and identified bursts of regularly sampled GPS fixes.
-
-\subsection{Covariates}
-We represented the physical landscape in our study area using a set of habitat
-covariates, including water-cover, distance-to-water, woodland-cover, and
-shrub/grassland-cover. Because water-cover greatly changes within and between
-years in the Okavango Delta, we applied a remote sensing algorithm and obtained
-weekly updated water-cover layers and corresponding distance-to-water layers
-from MODIS satellite imagery (see \citealp{Wolski.2017} and Appendix A3 in
-\citealp{Hofmann.2021}). Resulting layers thus temporally aligned with each
-dispersal event, ensuring minimal temporal lag between observed movements and
-corresponding water layers. We furthermore included a proxy for human influence,
-rendering anthropogenic pressures stemming from human-density, agricultural
-sites, and roads. All spatial layers were coarsened or interpolated to a target
-resolution of 250 m by 250 m. A detailed description of the preparation of each
-habitat covariate is provided in \cite{Hofmann.2021}.
+\subsection{Data Collection and Preparation}
+\subsubsection{GPS Data}
+We collected GPS movement data on 16 dispersing wild dogs (7 females and 9
+males) between 2011 and 2019 from a free-ranging population in northern Botswana
+(details on the data collection can be found in \cite{Cozzi.2020} and
+\cite{Hofmann.2021}). Because behavior during dispersal is more pertinent to
+landscape connectivity than behavior during residence \citep{Elliot.2014,
+Abrahms.2017}, we discarded data collected when individauls were resident. We
+determined the exact timing of emigration and settlement using direct field
+observations and through visual inspection of the net squared displacement (NSD)
+metric. NSD measures the squared Euclidean distance of a GPS relocation to a
+reference point \citep{Borger.2012}, which we set to the center of each
+dispersers' natal home range. Thus, dispersal was deemed to have started once an
+individual left its natal home range and ended once the NSD metric ramined
+constant, indicating settlement. During dispersal, GPS collars recorded a fix
+every 4 hours and they regularly transmitted data over the Iridium satellite
+system. To ensure regular timespans between fixes, we removed any fix that was
+not obtained on the aspired 4-hourly schedule (\( \pm \) 15 minutes). We then
+converted the remaining fixes (n = 4'169) into steps, where each step
+represented the straight-line movement between two consecutive GPS relocations
+\citep{Turchin.1998}.
+
+\subsubsection{Habitat Covariates}
+We represented the physical landscape in our study area by the habitat
+covariates \textit{water-cover, distance-to-water, woodland-cover,
+shrub/grassland-cover, and human-influence}. To reflect the seasonality of water
+in the study area, we obtained weekly updated layers for water-cover and
+corresponding layers for distance-to-water from MODIS satellite imagery using a
+remote sensing algorithm \citealp{Wolski.2017, Hofmann.2021} that is implemented
+in the \textsf{floodmapr} package which is available on GitHub
+(\url{https://github.com/DavidDHofmann/floodmapr}). For each step, we thus used
+the temporally associated water layers, reflecting prevailing conditions at the
+time when the step was realized. To ensure a consistent resolution across
+habitat covariates, we coarsened or interpolated all layers to a resolution of
+250 m x 250 m. A more detailed description of how we prepared each habitat
+covariate is provided in \cite{Hofmann.2021}.
 
 Besides habitat covariates, we computed movement metrics that we used as
-movement covariates in our ISSF models \citep{Avgar.2016, Fieberg.2020}.
+movement covariates in our ISSF analysis \citep{Avgar.2016, Fieberg.2021}.
 Movement metrics were calculated for each step and included the step length
 (\textsf{sl}), its natural logarithm (\textsf{log(sl)}), and the cosine of the
-relative turning angle (\textsf{cos(ta)}). Because wild dogs follow a distinct
-activity pattern \citep{Castello.2018}, we also coded a binary variable
-(\textsf{LowActivity}) indicating whether a step was realized during periods of
-low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
-(17:00 to 09:00 local time). Handling and manipulation of all data, as well as
-all models and simulations were implemented with the statistical software {\tt
-R}, version 3.6.6 \citep{R.2020}. Some helper functions were written in {\tt
-C++} and imported into {\tt R} using the {\tt Rcpp} package
-\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
-
-\subsection{Movement Model}
+relative turning angle (\textsf{cos(ta)}).  Moreover, we created the binary
+variable \textsf{LowActivity} indicating whether a step was realized during
+periods of low wild dog activity (09:00 to 17:00 local time) or high wild dog
+activity (17:00 to 09:00 local time, \citealp{Cozzi.2012}). We performed all
+data preparations, spatial computations, and statistical analysis in R, version
+3.6.6 \citep{R.2020}. Some helper functions were written in {\tt C++} and
+imported into {\tt R} using the {\tt Rcpp} package \citep{Eddelbuettel.2011,
+Eddelbuettel.2013}. We provide annotated R-codes to reproduce the three-step
+workflow using simulated data through GitHub
+(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, codes
+for the African wild dog case study are available through an online repository.
+
+\subsection{Step 1 - Movement Model}
 We used ISSFs to parametrize a mechanistic movement model for dispersing wild
-dogs \citep{Avgar.2016}. To conduct ISSF analysis, we paired each realized step
-with 24 random steps, so that a realized step plus its 24 random steps formed a
-stratum that received a unique identifier. As suggested by \cite{Avgar.2016}, we
-generated random steps by sampling random turning angles from a von Mises
-distribution with concentration parameter \(\kappa = 0\) (i.e. a uniform
-distribution on (\(-\pi, +\pi\))) and step lengths from a gamma distribution
-that was fitted to realized steps (scale \(\theta\) = 6'308 and shape \(k\) =
-0.37)). Along each step, we extracted and averaged values from the habitat
-covariate layers using the {\tt velox} package \citep{Hunziker.2021}. We also
+dogs \citep{Avgar.2016}. More specifically, we paired each realized step with 24
+random steps, so that a realized step plus its 24 random steps formed a stratum
+that received a unique identifier. As suggested by \cite{Avgar.2016}, we
+generated random steps by sampling random turning angles from a uniform
+distribution (\(-\pi, +\pi\))) (which is equivalent to a von Mises distribution
+with concentration parameter \(\kappa = 0\)) and step lengths from a gamma
+distribution that was fitted to realized steps (scale \(\theta\) = 6'308 and
+shape \(k\) = 0.37)). Along each step, we extracted and averaged values from the
+habitat covariate layers using the {\tt velox} package \citep{Hunziker.2021} and
 calculated the movement metrics \textsf{sl}, \textsf{log(sl)}, and
@@
 that animals assigned a selection score \(w(x)\) of the exponential form to each
-step \citep{Fortin.2005}. The selection score \(w(x)\) of each step thus
-depended on the step's associated covariates (\(x_1, x_2, ..., x_n\)) and on the
-animal's preferences (i.e. relative selection strengths; \citealp{Avgar.2017})
-towards these covariates (\(\beta_1, \beta_2, ..., \beta_n\)):
+step \citep{Fortin.2005}. \(w(x)\) depended on the step's associated covariates
+(\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e. relative
+selection strengths; \citealp{Avgar.2017}) towards these covariates (\(\beta_1,
+\beta_2, ..., \beta_n\)):
 
@@
 
-\noindent To estimate preferences (i.e. the \(\beta\)'s), we used mixed effects
-conditional logistic regression analysis \citep{Muff.2020} that we implemented
-using the r-package {\tt glmmTMB} \citep{Brooks.2017}. We treated animal IDs as
-random effects and modeled random slopes for each covariate.
-
-The structure of the movement model was based on a habitat selection model that
-we previously developed \citep{Hofmann.2021}. In the original model (hereafter
-referred to as \textit{base model}), no interactions between habitat covariates
-(i.e. \textsf{Water, DistanceToWater\textsuperscript{0.5}, Woodland,
-Shrubs/Grassland, Human Influence}) and movement covariates (i.e. \textsf{sl,
-log(sl), cos(ta)}) were considered. Hence, we slightly expanded the base model
-and proposed interactions between all movement covariates and habitat
-covariates, assuming that movement behavior depends on habitat conditions. More
-specifically, we started with the base model and incrementally increased model
-complexity by adding all possible two-way interactions between habitat
-covariates and movement covariates. For instance, for the covariate
-\textsf{Water}, we proposed the interactions \textsf{Water:sl},
-\textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides these interactions,
-we also allowed for correlations between turning angles and step lengths by
-proposing the interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)}.
-Furthermore, we formed the interactions \textsf{sl:LowActivity} and
-\textsf{log(sl):LowActivity} to render that step lengths are likely to be
-shorter during periods of inactivity. To assess the most parsimonious movement
+\noindent To estimate preferences (i.e. the \(\beta\)-coefficients), we used
+mixed effects conditional logistic regression analysis \citep{Muff.2020} that we
+implemented using the r-package {\tt glmmTMB} \citep{Brooks.2017}. To capitalize
+on the poisson trick introduced by \cite{Muff.2020}, we defined random
+intercepts for each stratum and fixed the intercept variance to an arbitraty
+high value of \(10 ** 6\). We also used disperser identity to model random
+slopes for all covariates.
+
+The covariate structure of the movement model was based on a habitat selection
+model that we previously developed (hereafter referred to as based model,
+\citealp{Hofmann.2021}). In the base model, no interactions among habitat
+covariates and movement covariates were considered. Hence, we expanded the base
+model and allowed for interactions between all movement covariates and habitat
+covariates, thus reflecting that movement behavior may depend on habitat
+conditions (details in Appendix AX). To assess the most parsimonious movement
 model among model candidates, we ran stepwise forward model selection based on
-Akaike's Information Criterion (AIC, \citealp{Burnham.2002}).
-
-We validated the predictive power of the most parsimonious movement model using
-k-fold cross-validation for case-control studies \cite{Fortin.2009}.
-Specifically, we randomly assigned 80\% of the strata to a training set and the
-remaining 20\% to a testing set. With the training data we parametrized a
-movement model and predicted selection scores \(w(x)\) for all steps in the test
-data. Within each stratum, we then assigned ranks 1-25 to each step based on
-predicted selection scores, so that rank 1 was given to the step with the
-highest score \(w(x)\). Within each strata, we determined the realized step's
-rank and calculated rank frequencies of realized steps across all strata.
-Finally, we computed Spearman's rank correlation between ranks and associated
-frequencies \(r_{s, realized}\). We replicated this procedure 100 times and
-computed the mean correlation coefficient (\(\bar{r}_{s, realized}\)), as well
-as its 95\% confidence interval across all replicates. For comparison, we
-repeated the same procedure 100 times assuming random preferences. Random
-preferences were implemented by discarding the realized step from all strata and
-identifying the rank of a random step in each stratum. Again, we calculated
-Spearman's rank correlation coefficient (\(r_{s, random}\)), its mean across
-repetitions (\(\bar{r}_{s, random}\)), and its 95\% confidence interval.
-Ultimately, this validation proves a significant prediction in case the
-confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do
-not overlap \citep{Fortin.2009}.
-
-\subsection{Dispersal Simulation}
-We used the most parsimonious movement model to simulate 80'000 virtual
-dispersers moving across the KAZA-TFCA. The simulation resembled an
-``inverted''  ISSF and was set up as follows. (1) We defined a random source
-point and assumed a random initial orientation of the animal. (2) Departing from
-the source point, we generated 25 random steps by sampling turning angles from a
-uniform distribution (\(-\pi, +\pi\)) and step lengths from our fitted gamma
-distribution. Similar to the input data, each random step represented the
-straight line movement possible within 4 hours. To prevent unreasonably large
-steps, we restricted sampled step lengths to a maximum of 35 km, which
-corresponded to the farthest distance ever traveled within 4 hours by one of our
-dispersers. (3) Along each random step, we extracted and averaged values from
-the habitat covariate layers and we calculated movement covariates. To ensure
+Akaike's Information Criterion (AIC, \citealp{Burnham.2002}). Finally, we
+validated the predictive power of the most parsimonious model using k-fold
+cross-validation for case-control studies as desribed in \cite{Fortin.2009}. In
+brief, this validation proofs a significant prediction in case the Spearman rank
+correlation of predicted step ranks and associated frequencies under the
+movement model is significantly stronger than under random preferences (details
+in Appendix AX).
+
+\subsection{Step 2 - Dispersal Simulation}
+We used the most parsimonious movement model to simulate individual disperseral
+trajectories within the study area. The simulation of a dispersal trajectory
+resembled an ``inverted''  ISSF and was set up as follows. (1) We defined a
+random source point and assumed a random initial orientation of the animal. (2)
+Starting from the source point, we generated 25 random steps by sampling turning
+angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths from our
+fitted gamma distribution. Like in the empirical data, each random step
+represented the straight line movement possible within 4 hours. To prevent
+unreasonably large steps, we restricted sampled step lengths to a maximum of 35
+km (i.e. the farthest dispersal distance traveled within 4 hours in our data).
+(3) Along each random step, we extracted and averaged values from the respective
+habitat covariate layers and calculated movement covariates. To ensure
 compatible scales with the fitted movement model, we standardized extracted
-values using the same scaling parameters that we applied to standardize our
-input data. (4) We applied the parametrized movement model to predict the
-selection score \(w(x)\) for each step and we translated predicted scores into
-probabilities using \ref{EQ2}. (5) We sampled one of the random steps based on
-predicted probabilities and determined the animal's new position. We then
-repeated steps (2) to (5) until 2'000 steps were realized.
+values using the mean and standard deviation from the empirical data. (4) We
+applied the parametrized movement model to predict the selection score \(w(x)\)
+for each step and we translated predicted scores into probabilities using
+\ref{EQ2}. (5) We used predicted probabilities to sample one of the random steps
+and determined the animal's new position. We repeated steps (2) to (5) until
+2,000 steps (i.e. 400 consecutive dispersal days) were realized.\todo{Since we did not explain the 8-hourly gaps, this might be confusing}
 
@@
 buffer zone, we resampled transgressing steps until they fully lied within the
-buffer zone, forcing individuals to be repelled from these map borders.
-
-\subsection{Source Points}
-We released 80'000 virtual dispersers from 80'000 unique source points
-distributed across the study area. 50'000 virtual dispersers were released from
-randomly selected source points within contiguous protected areas that were
-large enough to sustain viable wild dog populations (\(>\) 700
-km\textsuperscript{2}; \citealp{Pomilia.2015}), whereas another 30'000
-dispersers were released at random locations inside the 100 km wide buffer to
-render potential immigrants into the study system. By distributing source points
-randomly, the number of source points per km\textsuperscript{2} was
-approximately equal.
-
-\subsection{Convergence}
-\textcolor{red}{Maybe this paragraph should go into another paper?} To verify
-that the number of simulated individuals sufficed to ensure reliable estimates
-of connectivity, we evaluated how the relative traversal frequency across the
-landscape changed as we increased the number of simulated individuals. For this,
-we distributed 1'000 rectangular ``checkpoints'', each with an extent of 5 km x
-5 km at random locations inside the main study area (excluding the buffer). At
-each checkpoint, we then determined the relative traversal frequency by
-simulated dispersers as we increased the number of simulated individuals from 1
-to 50'000. To assess variability in the relative traversal frequency, we
-repeatedly subsampled 100 times from all 50'000 and computed the mean traversal
-frequency across replicates, as well as its 95\% confidence-interval. We deemed
-that a checkpoint converged as soon the width of the confidence-interval dropped
-below a value of 0.01.
-
-\subsection{Heatmap}
-To identify dispersal hotspots across our study area, we created a heatmap
-indicating the absolute frequency at which each raster-cell in our study area
-was visited by virtual dispersers (e.g. \citealp{Peer.2008, Hauenstein.2019,
-Zeller.2020}). For this, we rasterized all simulated trajectories and tallied
-them into a single map. If the same trajectory crossed a raster-cell twice, we
-only counted it once, thereby mitigating biases from individuals that were
-trapped and moved in circles. To achieve high performance rasterization, we used
-the R-package {\tt terra} \citep{Hijmans.2021b}.
-
-\subsection{Betweenness}
+buffer zone, forcing individuals to remain within the expanded study area.
+
+For the simulation, we distributed 80,000 unique source points within the study
+area. Of these, 50,000 were random locations inside protected areas that were
+larger than the average home range size of wild dogs (i.e. \(>\) 700
+km\textsuperscript{2}; \citealp{Pomilia.2015}), while the remaining 30,000
+source points were placed randomly inside the buffer zone, mimicing potential
+immigration into the study area (Figure S1). Due to the random distribution of
+source points, the number of source points per km\textsuperscript{2} in selected
+areas was approximately equal.
+
+To ensure reliable connectivity estimates, we determined the number of simulated
+dispersal trajectories required for connectivity to reach a  ``steady state''
+across the entire study area. For this purpose, we distributed 1,000 rectangular
+``checkpoints'', each with an extent of 5 km x 5 km at random coordinates within
+the study area (excluding the buffer). We then determined the relative frequency
+at which each checkpoint was traversed by simulated dispersers (hereafter
+referred to as relative traversal frequency) as we gradually increased the
+number of simulated trajectories from 1 to 50,000. To assess variability in the
+relative traversal frequency, we repeatedly subsampled 100 times from all 50'000
+dispersal trajectories and computed the mean traversal frequency across
+replicates, as well as its 95\% confidence-interval for each checkpoint. We
+considered connectivity to have reached a steady state once the width of the
+prediction-interval dropped below a value of 0.01 for all checkpoints.
+
+\subsection{Step 3 - Connectivity Maps}
+\subsubsection{Heatmap}
+To identify dispersal hotspots within the study area, we created a heatmap
+indicating the absolute frequency at which each raster-cell was traversed by
+simulated dispersers (e.g. \citealp{Peer.2008, Hauenstein.2019, Zeller.2020}).
+That is, we rasterized all simulated trajectories onto a raster with 250 m x 250
+m resolution and tallied them into a single map. If the same trajectory crossed
+a raster-cell twice, we only counted it once, thereby mitigating biases from
+individuals that were trapped and moved in circles. To achieve high performance
+rasterization, we used the R-package {\tt terra} \citep{Hijmans.2021b}.
+
+\subsubsection{Betweenness Map}
 To pinpoint discrete movement corridors and bottlenecks, we converted simulated
-trajectories into a network and calculated betweenness scores for each
-raster-cell in the study area \citep{BastilleRousseau.2018}. Betweenness
+trajectories into a network and calculated betweenness scores for all
+raster-cells in the study area \citep{BastilleRousseau.2018}. Betweenness
 measures how often a specific network-node (i.e. raster-cell) lies on a shortest
@@
 \citep{BastilleRousseau.2018}. To transform simulated trajectories into a
-network, we overlaid the study area (including the buffer) with a raster
-resolved at 5 km x 5 km, where the center of each raster-cell served as node in
-the final network. To identify edges (connections) between the nodes, we used
-the simulated trajectories and determined all transitions occurring from one
-node to another, as well as the frequency at which those transitions occurred.
-This resulted in an edge-list that we translated into a weighted network using
-the r-package {\tt igraph} \citep{Gabor.2006}. The weight of each edge was
-determined by the frequency of transitions, yet because {\tt igraph} handles
-edge weights (\(\omega\)) as costs, we inverted the traversal frequency in each
-cell by applying \(\omega = \frac{mean(Traversal Frequency)}{Traversal
-Frequency_i}\). Consequently, edges that were traversed frequently were assigned
-small weights and vice versa. Finally, we used the weighted network to calculate
-betweenness scores for all network nodes.
-
-\subsection{Inter-Patch Connectivity}
-To examine functional links between distinct patches in the KAZA-TFCA, we
-calculated inter-patch connectivity between national parks. The decision to
-focus on national parks was purely out of simplicity and does not imply that
+network, we followed \citep{BastilleRousseau.2018} and overlaid the study area
+(including the buffer) with a 5 km x 5 km raster, where the center of each
+raster-cell served as node in the final network. To identify edges (i.e.
+connections) between the nodes, we used the simulated trajectories and
+determined all transitions occurring from one node to another, as well as the
+frequency at which those transitions occurred. This resulted in an edge-list
+that we translated into a weighted network using the r-package {\tt igraph}
+\citep{Gabor.2006}. The weight of each edge was determined by the frequency of
+transitions, yet because {\tt igraph} handles edge weights (\(\omega\)) as
+costs, we had to invert the traversal-frequency thorugh each raster-cell by
+applying \(\omega = \frac{mean(Traversal Frequency)}{Traversal Frequency_i}\).
+Consequently, edges that were traversed frequently were assigned small weights
+and vice versa. Finally, we used the weighted network to calculate betweenness
+scores at all network nodes.
+
+\subsubsection{Inter-Patch Connectivity Map}
+To examine the presence and intensity of functional links (i.e. connections)
+between specific patches inside the KAZA-TFCA, we calculated inter-patch
+connectivity between national parks. The decision to only consider national
+parks as our  ``patches'' was purely out of simplicity and does not imply that
 connections between other protected areas are impossible. To quantify
@@
 originating from one national park successfully moved into another national
-park. Successful movement was said to be achieved if the individuals' trajectory
-intersected with the corresponding national park at least once. We also recorded
-the number of steps required until the first intersection with the respective
-national park, allowing us to compute average dispersal durations from one park
-to another. In summary, we determined \textit{if} and \textit{how often}
-dispersers moved between certain national parks, as well as \textit{how long}
-dispersers had to move to realize those connections.
+park. We considered movements successful if the individuals' dispersal
+trajectory intersected with the target national park at least once. We also
+recorded the number of steps required to reach the first intersection with the
+respective national park, allowing us to compute the average dispersal durations
+from one park to another. In summary, we determined \textit{if} and \textit{how
+often} dispersers moved between certain national parks, as well as \textit{how
+long} individuals had to move to make these connections.
 
@@
 \subsection{Movement Model}
-Our most parsimonious movement model resulting from the ISSF analysis consisted
-of a movement kernel, a habitat kernel, and their interactions. Compared to the
-base model reported in \citep{Hofmann.2021}, the movement model retained several
-additional interactions between habitat covariates and movement covariates
-(\Cref{MovementModel} and \Cref{MovementModelNumbers}). Although multiple models
-received an AIC weight above zero (Table 1 in Appendix S1), we only considered
-results from the most parsimonious model for simplicity. Since all models with
-positive AIC weight contained similar covariates (Table S1), this decision only
-marginally influenced subsequent steps. To aid with the interpretation of the
-final model, we followed \cite{Fieberg.2020} and produced a collection of
-auxiliary plots that are provided in Appendix S2.
-
-When all other covariates held constant at their means, the habitat kernel
-reveals that dispersing wild dogs avoid water but prefer its proximity, that
-dispersers avoid areas that are covered by woodlands, yet prefer regions covered
-by shrublands or grasslands, and that dispersers avoid movement through
-landscapes that are dominated by humans. All of these effects are strong and
-significant, except for the effect of \textsf{distance to water}, which is
-insignificant on the 5\% significance level.
-
-With regards to the movement kernel, the positive estimate for \textsf{cos(ta)}
-indicates that dispersers move with directional persistence, unlike what was
-proposed by the uniform turning angle distribution. This preference is
-particularly pronounced when dispersers realize large steps (i.e. move quickly),
-as indicated by the positive estimates for \textsf{cos(ta):sl} and
-\textsf{cos(ta):log(sl)}. Finally, the negative estimate for the interaction
-\textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (i.e. move
-slower) between 09:00 and 17:00 o'clock than during the rest of the day. Aside
-from the interaction \textsf{sl:LowActivity}, which has a very strong effect,
-effects are moderate strength yet significant on the 5\% significance level;
-only the interaction \textsf{cos(ta):sl} appears as insignificant on the 5\%
-significance level.
-
-Finally, several significant interactions between movement and habitat
-covariates suggest that movement behavior is contingent on habitat conditions.
-For example, there's strong evidence that dispersers realize shorter steps in
-areas covered by water or in forested areas. Furthermore, it appears that
-dispersers realize larger steps in areas that are dominated by shrubs/grassland,
-but shorter steps in areas that are distant to water. Finally, it seems that the
-preference for directional persistence is less pronounced when dispersers cross
-human-dominated landscapes, yet more pronounced at great distance from water.
-Nevertheless, some of these effects are only weakly significant and exhibit
-small effect sizes.
-
-The k-fold cross-validation of the movement model shows that the model
-substantially outperforms a random guess (\Cref{MovementModel}b) and correctly
-assigns high selection scores to realized steps. Confidence intervals of
-\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not overlap and
-therefore prove a reliable prediction. Furthermore, the significant correlation
-between ranks and corresponding frequencies for realized steps indicates a good
-fit between predictions and observations (\Cref{MovementModel}b). In comparison
-to the base model (\(\bar{r}_{s, realized} = -0.55\); \citealp{Hofmann.2021}),
-the inclusion of interactions between movement and habitat covariates slightly
-improved model performance (\(\bar{r}_{s, realized} = -0.65\).
+The most parsimonious movement model consisted of movement covariates, habitat
+covariates and several of their interactions, suggesting that movement behavior
+during dispersal depends on habitat conditions (\Cref{MovementModel}, Table S1
+and Table S2). Although multiple models received an AIC weight > 0 (Table S1),
+we only considered results from the most parsimonious model for simplicity. This
+decision only marginally influenced subsequent steps as all models with positive
+AIC weights contained similar covarites (Table S1). Aids for interpreting main
+effects and interactions are provided in Figure S2. Under average conditions,
+dispersing wild dogs avoided moving through water, woodlands, and areas
+dominated by humans, but preferred shrublands or grasslands
+(\Cref{MovementModel}). Dispersers realized shorter steps (indicating slower
+movements) in areas covered by water or woodland, while realizing larger steps
+in areas dominated by shrubs or grass. Moreover, dispersing wild dogs moved
+faster during twilight and at night (i.e. between 17:00 and 09:00 o'clock) than
+during the rest of the day. Although dispersers showed a preference for
+directional movements, especially when moving quickly, they did less so in
+proximity to humans or water, resulting in more tortuous movements in such
+areas.
+
+% When all other covariates held constant at their means, the habitat kernel
+% reveals that dispersing wild dogs avoid water but prefer its proximity, that
+% dispersers avoid areas that are covered by woodlands, yet prefer regions covered
+% by shrublands or grasslands, and that dispersers avoid movement through
+% landscapes that are dominated by humans. All of these effects are strong and
+% significant, except for the effect of \textsf{distance to water}, which is
+% insignificant on the 5\% significance level.
+%
+% With regards to the movement kernel, the positive estimate for \textsf{cos(ta)}
+% indicates that dispersers move with directional persistence, unlike what was
+% proposed by the uniform turning angle distribution. This preference is
+% particularly pronounced when dispersers realize large steps (i.e. move quickly),
+% as indicated by the positive estimates for \textsf{cos(ta):sl} and
+% \textsf{cos(ta):log(sl)}. Finally, the negative estimate for the interaction
+% \textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (i.e. move
+% slower) between 09:00 and 17:00 o'clock than during the rest of the day. Aside
+% from the interaction \textsf{sl:LowActivity}, which has a very strong effect,
+% effects are moderate strength yet significant on the 5\% significance level;
+% only the interaction \textsf{cos(ta):sl} appears as insignificant on the 5\%
+% significance level.
+%
+% Finally, several significant interactions between movement and habitat
+% covariates suggest that movement behavior is contingent on habitat conditions.
+% For example, there's strong evidence that dispersers realize shorter steps in
+% areas covered by water or in forested areas. Furthermore, it appears that
+% dispersers realize larger steps in areas that are dominated by shrubs/grassland,
+% but shorter steps in areas that are distant to water. Finally, it seems that the
+% preference for directional persistence is less pronounced when dispersers cross
+% human-dominated landscapes, yet more pronounced at great distance from water.
+% Nevertheless, some of these effects are only weakly significant and exhibit
+% small effect sizes.
+
+The k-fold cross-validation of the movement model showed that the final model
+substantially outperforms a random guess , suggesting reliable predictions
+(confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\)
+do not overlap). Moreover, the model correctly assigns high selection scores to
+realized steps (\Cref{MovementModel}b), indicating a good fit between
+predictions and observations. Compared to the base model (\(\bar{r}_{s,
+realized} = -0.55; 95\%-CI = [-0.57, -0.52]\); \citealp{Hofmann.2021}), the
+inclusion of several interactions between movement and habitat covariates
+significantly improved model performance (\(\bar{r}_{s, realized} = -0.65;
+95\%-CI = [-0.67, -0.64]\).
 
@@
 \subsection{Dispersal Simulation}
-On a machine with an octa-core AMD Ryzen 7 2700X processor (8 x 3.6 GHz, 16
-logical cores) and 64 GB of RAM, a batch of 1'000 simulated dispersers moving
-over 2'000 steps required 90 minutes to compute (\(\mu = 88.90\), \(\sigma =
-1.87\)). Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
-terminated after 120 hours (i.e. five days). Comparable simulations will be
-substantially faster for smaller study areas and lower resolution covariates, as
-the covariate extraction from large and high-resolution rasters was
-computationally the most demanding task. Out of the 50'000 dispersers initiated
-inside the main source area, only 4.5\% were eventually repelled by a map
-boundary, suggesting that biases due to boundary effects should be minimal.
-
-\subsection{Convergence}
-Our examination of the traversal frequencies across checkpoints shows that we
-reached convergence across all checkpoints after 10'500 simulated individuals
-(\Cref{Convergence}a and b). Even though variability kept decreasing as we
-increased the number of simulated individuals, the marginal benefit of each
-additional disperser diminished quickly (\Cref{Convergence}c).
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\textwidth]{99_Convergence}
-    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5
-    km) distributed randomly across the study area. The relative traversal
-    frequency is plotted against the number of simulated individuals to
-    visualize how quickly the metric converges to a steady state. (a) Replicated
-    (100 times) relative traversal frequencies across three randomly chosen
-    checkpoints as well as the corresponding 95\% prediction interval (PI). (b)
-    Averaged relative traversal frequency across all checkpoints and replicates
-    including a 95\% PI. (c) Width of the PI in relation to the number of
-    simulated dispersers.}
-    \label{Convergence}
-  \end{center}
-\end{figure}
+Our dispersal simulations based on the most parsimonious movement model proved
+useful for assessing landscape connectivity. Of the 50,000 simulated dispersers
+initiated within the main study area, only 4.5\% were eventually repelled by a
+map boundary, suggesting minimal biases due to boundary effects. Moreover, our
+examination of relative traversal frequencies across all checkpoints suggested
+that connectivity reached a steady state after 10,500 simulated dispersal
+trajectories (Figure S3). Although variability in relative traversal frequencies
+kept decreasing as we increased the number of simulated dispersers, the marginal
+benefit of additional trajectories diminished quickly (Figure S3).
 
 \subsection{Heatmap}
-\Cref{Heatmap} depicts the heatmap from all simulated individuals after 2'000
-steps (additional maps in Appendix SX). The map shows that large portions of
-land beyond the borders of the KAZA-TFCA are only infrequently visited by
-dispersers (dark blue areas), whereas within the KAZA-TFCA several extensive
-regions are regularly traversed (bright yellow and red areas). Most notably, the
-region in northern Botswana south of the Linyanti swamp stands out as a highly
-frequented dispersal hotspot. Still, the presence of several massive water
-bodies, such as the Okavango Delta, the Makgadikgadi Pan, and the Linyanti
-swamp, restricts dispersal movements and limits realized connectivity within the
-KAZA-TFCA. Similarly, the heatmap shows that dispersal across Zambia's and
-Zimbabwe's part of the KAZA-TFCA is limited, which can largely be attributed to
-substantial human influences resulting from high human density, roads, and
-agricultural activities in these areas. Outside the KAZA-TFCA, the most heavily
-used regions include the areas inside the Central Kalahari National Park in
-Botswana, the area south-west of the Khaudum National Park in Namibia, and the
-area around the Liuwa Plains National Park in Zambia. Even though the heatmap
-facilitates the identification of areas that are intensely used by virtual
-dispersers, it is not straightforward to spot dispersal corridors.
-Consequently, we turn to the betweenness map.
+The heatmap (\Cref{Heatmap}), which resulted from the summation of all simulated
+dispersal trajectories, showed that several extensive regions within the
+KAZA-TFCA were frequently traversed by dispersing wild dogs (mean traversal
+frequency = 166, IQR = 274, Figure S6a), while areas beyond the KAZA-TFCA
+boundary were rarely visited (mean traversal frequency = 61, IQR = 133, Figure
+S6a). Most notably, the region in northern Botswana south of the Linyanti swamp
+appeared as highly frequented dispersal hotspot (mean traversal frequency = 987,
+IQR = 558). Nevertheless, the presence of extensive water bodies, such as the
+Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, restricted
+dispersal movements and limited realized connectivity within the KAZA-TFCA.
+Similarly, high human densities, roads, and agricultural activities in Zambia's
+and Zimbabwe's part of the KAZA-TFCA limited dispersal movements. Outside the
+KAZA-TFCA, the most heavily used regions included the areas inside the Central
+Kalahari National Park in Botswana, the area south-west of the Khaudum National
+Park in Namibia, and the area surrounding the Liuwa Plains National Park in
+Zambia. Although the heatmap facilitated the identification of areas frequently
+traversed by simulated dispersers, it seemed impractical to pinpoint dispersal
+corridors.
 
@@
   tallied all simulated trajectories. Consequently, the map highlights areas
-  that are frequently traversed by virtual dispersers. Additional heatmaps
-  showing the traversal frequency when individuals move fewer than 2'000 steps
-  are provided in Appendix S3. For spatial reference we plotted a few selected
-  national parks (dark gray).}
+  that are frequently traversed by virtual dispersers. For spatial reference we
+  plotted a few selected national parks (dark gray). Additional heatmaps showing
+  the traversal frequency when individuals move fewer than 2'000 steps are
+  provided in Figure S4.}
   \label{Heatmap}
@@
 \subsection{Betweenness}
-Betweenness scores emerging after 2'000 simulated steps are presented in
-\Cref{Betweenness} and reveal a set of discrete dispersal corridors (additional
-maps in Appendix SX). Again, the region in northern Botswana stands out as a
-central dispersal hub that connects more remote regions in the study system.
-Towards east, the extension of this corridor runs through the Chobe National
-Park into the Hwange National Park. From there, a further extension connects to
-the distant Matusadona National Park in Zimbabwe. Northwest of the Linyanty
-ecosystem, a major corridor expands into Angola, where it splits and finally
-traverses over a long stretch of unprotected area into the Kafue National Park
-in Zambia. Several additional corridors with slightly lower betweenness scores
-exist, yet most of them run within the boundaries of the KAZA-TFCA. In general,
-only a few corridors directly link the peripheral regions of the KAZA-TFCA. For
-instance, there are few corridors between the Matusadona National Park in
-Zimbabwe and the Kafue National Park in Zimbabwe. Similarly, there are no direct
-links between the Zimbabwean and Angolan ``spikes'' of the KAZA-TFCA. In
-comparison to the heatmap, the betweenness map facilitates the identification of
-linear corridors and bottlenecks between habitat patches.
+The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
+that run within the KAZA-TFCA . Again, the region in northern Botswana emerged
+as wild dog dispersal hub that connected more remote regions in the study area.
+Towards east, the extension of this corridor ran through Chobe National Park
+into Hwange National Park. From there, a further extension connected to
+Matusadona National Park in Zimbabwe. Northwest of the Linyanty ecosystem, a
+major corridor expanded into Angola, where it splitted and finally traversed
+over a long stretch of unprotected area into Zambia's Kafue National Park.
+Several additional corridors with lower betweenness scores emerged, yet most of
+them ran within the KAZA-TFCA boundaries (median betweenness inside KAZA-TFCA =
+6947\(k\), IQR = 54311\(k\), Figure S6b). In general, there were few corridors
+that directly linked the peripheral regions of the KAZA-TFCA and passed through
+unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA =
+2685\(k\), IQR = 9891\(k\), Figure S6b). Compared to the heatmap, the
+betweenness map facilitated the identification of dispersal corridors between
+habitat patches.
+
+%  9 2000  Betweenness Inside KAZA-TFCA  6946817.   54311358.
+% 10 2000  Betweenness Outside KAZA-TFCA 2685174.    9890504.
 
@@
   \citep{BastilleRousseau.2018}. Note that we square-rooted betweenness scores
-  to improve visibility of corridors with low scores.}
+  to improve visibility of corridors with low scores. Additional betweenness
+  maps showing betweenness scores when individuals move fewer than 2'000 steps
+  are provided in Figure S4.}
   \label{Betweenness}
@@
 \subsection{Inter-Patch Connectivity}
+The map of inter-patch connectivity showed that the relative frequency at which
+simulated dispersers moved from one national park to another varied, as did the
+average dispersal duration required to make these connections
+(\Cref{InterpatchConnectivity}). Overall, interpatch connectivity between
+national parks in Angola, Namibia, and Botswana appeared to be high (mean
+relative frequency = +- ) with short dispersal durations (mean duration = +-)
+
 Results from the analysis of inter-patch connectivity are given in
-\Cref{AreasReached}, which shows all realized links by simulated dispersers
-between national parks. The map furthermore indicates the relative frequency at
-which dispersers originating from one national park successfully reached another
-national park, as well as the average duration dispersers had to move to realize
-those links. It is again worth pointing out that the figure is only intended as
-an illustration, as for clarity we only considered inter-patch connectivity
-between national parks (NPs), albeit plenty of links between other protected
-areas exist. Overall, inter-patch connectivity between NPs in Angola, Namibia,
-and Botswana appears to be high, with relatively short dispersal durations
-between national parks. In contrast, we observe that connections from the
-central region into the Kafue NP in Zambia require rather long dispersal events
-and are rather infrequent. Similarly, relatively few connections lead into
-Zimbabwe's Chizarira and Matusadona NP. In some cases, we also find imbalances
-between ingoing and outgoing links, hinting at potential source-sink dynamics
-that occur due to asymmetries in disperser's willingness to cross an area,
-depending on the direction in which the area is traversed. For instance, while a
-fair share of dispersers originating from the Chizaria NP in Zimbabwe manages to
-move into the Hwange NP, there are comparably few dispersers that succeed in the
-opposite direction.
+\Cref{InterpatchConnectivity}, which shows all realized links by simulated
+dispersers between national parks. The map furthermore indicates the relative
+frequency at which dispersers originating from one national park successfully
+reached another national park, as well as the average duration dispersers had to
+move to realize those links. It is again worth pointing out that the figure is
+only intended as an illustration, as for clarity we only considered inter-patch
+connectivity between national parks (NPs), albeit plenty of links between other
+protected areas exist. Overall, inter-patch connectivity between NPs in Angola,
+Namibia, and Botswana appears to be high, with relatively short dispersal
+durations between national parks. In contrast, we observe that connections from
+the central region into the Kafue NP in Zambia require rather long dispersal
+events and are rather infrequent. Similarly, relatively few connections lead
+into Zimbabwe's Chizarira and Matusadona NP. In some cases, we also find
+imbalances between ingoing and outgoing links, hinting at potential source-sink
+dynamics that occur due to asymmetries in disperser's willingness to cross an
+area, depending on the direction in which the area is traversed. For instance,
+while a fair share of dispersers originating from the Chizaria NP in Zimbabwe
+manages to move into the Hwange NP, there are comparably few dispersers that
+succeed in the opposite direction.
 
@@
 % identified a potential dispersal corridor between Angola's NPs and the Kafue NP
-% in Zambia, \Cref{AreasReached} suggests that this link is only rarely realized
-% and requires very long dispersal events. In contrast, we find that dispersal
-% between the Moremi NP and Chobe NP are relatively frequent and require fewer
-% steps, which can be expected given that the areas are located close to each
-% other.
+% in Zambia, \Cref{InterpatchConnectivity} suggests that this link is only
+% rarely realized and requires very long dispersal events. In contrast, we find
+% that dispersal between the Moremi NP and Chobe NP are relatively frequent and
+% require fewer steps, which can be expected given that the areas are located
+% close to each other.
 
 \begin{figure}
-  \includegraphics[width=\textwidth]{99_AreasReached.png}
+  \includegraphics[width=\textwidth]{99_InterpatchConnectivity.png}
   \caption{Map of inter-patch connectivity, highlighting connections between
@@
   be restricted to national parks.}
-  \label{AreasReached}
+  \label{InterpatchConnectivity}
 \end{figure}
@@
 could also be rendered by sampling them from copula probability distributions
-\citep{Hodel.2021a}, the ISSF framework allowed us to directly incorporate them
-in the movement model. In addition, by forming interactions between habitat
-covariates and movement covariates, we could render potential dependencies
-between movement and habitat preferences. For example, our final model contained
-an interaction between water-cover and step length, showing that dispersers
-realize shorter steps in areas covered by water. Likewise, we found that
-dispersers move more tortuously across water bodies than over dryland, which is
-clearly to be expected, given that wild dogs have to wade or swim when
+\citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to directly
+incorporate them in the movement model. In addition, by forming interactions
+between habitat covariates and movement covariates, we could render potential
+dependencies between movement and habitat preferences. For example, our final
+model contained an interaction between water-cover and step length, showing that
+dispersers realize shorter steps in areas covered by water. Likewise, we found
+that dispersers move more tortuously across water bodies than over dryland,
+which is clearly to be expected, given that wild dogs have to wade or swim when
 traversing waterbodies. The ability of accompanying such effects in a single
-model is one of the great strengths of ISSFs \citep{Avgar.2016, Fieberg.2020},
+model is one of the great strengths of ISSFs \citep{Avgar.2016, Fieberg.2021},
 which is why we believe the method offers a suitable framework for simulating
@@
 Our simulation of 80'000 dispersers moving 2'000 steps across the landscapes of
-the KAZA-TFCA required five days of computation on a modern desktop machine. The
+the KAZA-TFCA required five days of computation on a modern desktop machine.
+
+On a machine with an octa-core AMD Ryzen 7 2700X processor (8 x 3.6 GHz, 16
+logical cores) and 64 GB of RAM, a batch of 1'000 simulated dispersers moving
+over 2'000 steps required 90 minutes to compute (\(\mu = 88.90\), \(\sigma =
+1.87\)). Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
+terminated after 120 hours (i.e. five days). Comparable simulations will be
+substantially faster for smaller study areas and lower resolution covariates, as
+the covariate extraction from large and high-resolution rasters was
+computationally the most demanding task. Out of the 50'000 dispersers initiated
+inside the main source area, only 4.5\% were eventually repelled by a map
+boundary, suggesting that biases due to boundary effects should be minimal.
+
+ The
 long simulation time was primarily caused by the massive extent considered (ca.
@@
 varying environmental conditions is one of the key strengths of ISSFs
-\citep{Avgar.2016, Prokopenko.2017, Fieberg.2020}, which is why we believe that
+\citep{Avgar.2016, Prokopenko.2017, Fieberg.2021}, which is why we believe that
 the framework is exceptionally well suited for simulating dispersal and
@@
 area. In theoretical applications, this issue can be circumvented by simulating
-movement on a torus \citep{Hodel.2021b}. For real data, however, alternative
+movement on a torus \citep{Hodel.2022}. For real data, however, alternative
 solutions are needed. One approach would be to terminate the simulation,

### 2021-09-24 Chapter_1/04_Manuscript/Manuscript_Support.tex
Worked on Dom's comments

@@
 
-% % Create linenumbers
-% \linenumbers
+% Create linenumbers
+\linenumbers
 
@@
 %------------------------------------------------------------------------------
-%	Appendix S1: Source Points
+%	Appendix S1: Candiate Interactions
+%------------------------------------------------------------------------------
+\newpage
+\section{Candidate Interactions}
+We started with the base model and incrementally increased model complexity by
+adding all possible two-way interactions between habitat covariates and movement
+covariates. For instance, for the covariate \textsf{Water}, we proposed the
+interactions \textsf{Water:sl}, \textsf{Water:log(sl)}, and
+\textsf{Water:cos(ta)}. Besides these interactions, we also allowed for
+correlations between turning angles and step lengths by proposing the
+interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)}. Furthermore, we
+formed the interactions \textsf{sl:LowActivity} and \textsf{log(sl):LowActivity}
+to render that step lengths are likely to be shorter during periods of
+inactivity.
+
+%------------------------------------------------------------------------------
+%	Appendix S2: K-Fold Cross Validation Procedure
+%------------------------------------------------------------------------------
+\newpage
+\section{K-Fold Cross Validation Procedure}
+We validated the predictive power of the most parsimonious movement model using
+k-fold cross-validation for case-control studies \cite{Fortin.2009}.
+Specifically, we randomly assigned 80\% of the strata to a training set and the
+remaining 20\% to a testing set. Using the training set, we parametrized a
+movement model and predicted selection scores \(w(x)\) for all steps in the
+testing set. Within each stratum, we then assigned ranks 1-25 to each step based
+on predicted selection scores, so that rank 1 was given to the step with the
+highest score \(w(x)\). Within each strata, we determined the realized step's
+rank and calculated rank frequencies of realized steps across all strata.
+Finally, we computed Spearman's rank correlation between ranks and associated
+frequencies \(r_{s, realized}\). We replicated this procedure 100 times and
+computed the mean correlation coefficient (\(\bar{r}_{s, realized}\)), as well
+as its 95\% confidence interval across all replicates. For comparison, we
+repeated the same procedure 100 times assuming random preferences. Random
+preferences were implemented by discarding the realized step from all strata and
+identifying the rank of a random step in each stratum. Again, we calculated
+Spearman's rank correlation coefficient (\(r_{s, random}\)), its mean across
+repetitions (\(\bar{r}_{s, random}\)), and its 95\% confidence interval.
+Ultimately, this validation proves a significant prediction in case the
+confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do
+not overlap \citep{Fortin.2009}.
+
+%------------------------------------------------------------------------------
+%	Appendix S3: Source Points
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S2: All Movement Models
+%	Appendix S4: All Movement Models
 %------------------------------------------------------------------------------
@@
       \\
-          \textbf{Base Model}: cos(ta) + sl + log(sl) + WA + WO + DTW + HI + SH & 90091.40 & 787.67 & 0.00 & -45030.70\\
+      \textbf{Base Model}: cos(ta) + sl + log(sl) + WA + WO + DTW + HI + SH                                                                & 90091.40 & 787.67        & 0.00   & -45030.70 \\
       \\
@@
 %------------------------------------------------------------------------------
-%	Appendix S3: Movement Model interpretation
+%	Appendix S5: Most Parsimonious Movement Models
+%------------------------------------------------------------------------------
+\newpage
+\section{Movement Model}
+\begin{table}[htbp]
+ \begin{center}
+  \caption{Most parsimonious movement model for dispersing wild dogs. The model
+   consists of a movement kernel, a habitat kernel, and their interactions. The
+   movement kernel describes preferences with regards to movement behavior,
+   whereas the habitat kernel describes preferences with respect to habitat
+   conditions. Interactions between the two kernels indicate that movement
+   preferences are contingent on habitat conditions. Note that all covariates
+   were standardized to a mean of zero and standard deviation of 1. Plots to aid
+   with the interpretation of this model are given in Appendix S2.}
+  \label{MovementModelNumbers}
+  \resizebox{\textwidth}{!} {
+   \begin{threeparttable}
+    \begin{tabular}{llrrrc}
+     \toprule
+     Kernel & Covariate                                       & Coefficient & SE    & p-value     & Sign. \\
+     \midrule
+     \multirow{5}{*}{Habitat Kernel}
+            & Water                                           & -0.546      & 0.112 & \(<\) 0.001 & ***   \\
+            & DistanceToWater \textsuperscript{0.5}           & -0.390      & 0.231 & 0.092       & *     \\
+            & Woodland                                        & -0.364      & 0.086 & \(<\) 0.001 & ***   \\
+            & Shrubs/Grassland                                & 0.288       & 0.092 & 0.002       & ***   \\
+            & HumanInfluence                                  & -0.535      & 0.229 & 0.019       & **    \\
+     \hdashline
+     \multirow{6}{*}{Movement Kernel}
+            & sl                                              & 0.075       & 0.037 & 0.042       & **    \\
+            & cos(ta)                                         & 0.105       & 0.031 & 0.001       & ***   \\
+            & log(sl)                                         & 0.146       & 0.051 & 0.004       & ***   \\
+            & cos(ta) : sl                                    & 0.049       & 0.026 & 0.064       & *     \\
+            & cos(ta) : log(sl)                               & 0.076       & 0.026 & 0.003       & ***   \\
+            & sl : LowActivity                                & -0.917      & 0.113 & \(<\) 0.001 & ***   \\
+     \hdashline
+     \multirow{5}{*}{Interactions}
+            & sl : Water                                      & -0.305      & 0.076 & \(<\) 0.001 & ***   \\
+            & sl : Woodland                                   & -0.089      & 0.039 & 0.023       & **    \\
+            & sl : Shrubs/Grassland                           & 0.124       & 0.058 & 0.032       & **    \\
+            & sl : DistanceToWater \textsuperscript{0.5}      & -0.058      & 0.031 & 0.056       & *     \\
+            & cos(ta) : HumanInfluence                        & -0.040      & 0.022 & 0.070       & *     \\
+            & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063       & 0.026 & 0.017       & **    \\
+     \bottomrule
+    \end{tabular}
+    \begin{tablenotes}
+     \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
+      \quad *** \(p < 0.01\)}
+    \end{tablenotes}
+   \end{threeparttable}
+  }
+ \end{center}
+\end{table}
+
+%------------------------------------------------------------------------------
+%	Appendix S6: Movement Model interpretation
 %------------------------------------------------------------------------------
@@
 To ease with the interpretation of the most parsimonious movement model, we
-followed recommendations published in \cite{Fieberg.2020} and produced a series
+followed recommendations published in \cite{Fieberg.2021} and produced a series
 of plots highlighting how the habitat and movement kernel depended on covariate
@@
    parsimonious movement model from Table S1. The plots were generated
-    following recommendations reported in \cite{Fieberg.2020}. Subplots
+   following recommendations reported in \cite{Fieberg.2021}. Subplots
    \textbf{a1} to \textbf{a8} highlight dispersing wild dogs' movement kernel
@@
 %------------------------------------------------------------------------------
-%	Appendix S4: Heatmaps
+%	Appendix S7: Convergence
+%------------------------------------------------------------------------------
+\newpage
+\section{Convergence}
+\begin{figure}[htbp]
+  \begin{center}
+    \includegraphics[width=\textwidth]{99_Convergence}
+    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5
+    km) distributed randomly across the study area. The relative traversal
+    frequency is plotted against the number of simulated individuals to
+    visualize how quickly the metric converges to a steady state. (a) Replicated
+    (100 times) relative traversal frequencies across three randomly chosen
+    checkpoints as well as the corresponding 95\% prediction interval (PI). (b)
+    Averaged relative traversal frequency across all checkpoints and replicates
+    including a 95\% PI. (c) Width of the PI in relation to the number of
+    simulated dispersers.}
+    \label{Convergence}
+  \end{center}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%	Appendix S8: Heatmaps
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S5: Evolution of Betweenness Maps
+%	Appendix S9: Evolution of Betweenness Maps
 %------------------------------------------------------------------------------
@@
 
+%------------------------------------------------------------------------------
+%	Appendix S10: Map Comparison
+%------------------------------------------------------------------------------
+\newpage
+\section{Comparison of Values Inside and Outside of KAZA-TFCA}
+\begin{figure}[hbtp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_MapComparison.png}
+  \caption{Comparison of values from the heatmap and betweenness map inside
+  (blue) and outside (orange) the KAZA-TFCA borders.}
+  \label{MapComparison}
+ \end{center}
+\end{figure}
+
 \newpage

### 2021-09-24 Chapter_1/02_R-Scripts/02_Analysis/99_BacktransformModel.R

@@ -1,514 +1,0 @@
-################################################################################
-#### Backtransform Scaled Model
-################################################################################
-# Description: Here, I want to backtransform the scaled model to retrieve
-# coefficients for the movement metrics on the untransformed scale. However,
-# this script is only supposed to make sure the function that I use to
-# backtansform the data actually works. I can verify this by fitting an unscaled
-# model and then compare the coefficients to a model with scaled covariates that
-# I backtransform.
-
-# Clear R's brain
-rm(list = ls())
-
-# Surpress scientific notation
-options(scipen = 999)
-
-# Change the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)    # For data wrangling
-library(davidoff)     # Custom functions
-library(glmmTMB)      # For modelling
-
-################################################################################
-#### Loading Data
-################################################################################
-# Load step selection data
-dat <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF_Extracted).csv")
-
-# Keep only desired columns
-dat <- dat %>% select(c(
-      id
-    , step_id_
-    , case_
-    , sl_
-    , ta_
-    , inactive
-    , Water
-    , DistanceToWater
-    , Trees
-    , Shrubs
-    , HumansBuff5000 = HumanInfluenceBuffer_5000
-))
-
-# This is a crucial step. In the original model, step lengths were included in
-# meters. This resulted in convergence issues, which I overcame by standardizing
-# also the movement metrics. Alternatively, I could have also converted step
-# lengths to kilometers, yet I realized that the model fit was better using
-# standardized step lengths. Here, I will convert the step lengths to kilometers
-# which will allow me to refit the full model and see if my
[…]

### 2021-09-24 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/MapComparisons.R

@@ -1,0 +1,141 @@
+################################################################################
+#### Map Comparisons
+################################################################################
+# Description: In this script, we compare connectivity metrics inside and
+# outside the KAZA-TFCA and produce a nice plot to visualize the results
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(raster)       # For general raster manipulation
+library(rgdal)        # To read spatial data
+library(tidyverse)    # For data wrangling
+library(viridis)      # For nicer colors
+library(ggpubr)       # To arrange multiple plots
+
+# Reload connectivity maps
+heatmaps <- read_rds("03_Data/03_Results/99_Heatmaps.rds")
+betweenness <- stack("03_Data/03_Results/99_Betweenness.grd")
+
+# Apply focal filter to buffer/smooth maps
+betweenness <- lapply(1:nlayers(betweenness), function(x){
+  focal(betweenness[[x]], w = matrix(1, 3, 3), fun = mean)
+}) %>% stack()
+
+# Also, we want to square root
+betweenness <- sqrt(betweenness)
+
+# Subset to steps of interest
+heatmaps <- subset(heatmaps, steps %in% c(125, 500, 2000))
+
+# We also need to add the main area and the buffer
+heatmaps <- lapply(unique(heatmaps$steps), function(x) {
+  sub <- subset(heatmaps, steps == x)
+  ras <- stack(sub$heatmap)
+  ras <- sum(ras)
+  result <- tibble(Steps = x, Heatmap = list(ras))
+  return(result)
+}) %>% do.call(rbind, .)
+
+# Convert the betweenness maps into a tibble
+betweenness <- tibble(
+    Steps       = c(125, 500, 2000)
+  , Betweenness = list(betweenness[[1]], betweenness[[2]], betweenness[[3]])
+)
+
+# Put the maps together
+maps <- bind_cols(heatmaps, betweenness[, 2]) %>%
+  pivot_longer(cols = 2:3, names_to = "Type", values_to = "Map")
+
+################################################################################
+#### Median Values Inside and Outside KAZA
[…]

### 2021-09-24 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/StudyArea.R

@@ -172,9 +172,9 @@ p1 <- tm_shape(hill_africa) +
     ) +
   tm_shape(kaza_ext) +
     tm_borders(
-        col = "black"
+        col = "blue"
       , lty = 1
-      , lwd = 1
+      , lwd = 0.5
     ) +
   # tm_shape(lines_countries) +
   #   tm_lines(
@@ -259,7 +259,8 @@ p2 <- tm_shape(hill_kaza) +
     , ticks               = TRUE
   ) +
   tm_layout(
-    , frame                   = "gray20"
+    # , frame                   = "gray20"
+    , frame                   = "blue"
     , frame.lwd               = 1
     , asp                     = 1.2
     , legend.outside          = TRUE

### 2021-09-25 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

@@
 The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
-that run within the KAZA-TFCA . Again, the region in northern Botswana emerged
+that run within the KAZA-TFCA. Again, the region in northern Botswana emerged
 as wild dog dispersal hub that connected more remote regions in the study area.
@@
 that directly linked the peripheral regions of the KAZA-TFCA and passed through
-unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA =
+unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA-TFCA =
 2685\(k\), IQR = 9891\(k\), Figure S6b). Compared to the heatmap, the
@@
 average dispersal duration required to make these connections
-(\Cref{InterpatchConnectivity}). Overall, interpatch connectivity between
-national parks in Angola, Namibia, and Botswana appeared to be high (mean
-relative frequency = +- ) with short dispersal durations (mean duration = +-)
+(\Cref{InterpatchConnectivity}).
+
+Overall, interpatch connectivity between national parks in Angola, Namibia,
+Botswana, and Zimbabwe appeared to be high with only short required dispersal
+durations. For example, xx\% of simulated individuals departing from a national
+park in xx managed to move into a national park in xx, with average dispersal
+durations of xx\% steps. Conversely, we found that connectivity from and into
+national parks in Zambia were limited. Only xx\% of the dispersers leaving from
+Zambia managed to move into a national park in another country and for those
+connections xx steps were required on average. In some cases, we also detected
+imbalances between the number of ingoing and outgoing links. Make an example...
+
+ suggesting the presence of source-sink dynamics that occure due to asymmetric behavior depending on the direction in which a certain dispersal route is taken.
 

### 2021-09-25 Chapter_1/04_Manuscript/Manuscript_Support.tex
Various updates

@@
 \restoregeometry
+%------------------------------------------------------------------------------
+%	Appendix S11: Areas Reached
+%------------------------------------------------------------------------------
+\newpage
+\section{Dispersal into other National Parks}
+\begin{figure}[hbtp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_AreasReached.png}
+  \caption{Relative number of simulated dispersal trajectories that successfully
+  moved from one national park into another that is at least as far away as
+  indicated on the x-axis. Percentages are given in relation to the number of
+  simulated individuals from the national parks in the respective countries. For
+  example, over 85\% of all individuals originating from a national park in
+  Angola moved from their natal national park into another one. However, the
+  percentage gradually decreases as only national parks at higher euclidean
+  distances are considered.}
+  \label{AreasReached}
+ \end{center}
+\end{figure}
 
 %------------------------------------------------------------------------------
-%	Appendix S10: Map Comparison
+%	Appendix S11: Map Comparison
 %------------------------------------------------------------------------------
 \newpage
-\section{Comparison of Values Inside and Outside of KAZA-TFCA}
+\section{Comparison of Traversal Frequencies and Betweenness Scores Inside and Outside KAZA-TFCA}
 \begin{figure}[hbtp]

### 2021-12-13 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

--- 2021-09-24 Chapter_1/04_Manuscript/Manuscript.tex
@@
 \begin{abstract}
-
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-% The ability to disperse is contingent a sufficient degree of landscape
-% connectivity, which is why the identification and preservation of movement
-% corridors that promote connectivity has become a task of extraordinary
-% importance. Currently, ecologists mainly rely on least-cost analysis or circuit
-% theory to investigate connectivity, albeit both methods make several assumptions
-% are hardly met in reality. To address these issues, simulations from
-% individual-based movement models have been proposed, yet a unified framework to
-% simulate dispersal and quantify connectivity is lacking.
-%
-% Here, we employ a simple three-step workflow that combines several
-% already-existing methods to assess connectivity using simulated dispersal
-% trajectories. In the first step, we use integrated step selection functions to
-% parametrize a mechanistic movement model rendering dispersal behavior. In a
-% second step, we apply the parametrized model as an individual-based movement
-% model to simulate dispersal trajectories. Finally, in the third step, we
-% consolidate simulated trajectories into three complementary connectivity maps,
-% each focusing on a different aspect of landscape connectivity.
-%
-% We apply these steps to assess landscape connectivity for dispersing African
-% wild dogs (\textit{Lycaon pictus}) southern Africa. We thereby demonstrate that
-% simulations from integrated step selection functions offer a simple, yet
-% powerful alternative to traditional connectivity modeling techniques.
-
+\textbf{Short Version }
+The ability to disperse is contingent on a sufficient degree of landscape
+connectivity, which is why the identification and preservation of movement
+corridors that promote connectivity has become a task of extraordinary
+importance. Currently, ecologists rely on least-cost analysis or circuit
+theory to investigate connectivity, albeit both methods make several
+assumptions that are hardly met in reality. To address these issues,
+simulations from individual-based movement models have been proposed, yet a
+unified framework to simulate dispersal and quantify connectivity is lacking.
+
+Here, we propose a simple three-step workflow that combines several
+already-existing methods to assess connectivity using simulated dispersal
+trajectories. In step one, we use integrated step selection functions to
+parametrize a mechanistic movement model rendering dispersal behavior. In step
+two, we apply the parametrized model as an individual-based movement model to
+simulate dispersal trajectories. In step three, we combine simulated
+trajectories into three complementary connectivity maps, each focusing on a
+different aspect of landscape connectivity.
+
+We showcase the application of the proposed workflow using data of dispersing
+African wild dogs (\textit{Lycaon pictus}) and assess landscape connectivity
+within the world's largest transboundary conservation area in Southern Africa.
+We thereby shed light into dispersing wild dogs' habitat and movement
+preferences, while also uncovering crucial dispersal corridors. With this
+analysis, we demonstrate that simulations from integrated step selection
+functions offer a simple, yet powerful alternative to traditional connectivity
+modeling techniques.
+\end{abstract}
+
+\newpage
+\begin{abstract}
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-% Importance of dispersal & connectivity and mow its measured
+\textbf{Long Version }
 Dispersal of individuals is crucial for long-term species persistence and
-depends on a sufficient degree of landscape connectivity. Thusfar, researchers
-have investigated connectivity primarily using least-cost analysis and circuit
-theory, although both methods make assumptions that are rarely met in reality.
-Least-cost analysis assumes that animals move towards a known endpoint and are
-knowledgeable about the least costly route to reach it. Circuit theory relies on
-the assumption of a complete random walk without any directional persistence
-along the dispersing route. Albeit these shortcomings can be overcome by
-spatio-temporally explicitly simulating dispersing trajectories across the
-landscape, a unified approach for such a simulation is lacking.
-
-% What we do
+depends on a sufficient degree of landscape connectivity. To date, researchers
+primarily investigate connectivity using least-cost analysis and circuit theory,
+both methods that make assumptions which are hardly met in reality. Least-cost
+analysis assumes that animals move towards a known endpoint and are
+knowledgeable about the most favorable route to reach it. Circuit theory
+presumes a complete random walk and therefore fails to incorporate directional
+persistence. These shortcomings can be overcome by spatio-temporally explicitly
+simulating dispersal, yet a unified approach for such simulations is lacking.
+
 Here, we present a simple three-step workflow to simulate dispersal and assess
-connectivity starting from empirical GPS movement data. To exemplify its
-application, we employ the workflow to dispersal data collected on the
-endangered African wild dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi
+connectivity starting from empirical GPS movement data. In step one, we use
+integrated step selection functions to fit a mechanistic movement model
+describing habitat and movement preferences of dispersers. In step two, we apply
+the parameterized model to simulate individual dispersal trajectories. In step
+three, we combine the simulated trajectories into three complementary
+connectivity maps: a heatmap that highlights frequently traversed areas, a
+betweenness map that pinpoints dispersal corridors, and a map of inter-patch
+connectivity that indicates the presence and intensity of functional links
+between habitat patches. As a case study, we analyse data on the endangered
+African wild dog and assess landscape connectivity in the Kavango-Zambezi
 Transfrontier Conservation Area, the world’s largest transboundary conservation
-area. In step one, we use integrated step selection functions to fit a
-mechanistic movement model describing habitat and movement preferences of
-dispersers. In step two, we apply the parameterized model to simulate individual
-dispersal trajectories. In step three, we compress the simulated trajectories
-into three complementary connectivity maps: a heatmap that highlights frequently
-traversed areas, a betweenness map that pinpoints dispersal corridors, and a map
-of interpatch connectivity, indicating the presence or absence of functional
-links between habitat patches.
-
-% Some results
-Our results show that dispersing African wild dogs move with directional
-persistence in the vicinity to waterbodies and preferrably through areas with
-little human influence. This is reflected in the simulation with resulted in
-several dispersal hotspots and corridors along waterbodies through areas with
-little or no anthropogenic pressure. Connectivity between national parks inside
-the KAZA-TFCA appears to be good, albeit the parks in Zambia and Zimbabwe are
-comparably less well connected to the rest of KAZA-TFCA, suggesting the
-necessity for additional protected areas that serve as ``stepping stones'' into
-these regions.
-
-% Case study
+area.
+
+Our results show that wild dogs preferably disperse with directional persistence
+in the vicinity to water bodies and through areas with little human influence and
+sparse woodland cover. Dispersal simulations and subsequent connectivity maps
+revealed several dispersal hotspots and corridors across the extent of the
+KAZA-TFCA. Connectivity between NPs inside the KAZA-TFCA was good, yet few
+dispersers successfully moved from Zambia's NPs into neighboring areas.
+
 We show that simulations from step-selection functions offer a simple yet
-powerful alternative to traditional connectivity modeling techniques, although
-necessitating informed decisions about the number of simulated individuals and
-the duration of simulated dispersal events. Ultimately, our workflow permits a
-more mechanistic understanding of dispersal behaviour and landscape connectivity
-making it useful for a variety of applications in ecological, evolutionary, and
-conservation research.
+powerful alternative to traditional connectivity modeling techniques. In
+contrast to traditional connect modeling techniques, simulations not only make
+fewer unrealistic assumptions but also permit a more mechanistic understanding
+of dispersal and landscape connectivity. This makes our workflow useful for a
+variety of applications in ecological, evolutionary, and conservation research.
 
@@
 Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and colonize
-unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
-to disperse depends on a sufficient degree of landscape connectivity
+unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. Importantly, the
+ability to disperse depends on a sufficient degree of landscape connectivity
 \citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
@@
 why simulating dispersal from individual-based movement models provides a
-promising alternative approach to modeling connectivity \citep{Diniz.2019}.
+promising alternative approach to investigate connectivity \citep{Diniz.2019}.
 
@@
 migrating animals, it is unlikely to hold for dispersers, as dispersers
-typically move across unfamiliar territory and are therefore less aware of
-associated movement costs and potential endpoints \citep{Koen.2014, Cozzi.2020}.
-CT, on the contrary, posits that animals move according to a random walk,
-entailing that autocorrelation between subsequent movements cannot be rendered
+typically move across unfamiliar territory, so the associated movement costs and
+potential endpoints are unknown to them \citep{Koen.2014, Cozzi.2020}. CT, on
+the contrary, posits that animals move according to a random walk, entailing
+that autocorrelation between subsequent movements cannot be rendered
 \citep{Diniz.2019}. For dispersers, however, autocorrelated movements are
-regularly observed \citep{Cozzi.2020, Hofmann.2021}, implying that subsequent
-relocations tend to fall onto a straight line. Moreover, graph-based methods are
-unable to reflect the temporal dimension of dispersal, meaning that statements
-about the approximate duration required to move between habitats impossible
+regularly observed \citep{Cozzi.2020, Hofmann.2021}, implying that dispersal
+trajectories are usually strongly directional. Moreover, graph-based methods are
+unable to reflect the temporal dimension of dispersal, so that statements about
+the expected duration for moving between habitats are impossible
 \citep{Martensen.2017, Diniz.2019}.
@@
 The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
-trajectories from individual-based movement models (IBMMs) and by converting
-simulated trajectories into meaningful measures of connectivity
-\citep{Diniz.2019}). IBMMs allow to explicitly simulate how individuals move
-across and interact with the encountered landscape \citep{Kanagaraj.2013,
-Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020} and to render potential
-interactions between movement behavior and habitat conditions. This shifts the
-focus from a structural to a more functional view on connectivity
-\citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}. Furthermore,
-simulations from IBMMs generate movement sequentially, making the temporal
-dimension of dispersal explicit and allowing to model autocorrelation between
-subsequent movements \citep{Diniz.2019}. Finally, simulations from IBMMs do not
-enforce connections towards preconceived endpoints, thereby mitigating biases
-arising from misplaced endpoints. Despite these advantages, a unified framework
-to simulate dispersal and assess connectivity using IBMMs is lacking.
+from individual-based movement models (IBMMs) and converting simulated
+trajectories into meaningful measures of connectivity \citep{Diniz.2019}). In
+contrast to graph-based methods, IBMMs allow to explicitly simulate how
+individuals move across and interact with the encountered landscape
+\citep{Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020}, as
+well as to render potential interactions between movement behavior and habitat
+conditions, thus shifting the focus from a structural to a more functional view
+on connectivity \citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}.
+Furthermore, because simulations from IBMMs generate movement sequentially, the
+temporal dimension of dispersal becomes explicit and allows to model
+autocorrelation between subsequent movements \citep{Diniz.2019}. Finally,
+simulations from IBMMs do not enforce connections towards preconceived
+endpoints, thereby mitigating biases arising from misplaced endpoints. Despite
+these advantages, a unified framework to simulate dispersal and assess
+connectivity using IBMMs is lacking.
 
@@
 \subsection{Proposed Solution: Three-Step Workflow}
-Here, we draw on recent advancements in the fields of movement ecology and
-network theory to propose a simple three-step workflow for simulating dispersal
-movements and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
-step one, we combine GPS movement data of dispersing individuals with relevant
-habitat covariates to fit a mechanistic movement model using integrated step
-selection functions (ISSFs, \citealp{Avgar.2016}). ISSFs permit
-simultaneous inference on the focal species' habitat kernel (i.e. habitat
-preferences), its movement kernel (i.e. movement preferences/capabilities), and
-potential interactions between the two \citep{Avgar.2016, Fieberg.2021}. In step
-two,  we simulate individual dispersal trajectories using the movement model
-parametrized in step one. Finally, in step three, we convert simulated
-trajectories into three complementary connectivity maps, each highlighting a
-different aspect of connectivity. The proposed set of maps includes (i) a
+Here, we draw on recent advancements in movement ecology and propose a simple
+three-step workflow for simulating dispersal and assessing landscape
+connectivity (\Cref{GraphicalAbstract}). In step one, we combine GPS movement
+data of dispersing individuals with relevant habitat covariates to fit a
+mechanistic movement model using integrated step selection functions (ISSFs,
+\citealp{Avgar.2016}). ISSFs permit inference on the focal species' habitat
+kernel (i.e. habitat preferences), its movement kernel (i.e. movement
+preferences/capabilities), and potential interactions between the two
+\citep{Avgar.2016, Fieberg.2021}. In step two, we simulate individual dispersal
+trajectories using the parametrized movement from step one. Comparable
+simulations have already been employed to estimate steady-state utilization
+distributions of residents \citep{Potts.2013, Signer.2017} and to model
+landscape connectivity, yet disregarding interdependencies between habitat and
+movement kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we
+convert simulated trajectories into three complementary connectivity maps, each
+highlighting a different aspect of connectivity. The set of maps includes (i) a
 heatmap revealing areas that are frequently traversed by simulated dispersers
@@
 \citealp{BastilleRousseau.2018}), (iii) and a map of inter-patch connectivity,
-depicting the presence or absence and intensity of use of specific connections,
-and the average dispersal duration between corresponding habitat patches (e.g.
-\citealp{Kanagaraj.2013}).
+depicting the presence and intensity of use of specific connections, as well as
+the average dispersal duration required to connect corresponding habitat patches
+(e.g. \citealp{Gustafson.1996, Kanagaraj.2013}).
 
 % Introduction of the study species
-\subsection{Introduction of the Study Species}
-The endangered African wild dog (\textit{Lycaon pictus}) represents a highly
-mobile species whose persistence hinges on a  sufficient degree of landscape
+\subsection{Case Study}
+We showcase the application of the proposed workflow (\Cref{GraphicalAbstract})
+using dispersal data collected on the endangered African wild dog
+(\textit{Lycaon Pictus}). The African wild dog represents a highly mobile
+species whose persistence hinges on a sufficient degree of landscape
 connectivity. Once common throughout sub-Saharan Africa, this species has
@@
 habitat fragmentation, and disease outbreaks \citep{Woodroffe.2012}. Wild dogs
-typically emigrate in single-sex coalitions \citep{McNutt.1996, Behr.2020}
-capable of dispersing several hundred kilometers \citep{DaviesMostert.2012,
+typically disperse in single-sex coalitions \citep{McNutt.1996, Behr.2020} and
+are capable of dispersing several hundred kilometers \citep{DaviesMostert.2012,
 Masenga.2016, Cozzi.2020}. Although previous studies have investigated
@@
 information on landscape connectivity is essential for the conservation of this
-endangered carnivore.
-
-% % Some Background on SSFs
-% \subsection{Some Background on SSFs}
-% To fit a movement model, we propose to employ the framework of step selection
-% functions (SSFs; \citealp{Fortin.2005, Thurfjell.2014, Avgar.2016}). Originally,
-% SSFs were developed to learn about habitat preferences of the focal species
-% using GPS data \citep{Fortin.2005}. The method works by comparing covariates
-% along realized steps (the straight-line connecting two consecutive GPS
-% relocations; \citealp{Turchin.1998}) to the same covariates along randomly
-% generated alternative steps \citep{Fortin.2005, Thurfjell.2014} using
-% conditional logistic regression \citep{Fortin.2005, Muff.2020}. Recently,
-% \cite{Avgar.2016} have introduced \textit{integrated} SSFs (ISSFs), which is a
-% generalization of regular SSFs and requires that movement metrics are included
-% in the corresponding regression model. This not only reduces biases in estimated
-% preferences, but also permits simultaneous inference on a species' habitat
-% kernel (i.e. habitat preferences), its movement kernel (i.e. movement
-% preferences/capabilities), and potential interactions between the two kernels
-% \citep{Avgar.2016, Fieberg.2021}. Importantly, a model that was fitted using
-% ISSFs resembles a fully mechanistic IBMM, based on which movement can be
-% simulated \citep{Avgar.2016, Signer.2017}. Previously, ISSFs have been applied
-% to simulate utilization distributions of resident animals \citep{Avgar.2016,
-% Signer.2017}, yet the method remains underutilized for the simulation of
-% dispersal and detection of movement corridors.
-
-% % Some Background on Connectivity "Metrics"
-% \subsection{Some Background on  Connectivity ``Metrics'' }
-% Irrespective of the chosen modeling framework, simulated dispersal trajectories
-% need to be processed into measures of connectivity. To date, the most common
-% practice to translate simulations into connectivity measures is to compute
-% inter-patch connectivity between distinct habitats \citep{Diniz.2019}.
-% Inter-patch connectivity is a simple ratio between the number of dispersers
-% successfully moving between habitat patches and the total number of simulated
-% individuals from each patch (e.g. \citealp{Kanagaraj.2013}). As an alternative,
-% some researchers have tallied overlapping trajectories into what might be called
-% ``heatmaps'', i.e. maps that depict the frequency at which each spatial unit in
-% the study area was traversed by simulated dispersers (e.g. \citealp{Peer.2008,
-% Hauenstein.2019, Zeller.2020}). More recently, \cite{BastilleRousseau.2018}
-% proposed a method to convert trajectories into a network, so that metrics from
-% network theory can be computed. With regards to connectivity, the betweenness
-% metric seems promising, as it indicates the number of shortest paths going
-% through each node in the network and therefore highlights areas of exceptional
-% importance for connecting different regions in the study area. Because each of
-% the three metrics metrics focuses on a very different aspect of landscape
-% connectivity, we propose to view them as complementary and to use them in
-% concert when assessing connectivity.
+endangered carnivore. Here, we use GPS data of 16 dispersing wild dogs
+originating from a free-ranging population in northern Botswana to parametrize a
+mechanistic movement model, which we then employed to simulate 80,000 dispersal
+trajectories across the landscape of the world's largest transboundary
+conservation area, the Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA). We anticipated that simulations based on our three-step workflow
+would overcome several of the above highlighted conceptual shortcomings of
+traditional connectivity models and provide a more comprehensive view of
+landscape connectivity
 
@@
     \caption{Flowchart of the simulation-based connectivity analysis as proposed
-    in this article. First, GPS data and habitat covariates have to be
+    in this article. First, GPS data and habitat covariates must be
     collected. The combined data is then analyzed in an integrated step
@@
 
-% Case Study
-\subsection{Case Study}
-In this study, we illustrate the proposed three-step workflow
-(\Cref{GraphicalAbstract}) to assess landscape connectivity for African wild
-dogs within the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), the
-world's largest transboundary conservation area. To this end, we use GPS
-movement data of 16 dispersing wild dogs originating from a free-ranging
-population in northern Botswana to parametrize a mechanstic movement model,
-which we then employed to simulate 80,000 dispersal trajectories across the
-landscape of the KAZA-TFCA. We anticipated that simulations based on our
-three-step workflow would overcome several of the above highlighted conceptual
-shortcomings of traditional connectivity models and provide a more comprehensive
-view of landscape connectivity.
-
 \section{Methods}
@@
 The study area centered at -17\degree 13'9''S, 23\degree 56'4''E
-(\Cref{StudyArea}a) is located in southern Africa (\Cref{StudyArea}a) and spans
-more than 1.3 Mio. km\textsuperscript{2}, encompassing the entire KAZA-TFCA
-(\Cref{StudyArea}b). The KAZA-TFCA comprises parts of Angola, Botswana, Namibia,
-Zimbabwe, and Zambia and hosts diverse landscapes; ranging from savannah to
-grassland and from dry to moist woodland habitats. In its center lies the
-Okavango Delta, a dominant hydrogeographical feature and the world's largest
-flood-pulsing inland delta. Although large portions of the KAZA-TFCA are part of
-designated national parks and other protected areas, considerable human
-influence from roads, agricultural sites, and settlements remains.
+(\Cref{StudyArea}a) was located in southern Africa (\Cref{StudyArea}a) and
+spanned more than 1.3 Mio. km\textsuperscript{2}, encompassing the entire
+KAZA-TFCA (\Cref{StudyArea}b). The KAZA-TFCA comprises parts of Angola,
+Botswana, Namibia, Zimbabwe, and Zambia and hosts a rich diversity of
+landscapes, ranging from savannah to grassland and from dry to moist woodland
+habitats. In its center lies the Okavango Delta, a dominant hydro-geographical
+feature and the world's largest flood-pulsing inland delta. Large portions of
+the KAZA-TFCA are part of designated national parks (NPs) and other protected
+areas, yet considerable human influence through roads, agricultural sites, and
+settlements remains.
 
@@
     \caption{Illustration of the study area located in southern Africa. (a) The
-    study area was confined by a bounding box spanning the entire KAZA-TFCA and
-    comprised parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. (b)
-    The KAZA-TFCA represents the world's largest terrestrial conservation area
-    and covers a total of 520'000 km\textsuperscript{2}. Its main purpose is to
-    re-establish connectivity between already-existing national parks (dark
-    green) and other protected areas (light green).}
+    study area was confined by a bounding box spanning the entire KAZA-TFCA
+    which comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia.
+    (b) The KAZA-TFCA currently represents the world's largest terrestrial
+    transfrontier conservation area, covering a total area of 520'000
+    km\textsuperscript{2}. Its main purpose is to re-establish connectivity
+    between already-existing NPs (dark green) and other protected areas (light
+    green).}
     \label{StudyArea}
@@
 landscape connectivity than behavior during residence \citep{Elliot.2014,
-Abrahms.2017}, we discarded data collected when individauls were resident. We
+Abrahms.2017}, we discarded data collected when individuals were resident. We
 determined the exact timing of emigration and settlement using direct field
@@
 dispersers' natal home range. Thus, dispersal was deemed to have started once an
-individual left its natal home range and ended once the NSD metric ramined
+individual left its natal home range and ended once the NSD metric remained
 constant, indicating settlement. During dispersal, GPS collars recorded a fix
-every 4 hours and they regularly transmitted data over the Iridium satellite
-system. To ensure regular timespans between fixes, we removed any fix that was
-not obtained on the aspired 4-hourly schedule (\( \pm \) 15 minutes). We then
-converted the remaining fixes (n = 4'169) into steps, where each step
-represented the straight-line movement between two consecutive GPS relocations
-\citep{Turchin.1998}.
+every 4 hours and regularly transmitted data over the Iridium satellite system.
+To ensure regular timespans between GPS fixes, we removed any fix that was not
+successfully obtained on the aspired 4-hourly schedule (allowing for a tolerance
+of \( \pm \) 15 minutes). We then converted the remaining fixes (n = 4'169) into
+steps, where each step represented the straight-line movement between two
+consecutive GPS fixes \citep{Turchin.1998}.
 
@@
 covariates \textit{water-cover, distance-to-water, woodland-cover,
-shrub/grassland-cover, and human-influence}. To reflect the seasonality of water
-in the study area, we obtained weekly updated layers for water-cover and
-corresponding layers for distance-to-water from MODIS satellite imagery using a
-remote sensing algorithm \citealp{Wolski.2017, Hofmann.2021} that is implemented
-in the \textsf{floodmapr} package which is available on GitHub
-(\url{https://github.com/DavidDHofmann/floodmapr}). For each step, we thus used
-the temporally associated water layers, reflecting prevailing conditions at the
-time when the step was realized. To ensure a consistent resolution across
-habitat covariates, we coarsened or interpolated all layers to a resolution of
-250 m x 250 m. A more detailed description of how we prepared each habitat
-covariate is provided in \cite{Hofmann.2021}.
+shrub/grassland-cover, and human-influence}. To correctly render seasonality in
+water-cover, we applied a remote sensing algorithm that enabled us to obtain
+weekly updated raster-layers for water-cover and corresponding layers for
+distance-to-water from MODIS satellite imagery \citealp{Wolski.2017,
+Hofmann.2021}. This algorithm is now implemented in the \textsf{floodmapr}
+package which is available on GitHub
+(\url{https://github.com/DavidDHofmann/floodmapr}). Using the frequently updated
+floodmpas we were able to correctly assign the most appropriate covariate layer
+to each observed movement step. To ensure a consistent resolution across habitat
+covariates, we coarsened or interpolated all layers to a resolution of 250 m x
+250 m. A detailed description of how we prepared each habitat covariate is
+provided in \cite{Hofmann.2021}.
 
 Besides habitat covariates, we computed movement metrics that we used as
-movement covariates in our ISSF analysis \citep{Avgar.2016, Fieberg.2021}.
-Movement metrics were calculated for each step and included the step length
+movement covariates in ISSF models \citep{Avgar.2016, Fieberg.2021}. Movement
+metrics were calculated for each step and included the step length
 (\textsf{sl}), its natural logarithm (\textsf{log(sl)}), and the cosine of the
@@
 imported into {\tt R} using the {\tt Rcpp} package \citep{Eddelbuettel.2011,
-Eddelbuettel.2013}. We provide annotated R-codes to reproduce the three-step
-workflow using simulated data through GitHub
-(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, codes
-for the African wild dog case study are available through an online repository.
+Eddelbuettel.2013}.
 
@@
 dogs \citep{Avgar.2016}. More specifically, we paired each realized step with 24
-random steps, so that a realized step plus its 24 random steps formed a stratum
-that received a unique identifier. As suggested by \cite{Avgar.2016}, we
-generated random steps by sampling random turning angles from a uniform
-distribution (\(-\pi, +\pi\))) (which is equivalent to a von Mises distribution
-with concentration parameter \(\kappa = 0\)) and step lengths from a gamma
-distribution that was fitted to realized steps (scale \(\theta\) = 6'308 and
-shape \(k\) = 0.37)). Along each step, we extracted and averaged values from the
-habitat covariate layers using the {\tt velox} package \citep{Hunziker.2021} and
-calculated the movement metrics \textsf{sl}, \textsf{log(sl)}, and
-\textsf{cos(ta)} for each realized and random step. To facilitate model
-convergence, we standardized all continuous covariates to a mean of zero and a
-standard deviation of one. Since correlation among covariates was low (\(|r| <
-0.6\); \citealp{Latham.2011}), we retained all of them for modeling.
+random steps, so that a realized step plus its 24 random steps formed a
+25-step-stratum that received a unique identifier. As suggested by
+\cite{Avgar.2016}, we generated random steps by sampling random turning angles
+from a uniform distribution (\(-\pi, +\pi\)) (which is equivalent to a von Mises
+distribution with location and concentration parameters; \(\mu = \kappa = 0\))
+and step lengths from a gamma distribution that was fitted to realized steps
+(scale \(\theta\) = 6'308 and shape \(k\) = 0.37). Along each step, we extracted
+and averaged values from the habitat covariate layers using the {\tt velox}
+package \citep{Hunziker.2021} and calculated the movement metrics \textsf{sl},
+\textsf{log(sl)}, and \textsf{cos(ta)} for each realized and random step. To
+facilitate model convergence, we standardized all continuous covariates to a
+mean of zero and a standard deviation of one. Since correlation among covariates
+was low (\(|r| < 0.6\); \citealp{Latham.2011}), we retained all of them for
+modeling.
 
@@
 that animals assigned a selection score \(w(x)\) of the exponential form to each
-step \citep{Fortin.2005}. \(w(x)\) depended on the step's associated covariates
-(\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e. relative
-selection strengths; \citealp{Avgar.2017}) towards these covariates (\(\beta_1,
-\beta_2, ..., \beta_n\)):
+step (\ref{EQ2}; \citealp{Fortin.2005}). \(w(x)\) depended on the step's
+associated covariates (\(x_1, x_2, ..., x_n\)) and on the animal's preferences
+(i.e. relative selection strengths; \citealp{Avgar.2017}) towards these
+covariates (\(\beta_1, \beta_2, ..., \beta_n\)):
 
@@
 
-The probability of a step being realized was then contingent on the step's
-selection score, as well as on the selection scores of all other step in the
-same stratum:
+\noindent The probability of a step being realized was then contingent on the
+step's selection score, as well as on the selection scores of all other step in
+the same stratum:
 
@@
 implemented using the r-package {\tt glmmTMB} \citep{Brooks.2017}. To capitalize
-on the poisson trick introduced by \cite{Muff.2020}, we defined random
-intercepts for each stratum and fixed the intercept variance to an arbitraty
-high value of \(10 ** 6\). We also used disperser identity to model random
-slopes for all covariates.
+on the poisson trick proposed by \cite{Muff.2020}, we defined random intercepts
+for each stratum and fixed the intercept variance to an arbitrary high value of
+\(10 ^ 6\). We also used disperser identity to model random slopes for all
+covariates.
 
 The covariate structure of the movement model was based on a habitat selection
-model that we previously developed (hereafter referred to as based model,
-\citealp{Hofmann.2021}). In the base model, no interactions among habitat
-covariates and movement covariates were considered. Hence, we expanded the base
-model and allowed for interactions between all movement covariates and habitat
-covariates, thus reflecting that movement behavior may depend on habitat
-conditions (details in Appendix AX). To assess the most parsimonious movement
-model among model candidates, we ran stepwise forward model selection based on
-Akaike's Information Criterion (AIC, \citealp{Burnham.2002}). Finally, we
-validated the predictive power of the most parsimonious model using k-fold
-cross-validation for case-control studies as desribed in \cite{Fortin.2009}. In
-brief, this validation proofs a significant prediction in case the Spearman rank
-correlation of predicted step ranks and associated frequencies under the
-movement model is significantly stronger than under random preferences (details
-in Appendix AX).
+model that we previously developed for dispersing wild dogs (hereafter referred
+to as \textit{based model}, \citealp{Hofmann.2021}). In the base model, no
+interactions among habitat covariates and movement covariates were considered.
+Hence, we expanded the base model and allowed for interactions between all
+movement covariates and habitat covariates, thus reflecting that movement
+behavior may depend on habitat conditions (details in Appendix A1). To determine
+the most parsimonious movement model among model candidates, we ran stepwise
+forward model selection based on Akaike's Information Criterion (AIC,
+\citealp{Burnham.2002}). Finally, we validated the predictive power of the most
+parsimonious model using k-fold cross-validation for case-control studies as
+described in \cite{Fortin.2009}. This validation proofs a significant prediction
+in case the Spearman rank correlation of predicted step-ranks and associated
+frequencies under the movement model is significantly greater than under the
+assumption of random preferences (details in Appendix A2).
 
 \subsection{Step 2 - Dispersal Simulation}
-We used the most parsimonious movement model to simulate individual disperseral
+We used the most parsimonious movement model to simulate individual dispersal
 trajectories within the study area. The simulation of a dispersal trajectory
@@
 fitted gamma distribution. Like in the empirical data, each random step
-represented the straight line movement possible within 4 hours. To prevent
+represented the straight-line movement possible within 4 hours. To prevent
 unreasonably large steps, we restricted sampled step lengths to a maximum of 35
 km (i.e. the farthest dispersal distance traveled within 4 hours in our data).
-(3) Along each random step, we extracted and averaged values from the respective
+(3) Along each random step, we extracted and averaged values from the different
 habitat covariate layers and calculated movement covariates. To ensure
 compatible scales with the fitted movement model, we standardized extracted
-values using the mean and standard deviation from the empirical data. (4) We
+values using means and standard deviations from the empirical data. (4) We
 applied the parametrized movement model to predict the selection score \(w(x)\)
-for each step and we translated predicted scores into probabilities using
-\ref{EQ2}. (5) We used predicted probabilities to sample one of the random steps
-and determined the animal's new position. We repeated steps (2) to (5) until
-2,000 steps (i.e. 400 consecutive dispersal days) were realized.\todo{Since we did not explain the 8-hourly gaps, this might be confusing}
+for each step using \ref{EQ1} and we translated predicted scores into
+probabilities using \ref{EQ2}. (5) We used predicted probabilities to sample one
+of the random steps and determined the animal's new position. We repeated steps
+(2) to (5) until 2,000 steps (i.e. 400 consecutive dispersal days) were
+realized.\todo{Since we did not explain the 8-hourly gaps, this might be
+confusing}
 
@@
 km\textsuperscript{2}; \citealp{Pomilia.2015}), while the remaining 30,000
-source points were placed randomly inside the buffer zone, mimicing potential
+source points were placed randomly inside the buffer zone, mimicking potential
 immigration into the study area (Figure S1). Due to the random distribution of
@@
 dispersal trajectories and computed the mean traversal frequency across
-replicates, as well as its 95\% confidence-interval for each checkpoint. We
+replicates, as well as its 95\% prediction-interval for each checkpoint. We
 considered connectivity to have reached a steady state once the width of the
@@
 simulated dispersers (e.g. \citealp{Peer.2008, Hauenstein.2019, Zeller.2020}).
-That is, we rasterized all simulated trajectories onto a raster with 250 m x 250
-m resolution and tallied them into a single map. If the same trajectory crossed
-a raster-cell twice, we only counted it once, thereby mitigating biases from
-individuals that were trapped and moved in circles. To achieve high performance
-rasterization, we used the R-package {\tt terra} \citep{Hijmans.2021b}.
+Specifically, we rasterized all simulated trajectories onto a raster with 1 km x
+1 km resolution and tallied resulting layers into a single map. If the same
+trajectory crossed a raster-cell twice, we only counted it once, thereby
+mitigating biases from individuals that moved in circles because they were
+surrounded by unfavorable habitat. To achieve high performance rasterization, we
+used the R-package {\tt terra} \citep{Hijmans.2021b}.
 
@@
 trajectories into a network and calculated betweenness scores for all
-raster-cells in the study area \citep{BastilleRousseau.2018}. Betweenness
-measures how often a specific network-node (i.e. raster-cell) lies on a shortest
-path between any other pair of nodes and is therefore pertinent to connectivity
-\citep{BastilleRousseau.2018}. To transform simulated trajectories into a
+raster-cells in the study area \citep{BastilleRousseau.2018}. Betweenness is a
+pertinent metric for connectivity as it measures how often a specific
+network-node (i.e. raster-cell) lies on a shortest path between any other pair
+of nodes \citep{BastilleRousseau.2018}. To convert simulated trajectories into a
 network, we followed \citep{BastilleRousseau.2018} and overlaid the study area
@@
 transitions, yet because {\tt igraph} handles edge weights (\(\omega\)) as
-costs, we had to invert the traversal-frequency thorugh each raster-cell by
+costs, we had to invert the traversal-frequency through each raster-cell by
 applying \(\omega = \frac{mean(Traversal Frequency)}{Traversal Frequency_i}\).
-Consequently, edges that were traversed frequently were assigned small weights
-and vice versa. Finally, we used the weighted network to calculate betweenness
-scores at all network nodes.
+Consequently, regularly used edges received small weights (i.e. low costs) and
+vice versa. Finally, we used the weighted network to calculate betweenness
+scores for all network nodes.
 
@@
 between specific patches inside the KAZA-TFCA, we calculated inter-patch
-connectivity between national parks. The decision to only consider national
-parks as our  ``patches'' was purely out of simplicity and does not imply that
-connections between other protected areas are impossible. To quantify
-inter-patch connectivity, we computed the relative frequency at which dispersers
-originating from one national park successfully moved into another national
-park. We considered movements successful if the individuals' dispersal
-trajectory intersected with the target national park at least once. We also
-recorded the number of steps required to reach the first intersection with the
-respective national park, allowing us to compute the average dispersal durations
-from one park to another. In summary, we determined \textit{if} and \textit{how
-often} dispersers moved between certain national parks, as well as \textit{how
-long} individuals had to move to make these connections.
+connectivity between NPs (e.g. \citep{Gustafson.1996, Kanagaraj.2013}). The
+decision to only consider NPs as potential ``patches'' was purely out of
+simplicity and does not imply that connections between other protected areas are
+impossible. To quantify inter-patch connectivity, we computed the relative
+frequency at which dispersers originating from one NP successfully moved into
+another NP. We considered movements successful if the individuals' dispersal
+trajectory intersected with the target NP at least once. We also recorded the
+number of steps required to reach the first intersection with the respective NP,
+allowing us to compute the average dispersal durations from one park to another.
+In summary, we determined \textit{if} and \textit{how often} dispersers moved
+between certain NPs, as well as \textit{how long} individuals had to move to
+make these connections.
 
@@
 covariates and several of their interactions, suggesting that movement behavior
-during dispersal depends on habitat conditions (\Cref{MovementModel}, Table S1
+during dispersal depended on habitat conditions (\Cref{MovementModel}, Table S1
 and Table S2). Although multiple models received an AIC weight > 0 (Table S1),
@@
 decision only marginally influenced subsequent steps as all models with positive
-AIC weights contained similar covarites (Table S1). Aids for interpreting main
-effects and interactions are provided in Figure S2. Under average conditions,
-dispersing wild dogs avoided moving through water, woodlands, and areas
-dominated by humans, but preferred shrublands or grasslands
-(\Cref{MovementModel}). Dispersers realized shorter steps (indicating slower
-movements) in areas covered by water or woodland, while realizing larger steps
-in areas dominated by shrubs or grass. Moreover, dispersing wild dogs moved
-faster during twilight and at night (i.e. between 17:00 and 09:00 o'clock) than
-during the rest of the day. Although dispersers showed a preference for
-directional movements, especially when moving quickly, they did less so in
-proximity to humans or water, resulting in more tortuous movements in such
-areas.
-
-% When all other covariates held constant at their means, the habitat kernel
-% reveals that dispersing wild dogs avoid water but prefer its proximity, that
-% dispersers avoid areas that are covered by woodlands, yet prefer regions covered
-% by shrublands or grasslands, and that dispersers avoid movement through
-% landscapes that are dominated by humans. All of these effects are strong and
-% significant, except for the effect of \textsf{distance to water}, which is
-% insignificant on the 5\% significance level.
-%
-% With regards to the movement kernel, the positive estimate for \textsf{cos(ta)}
-% indicates that dispersers move with directional persistence, unlike what was
-% proposed by the uniform turning angle distribution. This preference is
-% particularly pronounced when dispersers realize large steps (i.e. move quickly),
-% as indicated by the positive estimates for \textsf{cos(ta):sl} and
-% \textsf{cos(ta):log(sl)}. Finally, the negative estimate for the interaction
-% \textsf{sl:LowActivity} reveals that wild dogs realize shorter steps (i.e. move
-% slower) between 09:00 and 17:00 o'clock than during the rest of the day. Aside
-% from the interaction \textsf{sl:LowActivity}, which has a very strong effect,
-% effects are moderate strength yet significant on the 5\% significance level;
-% only the interaction \textsf{cos(ta):sl} appears as insignificant on the 5\%
-% significance level.
-%
-% Finally, several significant interactions between movement and habitat
-% covariates suggest that movement behavior is contingent on habitat conditions.
-% For example, there's strong evidence that dispersers realize shorter steps in
-% areas covered by water or in forested areas. Furthermore, it appears that
-% dispersers realize larger steps in areas that are dominated by shrubs/grassland,
-% but shorter steps in areas that are distant to water. Finally, it seems that the
-% preference for directional persistence is less pronounced when dispersers cross
-% human-dominated landscapes, yet more pronounced at great distance from water.
-% Nevertheless, some of these effects are only weakly significant and exhibit
-% small effect sizes.
+AIC weights retained similar covariates (Table S1). Plots that facilitate model
+interpretation are provided in Figure S2. Under average conditions, dispersing
+wild dogs avoided moving through water, woodlands, and areas dominated by
+humans, but preferred shrublands or grasslands (\Cref{MovementModel}).
+Dispersers realized shorter steps (indicating slower movements) in areas covered
+by water or woodland, while realizing larger steps in areas dominated by shrubs
+or grass. Moreover, dispersing wild dogs moved faster during twilight and at
+night (i.e. between 17:00 and 09:00 o'clock) than during the rest of the day.
+Although dispersers showed a preference for directional movements (i.e. low
+turning angles), especially when moving quickly, they did less so in proximity
+to humans or water, resulting in more tortuous movements in such areas.
 
 The k-fold cross-validation of the movement model showed that the final model
-substantially outperforms a random guess , suggesting reliable predictions
+substantially outperformed a random guess and suggested reliable predictions
 (confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\)
-do not overlap). Moreover, the model correctly assigns high selection scores to
+do not overlap). Moreover, the model correctly assigned high selection scores to
 realized steps (\Cref{MovementModel}b), indicating a good fit between
@@
 \subsection{Dispersal Simulation}
-Our dispersal simulations based on the most parsimonious movement model proved
+Dispersal simulations based on the most parsimonious movement model proved
 useful for assessing landscape connectivity. Of the 50,000 simulated dispersers
@@
 map boundary, suggesting minimal biases due to boundary effects. Moreover, our
-examination of relative traversal frequencies across all checkpoints suggested
+examination of the relative traversal frequency across all checkpoints showed
 that connectivity reached a steady state after 10,500 simulated dispersal
-trajectories (Figure S3). Although variability in relative traversal frequencies
+trajectories (Figure S3). Although variability in relative traversal frequency
 kept decreasing as we increased the number of simulated dispersers, the marginal
@@
 KAZA-TFCA were frequently traversed by dispersing wild dogs (mean traversal
-frequency = 166, IQR = 274, Figure S6a), while areas beyond the KAZA-TFCA
-boundary were rarely visited (mean traversal frequency = 61, IQR = 133, Figure
-S6a). Most notably, the region in northern Botswana south of the Linyanti swamp
-appeared as highly frequented dispersal hotspot (mean traversal frequency = 987,
-IQR = 558). Nevertheless, the presence of extensive water bodies, such as the
-Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, restricted
-dispersal movements and limited realized connectivity within the KAZA-TFCA.
-Similarly, high human densities, roads, and agricultural activities in Zambia's
-and Zimbabwe's part of the KAZA-TFCA limited dispersal movements. Outside the
-KAZA-TFCA, the most heavily used regions included the areas inside the Central
-Kalahari National Park in Botswana, the area south-west of the Khaudum National
-Park in Namibia, and the area surrounding the Liuwa Plains National Park in
-Zambia. Although the heatmap facilitated the identification of areas frequently
-traversed by simulated dispersers, it seemed impractical to pinpoint dispersal
-corridors.
+frequency inside KAZA-TFCA = 166, IQR = 274, Figure S6a), whereas areas beyond
+the KAZA-TFCA boundary were rarely visited (mean traversal frequency outside
+KAZA-TFCA = 61, IQR = 133, Figure S6a). Most notably, the region in northern
+Botswana south of the Linyanti swamp appeared as highly frequented dispersal
+hotspot (mean traversal frequency = 987, IQR = 558). Nevertheless, the presence
+of extensive water bodies, such as the Okavango Delta, the Makgadikgadi Pan, and
+the Linyanti swamp, restricted dispersal movements and limited realized
+connectivity within the KAZA-TFCA. Similarly, high human density, roads, and
+agricultural activities in Zambia's and Zimbabwe's part of the KAZA-TFCA limited
+dispersal movements in those countries. Outside the KAZA-TFCA, the most heavily
+used regions included the areas inside the Central Kalahari National Park in
+Botswana, the area south-west of the Khaudum National Park in Namibia, and the
+area surrounding the Liuwa Plains National Park in Zambia. Although the heatmap
+facilitated the identification of areas frequently traversed by simulated
+dispersers, it seemed impractical to pinpoint dispersal corridors.
 
@@
   that are frequently traversed by virtual dispersers. For spatial reference we
-  plotted a few selected national parks (dark gray). Additional heatmaps showing
-  the traversal frequency when individuals move fewer than 2'000 steps are
-  provided in Figure S4.}
+  plotted a few selected NPs (dark gray). Additional heatmaps showing the
+  traversal frequency when individuals move fewer than 2'000 steps are provided
+  in Figure S4.}
   \label{Heatmap}
@@
 The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
-that run within the KAZA-TFCA . Again, the region in northern Botswana emerged
-as wild dog dispersal hub that connected more remote regions in the study area.
+that run within the KAZA-TFCA. Again, the region in northern Botswana emerged as
+a wild dog dispersal hub that connected more remote regions in the study area.
 Towards east, the extension of this corridor ran through Chobe National Park
@@
 Matusadona National Park in Zimbabwe. Northwest of the Linyanty ecosystem, a
-major corridor expanded into Angola, where it splitted and finally traversed
+major corridor expanded into Angola, where it split and finally traversed
 over a long stretch of unprotected area into Zambia's Kafue National Park.
@@
 them ran within the KAZA-TFCA boundaries (median betweenness inside KAZA-TFCA =
-6947\(k\), IQR = 54311\(k\), Figure S6b). In general, there were few corridors
+6.947 Mio, IQR = 54.311 Mio, Figure S6b). In general, there were few corridors
 that directly linked the peripheral regions of the KAZA-TFCA and passed through
-unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA =
-2685\(k\), IQR = 9891\(k\), Figure S6b). Compared to the heatmap, the
+unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA-TFCA =
+2.685 Mio, IQR = 9.891 Mio, Figure S6b). Compared to the heatmap, the
 betweenness map facilitated the identification of dispersal corridors between
@@
   \caption{Map of betweenness scores, highlighting distinct dispersal corridors
-  and potential bottlenecks across the extent of the KAZA-TFCA. A high
-  betweenness score indicates that the respective area is exceptionally
-  important for connecting different regions in the study area. In this sense
-  the metric can be used to pinpoint discrete movement corridors
+  and potential bottlenecks across the extent of the KAZA-TFCA. Betweenness
+  measures the number of shortest paths traversing through each node
+  (raster-cell). Hence, a high betweenness score indicates that the respective
+  area is exceptionally important for connecting different regions in the study
+  area. The metric is therefore useful to pinpoint discrete movement corridors
   \citep{BastilleRousseau.2018}. Note that we square-rooted betweenness scores
-  to improve visibility of corridors with low scores. Additional betweenness
-  maps showing betweenness scores when individuals move fewer than 2'000 steps
-  are provided in Figure S4.}
+  to improve visibility of corridors with comparably low scores. Additional
+  betweenness maps showing betweenness scores when individuals move fewer than
+  2'000 steps are provided in Figure S4.}
   \label{Betweenness}
@@
 The map of inter-patch connectivity showed that the relative frequency at which
-simulated dispersers moved from one national park to another varied, as did the
-average dispersal duration required to make these connections
-(\Cref{InterpatchConnectivity}). Overall, interpatch connectivity between
-national parks in Angola, Namibia, and Botswana appeared to be high (mean
-relative frequency = +- ) with short dispersal durations (mean duration = +-)
-
-Results from the analysis of inter-patch connectivity are given in
-\Cref{InterpatchConnectivity}, which shows all realized links by simulated
-dispersers between national parks. The map furthermore indicates the relative
-frequency at which dispersers originating from one national park successfully
-reached another national park, as well as the average duration dispersers had to
-move to realize those links. It is again worth pointing out that the figure is
-only intended as an illustration, as for clarity we only considered inter-patch
-connectivity between national parks (NPs), albeit plenty of links between other
-protected areas exist. Overall, inter-patch connectivity between NPs in Angola,
-Namibia, and Botswana appears to be high, with relatively short dispersal
-durations between national parks. In contrast, we observe that connections from
-the central region into the Kafue NP in Zambia require rather long dispersal
-events and are rather infrequent. Similarly, relatively few connections lead
-into Zimbabwe's Chizarira and Matusadona NP. In some cases, we also find
-imbalances between ingoing and outgoing links, hinting at potential source-sink
-dynamics that occur due to asymmetries in disperser's willingness to cross an
-area, depending on the direction in which the area is traversed. For instance,
-while a fair share of dispersers originating from the Chizaria NP in Zimbabwe
-manages to move into the Hwange NP, there are comparably few dispersers that
-succeed in the opposite direction.
-
-% The map shows all realized links by simulated dispersers between national parks
-% and indicates the average duration a disperser had to move to realize those
-% links. For instance, 6.8 \% of the simulated dispersers originating from the
-% Moremi National Park successfully reached the Chobe National Park and 4.2 \% the
-% Hwange National Park in Zimbabwe. On average, dispersers moved for 623 steps
-% before arriving at Chobe (SD = 520) and for 1'413 steps before arriving at
-% Hwange (SD = 371). While the map only depicts connections between national
-% parks, this does not imply that no connections to other protected areas exist
-% and is the result of our simplifying assumptions above. Interestingly, while we
-% identified a potential dispersal corridor between Angola's NPs and the Kafue NP
-% in Zambia, \Cref{InterpatchConnectivity} suggests that this link is only
-% rarely realized and requires very long dispersal events. In contrast, we find
-% that dispersal between the Moremi NP and Chobe NP are relatively frequent and
-% require fewer steps, which can be expected given that the areas are located
-% close to each other.
+simulated dispersers moved from one NP to another varied, as did the average
+dispersal duration required to make these connections
+(\Cref{InterpatchConnectivity}). Overall, inter-patch connectivity between NPs
+in Angola, Namibia, Botswana, and Zimbabwe appeared to be high; between 54\% and
+87\% of individuals originating from a NP in these countries successfully moved
+into some other NP (Figure S6a). Conversely, only 19\% of the dispersers leaving
+from a NP in Zambia managed to find a route into some other NP (Figure S6b).
+Prior to reaching another NP, individuals from Angola, Namibia, Botswana,
+Zimbabwe, and Zambia had to move for an average of 630, 640, 940, 1045, and 890
+steps, respectively. For some NPs, we also detected imbalances between the
+number of ingoing and outgoing links, hinting at possible source-sink dynamics.
+From Chobe NP, for instance, 510 individuals reached Moremi NP, yet the opposite
+route was only realized by 340 individuals. However, relative to the number of
+simulated individuals in these NPs, these numbers imply fractions of 50\% and
+68\%. Interestingly, it also appears that the dispersal corridor between
+Angola's NPs and the Kafue NP in Zambia identified in
+\Cref{InterpatchConnectivity} is only rarely realized.
 
@@
   \includegraphics[width=\textwidth]{99_InterpatchConnectivity.png}
-  \caption{Map of inter-patch connectivity, highlighting connections between
-  national parks (dark green). Yellow bubbles represent the center of the
-  different national parks and are sized in relation to the number of simulated
-  dispersers originating from each park. Black dots represent national parks
-  that were smaller than 700 km\textsuperscript{2} and therefore did not serve
-  as source areas. Arrows between national parks illustrate between which
-  national parks the simulated dispersers successfully moved and the color of
-  each arrow shows the average number of steps (4-hourly movements) that were
-  necessary to realize those connections. Additionally, the line thickness
-  indicates the relative number of dispersers originating from a national park
-  that realized those connections. Note that a similar network view could be
+  \caption{Map of inter-patch connectivity, highlighting connections between NPs
+  (dark green). Yellow bubbles represent the center of the different NPs and are
+  sized in relation to the number of simulated dispersers originating from each
+  park. Black dots represent NPs that were smaller than 700
+  km\textsuperscript{2} and therefore did not serve as source areas. Arrows
+  between NPs illustrate between which NPs the simulated dispersers successfully
+  moved and the color of each arrow shows the average number of steps (4-hourly
+  movements) that were necessary to realize those connections. Additionally, the
+  line thickness indicates the relative number of dispersers originating from a
+  NP that realized those connections. Note that a similar network view could be
   adopted to investigate connectivity between other protected areas need not to
-  be restricted to national parks.}
+  be restricted to NPs.}
   \label{InterpatchConnectivity}
@@
 \subsection{Short Summary}
-We used ISSFs to analyse data of dispersing wild dogs and to parametrize a fully
-mechanistic movement model describing how dispersers move through the available
-landscape. We employed the parametrized model as an individual-based movement
-model to simulate 80'000 dispersing wild dogs moving 2'000 steps across the
-extent of the KAZA-TFCA, the world's largest transboundary conservation area.
-Based on simulated dispersal trajectories, we derived three complementary maps,
-each geared towards a better understanding of dispersal and landscape
-connectivity. The set of maps included a heatmap, revealing frequently traversed
-areas, a betweenness-map, delineating critical dispersal corridors, and a map of
-inter-patch connectivity, indicating the presence or absence of functional links
-between national parks as well as the average dispersal duration required to
-realize those links. We thereby showcase that ISSFs offer a simple, yet powerful
-framework to parametrize movement models and simulate dispersal to assess
-landscape connectivity. Importantly, individual-based simulations from ISSFs
-overcome several conceptual shortcomings inherent to more traditional
-connectivity modeling techniques, such as LCPA and CT.
+We presented a three-step workflow to simulate dispersal and assess landscape
+connectivity. We also demonstrated application of the three steps using
+empirical dispersal data from a free-ranging population of African wild dogs. In
+step one, we used ISSFs to parametrize a fully mechanistic movement model
+describing how dispersing wild dogs move through the available landscape. In
+step two, we employed the fitted model to simulate 80,000 dispersing wild dogs
+moving 2,000 steps across the extent of the KAZA-TFCA, the world's largest
+transboundary conservation area. In step three, we converted simulations into
+three complementary connectivity maps, each emphasizing a different aspect of
+landscape connectivity. The set of maps included a heatmap revealing frequently
+traversed areas, a betweenness-map delineating critical dispersal corridors, and
+a map of inter-patch connectivity indicating the presence and intensity of
+functional links between NPs and the average dispersal duration required to make
+those links. With this, we showed that simulations from ISSFs overcome several
+conceptual shortcomings inherent to more traditional connectivity modeling
+techniques, such as LCPA and CT.
 
@@
 \subsection{Movement Model}
-Because our movement model of dispersing wild dogs comprised two interacting
-kernels, it effectively rendered habitat and movement preferences of dispersers,
-as well as how preferences depended on habitat conditions. Results from the
-habitat kernel were largely in concert with previous studies that investigated
-habitat selection by dispersing wild dogs \citep{DaviesMostert.2012,
-Masenga.2016, Woodroffe.2019, Oneill.2020, Hofmann.2021}. However, by also
-incorporating a movement kernel, we were able to model several additional
-complexities inherent to dispersal. For instance, we could accommodate that
-dispersers move with directional persistence \citep{Cozzi.2018, Hofmann.2021}
-and exhibit step lengths that are correlated with turning angles
-\citep{Morales.2004, Borger.2012} by including the appropriate interactions in
-the movement model. While correlations between step lengths and turning angles
-could also be rendered by sampling them from copula probability distributions
-\citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to directly
-incorporate them in the movement model. In addition, by forming interactions
-between habitat covariates and movement covariates, we could render potential
-dependencies between movement and habitat preferences. For example, our final
-model contained an interaction between water-cover and step length, showing that
-dispersers realize shorter steps in areas covered by water. Likewise, we found
-that dispersers move more tortuously across water bodies than over dryland,
-which is clearly to be expected, given that wild dogs have to wade or swim when
-traversing waterbodies. The ability of accompanying such effects in a single
-model is one of the great strengths of ISSFs \citep{Avgar.2016, Fieberg.2021},
-which is why we believe the method offers a suitable framework for simulating
-movement and assessing connectivity.
+Our movement model of dispersing wild dogs, which included interactions between
+habitat and movement covariates, effectively rendered habitat and movement
+preferences of dispersers, leading to a significant improvement in predictive
+performance compared to an earlier model that omitted such interactions
+\citep{Hofmann.2021}. Results on habitat preferences largely complied with
+previous studies that investigated habitat selection by dispersing wild dogs
+\citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020,
+Hofmann.2021}, yet by also rendering movement preferences, we were able to model
+several additional complexities common to dispersal. For instance, by including
+the appropriate interactions in the movement model we could accommodate that
+dispersers move with directional persistence and exhibit step lengths that are
+correlated with turning angles. Albeit the same autocorrelation could be
+rendered by sampling step lengths and turning angles from copula probability
+distributions \citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to
+incorporate correlations directly in the movement model. We only considered
+first order autocorrelation, i.e. correlation between two consecutive steps,
+although higher order autocorrelation is conceivable and may be desirable to
+model \citep{Dray.2010, McClintock.2012}. This will require vast amounts of GPS
+data that are not interrupted by missing fixes; something that is rarely
+achieved in reality \citep{Graves.2006}. Aside from modeling autocorrelation in
+consecutive movements, we also rendered potential dependencies between movement
+and habitat preferences by forming interactions between habitat covariates and
+movement covariates. For example, our final model contained an interaction
+between water-cover and step lengths showing that dispersers realized shorter
+steps in areas covered by water or an interaction between turning angles and
+water-cover highlighting that dispersers moved more tortuous in areas covered by
+water. The results themselves are not surprising, considering that wild dogs can
+only cross water by swimming or wading. Nevertheless, the ability of
+accompanying such effects in a single model is what makes ISSFs such a great and
+flexible tool for rendering a variety of different movement behaviors
+\citep{Avgar.2016, Fieberg.2021}. For this reason, we believe the method is a
+perfectly suited framework for individual-based simulations that serve to
+examine connectivity.
 
 \subsection{Simulation}
-Our simulation of 80'000 dispersers moving 2'000 steps across the landscapes of
-the KAZA-TFCA required five days of computation on a modern desktop machine.
-
-On a machine with an octa-core AMD Ryzen 7 2700X processor (8 x 3.6 GHz, 16
-logical cores) and 64 GB of RAM, a batch of 1'000 simulated dispersers moving
-over 2'000 steps required 90 minutes to compute (\(\mu = 88.90\), \(\sigma =
-1.87\)). Consequently, the simulation of all 80'000 dispersers (160 Mio. steps)
-terminated after 120 hours (i.e. five days). Comparable simulations will be
-substantially faster for smaller study areas and lower resolution covariates, as
-the covariate extraction from large and high-resolution rasters was
-computationally the most demanding task. Out of the 50'000 dispersers initiated
-inside the main source area, only 4.5\% were eventually repelled by a map
-boundary, suggesting that biases due to boundary effects should be minimal.
-
- The
-long simulation time was primarily caused by the massive extent considered (ca.
-1.8 Mio. km\textsuperscript{2} when including the buffer) and the large number
-of dispersers simulated. Most connectivity studies are limited to much smaller
-extents (e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017,
-Zeller.2020}) and will therefore achieve faster simulation times. We also
-believe that fewer simulated dispersers will often suffice, as the relative
-traversal frequency by simulated individuals through randomly placed checkpoints
-in the study area converged already after 10'500 simulated individuals. However,
-the required number of simulated individuals to achieve reliable estimates of
-connectivity will vary depending on the structure of the landscape and the
-dispersal ability of the focal species.
+Our simulation of 80'000 dispersers moving 2'000 steps across the KAZA-TFCA
+required five days of computation on a modern desktop machine (AMD Ryzen 7 2700X
+processor with 8 x 3.6 GHz and 16 logical cores, 64 GB of RAM). The long
+simulation time was primarily caused by the massive extent considered (ca. 1.8
+Mio km\textsuperscript{2}) and the large number of dispersers simulated. Most
+connectivity studies are limited to smaller extents (e.g.
+\citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020})
+and will therefore achieve faster simulation times. We also believe that fewer
+simulated dispersers will often suffice, as the relative traversal frequency by
+simulated individuals through randomly placed checkpoints across our study area
+converged already after 10,500 simulated dispersal trajectories. The number of
+required simulations to achieve reliable estimates of connectivity will,
+however, vary depending on the structure of the landscape and the dispersal
+ability of the focal species \citep{Gustafson.1996}. For species that disperse
+short distances through homogeneous environments, already few simulations may
+suffice go gauge connectivity, whereas for species that disperse over long
+distances through heterogeneous habitats, a large number of simulations will be
+required to sufficiently explore the spectrum of possible routes.
 
 \subsection{Maps}
-The heatmap generated from simulated dispersal trajectories highlighted that
-numerous individuals traversed the Moremi NP and the Chobe NP in northern
-Botswana. We previously uncovered the same area as potential dispersal hotspot
-using least-cost methods \citep{Hofmann.2021}, yet it has been questioned
-whether this was a consequence of the region being located in the center of the
-study area and connections being enforced between predefined start and
-endpoints. With the current simulations, connections were no longer enforced and
-simulated individuals were able to leave the main study area. Nonetheless, a
-majority of simulated individuals traversed the area in northern Botswana. This
-suggests that the dispersal hotspot is not merely an artifact of the applied
-method but truly results from landscape characteristics. The same region is also
-pronounced on the betweenness map, implying that it facilitates the relocation
-of individuals into more remote regions of the KAZA-TFCA. While this is an
-example of an area where both the heatmap and the betweenness map attest great
-importance, there are other instances where this is not true. For example, while
-the area between the Lengue-Luiana NP in Angola and the Kafue NP in Zambia
-receives a high betweenness-score, the heatmap shows that the area is only
-rarely traversed by dispersers. Hence, despite the region's importance for
-linking Angola's NPs to Zambia's NP, only few simulated dispersers actually used
-the corridor. Conversely, we find that the Central Kalahari NP receives a low
-betweenness score, despite being highly frequented by simulated dispersers.
-Besides highlighting frequently traversed areas and dispersal corridors, we also
-also consulted inter-patch connectivity between NPs, depicting the presence or
-absence of functional links between national parks. Because dispersal movements
-were rendered sequentially, we could also derive the average dispersal duration
-that was required to realize those links. This enabled us to show that movements
-from Angola into Zambia's Kafue NP are rare and require a large number of steps
-, whereas dispersal between the Moremi NP and Chobe NP occurs frequently and
-requires relatively few steps. All in all, the rich inference that can be drawn
-from the ensemble of proposed connectivity maps.
-
-% Other studies that use similar simulations
-\subsection{Related Literature}
-Our approach of simulating movement to assess connectivity is related to a
-series of previously published papers. \cite{Clark.2015}, for instance,
-collected GPS data on American black bears (\textit{Ursus americanus}) and used
-SSFs to fit a habitat selection model. They then used the fitted model to
-simulate movement and identify likely movement corridors between four distinct
-habitat patches. For the same species, \cite{Zeller.2020} used SSFs and
-forecasted seasonal habitat connectivity under changing land-use. Because both
-of these studies relied on \textit{regular} SSFs, rather than
-\textit{integrated} SSFs, they could not account for interdependencies among
-habitat and movement preferences \citep{Avgar.2016}. In addition, both studies
-employed data collected on resident individuals instead of dispersers, although
-evidence suggests that residents are more reluctant to cross areas that are
-readily traversed by dispersers \citep{Elliot.2014, Gaston.2016, Abrahms.2017,
-Keeley.2017}. The application of data collected on resident animals may
-therefore result in an underestimate of connectivity \citep{Elliot.2014}. Two
-further studies that used SSFs to simulate animal movement have been conducted
-by \cite{Potts.2013} and \cite{Signer.2017}, yet the primary purpose here was to
-estimate steady-state utilization distributions of resident animals and not on
-the analysis of dispersal and connectivity.
-
-\subsection{Advantages of ISSF Simulations}
-% Does not enforce route towards known endpoint
-A simulation-based approach as proposed in this article offers several
-advantages over LCPA and CT. In contrast to LCPA, for instance, an
-individual-based simulation does not require to assume known endpoints. Instead,
-each endpoint emerges naturally from a simulated dispersal trajectory. The
-ability of not needing to provide endpoints is particularly valuable for
-dispersal studies, because dispersers often venture into unfamiliar territory
-and are therefore unlikely to know the destination of their journey
-\citep{Elliot.2014, Abrahms.2017, Cozzi.2020}. Moreover, LCPA always enforces a
-connection towards the pre-defined endpoints, even if associated movement costs
-are unreasonably high. With simulations from ISSFs this is no longer the case. A
-connectivity model that does not require pre-defined endpoints also ensures that
-movement corridors are not enforced between certain start- and endpoints, which
-permits to detect potential routes that do not lead into suitable habitats but
-into ecological traps \citep{Dwernychuk.1972, VanDerMeer.2014} or areas with a
+Each of the three connectivity maps derived from simulated dispersal
+trajectories highlighted a different aspect of landscape connectivity. The
+heatmap was most suitable for pinpointing frequently traversed areas and showed
+that an exceptionally large number of dispersers moved through the regions of
+the Moremi NP and the Chobe NP in northern Botswana. We previously identified
+the same area as potential dispersal hotspot using LCPA \citep{Hofmann.2021},
+yet it was not clear whether this was a consequence of the region being located
+in the center of the study area and connections being enforced between
+predefined start and endpoints. Contrary to LCPA, a simulation-based approach as
+presented in this paper does not require predefined endpoints because endpoints
+emerge naturally as the result from a simulated dispersal trajectories. Not
+having to pre-determine endpoints is especially useful for dispersal studies, as
+known endpoints are usually an unrealistic assumption \citep{Elliot.2014,
+Abrahms.2017, Cozzi.2020}. Simultaneously, simulations permit to detect
+potential routes that do not necessarily lead into suitable habitats but into
+ecological traps \citep{Dwernychuk.1972, VanDerMeer.2014} or into areas with a
 high susceptibility for human wildlife conflicts \citep{Cushman.2018}.
-
-% Explicit Representation of Time
-In contrast to LCPA and CT, simulations from ISSFs furthermore yield the
-advantage of an explicit representation of time, which enables to answer
-questions such as: ``\textit{How long will it take a disperser to move from A to
-B?}'' or ``\textit{Is it possible for a disperser to move from A to B within X
-days?}''. An explicit representation of time also yields opportunities
-for studying how seasonality affects connectivity and to investigate whether
-some dispersal corridors are only available temporarily (\textit{dynamic
-connectivity}; \citealp{Zeller.2020}). With LCPA or CT, incorporating
-seasonality is currently impractical, as both methods require a static
-permeability surface as inputs. Hence, the only possibility to study seasonality
-effects is to repeat the same analysis using different permeability surfaces,
-each rendering the environment at a different point in time (e.g.
-\citealp{Benz.2016, Osipova.2019}). With simulations from ISSFs, on the other
-hand, the environment can be rendered dynamically ``as the dispersers move'',
-such that simulated individuals can respond to seasonal factors directly within
-the simulation. Hence, rather than employing a set of static habitat layers,
-each layer would be updated as the dispersers move, thus correctly rendering
-seasonal changes in the environment.
-
-% Requires Detailed Knowledge on Step Lengths and Turning Angles
-While an explicit representation of time provides several benefits, it requires
-that step lengths and turning angles are modeled properly
-\citep{Kanagaraj.2013}, so that dispersal durations between areas can be
-estimated reliably. Correctly rendering step lengths and turning angles under
-varying environmental conditions is one of the key strengths of ISSFs
-\citep{Avgar.2016, Prokopenko.2017, Fieberg.2021}, which is why we believe that
-the framework is exceptionally well suited for simulating dispersal and
-assessing landscape connectivity. In addition, the framework enables to model
-autocorrelation between step lengths and turning angles, thereby incorporating
-directional persistence. Here, we only considered first order autocorrelation,
-i.e. correlation between two consecutive steps. Although higher order
-autocorrelation is conceivable and might be desirable to model, this requires
-vast amounts of GPS data that is not intercepted by missing fixes and is
-therefore often impractical to model in reality.
+Irrespective of the fact that simulated individuals were no longer enforced to
+move towards known endpoints, numerous simulated individuals traversed
+the region in northern Botswana, demonstrating that this hotspot is not merely
+an artifact but truly results from its favorable landscape characteristics.
+
+In contrast to the heatmap, the betweenness map emphasized relatively narrow and
+linear movement routes and thus facilitated the identification of discrete
+movement corridors in the study area. The resulting map reinforced our notion
+that Botswana plays a central role for the establishment of connections into
+more remote regions of the KAZA-TFCA. While in this case both the heatmap and
+the betweenness map attribute high importance to the region in northern
+Botswana, there are other regions in which the two metrics do not necessarily
+coincide. The stretch of unprotected land between Luengue-Luiana NP in Angola
+and the Kafue NP in Zambia, for instance, receives a high betweenness-score, yet
+is only rarely traversed by dispersers according to the heatmap. This implies
+that the corridor takes on a crucial role for connecting Luengue-Luiana NP to
+Kafue NP, yet is in reality only traversed rarely. These contrasts highlight the
+value of consulting multiple metrics when assessing connectivity.
+
+As a final piece to the trinity of connectivity maps, we produced a map of
+inter-patch connectivity that strictly focused on NPs. The map depicted the
+frequency at which simulated individuals moved between NPs as well as the
+average duration (in steps) required to realize them. Calculating dispersal
+durations was only possible because dispersal trajectories were simulated
+spatio-temporally explicitly, something that is currently impossible using LCPA
+or CT. An explicit representation of time enables to answer questions such as:
+``\textit{How long will it take a disperser to move from A to B?}'' or
+``\textit{Is it possible for a disperser to move from A to B within X days?}''.
+Moreover, rendering time explicitly yields opportunities to study how
+seasonality affects connectivity and to investigate whether dispersal corridors
+are only available temporarily or all-year round (\textit{dynamic connectivity};
+\citealp{Zeller.2020}). With LCPA or CT, seasonality can currently only be
+incorporated by generating seasonal permeability surfaces and repeating the
+connectivity analysis for each surface individually (e.g. \citealp{Benz.2016,
+Osipova.2019}). With simulations from ISSFs, on the other hand, seasonal changes
+in the environment could be rendered dynamically ``as the dispersers move'', so
+that simulated individuals could directly respond to such fluctuations.
 
 \subsection{Disadvantages of ISSF Simulations}
-Despite the benefits that simulations from ISSFs offer, we also want to confer
-some of the non-trivial modeling decisions involved.
-
-In particular, it is worth pointing out five modeling decisions: (1) number of
-simulated individuals, (2) location of source points, (3) dispersal duration,
-(4) boundary behavior, and (5) how to handle uncertainty and individual
-variability.
+Despite the many benefits and great flexibility offered by simulations from
+ISSFs, one also must be aware of the associated non-trivial but important
+modeling decisions. Here, we will elaborate on four modeling decisions
+concerning: (1) the number of simulated individuals, (2) the location of source
+points, (3) the simulated dispersal duration, and (4) behavior at map
+boundaries.
 
@@
 simulated individuals. A higher number is always desirable, as each additional
-disperser provides novel information about landscape connectivity. However, this
-comes at the cost of computational efficiency, implying that a trade-off needs
-to be managed. As noted by \cite{Signer.2017}, the trade-off can be handled by
-simulating additional individuals only until estimated metrics converge. Here,
-we employed the relative traversal frequency across checkpoints as target metric
-and found that convergence across all checkpoints was already achieved after
-10'500 simulated individuals, yet we recognize that this strongly depends on the
-focal species dispersal ability and landscape characteristics.
-
-(2) Aside from specifying the absolute number of simulated individuals, one also
-needs to define a source point within a suitable source area for each
-individual. Here, we placed source points within protected areas large enough to
-sustain viable wild dog populations. Given that wild dogs primarily survive in
-formally protected areas \citep{Woodroffe.1999, DaviesMostert.2012,
-Woodroffe.2012, VanDerMeer.2014} we considered this decision to be appropriate.
-Due to a lack of precise knowledge about wild dog abundances in the different
-protected areas, we distributed source points randomly within them. If, however,
-corresponding data is available, source points can be distributed accordingly,
-reflecting the fact that source areas do not necessarily produce an identical
-number of dispersers. Alternatively, source points can be distributed
-homogeneously, but be weighted afterwards according to estimated densities in
-the respective source area. In cases where knowledge about suitable source areas
-is lacking, these could also be delineated using habitat suitability models
-(e.g. \citealp{Squires.2013}). After all, the challenge of selecting meaningful
-source areas and source points is not unique to individual-based simulations and
-also applies to LCPA or CT.
-
-(3) When employing ISSFs to simulate dispersers, it is also required to decide
-on meaningful dispersal durations (i.e. number of simulated steps). When
-sufficient dispersal data of the focal species is available, dispersal durations
-can be sampled from observed dispersal events. Due to the low number of observed
-dispersal events and due to the great variability in wild dogs' dispersal
-distances \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020} we opted against
-this approach. Instead, we simulated individuals for 2'000 steps, which is at
-the upper end of observed dispersal durations. Once the trajectories have been
-simulated, it is straightforward to subsample them and investigate the
-sensitivity of derived results with regards to different dispersal durations.
+disperser provides information about landscape connectivity, yet every
+individual also entails computational costs. Consequently, a trade-off needs to
+be managed and the benefits of simulating additional dispersers needs to be
+gauged against the cost of computation. \cite{Signer.2017} suggests handling
+this trade-off by simulating additional individuals only until the metrics of
+interest converge towards their steady state. We incorporated this idea and
+employed the relative traversal frequency across randomly placed checkpoints to
+verify that the number of simulated individuals was large enough. However, the
+exact number of required individuals might vary depending on the target metric
+and the anticipated connectivity map. More sophisticated target metrics tailored
+for each connectivity map will need to be developed in the future.
+
+(2) To initiate dispersers, a modeler needs to provide a set of source points.
+We placed source points within protected areas large enough to sustain viable
+wild dog populations, implicitly assuming wild dogs would primarily survive in
+large, formally protected areas \citep{Woodroffe.1999, DaviesMostert.2012,
+Woodroffe.2012, VanDerMeer.2014}. We lacked precise knowledge about wild dog
+densities in the different protected areas, so we distributed source points
+within them randomly. In cases where such data is available (e.g. thanks to a
+habitat suitability model; \citealp{Squires.2013}), source points can be
+distributed accordingly, reflecting that the number of dispersers may not
+strictly scale with the size of the source area. Alternatively, source points
+can be distributed homogeneously across suitable habitats and only later be
+weighted when deriving connectivity metrics. After all, the challenge of
+selecting meaningful source points is not unique to individual-based simulations
+but also applies to LCPA or CT.
+
+(3) When employing ISSFs to simulate dispersers, it is required to decide on the
+number of simulated steps (i.e. dispersal durations). If sufficient dispersal
+data of the focal species has been collected, dispersal durations could be
+sampled from observed dispersal events. In our case, the low number of observed
+dispersal events and the high variability in wild dog dispersal durations (see
+\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}) prevented us
+from sampling from observed dispersal events. Instead, we simulated all
+individuals for 2'000 steps, which is at the upper end of observed dispersal
+durations. This allowed us to subsample trajectories to shorter durations after
+their simulations and to investigate the sensitivity of our connectivity maps
+with respect to exact dispersal durations (Figures S4 and S5).
 
 (4) Unless simulated individuals are strongly drawn towards a point of
-attraction (e.g. \cite{Signer.2017}), dispersers eventually approach a map
-boundary, so that one or several of the proposed random steps leave the study
-area. In theoretical applications, this issue can be circumvented by simulating
-movement on a torus \citep{Hodel.2022}. For real data, however, alternative
-solutions are needed. One approach would be to terminate the simulation,
-assuming that the simulated animal left the study area forever. This can be
-problematic for individuals that are initiated in areas that are located close
-to map boundaries, especially since already a single random step leaving the
-study area breaks the simulation. Here, instead of breaking the simulation loop,
-we simply resampled transgressing random steps until they fully lied within the
-study area. This enforced simulated dispersers to be repelled by map boundaries
-and to remain within the main study area. Additionally, we artificially
-increased the study area using a buffer zone with randomized covariate values.
-This enabled virtual dispersers to leave and re-enter the main study area. The
-same method has been shown to effectively mitigate edge effects for graph-based
-connectivity models \citep{Koen.2010}.
-
-(5) We simulated dispersal using point estimates from our most parsimonious
-movement model. Depending on the amount of data and individual variability,
-these estimates can be subject to substantial uncertainty. For dispersal studies
-in particular, the low amount of data typically results in model coefficients
-with large confidence intervals \citep{Wiegand.2003, KramerSchadt.2007}. In
-these cases, point estimates may lead to biased connectivity estimates, which is
-why we urge future studies to investigate the sensitivity of ISSF simulations
-with respect to employed model parameters and to simulate dispersal based on
-parameters that encapsulate model uncertainty.
+attraction (e.g. \cite{Signer.2017}), some individuals will inevitably approach
+a map boundary such that one of their proposed random steps leaves the study
+area. In theoretical applications, this problem can be circumvented by
+simulating movement on a torus where no map boundaries exist \citep{Hodel.2022}.
+For real world data, however, other approaches are needed. One possible solution
+would be to simply terminate the simulation of the respective individual as soon
+as one of its random steps transgresses a map border. This approach implicitly
+assumes that the respective animal left the study area and will never return. In
+cases where individuals are released near map borders, this approach will be
+problematic, as already a single transgressing random step will break the
+simulation, thus strongly biasing dispersal from patches near map borders. We
+tried to mitigate this issue in two complementary ways. First, we artificially
+increased covariate layers (and therefore the study area) by a buffer zone with
+randomized covariate values. The same method has already proven effective in
+mitigating edge-effects for graph-based methods \citep{Koen.2010}. In our case,
+the buffer enabled dispersers to leave and re-enter the main study area, as well
+as to initiate potential immigrants into our study system. As a second
+mitigation measure, we resampled any transgressing random steps until they fully
+lied within the study area, thus enforcing individuals to always remain within
+the study area. Given that only very few simulated individuals were repelled by
+a virtual map boundary, we conclude that the two measures effectively obviated
+boundary effects.
 
 \subsection{Conclusion}
-To this end, we proposed and applied a simple three-step workflow that uses
-ISSFs to parametrize an individual-based movement model from which dispersal can
-be simulated with the purpose of informing about landscape connectivity. By
-explicitly simulating dispersal trajectories, simulations enable a more
-mechanistic understanding of connectivity, and overcome several of the
-unrealistic assumptions inherent to graph-based connectivity models, such as
-leas-cost analysis or circuit theory. We exemplified the application of the
-proposed workflow using data of dispersing wild dogs to assess landscape
-connectivity for the species within the KAZA-TFCA ecosystem. With this, we hope
-to have sparked interest in the powerful framework of step selection functions
-for investigating dispersal behavior and landscape connectivity. Nevertheless,
-we propose to view simulations from ISSF-models as complementary and not as
-substitutes to traditional connectivity modeling techniques.
-
-%------------------------------------------------------------------------------
-%  Further Results to Report
-%------------------------------------------------------------------------------
-% % Some step length, turning angle comparisons
-% On average, step lengths realized by the simulated dispersers (\(\mu_{sl} =
-% 2'093\) m, \(\sigma_{sl} = 3'067\)) were slightly shorter than those by observed
-% dispersers (\(\mu_{sl} = 2'326\) m, \(\sigma_{sl} = 3'323\)) and simulated
-% dispersers moved marginally less directional (\(\mu_{cos(ta)} = 0.057\),
-% \(\sigma_{cos(ta)} = 0.071\)) compared to observed dispersers (\(\mu_{cos(ta)} =
-% 0.078\), \(\sigma_{cos(ta)} = 0.072\)). These differences in step lengths and
-% turning angles can be attributed to minor disparities between habitat conditions
-% at the area within which we collected training data and habitat conditions
-% within the entire study area.
-
-%------------------------------------------------------------------------------
-%  Further Topics to Discuss
-%------------------------------------------------------------------------------
-
-% % Simulations for an early warning system?
-% Finally, simulations from ISSFs could be utilized as forecasting tool to predict
-% the likely whereabouts of GPS collared animals into the near future. In some
-% European countries, the reintroduction of large predators, such as bears, lynx,
-% and wolves, has triggered emotional discussions and raised public concern
-% \citep{Behr.2017}, particularly in areas where locals depend on income from
-% free-roaming livestock. An early warning system based on simulations could thus
-% serve to inform about potential encounters with large carnivores, such that
-% farmers can secure their livestock accordingly. This would hopefully increase
-% public acceptance of large predators.
-
-% % Downsides of connectivity?
-% Even though connectivity is in general thought to promote population viability,
-% it is also related to various aspects that may cause ecological damage, such as
-% increased connectivity into human-dominated areas or facilitated spread of
-% deadly diseases \citep{Brearly.2013}. Nevertheless, information on the
-% relationship between connectivity and the prevalence of deadly diseases in
-% African wild dogs is currently lacking.
-
-% % Social landscape
-% We have previously attributed the weak significance of distance to water to the
-% fact that we did not control for the presence or absence of conspecifics. We
-% stick to this reasoning as our expanded model still shows a rather large
-% uncertainty around the respective beta coefficients. To better gauge the
-% importance and influence of this covariate, future studies will need to control
-% for inter- and intra-specific interactions that may explain why and when
-% dispersers are attracted to or afraid of waterbodies. \cite{Fortin.2005}, for
-% instance, found that elk movement was significantly impacted by the density of
-% wolf in the area, such that habitat preferences strictly differed depending on
-% the presence or absence of wolves. The decision to settle is likely related to
-% the presence or absence of conspecifics. Hence, the exact dispersal duration and
-% distance will not be independent of current wild dog densities. In dispersing
-% wolves, for instance, the longest dispersal distances have been observed in
-% low-density populations \citep{Boyd.2005, Wabakken.2007}. The dispersal duration
-% may thus be determined by the by the amount of isolation between subpopulations
-% and population densities \citep{DaviesMostert.2012}
-
-% % Can generate multiple metrics
-% Another benefit of simulations from ISSFs is that their output can be processed
-% using a diverse suite of approaches, each enabling to focus on a different
-% aspect of connectivity. the results from LCPA and CT, on the other hand, are
-% usually restricted to single ``conductance'' maps. By turning simulated
-% trajectories into a network, for instance, researchers can capitalize on
-% concepts from network theory and calculate network metrics that are pertinent to
-% landscape connectivity \citep{BastilleRousseau.2018}. Here, we exemplified
-% this by calculating the betweenness metric, highlighting major movement
-% corridors. Alternatively, can learn about inter-patch connectivity and potential
-% source-sink dynamics between patches \citep{Ferreras.2001, Revilla.2004,
-% Kanagaraj.2013} based on simulated dispersal trajectories. Ultimately, by
-% accounting for habitat- and movement preferences of the focal species,
-% simulations from ISSFs are capable of rendering the behavioral ecology of the
-% focal species in detail, thereby permitting a more functional view on
-% connectivity that is not solely driven by the landscape structure and distances
-% between patches \citep{Gustafson.1996, Gardner.2004, Graf.2007,
-% KramerSchadt.2004, Revilla.2004, Revilla.2008, Kanagaraj.2013}.
-
-% % We did not validate all of our predictions
-% Even though our simulations generated several interesting results, most of them
-% need to be validated using independent data. We learned that between some
-% national parks asymmetries between the number of ingoing and outgoing dispersers
-% exist, suggesting that some patches act as sinks, whereas others serve as
-% sources for dispersers.
-
-% % Modeling Mortality
-% Although we did not render mortality, animals regularly die during dispersal,
-% mainly due to deadly encounters with predators, road kills, and persecution by
-% humans \citep{Bonnet.1999, Woodroffe.2012, Behr.2021b}. Mortality during
-% dispersal could therefore substantially limit functional connectivity
-% \citep{Bowler.2009}, especially in areas where the likelihood of encountering
-% competitors and humans is high \citep{Cozzi.2020}. If corresponding information
-% is available, mortality can and should be included in ISSF simulations.
-
-% % Spatially Realistic Population Models
-% The ability to realistically render movement during dispersal not only serves to
-% investigate landscape connectivity, but also forms the foundation for more
-% realistic spatially explicit population models in which dispersal is not merely
-% rendered through dispersal kernels or cellular automata movements
-% \citep{Visintin.2020}, but mechanistically based on observed movement and
-% habitat preferences (e.g. \citealp{Revilla.2008}, \citealp{Kleinmann.2017}).
-% Such models can ultimately be employed to conduct population viability analyses
-% \citep{Boyce.1992} in which species' dispersal abilities are taken into account.
-
-% % Modeling Decisions
-% \begin{figure}
-%   \includegraphics[width=\textwidth]{99_ModelingDecisions.png}
-%   \caption{Five major modeling decisions that a researcher needs to consider
-%   when simulating dispersers to assess landscape connectivity. (1) Number of
-%   simulated trajectories. (2) Handling individual variation: we used point
-%   estimates when simulating dispersers, assuming that there was no individual
-%   variation. Alternatively, however, one could draw preferences for each
-%   simulated individually based on model uncertainty. (3) Dispersal duration.
-%   While one could draw the number of simulated steps based on observed dispersal
-%   events, we opted for an alternative and simulated each individual for 2'000
-%   steps, which was at the upper end of observed dispersal durations. This allows
-%   to easily shorten the generated trajectories afterwards and to investigate the
-%   sensitivity of results with regards to the dispersal duration. (4) Boundary
-%   behavior. We allowed dispersers to leave and re-enter the main study area
-%   through a buffer zone with randomized covariate values. Alternatively, one
-%   could discard transgressing random steps, thereby forcing dispersers to remain
-%   within the study area or simply terminate the simulation, assuming the
-%   individual has disappeared. (5) Source point location. The location of source
-%   points should optimally be biologically informed. For our simulation, we
-%   placed source points within protected areas large enough to sustain viable
-%   wild dog populations. Conceivable alternatives include placing source points
-%   according to a habitat suitability model or based on abundance information.
-%   Importantly, one must also consider potential immigrants from outside the main
-%   study area.}
-%   \label{ModelingDecisions}
-% \end{figure}
+To this end, we proposed and applied a simple three-step workflow that relies on
+ISSF-analysis and enables to simulate dispersal and assess landscape
+connectivity. The proposed workflow overcomes several of the conceptual
+shortcomings inherent to LCPA and CT, such as the assumption of known endpoints,
+and provides a highly flexible tool for investigating connectivity. In our case
+study, the workflow proved useful to investigate landscape connectivity for the
+endangered African wild dog within the KAZA-TFCA ecosystem and allowed us to
+pinpoint frequently used corridors and dispersal hubs. We hope to have sparked
+interest in the powerful framework of ISSFs for investigating dispersal
+connectivity, while at the same time conferring some of the non-trivial modeling
+decisions involved.
 
@@
 \section{Data Availability}
-GPS movement data of dispersing wild dogs will be made available on dryad at the
-time of publication. Access to R-scripts that exemplify the application of the
-proposed framework to simulated data are provided through Github.
+GPS movement data of dispersing wild dogs is available on dryad
+\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
+proposed workflow using simulated data are provided through Github
+(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
+codes required for the African wild dog case study will be made available
+through an online repository at the time of publication.
 
@@
 Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
-to Johannes Signer for assisting with the simulation algorithm. This study was
-funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
+for sharing her data of three dispersing wild dogs. Furthermore, we would like
+to thank Johannes Signer for assisting with the simulation algorithm. This study
+was funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
 Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung

### 2021-12-13 Chapter_2/02_R-Scripts/01_DataCleaning/02_LandCover.r

@@ -1,0 +1,123 @@
+################################################################################
+#### Preparation of Globeland Land Cover Map
+################################################################################
+# Description: Simplify and resample of the Globeland Land Cover Dataset
+
+# Clear R's brain
+rm(list = ls())
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Load packages
+library(terra)       # To handle raster data
+library(tidyverse)   # For data wrangling
+
+# Load the globeland land cover map
+glob <- rast("03_Data/01_RawData/GLOBELAND/Globeland.tif")
+
+# Load the reference shapefile and raster
+s <- vect("03_Data/02_CleanData/00_General_Shapefile.shp")
+r <- rast("03_Data/02_CleanData/00_General_Raster.tif")
+
+# Crop the merged tiles to our reference shapefile
+glob <- crop(glob, s)
+
+# Aggregate to coarser resolution
+fact <- res(r)[1] / res(glob)[1]
+glob <- terra::aggregate(glob, fact = round(fact), fun = "modal")
+
+# Check out the distribution of values
+freq(glob)
+
+# Prepare reclassification table
+info <- data.frame(
+    Code = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 255)
+  , Class = c(
+      "CultivatedLand"
+    , "Forest"
+    , "Grassland"
+    , "Shrubland"
+    , "Wetland"
+    , "WaterBodies"
+    , "Tundra"
+    , "ArtificialSurfaces"
+    , "Bareland"
+    , "PermanentIceSnow"
+    , "NA"
+  )
+  , CodeNew = c(3, 4, 6, 5, 1, 1, 6, 2, 7, 7, 6)
+  , ClassNew = c(
+      "Cropland"
+    , "Forest"
+    , "Grassland"
+    , "Shrubs"
+    , "Water"
+    , "Water"
+    , "Grassland"
+    , "Urban"
+    , "Bare"
+    , "Bare"
+    , "Grassland"
+  )
+)
+
+# Arrange
+info <- arrange(info, CodeNew)
+
+# Assign a color to the new classes
+info$Color[info$CodeNew == 1] <- "#0e16ed"
+info$Color[info$CodeNew == 2] <- "#e6494b"
+info$Color[info$CodeNew == 3] <- "#f8f261"
+info$Color[info$CodeNew == 4] <- "#1e7b40"
+info$Color[info$CodeNew == 5] <- "#cca12b"
[…]

### 2021-12-13 Chapter_2/02_R-Scripts/01_DataCleaning/12_Agriculture(Globeland).r

@@ -1,49 +1,0 @@
-################################################################################
-#### Resampling and Cropping of the Globeland Agriculture Data
-################################################################################
-# Description: We can use globeland's agricultural fields to augment our
-# croplands layer. In this script I therefore extract agricultural fields so we
-# can merge them.
-
-# Clear R's brain
-rm(list = ls())
-
-# Set the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(raster)   # To handle raster data
-
-# Make use of multiple cores
-beginCluster()
-
-# Import the Globelands dataset
-crops <- raster("03_Data/01_RawData/GLOBELAND/Globeland.tif")
-
-# Load the reference raster
-r <- raster("03_Data/02_CleanData/00_General_Raster.tif")
-
-# Keep only crops
-crops <- crops == 10
-
-# Aggregate the globelands dataset to match the resolution of the reference
-# raster
-fact <- res(r)[1] / res(crops)[1]
-crops <- aggregate(crops, fact = round(fact), fun = max)
-
-# Resample the layer to match the reference raster
-crops_res <- resample(crops, r, "ngb")
-
-# Check NAs
-sum(is.na(values(crops_res)))
-
-# Save the result to file
-writeRaster(
-    x         = crops_res
-  , filename  = "03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_GLOBELAND.tif"
-  , overwrite = TRUE
-)
-
-# End cluster
-endCluster()

### 2021-12-13 Chapter_2/02_R-Scripts/01_DataCleaning/16_HumanDensity(WorldPop).r

@@ -1,71 +1,0 @@
-################################################################################
-#### Cleaning Worldpop Human Density Data
-################################################################################
-# Description: In this script I'll clean the population density estimates
-# downloaded from here: https://www.worldpop.org/
-
-# Clear R's brain
-rm(list = ls())
-
-# Set working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(raster)     # To handle raster data
-library(rgdal)      # To handle spatial data
-
-# Make use of multicore abilities
-beginCluster()
-
-################################################################################
-#### Stitch Tiles
-################################################################################
-# Load the data
-dat <- dir(
-    path       = "03_Data/01_RawData/WORLDPOP"
-  , full.names = T
-  , pattern    = ".tif$"
-)
-dat <- lapply(dat, raster)
-
-# Merge all tiles together
-merged <- do.call(merge, dat)
-
-# Load the reference shapefile and raster
-s <- shapefile("03_Data/02_CleanData/00_General_Shapefile.shp")
-r <- raster("03_Data/02_CleanData/00_General_Raster.tif")
-
-# Crop the merged tiles to our reference shapefile
-merged <- crop(merged, s)
-
-# Store the raster to file
-writeRaster(
-    x         = merged
-  , filename  = "03_Data/01_RawData/WORLDPOP/HumanDensity.tif"
-  , overwrite = TRUE
-)
-
-################################################################################
-#### Aggregate and Resample
-################################################################################
-# Aggregate the layer to 250m
-fact <- res(r)[1] / res(merged)[1]
-coarse <- aggregate(merged, fact = round(fact), fun = sum)
-
-# Resample the population density layer to the reference raster
-coarse <- resample(coarse, r, "bilinear")
-
-# Replace NAs and values below 0 with 0s
-coarse <- reclassify(coarse, rcl = c(NA, NA, 0))
-coarse <- reclassify(coarse, rcl = c(-Inf, 0, 0))
-
-# Store the result
-writeRaster(
-    x         = coarse
-  , filename  = "03_Data/02_CleanData/04_AnthropogenicFeatures_HumanDensity_WORLDPOP.tif"
-  , overwrite = TRUE
-)
-
-# End cluster
-endCluster()

### 2021-12-13 Chapter_2/02_R-Scripts/01_DataCleaning/18_Villages(Facebook).r

@@ -1,61 +1,0 @@
-################################################################################
-#### Identify Villages from Facebook Data
-################################################################################
-# Description: In this script I'll use the worldpop human density data to
-# identify villages
-
-# Clear R's brain
-rm(list = ls())
-
-# Set working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(raster)     # To handle raster data
-library(terra)      # To handle raster data
-library(rgdal)      # To handle vector data
-library(rgeos)      # To manipulate vector data
-
-# Load the human density estimates from worldpop
-dens <- rast("03_Data/01_RawData/FACEBOOK/HumanDensity.tif")
-
-# Remove NAs
-dens <- classify(dens, matrix(c(NA, 0), ncol = 2))
-
-# Aggregate to a coarser resolution (should be equal to worldpop)
-dens <- aggregate(dens, fact = 3, fun = sum)
-
-# Remove pixels that are inhabited by less than two people
-dens_bin <- dens > 2
-
-# Aggregate to even coarser resolution
-dens_bin <- aggregate(dens_bin, fact = 3, fun = max, na.rm = T)
-
-# Apply focal filter
-dens_bin <- focal(dens_bin, w = 3, fun = modal)
-
-# Remove 0s
-dens_bin[dens_bin == 0] <- NA
-
-# Coerce raster to polygon
-pols <- as.polygons(dens_bin)
-
-# Store polygons to a file and then reload
-file <- tempfile(fileext = ".shp")
-writeVector(pols, file, overwrite = T)
-pols <- readOGR(file)
-
-# Apply small buffer
-pols <- gBuffer(pols, width = 0.5 / 111)
-
-# Rasterize them
-r <- rast("03_Data/02_CleanData/00_General_Raster.tif")
-pols <- rasterize(vect(pols), r, field = 1, background = 0)
-
-# Store to file
-writeRaster(
-    x         = raster(pols)
-  , filename  = "03_Data/02_CleanData/04_AnthropogenicFeatures_Villages_FACEBOOK.tif"
-  , overwrite = T
-)

### 2022-01-05 Chapter_1/04_Manuscript/Manuscript.tex
Many updates

@@
 %------------------------------------------------------------------------------
-\title{Step by Step: Using Step Selection Functions to Simulate Dispersal and
-Assess Landscape Connectivity}
-
-% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
-% Connectivity}
+\title{A Three Steps Workflow for Simulating Dispersal Trajectories and
+Assessing Landscape Connectivity: A Case Study on the Endangered African Wild
+Dog}
 
@@
 % Current Date
-\date{\today}
+% \date{\today}
 
@@
 
-\textsuperscript{2} Botswana Predator Conservation, Private Bag 13, Maun,
-Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
+Maun, Botswana.
 
@@
 
-\textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
-Assess Landscape Connectivity
+\textbf{Running Title:} Simulating Wild Dog Dispersal Trajectories to Assess
+Landscape Connectivity
 
@@
 %------------------------------------------------------------------------------
+% \newpage
+% \begin{abstract}
+% %------------------------------------------------------------------------------
+% % Short Version
+% %------------------------------------------------------------------------------
+% \textbf{Short Version }
+% The ability to disperse is contingent on a sufficient degree of landscape
+% connectivity, which is why the identification and preservation of movement
+% corridors that promote connectivity has become a task of extraordinary
+% importance. Currently, ecologists rely on least-cost analysis or circuit theory
+% to investigate connectivity, although both methods make several assumptions that
+% are hardly met in reality. To address these issues, simulations from
+% individual-based movement models have been proposed, yet a unified framework to
+% simulate dispersal and quantify connectivity is lacking.
+%
+% Here, we propose a simple three-step workflow that combines several
+% already-existing methods to assess connectivity using simulated dispersal
+% trajectories. In step one, we use integrated step selection functions to
+% parametrize a mechanistic movement model rendering dispersal behavior. In step
+% two, we apply the parametrized model as an individual-based movement model to
+% simulate dispersal trajectories. In step three, we combine simulated
+% trajectories into three complementary connectivity maps, each focusing on a
+% different aspect of landscape connectivity.
+%
+% We showcase the application of the proposed workflow using empirical data of
+% dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
+% connectivity within the world's largest transboundary conservation area in
+% Southern Africa. We thereby shed light into dispersing wild dogs' habitat and
+% movement preferences, while also uncovering crucial dispersal corridors. With
+% this analysis, we demonstrate that simulations from integrated step selection
+% functions offer a simple, yet powerful alternative to traditional connectivity
+% modeling techniques.
+% \end{abstract}
+
 \newpage
@@
 %------------------------------------------------------------------------------
-% Short Version
-%------------------------------------------------------------------------------
-\textbf{Short Version }
-The ability to disperse is contingent on a sufficient degree of landscape
-connectivity, which is why the identification and preservation of movement
-corridors that promote connectivity has become a task of extraordinary
-importance. Currently, ecologists rely on least-cost analysis or circuit
-theory to investigate connectivity, albeit both methods make several
-assumptions that are hardly met in reality. To address these issues,
-simulations from individual-based movement models have been proposed, yet a
-unified framework to simulate dispersal and quantify connectivity is lacking.
-
-Here, we propose a simple three-step workflow that combines several
-already-existing methods to assess connectivity using simulated dispersal
-trajectories. In step one, we use integrated step selection functions to
-parametrize a mechanistic movement model rendering dispersal behavior. In step
-two, we apply the parametrized model as an individual-based movement model to
-simulate dispersal trajectories. In step three, we combine simulated
-trajectories into three complementary connectivity maps, each focusing on a
-different aspect of landscape connectivity.
-
-We showcase the application of the proposed workflow using data of dispersing
-African wild dogs (\textit{Lycaon pictus}) and assess landscape connectivity
-within the world's largest transboundary conservation area in Southern Africa.
-We thereby shed light into dispersing wild dogs' habitat and movement
-preferences, while also uncovering crucial dispersal corridors. With this
-analysis, we demonstrate that simulations from integrated step selection
-functions offer a simple, yet powerful alternative to traditional connectivity
-modeling techniques.
-\end{abstract}
-
-\newpage
-\begin{abstract}
-%------------------------------------------------------------------------------
 %  LONG VERSION
 %------------------------------------------------------------------------------
-\textbf{Long Version }
-Dispersal of individuals is crucial for long-term species persistence and
-depends on a sufficient degree of landscape connectivity. To date, researchers
-primarily investigate connectivity using least-cost analysis and circuit theory,
-both methods that make assumptions which are hardly met in reality. Least-cost
-analysis assumes that animals move towards a known endpoint and are
+Dispersal of individuals contributes to long-term population persistence, yet
+requires a sufficient degree of landscape connectivity. To date, connectivity
+has mainly been investigated using least-cost analysis and circuit theory, two
+methods that make assumptions that are hardly applicable during dispersal.
+Least-cost analysis assumes that animals move towards a known endpoint and are
 knowledgeable about the most favorable route to reach it. Circuit theory
-presumes a complete random walk and therefore fails to incorporate directional
-persistence. These shortcomings can be overcome by spatio-temporally explicitly
-simulating dispersal, yet a unified approach for such simulations is lacking.
-
-Here, we present a simple three-step workflow to simulate dispersal and assess
-connectivity starting from empirical GPS movement data. In step one, we use
-integrated step selection functions to fit a mechanistic movement model
-describing habitat and movement preferences of dispersers. In step two, we apply
-the parameterized model to simulate individual dispersal trajectories. In step
-three, we combine the simulated trajectories into three complementary
-connectivity maps: a heatmap that highlights frequently traversed areas, a
-betweenness map that pinpoints dispersal corridors, and a map of inter-patch
-connectivity that indicates the presence and intensity of functional links
-between habitat patches. As a case study, we analyse data on the endangered
-African wild dog and assess landscape connectivity in the Kavango-Zambezi
-Transfrontier Conservation Area, the world’s largest transboundary conservation
-area.
-
-Our results show that wild dogs preferably disperse with directional persistence
-in the vicinity to water bodies and through areas with little human influence and
-sparse woodland cover. Dispersal simulations and subsequent connectivity maps
-revealed several dispersal hotspots and corridors across the extent of the
-KAZA-TFCA. Connectivity between NPs inside the KAZA-TFCA was good, yet few
-dispersers successfully moved from Zambia's NPs into neighboring areas.
-
-We show that simulations from step-selection functions offer a simple yet
-powerful alternative to traditional connectivity modeling techniques. In
-contrast to traditional connect modeling techniques, simulations not only make
-fewer unrealistic assumptions but also permit a more mechanistic understanding
-of dispersal and landscape connectivity. This makes our workflow useful for a
-variety of applications in ecological, evolutionary, and conservation research.
+presumes a complete random walk and fails to incorporate directional
+persistence. While these assumptions can be relaxed by simulating dispersal
+movements spatio-temporally explicitly across the landscape of interest, a
+unified approach for such simulations is lacking.
+
+Here, we present a simple three-step workflow to simulate dispersal movements
+and to assess connectivity starting from empirical GPS movement data. In step
+one, we use integrated step selection functions to fit a mechanistic movement
+model describing habitat and movement preferences of dispersing individuals. In
+step two, we apply the parameterized movement model to simulate dispersal
+trajectories. In step three, we derive three complementary connectivity maps: a
+heatmap that highlights frequently traversed areas, a betweenness map that
+pinpoints dispersal corridors, and a map of inter-patch connectivity that
+indicates the presence and intensity of functional links between habitat
+patches. As a case study, we applied the workflow to GPS data collected during
+dispersal of the endangered African wild dog (\textit{Lycaon pictus}) inhabiting
+the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
+
+Results from the case study show that wild dogs preferably disperse with
+directional persistence in the vicinity to water bodies, through areas with
+little human influence and sparse woodland cover. Dispersal simulations and
+subsequent connectivity maps revealed several dispersal hotspots and corridors
+across the extent of the KAZA-TFCA. Connectivity between NPs inside the
+KAZA-TFCA was good, yet few dispersers successfully moved from Zambia's NPs into
+neighboring areas. By rendering interactions between movement preferences and
+habitat preferences, our movement model substantially outperformeed a previous
+model that omitted such interactions. While we simulated 80,000 dispersal
+trajectories in total, we found that metrics of connectivity only marginally
+changed beyond 10,500 simulations, suggesting that fewer simulations would have
+sufficed.
+
+Ultimately, we show that a simulation-based approach that leverages on
+step-selection functions offers a simple yet powerful alternative to traditional
+connectivity modeling techniques. Such an approach not only makes fewer
+biologically unrealistic assumptions but also permits a more mechanistic
+understanding of dispersal and landscape connectivity. Our workflow is thus
+useful for a variety of applications in ecological, evolutionary, and
+conservation research.
 
@@
 dispersal corridors that promote connectivity a task of fundamental importance
-\citep{Nathan.2008, Doerr.2011, Rudnick.2012}. The identification of dispersal
-corridors not only necessitates a comprehensive understanding of the factors
-that limit dispersal, but also an appropriate model to estimate connectivity
-\citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the two most
-commonly used connectivity models are least-cost path analysis (LCPA;
+\citep{Nathan.2008, Doerr.2011, Rudnick.2012}. Identifying dispersal corridors
+not only necessitates a comprehensive understanding of the factors that limit
+dispersal movements, but also an appropriate model to estimate connectivity
+\citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the most commonly
+used models to assess connectivity are least-cost path analysis (LCPA;
 \citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
-McRae.2008}), both graph-based methods that quantify conductance of the
-landscape based on habitat permeability \citep{Zeller.2012}. However, both
-approaches rest on assumptions that appear unsuitable for dispersers, which is
-why simulating dispersal from individual-based movement models provides a
-promising alternative approach to investigate connectivity \citep{Diniz.2019}.
+McRae.2008}). Unfortunately, both models rest on assumptions that appear
+unsuitable for dispersers, calling for the development of alternative
+approaches. One promising alternative to investigate dispersal and assess
+connectivity is to simulate dispersal from individual-based movement models
+\citep{Diniz.2019}.
 
@@
 preconceived endpoint and choose a cost-minimizing route accordingly
-\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be fulfilled by
+\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be justifiable for
 migrating animals, it is unlikely to hold for dispersers, as dispersers
-typically move across unfamiliar territory, so the associated movement costs and
-potential endpoints are unknown to them \citep{Koen.2014, Cozzi.2020}. CT, on
-the contrary, posits that animals move according to a random walk, entailing
-that autocorrelation between subsequent movements cannot be rendered
-\citep{Diniz.2019}. For dispersers, however, autocorrelated movements are
-regularly observed \citep{Cozzi.2020, Hofmann.2021}, implying that dispersal
-trajectories are usually strongly directional. Moreover, graph-based methods are
-unable to reflect the temporal dimension of dispersal, so that statements about
-the expected duration for moving between habitats are impossible
-\citep{Martensen.2017, Diniz.2019}.
+typically move across unfamiliar territory towards an unknown endpoint
+\citep{Koen.2014, Cozzi.2020}. CT, on the contrary, posits that animals move
+according to a random walk, entailing that autocorrelation between subsequent
+movements cannot be rendered \citep{Diniz.2019}. For dispersers, however,
+autocorrelated movements are regularly observed \citep{Cozzi.2020,
+Hofmann.2021}, implying that dispersal trajectories are usually strongly
+directional. Moreover, because both models require static permeability or
+resistance surfaces as input, they are unable to reflect the temporal dimension
+of dispersal, meaning that statements about the expected duration for moving
+between habitat patches are impossible \citep{Martensen.2017, Diniz.2019}.
 
@@
 The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
-from individual-based movement models (IBMMs) and converting simulated
+from individual-based movement models (IBMMs) and by converting simulated
 trajectories into meaningful measures of connectivity \citep{Diniz.2019}). In
-contrast to graph-based methods, IBMMs allow to explicitly simulate how
-individuals move across and interact with the encountered landscape
-\citep{Kanagaraj.2013, Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020}, as
-well as to render potential interactions between movement behavior and habitat
-conditions, thus shifting the focus from a structural to a more functional view
-on connectivity \citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}.
-Furthermore, because simulations from IBMMs generate movement sequentially, the
-temporal dimension of dispersal becomes explicit and allows to model
-autocorrelation between subsequent movements \citep{Diniz.2019}. Finally,
-simulations from IBMMs do not enforce connections towards preconceived
-endpoints, thereby mitigating biases arising from misplaced endpoints. Despite
-these advantages, a unified framework to simulate dispersal and assess
-connectivity using IBMMs is lacking.
+contrast to LCPA and CT, IBMMs allow to explicitly simulate how individuals move
+across and interact with the encountered landscape \citep{Kanagaraj.2013,
+Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020}, as well as to render
+potential interactions between movement behavior and habitat conditions
+\citep{Avgar.2016}. This strictly shifts the focus from a structural to a more
+functional view on connectivity \citep{Tischendorf.2000, Kanagaraj.2013,
+Hauenstein.2019}. Furthermore, IBMMs generate movement sequentially, i.e. they
+generate a series of steps, the temporal dimension of dispersal movements
+becomes explicit and allows to model autocorrelation between consecutive steps
+\citep{Diniz.2019}. Finally, simulations from IBMMs do not enforce movement or
+connections towards preconceived endpoints, thereby preventing biases arising
+from misplaced endpoints. Despite these advantages, a unified framework to
+simulate dispersal and assess connectivity using IBMMs is lacking.
 
@@
 \subsection{Proposed Solution: Three-Step Workflow}
-Here, we draw on recent advancements in movement ecology and propose a simple
-three-step workflow for simulating dispersal and assessing landscape
-connectivity (\Cref{GraphicalAbstract}). In step one, we combine GPS movement
-data of dispersing individuals with relevant habitat covariates to fit a
-mechanistic movement model using integrated step selection functions (ISSFs,
-\citealp{Avgar.2016}). ISSFs permit inference on the focal species' habitat
-kernel (i.e. habitat preferences), its movement kernel (i.e. movement
-preferences/capabilities), and potential interactions between the two
+Here, we propose a simple three-step workflow for simulating dispersal and
+assessing landscape connectivity (\Cref{GraphicalAbstract}). In step one, we
+combine GPS movement data of dispersing individuals with relevant habitat
+covariates to fit a mechanistic movement model using integrated step selection
+functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the study
+species' habitat kernel (i.e. habitat preferences), its movement kernel (i.e.
+movement preferences/capabilities), and potential interactions between the two
 \citep{Avgar.2016, Fieberg.2021}. In step two, we simulate individual dispersal
-trajectories using the parametrized movement from step one. Comparable
-simulations have already been employed to estimate steady-state utilization
-distributions of residents \citep{Potts.2013, Signer.2017} and to model
-landscape connectivity, yet disregarding interdependencies between habitat and
-movement kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we
-convert simulated trajectories into three complementary connectivity maps, each
+trajectories using the parametrized movement model. Comparable simulations have
+already been applied to estimate steady-state utilization distributions of
+resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
+connectivity, yet disregarding interdependencies between habitat and movement
+kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert
+simulated trajectories into three complementary connectivity maps, each
 highlighting a different aspect of connectivity. The set of maps includes (i) a
-heatmap revealing areas that are frequently traversed by simulated dispersers
-(e.g. \citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map
-delineating dispersal corridors and bottlenecks (e.g.
-\citealp{BastilleRousseau.2018}), (iii) and a map of inter-patch connectivity,
-depicting the presence and intensity of use of specific connections, as well as
-the average dispersal duration required to connect corresponding habitat patches
-(e.g. \citealp{Gustafson.1996, Kanagaraj.2013}).
+heatmap revealing areas that are frequently traversed by dispersers (e.g.
+\citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map delineating
+dispersal corridors and bottlenecks (e.g. \citealp{BastilleRousseau.2018}),
+(iii) and a map of inter-patch connectivity, depicting the presence and
+intensity of specific connections, as well as the average dispersal duration
+required to connect corresponding habitat patches (e.g. \citealp{Gustafson.1996,
+Kanagaraj.2013}).
 
@@
 We showcase the application of the proposed workflow (\Cref{GraphicalAbstract})
-using dispersal data collected on the endangered African wild dog
-(\textit{Lycaon Pictus}). The African wild dog represents a highly mobile
-species whose persistence hinges on a sufficient degree of landscape
-connectivity. Once common throughout sub-Saharan Africa, this species has
-disappeared from much of its historic range, largely due to human persecution,
-habitat fragmentation, and disease outbreaks \citep{Woodroffe.2012}. Wild dogs
-typically disperse in single-sex coalitions \citep{McNutt.1996, Behr.2020} and
-are capable of dispersing several hundred kilometers \citep{DaviesMostert.2012,
-Masenga.2016, Cozzi.2020}. Although previous studies have investigated
-connectivity using LCPA \citep{Hofmann.2021} or CT \citep{Brennan.2020}, a more
-comprehensive and mechanistic understanding of connectivity is missing for this
-species (but see \citealp{Creel.2020}). With fewer than 6,000 free-ranging wild
-dogs remaining in fragmented populations \citep{Woodroffe.2012}, reliable
-information on landscape connectivity is essential for the conservation of this
-endangered carnivore. Here, we use GPS data of 16 dispersing wild dogs
-originating from a free-ranging population in northern Botswana to parametrize a
-mechanistic movement model, which we then employed to simulate 80,000 dispersal
-trajectories across the landscape of the world's largest transboundary
-conservation area, the Kavango-Zambezi Transfrontier Conservation Area
-(KAZA-TFCA). We anticipated that simulations based on our three-step workflow
-would overcome several of the above highlighted conceptual shortcomings of
-traditional connectivity models and provide a more comprehensive view of
-landscape connectivity
+using GPS movement data collected on dispersing individuals of the endangered
+African wild dog (\textit{Lycaon Pictus}). The African wild dog is a highly
+mobile species whose population persistence heavily relies on the availability
+of large, natural or semi-natural landscapes and a sufficient degree of
+connectivity among remaining subpopulations. Once common throughout sub-Saharan
+Africa, this species has disappeared from much of its historic range, largely
+due to human persecution, habitat fragmentation, and disease outbreaks
+\citep{Woodroffe.2012}. Wild dogs typically disperse in single-sex coalitions
+\citep{McNutt.1996, Behr.2020} and are capable of dispersing several hundred
+kilometers \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}. Although
+previous studies have investigated connectivity using LCPA \citep{Hofmann.2021}
+or CT \citep{Brennan.2020}, a more comprehensive and mechanistic understanding
+of connectivity is missing for this species (but see \citealp{Creel.2020}). With
+fewer than 6,000 free-ranging wild dogs remaining in fragmented subpopulations
+\citep{Woodroffe.2012}, reliable information on landscape connectivity is
+essential for the conservation of this endangered carnivore. Here, we use GPS
+data of 16 dispersing wild dogs originating from a free-ranging population in
+northern Botswana to parametrize a mechanistic movement model, which we then
+employed to simulate 80,000 dispersal trajectories across the landscape of the
+world's largest transboundary conservation area, the Kavango-Zambezi
+Transfrontier Conservation Area (KAZA-TFCA). We anticipated that simulations
+based on our three-step workflow would overcome several of the above highlighted
+conceptual shortcomings of traditional connectivity models and provide a more
+comprehensive view on dispersal behavior and landscape connectivity.
 
@@
 \subsection{Study Area}
-The study area centered at -17\degree 13'9''S, 23\degree 56'4''E
-(\Cref{StudyArea}a) was located in southern Africa (\Cref{StudyArea}a) and
-spanned more than 1.3 Mio. km\textsuperscript{2}, encompassing the entire
-KAZA-TFCA (\Cref{StudyArea}b). The KAZA-TFCA comprises parts of Angola,
-Botswana, Namibia, Zimbabwe, and Zambia and hosts a rich diversity of
-landscapes, ranging from savannah to grassland and from dry to moist woodland
-habitats. In its center lies the Okavango Delta, a dominant hydro-geographical
-feature and the world's largest flood-pulsing inland delta. Large portions of
-the KAZA-TFCA are part of designated national parks (NPs) and other protected
-areas, yet considerable human influence through roads, agricultural sites, and
-settlements remains.
+Our simulation of dispersal trajectories and assessment of connectivity spanned
+across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA),
+an area of approximately 1.3 Mio. km\textsuperscript{2} (\Cref{StudyArea}a and
+b). The KAZA-TFCA is the world's largest transboundary conservation area and
+comprises parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia, thus hosting
+a rich diversity of landscapes, ranging from savannah to grassland and from dry
+to moist woodland habitats. In its center lies the Okavango Delta, a dominant
+hydro-geographical feature and the world's largest flood-pulsing inland delta.
+Large portions of the KAZA-TFCA are formally protected in the form of national
+parks (NPs) or other protected areas, yet considerable human-dominated
+landscapes (e.g. roads, agricultural sites, and settlements) remain.
 
@@
 \subsubsection{GPS Data}
-We collected GPS movement data on 16 dispersing wild dogs (7 females and 9
-males) between 2011 and 2019 from a free-ranging population in northern Botswana
-(details on the data collection can be found in \cite{Cozzi.2020} and
-\cite{Hofmann.2021}). Because behavior during dispersal is more pertinent to
-landscape connectivity than behavior during residence \citep{Elliot.2014,
-Abrahms.2017}, we discarded data collected when individuals were resident. We
-determined the exact timing of emigration and settlement using direct field
-observations and through visual inspection of the net squared displacement (NSD)
-metric. NSD measures the squared Euclidean distance of a GPS relocation to a
-reference point \citep{Borger.2012}, which we set to the center of each
-dispersers' natal home range. Thus, dispersal was deemed to have started once an
-individual left its natal home range and ended once the NSD metric remained
-constant, indicating settlement. During dispersal, GPS collars recorded a fix
-every 4 hours and regularly transmitted data over the Iridium satellite system.
-To ensure regular timespans between GPS fixes, we removed any fix that was not
-successfully obtained on the aspired 4-hourly schedule (allowing for a tolerance
-of \( \pm \) 15 minutes). We then converted the remaining fixes (n = 4'169) into
+We collected GPS movement data on 16 coalitions of dispersing African wild dogs
+(7 female and 9 male coalitions) between 2011 and 2019 from a free-ranging
+population in northern Botswana (details on the data collection can be found in
+\cite{Cozzi.2020} and \cite{Hofmann.2021}). To delineate periods of dispersal,
+we determined the exact timing of emigration from the natal pack and settlement
+in a new territory using direct field observations and through visual inspection
+of the net squared displacement (NSD) metric. The NSD metric measures the
+squared Euclidean distance of a GPS relocation to a reference point
+\citep{Borger.2012}, which we set to the center of the territory of each
+dispersers' natal pack. Thus, dispersal was deemed to have started once an
+individual left its natal territory and ended once the NSD metric remained
+constant, indicating settlement. Because behavior during dispersal is more
+pertinent to landscape connectivity than behavior during residence
+\citep{Elliot.2014, Abrahms.2017}, we only considered data collected during
+dispersal for our analysis. During dispersal, GPS collars recorded a fix every 4
+hours and regularly transmitted data over the Iridium satellite system. To
+ensure regular time intervals between GPS fixes, we removed any fixes that were
+not successfully obtained at the desired 4-hour schedule (allowing for a
+tolerance of \( \pm \) 15 minutes). We then converted the fixes (n = 4'169) into
 steps, where each step represented the straight-line movement between two
@@
 covariates \textit{water-cover, distance-to-water, woodland-cover,
-shrub/grassland-cover, and human-influence}. To correctly render seasonality in
-water-cover, we applied a remote sensing algorithm that enabled us to obtain
-weekly updated raster-layers for water-cover and corresponding layers for
-distance-to-water from MODIS satellite imagery \citealp{Wolski.2017,
-Hofmann.2021}. This algorithm is now implemented in the \textsf{floodmapr}
-package which is available on GitHub
-(\url{https://github.com/DavidDHofmann/floodmapr}). Using the frequently updated
-floodmpas we were able to correctly assign the most appropriate covariate layer
-to each observed movement step. To ensure a consistent resolution across habitat
-covariates, we coarsened or interpolated all layers to a resolution of 250 m x
-250 m. A detailed description of how we prepared each habitat covariate is
-provided in \cite{Hofmann.2021}.
+shrub/grassland-cover, and human-influence}. To render the seasonal dynamics in
+water-cover of major water bodies for the extent of the Okavango Delta, we
+applied an algorithm that enabled us to obtain weekly updated raster-layers for
+water-cover and distance-to-water from MODIS satellite imagery
+\citep{Wolski.2017, Hofmann.2021}. This algorithm is now implemented in the
+\textsf{floodmapr} package (available on GitHub;
+\url{https://github.com/DavidDHofmann/floodmapr}). To ensure a consistent
+resolution across habitat covariates, we coarsened or interpolated all layers to
+a resolution of 250 m x 250 m. A detailed description of how we prepared each
+habitat covariate is provided in \cite{Hofmann.2021}.
 
 Besides habitat covariates, we computed movement metrics that we used as
-movement covariates in ISSF models \citep{Avgar.2016, Fieberg.2021}. Movement
-metrics were calculated for each step and included the step length
-(\textsf{sl}), its natural logarithm (\textsf{log(sl)}), and the cosine of the
-relative turning angle (\textsf{cos(ta)}). Moreover, we created the binary
-variable \textsf{LowActivity} indicating whether a step was realized during
-periods of low wild dog activity (09:00 to 17:00 local time) or high wild dog
-activity (17:00 to 09:00 local time, \citealp{Cozzi.2012}). We performed all
+movement covariates in the ISSF models \citep{Avgar.2016, Fieberg.2021}.
+Specifically, we computed for each step the step length (\textsf{sl}), its
+natural logarithm (\textsf{log(sl)}), and the cosine of the relative turning
+angle (\textsf{cos(ta)}). Moreover, we created the binary variable
+\textsf{LowActivity}, indicating whether a step was realized during periods of
+low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
+(17:00 to 09:00 local time, following \citealp{Cozzi.2012}). We performed all
 data preparations, spatial computations, and statistical analysis in R, version
@@
 \subsection{Step 1 - Movement Model}
-We used ISSFs to parametrize a mechanistic movement model for dispersing wild
-dogs \citep{Avgar.2016}. More specifically, we paired each realized step with 24
-random steps, so that a realized step plus its 24 random steps formed a
-25-step-stratum that received a unique identifier. As suggested by
+We used ISSFs \citep{Avgar.2016} to parametrize a mechanistic movement model for
+dispersing wild dogs. More specifically, we paired each realized (i.e. observed)
+step with 24 random steps, so that a realized step plus its 24 random steps
+formed a 25-step-stratum that received a unique identifier. As suggested by
 \cite{Avgar.2016}, we generated random steps by sampling random turning angles
@@
 and step lengths from a gamma distribution that was fitted to realized steps
-(scale \(\theta\) = 6'308 and shape \(k\) = 0.37). Along each step, we extracted
-and averaged values from the habitat covariate layers using the {\tt velox}
-package \citep{Hunziker.2021} and calculated the movement metrics \textsf{sl},
-\textsf{log(sl)}, and \textsf{cos(ta)} for each realized and random step. To
-facilitate model convergence, we standardized all continuous covariates to a
-mean of zero and a standard deviation of one. Since correlation among covariates
-was low (\(|r| < 0.6\); \citealp{Latham.2011}), we retained all of them for
+(scale \(\theta\) = 6'308 and shape \(k\) = 0.37).
+
+Along each realized and random step, we extracted values from the habitat
+covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
+computed averages of each covariate along the steps. We further calculated the
+movement metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)} for each
+step. To facilitate model convergence, we standardized all continuous covariates
+to a mean of zero and a standard deviation of one. Correlations among covariates
+were low (\(|r| < 0.6\); \citealp{Latham.2011}), so we retained all of them for
 modeling.
@@
 To contrast realized steps (scored 1) and random steps (scored 0), we assumed
-that animals assigned a selection score \(w(x)\) of the exponential form to each
-step (\ref{EQ2}; \citealp{Fortin.2005}). \(w(x)\) depended on the step's
-associated covariates (\(x_1, x_2, ..., x_n\)) and on the animal's preferences
-(i.e. relative selection strengths; \citealp{Avgar.2017}) towards these
-covariates (\(\beta_1, \beta_2, ..., \beta_n\)):
+that animals assigned a selection score \(w(x)\) to each step (\ref{EQ1};
+\citealp{Fortin.2005}). \(w(x)\) depended on the step's associated covariates
+(\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e. relative
+selection strengths; \citealp{Avgar.2017}) towards these covariates (\(\beta_1,
+\beta_2, ..., \beta_n\)):
 
@@
 
-\noindent The probability of a step being realized was then contingent on the
-step's selection score, as well as on the selection scores of all other step in
-the same stratum:
+\noindent The probability of a step \(i\) being realized was then contingent on
+the step's selection score, as well as on the selection scores of all other step
+in the same stratum:
 
@@
 mixed effects conditional logistic regression analysis \citep{Muff.2020} that we
-implemented using the r-package {\tt glmmTMB} \citep{Brooks.2017}. To capitalize
-on the poisson trick proposed by \cite{Muff.2020}, we defined random intercepts
-for each stratum and fixed the intercept variance to an arbitrary high value of
-\(10 ^ 6\). We also used disperser identity to model random slopes for all
-covariates.
+implemented using the r-package {\tt glmmTMB} \citep{Brooks.2017}. The method
+introduced by \cite{Muff.2020} allows to model random slopes, yet requires to
+fix the variance of the stratum specific intercept to a large value. Hence we
+fixed the stratum specific intercept variance to an arbitrary high value of \(10
+^ 6\). We used disperser identity to model random slopes for all covariates.
 
 The covariate structure of the movement model was based on a habitat selection
-model that we previously developed for dispersing wild dogs (hereafter referred
+model that was previously developed for dispersing wild dogs (hereafter referred
 to as \textit{based model}, \citealp{Hofmann.2021}). In the base model, no
@@
 parsimonious model using k-fold cross-validation for case-control studies as
-described in \cite{Fortin.2009}. This validation proofs a significant prediction
+described in \cite{Fortin.2009}. This validation proves a significant prediction
 in case the Spearman rank correlation of predicted step-ranks and associated
@@
 resembled an ``inverted'' ISSF and was set up as follows. (1) We defined a
-random source point and assumed a random initial orientation of the animal. (2)
-Starting from the source point, we generated 25 random steps by sampling turning
-angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths from our
-fitted gamma distribution. Like in the empirical data, each random step
-represented the straight-line movement possible within 4 hours. To prevent
+random source point and assumed a random initial orientation of the simulated
+animal. (2) Starting from the source point, we generated 25 random steps by
+sampling turning angles from a uniform distribution (\(-\pi, +\pi\)) and step
+lengths from our fitted gamma distribution. Similarly to the empirical data,
+each random step represented a 4 hours straight-line movement. To prevent
 unreasonably large steps, we restricted sampled step lengths to a maximum of 35
@@
 (3) Along each random step, we extracted and averaged values from the different
-habitat covariate layers and calculated movement covariates. To ensure
-compatible scales with the fitted movement model, we standardized extracted
-values using means and standard deviations from the empirical data. (4) We
-applied the parametrized movement model to predict the selection score \(w(x)\)
-for each step using \ref{EQ1} and we translated predicted scores into
-probabilities using \ref{EQ2}. (5) We used predicted probabilities to sample one
-of the random steps and determined the animal's new position. We repeated steps
-(2) to (5) until 2,000 steps (i.e. 400 consecutive dispersal days) were
-realized.\todo{Since we did not explain the 8-hourly gaps, this might be
-confusing}
+habitat covariate layers and calculated movement covariates (as explained under
+Step 1). To ensure compatible scales with the fitted movement model, we
+standardized extracted values using means and standard deviations from the
+empirical data. (4) We applied the parametrized movement model to predict the
+selection score \(w(x)\) for each step using \ref{EQ1} and we converted
+predicted scores into probabilities using \ref{EQ2}. (5) We randomly sampled one
+of the generated random steps based on assigned probabilities and determined the
+animal's new position. We repeated steps (2) to (5) until 2,000 steps were
+realized.
 
 To mitigate edge effects and to deal with random steps leaving the study area,
-we followed \cite{Koen.2010} and artificially expanded all covariate layers by
-a 100 km wide buffer zone. Within the buffer zone, we randomized covariate
-values by resampling values from the original covariate layers. Through this
-buffer zone, simulated dispersers were able to leave and re-enter the main study
-area. In cases where proposed random steps transgressed the outer border of this
-buffer zone, we resampled transgressing steps until they fully lied within the
-buffer zone, forcing individuals to remain within the expanded study area.
-
-For the simulation, we distributed 80,000 unique source points within the study
-area. Of these, 50,000 were random locations inside protected areas that were
-larger than the average home range size of wild dogs (i.e. \(>\) 700
-km\textsuperscript{2}; \citealp{Pomilia.2015}), while the remaining 30,000
-source points were placed randomly inside the buffer zone, mimicking potential
-immigration into the study area (Figure S1). Due to the random distribution of
-source points, the number of source points per km\textsuperscript{2} in selected
-areas was approximately equal.
+we followed \cite{Koen.2010} and artificially expanded all covariate layers by a
+100 km wide buffer zone. Within the buffer zone, we randomized covariate values
+by resampling values from the original covariate layers. Through this buffer
+zone, simulated dispersers were able to leave and re-enter the main study area.
+In cases where random steps crossed the outer border of this buffer zone, we
+resampled steps until they fully lied within the buffer zone, essentially
+forcing individuals to remain within the expanded study area.
+
+For the simulation, we distributed 80,000 source points within the study area.
+Of these, 50,000 were located inside protected areas that were larger than the
+average home range of resident wild dog packs (i.e. \(>\) 700
+km\textsuperscript{2}; \citealp{Pomilia.2015}), while the remaining 30,000 were
+placed randomly inside the buffer zone, mimicking potential immigration into the
+study area (Figure S1).
 
@@
 across the entire study area. For this purpose, we distributed 1,000 rectangular
-``checkpoints'', each with an extent of 5 km x 5 km at random coordinates within
-the study area (excluding the buffer). We then determined the relative frequency
-at which each checkpoint was traversed by simulated dispersers (hereafter
-referred to as relative traversal frequency) as we gradually increased the
-number of simulated trajectories from 1 to 50,000. To assess variability in the
-relative traversal frequency, we repeatedly subsampled 100 times from all 50'000
-dispersal trajectories and computed the mean traversal frequency across
-replicates, as well as its 95\% prediction-interval for each checkpoint. We
-considered connectivity to have reached a steady state once the width of the
+``checkpoints'', each with an arbitrary extent of 5 km x 5 km at random
+coordinates within the study area (excluding the buffer). We then determined the
+relative frequency at which each checkpoint was traversed by simulated dispersal
+trajectories (hereafter referred to as relative traversal frequency) as we
+gradually increased the number of simulated trajectories from 1 to 50,000. To
+assess variability in the relative traversal frequency, we repeatedly subsampled
+100 times from all 50'000 trajectories and computed the mean traversal frequency
+across replicates, as well as its 95\% prediction-interval for each checkpoint.
+We considered connectivity to have reached a steady state once the width of the
 prediction-interval dropped below a value of 0.01 for all checkpoints.
@@
 indicating the absolute frequency at which each raster-cell was traversed by
-simulated dispersers (e.g. \citealp{Peer.2008, Hauenstein.2019, Zeller.2020}).
-Specifically, we rasterized all simulated trajectories onto a raster with 1 km x
-1 km resolution and tallied resulting layers into a single map. If the same
-trajectory crossed a raster-cell twice, we only counted it once, thereby
-mitigating biases from individuals that moved in circles because they were
-surrounded by unfavorable habitat. To achieve high performance rasterization, we
-used the R-package {\tt terra} \citep{Hijmans.2021b}.
+simulated dispersal trajectories (e.g. \citealp{Peer.2008, Hauenstein.2019,
+Zeller.2020}). Specifically, we rasterized all simulated trajectories onto a
+raster with 1 km x 1 km resolution and tallied resulting layers into a single
+map. If the same trajectory crossed a raster-cell twice, we only counted it once
+to reduce biases due to individuals that were surrounded by unfavorable habitat
+and therefore ``moved in circles''. To achieve high performance rasterization,
+we used the R-package {\tt terra} \citep{Hijmans.2021b}.
 
 \subsubsection{Betweenness Map}
-To pinpoint discrete movement corridors and bottlenecks, we converted simulated
+To pinpoint movement corridors and bottlenecks, we converted simulated
 trajectories into a network and calculated betweenness scores for all
@@
 pertinent metric for connectivity as it measures how often a specific
-network-node (i.e. raster-cell) lies on a shortest path between any other pair
-of nodes \citep{BastilleRousseau.2018}. To convert simulated trajectories into a
-network, we followed \citep{BastilleRousseau.2018} and overlaid the study area
-(including the buffer) with a 5 km x 5 km raster, where the center of each
-raster-cell served as node in the final network. To identify edges (i.e.
-connections) between the nodes, we used the simulated trajectories and
-determined all transitions occurring from one node to another, as well as the
-frequency at which those transitions occurred. This resulted in an edge-list
+network-node (in our case a raster-cell) lies on a shortest path between any
+other pair of nodes \citep{BastilleRousseau.2018}. To convert simulated
+trajectories into a network, we followed \cite{BastilleRousseau.2018} and
+overlaid the study area (including the buffer) with a 5 km x 5 km raster, where
+the center of each raster-cell served as node in the final network. To identify
+edges (i.e. connections) between the nodes, we used the simulated trajectories
+and determined all transitions occurring from one node to another, as well as
+the frequency at which those transitions occurred. This resulted in an edge-list
 that we translated into a weighted network using the r-package {\tt igraph}
@@
 transitions, yet because {\tt igraph} handles edge weights (\(\omega\)) as
-costs, we had to invert the traversal-frequency through each raster-cell by
-applying \(\omega = \frac{mean(Traversal Frequency)}{Traversal Frequency_i}\).
+costs, we inverted the traversal-frequency through each raster-cell by applying
+\(\omega = \frac{mean(Traversal Frequency)}{Traversal Frequency_i}\).
 Consequently, regularly used edges received small weights (i.e. low costs) and
@@
 decision to only consider NPs as potential ``patches'' was purely out of
-simplicity and does not imply that connections between other protected areas are
-impossible. To quantify inter-patch connectivity, we computed the relative
+simplicity and does not imply that connections between other protected areas do
+not occur. To quantify inter-patch connectivity, we computed the relative
 frequency at which dispersers originating from one NP successfully moved into
-another NP. We considered movements successful if the individuals' dispersal
-trajectory intersected with the target NP at least once. We also recorded the
-number of steps required to reach the first intersection with the respective NP,
-allowing us to compute the average dispersal durations from one park to another.
-In summary, we determined \textit{if} and \textit{how often} dispersers moved
-between certain NPs, as well as \textit{how long} individuals had to move to
-make these connections.
+another NP. We considered movements as successful if an individual's dispersal
+trajectory intersected with the target NP at least once. For each trajectory we
+also recorded the number of steps required to reach the first intersection with
+the respective NP, allowing us to compute the average dispersal durations from
+one NP to another. In summary, we determined \textit{if} and \textit{how often}
+dispersers moved between certain NPs, as well as \textit{how long} individuals
+had to move to make these connections.
 
@@
 covariates and several of their interactions, suggesting that movement behavior
-during dispersal depended on habitat conditions (\Cref{MovementModel}, Table S1
+during dispersal depended on habitat conditions (\Cref{MovementModel}a, Table S1
 and Table S2). Although multiple models received an AIC weight > 0 (Table S1),
@@
 decision only marginally influenced subsequent steps as all models with positive
-AIC weights retained similar covariates (Table S1). Plots that facilitate model
+AIC weights retained similar covariates (Table S1). Plots that aid with model
 interpretation are provided in Figure S2. Under average conditions, dispersing
 wild dogs avoided moving through water, woodlands, and areas dominated by
-humans, but preferred shrublands or grasslands (\Cref{MovementModel}).
+humans, but preferred shrublands or grasslands (\Cref{MovementModel}a).
 Dispersers realized shorter steps (indicating slower movements) in areas covered
 by water or woodland, while realizing larger steps in areas dominated by shrubs
-or grass. Moreover, dispersing wild dogs moved faster during twilight and at
-night (i.e. between 17:00 and 09:00 o'clock) than during the rest of the day.
-Although dispersers showed a preference for directional movements (i.e. low
-turning angles), especially when moving quickly, they did less so in proximity
-to humans or water, resulting in more tortuous movements in such areas.
+or grass (\Cref{MovementModel}a). Moreover, dispersing wild dogs moved faster
+during twilight and at night (i.e. between 17:00 and 09:00 o'clock) than during
+the rest of the day (\Cref{MovementModel}a). Although dispersers showed a
+preference for directional movements (i.e. low turning angles), especially when
+moving quickly, they did less so in proximity to humans or water, resulting in
+more tortuous movements in such areas (\Cref{MovementModel}a).
 
@@
 Dispersal simulations based on the most parsimonious movement model proved
-useful for assessing landscape connectivity. Of the 50,000 simulated dispersers
-initiated within the main study area, only 4.5\% were eventually repelled by a
-map boundary, suggesting minimal biases due to boundary effects. Moreover, our
+useful for assessing landscape connectivity. Of the 50,000 simulated dispersal
+trajectories with starting point within the main study area, only 4.5\% reached
+a map boundary, suggesting minimal biases due to boundary effects. Moreover, our
 examination of the relative traversal frequency across all checkpoints showed
@@
 \subsection{Heatmap}
-The heatmap (\Cref{Heatmap}), which resulted from the summation of all simulated
+The heatmap (\Cref{Heatmap}), which resulted from the sum of all simulated
 dispersal trajectories, showed that several extensive regions within the
-KAZA-TFCA were frequently traversed by dispersing wild dogs (mean traversal
+KAZA-TFCA were frequently traversed by dispersing wild dogs (median traversal
 frequency inside KAZA-TFCA = 166, IQR = 274, Figure S6a), whereas areas beyond
-the KAZA-TFCA boundary were rarely visited (mean traversal frequency outside
+the KAZA-TFCA boundary were rarely visited (median traversal frequency outside
 KAZA-TFCA = 61, IQR = 133, Figure S6a). Most notably, the region in northern
 Botswana south of the Linyanti swamp appeared as highly frequented dispersal
-hotspot (mean traversal frequency = 987, IQR = 558). Nevertheless, the presence
-of extensive water bodies, such as the Okavango Delta, the Makgadikgadi Pan, and
-the Linyanti swamp, restricted dispersal movements and limited realized
+hotspot (median traversal frequency = 987, IQR = 558). Nevertheless, the
+presence of extensive water bodies, such as the Okavango Delta, the Makgadikgadi
+Pan, and the Linyanti swamp, restricted dispersal movements and limited realized
 connectivity within the KAZA-TFCA. Similarly, high human density, roads, and
@@
 dispersal movements in those countries. Outside the KAZA-TFCA, the most heavily
-used regions included the areas inside the Central Kalahari National Park in
-Botswana, the area south-west of the Khaudum National Park in Namibia, and the
-area surrounding the Liuwa Plains National Park in Zambia. Although the heatmap
-facilitated the identification of areas frequently traversed by simulated
-dispersers, it seemed impractical to pinpoint dispersal corridors.
+used regions included the areas inside the Central Kalahari NP in Botswana, the
+area south-west of the Khaudum NP in Namibia, and the area surrounding the Liuwa
+Plains NP in Zambia. Although the heatmap facilitated the identification of
+areas frequently traversed by simulated dispersers, it seemed impractical to
+pinpoint dispersal corridors.
 
@@
 The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
-that run within the KAZA-TFCA. Again, the region in northern Botswana emerged as
-a wild dog dispersal hub that connected more remote regions in the study area.
-Towards east, the extension of this corridor ran through Chobe National Park
-into Hwange National Park. From there, a further extension connected to
-Matusadona National Park in Zimbabwe. Northwest of the Linyanty ecosystem, a
-major corridor expanded into Angola, where it split and finally traversed
-over a long stretch of unprotected area into Zambia's Kafue National Park.
-Several additional corridors with lower betweenness scores emerged, yet most of
-them ran within the KAZA-TFCA boundaries (median betweenness inside KAZA-TFCA =
-6.947 Mio, IQR = 54.311 Mio, Figure S6b). In general, there were few corridors
-that directly linked the peripheral regions of the KAZA-TFCA and passed through
+that run within the KAZA-TFCA. Again, northern Botswana emerged as a wild dog
+dispersal hub that connected more remote regions in the study area. Towards
+east, the extension of this corridor ran through Chobe NP into Hwange NP. From
+there, a further extension connected to Matusadona NP in Zimbabwe. Northwest of
+the Linyanti ecosystem, a major corridor expanded into Angola, where it split
+and finally traversed over a long stretch of unprotected area into Zambia's
+Kafue NP. Several additional corridors with lower betweenness scores emerged,
+yet most of them ran within the KAZA-TFCA boundaries (median betweenness inside
+KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR = 54.311 \(\times\)
+10\textsuperscript{6}, Figure S6b). In general, there were few corridors that
+directly linked the peripheral regions of the KAZA-TFCA and passed through
 unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA-TFCA =
-2.685 Mio, IQR = 9.891 Mio, Figure S6b). Compared to the heatmap, the
-betweenness map facilitated the identification of dispersal corridors between
-habitat patches.
+2.685 \(\times\) 10\textsuperscript{6}, IQR = 9.891 \(\times\)
+10\textsuperscript{6}, Figure S6b). Compared to the heatmap, the betweenness map
+facilitated the identification of dispersal corridors between habitat patches.
 
@@
 \subsection{Inter-Patch Connectivity}
-The map of inter-patch connectivity showed that the relative frequency at which
-simulated dispersers moved from one NP to another varied, as did the average
-dispersal duration required to make these connections
+The inter-patch connectivity map showed that the relative frequency at which
+simulated dispersal trajectories moved from one NP to another varied, as did the
+average dispersal duration required to make these connections
 (\Cref{InterpatchConnectivity}). Overall, inter-patch connectivity between NPs
@@
 87\% of individuals originating from a NP in these countries successfully moved
-into some other NP (Figure S6a). Conversely, only 19\% of the dispersers leaving
-from a NP in Zambia managed to find a route into some other NP (Figure S6b).
+into some other NP (Figure S7a). Conversely, only 19\% of the dispersers leaving
+from a NP in Zambia managed to find a route into some other NP (Figure S7b).
 Prior to reaching another NP, individuals from Angola, Namibia, Botswana,
@@
 route was only realized by 340 individuals. However, relative to the number of
-simulated individuals in these NPs, these numbers imply fractions of 50\% and
-68\%. Interestingly, it also appears that the dispersal corridor between
+simulated individuals, these numbers imply fractions of 50\% and 68\%,
+respectively. Furthermore, it appeared that the dispersal corridor between
 Angola's NPs and the Kafue NP in Zambia identified in
@@
 \subsection{Short Summary}
-We presented a three-step workflow to simulate dispersal and assess landscape
-connectivity. We also demonstrated application of the three steps using
-empirical dispersal data from a free-ranging population of African wild dogs. In
-step one, we used ISSFs to parametrize a fully mechanistic movement model
-describing how dispersing wild dogs move through the available landscape. In
-step two, we employed the fitted model to simulate 80,000 dispersing wild dogs
-moving 2,000 steps across the extent of the KAZA-TFCA, the world's largest
-transboundary conservation area. In step three, we converted simulations into
-three complementary connectivity maps, each emphasizing a different aspect of
-landscape connectivity. The set of maps included a heatmap revealing frequently
-traversed areas, a betweenness-map delineating critical dispersal corridors, and
-a map of inter-patch connectivity indicating the presence and intensity of
-functional links between NPs and the average dispersal duration required to make
-those links. With this, we showed that simulations from ISSFs overcome several
-conceptual shortcomings inherent to more traditional connectivity modeling
-techniques, such as LCPA and CT.
+We used empirical data from a free-ranging population of African wild dogs to
+demonstrate the application of a three-step workflow to simulate dispersal
+trajectories and assess landscape connectivity. In step one, we used ISSFs to
+parametrize a fully mechanistic movement model describing how dispersing wild
+dogs move through the landscape. In step two, we employed the movement model to
+simulate 80,000 dispersal trajectories across the extent of the KAZA-TFCA, the
+world's largest transboundary conservation area. In step three, we translated
+the simulated trajectories into three complementary connectivity maps, each
+emphasizing a different aspect of landscape connectivity (e.g. frequently
+traversed areas, critical dispersal corridors and bottlenecks, and the presence
+and intensity of functional links between suitable patches). Importantly, such
+simulations from ISSFs overcome several conceptual shortcomings inherent to more
+traditional connectivity modeling techniques, such as LCPA and CT.
 
@@
 \subsection{Movement Model}
-Our movement model of dispersing wild dogs, which included interactions between
-habitat and movement covariates, effectively rendered habitat and movement
-preferences of dispersers, leading to a significant improvement in predictive
-performance compared to an earlier model that omitted such interactions
-\citep{Hofmann.2021}. Results on habitat preferences largely complied with
-previous studies that investigated habitat selection by dispersing wild dogs
+Our results on habitat preferences showed that dispersers avoid areas dominated
+by humans and covered by water, but select for regions with open grassland in
+the vicinity to water bodies. This largely complied with previous studies that
+investigated habitat selection by dispersing wild dogs
 \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020,
-Hofmann.2021}, yet by also rendering movement preferences, we were able to model
-several additional complexities common to dispersal. For instance, by including
-the appropriate interactions in the movement model we could accommodate that
-dispersers move with directional persistence and exhibit step lengths that are
-correlated with turning angles. Albeit the same autocorrelation could be
-rendered by sampling step lengths and turning angles from copula probability
-distributions \citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to
-incorporate correlations directly in the movement model. We only considered
-first order autocorrelation, i.e. correlation between two consecutive steps,
-although higher order autocorrelation is conceivable and may be desirable to
-model \citep{Dray.2010, McClintock.2012}. This will require vast amounts of GPS
-data that are not interrupted by missing fixes; something that is rarely
-achieved in reality \citep{Graves.2006}. Aside from modeling autocorrelation in
-consecutive movements, we also rendered potential dependencies between movement
-and habitat preferences by forming interactions between habitat covariates and
-movement covariates. For example, our final model contained an interaction
-between water-cover and step lengths showing that dispersers realized shorter
-steps in areas covered by water or an interaction between turning angles and
-water-cover highlighting that dispersers moved more tortuous in areas covered by
-water. The results themselves are not surprising, considering that wild dogs can
-only cross water by swimming or wading. Nevertheless, the ability of
-accompanying such effects in a single model is what makes ISSFs such a great and
-flexible tool for rendering a variety of different movement behaviors
-\citep{Avgar.2016, Fieberg.2021}. For this reason, we believe the method is a
-perfectly suited framework for individual-based simulations that serve to
-examine connectivity.
+Hofmann.2021}. However, by also accounting for movement preferences, we were
+able to model several additional complexities common to dispersal. For instance,
+by including an interaction between turning angle and step length we could
+accommodate that dispersers exhibit step lengths that are correlated with
+turning angles, meaning turning angles are smaller when individuals move fast.
+Although similar autocorrelations could be incorporated by sampling step lengths
+and turning angles from copula probability distributions \citep{Hodel.2021a,
+Hodel.2021b}, the ISSF framework allowed us to conventiently include
+correlations in the movement model. We only considered first order
+autocorrelation, i.e. correlation between two consecutive steps, although higher
+order autocorrelation is conceivable and may be desirable to model
+\citep{Dray.2010, McClintock.2012}. This will, however, require vast amounts of
+GPS data that are not interrupted by missing fixes; something that is rarely
+achieved in reality \citep{Graves.2006}. The power and flexibility of ISSFs to
+model additive effects between habitat and movement covariates
+\citep{Avgar.2016, Signer.2021} furthermore allowed us to formally capture that
+dispersing wild dogs move slower and more tortuous in areas covered by water,
+something that surely was to be expected. Overall, the inclusion of interactions
+between habitat and movement covariates in our movement model lead to a
+significant improvement in predictive performance compared to an earlier model
+that omitted such interactions \citep{Hofmann.2021}.
 
 \subsection{Simulation}
-Our simulation of 80'000 dispersers moving 2'000 steps across the KAZA-TFCA
-required five days of computation on a modern desktop machine (AMD Ryzen 7 2700X
-processor with 8 x 3.6 GHz and 16 logical cores, 64 GB of RAM). The long
-simulation time was primarily caused by the massive extent considered (ca. 1.8
-Mio km\textsuperscript{2}) and the large number of dispersers simulated. Most
-connectivity studies are limited to smaller extents (e.g.
-\citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020})
-and will therefore achieve faster simulation times. We also believe that fewer
-simulated dispersers will often suffice, as the relative traversal frequency by
-simulated individuals through randomly placed checkpoints across our study area
-converged already after 10,500 simulated dispersal trajectories. The number of
-required simulations to achieve reliable estimates of connectivity will,
-however, vary depending on the structure of the landscape and the dispersal
-ability of the focal species \citep{Gustafson.1996}. For species that disperse
-short distances through homogeneous environments, already few simulations may
-suffice go gauge connectivity, whereas for species that disperse over long
-distances through heterogeneous habitats, a large number of simulations will be
-required to sufficiently explore the spectrum of possible routes.
+Our simulation of 80,000 dispersal trajectories moving 2'000 steps across the
+KAZA-TFCA required five days of computation on a desktop machine (AMD Ryzen 7
+2700X processor with 8 x 3.6 GHz and 16 logical cores, 64 GB of RAM). The long
+simulation time was primarily caused by the massive extent of the study area
+considered (ca. 1.8 Mio km\textsuperscript{2}) and the large number of
+trajectories simulated. Most connectivity studies focus on smaller study areas
+(e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017,
+Zeller.2020}) and will therefore achieve faster simulation times (given the same
+spatial resolution across covariates). We also believe that fewer simulated
+trajectories will often suffice, as the relative traversal frequency by
+simulated trajectories through randomly placed checkpoints across our study area
+converged already after 10,500 runs. The number of required simulations to
+achieve reliable estimates of connectivity will, however, vary depending on the
+structure of the landscape and the dispersal behavior of the focal species
+\citep{Gustafson.1996}. For species that disperse short distances through
+homogeneous environments, few simulations may suffice to gauge connectivity,
+whereas for species that disperse over long distances through heterogeneous
+habitats, a large number of simulations will be required to sufficiently explore
+the spectrum of possible routes.
 
@@
 that an exceptionally large number of dispersers moved through the regions of
-the Moremi NP and the Chobe NP in northern Botswana. We previously identified
-the same area as potential dispersal hotspot using LCPA \citep{Hofmann.2021},
-yet it was not clear whether this was a consequence of the region being located
-in the center of the study area and connections being enforced between
-predefined start and endpoints. Contrary to LCPA, a simulation-based approach as
-presented in this paper does not require predefined endpoints because endpoints
-emerge naturally as the result from a simulated dispersal trajectories. Not
-having to pre-determine endpoints is especially useful for dispersal studies, as
-known endpoints are usually an unrealistic assumption \citep{Elliot.2014,
-Abrahms.2017, Cozzi.2020}. Simultaneously, simulations permit to detect
-potential routes that do not necessarily lead into suitable habitats but into
-ecological traps \citep{Dwernychuk.1972, VanDerMeer.2014} or into areas with a
-high susceptibility for human wildlife conflicts \citep{Cushman.2018}.
-Irrespective of the fact that simulated individuals were no longer enforced to
-move towards known endpoints, numerous simulated individuals traversed
-the region in northern Botswana, demonstrating that this hotspot is not merely
-an artifact but truly results from its favorable landscape characteristics.
+the Moremi NP and the Chobe NP in northern Botswana. \citep{Hofmann.2021}
+previously identified the same area as potential dispersal hotspot using LCPA,
+however, it was not clear whether this was the consequence of the central
+location of the region and connections being enforced between predefined start
+and endpoints. Contrary to LCPA, a simulation-based approach as presented here
+does not require predefined endpoints because endpoints, as endpoints emerge
+naturally from the simulated dispersal trajectories. Not having to predefine
+endpoints is especially useful for dispersal studies, as known endpoints are
+usually an unrealistic assumption \citep{Elliot.2014, Abrahms.2017, Cozzi.2020}.
+At the same time, simulations permit to detect potential routes that do not
+necessarily lead into suitable habitats \citep{Dwernychuk.1972, VanDerMeer.2014}
+and into areas with a high susceptibility for human wildlife conflicts
+\citep{Cushman.2018}.
 
@@
 linear movement routes and thus facilitated the identification of discrete
-movement corridors in the study area. The resulting map reinforced our notion
-that Botswana plays a central role for the establishment of connections into
-more remote regions of the KAZA-TFCA. While in this case both the heatmap and
-the betweenness map attribute high importance to the region in northern
-Botswana, there are other regions in which the two metrics do not necessarily
-coincide. The stretch of unprotected land between Luengue-Luiana NP in Angola
-and the Kafue NP in Zambia, for instance, receives a high betweenness-score, yet
-is only rarely traversed by dispersers according to the heatmap. This implies
-that the corridor takes on a crucial role for connecting Luengue-Luiana NP to
-Kafue NP, yet is in reality only traversed rarely. These contrasts highlight the
-value of consulting multiple metrics when assessing connectivity.
-
-As a final piece to the trinity of connectivity maps, we produced a map of
-inter-patch connectivity that strictly focused on NPs. The map depicted the
+movement corridors. The resulting map reinforced our notion that Botswana plays
+a central role for the establishment of connections into more remote regions of
+the KAZA-TFCA. While, in this case, both the heatmap and the betweenness map
+attributed a high importance to northern Botswana, little consensus was found
+for other regions. For instance, the stretch of unprotected land between
+Luengue-Luiana NP in Angola and the Kafue NP in Zambia was characterized by a
+high betweenness-score, but at the same time received a low heatmap score. These
+contrasts highlight the complementary nature of the presented connectivity maps
+and emphasize the value of consulting multiple metrics when assessing
+connectivity.
+
+Finally, we produced a map of inter-patch connectivity. The map depicted the
 frequency at which simulated individuals moved between NPs as well as the
 average duration (in steps) required to realize them. Calculating dispersal
-durations was only possible because dispersal trajectories were simulated
-spatio-temporally explicitly, something that is currently impossible using LCPA
-or CT. An explicit representation of time enables to answer questions such as:
+durations was possible because dispersal trajectories were simulated spatially
+and temporally explicitly, something that is currently impossible with LCPA or
+CT. An explicit representation of time enables answerings questions such as:
 ``\textit{How long will it take a disperser to move from A to B?}'' or
 ``\textit{Is it possible for a disperser to move from A to B within X days?}''.
-Moreover, rendering time explicitly yields opportunities to study how
-seasonality affects connectivity and to investigate whether dispersal corridors
-are only available temporarily or all-year round (\textit{dynamic connectivity};
-\citealp{Zeller.2020}). With LCPA or CT, seasonality can currently only be
-incorporated by generating seasonal permeability surfaces and repeating the
-connectivity analysis for each surface individually (e.g. \citealp{Benz.2016,
-Osipova.2019}). With simulations from ISSFs, on the other hand, seasonal changes
-in the environment could be rendered dynamically ``as the dispersers move'', so
-that simulated individuals could directly respond to such fluctuations.
+Moreover, it yields opportunities to study how seasonality affects connectivity
+and to investigate whether dispersal corridors exist seasonally or all-year
+round (\textit{dynamic connectivity}; \citealp{Zeller.2020}). With LCPA or CT,
+seasonality can currently only be incorporated by repeatedly running a
+connectivity analysis using an array of seasonal permeability surfaces (e.g.
+\citealp{Benz.2016, Osipova.2019}). In contrast, simulations from ISSFs allow
+the environment to change ``as the dispersers move'', so that simulated
+trajectories can dynamically respond to seasonal fluctuations in the
+environment.
 
@@
 ISSFs, one also must be aware of the associated non-trivial but important
-modeling decisions. Here, we will elaborate on four modeling decisions
+modeling decisions. Here, we will further elaborate on four modeling decisions
 concerning: (1) the number of simulated individuals, (2) the location of source
@@
 
-(1) When simulating dispersal, the modeler needs to decide on the number of
-simulated individuals. A higher number is always desirable, as each additional
-disperser provides information about landscape connectivity, yet every
-individual also entails computational costs. Consequently, a trade-off needs to
-be managed and the benefits of simulating additional dispersers needs to be
-gauged against the cost of computation. \cite{Signer.2017} suggests handling
-this trade-off by simulating additional individuals only until the metrics of
-interest converge towards their steady state. We incorporated this idea and
-employed the relative traversal frequency across randomly placed checkpoints to
-verify that the number of simulated individuals was large enough. However, the
-exact number of required individuals might vary depending on the target metric
-and the anticipated connectivity map. More sophisticated target metrics tailored
-for each connectivity map will need to be developed in the future.
-
-(2) To initiate dispersers, a modeler needs to provide a set of source points.
-We placed source points within protected areas large enough to sustain viable
-wild dog populations, implicitly assuming wild dogs would primarily survive in
-large, formally protected areas \citep{Woodroffe.1999, DaviesMostert.2012,
-Woodroffe.2012, VanDerMeer.2014}. We lacked precise knowledge about wild dog
-densities in the different protected areas, so we distributed source points
-within them randomly. In cases where such data is available (e.g. thanks to a
-habitat suitability model; \citealp{Squires.2013}), source points can be
-distributed accordingly, reflecting that the number of dispersers may not
-strictly scale with the size of the source area. Alternatively, source points
-can be distributed homogeneously across suitable habitats and only later be
-weighted when deriving connectivity metrics. After all, the challenge of
-selecting meaningful source points is not unique to individual-based simulations
-but also applies to LCPA or CT.
-
-(3) When employing ISSFs to simulate dispersers, it is required to decide on the
-number of simulated steps (i.e. dispersal durations). If sufficient dispersal
-data of the focal species has been collected, dispersal durations could be
-sampled from observed dispersal events. In our case, the low number of observed
-dispersal events and the high variability in wild dog dispersal durations (see
-\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}) prevented us
-from sampling from observed dispersal events. Instead, we simulated all
-individuals for 2'000 steps, which is at the upper end of observed dispersal
-durations. This allowed us to subsample trajectories to shorter durations after
-their simulations and to investigate the sensitivity of our connectivity maps
-with respect to exact dispersal durations (Figures S4 and S5).
-
-(4) Unless simulated individuals are strongly drawn towards a point of
-attraction (e.g. \cite{Signer.2017}), some individuals will inevitably approach
-a map boundary such that one of their proposed random steps leaves the study
-area. In theoretical applications, this problem can be circumvented by
-simulating movement on a torus where no map boundaries exist \citep{Hodel.2022}.
-For real world data, however, other approaches are needed. One possible solution
-would be to simply terminate the simulation of the respective individual as soon
-as one of its random steps transgresses a map border. This approach implicitly
-assumes that the respective animal left the study area and will never return. In
-cases where individuals are released near map borders, this approach will be
-problematic, as already a single transgressing random step will break the
-simulation, thus strongly biasing dispersal from patches near map borders. We
-tried to mitigate this issue in two complementary ways. First, we artificially
-increased covariate layers (and therefore the study area) by a buffer zone with
-randomized covariate values. The same method has already proven effective in
-mitigating edge-effects for graph-based methods \citep{Koen.2010}. In our case,
-the buffer enabled dispersers to leave and re-enter the main study area, as well
-as to initiate potential immigrants into our study system. As a second
-mitigation measure, we resampled any transgressing random steps until they fully
-lied within the study area, thus enforcing individuals to always remain within
-the study area. Given that only very few simulated individuals were repelled by
-a virtual map boundary, we conclude that the two measures effectively obviated
-boundary effects.
+(1) When simulating dispersal trajectories, the modeler needs to decide on the
+number of simulated individuals. A higher number is always desirable, as each
+additional trajectory provides information about landscape connectivity.
+However, each additional simulation entails computational costs, so a trade-off
+needs to be managed. Here, we followed \cite{Signer.2017} who suggested to
+simulate additional individuals only until the metrics of interest converge
+towards a steady state. The exact number of required individuals might, however,
+vary depending on the target metric and the anticipated connectivity map.
+Consequently, more sophisticated target metrics tailored for each of the
+presented connectivity maps need to be developed in the future.
+
+(2) To initiate dispersers, a modeler needs to provide a set of source points at
+which the virtual disperser will be released. We placed source points within
+protected areas large enough to sustain viable wild dog populations, implicitly
+assuming wild dogs primarily survive in large, formally protected areas
+\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. We
+lacked precise knowledge about the presence and abundance of wild dogs in the
+different protected areas, so we distributed source randomly points within them.
+In cases where such data is available, source points could be distributed
+accordingly, reflecting that the number of simulated trajectories does not
+necessarily scales with the size of the source area. Alternatively, source
+points could be distributed homogeneously and only later be weighted when
+deriving connectivity metrics. In any case, the challenge of selecting
+meaningful source points is not unique to individual-based simulations but also
+applies to LCPA or CT.
+
+(3) The use of ISSFs to simulate dispersers requires deciding on the number of
+simulated steps (i.e. dispersal durations). If sufficient dispersal data of the
+focal species has been collected, dispersal durations can be sampled from
+observed dispersal events. Due to the low number of observed dispersal events,
+we opted against this solution and instead simulated all individuals for 2,000
+steps, which is at the upper end of observed dispersal durations
+\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. This
+approach also has the advantage that it allowed us to subsample simulated
+trajectories to shorter durations after their simulation and to investigate the
+sensitivity of our connectivity maps with respect to exact dispersal durations
+(Figures S4 and S5).
+
+(4) Unless simulated dispersal trajectories are strongly drawn towards a point
+of attraction (e.g. \cite{Signer.2017}), some trajectories will inevitably
+approach a map boundary where one of the generated random steps might leave the
+study area. In this case, one solution would be to simply terminate the
+simulation of the affected trajectory, implicitly assuming that the respective
+animal left the study area permanently. However, this approach might become
+problematic in cases where individuals are released near map borders, as already
+a single random step leaving the study area will break the simulation, thus
+biasing connectivity along map borders. Instead of breaking the simulation, we
+created a buffer zone \citep{Koen.2010} and resampled random steps until they
+fully lied within the study area. This approach proved an effective solution to
+overcome problems with boundary effects.
 
 \subsection{Conclusion}
-To this end, we proposed and applied a simple three-step workflow that relies on
-ISSF-analysis and enables to simulate dispersal and assess landscape
-connectivity. The proposed workflow overcomes several of the conceptual
-shortcomings inherent to LCPA and CT, such as the assumption of known endpoints,
-and provides a highly flexible tool for investigating connectivity. In our case
-study, the workflow proved useful to investigate landscape connectivity for the
-endangered African wild dog within the KAZA-TFCA ecosystem and allowed us to
-pinpoint frequently used corridors and dispersal hubs. We hope to have sparked
-interest in the powerful framework of ISSFs for investigating dispersal
-connectivity, while at the same time conferring some of the non-trivial modeling
-decisions involved.
+In summary, we proposed and applied a simple three-step workflow that relies on
+ISSF-analysis and enables the simulation of dispersal trajectories and the
+assessment of landscape connectivity. The proposed workflow overcomes several of
+the conceptual shortcomings inherent to LCPA and CT, such as the assumption of
+known endpoints, and provides a highly flexible tool for investigating
+connectivity. With this work, we hope to have sparked interest in the
+application, optimization, or creation of methods to investigate dispersal
+movements and connectivity via individual-based simulations, while at the same
+time stressing some of the non-trivial modeling decisions involved.
 
@@
 to thank Johannes Signer for assisting with the simulation algorithm. This study
-was funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
-Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung
-Temperatio, Wilderness Wildlife Trust Foundation, Forschungkredit der
-Universität Zürich, and a Swiss National Science Foundation Grant
-(31003A\_182286) to A. Ozgul.
+was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
+Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
+Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
+Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
+Grant (31003A\_182286) to A. Ozgul.
 

### 2022-01-05 Chapter_1/04_Manuscript/Manuscript_Support.tex
Many updates

@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Step by Step: Using Integrated Step Selection
-Analysis to Simulate Wild Dog Dispersal and Assess Landscape Connectivity}
+\title{\textbf{Appendix}\\ A Three Steps Workflow for Simulating Dispersal
+Trajectories and Assessing Landscape Connectivity: A Case Study on the
+Endangered African Wild Dog}
 
@@
 
- \textsuperscript{2} Botswana Predator Conservation, Private Bag 13, Maun,
- Botswana.
+ \textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
+ Maun, Botswana.
 
@@
 
- \textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
- Assess Landscape Connectivity
+\textbf{Running Title:} Simulating Wild Dog Dispersal Trajectories to Assess
+Landscape Connectivity
 
@@
   \caption{Auxiliary plots that help with the interpretation of the most
-   parsimonious movement model from Table S1. The plots were generated
-   following recommendations reported in \cite{Fieberg.2021}. Subplots
-   \textbf{a1} to \textbf{a8} highlight dispersing wild dogs' movement kernel
-   and indicate how the kernel is influenced by interactions with other
-   covariates. Subplots \textbf{b1} to \textbf{b5} depict results from
-   dispersing wild dogs' habitat kernel and highlight differences in predicted
-   relative selection scores (RSS) when varying values of the covariate of
-   interest. For each covariate, predictions were made on the range of values
+  parsimonious movement model from Table S1. The plots were generated following
+  recommendations reported in \cite{Fieberg.2021}. Subplots a1 to a8 highlight
+  dispersing wild dogs' movement kernel and indicate how the kernel is
+  influenced by interactions with other covariates. Subplots b1 to b5 depict
+  results from dispersing wild dogs' habitat kernel and highlight differences in
+  predicted relative selection scores (RSS) when varying values of the covariate
+  of interest. For each covariate, predictions were made on the range of values
   that was observed in the real data, assuming that all other covariates were
   centered and that steps were realized during periods of ``high'' wild dog
-   activity. Plot a1, for example, can be interpreted as follows: the
-   probability of realizing a step with a low turning angle is much higher when
-   the corresponding step is large. Moreover, b1 can be interpreted as follows:
+  activity. Plot a1, for example, can be interpreted as follows: the probability
+  of realizing a step with a low turning angle is much higher when the
+  corresponding step is large. Moreover, b1 can be interpreted as follows:
   relative probability of using a step decreases as the amount of water cover
@@
 \newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
-\section{Heatmaps in Relation to Number of Simulated Steps}
+\section{Heatmaps in Relation to the Number of Simulated Steps}
 \begin{figure}[hbtp]
@@
 \newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
-\section{Betweenness Maps in Relation to Number of Simulated Steps}
+\section{Betweenness Maps in Relation to the Number of Simulated Steps}
 \begin{figure}[hbtp]
@@
 \restoregeometry
+
 %------------------------------------------------------------------------------
-%	Appendix S11: Areas Reached
+%	Appendix S11: Map Comparison
+%------------------------------------------------------------------------------
+\newpage
+\section{Comparison of Traversal Frequencies and Betweenness Scores Inside and Outside KAZA-TFCA}
+\begin{figure}[hbtp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_MapComparison.png}
+  \caption{Comparison of values from the heatmap and betweenness map inside
+  (blue) and outside (orange) the KAZA-TFCA borders.}
+  \label{MapComparison}
+ \end{center}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%	Appendix S12: Areas Reached
 %------------------------------------------------------------------------------
@@
 
-%------------------------------------------------------------------------------
-%	Appendix S11: Map Comparison
-%------------------------------------------------------------------------------
-\newpage
-\section{Comparison of Traversal Frequencies and Betweenness Scores Inside and Outside KAZA-TFCA}
-\begin{figure}[hbtp]
- \begin{center}
-  \includegraphics[width = \textwidth]{99_MapComparison.png}
-  \caption{Comparison of values from the heatmap and betweenness map inside
-  (blue) and outside (orange) the KAZA-TFCA borders.}
-  \label{MapComparison}
- \end{center}
-\end{figure}
-
 \newpage

### 2022-01-05 Chapter_2/02_R-Scripts/01_DataCleaning/08_DynamicWaterAndVegetation.r

@@ -1,0 +1,84 @@
+################################################################################
+####
+################################################################################
+# Description: Use floodmaps to mask out anything that is covered by water in
+# the vegetation maps. Specifically, we will for each floodmap create a
+# corresponding vegetation map where water has been masked out. This requires us
+# to match the date of the floodmaps with the date of the vegetation maps. Note:
+# vegetation maps are based on data from day 65 of one year to day 64 of the
+# next year -> From the MODIS documentation
+# (https://lpdaac.usgs.gov/documents/112/MOD44B_User_Guide_V6.pdf): The product
+# date (of the vegetation data) refers to the start date of the annual period so
+# a product with ID “2005065” was produced with data from 2005065 – 2006064. The
+# start date of all MODIS VCF products is yyyy065 (where yyyy refers to the 4
+# digit year). This originally derived from the first full 16-day composite
+# period in the MODIS data record which begins with day of year 2000065. Hence,
+# the layername of each MODIS vegetation map refers to the period that starts on
+# day 65 of that year!
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory.
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Load required packages
+library(terra)       # To handle raster data
+library(lubridate)   # To handle dates
+library(tidyverse)   # To wrangle data
+
+# Let's prepare tibbles for the floodmaps and the vegetation maps so that we can
+# more easily subset the data. First for the flood data
+flood_dates <- dir(
+      path       = "03_Data/02_CleanData/00_Floodmaps/02_Resampled"
+    , pattern    = ".tif$"
+  ) %>% ymd()
+
+# Now for the shrubs/grassland cover
+grass_dates <- "03_Data/02_CleanData/00_Vegmaps/ShrubCover.grd" %>%
+  rast() %>%
+  names() %>%
+  tibble(
+      Year     = .
+    , FromDate = as.Date(64, origin = paste0(., "-01-01"))
+    , ToDate = as.Date(63, origin = paste0(., "-01-01")) + years(1)
+  )
+
+# Same for tree cover
+tree_dates <- "03_Data/02_CleanData/00_Vegmaps/TreeCover.grd" %>%
[…]

### 2022-01-05 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/AreasReached.R

@@ -1,177 +1,0 @@
-################################################################################
-#### Plot Reached Areas
-################################################################################
-# Description: Plot the areas reached from a source point
-
-# Clear R's brain
-rm(list = ls())
-
-# Set the working directory
-wd <- "/media/david/My Passport/Backups/WildDogs/15. PhD/00_WildDogs"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)      # For data wrangling
-library(raster)         # To handle spatial data
-library(pbmcapply)      # To run on multiple cores with progress bar
-library(rgeos)          # For spatial ananylsis
-library(igraph)         # For network analysis
-library(davidoff)       # Custom functions
-
-################################################################################
-#### Prepare Data
-################################################################################
-# Load source areas and points
-areas <- shapefile("03_Data/03_Results/99_SourceAreas2.shp")
-points <- shapefile("03_Data/03_Results/99_SourcePoints2.shp")
-
-# Visualize
-plot(areas)
-text(x = points, labels = points$ID, cex = 1)
-
-# Load the simulated dispersal trajectories
-sims <- read_rds("03_Data/03_Results/99_DispersalSimulationSub.rds")
-
-# Subset to a single start point
-sims <- subset(sims, StartPoint == 38)
-
-# Function to identify all areas reached by a trajectory
-getReached <- function(trajectories = NULL, areas = NULL){
-
-    # Identify reached areas
-    reached <- trajectories %>%
-      gIntersects(areas, ., byid = T) %>%
-      as.data.frame() %>%
-      setNames(areas$ID) %>%
-      cbind(trajectories@data, .) %>%
-      dplyr::select(ID, StartPoint, everything()) %>%
-      gather(key = EndPoint, value = Reached, 3:ncol(.)) %>%
-      subset(Reached) %>%
-      dplyr::select(-Reached) %>%
-      mutate(EndPoint = as.numeric(EndPoint)) %>%
-      arrange(ID, StartPoint, EndPoint)
-
-  # Return reached areas
-  return(reached)
-}
-
-# Prepare a tibble indicating the points in time and point sampling regimes for
-# which we want to identify the areas reached.
-reached <- as_tibble(
-  expand.grid(
-      steps     = c(68, 125, 250, 500, 1000, 2000)
[…]

### 2022-01-05 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/DispersalDuration.R

@@ -1,64 +1,0 @@
-################################################################################
-#### Plot of Dispersal Durations
-################################################################################
-# Description: A simple plot of the Dispersal Durations
-
-# Clear R's brain
-rm(list = ls())
-
-# Set the working directory
-wd <- "C:/Users/david/switchdrive/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)    # For data wrangling
-library(lubridate)    # To handle dates nicely
-library(ggpubr)       # For nice plots
-library(fitdistrplus) # To fit a distribution
-
-################################################################################
-#### Data Preparation
-################################################################################
-# Load required data
-cut <- read_csv("03_Data/02_CleanData/00_General_Dispersers_Popecol_CutoffDates.csv")
-
-# Add Two hours to the timestamps
-cut$StartDate <- cut$StartDate + hours(2)
-cut$EndDate <- cut$EndDate + hours(2)
-
-# Calculate dispersal duration
-durations <- cut %>%
-  mutate(DispersalDuration = EndDate - StartDate) %>%
-  group_by(DogName) %>%
-  summarize(DispersalDuration = as.numeric(sum(DispersalDuration)))
-
-# Fit an exponential distribution to the durations
-dist <- fitdist(durations$DispersalDuration, distr = "gamma")
-
-# Create new data from the distribution
-x <- seq(0, 300, by = 0.1)
-y <- dgamma(x, shape = dist$estimate[["shape"]], rate = dist$estimate[["rate"]])
-
-# Visualize
-p <- ggdensity(
-    durations
-  , x     = "DispersalDuration"
-  , fill  = "orange"
-  , color = "orange"
-  # , add   = "mean"
-  , rug   = TRUE
-  , xlab  = "Dispersal Duration (days)"
-  , ylab  = "Density"
-  , font.label = list(color = "black")
-  ) +
-  geom_line(data = data.frame(x, y), aes(x = x, y = y), lty = 2, col = "gray50") +
-  labs(color = "white") +
-  theme(
-      axis.ticks       = element_line(color = "white")
-    , axis.line        = element_line(color = "white")
-    , axis.text        = element_text(color = "white", size = 10)
-    , text             = element_text(color = "white")
-    , panel.background = element_rect(fill = "black")
-    , plot.background  = element_rect(fill = "black", color = "black")
-  )
-ggsave(p, device = "png", filename = "test.png", bg = "transparent", width = 3, height = 2)

### 2022-01-05 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/FloodAnimation.R

@@ -1,88 +1,0 @@
-############################################################
-#### Preparation Flood Animation
-############################################################
-# Clear R's brain
-rm(list = ls())
-
-# Set the working directory
-setwd("/home/david/ownCloud/University/15. PhD/Chapter_1")
-
-# Set a seed
-set.seed(12345)
-
-# Load required packages
-library(terra)
-library(raster)
-library(tidyverse)
-library(davidoff)
-library(viridis)
-library(animation)
-
-################################################################################
-#### Animation of the Flood
-################################################################################
-# Load required data
-maps <- dir(
-    path       = "03_Data/02_CleanData/00_Floodmaps/01_Original"
-  , pattern    = "*.tif$"
-  , full.names = T
-) %>% rast(., bands = 1)
-
-# Get the dates from the map descriptions
-dates <- as.Date(names(maps))
-
-# We only want to plot relatively cloud free images. Identify the cloud cover in
-# each image for this
-clouds <- c()
-for (i in 1:nlyr(maps)){
-  clouds[i] <- sum(values(maps[[i]]) == 127)
-  clouds[i] <- clouds[i] / ncell(maps[[i]])
-}
-
-# Let's identify the number of maps with a cloud coverage above 0.1
-sum(clouds > 0.1)
-
-# Keep only those images with a cloud coverage below 10%
-indices <- which(clouds < 0.1)
-maps <- maps[[indices]]
-dates <- dates[indices]
-
-# Load mask to cover any erronous pixels
-mask <- vect("03_Data/01_RawData/MODIS/MCD43A4/02_Masks/MaskNoWater.shp")
-
-# Reclassify the pixels
-rcl <- data.frame(old = c(0, 127, 255), new = c(1, 0, 0))
-maps <- classify(maps, rcl)
-
-# Mask pixels below the nowater mask
-maps <- mask(maps, mask, updatevalue = 0, inverse = T)
-
-# Convert to regular raster
-maps <- stack(maps)
-
-# Subset
-maps2 <- maps[[seq(1, 100, length.out = 5)]]
-for (i in 1:5){
-  png(paste0(i, ".png"), width = 1980, height = 1980, bg = "transparent")
-  plot(maps2[[i]]
-    , col    = c("gray25", "cornflowerblue")[c(1, 2)]
-    , legend = FALSE
-    , axes   = FALSE
-    , box    = FALSE
-  )
[…]

### 2022-01-05 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/MovementModel.R

@@ -1,232 +1,0 @@
-################################################################################
-#### Plot Movement Model Results
-################################################################################
-# Description: A simple plot of the Dispersal Durations
-
-# Clear R's brain
-rm(list = ls())
-
-# Set the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)  # For data wrangling
-library(lubridate)  # To handle dates nicely
-library(ggpubr)     # For nice plots
-library(davidoff)   # Custom functions
-library(glmmTMB)    # To handle model results
-library(cowplot)    # To add image to ggplot
-library(lemon)      # For nice axis labels
-library(imager)     # To import images
-library(viridis)    # For nice colors
-library(raster)     # To create rasters (visInt)
-library(ggdark)     # Dark ggplot theme
-
-################################################################################
-#### Prepare Data
-################################################################################
-# Old directory
-wd <- "/media/david/My Passport/Backups/WildDogs/15. PhD/00_WildDogs"
-setwd(wd)
-
-# Load the movement model
-move.mod <- readRDS("03_Data/03_Results/99_MovementModel.rds")
-
-# Select best model
-best <- move.mod$Model[[1]]
-
-# Summary of the best model
-summary(best)
-
-# Calculate Confidence intervals
-coeffs <- getCoeffs(best, pvalue = TRUE)[-1, ] %>%
-  mutate(
-      LCI = Coefficient - 1.96 * SE
-    , UCI = Coefficient + 1.96 * SE
-  )
-
-# Add stars indicating the significance
-coeffs$Significance <- sapply(1:nrow(coeffs), function(x){
-  if (coeffs$pvalue[x] <= 0.01){
-    return("***")
-  } else if (coeffs$pvalue[x] <= 0.05){
-    return("**")
-  } else if (coeffs$pvalue[x] <= 0.1){
-    return("*")
-  }
-})
-
-# Rename covariates
-coeffs$Covariate[coeffs$Covariate == "Shrubs"] <- "Shrubs/Grassland"
[…]

### 2022-01-05 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/Seasonality.r

@@ -1,0 +1,205 @@
+################################################################################
+#### Seasonality
+################################################################################
+# Description: Visualization of seasonality patterns in our study area.
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory.
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Load required packages
+library(terra)          # To handle raster files
+library(tidyverse)      # To wrangle data
+library(lubridate)      # To handle dates
+library(ggpubr)         # To arrange multiple plots
+
+# Load temperature data
+temp <- "03_Data/02_CleanData/00_Tempmaps" %>%
+  dir(pattern = ".grd$", full.names = T) %>%
+  rast()
+
+# Load precipitation data
+prec <- "03_Data/02_CleanData/00_Rainmaps" %>%
+  dir(pattern = ".grd$", full.names = T) %>%
+  rast()
+
+# Load floodmaps
+flood <- "03_Data/02_CleanData/00_Floodmaps/02_Resampled" %>%
+  dir(pattern = ".tif$", full.names = T) %>%
+  rast()
+
+# Load NDVI
+ndvi <- "03_Data/02_CleanData/01_LandCover_NDVI.grd" %>%
+  rast()
+
+# Identify dates for the temperature data
+temp_dates <- names(temp) %>%
+  tibble(
+      Date     = substr(., start = 1, stop  = 10)
+    , Hour     = substr(., start = 12, stop = 13)
+    , Datetime = make_datetime(year(Date), month(Date), day(Date), Hour, 0, 0)
+  ) %>%
+  pull(Datetime)
+
+# Identify dates for the precipitation data
+prec_dates <- names(prec) %>%
+  tibble(
+      Date     = substr(., start = 1, stop  = 10)
+    , Hour     = substr(., start = 12, stop = 13)
+    , Datetime = make_datetime(year(Date), month(Date), day(Date), Hour, 0, 0)
+  ) %>%
+  pull(Datetime)
+
+# Identify dates for the floodmaps
+flood_dates <- "03_Data/02_CleanData/00_Floodmaps/02_Resampled" %>%
+  dir(pattern = ".tif$") %>%
+  basename() %>%
+  ymd()
+
+# Identify dates for the NDVI maps
+ndvi_dates <- names(ndvi) %>% ymd()
+
+# Assign dates as layernames
+names(temp) <- temp_dates
[…]

### 2022-01-24 Chapter_1/04_Manuscript/Manuscript.tex
updates

@@
 %------------------------------------------------------------------------------
-\title{A Three Steps Workflow for Simulating Dispersal Trajectories and
-Assessing Landscape Connectivity: A Case Study on the Endangered African Wild
-Dog}
+\title{A Thee-Step Approach for Assessing Landscape Connectivity via Simulated
+Dispersal: African Wild Dog Case Study}
 
@@
 %
-% Here, we propose a simple three-step workflow that combines several
+% Here, we propose a simple three-step approach that combines several
 % already-existing methods to assess connectivity using simulated dispersal
@@
 %
-% We showcase the application of the proposed workflow using empirical data of
+% We showcase the application of the proposed approach using empirical data of
 % dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
@@
 has mainly been investigated using least-cost analysis and circuit theory, two
-methods that make assumptions that are hardly applicable during dispersal.
+methods that make assumptions that are hardly applicable to dispersal.
 Least-cost analysis assumes that animals move towards a known endpoint and are
-knowledgeable about the most favorable route to reach it. Circuit theory
-presumes a complete random walk and fails to incorporate directional
-persistence. While these assumptions can be relaxed by simulating dispersal
-movements spatio-temporally explicitly across the landscape of interest, a
-unified approach for such simulations is lacking.
-
-Here, we present a simple three-step workflow to simulate dispersal movements
-and to assess connectivity starting from empirical GPS movement data. In step
-one, we use integrated step selection functions to fit a mechanistic movement
-model describing habitat and movement preferences of dispersing individuals. In
-step two, we apply the parameterized movement model to simulate dispersal
-trajectories. In step three, we derive three complementary connectivity maps: a
-heatmap that highlights frequently traversed areas, a betweenness map that
-pinpoints dispersal corridors, and a map of inter-patch connectivity that
-indicates the presence and intensity of functional links between habitat
-patches. As a case study, we applied the workflow to GPS data collected during
-dispersal of the endangered African wild dog (\textit{Lycaon pictus}) inhabiting
-the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
+knowledgeable about the most favorable route. Circuit theory presumes a complete
+random walk and fails to incorporate directional persistence. While these
+assumptions can be relaxed by simulating dispersal movements explicitly across
+the landscape of interest, a unified approach for such simulations is lacking.
+
+Here, we present a simple three-step approach to simulate dispersal movements
+and to assess connectivity using empirical GPS movement data and environmental
+habitat layers. In step one, we use integrated step selection functions to fit a
+mechanistic movement model describing habitat and movement preferences of
+dispersing individuals. In step two, we apply the parameterized movement model
+to simulate dispersal trajectories. In step three, we derive three complementary
+connectivity maps: a heatmap that highlights frequently traversed areas, a
+betweenness map that pinpoints dispersal corridors, and a map of inter-patch
+connectivity that indicates the presence and intensity of functional links
+between habitat patches. As a case study, we applied the approach to GPS data
+from dispersing individuals of the endangered African wild dog (\textit{Lycaon
+pictus}) inhabiting the Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA).
 
@@
 neighboring areas. By rendering interactions between movement preferences and
-habitat preferences, our movement model substantially outperformeed a previous
-model that omitted such interactions. While we simulated 80,000 dispersal
-trajectories in total, we found that metrics of connectivity only marginally
-changed beyond 10,500 simulations, suggesting that fewer simulations would have
-sufficed.
+habitat preferences, our movement model substantially outperformed a model that
+omitted such interactions.
 
@@
 biologically unrealistic assumptions but also permits a more mechanistic
-understanding of dispersal and landscape connectivity. Our workflow is thus
+understanding of dispersal and landscape connectivity. Our approach is thus
 useful for a variety of applications in ecological, evolutionary, and
@@
 unsuitable for dispersers, calling for the development of alternative
-approaches. One promising alternative to investigate dispersal and assess
-connectivity is to simulate dispersal from individual-based movement models
-\citep{Diniz.2019}.
+approaches. One promising alternative is to assess landscape connectivity via
+simulated dispersal trajectories generated from individual-based movement models
+(IBMMs, \citep{Diniz.2019}).
 
@@
 The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
-from individual-based movement models (IBMMs) and by converting simulated
-trajectories into meaningful measures of connectivity \citep{Diniz.2019}). In
-contrast to LCPA and CT, IBMMs allow to explicitly simulate how individuals move
-across and interact with the encountered landscape \citep{Kanagaraj.2013,
-Clark.2015, Allen.2016, Hauenstein.2019, Zeller.2020}, as well as to render
-potential interactions between movement behavior and habitat conditions
-\citep{Avgar.2016}. This strictly shifts the focus from a structural to a more
-functional view on connectivity \citep{Tischendorf.2000, Kanagaraj.2013,
-Hauenstein.2019}. Furthermore, IBMMs generate movement sequentially, i.e. they
-generate a series of steps, the temporal dimension of dispersal movements
-becomes explicit and allows to model autocorrelation between consecutive steps
+using IBMMs and by converting simulated trajectories into meaningful measures of
+connectivity \citep{Diniz.2019}. In contrast to LCPA and CT, IBMMs allow to
+explicitly simulate how individuals move across and interact with the
+encountered landscape \citep{Kanagaraj.2013, Clark.2015, Allen.2016,
+Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
+between movement behavior and habitat conditions \citep{Avgar.2016}. This
+strictly shifts the focus from a structural to a more functional view on
+connectivity \citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}.
+Furthermore, IBMMs generate movement sequentially, i.e. they generate a series
+of steps, so that the temporal dimension of dispersal movements becomes explicit
+and allows modeling autocorrelation between consecutive steps
 \citep{Diniz.2019}. Finally, simulations from IBMMs do not enforce movement or
 connections towards preconceived endpoints, thereby preventing biases arising
-from misplaced endpoints. Despite these advantages, a unified framework to
-simulate dispersal and assess connectivity using IBMMs is lacking.
-
-% Proposed Solution: Three-Step Workflow
-\subsection{Proposed Solution: Three-Step Workflow}
-Here, we propose a simple three-step workflow for simulating dispersal and
-assessing landscape connectivity (\Cref{GraphicalAbstract}). In step one, we
-combine GPS movement data of dispersing individuals with relevant habitat
-covariates to fit a mechanistic movement model using integrated step selection
-functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the study
-species' habitat kernel (i.e. habitat preferences), its movement kernel (i.e.
-movement preferences/capabilities), and potential interactions between the two
-\citep{Avgar.2016, Fieberg.2021}. In step two, we simulate individual dispersal
-trajectories using the parametrized movement model. Comparable simulations have
-already been applied to estimate steady-state utilization distributions of
-resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
-connectivity, yet disregarding interdependencies between habitat and movement
-kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert
-simulated trajectories into three complementary connectivity maps, each
-highlighting a different aspect of connectivity. The set of maps includes (i) a
-heatmap revealing areas that are frequently traversed by dispersers (e.g.
-\citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map delineating
-dispersal corridors and bottlenecks (e.g. \citealp{BastilleRousseau.2018}),
-(iii) and a map of inter-patch connectivity, depicting the presence and
-intensity of specific connections, as well as the average dispersal duration
-required to connect corresponding habitat patches (e.g. \citealp{Gustafson.1996,
+from misplaced endpoints. Despite these advantages, a unifying framework to
+simulate dispersal and assess connectivity using IBMMs is lacking. Considering
+the large number of subjective decisions entailed by IBMMs, an approach that
+streamlines and unifies the application of dispersal simulations to assess
+connectivity will, however, be critical to safeguard interspecfic comparability
+and reproducability.
+
+% Proposed Solution: Three-Step approach
+\subsection{Proposed Solution: Three-Step Approach}
+Here, we propose and exemplify a simple three-step approach for simulating
+dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
+step one, we combine GPS movement data of dispersing individuals with relevant
+habitat covariates to fit a mechanistic movement model using integrated step
+selection functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the
+study species' habitat kernel (i.e. habitat preferences), its movement kernel
+(i.e. movement preferences/capabilities), and potential interactions between the
+two \citep{Avgar.2016, Fieberg.2021}. In step two, we simulate individual
+dispersal trajectories using the parametrized movement model. Comparable
+simulations have already been applied to estimate steady-state utilization
+distributions of resident individuals \citep{Potts.2013, Signer.2017} and to
+model landscape connectivity, yet disregarding interdependencies between habitat
+and movement kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we
+convert simulated trajectories into three complementary connectivity maps, each
+highlighting a different aspect of connectivity; (i) a heatmap revealing areas
+that are frequently traversed by dispersers (e.g. \citealp{Hauenstein.2019,
+Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
+bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
+inter-patch connectivity, depicting the presence and intensity of specific
+connections, as well as the average dispersal duration required for the
+realization of those connections (e.g. \citealp{Gustafson.1996,
 Kanagaraj.2013}).
@@
 \subsection{Case Study}
-We showcase the application of the proposed workflow (\Cref{GraphicalAbstract})
+We showcase the application of the proposed approach (\Cref{GraphicalAbstract})
 using GPS movement data collected on dispersing individuals of the endangered
-African wild dog (\textit{Lycaon Pictus}). The African wild dog is a highly
+African wild dog (\textit{Lycaon pictus}). The African wild dog is a highly
 mobile species whose population persistence heavily relies on the availability
@@
 Transfrontier Conservation Area (KAZA-TFCA). We anticipated that simulations
-based on our three-step workflow would overcome several of the above highlighted
+based on our three-step approach would overcome several of the above highlighted
 conceptual shortcomings of traditional connectivity models and provide a more
@@
 Large portions of the KAZA-TFCA are formally protected in the form of national
-parks (NPs) or other protected areas, yet considerable human-dominated
-landscapes (e.g. roads, agricultural sites, and settlements) remain.
+parks (NPs) or other protected areas, yet a considerable portion of the
+landscape is still human-dominated (e.g. roads, agricultural sites, and
+settlements).
 
@@
 \subsubsection{GPS Data}
-We collected GPS movement data on 16 coalitions of dispersing African wild dogs
+We collected GPS movement data on 16 dispersing coalitions of African wild dogs
 (7 female and 9 male coalitions) between 2011 and 2019 from a free-ranging
@@
 substantially outperformed a random guess and suggested reliable predictions
-(confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\)
-do not overlap). Moreover, the model correctly assigned high selection scores to
-realized steps (\Cref{MovementModel}b), indicating a good fit between
-predictions and observations. Compared to the base model (\(\bar{r}_{s,
-realized} = -0.55; 95\%-CI = [-0.57, -0.52]\); \citealp{Hofmann.2021}), the
-inclusion of several interactions between movement and habitat covariates
-significantly improved model performance (\(\bar{r}_{s, realized} = -0.65;
-95\%-CI = [-0.67, -0.64]\).
+(i.e. confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
+random}\) did not overlap). Moreover, the model correctly assigned high
+selection scores to realized steps (\Cref{MovementModel}b), indicating a good
+fit between predictions and observations. Compared to the base model
+(\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
+\citealp{Hofmann.2021}), the inclusion of several interactions between movement
+and habitat covariates significantly improved model performance (\(\bar{r}_{s,
+realized} = -0.65; 95\%-CI = [-0.67, -0.64]\).
 
@@
   \includegraphics[width=\textwidth]{99_InterpatchConnectivity.png}
-  \caption{Map of inter-patch connectivity, highlighting connections between NPs
-  (dark green). Yellow bubbles represent the center of the different NPs and are
-  sized in relation to the number of simulated dispersers originating from each
-  park. Black dots represent NPs that were smaller than 700
-  km\textsuperscript{2} and therefore did not serve as source areas. Arrows
-  between NPs illustrate between which NPs the simulated dispersers successfully
-  moved and the color of each arrow shows the average number of steps (4-hourly
-  movements) that were necessary to realize those connections. Additionally, the
-  line thickness indicates the relative number of dispersers originating from a
-  NP that realized those connections. Note that a similar network view could be
-  adopted to investigate connectivity between other protected areas need not to
-  be restricted to NPs.}
+  \caption{Map of inter-patch connectivity in relation to dispersal duration,
+  highlighting connections between NPs (dark green). Yellow bubbles represent
+  the center of the different NPs and are sized in relation to the number of
+  simulated dispersers originating from each park. Black dots represent NPs that
+  were smaller than 700 km\textsuperscript{2} and therefore did not serve as
+  source areas. Arrows between NPs illustrate between which NPs the simulated
+  dispersers successfully moved and the color of each arrow shows the average
+  number of steps (4-hourly movements) that were necessary to realize those
+  connections. Additionally, the line thickness indicates the relative number of
+  dispersers originating from a NP that realized those connections. Note that a
+  similar network view could be adopted to investigate connectivity between
+  other protected areas need not to be restricted to NPs.}
   \label{InterpatchConnectivity}
@@
 \subsection{Short Summary}
-We used empirical data from a free-ranging population of African wild dogs to
-demonstrate the application of a three-step workflow to simulate dispersal
-trajectories and assess landscape connectivity. In step one, we used ISSFs to
-parametrize a fully mechanistic movement model describing how dispersing wild
-dogs move through the landscape. In step two, we employed the movement model to
-simulate 80,000 dispersal trajectories across the extent of the KAZA-TFCA, the
-world's largest transboundary conservation area. In step three, we translated
-the simulated trajectories into three complementary connectivity maps, each
-emphasizing a different aspect of landscape connectivity (e.g. frequently
-traversed areas, critical dispersal corridors and bottlenecks, and the presence
-and intensity of functional links between suitable patches). Importantly, such
-simulations from ISSFs overcome several conceptual shortcomings inherent to more
-traditional connectivity modeling techniques, such as LCPA and CT.
+Here, we proposed a three-step approach to assess landscape connectivity via
+simulated dispersal trajectories, and demonstrated its application using
+empirical data from a free-ranging population of African wild dogs. In step one,
+we used ISSFs to parametrize a fully mechanistic movement model describing how
+individuals move through the landscape. In step two, we employed the movement
+model to simulate dispersal trajectories across the landscape. In step three, we
+translated the simulated trajectories into three complementary connectivity
+maps, each emphasizing a different aspect of landscape connectivity (e.g.
+frequently traversed areas, critical dispersal corridors and bottlenecks, and
+the presence and intensity of functional links between suitable patches).
+Importantly, such simulations from ISSFs overcome several conceptual
+shortcomings inherent to more traditional connectivity modeling techniques, such
+as LCPA and CT.
 
@@
 model additive effects between habitat and movement covariates
-\citep{Avgar.2016, Signer.2021} furthermore allowed us to formally capture that
+\citep{Avgar.2016, Signer.2017} furthermore allowed us to formally capture that
 dispersing wild dogs move slower and more tortuous in areas covered by water,
@@
 \subsection{Simulation}
-Our simulation of 80,000 dispersal trajectories moving 2'000 steps across the
-KAZA-TFCA required five days of computation on a desktop machine (AMD Ryzen 7
-2700X processor with 8 x 3.6 GHz and 16 logical cores, 64 GB of RAM). The long
-simulation time was primarily caused by the massive extent of the study area
-considered (ca. 1.8 Mio km\textsuperscript{2}) and the large number of
-trajectories simulated. Most connectivity studies focus on smaller study areas
-(e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017,
-Zeller.2020}) and will therefore achieve faster simulation times (given the same
-spatial resolution across covariates). We also believe that fewer simulated
-trajectories will often suffice, as the relative traversal frequency by
-simulated trajectories through randomly placed checkpoints across our study area
-converged already after 10,500 runs. The number of required simulations to
-achieve reliable estimates of connectivity will, however, vary depending on the
-structure of the landscape and the dispersal behavior of the focal species
-\citep{Gustafson.1996}. For species that disperse short distances through
-homogeneous environments, few simulations may suffice to gauge connectivity,
-whereas for species that disperse over long distances through heterogeneous
-habitats, a large number of simulations will be required to sufficiently explore
-the spectrum of possible routes.
+While our approach of simulating dispersal proved usedful to assess landscape
+connectivity, it was computationally very costly. Our simulation of 80,000
+dispersal trajectories moving 2'000 steps across the KAZA-TFCA required five
+days of computation on a regular desktop machine (AMD Ryzen 7 2700X processor
+with 8 x 3.6 GHz and 16 logical cores, 64 GB of RAM). The long simulation time
+was primarily caused by the massive extent of the study area considered (ca. 1.8
+Mio km\textsuperscript{2}) and the large number of trajectories simulated. Most
+connectivity studies focus on smaller study areas (e.g. \citealp{Kanagaraj.2013,
+Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020}) and will therefore achieve
+faster simulation times (given the same spatial resolution across covariates).
+We also believe that fewer simulated trajectories will often suffice, as the
+relative traversal frequency by simulated trajectories through randomly placed
+checkpoints across our study area converged already after 10,500 runs. The
+number of required simulations to achieve reliable estimates of connectivity
+will, however, vary depending on the structure of the landscape and the
+dispersal behavior of the focal species \citep{Gustafson.1996}. For species that
+disperse short distances through homogeneous environments, few simulations may
+suffice to gauge connectivity, whereas for species that disperse over long
+distances through heterogeneous habitats, a large number of simulations will be
+required to sufficiently explore the spectrum of possible routes.
 
@@
 usually an unrealistic assumption \citep{Elliot.2014, Abrahms.2017, Cozzi.2020}.
-At the same time, simulations permit to detect potential routes that do not
-necessarily lead into suitable habitats \citep{Dwernychuk.1972, VanDerMeer.2014}
-and into areas with a high susceptibility for human wildlife conflicts
-\citep{Cushman.2018}.
+Simulations also enable the detection of potential routes that do not lead into
+suitable habitats \citep{Dwernychuk.1972, VanDerMeer.2014} but into areas with a
+high susceptibility for human wildlife conflicts \citep{Cushman.2018}.
 
@@
 Despite the many benefits and great flexibility offered by simulations from
-ISSFs, one also must be aware of the associated non-trivial but important
-modeling decisions. Here, we will further elaborate on four modeling decisions
-concerning: (1) the number of simulated individuals, (2) the location of source
-points, (3) the simulated dispersal duration, and (4) behavior at map
-boundaries.
+ISSFs, one must also be aware of the associated non-trivial but important
+modeling decisions. Here, we will further elaborate on four modeling decisions:
+(1) the number of simulated individuals, (2) the location of source points, (3)
+the simulated dispersal duration, and (4) behavior at map boundaries.
 
@@
 points could be distributed homogeneously and only later be weighted when
-deriving connectivity metrics. In any case, the challenge of selecting
-meaningful source points is not unique to individual-based simulations but also
-applies to LCPA or CT.
+computing the heatmap, betweenness map, or inter-patch connectivity map. In any
+case, the challenge of selecting meaningful source points is not unique to
+individual-based simulations but also applies to LCPA or CT.
 
@@
 \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. This
-approach also has the advantage that it allowed us to subsample simulated
-trajectories to shorter durations after their simulation and to investigate the
-sensitivity of our connectivity maps with respect to exact dispersal durations
-(Figures S4 and S5).
+approach also had the advantage that it allowed us to subsample simulated
+trajectories to shorter durations after their simulation and thereby to
+investigate the sensitivity of our results with respect to exact dispersal
+durations (Figures S4 and S5).
 
@@
 of attraction (e.g. \cite{Signer.2017}), some trajectories will inevitably
-approach a map boundary where one of the generated random steps might leave the
-study area. In this case, one solution would be to simply terminate the
-simulation of the affected trajectory, implicitly assuming that the respective
-animal left the study area permanently. However, this approach might become
-problematic in cases where individuals are released near map borders, as already
-a single random step leaving the study area will break the simulation, thus
-biasing connectivity along map borders. Instead of breaking the simulation, we
-created a buffer zone \citep{Koen.2010} and resampled random steps until they
-fully lied within the study area. This approach proved an effective solution to
-overcome problems with boundary effects.
+approach a map boundary and one or more of the generated random steps will leave
+the study area. One solution to deal with such cases would be to simply
+terminate the simulation of the affected trajectory, implicitly assuming that
+the respective animal left the study area. However, this approach might produce
+ambiguous results in cases where individuals are released near map borders,
+since already a single random step leaving the study area will break the
+simulation, thus resulting in biased connectivity estimates along map borders.
+Rather than halting the simulation, we created a buffer zone \citep{Koen.2010}
+and resampled random steps until they fully lied within the study area. This
+proved to be an effective solution to overcome problems with boundary effects.
 
 \subsection{Conclusion}
-In summary, we proposed and applied a simple three-step workflow that relies on
+In summary, we proposed and applied a simple three-step approach that relies on
 ISSF-analysis and enables the simulation of dispersal trajectories and the
-assessment of landscape connectivity. The proposed workflow overcomes several of
+assessment of landscape connectivity. The proposed approach overcomes several of
 the conceptual shortcomings inherent to LCPA and CT, such as the assumption of
@@
 movements and connectivity via individual-based simulations, while at the same
-time stressing some of the non-trivial modeling decisions involved.
+time stressing some of the non-trivial modeling decisions involved. We also hope
+to provide a useful framework that helps unifying and streamlining the
+application of individual-based simulation for assessing landscape connectivity.
 
@@
 \citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
-proposed workflow using simulated data are provided through Github
+proposed approach using simulated data are provided through Github
 (\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all

### 2022-01-24 Chapter_1/04_Manuscript/Manuscript_Support.tex
updates

@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ A Three Steps Workflow for Simulating Dispersal
-Trajectories and Assessing Landscape Connectivity: A Case Study on the
-Endangered African Wild Dog}
+\title{\textbf{Appendix}\\ A Thee-Step Approach for Assessing Landscape
+Connectivity via Simulated Dispersal: African Wild Dog Case Study}
 

### 2022-01-24 Chapter_2/04_Manuscript/Manuscript_Support.tex
updates

--- 2020-08-21 Chapter_1/04_Manuscript/Manuscript_Support.tex
@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Step by Step: Using Integrated Step Selection
-Analysis to Simulate Wild Dog Dispersal and Assess Landscape Connectivity}
+\title{\textbf{Appendix}\\ Dispersing through Seasonal Landscapes}
 
@@
 \author{
-  David D. Hofmann\textsuperscript{1,\S} \and
+  David D. Hofmann\textsuperscript{1,2,\S} \and
+  Dominik M. Behr\textsuperscript{1,2} \and
   John W. McNutt\textsuperscript{2} \and
   Arpat Ozgul\textsuperscript{1} \and
-  Gabriele Cozzi\textsuperscript{1,2} \and
-  Dominik M. Behr\textsuperscript{1,2}
+  Gabriele Cozzi\textsuperscript{1,2}
 }
@@
 
-\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
-Botswana.
+ \textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
+ Maun, Botswana.
 
@@
 
-\textbf{Running Title:} Simulating Wild Dog Dispersal.
+\textbf{Running Title:} Dispersing through Seasonal Landscapes
 
@@
 
-% % Create linenumbers
+% Create linenumbers
 % \linenumbers
@@
 %------------------------------------------------------------------------------
-%	Appendix S1: All Movement Models
+%	Appendix S1: Pan Mapping
 %------------------------------------------------------------------------------
-\section{Movement Models}
-% latex table generated in R 3.6.3 by xtable 1.8-4 package
-% Thu Jul  9 09:17:26 2020
-\begin{table}[hbpt]
-  \caption{Results from the forward model selection procedure based on Akaike's
-  Information Criterion (AIC; \citealp{Burnham.2002}) for the movement model.
-  The base model upon which we based our movement model is depicted in the last
-  row. We omitted any other model with an AIC weight of zero.}
-  \label{ModelAICs}
-  \begin{center}
-    \resizebox{\textwidth}{!}{
-      \begin{threeparttable}
-        \begin{tabular}{llllll}
-        \toprule
-          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik & Model ID \\
-          \midrule
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 & 1\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 & 2\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 & 3\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 & 4\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 & 5\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 & 6\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 & 7\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 & 8\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 & 9\\
-          Base Model + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 & 10\\
-          \hdashline
-          ...\\
-          \hdashline
-          Base Model: cos(ta) + log(sl) + W + T + DTW + HI + S & 90091.40 & 787.67 & 0.00 & -45030.70 & NA\\
-         \bottomrule
+\newpage
+\section{Pan Mapping}
+We previously employed a floodmapping algorithm to generate weekly updated
+floodmaps, which allowed us to render the seasonal pulsing of the Okavango
+Delta. However, due to the coarse resolution of MODIS satellite imagery, based
+upon which the floodmaps were derived, prohibited a seasonal representation of
+more fine-scaled waterbodies, such as waterpans. In an attempt to overcome this
+limitation, we sought out and evaluated alternative satellite products that
+would provide better spatial resolution while still retaining sufficient
+temporal resolution to meaningfully render seasonal patterns. Candidate products
+comprised Landsat 7, Landsat 8, and Sentinel 2 satellite imagery, mainly because
+their data is freely accessible, provides a spatial resolution between 10 and 30
+meters, and because the satellites have revisit times between 5 and 16 days,
+thus allowing to produce monthly updated images. A brief overview of the
+characteristics of each product is presented in \Cref{SatelliteData}.
+
+\begin{table}
+  \caption{Comparison}
+  \label{SatelliteData}
+  \begin{tabular}{llll}
+             & Availability   & Spatial Resolution & Temporal Resolution                             \\
+  Landsat 7  & 1999 - Present & 15-60m             & Every 16th day                                  \\
+  Landsat 8  & 2013 - Present & 15-100m            & Every 16th day                                  \\
+  Sentinel 2 & 2015 - Present & 10-60m             & Every 10th day / every 5th day since March 2017
   \end{tabular}
-       \begin{tablenotes}
-         \item \textit{Note:} W = Water, DTW = Distance To Water, S =
-         Shrubs/Grassland, T = Trees, P = Protected, HI = Human Influence, RC =
-         Road Crossing, DTR = Distance To Roads, MA = Main Activity.
-       \end{tablenotes}
-     \end{threeparttable}
-    }
-  \end{center}
 \end{table}
 
-\begin{figure}[htpb]
+While Landsat 7's temporal availability would have aligned with the time window
+during which we collected GPS data on dispersing wild dogs, Landsat 7's scan
+line detector has been failing since May 31 2003, resulting in data gaps between
+adjacent tiles, thus rendering this product unusable for our purposes.
+
+With Landsat 8 and Sentinel 2 imagery remaining as viable candidates, we
+prepared a set of training polygons, consisting of the land cover classes
+\textit{Dryland}, \textit{Water}, and \textit{Wet Pans}. Although our primary
+goal was to detect wet pans, we deliberately added additional classes,
+anticipating that their inclusion would facilitate the separation of reflectance
+values. Due to the large extent considered and limited time available in the
+field, \textit{in situ} collection of training data was unfeasible and we
+instead opted for \textit{on-screen selection} of training data. More
+specifically, we used Google Earth to digitize areas that were clearly
+identifiable as either dryland, water, or wet pan. At the highest zoom level
+that is allowed in Google Earth, these categories are visually easy to
+distinguish \Cref{PanMapping}. To verify a sharp distinction between wet and dry
+pans, we placed several dryland polygons on areas that are seasonally covered by
+water \Cref{PanMapping}. Depending on the area of interest, Google Earth
+displays high-resolution satellite imagery collected on different dates, so
+assigned to each training polygon the date of acquisition of the displayed
+satellite image. This later allowed us to train the land cover classifier using
+Sentinel 2 or Landsat 8 data imagery acquired on the same dates. This procedure
+resulted in xx polygons (xx dryland, xx water, xx dry-pan, and xx wet-pan)
+distributed across xx dates.
+
+\begin{figure}[htbp]
  \begin{center}
-    \includegraphics[width = \textwidth]{99_AllMovementModels}
-    \caption{Coefficients of all movement models with an AIC weight above one.}
-    \label{AllMovementModels}
+  \includegraphics[width = \textwidth]{99_PanMapping.pdf}
+  \caption{Example of a \textit{wet pan} and \textit{dryland} training polygon
+  digitized on Google Earth. The extent of water can easily be gauged at this
+  zoom level in Google Earth. While the dryland polygon in this case covered an
+  area that is seasonally covered by water, other dryland polygons comprised
+  areas that are never inundated. However, to ensure reliable differentiation
+  between wet and dry pans, we included several dryland polygons located in dry
+  pans.}
+  \label{PanMapping}
  \end{center}
 \end{figure}
+
+% Download of data and preprocessing of satellite data (atmospheric correction,
+% cloud cover etc.). Computation of indices.
+
+To train a land-cover classifier, we selected 3'000 training points within the
+digitized training polygons following a stratified equal random sampling scheme
+\citep{Shetty.2021}. That is, we ensured that a total of 1'000 random points per
+land cover category were generated. This was necessary to ensure that pans,
+which only accounted for a very small fraction of the study area, were
+sufficiently well sampled. In fact, stratified equal random sampling has been
+shown to provide good class-level accuracy for minority classes, so we deemed
+this approach suitable for our purposes. At each generated random point we then
+extracted reflectance values of Sentinel 2 and Landsat 8 bands that temporally
+aligned with the date assigned to the respective random point
+\Cref{Reflectances}. For instance, if a random point fell into a polygon that
+was digitized using a Google Earth image generated on Dec 18 2018, we extracted
+values from the Sentinel 2 and Landsat 8 composite for December 2018. Using the
+extracted values, we parametrized a Random Forest (RF) classification model.
+
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_Reflectances.png}
+  \caption{}
+  \label{Reflectances}
+ \end{center}
+\end{figure}
+
+To validate the predictive power of the RF classifier, we employed 5-fold
+cross-validation. For this, we split the data into 5 groups and repeatedly
+fitted the RF model using 80\% of the data while predicting land cover
+categories on the remaining 20\%. We then contrasted predictions with true
+classes and computed confusion matrices to obtain estimates of the classifier's
+specificity, sensitivity, and overall accuracy. The results from this validation
+show that the RF classifier achieves very high overall accuracy, for both
+Landsat 8 and Sentinel 2 imagery \Cref{ClassificationValidation}.
+
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_ClassificationValidation.png}
+  \caption{}
+  \label{ClassificationValidation}
+ \end{center}
+\end{figure}
+
+While both Landsat 8 and Sentinel provided very good results, we opted for
+Sentinel 2 data, mainly due Sentinel 2's better temporal coverage and marginally
+higher resolution, which we thought was crucial for pan detection. To download
+satellite imagery of interest, we first generated a spatio-temporal moving
+window comprising all GPS locations from dispersing wild dogs of a given month,
+buffered by 100km. We checked for intersections of each month's moving window
+with Sentinel 2's tile-grid and determined the tiles to be downloaded for every
+month. For the intersecting tiles, we then downloaded all available Sentinel 2
+images falling into the respective month. Download and pre-processing of all
+Sentinel 2 images was achieved using the \textsf{sen2r} package in R. While
+atmospherically corrected Sentinel 2 data is currently being fed into ESA's
+database, a large portion has not yet been processed from level 1C (top of the
+atmosphere, TOA) to level 2C (bottom of the atmosphere, BOA). Thus, whenever
+level 2C data was not available, we downloaded level 1C data and pre-processed
+the imagery into level 2C using sen2cor \citep{MainKnorn.2017}, again invoked
+through \textsf{sen2r}.
 

### 2022-01-24 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/Betweenness.R

@@ -160,8 +160,8 @@ p1 <- ggplot() +
       x        = NULL
     , y        = NULL
     , fill     = NULL
-    , title    = "Betweenness"
-    , subtitle = "After 2'000 Steps"
+    # , title    = "Betweenness"
+    # , subtitle = "After 2'000 Steps"
   ) +
   theme(
       legend.position  = "bottom"

### 2022-01-24 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/InterpatchConnectivity.R

@@ -195,8 +195,8 @@ p1 <- ggplot() +
       x        = NULL
     , y        = NULL
     , fill     = NULL
-    , title    = "Interpatch Connectivity"
-    , subtitle = "In Relation to Dispersal Duration"
+    # , title    = "Interpatch Connectivity"
+    # , subtitle = "In Relation to Dispersal Duration"
   ) +
   guides(
       size  = guide_legend(title.position = "top", order = 2)
@@ -315,8 +315,8 @@ p3 <- ggplot() +
       x        = NULL
     , y        = NULL
     , fill     = NULL
-    , title    = "Interpatch Connectivity"
-    , subtitle = "In Relation to Dispersal Duration"
+    # , title    = "Interpatch Connectivity"
+    # , subtitle = "In Relation to Dispersal Duration"
   ) +
   guides(
     size = guide_legend(title = "Number of Simulations", title.position = "top")

### 2022-01-24 Chapter_2/02_R-Scripts/01_DataCleaning/12_SentinelDownload.R

@@ -1,0 +1,153 @@
+################################################################################
+#### Sentinel Data Download
+################################################################################
+# Description: Download of the files specified in the previous script
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(sen2r)         # To automate the download of Sentinel 2 data
+library(tidyverse)     # For data wrangling
+library(pbmcapply)     # For multicore abilities with progress bar
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Note output directories
+outdir_l1c <- "/media/david/Elements/L1C"
+outdir_l2a <- "/media/david/Elements/L2A"
+
+# Create directory into which we will download the data
+dir.create(outdir_l1c, showWarnings = F)
+dir.create(outdir_l2a, showWarnings = F)
+
+# Make sure the directories exist
+dir.exists(outdir_l1c)
+dir.exists(outdir_l2a)
+
+# Login to scihub
+write_scihub_login("dodx9", "Scihubbuster69_")
+
+# Read the file containing all data that we need to download
+todownload <- read_rds("/media/david/Elements/Todownload.rds")
+
+# Depending on the availability of L1C or L2A products, we put the downloaded
+# data into two different directories
+todownload$outdir <- ifelse(todownload$level == "1C", outdir_l1c, outdir_l2a)
+
+# Give each job a unique ID
+todownload$JobID <- 1:nrow(todownload)
+
+# Instead of requesting all downloads at once, I'd like to go through the files
+# in groups. Thus, let's create groups with around 20 members in each group
+n <- 20
+ngroups <- round(nrow(todownload) / n, -1)
+todownload$GroupID <- sort(rep(1:260, length.out = nrow(todownload)))
+print(todownload)
+
+# Depending on the group we will use different login credentials
+todownload$Username <- ifelse(todownload$GroupID %% 2 == 1, "dodx9", "dodx92")
+todownload$Password <- "Scihubbuster69_"
+
+# We only need to download files that haven't been downloaded yet. Thus, let's
+# check which files are already present
+exists <-
+  dir.exists(paste0(todownload$outdir, "/", todownload$name)) |
+  dir.exists(paste0(todownload$outdir, "/", todownload$name))
+todownload <- subset(todownload, !exists)
+
+# Order jobs so that the newest dates are coming first -> not yet archvied in
+# the long-term archive
[…]

### 2022-01-31 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

@@
 
-\textbf{Keywords:} dispersal, simulation, movement, integrated step selection
+\textbf{Keywords:} dispersal, simulation, movement, integrated step-selection
 function, Kavango-Zambezi Transfrontier Conservation Area, landscape
@@
 % already-existing methods to assess connectivity using simulated dispersal
-% trajectories. In step one, we use integrated step selection functions to
+% trajectories. In step one, we use integrated step-selection functions to
 % parametrize a mechanistic movement model rendering dispersal behavior. In step
@@
 % movement preferences, while also uncovering crucial dispersal corridors. With
-% this analysis, we demonstrate that simulations from integrated step selection
+% this analysis, we demonstrate that simulations from integrated step-selection
 % functions offer a simple, yet powerful alternative to traditional connectivity
@@
 and to assess connectivity using empirical GPS movement data and environmental
-habitat layers. In step one, we use integrated step selection functions to fit a
+habitat layers. In step one, we use integrated step-selection functions to fit a
 mechanistic movement model describing habitat and movement preferences of
@@
   moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
-  integrated step selection model that we fitted to the movement data of
+  integrated step-selection model that we fitted to the movement data of
   dispersing African wild dogs. To generate the heatmap, we rasterized and

### 2022-02-07 Chapter_1/04_Manuscript/Manuscript.tex
updates

@@
 assumptions can be relaxed by simulating dispersal movements explicitly across
-the landscape of interest, a unified approach for such simulations is lacking.
+the landscape, a unified approach for such simulations is lacking.
 
 Here, we present a simple three-step approach to simulate dispersal movements
-and to assess connectivity using empirical GPS movement data and environmental
-habitat layers. In step one, we use integrated step-selection functions to fit a
-mechanistic movement model describing habitat and movement preferences of
-dispersing individuals. In step two, we apply the parameterized movement model
-to simulate dispersal trajectories. In step three, we derive three complementary
-connectivity maps: a heatmap that highlights frequently traversed areas, a
-betweenness map that pinpoints dispersal corridors, and a map of inter-patch
-connectivity that indicates the presence and intensity of functional links
-between habitat patches. As a case study, we applied the approach to GPS data
-from dispersing individuals of the endangered African wild dog (\textit{Lycaon
-pictus}) inhabiting the Kavango-Zambezi Transfrontier Conservation Area
-(KAZA-TFCA).
-
-Results from the case study show that wild dogs preferably disperse with
-directional persistence in the vicinity to water bodies, through areas with
-little human influence and sparse woodland cover. Dispersal simulations and
-subsequent connectivity maps revealed several dispersal hotspots and corridors
-across the extent of the KAZA-TFCA. Connectivity between NPs inside the
-KAZA-TFCA was good, yet few dispersers successfully moved from Zambia's NPs into
-neighboring areas. By rendering interactions between movement preferences and
-habitat preferences, our movement model substantially outperformed a model that
-omitted such interactions.
-
-Ultimately, we show that a simulation-based approach that leverages on
-step-selection functions offers a simple yet powerful alternative to traditional
-connectivity modeling techniques. Such an approach not only makes fewer
-biologically unrealistic assumptions but also permits a more mechanistic
-understanding of dispersal and landscape connectivity. Our approach is thus
-useful for a variety of applications in ecological, evolutionary, and
-conservation research.
+and to assess connectivity using empirical GPS movement data and a set of
+habitat covariates. In step one, integrated step-selection functions are used to
+fit a mechanistic movement model describing habitat and movement preferences of
+dispersing individuals. In step two, the parameterized model is used to simulate
+dispersal trajectories across the study area. In step three, a set of three
+complementary connectivity maps is derived: a heatmap that highlights frequently
+traversed areas, a betweenness map that pinpoints dispersal corridors, and a map
+of inter-patch connectivity that indicates the presence and intensity of
+functional links between habitat patches. As a case study, we applied the
+approach to GPS data collected on dispersing individuals of the endangered
+African wild dog (\textit{Lycaon pictus}) inhabiting the Kavango-Zambezi
+Transfrontier Conservation Area (KAZA-TFCA) in southern Africa.
+
+Besides providing insights into the movement and habitat preferences of
+dispersing individuals, our movement model also rendered how movement behavior
+depended on habitat conditions. The model thereby substantially outperformed a
+model that omitted such interactions. While we simulated a total of 80,000
+dispersal trajectories, metrics of connectivity only marginally changed beyond
+10,500 simulations, suggesting that fewer simulations would have sufficed. The
+three connectivity maps derived from simulations revealed several dispersal
+hotspots and corridors across the extent of our study area. Each map emphasized
+a different aspect of landscape connectivity, thus highlighting their
+complementary nature. Overall, our case study demonstrates that a
+simulation-based approach utilizing step-selection functions offers a simple yet
+powerful alternative to traditional connectivity modeling techniques. Our
+approach not only makes fewer unrealistic assumptions about dispersal but also
+permits a more mechanistic understanding of dispersal movements and landscape
+connectivity. It is thus useful for a variety of applications in ecological,
+evolutionary, and conservation research.
 
@@
 Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and colonize
-unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. Importantly, the
-ability to disperse depends on a sufficient degree of landscape connectivity
+unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
+to disperse depends on a sufficient degree of landscape connectivity
 \citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
@@
 not only necessitates a comprehensive understanding of the factors that limit
-dispersal movements, but also an appropriate model to estimate connectivity
+dispersal, but also an appropriate model to estimate connectivity
 \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the most commonly
-used models to assess connectivity are least-cost path analysis (LCPA;
-\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
-McRae.2008}). Unfortunately, both models rest on assumptions that appear
-unsuitable for dispersers, calling for the development of alternative
-approaches. One promising alternative is to assess landscape connectivity via
-simulated dispersal trajectories generated from individual-based movement models
-(IBMMs, \citep{Diniz.2019}).
+used models are least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and
+circuit theory (CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both
+models rest on assumptions that appear unsuitable for dispersers, thus calling
+for the development of alternative approaches. One promising alternative is to
+assess landscape connectivity via simulated dispersal trajectories generated
+from individual-based movement models (IBMMs, \citealp{Diniz.2019}). However,
+the large number of subjective modeling decisions required by IBMMs renders
+among-system comparisons difficult.
 
@@
 autocorrelated movements are regularly observed \citep{Cozzi.2020,
-Hofmann.2021}, implying that dispersal trajectories are usually strongly
+Hofmann.2021}, meaning that dispersal trajectories are usually strongly
 directional. Moreover, because both models require static permeability or
 resistance surfaces as input, they are unable to reflect the temporal dimension
-of dispersal, meaning that statements about the expected duration for moving
+of dispersal, such that statements about the expected duration for moving
 between habitat patches are impossible \citep{Martensen.2017, Diniz.2019}.
@@
 Furthermore, IBMMs generate movement sequentially, i.e. they generate a series
-of steps, so that the temporal dimension of dispersal movements becomes explicit
-and allows modeling autocorrelation between consecutive steps
-\citep{Diniz.2019}. Finally, simulations from IBMMs do not enforce movement or
-connections towards preconceived endpoints, thereby preventing biases arising
-from misplaced endpoints. Despite these advantages, a unifying framework to
-simulate dispersal and assess connectivity using IBMMs is lacking. Considering
-the large number of subjective decisions entailed by IBMMs, an approach that
-streamlines and unifies the application of dispersal simulations to assess
-connectivity will, however, be critical to safeguard interspecfic comparability
-and reproducability.
+of steps, so that the temporal dimension of dispersal becomes explicit and
+allows modeling autocorrelation between consecutive steps \citep{Diniz.2019}.
+Finally, simulations from IBMMs do not enforce movement or connections towards
+preconceived endpoints, thereby preventing biases arising from misplaced
+endpoints. Despite these advantages, a unifying framework to simulate dispersal
+and assess connectivity using IBMMs is lacking. Considering the large number of
+subjective decisions entailed by IBMMs, an approach that streamlines and unifies
+the application of dispersal simulations to assess connectivity will, however,
+be critical to safeguard interspecfic comparability and reproducability among
+studies.
 
@@
 dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
-step one, we combine GPS movement data of dispersing individuals with relevant
-habitat covariates to fit a mechanistic movement model using integrated step
-selection functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the
-study species' habitat kernel (i.e. habitat preferences), its movement kernel
-(i.e. movement preferences/capabilities), and potential interactions between the
-two \citep{Avgar.2016, Fieberg.2021}. In step two, we simulate individual
-dispersal trajectories using the parametrized movement model. Comparable
-simulations have already been applied to estimate steady-state utilization
-distributions of resident individuals \citep{Potts.2013, Signer.2017} and to
-model landscape connectivity, yet disregarding interdependencies between habitat
-and movement kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we
-convert simulated trajectories into three complementary connectivity maps, each
-highlighting a different aspect of connectivity; (i) a heatmap revealing areas
-that are frequently traversed by dispersers (e.g. \citealp{Hauenstein.2019,
-Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
-bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
-inter-patch connectivity, depicting the presence and intensity of specific
-connections, as well as the average dispersal duration required for the
-realization of those connections (e.g. \citealp{Gustafson.1996,
-Kanagaraj.2013}).
+step one, GPS movement data of dispersing individuals is combined with habitat
+covariates to fit a mechanistic movement model via integrated step-selection
+functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the study
+species' habitat kernel (i.e. habitat preferences), its movement kernel (i.e.
+movement preferences/capabilities), as well as potential interactions between
+the two \citep{Avgar.2016, Fieberg.2021}. In step two, the parametrized movement
+model is used to simulate individual dispersal trajectories starting from
+randomly placed source points. Comparable simulations have already been applied
+to estimate steady-state utilization distributions of resident individuals
+\citep{Potts.2013, Signer.2017} and to model landscape connectivity, yet
+disregarding interdependencies between habitat and movement kernels
+\citep{Clark.2015, Zeller.2020}. Finally, in step three, the simulated
+trajectories are converted into three complementary connectivity maps; (i) a
+heatmap revealing areas that are frequently traversed by dispersers (e.g.
+\citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map delineating
+dispersal corridors and bottlenecks (e.g. \citealp{BastilleRousseau.2018}),
+(iii) and a map of inter-patch connectivity, depicting the presence and
+intensity of functional links between habitat patches, as well as the average
+dispersal duration required for the realization of those connections (e.g.
+\citealp{Gustafson.1996, Kanagaraj.2013}).
+
+% Here, we use GPS data of 16 dispersing wild dogs originating from a
+% free-ranging population in northern Botswana to parametrize a mechanistic
+% movement model, which we then employed to simulate 80,000 dispersal
+% trajectories across the landscape of the world's largest transboundary
+% conservation area, the Kavango-Zambezi Transfrontier Conservation Area
+% (KAZA-TFCA).
 
@@
 kilometers \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}. Although
-previous studies have investigated connectivity using LCPA \citep{Hofmann.2021}
-or CT \citep{Brennan.2020}, a more comprehensive and mechanistic understanding
-of connectivity is missing for this species (but see \citealp{Creel.2020}). With
-fewer than 6,000 free-ranging wild dogs remaining in fragmented subpopulations
-\citep{Woodroffe.2012}, reliable information on landscape connectivity is
-essential for the conservation of this endangered carnivore. Here, we use GPS
-data of 16 dispersing wild dogs originating from a free-ranging population in
-northern Botswana to parametrize a mechanistic movement model, which we then
-employed to simulate 80,000 dispersal trajectories across the landscape of the
-world's largest transboundary conservation area, the Kavango-Zambezi
-Transfrontier Conservation Area (KAZA-TFCA). We anticipated that simulations
-based on our three-step approach would overcome several of the above highlighted
-conceptual shortcomings of traditional connectivity models and provide a more
-comprehensive view on dispersal behavior and landscape connectivity.
+previous studies have investigated connectivity for this species using LCPA
+\citep{Hofmann.2021} or CT \citep{Brennan.2020}, a more comprehensive and
+mechanistic understanding of dispersal and connectivity is missing (but see
+\citealp{Creel.2020}). Nevertheless, with about 6,000 free-ranging wild dogs
+remaining in fragmented subpopulations \citep{Woodroffe.2012}, reliable
+information on dispersal behavior and landscape connectivity is essential for
+the conservation of this endangered carnivore. We anticipated that a
+connectivity assessment based on our three-step approach would overcome several
+of the above highlighted conceptual shortcomings of traditional connectivity
+models, while providing a more detailed view on movement behavior during
+dispersal.
 
@@
 \section{Methods}
-\subsection{Study Area}
+\subsection{Case Study}
+\subsubsection{GPS Data}
+We applied the three step approach mentioned above to GPS movement data from 16
+dispersing African wild dog coalitions (7 female and 9 male coalitions). This
+data has been collected between 2011 and 2019 from a free-ranging wild dog
+population in northern Botswana (details on the data collection can be found in
+\cite{Cozzi.2020} and \cite{Hofmann.2021}). During dispersal, GPS collars
+recorded a fix every 4 hours and regularly transmitted data over the Iridium
+satellite system. To ensure regular time intervals between GPS fixes, we removed
+any fixes that were not successfully obtained at the desired 4-hour schedule
+(allowing for a tolerance of \( \pm \) 15 minutes). To prepare the data for
+step-selection analysis, we converted the fixes (n = 4'169) into steps, where
+each step represented the straight-line movement between two consecutive GPS
+fixes \citep{Turchin.1998}.
+
+% To delineate periods of dispersal, we determined the exact timing of emigration
+% from the natal pack and settlement in a new territory using direct field
+% observations and through visual inspection of the net squared displacement (NSD)
+% metric. The NSD metric measures the squared Euclidean distance of a GPS
+% relocation to a reference point \citep{Borger.2012}, which we set to the center
+% of the territory of each dispersers' natal pack. Thus, dispersal was deemed to
+% have started once an individual left its natal territory and ended once the NSD
+% metric remained constant, indicating settlement. Because behavior during
+% dispersal is more pertinent to landscape connectivity than behavior during
+% residence \citep{Elliot.2014, Abrahms.2017}, we only considered data collected
+% during dispersal for our analysis.
+
+\subsubsection{Study Area}
 Our simulation of dispersal trajectories and assessment of connectivity spanned
-across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA),
-an area of approximately 1.3 Mio. km\textsuperscript{2} (\Cref{StudyArea}a and
-b). The KAZA-TFCA is the world's largest transboundary conservation area and
-comprises parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia, thus hosting
-a rich diversity of landscapes, ranging from savannah to grassland and from dry
-to moist woodland habitats. In its center lies the Okavango Delta, a dominant
-hydro-geographical feature and the world's largest flood-pulsing inland delta.
-Large portions of the KAZA-TFCA are formally protected in the form of national
-parks (NPs) or other protected areas, yet a considerable portion of the
-landscape is still human-dominated (e.g. roads, agricultural sites, and
-settlements).
+across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA,
+\Cref{StudyArea}a and b). The KAZA-TFCA is the world's largest transboundary
+conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
+and Zambia, thus hosting a rich diversity of landscapes, ranging from savannah
+to grassland and from dry to moist woodland habitats. In its center lies the
+Okavango Delta, a dominant hydro-geographical feature and the world's largest
+flood-pulsing inland delta. Large portions of the KAZA-TFCA are formally
+protected in the form of national parks (NPs) or other protected areas, yet a
+considerable portion of the landscape is still human-dominated (e.g. roads,
+agricultural sites, and settlements). We used NPs as patches to calculate
+inter-patch connectivity (cf. Section 2.4.3). The decision to only consider NPs
+as patches was purely out of simplicity and does not imply that dispersal
+between other protected areas does not occur.
 
@@
 
-\subsection{Data Collection and Preparation}
-\subsubsection{GPS Data}
-We collected GPS movement data on 16 dispersing coalitions of African wild dogs
-(7 female and 9 male coalitions) between 2011 and 2019 from a free-ranging
-population in northern Botswana (details on the data collection can be found in
-\cite{Cozzi.2020} and \cite{Hofmann.2021}). To delineate periods of dispersal,
-we determined the exact timing of emigration from the natal pack and settlement
-in a new territory using direct field observations and through visual inspection
-of the net squared displacement (NSD) metric. The NSD metric measures the
-squared Euclidean distance of a GPS relocation to a reference point
-\citep{Borger.2012}, which we set to the center of the territory of each
-dispersers' natal pack. Thus, dispersal was deemed to have started once an
-individual left its natal territory and ended once the NSD metric remained
-constant, indicating settlement. Because behavior during dispersal is more
-pertinent to landscape connectivity than behavior during residence
-\citep{Elliot.2014, Abrahms.2017}, we only considered data collected during
-dispersal for our analysis. During dispersal, GPS collars recorded a fix every 4
-hours and regularly transmitted data over the Iridium satellite system. To
-ensure regular time intervals between GPS fixes, we removed any fixes that were
-not successfully obtained at the desired 4-hour schedule (allowing for a
-tolerance of \( \pm \) 15 minutes). We then converted the fixes (n = 4'169) into
-steps, where each step represented the straight-line movement between two
-consecutive GPS fixes \citep{Turchin.1998}.
-
-\subsubsection{Habitat Covariates}
+\subsubsection{Covariates}
 We represented the physical landscape in our study area by the habitat
-covariates \textit{water-cover, distance-to-water, woodland-cover,
+covariates \textsf{water-cover, distance-to-water, woodland-cover,
 shrub/grassland-cover, and human-influence}. To render the seasonal dynamics in
@@
 applied an algorithm that enabled us to obtain weekly updated raster-layers for
-water-cover and distance-to-water from MODIS satellite imagery
-\citep{Wolski.2017, Hofmann.2021}. This algorithm is now implemented in the
-\textsf{floodmapr} package (available on GitHub;
+\textsf{water-cover} and \textsf{distance-to-water} from MODIS satellite imagery
+\citep{Wolski.2017, Hofmann.2021}. This algorithm is now implemented in the {\tt
+floodmapr} package (available on GitHub;
 \url{https://github.com/DavidDHofmann/floodmapr}). To ensure a consistent
@@
 natural logarithm (\textsf{log(sl)}), and the cosine of the relative turning
-angle (\textsf{cos(ta)}). Moreover, we created the binary variable
+angle (\textsf{cos(ta)}), which is a measure of directionality
+\citep{Turchin.1998}. Because wild dog activity is low during the hot midday
+hours \citep{Cozzi.2012}, we additionally created the variable
 \textsf{LowActivity}, indicating whether a step was realized during periods of
 low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
-(17:00 to 09:00 local time, following \citealp{Cozzi.2012}). We performed all
-data preparations, spatial computations, and statistical analysis in R, version
-3.6.6 \citep{R.2020}. Some helper functions were written in {\tt C++} and
-imported into {\tt R} using the {\tt Rcpp} package \citep{Eddelbuettel.2011,
-Eddelbuettel.2013}.
+(17:00 to 09:00 local time). We performed all data preparations, spatial
+computations, and statistical analysis in {\tt R}, version 3.6.6 \citep{R.2020}.
+Some helper functions were written in {\tt C++} and imported into {\tt R} using
+the {\tt Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}.
 
 \subsection{Step 1 - Movement Model}
-We used ISSFs \citep{Avgar.2016} to parametrize a mechanistic movement model for
-dispersing wild dogs. More specifically, we paired each realized (i.e. observed)
-step with 24 random steps, so that a realized step plus its 24 random steps
-formed a 25-step-stratum that received a unique identifier. As suggested by
+We used ISSFs \citep{Avgar.2016} to parametrize a mechanistic movement model
+starting from GPS movement data collected on dispersing individuals. More
+specifically, we paired each realized (i.e. observed) step with a set of 24
+randomly generated alternative steps. A realized step plus its 24 random steps
+formed a 25-step-stratum and received a unique identifier. As suggested by
 \cite{Avgar.2016}, we generated random steps by sampling random turning angles
 from a uniform distribution (\(-\pi, +\pi\)) (which is equivalent to a von Mises
-distribution with location and concentration parameters; \(\mu = \kappa = 0\))
+distribution with location and concentration parameters \(\mu = \kappa = 0\))
 and step lengths from a gamma distribution that was fitted to realized steps
@@
 
-Along each realized and random step, we extracted values from the habitat
+Along each realized and random step, we extracted values from underlying habitat
 covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
-computed averages of each covariate along the steps. We further calculated the
-movement metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)} for each
-step. To facilitate model convergence, we standardized all continuous covariates
-to a mean of zero and a standard deviation of one. Correlations among covariates
-were low (\(|r| < 0.6\); \citealp{Latham.2011}), so we retained all of them for
-modeling.
+computed averages of each covariate along the steps. For each step, we further
+calculated the movement metrics \textsf{sl}, \textsf{log(sl)}, and
+\textsf{cos(ta)}. To facilitate model convergence, we standardized all
+continuous covariates to a mean of zero and a standard deviation of one.
+Correlations among covariates were low (\(|r| < 0.6\); \citealp{Latham.2011}),
+so we retained all of them for modeling.
 
@@
 that animals assigned a selection score \(w(x)\) to each step (\ref{EQ1};
-\citealp{Fortin.2005}). \(w(x)\) depended on the step's associated covariates
-(\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e. relative
-selection strengths; \citealp{Avgar.2017}) towards these covariates (\(\beta_1,
-\beta_2, ..., \beta_n\)):
+\citealp{Fortin.2005}), where \(w(x)\) depended on the step's associated
+covariates (\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e.
+relative selection strengths; \citealp{Avgar.2017}) towards these covariates
+(\(\beta_1, \beta_2, ..., \beta_n\)):
 
@@
 \noindent To estimate preferences (i.e. the \(\beta\)-coefficients), we used
-mixed effects conditional logistic regression analysis \citep{Muff.2020} that we
-implemented using the r-package {\tt glmmTMB} \citep{Brooks.2017}. The method
+mixed effects conditional logistic regression analysis \citep{Muff.2020},
+implemented through the r-package {\tt glmmTMB} \citep{Brooks.2017}. The method
 introduced by \cite{Muff.2020} allows to model random slopes, yet requires to
-fix the variance of the stratum specific intercept to a large value. Hence we
-fixed the stratum specific intercept variance to an arbitrary high value of \(10
-^ 6\). We used disperser identity to model random slopes for all covariates.
-
-The covariate structure of the movement model was based on a habitat selection
-model that was previously developed for dispersing wild dogs (hereafter referred
-to as \textit{based model}, \citealp{Hofmann.2021}). In the base model, no
-interactions among habitat covariates and movement covariates were considered.
-Hence, we expanded the base model and allowed for interactions between all
-movement covariates and habitat covariates, thus reflecting that movement
-behavior may depend on habitat conditions (details in Appendix A1). To determine
-the most parsimonious movement model among model candidates, we ran stepwise
-forward model selection based on Akaike's Information Criterion (AIC,
-\citealp{Burnham.2002}). Finally, we validated the predictive power of the most
-parsimonious model using k-fold cross-validation for case-control studies as
-described in \cite{Fortin.2009}. This validation proves a significant prediction
-in case the Spearman rank correlation of predicted step-ranks and associated
-frequencies under the movement model is significantly greater than under the
-assumption of random preferences (details in Appendix A2).
+fix the variance of the stratum specific intercept to a large value. We
+therefore fixed the variance of the stratum specific intercept to an arbitrary
+high value of \(10 ^ 6\) and used disperser identity to model random slopes for
+all covariates.
+
+The structure of our movement model was based on a habitat selection model that
+was previously developed for dispersing wild dogs (hereafter referred to as
+\textit{base model}, \citealp{Hofmann.2021}). In the base model, no interactions
+among habitat covariates and movement covariates were considered, so we here
+expanded the model and allowed for interactions between all movement covariates
+and habitat covariates, thus reflecting that movement behavior could depend on
+habitat conditions (details in Appendix A1). To determine the most parsimonious
+movement model among model candidates, we ran stepwise forward model selection
+based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}). We
+validated the predictive power of the most parsimonious model using k-fold
+cross-validation for case-control studies as described in \cite{Fortin.2009}.
+This validation proves a significant prediction in case the Spearman rank
+correlation coefficient of predicted step-ranks and associated frequencies under
+the movement model is significantly greater than under the assumption of random
+preferences (details in Appendix A2).
 
 \subsection{Step 2 - Dispersal Simulation}
-We used the most parsimonious movement model to simulate individual dispersal
-trajectories within the study area. The simulation of a dispersal trajectory
-resembled an ``inverted'' ISSF and was set up as follows. (1) We defined a
-random source point and assumed a random initial orientation of the simulated
+We used the most parsimonious movement model to simulate 80'000 individual
+dispersal trajectories across the study area. The simulation of a dispersal
+trajectory resembled an ``inverted'' ISSF and was set up as follows. (1) We
+defined a source point and assumed a random initial orientation of the simulated
 animal. (2) Starting from the source point, we generated 25 random steps by
 sampling turning angles from a uniform distribution (\(-\pi, +\pi\)) and step
-lengths from our fitted gamma distribution. Similarly to the empirical data,
-each random step represented a 4 hours straight-line movement. To prevent
-unreasonably large steps, we restricted sampled step lengths to a maximum of 35
-km (i.e. the farthest dispersal distance traveled within 4 hours in our data).
-(3) Along each random step, we extracted and averaged values from the different
-habitat covariate layers and calculated movement covariates (as explained under
-Step 1). To ensure compatible scales with the fitted movement model, we
-standardized extracted values using means and standard deviations from the
+lengths from our fitted gamma distribution. (3) Along each random step, we
+extracted and averaged values from the different habitat covariate layers and
+computed the movement metrics \textsf{sl}, \textsf{log(sl)}, and
+\textsf{cos(ta)}. To ensure compatible scales with the fitted movement model, we
+standardized covariate values using means and standard deviations from the
 empirical data. (4) We applied the parametrized movement model to predict the
@@
 realized.
+
+As source points for the simulations, we distributed 50,000 points at random
+locations inside protected areas that were larger than the average home range of
+resident wild dog packs (i.e. \(>\) 700 km\textsuperscript{2};
+\citealp{Pomilia.2015}). We placed another 30,000 points randomly inside the
+buffer zone, mimicking potential immigration into the study area (Figure S1).
 
@@
 
-For the simulation, we distributed 80,000 source points within the study area.
-Of these, 50,000 were located inside protected areas that were larger than the
-average home range of resident wild dog packs (i.e. \(>\) 700
-km\textsuperscript{2}; \citealp{Pomilia.2015}), while the remaining 30,000 were
-placed randomly inside the buffer zone, mimicking potential immigration into the
-study area (Figure S1).
-
 To ensure reliable connectivity estimates, we determined the number of simulated
-dispersal trajectories required for connectivity to reach a ``steady state''
-across the entire study area. For this purpose, we distributed 1,000 rectangular
-``checkpoints'', each with an arbitrary extent of 5 km x 5 km at random
-coordinates within the study area (excluding the buffer). We then determined the
-relative frequency at which each checkpoint was traversed by simulated dispersal
-trajectories (hereafter referred to as relative traversal frequency) as we
-gradually increased the number of simulated trajectories from 1 to 50,000. To
-assess variability in the relative traversal frequency, we repeatedly subsampled
-100 times from all 50'000 trajectories and computed the mean traversal frequency
-across replicates, as well as its 95\% prediction-interval for each checkpoint.
-We considered connectivity to have reached a steady state once the width of the
+dispersal trajectories required for connectivity to reach a ``steady state''.
+For this purpose, we distributed 1,000 rectangular ``checkpoints'', each with an
+arbitrary extent of 5 km x 5 km, at random coordinates within the study area
+(excluding the buffer). We then determined the relative frequency at which each
+checkpoint was traversed by simulated dispersal trajectories (hereafter referred
+to as relative traversal frequency) as we gradually increased the number of
+simulated trajectories from 1 to 50,000. To assess variability in the relative
+traversal frequency, we repeatedly subsampled 100 times from all 50'000
+trajectories and computed the mean traversal frequency across replicates, as
+well as its 95\% prediction-interval for each checkpoint. We considered
+connectivity to have reached a steady state once the width of the
 prediction-interval dropped below a value of 0.01 for all checkpoints.
@@
 To identify dispersal hotspots within the study area, we created a heatmap
-indicating the absolute frequency at which each raster-cell was traversed by
+indicating the absolute frequency at which different areas were traversed by
 simulated dispersal trajectories (e.g. \citealp{Peer.2008, Hauenstein.2019,
@@
 raster with 1 km x 1 km resolution and tallied resulting layers into a single
-map. If the same trajectory crossed a raster-cell twice, we only counted it once
-to reduce biases due to individuals that were surrounded by unfavorable habitat
-and therefore ``moved in circles''. To achieve high performance rasterization,
-we used the R-package {\tt terra} \citep{Hijmans.2021b}.
+map. This procedure ensured that every trajectory was only counted once, even if
+it traversed the same raster-cell multiple times, thus reducing potential biases
+caused by individuals that were surrounded by unfavorable habitat and ``moved in
+circles''. To achieve high performance rasterization, we used the R-package {\tt
+terra} \citep{Hijmans.2021b}.
 
@@
 trajectories into a network, we followed \cite{BastilleRousseau.2018} and
-overlaid the study area (including the buffer) with a 5 km x 5 km raster, where
-the center of each raster-cell served as node in the final network. To identify
-edges (i.e. connections) between the nodes, we used the simulated trajectories
-and determined all transitions occurring from one node to another, as well as
-the frequency at which those transitions occurred. This resulted in an edge-list
-that we translated into a weighted network using the r-package {\tt igraph}
-\citep{Gabor.2006}. The weight of each edge was determined by the frequency of
-transitions, yet because {\tt igraph} handles edge weights (\(\omega\)) as
-costs, we inverted the traversal-frequency through each raster-cell by applying
-\(\omega = \frac{mean(Traversal Frequency)}{Traversal Frequency_i}\).
-Consequently, regularly used edges received small weights (i.e. low costs) and
-vice versa. Finally, we used the weighted network to calculate betweenness
-scores for all network nodes.
+overlaid the study area (including the buffer) with a raster containing 5 km x 5
+km raster-cells, where the center of each raster-cell served as node in the
+final network. To identify edges (i.e. connections) between the nodes, we used
+the simulated trajectories and determined all transitions occurring from one
+cell to another, as well as the frequency at which those transitions occurred.
+This resulted in an edge-list that we translated into a weighted network using
+the r-package {\tt igraph} \citep{Gabor.2006}. The final weight of each edge was
+determined by the frequency of transitions, yet because {\tt igraph} handles
+edge weights (\(\omega\)) as costs, we inverted the traversal-frequency through
+each raster-cell by applying \(\omega = \frac{mean(Traversal
+Frequency)}{Traversal Frequency_i}\). Consequently, regularly used edges
+received small weights (i.e. low costs) and vice versa. Finally, we used the
+weighted network to calculate betweenness scores for all network nodes.
 
@@
 To examine the presence and intensity of functional links (i.e. connections)
-between specific patches inside the KAZA-TFCA, we calculated inter-patch
-connectivity between NPs (e.g. \citep{Gustafson.1996, Kanagaraj.2013}). The
-decision to only consider NPs as potential ``patches'' was purely out of
-simplicity and does not imply that connections between other protected areas do
-not occur. To quantify inter-patch connectivity, we computed the relative
-frequency at which dispersers originating from one NP successfully moved into
-another NP. We considered movements as successful if an individual's dispersal
-trajectory intersected with the target NP at least once. For each trajectory we
-also recorded the number of steps required to reach the first intersection with
-the respective NP, allowing us to compute the average dispersal durations from
-one NP to another. In summary, we determined \textit{if} and \textit{how often}
-dispersers moved between certain NPs, as well as \textit{how long} individuals
-had to move to make these connections.
+between patches within the study area, we calculated inter-patch connectivity
+(e.g. \citealp{Gustafson.1996}, \citealp{Kanagaraj.2013}). For this, we computed
+the relative frequency at which dispersers originating from one patch
+successfully moved into another patch. We considered movements between patches
+as successful if an individual's dispersal trajectory intersected with the
+target patch at least once. For each trajectory we also recorded the number of
+steps required to reach the first intersection with the respective patch,
+allowing us to compute the average dispersal durations from one patch to
+another. In summary, we determined \textit{if} and \textit{how often} dispersers
+moved between certain patches, as well as \textit{how long} individuals had to
+move to make these connections.
 
@@
 during dispersal depended on habitat conditions (\Cref{MovementModel}a, Table S1
-and Table S2). Although multiple models received an AIC weight > 0 (Table S1),
-we only considered results from the most parsimonious model for simplicity. This
-decision only marginally influenced subsequent steps as all models with positive
-AIC weights retained similar covariates (Table S1). Plots that aid with model
-interpretation are provided in Figure S2. Under average conditions, dispersing
+and Table S2). Although multiple models received an AIC weight \(>\) 0 (Table
+S1), we only considered results from the most parsimonious model for simplicity.
+This decision only marginally influenced subsequent steps as all models with
+positive AIC weights retained similar covariates (Table S1). The k-fold
+cross-validation showed that the final model substantially outperformed a random
+guess and provided reliable predictions (i.e. confidence intervals of
+\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap).
+Moreover, the model correctly assigned high selection scores to realized steps
+(\Cref{MovementModel}b), indicating a good fit between predictions and
+observations. Compared to the base model (\(\bar{r}_{s, realized} = -0.55;
+95\%-CI = [-0.57, -0.52]\); \citealp{Hofmann.2021}), the inclusion of several
+interactions between movement and habitat covariates significantly improved
+model performance (\(\bar{r}_{s, realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)).
+
+Plots that aid with the interpretation of the most parsimonious movement model
+are provided in Figure S2 and suggest that, under average conditions, dispersing
 wild dogs avoided moving through water, woodlands, and areas dominated by
-humans, but preferred shrublands or grasslands (\Cref{MovementModel}a).
-Dispersers realized shorter steps (indicating slower movements) in areas covered
-by water or woodland, while realizing larger steps in areas dominated by shrubs
-or grass (\Cref{MovementModel}a). Moreover, dispersing wild dogs moved faster
-during twilight and at night (i.e. between 17:00 and 09:00 o'clock) than during
-the rest of the day (\Cref{MovementModel}a). Although dispersers showed a
-preference for directional movements (i.e. low turning angles), especially when
-moving quickly, they did less so in proximity to humans or water, resulting in
-more tortuous movements in such areas (\Cref{MovementModel}a).
-
-The k-fold cross-validation of the movement model showed that the final model
-substantially outperformed a random guess and suggested reliable predictions
-(i.e. confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-random}\) did not overlap). Moreover, the model correctly assigned high
-selection scores to realized steps (\Cref{MovementModel}b), indicating a good
-fit between predictions and observations. Compared to the base model
-(\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
-\citealp{Hofmann.2021}), the inclusion of several interactions between movement
-and habitat covariates significantly improved model performance (\(\bar{r}_{s,
-realized} = -0.65; 95\%-CI = [-0.67, -0.64]\).
+humans, but preferred moving across shrublands or grasslands
+(\Cref{MovementModel}a). Dispersers realized shorter steps (indicating slower
+movements) in areas covered by water or woodland, while realizing larger steps
+in areas dominated by shrubs or grass (\Cref{MovementModel}a). We found a
+particularly strong effect for the variable \textsf{LowActivity}, suggesting
+that dispersing wild dogs moved faster during twilight and at night (i.e.
+between 17:00 and 09:00 o'clock) in comparison to the rest of the day
+(\Cref{MovementModel}a). Although dispersers revealed a preference for
+directional movements (i.e. low turning angles), especially when moving quickly,
+they did less so in proximity to humans or water, resulting in more tortuous
+movements in such areas (\Cref{MovementModel}a).
 
@@
 kept decreasing as we increased the number of simulated dispersers, the marginal
-benefit of additional trajectories diminished quickly (Figure S3).
+benefit of simulating additional trajectories diminished quickly (Figure S3).
 
@@
 The heatmap (\Cref{Heatmap}), which resulted from the sum of all simulated
-dispersal trajectories, showed that several extensive regions within the
-KAZA-TFCA were frequently traversed by dispersing wild dogs (median traversal
-frequency inside KAZA-TFCA = 166, IQR = 274, Figure S6a), whereas areas beyond
-the KAZA-TFCA boundary were rarely visited (median traversal frequency outside
-KAZA-TFCA = 61, IQR = 133, Figure S6a). Most notably, the region in northern
-Botswana south of the Linyanti swamp appeared as highly frequented dispersal
-hotspot (median traversal frequency = 987, IQR = 558). Nevertheless, the
-presence of extensive water bodies, such as the Okavango Delta, the Makgadikgadi
-Pan, and the Linyanti swamp, restricted dispersal movements and limited realized
-connectivity within the KAZA-TFCA. Similarly, high human density, roads, and
-agricultural activities in Zambia's and Zimbabwe's part of the KAZA-TFCA limited
-dispersal movements in those countries. Outside the KAZA-TFCA, the most heavily
-used regions included the areas inside the Central Kalahari NP in Botswana, the
-area south-west of the Khaudum NP in Namibia, and the area surrounding the Liuwa
-Plains NP in Zambia. Although the heatmap facilitated the identification of
-areas frequently traversed by simulated dispersers, it seemed impractical to
+dispersal trajectories, allowed us to pinpoint areas that were intensely used by
+simulated dispersers. This enabled us to compare areas inside and outside the
+KAZA-TFCA borders with respect to the frequency at which they were visited. For
+instance, we could deduct that areas inside the KAZA-TFCA were frequently
+traversed by dispersers (median traversal frequency inside KAZA-TFCA = 166, IQR
+= 274, Figure S6a), whereas areas beyond the KAZA-TFCA boundary were
+comparatively rarely visited (median traversal frequency outside KAZA-TFCA = 61,
+IQR = 133, Figure S6a). Most notably, the region in northern Botswana south of
+the Linyanti swamp appeared to serve as highly frequented dispersal hotspot
+(median traversal frequency = 987, IQR = 558). Aside from revealing movement
+hotspots, the heatmap also provided information on areas that act as movement
+barriers. The presence of extensive water bodies, such as, for example, the
+Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, substantially
+restricted dispersal movements and limited realized connectivity inside the
+KAZA-TFCA. Although the heatmap improved our understanding of the frequency at
+which areas were traversed by simulated dispersers, it seemed impractical to
 pinpoint dispersal corridors.
@@
 The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
-that run within the KAZA-TFCA. Again, northern Botswana emerged as a wild dog
-dispersal hub that connected more remote regions in the study area. Towards
-east, the extension of this corridor ran through Chobe NP into Hwange NP. From
-there, a further extension connected to Matusadona NP in Zimbabwe. Northwest of
-the Linyanti ecosystem, a major corridor expanded into Angola, where it split
-and finally traversed over a long stretch of unprotected area into Zambia's
-Kafue NP. Several additional corridors with lower betweenness scores emerged,
-yet most of them ran within the KAZA-TFCA boundaries (median betweenness inside
-KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR = 54.311 \(\times\)
-10\textsuperscript{6}, Figure S6b). In general, there were few corridors that
-directly linked the peripheral regions of the KAZA-TFCA and passed through
-unprotected areas outside the KAZA-TFCA (mean betweenness outside KAZA-TFCA =
-2.685 \(\times\) 10\textsuperscript{6}, IQR = 9.891 \(\times\)
-10\textsuperscript{6}, Figure S6b). Compared to the heatmap, the betweenness map
-facilitated the identification of dispersal corridors between habitat patches.
+that run across the study area. In comparison to the heatmap, the betweenness
+map thus facilitated the identification of narrower and more linear routes that
+are used to disperse between distinct regions. Again, northern Botswana emerged
+as a wild dog dispersal corridor that connected more remote regions in the study
+area. Towards east, the extension of this corridor ran through Chobe NP into
+Hwange NP. From there, a further extension connected to Matusadona NP in
+Zimbabwe. Northwest of the Linyanti ecosystem, a major corridor expanded into
+Angola, where it split and finally traversed over a long stretch of unprotected
+area into Zambia's Kafue NP. Several additional corridors with lower betweenness
+scores emerged, yet most of them ran within the KAZA-TFCA boundaries (median
+betweenness inside KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR =
+54.311 \(\times\) 10\textsuperscript{6}, Figure S6b). In general, there were few
+corridors that directly linked the peripheral regions of the KAZA-TFCA and
+passed through unprotected areas outside the KAZA-TFCA (mean betweenness outside
+KAZA-TFCA = 2.685 \(\times\) 10\textsuperscript{6}, IQR = 9.891 \(\times\)
+10\textsuperscript{6}, Figure S6b).
 
@@
 average dispersal duration required to make these connections
-(\Cref{InterpatchConnectivity}). Overall, inter-patch connectivity between NPs
-in Angola, Namibia, Botswana, and Zimbabwe appeared to be high; between 54\% and
+(\Cref{InterpatchConnectivity}). For some NPs, we also detected imbalances
+between the number of ingoing and outgoing links, hinting at possible
+source-sink dynamics. From Chobe NP, for instance, 510 individuals reached
+Moremi NP, yet the opposite route was only realized by 340 individuals. However,
+relative to the number of simulated individuals, these numbers imply fractions
+of 50\% and 68\%, respectively. Overall, inter-patch connectivity between NPs in
+Angola, Namibia, Botswana, and Zimbabwe appeared to be high; between 54\% and
 87\% of individuals originating from a NP in these countries successfully moved
 into some other NP (Figure S7a). Conversely, only 19\% of the dispersers leaving
-from a NP in Zambia managed to find a route into some other NP (Figure S7b).
+from a NP in Zambia managed to find their way into some other NP (Figure S7b).
 Prior to reaching another NP, individuals from Angola, Namibia, Botswana,
 Zimbabwe, and Zambia had to move for an average of 630, 640, 940, 1045, and 890
-steps, respectively. For some NPs, we also detected imbalances between the
-number of ingoing and outgoing links, hinting at possible source-sink dynamics.
-From Chobe NP, for instance, 510 individuals reached Moremi NP, yet the opposite
-route was only realized by 340 individuals. However, relative to the number of
-simulated individuals, these numbers imply fractions of 50\% and 68\%,
-respectively. Furthermore, it appeared that the dispersal corridor between
-Angola's NPs and the Kafue NP in Zambia identified in
+steps, respectively. Furthermore, it appeared that the dispersal corridor
+between Angola's NPs and the Kafue NP in Zambia identified in
 \Cref{InterpatchConnectivity} is only rarely realized.
@@
 frequently traversed areas, critical dispersal corridors and bottlenecks, and
-the presence and intensity of functional links between suitable patches).
-Importantly, such simulations from ISSFs overcome several conceptual
-shortcomings inherent to more traditional connectivity modeling techniques, such
-as LCPA and CT.
+the presence and intensity of functional links between suitable patches). Such
+simulations from ISSFs overcome several conceptual shortcomings inherent to
+traditional connectivity modeling techniques, such as LCPA and CT, and enable a
+more detailed view on landscape connectivity.
 
@@
 able to model several additional complexities common to dispersal. For instance,
-by including an interaction between turning angle and step length we could
-accommodate that dispersers exhibit step lengths that are correlated with
-turning angles, meaning turning angles are smaller when individuals move fast.
-Although similar autocorrelations could be incorporated by sampling step lengths
-and turning angles from copula probability distributions \citep{Hodel.2021a,
-Hodel.2021b}, the ISSF framework allowed us to conventiently include
-correlations in the movement model. We only considered first order
-autocorrelation, i.e. correlation between two consecutive steps, although higher
-order autocorrelation is conceivable and may be desirable to model
-\citep{Dray.2010, McClintock.2012}. This will, however, require vast amounts of
-GPS data that are not interrupted by missing fixes; something that is rarely
-achieved in reality \citep{Graves.2006}. The power and flexibility of ISSFs to
-model additive effects between habitat and movement covariates
+our model attributed a strong effect to the binary variable
+\textsf{LowActivity}, demonstrating the necessity to consider different activity
+period of the focal species. By including an interaction between turning angle
+and step length we could furthermore accommodate that dispersers exhibit step
+lengths that are correlated with turning angles, meaning turning angles are
+smaller when individuals move fast. Although similar autocorrelations could be
+incorporated by sampling step lengths and turning angles from copula probability
+distributions \citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to
+conveniently include correlations in the movement model. We only considered
+first order autocorrelation, i.e. correlation between two consecutive steps,
+although higher order autocorrelation is conceivable and may be desirable to
+model \citep{Dray.2010, McClintock.2012}. This will, however, require vast
+amounts of GPS data that are not interrupted by missing fixes; something that is
+rarely achieved in reality \citep{Graves.2006}. The power and flexibility of
+ISSFs to model additive effects between habitat and movement covariates
 \citep{Avgar.2016, Signer.2017} furthermore allowed us to formally capture that
@@
 
-\subsection{Simulation}
-While our approach of simulating dispersal proved usedful to assess landscape
-connectivity, it was computationally very costly. Our simulation of 80,000
-dispersal trajectories moving 2'000 steps across the KAZA-TFCA required five
-days of computation on a regular desktop machine (AMD Ryzen 7 2700X processor
-with 8 x 3.6 GHz and 16 logical cores, 64 GB of RAM). The long simulation time
-was primarily caused by the massive extent of the study area considered (ca. 1.8
-Mio km\textsuperscript{2}) and the large number of trajectories simulated. Most
-connectivity studies focus on smaller study areas (e.g. \citealp{Kanagaraj.2013,
-Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020}) and will therefore achieve
-faster simulation times (given the same spatial resolution across covariates).
-We also believe that fewer simulated trajectories will often suffice, as the
-relative traversal frequency by simulated trajectories through randomly placed
-checkpoints across our study area converged already after 10,500 runs. The
-number of required simulations to achieve reliable estimates of connectivity
-will, however, vary depending on the structure of the landscape and the
-dispersal behavior of the focal species \citep{Gustafson.1996}. For species that
-disperse short distances through homogeneous environments, few simulations may
-suffice to gauge connectivity, whereas for species that disperse over long
-distances through heterogeneous habitats, a large number of simulations will be
-required to sufficiently explore the spectrum of possible routes.
-
 \subsection{Maps}
@@
 that an exceptionally large number of dispersers moved through the regions of
-the Moremi NP and the Chobe NP in northern Botswana. \citep{Hofmann.2021}
+the Moremi NP and the Chobe NP in northern Botswana. \cite{Hofmann.2021}
 previously identified the same area as potential dispersal hotspot using LCPA,
-however, it was not clear whether this was the consequence of the central
-location of the region and connections being enforced between predefined start
-and endpoints. Contrary to LCPA, a simulation-based approach as presented here
-does not require predefined endpoints because endpoints, as endpoints emerge
-naturally from the simulated dispersal trajectories. Not having to predefine
-endpoints is especially useful for dispersal studies, as known endpoints are
-usually an unrealistic assumption \citep{Elliot.2014, Abrahms.2017, Cozzi.2020}.
-Simulations also enable the detection of potential routes that do not lead into
-suitable habitats \citep{Dwernychuk.1972, VanDerMeer.2014} but into areas with a
-high susceptibility for human wildlife conflicts \citep{Cushman.2018}.
+however, following their analysis it was not clear whether this was the
+consequence of the central location of the region and connections being enforced
+between predefined start and endpoints. Contrary to LCPA, a simulation-based
+approach as presented here does not require predefined endpoints, as endpoints
+emerge naturally from the simulated dispersal trajectories. Not having to
+predefine endpoints is especially useful for dispersal studies, as known
+endpoints are usually an unrealistic assumption \citep{Elliot.2014,
+Abrahms.2017, Cozzi.2020}. Simulations also enable the detection of potential
+routes that do not lead into suitable habitats \citep{Dwernychuk.1972,
+VanDerMeer.2014} but into areas with a high susceptibility for human wildlife
+conflicts \citep{Cushman.2018}.
 
@@
 linear movement routes and thus facilitated the identification of discrete
-movement corridors. The resulting map reinforced our notion that Botswana plays
-a central role for the establishment of connections into more remote regions of
-the KAZA-TFCA. While, in this case, both the heatmap and the betweenness map
-attributed a high importance to northern Botswana, little consensus was found
-for other regions. For instance, the stretch of unprotected land between
-Luengue-Luiana NP in Angola and the Kafue NP in Zambia was characterized by a
-high betweenness-score, but at the same time received a low heatmap score. These
-contrasts highlight the complementary nature of the presented connectivity maps
-and emphasize the value of consulting multiple metrics when assessing
-connectivity.
-
-Finally, we produced a map of inter-patch connectivity. The map depicted the
-frequency at which simulated individuals moved between NPs as well as the
-average duration (in steps) required to realize them. Calculating dispersal
-durations was possible because dispersal trajectories were simulated spatially
-and temporally explicitly, something that is currently impossible with LCPA or
-CT. An explicit representation of time enables answerings questions such as:
-``\textit{How long will it take a disperser to move from A to B?}'' or
-``\textit{Is it possible for a disperser to move from A to B within X days?}''.
-Moreover, it yields opportunities to study how seasonality affects connectivity
-and to investigate whether dispersal corridors exist seasonally or all-year
-round (\textit{dynamic connectivity}; \citealp{Zeller.2020}). With LCPA or CT,
-seasonality can currently only be incorporated by repeatedly running a
-connectivity analysis using an array of seasonal permeability surfaces (e.g.
-\citealp{Benz.2016, Osipova.2019}). In contrast, simulations from ISSFs allow
-the environment to change ``as the dispersers move'', so that simulated
-trajectories can dynamically respond to seasonal fluctuations in the
-environment.
+movement corridors. While in some cases both the heatmap and the betweenness map
+attributed a high importance to the same areas (e.g. northern Botswana), little
+consensus was found for other regions. For instance, while the stretch of
+unprotected land between Luengue-Luiana NP in Angola and the Kafue NP in Zambia
+was characterized by a high betweenness-score, it only received a low score on
+the heatmap. The reason for this differences can be found in the differential
+way in which the maps view connectivity. While the heatmap assigns high
+connectivity to areas that are frequently traversed, it does not distinguish
+between areas that truly bring individuals into different parts of the study
+area and regions that lead into ecological traps. The converse is true on the
+betweenness map, which strictly highlights regions that promote movement into
+other areas of the landscape and thus might promote gene-flow. Because neither
+of the two maps provides insights into functional links between distinct habitat
+patches or how connections depend on the dispersal duration, we also produced a
+map of inter-patch connectivity. The map depicted the frequency at which
+simulated individuals moved between NPs as well as the average duration (in
+steps) required to realize them. Calculating dispersal durations was possible
+because dispersal trajectories were simulated spatially and temporally
+explicitly, something that is currently unfeasible with LCPA or CT. An explicit
+representation of time enables answerings questions such as: ``\textit{How long
+will it take a disperser to move from A to B?}'' or ``\textit{Is it possible for
+a disperser to move from A to B within X days?}''. Moreover, it yields
+opportunities to incorporate seasonality and to investigate whether dispersal
+corridors exist seasonally or all-year round (\textit{dynamic connectivity};
+\citealp{Zeller.2020}). With LCPA or CT, seasonality can currently only be
+incorporated by repeatedly running a connectivity analysis using an array of
+seasonal permeability surfaces (e.g. \citealp{Benz.2016, Osipova.2019}). In
+contrast, simulations from ISSFs allow the environment to change ``as the
+dispersers move'', so that simulated trajectories can dynamically respond to
+seasonal fluctuations in the environment.
+
+Overall, the contrasts highlight the complementary nature of the presented
+connectivity maps and emphasize the value of consulting multiple metrics when
+assessing connectivity.
 
@@
 Despite the many benefits and great flexibility offered by simulations from
-ISSFs, one must also be aware of the associated non-trivial but important
-modeling decisions. Here, we will further elaborate on four modeling decisions:
-(1) the number of simulated individuals, (2) the location of source points, (3)
-the simulated dispersal duration, and (4) behavior at map boundaries.
+ISSFs, one must also be aware of the associated limitations. For example, while
+our approach of simulating dispersal proved usedful to assess landscape
+connectivity, it was computationally costly. Simulating 80,000 dispersal
+trajectories for 2'000 steps across the KAZA-TFCA required five days of
+computation on a regular desktop machine (AMD Ryzen 7 2700X processor with 8 x
+3.6 GHz and 16 logical cores, 64 GB of RAM). The long simulation time was
+primarily caused by the massive extent of the study area considered (ca. 1.8 Mio
+km\textsuperscript{2}) and the large number of trajectories simulated. Most
+connectivity studies focus on smaller study areas (e.g. \citealp{Kanagaraj.2013,
+Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020}) and will therefore require
+fewer simulations and achieve faster simulation times (given the same spatial
+resolution). We also believe that fewer simulated trajectories will often
+suffice, as the relative traversal frequency by simulated trajectories through
+randomly placed checkpoints across our study area converged already after 10,500
+runs. The number of required simulations to achieve reliable estimates of
+connectivity will, however, vary depending on the structure of the landscape and
+the dispersal capabilities of the focal species \citep{Gustafson.1996}. For
+species that disperse short distances through homogeneous environments, few
+simulations may suffice to gauge connectivity, whereas for species that disperse
+over long distances through heterogeneous habitats, a large number of
+simulations will be necessary to sufficiently explore the spectrum of possible
+routes.
+
+Aside from the computational requirements, simulations further entail several
+non-trivial but important modeling decisions, on four of which we would like to
+further elaborate: (1) the number of simulated individuals, (2) the location of
+source points, (3) the simulated dispersal duration, and (4) behavior at map
+boundaries.
 
@@
 towards a steady state. The exact number of required individuals might, however,
-vary depending on the target metric and the anticipated connectivity map.
-Consequently, more sophisticated target metrics tailored for each of the
-presented connectivity maps need to be developed in the future.
+vary depending on the target metric and the anticipated connectivity map. More
+sophisticated target metrics tailored to each of the presented connectivity
+maps need to be developed in the future.
 
@@
 (4) Unless simulated dispersal trajectories are strongly drawn towards a point
-of attraction (e.g. \cite{Signer.2017}), some trajectories will inevitably
+of attraction (e.g. \citealp{Signer.2017}), some trajectories will inevitably
 approach a map boundary and one or more of the generated random steps will leave
@@
 (\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
-codes required for the African wild dog case study will be made available
-through an online repository at the time of publication.
+codes required to reproduce the African wild dog case study will be made
+available through an online repository at the time of publication.
 

### 2022-02-07 Chapter_1/04_Manuscript/Manuscript_Support.tex
updates

@@
 \section{Candidate Interactions}
-We started with the base model and incrementally increased model complexity by
-adding all possible two-way interactions between habitat covariates and movement
-covariates. For instance, for the covariate \textsf{Water}, we proposed the
-interactions \textsf{Water:sl}, \textsf{Water:log(sl)}, and
-\textsf{Water:cos(ta)}. Besides these interactions, we also allowed for
-correlations between turning angles and step lengths by proposing the
-interactions \textsf{sl:cos(ta)} and \textsf{log(sl):cos(ta)}. Furthermore, we
-formed the interactions \textsf{sl:LowActivity} and \textsf{log(sl):LowActivity}
-to render that step lengths are likely to be shorter during periods of
-inactivity.
+We started with the base model developed by \cite{Hofmann.2021} and
+incrementally increased model complexity by adding all possible two-way
+interactions between habitat covariates and movement covariates. For instance,
+for the covariate \textsf{Water}, we proposed the interactions
+\textsf{Water:sl}, \textsf{Water:log(sl)}, and \textsf{Water:cos(ta)}. Besides
+these interactions, we also allowed for correlations between turning angles and
+step lengths by proposing the interactions \textsf{sl:cos(ta)} and
+\textsf{log(sl):cos(ta)}. Furthermore, we formed the interactions
+\textsf{sl:LowActivity} and \textsf{log(sl):LowActivity} to render that step
+lengths are likely to be shorter during periods of inactivity.
 
@@
   was the model that we used to simulate movement of dispersers. The base model
-   upon which we based our movement model is depicted in the last row. We omitted
-   all models with an AIC weight of zero from the table.}
+  upon which we based our movement model is depicted in the last row and was
+  originally presented in \cite{Hofmann.2021}. We omitted all models with an AIC
+  weight of zero from the table.}
   \label{ModelAICs}

### 2022-02-07 Chapter_4/04_Manuscript/Manuscript.tex
updates

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Applying Step Selection Functions to Data with Data Gaps: Simulation and
Case Study on African wild dogs}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \and
  Gabriele Cozzi\textsuperscript{1,2} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Dominik M. Behr\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
Maun, Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\vspace{4cm}

\textbf{Running Title:} None

\vspace{0.5cm}

\textbf{Keywords:} step selection functions, gps data, animal movement, Lycaon
pictus, irregular data

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
This is an abstract
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

\section{Introduction}
GPS data is usually collected on fixed temporal intervals (could do a quick and
dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
even days between subsequent reolocations. Nevertheless, GPS regularly fail to
adhere to the anticipated fixrate schedule due to cloud cover or other
unfavorable conditions, thus resulting in irregular data. In other cases, the
GPS schedule is not strictly fixed but may be subject to change. For example,
\cite{Cozzi.2020} increase the fixrate of their GPS collars once per week from 4
to 8 fixes per day. We chose to coin this type of data irregular, because it is
not collected on the anticipated fixrate schedule and therefore not directly
comparable to the vast majority of collected data.

Step selection analysis is a powerful framework that enables researchers to
study habitat and movement preferences of their focal species. The method works
by comparing spatial covariates at locations where the studied animal has been
observed, to a set of locations where the animal was presumably absent. For
this, the collected GPS data is converted into steps, where a step represents
the straight line movement between two consecutive GPS fixes. One can then
compare covariates experienced along observed steps to covariates along
potential alternative steps and thus infer preferred or avoided features. By
restricting the availability domain to a set of steps, SSA is specifically
intended to account for spatio-temporal autocorrelation inherent to data
collected using GPS collars and is therefore one of the preferred methods for
analysing such data. However, it is considered good practice to only consider
steps with similar step durations in step-selection models. Depending on the
amount of available data and depending on the number of missing GPS fixes, this
might entail little sacrifice. In some situations, however, removing irregular
data implies a substantial loss of information. Additionally, recent
developments in the SSA realm have brought forward several improvements to SSA
that might allow to relax the assumption of regular step-durations. This
includes the \textit{integrated} SSA approach presented by \citep{Avgar.2016},
in which inference on habitat \textit{and} movement preferences are possible
thanks to the inclusion of movement descriptors in the respective model. More
recently, the method has been further refined by \citep{Munden.2021}, who coined
the term \textit{time-varying integrated} SSA. This method was developed for
high frequency data which has been rarified using a change-point detection
algorithm, thus also resulting in irregular step-durations.

Here, we questioned the practice of removing any irregular data in SSA and
investigated whether such data could be used to inform step selection models.
For this, we conducted a simulation study where we simulated movement
trajectories with known preferences across a virtual landscape. We then
artifically rarified the ``observed'' data by removing GPS fixes and we analysed
this data using SSA. This allowed us to investigate if and how the inclusion of
irregular data affected model estimates. We also varied the degree of
irregularity in the data and we tested for the effects of adjusting the
parametric step-length and turning distributions to different step-durations.
Besides the simulation study, we also analysed a dataset collected on dispersing
African wild dogs and examined the implications considering or discarding
irregular data.

We hypothesized that the precision of model estimates would decrease as we
increased the missigngess in the data. We also expected that the inclusion of
irregular data would not improve the precision of estimates and rather lead to
biased values. However, we anticipated that such biases, at least in estimated
movement kernel parameters, could be reduced by appropriately adjusting the
availability domain.

\section{Methods}
\subsection{Simulation Study}
\subsubsection{Spatial Covariates}
We simulated a virtual landscape comprising three spatial covariate layers, each
with a resolution of 300 x 300 pixels (\Cref{Covariates}) spanning across x- and
y-coordinates from 0 to 300. The first layer (\textsf{water}) represented
water-bodies and was simulated using a random cluster nearest‐neighbour neutral
landscape model \citep{Saura.2000}, with the cluster-proportion set to 0.5 and
the patch occupancy of water fixed to 20\%. The second layer (\textsf{elev})
resembled an elevation layer and was simulated using a Gaussian random fields
neutral landscape model \citep{Schlather.2015}, with an autocorrelation range of
10, magnitude of variation of the landscape of 1, and a magnitude of variation
in scale of 0. We simulated both of these layers using the r-package {\tt NLMR}
\citep{Sciaini.2018}. The third layer (\textsf{dist}) simply indicated the
distance (in pixels) to the center of the virtual landscape \((x = 150, y =
150)\), and can be understood as the distance to the center of an animal's
home-range. We computed spatial distances using the r-package {\tt raster}
\citep{Hijmans.2022}. We normalized the values from all layers to a range
between zero and one.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Covariates.png}
  \caption{Virtual landscape across which we simulated movement trajectories.
  All layers have a resolution of 300 x 300 pixels and were generated randomly.
  Simulated individuals were initiated within the white dashed rectangle, which
  ensured that they would not be released directly at a map border.}
  \label{Covariates}
  \end{center}
\end{figure}

\subsubsection{Movement Simulation}
To simulate movement across the above generated virtual lanscape, we employed an
``inverted'' step-selection function, which proceeded as follows. First, we
generated a random starting location. To prevent starting points in the vicinity
of map borders, we restricted sampled locations to x- and y-coordinates between
50 and 250 (white dotted rectangle in \Cref{Covariates}). Second, we generated a
set of 10 random steps originating from the sampled starting point. We then
generated random steps by sampling turing angles from a von Mises distribution
with concentration parameter \(\kappa = 0.5\) and location parameter \( \mu = 0
\), and step lengths from a gamma distribution with shape parameter \(k = 3 \)
and scale parameter \(\theta = 1\). Third, we extracted covariate values along
each random step from the underlying covariate layers and compute the average of
each covariate along every step. Fourth, we predicted the probability of being
chosen for each step by applying \Cref{EQ1}. The vector of relative habitat
preferences used to make predictions was set to \(\beta_{water} = -1\),
\(\beta_{elev} = 0.5\), and \(\beta_{dist} = -15\). Fifth, we sampled one of the
random steps based on predicted probabilities and computed the new position of
the simulated individual. We the repated these steps until a total of 100 steps
were realized and we replicated the simulation 100 times, thus resulting in 100
independent movement trajectories (\Cref{Simulations}). Note that by excluding
interactions among habitat covariates and step metrics we implicitly assume
independent habitat and movement kernels \citep{Avgar.2016}.

\begin{figure}
  \begin{center}
  \includegraphics[width = 0.5\textwidth]{99_Simulations.png}
  \caption{100 Simulated movement trajectories, each comprising of 100 steps.
  Simulated individual were initiated within the blue rectangle to mitigate edge
  effects.}
  \label{Simulations}
  \end{center}
\end{figure}

\subsubsection{Step Selection Analysis}
To assess the consequences of missing GPS data in when employing step-selection
functions, we analysed the simulated GPS data under different conditions
\Cref{Design}. Specifically, we varied the degree of data-missingness (0\% to
80\%), the forgiveness when analysing the data (1 to 5 fixes), and the
parametric distributions from which random steps were generated (static vs.
dynamic distributions). We replicated each combination of conditions 100 times
and computed averages and standard deviations from the resulting model
estimates.

In this context, missingness and forgiveness are complementary terms. While
missingness describes the fraction of data that has not been collected, we use
the term forgiveness to describe how much missingness a modeler is willing to
accept. A modeler with forgiveness of one, for instance, only considers steps
with a step-duration of one (i.e. fixes that are regularly spaced). In contrast,
a modeler with forgiveness of two would also accept steps with a step-duration
of two, thus pardoning a single missing fix.

\paragraph{Data Rarification}
We rarified the simulated GPS data by randomly removing a fixed fraction of GPS
fixes. To assess the impact of different degrees of  ``missingness'', we
incrementally increased the fraction of missing fixes from 0.0 (complete
dataset) to 0.8 (80\% missing) with 10\% increments. The removal of GPS fixes
introduced irregular temporal intervals between remaining fixes; we will refer
to these intervals as ``step-times'', where the  ``step'' refers to the straight
line movement between two consecutive GPS fixes \citep{Turchin.1998}. In the
complete dataset, a step-time of one was assumed, whereas the step-time
increased by one for every fix that was missing between subsequent fixes. For
instance, a single missing fix between two other fixes directly increased the
step time from one to two.

\paragraph{Identifying Bursts and Computing Step Metrics}
We used the generated data to compute movement bursts. A movement burst
consisted of a sequence consecutive GPS fixes with step-times that did not
exceed the accepted forgiveness. For instance, if the forgiveness was one,
already a single missing GPS fix introduced a new burst. In contrast, if the
forgiveness was set to two, step-times of two units were allowed, without
introducing a new step.

We will use the term  ``forgiveness'' to indicate how many missing fixes were
accepted before a new burst was enforced. For instance, a forgiveness of one
implied that a single missing fix did not result in a new burst. We then
converted the GPS fixes belonging to the same burst into steps, where each step
resembled the straight line movement between two GPS fixes \citep{Turchin.1998}.
For each step we computed a set of step metrics. This includd the step length,
its logarithm, the turning angle and its cosine.

\paragraph{Step Selection Function}
To conduct step selection analysis, we paired each observed step with 10 random
steps. We generated random steps by sampling turning angles from a von Mises
distribution and step lengths from a gamma distribution. Depending on the study
design, we either sampled turning angles and step lengths from fixed or fixed or
dynamic distributions (cfr. Section ...). Along observed and random steps, we
extracted underlying covariates and computed their averages along steps.

\paragraph{Step Length and Turning Angle Distributions}
To generate random steps, we employed two competing approaches. In the first
approach, we sampled step-lengths and turning-angles from the same distributions
that were used to simulate GPS data (cfr. Section ...). To adjust sampled step
lengths to the different step durations of observed steps, we multiplied sampled
step lengths with the step-duration of the respective observed step. However, in
a pre-analysis we found that step-lengths do not linearly scale with step
duration, hence we also tested an alternative approach. Specifically, we fitted
separate step-length and turning angle distributions for each step duration.

\paragraph{Regression Model}
We estimated ... using conditional logistic regression, implemented using the
r-package \textsf{survival} \citep{Therneau.2021}. We followed \cite{Avgar.2016}
and \cite{Fieberg.2021} and employed \textit{integrated} SSA. That is, aside
from the habitat covariates (\textsf{water, elev, dist}), we also included
descriptors of the step length and turning angle (\textsf{sl, log(sl), cos(ta)})
in our regression model. The model call was as follows:

$$
case\_ \sim water + elev + dist + sl + log(sl) + cos(ta)
$$

\subsection{Case Study}

because wild dogs move comparably little in the hours between xx and xx, we made
the simplyfing assumption that the 8-hour period from xx to xx is comparable to
a 4-hour period.

\section{Results}
\section{Discussion}
Results from our simulation study demonstrate that the inclusion of GPS fixes
can be used to gain information on relative habitat preferences of the focal
species. While the inclusion of steps from irregular sampling schemes resulted
in substantial biases for estimates related to step metrics, we showed that
these biases can effectively be mitigated by generating step lengths and turning
angles from distributions that are separately fit to steps from different step
durations. Regardless of biases in beta coefficients for step metrics, we found
that habitat selection estimates were unbiased, even when including irregular
steps. Even more, the inclusion of irregular steps drastially reduced model
uncertainty, thus suggesting that such data can be used to gain information that
would otherwise be lost.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing wild dogs is available on dryad
\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
proposed approach using simulated data are provided through Github
(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
codes required to reproduce the African wild dog case study will be made
available through an online repository at the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we would like
to thank Johannes Signer for assisting with the simulation algorithm. This study
was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}

### 2022-02-07 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/InterpatchConnectivityFinal.R

@@ -450,7 +450,7 @@ legend3 <- get_legend(p3)
 # Put legend with protected areas into main plot
 p4 <- p1 + annotation_custom(
       grob = legend2
-    , xmin = 18.75
+    , xmin = 18.35
     , xmax = 21
     , ymin = -13
     , ymax = -14
@@ -479,7 +479,7 @@ p6 <- ggplotify::as.ggplot(p6)
 p6
 
 # Store
-ggsave("04_Manuscript/99_InterpatchConnectivity.png", plot = p6)
+ggsave("04_Manuscript/99_InterpatchConnectivity.png", plot = p6, bg = "white")
 
 ################################################################################
 #### Network Plot for Countries
@@ -704,7 +704,7 @@ legend3 <- get_legend(p3)
 # Put legend with protected areas into main plot
 p4 <- p1 + annotation_custom(
       grob = legend2
-    , xmin = 18.75
+    , xmin = 18.35
     , xmax = 21
     , ymin = -13
     , ymax = -14
@@ -733,4 +733,4 @@ p6 <- ggplotify::as.ggplot(p6)
 p6
 
 # Store
-ggsave("04_Manuscript/99_InterpatchConnectivityCountries.png", plot = p6)
+ggsave("04_Manuscript/99_InterpatchConnectivityCountries.png", plot = p6, bg = "white")

### 2022-02-07 Chapter_2/02_R-Scripts/01_DataCleaning/12_SentinelDownload.R

@@ -68,7 +68,8 @@ todownload <- todownload %>%
   nest()
 
 # Remove the first couple of entries as there appears to be an issue
-todownload <- todownload[-(1:14), ]
+# todownload <- todownload[-(1:20), ]
+todownload <- todownload[-2, ]
 
 # Check out what we need to download
 print(todownload)

### 2022-02-10 Chapter_1/04_Manuscript/Manuscript.tex
Various updates

@@
 random walk and fails to incorporate directional persistence. While these
-assumptions can be relaxed by simulating dispersal movements explicitly across
-the landscape, a unified approach for such simulations is lacking.
-
-Here, we present a simple three-step approach to simulate dispersal movements
-and to assess connectivity using empirical GPS movement data and a set of
-habitat covariates. In step one, integrated step-selection functions are used to
-fit a mechanistic movement model describing habitat and movement preferences of
-dispersing individuals. In step two, the parameterized model is used to simulate
-dispersal trajectories across the study area. In step three, a set of three
-complementary connectivity maps is derived: a heatmap that highlights frequently
-traversed areas, a betweenness map that pinpoints dispersal corridors, and a map
-of inter-patch connectivity that indicates the presence and intensity of
-functional links between habitat patches. As a case study, we applied the
-approach to GPS data collected on dispersing individuals of the endangered
-African wild dog (\textit{Lycaon pictus}) inhabiting the Kavango-Zambezi
-Transfrontier Conservation Area (KAZA-TFCA) in southern Africa.
-
-Besides providing insights into the movement and habitat preferences of
-dispersing individuals, our movement model also rendered how movement behavior
-depended on habitat conditions. The model thereby substantially outperformed a
-model that omitted such interactions. While we simulated a total of 80,000
-dispersal trajectories, metrics of connectivity only marginally changed beyond
-10,500 simulations, suggesting that fewer simulations would have sufficed. The
-three connectivity maps derived from simulations revealed several dispersal
-hotspots and corridors across the extent of our study area. Each map emphasized
-a different aspect of landscape connectivity, thus highlighting their
-complementary nature. Overall, our case study demonstrates that a
-simulation-based approach utilizing step-selection functions offers a simple yet
-powerful alternative to traditional connectivity modeling techniques. Our
-approach not only makes fewer unrealistic assumptions about dispersal but also
-permits a more mechanistic understanding of dispersal movements and landscape
-connectivity. It is thus useful for a variety of applications in ecological,
-evolutionary, and conservation research.
+assumptions can be relaxed by explicitly simulating dispersal trajectories
+across the landscape, a unified approach for such simulations is lacking.
+
+Here, we present a simple three-step approach to simulate dispersal and to
+assess connectivity using empirical GPS movement data and a set of habitat
+covariates. In step one, we use integrated step-selection functions to fit a
+mechanistic movement model describing habitat and movement preferences of
+dispersing individuals. In step two, we apply the parameterized model to
+simulate dispersal across the study area. In step three, we derive a set of
+three complementary connectivity maps; a heatmap highlighting frequently
+traversed areas, a betweenness map pinpointing dispersal corridors, and a map of
+inter-patch connectivity indicating the presence and intensity of functional
+links between habitat patches. As a case study, we apply the three-step
+approach to GPS data collected on dispersing African wild dogs (\textit{Lycaon
+pictus}) inhabiting the Kavango-Zambezi Transfrontier Conservation Area
+(KAZA-TFCA) in southern Africa.
+
+Using step-selection functions we successfully parametrized a detailed dispersal
+model that described dispersing individuals habitat and movement preferences, as
+well as potential interactions among the two. The model substantially
+outperformed a model that omitted such interactions and enabled us to simulate
+80,000 dispersal trajectories across the study area. By explicitly simulating
+individuals, this approach not only requires fewer unrealistic assumptions about
+dispersal, but also permits the calculation of a rich suite of connectivity
+metrics that together provide a comprehensive view on landscape connectivity.
+The three derived connectivity maps revealed several dispersal hotspots and
+corridors across the extent of our study area, yet each map highlighted a
+different aspect of landscape connectivity, thus emphasizing their complementary
+nature. Overall, our case study demonstrates that a simulation-based approach
+offers a simple yet powerful alternative to traditional connectivity modeling
+techniques. The approach is therefore useful for a variety of applications in
+ecological, evolutionary, and conservation research.
 
@@
 genetic diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
-Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and colonize
-unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
-to disperse depends on a sufficient degree of landscape connectivity
+Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and to
+colonize unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the
+ability to disperse depends on a sufficient degree of landscape connectivity
 \citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
@@
 \citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the most commonly
-used models are least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and
-circuit theory (CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both
-models rest on assumptions that appear unsuitable for dispersers, thus calling
-for the development of alternative approaches. One promising alternative is to
-assess landscape connectivity via simulated dispersal trajectories generated
-from individual-based movement models (IBMMs, \citealp{Diniz.2019}). However,
-the large number of subjective modeling decisions required by IBMMs renders
-among-system comparisons difficult.
+used connectivity models are least-cost path analysis (LCPA;
+\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
+McRae.2008}). Unfortunately, both models rest on assumptions that appear
+unsuitable for dispersers, thus calling for the development of alternative
+approaches. One promising alternative is to assess landscape connectivity via
+simulated dispersal trajectories generated from individual-based movement models
+(IBMMs, \citealp{Diniz.2019}). However, IBMMs require a large number of
+subjective modeling decisions, thus rendering among-system comparisons
+difficult.
 
@@
 resistance surfaces as input, they are unable to reflect the temporal dimension
-of dispersal, such that statements about the expected duration for moving
-between habitat patches are impossible \citep{Martensen.2017, Diniz.2019}.
+of dispersal and therefore fail to allow statements about the expected duration
+for moving between habitat patches \citep{Martensen.2017, Diniz.2019}.
 
@@
 of steps, so that the temporal dimension of dispersal becomes explicit and
-allows modeling autocorrelation between consecutive steps \citep{Diniz.2019}.
+allows modeling autocorrelation between successive steps \citep{Diniz.2019}.
 Finally, simulations from IBMMs do not enforce movement or connections towards
-preconceived endpoints, thereby preventing biases arising from misplaced
-endpoints. Despite these advantages, a unifying framework to simulate dispersal
-and assess connectivity using IBMMs is lacking. Considering the large number of
-subjective decisions entailed by IBMMs, an approach that streamlines and unifies
-the application of dispersal simulations to assess connectivity will, however,
-be critical to safeguard interspecfic comparability and reproducability among
-studies.
+preconceived endpoints but allow individuals to adjust their route  ``on the
+go'', thereby preventing biases arising from misplaced endpoints. Despite these
+advantages, a unifying approach to simulate dispersal and assess connectivity
+using IBMMs is lacking. Considering the large number of subjective decisions
+entailed by IBMMs, an approach that streamlines and standardizes the application
+of dispersal simulations to assess connectivity will, however, be critical to
+safeguard comparability among studies.
 
@@
 dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
-step one, GPS movement data of dispersing individuals is combined with habitat
+step one, we combine GPS movement data of dispersing individuals with habitat
 covariates to fit a mechanistic movement model via integrated step-selection
-functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the study
-species' habitat kernel (i.e. habitat preferences), its movement kernel (i.e.
-movement preferences/capabilities), as well as potential interactions between
-the two \citep{Avgar.2016, Fieberg.2021}. In step two, the parametrized movement
-model is used to simulate individual dispersal trajectories starting from
-randomly placed source points. Comparable simulations have already been applied
-to estimate steady-state utilization distributions of resident individuals
-\citep{Potts.2013, Signer.2017} and to model landscape connectivity, yet
-disregarding interdependencies between habitat and movement kernels
-\citep{Clark.2015, Zeller.2020}. Finally, in step three, the simulated
-trajectories are converted into three complementary connectivity maps; (i) a
-heatmap revealing areas that are frequently traversed by dispersers (e.g.
-\citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map delineating
-dispersal corridors and bottlenecks (e.g. \citealp{BastilleRousseau.2018}),
-(iii) and a map of inter-patch connectivity, depicting the presence and
-intensity of functional links between habitat patches, as well as the average
-dispersal duration required for the realization of those connections (e.g.
-\citealp{Gustafson.1996, Kanagaraj.2013}).
+functions (ISSFs, \citealp{Avgar.2016}). We chose to use ISSFs because the
+framework not only allows inference on the study species' habitat kernel (i.e.
+its habitat preferences), but also its movement kernel (i.e. its movement
+preferences/capabilities) and potential interactions among the two
+\citep{Avgar.2016, Fieberg.2021}. In step two, we use the parametrized movement
+model to simulate dispersal across the study area. Comparable simulations have
+already been applied to estimate steady-state utilization distributions of
+resident individuals \citep{Potts.2013, Signer.2017} and to model landscape
+connectivity, yet disregarding interdependencies between habitat and movement
+kernels \citep{Clark.2015, Zeller.2020}. Finally, in step three, we convert the
+simulated trajectories into three complementary connectivity maps; (i) a heatmap
+revealing frequently traversed areas (e.g. \citealp{Hauenstein.2019,
+Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
+bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
+inter-patch connectivity, depicting the presence and intensity of functional
+links between habitat patches, as well as the average dispersal duration
+required to realize those connections (e.g. \citealp{Gustafson.1996,
+Kanagaraj.2013}).
 
@@
 \subsection{Case Study}
-We showcase the application of the proposed approach (\Cref{GraphicalAbstract})
-using GPS movement data collected on dispersing individuals of the endangered
-African wild dog (\textit{Lycaon pictus}). The African wild dog is a highly
-mobile species whose population persistence heavily relies on the availability
-of large, natural or semi-natural landscapes and a sufficient degree of
-connectivity among remaining subpopulations. Once common throughout sub-Saharan
-Africa, this species has disappeared from much of its historic range, largely
-due to human persecution, habitat fragmentation, and disease outbreaks
-\citep{Woodroffe.2012}. Wild dogs typically disperse in single-sex coalitions
-\citep{McNutt.1996, Behr.2020} and are capable of dispersing several hundred
-kilometers \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}. Although
-previous studies have investigated connectivity for this species using LCPA
-\citep{Hofmann.2021} or CT \citep{Brennan.2020}, a more comprehensive and
-mechanistic understanding of dispersal and connectivity is missing (but see
-\citealp{Creel.2020}). Nevertheless, with about 6,000 free-ranging wild dogs
-remaining in fragmented subpopulations \citep{Woodroffe.2012}, reliable
-information on dispersal behavior and landscape connectivity is essential for
-the conservation of this endangered carnivore. We anticipated that a
-connectivity assessment based on our three-step approach would overcome several
-of the above highlighted conceptual shortcomings of traditional connectivity
-models, while providing a more detailed view on movement behavior during
-dispersal.
+We showcase the application of the proposed approach using GPS movement data
+collected on dispersing African wild dogs (\textit{Lycaon pictus}). The African
+wild dog is a highly mobile species whose population persistence heavily relies
+on the availability of large, natural or semi-natural landscapes and a
+sufficient degree of connectivity among remaining subpopulations. Once common
+throughout sub-Saharan Africa, this species has disappeared from much of its
+historic range, largely due to human persecution, habitat fragmentation, and
+disease outbreaks \citep{Woodroffe.2012}. Wild dogs typically disperse in
+single-sex coalitions \citep{McNutt.1996, Behr.2020} and are capable of
+dispersing several hundred kilometers \citep{DaviesMostert.2012, Masenga.2016,
+Cozzi.2020}. Although previous studies have investigated connectivity for this
+species using LCPA \citep{Hofmann.2021} and CT \citep{Brennan.2020}, a more
+comprehensive and mechanistic understanding of dispersal and connectivity is
+missing (but see \citealp{Creel.2020}). Nevertheless, with about 6,000
+free-ranging wild dogs remaining in fragmented subpopulations
+\citep{Woodroffe.2012}, reliable information on dispersal behavior and landscape
+connectivity is essential for the conservation of this endangered carnivore. We
+anticipated that a connectivity assessment based upon our three-step approach
+would overcome several of the conceptual shortcomings of traditional
+connectivity models, while providing a more detailed view on movement behavior
+during dispersal its implications for landscape connectivity.
 
@@
 \subsubsection{GPS Data}
-We applied the three step approach mentioned above to GPS movement data from 16
-dispersing African wild dog coalitions (7 female and 9 male coalitions). This
-data has been collected between 2011 and 2019 from a free-ranging wild dog
-population in northern Botswana (details on the data collection can be found in
-\cite{Cozzi.2020} and \cite{Hofmann.2021}). During dispersal, GPS collars
-recorded a fix every 4 hours and regularly transmitted data over the Iridium
-satellite system. To ensure regular time intervals between GPS fixes, we removed
-any fixes that were not successfully obtained at the desired 4-hour schedule
-(allowing for a tolerance of \( \pm \) 15 minutes). To prepare the data for
-step-selection analysis, we converted the fixes (n = 4'169) into steps, where
-each step represented the straight-line movement between two consecutive GPS
-fixes \citep{Turchin.1998}.
+We applied the three step approach presented in \Cref{GraphicalAbstract} to GPS
+movement data from 16 dispersing African wild dog coalitions (7 female and 9
+male coalitions). This data has been collected between 2011 and 2019 from a
+free-ranging wild dog population in northern Botswana. During dispersal, GPS
+collars recorded a fix every 4 hours and regularly transmitted data over the
+Iridium satellite system. To ensure comparable time intervals between GPS fixes,
+we removed any fixes that were not successfully obtained at the desired 4-hour
+schedule (allowing for a tolerance of \( \pm \) 15 minutes). To prepare the data
+for step-selection analysis, we converted the fixes (n = 4'169) into steps,
+where each step represented the straight-line movement between two consecutive
+GPS fixes \citep{Turchin.1998}. We only considered steps with equal
+step-durations (i.e. 4 hours) for further analysis. We will refer to these steps
+as ``realized steps''. Additional details on the data collection and preparation
+can be found in \cite{Cozzi.2020} and \cite{Hofmann.2021}.
 
@@
 protected in the form of national parks (NPs) or other protected areas, yet a
-considerable portion of the landscape is still human-dominated (e.g. roads,
-agricultural sites, and settlements). We used NPs as patches to calculate
-inter-patch connectivity (cf. Section 2.4.3). The decision to only consider NPs
-as patches was purely out of simplicity and does not imply that dispersal
-between other protected areas does not occur.
+considerable portion of the landscape remains human-dominated (e.g. roads,
+agricultural sites, and settlements).
 
@@
     which comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia.
-    (b) The KAZA-TFCA currently represents the world's largest terrestrial
-    transfrontier conservation area, covering a total area of 520'000
-    km\textsuperscript{2}. Its main purpose is to re-establish connectivity
-    between already-existing NPs (dark green) and other protected areas (light
-    green).}
+    Data on remaining wild dog populations has been sourced from
+    \cite{Woodroffe.2012}. (b) The KAZA-TFCA currently represents the world's
+    largest terrestrial transfrontier conservation area, covering a total area
+    of 520'000 km\textsuperscript{2}. Its main purpose is to re-establish
+    connectivity between already-existing NPs (dark green) and other protected
+    areas (light green).}
     \label{StudyArea}
@@
 
-\subsubsection{Covariates}
+\subsubsection{Habitat Covariates}
 We represented the physical landscape in our study area by the habitat
 covariates \textsf{water-cover, distance-to-water, woodland-cover,
-shrub/grassland-cover, and human-influence}. To render the seasonal dynamics in
-water-cover of major water bodies for the extent of the Okavango Delta, we
-applied an algorithm that enabled us to obtain weekly updated raster-layers for
-\textsf{water-cover} and \textsf{distance-to-water} from MODIS satellite imagery
-\citep{Wolski.2017, Hofmann.2021}. This algorithm is now implemented in the {\tt
-floodmapr} package (available on GitHub;
-\url{https://github.com/DavidDHofmann/floodmapr}). To ensure a consistent
-resolution across habitat covariates, we coarsened or interpolated all layers to
-a resolution of 250 m x 250 m. A detailed description of how we prepared each
-habitat covariate is provided in \cite{Hofmann.2021}.
-
-Besides habitat covariates, we computed movement metrics that we used as
-movement covariates in the ISSF models \citep{Avgar.2016, Fieberg.2021}.
-Specifically, we computed for each step the step length (\textsf{sl}), its
+shrub/grassland-cover, and human-influence}. To render the seasonal dynamics of
+water-cover for the extent of the Okavango Delta, we applied an algorithm that
+enabled us to obtain weekly updated raster-layers for \textsf{water-cover} and
+\textsf{distance-to-water} from MODIS satellite imagery \citep{Wolski.2017,
+Hofmann.2021}. This algorithm is now implemented in the {\tt floodmapr} package
+(available on GitHub; \url{https://github.com/DavidDHofmann/floodmapr}). To
+ensure a consistent resolution across habitat covariates, we coarsened or
+interpolated all layers to a resolution of 250 m x 250 m. A detailed description
+of how we prepared each habitat covariate is provided in \cite{Hofmann.2021}.
+
+We performed all data preparations, spatial computations, and statistical
+analysis in {\tt R}, version 3.6.6 \citep{R.2020}. Some helper functions were
+written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
+\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
+
+\subsection{Step 1 - Movement Model}
+We combined the collected GPS data with habitat covariates and used ISSFs
+\citep{Avgar.2016} to parametrize a mechanistic movement model. More
+specifically, we paired each realized step with a set of 24 randomly generated
+alternative steps. A realized and its 24 random steps together formed a stratum
+that received a unique identifier. As suggested by \cite{Avgar.2016}, we
+generated random steps by sampling random turning angles from a uniform
+distribution (\(-\pi, +\pi\)) and step lengths from a gamma distribution that
+was fitted to realized steps (scale \(\theta\) = 6'308 and shape \(k\) = 0.37).
+Note that our approach of sampling turning angles from a uniform distribution
+does not imply that we assume uniform turning angles, as we will account for
+directionality later in the model \citep{Avgar.2016, Fieberg.2021}.
+
+Along each realized and random step, we extracted values from underlying habitat
+covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
+computed averages of each covariate along the steps. Besides extracting
+\textit{habitat covariates}, we also computed movement metrics that we used as
+\textit{movement covariates} in the ISSF models \citep{Avgar.2016,
+Fieberg.2021}. Specifically, we computed the step length (\textsf{sl}), its
 natural logarithm (\textsf{log(sl)}), and the cosine of the relative turning
 angle (\textsf{cos(ta)}), which is a measure of directionality
-\citep{Turchin.1998}. Because wild dog activity is low during the hot midday
-hours \citep{Cozzi.2012}, we additionally created the variable
+\citep{Turchin.1998}, for each step. Because wild dog activity is low during the
+hot midday hours \citep{Cozzi.2012}, we additionally created the variable
 \textsf{LowActivity}, indicating whether a step was realized during periods of
 low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
-(17:00 to 09:00 local time). We performed all data preparations, spatial
-computations, and statistical analysis in {\tt R}, version 3.6.6 \citep{R.2020}.
-Some helper functions were written in {\tt C++} and imported into {\tt R} using
-the {\tt Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}.
-
-\subsection{Step 1 - Movement Model}
-We used ISSFs \citep{Avgar.2016} to parametrize a mechanistic movement model
-starting from GPS movement data collected on dispersing individuals. More
-specifically, we paired each realized (i.e. observed) step with a set of 24
-randomly generated alternative steps. A realized step plus its 24 random steps
-formed a 25-step-stratum and received a unique identifier. As suggested by
-\cite{Avgar.2016}, we generated random steps by sampling random turning angles
-from a uniform distribution (\(-\pi, +\pi\)) (which is equivalent to a von Mises
-distribution with location and concentration parameters \(\mu = \kappa = 0\))
-and step lengths from a gamma distribution that was fitted to realized steps
-(scale \(\theta\) = 6'308 and shape \(k\) = 0.37).
-
-Along each realized and random step, we extracted values from underlying habitat
-covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
-computed averages of each covariate along the steps. For each step, we further
-calculated the movement metrics \textsf{sl}, \textsf{log(sl)}, and
-\textsf{cos(ta)}. To facilitate model convergence, we standardized all
-continuous covariates to a mean of zero and a standard deviation of one.
+(17:00 to 09:00 local time). To facilitate model convergence, we standardized
+all continuous covariates to a mean of zero and a standard deviation of one.
 Correlations among covariates were low (\(|r| < 0.6\); \citealp{Latham.2011}),
@@
 \citealp{Fortin.2005}), where \(w(x)\) depended on the step's associated
-covariates (\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e.
-relative selection strengths; \citealp{Avgar.2017}) towards these covariates
-(\(\beta_1, \beta_2, ..., \beta_n\)):
+covariates (\(x_1, x_2, ..., x_n\)) and on the animal's relative selection
+strengths \citealp{Avgar.2017} towards these covariates (\(\beta_1, \beta_2,
+..., \beta_n\)):
 
@@
 
-\noindent To estimate preferences (i.e. the \(\beta\)-coefficients), we used
-mixed effects conditional logistic regression analysis \citep{Muff.2020},
-implemented through the r-package {\tt glmmTMB} \citep{Brooks.2017}. The method
-introduced by \cite{Muff.2020} allows to model random slopes, yet requires to
-fix the variance of the stratum specific intercept to a large value. We
-therefore fixed the variance of the stratum specific intercept to an arbitrary
-high value of \(10 ^ 6\) and used disperser identity to model random slopes for
-all covariates.
-
-The structure of our movement model was based on a habitat selection model that
-was previously developed for dispersing wild dogs (hereafter referred to as
-\textit{base model}, \citealp{Hofmann.2021}). In the base model, no interactions
-among habitat covariates and movement covariates were considered, so we here
-expanded the model and allowed for interactions between all movement covariates
-and habitat covariates, thus reflecting that movement behavior could depend on
-habitat conditions (details in Appendix A1). To determine the most parsimonious
-movement model among model candidates, we ran stepwise forward model selection
-based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}). We
-validated the predictive power of the most parsimonious model using k-fold
+\noindent To estimate relative selection strenghts (i.e. the
+\(\beta\)-coefficients), we used mixed effects conditional logistic regression
+analysis, implemented through the r-package {\tt glmmTMB} \citep{Brooks.2017}.
+The implementation of conditional logistic regression has been proposed by
+\cite{Muff.2020} and allows to model random slopes. The method requires to fix
+the variance of the stratum specific intercept to a large value, so we fixed it
+to an arbitrary high value of \(10 ^ 6\) and used disperser identity to model
+random slopes for all covariates.
+
+Our movement model was based on a habitat selection model that was previously
+developed for dispersing wild dogs (hereafter referred to as \textit{base
+model}, \citealp{Hofmann.2021}). In the base model, no interactions among
+habitat covariates and movement covariates were considered, so we here expanded
+the model and allowed for such interactions, acknowledging that movement
+preferences during dispersal could depend on habitat conditions (details in
+Appendix A1). To determine the most parsimonious movement model among model
+candidates, we ran stepwise forward model selection based on Akaike's
+Information Criterion (AIC, \citealp{Burnham.2002}). More specifically, we
+started with the base model and iteratively increased model complexity by adding
+all possible interactions between movement and habitat covariates. We validated
+the predictive power of the most parsimonious model using k-fold
 cross-validation for case-control studies as described in \cite{Fortin.2009}.
@@
 \subsection{Step 2 - Dispersal Simulation}
-We used the most parsimonious movement model to simulate 80'000 individual
-dispersal trajectories across the study area. The simulation of a dispersal
-trajectory resembled an ``inverted'' ISSF and was set up as follows. (1) We
-defined a source point and assumed a random initial orientation of the simulated
-animal. (2) Starting from the source point, we generated 25 random steps by
-sampling turning angles from a uniform distribution (\(-\pi, +\pi\)) and step
-lengths from our fitted gamma distribution. (3) Along each random step, we
-extracted and averaged values from the different habitat covariate layers and
-computed the movement metrics \textsf{sl}, \textsf{log(sl)}, and
-\textsf{cos(ta)}. To ensure compatible scales with the fitted movement model, we
-standardized covariate values using means and standard deviations from the
-empirical data. (4) We applied the parametrized movement model to predict the
-selection score \(w(x)\) for each step using \ref{EQ1} and we converted
-predicted scores into probabilities using \ref{EQ2}. (5) We randomly sampled one
-of the generated random steps based on assigned probabilities and determined the
-animal's new position. We repeated steps (2) to (5) until 2,000 steps were
-realized.
+We used the most parsimonious movement model to simulate individual dispersal
+trajectories across the study area. The simulation of a dispersal trajectory
+resembled an ``inverted'' ISSF and was set up as follows. (1) We defined a
+source point and assumed a random initial orientation of the simulated animal.
+(2) Starting from the source point, we generated 25 random steps by sampling
+turning angles from a uniform distribution (\(-\pi, +\pi\)) and step lengths
+from our fitted gamma distribution. (3) Along each random step, we extracted and
+averaged values from the habitat covariate layers and we computed the movement
+metrics \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)}. To ensure
+compatible scales with the fitted movement model, we standardized covariate
+values using means and standard deviations from the empirical data. (4) We
+applied the parametrized movement model to predict the selection score \(w(x)\)
+for each step using \ref{EQ1} and we converted predicted scores into
+probabilities using \ref{EQ2}. (5) We randomly sampled one of the generated
+random steps based on assigned probabilities and determined the animal's new
+position. We repeated steps (2) to (5) until 2,000 steps were realized and we
+repeated the simulation until a total of 80'000 dispersal trajectories was
+reached.
 
 As source points for the simulations, we distributed 50,000 points at random
-locations inside protected areas that were larger than the average home range of
-resident wild dog packs (i.e. \(>\) 700 km\textsuperscript{2};
+locations inside protected areas that were large enough to host an average size
+wild dog home range (i.e. \(>\) 700 km\textsuperscript{2};
 \citealp{Pomilia.2015}). We placed another 30,000 points randomly inside the
@@
 resampled steps until they fully lied within the buffer zone, essentially
-forcing individuals to remain within the expanded study area.
+forcing simulated individuals to remain within the expanded study area.
 
 To ensure reliable connectivity estimates, we determined the number of simulated
-dispersal trajectories required for connectivity to reach a ``steady state''.
-For this purpose, we distributed 1,000 rectangular ``checkpoints'', each with an
-arbitrary extent of 5 km x 5 km, at random coordinates within the study area
-(excluding the buffer). We then determined the relative frequency at which each
-checkpoint was traversed by simulated dispersal trajectories (hereafter referred
-to as relative traversal frequency) as we gradually increased the number of
-simulated trajectories from 1 to 50,000. To assess variability in the relative
-traversal frequency, we repeatedly subsampled 100 times from all 50'000
-trajectories and computed the mean traversal frequency across replicates, as
-well as its 95\% prediction-interval for each checkpoint. We considered
-connectivity to have reached a steady state once the width of the
-prediction-interval dropped below a value of 0.01 for all checkpoints.
+dispersal trajectories required to reach a ``steady state''. For this purpose,
+we distributed 1,000 rectangular ``checkpoints'', each with an arbitrary extent
+of 5 km x 5 km, at random coordinates within the study area (excluding the
+buffer). We then determined the relative frequency at which each checkpoint was
+traversed by simulated dispersal trajectories (hereafter referred to as relative
+traversal frequency) as we gradually increased the number of simulated
+trajectories from 1 to 50,000. To assess variability in the relative traversal
+frequency, we repeatedly subsampled 100 times from all 50'000 trajectories and
+computed the mean traversal frequency across replicates, as well as its 95\%
+prediction-interval for each checkpoint. We considered connectivity to have
+reached a steady state once the width of the prediction-interval dropped below a
+value of 0.01 for all checkpoints.
 
@@
 Frequency)}{Traversal Frequency_i}\). Consequently, regularly used edges
-received small weights (i.e. low costs) and vice versa. Finally, we used the
-weighted network to calculate betweenness scores for all network nodes.
+received small weights (i.e. low costs) and vice versa. We used the weighted
+network to calculate betweenness scores for all network nodes.
 
@@
 successfully moved into another patch. We considered movements between patches
-as successful if an individual's dispersal trajectory intersected with the
-target patch at least once. For each trajectory we also recorded the number of
-steps required to reach the first intersection with the respective patch,
-allowing us to compute the average dispersal durations from one patch to
-another. In summary, we determined \textit{if} and \textit{how often} dispersers
-moved between certain patches, as well as \textit{how long} individuals had to
-move to make these connections.
+as successful if an individual's dispersal trajectory originating from the
+source patch intersected with the target patch at least once. For each
+trajectory we also recorded the number of steps required to reach the first
+intersection with the respective patch, allowing us to compute the average
+dispersal durations from one patch to another. In summary, we determined
+\textit{if} and \textit{how often} dispersers moved between certain patches, as
+well as \textit{how long} individuals had to move to make these connections. In
+our case study, we used NPs as patches to determine inter-patch connectivity,
+hence we'll use the terms interchangeably from here on. The decision to focus on
+NPs was purely out of simplicity and should not imply that dispersal between
+other areas is impossible.
 
@@
 The most parsimonious movement model consisted of movement covariates, habitat
-covariates and several of their interactions, suggesting that movement behavior
-during dispersal depended on habitat conditions (\Cref{MovementModel}a, Table S1
-and Table S2). Although multiple models received an AIC weight \(>\) 0 (Table
-S1), we only considered results from the most parsimonious model for simplicity.
-This decision only marginally influenced subsequent steps as all models with
-positive AIC weights retained similar covariates (Table S1). The k-fold
-cross-validation showed that the final model substantially outperformed a random
-guess and provided reliable predictions (i.e. confidence intervals of
-\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap).
-Moreover, the model correctly assigned high selection scores to realized steps
-(\Cref{MovementModel}b), indicating a good fit between predictions and
-observations. Compared to the base model (\(\bar{r}_{s, realized} = -0.55;
-95\%-CI = [-0.57, -0.52]\); \citealp{Hofmann.2021}), the inclusion of several
-interactions between movement and habitat covariates significantly improved
-model performance (\(\bar{r}_{s, realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)).
+covariates, as well as several of their interactions, thus suggesting that
+movement behavior during dispersal depended on habitat conditions
+(\Cref{MovementModel}a, Table S1 and Table S2). Although multiple models
+received an AIC weight \(>\) 0 (Table S1), we only considered results from the
+most parsimonious model for simplicity. This decision only marginally influenced
+subsequent steps as all models with positive AIC weights retained similar
+covariates (Table S1). The k-fold cross-validation showed that the final model
+substantially outperformed a random guess and provided reliable predictions
+(i.e. confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
+random}\) did not overlap). Moreover, the model correctly assigned high
+selection scores to realized steps (\Cref{MovementModel}b), indicating a good
+fit between predictions and observations. As can be taken from the Spearman rank
+correlation coefficient, the inclusion of several interactions between movement
+and habitat covariates significantly improved model performance (\(\bar{r}_{s,
+realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)), compared to the base model
+(\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
+\citealp{Hofmann.2021})
 
@@
 in areas dominated by shrubs or grass (\Cref{MovementModel}a). We found a
-particularly strong effect for the variable \textsf{LowActivity}, suggesting
-that dispersing wild dogs moved faster during twilight and at night (i.e.
-between 17:00 and 09:00 o'clock) in comparison to the rest of the day
-(\Cref{MovementModel}a). Although dispersers revealed a preference for
-directional movements (i.e. low turning angles), especially when moving quickly,
-they did less so in proximity to humans or water, resulting in more tortuous
-movements in such areas (\Cref{MovementModel}a).
+particularly large effect for the variable \textsf{LowActivity}, suggesting that
+dispersing wild dogs moved substantially faster during twilight and at night
+(i.e. between 17:00 and 09:00 o'clock; \Cref{MovementModel}a). Although
+dispersers revealed a preference for directional movements (i.e. low turning
+angles), especially when moving quickly, they did less so in proximity to humans
+or water, resulting in more tortuous movements in such areas
+(\Cref{MovementModel}a).
 
@@
 trajectories with starting point within the main study area, only 4.5\% reached
-a map boundary, suggesting minimal biases due to boundary effects. Moreover, our
-examination of the relative traversal frequency across all checkpoints showed
-that connectivity reached a steady state after 10,500 simulated dispersal
-trajectories (Figure S3). Although variability in relative traversal frequency
-kept decreasing as we increased the number of simulated dispersers, the marginal
-benefit of simulating additional trajectories diminished quickly (Figure S3).
+a map boundary, suggesting that we successfully mitigated biases from boundary
+effects. Moreover, our examination of the relative traversal frequency across
+all checkpoints showed that the relative traversal frequency reached a steady
+state after 10,500 simulated dispersal trajectories (Figure S3). Although
+variability in the relative traversal frequency kept decreasing as we increased
+the number of simulated dispersers, the marginal benefit of simulating
+additional trajectories diminished quickly (Figure S3).
 
 \subsection{Heatmap}
-The heatmap (\Cref{Heatmap}), which resulted from the sum of all simulated
-dispersal trajectories, allowed us to pinpoint areas that were intensely used by
-simulated dispersers. This enabled us to compare areas inside and outside the
-KAZA-TFCA borders with respect to the frequency at which they were visited. For
+The heatmap (\Cref{Heatmap}), which resulted from the summation of all simulated
+dispersal trajectories, allowed us to pinpoint areas that were frequently
+visited and enabled us to compare areas inside and outside the KAZA-TFCA borders
+with respect to the intensity at which they were used for dispersal. For
 instance, we could deduct that areas inside the KAZA-TFCA were frequently
@@
 (median traversal frequency = 987, IQR = 558). Aside from revealing movement
-hotspots, the heatmap also provided information on areas that act as movement
-barriers. The presence of extensive water bodies, such as, for example, the
-Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, substantially
-restricted dispersal movements and limited realized connectivity inside the
-KAZA-TFCA. Although the heatmap improved our understanding of the frequency at
-which areas were traversed by simulated dispersers, it seemed impractical to
-pinpoint dispersal corridors.
+hotspots, the heatmap also provided information on areas that appeared to hinder
+movement. For example, extensive water bodies, such as the Okavango Delta, the
+Makgadikgadi Pan, and the Linyanti swamp, substantially restricted dispersal
+movements and limited realized connectivity inside the KAZA-TFCA. Similarly, the
+landscapes of Zambia and Zimbabwe were only rarely used for dispersal, even
+within the KAZA-TFCA boundaries (Figure S7a). Despite the fact that the heatmap
+improved our understanding of the frequency at which areas were traversed by
+simulated dispersers, it seemed impractical to pinpoint dispersal corridors.
 
@@
 \subsection{Betweenness}
-The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
-that run across the study area. In comparison to the heatmap, the betweenness
-map thus facilitated the identification of narrower and more linear routes that
-are used to disperse between distinct regions. Again, northern Botswana emerged
-as a wild dog dispersal corridor that connected more remote regions in the study
-area. Towards east, the extension of this corridor ran through Chobe NP into
-Hwange NP. From there, a further extension connected to Matusadona NP in
-Zimbabwe. Northwest of the Linyanti ecosystem, a major corridor expanded into
-Angola, where it split and finally traversed over a long stretch of unprotected
-area into Zambia's Kafue NP. Several additional corridors with lower betweenness
-scores emerged, yet most of them ran within the KAZA-TFCA boundaries (median
-betweenness inside KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR =
-54.311 \(\times\) 10\textsuperscript{6}, Figure S6b). In general, there were few
-corridors that directly linked the peripheral regions of the KAZA-TFCA and
-passed through unprotected areas outside the KAZA-TFCA (mean betweenness outside
-KAZA-TFCA = 2.685 \(\times\) 10\textsuperscript{6}, IQR = 9.891 \(\times\)
-10\textsuperscript{6}, Figure S6b).
+The betweenness map (\Cref{Betweenness}) revealed several distinct dispersal
+corridors that run across the study area. In comparison to the heatmap, the
+betweenness map was less biased towards areas with many dispersers and
+pronounced narrower, more linear routes that were used by simulated individuals
+to move between regions. Again, northern Botswana emerged as a wild dog
+dispersal corridor that connected more remote regions in the study area. Towards
+east, the extension of this corridor ran through Chobe NP into Hwange NP. From
+there, a further extension connected to Matusadona NP in Zimbabwe. Northwest of
+the Linyanti ecosystem, a major corridor expanded into Angola, where it split
+and finally traversed over a long stretch of unprotected area into Zambia's
+Kafue NP. Several additional corridors with lower betweenness scores emerged,
+yet most of them ran within the KAZA-TFCA boundaries (median betweenness inside
+KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR = 54.311 \(\times\)
+10\textsuperscript{6}, Figure S6b). Consequently, only few corridors directly
+linked the peripheral regions of the KAZA-TFCA and passed through unprotected
+areas outside its borders (mean betweenness outside KAZA-TFCA = 2.685 \(\times\)
+10\textsuperscript{6}, IQR = 9.891 \(\times\) 10\textsuperscript{6}, Figure
+S6b).
 
@@
 The inter-patch connectivity map showed that the relative frequency at which
-simulated dispersal trajectories moved from one NP to another varied, as did the
-average dispersal duration required to make these connections
-(\Cref{InterpatchConnectivity}). For some NPs, we also detected imbalances
-between the number of ingoing and outgoing links, hinting at possible
-source-sink dynamics. From Chobe NP, for instance, 510 individuals reached
-Moremi NP, yet the opposite route was only realized by 340 individuals. However,
-relative to the number of simulated individuals, these numbers imply fractions
-of 50\% and 68\%, respectively. Overall, inter-patch connectivity between NPs in
-Angola, Namibia, Botswana, and Zimbabwe appeared to be high; between 54\% and
-87\% of individuals originating from a NP in these countries successfully moved
-into some other NP (Figure S7a). Conversely, only 19\% of the dispersers leaving
-from a NP in Zambia managed to find their way into some other NP (Figure S7b).
-Prior to reaching another NP, individuals from Angola, Namibia, Botswana,
+simulated dispersal trajectories moved from one patch to another varied, as did
+the average dispersal duration between patches (\Cref{InterpatchConnectivity}).
+The map thereby completed the picture on connectivity and provided valuable
+insights into the frequency and duration of connections between patches. For
+some patches, we also detected imbalances between the number of incoming and
+outgoing links, hinting at possible source-sink dynamics. From Chobe NP, for
+instance, 510 individuals reached the Moremi NP, yet the opposite route was only
+realized by 340 individuals. Relative to the number of simulated individuals,
+however, these numbers correspond to fractions of 50\% and 68\%, respectively.
+Overall, inter-patch connectivity between patches in Angola, Namibia, Botswana,
+and Zimbabwe appeared to be high; between 54\% and 87\% of individuals
+originating from a patch in these countries successfully moved into at least on
+other patch (Figure S8a). Conversely, only 19\% of the dispersers leaving from a
+patch in Zambia managed to find their way into some other patch (Figure S8b).
+Prior to reaching another patch, individuals from Angola, Namibia, Botswana,
 Zimbabwe, and Zambia had to move for an average of 630, 640, 940, 1045, and 890
-steps, respectively. Furthermore, it appeared that the dispersal corridor
-between Angola's NPs and the Kafue NP in Zambia identified in
-\Cref{InterpatchConnectivity} is only rarely realized.
+steps, respectively. Furthermore, it appeared that the corridor previously
+identified on \Cref{InterpatchConnectivity} between Angola's NPs and the Kafue
+NP in Zambia is only rarely realized.
 
@@
 \subsection{Short Summary}
-Here, we proposed a three-step approach to assess landscape connectivity via
-simulated dispersal trajectories, and demonstrated its application using
+Here, we presented a simple three-step approach to assess landscape connectivity
+via simulated dispersal trajectories and we demonstrated its application using
 empirical data from a free-ranging population of African wild dogs. In step one,
 we used ISSFs to parametrize a fully mechanistic movement model describing how
-individuals move through the landscape. In step two, we employed the movement
-model to simulate dispersal trajectories across the landscape. In step three, we
-translated the simulated trajectories into three complementary connectivity
-maps, each emphasizing a different aspect of landscape connectivity (e.g.
-frequently traversed areas, critical dispersal corridors and bottlenecks, and
-the presence and intensity of functional links between suitable patches). Such
-simulations from ISSFs overcome several conceptual shortcomings inherent to
-traditional connectivity modeling techniques, such as LCPA and CT, and enable a
-more detailed view on landscape connectivity.
+individuals move through the landscape. Aside from rendering habitat
+preferences, the model also encapsulated movement preferences and potential
+interactions between movement and habitat preferences. In step two, we employed
+the movement model and simulated dispersal trajectories across the landscape. In
+comparison to more traditional connectivity modeling techniques, such
+simulations require fewer unrealistic assumptions about dispersal and enable the
+derivation of multiple connectivity metrics. Hence, in step three, we translated
+the simulated trajectories into three complementary connectivity maps, each
+emphasizing a different aspect of landscape connectivity (e.g. frequently
+traversed areas, critical dispersal corridors and bottlenecks, and the presence
+and intensity of functional links between suitable patches).
 
@@
 \subsection{Movement Model}
-Our results on habitat preferences showed that dispersers avoid areas dominated
-by humans and covered by water, but select for regions with open grassland in
-the vicinity to water bodies. This largely complied with previous studies that
-investigated habitat selection by dispersing wild dogs
+Results on the habitat kernel from our model showed that dispersers avoided
+areas dominated by humans and covered by water, but selected for regions with
+open grassland in the vicinity to water bodies. This largely complied with
+previous studies that investigated habitat selection by dispersing wild dogs
 \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020,
-Hofmann.2021}. However, by also accounting for movement preferences, we were
-able to model several additional complexities common to dispersal. For instance,
-our model attributed a strong effect to the binary variable
-\textsf{LowActivity}, demonstrating the necessity to consider different activity
-period of the focal species. By including an interaction between turning angle
-and step length we could furthermore accommodate that dispersers exhibit step
-lengths that are correlated with turning angles, meaning turning angles are
-smaller when individuals move fast. Although similar autocorrelations could be
-incorporated by sampling step lengths and turning angles from copula probability
-distributions \citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to
-conveniently include correlations in the movement model. We only considered
-first order autocorrelation, i.e. correlation between two consecutive steps,
-although higher order autocorrelation is conceivable and may be desirable to
-model \citep{Dray.2010, McClintock.2012}. This will, however, require vast
-amounts of GPS data that are not interrupted by missing fixes; something that is
-rarely achieved in reality \citep{Graves.2006}. The power and flexibility of
-ISSFs to model additive effects between habitat and movement covariates
+Hofmann.2021}. However, instead of merely generating insights on dispersers'
+habitat preferences, the ISSF framework also permitted us to model several
+additional complexities common to dispersal. For instance, by including the
+interactions \textsf{cos(ta):sl} and \textsf{cos(ta):log(sl)}, we could
+accommodate that dispersers exhibit turning angles that are correlated with step
+lengths, meaning that turning angles tend to be smaller when individuals move
+fast. Although similar autocorrelations could be incorporated by sampling step
+lengths and turning angles from copula probability distributions
+\citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to conveniently
+model such peculiarities directly in the movement model. While we only
+considered first order autocorrelation, i.e. correlation between two consecutive
+steps, higher order autocorrelation is conceivable and may be desirable to model
+\citep{Dray.2010, McClintock.2012}. However, this will require vast amounts of
+GPS data that are not interrupted by missing fixes; something that is rarely
+achieved in reality \citep{Graves.2006}. The power and flexibility of ISSFs to
+model additive effects between habitat and movement covariates
 \citep{Avgar.2016, Signer.2017} furthermore allowed us to formally capture that
-dispersing wild dogs move slower and more tortuous in areas covered by water,
-something that surely was to be expected. Overall, the inclusion of interactions
-between habitat and movement covariates in our movement model lead to a
-significant improvement in predictive performance compared to an earlier model
-that omitted such interactions \citep{Hofmann.2021}.
+dispersing wild dogs move slower and more tortuous in areas covered by water.
+Such effects may be of limited interest and novelty from a biological
+perspective, yet they are important to be considered when simulating dispersal,
+in particular if one is interested in estimating dispersal durations between
+habitat patches. Overall, the inclusion of interactions between habitat and
+movement covariates in our movement model lead to a significant improvement in
+predictive performance compared to an earlier model that omitted such
+interactions \citep{Hofmann.2021}.
+
+% The model also enabled us to include the interaction \textsf{sl:LowActivity} to
+% render that wild dogs primarily move at dask and dawn. The resulting beta
+% estimate was large, demonstrating the necessity to consider different phases of
+% activities of the focal species.
 
@@
 approach as presented here does not require predefined endpoints, as endpoints
-emerge naturally from the simulated dispersal trajectories. Not having to
-predefine endpoints is especially useful for dispersal studies, as known
-endpoints are usually an unrealistic assumption \citep{Elliot.2014,
-Abrahms.2017, Cozzi.2020}. Simulations also enable the detection of potential
-routes that do not lead into suitable habitats \citep{Dwernychuk.1972,
-VanDerMeer.2014} but into areas with a high susceptibility for human wildlife
-conflicts \citep{Cushman.2018}.
+emerge naturally from the simulated dispersal trajectories. This is especially
+useful for dispersal studies, where known endpoints are usually an unrealistic
+assumption \citep{Elliot.2014, Abrahms.2017, Cozzi.2020}. The fact that the same
+region was emphasized using vastly different methods to model connectivity thus
+reinforces our notion that the area is of exceptional importance to dispersing
+wild dogs. Because simulated individuals are not forced to move towards certain
+endpoints, a simulation-based approach not only lends itself to study landscape
+connectivity, but also to uncover potential dispersal traps
+\citep{Dwernychuk.1972, VanDerMeer.2014} or areas with a high susceptibility for
+human wildlife conflicts \citep{Cushman.2018}.
 
 In contrast to the heatmap, the betweenness map emphasized relatively narrow and
-linear movement routes and thus facilitated the identification of discrete
+linear movement routes. It thus facilitated the identification of discrete
 movement corridors. While in some cases both the heatmap and the betweenness map
 attributed a high importance to the same areas (e.g. northern Botswana), little
-consensus was found for other regions. For instance, while the stretch of
-unprotected land between Luengue-Luiana NP in Angola and the Kafue NP in Zambia
-was characterized by a high betweenness-score, it only received a low score on
-the heatmap. The reason for this differences can be found in the differential
-way in which the maps view connectivity. While the heatmap assigns high
-connectivity to areas that are frequently traversed, it does not distinguish
-between areas that truly bring individuals into different parts of the study
-area and regions that lead into ecological traps. The converse is true on the
-betweenness map, which strictly highlights regions that promote movement into
-other areas of the landscape and thus might promote gene-flow. Because neither
-of the two maps provides insights into functional links between distinct habitat
-patches or how connections depend on the dispersal duration, we also produced a
-map of inter-patch connectivity. The map depicted the frequency at which
-simulated individuals moved between NPs as well as the average duration (in
-steps) required to realize them. Calculating dispersal durations was possible
-because dispersal trajectories were simulated spatially and temporally
-explicitly, something that is currently unfeasible with LCPA or CT. An explicit
+consensus was found for other regions. For instance, the stretch of unprotected
+land between Luengue-Luiana NP in Angola and the Kafue NP in Zambia was
+characterized by a high betweenness-scores, yet it only received low scores on
+the heatmap. This is due to the differential way in which the maps view
+connectivity. While the heatmap attributes a high connectivity to areas that are
+frequently traversed, it does not distinguish between areas that truly bring
+individuals into other regions of the study area and regions that lead into
+ecological traps. The converse is true on the betweenness map, as it strictly
+highlights regions that promote movement into other areas of the landscape and
+thus promote gene-flow. However, neither of the two maps provides insights into
+functional links between distinct habitat patches or how connections depend on
+the dispersal duration. For this reason, we also produced a map of inter-patch
+connectivity. This map depicted the frequency at which simulated individuals
+moved between patches as well as the average dispersal duration (in steps)
+required to realize them. Calculating dispersal durations was only possible
+because trajectories were simulated spatially and temporally explicitly,
+something that is currently unfeasible with LCPA or CT. An explicit
 representation of time enables answerings questions such as: ``\textit{How long
@@
 \citealp{Zeller.2020}). With LCPA or CT, seasonality can currently only be
-incorporated by repeatedly running a connectivity analysis using an array of
-seasonal permeability surfaces (e.g. \citealp{Benz.2016, Osipova.2019}). In
-contrast, simulations from ISSFs allow the environment to change ``as the
-dispersers move'', so that simulated trajectories can dynamically respond to
-seasonal fluctuations in the environment.
-
-Overall, the contrasts highlight the complementary nature of the presented
-connectivity maps and emphasize the value of consulting multiple metrics when
-assessing connectivity.
+incorporated through the preparation of multiple permeability surfaces on which
+the same connectivity model is repeatedly applied (e.g. \citealp{Benz.2016,
+Osipova.2019}). With simulations from ISSFs, in contrast, the environment could
+change ``as the dispersers move'', so that simulated trajectories would
+dynamically respond to seasonal fluctuations in the environment.
 
@@
 trajectories for 2'000 steps across the KAZA-TFCA required five days of
-computation on a regular desktop machine (AMD Ryzen 7 2700X processor with 8 x
-3.6 GHz and 16 logical cores, 64 GB of RAM). The long simulation time was
-primarily caused by the massive extent of the study area considered (ca. 1.8 Mio
-km\textsuperscript{2}) and the large number of trajectories simulated. Most
+computation on a regular desktop machine (AMD Ryzen 7 2700X octa-core processor
+with 3.6 GHz, 64 GB of RAM). The long simulation time was primarily caused by
+the massive extent of the study area considered (ca. 1.8 Mio
+km\textsuperscript{2}) and the large number of simulated trajectories. Most
 connectivity studies focus on smaller study areas (e.g. \citealp{Kanagaraj.2013,
@@
 randomly placed checkpoints across our study area converged already after 10,500
-runs. The number of required simulations to achieve reliable estimates of
-connectivity will, however, vary depending on the structure of the landscape and
-the dispersal capabilities of the focal species \citep{Gustafson.1996}. For
+runs. The exact number of required simulations to achieve reliable estimates of
+connectivity will, of course, vary depending on the structure of the landscape
+and the dispersal capabilities of the focal species \citep{Gustafson.1996}. For
 species that disperse short distances through homogeneous environments, few
@@
 over long distances through heterogeneous habitats, a large number of
-simulations will be necessary to sufficiently explore the spectrum of possible
+simulations will be required to sufficiently explore the spectrum of possible
 routes.
@@
 Aside from the computational requirements, simulations further entail several
-non-trivial but important modeling decisions, on four of which we would like to
-further elaborate: (1) the number of simulated individuals, (2) the location of
-source points, (3) the simulated dispersal duration, and (4) behavior at map
-boundaries.
+non-trivial but important modeling decisions. On four such decisions we would
+like to further elaborate: (1) the number of simulated individuals, (2) the
+location of source points, (3) the simulated dispersal duration, and (4)
+the behavior at map boundaries.
 
@@
 additional trajectory provides information about landscape connectivity.
-However, each additional simulation entails computational costs, so a trade-off
-needs to be managed. Here, we followed \cite{Signer.2017} who suggested to
-simulate additional individuals only until the metrics of interest converge
-towards a steady state. The exact number of required individuals might, however,
-vary depending on the target metric and the anticipated connectivity map. More
-sophisticated target metrics tailored to each of the presented connectivity
-maps need to be developed in the future.
+However, each additional simulation imposes computational costs, so a trade-off
+needs to be managed. \cite{Signer.2017} proposed to handle the trade-off by
+simulating additional individuals only until the metrics of interest converge
+towards a steady state. Here, we used the relative traversal frequency as target
+metric and found that it converged already after 10'500 simulated individuals.
+The exact number of required individuals might, however, vary depending on the
+employed target metric and the anticipated connectivity map. More sophisticated
+target metrics than the relative traversal frequency, preferably tailored to
+different connectivity maps, need to be developed in the future.
 
 (2) To initiate dispersers, a modeler needs to provide a set of source points at
-which the virtual disperser will be released. We placed source points within
+which the virtual dispersers are released. We placed source points within
 protected areas large enough to sustain viable wild dog populations, implicitly
-assuming wild dogs primarily survive in large, formally protected areas
-\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. We
-lacked precise knowledge about the presence and abundance of wild dogs in the
-different protected areas, so we distributed source randomly points within them.
-In cases where such data is available, source points could be distributed
-accordingly, reflecting that the number of simulated trajectories does not
-necessarily scales with the size of the source area. Alternatively, source
-points could be distributed homogeneously and only later be weighted when
-computing the heatmap, betweenness map, or inter-patch connectivity map. In any
-case, the challenge of selecting meaningful source points is not unique to
-individual-based simulations but also applies to LCPA or CT.
+assuming that wild dogs primarily survive in large, formally protected areas
+\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}.
+Moreover, we lacked precise knowledge about the distribution and abundance of
+wild dogs across protected areas, so we placed source points randomly within
+them. In cases where more detailed data distribution and abundance are
+available, source points could be distributed accordingly. Alternatively, source
+points could be distributed homogeneously but later be weighted when computing
+connectivity metrics. In any case, the challenge of selecting meaningful source
+points is not unique to individual-based simulations but also applies to LCPA
+and CT.
 
 (3) The use of ISSFs to simulate dispersers requires deciding on the number of
-simulated steps (i.e. dispersal durations). If sufficient dispersal data of the
-focal species has been collected, dispersal durations can be sampled from
-observed dispersal events. Due to the low number of observed dispersal events,
-we opted against this solution and instead simulated all individuals for 2,000
-steps, which is at the upper end of observed dispersal durations
-\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. This
-approach also had the advantage that it allowed us to subsample simulated
-trajectories to shorter durations after their simulation and thereby to
-investigate the sensitivity of our results with respect to exact dispersal
-durations (Figures S4 and S5).
+simulated steps (i.e. the simulated dispersal durations). If sufficient
+dispersal data of the focal species has been collected, dispersal durations
+could be sampled from observed dispersal events or from parametric distributions
+fit to observed data. Due to the low number of observed dispersal events, we
+opted against this solution and instead simulated all individuals for 2,000
+steps, which was at the upper end of observed dispersal durations in African
+wild dogs \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
+This approach had the advantage that it allowed us to systematically shorten the
+simulated trajectories after their simulation and thereby to investigate the
+sensitivity of our results with respect to exact dispersal durations (Figures S4
+and S5).
 
 (4) Unless simulated dispersal trajectories are strongly drawn towards a point
-of attraction (e.g. \citealp{Signer.2017}), some trajectories will inevitably
-approach a map boundary and one or more of the generated random steps will leave
-the study area. One solution to deal with such cases would be to simply
-terminate the simulation of the affected trajectory, implicitly assuming that
-the respective animal left the study area. However, this approach might produce
-ambiguous results in cases where individuals are released near map borders,
-since already a single random step leaving the study area will break the
-simulation, thus resulting in biased connectivity estimates along map borders.
-Rather than halting the simulation, we created a buffer zone \citep{Koen.2010}
-and resampled random steps until they fully lied within the study area. This
-proved to be an effective solution to overcome problems with boundary effects.
+of attraction inside the study area(e.g. \citealp{Signer.2017}), some
+trajectories will inevitably approach one of the map boundaries. In this case,
+one or more of the generated random steps might leave the study area, making it
+impossible to compute a selection score. A possible solution is to simply
+terminate the simulation affected trajectory, assuming that the simulated
+individual has left the study area. However, this approach might produce
+ambiguous results in cases where many individuals are released near map borders,
+especially because already a single random step leaving the study area will
+break the simulation, thus resulting in biased connectivity estimates along map
+borders. Rather than breaking the simulation, we created a buffer zone
+\citep{Koen.2010} and resampled random steps until they fully lied within the
+study area. This proved to be an effective solution to overcome problems with
+boundary effects.
 
@@
 known endpoints, and provides a highly flexible tool for investigating
-connectivity. With this work, we hope to have sparked interest in the
-application, optimization, or creation of methods to investigate dispersal
-movements and connectivity via individual-based simulations, while at the same
-time stressing some of the non-trivial modeling decisions involved. We also hope
-to provide a useful framework that helps unifying and streamlining the
-application of individual-based simulation for assessing landscape connectivity.
+connectivity. Moreover, the simulation of dispersal opens up new avenues for
+incorporating interactions between habitat and movement covariates and provides
+the foundation for a rich suite of complementary connectivity measures. With
+this work, we hope to have sparked interest in the application, optimization, or
+creation of methods to investigate dispersal and connectivity via
+individual-based simulations, while at the same time stressing some of the
+non-trivial modeling decisions involved. We also hope to provide a useful
+framework that helps unifying and streamlining the application of
+individual-based simulations for assessing landscape connectivity.
 

### 2022-02-10 Chapter_1/04_Manuscript/Manuscript_2022-02-07.tex
Various updates

--- 2022-02-07 Chapter_1/04_Manuscript/Manuscript.tex
@@
 has mainly been investigated using least-cost analysis and circuit theory, two
-methods that make assumptions that are hardly applicable to dispersal.
-Least-cost analysis assumes that animals move towards a known endpoint and are
-knowledgeable about the most favorable route. Circuit theory presumes a complete
-random walk and fails to incorporate directional persistence. While these
-assumptions can be relaxed by simulating dispersal movements explicitly across
-the landscape, a unified approach for such simulations is lacking.
+methods that make assumptions that are hardly applicable to dispersal. While
+these assumptions can be relaxed by simulating dispersal trajectories across the
+landscape, a unified approach for such simulations is lacking.
+
+% Least-cost analysis assumes that animals move towards a known endpoint and are
+% knowledgeable about the most favorable route. Circuit theory presumes a
+% complete random walk and fails to incorporate directional persistence.
 
@@
 \cite{Avgar.2016}, we generated random steps by sampling random turning angles
-from a uniform distribution (\(-\pi, +\pi\)) (which is equivalent to a von Mises
-distribution with location and concentration parameters \(\mu = \kappa = 0\))
-and step lengths from a gamma distribution that was fitted to realized steps
-(scale \(\theta\) = 6'308 and shape \(k\) = 0.37).
+from a uniform distribution (\(-\pi, +\pi\)) and step lengths from a gamma
+distribution that was fitted to realized steps (scale \(\theta\) = 6'308 and
+shape \(k\) = 0.37).
 

### 2022-02-10 Chapter_4/02_R-Scripts/02_AnalyseData.r

@@ -26,9 +26,10 @@ options(scipen = 999)
 # Load observed movement data
 obs <- read_csv("03_Data/01_RawData/ObservedMovements.csv")
 
-# Define original distributions
-sl_dist <- c(shape = 3, scale = 1)
-ta_dist <- c(kappa = 0.5, mu = 0)
+# Define the original distributions
+sl_dist <- list(shape = 3, scale = 1)
+ta_dist <- list(kappa = 0.5, mu = 0)
+dists <- list(uncorrected = list(sl = sl_dist, ta = ta_dist))
 
 # Load covariate layers
 covars <- stack("03_Data/01_RawData/CovariateLayers.grd")
@@ -37,10 +38,10 @@ covars <- readAll(covars)
 # Specify the different design combinations through which we want to run. Note
 # that a forgiveness of 1 refers to a regular step selection function
 dat <- expand_grid(
-    Missingness = seq(0, 0.8, by = 0.1)  # Fraction of the fixes that is removed
-  , Forgiveness = 1:5                    # Allowed lag of steps (in steps)
-  , Replicate   = 1:100                  # Number of replicates for each combination
-  , AdjustDists = c(T, F)                # Whether to use a dynamic distribution for step lengths
+    Missingness    = seq(0, 0.8, by = 0.1)                  # Fraction of the fixes that is removed
+  , Forgiveness    = 1:5                                    # Allowed lag of steps (in steps)
+  , Replicate      = 1:100                                  # Number of replicates for each combination
+  , Distributions  = c("uncorrected", "naive", "dynamic")   # Which distributions to use
 )
 
 ################################################################################
@@ -102,7 +103,7 @@ computeMetrics <- function(data) {
 }
 
 # Function to generate random steps
-computeSSF <- function(data, n_rsteps, adjust_dists) {
+computeSSF <- function(data, n_rsteps, distributions) {
 
   # Indicate case steps
   data$case <- 1
@@ -116,40 +117,49 @@ computeSSF <- function(data, n_rsteps, adjust_dists) {
   # Indicate that they are control steps (case = 0)
   rand$case <- 0
 
-  # Sample new step lengths and turning angles
-  if (adjust_dists) {
[…]

### 2022-02-22 Chapter_1/04_Manuscript/Manuscript.tex
updates

@@
 %------------------------------------------------------------------------------
-Dispersal of individuals contributes to long-term population persistence, yet
-requires a sufficient degree of landscape connectivity. To date, connectivity
-has mainly been investigated using least-cost analysis and circuit theory, two
-methods that make assumptions that are hardly applicable to dispersal.
-Least-cost analysis assumes that animals move towards a known endpoint and are
-knowledgeable about the most favorable route. Circuit theory presumes a complete
-random walk and fails to incorporate directional persistence. While these
-assumptions can be relaxed by explicitly simulating dispersal trajectories
-across the landscape, a unified approach for such simulations is lacking.
-
-Here, we present a simple three-step approach to simulate dispersal and to
-assess connectivity using empirical GPS movement data and a set of habitat
-covariates. In step one, we use integrated step-selection functions to fit a
-mechanistic movement model describing habitat and movement preferences of
+\noindent 1. Dispersal of individuals contributes to long-term population
+persistence, yet requires a sufficient degree of landscape connectivity. To
+date, connectivity has mainly been investigated using least-cost analysis and
+circuit theory, two methods that make assumptions that are hardly applicable to
+dispersal. Least-cost analysis assumes that animals move towards a known
+endpoint and are knowledgeable about the most favorable route. Circuit theory
+presumes a complete random walk and fails to incorporate directional
+persistence. While these assumptions can be relaxed by explicitly simulating
+dispersal trajectories across the landscape, a unified approach for such
+simulations is lacking.
+
+\noindent 2. Here, we present a simple three-step approach to simulate dispersal
+and to assess connectivity using empirical GPS movement data and a set of
+habitat covariates. In step one, we use integrated step-selection functions to
+fit a mechanistic movement model describing habitat and movement preferences of
 dispersing individuals. In step two, we apply the parameterized model to
@@
 inter-patch connectivity indicating the presence and intensity of functional
-links between habitat patches. As a case study, we apply the three-step
-approach to GPS data collected on dispersing African wild dogs (\textit{Lycaon
-pictus}) inhabiting the Kavango-Zambezi Transfrontier Conservation Area
-(KAZA-TFCA) in southern Africa.
-
-Using step-selection functions we successfully parametrized a detailed dispersal
-model that described dispersing individuals habitat and movement preferences, as
-well as potential interactions among the two. The model substantially
-outperformed a model that omitted such interactions and enabled us to simulate
-80,000 dispersal trajectories across the study area. By explicitly simulating
-individuals, this approach not only requires fewer unrealistic assumptions about
-dispersal, but also permits the calculation of a rich suite of connectivity
-metrics that together provide a comprehensive view on landscape connectivity.
-The three derived connectivity maps revealed several dispersal hotspots and
+links between habitat patches. We demonstrate the applicability of the proposed
+three-step approach in a case study in which we use GPS data collected on
+dispersing African wild dogs (\textit{Lycaon pictus}) inhabiting the
+Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA) in southern Africa.
+
+\noindent 3. Using step-selection functions we successfully parametrized a
+detailed dispersal model that described dispersing individuals habitat and
+movement preferences, as well as potential interactions among the two. The model
+substantially outperformed a model that omitted such interactions and enabled us
+to simulate 80,000 dispersal trajectories across the study area.
+
+\noindent 4. By explicitly simulating individuals, our approach not only
+requires fewer unrealistic assumptions about dispersal, but also permits the
+calculation of a variety of connectivity metrics that together provide a
+comprehensive view of landscape connectivity. In our case study, the three
+derived connectivity maps revealed several wild dog dispersal hotspots and
 corridors across the extent of our study area, yet each map highlighted a
@@
 offers a simple yet powerful alternative to traditional connectivity modeling
-techniques. The approach is therefore useful for a variety of applications in
+techniques. Our approach is therefore useful for a variety of applications in
 ecological, evolutionary, and conservation research.
@@
 (IBMMs, \citealp{Diniz.2019}). However, IBMMs require a large number of
-subjective modeling decisions, thus rendering among-system comparisons
-difficult.
+subjective modeling decisions, thus making among-system comparisons difficult.
 
@@
 useful for assessing landscape connectivity. Of the 50,000 simulated dispersal
-trajectories with starting point within the main study area, only 4.5\% reached
-a map boundary, suggesting that we successfully mitigated biases from boundary
+trajectories that originated from the main study area, only 4.5\% reached a map
+boundary, suggesting that we successfully mitigated biases from boundary
 effects. Moreover, our examination of the relative traversal frequency across

### 2022-02-22 Chapter_4/04_Manuscript/Manuscript.tex
updates

@@
 high frequency data which has been rarified using a change-point detection
-algorithm, thus also resulting in irregular step-durations.
+algorithm, thus also resulting in irregular step-durations. In case of missing
+steps, in contrast, steps durations are not entirely random but usually
+clustered.
 
@@
 set of 10 random steps originating from the sampled starting point. We then
-generated random steps by sampling turing angles from a von Mises distribution
+generated random steps by sampling turning angles from a von Mises distribution
 with concentration parameter \(\kappa = 0.5\) and location parameter \( \mu = 0
@@
 and scale parameter \(\theta = 1\). Third, we extracted covariate values along
-each random step from the underlying covariate layers and compute the average of
-each covariate along every step. Fourth, we predicted the probability of being
-chosen for each step by applying \Cref{EQ1}. The vector of relative habitat
-preferences used to make predictions was set to \(\beta_{water} = -1\),
+each random step from the underlying covariate layers and computed the average
+of each covariate along every step. Fourth, we predicted for each step the
+probability of being chosen by applying \Cref{EQ1}. The vector of relative
+habitat preferences used to make predictions was set to \(\beta_{water} = -1\),
 \(\beta_{elev} = 0.5\), and \(\beta_{dist} = -15\). Fifth, we sampled one of the
@@
   \caption{100 Simulated movement trajectories, each comprising of 100 steps.
-  Simulated individual were initiated within the blue rectangle to mitigate edge
-  effects.}
+  Simulated individuals were initiated within the blue rectangle to mitigate
+  edge effects.}
   \label{Simulations}
@@
 
-\paragraph{Data Rarification}
-We rarified the simulated GPS data by randomly removing a fixed fraction of GPS
+\paragraph{Data Rarefication}
+We rarefied the simulated GPS data by randomly removing a fixed fraction of GPS
 fixes. To assess the impact of different degrees of  ``missingness'', we
@@
 \paragraph{Step Length and Turning Angle Distributions}
-To generate random steps, we employed two competing approaches. In the first
-approach, we sampled step-lengths and turning-angles from the same distributions
-that were used to simulate GPS data (cfr. Section ...). To adjust sampled step
-lengths to the different step durations of observed steps, we multiplied sampled
-step lengths with the step-duration of the respective observed step. However, in
-a pre-analysis we found that step-lengths do not linearly scale with step
-duration, hence we also tested an alternative approach. Specifically, we fitted
-separate step-length and turning angle distributions for each step duration.
+To generate random steps, we employed three competing approaches, coined
+\textit{uncorrected}, \textit{na\"ive}, and \textit{dynamic} approach,
+respectively. In the uncorrected approach, we sampled step-lengths and
+turning-angles from parametric gamma and von Mises distributions that were
+fitted to observed steps with step-duration of one. In the \textit{na\"ive}
+approach, we sampled step lengths and turning angles from the same
+distributions, yet we linearly scaled sampled step lengths to the actual
+step-duration. For instance, for any step with a step-duration of two, we
+doubled the sampled step length. This approach is representative of the
+time-varying SSF approach proposed by \cite{Munden.2021}. Finally, in the
+\textit{dynamic} approach, we sampled step lengths and turning angles from
+distributions that were fit to different step-durations. That is, we fitted
+separate step length and turning angles distributions to every possible step
+duration. To achieve this, we subsampled the observed data to 50\% of the fixes,
+which resulted in varying step-durations. We then and fitted step length and
+turning angle distributions to the steps from different step-durations. We
+repeated this procedure 1000 times and averaged the parameter estimates across
+replicates.
 

### 2022-04-29 Chapter_1/04_Manuscript/Manuscript.tex
upd

@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{scalerel}
+\usepackage{tikz}
+\usetikzlibrary{svg.path}
 
@@
 
+% To be able to put an ORCID
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{
+  orcidlogo/.pic={
+    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
+    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
+                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+  }
+}
+\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
+\begin{tikzpicture}[yscale=-1,transform shape]
+  \pic{orcidlogo};
+\end{tikzpicture}
+}{|}}}}
+
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-\title{A Thee-Step Approach for Assessing Landscape Connectivity via Simulated
+\title{A Three-Step Approach for Assessing Landscape Connectivity via Simulated
 Dispersal: African Wild Dog Case Study}
@@
 \author{
-  David D. Hofmann\textsuperscript{1,2,\S} \and
-  Gabriele Cozzi\textsuperscript{1,2} \and
+  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
+  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
   John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \and
-  Dominik M. Behr\textsuperscript{1,2}
+  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
+  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
 }
@@
 %------------------------------------------------------------------------------
-\noindent 1. Dispersal of individuals contributes to long-term population
-persistence, yet requires a sufficient degree of landscape connectivity. To
-date, connectivity has mainly been investigated using least-cost analysis and
-circuit theory, two methods that make assumptions that are hardly applicable to
-dispersal. Least-cost analysis assumes that animals move towards a known
-endpoint and are knowledgeable about the most favorable route. Circuit theory
-presumes a complete random walk and fails to incorporate directional
-persistence. While these assumptions can be relaxed by explicitly simulating
-dispersal trajectories across the landscape, a unified approach for such
-simulations is lacking.
-
-\noindent 2. Here, we present a simple three-step approach to simulate dispersal
-and to assess connectivity using empirical GPS movement data and a set of
-habitat covariates. In step one, we use integrated step-selection functions to
-fit a mechanistic movement model describing habitat and movement preferences of
-dispersing individuals. In step two, we apply the parameterized model to
-simulate dispersal across the study area. In step three, we derive a set of
-three complementary connectivity maps; a heatmap highlighting frequently
-traversed areas, a betweenness map pinpointing dispersal corridors, and a map of
-inter-patch connectivity indicating the presence and intensity of functional
-links between habitat patches. We demonstrate the applicability of the proposed
-three-step approach in a case study in which we use GPS data collected on
-dispersing African wild dogs (\textit{Lycaon pictus}) inhabiting the
-Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA) in southern Africa.
-
-\noindent 3. Using step-selection functions we successfully parametrized a
-detailed dispersal model that described dispersing individuals habitat and
-movement preferences, as well as potential interactions among the two. The model
-substantially outperformed a model that omitted such interactions and enabled us
-to simulate 80,000 dispersal trajectories across the study area.
-
-\noindent 4. By explicitly simulating individuals, our approach not only
-requires fewer unrealistic assumptions about dispersal, but also permits the
-calculation of a variety of connectivity metrics that together provide a
-comprehensive view of landscape connectivity. In our case study, the three
-derived connectivity maps revealed several wild dog dispersal hotspots and
-corridors across the extent of our study area, yet each map highlighted a
-different aspect of landscape connectivity, thus emphasizing their complementary
-nature. Overall, our case study demonstrates that a simulation-based approach
-offers a simple yet powerful alternative to traditional connectivity modeling
-techniques. Our approach is therefore useful for a variety of applications in
-ecological, evolutionary, and conservation research.
+\noindent \textbf{Context} \quad Dispersal of individuals contributes to
+long-term population persistence, yet requires a sufficient degree of landscape
+connectivity. To date, connectivity has mainly been investigated using
+least-cost analysis and circuit theory, two methods that make assumptions that
+are hardly applicable to dispersal. While these assumptions can be relaxed by
+explicitly simulating dispersal trajectories across the landscape, a unified
+approach for such simulations is lacking.
+
+\noindent \textbf{Objectives} \quad Here, we propose and apply a simple
+three-step approach to simulate dispersal and to assess connectivity using
+empirical GPS movement data and a set of habitat covariates.
+
+\noindent \textbf{Methods} \quad In step one of the proposed approach, we use
+integrated step-selection functions to fit a mechanistic movement model
+describing habitat and movement preferences of dispersing individuals. In step
+two, we apply the parameterized model to simulate dispersal across the study
+area. In step three, we derive three complementary connectivity maps; a heatmap
+highlighting frequently traversed areas, a betweenness map pinpointing dispersal
+corridors, and a map of inter-patch connectivity indicating the presence and
+intensity of functional links between habitat patches. We demonstrate the
+applicability of the proposed three-step approach in a case study in which we
+use GPS data collected on dispersing African wild dogs (\textit{Lycaon pictus})
+inhabiting northern Botswana.
+
+\noindent \textbf{Results} \quad Using step-selection functions we successfully
+parametrized a detailed dispersal model that described dispersing individuals'
+habitat and movement preferences, as well as potential interactions among the
+two. The model substantially outperformed a model that omitted such interactions
+and enabled us to simulate 80,000 dispersal trajectories across the study area.
+
+\noindent \textbf{Conclusion} \quad By explicitly simulating dispersal
+trajectories, our approach not only requires fewer unrealistic assumptions about
+dispersal, but also permits the calculation of multiple connectivity metrics
+that together provide a comprehensive view of landscape connectivity. In our
+case study, the three derived connectivity maps revealed several wild dog
+dispersal hotspots and corridors across the extent of our study area. Each map
+highlighted a different aspect of landscape connectivity, thus emphasizing their
+complementary nature. Overall, our case study demonstrates that a
+simulation-based approach offers a simple yet powerful alternative to
+traditional connectivity modeling techniques. It is therefore useful for a
+variety of applications in ecological, evolutionary, and conservation research.
 
@@
 
-\onehalfspacing
-\tableofcontents
+% \onehalfspacing
+% \tableofcontents
 \doublespacing
@@
 % Importance of Dispersal & Connectivity
-\subsection{Importance of Connectivity \& Connectivity Models}
+% \subsection{Importance of Connectivity \& Connectivity Models}
 Dispersal of individuals is a vital process that allows species to maintain
-genetic diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
-Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and to
-colonize unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the
-ability to disperse depends on a sufficient degree of landscape connectivity
-\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
-dispersal corridors that promote connectivity a task of fundamental importance
-\citep{Nathan.2008, Doerr.2011, Rudnick.2012}. Identifying dispersal corridors
-not only necessitates a comprehensive understanding of the factors that limit
-dispersal, but also an appropriate model to estimate connectivity
-\citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the most commonly
-used connectivity models are least-cost path analysis (LCPA;
-\citealp{Adriaensen.2003}) and circuit theory (CT; \citealp{McRae.2006,
-McRae.2008}). Unfortunately, both models rest on assumptions that appear
-unsuitable for dispersers, thus calling for the development of alternative
-approaches. One promising alternative is to assess landscape connectivity via
-simulated dispersal trajectories generated from individual-based movement models
-(IBMMs, \citealp{Diniz.2019}). However, IBMMs require a large number of
-subjective modeling decisions, thus making among-system comparisons difficult.
+genetic diversity \citep{Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013},
+rescue non-viable populations \citep{Brown.1977}, and to colonize unoccupied
+habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability to disperse
+depends on a sufficient degree of landscape connectivity \citep{Fahrig.2003,
+Clobert.2012}, making the identification and protection of dispersal corridors
+that promote connectivity a task of fundamental importance \citep{Doerr.2011,
+Rudnick.2012}. Identifying dispersal corridors not only necessitates a
+comprehensive understanding of the factors that limit dispersal, but also an
+appropriate model to estimate connectivity \citep{Baguette.2013, Vasudev.2015,
+Hofmann.2021}. To date, the most commonly used connectivity models are
+least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory
+(CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both models rest on
+assumptions that appear unsuitable for dispersers, thus calling for the
+development of alternative approaches. One promising alternative is to assess
+landscape connectivity via simulated dispersal trajectories generated from
+individual-based movement models (IBMMs, \citealp{Diniz.2019}). However, IBMMs
+require a large number of subjective modeling decisions, thus making
+among-system comparisons difficult.
 
 % Issues with Both Methods
-\subsection{Issues with Traditional Connectivity Models}
+% \subsection{Issues with Traditional Connectivity Models}
 Traditional connectivity models make assumptions that are rarely met for
@@
 % What about IBMMs?
-\subsection{What about IBMMs?}
+% \subsection{What about IBMMs?}
 The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
@@
 strictly shifts the focus from a structural to a more functional view on
-connectivity \citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}.
-Furthermore, IBMMs generate movement sequentially, i.e. they generate a series
-of steps, so that the temporal dimension of dispersal becomes explicit and
-allows modeling autocorrelation between successive steps \citep{Diniz.2019}.
-Finally, simulations from IBMMs do not enforce movement or connections towards
-preconceived endpoints but allow individuals to adjust their route  ``on the
-go'', thereby preventing biases arising from misplaced endpoints. Despite these
-advantages, a unifying approach to simulate dispersal and assess connectivity
-using IBMMs is lacking. Considering the large number of subjective decisions
-entailed by IBMMs, an approach that streamlines and standardizes the application
-of dispersal simulations to assess connectivity will, however, be critical to
-safeguard comparability among studies.
+connectivity \citep{Tischendorf.2000}. Furthermore, IBMMs generate movement
+sequentially, i.e. they generate a series of steps, so that the temporal
+dimension of dispersal becomes explicit and allows modeling autocorrelation
+between successive steps \citep{Diniz.2019}. Finally, simulations from IBMMs do
+not enforce movement or connections towards preconceived endpoints but allow
+individuals to adjust their route  ``on the go'', thereby preventing biases
+arising from misplaced endpoints. Despite these advantages, a unifying approach
+to simulate dispersal and assess connectivity using IBMMs is lacking.
+Considering the large number of subjective decisions entailed by IBMMs, an
+approach that streamlines and standardizes the application of dispersal
+simulations to assess connectivity will, however, be critical to safeguard
+comparability among studies.
 
 % Proposed Solution: Three-Step approach
-\subsection{Proposed Solution: Three-Step Approach}
+% \subsection{Proposed Solution: Three-Step Approach}
 Here, we propose and exemplify a simple three-step approach for simulating
@@
 % Introduction of the study species
-\subsection{Case Study}
+% \subsection{Case Study}
 We showcase the application of the proposed approach using GPS movement data
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_GraphicalAbstract2.pdf}
-    \caption{Flowchart of the simulation-based connectivity analysis as proposed
-    in this article. First, GPS data and habitat covariates must be
-    collected. The combined data is then analyzed in an integrated step
-    selection model, which enables the parametrization of the focal species'
-    habitat and movement kernels and results in a mechanistic movement model.
-    The parametrized model is then treated as an individual-based movement model
-    and used to simulate dispersal trajectories. Ultimately, simulated
-    trajectories serve to produce a set of maps that are pertinent to landscape
-    connectivity. This includes a heatmap, indicating the traversal frequency
-    across each spatial unit of the study area, a betweenness map, highlighting
-    movement corridors and bottlenecks, and, finally, an inter-patch
-    connectivity map, where the frequency of connections and their average
-    duration can be depicted.}
+    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
+    \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
+    data and habitat covariates must be collected. The combined data is then
+    analyzed using an integrated step selection model. The parametrized model is
+    then treated as an individual-based movement model and used to simulate
+    dispersal trajectories. Ultimately, simulated trajectories are translated
+    into a set of maps that are pertinent to landscape connectivity. This
+    includes a heatmap, indicating the traversal frequency across each spatial
+    unit of the study area, a betweenness map, highlighting movement corridors
+    and bottlenecks, and, finally, an inter-patch connectivity map, where the
+    frequency of connections and their average duration can be depicted.}
     \label{GraphicalAbstract}
@@
     \end{tikzpicture}
-    \caption{Illustration of the study area located in southern Africa. (a) The
-    study area was confined by a bounding box spanning the entire KAZA-TFCA
-    which comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia.
-    Data on remaining wild dog populations has been sourced from
-    \cite{Woodroffe.2012}. (b) The KAZA-TFCA currently represents the world's
-    largest terrestrial transfrontier conservation area, covering a total area
-    of 520'000 km\textsuperscript{2}. Its main purpose is to re-establish
+    \caption{Illustration of the study area in southern Africa. (a) The study
+    area was confined by a bounding box spanning the entire KAZA-TFCA which
+    comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia. Data on
+    remaining wild dog populations (orange) has been sourced from
+    \cite{Woodroffe.2012}. (b) The KAZA-TFCA represents the world's largest
+    terrestrial transfrontier conservation area and covers a total area of
+    520'000 km\textsuperscript{2}. Its main purpose is to re-establish
     connectivity between already-existing NPs (dark green) and other protected
@@
 written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
-\citep{Eddelbuettel.2011, Eddelbuettel.2013}.
+\citep{Eddelbuettel.2011, Eddelbuettel.2013, Eddelbuettel.2018}.
 
@@
 Along each realized and random step, we extracted values from underlying habitat
-covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
-computed averages of each covariate along the steps. Besides extracting
-\textit{habitat covariates}, we also computed movement metrics that we used as
-\textit{movement covariates} in the ISSF models \citep{Avgar.2016,
-Fieberg.2021}. Specifically, we computed the step length (\textsf{sl}), its
-natural logarithm (\textsf{log(sl)}), and the cosine of the relative turning
-angle (\textsf{cos(ta)}), which is a measure of directionality
+covariate layers and we computed averages of each covariate along the steps.
+Besides extracting \textit{habitat covariates}, we also computed movement
+metrics that we used as \textit{movement covariates} in the ISSF models
+\citep{Avgar.2016, Fieberg.2021}. Specifically, we computed the step length
+(\textsf{sl}), its natural logarithm (\textsf{log(sl)}), and the cosine of the
+relative turning angle (\textsf{cos(ta)}), which is a measure of directionality
 \citep{Turchin.1998}, for each step. Because wild dog activity is low during the
@@
 indicating the absolute frequency at which different areas were traversed by
-simulated dispersal trajectories (e.g. \citealp{Peer.2008, Hauenstein.2019,
-Zeller.2020}). Specifically, we rasterized all simulated trajectories onto a
-raster with 1 km x 1 km resolution and tallied resulting layers into a single
-map. This procedure ensured that every trajectory was only counted once, even if
-it traversed the same raster-cell multiple times, thus reducing potential biases
-caused by individuals that were surrounded by unfavorable habitat and ``moved in
+simulated dispersal trajectories (e.g. \citealp{Hauenstein.2019, Zeller.2020}).
+Specifically, we rasterized all simulated trajectories onto a raster with 1 km x
+1 km resolution and tallied resulting layers into a single map. This procedure
+ensured that every trajectory was only counted once, even if it traversed the
+same raster-cell multiple times, thus reducing potential biases caused by
+individuals that were surrounded by unfavorable habitat and ``moved in
 circles''. To achieve high performance rasterization, we used the R-package {\tt
@@
 \section{Results}
-\subsection{Movement Model}
+% \subsection{Movement Model}
 The most parsimonious movement model consisted of movement covariates, habitat
@@
     associated frequency in (b1) highlights that the most parsimonious model
-    successfully outperforms a random guess (b2) and assigns comparably high
-    selection scores to realized steps.}
+    successfully outperformed a random guess (b2) and assigned high selection
+    scores to realized steps.}
     \label{MovementModel}
@@
   tallied all simulated trajectories. Consequently, the map highlights areas
-  that are frequently traversed by virtual dispersers. For spatial reference we
-  plotted a few selected NPs (dark gray). Additional heatmaps showing the
-  traversal frequency when individuals move fewer than 2'000 steps are provided
-  in Figure S4.}
+  that are frequently traversed. For spatial reference we plotted a few selected
+  NPs (dark gray). Additional heatmaps showing the traversal frequency when
+  individuals move fewer than 2'000 steps are provided in Figure S4.}
   \label{Heatmap}
@@
   simulated dispersers originating from each park. Black dots represent NPs that
-  were smaller than 700 km\textsuperscript{2} and therefore did not serve as
+  were smaller than 700 km\textsuperscript{2} and therefore were not used as
   source areas. Arrows between NPs illustrate between which NPs the simulated
   dispersers successfully moved and the color of each arrow shows the average
-  number of steps (4-hourly movements) that were necessary to realize those
+  number of steps (i.e. 4-hourly movements) that were necessary to realize those
   connections. Additionally, the line thickness indicates the relative number of
-  dispersers originating from a NP that realized those connections. Note that a
-  similar network view could be adopted to investigate connectivity between
-  other protected areas need not to be restricted to NPs.}
+  dispersers originating from a NP that realized those connections.}
   \label{InterpatchConnectivity}
@@
 % Short Summary
-\subsection{Short Summary}
+% \subsection{Short Summary}
 Here, we presented a simple three-step approach to assess landscape connectivity
@@
 % Movement Model
-\subsection{Movement Model}
+% \subsection{Movement Model}
 Results on the habitat kernel from our model showed that dispersers avoided
@@
 lengths and turning angles from copula probability distributions
-\citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to conveniently
-model such peculiarities directly in the movement model. While we only
-considered first order autocorrelation, i.e. correlation between two consecutive
-steps, higher order autocorrelation is conceivable and may be desirable to model
+\citep{Hodel.2022}, the ISSF framework allowed us to conveniently model such
+peculiarities directly in the movement model. While we only considered first
+order autocorrelation, i.e. correlation between two consecutive steps, higher
+order autocorrelation is conceivable and may be desirable to model
 \citep{Dray.2010, McClintock.2012}. However, this will require vast amounts of
@@
 
-\subsection{Maps}
+% \subsection{Maps}
 Each of the three connectivity maps derived from simulated dispersal
@@
 connectivity, but also to uncover potential dispersal traps
-\citep{Dwernychuk.1972, VanDerMeer.2014} or areas with a high susceptibility for
-human wildlife conflicts \citep{Cushman.2018}.
+\citep{VanDerMeer.2014} or areas with a high susceptibility for human wildlife
+conflicts \citep{Cushman.2018}.
 
@@
 incorporated through the preparation of multiple permeability surfaces on which
-the same connectivity model is repeatedly applied (e.g. \citealp{Benz.2016,
-Osipova.2019}). With simulations from ISSFs, in contrast, the environment could
-change ``as the dispersers move'', so that simulated trajectories would
-dynamically respond to seasonal fluctuations in the environment.
-
-\subsection{Disadvantages of ISSF Simulations}
+the same connectivity model is repeatedly applied (e.g. \citealp{Osipova.2019}).
+With simulations from ISSFs, in contrast, the environment could change ``as the
+dispersers move'', so that simulated trajectories would dynamically respond to
+seasonal fluctuations in the environment.
+
+% \subsection{Disadvantages of ISSF Simulations}
 Despite the many benefits and great flexibility offered by simulations from
@@
 assuming that wild dogs primarily survive in large, formally protected areas
-\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}.
-Moreover, we lacked precise knowledge about the distribution and abundance of
-wild dogs across protected areas, so we placed source points randomly within
-them. In cases where more detailed data distribution and abundance are
-available, source points could be distributed accordingly. Alternatively, source
-points could be distributed homogeneously but later be weighted when computing
-connectivity metrics. In any case, the challenge of selecting meaningful source
-points is not unique to individual-based simulations but also applies to LCPA
-and CT.
+\citep{DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. Moreover, we lacked
+precise knowledge about the distribution and abundance of wild dogs across
+protected areas, so we placed source points randomly within them. In cases where
+more detailed data distribution and abundance are available, source points could
+be distributed accordingly. Alternatively, source points could be distributed
+homogeneously but later be weighted when computing connectivity metrics. In any
+case, the challenge of selecting meaningful source points is not unique to
+individual-based simulations but also applies to LCPA and CT.
 
@@
 
-\subsection{Conclusion}
+% \subsection{Conclusion}
 In summary, we proposed and applied a simple three-step approach that relies on
@@
 Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
-Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
-Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
-Grant (31003A\_182286) to A. Ozgul.
+Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust,
+Forschungskredit of the University of Zürich (grant no. FK-21-091), and a Swiss
+National Science Foundation grant (no. 31003A\_182286) to A. Ozgul.
+
+\section{Conflict of Interest}
+All authors declare that they have no conflicts of interest.
 

### 2022-04-29 Chapter_1/04_Manuscript/Manuscript_Support.tex
Various updates
upd

@@
 \usepackage{pdflscape}
+\usepackage{scalerel}
+\usepackage{tikz}
+\usetikzlibrary{svg.path}
 
@@
 
+% To be able to put an ORCID
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{
+  orcidlogo/.pic={
+    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
+    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
+                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+  }
+}
+\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
+\begin{tikzpicture}[yscale=-1,transform shape]
+  \pic{orcidlogo};
+\end{tikzpicture}
+}{|}}}}
+
 % For supplementary material
@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ A Thee-Step Approach for Assessing Landscape
+\title{\textbf{Appendix}\\ A Three-Step Approach for Assessing Landscape
 Connectivity via Simulated Dispersal: African Wild Dog Case Study}
@@
 
+\begin{figure}[hbtp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_MapComparisonCountries.png}
+  \caption{Comparison of values from the heatmap and betweenness map inside
+  (blue) and outside (orange) the KAZA-TFCA borders within different countries.}
+  \label{MapComparisonCountries}
+ \end{center}
+\end{figure}
+
 %------------------------------------------------------------------------------

### 2022-04-29 Chapter_2/02_R-Scripts/00_Functions.R

@@ -8,6 +8,17 @@ correctedName <- function(x) {
 }
 
 ################################################################################
+#### Function to Compute Normalized Difference Index
+################################################################################
+# Function to compute the normalized difference (nd) index of two bands
+nd <- function(img, band_x, band_y) {
+  x <- img[[band_x]]
+  y <- img[[band_y]]
+  nd <- (x - y) / (x + y)
+  return(nd)
+}
+
+################################################################################
 #### Function to calculate Distances on a Raster Efficiently
 ################################################################################
 # Compute distance to a certain value on a raster
@@ -46,7 +57,7 @@ distanceTo <- function(x, value = 1) {
 #### Function to Resample GPS Data to a Desired Resolution
 ################################################################################
 # Function to resample fixes to a desired resolution
-.resFix <- function(data, hours, start, tol = 0.5) {
+resampleFixes <- function(data, hours, start, tol = 0.5) {
 
   # Identify the first date at which a fix was taken and set the time to the
   # desired start time
@@ -81,18 +92,23 @@ distanceTo <- function(x, value = 1) {
   return(data[closest, ])
 }
 
-# Wrapper to run the function on multiple individuals and multiple cores
-resampleFixes <- function(data, ID = NULL, hours, start, tol = 0.5, cores = 1) {
-  if (is.null(ID)) {
-    data_resampled <- .resFix(data, hours, start, tol)
-  } else {
-    data_resampled <- data %>% nest(data = -ID)
-    data_resampled$data <- pbmclapply(1:nrow(data_resampled), ignore.interactive = T, mc.cores = cores, function(x) {
-      .resFix(data_resampled$data[[x]], hours = hours, start = start, tol = tol)
-    })
-    data_resampled <- unnest(data_resampled, data)
-  }
-  return(data_resampled)
+################################################################################
+#### Function to Compute Bursts
+################################################################################
+# Function to compute bursts
+computeBursts <- function(data) {
+  bursted <- data %>%
+    mutate(dt_ = Timestamp - lag(Timestamp)) %>%
+    mutate(dt_ = as.numeric(dt_, units = "hours")) %>%
+    ungroup() %>%
+    mutate(NewBurst = ifelse(
+      dt_ > 8.25 |
[…]

### 2022-04-29 Chapter_2/02_R-Scripts/01_DataCleaning/02_LandCover.R

@@ -6,13 +6,13 @@
 # Clear R's brain
 rm(list = ls())
 
-# Set working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
-setwd(wd)
-
 # Load packages
 library(terra)       # To handle raster data
 library(tidyverse)   # For data wrangling
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
 
 # Load the globeland land cover map
 glob <- rast("03_Data/01_RawData/GLOBELAND/Globeland.tif")

### 2022-07-13 Chapter_4/04_Manuscript/Manuscript.tex
upd

@@
 \begin{abstract}
-This is an abstract
+Animal movement data collected in the field is often noisy and incomplete, thus
+confronting movement ecologists with temporal irregularities in their data. Step
+selection functions have become a popular tool to analyse movement data and to
+infer a species movement or preferences, yet they are based on the assumption of
+equal temporal gaps between consecutive datapoints. To adhere to this
+requirement, researchers typically only consider bursts of consecutive
+relocations with equal step-durations, thus discarding large portions of their
+data. Here, we use simulated data to question this practice and we examine
+whether including temporally irregular can improve estimates from step-selection
+models. In addition, we explore the usefulness of multiple methods to account
+for temporal irregularity in movement data. Our results show that including
+irregular data improves model estimates substantially. By applying methods that
+account for the irregularity in the data, the precision of model estimates can
+further be improved, albeit only marginally.
 \end{abstract}
@@
 dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
-even days between subsequent reolocations. Nevertheless, GPS regularly fail to
-adhere to the anticipated fixrate schedule due to cloud cover or other
+even days between subsequent reolocations. Nevertheless, GPS devices regularly
+fail to adhere to the anticipated fixrate schedule due to cloud cover or other
 unfavorable conditions, thus resulting in irregular data. In other cases, the
 GPS schedule is not strictly fixed but may be subject to change. For example,
-\cite{Cozzi.2020} increase the fixrate of their GPS collars once per week from 4
-to 8 fixes per day. We chose to coin this type of data irregular, because it is
-not collected on the anticipated fixrate schedule and therefore not directly
-comparable to the vast majority of collected data.
+\cite{Cozzi.2020} increased the fixrate of their GPS collars once per week from
+4 to 8 fixes per day, aiming to collect more fine scaled data for a limited
+time. We chose to coin this type of data \textit{irregular}, because it is not
+collected on the anticipated fixrate schedule and therefore not directly
+comparable to the vast majority of the collected data.
 
@@
 amount of available data and depending on the number of missing GPS fixes, this
-might entail little sacrifice. In some situations, however, removing irregular
-data implies a substantial loss of information. Additionally, recent
+practice might entail little sacrifice. In some situations, however, removing
+irregular data implies a substantial loss of information. Additionally, recent
 developments in the SSA realm have brought forward several improvements to SSA
@@
 irregular data.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
+  \caption{}
+  \label{Covariates}
+  \end{center}
+\end{figure}
 
@@
 interactions among habitat covariates and step metrics we implicitly assume
-independent habitat and movement kernels \citep{Avgar.2016}.
+independent habitat and movement kernels \citep{Avgar.2016}. Although
+measurement error would have been straight forward to implement, this was not
+the main focus of our analysis and so we assumed GPS locations to be 100 percent
+accurate.
 
@@
 of two, thus pardoning a single missing fix.
+
+The idea of the multiple imputation approach is as follows...
 

### 2022-07-20 Chapter_4/04_Manuscript/Manuscript.tex
updated scripts

--- 2022-02-22 Chapter_4/04_Manuscript/Manuscript.tex
@@
 
-\textbf{Keywords:} step selection functions, gps data, animal movement, Lycaon
-pictus, irregular data
+\textbf{Keywords:} step selection analysis, gps data, animal movement, irregular
+data
 
@@
 \begin{abstract}
-This is an abstract
+Animal movement data collected using GPS collars is often noisy and incomplete,
+thus confronting movement ecologists with data that is not always uniformly
+distributed in time. Step selection analysis has become a popular tool to
+analyse such data and to infer a species movement and habitat preferences, yet
+it requires that consecutive GPS relocations are collected at regular temporal
+intervals. To adhere to this requirement, researchers typically discard large
+portions of their data and only consider GPS bursts that were collected at
+comparable intervals. Here, we assess the appropriateness of this practice and
+use simulated data to examine if including temporally irregular data could, in
+fact, improve estimates from step-selection analyses. Moreover, we explore the
+usefulness of several methods to account for the temporal irregularity. Our
+results suggest that utilizing a full dataset, despite potential temporal
+irregularities, is beneficial, especially when an appropriate method is chosen
+to account for such irregularities.
 \end{abstract}
@@
 dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
-even days between subsequent reolocations. Nevertheless, GPS regularly fail to
-adhere to the anticipated fixrate schedule due to cloud cover or other
-unfavorable conditions, thus resulting in irregular data. In other cases, the
-GPS schedule is not strictly fixed but may be subject to change. For example,
-\cite{Cozzi.2020} increase the fixrate of their GPS collars once per week from 4
-to 8 fixes per day. We chose to coin this type of data irregular, because it is
-not collected on the anticipated fixrate schedule and therefore not directly
-comparable to the vast majority of collected data.
+even days between subsequent reolocations. Nevertheless, GPS devices frequently
+fail to adhere to the anticipated schedule, for instance due to cloud cover or
+other unfavorable conditions, thus resulting in irregular data. In other cases,
+the GPS schedule is not strictly fixed but may be subject to change. For
+example, \cite{Cozzi.2020} increased the fixrate of their GPS collars once per
+week from 4 to 8 fixes per day, aiming to collect more fine scaled data for a
+limited time. We chose to coin this type of data \textit{irregular}, because it
+is not collected on the anticipated fixrate schedule and therefore not directly
+comparable to the vast majority of the collected data.
 
@@
 irregular data.
+
+Specifically, we employed six competing approaches to handle temporal
+irregularity in the data.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
+  \caption{}
+  \label{Covariates}
+  \end{center}
+\end{figure}
 
@@
 \section{Results}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_DistributionParameters.png}
+  \caption{}
+  \label{DistributionParameters}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_ResultsHabitatKernel.png}
+  \caption{}
+  \label{ResultsHabitat}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_ResultsMovementKernel.png}
+  \caption{}
+  \label{ResultsMovement}
+  \end{center}
+\end{figure}
+
 \section{Discussion}

### 2022-07-20 Chapter_4/02_R-Scripts/02_Visualization/Covariates.R

@@ -1,0 +1,59 @@
+################################################################################
+#### Covariates
+################################################################################
+# Description: Plot of the covariates
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(raster)        # To handle spatial data
+library(tidyverse)     # For data wrangling
+library(sf)            # For plotting
+library(pbmcapply)     # For multicore abilities
+
+# Set working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
+
+# Load observed movement data and covariates
+obs <- read_csv("03_Data/01_RawData/ObservedMovements.csv")
+covars <- stack("03_Data/01_RawData/CovariateLayers.grd")
+
+# Extent in which animals were released
+ext <- extent(c(50, 250, 50, 250))
+ext <- as(ext, "SpatialPolygons")
+
+################################################################################
+#### Plot of Covariates and Simulated Trajectories
+################################################################################
+# Plot the covariates
+p1 <- as.data.frame(covars, xy = T) %>%
+  gather(key = covariate, value = value, 3:5) %>%
+  ggplot(aes(x = x, y = y, fill = value)) +
+    geom_raster() +
+    geom_sf(data = st_as_sf(ext), inherit.aes = F, fill = NA, col = "white", lty = 2) +
+    scale_fill_viridis_c(option = "viridis") +
+    coord_sf() +
+    theme_minimal() +
+    facet_wrap("covariate") +
+    scale_x_continuous(breaks = seq(0, 300, by = 100)) +
+    scale_y_continuous(breaks = seq(0, 300, by = 100)) +
+    theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) +
+    theme(strip.background = element_rect(fill = "gray95", color = NA))
+
+# Plot of trajectories
+p2 <- ggplot(obs, aes(x = x, y = y, col = as.factor(ID))) +
+  geom_sf(data = st_as_sf(ext), col = "cornflowerblue", fill = NA, inherit.aes = F, lwd = 1) +
+  geom_path(size = 0.1) +
+  geom_point(size = 0.1) +
+  coord_sf(xlim = c(0, 300), ylim = c(0, 300)) +
+  theme_minimal() +
+  scale_color_viridis_d() +
+  theme(
+      legend.position = "none"
+    , axis.title.y    = element_text(angle = 0, vjust = 0.5)
+  )
+
[…]

### 2022-07-26 Chapter_4/04_Manuscript/Manuscript.tex
Various updates

@@
 %------------------------------------------------------------------------------
-\title{Applying Step Selection Functions to Data with Data Gaps: Simulation and
-Case Study on African wild dogs}
+\title{Applying Step Selection Functions to Temporally Irregular GPS Data - A
+Simulation Study}
 
@@
 thus confronting movement ecologists with data that is not always uniformly
-distributed in time. Step selection analysis has become a popular tool to
-analyse such data and to infer a species movement and habitat preferences, yet
-it requires that consecutive GPS relocations are collected at regular temporal
-intervals. To adhere to this requirement, researchers typically discard large
-portions of their data and only consider GPS bursts that were collected at
-comparable intervals. Here, we assess the appropriateness of this practice and
-use simulated data to examine if including temporally irregular data could, in
-fact, improve estimates from step-selection analyses. Moreover, we explore the
-usefulness of several methods to account for the temporal irregularity. Our
-results suggest that utilizing a full dataset, despite potential temporal
-irregularities, is beneficial, especially when an appropriate method is chosen
-to account for such irregularities.
+distributed in time. Step selection analysis has become a popular tool to use
+such data and infer a species movement and habitat preferences. However, in
+order to ensure comparability among steps, the method requires that consecutive
+GPS relocations are collected at regular temporal intervals. To adhere to this
+requirement, researchers typically discard large portions of their data and only
+consider GPS bursts that were collected at similar intervals. Here, we assess
+the appropriateness of this practice and use simulated data to examine if
+including temporally irregular data could, in fact, improve estimates from
+step-selection analyses. We also explore several alternative methods to account
+for temporal irregularity. Our results suggest that utilizing a full dataset,
+even if it contains temporal irregularities, improves the accuracy of step
+selection estimates, especially when an appropriate method is chosen to account
+for temporal irregularities.
 \end{abstract}
@@
 \section{Introduction}
-GPS data is usually collected on fixed temporal intervals (could do a quick and
-dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
-even days between subsequent reolocations. Nevertheless, GPS devices frequently
-fail to adhere to the anticipated schedule, for instance due to cloud cover or
-other unfavorable conditions, thus resulting in irregular data. In other cases,
-the GPS schedule is not strictly fixed but may be subject to change. For
-example, \cite{Cozzi.2020} increased the fixrate of their GPS collars once per
-week from 4 to 8 fixes per day, aiming to collect more fine scaled data for a
-limited time. We chose to coin this type of data \textit{irregular}, because it
-is not collected on the anticipated fixrate schedule and therefore not directly
-comparable to the vast majority of the collected data.
+Researchers usually collect GPS on fixed temporal intervals (could do a quick
+and dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
+even days between subsequent reolocations. Despite this, GPS devices frequently
+fail to adhere to the anticipated schedule, for instance due to cloud cover,
+thus resulting in GPS data with irregular temporal intervals. In some cases, the
+GPS schedule may intentionally be subject to change to maximize battery
+lifetime. For example, \cite{Cozzi.2020} increased the fixrate of their GPS
+collars once per week from 4 to 8 fixes per day, aiming to collect more fine
+scaled data for a limited time. In either case, the data at hand will be patchy
+in terms of the temporal lag between consecutive recordings. We chose to coin
+this type of data \textit{irregular}, because it is not collected on the
+anticipated fixrate schedule and therefore not directly comparable to the vast
+majority of the collected data.
 
-Step selection analysis is a powerful framework that enables researchers to
-study habitat and movement preferences of their focal species. The method works
-by comparing spatial covariates at locations where the studied animal has been
-observed, to a set of locations where the animal was presumably absent. For
-this, the collected GPS data is converted into steps, where a step represents
-the straight line movement between two consecutive GPS fixes. One can then
-compare covariates experienced along observed steps to covariates along
-potential alternative steps and thus infer preferred or avoided features. By
-restricting the availability domain to a set of steps, SSA is specifically
-intended to account for spatio-temporal autocorrelation inherent to data
-collected using GPS collars and is therefore one of the preferred methods for
-analysing such data. However, it is considered good practice to only consider
-steps with similar step durations in step-selection models. Depending on the
-amount of available data and depending on the number of missing GPS fixes, this
-practice might entail little sacrifice. In some situations, however, removing
-irregular data implies a substantial loss of information. Additionally, recent
-developments in the SSA realm have brought forward several improvements to SSA
-that might allow to relax the assumption of regular step-durations. This
-includes the \textit{integrated} SSA approach presented by \citep{Avgar.2016},
-in which inference on habitat \textit{and} movement preferences are possible
-thanks to the inclusion of movement descriptors in the respective model. More
-recently, the method has been further refined by \citep{Munden.2021}, who coined
-the term \textit{time-varying integrated} SSA. This method was developed for
-high frequency data which has been rarified using a change-point detection
-algorithm, thus also resulting in irregular step-durations. In case of missing
-steps, in contrast, steps durations are not entirely random but usually
-clustered.
+Step selection analysis (SSA) is a powerful framework that enables researchers
+to investigate the habitat and movement kernel of their focal species. While the
+habitat kernel describes relative habitat preferences of the study species, the
+movement kernel describes its movement patterns. SSA works by comparing spatial
+covariates at locations where the studied animal has been observed, to a set of
+locations where the animal was presumably absent. For this, the collected GPS
+data is converted into steps, where a step represents the straight line movement
+between two consecutive GPS fixes \citep{Turchin.1998}. One can then compare
+covariates experienced along observed steps to covariates along potential
+alternative steps and thus infer preferred or avoided features. By restricting
+the availability domain to a set of steps, SSA is specifically intended to
+account for spatio-temporal autocorrelation inherent to data collected using GPS
+collars and is therefore one of the preferred methods for analysing such data.
+However, to ensure that steps are comparable, researchers usually only consider
+steps with similar step-durations in their models. We will coin a step with
+regular step-duration and for which a relative turning angle can be computed as
+a ``valid'' step. Note that in order for a step to be valid it is necessary to
+have at least three fixes collected at regular intervals. With only two fixes,
+the relative turning angle could not be computed.
+
+At this stage, it is worth clearly defining some of the terms that will be used
+repeatedly throughout this manuscript.
+
+\begin{itemize}
+  \item Regular GPS data: GPS data were all GPS fixes have been successfully
+  obtained on the aspired GPS schedule, e.g. every 2 hours, every day, etc.
+  \item Irregular GPS data: GPS data were some GPS fixes failed to be obtained
+  on the aspired GPS schedule, e.g. due to cloud cover or communication issues.
+  \item Step: Straight line connecting two consecutive GPS locations
+  \item Step-length: Euclidean distance covered by a step
+  \item Step-duration: Temporal duration of the step, i.e. the temporal
+  difference between the start- and endpoint of the step.
+  \item (Relative) turning angle: Orientation of a step, relative to the
+  preceeding step. Positive values indicate left-turns, negative values indicate
+  right turns.
+  \item Valid step: A step for which a step-length, step-duration, as well as
+  turning angle can be computed. These steps can be used for step-selection
+  analysis. Note that the first step of a trajectory will always be invalid, as
+  no turning angle can be computed for it.
+  \item Missingness: Fraction of the data that should have been collected, but
+  for some reason was not. For example, if ten fixes should have been collected
+  but two are missing, there is a missingness of 0.2 (i.e. 20\%).
+  \item Forgiveness: How forgivin a modeler is to accept a step that exceeds the
+  aspired step-duration. A modeler with forgiveness of one, for instance, only
+  considers steps with a step-duration of one (i.e. steps resulting from fixes
+  that are regularly spaced in time). In contrast, a modeler with forgiveness of
+  two would also accept steps with a step-duration of two, thus pardoning a
+  single missing fix.
+  \item Burst: Sequence of consecutive GPS fixes where the step-duration does
+  not exceed the forgiveness of the modeler.
+\end{itemize}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Overview.png}
+  \caption{Figure demonstrating how a missing fix affects the number of valid
+  steps that can be used for step selection analysis. In subfigure (a), a
+  complete trajectory without any missing fixes is given. This trajectory
+  results in a total of four valid steps that can be analysed (the first step is
+  lacking a relative turning angle and thus invalid). Each of the steps in (a)
+  has a step-duration of one, which is indicated in. In subfigure (b), the
+  fourth fix was removed. If the modeler does excludes steps with step-durations
+  above one (i.e. has a forgiveness of one), this results in only a single valid
+  step for analysis. Finally, subfigure (c) depicts the same situation, yet with
+  a modeler that has a forgiveness of two, and is thus willing to accept steps
+  with a step-duration of two. Here, we end up with three valid steps.}
+  \label{Overview}
+  \end{center}
+\end{figure}
+
+Depending on the amount of available data and the number of missing GPS fixes,
+only considering regular bursts might entail little sacrifice. In some
+situations, however, removing irregular data will imply a substantial loss of
+information. For instance, consider the trajectory depicted in \Cref{Overview}a.
+Assuming that the fixes 1-5 were collected on a regular interval, this sequence
+will result in 6 steps, 5 of which would be valid (for the first step we can't
+compute a relative turning angle). If, however, the 4th fix is missing, and the
+modeler is not willing to accept steps with a duration above one, then we are
+down to a single valid step. That is, one missing fix reduced the number of
+valid steps by three. If, on the other hand, the modeler deems a step-duration
+of two as acceptable, we achieve three valid steps. From this example it is
+clear that already a relatively small missingness implies a subsantial loss in
+the number of valid steps that can be used for step selection analysis (see also
+\Cref{ValidSteps}).
+
+Additionally, recent developments in the SSA realm have brought
+forward several improvements to SSA that might allow to relax the assumption of
+regular step-durations. This includes the \textit{integrated} SSA approach
+presented by \citep{Avgar.2016}, in which inference on habitat \textit{and}
+movement preferences are possible thanks to the inclusion of movement
+descriptors in the respective model. More recently, the method has been further
+refined by \citep{Munden.2021}, who coined the term \textit{time-varying
+integrated} SSA. This method was developed for high frequency data which has
+been rarified using a change-point detection algorithm, thus also resulting in
+irregular step-durations. In case of missing steps, in contrast, steps durations
+are not entirely random but usually clustered.
 
@@
   \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
-  \caption{}
-  \label{Covariates}
+  \caption{Illustration of how missingness in GPS data reduces the number of
+  steps that can be used in step selection analysis. At a missingness of 0, all
+  1'000 theoretical GPS points could be used to compute steps, thus resulting in
+  998 steps. However, if the missingness increases, data gaps will prevent the
+  computation of some steps. However, the higher the forgiveness of a modeler,
+  the more missing fixes a modeler will accept to still be able to compute a
+  step.}
+  \label{ValidSteps}
   \end{center}
@@
 \section{Methods}
-\subsection{Simulation Study}
-\subsubsection{Spatial Covariates}
+\subsection{Spatial Covariates}
 We simulated a virtual landscape comprising three spatial covariate layers, each
 with a resolution of 300 x 300 pixels (\Cref{Covariates}) spanning across x- and
-y-coordinates from 0 to 300. The first layer (\textsf{water}) represented
-water-bodies and was simulated using a random cluster nearest‐neighbour neutral
-landscape model \citep{Saura.2000}, with the cluster-proportion set to 0.5 and
-the patch occupancy of water fixed to 20\%. The second layer (\textsf{elev})
-resembled an elevation layer and was simulated using a Gaussian random fields
-neutral landscape model \citep{Schlather.2015}, with an autocorrelation range of
-10, magnitude of variation of the landscape of 1, and a magnitude of variation
-in scale of 0. We simulated both of these layers using the r-package {\tt NLMR}
-\citep{Sciaini.2018}. The third layer (\textsf{dist}) simply indicated the
-distance (in pixels) to the center of the virtual landscape \((x = 150, y =
-150)\), and can be understood as the distance to the center of an animal's
-home-range. We computed spatial distances using the r-package {\tt raster}
-\citep{Hijmans.2022}. We normalized the values from all layers to a range
-between zero and one.
+y-coordinates from 0 to 300. The first layer (\textsf{forest}) represented areas
+covered by forest and was simulated using a random cluster nearest‐neighbour
+neutral landscape model \citep{Saura.2000}, with the cluster-proportion set to
+0.5 and the patch occupancy of forest fixed to 20\%. The second layer
+(\textsf{elev}) resembled an elevation layer and was simulated using a Gaussian
+random fields neutral landscape model \citep{Schlather.2015}, with an
+autocorrelation range of 10, magnitude of variation of the landscape of 1, and a
+magnitude of variation in scale of 0. We simulated both of these layers using
+the r-package {\tt NLMR} \citep{Sciaini.2018}. The third layer (\textsf{dist})
+indicated the distance (in pixels) to the center of the virtual landscape \((x =
+150, y = 150)\), and can be understood as a point of attraction, such as, for
+instance, the center of an animal's home-range. We computed spatial distances
+using the r-package {\tt raster} \citep{Hijmans.2022}. We normalized the values
+from all layers to a range between zero and one.
 
@@
 
-\subsubsection{Movement Simulation}
-To simulate movement across the above generated virtual lanscape, we employed an
-``inverted'' step-selection function, which proceeded as follows. First, we
-generated a random starting location. To prevent starting points in the vicinity
-of map borders, we restricted sampled locations to x- and y-coordinates between
-50 and 250 (white dotted rectangle in \Cref{Covariates}). Second, we generated a
-set of 10 random steps originating from the sampled starting point. We then
-generated random steps by sampling turning angles from a von Mises distribution
-with concentration parameter \(\kappa = 0.5\) and location parameter \( \mu = 0
-\), and step lengths from a gamma distribution with shape parameter \(k = 3 \)
-and scale parameter \(\theta = 1\). Third, we extracted covariate values along
-each random step from the underlying covariate layers and computed the average
-of each covariate along every step. Fourth, we predicted for each step the
-probability of being chosen by applying \Cref{EQ1}. The vector of relative
-habitat preferences used to make predictions was set to \(\beta_{water} = -1\),
-\(\beta_{elev} = 0.5\), and \(\beta_{dist} = -15\). Fifth, we sampled one of the
-random steps based on predicted probabilities and computed the new position of
-the simulated individual. We the repated these steps until a total of 100 steps
-were realized and we replicated the simulation 100 times, thus resulting in 100
-independent movement trajectories (\Cref{Simulations}). Note that by excluding
-interactions among habitat covariates and step metrics we implicitly assume
-independent habitat and movement kernels \citep{Avgar.2016}. Although
-measurement error would have been straight forward to implement, this was not
-the main focus of our analysis and so we assumed GPS locations to be 100 percent
-accurate.
+\subsection{Movement Simulation}
+To simulate movement across the virtual landscape, we employed an ``inverted''
+step-selection function that proceeded as follows. First, we generated a random
+starting location by sampling random x- and y-coordinates. To prevent starting
+points in the vicinity of map borders, we restricted sampled locations to x- and
+y-coordinates between 50 and 250 (white dotted rectangle in \Cref{Covariates}).
+Second, we generated a set of 10 random steps originating from the randomly
+selected starting point. We then generated random steps by sampling turning
+angles from a von Mises distribution with concentration parameter \(\kappa =
+0.5\) and location parameter \( \mu = 0 \), and step lengths from a gamma
+distribution with shape parameter \(k = 3 \) and scale parameter \(\theta = 1\).
+Third, we extracted covariate values along each random step from the underlying
+covariate layers and computed the mean of each covariate along every step.
+Fourth, we predicted for each step the probability of being chosen by applying
+\Cref{EQ1}. The vector of relative selection strengths (i.e. the \(\beta\)s used
+to make predictions was set to \(\beta_{forest} = -1\), \(\beta_{elev} = 0.5\),
+and \(\beta_{dist} = -20\). Fifth, we sampled one of the random steps based on
+predicted probabilities and computed the new position of the simulated
+individual. We then repeated these steps until a total of 100 steps were. We
+replicated the simulation 10'000 times, thus resulting in 10'000 independent
+movement trajectories (\Cref{Simulations}). Although measurement error would
+have been straight forward to implement, this was not the main focus of our
+analysis and so we assumed GPS locations to be 100 percent accurate with regards
+to an animal's true x- and y-locations. The simulated trajectories are depicted
+in \Cref{Simulations}.
 
@@
   \includegraphics[width = 0.5\textwidth]{99_Simulations.png}
-  \caption{100 Simulated movement trajectories, each comprising of 100 steps.
+  \caption{10'000 Simulated movement trajectories, each comprising of 100 steps.
   Simulated individuals were initiated within the blue rectangle to mitigate
@@
 
-\subsubsection{Step Selection Analysis}
+\subsection{Treatments}
 To assess the consequences of missing GPS data in when employing step-selection
-functions, we analysed the simulated GPS data under different conditions
-\Cref{Design}. Specifically, we varied the degree of data-missingness (0\% to
-80\%), the forgiveness when analysing the data (1 to 5 fixes), and the
-parametric distributions from which random steps were generated (static vs.
-dynamic distributions). We replicated each combination of conditions 100 times
-and computed averages and standard deviations from the resulting model
-estimates.
+functions, we analysed the simulated GPS data under different conditions and
+using different methods \Cref{Design}. Specifically, we varied the degree of
+missingness (from 0\% to 50\%), the forgiveness with regard to step-durations
+(from 1 to 5), and the way in which random steps were generated. We replicated
+each treatment combination 100 times, estimated step-selection parameters and
+computed averages and standard deviations for each treatment.
 
-In this context, missingness and forgiveness are complementary terms. While
-missingness describes the fraction of data that has not been collected, we use
-the term forgiveness to describe how much missingness a modeler is willing to
-accept. A modeler with forgiveness of one, for instance, only considers steps
-with a step-duration of one (i.e. fixes that are regularly spaced). In contrast,
-a modeler with forgiveness of two would also accept steps with a step-duration
-of two, thus pardoning a single missing fix.
+\subsection{Data Rarefication}
+We rarefied the simulated GPS data by randomly removing a fixed fraction of GPS
+fixes. To assess the impact of different degrees of ``missingness'', we varied
+the fraction of removed data from 0 (complete dataset) to 0.5 (50\% missing)
+with increments of 0.1. The random removal of GPS fixes introduced temporal
+irregularity of the GPS fixes and so the resulting step-durations differed
+depending on the time elapsed between remaining fixes. In the complete dataset,
+a step-duration of one was assumed, whereas the step-duration increased by one
+for every missing fix.
 
-The idea of the multiple imputation approach is as follows...
+\subsection{Identifying Bursts and Computing Step Metrics}
+We used the rarefied data to compute movement bursts. A movement burst consisted
+of a sequence consecutive GPS fixes with step-durations that did not exceed the
+forgiveness. For instance, if the forgiveness was one, already a single missing
+GPS fix introduced a new burst. In contrast, if the forgiveness was two,
+step-durations up to two were allowed without enforcing a new burst. For each
+step in a burst we then computed the step length (\textsf{sl}), its natural
+logarithm (\textsf{log(sl)}), and the cosine of the relative turning angle
+(\textsf{cos(ta)}).
 
-\paragraph{Data Rarefication}
-We rarefied the simulated GPS data by randomly removing a fixed fraction of GPS
-fixes. To assess the impact of different degrees of  ``missingness'', we
-incrementally increased the fraction of missing fixes from 0.0 (complete
-dataset) to 0.8 (80\% missing) with 10\% increments. The removal of GPS fixes
-introduced irregular temporal intervals between remaining fixes; we will refer
-to these intervals as ``step-times'', where the  ``step'' refers to the straight
-line movement between two consecutive GPS fixes \citep{Turchin.1998}. In the
-complete dataset, a step-time of one was assumed, whereas the step-time
-increased by one for every fix that was missing between subsequent fixes. For
-instance, a single missing fix between two other fixes directly increased the
-step time from one to two.
+\subsection{Step Selection Function}
+To conduct step selection analysis, we paired each observed step with 10 random
+steps. In general, we generated random steps by sampling turning angles from a
+von Mises distribution and step lengths from a gamma distribution, both fitted
+to the observed data. However, depending on the chosen approach, the generation
+of random steps differed slightly. Specifically, we employed six different
+approaches:
+\begin{itemize}
+  \item \textit{Uncorrected}: In the \textit{uncorrected} approach, we sampled
+  step-lengths and turning-angles from parametric distributions fitted to
+  observed steps with step-durations of one. This approach ignores the fact that
+  observed steps exhibit different step-durations.
+  \item \textit{Na\"ive}: In the \textit{na\"ive} approach, we sampled
+  step-lengths and turning-angles from parametric distributions fitted to
+  observed steps with step-durations of one. However, we linearly scaled sampled
+  step lengths to the step-duration of the observed step. For instance, for any
+  step with a step-duration of two, we doubled sampled step lengths. This
+  approach implicitly assumes that step lengths linearly scale with
+  step-durations.
+  \item \textit{Dynamic}: In the \textit{dynamic} approach, we sampled step
+  lengths and turning angles from distributions that were fit to different
+  step-durations. That is, for observed steps with step duration of two, we
+  sampled step lengths and turning angles from distirbutions fit to observed
+  steps of duration of two. To parametrize all the necessariy distributions, we
+  randomly removed 50\% fixes from the simulated data and generated steps
+  between the remaining fixes. This resulted in steps with varying
+  step-durations. Based on the resulting step lengths and turning angles we
+  fitted separate distributions for all step-durations. We replicated the
+  subsampling procedure 1000 times and averaged the estimates for distribution
+  parameters across replicates.
+  \item \textit{Multistep}: In the \textit{multistep} approach, we sampled
+  step-lengths and turning-angles from parametric distributions fitted to
+  observed steps with step-durations of one. However, if an observed step had a
+  step-durations larger than one, we augmented each random step with yet another
+  random step. We repeated this augmentation until the number of consecutive
+  steps matched the desired step-duration. For instance, for an observed step
+  with step-duration of two, we sampled random steps twice and concatenated them
+  into ``random paths''. The resulting paths were then simplified to a straight
+  line connecting the first and last coordinate of each path.
+  \item \textit{Model}: In the \textit{model} approach we sampled step lengths
+  and turning angles from parametric distributions fitted to observed steps with
+  step-durations of one. However, we accounted for potential differences in
+  distribution parameters by including predictor variables in the model to
+  update the distribution parameters depending on the step-duration (cfr
+  \Cref{RegressionModel}).
+  \item \textit{Imputed}: In the \textit{imputed} approach, we imputed missing
+  fixes using predictions from a simple movement model. Specifically, we fitted
+  fitted Johnson's single-state movement model \citep{Johnson.2008} to each
+  observed trajectory and used the parameterized model to predict locations for
+  the missing fixes. This resulted in a complete dataset without any missing
+  fixes. We than sampled step-lengths and turning angles from distributions
+  fitted to observed steps with step-duration of one.
+\end{itemize}
 
-\paragraph{Identifying Bursts and Computing Step Metrics}
-We used the generated data to compute movement bursts. A movement burst
-consisted of a sequence consecutive GPS fixes with step-times that did not
-exceed the accepted forgiveness. For instance, if the forgiveness was one,
-already a single missing GPS fix introduced a new burst. In contrast, if the
-forgiveness was set to two, step-times of two units were allowed, without
-introducing a new step.
-
-We will use the term  ``forgiveness'' to indicate how many missing fixes were
-accepted before a new burst was enforced. For instance, a forgiveness of one
-implied that a single missing fix did not result in a new burst. We then
-converted the GPS fixes belonging to the same burst into steps, where each step
-resembled the straight line movement between two GPS fixes \citep{Turchin.1998}.
-For each step we computed a set of step metrics. This includd the step length,
-its logarithm, the turning angle and its cosine.
-
-\paragraph{Step Selection Function}
-To conduct step selection analysis, we paired each observed step with 10 random
-steps. We generated random steps by sampling turning angles from a von Mises
-distribution and step lengths from a gamma distribution. Depending on the study
-design, we either sampled turning angles and step lengths from fixed or fixed or
-dynamic distributions (cfr. Section ...). Along observed and random steps, we
-extracted underlying covariates and computed their averages along steps.
-
-\paragraph{Step Length and Turning Angle Distributions}
-To generate random steps, we employed three competing approaches, coined
-\textit{uncorrected}, \textit{na\"ive}, and \textit{dynamic} approach,
-respectively. In the uncorrected approach, we sampled step-lengths and
-turning-angles from parametric gamma and von Mises distributions that were
-fitted to observed steps with step-duration of one. In the \textit{na\"ive}
-approach, we sampled step lengths and turning angles from the same
-distributions, yet we linearly scaled sampled step lengths to the actual
-step-duration. For instance, for any step with a step-duration of two, we
-doubled the sampled step length. This approach is representative of the
-time-varying SSF approach proposed by \cite{Munden.2021}. Finally, in the
-\textit{dynamic} approach, we sampled step lengths and turning angles from
-distributions that were fit to different step-durations. That is, we fitted
-separate step length and turning angles distributions to every possible step
-duration. To achieve this, we subsampled the observed data to 50\% of the fixes,
-which resulted in varying step-durations. We then and fitted step length and
-turning angle distributions to the steps from different step-durations. We
-repeated this procedure 1000 times and averaged the parameter estimates across
-replicates.
-
-\paragraph{Regression Model}
-We estimated ... using conditional logistic regression, implemented using the
-r-package \textsf{survival} \citep{Therneau.2021}. We followed \cite{Avgar.2016}
-and \cite{Fieberg.2021} and employed \textit{integrated} SSA. That is, aside
-from the habitat covariates (\textsf{water, elev, dist}), we also included
-descriptors of the step length and turning angle (\textsf{sl, log(sl), cos(ta)})
-in our regression model. The model call was as follows:
-
-$$
-case\_ \sim water + elev + dist + sl + log(sl) + cos(ta)
-$$
-
-\subsection{Case Study}
-
-because wild dogs move comparably little in the hours between xx and xx, we made
-the simplyfing assumption that the 8-hour period from xx to xx is comparable to
-a 4-hour period.
-
-\section{Results}
+Together, an observed and its 10 associated random steps formed a stratum that
+received a unique ID. Along all steps we then extracted covariate values from
+the underlying covariate layers and computed the mean of each covariate along
+every step.
 
@@
   \includegraphics[width = \textwidth]{99_DistributionParameters.png}
-  \caption{}
+  \caption{Parameter estimates for the von Mises distribution (top row) and
+  gamma distribution (bottom row) fitted to steps with different durations. The
+  von Mises distribution requires two parameters, namely a concentration
+  parameter and a location parameter. Here, the conentration parameter
+  diminishes as the step-duration increases, indicating that steps with higher
+  step duration tend to be less directional. The location parameter, on the
+  other hand, remains unchanged at 0, indicating that movement tends to be
+  forward directed and is not biased towards left or right turns. The gamma
+  distribution requires a shape and a scale parameter. Here, the shape parameter
+  linearly increases with increasing step-duration, yet the shape parameter
+  levels off at a step-duration of two.}
   \label{DistributionParameters}
@@
 
+\subsection{Regression Model}
+\label{RegressionModel}
+We estimated relative selection strengths (i.e. the \(\beta\)s in \Cref{EQ1})
+using conditional logistic regression, implemented using the r-package
+\textsf{survival} \citep{Therneau.2021}. That is, our response variable was a
+binary variable indicating if a step was an observed (i.e. a case step, scored
+1) or random step (i.e. a control step, scored 0). Following \cite{Avgar.2016}
+and \cite{Fieberg.2021}, we employed \textit{integrated} SSA and besides habitat
+covariates (\textsf{forest, elev, dist}), also included descriptors of the step
+length and turning angle (\textsf{sl, log(sl), cos(ta)}) as predictors in our
+regression model. Including those predictors allowed us to learn more about the
+movement characteristics of the focal species and to update our tentative
+parameters for the step-length and turning angle distirbutions (for further
+details see \citealp{Fieberg.2021}). The model call was as follows:
+
+$$
+case \sim forest + elev + dist + sl + log(sl) + cos(ta)
+$$
+
+For the \textit{model} approach, however, the model call was slightly adjusted
+and also included interactions between the factor covariate step-duration and
+the movement descriptors. It thus looked as follows:
+
+\begin{align*}
+case \sim forest &+ elev + dist + sl + log(sl) + cos(ta) + sl:duration \\
+  &+ log(sl):duration + cos(ta):duration
+\end{align*}
+
+We recorded beta estimates across all replicates, computed their means and
+standard deviations for the different treatment combinations.
+
+\section{Results}
+\subsection{Habitat Kernel}
+
+Model estimates for the habitat kernel are illustrated in \Cref{ResultsHabitat}.
 \begin{figure}
@@
   \includegraphics[width = \textwidth]{99_ResultsHabitatKernel.png}
-  \caption{}
+  \caption{Parameter estimates with regards to the habitat covariates
+  \textsf{dist}, \textsf{elev}, and \textsf{forest}. Dots indicate the mean
+  estimates across 100 replicates, whereas the vertical line delineate $\pm 2SD$
+  from the mean. The black dotted lines indicate the ``true'' values that were
+  used to simulate movement. Colors separate different approaches to account for
+  varying step-durations in the data. For clarity, we only show results for the
+  treatments with data missingness of 0.5.}
   \label{ResultsHabitat}
@@
 
+\subsection{Movement Kernel}
+
 \begin{figure}
@@
   \includegraphics[width = \textwidth]{99_ResultsMovementKernel.png}
-  \caption{}
+  \caption{Parameter estimates with regards to the movement covariates
+  \textsf{sl}, \textsf{log(sl)}, and \textsf{cos(ta)}. Dots indicate the mean
+  estimates across 100 replicates, whereas the vertical line delineate $\pm 2SD$
+  from the mean. The black dotted lines indicate the ``true'' values that were
+  used to simulate movement. Colors separate different approaches to account for
+  varying step-durations in the data. For clarity, we only show results for the
+  treatments with data missingness of 0.5.}
   \label{ResultsMovement}

### 2022-07-26 Chapter_8/04_Manuscript/Manuscript.tex
Various updates

--- 2022-07-13 Chapter_4/04_Manuscript/Manuscript.tex
@@
 %------------------------------------------------------------------------------
-\title{Applying Step Selection Functions to Data with Data Gaps: Simulation and
-Case Study on African wild dogs}
+\title{Flooding of the Okavango Delta influences Connectivity for Dispersing
+African Wild Dogs}
 
@@
 
-\textbf{Keywords:} step selection functions, gps data, animal movement, Lycaon
-pictus, irregular data
+\textbf{Keywords:} Seasonal Flooding of the Okavango Delta and its Consequences
+for African Wild Dog Connectivity
 
@@
 \begin{abstract}
-Animal movement data collected in the field is often noisy and incomplete, thus
-confronting movement ecologists with temporal irregularities in their data. Step
-selection functions have become a popular tool to analyse movement data and to
-infer a species movement or preferences, yet they are based on the assumption of
-equal temporal gaps between consecutive datapoints. To adhere to this
-requirement, researchers typically only consider bursts of consecutive
-relocations with equal step-durations, thus discarding large portions of their
-data. Here, we use simulated data to question this practice and we examine
-whether including temporally irregular can improve estimates from step-selection
-models. In addition, we explore the usefulness of multiple methods to account
-for temporal irregularity in movement data. Our results show that including
-irregular data improves model estimates substantially. By applying methods that
-account for the irregularity in the data, the precision of model estimates can
-further be improved, albeit only marginally.
+Many ecosystems experience drastic changes in environmental conditions due to
+seasonality. While such seasonal changes may drastically alter connectivity for
+endangered species, most studies represent the environment by a static set of
+spatial layers, thus ignoring seasonal changes. Here, we address this
+shortcoming and employ individual-based simulations to investigate how seasonal
+flooding of the Okavango delta influences connectivity for dispersing African
+wild dogs (\textit{Lycaon pictus}). Our results show that the Okavango delta
+poses a substantial dispersal barrier when the flood is at maximum extent, yet
+that viable dispersal corridors exist when the flood is at a minimum level.
+Despite a better understanding of the conservation needs for African wild dogs,
+our study also provides evidence that incorporating seasonality in studies of
+connectivity is imperative to more accurately predict dispersal ability of
+endangered species.
 \end{abstract}
@@
 \section{Introduction}
-GPS data is usually collected on fixed temporal intervals (could do a quick and
-dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
-even days between subsequent reolocations. Nevertheless, GPS devices regularly
-fail to adhere to the anticipated fixrate schedule due to cloud cover or other
-unfavorable conditions, thus resulting in irregular data. In other cases, the
-GPS schedule is not strictly fixed but may be subject to change. For example,
-\cite{Cozzi.2020} increased the fixrate of their GPS collars once per week from
-4 to 8 fixes per day, aiming to collect more fine scaled data for a limited
-time. We chose to coin this type of data \textit{irregular}, because it is not
-collected on the anticipated fixrate schedule and therefore not directly
-comparable to the vast majority of the collected data.
+Hello, I'm an introduction.
 
-Step selection analysis is a powerful framework that enables researchers to
-study habitat and movement preferences of their focal species. The method works
-by comparing spatial covariates at locations where the studied animal has been
-observed, to a set of locations where the animal was presumably absent. For
-this, the collected GPS data is converted into steps, where a step represents
-the straight line movement between two consecutive GPS fixes. One can then
-compare covariates experienced along observed steps to covariates along
-potential alternative steps and thus infer preferred or avoided features. By
-restricting the availability domain to a set of steps, SSA is specifically
-intended to account for spatio-temporal autocorrelation inherent to data
-collected using GPS collars and is therefore one of the preferred methods for
-analysing such data. However, it is considered good practice to only consider
-steps with similar step durations in step-selection models. Depending on the
-amount of available data and depending on the number of missing GPS fixes, this
-practice might entail little sacrifice. In some situations, however, removing
-irregular data implies a substantial loss of information. Additionally, recent
-developments in the SSA realm have brought forward several improvements to SSA
-that might allow to relax the assumption of regular step-durations. This
-includes the \textit{integrated} SSA approach presented by \citep{Avgar.2016},
-in which inference on habitat \textit{and} movement preferences are possible
-thanks to the inclusion of movement descriptors in the respective model. More
-recently, the method has been further refined by \citep{Munden.2021}, who coined
-the term \textit{time-varying integrated} SSA. This method was developed for
-high frequency data which has been rarified using a change-point detection
-algorithm, thus also resulting in irregular step-durations. In case of missing
-steps, in contrast, steps durations are not entirely random but usually
-clustered.
+\section{Methods}
 
-Here, we questioned the practice of removing any irregular data in SSA and
-investigated whether such data could be used to inform step selection models.
-For this, we conducted a simulation study where we simulated movement
-trajectories with known preferences across a virtual landscape. We then
-artifically rarified the ``observed'' data by removing GPS fixes and we analysed
-this data using SSA. This allowed us to investigate if and how the inclusion of
-irregular data affected model estimates. We also varied the degree of
-irregularity in the data and we tested for the effects of adjusting the
-parametric step-length and turning distributions to different step-durations.
-Besides the simulation study, we also analysed a dataset collected on dispersing
-African wild dogs and examined the implications considering or discarding
-irregular data.
+\subsection{Study Area}
+As depicted in \Cref{StudyArea}, the study area encomassed ...
 
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
-  \caption{}
-  \label{Covariates}
+  \includegraphics[width = \textwidth]{99_StudyArea.png}
+  \caption{Study area across which we simulated dispersal events. Virtual
+  dispersers were released at random locations within the orange source areas.}
+  \label{StudyArea}
   \end{center}
@@
 
-We hypothesized that the precision of model estimates would decrease as we
-increased the missigngess in the data. We also expected that the inclusion of
-irregular data would not improve the precision of estimates and rather lead to
-biased values. However, we anticipated that such biases, at least in estimated
-movement kernel parameters, could be reduced by appropriately adjusting the
-availability domain.
-
-\section{Methods}
-\subsection{Simulation Study}
-\subsubsection{Spatial Covariates}
-We simulated a virtual landscape comprising three spatial covariate layers, each
-with a resolution of 300 x 300 pixels (\Cref{Covariates}) spanning across x- and
-y-coordinates from 0 to 300. The first layer (\textsf{water}) represented
-water-bodies and was simulated using a random cluster nearest‐neighbour neutral
-landscape model \citep{Saura.2000}, with the cluster-proportion set to 0.5 and
-the patch occupancy of water fixed to 20\%. The second layer (\textsf{elev})
-resembled an elevation layer and was simulated using a Gaussian random fields
-neutral landscape model \citep{Schlather.2015}, with an autocorrelation range of
-10, magnitude of variation of the landscape of 1, and a magnitude of variation
-in scale of 0. We simulated both of these layers using the r-package {\tt NLMR}
-\citep{Sciaini.2018}. The third layer (\textsf{dist}) simply indicated the
-distance (in pixels) to the center of the virtual landscape \((x = 150, y =
-150)\), and can be understood as the distance to the center of an animal's
-home-range. We computed spatial distances using the r-package {\tt raster}
-\citep{Hijmans.2022}. We normalized the values from all layers to a range
-between zero and one.
+\subsection{Flood}
+In order to generate spatial layers that depict the flood extent at different
+points in time, we ... \Cref{FloodExtent}.
 
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Covariates.png}
-  \caption{Virtual landscape across which we simulated movement trajectories.
-  All layers have a resolution of 300 x 300 pixels and were generated randomly.
-  Simulated individuals were initiated within the white dashed rectangle, which
-  ensured that they would not be released directly at a map border.}
-  \label{Covariates}
+  \includegraphics[width = \textwidth]{99_FloodExtent.png}
+  \caption{}
+  \label{FloodExtent}
   \end{center}
 \end{figure}
-
-\subsubsection{Movement Simulation}
-To simulate movement across the above generated virtual lanscape, we employed an
-``inverted'' step-selection function, which proceeded as follows. First, we
-generated a random starting location. To prevent starting points in the vicinity
-of map borders, we restricted sampled locations to x- and y-coordinates between
-50 and 250 (white dotted rectangle in \Cref{Covariates}). Second, we generated a
-set of 10 random steps originating from the sampled starting point. We then
-generated random steps by sampling turning angles from a von Mises distribution
-with concentration parameter \(\kappa = 0.5\) and location parameter \( \mu = 0
-\), and step lengths from a gamma distribution with shape parameter \(k = 3 \)
-and scale parameter \(\theta = 1\). Third, we extracted covariate values along
-each random step from the underlying covariate layers and computed the average
-of each covariate along every step. Fourth, we predicted for each step the
-probability of being chosen by applying \Cref{EQ1}. The vector of relative
-habitat preferences used to make predictions was set to \(\beta_{water} = -1\),
-\(\beta_{elev} = 0.5\), and \(\beta_{dist} = -15\). Fifth, we sampled one of the
-random steps based on predicted probabilities and computed the new position of
-the simulated individual. We the repated these steps until a total of 100 steps
-were realized and we replicated the simulation 100 times, thus resulting in 100
-independent movement trajectories (\Cref{Simulations}). Note that by excluding
-interactions among habitat covariates and step metrics we implicitly assume
-independent habitat and movement kernels \citep{Avgar.2016}. Although
-measurement error would have been straight forward to implement, this was not
-the main focus of our analysis and so we assumed GPS locations to be 100 percent
-accurate.
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = 0.5\textwidth]{99_Simulations.png}
-  \caption{100 Simulated movement trajectories, each comprising of 100 steps.
-  Simulated individuals were initiated within the blue rectangle to mitigate
-  edge effects.}
-  \label{Simulations}
-  \end{center}
-\end{figure}
-
-\subsubsection{Step Selection Analysis}
-To assess the consequences of missing GPS data in when employing step-selection
-functions, we analysed the simulated GPS data under different conditions
-\Cref{Design}. Specifically, we varied the degree of data-missingness (0\% to
-80\%), the forgiveness when analysing the data (1 to 5 fixes), and the
-parametric distributions from which random steps were generated (static vs.
-dynamic distributions). We replicated each combination of conditions 100 times
-and computed averages and standard deviations from the resulting model
-estimates.
-
-In this context, missingness and forgiveness are complementary terms. While
-missingness describes the fraction of data that has not been collected, we use
-the term forgiveness to describe how much missingness a modeler is willing to
-accept. A modeler with forgiveness of one, for instance, only considers steps
-with a step-duration of one (i.e. fixes that are regularly spaced). In contrast,
-a modeler with forgiveness of two would also accept steps with a step-duration
-of two, thus pardoning a single missing fix.
-
-The idea of the multiple imputation approach is as follows...
-
-\paragraph{Data Rarefication}
-We rarefied the simulated GPS data by randomly removing a fixed fraction of GPS
-fixes. To assess the impact of different degrees of  ``missingness'', we
-incrementally increased the fraction of missing fixes from 0.0 (complete
-dataset) to 0.8 (80\% missing) with 10\% increments. The removal of GPS fixes
-introduced irregular temporal intervals between remaining fixes; we will refer
-to these intervals as ``step-times'', where the  ``step'' refers to the straight
-line movement between two consecutive GPS fixes \citep{Turchin.1998}. In the
-complete dataset, a step-time of one was assumed, whereas the step-time
-increased by one for every fix that was missing between subsequent fixes. For
-instance, a single missing fix between two other fixes directly increased the
-step time from one to two.
-
-\paragraph{Identifying Bursts and Computing Step Metrics}
-We used the generated data to compute movement bursts. A movement burst
-consisted of a sequence consecutive GPS fixes with step-times that did not
-exceed the accepted forgiveness. For instance, if the forgiveness was one,
-already a single missing GPS fix introduced a new burst. In contrast, if the
-forgiveness was set to two, step-times of two units were allowed, without
-introducing a new step.
-
-We will use the term  ``forgiveness'' to indicate how many missing fixes were
-accepted before a new burst was enforced. For instance, a forgiveness of one
-implied that a single missing fix did not result in a new burst. We then
-converted the GPS fixes belonging to the same burst into steps, where each step
-resembled the straight line movement between two GPS fixes \citep{Turchin.1998}.
-For each step we computed a set of step metrics. This includd the step length,
-its logarithm, the turning angle and its cosine.
-
-\paragraph{Step Selection Function}
-To conduct step selection analysis, we paired each observed step with 10 random
-steps. We generated random steps by sampling turning angles from a von Mises
-distribution and step lengths from a gamma distribution. Depending on the study
-design, we either sampled turning angles and step lengths from fixed or fixed or
-dynamic distributions (cfr. Section ...). Along observed and random steps, we
-extracted underlying covariates and computed their averages along steps.
-
-\paragraph{Step Length and Turning Angle Distributions}
-To generate random steps, we employed three competing approaches, coined
-\textit{uncorrected}, \textit{na\"ive}, and \textit{dynamic} approach,
-respectively. In the uncorrected approach, we sampled step-lengths and
-turning-angles from parametric gamma and von Mises distributions that were
-fitted to observed steps with step-duration of one. In the \textit{na\"ive}
-approach, we sampled step lengths and turning angles from the same
-distributions, yet we linearly scaled sampled step lengths to the actual
-step-duration. For instance, for any step with a step-duration of two, we
-doubled the sampled step length. This approach is representative of the
-time-varying SSF approach proposed by \cite{Munden.2021}. Finally, in the
-\textit{dynamic} approach, we sampled step lengths and turning angles from
-distributions that were fit to different step-durations. That is, we fitted
-separate step length and turning angles distributions to every possible step
-duration. To achieve this, we subsampled the observed data to 50\% of the fixes,
-which resulted in varying step-durations. We then and fitted step length and
-turning angle distributions to the steps from different step-durations. We
-repeated this procedure 1000 times and averaged the parameter estimates across
-replicates.
-
-\paragraph{Regression Model}
-We estimated ... using conditional logistic regression, implemented using the
-r-package \textsf{survival} \citep{Therneau.2021}. We followed \cite{Avgar.2016}
-and \cite{Fieberg.2021} and employed \textit{integrated} SSA. That is, aside
-from the habitat covariates (\textsf{water, elev, dist}), we also included
-descriptors of the step length and turning angle (\textsf{sl, log(sl), cos(ta)})
-in our regression model. The model call was as follows:
-
-$$
-case\_ \sim water + elev + dist + sl + log(sl) + cos(ta)
-$$
-
-\subsection{Case Study}
-
-because wild dogs move comparably little in the hours between xx and xx, we made
-the simplyfing assumption that the 8-hour period from xx to xx is comparable to
-a 4-hour period.
-
-\section{Results}
-\section{Discussion}
-Results from our simulation study demonstrate that the inclusion of GPS fixes
-can be used to gain information on relative habitat preferences of the focal
-species. While the inclusion of steps from irregular sampling schemes resulted
-in substantial biases for estimates related to step metrics, we showed that
-these biases can effectively be mitigated by generating step lengths and turning
-angles from distributions that are separately fit to steps from different step
-durations. Regardless of biases in beta coefficients for step metrics, we found
-that habitat selection estimates were unbiased, even when including irregular
-steps. Even more, the inclusion of irregular steps drastially reduced model
-uncertainty, thus suggesting that such data can be used to gain information that
-would otherwise be lost.
 

### 2022-07-26 Chapter_8/02_R-Scripts/01_Analysis/06_Betweenness.R

@@ -23,11 +23,10 @@ source("02_R-Scripts/00_Functions.R")
 # Load dispersal simulations
 sims <- read_rds("03_Data/03_Results/DispersalSimulation.rds")
 sims <- subset(sims, FloodLevel != "Mean")
+sims <- subset(sims, Area != 2)
 
-# Keep only columns
+# Keep only certain columns
 sims <- sims[, c("x", "y", "TrackID", "StepNumber", "Area", "FloodLevel")]
-
-sims <- subset(sims, TrackID %in% sample(sims$TrackID, 1000))
 
 # Function to retrieve the visitation history from a sequence of values
 visitHist <- function(x, singlecount = F) {
@@ -158,7 +157,7 @@ combined <- stack(maps)
 writeRaster(combined, "03_Data/03_Results/BetweennessMaps.tif", overwrite = T)
 
 # Add maps to the tibble
-design <- mutate(design, betweenness = lapply(1:nlayers(combined), function(x){
+design <- mutate(design, betweenness = lapply(1:nlayers(combined), function(x) {
   combined[[x]]
 }))
 

### 2022-08-08 Chapter_8/04_Manuscript/Manuscript.tex
Various updates

@@
 \begin{abstract}
-Many ecosystems experience drastic changes in environmental conditions due to
-seasonality. While such seasonal changes may drastically alter connectivity for
-endangered species, most studies represent the environment by a static set of
-spatial layers, thus ignoring seasonal changes. Here, we address this
+Many ecosystems experience substantial changes in environmental conditions due
+to seasonality. While such seasonal changes may drastically alter connectivity
+for endangered species, most studies represent the environment by a static set
+of spatial layers, thus ignoring seasonal variation. Here, we address this
 shortcoming and employ individual-based simulations to investigate how seasonal
@@
 our study also provides evidence that incorporating seasonality in studies of
-connectivity is imperative to more accurately predict dispersal ability of
+connectivity is imperative to more accurately predict the dispersal ability of
 endangered species.
@@
 
+However, \cite{Hofmann.2021} and \cite{Hofmann.2022} used a single floodmap...
+
 \section{Methods}
@@
 
-\subsection{Flood}
+\subsection{Spatial Habitat Layers}
 In order to generate spatial layers that depict the flood extent at different
@@
 \end{figure}
+
+\subsection{Dispersal Simulation}
+We used a previously parametrized movement model to simulate dispersal of
+African wild dogs. To simulate dispersing wild dogs, we employed the simulation
+framework presented in \cite{Hofmann.2022}. In this framework, a movement model
+that renders habitat and movement preferences of dispersing individuals is
+parametrized using step-selection functions. Once parametrized, the model can be
+employed to simulate virtual dispersers moving across the landscape.
+
+We released virtual dispersers at random locations within the four distinct
+source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
+individuals moving across the landscape at minimum flood and another xx
+individuals at maximum flood. For comparison, we also ran the simulations for a
+medium flood extent, yet the results for this will be presented in the appendix.
+
+\subsection{Connectivity}
+To gain insights into landscape connectivity from the simulated dispersal
+events, we prepared three complementary connectivity maps (see again
+\citep{Hofmann.2022}); a heatmap, indicating the absolute frequency at which
+each pixel in the study area was traversed by a virtual disperser, a betweenness
+map, highlighting pixels of particular importance to facilitate movement from
+one area to another, and a map of interpatch connectivity, depicting the
+presence and intensity of functional links between source areas. To compute the
+heatmap, we counted how many times each pixel in the study area was traversed by
+a virtual disperser. To calculate betweenness, we overlayed the study area with
+a regular grid and used the simulated trajectories to determine movements from
+one grid-cell to another. Specifically, we counted how often transitions from
+one grid-cell to any other occured. The centerpoint of each grid-cell then
+served as network node, and the number of transitions from one cell to another
+as weights between the respective nodes. Based on the so generated network, we
+then computed betweenness scores for all network nodes (i.e. for all grid-cells
+across the study area). Lastly, to compute interpatch connectivity, we
+determined the frequency at which virtual dispersers originating from one source
+area reached any of the other source areas. We also calculated the average
+dispersal duration needed until individuals reached the respective area.
+
+\section{Results}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Heatmaps2000.png}
+  \caption{}
+  \label{Heatmaps}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_DifferenceHeatmaps.png}
+  \caption{}
+  \label{DifferenceHeatmaps}
+  \end{center}
+\end{figure}
+
+\section{Discussion}
 

### 2022-08-08 Chapter_8/02_R-Scripts/01_Analysis/07_Interpatch.R

@@ -259,36 +259,3 @@ p1 <- ggplot() +
 
 # Store the plot
 ggsave("04_Manuscript/Interpatch.png", plot = p, width = 8, height = 6)
-
-# ################################################################################
-# #### Identify Direct Connections between National Parks
-# ################################################################################
-# # Note: Here we are going to look at all direct connections between national
-# # parks. That is, if a trajectory moves from A through B to C, we generate an
-# # edge between A and B and between A and C, but not between B and C.
-#
-# # Identify how long it takes on average to reach the different areas
-# visits <- sims %>%
-#   rename(From = SourceArea, To = CurrentPark) %>%
-#   group_by(TrackID, From, To) %>%
-#   summarize(
-#       StepNumber        = min(StepNumber)
-#     , DistanceFromFirst = min(DistanceFromFirst)
-#     , .groups           = "drop"
-#   ) %>%
-#   subset(!is.na(From) & !is.nan(To) & !is.na(To)) %>%
-#   arrange(TrackID, StepNumber)
-#
-# # Replace national park IDs with proper park names. Also identify the country
-# # for each of the national parks.
-# visits$FromPark    <- as.character(prot$Name[match(visits$From, prot$ID)])
-# visits$ToPark      <- as.character(prot$Name[match(visits$To, prot$ID)])
-# visits$FromCountry <- as.character(prot$Country[match(visits$From, prot$ID)])
-# visits$ToCountry   <- as.character(prot$Country[match(visits$To, prot$ID)])
-#
-# # Store visits to file
-# write_rds(visits, "03_Data/03_Results/99_InterpatchConnectivity.rds")
-#
-# # Also store the number of simulated individuals
-# write_rds(number_simulated_country, "03_Data/03_Results/99_NumberSimulatedCountry.rds")
-# write_rds(number_simulated_park, "03_Data/03_Results/99_NumberSimulatedPark.rds")

### 2022-09-05 Chapter_8/04_Manuscript/Manuscript.tex
Updated manuscript

@@
 \begin{abstract}
-Many ecosystems experience substantial changes in environmental conditions due
-to seasonality. While such seasonal changes may drastically alter connectivity
-for endangered species, most studies represent the environment by a static set
-of spatial layers, thus ignoring seasonal variation. Here, we address this
-shortcoming and employ individual-based simulations to investigate how seasonal
-flooding of the Okavango delta influences connectivity for dispersing African
-wild dogs (\textit{Lycaon pictus}). Our results show that the Okavango delta
-poses a substantial dispersal barrier when the flood is at maximum extent, yet
-that viable dispersal corridors exist when the flood is at a minimum level.
-Despite a better understanding of the conservation needs for African wild dogs,
-our study also provides evidence that incorporating seasonality in studies of
-connectivity is imperative to more accurately predict the dispersal ability of
-endangered species.
+Climate change is expected to considerably impact vital rates of endangered
+species. In particular dispersal, the phase during which species leave their
+natal area and attempt to find a new territory to settle, may
+
+We utilize seasonal changes in the flooding regime across the Okavango delta as
+a large scale natural experiment to investigate how environmental changes impact
+species' ability to disperse, and utlimately connectivity between remaining
+subpopulations. Specifically, we apply individual-based dispersal simualtions to
+simulate dispersal trajectories under two extreme scenarios, assuming minimal
+and maximum flood extents.
+
+Our results show that during years of minimum flood extent, the Okavango delta
+releases vast areas that dispersers use as movement corridors, whereas during
+years of extreme flood the delta's floodwaters pose a considerable dispersal
+barrier that forces dispersing individuals closer towards human settlements.
+Besides a better understanding of the conservation needs for the African wild
+dog, our study also provides evidence that incorporating seasonality in studies
+of connectivity is imperative to more accurately predict the dispersal ability
+of endangered species.
+
 \end{abstract}
@@
 \section{Introduction}
-Hello, I'm an introduction.
+\subsection{Climate Change}
+\subsection{Vital Rates}
+\subsection{Dispersal}
+\subsection{Connectivity}
+Dispersal is an important, if not the important, driver of landscape
+connectivity and therefore of major interest to conservation authorities.
 
-However, \cite{Hofmann.2021} and \cite{Hofmann.2022} used a single floodmap...
+\subsection{Okavango Delta and African Wild Dogs}
+The Okavango delta in Southern Africa poses a unique opportunity to study the
+impacts of environmetal change on species dispersal ability and connectivity in
+large scale natural experiment setup.
 
 \section{Methods}
+Programming language, packages, etc.
 
 \subsection{Study Area}
-As depicted in \Cref{StudyArea}, the study area encomassed ...
+The study area for this analysis was focused on the Okavango delta and its
+surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
+Zimbabwe, and Zambia \Cref{StudyArea}. To accomodate for the long distance
+dispersal events commonly observed in the African wild dog, we considered a
+large rectangular extent stretching from ... to ...., totalling to an area of xx
+km. The dominant geographical feature in this region is the Okavango delta, the
+world's largest inland delta and main driver of seasonal environmental change in
+the region. While local rainfalls marginally impact water levels across the
+delta, the main flood-pulsing rythm is driven by the water-influx from the
+Okavango river. The Okavango carries rain waters collected the Angolian
+highlands into the delta's distributaries. Although precipitation in Angola
+peaks around ..., the water arrives in the delta with a delay of ... It only
+slowly descends through the delta and reaches the distal ends (at the faults)
+around ... At minimum extent, the flood covers an area of around xx km, whereas
+during maximum flood it covers ... Vegetation in the study area can largely be
+categorized into dense mopane forest, mixed acacia woodland, and plain
+grassland. Human influence in the study area is generally low and mainly
+concentrated to small villages at the western periphery of the delta, as well as
+to the city of Maun at the southern tip of the delta. Large portions of land are
+dedicated national parks, game reserves or forest reserves and the area is part
+of the world's largest transboundary conservation initiative, the
+Kavango-Zambezi Transfrontier Conservation Area.
 
@@
 \subsection{Spatial Habitat Layers}
-In order to generate spatial layers that depict the flood extent at different
-points in time, we ... \Cref{FloodExtent}.
+We represented the physical landscape through which dispersers could move by a
+set of spatially referenced habitat layers each resolve at 250m x 250m. The set
+of layers included water-cover, distance-to-water, tree-cover,
+shrub/grassland-cover, and a human influence layer depicting anthropogenic
+influences through villages, roads, and agriculture. A detailed description of
+the different layers is provided in \Cref{Hofmann.2021}. Importantly, the
+water-cover and the derived distance-to water layers were generated using MODIS
+Terra satellite imagery, which enabled us to generate weekly updated layers that
+provided detailed information about the flood-extent at any given point in time.
+We had a total of 8xx remote sensed floodmaps at our disposal and used them to
+generate two extreme scenarios; a minimal and maximum flood scenario. To create
+the minimum flood scenario, we tallied the 50 floodmaps with smallest flood
+extent into an average image. Finally, we created a binary map... Similarly, we
+created an average image for high flood using the 50 most flooded maps. The
+resulting maps are depicted in \Cref{FloodExtent}. For completeness we also
+generated an  ``average'' floodmap by averaging across all 800 floodmaps
+available to us. The results based upond this layer are presented in the
+Appendix.
 
@@
 
+\subsection{Dispersal Model}
+Our dispersal model was based on a previously parametrized and validated
+\textit{integrated step-selection function (iSSF)} applied to the dispersal data
+of 16 dispersing African wild dogs inhabiting the surroundings of the Moremi
+National Park in the Okavango delta \citep{Hofmann.2021}. The iSSF model
+consists of two complementary ``kernels'' plus their interactions. The first
+kernel is a movement kernel and describes general movement behavior of
+dispersing AWDs, irrespective of habitat conditions. The second kernel is a
+habitat kernel and describes preferences of AWDs with regards to environmental
+conditions. Finally, the model also includes interactions among the two kernels
+and therefore allows to render how movement behavior changes depending on
+habitat conditions.
+
+\subsection{Source Areas}
+We simulated dispersing AWDs originating from nine distinct source patches
+located in the vicinity of the Okavango delta. The source areas were generated
+as follows. First, we overlayed the OD with an oval that was bound by
+geographical landmarks; in the north, the oval was bound by the Inflow of the
+Okavango river into the ``pan-handle'' of the OD, north-east the oval was bound
+by the Selinda-Spillway and the Linyanty swamp, towards South-East it was bound
+by the Boteti river, and towards South-West by Lake Ngami. We then dissected the
+polygon into five distinct patches using the same natural landmarks
+(\Cref{StudyArea}). Patch one was given by the area south of the Boteti river
+and area two by the area north of it up until the panhandle. Area three
+stretched from the region easf of Maun towards north until the Selinda-spillway,
+whereas area four stretched north of the spillway until west towards the
+panhandle. Finally, a fifrth source area marked the peninsuale in the center of
+the OD. In addition to these five source patches, we also distributed small
+peripheral patches that we used to simulate individuals immigrating
+\textit{into} system and to keep track of individuals leaving it.
+
 \subsection{Dispersal Simulation}
-We used a previously parametrized movement model to simulate dispersal of
-African wild dogs. To simulate dispersing wild dogs, we employed the simulation
-framework presented in \cite{Hofmann.2022}. In this framework, a movement model
-that renders habitat and movement preferences of dispersing individuals is
-parametrized using step-selection functions. Once parametrized, the model can be
-employed to simulate virtual dispersers moving across the landscape.
+For both environmental scenarios we simulated 1'000 individuals dispersing from
+each of the nine source area depicted in \Cref{StudyArea}. The simulation
+algorithm was based on the algorithm described in \Cref{Hofmann.2022} and works
+as follows. A random location within the source area is chosen as a starting
+point. Originating from the starting point, a set of 25 random steps are
+generated by sampling step lengths from a gamma distribution fitted to observed
+steps (what is a step?) (shape = , scale = ) and turning angles from a uniform
+distibution on ($-\pi, +\pi$). Along each random step the underlying spatial
+covariates are extracted and relevant movement metrics are computed (i.e.
+log(sl), cos(ta), ta). The parametrized dispersal model is used to predict the
+probability of each step for being chosen given the steps covariate values. One
+of the steps is sampled based on assigned probabilities and the location of the
+animal is updated. The procedure is then repated until the desired number of
+steps (in our case 2000) is realized. Resulting trajectories resemble correlated
+random walks. As the original model was trained using 4-hourly steps, a
+simulated step also resembled straight line movements within four hours.
 
-We released virtual dispersers at random locations within the four distinct
-source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
-individuals moving across the landscape at minimum flood and another xx
-individuals at maximum flood. For comparison, we also ran the simulations for a
-medium flood extent, yet the results for this will be presented in the appendix.
+% We used a previously parametrized dispersal model to simulate dispersal
+% trajectories of African wild dogs across the study area. To simulate dispersing
+% wild dogs, we employed the simulation framework presented in
+% \cite{Hofmann.2022}. In this framework, a movement model that renders habitat
+% and movement preferences of dispersing individuals is parametrized using
+% step-selection functions. Once parametrized, the model can be employed to
+% simulate virtual dispersers moving across the landscape.
+%
+% We released virtual dispersers at random locations within the four distinct
+% source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
+% individuals moving across the landscape at minimum flood and another xx
+% individuals at maximum flood. For comparison, we also ran the simulations for a
+% medium flood extent, yet the results for this will be presented in the appendix.
 
 \subsection{Connectivity}
+Based on simulated dispersal trajectories we quantified connectivity across the
+OD for both environmental scenaiors in three complementary way. First, we
+generated heatmaps showing how often different areas in the landscape were
+visited by simulated dispersers. Such heatmaps serve to detect dispersal
+hotspots, yet are less suitable to detect pinchpoints and bottlenecks. Hence, we
+also computed spatially explicit betweenness scores which are useful to
+highlight movement corridors (Bastille). For this, we overlayed the study area
+with a regular grid with 2.5 km x 2.5 km grid cells and determined how often
+simulated individuals transitioned from one grid-cell to another. Note that we
+didn't consider revisits to avoid pronouncing regions in which individuals were
+encroached and moved in circles.
+
 To gain insights into landscape connectivity from the simulated dispersal
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Heatmaps2000.png}
+  \includegraphics[width = \textwidth]{99_Heatmaps.png}
   \caption{}
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_DifferenceHeatmaps.png}
+  \includegraphics[width = \textwidth]{99_DifferenceMap.png}
   \caption{}

### 2022-09-05 Chapter_8/02_R-Scripts/01_Analysis/01_Setup.R

@@ -131,6 +131,15 @@ writeRaster(tre, file.path(new_chap, "03_Data/01_RawData/MODIS/Trees.tif"), over
 writeRaster(gra, file.path(new_chap, "03_Data/01_RawData/MODIS/Shrubs.tif"), overwrite = T)
 
 ################################################################################
+#### Dispersal Data
+################################################################################
+# Copy data of the dispersers
+file.copy(
+    from = file.path(old_chap, "03_Data/02_CleanData/00_General_Dispersers_POPECOL.csv")
+  , to   = file.path(new_chap, "03_Data/02_CleanData/Dispersers.csv")
+)
+
+################################################################################
 #### Reference Layers
 ################################################################################
 # Print to terminal
@@ -168,6 +177,7 @@ vills <- subset(vills, !is.na(place)) %>%
 
 # Make a regular shapefile
 st_write(vills, "03_Data/02_CleanData/Villages.shp", delete_layer = T)
+
 
 ################################################################################
 #### Session Information

### 2022-09-05 Chapter_8/02_R-Scripts/01_Analysis/07_Interpatch.R

@@ -33,10 +33,10 @@ area <- vect("03_Data/02_CleanData/SourceAreas.shp")
 # Rasterize them to the reference raster
 area_r <- terra::rasterize(area, y = r, field = "ID")
 
-# Visualize them
-plot(area_r, main = "Source Areas")
-plot(area, add = T)
-text(area, "ID", cex = 0.5, halo = T)
+# # Visualize them
+# plot(area_r, main = "Source Areas")
+# plot(area, add = T)
+# text(area, "ID", cex = 0.5, halo = T)
 
 ################################################################################
 #### Prepare Simulations
@@ -46,6 +46,14 @@ sims <- read_rds("03_Data/03_Results/DispersalSimulation.rds")
 
 # Keep only desired columns
 sims <- sims[, c("x", "y", "TrackID", "StepNumber", "SourceArea", "FloodLevel")]
+
+# Count number of simulations per source area
+n_sample <- sims %>%
+  select(TrackID, SourceArea, FloodLevel) %>%
+  distinct() %>%
+  count(SourceArea, FloodLevel) %>%
+  pull(n) %>%
+  unique()
 
 # Make coordinates of simulated trajectories spatial
 coordinates(sims) <- c("x", "y")
@@ -94,8 +102,8 @@ visits <- visits %>%
 
 # Identify how long it takes to reach the different areas
 visits <- visits %>%
-  rename(From = SourceArea, To = Area) %>%
-  group_by(TrackID, FloodLevel, From, To) %>%
+  rename(SourceArea = SourceArea, CurrentArea = Area) %>%
+  group_by(TrackID, FloodLevel, SourceArea, CurrentArea) %>%
   summarize(
       StepNumber        = min(StepNumber)
     , DistanceFromFirst = min(DistanceFromFirst)
@@ -107,7 +115,7 @@ visits <- visits %>%
 # Compute summary statistics by source area
 summarizeVisits <- function(visits) {
   visits %>%
-    group_by(FloodLevel, From, To) %>%
+    group_by(FloodLevel, SourceArea, CurrentArea) %>%
     summarize(
         MeanStepNumber = mean(StepNumber)
       , SDStepNumber   = sd(StepNumber)
@@ -119,25 +127,12 @@ summarizeVisits <- function(visits) {
 # Try it
 summarizeVisits(visits)
 
-# This is cool. Let's now use the function in a bootstrapping approach to get
[…]

### 2022-09-05 Chapter_8/02_R-Scripts/01_Analysis/08_ExtentFlood.R

@@ -1,0 +1,41 @@
+################################################################################
+#### Computing the Flood Extent
+################################################################################
+# Description: Computing the flood extent of the Okavango Delta at minimum and
+# minimum flood
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
+setwd(wd)
+
+# Change the working directory
+library(terra)
+
+# Load the watermaps
+dat <- rast("03_Data/02_CleanData/WaterCover.tif")
+dat <- dat[[c("min", "max")]]
+
+# We only want to focus on the Okavango delta, thus, let's load a shapefile of
+# it
+oka <- vect("03_Data/02_CleanData/MajorWaters.shp")
+oka <- oka[oka$name == "Okavango Delta", ]
+
+# Buffer it slightly
+oka <- buffer(oka, width = 7500)
+oka <- fillHoles(oka)
+
+# Let's take a look at the overlap
+plot(dat[["max"]])
+plot(oka, add = T)
+
+# This looks good. We now mask anything outside that polygon
+dat <- mask(dat, oka)
+dat <- trim(dat)
+dat <- subst(dat, 0, NA)
+
+# Now let's compute the total area covered by the flood in the two cases
+areas <- expanse(dat, unit = "km")
+plot(dat, col = "cornflowerblue", main = paste0(names(dat), ": ", round(areas), " km2"))

### 2022-09-13 Chapter_4/02_R-Scripts/03_CaseStudy.r
Minor updates

@@ -1,621 +1,0 @@
-################################################################################
-#### Integrated Step Selection Analysis of African Wild Dog Data
-################################################################################
-# Description: Use ISSF analysis estimate preferences on the simulated data
-
-# Clear R's brain
-rm(list = ls())
-
-# Load required packages
-library(raster)        # To handle spatial data
-library(terra)         # To handle spatial data
-library(pbmcapply)     # For parallel computing
-library(tidyverse)     # For data wrangling
-library(survival)      # To run conditional logistic regression
-library(lubridate)     # To handle dates
-library(amt)           # To fit distributions
-library(maptools)      # Access to spatstat functions
-library(spatstat)      # Access to spatstat functions
-library(glmmTMB)       # To fit mixed effects clogit
-
-# Set working directory
-setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
-
-# Load custom functions
-source("02_R-Scripts/00_Functions.r")
-
-# Suppress scientific notation
-options(scipen = 999)
-
-################################################################################
-#### GPS Data Cleaning
-################################################################################
-# Load observed movement data of dispersing wild dogs and nest data by ID
-obs <- "03_Data/01_RawData/WildDogs.csv" %>%
-  read_csv() %>%
-  subset(State == "Disperser") %>%
-  mutate(TimestampRounded = round_date(Timestamp, "1 hour")) %>%
-  group_by(DogName) %>%
-  rename(ID = DogName) %>%
-  nest()
-
-# Resample the GPS data to 1 hour
-obs$data <- suppressMessages(
-  pbmclapply(1:nrow(obs)
-    , ignore.interactive = T
-    , mc.cores           = detectCores() - 1
-    , FUN                = function(x){
-      resFix(obs$data[[x]], hours = 1, start = 1, tol = 0.5)
-    }
-  )
-)
-
-# Unnest and keep only fixes on the attempted schedule
-obs <- obs %>%
-  unnest(data) %>%
-  subset(hour(TimestampRounded) %in% c(3, 7, 15, 19, 23)) %>%
-  ungroup()
-
-# Determine "step_number" and "step_ids"
-obs <- obs %>%
-  mutate(step_id = row_number()) %>%
-  group_by(ID) %>%
-  mutate(step_number = row_number()) %>%
-  ungroup()
-
-# Let's define an extent from the GPS locations
[…]

### 2022-09-13 Chapter_8/02_R-Scripts/02_Visualization/Interpatch.Rmd
Major update. Adjusted generation of source areas.

---
title: "Interpatch"
author: "David Hofmann"
date: "2022-09-06"
output: pdf_document
---
This is just a test document to prepare the table of inter-patch connectivity.

```{r, echo = F, warning = F, message = F}
# Clear R's brain
rm(list = ls())

# Load required packages
library(tidyverse)    # To wrangle data
library(kableExtra)   # To generate a nice table

# Load results on interpatch connectivity
dat <- read_rds("/home/david/ownCloud/University/15. PhD/Chapter_8/03_Data/03_Results/BootstrappedInterpatchConnectivity.rds")
dat <- subset(dat, FloodLevel != "Mean")

################################################################################
#### Frequency
################################################################################
# Reshape the data
dat_arr_freq <- dat %>%
  select(-c(StepNumber, StepNumberSE)) %>%
  mutate(Combined = paste(round(dat$Freq), "$\\pm$", sprintf("%.2f", round(dat$FreqSE, 2)))) %>%
  select(-c(Freq, FreqSE)) %>%
  pivot_wider(names_from = SourceArea, values_from = Combined) %>%
  arrange(CurrentArea, desc(FloodLevel)) %>%
  select(To = CurrentArea, everything()) %>%
  select(-FloodLevel)

# Make a nice table
table <- kbl(dat_arr_freq, booktabs = T) %>%
  # add_header_above(c("", "1", "2", "3", "4", "5", "6", "7", "8", "9")) %>%
  # add_header_above(c("", "Main" = 5, "Buffer" = 4)) %>%
  add_header_above(c("", "From" = 9)) %>%
  column_spec(1, border_right = T) %>%
  collapse_rows(1, latex_hline = c("custom"), custom_latex_hline = 1)

```

### 2022-09-13 Chapter_8/04_Manuscript/Manuscript.tex
Major update. Adjusted generation of source areas.

@@
 \begin{abstract}
-Climate change is expected to considerably impact vital rates of endangered
-species. In particular dispersal, the phase during which species leave their
-natal area and attempt to find a new territory to settle, may
+Climate change is expected to profoundly impact vital rates of wild living
+animal populations. While the impact of climate change on the demographics of
+local subpopulations has been studied repeatedly, little is known about the
+consequences of environmental change on dispersal and connectivity between such
+subpopulations.
 
-We utilize seasonal changes in the flooding regime across the Okavango delta as
-a large scale natural experiment to investigate how environmental changes impact
-species' ability to disperse, and utlimately connectivity between remaining
-subpopulations. Specifically, we apply individual-based dispersal simualtions to
-simulate dispersal trajectories under two extreme scenarios, assuming minimal
-and maximum flood extents.
+Here, we capitalize on natural, flood-pulse driven environmental change across
+the Okavando Deltas in northern Botswana to investigate the impact of changing
+environmental conditions on the dispersal ability of the endangered African wild
+dog. For this, we utilize a previously parametrized dispersal model and apply it
+to simulate dispersal under two extreme environmental scenarios; one assuming
+minimal flood extent, one assuming maximum flood extent.
 
-Our results show that during years of minimum flood extent, the Okavango delta
-releases vast areas that dispersers use as movement corridors, whereas during
-years of extreme flood the delta's floodwaters pose a considerable dispersal
-barrier that forces dispersing individuals closer towards human settlements.
-Besides a better understanding of the conservation needs for the African wild
-dog, our study also provides evidence that incorporating seasonality in studies
-of connectivity is imperative to more accurately predict the dispersal ability
-of endangered species.
+In genral, our results suggest only a marginal impact of the flood extent on
+dispersal prospects. Even during maximum flood, dispersal between neighboring
+patches remained high. Between a few selected patches, however, dispersal was
+substantially restricted during times of high flood. That is the Okavango delta
+posed a substantial dispersal barrier that reduced connectivity between patches
+during maximum flood, but revealed vital dispersal corridors that facilitated
+dispersal into neighboring areas during minimum flood. Besides a better
+understanding of the conservation needs for the African wild dog, our study also
+presents a first step towards incorporating environmental change due to
+seasonality or climate change. It also emphasizes the usefulness of
+individual-based dispersal simulations as a conservation tool to study impacts
+of environmental change in a spatially realistic way.
 
@@
 \subsection{Climate Change}
+
+Despite the importance of climate change in determining species viability,
+predicting the way and extent to which it will affect species remains
+challenging. One exciting avenue to gain insights into the implications of
+environmental change is to utilize seasonal variation in the environment.
+
+Cite IPCC, Planetary boundaries, Tucker (step length). Phenological shifts. Cite
+Robynne, Arpat, Megan, Dominik, Rabaiotti etc.
+
 \subsection{Vital Rates}
+Several studies investigating the impacts of climate change on life-history
+traits have revealed that population persistence may critically depends on
+climate conditions \citep{Paniw.2019}. While many studies examine climate
+impacts on survival and reproductive rates, the influence of global change or
+climatic extreme events on dispersal prospects has received little attention.
+
+Ecosystems that undergoe seasonal environmental variation, in particular arid
+regions that experience extreme climatic conditions, are among the most
+vulnerable to climate change.
+
 \subsection{Dispersal}
@@
 impacts of environmetal change on species dispersal ability and connectivity in
-large scale natural experiment setup.
+large scale natural experiment setup. According to projections by the IPCC,
+Botswana and its surroundings are among the most vulnerable to climate change.
 
 \section{Methods}
-Programming language, packages, etc.
+We conducted all analyses using the programming language R.
 
@@
 surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
-Zimbabwe, and Zambia \Cref{StudyArea}. To accomodate for the long distance
-dispersal events commonly observed in the African wild dog, we considered a
-large rectangular extent stretching from ... to ...., totalling to an area of xx
-km. The dominant geographical feature in this region is the Okavango delta, the
-world's largest inland delta and main driver of seasonal environmental change in
-the region. While local rainfalls marginally impact water levels across the
-delta, the main flood-pulsing rythm is driven by the water-influx from the
-Okavango river. The Okavango carries rain waters collected the Angolian
-highlands into the delta's distributaries. Although precipitation in Angola
-peaks around ..., the water arrives in the delta with a delay of ... It only
-slowly descends through the delta and reaches the distal ends (at the faults)
-around ... At minimum extent, the flood covers an area of around xx km, whereas
-during maximum flood it covers ... Vegetation in the study area can largely be
-categorized into dense mopane forest, mixed acacia woodland, and plain
-grassland. Human influence in the study area is generally low and mainly
-concentrated to small villages at the western periphery of the delta, as well as
-to the city of Maun at the southern tip of the delta. Large portions of land are
-dedicated national parks, game reserves or forest reserves and the area is part
-of the world's largest transboundary conservation initiative, the
-Kavango-Zambezi Transfrontier Conservation Area.
+Zimbabwe, and Zambia (\Cref{StudyArea}). To accommodate for the long distance
+dispersal events commonly observed in African wild dogs (e.g.
+\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}), we considered a large
+rectangular extent stretching from 20\degree 30' E to 26\degree E. totaling to
+an area of 300'000 km\textsuperscript{2}. The dominant geographical feature in
+this region is the Okavango delta (OD), the world's largest inland delta and the
+main driver of seasonal environmental change in the region. The flood-pulsing
+rhythm of the OD is mainly dictated by precipitation in the Angolian highlands,
+where rainfalls are channeled into the OD through the Okavango River. Although
+precipitation in Angola peaks in ..., water only slowly descends through the
+Okavango river and its tributaries, reaching the distal ends (i.e. the faults)
+of the delta in July or August. At minimum extent, the flood covers an area of
+3'600 km\textsuperscript{2}, whereas during maximum flood it covers more than
+9'000 km\textsuperscript{2}. Vegetation in the study area is dominated by dense
+mopane forest, mixed acacia woodland, and plain grassland. Human influence in
+the study area is low and mainly concentrated around small villages at the
+western periphery of the delta and to the city of Maun at the southern tip of
+the delta. Large portions of land are dedicated national parks, game reserves or
+forest reserves. The study area is also part of the world's largest
+transboundary conservation initiative, the Kavango-Zambezi Transfrontier
+Conservation Area.
 
@@
 \subsection{Connectivity}
-Based on simulated dispersal trajectories we quantified connectivity across the
-OD for both environmental scenaiors in three complementary way. First, we
-generated heatmaps showing how often different areas in the landscape were
-visited by simulated dispersers. Such heatmaps serve to detect dispersal
-hotspots, yet are less suitable to detect pinchpoints and bottlenecks. Hence, we
-also computed spatially explicit betweenness scores which are useful to
-highlight movement corridors (Bastille). For this, we overlayed the study area
-with a regular grid with 2.5 km x 2.5 km grid cells and determined how often
-simulated individuals transitioned from one grid-cell to another. Note that we
-didn't consider revisits to avoid pronouncing regions in which individuals were
-encroached and moved in circles.
-
-To gain insights into landscape connectivity from the simulated dispersal
-events, we prepared three complementary connectivity maps (see again
-\citep{Hofmann.2022}); a heatmap, indicating the absolute frequency at which
-each pixel in the study area was traversed by a virtual disperser, a betweenness
-map, highlighting pixels of particular importance to facilitate movement from
-one area to another, and a map of interpatch connectivity, depicting the
-presence and intensity of functional links between source areas. To compute the
-heatmap, we counted how many times each pixel in the study area was traversed by
-a virtual disperser. To calculate betweenness, we overlayed the study area with
-a regular grid and used the simulated trajectories to determine movements from
-one grid-cell to another. Specifically, we counted how often transitions from
-one grid-cell to any other occured. The centerpoint of each grid-cell then
-served as network node, and the number of transitions from one cell to another
-as weights between the respective nodes. Based on the so generated network, we
-then computed betweenness scores for all network nodes (i.e. for all grid-cells
-across the study area). Lastly, to compute interpatch connectivity, we
-determined the frequency at which virtual dispersers originating from one source
-area reached any of the other source areas. We also calculated the average
-dispersal duration needed until individuals reached the respective area.
+Based on simulated dispersal trajectories in the two scenarios we quantified
+connectivity in three complementary ways. First, we generated heatmaps depicting
+the frequency at which different areas in the landscape were visited by
+simulated dispersers. Such heatmaps serve to detect dispersal hotspots but are
+less suitable to detect pinchpoints and bottlenecks. Hence, we also computed
+spatially explicit betweenness scores which are useful to highlight such
+pinchpoints (Bastille). To compute betweenness, we overlayed the study area with
+a regular grid with 2.5 km x 2.5 km grid cells and determined how often
+simulated individuals transitioned from one grid-cell to another. In cases where
+the same individual repeatedly realized the same cell-transition, we only
+counted a single transition to avoid emphasizing regions regions in which
+individuals moved in circles. Finally, we calculated the number of successful
+dispersal events between the different source areas. We coin this type of
+connectivity ``inter-patch connectivity''. Dispersal between two areas was said
+to be successful whenever a trajectory leaving one area intersected with the
+target area. To gauge the dispersal duration needed to move between patches (be
+consistent with ``patches'', ``source-areas'' etc.), we also recorded the
+minimum number of steps that individuals had to move to move between areas.
 
 \section{Results}
+Heatmaps produced from simulated dispersal trajectories reveal that the OD acts
+as a major dispersal barrier during periods of high flood, but reveals viable
+dispersal corridors during periods of low flood. In particular, our analysis
+reveals that the area north-west of Maun is frequently visited by dispersers at
+in the minimum-flood scenario, yet that an extended flood, coupled with high
+human influence around the city of Maun, enforces simulated individuals to
+either detour around Maun, or to slip through the narrow window between village
+and flood. Similarly, the betweenness map indicates the presence of a dispersal
+corridor extending from the western section of the delta leading across the
+delta's center into its western part, yet only during low-flood. During maximum
+flood, the same corridor is pushed south and enforces dispersers to move closer
+to the city of Maun. In general, connectivity during maxium flood is lower
+compared to the minimum flood scenario (number links vs. number liks). Most
+notably, source area 5 is almost inaccessible during maximum flood.
+
+ Most notably, the
+area north-west of Maun acts as dispersal habitat during phases of low flood,
+yet is almost inaccessible during phases of maxium flood. This is reflected in
+the heatmap, where connectivity across the western part of the delta is low
+during maximum flood, but also in the betweenness map, where an important
+movement corridors through the delta is only visible during minimum flood.
+Although connectivity across large parts of the landscape remains unchanged
+across environmental scenarios, one critical movement corridor shifts in
+response to an expanding flood. Most notably, a corridor that links the western
+and eastern part of the delta during minimum flood gets redirected further
+south, thus leading more closely to the city of Maun. Unsurprisingly, the
+differences are more pronounced for areas directly affected by the flood, yet it
+is worth noting that even some areas not directly affected by the flood.
 
@@
 
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Betweenness.png}
+  \caption{}
+  \label{Betweenness}
+  \end{center}
+\end{figure}
+
+\input{IPC_Frequency.tex}
+\input{IPC_Duration.tex}
+
 \section{Discussion}
+According to our simulations, the propensity to move between the eastern and
+western part of the delta is much lower during maximum extent. This is mainly
+due to the flood-waters and the city of Maun acting as dispersal barriers.
+During maximum extent, the floodwaters of the delta close a gap between the
+delta and Maun that otherwise would serve as dispersal corridor. Anecdotal
+evidence supports this hypothesis, for the only dispersing individuals recorded
+to move from the eastern to the western part of the delta moved at times of low
+flood. In line with this, it appears that a large flood extent pushes dispersing
+individuals to move closer to human inhabited areas such as the village of Maun.
+
+Predicting how climate change will impact the dispersal ability of AWDs is
+challenging for multiple reasons. First of all, predicting the flooding patterns
+of the OD under climate change is merely impossible due to the complex feedbacks
+between surface-temperature, soil conditions, precipitation patterns and the
+associated changes in vegetation. Second, the delta is not only prone to changes
+in environmental conditions, but also to changes in anthropogenic use of the
+inflowing water. Finally, it is unclear how AWDs, in fact, how any
+species, will cope with environmental change due to global warming. Even though
+some studies predict that AWD populations are likely to decline under increasing
+temperatures, these studies fail to account for the behavioral plasticity of
+their focal species. AWDs respiratory system, for instance, has evolved as a
+perfect adaptation to high temperatures and AWDs may, in fact, profit from a
+comparative advantage (cite an economist) over their competitors and prey under
+rising temperatures. Although the theory of comparative advantages is a
+fundamental concept in economics, it has yet to find its way into ecological
+studies.
+
+\cite{MurrayHudson.2006} predicted that increased temperatures, additional human
+abstractions, and reduced river flows might lead to a ``Delta dying'' and that
+the impact of climate will be much more pronounced than the impact of
+anthropogenic water use.
+
+Although local rainfalls in Botswana are expected to increase in terms of
+intensity, simulations show that the length of the rainy season will decline,
+more than offsetting the incline in precipitation \cite{Akinyemi.2019}.
+
+According to our simulations, dispersers are able to cover larger distances
+during periods of low flood. This finding is little surprising, considering that
+inundated areas act as dispersal barriers and force dispersers to detour and
+circumvent water-covered areas. However, it still leads to an interesting
+hypothesis. Previous studies have shown that the euclidean dispersal distance of
+female coalitions is larger than that of male coalitions. This has led to ...
+However, demographic analyses have also revealed that female offspring tend to
+emigrate from their pack at younger ages and earlier in the year, when
+floodwaters are still at a relatively low level (Behr). It is thus conceivable
+that the sex-differences in reported dispersal distances is mainly a consequence
+of environmental conditions during dispersal, rather than owed to physiological
+differences between sexes.
+
+While our analysis marks an important step into incorporating environmental
+change into studies of connectivity, there are several critical additions that
+should be considered in the future. We studied dispersal and connectivity under
+two different environmental scenarios, yet our movement model assumed that
+dispersers had identical habitat and movement preferences in both scenarios. In
+reality, however, it can be expected that movement and habitat kernels of
+dispersers differ depending on the season considered (examples).
+
+An additional complication arises when species movement is not solely driven by
+environmental conditions, but also affected by itra- and inter-specific factors.
+For instance, ... has shown that dispersers...
+Rendering such conditions alone is challenging, yet rendering the conditions
+under changing environmental conditions is merely impossible.
+
+To address such differences, researchers could model
+habitat and movement preferences using season-dependent models, or,
+alternatively, by combining hidden markov models with step-selection functions.
+(cite papers that fieberg sent)
+
+Only recently, it has been discovered that AWDs communicate using shared marking
+sites. The role of such marking sites for dispersing coalitions remains to be
+investigated, yet it is is likely that, akin to resident packs, use SMS as
+navigation waypoint and demarcation lines. Chemical analyses suggest that the
+compounds used for communication are highly volatile and may not persist in
+extreme climate conditions. In result, dispersers may lose their ability to
+effectively navigate across the landscape and locate potential mates with whom
+to settle. This would reduce pack-formation prospects and undermine...
+
+Validating predictions from individual-based dispersal models is challenging and
+requires additional dispersal data, which is inherently scarce anyways. Sciticen
+science may help to fill this gap by augmenting observed GPS data with
+occasional sightings of uncollared dispersing coalitions. This is especially
+ciritical for species that disperse across borders and beyond confined study
+areas. The African carnivore wildbook offers ...
+
+The OD is an important driver of species distribution and it has been found that
+an expanding flood limits available habitat, thus leading to more inter-specific
+competition, particularly between AWDs and lions.
+
+\subsection{Conclusion}
+Our dispersal simulations across two extreme climatic scenarios reveal striking
+differences in dispersal prospects and landscape connectivity for dispersing
+AWDs. This implies that (1) climatic variation, be it due to seasonality or
+climate change, must be included in anlayses dealing with dispersal and (2) that
+projected climate change is likely to have profound impacts on landscape
+connectivity.
 

### 2022-09-13 Chapter_8/02_R-Scripts/01_Analysis/03_SourceAreas.R

@@ -8,7 +8,6 @@ rm(list = ls())
 
 # Change the working directory
 setwd("/home/david/ownCloud/University/15. PhD/Chapter_8")
-# setwd("C:/Users/david/switchdrive/University/15. PhD/Chapter_8")
 
 # Load required packages
 library(raster)         # To handle spatial data
@@ -17,96 +16,88 @@ library(terra)          # To handle spatial data
 # Load custom functions
 source("02_R-Scripts/00_Functions.R")
 
-# We'll use geographical, in particular hydrological features to separate source
-# areas
-river <- vect("03_Data/02_CleanData/MajorRivers.shp")
-river <- river[-2, ]
-water <- vect("03_Data/02_CleanData/MajorWaters.shp")
-water <- water[c(1, 2, 4), ]
-roads <- vect("03_Data/02_CleanData/Roads.shp")
+# Load some landmarks according to which we will delineate the source areas
+wat <- vect("03_Data/02_CleanData/MajorWaters.shp")
+wat <- wat[c(1, 2, 4), ]
+roa <- vect("03_Data/02_CleanData/Roads.shp")
+riv <- vect("03_Data/02_CleanData/MajorRivers.shp")
+riv <- riv[-2, ]
 
-# Buffer polygons slightly
-river_buff <- buffer(river, width = 2000)
-water_buff <- buffer(water, width = 500)
+# Buffer slightly
+wat_buff <- buffer(wat, width = 500)
+wat_buff <- makeValid(wat_buff)
 
-# Simplify geometries
-river_buff <- aggregate(river_buff)
-water_buff <- aggregate(water_buff)
+# Specify centroids for the different source areas
+src <- data.frame(
+    x  = c(22.7, 23.75, 23.4, 22.7, 21.8)
+  , y  = c(-20.1, -19.4, -18.8, -18.3, -19.2)
+  , ID = 1:5
+)
+src <- vect(as.matrix(src[, c("x", "y")]), crs = crs(wat), atts = src)
+src <- buffer(src, width = 20000)
 
-# Make sure geometries are valid
-river_buff <- makeValid(river_buff)
-water_buff <- makeValid(water_buff)
+# Let's also generate target zones. One target zone is the center of the delta,
+# the remaining targets are placed on an ellipse around the delta. We'll make
+# use of an ellipse to generate said areas
[…]

### 2022-09-13 Chapter_8/02_R-Scripts/01_Analysis/07_Interpatch.R

@@ -1,156 +1,0 @@
-################################################################################
-#### Interpatch Connectivity
-################################################################################
-# Description: Computing interpatch connectivity from simulated dispersal paths
-
-# Clear R's brain
-rm(list = ls())
-
-# Change the working directory
-setwd("/home/david/ownCloud/University/15. PhD/Chapter_8")
-
-# Load required packages
-library(raster)         # To handle spatial data
-library(terra)          # To handle spatial data
-library(tidyverse)      # To wrangle data
-library(lubridate)      # To handle dates
-library(pbmcapply)      # To run stuff in parallel
-library(igraph)         # For network analysis
-library(rgeos)          # To manipulate spatial objects
-library(ggnetwork)      # To plot network using ggplot
-library(sf)             # To plot spatial features
-library(ggspatial)      # To add scale bars etc to plots
-
-# Load custom functions
-source("02_R-Scripts/00_Functions.R")
-
-# Load reference raster
-r <- rast("03_Data/02_CleanData/ReferenceRaster.tif")
-
-# Load source areas
-area <- vect("03_Data/02_CleanData/SourceAreas.shp")
-
-# Rasterize them to the reference raster
-area_r <- terra::rasterize(area, y = r, field = "ID")
-
-# # Visualize them
-# plot(area_r, main = "Source Areas")
-# plot(area, add = T)
-# text(area, "ID", cex = 0.5, halo = T)
-
-################################################################################
-#### Prepare Simulations
-################################################################################
-# Load dispersal simulations
-sims <- read_rds("03_Data/03_Results/DispersalSimulation.rds")
-
-# Keep only desired columns
-sims <- sims[, c("x", "y", "TrackID", "StepNumber", "SourceArea", "FloodLevel")]
-
-# Count number of simulations per source area
-n_sample <- sims %>%
-  select(TrackID, SourceArea, FloodLevel) %>%
-  distinct() %>%
-  count(SourceArea, FloodLevel) %>%
-  pull(n) %>%
-  unique()
-
-# Make coordinates of simulated trajectories spatial
-coordinates(sims) <- c("x", "y")
-crs(sims) <- CRS("+init=epsg:4326")
-
-# Identify through which national parks the dispersers moved
[…]

### 2022-09-14 Chapter_8/Backups/02_Visualization/StudyArea.R
Various Updates

@@ -1,0 +1,296 @@
+################################################################################
+#### Study Area
+################################################################################
+# Description: Plots of the study area
+
+# Clear R's brain
+rm(list = ls())
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(raster)       # To handle raster files
+library(sf)           # To handle spatial data
+library(ggplot2)      # For nice plots
+library(ggspatial)    # To add scale bars and north arrows
+library(ggpubr)       # To extract plot legends
+library(grid)         # To arrange plots
+library(gridExtra)    # To arrange plots
+library(gtable)       # To arrange plots
+library(RColorBrewer) # For colors
+library(ggpattern)    # For hatched fills
+
+# Load custom functions
+source("02_R-Scripts/00_Functions.R")
+
+# Get the reference raster and derive an extent from it
+r <- raster("03_Data/02_CleanData/ReferenceRaster.tif")
+extent <- as(extent(r), "SpatialPolygons")
+crs(extent) <- CRS("+init=epsg:4326")
+extent <- st_as_sf(extent)
+r <- as.data.frame(r, xy = T)
+
+# Load stuff that we would like to plot
+africa <- read_sf("03_Data/02_CleanData/Africa.shp")
+prot   <- read_sf("03_Data/02_CleanData/Protected.shp")
+water  <- read_sf("03_Data/02_CleanData/MajorWaters.shp")
+river  <- read_sf("03_Data/02_CleanData/MajorRivers.shp")
+areas  <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
+roads  <- read_sf("03_Data/02_CleanData/Roads.shp")
+fault  <- read_sf("/home/david/ownCloud/University/15. PhD/Chapter_1/03_Data/01_RawData/DAVID/Faults.shp")
+vills  <- read_sf("03_Data/02_CleanData/Villages.shp")
+vills  <- cbind(st_drop_geometry(vills), st_coordinates(vills)) %>%
+  rename(x = X, y = Y)
+
+# I also want to add the Okavango river
[…]

### 2022-11-03 Chapter_8/04_Manuscript/Manuscript.tex
Various Updates
Updates addressing reviewer comments

@@
   David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
+  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
   John W. McNutt\textsuperscript{2} \and
   Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
+  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
 }
@@
 
-\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)
+\textsuperscript{\S} Corresponding author: david.hofmann2@uzh.ch
 
@@
 
-\textbf{Running Title:} None
+\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
+Consequences for African Wild Dog Dispersal and Connectivity
 
@@
 
-\textbf{Keywords:} Seasonal Flooding of the Okavango Delta and its Consequences
-for African Wild Dog Connectivity
+\textbf{Keywords:} movement, connectivity, seasonality, dispersal, conservation,
+permeability
 
@@
 \begin{abstract}
-Climate change is expected to profoundly impact vital rates of wild living
+Climate change is expected to profoundly impact the life history of wild-living
 animal populations. While the impact of climate change on the demographics of
 local subpopulations has been studied repeatedly, little is known about the
-consequences of environmental change on dispersal and connectivity between such
-subpopulations.
+consequences of environmental change on dispersal and connectivity.
 
-Here, we capitalize on natural, flood-pulse driven environmental change across
-the Okavando Deltas in northern Botswana to investigate the impact of changing
-environmental conditions on the dispersal ability of the endangered African wild
-dog. For this, we utilize a previously parametrized dispersal model and apply it
-to simulate dispersal under two extreme environmental scenarios; one assuming
-minimal flood extent, one assuming maximum flood extent.
+We capitalize on a ``natural experimental setup'', the flood-pulse driven
+environmental change across the Okavango Delta in northern Botswana, to
+investigate the impact of changing environmental conditions on dispersal
+patterns and connectivity of the endangered African wild dog (\textit{Lycaon
+pictus}). For this, we simulate dispersal trajectories across the landscape
+under two extreme environmental scenarios; one assuming a maximum flood extent,
+one assuming a minimum flood extent.
 
-In genral, our results suggest only a marginal impact of the flood extent on
-dispersal prospects. Even during maximum flood, dispersal between neighboring
-patches remained high. Between a few selected patches, however, dispersal was
-substantially restricted during times of high flood. That is the Okavango delta
-posed a substantial dispersal barrier that reduced connectivity between patches
-during maximum flood, but revealed vital dispersal corridors that facilitated
-dispersal into neighboring areas during minimum flood. Besides a better
-understanding of the conservation needs for the African wild dog, our study also
-presents a first step towards incorporating environmental change due to
-seasonality or climate change. It also emphasizes the usefulness of
-individual-based dispersal simulations as a conservation tool to study impacts
-of environmental change in a spatially realistic way.
+During maximum flood, we observe a reduction in connectivity and an increase in
+dispersal durations between distinct habitat patches. Notably, dispersal into
+the center of the Okavango delta is reduced by 80\% during maximum flooding. At
+minimum flooding, conversely, the delta reveals vital dispersal corridors and
+increases chances of successful dispersal into neighboring habitat patches.
+Concomittantly, dispersal durations to move between patches is reduced at low
+flood levels.
+
+Climate change is expected to critically impact the flood-pulsing cycles of the
+Okavango delta and may therefore critically alter landscape connectivity in this
+area. Besides a better understanding of the conservation needs for the African
+wild dog, our study also presents a first step towards incorporating
+environmental changes due to seasonality or climate change in dispersal and
+connectivity analyses. Finally, our analysis demonstrates the usefulness of
+individual-based dispersal simulations as a pertinent conservation tool to study
+impacts of environmental change while accounting for dispersal movements.
 
@@
 \subsection{Climate Change}
+Climate change is expected to profoundly impact ecosystems worldwide, with
+far-reaching consequences for species living therein (cite IPCC)
+\citep{Ozgul.2010, Radchuk.2019}. However, empirical information on species' responses to climate change remains
+scarce, especially for regions most vulnerable to altered conditions
+\citep{Paniw.2021}.
 
-Despite the importance of climate change in determining species viability,
-predicting the way and extent to which it will affect species remains
-challenging. One exciting avenue to gain insights into the implications of
-environmental change is to utilize seasonal variation in the environment.
+\subsection{Dispersal}
+The lack of knowledge in terms of species ability to cope with climate change is
+further compounded by a lack of information on dispersal under changing
+environmental conditions \citep{Travis.2013}. By bringing individuals away from
+their natal location, dispersal strongly influences population dynamics
+\citep{Clobert.2012}; it enables genetic exchange \citep{Frankham.2002,
+Leigh.2012, Baguette.2013}, facilitates the colonization of empty habitats
+\citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and promotes reinforcement
+of weakened and small subpopulations \citep{Brown.1977}. Dispersal also serves
+as a means to track suitable habitat \citep{Raia.2012} and may thus serve as a
+means to mediate the demographic consequences of environmental change
+\citep{Hodgson.2009, Travis.2013}. Dispersal is also inextricably linked to the
+concept of landscape connectivity, which is generally understood as the degree
+by which the landscape facilitates or impedes movements among habitat patches
+\citep{Tylor.1993}.
+
+As previously highlighted by \cite{Travis.2013}, climate may affect dispersal
+directly, by altering the propensity of individuals to disperse, or indirectly,
+through changes in the biophysical environment. Here, we will focus on the
+latter and study dispersal prospects under changing environmental conditions.
+
+In spite of its importance for determining population dynamics and landscape
+connectivity, our understanding of dispersal and how dispersal is affected by
+changing environmental conditions is limited. This is owed to the difficulty of
+obtaining data collected on dispersing animals at the appropriate temporal and
+spatial scale \citep{Graves.2014, Vasudev.2015} and our inability to project
+dispersal prospects under changing environmental conditions into the future.
+
+The ability of species
+to persist under these changing conditions depends on the amplitude and speed of
+environmental change and species' responses to such changes (cite someone).
+
+It is, in fact, presumed that humanity has already transgressed the planetary
+boundary for climate change, therefore triggering irreversible feedbacks in the
+biosphere \citep{Rockstrom.2009}.
 
@@
 
-\subsection{Vital Rates}
-Several studies investigating the impacts of climate change on life-history
-traits have revealed that population persistence may critically depends on
-climate conditions \citep{Paniw.2019}. While many studies examine climate
-impacts on survival and reproductive rates, the influence of global change or
-climatic extreme events on dispersal prospects has received little attention.
-
-Ecosystems that undergoe seasonal environmental variation, in particular arid
-regions that experience extreme climatic conditions, are among the most
-vulnerable to climate change.
-
-\subsection{Dispersal}
 \subsection{Connectivity}
 Dispersal is an important, if not the important, driver of landscape
-connectivity and therefore of major interest to conservation authorities.
+connectivity and therefore of major interest to conservation authorities. It has
+also been demonstrated that dispersal may lead to population declines in areas
+where anthropogenic mortality is high and dispersal prospects low
+\citep{Leigh.2012}.
 
@@
 Botswana and its surroundings are among the most vulnerable to climate change.
+While global temperatures are expected to increase by xx degrees until the end
+of the 21st century, For example, \cite{Engelbrecht.2015} predicts an increase
+between 4 and 6\degree C for regions in southern Africa. One species that
+heavily relies on dispersal for long-term persistence and that inhabits such an
+arid ecosystem in Southern Africa is the African wild dog (\textit{Lycaon
+pictus}). The African wild dog is a lightly built canid lives in cohesive packs
+comprising up to 20 individuals. Wild dogs are social breeders and the majority
+of reproduction is monopolized by the dominant alpha couple who are supported by
+subordinates in taking care of their pups. The species exhibits strong
+inbreeding avoidance, hence juvenile individuals born into the pack disperse in
+same-sex sibling coalitions after reaching sexual maturity \citep{McNutt.1996,
+Behr.2020}. Notably, the African wild dog is the sole extant representative of
+its genus and thus considered as high-priority species for conservation efforts
+\citep{Leigh.2012}. However, fragmentation and destruction of their historic
+habitat may have decreased dispersal success for this species and thus
+jeopardizes connectivity \citep{Leigh.2012}. Since the impacts of climate change
+are not uniformly distributed across the planet, but particularly pronounced at
+high latitudes and arid ecosystems, the species may be seen as most vulnerable
+to such ecosystem shifts.
 
-\section{Methods}
-We conducted all analyses using the programming language R.
+\subsection{Our Study}
+Here, we capitalize on a previously parametrized dispersal model and use
+dispersal simulations to investigate connectivity patterns for African wild dogs
+under two extreme scenarios; one assuming maximum flooding of the Okavango delta
+and one assuming minimum flooding of the Okavango delta. Given that water poses
+a barrier to dispersing individuals, we anticipate that dispersal prospects and
+connectivity during maximum flood are low. During minimum flood, in contrast, we
+expect to reveal the presence of several dispersal corridors that may be used to
+freely and quickly move between remaining habitat patches.
+
+\section{Materials and Methods}
+We conducted all analyses using the programming language \textsf{R}. Any spatial
+data manipulation was completed using the \textsf{terra} and \textsf{spatstat}
+packages. Several helper functions for the simulation algorithm were written in
+\textsf{C++} and imported to R using the \textsf{Rcpp} package. Figures were
+generated using \textsf{ggplot2}, \textsf{igraph}, and \textsf{ggnetwork}. All
+R-scripts used to conduct our analyses are provided through an online
+repository.
 
 \subsection{Study Area}
-The study area for this analysis was focused on the Okavango delta and its
+The study area for this analysis was focused on the Okavango delta (OD) and its
 surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
-Zimbabwe, and Zambia (\Cref{StudyArea}). To accommodate for the long distance
-dispersal events commonly observed in African wild dogs (e.g.
-\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}), we considered a large
-rectangular extent stretching from 20\degree 30' E to 26\degree E. totaling to
-an area of 300'000 km\textsuperscript{2}. The dominant geographical feature in
-this region is the Okavango delta (OD), the world's largest inland delta and the
-main driver of seasonal environmental change in the region. The flood-pulsing
-rhythm of the OD is mainly dictated by precipitation in the Angolian highlands,
-where rainfalls are channeled into the OD through the Okavango River. Although
-precipitation in Angola peaks in ..., water only slowly descends through the
-Okavango river and its tributaries, reaching the distal ends (i.e. the faults)
-of the delta in July or August. At minimum extent, the flood covers an area of
-3'600 km\textsuperscript{2}, whereas during maximum flood it covers more than
-9'000 km\textsuperscript{2}. Vegetation in the study area is dominated by dense
-mopane forest, mixed acacia woodland, and plain grassland. Human influence in
-the study area is low and mainly concentrated around small villages at the
-western periphery of the delta and to the city of Maun at the southern tip of
-the delta. Large portions of land are dedicated national parks, game reserves or
-forest reserves. The study area is also part of the world's largest
-transboundary conservation initiative, the Kavango-Zambezi Transfrontier
-Conservation Area.
+Zimbabwe, and Zambia (\Cref{StudyArea}). The OD is the world's largest inland
+delta and the main driver of seasonal environmental change in the region. While
+our primary focus lied on the immediate surroundings of the Okavango Delta, we
+considered a large rectangular extent stretching from 20\degree 30' E to
+26\degree E. (totaling to an area of 300'000 km\textsuperscript{2}) to
+accommodate for the long distance dispersal events commonly observed in African
+wild dogs (e.g. \citealp{Davies-Mostert.2012, Masenga.2016, Cozzi.2020}). The
+flood-pulsing rhythm of the OD is mainly dictated by precipitation in the
+catchment areas of the Angolian highlands, where rainwater is collected and
+channeled into the OD through the Okavango River. Although precipitation in
+Angola peaks between December and March, water only slowly descends through the
+Okavango river and its distributaries, reaching the distal ends of the delta in
+July or August, where the water percolates at the Thamalakane and Kunyere
+Faults. At minimum extent, the flood covers an area of 3'600
+km\textsuperscript{2}, during maximum flood more than 9'000
+km\textsuperscript{2}. Vegetation in the study area is dominated by mopane
+forest, mixed acacia woodland, and grassland. Human influence is low and mainly
+concentrated around small villages at the western periphery of the delta as well
+as the city of Maun at the south-eastern tip of the OD. Large portions of land
+are dedicated national parks, game reserves or forest reserves. The study area
+is also part of the world's largest transboundary conservation initiative, the
+Kavango-Zambezi Transfrontier Conservation Area, which aims to restore
+connectivity between protected areas in Southern Africa.
 
@@
   \includegraphics[width = \textwidth]{99_StudyArea.png}
-  \caption{Study area across which we simulated dispersal events. Virtual
-  dispersers were released at random locations within the orange source areas.}
+  \caption{Study area across which we simulated dispersal. Simulated dispersers
+  were released at random locations within the orange source areas distributed
+  across the delta. Emigration zones (purple) served as checkpoints and enabled
+  us to identify if and where simulated dispersers left the close surroundings
+  of the Okavango delta. These zones were generated using a set of cutlines
+  originating from the center of the delta and roughly cutting an elliptical
+  buffer zone into sections of equal size.}
   \label{StudyArea}
@@
 We represented the physical landscape through which dispersers could move by a
-set of spatially referenced habitat layers each resolve at 250m x 250m. The set
-of layers included water-cover, distance-to-water, tree-cover,
+set of spatially referenced habitat layers, each resolved at 250m x 250m. The
+set of layers included water-cover, distance-to-water, tree-cover,
 shrub/grassland-cover, and a human influence layer depicting anthropogenic
 influences through villages, roads, and agriculture. A detailed description of
-the different layers is provided in \Cref{Hofmann.2021}. Importantly, the
-water-cover and the derived distance-to water layers were generated using MODIS
-Terra satellite imagery, which enabled us to generate weekly updated layers that
-provided detailed information about the flood-extent at any given point in time.
-We had a total of 8xx remote sensed floodmaps at our disposal and used them to
-generate two extreme scenarios; a minimal and maximum flood scenario. To create
-the minimum flood scenario, we tallied the 50 floodmaps with smallest flood
-extent into an average image. Finally, we created a binary map... Similarly, we
-created an average image for high flood using the 50 most flooded maps. The
-resulting maps are depicted in \Cref{FloodExtent}. For completeness we also
-generated an  ``average'' floodmap by averaging across all 800 floodmaps
-available to us. The results based upond this layer are presented in the
-Appendix.
+the different habitat layers is provided in \Cref{Hofmann.2021}. Importantly,
+the water-cover and derived distance-to water layers were generated using MODIS
+Terra MCD43A4 satellite imagery that was classified using a ``floodmapping''
+algorithm developed by \citep{Wolski.2017} and available through the R-package
+\textsf{floodmapr}. The algorithm allowed us to generate almost weekly updated
+``floodmaps'', thus providing detailed information about the flood-extent at any
+given point in time. In total, we generated 700 floodmaps between the years 2000
+and 2019. Based on these maps, we generated a minimal and maximum flood
+scenario. To create the minimum flood scenario, we averaged the 50 floodmaps
+with smallest flood extent and generated a binary image using areas that were
+inundated in at least 50\% of the maps. Similarly, we created an average image
+for high flood using the 50 most flooded maps. The final maps are depicted in
+\Cref{FloodExtent}.
 
@@
   \includegraphics[width = \textwidth]{99_FloodExtent.png}
-  \caption{}
+  \caption{Flood extent in the two scenarios considered. In the left panel, the
+  flood is at an extremely low level, stretching across 3'558
+  km\textsuperscript{2}, whereas in the right panel the flood is at an extremely
+  high level and covers 9'397 km\textsuperscript{2}. The two maps were generated
+  using 700 remote sensed MODIS MCD43A4 satellite images spanning the years 2000
+  to 2019.}
   \label{FloodExtent}
@@
 Our dispersal model was based on a previously parametrized and validated
-\textit{integrated step-selection function (iSSF)} applied to the dispersal data
-of 16 dispersing African wild dogs inhabiting the surroundings of the Moremi
-National Park in the Okavango delta \citep{Hofmann.2021}. The iSSF model
-consists of two complementary ``kernels'' plus their interactions. The first
-kernel is a movement kernel and describes general movement behavior of
-dispersing AWDs, irrespective of habitat conditions. The second kernel is a
-habitat kernel and describes preferences of AWDs with regards to environmental
-conditions. Finally, the model also includes interactions among the two kernels
-and therefore allows to render how movement behavior changes depending on
-habitat conditions.
+integrated step-selection function (iSSF, \citealp{Avgar.2016}) fitted to GPS
+data of dispersing AWDs \citep{Hofmann.2021}. In step selection functions (SSFs,
+\citealp{Fortin.2005}), observed GPS locations are converted into steps (the
+straight-line traveled between two GPS recordings \citep{Turchin.1998a}) and
+compared to a set of \textit{random} steps in a (mixed effects) conditional
+logistic regression framework \citep{Fortin.2005, Thurfjell.2014, Muff.2020,
+Fieberg.2021}. The model presented in \citep{Hofmann.2021} used dispersal data
+of 16 dispersing African wild dogs from a free-ranging wild dog population in
+northern Botswana. GPS data during dispersal was collected at 4-hourly intervals
+and translated into steps of similar duration. Observed steps were then paired
+step with 24 random steps that were generated using a uniform distribution for
+turning angles (\(-\pi, +\pi\)) and step lengths from a gamma distribution
+fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\) = 0.37). It
+was then assumed that animals assigned to each observed and random step a
+selection score of the form \citep{Fortin.2005}:
 
-\subsection{Source Areas}
-We simulated dispersing AWDs originating from nine distinct source patches
-located in the vicinity of the Okavango delta. The source areas were generated
-as follows. First, we overlayed the OD with an oval that was bound by
-geographical landmarks; in the north, the oval was bound by the Inflow of the
-Okavango river into the ``pan-handle'' of the OD, north-east the oval was bound
-by the Selinda-Spillway and the Linyanty swamp, towards South-East it was bound
-by the Boteti river, and towards South-West by Lake Ngami. We then dissected the
-polygon into five distinct patches using the same natural landmarks
-(\Cref{StudyArea}). Patch one was given by the area south of the Boteti river
-and area two by the area north of it up until the panhandle. Area three
-stretched from the region easf of Maun towards north until the Selinda-spillway,
-whereas area four stretched north of the spillway until west towards the
-panhandle. Finally, a fifrth source area marked the peninsuale in the center of
-the OD. In addition to these five source patches, we also distributed small
-peripheral patches that we used to simulate individuals immigrating
-\textit{into} system and to keep track of individuals leaving it.
+\begin{equation}
+\label{EQ1}
+  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
+\end{equation}
+
+Where (\(x_1, x_2, ..., x_n\)) represent the covariate values along each of the
+steps and the (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's relative
+selection strengths \citealp{Avgar.2017} towards these covariates. The benefit
+of \textit{integrated} SSFs over regular SSFs is that they provide a means to
+render two complementary ``kernels''. A movement kernel that describes general
+movement behavior of dispersing AWDs and a habitat kernel that describes
+preferences of AWDs with regards to environmental conditions
+\citep{Fieberg.2021}. iSSFs also allow interactions among the two kernels and
+are thus suitable to render that movement behavior may change depending on
+habitat conditions. A brief summary of the fitted dispersal model is provided in
+Appendix xx.
+
+\subsection{Source Areas and Emigration Zones}
+We simulated dispersing AWDs originating from six distinct source areas located
+in the vicinity of the Okavango delta (\Cref{StudyArea}). For areas one to six
+we selected locations at the delta's periphery that remained dry in both
+scenarios. At those locations, we generated circular buffers with a radius of 20
+km. For source area six, we isolated a polygon covering Chief's Island, a
+peninsula located at the OD's center. Besides source areas, we also generated
+``zones of emigration'' that we used as checkpoints to determine if and where
+simulated individuals left the delta's vicinity (\Cref{StudyArea}). We generated
+emigration zones by first overlaying the OD with an elliptic buffer zone that we
+dissected using a set of cutlines that originated from the ODs center and spread
+according to cardinal points (\Cref{StudyArea}).
 
 \subsection{Dispersal Simulation}
-For both environmental scenarios we simulated 1'000 individuals dispersing from
-each of the nine source area depicted in \Cref{StudyArea}. The simulation
-algorithm was based on the algorithm described in \Cref{Hofmann.2022} and works
-as follows. A random location within the source area is chosen as a starting
-point. Originating from the starting point, a set of 25 random steps are
-generated by sampling step lengths from a gamma distribution fitted to observed
-steps (what is a step?) (shape = , scale = ) and turning angles from a uniform
-distibution on ($-\pi, +\pi$). Along each random step the underlying spatial
-covariates are extracted and relevant movement metrics are computed (i.e.
-log(sl), cos(ta), ta). The parametrized dispersal model is used to predict the
-probability of each step for being chosen given the steps covariate values. One
-of the steps is sampled based on assigned probabilities and the location of the
-animal is updated. The procedure is then repated until the desired number of
-steps (in our case 2000) is realized. Resulting trajectories resemble correlated
-random walks. As the original model was trained using 4-hourly steps, a
-simulated step also resembled straight line movements within four hours.
+For each source area we simulated 1'000 individuals, once assuming a minimum
+flood, once assuming a maximum flood. This resulted in the simulation of 6'000
+individuals for each environmental scenario, hence 12'000 individuals in total.
+The simulation algorithm was based on the algorithm described in
+\Cref{Hofmann.2022} and works as follows. A random location within the source
+area is chosen as a starting point. Originating from the starting point, a set
+of 25 random steps is generated by sampling step lengths from a gamma
+distribution fitted to observed steps (what is a step?) (shape = , scale = ) and
+turning angles from a uniform distibution ($-\pi, +\pi$). Along each random step
+the underlying spatial covariates are extracted and relevant movement metrics
+are computed (i.e. log(sl), cos(ta), ta). The parametrized dispersal model is
+used to predict the probability of each step for being chosen given the steps
+covariate values. One of the steps is sampled based on assigned probabilities
+and the location of the animal is updated. The procedure is then repated until
+the desired number of steps is realized. Here, we simulated each individual for
+2'000 steps, which correspons to the longest dispersal duration recorded in our
+data. The original model was trained using 4-hourly steps, thus a simulated step
+also resembled the movement conducted within four hours. Trajectories resulting
+from such a simulation can be understood as correlated random walks that take
+into account both habitat and movement preferences of dispersing individuals.
 
@@
 
-\subsection{Connectivity}
+\subsection{Dispersal Prospects and Connectivity}
 Based on simulated dispersal trajectories in the two scenarios we quantified
-connectivity in three complementary ways. First, we generated heatmaps depicting
-the frequency at which different areas in the landscape were visited by
-simulated dispersers. Such heatmaps serve to detect dispersal hotspots but are
-less suitable to detect pinchpoints and bottlenecks. Hence, we also computed
-spatially explicit betweenness scores which are useful to highlight such
-pinchpoints (Bastille). To compute betweenness, we overlayed the study area with
-a regular grid with 2.5 km x 2.5 km grid cells and determined how often
-simulated individuals transitioned from one grid-cell to another. In cases where
-the same individual repeatedly realized the same cell-transition, we only
-counted a single transition to avoid emphasizing regions regions in which
-individuals moved in circles. Finally, we calculated the number of successful
-dispersal events between the different source areas. We coin this type of
-connectivity ``inter-patch connectivity''. Dispersal between two areas was said
-to be successful whenever a trajectory leaving one area intersected with the
-target area. To gauge the dispersal duration needed to move between patches (be
-consistent with ``patches'', ``source-areas'' etc.), we also recorded the
-minimum number of steps that individuals had to move to move between areas.
+dispersal sucess and connectivity using three complementary connectivity metrics
+as outlined in \cite{Hofmann.2022}. First, we generated heatmaps depicting the
+frequency at which different areas in the landscape were visited by simulated
+dispersers. Such heatmaps serve to detect dispersal hotspots and areas of
+intense use. However, they are less suitable for detecting pinchpoints and
+bottlenecks that are critical in linking distinct patches. Hence, we also
+computed spatially explicit betweenness scores which are useful to highlight
+exactly such pinchpoints \citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}.
+To compute betweenness, we overlayed the study area with a regular grid with 2.5
+km x 2.5 km grid cells and determined how often simulated individuals
+transitioned from one grid-cell to another. Note that in case the same
+individual repeatedly realized the same cell-transition (e.g. repeatedly moved
+between A-B-A-B...), we only counted a single transition to avoid emphasis on
+regions where individuals moved in circles. With on the so generated
+transitions, we generated a network using the centers of all grid-cells as nodes
+and cell-transitions as weighted connections between the nodes. Based on this
+network we computed weighted betweenness scores using the R-package
+\textsf{igraph}. As a final connectivity metric and metric of dispersal success,
+we calculated the number of successful dispersal events between the different
+source areas as well as towards the emigration zones. We coin this type of
+connectivity ``inter-patch connectivity'' as it relates to the movement between
+distinct patches. Dispersal between two areas was said to be successful whenever
+a trajectory leaving one area intersected with the target area. To gauge the
+dispersal duration needed to move between patches (be consistent with
+``patches'', ``source-areas'' etc.), we also recorded the minimum number of
+steps that individuals moved before arriving at the respective patch.
 
 \section{Results}
+\subsection{Heatmaps}
 Heatmaps produced from simulated dispersal trajectories reveal that the OD acts
 as a major dispersal barrier during periods of high flood, but reveals viable
-dispersal corridors during periods of low flood. In particular, our analysis
-reveals that the area north-west of Maun is frequently visited by dispersers at
-in the minimum-flood scenario, yet that an extended flood, coupled with high
-human influence around the city of Maun, enforces simulated individuals to
-either detour around Maun, or to slip through the narrow window between village
-and flood. Similarly, the betweenness map indicates the presence of a dispersal
-corridor extending from the western section of the delta leading across the
-delta's center into its western part, yet only during low-flood. During maximum
-flood, the same corridor is pushed south and enforces dispersers to move closer
-to the city of Maun. In general, connectivity during maxium flood is lower
-compared to the minimum flood scenario (number links vs. number liks). Most
-notably, source area 5 is almost inaccessible during maximum flood.
-
- Most notably, the
-area north-west of Maun acts as dispersal habitat during phases of low flood,
-yet is almost inaccessible during phases of maxium flood. This is reflected in
-the heatmap, where connectivity across the western part of the delta is low
-during maximum flood, but also in the betweenness map, where an important
-movement corridors through the delta is only visible during minimum flood.
-Although connectivity across large parts of the landscape remains unchanged
-across environmental scenarios, one critical movement corridor shifts in
-response to an expanding flood. Most notably, a corridor that links the western
-and eastern part of the delta during minimum flood gets redirected further
-south, thus leading more closely to the city of Maun. Unsurprisingly, the
-differences are more pronounced for areas directly affected by the flood, yet it
-is worth noting that even some areas not directly affected by the flood.
+dispersal corridors during periods of low flood (\Cref{Heatmaps}). During
+minimum flood, the area north-west of Maun appears to serve as vital dispersal
+habitat. The same area is entirely avoided during maximum flood. Besides
+striking differences in connectivity for the close vicinity of the delta, the
+remainder of the study area shows only marginal differences in connectivity
+between the two scenarios. For instance, in both scenarios the area south of the
+Linyanti swamp appears as frequently visited dispersal habitat. Additional
+heatmaps highlighting differences in connectivity for each source area
+separately are provided in Appendix SX.
 
@@
   \includegraphics[width = \textwidth]{99_Heatmaps.png}
-  \caption{}
+  \caption{Heatmaps depicting the number of simulated dispersal trajectories
+  traversing each grid-cell in the study area. The left panel shows results for
+  the minimum flood scenario, whereas the right panel shows results for the
+  maximum flood scenario. Source areas (numbered 1-6) from which dispresers were
+  released, and emigration zones (numbered 7-14) are shaded in dark gray.}
   \label{Heatmaps}
+  \end{center}
+\end{figure}
+
+\subsection{Betweenness}
+The betweenness maps reveal a similar pattern in that connectivity through the
+OD is only pronounced during periods of low floods and vanishes entirely during
+maximum flood (\Cref{Betweenness}). A set of four dispersal corridors meets on
+the central peninsula (source area 5, \Cref{Betweenness}) at minimum flood but
+the same corridors are absent when the flood reaches a maximum extent. Instead,
+a narrow corridor runs north west of Maun, connecting source areas one and two.
+Again, the remainder of the study area is only marginally affected by flooding
+patterns. Additional betweenness maps highlighting differences in connectivity
+for each source area separately are provided in Appendix SX.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Betweenness.png}
+  \caption{Betweenness scores highlighting potential dispersal corridors for
+  both the minimum and maximum flood scenario. Areas with high betweenness
+  scores (bright yellow) are used by many simulated individuals to move into
+  adjacent regions and can thus be understood as critical pinch-points. Source
+  areas (numbered 1-6) from which dispresers were released, and emigration zones
+  (numbered 7-14) are shaded in light gray.}
+  \label{Betweenness}
+  \end{center}
+\end{figure}
+
+\subsection{Inter-Patch Connectivity}
+Our analysis of interpatch connectivity demonstrates notable differences in
+dispersal prospects and duration depending on the extent of the flood
+(\Cref{IPCTable}). While $\input{99_NumberReachingOthersMinFlood.tex}$ simulated
+dispersers reach another source area during minimum extent, only
+$\input{99_NumberReachingOthersMaxFlood.tex}$ do so during maximum extent. The
+differences are particularly pronounced for individuals dispersing from or into
+the source area located at the OD's center (\Cref{Interpatch}). While the area
+is reached by $\input{99_NumberReachingArea6MinFlood.tex}$ simulated individuals
+during minimum flood, only $\input{99_NumberReachingArea6MaxFlood.tex}$
+dispersers arrive there during maximum flood. Furthermore, the dispersal
+duration into source area six from any other source area increases from
+$\input{99_StepsToReachingArea6MinFlood.tex}$ to
+$\input{99_StepsToReachingArea6MaxFlood.tex}$. Across all simulations, the
+average dispersal duration before reaching another source area increases from xx
+to xx from the minimum to the maximum flood scenario. Nevertheless, connectivity
+into some areas increases during maximum flooding. Emigration increased slightly
+from xx to xx
+
+For instance, while dispersers .. Additional maps highlighting differences in
+inter-patch connectivty for each source area separately are provided in Appendix
+SX.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_IPCTable.png}
+  \caption{Dispersal frequency (a) and duration (b) (in steps) between source
+  areas and emigration zones during minimum and maximum flood.}
+  \label{IPCTable}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_DifferenceMap.png}
-  \caption{}
-  \label{DifferenceHeatmaps}
+  \includegraphics[width = \textwidth]{99_Interpatch.png}
+  \caption{Inter-patch connectivity into and from source area number six. This
+  patch experienced the most drastic reduction in connectivity in result to a
+  high flood level. During minimum flood there is ample connectivity from and
+  into source areas 1-3. During maximum flood, connectivity is lower and mainly
+  limited to source area 2.}
+  \label{Interpatch}
   \end{center}
@@
 
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_Betweenness.png}
-  \caption{}
-  \label{Betweenness}
-  \end{center}
-\end{figure}
-
-\input{IPC_Frequency.tex}
-\input{IPC_Duration.tex}
+% \input{IPC_Frequency.tex}
+% \input{IPC_Duration.tex}
+% \input{IPC.tex}
 
@@
 
-\cite{MurrayHudson.2006} predicted that increased temperatures, additional human
+\cite{Murray-Hudson.2006} predicted that increased temperatures, additional human
 abstractions, and reduced river flows might lead to a ``Delta dying'' and that
@@
 more than offsetting the incline in precipitation \cite{Akinyemi.2019}.
+
+We studied a population of African wild dogs that resides in a natural
+environment with little human influence. This is only representative for a small
+share of the extant wild dog populations, as most individuals reside in areas
+that are prone to substantial edge effects. For these populations, the benefit
+of dispersal is disputed, as dispersal under high anthropogenic mortality may
+lead to a net-loss of individuals \citep{Leigh.2012}.
+
+We assessed the implications of environmental change on the dispersal prospects,
+yet we did not consider how changing conditions alter dispersal propensity.
 
@@
 \singlespacing
-\bibliography{Literatur}
+\bibliography{Literature}
 \endgroup

### 2022-11-03 Chapter_8/04_Manuscript/Manuscript_Support.tex
Updates addressing reviewer comments

--- 2022-09-05 Chapter_8/04_Manuscript/Manuscript.tex
@@
 \usepackage[flushleft]{threeparttable}
-\usepackage[nomarkers, nolists]{endfloat}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \captionsetup{labelfont=bf, format=plain, font=small}
+
+% For supplementary material
+\newcommand{\beginappendix}{%
+  \setcounter{table}{0}
+  \renewcommand{\thetable}{S\arabic{table}}%
+  \setcounter{figure}{0}
+  \renewcommand{\thefigure}{S\arabic{figure}}%
+  \setcounter{equation}{0}
+  \renewcommand{\theequation}{Equation S\arabic{equation}}%
+  \setcounter{section}{0}
+  \renewcommand{\thesection}{A.\arabic{section}}%
+}
 
@@
 %------------------------------------------------------------------------------
-\title{Flooding of the Okavango Delta influences Connectivity for Dispersing
-African Wild Dogs}
+\title{\textbf{Appendix}\\ Flooding of the Okavango Delta influences
+Connectivity for Dispersing African Wild Dogs}
 
@@
 
-\textbf{Keywords:} Seasonal Flooding of the Okavango Delta and its Consequences
-for African Wild Dog Connectivity
+\textbf{Keywords:} movement, connectivity, seasonality, dispersal, conservation,
+permeability
 
 \end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Abstract
-%------------------------------------------------------------------------------
-\newpage
-\begin{abstract}
-Climate change is expected to considerably impact vital rates of endangered
-species. In particular dispersal, the phase during which species leave their
-natal area and attempt to find a new territory to settle, may
-
-We utilize seasonal changes in the flooding regime across the Okavango delta as
-a large scale natural experiment to investigate how environmental changes impact
-species' ability to disperse, and utlimately connectivity between remaining
-subpopulations. Specifically, we apply individual-based dispersal simualtions to
-simulate dispersal trajectories under two extreme scenarios, assuming minimal
-and maximum flood extents.
-
-Our results show that during years of minimum flood extent, the Okavango delta
-releases vast areas that dispersers use as movement corridors, whereas during
-years of extreme flood the delta's floodwaters pose a considerable dispersal
-barrier that forces dispersing individuals closer towards human settlements.
-Besides a better understanding of the conservation needs for the African wild
-dog, our study also provides evidence that incorporating seasonality in studies
-of connectivity is imperative to more accurately predict the dispersal ability
-of endangered species.
-
-\end{abstract}
 
@@
 
-\onehalfspacing
-\tableofcontents
-\doublespacing
-
 % Change page numbering
-\newpage
 \pagenumbering{arabic}
@@
 
-\section{Introduction}
-\subsection{Climate Change}
-\subsection{Vital Rates}
-\subsection{Dispersal}
-\subsection{Connectivity}
-Dispersal is an important, if not the important, driver of landscape
-connectivity and therefore of major interest to conservation authorities.
+% Change to appendix counters
+\appendix
+\beginappendix
 
-\subsection{Okavango Delta and African Wild Dogs}
-The Okavango delta in Southern Africa poses a unique opportunity to study the
-impacts of environmetal change on species dispersal ability and connectivity in
-large scale natural experiment setup.
-
-\section{Methods}
-Programming language, packages, etc.
-
-\subsection{Study Area}
-The study area for this analysis was focused on the Okavango delta and its
-surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
-Zimbabwe, and Zambia \Cref{StudyArea}. To accomodate for the long distance
-dispersal events commonly observed in the African wild dog, we considered a
-large rectangular extent stretching from ... to ...., totalling to an area of xx
-km. The dominant geographical feature in this region is the Okavango delta, the
-world's largest inland delta and main driver of seasonal environmental change in
-the region. While local rainfalls marginally impact water levels across the
-delta, the main flood-pulsing rythm is driven by the water-influx from the
-Okavango river. The Okavango carries rain waters collected the Angolian
-highlands into the delta's distributaries. Although precipitation in Angola
-peaks around ..., the water arrives in the delta with a delay of ... It only
-slowly descends through the delta and reaches the distal ends (at the faults)
-around ... At minimum extent, the flood covers an area of around xx km, whereas
-during maximum flood it covers ... Vegetation in the study area can largely be
-categorized into dense mopane forest, mixed acacia woodland, and plain
-grassland. Human influence in the study area is generally low and mainly
-concentrated to small villages at the western periphery of the delta, as well as
-to the city of Maun at the southern tip of the delta. Large portions of land are
-dedicated national parks, game reserves or forest reserves and the area is part
-of the world's largest transboundary conservation initiative, the
-Kavango-Zambezi Transfrontier Conservation Area.
-
-\begin{figure}
+%------------------------------------------------------------------------------
+%  Appendix S1: Source-Specific Heatmaps
+%------------------------------------------------------------------------------
+\newpage
+\section{Source-Specific Heatmaps}
+\begin{figure}[htbp]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_StudyArea.png}
-  \caption{Study area across which we simulated dispersal events. Virtual
-  dispersers were released at random locations within the orange source areas.}
-  \label{StudyArea}
-  \end{center}
-\end{figure}
-
-\subsection{Spatial Habitat Layers}
-We represented the physical landscape through which dispersers could move by a
-set of spatially referenced habitat layers each resolve at 250m x 250m. The set
-of layers included water-cover, distance-to-water, tree-cover,
-shrub/grassland-cover, and a human influence layer depicting anthropogenic
-influences through villages, roads, and agriculture. A detailed description of
-the different layers is provided in \Cref{Hofmann.2021}. Importantly, the
-water-cover and the derived distance-to water layers were generated using MODIS
-Terra satellite imagery, which enabled us to generate weekly updated layers that
-provided detailed information about the flood-extent at any given point in time.
-We had a total of 8xx remote sensed floodmaps at our disposal and used them to
-generate two extreme scenarios; a minimal and maximum flood scenario. To create
-the minimum flood scenario, we tallied the 50 floodmaps with smallest flood
-extent into an average image. Finally, we created a binary map... Similarly, we
-created an average image for high flood using the 50 most flooded maps. The
-resulting maps are depicted in \Cref{FloodExtent}. For completeness we also
-generated an  ``average'' floodmap by averaging across all 800 floodmaps
-available to us. The results based upond this layer are presented in the
-Appendix.
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_FloodExtent.png}
-  \caption{}
-  \label{FloodExtent}
-  \end{center}
-\end{figure}
-
-\subsection{Dispersal Model}
-Our dispersal model was based on a previously parametrized and validated
-\textit{integrated step-selection function (iSSF)} applied to the dispersal data
-of 16 dispersing African wild dogs inhabiting the surroundings of the Moremi
-National Park in the Okavango delta \citep{Hofmann.2021}. The iSSF model
-consists of two complementary ``kernels'' plus their interactions. The first
-kernel is a movement kernel and describes general movement behavior of
-dispersing AWDs, irrespective of habitat conditions. The second kernel is a
-habitat kernel and describes preferences of AWDs with regards to environmental
-conditions. Finally, the model also includes interactions among the two kernels
-and therefore allows to render how movement behavior changes depending on
-habitat conditions.
-
-\subsection{Source Areas}
-We simulated dispersing AWDs originating from nine distinct source patches
-located in the vicinity of the Okavango delta. The source areas were generated
-as follows. First, we overlayed the OD with an oval that was bound by
-geographical landmarks; in the north, the oval was bound by the Inflow of the
-Okavango river into the ``pan-handle'' of the OD, north-east the oval was bound
-by the Selinda-Spillway and the Linyanty swamp, towards South-East it was bound
-by the Boteti river, and towards South-West by Lake Ngami. We then dissected the
-polygon into five distinct patches using the same natural landmarks
-(\Cref{StudyArea}). Patch one was given by the area south of the Boteti river
-and area two by the area north of it up until the panhandle. Area three
-stretched from the region easf of Maun towards north until the Selinda-spillway,
-whereas area four stretched north of the spillway until west towards the
-panhandle. Finally, a fifrth source area marked the peninsuale in the center of
-the OD. In addition to these five source patches, we also distributed small
-peripheral patches that we used to simulate individuals immigrating
-\textit{into} system and to keep track of individuals leaving it.
-
-\subsection{Dispersal Simulation}
-For both environmental scenarios we simulated 1'000 individuals dispersing from
-each of the nine source area depicted in \Cref{StudyArea}. The simulation
-algorithm was based on the algorithm described in \Cref{Hofmann.2022} and works
-as follows. A random location within the source area is chosen as a starting
-point. Originating from the starting point, a set of 25 random steps are
-generated by sampling step lengths from a gamma distribution fitted to observed
-steps (what is a step?) (shape = , scale = ) and turning angles from a uniform
-distibution on ($-\pi, +\pi$). Along each random step the underlying spatial
-covariates are extracted and relevant movement metrics are computed (i.e.
-log(sl), cos(ta), ta). The parametrized dispersal model is used to predict the
-probability of each step for being chosen given the steps covariate values. One
-of the steps is sampled based on assigned probabilities and the location of the
-animal is updated. The procedure is then repated until the desired number of
-steps (in our case 2000) is realized. Resulting trajectories resemble correlated
-random walks. As the original model was trained using 4-hourly steps, a
-simulated step also resembled straight line movements within four hours.
-
-% We used a previously parametrized dispersal model to simulate dispersal
-% trajectories of African wild dogs across the study area. To simulate dispersing
-% wild dogs, we employed the simulation framework presented in
-% \cite{Hofmann.2022}. In this framework, a movement model that renders habitat
-% and movement preferences of dispersing individuals is parametrized using
-% step-selection functions. Once parametrized, the model can be employed to
-% simulate virtual dispersers moving across the landscape.
-%
-% We released virtual dispersers at random locations within the four distinct
-% source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
-% individuals moving across the landscape at minimum flood and another xx
-% individuals at maximum flood. For comparison, we also ran the simulations for a
-% medium flood extent, yet the results for this will be presented in the appendix.
-
-\subsection{Connectivity}
-Based on simulated dispersal trajectories we quantified connectivity across the
-OD for both environmental scenaiors in three complementary way. First, we
-generated heatmaps showing how often different areas in the landscape were
-visited by simulated dispersers. Such heatmaps serve to detect dispersal
-hotspots, yet are less suitable to detect pinchpoints and bottlenecks. Hence, we
-also computed spatially explicit betweenness scores which are useful to
-highlight movement corridors (Bastille). For this, we overlayed the study area
-with a regular grid with 2.5 km x 2.5 km grid cells and determined how often
-simulated individuals transitioned from one grid-cell to another. Note that we
-didn't consider revisits to avoid pronouncing regions in which individuals were
-encroached and moved in circles.
-
-To gain insights into landscape connectivity from the simulated dispersal
-events, we prepared three complementary connectivity maps (see again
-\citep{Hofmann.2022}); a heatmap, indicating the absolute frequency at which
-each pixel in the study area was traversed by a virtual disperser, a betweenness
-map, highlighting pixels of particular importance to facilitate movement from
-one area to another, and a map of interpatch connectivity, depicting the
-presence and intensity of functional links between source areas. To compute the
-heatmap, we counted how many times each pixel in the study area was traversed by
-a virtual disperser. To calculate betweenness, we overlayed the study area with
-a regular grid and used the simulated trajectories to determine movements from
-one grid-cell to another. Specifically, we counted how often transitions from
-one grid-cell to any other occured. The centerpoint of each grid-cell then
-served as network node, and the number of transitions from one cell to another
-as weights between the respective nodes. Based on the so generated network, we
-then computed betweenness scores for all network nodes (i.e. for all grid-cells
-across the study area). Lastly, to compute interpatch connectivity, we
-determined the frequency at which virtual dispersers originating from one source
-area reached any of the other source areas. We also calculated the average
-dispersal duration needed until individuals reached the respective area.
-
-\section{Results}
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_Heatmaps.png}
+  \includegraphics[width = \textwidth]{99_HeatmapsIndividual.png}
   \caption{}
@@
 
-\begin{figure}
+%------------------------------------------------------------------------------
+%  Appendix S2: Source-Specific Betweenness
+%------------------------------------------------------------------------------
+\newpage
+\section{Source-Specific Betweenness}
+\begin{figure}[htbp]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_DifferenceMap.png}
+  \includegraphics[width = \textwidth]{99_BetweennessIndividual.png}
   \caption{}
-  \label{DifferenceHeatmaps}
+  \label{Betweenness}
   \end{center}
@@
 
-\section{Discussion}
+%------------------------------------------------------------------------------
+%  Appendix S3: Source-Specific Betweenness
+%------------------------------------------------------------------------------
+\newpage
+\section{Source-Specific Inter-Patch Connectivity}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_IPCMain.png}
+  \caption{}
+  \label{IPCMain}
+  \end{center}
+\end{figure}
 
-\section{Authors' Contributions}
-D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
-D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
-first draft of the manuscript and all authors contributed to the drafts at
-several stages and gave final approval for publication.
-
-\section{Data Availability}
-GPS movement data of dispersing wild dogs is available on dryad
-\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
-proposed approach using simulated data are provided through Github
-(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
-codes required to reproduce the African wild dog case study will be made
-available through an online repository at the time of publication.
-
-\section{Acknowledgements}
-We thank the Ministry of Environment and Tourism of Botswana for granting
-permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
-Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we would like
-to thank Johannes Signer for assisting with the simulation algorithm. This study
-was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
-Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
-Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
-Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
-Grant (31003A\_182286) to A. Ozgul.
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_IPCBuffer.png}
+  \caption{}
+  \label{IPCBuffer}
+  \end{center}
+\end{figure}
 

### 2022-11-03 Chapter_8/02_R-Scripts/02_Visualization/Emigration.R

@@ -1,0 +1,125 @@
+################################################################################
+#### Emigration
+################################################################################
+# Description: Visualization of emigration into different emigration zones
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
+setwd(wd)
+
+# Load required packages
+library(raster)         # To handle spatial data
+library(tidyverse)      # To wrangle data
+library(sf)             # To plot spatial features
+library(geomtextpath)   # For angled labels in ggplot
+library(ggspatial)      # To add scale bars and north arrows
+library(ggpubr)         # To extract legends
+
+# Reload emigration data
+emigration <- read_rds("03_Data/03_Results/99_Emigration.rds")
+
+# Load stuff that we would like to plot
+water  <- read_sf("03_Data/02_CleanData/MajorWaters.shp")
+river  <- read_sf("03_Data/02_CleanData/MajorRivers.shp")
+areas  <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
+cutli  <- read_sf("03_Data/02_CleanData/Cutlines.shp")
+
+# I also want to add the Okavango river
+kava <- raster("03_Data/01_RawData/MERIT/Rivers.tif")
+kava <- crop(kava, extent(c(20.5, 22, -18.2, -17.5)))
+kava <- rasterToPolygons(kava, fun = function(x) {x == 1})
+kava <- aggregate(kava)
+kava <- buffer(kava, width = 0.1 / 111, dissolve = T)
+kava <- st_as_sf(kava)
+
+# Generate area labels
+labels_areas <- st_coordinates(st_point_on_surface(areas))
+labels_areas <- cbind(labels_areas, st_drop_geometry(areas))
+
+# Visualize emigration using a circular plot
+p1 <- ggplot(emigration, aes(x = as.factor(CurrentArea), y = n, fill = FloodLevel, col = FloodLevel)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 1), size = 0.3, width = 0.75, alpha = 0.8) +
+  geom_textpath(aes(y = n + 200, label = format(n, big.mark = "'", scientific = FALSE)), position = position_dodge(width = 1), col = "black", size = 2, angle = 90) +
[…]

### 2022-11-18 Chapter_1/04_Manuscript/Manuscript.tex
Revisions done

@@
 Dispersal of individuals is a vital process that allows species to maintain
-genetic diversity \citep{Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013},
-rescue non-viable populations \citep{Brown.1977}, and to colonize unoccupied
-habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability to disperse
-depends on a sufficient degree of landscape connectivity \citep{Fahrig.2003,
-Clobert.2012}, making the identification and protection of dispersal corridors
-that promote connectivity a task of fundamental importance \citep{Doerr.2011,
-Rudnick.2012}. Identifying dispersal corridors not only necessitates a
-comprehensive understanding of the factors that limit dispersal, but also an
-appropriate model to estimate connectivity \citep{Baguette.2013, Vasudev.2015,
-Hofmann.2021}. To date, the most commonly used connectivity models are
-least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit theory
-(CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both models rest on
-assumptions that appear unsuitable for dispersers, thus calling for the
+genetic diversity \citep{Perrin.2000, Frankham.2002, Leigh.2012, Baguette.2013,
+LaPoint.2013}, rescue non-viable populations \citep{Brown.1977}, and to colonize
+unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
+to disperse depends on a sufficient degree of landscape connectivity
+\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
+dispersal corridors that promote connectivity a task of fundamental importance
+\citep{Doerr.2011, Rudnick.2012}. Identifying dispersal corridors not only
+necessitates a comprehensive understanding of the factors that limit dispersal,
+but also an appropriate model to estimate connectivity \citep{Baguette.2013,
+Vasudev.2015, Hofmann.2021}. To date, the most commonly used connectivity models
+are least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and circuit
+theory (CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both models rest
+on assumptions that appear unsuitable for dispersers, thus calling for the
 development of alternative approaches. One promising alternative is to assess
@@
 Hofmann.2021}, meaning that dispersal trajectories are usually strongly
-directional. Moreover, because both models require static permeability or
-resistance surfaces as input, they are unable to reflect the temporal dimension
-of dispersal and therefore fail to allow statements about the expected duration
-for moving between habitat patches \citep{Martensen.2017, Diniz.2019}.
+directional. An interesting generalization that bridges the continuum between
+LCPA and CT has been proposed by \citep{Panzacchi.2016} and enables to
+capitalize on the merits of both approaches. Despite these and several other
+generalizations of LCPA and CT (e.g. \citealp{Pinto.2009, Landguth.2012,
+Panzacchi.2016, Brennan.2020}), some shortcomings remain. Most notably, all of
+these methods rely on static permeability or resistance surfaces that can't
+reflect the temporal dimension of dispersal. This permits statements about the
+expected duration for moving between habitat patches \citep{Martensen.2017,
+Diniz.2019}.
 
@@
 Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
-between movement behavior and habitat conditions \citep{Avgar.2016}. This
-strictly shifts the focus from a structural to a more functional view on
-connectivity \citep{Tischendorf.2000}. Furthermore, IBMMs generate movement
-sequentially, i.e. they generate a series of steps, so that the temporal
-dimension of dispersal becomes explicit and allows modeling autocorrelation
-between successive steps \citep{Diniz.2019}. Finally, simulations from IBMMs do
-not enforce movement or connections towards preconceived endpoints but allow
+between movement behavior and habitat conditions \citep{Avgar.2016}. This shifts
+the focus towards a more functional view on connectivity
+\citep{Tischendorf.2000}. Furthermore, IBMMs generate movement sequentially,
+i.e. they generate a series of steps, so that the temporal dimension of
+dispersal becomes explicit and allows modeling autocorrelation between
+successive steps \citep{Diniz.2019}. Finally, simulations from IBMMs do not
+enforce movement or connections towards preconceived endpoints but allow
 individuals to adjust their route  ``on the go'', thereby preventing biases
@@
     data and habitat covariates must be collected. The combined data is then
-    analyzed using an integrated step selection model. The parametrized model is
-    then treated as an individual-based movement model and used to simulate
-    dispersal trajectories. Ultimately, simulated trajectories are translated
-    into a set of maps that are pertinent to landscape connectivity. This
-    includes a heatmap, indicating the traversal frequency across each spatial
-    unit of the study area, a betweenness map, highlighting movement corridors
-    and bottlenecks, and, finally, an inter-patch connectivity map, where the
-    frequency of connections and their average duration can be depicted.}
+    analyzed using an integrated step selection model (step 1). The parametrized
+    model is then treated as an individual-based movement model and used to
+    simulate dispersal trajectories (step 2). Ultimately, simulated trajectories
+    are translated into a set of maps that are pertinent to landscape
+    connectivity (step 3). This includes a heatmap, indicating the traversal
+    frequency across each spatial unit of the study area, a betweenness map,
+    highlighting movement corridors and bottlenecks, and, finally, an
+    inter-patch connectivity map, where the frequency of connections and their
+    average duration can be depicted.}
     \label{GraphicalAbstract}
@@
 step-durations (i.e. 4 hours) for further analysis. We will refer to these steps
-as ``realized steps''. Additional details on the data collection and preparation
-can be found in \cite{Cozzi.2020} and \cite{Hofmann.2021}.
+as ``realized steps''. We did not differentiate between sexes, for previous
+research found little differences between sexes during dispersal
+\citep{Woodroffe.2019, Cozzi.2020}. Additional details on the data collection
+and preparation can be found in \cite{Cozzi.2020} and \cite{Hofmann.2021}.
 
@@
 across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA,
-\Cref{StudyArea}a and b). The KAZA-TFCA is the world's largest transboundary
-conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
-and Zambia, thus hosting a rich diversity of landscapes, ranging from savannah
-to grassland and from dry to moist woodland habitats. In its center lies the
-Okavango Delta, a dominant hydro-geographical feature and the world's largest
-flood-pulsing inland delta. Large portions of the KAZA-TFCA are formally
-protected in the form of national parks (NPs) or other protected areas, yet a
-considerable portion of the landscape remains human-dominated (e.g. roads,
-agricultural sites, and settlements).
+\Cref{StudyArea}a and b) and encompassed a rectangular extent of roughly 1.3
+Mio. km\textsuperscript{2}. With an area of 520'000 km\textsuperscript{2}, the
+KAZA-TFCA is the world's largest transboundary conservation area and comprises
+parts of Angola, Botswana, Namibia, Zimbabwe, and Zambia, thus hosting a rich
+diversity of landscapes, ranging from savannah to grassland and from dry to
+moist woodland habitats. In its center lies the Okavango Delta, a dominant
+hydro-geographical feature and the world's largest flood-pulsing inland delta.
+Large portions of the KAZA-TFCA are formally protected in the form of national
+parks (NPs) or other protected areas, yet a considerable portion of the
+landscape remains human-dominated (e.g. roads, agricultural sites, and
+settlements).
 
@@
 covariates (\(x_1, x_2, ..., x_n\)) and on the animal's relative selection
-strengths \citealp{Avgar.2017} towards these covariates (\(\beta_1, \beta_2,
+strengths \citep{Avgar.2017} towards these covariates (\(\beta_1, \beta_2,
 ..., \beta_n\)):
@@
 started with the base model and iteratively increased model complexity by adding
-all possible interactions between movement and habitat covariates. We validated
-the predictive power of the most parsimonious model using k-fold
-cross-validation for case-control studies as described in \cite{Fortin.2009}.
-This validation proves a significant prediction in case the Spearman rank
-correlation coefficient of predicted step-ranks and associated frequencies under
-the movement model is significantly greater than under the assumption of random
-preferences (details in Appendix A2).
+all possible interactions between movement and habitat covariates. Given that
+the focus of our analysis lied on predicting dispersal patterns and all model
+candidates were biologically intuitive, we deemed the use of model selection
+appropriate. However, caution should be employed if causal relationships are of
+interest, as model selection may lead to biased parameter estimate
+\citep{Whittingham.2006}. We validated the predictive power of the most
+parsimonious model using k-fold cross-validation for case-control studies as
+described in \cite{Fortin.2009}. This validation attests significant prediction
+ability to the movement model if the model outperforms a random guess and
+systematically assigns low ranks (high selection scores) to observed steps
+(details in Appendix A2).
 
@@
 trajectories into a network, we followed \cite{BastilleRousseau.2018} and
-overlaid the study area (including the buffer) with a raster containing 5 km x 5
-km raster-cells, where the center of each raster-cell served as node in the
+overlaid the study area (including the buffer) with a raster containing 2.5 km x
+2.5 km raster-cells, where the center of each raster-cell served as node in the
 final network. To identify edges (i.e. connections) between the nodes, we used
 the simulated trajectories and determined all transitions occurring from one
-cell to another, as well as the frequency at which those transitions occurred.
-This resulted in an edge-list that we translated into a weighted network using
-the r-package {\tt igraph} \citep{Gabor.2006}. The final weight of each edge was
-determined by the frequency of transitions, yet because {\tt igraph} handles
-edge weights (\(\omega\)) as costs, we inverted the traversal-frequency through
-each raster-cell by applying \(\omega = \frac{mean(Traversal
-Frequency)}{Traversal Frequency_i}\). Consequently, regularly used edges
-received small weights (i.e. low costs) and vice versa. We used the weighted
-network to calculate betweenness scores for all network nodes.
+cell to another, as well as the frequency at which those transitions occurred
+(see also Appendix A4). This resulted in an edge-list that we translated into a
+weighted network using the r-package {\tt igraph} \citep{Gabor.2006}. The final
+weight of each edge was determined by the frequency of transitions, yet because
+{\tt igraph} handles edge weights (\(\omega\)) as costs, we inverted the
+traversal-frequency through each raster-cell by applying \(\omega =
+\frac{mean(Traversal Frequency)}{Traversal Frequency_i}\). Consequently,
+regularly used edges received small weights (i.e. low costs) and vice versa. We
+used the weighted network to calculate betweenness scores for all network nodes.
 
@@
 other areas is impossible.
+
+\subsubsection{Validation}
+To validate our predictions of connectivity, we utilized additional dispersal
+data that was collected on eight dispersing coalitions between the years 2019
+and 2022 (totalling to 2'668 GPS locations). We used path selection analysis
+(PSF, \citealp{Cushman.2010}) to assess if observed dispersal trajectories
+followed areas of high predicted connectivity. Similar to SSF, PSF enables to
+detect selection for certain features by comparing observed paths to randomly
+generated paths. Here, we paired each observed path with 50 random paths that we
+generated by randomly rotating and shifting observed paths by a random angle
+\(\alpha \sim U(-\pi, +\pi)\) and a random distance \(d \sim U(0 \) km, \(50\)
+km\(\)). Along each path, we then extracted connectivity values from the heatmap
+(see above) generated after 68, 125, 250, 500, and 2'000 simulated steps,
+respectively. Finally, we ran conditional logistic regression to contrast
+observed and random paths. In case of systematic selection for high-connectivity
+areas, the regression coefficients from the corresponding conditional logistic
+regression model should be positive.
 
@@
 (\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
-\citealp{Hofmann.2021})
+\citealp{Hofmann.2021}). Our validation of the resulting connectivity maps using
+independent dispersal data showed that dispersers preferentially followed areas
+of high predicted connectivity, as coefficients from the PSF models were all
+significantly greater than zero (\Cref{MovementModel}c). The movement model thus
+successfully predicted functional connectivity.
 
 Plots that aid with the interpretation of the most parsimonious movement model
-are provided in Figure S2 and suggest that, under average conditions, dispersing
+are provided in Figure S3 and suggest that, under average conditions, dispersing
 wild dogs avoided moving through water, woodlands, and areas dominated by
@@
     Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
-    Results from the k-fold cross validation procedure. The upper plot shows
-    rank frequencies of realized steps according to model predictions with known
-    preferences, whereas the lower plot shows rank frequencies of realized steps
+    Results from the k-fold cross validation procedure. Subfigure b1 shows rank
+    frequencies of realized steps according to model predictions with known
+    preferences, whereas subfigure b2 shows rank frequencies of realized steps
     when assuming random preferences. The blue ribbon shows the prediction
@@
     associated frequency in (b1) highlights that the most parsimonious model
-    successfully outperformed a random guess (b2) and assigned high selection
-    scores to realized steps.}
+    successfully outperformed a random guess (b2) and frequently assigned low
+    ranks (i.e. high selection scores) to realized steps but only rarely high
+    ranks (i.e. low selection scores). (c) Results from the PSF analysis using
+    independent dispersal data show that dispersers preferrably moved through
+    areas where our heatmaps predicted high connectivity. Results are shown for
+    heatmaps realized after 68, 125, 250, 500, and 2'000 simulated steps,
+    respectively.}
     \label{MovementModel}
@@
 all checkpoints showed that the relative traversal frequency reached a steady
-state after 10,500 simulated dispersal trajectories (Figure S3). Although
+state after 10,500 simulated dispersal trajectories (Figure S4). Although
 variability in the relative traversal frequency kept decreasing as we increased
 the number of simulated dispersers, the marginal benefit of simulating
-additional trajectories diminished quickly (Figure S3).
+additional trajectories diminished quickly (Figure S4).
 
@@
 traversed by dispersers (median traversal frequency inside KAZA-TFCA = 166, IQR
-= 274, Figure S6a), whereas areas beyond the KAZA-TFCA boundary were
+= 274, Figure S7a), whereas areas beyond the KAZA-TFCA boundary were
 comparatively rarely visited (median traversal frequency outside KAZA-TFCA = 61,
-IQR = 133, Figure S6a). Most notably, the region in northern Botswana south of
+IQR = 133, Figure S7a). Most notably, the region in northern Botswana south of
 the Linyanti swamp appeared to serve as highly frequented dispersal hotspot
@@
 landscapes of Zambia and Zimbabwe were only rarely used for dispersal, even
-within the KAZA-TFCA boundaries (Figure S7a). Despite the fact that the heatmap
+within the KAZA-TFCA boundaries (Figure S8a). Despite the fact that the heatmap
 improved our understanding of the frequency at which areas were traversed by
@@
   NPs (dark gray). Additional heatmaps showing the traversal frequency when
-  individuals move fewer than 2'000 steps are provided in Figure S4.}
+  individuals move fewer than 2'000 steps are provided in Figure S5.}
   \label{Heatmap}
@@
 KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR = 54.311 \(\times\)
-10\textsuperscript{6}, Figure S6b). Consequently, only few corridors directly
+10\textsuperscript{6}, Figure S7b). Consequently, only few corridors directly
 linked the peripheral regions of the KAZA-TFCA and passed through unprotected
@@
 10\textsuperscript{6}, IQR = 9.891 \(\times\) 10\textsuperscript{6}, Figure
-S6b).
+S7b).
 
@@
   betweenness maps showing betweenness scores when individuals move fewer than
-  2'000 steps are provided in Figure S4.}
+  2'000 steps are provided in Figure S6.}
   \label{Betweenness}
@@
 originating from a patch in these countries successfully moved into at least on
-other patch (Figure S8a). Conversely, only 19\% of the dispersers leaving from a
-patch in Zambia managed to find their way into some other patch (Figure S8b).
+other patch (Figure S9a). Conversely, only 19\% of the dispersers leaving from a
+patch in Zambia managed to find their way into some other patch (Figure S9b).
 Prior to reaching another patch, individuals from Angola, Namibia, Botswana,
@@
 
+% \subsection{Structural vs. Functional Connectivity}
+% Our approach enabled us to translate a simple set of small-scale behavioral
+% rules into large scale patterns of connectivity, something previously deemed
+% computationally unfeasible, yet critical for linking structural to functional
+% connectivity \citep{Doerr.2011}. Structural connectivity focuses purely on the
+% spatial arrangement of suitable habitat in the landscape, whereas functional
+% connectivity also takes into account a species dispersal ability and behavioral
+% response to the landscape \citep{Tischendorf.2000}. Functional connectivity is
+% of greater interest in conservation science, yet much more difficult to measure
+% \citep{Baguette.2013}, which is why structural connectivity often serves as
+% surrogate., by contrast, is easier to measure and contributes significantly to
+% functional connectivity, which is why it often serves as surrogate for
+% functional connectivity \cite{Doerr.2011, Fattebert.2015}. Through incorporation
+% of a species habitat preferences into a permeability surface, both LCPA and CT
+% allow
+
+Our approach enabled us to translate a simple set of small-scale behavioral
+rules into large scale patterns of connectivity, something previously deemed
+computationally unfeasible, yet critical for linking structural and functional
+connectivity \citep{Doerr.2011}. Structural connectivity focuses purely on the
+spatial arrangement of suitable habitat in the landscape, whereas functional
+connectivity also takes into account a species dispersal ability and behavioral
+response to the landscape \citep{Tischendorf.2000}. Functional connectivity is
+of greater interest to conservation scientists, yet is difficult to quantify
+\citep{Baguette.2013}, which is why structural connectivity often serves as
+surrogate \cite{Doerr.2011, Fattebert.2015}. LCPA and CT incorporate functional
+aspects of connectivity through the permeability surface, which reflects a
+species habitat preferences and thus renders behavioral impacts of the lanscape
+on the focal species. Aside from rendering habitat preferences, our model also
+integrates peculiarities of the focal species movement behavior, thus adding
+further insights on functional connectivity. In addition, we could use
+independent dispersal data to prove that our predictions of connectiviy aligned
+with observed functional connectivity patterns.
+
+% If I were to force a couple of comments it would be how the authors discussed
+% the shift from a structural to a functional view on connectivity (lines 38-39).
+% For one, using the integrative approach appears to me as a way where the
+% movement and habitat kernels (and the interaction between them) inform the
+% relationship between structural connectivity (the landscape) and functional
+% connectivity (dispersal), which I believe is the core of the Tishendorf & Fahrig
+% (2000) paper. Consequently, I don’t see this as a shift, but more an
+% approximation to understand how they are related. Two, because understanding the
+% relationship between structural and functional connectivity is a key goal in
+% landscape ecology, I feel the authors could have discussed the implication of
+% this approach to this end.
+
 % \subsection{Disadvantages of ISSF Simulations}
@@
 the massive extent of the study area considered (ca. 1.8 Mio
-km\textsuperscript{2}) and the large number of simulated trajectories. Most
-connectivity studies focus on smaller study areas (e.g. \citealp{Kanagaraj.2013,
-Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020}) and will therefore require
-fewer simulations and achieve faster simulation times (given the same spatial
-resolution). We also believe that fewer simulated trajectories will often
-suffice, as the relative traversal frequency by simulated trajectories through
-randomly placed checkpoints across our study area converged already after 10,500
-runs. The exact number of required simulations to achieve reliable estimates of
-connectivity will, of course, vary depending on the structure of the landscape
-and the dispersal capabilities of the focal species \citep{Gustafson.1996}. For
-species that disperse short distances through homogeneous environments, few
-simulations may suffice to gauge connectivity, whereas for species that disperse
-over long distances through heterogeneous habitats, a large number of
-simulations will be required to sufficiently explore the spectrum of possible
-routes.
+km\textsuperscript{2}), the large number of simulated trajectories, and the fact
+that we extracted covariates along each step, rather than just at their start or
+endpoints. Most connectivity studies focus on smaller study areas (e.g.
+\citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020})
+and will therefore require fewer simulations and achieve faster simulation times
+(given the same spatial resolution). We also believe that fewer simulated
+trajectories will often suffice, as the relative traversal frequency by
+simulated trajectories through randomly placed checkpoints across our study area
+converged already after 10,500 runs. The exact number of required simulations to
+achieve reliable estimates of connectivity will, of course, vary depending on
+the structure of the landscape and the dispersal capabilities of the focal
+species \citep{Gustafson.1996}. For species that disperse short distances
+through homogeneous environments, few simulations may suffice to gauge
+connectivity, whereas for species that disperse over long distances through
+heterogeneous habitats, a large number of simulations will be required to
+sufficiently explore the spectrum of possible routes. Finally, it may often
+suffice to extract covariates at each step's start or endpoints, thus
+considerably speeding up simulation times \citep{Signer.2017}.
 
@@
 protected areas, so we placed source points randomly within them. In cases where
-more detailed data distribution and abundance are available, source points could
-be distributed accordingly. Alternatively, source points could be distributed
-homogeneously but later be weighted when computing connectivity metrics. In any
-case, the challenge of selecting meaningful source points is not unique to
-individual-based simulations but also applies to LCPA and CT.
+more detailed data about the distribution and abundance of the focal species are
+available, source points could be distributed accordingly. Alternatively, source
+points could be distributed homogeneously but later be weighted when computing
+connectivity metrics. In any case, the challenge of selecting meaningful source
+points is not unique to individual-based simulations but also applies to LCPA
+and CT.
 
@@
 simulated trajectories after their simulation and thereby to investigate the
-sensitivity of our results with respect to exact dispersal durations (Figures S4
-and S5).
+sensitivity of our results with respect to exact dispersal durations (Figures S5
+and S6).
 
@@
 impossible to compute a selection score. A possible solution is to simply
-terminate the simulation affected trajectory, assuming that the simulated
+terminate the simulation of the affected trajectory, assuming that the simulated
 individual has left the study area. However, this approach might produce

### 2022-11-18 Chapter_1/04_Manuscript/Manuscript_2022-11-07.tex
Revisions done

--- 2022-04-29 Chapter_1/04_Manuscript/Manuscript.tex
@@
 \usepackage[colorinlistoftodos]{todonotes}
-\usepackage{scalerel}
-\usepackage{tikz}
-\usetikzlibrary{svg.path}
 
@@
 
-% To be able to put an ORCID
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-\title{A Three-Step Approach for Assessing Landscape Connectivity via Simulated
+\title{A Thee-Step Approach for Assessing Landscape Connectivity via Simulated
 Dispersal: African Wild Dog Case Study}
@@
 \author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
+  David D. Hofmann\textsuperscript{1,2,\S} \and
+  Gabriele Cozzi\textsuperscript{1,2} \and
   John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
+  Arpat Ozgul\textsuperscript{1} \and
+  Dominik M. Behr\textsuperscript{1,2}
 }
@@
 %------------------------------------------------------------------------------
-\noindent \textbf{Context} \quad Dispersal of individuals contributes to
-long-term population persistence, yet requires a sufficient degree of landscape
-connectivity. To date, connectivity has mainly been investigated using
-least-cost analysis and circuit theory, two methods that make assumptions that
-are hardly applicable to dispersal. While these assumptions can be relaxed by
-explicitly simulating dispersal trajectories across the landscape, a unified
-approach for such simulations is lacking.
-
-\noindent \textbf{Objectives} \quad Here, we propose and apply a simple
-three-step approach to simulate dispersal and to assess connectivity using
-empirical GPS movement data and a set of habitat covariates.
-
-\noindent \textbf{Methods} \quad In step one of the proposed approach, we use
-integrated step-selection functions to fit a mechanistic movement model
-describing habitat and movement preferences of dispersing individuals. In step
-two, we apply the parameterized model to simulate dispersal across the study
-area. In step three, we derive three complementary connectivity maps; a heatmap
-highlighting frequently traversed areas, a betweenness map pinpointing dispersal
-corridors, and a map of inter-patch connectivity indicating the presence and
-intensity of functional links between habitat patches. We demonstrate the
-applicability of the proposed three-step approach in a case study in which we
-use GPS data collected on dispersing African wild dogs (\textit{Lycaon pictus})
-inhabiting northern Botswana.
-
-\noindent \textbf{Results} \quad Using step-selection functions we successfully
-parametrized a detailed dispersal model that described dispersing individuals'
-habitat and movement preferences, as well as potential interactions among the
-two. The model substantially outperformed a model that omitted such interactions
-and enabled us to simulate 80,000 dispersal trajectories across the study area.
-
-\noindent \textbf{Conclusion} \quad By explicitly simulating dispersal
-trajectories, our approach not only requires fewer unrealistic assumptions about
-dispersal, but also permits the calculation of multiple connectivity metrics
-that together provide a comprehensive view of landscape connectivity. In our
-case study, the three derived connectivity maps revealed several wild dog
-dispersal hotspots and corridors across the extent of our study area. Each map
-highlighted a different aspect of landscape connectivity, thus emphasizing their
-complementary nature. Overall, our case study demonstrates that a
-simulation-based approach offers a simple yet powerful alternative to
-traditional connectivity modeling techniques. It is therefore useful for a
-variety of applications in ecological, evolutionary, and conservation research.
+\noindent 1. Dispersal of individuals contributes to long-term population
+persistence, yet requires a sufficient degree of landscape connectivity. To
+date, connectivity has mainly been investigated using least-cost analysis and
+circuit theory, two methods that make assumptions that are hardly applicable to
+dispersal. While these assumptions can be relaxed by explicitly simulating
+dispersal trajectories across the landscape, a unified approach for such
+simulations is lacking.
+
+\noindent 2. Here, we present a simple three-step approach to simulate dispersal
+and to assess connectivity using empirical GPS movement data and a set of
+habitat covariates. In step one, we use integrated step-selection functions to
+fit a mechanistic movement model describing habitat and movement preferences of
+dispersing individuals. In step two, we apply the parameterized model to
+simulate dispersal across the study area. In step three, we derive three
+complementary connectivity maps; a heatmap highlighting frequently traversed
+areas, a betweenness map pinpointing dispersal corridors, and a map of
+inter-patch connectivity indicating the presence and intensity of functional
+links between habitat patches. We demonstrate the applicability of the proposed
+three-step approach in a case study in which we use GPS data collected on
+dispersing African wild dogs (\textit{Lycaon pictus}) inhabiting northern
+Botswana.
+
+\noindent 3. Using step-selection functions we successfully parametrized a
+detailed dispersal model that described dispersing individuals habitat and
+movement preferences, as well as potential interactions among the two. The model
+substantially outperformed a model that omitted such interactions and enabled us
+to simulate 80,000 dispersal trajectories across the study area.
+
+\noindent 4. By explicitly simulating individuals, our approach not only
+requires fewer unrealistic assumptions about dispersal, but also permits the
+calculation of multiple connectivity metrics that together provide a
+comprehensive view of landscape connectivity. In our case study, the three
+derived connectivity maps revealed several wild dog dispersal hotspots and
+corridors across the extent of our study area. Each map highlighted a different
+aspect of landscape connectivity, thus emphasizing their complementary nature.
+Overall, our case study demonstrates that a simulation-based approach offers a
+simple yet powerful alternative to traditional connectivity modeling techniques.
+It is therefore useful for a variety of applications in ecological,
+evolutionary, and conservation research.
 

### 2022-11-18 Chapter_1/04_Manuscript/Manuscript_Support.tex
Revisions done

@@
 %------------------------------------------------------------------------------
-%	Appendix S4: All Movement Models
+%	Appendix S4: Betweenness Approaches
+%------------------------------------------------------------------------------
+\newpage
+\section{Betweenness Maps}
+To generate betweenness maps, we first overlaid the study area with a regular
+grid with a resolution of 2.5 km x 2.5 km per grid cell. The grid cells were
+labelled with unique numbers, which allowed us to then generate a ``visitation
+history'' for each simulated trajectory, showing through which grid cells each
+simulated invidual moved. Based on this history, we tallied the number of
+transitions that occured from one cell to another. Ultimately, the so generated
+list of transitions served to generate a network and to compute betweenness
+scores. The resolution of the overlaid grid thus dictates the resolution at
+which betweenness can be mapped and has important implications for the way in
+which cell-transitions need to be determined. In the simplest case,
+cell-transitions can be determined by only looking at the start and endpoint of
+each step. If it ends in a different grid-cell than it started, a transition has
+occured. This approach is computationally simple as it only requires a point
+extraction (at the start and endpoint of each step). However, relying on a point
+extraction has the disadvantage that transitions between non-adjacent cells can
+occur if a single step stretches across multiple grid-cells. Especially when the
+resolution of the grid is high, this problem becomes more pertinent. As an
+alternative, one could not only consider the start and endpoint of each step,
+but the entire line along the step, i.e. conduct a line extraction. However,
+this requires that extracted values (i.e. grid-cell labels) are ordered in
+accordance to the direction of movement. In R, such an extraction can be
+achieved e.g. using \textsf{raster::extract(raster, line, along = T)}, which is
+computationally very demanding and therefore often an unviable solution. Here,
+we employed yet another approach and generated interpolated coordinates along
+each simulated step. We then determined cell transition using those interpolated
+coordinates. Since this approach still uses a point extraction, it is
+considerably faster while still enabling to detect cell-transitions at greater
+resolution. In \Cref{BetweennessApproaches}, we use simulated movement paths to
+compare betweenness maps that are based on a (1) point extraction, a (2) line
+extraction, and a (3) point extraction with interpolated coordinates. As you can
+see, the approach using interpolated coordinates approximates the results from
+the line extraction very closely, even at higher resolutions. A detailed
+discussion about this is can also be found in \citep{BastilleRousseau.2018}.
+
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_BetweennessApproaches.png}
+  \caption{We simulated five trajectories where individuals moved through a
+  narrow corridor. We then overlaid those simulations with a coarse (a1) and
+  detailed (b1) grid that was used to compute betweenness scores. This required
+  us to determine cell-transitions, which we did using three competing
+  approaches. Once using a point extraction (a2 and b2), once using a line
+  extraction (a3 and b3), and once using a point extraction with interpolated
+  coordinates (a4 and b4). In brackets are the average computation times $\pm$
+  SD as computed from 10 repetitions.}
+  \label{BetweennessApproaches}
+ \end{center}
+\end{figure}
+
+%------------------------------------------------------------------------------
+%	Appendix S5: All Movement Models
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S5: Most Parsimonious Movement Models
+%	Appendix S6: Most Parsimonious Movement Models
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S6: Movement Model interpretation
+%	Appendix S7: Movement Model interpretation
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S7: Convergence
+%	Appendix S8: Convergence
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S8: Heatmaps
+%	Appendix S9: Heatmaps
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S9: Evolution of Betweenness Maps
+%	Appendix S10: Evolution of Betweenness Maps
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S11: Map Comparison
+%	Appendix S12: Map Comparison
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-%	Appendix S12: Areas Reached
+%	Appendix S13: Areas Reached
 %------------------------------------------------------------------------------

### 2022-11-18 Chapter_1/02_R-Scripts/03_Visualization/01_VisualizationsForManuscript/MovementModelResults.R

@@ -36,16 +36,20 @@ options(scipen = 999)
 #### Prepare Data
 ################################################################################
 # Load the movement model
-move.mod <- readRDS("03_Data/03_Results/99_MovementModel.rds")
+best <- readRDS("03_Data/03_Results/99_MovementModel.rds")
+best <- best$Model[[1]]
 
-# Select best model
-best <- move.mod$Model[[1]]
-
-# Summary of the best model
-summary(best)
+# Extract model results (note that newer versions of glmmTMB and TMB are not
+# compatible with older models and we need to manually extract the values of
+# interest)
+best <- tibble(
+    Covariate   = names(fixef(best)$cond)
+  , Coefficient = fixef(best)$cond
+  , SE          = sqrt(diag(vcov(best, full = T)))[1:length(Coefficient)]
+)
 
 # Calculate confidence intervals
-coeffs <- getCoeffs(best, pvalue = TRUE)[-1, ] %>%
+coeffs <- best[-1, ] %>%
   mutate(
       LCI_90 = Coefficient - qnorm(1 - (1 - 0.90) / 2) * SE
     , UCI_90 = Coefficient + qnorm(1 - (1 - 0.90) / 2) * SE
@@ -55,13 +59,16 @@ coeffs <- getCoeffs(best, pvalue = TRUE)[-1, ] %>%
     , UCI_99 = Coefficient + qnorm(1 - (1 - 0.99) / 2) * SE
   )
 
+# Compute p-values
+coeffs <- mutate(coeffs, pvalue = 2 * pnorm(-abs(Coefficient) / SE))
+
 # Add stars indicating the significance
 coeffs$Significance <- sapply(1:nrow(coeffs), function(x){
   if (coeffs$pvalue[x] <= 0.01){
     return("***")
-  } else if (coeffs$pvalue[x] <= 0.05){
+  } else if (coeffs$pvalue[x] <= 0.05) {
     return("**")
-  } else if (coeffs$pvalue[x] <= 0.1){
+  } else if (coeffs$pvalue[x] <= 0.1) {
     return("*")
   }
 })
@@ -372,9 +379,10 @@ print(xtable(tex, digits = 3)
 #### Validation
 ################################################################################
 # Load required data
-validation <- read_rds("03_Data/03_Results/99_ModelValidation.rds")
-dat_pref <- read_rds("03_Data/03_Results/99_ModelValidation(Data).rds")[[1]]
[…]

### 2022-11-29 Chapter_1/04_Manuscript/Manuscript.tex
Incorporated review feedback

@@
 We performed all data preparations, spatial computations, and statistical
-analysis in {\tt R}, version 3.6.6 \citep{R.2020}. Some helper functions were
+analysis in {\tt R}, version 4.2.2 \citep{R.2022}. Some helper functions were
 written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
@@
 circles''. To achieve high performance rasterization, we used the R-package {\tt
-terra} \citep{Hijmans.2021b}.
+terra} \citep{Hijmans.2021b}. For a subset of the study area, we also generated
+heatmaps at 250 m x 250 m, yet found little qualitative differences to the
+coarser resolution, thus suggesting the choice of 1 km x 1 km to be appropriate.
 
@@
 To validate our predictions of connectivity, we utilized additional dispersal
-data that was collected on eight dispersing coalitions between the years 2019
-and 2022 (totalling to 2'668 GPS locations). We used path selection analysis
-(PSF, \citealp{Cushman.2010}) to assess if observed dispersal trajectories
-followed areas of high predicted connectivity. Similar to SSF, PSF enables to
-detect selection for certain features by comparing observed paths to randomly
-generated paths. Here, we paired each observed path with 50 random paths that we
-generated by randomly rotating and shifting observed paths by a random angle
-\(\alpha \sim U(-\pi, +\pi)\) and a random distance \(d \sim U(0 \) km, \(50\)
-km\(\)). Along each path, we then extracted connectivity values from the heatmap
-(see above) generated after 68, 125, 250, 500, and 2'000 simulated steps,
-respectively. Finally, we ran conditional logistic regression to contrast
-observed and random paths. In case of systematic selection for high-connectivity
-areas, the regression coefficients from the corresponding conditional logistic
-regression model should be positive.
+data that was collected on eight dispersing coalitions between 2019 and 2022
+(totalling to 2'668 GPS locations). We used path selection analysis (PSF,
+\citealp{Cushman.2010}) to assess if observed dispersal trajectories followed
+areas of high predicted connectivity. Similar to SSF, PSF enables to detect
+selection for certain features by comparing observed paths to randomly generated
+paths. Here, we paired each observed path with 50 random paths that we generated
+by randomly rotating and shifting observed paths by a random angle \(\alpha \sim
+U(-\pi, +\pi)\) and a random distance \(d \sim U(0 \) km, \(50\) km\(\)). Along
+each path, we then extracted connectivity values from the heatmap (see above)
+generated after 68, 125, 250, 500, and 2'000 simulated steps, respectively.
+Finally, we ran conditional logistic regression to contrast observed and random
+paths. In case of systematic selection for high-connectivity areas, the
+regression coefficients from the corresponding conditional logistic regression
+model should be positive.
 
@@
 \citep{VanDerMeer.2014} or areas with a high susceptibility for human wildlife
-conflicts \citep{Cushman.2018}.
+conflicts \citep{Cushman.2018}. Using independent dispersal data we showed that
+dispersers indeed followed areas of high predicted connectivity. Importantly,
+however, these predictions were based on a scenario of a relatively extended
+flood, which may not have accurately represented environmental conditions for
+dispersers moving through areas affected by the flood. Accounting for such
+differences would have improved the predictive performance of our model.
 
@@
 \citep{Baguette.2013}, which is why structural connectivity often serves as
-surrogate \cite{Doerr.2011, Fattebert.2015}. LCPA and CT incorporate functional
+surrogate \citep{Doerr.2011, Fattebert.2015}. LCPA and CT incorporate functional
 aspects of connectivity through the permeability surface, which reflects a
@@
 integrates peculiarities of the focal species movement behavior, thus adding
-further insights on functional connectivity. In addition, we could use
+further insights on functional connectivity. In addition, we successfully used
 independent dispersal data to prove that our predictions of connectiviy aligned
@@
 with 3.6 GHz, 64 GB of RAM). The long simulation time was primarily caused by
-the massive extent of the study area considered (ca. 1.8 Mio
+the massive extent of the study area considered (ca. 1.3 Mio
 km\textsuperscript{2}), the large number of simulated trajectories, and the fact

### 2022-11-29 Chapter_8/04_Manuscript/Manuscript.tex
Updated betweenness maps

@@
 Climate change is expected to critically impact the flood-pulsing cycles of the
-Okavango delta and may therefore critically alter landscape connectivity in this
-area. Besides a better understanding of the conservation needs for the African
-wild dog, our study also presents a first step towards incorporating
+Okavango delta and may therefore substantially alter landscape connectivity in
+this area. Besides a better understanding of the conservation needs for the
+African wild dog, our study also presents a first step towards incorporating
 environmental changes due to seasonality or climate change in dispersal and

### 2022-12-08 Chapter_10/02_R-Scripts/01_Analysis/04_Simulation.R
Various

@@ -1,0 +1,439 @@
+################################################################################
+#### Simulation
+################################################################################
+# Description: Simulation of dispersal
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_10")
+
+# Load required packages
+library(raster)         # To handle spatial data
+library(terra)          # To handle spatial data
+library(velox)          # For quick extraction
+library(tidyverse)      # To wrangle data
+library(lubridate)      # To handle dates
+library(glmmTMB)        # To handle the movement model
+library(Rcpp)           # To load C++ functions
+library(pbmcapply)      # For multicore use with progress bar
+
+# Load custom functions
+source("02_R-Scripts/00_Functions.R")
+sourceCpp("02_R-Scripts/00_Functions.cpp")
+
+################################################################################
+#### Load Required Data
+################################################################################
+# Load the calibrated movement model
+mod <- read_rds("03_Data/02_CleanData/MovementModel.rds")
+
+# Load the scaling parameters
+scaling <- read_rds("03_Data/02_CleanData/Scaling.rds")
+
+# Load the fitted gamma distribution (for sampling step lengths)
+sl_dist <- read_rds("03_Data/02_CleanData/GammaDistribution.rds")
+
+# Load the required spatial layers layers
+layers <- rast("03_Data/02_CleanData/SpatialLayers.tif")
+
+# Simulate hrs
+
+
+
+hrs <- simHR(n = 500, m = , r = layers, buff = , min_area = 50, max_area = 1800)
+
+# Extend the covariates artificially (fill added buffer with random values)
+cat("Extending covariate layers...\n")
+ext <- ext(layers) * 1.2
+cov <- extendRaster(layers, ext)
+
+# Apply our custom functions to make everything ready for the simulation
+mod  <- prepareModel(mod)
+cov  <- prepareCovars(layers)
+
+################################################################################
+#### Simulation Function
+################################################################################
+# Function to simulate dispersal based on a step selection model that was fitted
+# in the glmmTMB framework
+disperse <- function(
+    source              = NULL    # Start Coordinates
[…]

### 2022-12-08 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/DispersalPaths.r

@@ -113,8 +113,8 @@ p3 <- ggplot(sat, aes(x = x, y = y, fill = rgb(red, green, blue, maxColorValue =
   geom_raster() +
   geom_point(data = dat, inherit.aes = F, aes(x = x, y = y, col = DogName), size = 0.4) +
   geom_path(data = dat, inherit.aes = F, aes(x = x, y = y, col = DogName), size = 0.3) +
-  geom_sf(data = subset(prot, Name %in% c("Hwange", "Moremi")), fill = "white", col = "white", alpha = 0.2, inherit.aes = F) +
-  geom_sf(data = afri, fill = NA, col = "white", inherit.aes = F, lty = 2) +
+  # geom_sf(data = subset(prot, Name %in% c("Hwange", "Moremi")), fill = "white", col = "white", alpha = 0.2, inherit.aes = F) +
+  geom_sf(data = afri, fill = NA, col = "white", inherit.aes = F, lty = 1) +
   scale_fill_identity() +
   coord_sf(crs = "+init=epsg:4326") +
   theme_minimal() +

### 2022-12-22 Chapter_8/04_Manuscript/Manuscript.tex
upd

@@
 patterns and connectivity of the endangered African wild dog (\textit{Lycaon
-pictus}). For this, we simulate dispersal trajectories across the landscape
-under two extreme environmental scenarios; one assuming a maximum flood extent,
-one assuming a minimum flood extent.
+pictus}). For this, we simulate dispersal trajectories across the landscape of
+the Okavango Delta under two extreme environmental scenarios; one assuming a
+maximum flood extent, one assuming a minimum flood extent.
 
@@
 minimum flooding, conversely, the delta reveals vital dispersal corridors and
-increases chances of successful dispersal into neighboring habitat patches.
-Concomittantly, dispersal durations to move between patches is reduced at low
-flood levels.
+increases chances of successful dispersal into neighboring habitat patches by xx
+percent. Concomitantly, dispersal durations to move between patches is reduced,
+on average by xx days,at low flood levels.
 
@@
 Okavango delta and may therefore substantially alter landscape connectivity in
-this area. Besides a better understanding of the conservation needs for the
-African wild dog, our study also presents a first step towards incorporating
-environmental changes due to seasonality or climate change in dispersal and
-connectivity analyses. Finally, our analysis demonstrates the usefulness of
-individual-based dispersal simulations as a pertinent conservation tool to study
-impacts of environmental change while accounting for dispersal movements.
+this area. Our analysis suggests that anticipating such alterations in
+environmental conditions is key when implementing conservation strategies.
+Besides a better understanding of the conservation needs for the African wild
+dog, our study also presents a first step towards incorporating environmental
+changes due to seasonality or climate change in dispersal and connectivity
+analyses and demonstrates the usefulness of individual-based dispersal
+simulations as a pertinent conservation tool to study impacts of environmental
+change on dispersal success.
 
@@
 \section{Introduction}
-\subsection{Climate Change}
-Climate change is expected to profoundly impact ecosystems worldwide, with
-far-reaching consequences for species living therein (cite IPCC)
-\citep{Ozgul.2010, Radchuk.2019}. However, empirical information on species' responses to climate change remains
-scarce, especially for regions most vulnerable to altered conditions
+Climate change is expected to profoundly impact ecosystems worldwide with
+far-reaching consequences for species living therein \citep{Ozgul.2010,
+Radchuk.2019, IPCC.2022}. By altering environmental conditions, climate change
+affects animal behavior \citep{Fuller.2016}, resource availability
+\citep{Durant.2007}, population dynamics \citep{Paniw.2021}, and the
+distribution of wild living animal populations \citep{Thomas.2004,
+Thuiller.2006}. One potential life-history pathway through which species may
+mediate the effects of environmental change is dispersal \citep{Anderson.2012}.
+By bringing individuals away from their natal location \citep{Clobert.2012},
+dispersal equips species with a means to adapt to climate change by tracking
+favorable habitat conditions \citep{Raia.2012} and to shift into a different
+part of their fundamental niches \citep{Kokko.2022}. Dispersal furthermore
+provides resilience against changing conditions by ensuring genetic diversity
+\citep{Frankham.2002, Leigh.2012, Baguette.2013}, facilitating the colonization
+empty habitats \citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and
+promoting the reinforcement of weakened and small subpopulations
+\citep{Brown.1977}. As such, it is little surpsing that improving connectivity
+is viewed as the number one strategy to combat the adverse effects of climate
+change \citep{Heller.2009}.
+
+While dispersal offers a means to offset the demographic consequences of climate
+change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a function of
+climatic and environmental conditions (e.g. \citep{Behr.2020}). The link between
+dispersal and climaticor environmental can either be indirect, for example if
+the propensity of individuals to disperse depends on environmental conditions,
+or direct, when the biophysical environment through which dispersers move
+affects dispersal prospects \citep{Travis.2013}. As the latter case indicates,
+dispersal is also inextricably linked to the concept landscape connectivity
+\citep{Baguette.2013}, which is understood as the degree by which the landscape
+facilitates or impedes movements \citep{Taylor.1993}. Only if a sufficient
+degree of landscape connectivity is maintained, successful dispersal of
+individuals can be guaranteed \citep{Fahrig.2003}. Unfortunately, our
+understanding how changing environmental conditions influence connectivity and
+dispersal is limited \citep{Kokko.2006, Travis.2013}. Because collecting
+information on dispersing animals is difficult and because it is difficult to
+assess how changing environmental conditions affect connectivity...
+
+The continued degradation and destruction habitats across the globe impedes
+dispersal, promotes inbreeding and genetic drift and increases the
+likelihood of local extirpation due to stochastic events \citep{Sawyer.2011}.
+Combined, these factors limit species ability to adapt to changing environmental
+conditions \citep{Fahrig.2003}.
+
+\cite{Ashrafzadeh.2019} found that cimate change is likely to substantially
+reduce connectivity for mountain newts in Iran. Similarly, \cite{Luo.2021}
+predicted significant reductions in connectivity for the giant spiny frog in
+South-East Asia.
+
+However, failing to take into account potential changes in climatic and
+environmental conditions imperils our ability to design cost-effective
+conservation strategies.
+
+Here, we used naturally occuring seasonal flood-cycles to approximate the likely
+changes in connectivity that will arise under amplified climatic conditions due
+to climate change.
+While a better understanding of the ODs flooding regime under global change is of major interest, reliable predictions remain challenging due to the complex interplay between climate,
+anthropogenic water use, animal behavior, and topography of the region.
+
+Climate change
+could lead to the disruption of existing dispersal corridors and the emergence
+of new dispersal barriers but may also provide possibilities to access
+previously unsuitable dispersal routes. Designing and implementing effective
+conservation strategies that aim to improve landscape connectivity, such as
+habitat restoration or the creation of wildlife corridors, requires better
+understanding of such dynamics.
+
+Albeit the impacts of climate change on connectivity cannot be directly
+estimated, one can study how changing environmental conditions due to
+seasonality influence patterns of dispersal and connectivity.
+
+Habitat degradation and fragmentation lead to isolated subpopulations that have
+a high associated risk of extinction due to stochastic events and generally
+decreased genetic variability \citep{Melbourne.2008}.
+
+More recently, individual-based movement models (IBMMs) have been uprising and
+are recommended for their ability to incorporate several crucial realities of
+animal behavior, such as, for example, autocorrelation, that are by omitted by
+necessity using resistance-based approaches \citep{Hofmann.2023,
+Unnithan.2022a}. IBMMs allow much greater flexibility and are thus considered
+powerful tools for examining connectivity under different landscape
+configurations \citep{Littlefield.2019}.
+
+Botswana is among the countries most heavily influenced by climate change .
+
+For multispecies approaches to connectivity within the KAZA-TFCA, see
+\cite{Brennan.2020} and \cite{Lines.2021}.
+
+Talk about connectivity models
+Talk about functional connectivity
+Talk about habitat fragmentation and destruction
+
+Osipova
+
+Unfortunately, little is known about the specific responses of different species
+to climate change, especially for regions most vulnerable to altered conditions
 \citep{Paniw.2021}.
 
-\subsection{Dispersal}
-The lack of knowledge in terms of species ability to cope with climate change is
-further compounded by a lack of information on dispersal under changing
-environmental conditions \citep{Travis.2013}. By bringing individuals away from
-their natal location, dispersal strongly influences population dynamics
-\citep{Clobert.2012}; it enables genetic exchange \citep{Frankham.2002,
-Leigh.2012, Baguette.2013}, facilitates the colonization of empty habitats
-\citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and promotes reinforcement
-of weakened and small subpopulations \citep{Brown.1977}. Dispersal also serves
-as a means to track suitable habitat \citep{Raia.2012} and may thus serve as a
-means to mediate the demographic consequences of environmental change
-\citep{Hodgson.2009, Travis.2013}. Dispersal is also inextricably linked to the
-concept of landscape connectivity, which is generally understood as the degree
-by which the landscape facilitates or impedes movements among habitat patches
-\citep{Tylor.1993}.
+Only if there
 
-As previously highlighted by \cite{Travis.2013}, climate may affect dispersal
-directly, by altering the propensity of individuals to disperse, or indirectly,
-through changes in the biophysical environment. Here, we will focus on the
-latter and study dispersal prospects under changing environmental conditions.
+The challenges associated with studying dispersing animals have thus far
+impaired the collection of data of dispersers at the appropriate temporal and
+spatial scale \citep{Graves.2014, Vasudev.2015} and weakened our ability to
+project dispersal prospects under changing environmental conditions into the
+future. Nevertheless, a more comprehensive understanding of dispersal under
+changing environmental conditions is critical for designing effective
+conservation strategies in the light of climate change. By predicting and
+modeling dispersal patterns under different scenarios, we can better anticipate
+the impacts of climate change on species and design effective conservation
+strategies to protect them. This is especially important for endangered species
+that are already at the verge of extinction.
 
-In spite of its importance for determining population dynamics and landscape
-connectivity, our understanding of dispersal and how dispersal is affected by
-changing environmental conditions is limited. This is owed to the difficulty of
-obtaining data collected on dispersing animals at the appropriate temporal and
-spatial scale \citep{Graves.2014, Vasudev.2015} and our inability to project
-dispersal prospects under changing environmental conditions into the future.
-
-The ability of species
-to persist under these changing conditions depends on the amplitude and speed of
-environmental change and species' responses to such changes (cite someone).
-
-It is, in fact, presumed that humanity has already transgressed the planetary
-boundary for climate change, therefore triggering irreversible feedbacks in the
-biosphere \citep{Rockstrom.2009}.
-
-Cite IPCC, Planetary boundaries, Tucker (step length). Phenological shifts. Cite
-Robynne, Arpat, Megan, Dominik, Rabaiotti etc.
-
-\subsection{Connectivity}
-Dispersal is an important, if not the important, driver of landscape
-connectivity and therefore of major interest to conservation authorities. It has
-also been demonstrated that dispersal may lead to population declines in areas
-where anthropogenic mortality is high and dispersal prospects low
-\citep{Leigh.2012}.
-
-\subsection{Okavango Delta and African Wild Dogs}
 The Okavango delta in Southern Africa poses a unique opportunity to study the
-impacts of environmetal change on species dispersal ability and connectivity in
+impacts of environmental change on species dispersal ability and connectivity in a
 large scale natural experiment setup. According to projections by the IPCC,
@@
 
-\subsection{Our Study}
-Here, we capitalize on a previously parametrized dispersal model and use
-dispersal simulations to investigate connectivity patterns for African wild dogs
-under two extreme scenarios; one assuming maximum flooding of the Okavango delta
-and one assuming minimum flooding of the Okavango delta. Given that water poses
-a barrier to dispersing individuals, we anticipate that dispersal prospects and
-connectivity during maximum flood are low. During minimum flood, in contrast, we
-expect to reveal the presence of several dispersal corridors that may be used to
-freely and quickly move between remaining habitat patches.
+Talk about Zeller und Osipova
+
+In a previous study, \citep{Hofmann.2022} have investigated landscape
+connectivity in the same ecosystem, yet assuming a static landscape.
+
+Here, we capitalize on a previously parameterized dispersal model and use
+dispersal simulations to investigate dispersal success and connectivity patterns
+for African wild dogs under two extreme scenarios; one assuming maximum flooding
+of the Okavango delta and one assuming minimum flooding of the Okavango delta.
+Given that water poses a barrier to dispersing individuals, we anticipated that
+dispersal prospects and connectivity during maximum flood are low. During
+minimum flood, in contrast, we expect to reveal the presence of several
+dispersal corridors that may be used to freely and quickly move between
+remaining habitat patches.
 
@@
 \section{Results}
+\subsection{General}
+See \Cref{Emigration}
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Emigration.png}
+  \caption{Caption}
+  \label{Emigration}
+  \end{center}
+\end{figure}
+
 \subsection{Heatmaps}
@@
 
+Although dispersal is generally believed to be beneficial for the long term viability of species, it can also lead to the net-loss of individuals in areas where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.
+
 We studied a population of African wild dogs that resides in a natural
@@
 lead to a net-loss of individuals \citep{Leigh.2012}.
+
+Dispersal is an important, if not the most important, driver of landscape
+connectivity and therefore of major interest to conservation authorities. It has
+also been demonstrated that dispersal may lead to population declines in areas
+where anthropogenic mortality is high and dispersal prospects low
+\citep{Leigh.2012}.
+
+Efforts aiming at restoring or improving landscape connectivity should consider
+and take into account changes in landscape characteristics due to climate
+change.
 
@@
 
+Climate change will not only change species distribution but also community
+composition \citep{Thuiller.2006}.
+
 To address such differences, researchers could model
@@
 (cite papers that fieberg sent)
+
+For the discussion: Through changing environmental conditions, climate change
+will also impact the distribution of various species with far-reaching
+consequences on inter- and intra-specific competition (Abrahms). Due to a lack
+of data, we omitted any inclusion of social factors in our dispersal model,
+albeit they are known to be important determinants of dispersal (Armansin).
+Studying how climate change influences species distribution and their
+interactions, and ultimately how this alters dispersal, will be challenging but
+necessary to more realistically simulate dispersal.
 
@@
 
+Although the OD is arguably the main driver of seasonal change across the
+studied ecosystem, there are several other factors that undergoe seasonal
+change, including vegetation and the abundance and distribution of prey or
+predators.
+
 The OD is an important driver of species distribution and it has been found that
@@
 competition, particularly between AWDs and lions.
+
+Although validating the predictions resulting from our dispersal model is
+challenging, the simulations enable us to generate several interesting
+hypotheses...
 

### 2022-12-22 Chapter_10/02_R-Scripts/00_Functions.R

@@ -874,7 +874,7 @@ splitPoly <- function(x) {
 #' @param x SpatVect Polygon to be split
 #' @param max_area numeric area above which polygon should be split
 #' @return SpatVect Split polygon
-polySplitUntil <- function(x, max_area = NULL) {
+splitPolyUntil <- function(x, max_area = NULL) {
   x$Area <- expanse(x, unit = "km")
   while (any(x$Area > max_area)) {
 
@@ -883,7 +883,7 @@ polySplitUntil <- function(x, max_area = NULL) {
     split <- list()
     for (i in 1:length(x)) {
       if (i %in% index) {
-        split[[i]] <- polySplit(x[i, ])
+        split[[i]] <- splitPoly(x[i, ])
       } else {
         split[[i]] <- x[i, ]
       }
@@ -940,7 +940,7 @@ simHR <- function(
 
   # Split polygons that are too big and remove polygons that are too small
   if (crs(pols) == "") { crs(pols) <- crs(r) }
-  pols      <- polySplitUntil(pols, max_area = max_area)
+  pols      <- splitPolyUntil(pols, max_area = max_area)
   pols      <- pols[pols$Area >= min_area, ]
   pols      <- disagg(pols)
   pols$ID   <- sample(1:length(pols))
@@ -970,4 +970,210 @@ metricsHR <- function(x) {
       , Number   = length(area)
     )
   return(metrics)
+}
+
+################################################################################
+#### Dispersal Simulation Function
+################################################################################
+#' Simulate Dispersal
+#'
+#' Function to simulate dispersal using a step selection function
+#' @export
+#' @param source matrix of start coordinates
+#' @param covars spatial covariates prepared using the custom function
+#' @param model issf model prepared using the custom function
+#' @param sl_dist Step Length Distribution
+#' @param sl_max What is the largest possible step?
+#' @param date Start date
+#' @param n_steps Number of steps simulated
+#' @param n_rsteps Number of random steps proposed
+#' @param scaling Dataframe to scale extracted covariates
+#' @param stop Should the simulation stop at a boundary?
+#' @param x SpatRaster of the home ranges
+#' @return tibble containing the metrics
+disperse <- function(
+    source              = NULL    # Start Coordinates
+  , covars              = NULL    # Spatial Covariates, prepared with our funct.
[…]

### 2022-12-22 Chapter_10/02_R-Scripts/01_Analysis/04_Simulation.R

@@ -1,439 +1,0 @@
-################################################################################
-#### Simulation
-################################################################################
-# Description: Simulation of dispersal
-
-# Clear R's brain
-rm(list = ls())
-
-# Change the working directory
-setwd("/home/david/ownCloud/University/15. PhD/Chapter_10")
-
-# Load required packages
-library(raster)         # To handle spatial data
-library(terra)          # To handle spatial data
-library(velox)          # For quick extraction
-library(tidyverse)      # To wrangle data
-library(lubridate)      # To handle dates
-library(glmmTMB)        # To handle the movement model
-library(Rcpp)           # To load C++ functions
-library(pbmcapply)      # For multicore use with progress bar
-
-# Load custom functions
-source("02_R-Scripts/00_Functions.R")
-sourceCpp("02_R-Scripts/00_Functions.cpp")
-
-################################################################################
-#### Load Required Data
-################################################################################
-# Load the calibrated movement model
-mod <- read_rds("03_Data/02_CleanData/MovementModel.rds")
-
-# Load the scaling parameters
-scaling <- read_rds("03_Data/02_CleanData/Scaling.rds")
-
-# Load the fitted gamma distribution (for sampling step lengths)
-sl_dist <- read_rds("03_Data/02_CleanData/GammaDistribution.rds")
-
-# Load the required spatial layers layers
-layers <- rast("03_Data/02_CleanData/SpatialLayers.tif")
-
-# Simulate hrs
-
-
-
-hrs <- simHR(n = 500, m = , r = layers, buff = , min_area = 50, max_area = 1800)
-
-# Extend the covariates artificially (fill added buffer with random values)
-cat("Extending covariate layers...\n")
-ext <- ext(layers) * 1.2
-cov <- extendRaster(layers, ext)
-
-# Apply our custom functions to make everything ready for the simulation
-mod  <- prepareModel(mod)
-cov  <- prepareCovars(layers)
-
-################################################################################
-#### Simulation Function
-################################################################################
-# Function to simulate dispersal based on a step selection model that was fitted
-# in the glmmTMB framework
-disperse <- function(
-    source              = NULL    # Start Coordinates
[…]

### 2022-12-22 Chapter_4/02_R-Scripts/00_Functions.R

@@ -320,20 +320,19 @@ extract_covariates_along_interpolated <- function(x, covariates, by = 0.1){
 ################################################################################
 # Function that extends a covariate layer and fills the added border with values
 # sampled from the layer
-extendRaster <- function(x, y){
-  extended <- lapply(1:nlayers(x), function(z) {
-    layer <- x[[z]]
-    na_mask <- is.na(layer)
-    vals <- values(layer)
-    vals <- na.omit(vals)
-    r <- extend(layer, y)
-    na_mask <- extend(na_mask, y, value = 0)
-    indices <- which(is.na(values(r)))
-    values(r)[indices] <- vals[runif(length(indices), min = 1, max = length(vals))]
-    r <- mask(r, na_mask, maskvalue = 1, updatevalue = NA)
-    return(r)
-  })
-  return(stack(extended))
+extendRaster <- function(x, y) {
+  x_mask  <- is.na(x)
+  x_mask  <- extend(x_mask, y, fill = F)
+  x_large <- extend(x, y, fill = NA)
+  for (i in 1:nlyr(x_large)) {
+    indices <- which(is.na(x_large[[i]][]))
+    n       <- length(indices)
+    values  <- na.omit(x[[i]][])
+    x_large[[i]][indices] <- sample(values, size = n, replace = T)
+    x_large[[i]] <- mask(x_large[[i]], x_mask[[i]], maskvalues = 1, updatevalue = NA)
+  }
+  names(x_large) <- names(x)
+  return(x_large)
 }
 
 ################################################################################

### 2023-01-04 Chapter_2/04_Manuscript/Manuscript.tex
Various updates

@@
 \usepackage[nomarkers, nolists]{endfloat}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{scalerel}
+\usepackage{tikz}
+\usetikzlibrary{svg.path}
 
@@
 %------------------------------------------------------------------------------
-\title{Dispersal Through Seasonal Landscapes}
-
-% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
-% Connectivity}
+\title{Flooding of the Okavango Delta influences Connectivity for Dispersing
+African Wild Dogs}
 
@@
 \author{
-  David D. Hofmann\textsuperscript{1,\S} \and
+  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
+  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
   John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \and
-  Gabriele Cozzi\textsuperscript{1,2} \and
-  Dominik M. Behr\textsuperscript{1,2}
+  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
+  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
 }
@@
 
-\textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
-Assess Landscape Connectivity
+\textbf{Running Title:} None
 
@@
 
-\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
-Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
-pictus
+\textbf{Keywords:} Seasonal Flooding of the Okavango Delta and its Consequences
+for African Wild Dog Connectivity
 
@@
 \begin{abstract}
-This is the abstract.
+Many ecosystems experience substantial changes in environmental conditions due
+to seasonality. While such seasonal changes may drastically alter connectivity
+for endangered species, most studies represent the environment by a static set
+of spatial layers, thus ignoring seasonal variation. Here, we address this
+shortcoming and employ individual-based simulations to investigate how seasonal
+flooding of the Okavango delta influences connectivity for dispersing African
+wild dogs (\textit{Lycaon pictus}). Our results show that the Okavango delta
+poses a substantial dispersal barrier when the flood is at maximum extent, yet
+that viable dispersal corridors exist when the flood is at a minimum level.
+Despite a better understanding of the conservation needs for African wild dogs,
+our study also provides evidence that incorporating seasonality in studies of
+connectivity is imperative to more accurately predict the dispersal ability of
+endangered species.
 \end{abstract}
@@
 % Create linenumbers
-\linenumbers
+% \linenumbers
 
 \section{Introduction}
+Hello, I'm an introduction.
 
-% Taken from Zeller.2020
-Connectivity for wildlife is typically modeled in static environments and
-estimated for a single snapshot in time (e.g., [6,7]). However, wildlife
-responses to landscapes are inherently dynamic, as are the landscapes themselves
-(e.g., [8])
-
-Modeling connectivity in a static environment may mask the importance of connectivity for
-maintaining wildlife population dynamics and may result in identifying corridors that are only
-periodically functional [15,16,18]
+However, \cite{Hofmann.2021} and \cite{Hofmann.2022} used a single floodmap...
 
 \section{Methods}
+
+\subsection{Study Area}
+As depicted in \Cref{StudyArea}, the study area encomassed ...
+
+\subsection{Spatial Habitat Layers}
+We attempted to represent each spatial covariate at the highest temporal
+resolution available, which resulted in various degrees of temporal accuracy. In
+order to generate spatial layers that depict the flood extent at different
+points in time, we ... \Cref{FloodExtent}.
+
+\subsection{Dispersal Simulation}
+We used a previously parametrized movement model to simulate dispersal of
+African wild dogs. To simulate dispersing wild dogs, we employed the simulation
+framework presented in \cite{Hofmann.2022}. In this framework, a movement model
+that renders habitat and movement preferences of dispersing individuals is
+parametrized using step-selection functions. Once parametrized, the model can be
+employed to simulate virtual dispersers moving across the landscape.
+
+We released virtual dispersers at random locations within the four distinct
+source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
+individuals moving across the landscape at minimum flood and another xx
+individuals at maximum flood. For comparison, we also ran the simulations for a
+medium flood extent, yet the results for this will be presented in the appendix.
+
+\subsection{Connectivity}
+To gain insights into landscape connectivity from the simulated dispersal
+events, we prepared three complementary connectivity maps (see again
+\citep{Hofmann.2022}); a heatmap, indicating the absolute frequency at which
+each pixel in the study area was traversed by a virtual disperser, a betweenness
+map, highlighting pixels of particular importance to facilitate movement from
+one area to another, and a map of interpatch connectivity, depicting the
+presence and intensity of functional links between source areas. To compute the
+heatmap, we counted how many times each pixel in the study area was traversed by
+a virtual disperser. To calculate betweenness, we overlayed the study area with
+a regular grid and used the simulated trajectories to determine movements from
+one grid-cell to another. Specifically, we counted how often transitions from
+one grid-cell to any other occured. The centerpoint of each grid-cell then
+served as network node, and the number of transitions from one cell to another
+as weights between the respective nodes. Based on the so generated network, we
+then computed betweenness scores for all network nodes (i.e. for all grid-cells
+across the study area). Lastly, to compute interpatch connectivity, we
+determined the frequency at which virtual dispersers originating from one source
+area reached any of the other source areas. We also calculated the average
+dispersal duration needed until individuals reached the respective area.
 
@@
 \section{Data Availability}
-GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication. Access to all R-scripts for our analyses is provided
-through Github.
+GPS movement data of dispersing wild dogs is available on dryad
+\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
+proposed approach using simulated data are provided through Github
+(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
+codes required to reproduce the African wild dog case study will be made
+available through an online repository at the time of publication.
 
@@
 Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
-to Johannes Signer for assisting with the simulation algorithm. This study was
-funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
-Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung
-Temperatio, Wilderness Wildlife Trust Foundation, Forschungkredit der
-Universität Zürich, and a Swiss National Science Foundation Grant
-(31003A\_182286) to A. Ozgul.
+for sharing her data of three dispersing wild dogs. Furthermore, we would like
+to thank Johannes Signer for assisting with the simulation algorithm. This study
+was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
+Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
+Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
+Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
+Grant (31003A\_182286) to A. Ozgul.
 

### 2023-01-04 Chapter_8/04_Manuscript/Manuscript.tex
Various updates

@@
 Talk about habitat fragmentation and destruction
-
-Osipova
+Say that wild dogs are keystone predators and an umberella species for
+conservation
 
@@
 \citep{Paniw.2021}.
+
+Earlier studies demonstrating the importance of rendering seasonal change
+include \cite{Chetkiewicz.2009}, who use resource selection functions to demonstrate
+that connectivity differs for Ursus actor and Puma concolor and \cite{Mui.2017},
+who showed that connectivity for turtles changes depending on the season, and
+\cite{Kaszta.2021}, who show monthly differences in elephant's habitat selection
+using path-selection functions.
 

### 2023-01-05 Chapter_8/04_Manuscript/Manuscript.tex
various updates

@@
 dispersal durations between distinct habitat patches. Notably, dispersal into
-the center of the Okavango delta is reduced by 80\% during maximum flooding. At
-minimum flooding, conversely, the delta reveals vital dispersal corridors and
-increases chances of successful dispersal into neighboring habitat patches by xx
-percent. Concomitantly, dispersal durations to move between patches is reduced,
-on average by xx days,at low flood levels.
+the central habitats of the Okavango delta is reduced by 80\% during maximum
+flooding. At minimum flooding, conversely, the delta reveals vital dispersal
+corridors and increases chances of successful dispersal into neighboring habitat
+patches by xx percent. Furthermore, we find that dispersers get pushed into
+different zones of potential human-wildlife conflict depending on the flood
+extent.
 
-Climate change is expected to critically impact the flood-pulsing cycles of the
-Okavango delta and may therefore substantially alter landscape connectivity in
-this area. Our analysis suggests that anticipating such alterations in
-environmental conditions is key when implementing conservation strategies.
-Besides a better understanding of the conservation needs for the African wild
-dog, our study also presents a first step towards incorporating environmental
-changes due to seasonality or climate change in dispersal and connectivity
-analyses and demonstrates the usefulness of individual-based dispersal
-simulations as a pertinent conservation tool to study impacts of environmental
-change on dispersal success.
+Climate change is expected to critically impact water-cycles across the globe,
+including the flood-pulsing dynamics of the Okavango delta. Our results suggest
+that anticipating and taking into account such changes in environmental
+conditions is key when developing and implementing conservation strategies,
+especially when the aim is to improve connectivity. Besides this, our analysis
+also provides important insights towards a better understanding of the
+conservation needs for the African wild dog and demonstrates the usefulness of
+individual-based dispersal simulations as a pertinent conservation tool to study
+impacts of environmental change on dispersal success.
 
@@
 distribution of wild living animal populations \citep{Thomas.2004,
-Thuiller.2006}. One potential life-history pathway through which species may
-mediate the effects of environmental change is dispersal \citep{Anderson.2012}.
-By bringing individuals away from their natal location \citep{Clobert.2012},
-dispersal equips species with a means to adapt to climate change by tracking
-favorable habitat conditions \citep{Raia.2012} and to shift into a different
-part of their fundamental niches \citep{Kokko.2022}. Dispersal furthermore
-provides resilience against changing conditions by ensuring genetic diversity
-\citep{Frankham.2002, Leigh.2012, Baguette.2013}, facilitating the colonization
-empty habitats \citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and
-promoting the reinforcement of weakened and small subpopulations
-\citep{Brown.1977}. As such, it is little surpsing that improving connectivity
-is viewed as the number one strategy to combat the adverse effects of climate
-change \citep{Heller.2009}.
+Thuiller.2006}. One life-history pathway through which species potentially
+mediate the effects of environmental change is dispersal \citep{Anderson.2012},
+i.e. the movement of individuals away from their natal location to the site of
+first reproduction \citep{Clobert.2012}. The ability to disperse enables species
+to adapt to climate change by tracking favorable habitat conditions
+\citep{Raia.2012} or by shifting into a different region of their fundamental
+niches \citep{Kokko.2022}. Dispersal furthermore provides resilience against
+changing conditions by safeguarding genetic diversity \citep{Frankham.2002,
+Leigh.2012, Baguette.2013}, facilitating the colonization empty habitats
+\citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and promoting the
+reinforcement of weakened and small subpopulations \citep{Brown.1977}.
 
@@
 change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a function of
-climatic and environmental conditions (e.g. \citep{Behr.2020}). The link between
-dispersal and climaticor environmental can either be indirect, for example if
-the propensity of individuals to disperse depends on environmental conditions,
-or direct, when the biophysical environment through which dispersers move
-affects dispersal prospects \citep{Travis.2013}. As the latter case indicates,
-dispersal is also inextricably linked to the concept landscape connectivity
+climatic and environmental conditions (e.g. \citep{Elliot.2014, Behr.2020}). The
+link between dispersal and the environment can either be indirect, for example
+if the propensity of individuals to disperse depends on environmental
+conditions, or direct, when the biophysical environment through which dispersers
+move affects dispersal prospects \citep{Travis.2013}. The latter case highlights
+that dispersal is also inextricably linked to the concept landscape connectivity
 \citep{Baguette.2013}, which is understood as the degree by which the landscape
-facilitates or impedes movements \citep{Taylor.1993}. Only if a sufficient
-degree of landscape connectivity is maintained, successful dispersal of
-individuals can be guaranteed \citep{Fahrig.2003}. Unfortunately, our
+facilitates or impedes movements \citep{Taylor.1993}. A sufficient degree of
+landscape connectivity is therefore a critical prerequisited of successful
+dispersal \citep{Fahrig.2003}. Unfortunately, the continued degradation and
+destruction habitats worldwide continues to imperil the dispersal ability of
+many species and increases the prospects of inbreeding, genetic risk and of
+local extirpation due to stochastic events \citep{Melbourne.2008, Sawyer.2011}.
+Combined, these factors limit species ability to adapt to changing environmental
+conditions \citep{Fahrig.2003}. Conservation strategies that aim to facilitate
+dispersal by improving landscape connectivity are thus often viewed as pinacle
+of conservation strategies to combat the adverse effects of climate change
+\citep{Heller.2009}.
+
+In recent years, there has been a growing body of research that uses animal
+relocation data to identify movement corridors and assess connectivity at large
+scales \citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014,
+Benz.2016, Osipova.2019}. Identification of potential corridors relies on the
+estimation of permeability (respectively resistance) surfaces, which return the
+ease or willingness at which the focal species traverses a specific landscape
+\citep{Sawyer.2011}. Such surfaces are typically created on the basis of habitat
+preferences, which can be estimated using a suite of selection functions
+\citep{Boyce.2002, Fortin.2005, Cushman.2010}. In selection functions, habitat
+preferences are inferred by comparing spatial covariates (e.g. environmental and
+anthropogenic) at locations visited by the animal to the same spatial covariates
+at randomly selected locations in the animal's availability domain
+\citep{Zeller.2012, Thurfjell.2014}. Importantly, selection functions rely on
+adequate landscape and GPS relocation data that are representative of the
+process being studied \citep{Diniz.2020}. More specifically, GPS relocation data
+collected on dispersing individuals has been shown to outperform GPS relocation
+data collected on resident individuals when aiming at detecting large scale
+dispersal corridors \citep{Elliot.2014, Diniz.2020}. However, few connectivity
+studies have partitioned their GPS relocation data according to behavioral modes
+(e.g. resident vs. dispersing) of the studied species \citep{Wilson.2012,
+Vasudev.2015}. As such, most permeability surfaces upon which dispersal
+corridors are identified are created using GPS relocation data that were
+collected on resident individuals. This introduces severe biases and
+substantially reduces the power to reveal meaningful movement corridors, for
+dispersing individuals have different needs and drives compared to resident
+individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}. Consequently, these
+biases have limited our ability to assess the adequacy of the boundaries of
+protected areas in securing connectivity for their protected species.
+
+To study connectivity, various modeling techniques have emerged
+\cite{Etherington.2016, Diniz.2019}. Initially, the techniques were limited to
+studying strucutral aspects of connectivity by focusing on the composition and
+configuration of habitat patches, while ignoring species' responses to the
+landscape matrix \citep{Tischendorf.2000, Doerr.2011}. With the increasing
+availability of telemetry data and methods to study habitat and movement
+preferences \citep{Boyce.2002, Fortin.2009, Cushmann.2010, Avgar.2016}, however,
+the focus has shifted towards more functional aspects of connectivity
+\citep{Tischendorf.2000, Doerr.2011}. Currently, the most prominent
+\textit{functional} connectivity models are least-cost path analysis (LCPA,
+\citealp{Adriaensen.2003}) and circuit theory (CT, \citealp{McRae.2008}), two
+graph-based methods that estimate conductance of the landscape by means of a
+resistance (or inversely permeability) surface \citep{Zeller.2012}. Because both
+approaches make assumptions that are rarely met by dispersing individuals,
+individual-based movement models (IBMMs) have also gained some momentum recently
+\citep{Kanagaraj.2013, Allen.2016, Hauenstein.2019, Diniz.2019, Zeller.2020,
+Unnithan.2021, Unnithan.2022, Hofmann.2022}. In these models, movement of
+individuals is explicitly simulated and connectivity metrics are derived based
+on the simulated dispersal paths. IBMMs provide great modeling flexibility and
+are thus considered powerful tools for examining connectivity under different
+landscape configurations \citep{Littlefield.2019}.
+
+Irrespective of the chosen method, connectivity is often modeled as a snapshot
+in time that fails to render dynamic environmental change. However, some recent
+studies have shown that connectivity may differ substantially depending on
+season ... some examples.
+
+ Unfortunately, our
 understanding how changing environmental conditions influence connectivity and
@@
 assess how changing environmental conditions affect connectivity...
-
-The continued degradation and destruction habitats across the globe impedes
-dispersal, promotes inbreeding and genetic drift and increases the
-likelihood of local extirpation due to stochastic events \citep{Sawyer.2011}.
-Combined, these factors limit species ability to adapt to changing environmental
-conditions \citep{Fahrig.2003}.
 
@@
 changes in connectivity that will arise under amplified climatic conditions due
-to climate change.
-While a better understanding of the ODs flooding regime under global change is of major interest, reliable predictions remain challenging due to the complex interplay between climate,
-anthropogenic water use, animal behavior, and topography of the region.
+to climate change. While a better understanding of the ODs flooding regime under
+global change is of major interest, reliable predictions remain challenging due
+to the complex interplay between climate, anthropogenic water use, animal
+behavior, and topography of the region.
 
@@
 necessity using resistance-based approaches \citep{Hofmann.2023,
-Unnithan.2022a}. IBMMs allow much greater flexibility and are thus considered
-powerful tools for examining connectivity under different landscape
-configurations \citep{Littlefield.2019}.
-
+Unnithan.2022a}.
 Botswana is among the countries most heavily influenced by climate change .
@@
 
-Talk about connectivity models
-Talk about functional connectivity
-Talk about habitat fragmentation and destruction
 Say that wild dogs are keystone predators and an umberella species for
@@
 The Okavango delta in Southern Africa poses a unique opportunity to study the
-impacts of environmental change on species dispersal ability and connectivity in a
-large scale natural experiment setup. According to projections by the IPCC,
+impacts of environmental change on species dispersal ability and connectivity in
+a large scale natural experiment setup. According to projections by the IPCC,
 Botswana and its surroundings are among the most vulnerable to climate change.
@@
 of the 21st century, For example, \cite{Engelbrecht.2015} predicts an increase
-between 4 and 6\degree C for regions in southern Africa. One species that
-heavily relies on dispersal for long-term persistence and that inhabits such an
-arid ecosystem in Southern Africa is the African wild dog (\textit{Lycaon
-pictus}). The African wild dog is a lightly built canid lives in cohesive packs
-comprising up to 20 individuals. Wild dogs are social breeders and the majority
-of reproduction is monopolized by the dominant alpha couple who are supported by
-subordinates in taking care of their pups. The species exhibits strong
-inbreeding avoidance, hence juvenile individuals born into the pack disperse in
-same-sex sibling coalitions after reaching sexual maturity \citep{McNutt.1996,
-Behr.2020}. Notably, the African wild dog is the sole extant representative of
-its genus and thus considered as high-priority species for conservation efforts
-\citep{Leigh.2012}. However, fragmentation and destruction of their historic
-habitat may have decreased dispersal success for this species and thus
-jeopardizes connectivity \citep{Leigh.2012}. Since the impacts of climate change
-are not uniformly distributed across the planet, but particularly pronounced at
-high latitudes and arid ecosystems, the species may be seen as most vulnerable
-to such ecosystem shifts.
+between 4 and 6\degree C for regions in southern Africa. One of the keyspone
+predators and an umberella species for conservation efforts in this ecosystem is
+the African wild dog (\textit{Lycaon pictus}). This species heavily relies on
+dispersal for long-term persistence. The African wild dog is a lightly built
+canid lives in cohesive packs comprising up to 20 individuals. Wild dogs are
+social breeders and the majority of reproduction is monopolized by the dominant
+alpha couple who are supported by subordinates in taking care of their pups. The
+species exhibits strong inbreeding avoidance, hence juvenile individuals born
+into the pack disperse in same-sex sibling coalitions after reaching sexual
+maturity \citep{McNutt.1996, Behr.2020}. Notably, the African wild dog is the
+sole extant representative of its genus and thus considered as high-priority
+species for conservation efforts \citep{Leigh.2012}. However, fragmentation and
+destruction of their historic habitat may have decreased dispersal success for
+this species and thus jeopardizes connectivity \citep{Leigh.2012}. Since the
+impacts of climate change are not uniformly distributed across the planet, but
+particularly pronounced at high latitudes and arid ecosystems, the species may
+be seen as most vulnerable to such ecosystem shifts.
 
@@
 
-In a previous study, \citep{Hofmann.2022} have investigated landscape
-connectivity in the same ecosystem, yet assuming a static landscape.
-
-Here, we capitalize on a previously parameterized dispersal model and use
-dispersal simulations to investigate dispersal success and connectivity patterns
-for African wild dogs under two extreme scenarios; one assuming maximum flooding
-of the Okavango delta and one assuming minimum flooding of the Okavango delta.
-Given that water poses a barrier to dispersing individuals, we anticipated that
-dispersal prospects and connectivity during maximum flood are low. During
-minimum flood, in contrast, we expect to reveal the presence of several
-dispersal corridors that may be used to freely and quickly move between
-remaining habitat patches.
+Here, we capitalize on the previously parameterized dispersal model by
+\cite{Hofmann.2022} and use dispersal simulations to investigate dispersal
+success and connectivity patterns for African wild dogs under two extreme
+scenarios; one assuming maximum flooding of the Okavango delta and one assuming
+minimum flooding of the Okavango delta. Given that water poses a barrier to
+dispersing individuals, we anticipated that dispersal prospects and connectivity
+during maximum flood are low. During minimum flood, in contrast, we expect to
+reveal the presence of several dispersal corridors that may be used to freely
+and quickly move between remaining habitat patches.
 
 \section{Materials and Methods}
-We conducted all analyses using the programming language \textsf{R}. Any spatial
-data manipulation was completed using the \textsf{terra} and \textsf{spatstat}
-packages. Several helper functions for the simulation algorithm were written in
-\textsf{C++} and imported to R using the \textsf{Rcpp} package. Figures were
-generated using \textsf{ggplot2}, \textsf{igraph}, and \textsf{ggnetwork}. All
-R-scripts used to conduct our analyses are provided through an online
-repository.
+We conducted all analyses using the programming language \textsf{R}
+\citep{RCoreTeam.2022}. Any spatial data manipulation was completed using the
+\textsf{terra} \citep{Hijmans.2022} and \textsf{spatstat} packages
+\citep{Baddeley.2015}. Several helper functions for the simulation algorithm
+were written in \textsf{C++} and imported to R using the \textsf{Rcpp} package
+\citep{Eddelbuettel.2011}. Figures were generated using \textsf{ggplot2}
+\citep{Wickham.2016}, \textsf{igraph} \citep{Csardi.2006}, and
+\textsf{ggnetwork} \citep{Briatte.2021}. All R-scripts used to conduct our
+analyses are provided through an online repository.
 
@@
 
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Distance.png}
+  \caption{Caption}
+  \label{Distance}
+  \end{center}
+\end{figure}
+
 % \input{IPC_Frequency.tex}
@@
 
-Although dispersal is generally believed to be beneficial for the long term viability of species, it can also lead to the net-loss of individuals in areas where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.
+Although dispersal is generally believed to be beneficial for the long term
+viability of species, it can also lead to the net-loss of individuals in areas
+where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.
 

### 2023-01-08 Chapter_8/04_Manuscript/Manuscript.tex
upd

@@
 \section{Introduction}
+\subsection{Climate Change and Dispersal}
 Climate change is expected to profoundly impact ecosystems worldwide with
@@
 distribution of wild living animal populations \citep{Thomas.2004,
-Thuiller.2006}. One life-history pathway through which species potentially
-mediate the effects of environmental change is dispersal \citep{Anderson.2012},
-i.e. the movement of individuals away from their natal location to the site of
-first reproduction \citep{Clobert.2012}. The ability to disperse enables species
-to adapt to climate change by tracking favorable habitat conditions
-\citep{Raia.2012} or by shifting into a different region of their fundamental
-niches \citep{Kokko.2022}. Dispersal furthermore provides resilience against
-changing conditions by safeguarding genetic diversity \citep{Frankham.2002,
-Leigh.2012, Baguette.2013}, facilitating the colonization empty habitats
-\citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and promoting the
-reinforcement of weakened and small subpopulations \citep{Brown.1977}.
+Thuiller.2006}. One possible life-history pathway through which species may
+mediate the negative consequences of environmental change is dispersal
+\citep{Anderson.2012}, the movement of individuals away from their natal
+location to the site of first reproduction \citep{Clobert.2012}. Through
+dispersal, species can adapt to climate change by tracking favorable habitat
+conditions \citep{Raia.2012} and by shifting into a different region of their
+fundamental niche \citep{Kokko.2006}. Dispersal furthermore provides resilience
+against changing environmental conditions by safeguarding genetic diversity
+\citep{Frankham.2002, Leigh.2012, Baguette.2013}, facilitating the colonization
+empty habitats \citep{Gustafson.1996, Hanski.1999a, MacArthur.2001}, and
+promoting the reinforcement of weakened and small subpopulations
+\citep{Brown.1977}. Through these processes, dispersal crucially contributes to
+long term population viability \citep{Kokko.2006, Fahrig.2003}.
 
-While dispersal offers a means to offset the demographic consequences of climate
-change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a function of
-climatic and environmental conditions (e.g. \citep{Elliot.2014, Behr.2020}). The
-link between dispersal and the environment can either be indirect, for example
-if the propensity of individuals to disperse depends on environmental
-conditions, or direct, when the biophysical environment through which dispersers
-move affects dispersal prospects \citep{Travis.2013}. The latter case highlights
-that dispersal is also inextricably linked to the concept landscape connectivity
-\citep{Baguette.2013}, which is understood as the degree by which the landscape
-facilitates or impedes movements \citep{Taylor.1993}. A sufficient degree of
-landscape connectivity is therefore a critical prerequisited of successful
-dispersal \citep{Fahrig.2003}. Unfortunately, the continued degradation and
+\subsection{Connectivity}
+While dispersal offers a means to offset the negative demographic consequences
+of climate change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a
+function of climatic and environmental conditions (e.g. \citep{Elliot.2014,
+Behr.2020}). The link between dispersal and the environment can either be
+indirect, for example if the propensity of individuals to disperse depends on
+environmental conditions, or direct, when the biophysical environment through
+which dispersers move affects dispersal prospects \citep{Travis.2013}. The
+latter case highlights that dispersal is also inextricably linked to the concept
+landscape connectivity \citep{Baguette.2013}, which is understood as the degree
+to which the landscape facilitates or impedes movements \citep{Taylor.1993}. A
+sufficient degree of landscape connectivity is therefore a critical prerequisite
+for successful dispersal \citep{Fahrig.2003}. Conservation strategies that aim
+to facilitate dispersal by improving landscape connectivity are thus often
+viewed as pinacle of conservation strategies to combat the adverse effects of
+climate change \citep{Heller.2009}. Despite this, the continued degradation and
 destruction habitats worldwide continues to imperil the dispersal ability of
-many species and increases the prospects of inbreeding, genetic risk and of
-local extirpation due to stochastic events \citep{Melbourne.2008, Sawyer.2011}.
-Combined, these factors limit species ability to adapt to changing environmental
-conditions \citep{Fahrig.2003}. Conservation strategies that aim to facilitate
-dispersal by improving landscape connectivity are thus often viewed as pinacle
-of conservation strategies to combat the adverse effects of climate change
-\citep{Heller.2009}.
+many species and increases the prospects of inbreeding, genetic drift and local
+extirpation due to stochastic events \citep{Melbourne.2008, Sawyer.2011}
+alltogether limiting species' ability to adapt to changing environmental
+conditions \citep{Fahrig.2003}. Succssful conservation or restoration of areas
+that are of particular importance to dispersing animals thus demands a thorough
+understanding of the factors that promote or limit dispersal
+\citep{Baguette.2013}.
 
-In recent years, there has been a growing body of research that uses animal
-relocation data to identify movement corridors and assess connectivity at large
-scales \citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014,
-Benz.2016, Osipova.2019}. Identification of potential corridors relies on the
-estimation of permeability (respectively resistance) surfaces, which return the
-ease or willingness at which the focal species traverses a specific landscape
-\citep{Sawyer.2011}. Such surfaces are typically created on the basis of habitat
-preferences, which can be estimated using a suite of selection functions
-\citep{Boyce.2002, Fortin.2005, Cushman.2010}. In selection functions, habitat
-preferences are inferred by comparing spatial covariates (e.g. environmental and
-anthropogenic) at locations visited by the animal to the same spatial covariates
-at randomly selected locations in the animal's availability domain
-\citep{Zeller.2012, Thurfjell.2014}. Importantly, selection functions rely on
-adequate landscape and GPS relocation data that are representative of the
-process being studied \citep{Diniz.2020}. More specifically, GPS relocation data
-collected on dispersing individuals has been shown to outperform GPS relocation
-data collected on resident individuals when aiming at detecting large scale
-dispersal corridors \citep{Elliot.2014, Diniz.2020}. However, few connectivity
-studies have partitioned their GPS relocation data according to behavioral modes
-(e.g. resident vs. dispersing) of the studied species \citep{Wilson.2012,
-Vasudev.2015}. As such, most permeability surfaces upon which dispersal
-corridors are identified are created using GPS relocation data that were
-collected on resident individuals. This introduces severe biases and
-substantially reduces the power to reveal meaningful movement corridors, for
-dispersing individuals have different needs and drives compared to resident
-individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}. Consequently, these
-biases have limited our ability to assess the adequacy of the boundaries of
-protected areas in securing connectivity for their protected species.
-
+\subsection{Modeling Connectivity}
 To study connectivity, various modeling techniques have emerged
@@
 landscape matrix \citep{Tischendorf.2000, Doerr.2011}. With the increasing
-availability of telemetry data and methods to study habitat and movement
-preferences \citep{Boyce.2002, Fortin.2009, Cushmann.2010, Avgar.2016}, however,
-the focus has shifted towards more functional aspects of connectivity
-\citep{Tischendorf.2000, Doerr.2011}. Currently, the most prominent
-\textit{functional} connectivity models are least-cost path analysis (LCPA,
-\citealp{Adriaensen.2003}) and circuit theory (CT, \citealp{McRae.2008}), two
-graph-based methods that estimate conductance of the landscape by means of a
-resistance (or inversely permeability) surface \citep{Zeller.2012}. Because both
+availability of telemetry data and methods to study species' habitat and
+movement preferences \citep{Boyce.2002, Fortin.2009, Cushman.2010, Avgar.2016},
+however, the focus has shifted towards a more functional understaning of
+connectivity, which also takes into account how species interact with their
+surroundings \citep{Tischendorf.2000, Doerr.2011}. Currently, the most prominent
+\textit{functional} connectivity models are based on least-cost path analysis
+(LCPA, \citealp{Adriaensen.2003}) and circuit theory (CT, \citealp{McRae.2008}),
+two graph-based methods that estimate conductance of the landscape by means of a
+resistance (or inversely permeability) surface \citep{Zeller.2012}. Since both
 approaches make assumptions that are rarely met by dispersing individuals,
@@
 \citep{Kanagaraj.2013, Allen.2016, Hauenstein.2019, Diniz.2019, Zeller.2020,
-Unnithan.2021, Unnithan.2022, Hofmann.2022}. In these models, movement of
-individuals is explicitly simulated and connectivity metrics are derived based
-on the simulated dispersal paths. IBMMs provide great modeling flexibility and
-are thus considered powerful tools for examining connectivity under different
-landscape configurations \citep{Littlefield.2019}.
+UnnithanKumar.2022, UnnithanKumar.2022a, Hofmann.2022}. In these models,
+dispersal of individuals is explicitly simulated across the landscape of
+interest and connectivity metrics are derived based on the simulated dispersal
+paths \citep{Diniz.2019}. Because movement is simulated spatially and temporally
+explicitly, IBMMs provide great modeling flexibility and are considered powerful
+tools for examining connectivity under different landscape configurations
+\citep{Littlefield.2019, UnnithanKumar.2022a}.
 
-Irrespective of the chosen method, connectivity is often modeled as a snapshot
-in time that fails to render dynamic environmental change. However, some recent
-studies have shown that connectivity may differ substantially depending on
-season ... some examples.
+\subsection{Seasonality or Climate Change}
+While most studies examine connectivity assuming a static landscape, this hardly
+ever captures the true nature of most ecosystems, which typically experience
+dynamic environmental change, either due to seasonality or climatic change.
+Connectivity analyses that omit such dynamism not only fail to capture an
+important aspect of reality, but also prohibit studying the sensitivity of
+connectivity with regards to environmental change, such as the one expected
+under predicted climate change. Recent studies that have incorporated seasonal
+or climatic dynamism in their analysis have revealed vast differences in
+connectivity. For instance, \cite{Mui.2017} developed seasonal resistance maps
+for Blanding's turtle \textit{Emydoidea blandingii} and showed that connectivity
+was substantially lower in late summer compared to spring. Similarly,
+\cite{Osipova.2019} studied connectivity for African elephants
+(\textit{Loxodonta africana}) during wet and dry season, and found that ignoring
+seasonality resulted in an underestimation of connectivity during the wet
+season, but an overestimation during the dry season. For the same species,
+\citep{Kaszta.2021} generated monthly connectivity maps, demonstrating stark
+differences in connectivity across the year. Other studies have used
+species-distribution models in conjuction with climate predictions to highlight
+how connectivity might change in result to species shifting their ranges. For
+example, \cite{Ashrafzadeh.2019} found that cimate change is likely to lead to
+increased habitat fragmentation and therefore reduced connectivity for mountain
+newts in Iran. Similarly, \cite{Luo.2021} predicted significant reductions in
+connectivity for the giant spiny frog in South-East Asia. Climate change may
+thus lead to the disruption of existing dispersal corridors but may also provide
+possibilities to access previously unsuitable dispersal routes. Designing and
+implementing cost-effective conservation strategies that aim to improve
+landscape connectivity, such as habitat restoration or the creation of wildlife
+corridors, requires better understanding of such dynamics.
 
- Unfortunately, our
-understanding how changing environmental conditions influence connectivity and
-dispersal is limited \citep{Kokko.2006, Travis.2013}. Because collecting
-information on dispersing animals is difficult and because it is difficult to
-assess how changing environmental conditions affect connectivity...
-
-\cite{Ashrafzadeh.2019} found that cimate change is likely to substantially
-reduce connectivity for mountain newts in Iran. Similarly, \cite{Luo.2021}
-predicted significant reductions in connectivity for the giant spiny frog in
-South-East Asia.
-
-However, failing to take into account potential changes in climatic and
-environmental conditions imperils our ability to design cost-effective
-conservation strategies.
-
-Here, we used naturally occuring seasonal flood-cycles to approximate the likely
-changes in connectivity that will arise under amplified climatic conditions due
-to climate change. While a better understanding of the ODs flooding regime under
-global change is of major interest, reliable predictions remain challenging due
-to the complex interplay between climate, anthropogenic water use, animal
-behavior, and topography of the region.
-
-Climate change
-could lead to the disruption of existing dispersal corridors and the emergence
-of new dispersal barriers but may also provide possibilities to access
-previously unsuitable dispersal routes. Designing and implementing effective
-conservation strategies that aim to improve landscape connectivity, such as
-habitat restoration or the creation of wildlife corridors, requires better
-understanding of such dynamics.
-
-Albeit the impacts of climate change on connectivity cannot be directly
-estimated, one can study how changing environmental conditions due to
-seasonality influence patterns of dispersal and connectivity.
-
-Habitat degradation and fragmentation lead to isolated subpopulations that have
-a high associated risk of extinction due to stochastic events and generally
-decreased genetic variability \citep{Melbourne.2008}.
-
-More recently, individual-based movement models (IBMMs) have been uprising and
-are recommended for their ability to incorporate several crucial realities of
-animal behavior, such as, for example, autocorrelation, that are by omitted by
-necessity using resistance-based approaches \citep{Hofmann.2023,
-Unnithan.2022a}.
-Botswana is among the countries most heavily influenced by climate change .
+\subsection{What we Did}
+Here, we capitalize on naturally occuring seasonal changes to approximate the
+likely changes in connectivity that will arise under amplified climatic
+conditions due to climate change. Specifically, flood-cycles of the OD. While a
+better understanding of the ODs flooding regime under global change is of major
+interest, reliable predictions remain challenging due to the complex interplay
+between climate, anthropogenic water use, animal behavior, and topography of the
+region.
 
@@
 \citep{Paniw.2021}.
-
-Earlier studies demonstrating the importance of rendering seasonal change
-include \cite{Chetkiewicz.2009}, who use resource selection functions to demonstrate
-that connectivity differs for Ursus actor and Puma concolor and \cite{Mui.2017},
-who showed that connectivity for turtles changes depending on the season, and
-\cite{Kaszta.2021}, who show monthly differences in elephant's habitat selection
-using path-selection functions.
-
-Only if there
 
@@
 impacts of environmental change on species dispersal ability and connectivity in
-a large scale natural experiment setup. According to projections by the IPCC,
-Botswana and its surroundings are among the most vulnerable to climate change.
-While global temperatures are expected to increase by xx degrees until the end
-of the 21st century, For example, \cite{Engelbrecht.2015} predicts an increase
-between 4 and 6\degree C for regions in southern Africa. One of the keyspone
-predators and an umberella species for conservation efforts in this ecosystem is
-the African wild dog (\textit{Lycaon pictus}). This species heavily relies on
-dispersal for long-term persistence. The African wild dog is a lightly built
-canid lives in cohesive packs comprising up to 20 individuals. Wild dogs are
-social breeders and the majority of reproduction is monopolized by the dominant
-alpha couple who are supported by subordinates in taking care of their pups. The
-species exhibits strong inbreeding avoidance, hence juvenile individuals born
-into the pack disperse in same-sex sibling coalitions after reaching sexual
-maturity \citep{McNutt.1996, Behr.2020}. Notably, the African wild dog is the
-sole extant representative of its genus and thus considered as high-priority
-species for conservation efforts \citep{Leigh.2012}. However, fragmentation and
-destruction of their historic habitat may have decreased dispersal success for
-this species and thus jeopardizes connectivity \citep{Leigh.2012}. Since the
-impacts of climate change are not uniformly distributed across the planet, but
-particularly pronounced at high latitudes and arid ecosystems, the species may
-be seen as most vulnerable to such ecosystem shifts.
-
-Talk about Zeller und Osipova
+a large scale natural experiment setup. Talk about flood cycles. According to
+projections by the IPCC, Botswana and its surroundings are among the most
+vulnerable to climate change. While global temperatures are expected to increase
+by xx degrees until the end of the 21st century, For example,
+\cite{Engelbrecht.2015a} predicts an increase between 4 and 6\degree C for
+regions in southern Africa. One of the keyspone predators and an umberella
+species for conservation efforts in this ecosystem is the African wild dog
+(\textit{Lycaon pictus}). This species heavily relies on dispersal for long-term
+persistence. The African wild dog is a lightly built canid lives in cohesive
+packs comprising up to 20 individuals. Wild dogs are social breeders and the
+majority of reproduction is monopolized by the dominant alpha couple who are
+supported by subordinates in taking care of their pups. The species exhibits
+strong inbreeding avoidance, hence juvenile individuals born into the pack
+disperse in same-sex sibling coalitions after reaching sexual maturity
+\citep{McNutt.1996, Behr.2020}. Notably, the African wild dog is the sole extant
+representative of its genus and thus considered as high-priority species for
+conservation efforts \citep{Leigh.2012}. However, fragmentation and destruction
+of their historic habitat may have decreased dispersal success for this species
+and thus jeopardizes connectivity \citep{Leigh.2012}. Since the impacts of
+climate change are not uniformly distributed across the planet, but particularly
+pronounced at high latitudes and arid ecosystems, the species may be seen as
+most vulnerable to such ecosystem shifts.
 
 Here, we capitalize on the previously parameterized dispersal model by
-\cite{Hofmann.2022} and use dispersal simulations to investigate dispersal
+\cite{Hofmann.2023} and use IBMMs to simulate dispersal and assess dispersal
 success and connectivity patterns for African wild dogs under two extreme
@@
 with smallest flood extent and generated a binary image using areas that were
-inundated in at least 50\% of the maps. Similarly, we created an average image
-for high flood using the 50 most flooded maps. The final maps are depicted in
+inundated in at least 50\% of the maps. The resulting final maps are depicted in
 \Cref{FloodExtent}.
@@
 integrated step-selection function (iSSF, \citealp{Avgar.2016}) fitted to GPS
-data of dispersing AWDs \citep{Hofmann.2021}. In step selection functions (SSFs,
+data of dispersing AWDs \citep{Hofmann.2023}. In step selection functions (SSFs,
 \citealp{Fortin.2005}), observed GPS locations are converted into steps (the
@@
 logistic regression framework \citep{Fortin.2005, Thurfjell.2014, Muff.2020,
-Fieberg.2021}. The model presented in \citep{Hofmann.2021} used dispersal data
+Fieberg.2021}. The model presented in \citep{Hofmann.2023} used dispersal data
 of 16 dispersing African wild dogs from a free-ranging wild dog population in
@@
 in the vicinity of the Okavango delta (\Cref{StudyArea}). For areas one to six
-we selected locations at the delta's periphery that remained dry in both
-scenarios. At those locations, we generated circular buffers with a radius of 20
-km. For source area six, we isolated a polygon covering Chief's Island, a
-peninsula located at the OD's center. Besides source areas, we also generated
-``zones of emigration'' that we used as checkpoints to determine if and where
-simulated individuals left the delta's vicinity (\Cref{StudyArea}). We generated
-emigration zones by first overlaying the OD with an elliptic buffer zone that we
-dissected using a set of cutlines that originated from the ODs center and spread
-according to cardinal points (\Cref{StudyArea}).
+we selected locations at the delta's periphery that remained dry during minimum
+and maximum flooding. At those locations, we generated circular buffers with a
+radius of 20 km. For source area six, we isolated a polygon covering Chief's
+Island, a peninsula located at the OD's center. Besides source areas, we also
+generated ``zones of emigration'' that we used as checkpoints to determine if
+and where simulated individuals left the delta's vicinity (\Cref{StudyArea}). We
+generated emigration zones by first overlaying the OD with an elliptic buffer
+zone that we dissected into roughly equally sized pieces in accordance to
+cardinal points (\Cref{StudyArea}).
 
 \subsection{Dispersal Simulation}
-For each source area we simulated 1'000 individuals, once assuming a minimum
-flood, once assuming a maximum flood. This resulted in the simulation of 6'000
-individuals for each environmental scenario, hence 12'000 individuals in total.
-The simulation algorithm was based on the algorithm described in
-\Cref{Hofmann.2022} and works as follows. A random location within the source
-area is chosen as a starting point. Originating from the starting point, a set
-of 25 random steps is generated by sampling step lengths from a gamma
-distribution fitted to observed steps (what is a step?) (shape = , scale = ) and
-turning angles from a uniform distibution ($-\pi, +\pi$). Along each random step
-the underlying spatial covariates are extracted and relevant movement metrics
-are computed (i.e. log(sl), cos(ta), ta). The parametrized dispersal model is
-used to predict the probability of each step for being chosen given the steps
-covariate values. One of the steps is sampled based on assigned probabilities
-and the location of the animal is updated. The procedure is then repated until
-the desired number of steps is realized. Here, we simulated each individual for
-2'000 steps, which correspons to the longest dispersal duration recorded in our
-data. The original model was trained using 4-hourly steps, thus a simulated step
-also resembled the movement conducted within four hours. Trajectories resulting
-from such a simulation can be understood as correlated random walks that take
-into account both habitat and movement preferences of dispersing individuals.
+Originating from each source area we simulated 1'000 individuals dispersing for
+a total of 2000 steps (which equals to a dispersal duration of 400 days). Each
+individual was simulated once assuming a minimum flood, once assuming a maximum
+flood. This resulted in the simulation of a total of 12'000 individuals. The
+simulation algorithm was based on the algorithm described in \Cref{Hofmann.2023}
+and works as follows. A random location within the source area is chosen as a
+starting point. Originating from the starting point, a set of 25 random steps is
+generated by sampling step lengths from a gamma distribution fitted to observed
+steps (what is a step?) (shape = , scale = ) and turning angles from a uniform
+distibution ($-\pi, +\pi$). Along each random step the underlying spatial
+covariates are extracted and relevant movement metrics are computed (i.e.
+log(sl), cos(ta), ta). The parametrized dispersal model is used to predict the
+probability of each step for being chosen given the steps covariate values. One
+of the steps is sampled based on assigned probabilities and the location of the
+animal is updated. The procedure is then repated until the desired number of
+steps is realized. Here, we simulated each individual for 2'000 steps, which
+correspons to the longest dispersal duration recorded in our data. The original
+model was trained using 4-hourly steps, thus a simulated step also resembled the
+movement conducted within four hours. Trajectories resulting from such a
+simulation can be understood as correlated random walks that take into account
+both habitat and movement preferences of dispersing individuals.
 
@@
 
-\subsection{Dispersal Prospects and Connectivity}
-Based on simulated dispersal trajectories in the two scenarios we quantified
-dispersal sucess and connectivity using three complementary connectivity metrics
-as outlined in \cite{Hofmann.2022}. First, we generated heatmaps depicting the
-frequency at which different areas in the landscape were visited by simulated
-dispersers. Such heatmaps serve to detect dispersal hotspots and areas of
-intense use. However, they are less suitable for detecting pinchpoints and
-bottlenecks that are critical in linking distinct patches. Hence, we also
-computed spatially explicit betweenness scores which are useful to highlight
-exactly such pinchpoints \citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}.
-To compute betweenness, we overlayed the study area with a regular grid with 2.5
+\subsection{Dispersal Prospects and Connectivity and Distance To Humans}
+Based on simulated dispersal trajectories we quantified dispersal sucess and
+connectivity using three complementary connectivity metrics as outlined in
+\cite{Hofmann.2023}. First, we generated heatmaps depicting the frequency at
+which different areas in the landscape were visited by simulated dispersers.
+Such heatmaps serve to detect dispersal hotspots and areas of intense use.
+However, they are less suitable for detecting pinchpoints and bottlenecks that
+are critical in linking distinct patches. Hence, we also computed spatially
+explicit betweenness scores which are useful to highlight exactly such
+pinchpoints \citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}. To compute
+betweenness, we overlayed the study area with a regular grid comprising of 2.5
 km x 2.5 km grid cells and determined how often simulated individuals
-transitioned from one grid-cell to another. Note that in case the same
-individual repeatedly realized the same cell-transition (e.g. repeatedly moved
-between A-B-A-B...), we only counted a single transition to avoid emphasis on
-regions where individuals moved in circles. With on the so generated
-transitions, we generated a network using the centers of all grid-cells as nodes
-and cell-transitions as weighted connections between the nodes. Based on this
-network we computed weighted betweenness scores using the R-package
-\textsf{igraph}. As a final connectivity metric and metric of dispersal success,
-we calculated the number of successful dispersal events between the different
-source areas as well as towards the emigration zones. We coin this type of
-connectivity ``inter-patch connectivity'' as it relates to the movement between
-distinct patches. Dispersal between two areas was said to be successful whenever
-a trajectory leaving one area intersected with the target area. To gauge the
-dispersal duration needed to move between patches (be consistent with
-``patches'', ``source-areas'' etc.), we also recorded the minimum number of
-steps that individuals moved before arriving at the respective patch.
+transitioned from one grid-cell to another. To determine such transitions, we
+assessed the sequence of grid-cells thorugh which each individual moved. In case
+the same individual repeatedly realized the same cell-transition (e.g.
+repeatedly moved between A-B-A-B...), we only counted a single transition to
+avoid emphasis on regions where individuals moved in circles. With on the so
+generated transitions, we generated a network using the centers of all
+grid-cells as nodes and cell-transitions as weighted connections between the
+nodes (for details, see \citep{Hofmann.2023}). Based on this network we computed
+weighted betweenness scores. As a final connectivity metric, we quantified
+dispersal success and calculated the number of successful dispersal events
+between the different source areas. We coin this type of connectivity
+``inter-patch connectivity'' as it relates to the movement between distinct
+patches. Dispersal between two areas was said to be successful whenever a
+trajectory leaving one area intersected with the target area. We also tallied
+the number of individuals that left the OD's vicinity and moved through one of
+the emigration zones. To gauge the dispersal duration needed to move between
+patches (be consistent with ``patches'', ``source-areas'' etc.), we recorded the
+minimum number of steps that individuals moved before arriving at the respective
+patch. Finally, we identified zones with a high potential for human wildlife
+conflict. For this, we isolated all simulated locations that lied within a xx km
+buffer to the nearest human-influenced pixel in the study area. Based on the so
+isolated coordinates we generated heatmaps highlighting areas with a high
+density of coordinates within the vicinity of humans.
 
@@
 minimum flood, the area north-west of Maun appears to serve as vital dispersal
-habitat. The same area is entirely avoided during maximum flood. Besides
+habitat, yet the same area is entirely avoided during maximum flood. Besides
 striking differences in connectivity for the close vicinity of the delta, the
@@
 \end{figure}
+
+The density around Maun increases from xx to xx, the density around the pan
+handel from to xx
 

### 2023-01-19 Chapter_8/02_R-Scripts/02_Visualization/Difference.R
Various updates

@@ -1,48 +1,0 @@
-################################################################################
-#### Difference Plots
-################################################################################
-# Description: Put together the individual difference plots
-
-# Clear R's brain
-rm(list = ls())
-
-# Change the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
-setwd(wd)
-
-# Load required packages
-library(tidyverse)      # To wrangle data
-library(ggpubr)         # To arrange multiple plots
-
-# Load individual plots
-p1 <- read_rds("04_Manuscript/99_HeatmapsDifference.rds")
-p2 <- read_rds("04_Manuscript/99_BetweennessDifference.rds")
-p3 <- read_rds("04_Manuscript/99_DistanceDifference.rds")
-
-# We want to get a similar look to the "facets"
-p1 <- p1 + facet_wrap(~ "Heatmap")
-p2 <- p2 + facet_wrap(~ "Betweenness")
-p3 <- p3 + facet_wrap(~ "Distance")
-
-# # Extract legends
-# p1_legend <- get_legend(p1)
-# p2_legend <- get_legend(p2)
-# p3_legend <- get_legend(p3)
-#
-# # Remove legends from plots
-# p1 <- p1 + theme(legend.position = "none")
-# p2 <- p2 + theme(legend.position = "none")
-# p3 <- p3 + theme(legend.position = "none")
-
-# Also remove the axis from the second and third plot
-p2 <- p2 + rremove("y.text")
-p3 <- p3 + rremove("y.text")
-
-# Adjust margins so plots can be put closer together
-p1 <- p1 + theme(plot.margin = unit(c(0.1, 0, 0.1, 0), "cm"))
-p2 <- p2 + theme(plot.margin = unit(c(0.1, 0, 0.1, -0.75), "cm"))
-p3 <- p3 + theme(plot.margin = unit(c(0.1, 0, 0.1, -0.75), "cm"))
-
-# Arrange them and add legends pack
-p <- ggarrange(p1, p2, p3, nrow = 1, align = "hv")
-ggsave("04_Manuscript/99_Difference.png", p, width = 10, height = 5)

### 2023-01-19 Chapter_8/02_R-Scripts/02_Visualization/Heatmaps.R

@@ -1,365 +1,0 @@
-################################################################################
-#### Heatmaps
-################################################################################
-# Description: Visualization of the heatmaps
-
-# Clear R's brain
-rm(list = ls())
-
-# Change the working directory
-wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
-setwd(wd)
-
-# Load required packages
-library(terra)          # To handle spatial data
-library(raster)         # To handle spatial data
-library(tidyverse)      # To wrangle data
-library(lubridate)      # To handle dates
-library(ggspatial)      # For scale bar and north arrow
-library(rgdal)          # To handle spatial data
-library(sf)             # To handle spatial data
-library(scales)         # To squish oob values
-library(latex2exp)      # For easy latex code
-library(RColorBrewer)   # For custom colors
-library(ggpubr)         # To arrange multiple plots
-
-# Load custom functions
-source("02_R-Scripts/00_Functions.R")
-
-# Load shapefiles that we want to plot
-water <- read_sf("03_Data/02_CleanData/MajorWaters.shp")
-areas <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
-roads <- read_sf("03_Data/02_CleanData/Roads.shp")
-afric <- read_sf("03_Data/02_CleanData/Africa.shp")
-vills <- read_sf("03_Data/02_CleanData/Villages.shp")
-vills <- cbind(st_drop_geometry(vills), st_coordinates(vills)) %>%
-  rename(x = X, y = Y)
-
-# Prepare a custom color ramp (I don't like ggplots version of spectral)
-spectral <- colorRampPalette(rev(brewer.pal(11, name = "Spectral")))
-
-# Load the reference raster
-r <- raster("03_Data/02_CleanData/ReferenceRaster.tif")
-r <- as.data.frame(r, xy = T)
-
-# Create country labels
-labels_countries <- data.frame(
-    x     = c(25.5, 26, 25.7, 21.5, 23.5)
-  , y     = c(-19.3, -18.2, -17.6, -17.6, -17.8)
-  , Label = c("BOTSWANA", "ZIMBABWE", "ZAMBIA", "ANGOLA", "NAMIBIA")
-)
-
-# Create labels for some geographical landmarks
-labels_waters <- data.frame(
-    x     = c(22.6, 23.7, 27.1, 25.6)
[…]

### 2023-01-19 Chapter_8/Backups/02_Visualization/Heatmaps.R

@@ -8,7 +8,6 @@ rm(list = ls())
 
 # Change the working directory
 wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
-# wd <- "C:/Users/david/switchdrive/University/15. PhD/Chapter_8"
 setwd(wd)
 
 # Load required packages
@@ -22,12 +21,14 @@ library(sf)             # To handle spatial data
 library(scales)         # To squish oob values
 library(latex2exp)      # For easy latex code
 library(RColorBrewer)   # For custom colors
+library(ggpubr)         # To arrange multiple plots
 
 # Load custom functions
 source("02_R-Scripts/00_Functions.R")
 
 # Load shapefiles that we want to plot
-area  <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
+water <- read_sf("03_Data/02_CleanData/MajorWaters.shp")
+areas <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
 roads <- read_sf("03_Data/02_CleanData/Roads.shp")
 afric <- read_sf("03_Data/02_CleanData/Africa.shp")
 vills <- read_sf("03_Data/02_CleanData/Villages.shp")
@@ -55,23 +56,34 @@ labels_waters <- data.frame(
   , Label = c("Okavango\nDelta", "Linyanti\nSwamp", "Lake\nKariba", "Makgadikgadi\nPans")
 )
 
-# Add albels for source areas
-labels_source <- data.frame(
-    x     = c(23.4, 22, 23.5, 22.7, 23.1, 23.4, 21.55, 24.1, 22.6)
-  , y     = c(-20.3, -19.4, -18.9, -18.3, -19.4, -20.5, -19.5, -19.5, -18)
-  , Label = 1:9
-)
+# Generate area labels
+labels_areas <- st_coordinates(st_point_on_surface(areas))
+labels_areas <- cbind(labels_areas, st_drop_geometry(areas))
 
 # Load the heatmaps and keep only desired columns
-maps <- "03_Data/03_Results/HeatmapsBetweennessGlobal.rds" %>%
+maps1 <- "03_Data/03_Results/HeatmapsBetweennessGlobal.rds" %>%
   read_rds() %>%
-  select(Steps, FloodLevel, Data = Heatmap) %>%
-  subset(Steps == 2000 & FloodLevel != "Mean")
+  mutate(SourceArea = NA, Level = "Global") %>%
[…]

### 2023-03-14 Chapter_1/04_Manuscript/Manuscript.tex
upd
various updates

@@
 \textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
+Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
 Switzerland.
 
-\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
-Maun, Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+Private Bag 13, Maun, Botswana.
 
@@
 directional. An interesting generalization that bridges the continuum between
-LCPA and CT has been proposed by \citep{Panzacchi.2016} and enables to
+LCPA and CT has been proposed by \cite{Panzacchi.2016} and enables to
 capitalize on the merits of both approaches. Despite these and several other
@@
 traversal frequency) as we gradually increased the number of simulated
-trajectories from 1 to 50,000. To assess variability in the relative traversal
+trajectories from 1 to 50'000. To assess variability in the relative traversal
 frequency, we repeatedly subsampled 100 times from all 50'000 trajectories and
@@
 data that was collected on eight dispersing coalitions between 2019 and 2022
-(totalling to 2'668 GPS locations). We used path selection analysis (PSF,
+(totalling to 2'668 GPS locations). We used a path selection function (PSF,
 \citealp{Cushman.2010}) to assess if observed dispersal trajectories followed
@@
     Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
-    Results from the k-fold cross validation procedure. Subfigure b1 shows rank
+    Results from the k-fold cross-validation procedure. Subfigure b1 shows rank
     frequencies of realized steps according to model predictions with known
@@
 Dispersal simulations based on the most parsimonious movement model proved
-useful for assessing landscape connectivity. Of the 50,000 simulated dispersal
+useful for assessing landscape connectivity. Of the 50'000 simulated dispersal
 trajectories that originated from the main study area, only 4.5\% reached a map
@@
 all checkpoints showed that the relative traversal frequency reached a steady
-state after 10,500 simulated dispersal trajectories (Figure S4). Although
+state after 10'500 simulated dispersal trajectories (Figure S4). Although
 variability in the relative traversal frequency kept decreasing as we increased
@@
 our approach of simulating dispersal proved usedful to assess landscape
-connectivity, it was computationally costly. Simulating 80,000 dispersal
+connectivity, it was computationally costly. Simulating 80'000 dispersal
 trajectories for 2'000 steps across the KAZA-TFCA required five days of
@@
 simulated trajectories through randomly placed checkpoints across our study area
-converged already after 10,500 runs. The exact number of required simulations to
+converged already after 10'500 runs. The exact number of required simulations to
 achieve reliable estimates of connectivity will, of course, vary depending on

### 2023-03-14 Chapter_2/04_Manuscript/Manuscript.tex
various updates

@@
 \subsection{Spatial Habitat Layers}
-We attempted to represent each spatial covariate at the highest temporal
-resolution available, which resulted in various degrees of temporal accuracy. In
-order to generate spatial layers that depict the flood extent at different
+In order to generate spatial layers that depict the flood extent at different
 points in time, we ... \Cref{FloodExtent}.

### 2023-03-14 Chapter_8/04_Manuscript/Manuscript.tex
various updates

--- 2023-01-04 Chapter_8/04_Manuscript/Manuscript.tex
@@
 %------------------------------------------------------------------------------
-\title{Flooding of the Okavango Delta influences Connectivity for Dispersing
-African Wild Dogs}
+% \title{Flooding of the Okavango Delta influences Connectivity for Dispersing
+% African Wild Dogs}
+\title{African Wild Dog Dispersal and Connectivity under Climate Change -
+Lessons Learned from Seasonal Flood Extremes}
 
@@
   John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
+  Arpat Ozgul\textsuperscript{1,2} \orcid{0000-0001-7477-2642} \and
   Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
@@
 
-\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
-Maun, Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+Private Bag 13, Maun, Botswana.
 
@@
 
-\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
-Consequences for African Wild Dog Dispersal and Connectivity
+\textbf{Running Title:} African Wild Dog Dispersal in a Changing Climate:
+Lessons Learned from Seasonal Flood Extremes
 
@@
 
-\textbf{Keywords:} movement, connectivity, seasonality, dispersal, conservation,
-permeability
+\textbf{Keywords:} movement, connectivity, climate change, seasonality,
+dispersal, conservation, individual-based simulations, step-selection
 
@@
 Climate change is expected to profoundly impact the life history of wild-living
-animal populations. While the impact of climate change on the demographics of
-local subpopulations has been studied repeatedly, little is known about the
-consequences of environmental change on dispersal and connectivity.
+animal populations. While the impacts of climate change on the demographics of
+local subpopulations have been studied repeatedly, little is known about the
+consequences for dispersal and connectivity.
 
-We capitalize on a ``natural experimental setup'', the flood-pulse driven
-environmental change across the Okavango Delta in northern Botswana, to
-investigate the impact of changing environmental conditions on dispersal
-patterns and connectivity of the endangered African wild dog (\textit{Lycaon
-pictus}). For this, we simulate dispersal trajectories across the landscape of
-the Okavango Delta under two extreme environmental scenarios; one assuming a
-maximum flood extent, one assuming a minimum flood extent.
+We capitalize on a ``natural experimental setup'', the flood-pulse driven change
+in surface-water across the Okavango Delta in northern Botswana, to investigate
+the impact of changing environmental conditions on dispersal patterns and
+connectivity of the endangered African wild dog (\textit{Lycaon pictus}). For
+this, we simulate dispersal trajectories across the Okavango Delta under two
+extreme scenarios that serve to represent environmental conditions akin to those
+expected under continued climate change; one assuming a maximum flood, one
+assuming a minimum flood.
 
-During maximum flood, we observe a reduction in connectivity and an increase in
-dispersal durations between distinct habitat patches. Notably, dispersal into
-the center of the Okavango delta is reduced by 80\% during maximum flooding. At
-minimum flooding, conversely, the delta reveals vital dispersal corridors and
-increases chances of successful dispersal into neighboring habitat patches by xx
-percent. Concomitantly, dispersal durations to move between patches is reduced,
-on average by xx days,at low flood levels.
+During maximum flood, the Okavango Delta poses an important dispersal barrier
+that reduces dispersal prospects in increases dispersal durations between
+distinct areas. Across the entire study area, we observe
+$\input{99_NumberReachingOthersPercentage.tex}$\% lower dispersal success and
+$\input{99_StepsToReachingOthersPercentage.tex}$\% longer dispersal durations
+during maximum flood. Most notably, dispersal into the central habitats of the
+Okavango Delta is reduced by $\input{99_NumberReachingArea6Percentage.tex}$\%
+with an accompanied increase in dispersal durations of
+$\input{99_StepsToReachingArea6Percentage.tex}$\%. Depending on the flood,
+dispersal corridors traversed different areas and dispersers moved into
+proximity of different human-dominated areas.
 
-Climate change is expected to critically impact the flood-pulsing cycles of the
-Okavango delta and may therefore substantially alter landscape connectivity in
-this area. Our analysis suggests that anticipating such alterations in
-environmental conditions is key when implementing conservation strategies.
-Besides a better understanding of the conservation needs for the African wild
-dog, our study also presents a first step towards incorporating environmental
-changes due to seasonality or climate change in dispersal and connectivity
-analyses and demonstrates the usefulness of individual-based dispersal
-simulations as a pertinent conservation tool to study impacts of environmental
-change on dispersal success.
+Whilst the exact impacts of climate change on the flooding regime of the
+Okavango Delta remain unknown, our results suggest that connectivity will vastly
+differ depending on future flood conditions. Acknowledging such differences will
+be key to design effective conservation strategies, especially in light of
+ongoing climate change. Since we highlight critical dispersal corridors and
+human-wildlife conflict zones for two distinct future scenarios, our results
+will facilitate the evidence-based conservation of the endangered African wild
+dog.
 
@@
 \section{Introduction}
-Climate change is expected to profoundly impact ecosystems worldwide with
-far-reaching consequences for species living therein \citep{Ozgul.2010,
+\subsection{Climate Change and Dispersal}
+Climate change is expected to profoundly impact ecosystems across the globe with
+far-reaching consequences for the species living therein \citep{Ozgul.2010,
  Radchuk.2019, IPCC.2022}. By altering environmental conditions, climate change
@@
 distribution of wild living animal populations \citep{Thomas.2004,
-Thuiller.2006}. One potential life-history pathway through which species may
-mediate the effects of environmental change is dispersal \citep{Anderson.2012}.
-By bringing individuals away from their natal location \citep{Clobert.2012},
-dispersal equips species with a means to adapt to climate change by tracking
-favorable habitat conditions \citep{Raia.2012} and to shift into a different
-part of their fundamental niches \citep{Kokko.2022}. Dispersal furthermore
-provides resilience against changing conditions by ensuring genetic diversity
-\citep{Frankham.2002, Leigh.2012, Baguette.2013}, facilitating the colonization
-empty habitats \citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and
-promoting the reinforcement of weakened and small subpopulations
-\citep{Brown.1977}. As such, it is little surpsing that improving connectivity
-is viewed as the number one strategy to combat the adverse effects of climate
-change \citep{Heller.2009}.
+ Thuiller.2006}. An important life-history pathway through which species may
+mediate the negative consequences of climate change is dispersal
+\citep{Anderson.2012}, i.e. the movement of individuals away from their natal
+location to the site of first reproduction \citep{Clobert.2012}. Through
+dispersal, species may adapt to climate change by tracking favorable habitat
+conditions \citep{Raia.2012} and by shifting into a different region of their
+fundamental niche \citep{Kokko.2006}. Dispersal also facilitates the
+colonization empty habitats \citep{Gustafson.1996, Hanski.1999a,
+ MacArthur.2001}, promotes the reinforcement of weakened and small subpopulations
+\citep{Brown.1977}, and safeguards genetic diversity \citep{Frankham.2002,
+ Leigh.2012, Baguette.2013}, thus providing additional resilience against
+changing environmental conditions \citep{Kokko.2006, Fahrig.2003}.
 
-While dispersal offers a means to offset the demographic consequences of climate
-change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a function of
-climatic and environmental conditions (e.g. \citep{Behr.2020}). The link between
-dispersal and climaticor environmental can either be indirect, for example if
-the propensity of individuals to disperse depends on environmental conditions,
-or direct, when the biophysical environment through which dispersers move
-affects dispersal prospects \citep{Travis.2013}. As the latter case indicates,
-dispersal is also inextricably linked to the concept landscape connectivity
-\citep{Baguette.2013}, which is understood as the degree by which the landscape
-facilitates or impedes movements \citep{Taylor.1993}. Only if a sufficient
-degree of landscape connectivity is maintained, successful dispersal of
-individuals can be guaranteed \citep{Fahrig.2003}. Unfortunately, our
-understanding how changing environmental conditions influence connectivity and
-dispersal is limited \citep{Kokko.2006, Travis.2013}. Because collecting
-information on dispersing animals is difficult and because it is difficult to
-assess how changing environmental conditions affect connectivity...
+\subsection{Connectivity}
+While dispersal offers a means to offset the negative demographic consequences
+of climate change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a
+function of climatic and environmental conditions (e.g. \citealp{Elliot.2014,
+ Behr.2020}). The link between dispersal and the environment can either be
+indirect, for example if the propensity of individuals to disperse depends on
+environmental conditions, or direct, when the biophysical environment through
+which dispersers move affects dispersal prospects \citep{Travis.2013}. The
+latter highlights that dispersal is also inextricably linked to the concept
+landscape connectivity \citep{Baguette.2013}, which is understood as the degree
+to which the landscape facilitates or impedes movements \citep{Taylor.1993}. A
+sufficient degree of landscape connectivity is thus a critical prerequisite for
+successful dispersal \citep{Fahrig.2003}, yet the continued degradation and
+destruction habitats worldwide continues to imperil the dispersal prospects of
+many species \citep{Melbourne.2008, Sawyer.2011}. Conservation strategies that
+aim at facilitating dispersal by improving landscape connectivity are therefore
+often viewed as pinnacle of all conservation strategies \citep{Heller.2009}.
+Despite this, our understanding of dispersal and its implications for
+connectivity is limited, especially in light of changing environmental
+conditions.
 
-The continued degradation and destruction habitats across the globe impedes
-dispersal, promotes inbreeding and genetic drift and increases the
-likelihood of local extirpation due to stochastic events \citep{Sawyer.2011}.
-Combined, these factors limit species ability to adapt to changing environmental
-conditions \citep{Fahrig.2003}.
+\subsection{Modeling Connectivity}
+To study dispersal and connectivity, various modeling techniques have emerged
+(see e.g. \citealp{Etherington.2016} and \citealp{Diniz.2019} for overviews).
+Initially, the techniques were limited to examining structural aspects of
+connectivity by focusing on the composition and configuration of habitat
+patches, while ignoring species' responses to the landscape matrix
+\citep{Tischendorf.2000, Doerr.2011}. With the increasing availability of
+telemetry data and methods to study species' habitat and movement preferences
+\citep{Boyce.2002, Fortin.2009, Cushman.2010, Avgar.2016}, preferably during
+dispersal \citep{Elliot.2014}, however, the focus has shifted towards a more
+functional view on connectivity, which also takes into account how species
+interact with their surroundings \citep{Tischendorf.2000, Doerr.2011}.
+Currently, the most prominent \textit{functional} connectivity models are based
+on least-cost path analysis (LCPA, \citealp{Adriaensen.2003}) and circuit theory
+(CT, \citealp{McRae.2008}), two graph-based methods that estimate conductance of
+the landscape by means of a resistance (or inversely permeability) surface
+\citep{Zeller.2012}. Such a surface is meant to reflect the ease or willingness
+at which the focal species traverses a specific area and is generated by
+consolidating multiple habitat layers into a single layer of resistance
+\citep{Zeller.2012}. Since both LCPA and CT approaches make assumptions that are
+rarely met by dispersing individuals, individual-based movement models (IBMMs),
+in which dispersal movements are simulated explicitly, have also gained some
+momentum \citep{Kanagaraj.2013, Allen.2016, Hauenstein.2019, Diniz.2019,
+ Zeller.2020, UnnithanKumar.2022, UnnithanKumar.2022a, Hofmann.2023}. IBMMs
+provide great modeling flexibility and are thus considered powerful tools for
+examining connectivity under different landscape configurations
+\citep{Littlefield.2019, UnnithanKumar.2022a}. However, most connectivity
+studies focus on a snapshot in time and fail to account for changing
+environmental conditions, such as those akin to climate change. Moreover, the
+challenges associated with studying dispersing animals further impairs the
+collection of data during dispersal at the appropriate temporal and spatial
+scale \citep{Graves.2014, Vasudev.2015} and weakens our ability to project
+dispersal prospects under changing environmental conditions into the future.
 
-\cite{Ashrafzadeh.2019} found that cimate change is likely to substantially
-reduce connectivity for mountain newts in Iran. Similarly, \cite{Luo.2021}
-predicted significant reductions in connectivity for the giant spiny frog in
-South-East Asia.
+\subsection{Climate Change and Seasonality}
+Predicting the impacts of climate change on dispersal and connectivity is
+non-trivial and typically requires spatial information about future climatic or
+environmental conditions over the area of interest \citep{Littlefield.2019}.
+This information can then be used in various ways. \cite{Ashrafzadeh.2019}, for
+example, combined climatic predictions until 2070 with a species-distribution
+model for mountain newts (\textit{Neurergus kaiseri}) in Iran to demonstrate a
+decrease in connectivity due to increased habitat fragmentation. Similarly,
+\cite{Luo.2021} mapped the future distribution of the giant spiny frog
+(\textit{Quasipaa spinosa}) under different representative climate pathways and
+reported a reduction in connectivity for the species across South-East Asia. In
+these studies, the focus lies on the impacts of climate change on species
+distribution and subsequent changes in connectivity due to the configuration of
+habitat patches, yet less on the habitat matrix and its implications for
+dispersal. For martens, (\textit{Martes americana}), \cite{Wasserman.2012}
+developed several resistance layers emerging under different climate scenarios
+and find that already low warmings will result in increased isolation of
+remaining subpopulations. While not primarily focused on climate change, another
+body of literature captures environmental variability by generating resistance
+surfaces for different scenarios. \cite{Mui.2017}, for instance, developed
+seasonal resistance maps for Blanding's turtle \textit{Emydoidea blandingii}
+showing that connectivity was substantially lower in late summer compared to
+spring. Similarly, \cite{Osipova.2019} studied connectivity for African
+elephants (\textit{Loxodonta africana}) during wet and dry season and found that
+ignoring seasonality resulted in an underestimation of connectivity during the
+wet season and an overestimation during the dry season. For the same species,
+\cite{Kaszta.2021} provide monthly updated connectivity maps revealing that
+connectivity varies strongly across a typical year. Finally, \cite{Zeller.2020}
+use dynamic resistance surfaces showing differences in connectivity for black
+bears \textit{Ursus americano}. Altogether, the studies exemplify that
+connectivity should not be regarded as static in time, but dynamic across and
+within years.
 
-However, failing to take into account potential changes in climatic and
-environmental conditions imperils our ability to design cost-effective
-conservation strategies.
+In many cases, anticipating environmental conditions under climate change is not
+viable as relevant data is not available or entails major uncertainty
+\citep{Collins.2012}. This is particularly true for complex ecosystems with
+intricate feedback loops and in cases where one is interested in landscape
+characteristics, rather than climatic conditions. In general, it is accepted
+that aside from increasing temperatures, climate change will also amplify the
+frequency and magnitude of extreme events, such as severe droughts, heavy
+precipitation, floods, and storms \citep{Stott.2016, Ummenhofer.2017,
+ IPCC.2022}. Thus, instead of attempting to study the impacts of climate change
+directly, one may capitalize on naturally occurring fluctuations of the
+environment to gauge the likely consequences of shifting the system towards what
+is currently considered an extreme.
 
-Here, we used naturally occuring seasonal flood-cycles to approximate the likely
-changes in connectivity that will arise under amplified climatic conditions due
-to climate change.
-While a better understanding of the ODs flooding regime under global change is of major interest, reliable predictions remain challenging due to the complex interplay between climate,
-anthropogenic water use, animal behavior, and topography of the region.
+\subsection{Okavango Delta}
+The Okavango Delta (OD) in Southern Africa poses a unique opportunity to study
+the impacts of extreme environmental change on species dispersal ability and
+connectivity in a large scale natural experiment setup. The OD is the world's
+largest inland delta and characterized by substantial seasonal differences in
+surface-water cover. Throughout the course of a year, the area covered by the
+OD's floodwaters can fluctuate between 3'000 and 10'000 km\textsuperscript{2}
+with striking variability within and between years \citep{Gumbricht.2004,
+ Wolski.2017}. Importantly, the region is among the most vulnerable to climate
+change, as a temperature increase of 4 to 6\degree C above pre-industrial levels
+is expected within the 21\textsuperscript{st} century \citep{Engelbrecht.2015,
+ Akinyemi.2019}, which is far beyond the global average \citep{IPCC.2022}. A
+keystone predator in this ecosystem and an umbrella species for conservation
+efforts is the African wild dog (AWD, \textit{Lycaon pictus}). While the species
+was once widespread across across entire Sub-Saharan Africa, it has disappeared
+from a vast majority of its historic range, mainly due to human persecution,
+deadly diseases, and continued destruction and degradation of its habitats
+\citep{Woodroffe.2012}. AWDs are characterized by an unsurpassed dispersal
+ability, as young individuals that leave their natal pack can cover several
+hundred kilometers within a few days \citep{Davies-Mostert.2012, Masenga.2016,
+ Cozzi.2020}. Dispersal typically happens in dispersal coalitions of same-sex
+siblings  \citep{McNutt.1996}. Previous research has shown that the ODs
+floodwaters pose a major dispersal barrier, yet the analysis was centered around
+an average flooding scenario \citep{Hofmann.2021}.
 
-Climate change
-could lead to the disruption of existing dispersal corridors and the emergence
-of new dispersal barriers but may also provide possibilities to access
-previously unsuitable dispersal routes. Designing and implementing effective
-conservation strategies that aim to improve landscape connectivity, such as
-habitat restoration or the creation of wildlife corridors, requires better
-understanding of such dynamics.
-
-Albeit the impacts of climate change on connectivity cannot be directly
-estimated, one can study how changing environmental conditions due to
-seasonality influence patterns of dispersal and connectivity.
-
-Habitat degradation and fragmentation lead to isolated subpopulations that have
-a high associated risk of extinction due to stochastic events and generally
-decreased genetic variability \citep{Melbourne.2008}.
-
-More recently, individual-based movement models (IBMMs) have been uprising and
-are recommended for their ability to incorporate several crucial realities of
-animal behavior, such as, for example, autocorrelation, that are by omitted by
-necessity using resistance-based approaches \citep{Hofmann.2023,
-Unnithan.2022a}. IBMMs allow much greater flexibility and are thus considered
-powerful tools for examining connectivity under different landscape
-configurations \citep{Littlefield.2019}.
-
-Botswana is among the countries most heavily influenced by climate change .
-
-For multispecies approaches to connectivity within the KAZA-TFCA, see
-\cite{Brennan.2020} and \cite{Lines.2021}.
-
-Talk about connectivity models
-Talk about functional connectivity
-Talk about habitat fragmentation and destruction
-Say that wild dogs are keystone predators and an umberella species for
-conservation
-
-Unfortunately, little is known about the specific responses of different species
-to climate change, especially for regions most vulnerable to altered conditions
-\citep{Paniw.2021}.
-
-Earlier studies demonstrating the importance of rendering seasonal change
-include \cite{Chetkiewicz.2009}, who use resource selection functions to demonstrate
-that connectivity differs for Ursus actor and Puma concolor and \cite{Mui.2017},
-who showed that connectivity for turtles changes depending on the season, and
-\cite{Kaszta.2021}, who show monthly differences in elephant's habitat selection
-using path-selection functions.
-
-Only if there
-
-The challenges associated with studying dispersing animals have thus far
-impaired the collection of data of dispersers at the appropriate temporal and
-spatial scale \citep{Graves.2014, Vasudev.2015} and weakened our ability to
-project dispersal prospects under changing environmental conditions into the
-future. Nevertheless, a more comprehensive understanding of dispersal under
-changing environmental conditions is critical for designing effective
-conservation strategies in the light of climate change. By predicting and
-modeling dispersal patterns under different scenarios, we can better anticipate
-the impacts of climate change on species and design effective conservation
-strategies to protect them. This is especially important for endangered species
-that are already at the verge of extinction.
-
-The Okavango delta in Southern Africa poses a unique opportunity to study the
-impacts of environmental change on species dispersal ability and connectivity in a
-large scale natural experiment setup. According to projections by the IPCC,
-Botswana and its surroundings are among the most vulnerable to climate change.
-While global temperatures are expected to increase by xx degrees until the end
-of the 21st century, For example, \cite{Engelbrecht.2015} predicts an increase
-between 4 and 6\degree C for regions in southern Africa. One species that
-heavily relies on dispersal for long-term persistence and that inhabits such an
-arid ecosystem in Southern Africa is the African wild dog (\textit{Lycaon
-pictus}). The African wild dog is a lightly built canid lives in cohesive packs
-comprising up to 20 individuals. Wild dogs are social breeders and the majority
-of reproduction is monopolized by the dominant alpha couple who are supported by
-subordinates in taking care of their pups. The species exhibits strong
-inbreeding avoidance, hence juvenile individuals born into the pack disperse in
-same-sex sibling coalitions after reaching sexual maturity \citep{McNutt.1996,
-Behr.2020}. Notably, the African wild dog is the sole extant representative of
-its genus and thus considered as high-priority species for conservation efforts
-\citep{Leigh.2012}. However, fragmentation and destruction of their historic
-habitat may have decreased dispersal success for this species and thus
-jeopardizes connectivity \citep{Leigh.2012}. Since the impacts of climate change
-are not uniformly distributed across the planet, but particularly pronounced at
-high latitudes and arid ecosystems, the species may be seen as most vulnerable
-to such ecosystem shifts.
-
-Talk about Zeller und Osipova
-
-In a previous study, \citep{Hofmann.2022} have investigated landscape
-connectivity in the same ecosystem, yet assuming a static landscape.
-
-Here, we capitalize on a previously parameterized dispersal model and use
-dispersal simulations to investigate dispersal success and connectivity patterns
-for African wild dogs under two extreme scenarios; one assuming maximum flooding
-of the Okavango delta and one assuming minimum flooding of the Okavango delta.
-Given that water poses a barrier to dispersing individuals, we anticipated that
-dispersal prospects and connectivity during maximum flood are low. During
-minimum flood, in contrast, we expect to reveal the presence of several
-dispersal corridors that may be used to freely and quickly move between
-remaining habitat patches.
+\subsection{What We Did}
+Here, we utilize a previously parameterized and validated dispersal model as
+IBMM to simulate dispersal and assess dispersal success and connectivity
+patterns for African wild dogs under two extreme scenarios: one assuming maximum
+flooding of the Okavango delta and one assuming minimum flooding of the Okavango
+delta. The IBMM was trained using GPS data collected during dispersal and
+frequently updated environmental data, thus providing a high degree of realism
+\citep{Hofmann.2021}. Given that dispersers avoid crossing through water (albeit
+we do occasionally observe it in the field), we anticipated that dispersal
+prospects and connectivity during maximum flood would be low. Moreover,  when
+the flood extent of the OD is at its maximum, the water extends almost into the
+densely populated village of Maun. Since both the flood and humans are avoided
+by dispersing wild dogs, we anticipated that a fully flooded Delta would result
+in a total halt of movement between the Western and Eastern side of the OD.
+Information on habitat selection or connectivity is also suitable for predicting
+areas with an elevated potential for human wildlife conflict
+\citep{Buchholtz.2020}. Hence, also quantified areas with high potential for
+human wildlife conflict.
 
 \section{Materials and Methods}
-We conducted all analyses using the programming language \textsf{R}. Any spatial
-data manipulation was completed using the \textsf{terra} and \textsf{spatstat}
-packages. Several helper functions for the simulation algorithm were written in
-\textsf{C++} and imported to R using the \textsf{Rcpp} package. Figures were
-generated using \textsf{ggplot2}, \textsf{igraph}, and \textsf{ggnetwork}. All
-R-scripts used to conduct our analyses are provided through an online
-repository.
+We conducted all data preparation and analyses using the programming language
+\textsf{R} \citep{RCoreTeam.2022}. For any spatial data manipulation, we used
+the packages \textsf{terra} \citep{Hijmans.2022} and \textsf{spatstat}
+\citep{Baddeley.2015}. Several helper functions for the dispersal simulation
+algorithm were written in \textsf{C++} and imported to R using the \textsf{Rcpp}
+package \citep{Eddelbuettel.2011}. Network analysis was achieved in
+\textsf{igraph} \citep{Csardi.2006} and figures were generated using
+\textsf{ggplot2} \citep{Wickham.2016} and \textsf{ggnetwork}
+\citep{Briatte.2021}. All R-scripts required to replicate our analyses are
+provided through an online repository.
 
@@
 surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
-Zimbabwe, and Zambia (\Cref{StudyArea}). The OD is the world's largest inland
-delta and the main driver of seasonal environmental change in the region. While
-our primary focus lied on the immediate surroundings of the Okavango Delta, we
-considered a large rectangular extent stretching from 20\degree 30' E to
-26\degree E. (totaling to an area of 300'000 km\textsuperscript{2}) to
-accommodate for the long distance dispersal events commonly observed in African
-wild dogs (e.g. \citealp{Davies-Mostert.2012, Masenga.2016, Cozzi.2020}). The
-flood-pulsing rhythm of the OD is mainly dictated by precipitation in the
-catchment areas of the Angolian highlands, where rainwater is collected and
-channeled into the OD through the Okavango River. Although precipitation in
-Angola peaks between December and March, water only slowly descends through the
-Okavango river and its distributaries, reaching the distal ends of the delta in
-July or August, where the water percolates at the Thamalakane and Kunyere
-Faults. At minimum extent, the flood covers an area of 3'600
-km\textsuperscript{2}, during maximum flood more than 9'000
-km\textsuperscript{2}. Vegetation in the study area is dominated by mopane
-forest, mixed acacia woodland, and grassland. Human influence is low and mainly
-concentrated around small villages at the western periphery of the delta as well
-as the city of Maun at the south-eastern tip of the OD. Large portions of land
-are dedicated national parks, game reserves or forest reserves. The study area
-is also part of the world's largest transboundary conservation initiative, the
-Kavango-Zambezi Transfrontier Conservation Area, which aims to restore
-connectivity between protected areas in Southern Africa.
+Zimbabwe, and Zambia (\Cref{StudyArea}). While our primary focus lied on the
+immediate surroundings of the Okavango Delta, we considered an extended
+rectangular extent stretching from 20\degree 30' E to 26\degree E. This
+encompasses a total of 300'000 km\textsuperscript{2}) and comprises all long
+distance dispersal events previously recorded in this area \citep{Cozzi.2020,
+ Hofmann.2021}. The annual flood-pulsing rhythm of the OD is mainly dictated by
+precipitation in the Angolan highlands, which serve as catchment areas from
+which water is further channeled into the Okavango River and transported into
+the OD \citep{Wolski.2017}. Although precipitation reaches its maximum between
+December and March, the collected water only slowly descends through the
+Okavango River and its distributaries, reaching the distal ends of the delta
+only towards July or August. In result, peak flooding is out of sync with local
+precipitation, such that the flood usually arrives in the OD during the peak dry
+season. Once the water reaches the OD's distal ends, it percolates at the
+Thamalakane and Kunyere Faults, two natural faultlines at which the waterflow is
+hindered. During minimum extent, the flood covers an area of only 3'600
+km\textsuperscript{2}, whereas during maximum flood more than 9'000
+km\textsuperscript{2} are flooded. Vegetation in this study area is dominated by
+mopane forest, mixed acacia woodland, and grassland. Human influence is low and
+mainly concentrated around small villages at the western periphery of the OD as
+well as the city of Maun at the south-eastern tip of the OD. Large portions of
+land are dedicated national parks, game reserves or forest reserves. The study
+area is also part of the world's largest transboundary conservation initiative,
+the Kavango-Zambezi Transfrontier Conservation Area. Previous studies have
+attributed a high potential of this initiative for improving connectivity for
+various species \citep{Brennan.2020, Lines.2021, Hofmann.2021}.
 
@@
   \includegraphics[width = \textwidth]{99_StudyArea.png}
-  \caption{Study area across which we simulated dispersal. Simulated dispersers
-  were released at random locations within the orange source areas distributed
-  across the delta. Emigration zones (purple) served as checkpoints and enabled
-  us to identify if and where simulated dispersers left the close surroundings
-  of the Okavango delta. These zones were generated using a set of cutlines
-  originating from the center of the delta and roughly cutting an elliptical
-  buffer zone into sections of equal size.}
+  \caption{(a) Location of the study area in Southern Africa across which we
+   simulated dispersing African wild dogs. To mitigate edge effects, our study
+   area comprised a buffer zone (red polygon) within which we randomized
+   covariate values of the habitat layers. (b) The study area was centered on the
+   Okavango Delta and encompassed its immediate surroundings. We initiated
+   simulated dispersers at random locations within one of the six source areas
+   (orange polygons) that we distributed across the delta. Emigration zones
+   (purple polygons) served as checkpoints to identify if and where simulated
+   dispersers left the close surroundings of the Okavango delta. These zones were
+   generated using a set of cutlines (black dotted lines) originating from the
+   center of the delta that dissected an elliptical buffer surrounding the delta
+   into sections of equal size and in accordance with the cardinal directions.}
   \label{StudyArea}
@@
 We represented the physical landscape through which dispersers could move by a
-set of spatially referenced habitat layers, each resolved at 250m x 250m. The
+set of spatially referenced habitat layers, each with a resolution of 250 m. The
 set of layers included water-cover, distance-to-water, tree-cover,
@@
 influences through villages, roads, and agriculture. A detailed description of
-the different habitat layers is provided in \Cref{Hofmann.2021}. Importantly,
+the different habitat layers is provided in \citep{Hofmann.2021}. Importantly,
 the water-cover and derived distance-to water layers were generated using MODIS
-Terra MCD43A4 satellite imagery that was classified using a ``floodmapping''
-algorithm developed by \citep{Wolski.2017} and available through the R-package
-\textsf{floodmapr}. The algorithm allowed us to generate almost weekly updated
-``floodmaps'', thus providing detailed information about the flood-extent at any
-given point in time. In total, we generated 700 floodmaps between the years 2000
-and 2019. Based on these maps, we generated a minimal and maximum flood
-scenario. To create the minimum flood scenario, we averaged the 50 floodmaps
-with smallest flood extent and generated a binary image using areas that were
-inundated in at least 50\% of the maps. Similarly, we created an average image
-for high flood using the 50 most flooded maps. The final maps are depicted in
-\Cref{FloodExtent}.
+Terra MCD43A4 satellite imagery that was classified into binary water-cover maps
+using a ``floodmapping'' algorithm developed by \citep{Wolski.2017}. This
+allowed us to generate almost weekly updated ``floodmaps'', thus providing
+detailed information about the flood-extent at any given point in time. In
+total, we generated 700 floodmaps between the years 2000 and 2019, based on
+which we generated a minimal and maximum flood scenario. To create the minimum
+(maximum) flood scenario, we averaged the 100 floodmaps with the smallest
+(highest) flood extent and generated a binary layer by masking all pixels that
+were inundated in less than 50\% of the maps. The resulting maps are presented
+in \Cref{FloodExtent}. Ultimately, we combined the habitat layers into two
+stacks, one representing the minimum flood scenario, one representing the
+maximum flood scenario. To mitigate edge effects during the dispersal
+simulation, we followed \cite{Koen.2010} and expanded the spatial extent of the
+stacked layers by 20\% and randomized habitat values within the so created
+buffer zone (red rectangle in \Cref{StudyArea}a).
 
@@
   \includegraphics[width = \textwidth]{99_FloodExtent.png}
-  \caption{Flood extent in the two scenarios considered. In the left panel, the
-  flood is at an extremely low level, stretching across 3'558
-  km\textsuperscript{2}, whereas in the right panel the flood is at an extremely
-  high level and covers 9'397 km\textsuperscript{2}. The two maps were generated
-  using 700 remote sensed MODIS MCD43A4 satellite images spanning the years 2000
-  to 2019.}
+  \caption{Flood extent in the minimum (a) and maximum (b) flood scenarios. In
+   the minimum flood scenario (a), water stretched across 3'558
+   km\textsuperscript{2}, whereas during maximum flood (b) it covered 9'397
+   km\textsuperscript{2}. The two maps were generated using 700 remote sensed
+   MODIS MCD43A4 satellite images spanning the years 2000 to 2019.}
   \label{FloodExtent}
@@
 
-\subsection{Dispersal Model}
-Our dispersal model was based on a previously parametrized and validated
-integrated step-selection function (iSSF, \citealp{Avgar.2016}) fitted to GPS
-data of dispersing AWDs \citep{Hofmann.2021}. In step selection functions (SSFs,
-\citealp{Fortin.2005}), observed GPS locations are converted into steps (the
-straight-line traveled between two GPS recordings \citep{Turchin.1998a}) and
-compared to a set of \textit{random} steps in a (mixed effects) conditional
-logistic regression framework \citep{Fortin.2005, Thurfjell.2014, Muff.2020,
-Fieberg.2021}. The model presented in \citep{Hofmann.2021} used dispersal data
-of 16 dispersing African wild dogs from a free-ranging wild dog population in
-northern Botswana. GPS data during dispersal was collected at 4-hourly intervals
-and translated into steps of similar duration. Observed steps were then paired
-step with 24 random steps that were generated using a uniform distribution for
-turning angles (\(-\pi, +\pi\)) and step lengths from a gamma distribution
-fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\) = 0.37). It
-was then assumed that animals assigned to each observed and random step a
-selection score of the form \citep{Fortin.2005}:
-
-\begin{equation}
-\label{EQ1}
-  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
-\end{equation}
-
-Where (\(x_1, x_2, ..., x_n\)) represent the covariate values along each of the
-steps and the (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's relative
-selection strengths \citealp{Avgar.2017} towards these covariates. The benefit
-of \textit{integrated} SSFs over regular SSFs is that they provide a means to
-render two complementary ``kernels''. A movement kernel that describes general
-movement behavior of dispersing AWDs and a habitat kernel that describes
-preferences of AWDs with regards to environmental conditions
-\citep{Fieberg.2021}. iSSFs also allow interactions among the two kernels and
-are thus suitable to render that movement behavior may change depending on
-habitat conditions. A brief summary of the fitted dispersal model is provided in
-Appendix xx.
-
 \subsection{Source Areas and Emigration Zones}
 We simulated dispersing AWDs originating from six distinct source areas located
-in the vicinity of the Okavango delta (\Cref{StudyArea}). For areas one to six
-we selected locations at the delta's periphery that remained dry in both
-scenarios. At those locations, we generated circular buffers with a radius of 20
-km. For source area six, we isolated a polygon covering Chief's Island, a
-peninsula located at the OD's center. Besides source areas, we also generated
-``zones of emigration'' that we used as checkpoints to determine if and where
-simulated individuals left the delta's vicinity (\Cref{StudyArea}). We generated
-emigration zones by first overlaying the OD with an elliptic buffer zone that we
-dissected using a set of cutlines that originated from the ODs center and spread
-according to cardinal points (\Cref{StudyArea}).
+in the vicinity of the OD (\Cref{StudyArea}). We placed source areas in regions
+that remained dry during both the minimum and maximum flooding scenario and are
+known to host viable wild living wild dog populations. While source areas one to
+five were located across the delta's periphery, source area six laid in the OD's
+center. The selection of distinct source areas served to facilitate the
+identification and quantification of the number of successful dispersal events
+between different regions of the OD. Besides source areas, we also generated
+``emigration zones'' that we used as checkpoints to determine if and where
+simulated individuals left the delta's immediate surroundings
+(\Cref{StudyArea}). We generated these zones by first overlaying the OD with an
+elliptic that we dissected into roughly equally sized polygons in accordance
+with cardinal points (\Cref{StudyArea}).
 
 \subsection{Dispersal Simulation}
-For each source area we simulated 1'000 individuals, once assuming a minimum
-flood, once assuming a maximum flood. This resulted in the simulation of 6'000
-individuals for each environmental scenario, hence 12'000 individuals in total.
-The simulation algorithm was based on the algorithm described in
-\Cref{Hofmann.2022} and works as follows. A random location within the source
-area is chosen as a starting point. Originating from the starting point, a set
-of 25 random steps is generated by sampling step lengths from a gamma
-distribution fitted to observed steps (what is a step?) (shape = , scale = ) and
-turning angles from a uniform distibution ($-\pi, +\pi$). Along each random step
-the underlying spatial covariates are extracted and relevant movement metrics
-are computed (i.e. log(sl), cos(ta), ta). The parametrized dispersal model is
-used to predict the probability of each step for being chosen given the steps
-covariate values. One of the steps is sampled based on assigned probabilities
-and the location of the animal is updated. The procedure is then repated until
-the desired number of steps is realized. Here, we simulated each individual for
-2'000 steps, which correspons to the longest dispersal duration recorded in our
-data. The original model was trained using 4-hourly steps, thus a simulated step
-also resembled the movement conducted within four hours. Trajectories resulting
-from such a simulation can be understood as correlated random walks that take
-into account both habitat and movement preferences of dispersing individuals.
+We used a previously parameterized and validated dispersal model to simulate
+dispersal of AWDs. The dispersal model was trained using GPS data of 16 wild dog
+coalitions dispersing across northern Botswana \citep{Hofmann.2023} which was
+fed into an integrated step-selection function (iSSF, \citealp{Avgar.2016}). In
+iSSFs, consecutive GPS locations are converted into steps (the straight-line
+traveled between two GPS recordings \citep{Turchin.1998a}) and compared to a set
+of \textit{random} steps in a conditional logistic regression framework
+\citep{Fortin.2005, Thurfjell.2014, Muff.2020, Fieberg.2021}. Because iSSFs
+capitalize on the autocorrelated nature of the collected data, they provide
+better estimates of connectivity than traditional resource selection approaches
+\citep{Zeller.2016}. The model presented in \cite{Hofmann.2023} comprised of a
+movement kernel, describing how dispersers move across the landscape in the
+absence of habitat selection, a habitat kernel, indicating preferred or avoided
+habitat features, and interactions among the two, i.e. how movement behavior
+changes depending on habitat conditions. According to this model, the main
+characteristics of AWD dispersal movements are avoidance of water, avoidance of
+areas influenced by humans, and a preference for directional and fast movements.
+The model parameters are provided in Appendix SX and explained in
+\citealp{Hofmann.2023}.
 
-% We used a previously parametrized dispersal model to simulate dispersal
-% trajectories of African wild dogs across the study area. To simulate dispersing
-% wild dogs, we employed the simulation framework presented in
-% \cite{Hofmann.2022}. In this framework, a movement model that renders habitat
-% and movement preferences of dispersing individuals is parametrized using
-% step-selection functions. Once parametrized, the model can be employed to
-% simulate virtual dispersers moving across the landscape.
-%
-% We released virtual dispersers at random locations within the four distinct
-% source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
-% individuals moving across the landscape at minimum flood and another xx
-% individuals at maximum flood. For comparison, we also ran the simulations for a
-% medium flood extent, yet the results for this will be presented in the appendix.
+Originating from each of the six source areas, we simulated 2'000 individuals
+dispersing for a total of 2'000 steps. 1'000 individuals were simulated assuming
+a minimum flood, the remaining 1'000 assuming a maximum flood. This resulted in
+the simulation of a total of 12'000 individuals. The simulation procedure was
+based on the algorithm described in \cite{Hofmann.2023} and works as follows. A
+random location within the source area is defined as starting point. Originating
+from the starting point, a set of 25 random steps is generated by sampling step
+lengths from a gamma distribution fitted to observed steps (shape = 0.37, scale
+= 6'316) and turning angles from a uniform distribution ($-\pi, +\pi$). Along
+each random step the underlying spatial covariates are extracted, and relevant
+movement metrics are computed. \(\beta-\)estimates from the fitted model are
+used to predict the probability of each step for being chosen, given the steps
+associated covariates. Among the 25 proposed steps, one is chosen at random
+based on assigned probabilities. The location of the animal is updated, and the
+procedure is repeated until the desired number of steps is realized. Here, we
+simulated each individual for 2'000 steps, corresponding to a dispersal duration
+of 400 days and the longest dispersal duration recorded in this study area
+\citep{Cozzi.2020, Hofmann.2021}. The simulated trajectories can be understood
+as correlated random walks.
 
-\subsection{Dispersal Prospects and Connectivity}
-Based on simulated dispersal trajectories in the two scenarios we quantified
-dispersal sucess and connectivity using three complementary connectivity metrics
-as outlined in \cite{Hofmann.2022}. First, we generated heatmaps depicting the
-frequency at which different areas in the landscape were visited by simulated
-dispersers. Such heatmaps serve to detect dispersal hotspots and areas of
-intense use. However, they are less suitable for detecting pinchpoints and
-bottlenecks that are critical in linking distinct patches. Hence, we also
-computed spatially explicit betweenness scores which are useful to highlight
-exactly such pinchpoints \citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}.
-To compute betweenness, we overlayed the study area with a regular grid with 2.5
-km x 2.5 km grid cells and determined how often simulated individuals
-transitioned from one grid-cell to another. Note that in case the same
-individual repeatedly realized the same cell-transition (e.g. repeatedly moved
-between A-B-A-B...), we only counted a single transition to avoid emphasis on
-regions where individuals moved in circles. With on the so generated
-transitions, we generated a network using the centers of all grid-cells as nodes
-and cell-transitions as weighted connections between the nodes. Based on this
-network we computed weighted betweenness scores using the R-package
-\textsf{igraph}. As a final connectivity metric and metric of dispersal success,
-we calculated the number of successful dispersal events between the different
-source areas as well as towards the emigration zones. We coin this type of
-connectivity ``inter-patch connectivity'' as it relates to the movement between
-distinct patches. Dispersal between two areas was said to be successful whenever
-a trajectory leaving one area intersected with the target area. To gauge the
-dispersal duration needed to move between patches (be consistent with
-``patches'', ``source-areas'' etc.), we also recorded the minimum number of
-steps that individuals moved before arriving at the respective patch.
+\subsection{Derived Metrics}
+Based on simulated dispersal trajectories we quantified connectivity and
+identified areas of elevated potential for human wildlife conflict. Our
+assessment of connectivity was based on the three complementary connectivity
+metrics for IBMMs discussed in \cite{Hofmann.2023}. The set of metrics comprised
+of \textit{heatmaps}, depicting areas of intense use, \textit{betweenness maps},
+highlighting dispersal corridors and bottlenecks and \textit{maps of inter-patch
+ connectivity}, visualizing dispersal success, and duration into distinct habitat
+patches. We generated heatmaps by superimposing the study area with a grid with
+a spatial resolution of 1 km and quantifying the frequency of simulated
+trajectories traversing each grid cell. To compute spatially mapped betweenness
+scores, we overlaid the study area with a grid that had a resolution of 2.5 km
+and determined the frequency at which simulated individuals transitioned from
+one grid-cell to another. A cell-transition was said to occur whenever a
+simulated step crossed from one grid-cell across or into another. In case the
+same individual repeatedly realized the same cell-transition, we only counted a
+single transition to avoid emphasis on regions where individuals moved in
+circles. This resulted in a weighted edge-list that we used to compute weighted
+betweenness scores for each grid-cell, i.e. the importance of the respective
+grid-cell in facilitating movement into adjacent areas
+\citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}. Betweenness was computed
+using the \textsf{igraph} R-package \citep{Csardi.2006}. Because the
+computations associated with calculating betweenness scores are computationally
+more demanding, we deemed the grid size of 2.5 km a sensible compromise between
+efficiency and resolution. As a final connectivity metric, we computed the
+number of successful dispersal events between each of the six distinct source
+areas. We coin this type of connectivity ``inter-patch connectivity''. Dispersal
+between two areas was said to be successful whenever the trajectory of an
+individual leaving one area intersected with the polygon of another area. We
+also estimated the number of individuals that left the OD's vicinity and moved
+into an emigration zone. To quantify the dispersal durations required to move
+between source areas or into emigration zones, we recorded the minimum number of
+steps that individuals moved before arriving at the respective destination.
+Besides connectivity, we also identified zones with a high potential for human
+wildlife conflict. For this, we isolated all simulated locations where simulated
+individuals moved within 500 meters of the nearest human-influenced grid-cell.
+Based on the so isolated coordinates we generated density maps. To highlight
+differences between derived metrics during maximum and minimum flooding, we
+computed difference maps for the heatmap, betweenness map, and human wildlife
+conflict maps.
 
 \section{Results}
-\subsection{General}
-See \Cref{Emigration}
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_Emigration.png}
-  \caption{Caption}
-  \label{Emigration}
-  \end{center}
-\end{figure}
+Figures depicting the derived connectivity and human-wildlife conflict maps are
+provided in \Cref{Metrics}. Difference maps to visualize the differences between
+minimum and maximum flood are given in \Cref{Difference}. For brevity, we will
+here focus on system-wide connectivity patterns and only selectively point to
+regional results. Local connectivity maps derived for each source-area
+separately are presented in the Appendix. As the heatmaps in \Cref{Metrics}a
+reveal, the OD acts as major dispersal barrier during maximum flood yet reveals
+vital dispersal habitat during minimum flood. Differences between maximum and
+minimum flood (\Cref{Difference}a) are particularly pronounced for the region
+between source areas 1 and 2, where few dispersers occur during times of maximum
+flood. In fact, because the floodwaters of the OD reach almost into Maun, the OD
+creates a line of separation between its eastern and western sections. The
+separation is further amplified as the city of Maun is avoided by dispersers in
+both scenarios. Similar patterns are observed on the betweenness maps
+(\Cref{Metrics}b), where several pinch-points and bottlenecks linking source
+area 6 to the surrounding source areas exist during minimum flood. During
+maximum flood, however, these links vanish and instead a single corridor at the
+south-eastern tip of the OD emerges (\Cref{Difference}b). Despite its apparent
+importance in linking the eastern and western sections of the delta, it is
+evident from (\Cref{Metrics}a) that this corridor is only rarely used,
+especially during the maximum flood scenario. As for the potential for human
+wildlife conflict, two clusters emerge (\Cref{Metrics}c). The first cluster lies
+at the inflow of the Okavango Delta between source areas 4 and 5 and is most
+pronounced during minimum flood (\Cref{Difference}c). Another, albeit visually
+less distinct, cluster covers the area at the distal end of the OD, stretching
+from lake Ngami to Maun. This area appears particularly relevant at maximum
+flood (\Cref{Difference}c). Our analysis of inter-patch connectivity further
+demonstrates notable differences in dispersal prospects and dispersal durations
+depending on the extent of the flood (\Cref{Metrics}d and \Cref{IPCTable}).
+While \input{99_NumberReachingOthersMinFlood.tex} simulated dispersers reach
+another source area during minimum extent, only
+\input{99_NumberReachingOthersMaxFlood.tex} do so during maximum extent, thus
+indicating an overall decrease in dispersal success of
+\input{99_NumberReachingOthersPercentage.tex}\% during maximum flood.
+Concomitantly, the average minimum dispersal durations increases by
+\input{99_StepsToReachingOthersPercentage.tex}\%, i.e. from
+\input{99_StepsToReachingOthersMinFlood} steps to
+\input{99_StepsToReachingOthersMaxFlood} steps during maximum flood. These
+differences are particularly pronounced for individuals dispersing into source
+area 6 on Chief's Island. While the area is reached by
+\input{99_NumberReachingArea6MinFlood.tex} simulated individuals during minimum
+flood, only \input{99_NumberReachingArea6MaxFlood.tex}, i.e.
+\input{99_NumberReachingArea6Percentage.tex}\% less, arrive there during maximum
+flood. Furthermore, the dispersal duration into source area six from any other
+source area increases by \input{99_StepsToReachingArea6Percentage}\% from
+\input{99_StepsToReachingArea6MinFlood.tex} steps to
+\input{99_StepsToReachingArea6MaxFlood.tex} steps. In few occasions,
+connectivity between some areas increased during maximum flooding, for instance.
+Temporary emigration increased from \input{99_TemporaryEmigrationMinFlood.tex}
+to \input{99_TemporaryEmigrationMaxFlood.tex} trajectories (i.e.
+by \input{99_TemporaryEmigrationPercentage.tex}\%). Permanent emigration increased
+from \input{99_PermanentEmigrationMinFlood.tex} to
+\input{99_PermanentEmigrationMaxFlood.tex} trajectories (i.e.
+by \input{99_PermanentEmigrationPercentage.tex}\%).
 
-\subsection{Heatmaps}
-Heatmaps produced from simulated dispersal trajectories reveal that the OD acts
-as a major dispersal barrier during periods of high flood, but reveals viable
-dispersal corridors during periods of low flood (\Cref{Heatmaps}). During
-minimum flood, the area north-west of Maun appears to serve as vital dispersal
-habitat. The same area is entirely avoided during maximum flood. Besides
-striking differences in connectivity for the close vicinity of the delta, the
-remainder of the study area shows only marginal differences in connectivity
-between the two scenarios. For instance, in both scenarios the area south of the
-Linyanti swamp appears as frequently visited dispersal habitat. Additional
-heatmaps highlighting differences in connectivity for each source area
-separately are provided in Appendix SX.
+\begin{table}
+ \caption{(a) Dispersal frequency and (b) duration (in steps) between source
+  areas (labeled 1 to 6) and emigration zones (labeled 7 to 14) during minimum
+  and maximum flood.}
+ \label{IPCTable}
+ \includegraphics[width = \textwidth]{99_IPCTable.png}
+\end{table}
 
@@
  \begin{center}
-  \includegraphics[width = \textwidth]{99_Heatmaps.png}
-  \caption{Heatmaps depicting the number of simulated dispersal trajectories
-  traversing each grid-cell in the study area. The left panel shows results for
-  the minimum flood scenario, whereas the right panel shows results for the
-  maximum flood scenario. Source areas (numbered 1-6) from which dispresers were
-  released, and emigration zones (numbered 7-14) are shaded in dark gray.}
-  \label{Heatmaps}
-  \end{center}
-\end{figure}
-
-\subsection{Betweenness}
-The betweenness maps reveal a similar pattern in that connectivity through the
-OD is only pronounced during periods of low floods and vanishes entirely during
-maximum flood (\Cref{Betweenness}). A set of four dispersal corridors meets on
-the central peninsula (source area 5, \Cref{Betweenness}) at minimum flood but
-the same corridors are absent when the flood reaches a maximum extent. Instead,
-a narrow corridor runs north west of Maun, connecting source areas one and two.
-Again, the remainder of the study area is only marginally affected by flooding
-patterns. Additional betweenness maps highlighting differences in connectivity
-for each source area separately are provided in Appendix SX.
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_Betweenness.png}
-  \caption{Betweenness scores highlighting potential dispersal corridors for
-  both the minimum and maximum flood scenario. Areas with high betweenness
-  scores (bright yellow) are used by many simulated individuals to move into
-  adjacent regions and can thus be understood as critical pinch-points. Source
-  areas (numbered 1-6) from which dispresers were released, and emigration zones
-  (numbered 7-14) are shaded in light gray.}
-  \label{Betweenness}
-  \end{center}
-\end{figure}
-
-\subsection{Inter-Patch Connectivity}
-Our analysis of interpatch connectivity demonstrates notable differences in
-dispersal prospects and duration depending on the extent of the flood
-(\Cref{IPCTable}). While $\input{99_NumberReachingOthersMinFlood.tex}$ simulated
-dispersers reach another source area during minimum extent, only
-$\input{99_NumberReachingOthersMaxFlood.tex}$ do so during maximum extent. The
-differences are particularly pronounced for individuals dispersing from or into
-the source area located at the OD's center (\Cref{Interpatch}). While the area
-is reached by $\input{99_NumberReachingArea6MinFlood.tex}$ simulated individuals
-during minimum flood, only $\input{99_NumberReachingArea6MaxFlood.tex}$
-dispersers arrive there during maximum flood. Furthermore, the dispersal
-duration into source area six from any other source area increases from
-$\input{99_StepsToReachingArea6MinFlood.tex}$ to
-$\input{99_StepsToReachingArea6MaxFlood.tex}$. Across all simulations, the
-average dispersal duration before reaching another source area increases from xx
-to xx from the minimum to the maximum flood scenario. Nevertheless, connectivity
-into some areas increases during maximum flooding. Emigration increased slightly
-from xx to xx
-
-For instance, while dispersers .. Additional maps highlighting differences in
-inter-patch connectivty for each source area separately are provided in Appendix
-SX.
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_IPCTable.png}
-  \caption{Dispersal frequency (a) and duration (b) (in steps) between source
-  areas and emigration zones during minimum and maximum flood.}
-  \label{IPCTable}
+  \includegraphics[width = \textwidth]{99_ImmigrationEmigration.png}
+  \caption{Number of individuals emigrating from, or immigrating into a specific
+  source area (focal area). Colors indicate into which other areas emigrants
+  moved or from which other areas immigrants originate. For instance, the most
+  left plot in the upper panel shows the number of individuals moving from
+  source area 1 into the six other source areas during minimum and maximum
+  flood, respectively.}
+  \label{EmigrationImmigration}
  \end{center}
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Interpatch.png}
-  \caption{Inter-patch connectivity into and from source area number six. This
-  patch experienced the most drastic reduction in connectivity in result to a
-  high flood level. During minimum flood there is ample connectivity from and
-  into source areas 1-3. During maximum flood, connectivity is lower and mainly
-  limited to source area 2.}
-  \label{Interpatch}
+  \includegraphics[width = \textwidth]{99_Emigration.png}
+  \caption{Absolute number of simulated trajectories running into each of the
+  designated emigration zones (purple) during minimum and maximum flood.
+  Subfigure (a) depicts the total number of emigrating trajectories, including
+  temporary emigrants that eventually returned into the OD's vicinity. Subfigure
+  (b) only depicts permanent emigration events.}
+  \label{Emigration}
   \end{center}
@@
 
-% \input{IPC_Frequency.tex}
-% \input{IPC_Duration.tex}
-% \input{IPC.tex}
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = 0.8\textwidth]{99_Metrics.png}
+  \caption{(a) Heatmaps, (b) betweenness maps, (c) maps of human wildlife
+   conflict, and (d) maps of inter-patch connectivity, derived from simulated
+   dispersal events. Left panels were derived from the minimum flood scenario,
+   right panels from the maximum flood scenario. Source areas from which
+   dispersers were released are numbered 1-6.  The color scale for betweenness
+   scores in (b) was square-rooted to improve visibility of corridors with lower
+   values. Note that for clarity in (d) we only present links between adjacent
+   source areas. Additional, source-specific maps for each of the four metrics
+   are provided in the appendix.)}
+  \label{Metrics}
+ \end{center}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = \textwidth]{99_Difference.png}
+  \caption{Difference maps generated from the (a) heatmaps, (b) betweenness
+   maps, and (c) maps of human wildlife conflict in \Cref{Metrics}. The maps
+   depict the difference between maximum and minimum flood for each metric (i.e.
+   $Metric_{flood:max} - Metric_{flood:min}$). Orange regions indicate that the
+   respective metric was higher during minimum flood, blue regions that the
+   metric was higher during maximum flood.}
+  \label{Difference}
+ \end{center}
+\end{figure}
 
 \section{Discussion}
-According to our simulations, the propensity to move between the eastern and
-western part of the delta is much lower during maximum extent. This is mainly
-due to the flood-waters and the city of Maun acting as dispersal barriers.
-During maximum extent, the floodwaters of the delta close a gap between the
-delta and Maun that otherwise would serve as dispersal corridor. Anecdotal
-evidence supports this hypothesis, for the only dispersing individuals recorded
-to move from the eastern to the western part of the delta moved at times of low
-flood. In line with this, it appears that a large flood extent pushes dispersing
-individuals to move closer to human inhabited areas such as the village of Maun.
+\subsection{Brief Summary}
+In this study, we used a previously parameterized and validated movement model
+to simulate dispersal trajectories of AWDs across the OD under two extreme
+environmental scenarios: one representing minimal flooding and one representing
+maximum flooding. This approach allowed us to investigate connectivity patterns
+that emerge under extreme environmental conditions, similar to those projected
+under climate change. Predictions of flood conditions across the OD under
+climate change remain ambiguous, yet it is generally agreed that climate change
+will amplify climatic variability and result in either exacerbated or attenuated
+flood events. Our two reference scenarios served to approximate these
+conditions. By providing a comprehensive set of connectivity maps for both
+scenarios, we highlighted how dispersal routes and prospects of the endangered
+AWD changed depending on flood conditions. Our simulations revealed that the
+propensity to move between the eastern and western sections of the OD decreased
+significantly during maximum flood. This effect likely resulted from the
+combined influence of floodwaters and anthropogenic pressures, which together
+formed a dispersal barrier that limited connectivity. When flooding was at a
+minimum, on the other hand, the retracted floodwaters revealed vital dispersal
+habitats that facilitated movement between the western and eastern regions of
+the OD. Anecdotal evidence supports the notion that vital dispersal habitats are
+only available during periods of low flood, for the only dispersal coalition
+recorded to successfully move between the eastern and western delta using GPS
+data was observed at a time of minimal flooding  \citep{Cozzi.2020}. The lack of
+dispersal habitat during maximum flood also resulted in an almost complete
+isolation of Chief's Island, the OD's central peninsular (source area 6 in
+\Cref{StudyArea}b). The peninsular itself remains remains dry in both scenarios,
+yet becomes entirely surrounded by water during times of maximum flood. This
+limits pathways to emigrate or immigrate and results in low dispersal prospects
+for individuals moving towards this area. Despite the general reduction of
+connectivity and increased dispersal durations during maximum flood, the number
+of dispersers moving between some of the source areas increased during maximum
+flood. This was, for instance, the case for movements between sour areas 1 and 4
+and likely resulted from individuals that were deflected by the expanded flood
+and redirected into more accessible habitats. The deflection of individuals by
+the flood also had direct implications for the spatial arrangement of dispersal
+corridors as movement routes that traverse the central parts of the OD
+disappeared and more individuals were funneled through a corridor running
+along the southern fringes of the OD.
 
-Predicting how climate change will impact the dispersal ability of AWDs is
-challenging for multiple reasons. First of all, predicting the flooding patterns
-of the OD under climate change is merely impossible due to the complex feedbacks
-between surface-temperature, soil conditions, precipitation patterns and the
-associated changes in vegetation. Second, the delta is not only prone to changes
-in environmental conditions, but also to changes in anthropogenic use of the
-inflowing water. Finally, it is unclear how AWDs, in fact, how any
-species, will cope with environmental change due to global warming. Even though
-some studies predict that AWD populations are likely to decline under increasing
-temperatures, these studies fail to account for the behavioral plasticity of
-their focal species. AWDs respiratory system, for instance, has evolved as a
-perfect adaptation to high temperatures and AWDs may, in fact, profit from a
-comparative advantage (cite an economist) over their competitors and prey under
-rising temperatures. Although the theory of comparative advantages is a
-fundamental concept in economics, it has yet to find its way into ecological
-studies.
+\subsection{Validating Predictions}
+Although the movement model underlying our simulations was validated using
+independent dispersal data (see \citealp{Hofmann.2023}), assessing the
+reliability of our predictions for the two extreme flooding scenarios remains
+challenging. Firstly, because collecting data of dispersing individiauls is
+difficult \textit{per se} \citep{Fattebert.2015, Cozzi.2020} and, secondly,
+because flood-conditions akin to those studied here only occur rarely
+\citep{Wolski.2017}. Dispersers typically only make up for a small proportion of
+the entire population and predicting the timing of dispersal is often
+non-trivial. Coupled with the limited battery-lifetime of most conventional GPS
+collars, it is logistically unfeasible to collected large amounts of GPS data
+during dispersal. However, as an alternative to GPS data, genetic or
+observational data could also yield insight into functional connectivity.
+Genetic data is often viewed as ultimate measure of functional connectivity
+\citep{Baguette.2013} and thus often serves as validation of connectivity maps
+(e.g. \citealp{Cushman.2010} or \citealp{Spear.2010}). Genetic analysis across
+southern Africa revealed moderate levels of dispersal and identified a
+genetically particularly diverse population cluster located near northern
+Botswana \citep{Tensen.2022}. While such analyses provide valuable insights into
+long-term dispersal patterns, they are unlikely to yield insights on short-term
+connectivity patterns such as those studied here. Observational data, on the
+other hand, readily delivers information on seasonally changing dispersal
+patterns. Such data may not only be collected by trained field assistants, but
+could also be obtained through photographic evidence from tourists in a
+citicen's science approach. AWDs, as well as most other large carnivores, are
+individudally identifiable, either by their coat pattern or other unique
+features. Given a set of georeferenced images, individuals can thus be traced
+through space and time. \cite{Cozzi.2013} provide first evidence on the
+usefulness of this approach. As of today, however, the moderate amount of data
+collected and spatially biased sampling effort prohibits using such data for
+validation purposes. Finally, our representation of the flood was, by design,
+focused on rare extreme events. Since the beginning of our dataset.
+Consequently, the amount of data suitable for validation is limited.
 
-\cite{Murray-Hudson.2006} predicted that increased temperatures, additional human
-abstractions, and reduced river flows might lead to a ``Delta dying'' and that
-the impact of climate will be much more pronounced than the impact of
-anthropogenic water use.
+\subsection{Climate Change in the Okavango Delta}
+Despite the importance of the OD as a driver of ecosystem functioning, species
+distribution, and dispersal corridors, predicting flood patterns under future
+conditions has proven notoriously difficult \citep{Wolski.2008}. This is owed to
+the intricate interplay between climatic conditions, anthropogenic water usage,
+and the topographic peculiarities of the region. In regard to climatic
+conditions, Southern Africa is projected to face temperature rises above the
+global average \citep{Engelbrecht.2015}. According to predictions, this will
+cause a more intense but shorter rainy season in Botswana \citep{Akinyemi.2019}.
+Precipitation across the ODs catchment areas in Angola is expected to increase
+but it remains unclear whether elevated precipitation levels will be offset by
+increased temperatures and accelerated evapotranspiration \citep{Wolski.2008,
+Moses.2018}. In a recent study, \cite{Wolski.2008} used three competing climate
+models and predicted that conditions across the delta may range from ``much
+wetter'' to ``much drier''. Accurate predictions of future conditions across the
+OD are further hindered by multi-decadal oscillations in precipitation patterns
+in Angola that cause shifts between wet and dry periods and may offset or
+amplify long term trends over short periods \citep{Wolski.2008, Wolski.2012}.
+Besides climatic uncertainties, the OD's future is plagued by socio-economic
+uncertainties. The OD and its tributaries represent important water-sources for
+adjacent communities and are subject to intense developmental debates about
+future abstractions. These result from an ever-growing human population,
+increasing socio-economic needs, and resettlement in Angola following peace
+\citep{Kgathi.2006} and have culminated in large uncertainties regarding the
+dimensions of future water abstractions \citep{Hughes.2011}. Although upstream
+abstractions along the Okavango River are thought to have relatively little
+impact on the flooding pattern of the OD, the combined effects of climate change
+and anthropogenic abstractions could result in significant ``delta-drying''
+\citep{Murray-Hudson.2006}. A final complicating factor is the OD's shallow
+gradient (1:3300, \citealp{Gumbricht.2004}) in result to which water only slowly
+descends through the OD. Hippos dredge many of the slow-moving waterways and
+thereby ensure a steady flow of water, yet their behavior can also lead to the
+creation of new waterways and thus to a spatial redistribution of floodwater
+across the delta \citep{McCarthy.1998a}. In summary, the significant
+uncertainties regarding future flood conditions preclude clear predictions of
+connectivity and complicate the protection and preservation of important
+movement corridors, especially in light of climate change. Thus, instead of
+anticipating and preparing for a single, clearly defined scenario, conservation
+authorities must maintain flexibility and develop a multitude of strategies that
+can be applied to each case.
 
-Although local rainfalls in Botswana are expected to increase in terms of
-intensity, simulations show that the length of the rainy season will decline,
-more than offsetting the incline in precipitation \cite{Akinyemi.2019}.
+\subsection{Social Resistance}
+Here, we focused on the influence of environmental resistance on dispersal but
+disregarded the impact of social resistance. That is, while our simulation
+rendered how environmental features affect dispersal behavior, it neglected
+potential interactions between dispersing AWDs and their conspecifics,
+predators, or pray. This was a simplifying assumption and owed to a lack of data
+on sympatric species at the appropriate temporal and spatial scale. According to
+the social resistance hypothesis, however, dispersers' movement patterns are
+likely to be driven not only by environmental features, but by a combination of
+environmental-, intra-, and interspecific conditions \citep{Armansin.2019}.
+Together, intra- and inter-specific factors are sometimes referred to the
+\textit{social landscape} through which dispersers move and are thought to be
+important drivers of dispersal behavior \citep{Wey.2015}. Previously, this has
+been demonstrated for dispersing meerkats (\textit{Suricata suricatta}), a
+species surprisingly similar to the AWD in terms of its social organization and
+dispersal behavior \citep{Cozzi.2018}. Similarly, it has been discovered that
+AWDs rely on shared marking sites to advocate presence and reproductive status
+\citep{Apps.2022, Claase.2022}. Although the sites are primarily used by
+resident AWDs, even dispersers or competing species use them occasionally to
+obtain information on the presence of local packs. In fact, dispersers may use
+the marking sites to more effectively navigate the social landscape and avoid
+competitors or locateother-sex dispersal coalitions. Besides such intra-specific
+factors, also the distribution and abundance of prey can be expected to have
+direct consequences on dispersal routes, as dispersers are usually in search of
+suitable, i.e. prey-rich, territory to settle. This goes to show that accounting
+for the social landscape will no only add another level of realism, but also
+additional complications. Interacting species cannot be considered as
+independent of each other, meaning that the impact of climate change on one
+species will result in trophic cascades affecting several species at once with
+substantial alterations in the community composition \citep{Thuiller.2006}.
 
-Although dispersal is generally believed to be beneficial for the long term viability of species, it can also lead to the net-loss of individuals in areas where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.
+AWDs primarily prey on species that are either restricted by access to water or
+by the forage growing in its close proximity. Attraction of prey to water is
+particularly evident during the dry season, when pans and puddles dry up and
+forage becomes scarce across the landscape. During these periods, floodplains
+serve as fallback habitat and attract large herds of ungulates that forage near
+rivers and streams filled by floodwater \citep{Bonyongo.2005, Bennitt.2014}. We
+previously reported that dispersing AWDs also prefer moving near water sources,
+likely in an attempt to track their prey \citep{Hofmann.2021}. However, a
+reduction in the spatial extent of flood and associated losses in floodplain
+habitats could lead to increased concentration of prey and intensified
+competition among large carnivores. Within this carnivore community, AWDs are
+inferior predators and face challenges in competition with lions, spotted
+hyenas. Resident or dispersing AWDs that avoid superior competitors in space or
+time may get forced into areas of moderate prey-availability \citep{Droge.2017}.
+An increased flood-level, on the other hand, will likely limit the amount of
+suitable habitat across the OD's landscape. Recent research on lions has
+demonstrated that flooding reduces the carrying capacity of the system for
+lions, resulting in a crowing within remaining habitats and increased
+competition. It is, however, unclear how an incrased flood level will affect
+inter-specific competition and whether AWDs will persist under elevated levels
+of intra-guild competition.
 
-We studied a population of African wild dogs that resides in a natural
-environment with little human influence. This is only representative for a small
-share of the extant wild dog populations, as most individuals reside in areas
-that are prone to substantial edge effects. For these populations, the benefit
-of dispersal is disputed, as dispersal under high anthropogenic mortality may
-lead to a net-loss of individuals \citep{Leigh.2012}.
+\subsection{Anthropogenic Resistance}
+To this day, the social acceptance of AWDs among the local population of
+northern Botswana has not been investigated. This prohibits a deeper
+understanding of the anthropogenic resistance experienced by dispersing
+individuals \citep{Ghoddousi.2021}. Talk a bit about anthropogenic resistance.
+While our dispersal model rendered dispersers' behavior with regards to the
+presence of humans, it did not take human bheavior into account. This is
+generally referred to as anthropogenic resistance. The employed dispersal model
+rendered how biophysical elements and anthropogenic presence influence dispersal
+movements. It did not, however, account for social or anthropogenic resistance.
+Corridors that are estimated based on landscape features only may over- or
+under-estimate true connectivity. Overestimate in areas where there is strong
+anthropogenic resistance (e.g. hunting, trapping) - Underestimate in areas where
+humans facilitate movements (e.g. through supplementary food supplies). The
+flood might funnel individuals into unsafe areas with high risk of human-caused
+mortality \citep{Northrup.2012} (i.e. ecological traps). Depending on the flood,
+individuals get funneled towards different regions of high anthropogenic
+influence, suggesting that climate change may induce spatial shifts in regions
+with a high potential for human wildlife conflict. Depending on the level of
+anthropogenic resistance that dispersing wild dogs experience in the different
+areas, these regions may act as ecological traps into which individuals get
+funneled due to external conditions.
 
-Dispersal is an important, if not the most important, driver of landscape
-connectivity and therefore of major interest to conservation authorities. It has
-also been demonstrated that dispersal may lead to population declines in areas
-where anthropogenic mortality is high and dispersal prospects low
-\citep{Leigh.2012}.
+% Comparable observations have been made by \cite{Towns.2009}, who show that
+% annual fluctuations in sea-ice extent forces polar bears \textit{Ursus
+% maritimus} to move onshore and more proximal to humans, thus increasing the
+% likelihood for HWC. Consequently, human wildlife conflict may be viewed not as
+% static, but dynamic variable that is prone to seasonal fluctuations and long
+% term climatic trends. This is also what \cite{Kitratporn.2022} find. They report
+% that human wildlife conflict with elephants is expected to increase in some
+% areas (where habitat suitability increases) but to decrease in others (where
+% habitat suitability decreases).
 
-Efforts aiming at restoring or improving landscape connectivity should consider
-and take into account changes in landscape characteristics due to climate
-change.
+% For example, \cite{Michalski.2006} report elevated live-stock depredation by
+% \textit{Puma concolor} and \textit{Panthera onca} on cattle ranches in close
+% proximity to forest corridors. Similarly, changing water temperatures and
+% associated shifts in prey-availability have resulted in white sharks moving in
+% closer vicinity to humans ultimately resulting in an increased frequency of
+% shark attacks\citep{Chapman.2016}.
 
-We assessed the implications of environmental change on the dispersal prospects,
-yet we did not consider how changing conditions alter dispersal propensity.
-
-According to our simulations, dispersers are able to cover larger distances
-during periods of low flood. This finding is little surprising, considering that
-inundated areas act as dispersal barriers and force dispersers to detour and
-circumvent water-covered areas. However, it still leads to an interesting
-hypothesis. Previous studies have shown that the euclidean dispersal distance of
-female coalitions is larger than that of male coalitions. This has led to ...
-However, demographic analyses have also revealed that female offspring tend to
-emigrate from their pack at younger ages and earlier in the year, when
-floodwaters are still at a relatively low level (Behr). It is thus conceivable
-that the sex-differences in reported dispersal distances is mainly a consequence
-of environmental conditions during dispersal, rather than owed to physiological
-differences between sexes.
-
-While our analysis marks an important step into incorporating environmental
-change into studies of connectivity, there are several critical additions that
-should be considered in the future. We studied dispersal and connectivity under
-two different environmental scenarios, yet our movement model assumed that
-dispersers had identical habitat and movement preferences in both scenarios. In
-reality, however, it can be expected that movement and habitat kernels of
-dispersers differ depending on the season considered (examples).
-
-An additional complication arises when species movement is not solely driven by
-environmental conditions, but also affected by itra- and inter-specific factors.
-For instance, ... has shown that dispersers...
-Rendering such conditions alone is challenging, yet rendering the conditions
-under changing environmental conditions is merely impossible.
-
-Climate change will not only change species distribution but also community
-composition \citep{Thuiller.2006}.
-
-To address such differences, researchers could model
-habitat and movement preferences using season-dependent models, or,
-alternatively, by combining hidden markov models with step-selection functions.
-(cite papers that fieberg sent)
-
-For the discussion: Through changing environmental conditions, climate change
-will also impact the distribution of various species with far-reaching
-consequences on inter- and intra-specific competition (Abrahms). Due to a lack
-of data, we omitted any inclusion of social factors in our dispersal model,
-albeit they are known to be important determinants of dispersal (Armansin).
-Studying how climate change influences species distribution and their
-interactions, and ultimately how this alters dispersal, will be challenging but
-necessary to more realistically simulate dispersal.
-
-Only recently, it has been discovered that AWDs communicate using shared marking
-sites. The role of such marking sites for dispersing coalitions remains to be
-investigated, yet it is is likely that, akin to resident packs, use SMS as
-navigation waypoint and demarcation lines. Chemical analyses suggest that the
-compounds used for communication are highly volatile and may not persist in
-extreme climate conditions. In result, dispersers may lose their ability to
-effectively navigate across the landscape and locate potential mates with whom
-to settle. This would reduce pack-formation prospects and undermine...
-
-Validating predictions from individual-based dispersal models is challenging and
-requires additional dispersal data, which is inherently scarce anyways. Sciticen
-science may help to fill this gap by augmenting observed GPS data with
-occasional sightings of uncollared dispersing coalitions. This is especially
-ciritical for species that disperse across borders and beyond confined study
-areas. The African carnivore wildbook offers ...
-
-Although the OD is arguably the main driver of seasonal change across the
-studied ecosystem, there are several other factors that undergoe seasonal
-change, including vegetation and the abundance and distribution of prey or
-predators.
-
-The OD is an important driver of species distribution and it has been found that
-an expanding flood limits available habitat, thus leading to more inter-specific
-competition, particularly between AWDs and lions.
-
-Although validating the predictions resulting from our dispersal model is
-challenging, the simulations enable us to generate several interesting
-hypotheses...
+\subsection{Human Wildlife Conflict}
+It is well documented that a close proximity between humans and wildlife
+increases the likelihood of human-wildlife conflict (e.g.
+\citealp{Michalski.2006} or \citealp{Chapman.2016}). It can thus be expected
+that areas where dispersers move into proximity of human-dominated landscapes
+hold an increased potential for human-wildlife-conflict (HWC). It has been
+suggested that climate change will increase competition between humans and
+wildlife for scarce resources and thereby exacerbate HWC globally
+\cite{Abrahms.2021}. Our simulations suggest that dispersers may indeed utilize
+different routes depending on flood conditions. In our case, tthis did not
+result in an overall increase in HWC, but we observed a regional shift of where
+dispersers come into the vicinity of human-dominated areas. During minimum
+flood, these areas were most prominent along the OD's panhandle, where the
+Okavango River enters the alluvial fan of the OD. The panhandle is inhabited
+comparably densely ($\input{99_HumanDensityPanhandle.tex}$ inh. /
+km\textsuperscript{2}) and used for both agricultural farming
+($\input{99_AgriculturePanhandle.tex}$\% covered by agricultural fields) and
+livestock farming ($\input{99_CattleDensityPanhandle.tex}$ cattle /
+km\textsuperscript{2}). It has previously received attention as a hotspot for
+human-wildlife conflict due to livestock depredation by carnivores
+\citep{LeFlore.2019} and repeated elephant raids \citep{Buchholtz.2020}. During
+maximum flood, in contrast, a larger number of dispersers moved into proximity
+of Maun and the adjacent region of Lake Ngami at the southern fringes of the OD.
+Maun is the biggest and most densely populated city in the study area
+($\input{99_HumanDensityMaun.tex}$ inh. / km\textsuperscript{2}) and serves as
+hub for touristic excursions into the OD. It's surrounding area is
+agriculturally less intensively used ($\input{99_AgricultureMaun.tex}$\% covered
+by agricultural fields) but the livestock density is comparably high
+($\input{99_CattleDensityMaun.tex}$ cattle / km\textsuperscript{2}). Despite
+this and the fact less than 3\% of livestock depredations can be linked to AWDs,
+there have been numerous occasions where AWDs were harmed or killed within the
+city's proximity \citep{Gusset.2009, Cozzi.2020}. While the panhandle and the
+city of Maun themselves are unprotected, they are located near formally
+protected areas and may thus serve as ecological traps for wildlife leaving the
+surrounding protected areas \citep{Woodroffe.1998, Northrup.2012}. Dispersing
+individuals appear to be particularly at risk, as they readily venture outside
+protected areas into hostile landscapes \citep{Elliot.2014, Cozzi.2020}. In
+Zambia, dispersal and the associated mortality have caused deterioration of
+genetic diversity and resulted in the ned loss of individuals
+\citep{Leigh.2012}. Recent genetic analyses across southern Africa, in contrast,
+identified a genetically diverse population cluster near northern Botswana,
+suggesting moderate levels of dispersal \citep{Tensen.2022}. Nevertheless,
+anticipating how climate change and the associated changes in the biophysical
+landscape through which dispersers move will impact hWC will be paramount to
+prioritize efforts and more efficiently conserve and sustain dispersal pathways.
 
 \subsection{Conclusion}
-Our dispersal simulations across two extreme climatic scenarios reveal striking
-differences in dispersal prospects and landscape connectivity for dispersing
-AWDs. This implies that (1) climatic variation, be it due to seasonality or
-climate change, must be included in anlayses dealing with dispersal and (2) that
-projected climate change is likely to have profound impacts on landscape
-connectivity.
+Our dispersal simulations across two extreme environmental scenarios revealed
+striking differences in dispersal prospects and landscape connectivity for
+dispersing AWDs. We thereby showed that extreme environmental conditions, akin
+to those projected under climate change, will have important impacts on
+functional connectivity and may alter areas of HWC. Given the complexity of the
+studied ecosystem and its associated intricate feedback loops, predictions of
+future conditions remain challenging and plagued by uncertainty. Wildlife
+managers and conservation bodies therefore need to move beyond focusing on
+single scenarios and consider multiple possibilities to adequately respond to
+changes in the environment due to climate change, while also coping with the
+socio-economic needs of an ever-expanding human population. This will require
+the development of protection strategies that can accommodate both more extreme
+pronounced, or less intense flood. Successful conservation strategies will be of
+particular relevance for wide-ranging, endangered species that are already at
+the verge of extinction, such as the African wild dog.
 
 \section{Authors' Contributions}
-D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
-D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
+D.D.H., G.C., D.M.B., A.O. and conceived the study and designed methodology;
+D.D.H., G.C., D.M.B., and J.W.M. collected the data; D.D.H. analysed the data;
+G.C., D.M.B., and A.O. assisted with modeling; D.D.H., G.C., and D.M.B.wrote the
 first draft of the manuscript and all authors contributed to the drafts at
@@
 \section{Data Availability}
-GPS movement data of dispersing wild dogs is available on dryad
-\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
-proposed approach using simulated data are provided through Github
-(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
-codes required to reproduce the African wild dog case study will be made
-available through an online repository at the time of publication.
+Access to R-scripts to replicate our anlaysis will be provided through an online
+repository at the time of publication.
 
@@
 Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we would like
-to thank Johannes Signer for assisting with the simulation algorithm. This study
-was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
-Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
-Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
+for sharing her data of three dispersing wild dogs. This study was funded by
+Albert-Heim Foundation, Basler Stiftung für Biologische Forschung, Claraz
+Foundation, Idea Wild, Jacot Foundation, National Geographic Society, Parrotia
+Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
 Forschungkredit der Universität Zürich, and a Swiss National Science Foundation

### 2023-03-14 Chapter_8/04_Manuscript/Manuscript_2023-01-19.tex
various updates

--- 2023-03-14 Chapter_8/04_Manuscript/Manuscript.tex
@@
   John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1,2} \orcid{0000-0001-7477-2642} \and
+  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
   Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
@@
 
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
+Maun, Botswana.
 
@@
 
-\textbf{Running Title:} African Wild Dog Dispersal in a Changing Climate:
-Lessons Learned from Seasonal Flood Extremes
+\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
+Consequences for African Wild Dog Dispersal and Connectivity
+
+\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
+Consequences for African Wild Dog Connectivity under Climate Change
 
@@
 \textbf{Keywords:} movement, connectivity, climate change, seasonality,
-dispersal, conservation, individual-based simulations, step-selection
+dispersal, conservation, individual-based
 
@@
 local subpopulations have been studied repeatedly, little is known about the
-consequences for dispersal and connectivity.
+consequences of environmental change on dispersal and connectivity.
 
@@
 this, we simulate dispersal trajectories across the Okavango Delta under two
-extreme scenarios that serve to represent environmental conditions akin to those
-expected under continued climate change; one assuming a maximum flood, one
-assuming a minimum flood.
+extreme environmental scenarios; one assuming a maximum flood, one assuming a
+minimum flood. These scenarios serve to approximate environmental conditions
+under amplified climatic variability, as it is expected under continued climate
+change.
 
 During maximum flood, the Okavango Delta poses an important dispersal barrier
-that reduces dispersal prospects in increases dispersal durations between
-distinct areas. Across the entire study area, we observe
+that reduces dispersal prospects between distinct areas and increases dispersal
+durations. At maximum flood, dispersal into the central habitats of the Okavango
+Delta is reduced by $\input{99_NumberReachingArea6Percentage.tex}$\% and
+dispersal durations increased by
+$\input{99_StepsToReachingArea6Percentage.tex}$\%. Across the entire study area,
+the differences are less dramatic, with
 $\input{99_NumberReachingOthersPercentage.tex}$\% lower dispersal success and
 $\input{99_StepsToReachingOthersPercentage.tex}$\% longer dispersal durations
-during maximum flood. Most notably, dispersal into the central habitats of the
-Okavango Delta is reduced by $\input{99_NumberReachingArea6Percentage.tex}$\%
-with an accompanied increase in dispersal durations of
-$\input{99_StepsToReachingArea6Percentage.tex}$\%. Depending on the flood,
-dispersal corridors traversed different areas and dispersers moved into
-proximity of different human-dominated areas.
+during maximum flood. We also found that a maximum flood shifted dispersal
+bottlenecks and areas with a high potential for human-wildlife conflict closer
+towards densely populated villages. We provide a rich suite of maps to visualize
+the differences between minimum and maximum flood.
 
@@
 Okavango Delta remain unknown, our results suggest that connectivity will vastly
-differ depending on future flood conditions. Acknowledging such differences will
-be key to design effective conservation strategies, especially in light of
-ongoing climate change. Since we highlight critical dispersal corridors and
-human-wildlife conflict zones for two distinct future scenarios, our results
-will facilitate the evidence-based conservation of the endangered African wild
-dog.
+differ depending on future flood conditions. Considering such differences will
+be key to designing future proof conservation strategies, particularly in light
+of ongoing climate change. Our results also have direct applicability for the
+conservation of the endangered African wild dog as we highlight critical
+dispersal corridors and human-wildlife conflict zones in two distinct scenarios.
 
@@
 in a total halt of movement between the Western and Eastern side of the OD.
-Information on habitat selection or connectivity is also suitable for predicting
-areas with an elevated potential for human wildlife conflict
-\citep{Buchholtz.2020}. Hence, also quantified areas with high potential for
-human wildlife conflict.
 
@@
 which we generated a minimal and maximum flood scenario. To create the minimum
-(maximum) flood scenario, we averaged the 100 floodmaps with the smallest
+(maximum) flood scenario, we averaged the 50 floodmaps with the smallest
 (highest) flood extent and generated a binary layer by masking all pixels that
@@
 stacked layers by 20\% and randomized habitat values within the so created
-buffer zone (red rectangle in \Cref{StudyArea}a).
+buffer zone (red rectangle in \Cref{StudyArea} a).
 
@@
 of \textit{random} steps in a conditional logistic regression framework
-\citep{Fortin.2005, Thurfjell.2014, Muff.2020, Fieberg.2021}. Because iSSFs
-capitalize on the autocorrelated nature of the collected data, they provide
-better estimates of connectivity than traditional resource selection approaches
-\citep{Zeller.2016}. The model presented in \cite{Hofmann.2023} comprised of a
-movement kernel, describing how dispersers move across the landscape in the
-absence of habitat selection, a habitat kernel, indicating preferred or avoided
-habitat features, and interactions among the two, i.e. how movement behavior
-changes depending on habitat conditions. According to this model, the main
-characteristics of AWD dispersal movements are avoidance of water, avoidance of
-areas influenced by humans, and a preference for directional and fast movements.
-The model parameters are provided in Appendix SX and explained in
-\citealp{Hofmann.2023}.
+\citep{Fortin.2005, Thurfjell.2014, Muff.2020, Fieberg.2021}. The model
+presented in \cite{Hofmann.2023} comprised of a movement kernel, describing how
+dispersers move across the landscape in the absence of habitat selection, a
+habitat kernel, indicating preferred or avoided habitat features, and
+interactions among the two, i.e. how movement behavior changes depending on
+habitat conditions. According to this model, the main characteristics of AWD
+dispersal movements are avoidance of water, avoidance of areas influenced by
+humans, and a preference for directional and fast movements. The model
+parameters are provided in Appendix SX and explained in \citealp{Hofmann.2023}.
 
@@
 the simulation of a total of 12'000 individuals. The simulation procedure was
-based on the algorithm described in \cite{Hofmann.2023} and works as follows. A
+based on the algorithm described in \Cref{Hofmann.2023} and works as follows. A
 random location within the source area is defined as starting point. Originating
@@
 provided in \Cref{Metrics}. Difference maps to visualize the differences between
-minimum and maximum flood are given in \Cref{Difference}. For brevity, we will
-here focus on system-wide connectivity patterns and only selectively point to
-regional results. Local connectivity maps derived for each source-area
-separately are presented in the Appendix. As the heatmaps in \Cref{Metrics}a
-reveal, the OD acts as major dispersal barrier during maximum flood yet reveals
-vital dispersal habitat during minimum flood. Differences between maximum and
-minimum flood (\Cref{Difference}a) are particularly pronounced for the region
-between source areas 1 and 2, where few dispersers occur during times of maximum
-flood. In fact, because the floodwaters of the OD reach almost into Maun, the OD
+minimum and maximum flood are given in \Cref{Difference}. Source-specific maps
+are presented in the Appendix. As the heatmaps in \Cref{Metrics}a reveal, the OD
+acts as major dispersal barrier during maximum flood yet reveals vital
+dispersal habitat during minimum flood. Differences between maximum and minimum
+flood (\Cref{Difference}a) are particularly pronounced for the region between
+source areas 1 and 2, where few dispersers occur during times of maximum flood.
+In fact, because the floodwaters of the OD reach almost into Maun, the OD
 creates a line of separation between its eastern and western sections. The
@@
 both scenarios. Similar patterns are observed on the betweenness maps
-(\Cref{Metrics}b), where several pinch-points and bottlenecks linking source
-area 6 to the surrounding source areas exist during minimum flood. During
-maximum flood, however, these links vanish and instead a single corridor at the
+(\Cref{Metrics}b), where several pinch-points and bottlenecks linking source area
+6 to the surrounding source areas exist during minimum flood. During maximum
+flood, however, these links vanish and instead a single corridor at the
 south-eastern tip of the OD emerges (\Cref{Difference}b). Despite its apparent
@@
 depending on the extent of the flood (\Cref{Metrics}d and \Cref{IPCTable}).
-While \input{99_NumberReachingOthersMinFlood.tex} simulated dispersers reach
+While $\input{99_NumberReachingOthersMinFlood.tex}$ simulated dispersers reach
 another source area during minimum extent, only
-\input{99_NumberReachingOthersMaxFlood.tex} do so during maximum extent, thus
+$\input{99_NumberReachingOthersMaxFlood.tex}$ do so during maximum extent, thus
 indicating an overall decrease in dispersal success of
-\input{99_NumberReachingOthersPercentage.tex}\% during maximum flood.
+$\input{99_NumberReachingOthersPercentage.tex}$\% during maximum flood.
 Concomitantly, the average minimum dispersal durations increases by
-\input{99_StepsToReachingOthersPercentage.tex}\%, i.e. from
-\input{99_StepsToReachingOthersMinFlood} steps to
-\input{99_StepsToReachingOthersMaxFlood} steps during maximum flood. These
+$\input{99_StepsToReachingOthersPercentage.tex}$\%, i.e. from
+$\input{99_NumberReachingOthersMinFlood}$ steps to
+$\input{99_NumberReachingOthersMaxFlood}$ steps during maximum flood. These
 differences are particularly pronounced for individuals dispersing into source
-area 6 on Chief's Island. While the area is reached by
-\input{99_NumberReachingArea6MinFlood.tex} simulated individuals during minimum
-flood, only \input{99_NumberReachingArea6MaxFlood.tex}, i.e.
-\input{99_NumberReachingArea6Percentage.tex}\% less, arrive there during maximum
-flood. Furthermore, the dispersal duration into source area six from any other
-source area increases by \input{99_StepsToReachingArea6Percentage}\% from
-\input{99_StepsToReachingArea6MinFlood.tex} steps to
-\input{99_StepsToReachingArea6MaxFlood.tex} steps. In few occasions,
-connectivity between some areas increased during maximum flooding, for instance.
-Temporary emigration increased from \input{99_TemporaryEmigrationMinFlood.tex}
-to \input{99_TemporaryEmigrationMaxFlood.tex} trajectories (i.e.
-by \input{99_TemporaryEmigrationPercentage.tex}\%). Permanent emigration increased
-from \input{99_PermanentEmigrationMinFlood.tex} to
-\input{99_PermanentEmigrationMaxFlood.tex} trajectories (i.e.
-by \input{99_PermanentEmigrationPercentage.tex}\%).
-
-\begin{table}
- \caption{(a) Dispersal frequency and (b) duration (in steps) between source
-  areas (labeled 1 to 6) and emigration zones (labeled 7 to 14) during minimum
-  and maximum flood.}
- \label{IPCTable}
- \includegraphics[width = \textwidth]{99_IPCTable.png}
-\end{table}
-
-\begin{figure}
- \begin{center}
-  \includegraphics[width = \textwidth]{99_ImmigrationEmigration.png}
-  \caption{Number of individuals emigrating from, or immigrating into a specific
-  source area (focal area). Colors indicate into which other areas emigrants
-  moved or from which other areas immigrants originate. For instance, the most
-  left plot in the upper panel shows the number of individuals moving from
-  source area 1 into the six other source areas during minimum and maximum
-  flood, respectively.}
-  \label{EmigrationImmigration}
- \end{center}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_Emigration.png}
-  \caption{Absolute number of simulated trajectories running into each of the
-  designated emigration zones (purple) during minimum and maximum flood.
-  Subfigure (a) depicts the total number of emigrating trajectories, including
-  temporary emigrants that eventually returned into the OD's vicinity. Subfigure
-  (b) only depicts permanent emigration events.}
-  \label{Emigration}
-  \end{center}
-\end{figure}
+area 6. While the area is reached by
+$\input{99_NumberReachingArea6MinFlood.tex}$ simulated individuals during
+minimum flood, only $\input{99_NumberReachingArea6MaxFlood.tex}$, i.e.
+$\input{99_NumberReachingArea6Percentage.tex}$\% less, arrive there during
+maximum flood. Furthermore, the dispersal duration into source area six from any
+other source area increases by $\input{99_StepsToReachingArea6Percentage}$\%
+from $\input{99_StepsToReachingArea6MinFlood.tex}$ steps to
+$\input{99_StepsToReachingArea6MaxFlood.tex}$ steps.
 
@@
 
+\begin{table}
+  \caption{(a) Dispersal frequency and (b) duration (in steps) between source
+  areas (labeled 1 to 6) and emigration zones (labeled 7 to 14) during minimum
+  and maximum flood.}
+  \label{IPCTable}
+  \includegraphics[width = \textwidth]{99_IPCTable.png}
+\end{table}
+
 \section{Discussion}
 \subsection{Brief Summary}
-In this study, we used a previously parameterized and validated movement model
-to simulate dispersal trajectories of AWDs across the OD under two extreme
-environmental scenarios: one representing minimal flooding and one representing
-maximum flooding. This approach allowed us to investigate connectivity patterns
-that emerge under extreme environmental conditions, similar to those projected
-under climate change. Predictions of flood conditions across the OD under
-climate change remain ambiguous, yet it is generally agreed that climate change
-will amplify climatic variability and result in either exacerbated or attenuated
-flood events. Our two reference scenarios served to approximate these
-conditions. By providing a comprehensive set of connectivity maps for both
-scenarios, we highlighted how dispersal routes and prospects of the endangered
-AWD changed depending on flood conditions. Our simulations revealed that the
-propensity to move between the eastern and western sections of the OD decreased
-significantly during maximum flood. This effect likely resulted from the
-combined influence of floodwaters and anthropogenic pressures, which together
-formed a dispersal barrier that limited connectivity. When flooding was at a
-minimum, on the other hand, the retracted floodwaters revealed vital dispersal
-habitats that facilitated movement between the western and eastern regions of
-the OD. Anecdotal evidence supports the notion that vital dispersal habitats are
-only available during periods of low flood, for the only dispersal coalition
-recorded to successfully move between the eastern and western delta using GPS
-data was observed at a time of minimal flooding  \citep{Cozzi.2020}. The lack of
-dispersal habitat during maximum flood also resulted in an almost complete
-isolation of Chief's Island, the OD's central peninsular (source area 6 in
-\Cref{StudyArea}b). The peninsular itself remains remains dry in both scenarios,
-yet becomes entirely surrounded by water during times of maximum flood. This
-limits pathways to emigrate or immigrate and results in low dispersal prospects
-for individuals moving towards this area. Despite the general reduction of
-connectivity and increased dispersal durations during maximum flood, the number
-of dispersers moving between some of the source areas increased during maximum
-flood. This was, for instance, the case for movements between sour areas 1 and 4
-and likely resulted from individuals that were deflected by the expanded flood
-and redirected into more accessible habitats. The deflection of individuals by
-the flood also had direct implications for the spatial arrangement of dispersal
-corridors as movement routes that traverse the central parts of the OD
-disappeared and more individuals were funneled through a corridor running
-along the southern fringes of the OD.
+According to our simulations, the propensity to move between the eastern and
+western part of the delta is much lower during maximum extent. This is mainly
+due to the flood-waters and the city of Maun acting as dispersal barriers.
+During maximum extent, the floodwaters of the delta close a gap between the
+delta and Maun that otherwise would serve as dispersal corridor. Anecdotal
+evidence supports this hypothesis, for the only dispersing individuals recorded
+to move from the eastern to the western part of the delta moved at times of low
+flood. In line with this, it appears that a large flood extent pushes dispersing
+individuals to move closer to human inhabited areas such as the village of Maun.
 
-\subsection{Validating Predictions}
-Although the movement model underlying our simulations was validated using
-independent dispersal data (see \citealp{Hofmann.2023}), assessing the
-reliability of our predictions for the two extreme flooding scenarios remains
-challenging. Firstly, because collecting data of dispersing individiauls is
-difficult \textit{per se} \citep{Fattebert.2015, Cozzi.2020} and, secondly,
-because flood-conditions akin to those studied here only occur rarely
-\citep{Wolski.2017}. Dispersers typically only make up for a small proportion of
-the entire population and predicting the timing of dispersal is often
-non-trivial. Coupled with the limited battery-lifetime of most conventional GPS
-collars, it is logistically unfeasible to collected large amounts of GPS data
-during dispersal. However, as an alternative to GPS data, genetic or
-observational data could also yield insight into functional connectivity.
-Genetic data is often viewed as ultimate measure of functional connectivity
-\citep{Baguette.2013} and thus often serves as validation of connectivity maps
-(e.g. \citealp{Cushman.2010} or \citealp{Spear.2010}). Genetic analysis across
-southern Africa revealed moderate levels of dispersal and identified a
-genetically particularly diverse population cluster located near northern
-Botswana \citep{Tensen.2022}. While such analyses provide valuable insights into
-long-term dispersal patterns, they are unlikely to yield insights on short-term
-connectivity patterns such as those studied here. Observational data, on the
-other hand, readily delivers information on seasonally changing dispersal
-patterns. Such data may not only be collected by trained field assistants, but
-could also be obtained through photographic evidence from tourists in a
-citicen's science approach. AWDs, as well as most other large carnivores, are
-individudally identifiable, either by their coat pattern or other unique
-features. Given a set of georeferenced images, individuals can thus be traced
-through space and time. \cite{Cozzi.2013} provide first evidence on the
-usefulness of this approach. As of today, however, the moderate amount of data
-collected and spatially biased sampling effort prohibits using such data for
-validation purposes. Finally, our representation of the flood was, by design,
-focused on rare extreme events. Since the beginning of our dataset.
-Consequently, the amount of data suitable for validation is limited.
+Predicting how climate change will impact the dispersal ability of AWDs is
+challenging for multiple reasons. First of all, predicting the flooding patterns
+of the OD under climate change is non-trivial due to complex feedback-loops
+between surface-temperature, soil conditions, precipitation patterns, and the
+associated changes in vegetation.
 
-\subsection{Climate Change in the Okavango Delta}
-Despite the importance of the OD as a driver of ecosystem functioning, species
-distribution, and dispersal corridors, predicting flood patterns under future
-conditions has proven notoriously difficult \citep{Wolski.2008}. This is owed to
-the intricate interplay between climatic conditions, anthropogenic water usage,
-and the topographic peculiarities of the region. In regard to climatic
-conditions, Southern Africa is projected to face temperature rises above the
-global average \citep{Engelbrecht.2015}. According to predictions, this will
-cause a more intense but shorter rainy season in Botswana \citep{Akinyemi.2019}.
-Precipitation across the ODs catchment areas in Angola is expected to increase
-but it remains unclear whether elevated precipitation levels will be offset by
-increased temperatures and accelerated evapotranspiration \citep{Wolski.2008,
-Moses.2018}. In a recent study, \cite{Wolski.2008} used three competing climate
-models and predicted that conditions across the delta may range from ``much
-wetter'' to ``much drier''. Accurate predictions of future conditions across the
-OD are further hindered by multi-decadal oscillations in precipitation patterns
-in Angola that cause shifts between wet and dry periods and may offset or
-amplify long term trends over short periods \citep{Wolski.2008, Wolski.2012}.
-Besides climatic uncertainties, the OD's future is plagued by socio-economic
-uncertainties. The OD and its tributaries represent important water-sources for
-adjacent communities and are subject to intense developmental debates about
-future abstractions. These result from an ever-growing human population,
-increasing socio-economic needs, and resettlement in Angola following peace
-\citep{Kgathi.2006} and have culminated in large uncertainties regarding the
-dimensions of future water abstractions \citep{Hughes.2011}. Although upstream
-abstractions along the Okavango River are thought to have relatively little
-impact on the flooding pattern of the OD, the combined effects of climate change
-and anthropogenic abstractions could result in significant ``delta-drying''
-\citep{Murray-Hudson.2006}. A final complicating factor is the OD's shallow
-gradient (1:3300, \citealp{Gumbricht.2004}) in result to which water only slowly
-descends through the OD. Hippos dredge many of the slow-moving waterways and
-thereby ensure a steady flow of water, yet their behavior can also lead to the
-creation of new waterways and thus to a spatial redistribution of floodwater
-across the delta \citep{McCarthy.1998a}. In summary, the significant
-uncertainties regarding future flood conditions preclude clear predictions of
-connectivity and complicate the protection and preservation of important
-movement corridors, especially in light of climate change. Thus, instead of
-anticipating and preparing for a single, clearly defined scenario, conservation
-authorities must maintain flexibility and develop a multitude of strategies that
-can be applied to each case.
+Second, the delta is not only prone to changes
+in environmental conditions, but also to changes in anthropogenic use of the
+inflowing water. Finally, it is unclear how AWDs, in fact, how any
+species, will cope with environmental change due to global warming. Even though
+some studies predict that AWD populations are likely to decline under increasing
+temperatures, these studies fail to account for the behavioral plasticity of
+their focal species. AWDs respiratory system, for instance, has evolved as a
+perfect adaptation to high temperatures and AWDs may, in fact, profit from a
+comparative advantage (cite an economist) over their competitors and prey under
+rising temperatures. Although the theory of comparative advantages is a
+fundamental concept in economics, it has yet to find its way into ecological
+studies.
 
-\subsection{Social Resistance}
-Here, we focused on the influence of environmental resistance on dispersal but
-disregarded the impact of social resistance. That is, while our simulation
-rendered how environmental features affect dispersal behavior, it neglected
-potential interactions between dispersing AWDs and their conspecifics,
-predators, or pray. This was a simplifying assumption and owed to a lack of data
-on sympatric species at the appropriate temporal and spatial scale. According to
-the social resistance hypothesis, however, dispersers' movement patterns are
-likely to be driven not only by environmental features, but by a combination of
-environmental-, intra-, and interspecific conditions \citep{Armansin.2019}.
-Together, intra- and inter-specific factors are sometimes referred to the
-\textit{social landscape} through which dispersers move and are thought to be
-important drivers of dispersal behavior \citep{Wey.2015}. Previously, this has
-been demonstrated for dispersing meerkats (\textit{Suricata suricatta}), a
-species surprisingly similar to the AWD in terms of its social organization and
-dispersal behavior \citep{Cozzi.2018}. Similarly, it has been discovered that
-AWDs rely on shared marking sites to advocate presence and reproductive status
-\citep{Apps.2022, Claase.2022}. Although the sites are primarily used by
-resident AWDs, even dispersers or competing species use them occasionally to
-obtain information on the presence of local packs. In fact, dispersers may use
-the marking sites to more effectively navigate the social landscape and avoid
-competitors or locateother-sex dispersal coalitions. Besides such intra-specific
-factors, also the distribution and abundance of prey can be expected to have
-direct consequences on dispersal routes, as dispersers are usually in search of
-suitable, i.e. prey-rich, territory to settle. This goes to show that accounting
-for the social landscape will no only add another level of realism, but also
-additional complications. Interacting species cannot be considered as
-independent of each other, meaning that the impact of climate change on one
-species will result in trophic cascades affecting several species at once with
-substantial alterations in the community composition \citep{Thuiller.2006}.
+\cite{Murray-Hudson.2006} predicted that increased temperatures, additional human
+abstractions, and reduced river flows might lead to a ``Delta dying'' and that
+the impact of climate will be much more pronounced than the impact of
+anthropogenic water use. However, the paper also found that predicted conditions
+vary from much dryer to much wetter depending on the employed climate model.
 
-AWDs primarily prey on species that are either restricted by access to water or
-by the forage growing in its close proximity. Attraction of prey to water is
-particularly evident during the dry season, when pans and puddles dry up and
-forage becomes scarce across the landscape. During these periods, floodplains
-serve as fallback habitat and attract large herds of ungulates that forage near
-rivers and streams filled by floodwater \citep{Bonyongo.2005, Bennitt.2014}. We
-previously reported that dispersing AWDs also prefer moving near water sources,
-likely in an attempt to track their prey \citep{Hofmann.2021}. However, a
-reduction in the spatial extent of flood and associated losses in floodplain
-habitats could lead to increased concentration of prey and intensified
-competition among large carnivores. Within this carnivore community, AWDs are
-inferior predators and face challenges in competition with lions, spotted
-hyenas. Resident or dispersing AWDs that avoid superior competitors in space or
-time may get forced into areas of moderate prey-availability \citep{Droge.2017}.
-An increased flood-level, on the other hand, will likely limit the amount of
-suitable habitat across the OD's landscape. Recent research on lions has
-demonstrated that flooding reduces the carrying capacity of the system for
-lions, resulting in a crowing within remaining habitats and increased
-competition. It is, however, unclear how an incrased flood level will affect
-inter-specific competition and whether AWDs will persist under elevated levels
-of intra-guild competition.
+While precipitation in the OD's catchment areas are expected to increase, it
+remains unclear wheather the effect will be offset by increased
+evapotranspiration. \cite{Wolski.2018} predict changes that can range from much
+``wetter'' to much  ``drier'' conditions depending on the employed climate
+model. In conclusion, there is too much uncertainty to make reliable
+predictions, meaning that conservation practicioners will need to retain
+flexibility in their planning.
 
-\subsection{Anthropogenic Resistance}
-To this day, the social acceptance of AWDs among the local population of
-northern Botswana has not been investigated. This prohibits a deeper
-understanding of the anthropogenic resistance experienced by dispersing
-individuals \citep{Ghoddousi.2021}. Talk a bit about anthropogenic resistance.
-While our dispersal model rendered dispersers' behavior with regards to the
-presence of humans, it did not take human bheavior into account. This is
-generally referred to as anthropogenic resistance. The employed dispersal model
-rendered how biophysical elements and anthropogenic presence influence dispersal
-movements. It did not, however, account for social or anthropogenic resistance.
-Corridors that are estimated based on landscape features only may over- or
-under-estimate true connectivity. Overestimate in areas where there is strong
-anthropogenic resistance (e.g. hunting, trapping) - Underestimate in areas where
-humans facilitate movements (e.g. through supplementary food supplies). The
-flood might funnel individuals into unsafe areas with high risk of human-caused
-mortality \citep{Northrup.2012} (i.e. ecological traps). Depending on the flood,
-individuals get funneled towards different regions of high anthropogenic
-influence, suggesting that climate change may induce spatial shifts in regions
-with a high potential for human wildlife conflict. Depending on the level of
-anthropogenic resistance that dispersing wild dogs experience in the different
-areas, these regions may act as ecological traps into which individuals get
-funneled due to external conditions.
+Climate models consistently predict an above global average increase in
+temperatures for Southern Africa.
 
-% Comparable observations have been made by \cite{Towns.2009}, who show that
-% annual fluctuations in sea-ice extent forces polar bears \textit{Ursus
-% maritimus} to move onshore and more proximal to humans, thus increasing the
-% likelihood for HWC. Consequently, human wildlife conflict may be viewed not as
-% static, but dynamic variable that is prone to seasonal fluctuations and long
-% term climatic trends. This is also what \cite{Kitratporn.2022} find. They report
-% that human wildlife conflict with elephants is expected to increase in some
-% areas (where habitat suitability increases) but to decrease in others (where
-% habitat suitability decreases).
+There is large uncertainty regarding the future anthropogenic use and
+abstractions from the Okavango river \citep{Hughes.2011}.
 
-% For example, \cite{Michalski.2006} report elevated live-stock depredation by
-% \textit{Puma concolor} and \textit{Panthera onca} on cattle ranches in close
-% proximity to forest corridors. Similarly, changing water temperatures and
-% associated shifts in prey-availability have resulted in white sharks moving in
-% closer vicinity to humans ultimately resulting in an increased frequency of
-% shark attacks\citep{Chapman.2016}.
+It has been suggested that climate change will intensivy human wildlife conflict
+due to increased competition for scarce resources \cite{Abrahms.2021}. Instead
+of \textit{increased} human wildlife conflict, our results suggest that human
+wildlife conflict will shift. Given that these areas have had little exposure to
+such conflicts in the past, inexperience and lack of knowledge may explain the
+resulting apparent increases in conflict situations.
 
-\subsection{Human Wildlife Conflict}
-It is well documented that a close proximity between humans and wildlife
-increases the likelihood of human-wildlife conflict (e.g.
-\citealp{Michalski.2006} or \citealp{Chapman.2016}). It can thus be expected
-that areas where dispersers move into proximity of human-dominated landscapes
-hold an increased potential for human-wildlife-conflict (HWC). It has been
-suggested that climate change will increase competition between humans and
-wildlife for scarce resources and thereby exacerbate HWC globally
-\cite{Abrahms.2021}. Our simulations suggest that dispersers may indeed utilize
-different routes depending on flood conditions. In our case, tthis did not
-result in an overall increase in HWC, but we observed a regional shift of where
-dispersers come into the vicinity of human-dominated areas. During minimum
-flood, these areas were most prominent along the OD's panhandle, where the
-Okavango River enters the alluvial fan of the OD. The panhandle is inhabited
-comparably densely ($\input{99_HumanDensityPanhandle.tex}$ inh. /
-km\textsuperscript{2}) and used for both agricultural farming
-($\input{99_AgriculturePanhandle.tex}$\% covered by agricultural fields) and
-livestock farming ($\input{99_CattleDensityPanhandle.tex}$ cattle /
-km\textsuperscript{2}). It has previously received attention as a hotspot for
-human-wildlife conflict due to livestock depredation by carnivores
-\citep{LeFlore.2019} and repeated elephant raids \citep{Buchholtz.2020}. During
-maximum flood, in contrast, a larger number of dispersers moved into proximity
-of Maun and the adjacent region of Lake Ngami at the southern fringes of the OD.
-Maun is the biggest and most densely populated city in the study area
-($\input{99_HumanDensityMaun.tex}$ inh. / km\textsuperscript{2}) and serves as
-hub for touristic excursions into the OD. It's surrounding area is
-agriculturally less intensively used ($\input{99_AgricultureMaun.tex}$\% covered
-by agricultural fields) but the livestock density is comparably high
-($\input{99_CattleDensityMaun.tex}$ cattle / km\textsuperscript{2}). Despite
-this and the fact less than 3\% of livestock depredations can be linked to AWDs,
-there have been numerous occasions where AWDs were harmed or killed within the
-city's proximity \citep{Gusset.2009, Cozzi.2020}. While the panhandle and the
-city of Maun themselves are unprotected, they are located near formally
-protected areas and may thus serve as ecological traps for wildlife leaving the
-surrounding protected areas \citep{Woodroffe.1998, Northrup.2012}. Dispersing
-individuals appear to be particularly at risk, as they readily venture outside
-protected areas into hostile landscapes \citep{Elliot.2014, Cozzi.2020}. In
-Zambia, dispersal and the associated mortality have caused deterioration of
-genetic diversity and resulted in the ned loss of individuals
-\citep{Leigh.2012}. Recent genetic analyses across southern Africa, in contrast,
-identified a genetically diverse population cluster near northern Botswana,
-suggesting moderate levels of dispersal \citep{Tensen.2022}. Nevertheless,
-anticipating how climate change and the associated changes in the biophysical
-landscape through which dispersers move will impact hWC will be paramount to
-prioritize efforts and more efficiently conserve and sustain dispersal pathways.
+Besides the directional change in climate, it is agreed that there will be an
+increased frequence in extreme climate events.
+
+Along a similar vein, the shrinking of the sea-ice has resulted in a higher
+incidence of problem polar bears in Canada over the past decades
+\citep{Towns.2009}. Similarly, it has been found that the annual range shifts of
+polar bears, in result to seasonally melting sea-ice, results in increased
+frequency of HWC.
+
+Step-selection functions (or path selection functions) provide better estimates
+of connectivity than other resource selection approaches \citep{Zeller.2016}.
+
+Information on habitat selection or connectivity are suitable for predicting
+areas with an elevated potential for human wildlife conflict
+\citep{Buchholtz.2020}.
+
+Although local rainfalls in Botswana are expected to increase in terms of
+intensity, simulations show that the length of the rainy season will decline,
+more than offsetting the incline in precipitation \cite{Akinyemi.2019}.
+
+Although dispersal is generally believed to be beneficial for the long term
+viability of species, it can also lead to the net-loss of individuals in areas
+where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.
+
+We studied a population of African wild dogs that resides in a natural
+environment with little human influence. This is only representative for a small
+share of the extant wild dog populations, as most individuals reside in areas
+that are prone to substantial edge effects. For these populations, the benefit
+of dispersal is disputed, as dispersal under high anthropogenic mortality may
+lead to a net-loss of individuals \citep{Leigh.2012}.
+
+Dispersal is an important, if not the most important, driver of landscape
+connectivity and therefore of major interest to conservation authorities. It has
+also been demonstrated that dispersal may lead to population declines in areas
+where anthropogenic mortality is high and dispersal prospects low
+\citep{Leigh.2012}.
+
+Efforts aiming at restoring or improving landscape connectivity should consider
+and take into account changes in landscape characteristics due to climate
+change.
+
+Maybe cite the \cite{Gusset.2009} paper.
+
+We assessed the implications of environmental change on the dispersal prospects,
+yet we did not consider how changing conditions alter dispersal propensity.
+
+We studied the indirect (through the landscape), but not the direct (propensity
+to disperse) conesquences of climate change on dispersal.
+
+According to our simulations, dispersers are able to cover larger distances
+during periods of low flood. This finding is little surprising, considering that
+inundated areas act as dispersal barriers and force dispersers to detour and
+circumvent water-covered areas. However, it still leads to an interesting
+hypothesis. Previous studies have shown that the euclidean dispersal distance of
+female coalitions is larger than that of male coalitions. This has led to ...
+However, demographic analyses have also revealed that female offspring tend to
+emigrate from their pack at younger ages and earlier in the year, when
+floodwaters are still at a relatively low level (Behr). It is thus conceivable
+that the sex-differences in reported dispersal distances is mainly a consequence
+of environmental conditions during dispersal, rather than owed to physiological
+differences between sexes.
+
+Unsurprisingly, \cite{Michalski.2006} show that increased proximity between
+wildlife and humans significantly increases the likelihood of human-wildlife
+conflict.
+
+\subsection{Effect of Climate Change on the Okavango Delta}
+\cite{Akinyemi.2019} predicts fewer rain days in Botswana (may want to look at
+predictions for Angola though)
+
+While a better understanding of the ODs flooding regime under global change is
+of major interest, the specific impacts of climate change on the OD are not
+fully understood. This is owed to the complex interplay between climate,
+anthropogenic water use, animal behavior, and topography of the region. Talk a
+bit about anthropogenic water use (e.g. Namibia is building a hydropower station
+in the Zambezi Region) Talk a bit about droughts, etc. Rising sea-level ->
+Intrusion of salt water Simulations by \cite{Wolski.2006} reveal that upstream
+abstractions have a smaller infuence on flooding patterns of the OD than climate
+change. In combination, however, Simulated upstream abstractions and climate
+change resulted in significant ``delta-dying'', i.e....
+
+\subsection{Species Distribution}
+Climate change and anthropogenic use will result in trophic ripple effects
+\citep{Wolski.2006}. Climate change affects species distribution through habitat
+suitability, which determines structural connectivity which ultimately
+determines functional connectivity. How climate change will affect the
+distribution of African wild dogs is not known. Given that the species has
+occured across the entire Sub-Saharan continent suggests that this species is
+highly adaptable and limited mainly by the presence of humans, yet less by the
+habitat itself. In fact, wild dogs have been found on mount Kilimandjaro.
+
+\subsection{Social Landscape}
+The OD is an important driver of species distribution and it has been found that
+an expanding flood limits available habitat, thus leading to more inter-specific
+competition, particularly between AWDs and lions (Robynne).
+Although the OD is arguably the main driver of seasonal change across the
+studied ecosystem, there are several other factors that undergoe seasonal
+change, including vegetation and the abundance and distribution of prey or
+predators.An additional complication arises when species movement is not solely driven by
+environmental conditions, but also affected by itra- and inter-specific factors.
+For instance, ... has shown that dispersers...
+Rendering such conditions alone is challenging, yet rendering the conditions
+under changing environmental conditions is merely impossible.
+As highlighted in the papers presented in the intro, climate change change
+environmental conditions, and in return species distribution, which will, may
+ultimately culminate in alterations of community composition
+\citep{Thuiller.2006}. Previous studies have shown that intra-specific factors
+are imortant determinants of dispersal movements \citep{Cozzi.2018} and it can
+be expected that inter-specific factors are similarly important
+\citep{Armansin.2019}. Through changing environmental conditions, climate change
+will also impact the distribution of various species with far-reaching
+consequences on inter- and intra-specific competition (Abrahms). Due to a lack
+of data, we omitted any inclusion of social factors in our dispersal model,
+albeit they are known to be important determinants of dispersal (Armansin).
+Studying how climate change influences species distribution and their
+interactions, and ultimately how this alters dispersal, will be challenging but
+necessary to more realistically simulate dispersal.
+
+\subsection{Seasonality in Dispersal Models}
+While our analysis marks an important step into incorporating environmental
+change into studies of connectivity, there are several critical additions that
+should be considered in the future. We studied dispersal and connectivity under
+two different environmental scenarios, yet our movement model assumed that
+dispersers had identical habitat and movement preferences in both scenarios. In
+reality, however, it can be expected that movement and habitat kernels of
+dispersers differ depending on the season considered (examples).
+Can incporporate seasonality by keeping environmental conditions constant, but
+rendering changes in species habitat preferences. Or, keep species habitat
+preferences constant and vary the environemnt. Or both. To address such
+differences, researchers could model habitat and movement preferences using
+season-dependent models, or, alternatively, by combining hidden markov models
+with step-selection functions. (cite papers that fieberg sent)
+
+Connectivity analyses that omit dynamism, be it due to seasonality or climate
+change, not only fail to capture an important aspect of reality, but also forego
+the ability to gain an understanding of connectivity under ``non-average''
+conditions, such as those expected under climate change. Reqires more data on
+dispersers -> Difficult to get -> Link to citizen science
+
+\subsection{Citizen Science}
+Validating predictions from individual-based dispersal models is challenging and
+requires additional dispersal data, which is inherently scarce anyways. Sciticen
+science may help to fill this gap by augmenting observed GPS data with
+occasional sightings of uncollared dispersing coalitions. This is especially
+ciritical for species that disperse across borders and beyond confined study
+areas. The African carnivore wildbook offers ...
 
 \subsection{Conclusion}
-Our dispersal simulations across two extreme environmental scenarios revealed
-striking differences in dispersal prospects and landscape connectivity for
-dispersing AWDs. We thereby showed that extreme environmental conditions, akin
-to those projected under climate change, will have important impacts on
-functional connectivity and may alter areas of HWC. Given the complexity of the
-studied ecosystem and its associated intricate feedback loops, predictions of
-future conditions remain challenging and plagued by uncertainty. Wildlife
-managers and conservation bodies therefore need to move beyond focusing on
-single scenarios and consider multiple possibilities to adequately respond to
-changes in the environment due to climate change, while also coping with the
-socio-economic needs of an ever-expanding human population. This will require
-the development of protection strategies that can accommodate both more extreme
-pronounced, or less intense flood. Successful conservation strategies will be of
-particular relevance for wide-ranging, endangered species that are already at
-the verge of extinction, such as the African wild dog.
+Our dispersal simulations across two extreme climatic scenarios reveal striking
+differences in dispersal prospects and landscape connectivity for dispersing
+AWDs. This implies that (1) climatic variation, be it due to seasonality or
+climate change, must be included in anlayses dealing with dispersal and (2) that
+projected climate change is likely to have profound impacts on landscape
+connectivity.
+
+Nevertheless, a more comprehensive understanding of dispersal under
+changing environmental conditions is critical for designing effective
+conservation strategies in the light of climate change. By predicting and
+modeling dispersal patterns under different scenarios, we can better anticipate
+the impacts of climate change on species and design effective conservation
+strategies to protect them. This is especially important for endangered species
+that are already at the verge of extinction.
 

### 2023-03-14 Chapter_8/04_Manuscript/Manuscript_Support.tex
various updates
upd

@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Flooding of the Okavango Delta influences
-Connectivity for Dispersing African Wild Dogs}
+\title{\textbf{Appendix}\\ African Wild Dog Dispersal and Connectivity under
+Climate Change - Lessons Learned from Seasonal Flood Extremes}
 
@@
 
-%------------------------------------------------------------------------------
-%  Appendix S1: Source-Specific Heatmaps
-%------------------------------------------------------------------------------
+\newpage
+\section{Dispersal Model}
+The model employed to simulate dispersal was based on an integrated step
+selection function (iSSF). In (integrated) step selection functions (iSSFs,
+\citealp{Fortin.2005, Avgar.2016}), observed GPS locations are converted into
+steps (the straight-line traveled between two GPS recordings
+\citep{Turchin.1998}) and compared to a set of \textit{random} steps in a
+conditional logistic regression framework \citep{Fortin.2005, Thurfjell.2014,
+Muff.2020, Fieberg.2021}. The model presented in \citep{Hofmann.2023} used
+dispersal data collected on 16 dispersing AWDs from a free-ranging wild dog
+population in northern Botswana. GPS data during dispersal was collected at
+4-hourly intervals and translated into steps of similar duration. Observed steps
+were then paired step with 24 random steps that were generated using a uniform
+distribution for turning angles (\(-\pi, +\pi\)) and step lengths from a gamma
+distribution fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\)
+= 0.37). It was then assumed that animals assigned to each observed and random
+step a selection score of the form \citep{Fortin.2005}:
+
+\begin{equation}
+\label{EQ1}
+  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
+\end{equation}
+
+Where (\(x_1, x_2, ..., x_n\)) represent the covariate values along each of the
+steps and the (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's relative
+selection strengths \citealp{Avgar.2017} towards these covariates. The benefit
+of \textit{integrated} SSFs over regular SSFs is that they provide a means to
+render two complementary ``kernels''. A \textit{movement} kernel that describes
+general movement behavior of dispersing AWDs and a \textit{habitat} kernel that
+describes preferences of AWDs with regards to environmental conditions
+\citep{Fieberg.2021}. iSSFs also allow interactions among the two kernels and
+are thus suitable to render that movement behavior may change depending on
+habitat conditions. A fitted iSSF model can be used as an individual-based
+movement model to simulate dispersal \citep{Signer.2017, Hofmann.2023}
+
+\newpage
+\section{Dispersal Model Estimates}
+\Cref{Model} depicts the model estimates from \cite{Hofmann.2023} that we used
+to simulate dispersal trajectories in this study.
+
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_MovementModel.png}
+  \caption{Model parameters from the step-selection model implemented by
+  \cite{Hofmann.2023}. The model was fit to GPS data of dispersing African wild
+  dogs and comprises of a habitat kernel (light blue band), a movement kernel
+  (dark blue band), and their interactions (purple band). Abbreviations are as
+  follows: sl = step-length, log(sl) = natural logarithm of the step-length,
+  cos(ta) = cosine of the relative turning angle.}
+  \label{Model}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Source-Specific Inter-Patch Connectivity}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_IPCMain.png}
+  \caption{Spatial representation of inter-patch connectivity derived for each
+  source-area separately across the two extreme flood-scenarios. The focal
+  source area of each subfigure is highlighted by a red circle. Subfigure (a),
+  for instance, depicts inter-patch connectivity for source-area 1.}
+  \label{IPCMain}
+  \end{center}
+\end{figure}
+
+\newpage
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_IPCBuffer.png}
+  \caption{Spatial representation of emigration patterns derived for each
+  source-area separately across the two extreme flood-scenarios. The focal
+  source area of each subfigure is highlighted by a red circle. Subfigure (a),
+  for instance, depicts the number of emigrating individuals from source-area
+  1.}
+  \label{IPCBuffer}
+  \end{center}
+\end{figure}
+
 \newpage
@@
   \includegraphics[width = \textwidth]{99_HeatmapsIndividual.png}
-  \caption{}
+  \caption{Heatmaps prepared for each source-area separately across the two
+  extreme flood-scenarios. The focal source area of each subfigure is
+  highlighted by a red circle. Subfigure (a), for instance, depicts the heatmaps
+  for source-area 1.}
   \label{Heatmaps}
@@
 
-%------------------------------------------------------------------------------
-%  Appendix S2: Source-Specific Betweenness
-%------------------------------------------------------------------------------
 \newpage
@@
   \includegraphics[width = \textwidth]{99_BetweennessIndividual.png}
-  \caption{}
+  \caption{Betweenness maps prepared for each source-area separately across the
+  two extreme flood-scenarios. The focal source area of each subfigure is
+  highlighted by a red circle. Subfigure (a), for instance, depicts the
+  betweenness maps for source-area 1.}
   \label{Betweenness}
@@
 
-%------------------------------------------------------------------------------
-%  Appendix S3: Source-Specific Betweenness
-%------------------------------------------------------------------------------
 \newpage
-\section{Source-Specific Inter-Patch Connectivity}
+\section{Source-Specific Human Wildlife Conflict}
 \begin{figure}[htbp]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_IPCMain.png}
-  \caption{}
-  \label{IPCMain}
-  \end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_IPCBuffer.png}
-  \caption{}
-  \label{IPCBuffer}
+  \includegraphics[width = \textwidth]{99_HumanWildlifeConflictIndividual.png}
+  \caption{Human wildlife conflict maps prepared for each source-area separately
+  across the two extreme flood-scenarios. The focal source area of each
+  subfigure is highlighted by a red circle. Subfigure (a), for instance, depicts
+  the human wildlife conflict maps for source-area 1.}
+  \label{HWC}
   \end{center}

### 2023-03-14 Chapter_8/04_Manuscript/Manuscript_Support_2023-01-19.tex
various updates

--- 2022-11-03 Chapter_8/04_Manuscript/Manuscript_Support.tex
@@
 
+\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
+Consequences for African Wild Dog Connectivity under Climate Change
+
 \vspace{0.5cm}
@@
 
-%------------------------------------------------------------------------------
-%  Appendix S1: Source-Specific Heatmaps
-%------------------------------------------------------------------------------
+\newpage
+\section{Dispersal Model}
+The model employed to simulate dispersal was based on an integrated step
+selection function (iSSF). In (integrated) step selection functions (iSSFs,
+\citealp{Fortin.2005, Avgar.2016}), observed GPS locations are converted into
+steps (the straight-line traveled between two GPS recordings
+\citep{Turchin.1998}) and compared to a set of \textit{random} steps in a
+conditional logistic regression framework \citep{Fortin.2005, Thurfjell.2014,
+Muff.2020, Fieberg.2021}. The model presented in \citep{Hofmann.2023} used
+dispersal data collected on 16 dispersing AWDs from a free-ranging wild dog
+population in northern Botswana. GPS data during dispersal was collected at
+4-hourly intervals and translated into steps of similar duration. Observed steps
+were then paired step with 24 random steps that were generated using a uniform
+distribution for turning angles (\(-\pi, +\pi\)) and step lengths from a gamma
+distribution fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\)
+= 0.37). It was then assumed that animals assigned to each observed and random
+step a selection score of the form \citep{Fortin.2005}:
+
+\begin{equation}
+\label{EQ1}
+  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
+\end{equation}
+
+Where (\(x_1, x_2, ..., x_n\)) represent the covariate values along each of the
+steps and the (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's relative
+selection strengths \citealp{Avgar.2017} towards these covariates. The benefit
+of \textit{integrated} SSFs over regular SSFs is that they provide a means to
+render two complementary ``kernels''. A \textit{movement} kernel that describes
+general movement behavior of dispersing AWDs and a \textit{habitat} kernel that
+describes preferences of AWDs with regards to environmental conditions
+\citep{Fieberg.2021}. iSSFs also allow interactions among the two kernels and
+are thus suitable to render that movement behavior may change depending on
+habitat conditions. A fitted iSSF model can be used as an individual-based
+movement model to simulate dispersal \citep{Signer.2017, Hofmann.2023}
+
 \newpage
@@
 
-%------------------------------------------------------------------------------
-%  Appendix S3: Source-Specific Betweenness
-%------------------------------------------------------------------------------
+\newpage
+\section{Source-Specific Human Wildlife Conflict}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_HumanWildlifeConflictIndividual.png}
+  \caption{}
+  \label{HWC}
+  \end{center}
+\end{figure}
+
 \newpage
@@
   \includegraphics[width = \textwidth]{99_IPCMain.png}
-  \caption{}
+  \caption{Spatial representation of the tables 1a and 1b from the main
+  manuscript.}
   \label{IPCMain}
@@
 
+\newpage
 \begin{figure}[htbp]
@@
   \label{IPCBuffer}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Emigration}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = 0.8\textwidth]{99_Emigration.png}
+  \caption{Absolute number of simulated AWDs moving into each of the designated
+  emigration zones (purple), in the minimum and maximum flood scenarios.}
+  \label{Emigration}
   \end{center}

### 2023-03-14 Chapter_10/02_R-Scripts/01_Analysis/01_Setup.R

@@ -36,6 +36,8 @@ new_chap <- "/home/david/ownCloud/University/15. PhD/Chapter_10"
 dir.create("03_Data/01_RawData", showWarnings = F)
 dir.create("03_Data/02_CleanData", showWarnings = F)
 dir.create("03_Data/03_Results", showWarnings = F)
+dir.create("04_Manuscript", showWarnings = F)
+dir.create("05_Presentation", showWarnings = F)
 
 ################################################################################
 #### Copy Data from Chapter 1

### 2023-03-14 Chapter_2/02_R-Scripts/01_DataCleaning/04_Floodmaps.R

@@ -21,8 +21,9 @@ library(floodmapr)    # For floodmapping
 library(pbmcapply)    # For multicore use
 
 # Specify username and password to Earth Data account
-username <- "DoDx9"
-password <- "EarthData99"
+load("/home/david/ownCloud/Dokumente/Bibliothek/Wissen/R-Scripts/EarthDataLogin.rds")
+# username <- "username"
+# password <- "password"
 
 ################################################################################
 #### Reclassify and Resample ORI Floodmaps
@@ -119,7 +120,7 @@ pot_dates %>% mutate(Month = month(Date)) %>% count(Month, Floodmap) %>% subset(
 
 # Load dispersal data and identify first and last date. We can use those to
 # determine for timerange we should try to download additional floodmaps
-dis_dates <- "03_Data/02_CleanData/00_General_Dispersers.csv" %>%
+dis_dates <- "03_Data/02_CleanData/Dispersers.csv" %>%
   read_csv(show_col_types = F) %>%
   subset(State == "Disperser") %>%
   pull(Timestamp) %>%
@@ -337,7 +338,7 @@ newnames <- paste0("03_Data/02_CleanData/00_Floodmaps/02_Resampled/", basename(f
 
 # Resample floodmaps
 dir.create("03_Data/02_CleanData/00_Floodmaps/02_Resampled", showWarnings = F)
-r <- rast("03_Data/02_CleanData/00_General_Raster.tif")
+r <- rast("03_Data/02_CleanData/Raster.tif")
 if (length(files) > 0) {
   cat("Resampling floodmaps...\n")
   invisible(pbmclapply(1:length(files), ignore.interactive = T, mc.cores = 1, function(x) {

### 2023-03-14 Chapter_2/02_R-Scripts/02_Analysis/00_SSF.R

@@ -28,9 +28,9 @@ cores <- detectCores() - 1
 ################################################################################
 #### Computing Step Metrics
 ################################################################################
-# Load the gps data of dispersers and create rounded timestamps. Then nest the
+# Load the gps data of dispersers and create rounded timestamps. Then, nest the
 # data by individual
-data <- "03_Data/02_CleanData/00_General_Dispersers.csv" %>%
+data <- "03_Data/02_CleanData/Dispersers.csv" %>%
   read_csv(show_col_types = F) %>%
   subset(State == "Disperser") %>%
   mutate(TimestampRounded = round_date(Timestamp, "1 hour")) %>%
@@ -48,7 +48,13 @@ print(data, n = 30)
 cat("Computing bursts from GPS data...\n")
 data$Bursts <- pbmclapply(data$GPS, ignore.interactive = T, mc.cores = cores, function(x) {
   res <- resampleFixes(x, hours = 2, start = 1, tol = 0.5)
+  if (length(res) == 1) {
+    return(NA)
+  }
   sub <- subset(res, hour(TimestampRounded) %in% c(3, 7, 15, 19, 23))
+  if (length(sub) == 1) {
+    return(NA)
+  }
   bur <- computeBursts(sub)
   bur[, c("x", "y")] <- reprojectCoords(cbind(bur$x, bur$y)
     , from = "+init=epsg:4326"
@@ -73,8 +79,8 @@ steps <- data %>%
   mutate(Nrow = map_dbl(data, nrow)) %>%
   subset(Nrow > 3) %>%
   mutate(Steps = map(data, stepMetrics)) %>%
-  dplyr::select(ID, burst_id, Steps) %>%
-  mutate(burst_id = 1:n()) %>%
+  dplyr::select(ID, BurstID = burst_id, Steps) %>%
+  mutate(BurstID = 1:n()) %>%
   unnest(Steps)
 
 # Remove steps where the relative turning angle is NA
@@ -108,7 +114,7 @@ steps %>%
   plot(type = "l")
 
 # Visualize the steps
-ggplot(steps, aes(x = x, y = y, col = as.factor(burst_id), group = burst_id)) +
+ggplot(steps, aes(x = x, y = y, col = as.factor(BurstID), group = BurstID)) +
   geom_path(size = 0.1) +
   geom_point(size = 0.5) +
   coord_sf() +
[…]

### 2023-03-14 Chapter_2/02_R-Scripts/03_Visualization/02_VisualizationsForPresentation/CovariateAnimations.R

@@ -1,0 +1,361 @@
+################################################################################
+#### Animations of Dynamic Covariates
+################################################################################
+# Description: Preparation of animations of all dynamic covariates
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory.
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Load required packages
+library(terra)               # To handle rasters
+library(lubridate)           # To handle dates
+library(ggspatial)           # To add scale bar and north arrow to plots
+library(tidyverse)           # For data wrangling
+library(sf)                  # For plotting
+library(animation)           # To store animation
+library(pbmcapply)           # For multicore abilities with progress bar
+library(magick)              # To load images
+library(scales)              # To handle out of bounds (oob) values
+library(ggpubr)              # To grab legends
+
+# Load data to plot in the background
+r      <- rast("03_Data/02_CleanData/00_General_Raster.tif")
+africa <- vect("03_Data/02_CleanData/00_General_Africa.shp") %>% crop(r)
+
+# Coerce data to formats that are plottable with ggplot
+r <- as.data.frame(r, xy = T)
+africa <- st_as_sf(africa)
+
+################################################################################
+#### Precipitation, Temperature, NDVI
+################################################################################
+# Loop through the different covariates and prepare the plots
+covs <- c("Rainmaps", "Tempmaps", "NDVI")
+for (i in covs) {
+  # i <- covs[1]
+  filename <- paste0("05_Presentation/", i, ".gif")
+  if (file.exists(filename)) {
+    next
+  }
+
+  # Load the data
+  dat <- dir(
+      path    = paste0("03_Data/02_CleanData/00_", i)
+    , pattern = ".grd$", full.names = T
+  ) %>% rast()
+
+  # Retrieve dates from layernames
+  dates <- names(dat) %>%
+    substr(start = 1, stop = 10) %>%
+    ymd()
+
+  # Subset to a selected date range
+  if (i == "NDVI") {
+    subdat <- dat
+  } else {
+    subdat <- dat[[dates > "2019-10-24" & dates < "2019-11-10"]]
+  }
+
+  # Specify colors
[…]

### 2023-03-14 Chapter_8/02_R-Scripts/01_Analysis/01_Setup.R

@@ -90,6 +90,11 @@ writeVector(riv, file.path(new_chap, "03_Data/02_CleanData/MajorRivers.shp"), ov
 fau <- vect(file.path(old_chap, "03_Data/01_RawData/DAVID/Faults.shp"))
 writeVector(fau, file.path(new_chap, "03_Data/02_CleanData/Faults.shp"), overwrite = T)
 
+# Copy areas of interest
+aoi <- vect(file.path(old_chap, "03_Data/01_RawData/DAVID/AreasOfInterest.shp"))
+aoi <- smooth(aoi)
+writeVector(aoi, file.path(new_chap, "03_Data/02_CleanData/AreasOfInterest.shp"), overwrite = T)
+
 # Copy protected areas
 pro <- vect(file.path(old_chap, "03_Data/02_CleanData/02_LandUse_Protected_PEACEPARKS.shp"))
 pro <- crop(pro, ext)
@@ -135,6 +140,26 @@ dis <- crop(dis, ext)
 plot(dis)
 names(dis) <- "DistanceToHumans"
 writeRaster(dis, file.path(new_chap, "03_Data/02_CleanData/DistanceToHumans.tif"), overwrite = T)
+
+# Copy the raw "Human Density" layer
+hum <- rast(file.path(old_chap, "03_Data/02_CleanData/04_AnthropogenicFeatures_HumanInfluence_FACEBOOK.tif"))
+hum <- crop(hum, ext * 1.4)
+names(hum) <- "HumanDensity"
+writeRaster(hum, file.path(new_chap, "03_Data/02_CleanData/HumanDensity.tif"), overwrite = T)
+
+# Copy the raw "Agriculture" layer(s)
+cro1 <- rast(file.path(old_chap, "03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_CROPLANDS.tif"))
+cro2 <- rast(file.path(old_chap, "03_Data/02_CleanData/04_AnthropogenicFeatures_Agriculture_GLOBELAND.tif"))
+cro <- max(c(cro1, cro2))
+cro <- crop(cro, ext * 1.4)
+names(cro) <- "Agriculture"
+writeRaster(cro, file.path(new_chap, "03_Data/02_CleanData/Agriculture.tif"), overwrite = T)
+
+# Copy information on cattle
+cat <- vect(file.path(old_chap, "03_Data/01_RawData/GABRIELE/Cattle.shp"))
[…]

### 2023-03-14 Chapter_8/02_R-Scripts/01_Analysis/09_Distance.R

@@ -10,11 +10,12 @@ rm(list = ls())
 setwd("/home/david/ownCloud/University/15. PhD/Chapter_8")
 
 # Load required packages
-library(terra)     # To handle spatial data
-library(raster)    # To handle spatial data
-library(tidyverse) # To wrangle data
-library(maptools)  # To compute point density
-library(spatstat)  # To compute point density
+library(terra)      # To handle spatial data
+library(raster)     # To handle spatial data
+library(tidyverse)  # To wrangle data
+library(maptools)   # To compute point density
+library(spatstat)   # To compute point density
+library(pbmcapply) # For multicore progress bar
 
 # Load custom functions
 source("02_R-Scripts/00_Functions.R")
@@ -28,6 +29,9 @@ sims <- read_rds("03_Data/03_Results/DispersalSimulation.rds")
 # Keep only desired columns
 sims <- sims[, c("x", "y", "sl_", "ta_", "TrackID", "StepNumber", "SourceArea", "FloodLevel")]
 
+# For later, extract the unique combinations of Track_IDs and FloodLevels
+allid <- sims %>% select(TrackID, FloodLevel, SourceArea) %>% distinct()
+
 ################################################################################
 #### Distance to Humans
 ################################################################################
@@ -37,10 +41,16 @@ dir.create("03_Data/03_Results/99_Distances", showWarnings = F)
 # Calculate the distance to humans for each step
 sims$DistanceToHumans <- terra::extract(humans, cbind(sims$x, sims$y))[, 1]
 
+# We only care about locations that are closer than 500 meters to humans (this
+# will strictly reduce the number of datapoints we have to keep track off, but
+# also implies that we have to keep track of the original TrackIDs (allid) when
+# we bootstrap.
+sims <- subset(sims, DistanceToHumans < 500)
+
 # Design through which to loop
 design <- expand_grid(
-    FloodLevel = unique(sims$FloodLevel)
-  , SourceArea = unique(sims$SourceArea)
+    FloodLevel = unique(allid$FloodLevel)
+  , SourceArea = unique(allid$SourceArea)
 )
[…]

### 2023-03-14 Chapter_8/02_R-Scripts/02_Visualization/Manuscript/Distance.R

@@ -1,0 +1,59 @@
+################################################################################
+#### Distance Plot
+################################################################################
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)
+library(colorspace)
+
+# Reload the metrics
+metrics <- read_rds("03_Data/03_Results/DistanceAOI.rds")
+metrics <- mutate(metrics, LWR = Number - SE, UPR = Number + SE)
+metrics <- mutate(metrics, FloodLevel = paste0(FloodLevel, "-Flood"))
+
+# Visualize
+cols <- c("orange", "cornflowerblue")
+p <- ggplot(metrics, aes(x = FloodLevel, y = Number, ymin = LWR, ymax = UPR, fill = FloodLevel)) +
+  geom_col(color = "white", alpha = 0.8) +
+  geom_errorbar(width = 0.2, aes(color = FloodLevel)) +
+  facet_grid(~ Name) +
+  scale_color_manual(name = "Flood-Level", values = lighten(cols, 0.5)) +
+  scale_fill_manual(
+      values = cols
+    , guide  = guide_legend(
+      , title          = "Flood-Level"
+      , show.limits    = T
+      , title.position = "bottom"
+      , title.hjust    = 0.5
+      , label.position = "bottom"
+      , ticks          = F
+      , nrow           = 1
+    )
+  ) +
+  theme(
+      legend.position   = "none"
+    , legend.key.height = unit(0.2, 'cm')
+    , legend.key.width  = unit(1, 'cm')
+    , panel.grid.major  = element_line(color = "gray95")
+    , panel.grid.minor  = element_blank()
+    , axis.text         = element_text(size  = 6)
+    , axis.title        = element_text(size  = 9)
+    , panel.background  = element_blank()
+  ) +
+  xlab("") +
+  ylab("Density")
+
+# Store it
+ggsave("04_Manuscript/99_HWCDifferenceAOI.png"
+  , bg     = "white"
+  , width  = 4
+  , height = 3
+  , scale  = 1.2
+  , plot   = p
+)

### 2023-03-14 Chapter_8/02_R-Scripts/02_Visualization/Manuscript/Floodmaps.R

@@ -1,0 +1,189 @@
+################################################################################
+#### Dynamic Floodmaps
+################################################################################
+# Description: Plot of the two floodmaps used
+
+# Clear R's brain
+rm(list = ls())
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_8"
+setwd(wd)
+
+# Load required packages
+library(sf)           # To handle spatial data
+library(tidyverse)    # For data.wrangling
+library(terra)        # To handle spatial data
+library(ggplot2)      # For nice plots
+library(ggspatial)    # To add scale bars and north arrows
+library(ggpubr)       # To put plots together
+
+# Load custom functions
+source("02_R-Scripts/00_Functions.R")
+
+# Load stuff that we would like to plot
+africa <- read_sf("03_Data/02_CleanData/Africa.shp")
+water  <- rast("03_Data/02_CleanData/WaterCover.tif")
+roads  <- read_sf("03_Data/02_CleanData/Roads.shp")
+vills  <- read_sf("03_Data/02_CleanData/Villages.shp")
+vills  <- cbind(st_drop_geometry(vills), st_coordinates(vills)) %>%
+  rename(x = X, y = Y)
+
+# Reference raster
+r <- rast("03_Data/02_CleanData/ReferenceRaster.tif")
+r <- as.data.frame(r, xy = T)
+
+# We only care about min/max extents
+water  <- water[[c("min", "max")]]
+names(water) <- c("Min", "Max")
+
+# To compute the flood extent, we need to focus on the Okavango delta, thus,
+# let's load a shapefile of it
+oka <- vect("03_Data/02_CleanData/MajorWaters.shp")
+oka <- oka[oka$name == "Okavango Delta", ]
+
+# Buffer it slightly
+oka <- buffer(oka, width = 10000)
+oka <- fillHoles(oka)
+
+# This looks good. We now mask anything outside that polygon
+water_masked <- mask(water, oka)
+water_masked <- trim(water_masked)
+water_masked <- subst(water_masked, 0, NA)
+
+# Convert shape of okavango delta to sf
+crs(oka) <- "+init=epsg:4236"
+oka            <- st_as_sf(oka)
+oka            <- rbind(oka, oka)
[…]

### 2024-07-01 Chapter_1/04_Manuscript/Manuscript.tex
Publish PHD

@@
 \noindent The probability of a step \(i\) being realized was then contingent on
-the step's selection score, as well as on the selection scores of all other step
-in the same stratum:
+the step's selection score, as well as on the selection scores of all other
+steps in the same stratum:
 

### 2024-07-01 Chapter_1/04_Manuscript/Manuscript_Support.tex
upd
Publish PHD

@@
 \textsuperscript{1} Department of Evolutionary Biology and Environmental
- Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
+Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
 Switzerland.
 
- \textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
- Maun, Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+Private Bag 13, Maun, Botswana.
 

### 2024-07-01 Chapter_2/04_Manuscript/Manuscript.tex
Publish PHD

--- 2021-06-24 Chapter_2/04_Manuscript/Manuscript.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[round]{natbib}
 \usepackage{amsmath}
@@
 \usepackage{gensymb}
+\usepackage{textcomp}
 \usepackage{graphicx}
@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{hyperref}
 \usepackage[nameinlink]{cleveref}
-\usepackage{lineno}
 \usepackage{multirow}
@@
 \usepackage[nomarkers, nolists]{endfloat}
+\usepackage{scalerel}
+\usepackage{makecell}
+\usepackage{ifthen}
+\usepackage{tikz}
+\usepackage{lineno}
+\usetikzlibrary{svg.path}
+
+%------------------------------------------------------------------------------
+%  Bibtex vs Biblatex
+%------------------------------------------------------------------------------
+% While bibtex is very nice to use, it's insanely slow. I therefore work with
+% bibtex until the final stage. Set the value below to "true" or "false" to pick
+% which one should be used.
+\newboolean{usebiblatex}
+\setboolean{usebiblatex}{true}
+
+% To remove figures from the document, only keeping their captions
+\AtBeginDelayedFloats{%
+  \renewcommand{\includegraphics}[2][]{}% Remove \includegraphics functionality
+  \renewcommand{\resizebox}[3]{}% Remove \resizebox functionality
+}
+\setcounter{totalnumber}{10}% Allow up to 10 floats per page
+\setcounter{topnumber}{10}% Allow up to 10 floats per page
+\setcounter{bottomnumber}{10}% Allow up to 10 floats per page
+\renewcommand{\efloatseparator}{\mbox{}}% From the documentation
+
+% Depending on the question, the commands below will prepare the document
+% accordingly
+\ifthenelse{\boolean{usebiblatex}}{
+  %-----------------------------------------------------------------------------
+  %  If using biblatex
+  %-----------------------------------------------------------------------------
+  % Load required packages and refer to correct bibliography
+  \usepackage{csquotes}
+  \usepackage[
+      backend=biber
+    , style=apa          % To display Author-year in the text
+    , natbib=true        % Necessary to use citep, etc.
+    , hyperref=true      % To have clickable links in the references
+    , useeditor=false    % Don't add editor to publications
+    , sorting=ynt        % Sort by year, name, and title
+    , uniquename=false   % Avoid that initials are put
+  ]{biblatex}
+  \addbibresource{Literature.bib}
+
+  % To include author in link
+  \DeclareFieldFormat{citehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{#1}}
+  \DeclareFieldFormat{textcitehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{%
+    #1%
+    \ifbool{cbx:parens}
+      {\bibcloseparen\global\boolfalse{cbx:parens}}
+      {}}}
+  \savebibmacro{cite}
+  \savebibmacro{textcite}
+  \renewbibmacro*{cite}{%
+    \printtext[citehyperref]{%
+      \restorebibmacro{cite}%
+      \usebibmacro{cite}}}
+  \renewbibmacro*{textcite}{%
+    \ifboolexpr{
+      ( not test {\iffieldundef{prenote}} and
+        test {\ifnumequal{\value{citecount}}{1}} )
+      or
+      ( not test {\iffieldundef{postnote}} and
+        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
+    }
+      {\DeclareFieldAlias{textcitehyperref}{noformat}}
+      {}%
+    \printtext[textcitehyperref]{%
+      \restorebibmacro{textcite}%
+      \usebibmacro{textcite}}}
+
+  % Remove editor form everything but books
+  \AtEveryBibitem{%
+   \ifentrytype{book}{}{%
+    \clearname{editor}
+   }
+  }
+}{
+  %-----------------------------------------------------------------------------
+  %  If using bibtex
+  %-----------------------------------------------------------------------------
+  % Load required packages and select citation style
+  \usepackage[round]{natbib}
+  \bibliographystyle{apalike}
+}
 
@@
 %	Some Styling
+%------------------------------------------------------------------------------
+% Link colors
+% \definecolor{linkcolor}{HTML}{000000}
+% \hypersetup{
+%   colorlinks = true,
+%   linkcolor  = linkcolor,
+%   urlcolor   = linkcolor,
+%   citecolor  = linkcolor,
+% }
+
+% Changing the style of captions in figures etc.
+\captionsetup{labelfont=bf, format=plain, font=small}
+
+% Change how equations are referenced
+\renewcommand{\theequation}{Equation \arabic{equation}}%
+
+% Avoid skip when including text from external .tex file
+\newcommand{\inputy}[1]{\input{#1}\unskip}
+
+% For supplementary material
+\newcommand{\beginappendix}{%
+  \setcounter{table}{0}
+  \renewcommand{\thetable}{S\arabic{table}}%
+  \setcounter{figure}{0}
+  \renewcommand{\thefigure}{S\arabic{figure}}%
+  \setcounter{equation}{0}
+  \renewcommand{\theequation}{Equation S\arabic{equation}}%
+  \setcounter{section}{0}
+  \renewcommand{\thesection}{A.\arabic{section}}%
+}
+
+%------------------------------------------------------------------------------
+%  ORCID
 %------------------------------------------------------------------------------
@@
 
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
+% Create the ORCID logo
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{
+  orcidlogo/.pic={
+    \fill[orcidlogocol]
+      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
+        ,0,128,0C198.7,0,256,57.3,256,128z};
+    \fill[white]
+      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
+      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
+        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
+        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
+      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+  }
+}
 
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
+% Command to create the ORCID
+\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
+\begin{tikzpicture}[yscale=-1,transform shape]
+  \pic{orcidlogo};
+\end{tikzpicture}
+}{|}}}}
 
@@
 %------------------------------------------------------------------------------
-\title{Dispersal Through Seasonal Landscapes}
-
-% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
-% Connectivity}
+% \title{Flooding of the Okavango Delta influences Connectivity for Dispersing
+% African Wild Dogs}
+\title{Dispersal and Connectivity in Increasingly Extreme Climatic Conditions}
 
@@
 \author{
-  David D. Hofmann\textsuperscript{1,\S} \and
+  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
+  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
   John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \and
-  Gabriele Cozzi\textsuperscript{1,2} \and
-  Dominik M. Behr\textsuperscript{1,2}
+  Arpat Ozgul\textsuperscript{1,2} \orcid{0000-0001-7477-2642} \and
+  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
 }
@@
 \pagenumbering{gobble}
-
-% Required to be able to cite
-\bibliographystyle{apalike}
 
@@
 \textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
+Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
 Switzerland.
 
-\textsuperscript{2} Botswana Predator Conservation, Private Bag 13, Maun,
-Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+Private Bag 13, Maun, Botswana.
 
-\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)
+\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
 
@@
 
-\textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
-Assess Landscape Connectivity
+\textbf{Running Title:} Dispersal in a Changing Climate
 
@@
 
-\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
-Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
-pictus
+\textbf{Keywords:} climate change, conservation, dispersal, flood extremes,
+individual-based simulations, landscape connectivity, movement ecology,
+step-selection functions
 
@@
 \begin{abstract}
-This is the abstract.
+While climate change has been shown to impact several life-history traits of
+wild-living animal populations, little is known about its effects on dispersal
+and connectivity.
+
+Here, we capitalize on the highly variable flooding regime of the Okavango Delta
+to investigate impacts of changing environmental conditions on dispersal and
+connectivity of the endangered African wild dog (\textit{Lycaon pictus}). Based
+on remote sensed flood extents observed over 20 years, we derive two extreme
+flood scenarios: a minimum and a maximum flood extent; representative of very
+dry and very wet environmental periods. These conditions are akin to those
+anticipated under increased climatic variability, as it is expected under
+climate change. Using a movement model parametrized with GPS data from
+dispersing individuals, we simulate 12,000 individual dispersal trajectories
+across the ecosystem under both scenarios and investigate patterns of
+connectivity.
+
+Across the entire ecosystem, surface water coverage during maximum flood extent
+reduces dispersal success (i.e., the propensity of individuals to disperse
+between adjacent subpopulations) by 12\% and increases dispersal durations by
+17\%. Locally, however, dispersal success diminishes by as much as 78\%.
+Depending on the flood extent, alternative dispersal corridors emerge, some of
+which in the immediate vicinity of human-dominated landscapes. Notably, under
+maximum flood extent, the number of dispersing trajectories moving into
+human-dominated landscapes decreases by 41\% at the Okavango Delta's inflow, but
+increases by 126\% at the Delta's distal end. This may drive the amplification
+of human-wildlife conflict.
+
+Whilst predicting the impacts of climate change on environmental conditions
+on-the-ground remains challenging, our results highlight that environmental
+change may have significant consequences for dispersal patterns and
+connectivity, and ultimately, population viability. Acknowledging and
+anticipating such impacts will be key to effective conservation strategies and
+to preserve vital dispersal corridors in light of climate change and other
+human-related landscape alterations.
+
 \end{abstract}
@@
 
-\onehalfspacing
-\tableofcontents
-\doublespacing
+% \onehalfspacing
+% \tableofcontents
+% \doublespacing
 
@@
 
-% Taken from Zeller.2020
-Connectivity for wildlife is typically modeled in static environments and
-estimated for a single snapshot in time (e.g., [6,7]). However, wildlife
-responses to landscapes are inherently dynamic, as are the landscapes themselves
-(e.g., [8])
+Climate change is expected to impact ecosystems across the globe with
+far-reaching consequences for the species living therein \citep{Paniw.2019,
+Radchuk.2019, IPCC.2022, Ozgul.2023}. By altering environmental conditions,
+climate change has been shown to affect animal behavior \citep{Fuller.2016},
+resource availability \citep{Durant.2007}, demographic rates of local
+populations \citep{Paniw.2021}, range distribution of wild-living animals
+\citep{Thomas.2004, Thuiller.2006}, and the potential for human-wildlife
+conflict (hereafter referred to as HWC; \citealp{Abrahms.2023}), thus
+threatening the viability of most wildlife species. An important life-history
+trait known to influence the persistence of wild living populations across large
+spatial scales is dispersal \citep{Hanski.1999, Bowler.2005, Kokko.2006}, which
+is defined as the movement of individuals away from their natal location to the
+site of first reproduction \citep{Clobert.2012}. Despite the crucial role of
+dispersal in maintaining population viability, little is known about the impacts
+of climate change on long-distance dispersal and landscape connectivity.
+Nevertheless, assessing the consequences of climate change for dispersal is
+essential to anticipate potential shifts in range distributions and population
+dynamics.
 
-Modeling connectivity in a static environment may mask the importance of connectivity for
-maintaining wildlife population dynamics and may result in identifying corridors that are only
-periodically functional [15,16,18]
+Predicting how dispersal and connectivity respond to environmental change
+remains challenging \citep{Littlefield.2019}. This is mainly due to insufficient
+data on dispersing animals at the appropriate spatial and temporal scales
+\citep{Graves.2014, Vasudev.2015}, coupled with limited insights into potential
+differences in environmental conditions under climate change
+\citep{Scheiter.2009, IPCC.2022}. To address these shortcomings, one approach
+has been to combine climate projections and species distribution models with the
+aim of predicting future species ranges and investigating the impacts of such
+range shifts on structural connectivity \citep{Wasserman.2012, Ashrafzadeh.2019,
+Luo.2021}. However, this approach fails to translate predicted atmospheric
+conditions into ground-level landscape characteristics (e.g., vegetation cover,
+surface-water). An alternative approach, albeit not primarily focusing on
+climate change, is to investigate how seasonality affects functional
+connectivity (e.g., \citealp{Mui.2017, Osipova.2019, Zeller.2020, Kaszta.2021}).
+This is usually achieved using seasonally updated resistance surfaces that
+depict the ease or difficulty at which the focal species can traverse a
+particular environment in a specific season \citep{Zeller.2012}. Despite its
+biological relevance for understanding seasonal variability, this approach
+suffers from the short time-span at which processes are investigated and hinders
+drawing inferences on the long-term effects of climate change on dispersal and
+connectivity. Recently, alternative approaches for investigating landscape
+connectivity that combine empirical GPS movement data with a modeling framework
+to reconstruct and predict movement trajectories at the level of the individual
+have been suggested \citep{Signer.2017, Hofmann.2023, Signer.2023}. Such
+predicted trajectories are uncoupled from any temporal constraint and therefore
+suitable to investigate the effects of changing landscapes, as observed or
+predicted under the influence of climate change, on dispersal and connectivity.
+Given that climate change will shift systems towards conditions that currently
+represent extremes \citep{Stott.2016, Ummenhofer.2017, IPCC.2022}, we argue that
+a focus on extreme, rather than seasonal, variations could serve as a robust way
+to learn about the impacts of climate change on patterns of dispersal and
+connectivity. Such an approach appears particularly useful in cases where data
+on future environmental conditions are difficult to obtain or plagued by
+uncertainty \citep{Collins.2012}.
 
-\section{Methods}
+One ecosystem that offers a unique opportunity to study the impacts of extreme
+environmental conditions on dispersal and connectivity in a large-scale natural
+experiment is the Okavango Delta in Botswana. The Okavango Delta (henceforth
+``Delta'') is the world's largest inland delta and characterized by striking
+variability in its flood extent within and between years \citep{Gumbricht.2004,
+Wolski.2017}. The area covered by the Delta's floodwaters can fluctuate between
+3,500 km\textsuperscript{2} during particularly dry and 14,000
+km\textsuperscript{2} during particularly wet periods \citep{McCarthy.2003,
+Gumbricht.2004}. The region is among the most vulnerable to climate change, as
+temperature increases of 4 to 6\degree C above pre-industrial levels are
+expected within the 21\textsuperscript{st} century \citep{Engelbrecht.2015,
+Akinyemi.2019}; a prediction that goes far beyond the global average
+\citep{IPCC.2022}. The rise in ambient temperature is likely to elevate
+evapotranspiration rates across the Delta's alluvial fan, yet it is unclear
+whether water-losses to the atmosphere will be offset or exceeded by an increase
+in precipitation across the Delta's catchment areas in Angola
+\citep{Wolski.2008, IPCC.2022}. Despite the importance of the Delta as a driver
+of ecosystem functioning \citep{Wolski.2008}, species distribution
+\citep{Bonyongo.2005, Bennitt.2014}, and dispersal corridors
+\citep{Hofmann.2021, Hofmann.2023}, forecasting its flooding regime under
+climate change has proven notoriously difficult. Predictions of the flood extent
+under climate change range from ``much drier'' to ``much wetter'' depending on
+the employed model and climate scenario \citep{Murray-Hudson.2006, Wolski.2008}.
+
+Within this ecosystem a keystone predator and an umbrella species for
+conservation efforts is the African wild dog (\textit{Lycaon pictus}), a large
+carnivore that is characterized by a remarkable dispersal ability
+\citep{McNutt.1996, Davies-Mostert.2012, Masenga.2016, Cozzi.2020,
+Sandoval-Seres.2022}. Under favorable conditions, young individuals that leave
+their natal pack can cover up to several hundred kilometers within a few days,
+crossing a vast diversity of habitat-conditions (e.g., \citealp{Cozzi.2020}).
+While historically observed Euclidean dispersal distances are limited to 5-500
+km \citep{Davies-Mostert.2012, Cozzi.2020, Sandoval-Seres.2022}, cumulative
+dispersal distances of over 5,000 km have been recorded \citep{Masenga.2016}.
+Previous research on dispersing individuals revealed that the Delta's
+floodwater represents a major barrier to dispersal \citep{Hofmann.2021,
+Hofmann.2023} and it can be hypothesized that wild dog dispersal and
+connectivity will be profoundly influenced by future changes in the flood
+regime, as expected under the influence of climate change.
+
+Utilizing a previously validated individual-based movement model parametrized
+with empirical GPS data collected on dispersing wild dogs, we simulated 12,000
+individual dispersal trajectories under two extreme environmental scenarios: one
+assuming a maximum flood extent, representing above-average wet climatic
+conditions, and one assuming a minimum flood extent, representing acute dry
+conditions. Both scenarios reflect possible outcomes under the effect of climate
+change. For each scenario, we assessed dispersal success and dispersal durations
+of simulated trajectories between distinct regions, mapped the emergence of
+alternative dispersal corridors and bottlenecks, and investigated how differing
+corridor arrangements influenced the potential for HWC (\Cref{Concept}). We
+anticipated lower dispersal success and connectivity, as well as an increased
+propensity to disperse outside the main study area during maximum flood.
+Furthermore, we anticipated major dispersal corridors to differ between minimum
+and maximum flood, thus resulting in alternative hotspots for HWC. Ultimately,
+our goal was to highlight that altered climatic conditions and associated
+changes in landscape characteristics can substantially affect the spatial
+arrangement of movement corridors and, subsequently, the potential for HWC.
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = \textwidth]{Figures/GraphicalAbstract.pdf}
+  \caption{Conceptual illustration of the employed analytical framework to
+  investigate connectivity for dispersing African wild dogs under a minimum and
+  maximum flood extent in the Okavango Delta. (a) Based on a series of 700
+  floodmaps, obtained between the years 2000 and 2019, we derived extreme
+  scenarios for a minimum and maximum flood, using the minimum 100 and maximum
+  100 historically observed flood extremes. These layers were then joined with a
+  static set of covariates and fed into a dispersal simulation. (b) The
+  dispersal simulation was based on a previously parametrized integrated
+  step-selection function for dispersing wild dogs and allowed simulating
+  dispersal trajectories in the two extreme scenarios. (c) From simulated
+  dispersal trajectories we generated four summary maps, highlighting
+  connectivity patterns and areas potentially prone to human-wildlife conflict.}
+  \label{Concept}
+ \end{center}
+\end{figure}
+
+\section{Materials and Methods}
+
+\subsection{Study Area}
+
+To investigate the effects of flood extremes on dispersal and connectivity, we
+focused on a \textit{core study area} of 70,000 km\textsuperscript{2} that
+comprised the Delta and its surroundings (\Cref{StudyArea}b; purple circle). To
+accommodate for the long-distance dispersal events observed in this ecosystem
+\citep{Cozzi.2020, Hofmann.2021, Cozzi.2023}, we also considered an
+\textit{extended study area} of  300,000 km\textsuperscript{2} spanning from
+21\degree 30' S to 17\degree 30' S to and 20\degree 30' E to 26\degree E and
+comprising the Delta and parts of Angola, Namibia, Botswana, Zimbabwe, and
+Zambia (\Cref{StudyArea}a). Rainfall in this area averages at 450 mm and is
+mainly concentrated to the wet season between November and April
+\citep{Mendelsohn.2004}. Even though local rains and ground table levels
+resulting from floods of previous years may impact annual flood levels
+\citep{McCarthy.2006}, the Delta's flood extent is primarily driven by
+precipitation patterns in its catchment areas in the Angolan highlands, from
+where water is channeled into the Cubango and Cuito rivers and subsequently
+discharged into the Okavango Delta \citep{McCarthy.1997, McCarthy.2003}. Since
+water only slowly descends from the catchment areas into the Delta and across
+its shallow alluvial fan (gradient $\sim$ 1/3,400, \citealp{McCarthy.1997}), the
+flood usually reaches its maximum extent long after the local rains have ceased,
+during the peak dry season in July or August. Once the floodwater reaches the
+Delta's distal ends, it percolates at the Thamalakane and Kunyere Faults, two
+natural fault lines at which the water-flow is hindered (\Cref{StudyArea}b).
+After reaching its maximum extent, water evaporates or percolates and the flood
+steadily retracts over subsequent months. Such inner-annual dynamics are further
+amplified or buffered by multi-annual cycles in precipitation patterns across
+Angola \citep{Wolski.2012}. Due to the intricate interplay between precipitation
+patterns in Angola, local ground-table levels, evapotranspiration, and
+anthropogenic water abstractions, predictions of the flood extent under climate
+change have proven challenging and range from ``much drier'' to ``much wetter''
+depending on employed climate scenarios \citep{Kgathi.2006, Wolski.2008,
+Hughes.2011, Wolski.2012, Moses.2018}. In the core study area, the vegetation is
+dominated by mopane forest (\textit{Colophospermum mopane}), mixed woodland
+acacia-dominated (\textit{Acacia spp.}), and grassland \citep{Mendelsohn.2010}.
+Human influence is relatively low and mainly concentrated around small villages
+at the western and southern periphery of the Delta. The largest urban center is
+Maun, a spread-out settlement at the south-eastern tip of the Delta. Large
+portions of the core study area are designated as national parks, game reserves,
+or wildlife management areas. Some remaining sections of land are used for
+cattle farming and suffer from moderate levels livestock depredation and
+associated human-wildlife conflict \citep{Gusset.2009, McNutt.2017}. The core
+study area and its extended surroundings are part of the world's largest
+transboundary conservation initiative, the Kavango-Zambezi Transfrontier
+Conservation Area (KAZA-TFCA, \Cref{StudyArea}a), which attempts to re-establish
+connectivity among several core habitats. Previous studies have indicated that
+this initiative has high potential for improving connectivity among
+subpopulations of various species \citep{Brennan.2020, Lines.2021}, including
+African wild dogs \citep{Hofmann.2021}. However, a better understanding of
+connectivity within this ecosystem under climate change is lacking.
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = \textwidth]{Figures/StudyArea2.png} \caption{(a)
+  Location of the study area, which forms part of the Kavango-Zambezi
+  Transfrontier Conservation Area (KAZA-TFCA, red polygon) in Southern Africa.
+  (b) The division of the core study area (enclosed by the purple ellipse) into
+  sub-areas (1 to 6) was based on the hydrographic structure of the Okavango
+  Delta and its tributaries. We simulated dispersal trajectories starting at
+  random locations within each of the six source areas (orange circles). Purple
+  zones (7 to 14) represent zones that we used to identify if and where
+  simulated dispersers left the close surroundings of the Okavango Delta (thus
+  called egression zones). These zones were generated using a set of cutlines
+  (purple dotted lines) originating from the center of the Delta that dissected
+  an elliptical buffer surrounding the Delta into sections of equal size and in
+  accordance with cardinal directions. The yellow lines represent GPS
+  trajectories of dispersing wild dogs that were observed using GPS collars.
+  While not all of these individuals were used to parametrize the dispersal
+  model presented by \citet{Hofmann.2023}, they serve to highlight the spatial
+  scale at which the species disperses. Map lines delineate study areas and do
+  not necessarily depict accepted national boundaries.}
+  \label{StudyArea}
+ \end{center}
+\end{figure}
+
+\subsection{Spatial Habitat Layers}
+
+We represented the physical landscape through which dispersers could move by a
+set of spatially referenced habitat layers, each with a resolution of 250 m x
+250 m. The set of layers included water-cover, distance-to-water, tree-cover,
+shrub/grassland-cover, and a composite human influence layer representing
+settlements, roads, and agricultural areas. A detailed description of these
+layers is provided by \citet{Hofmann.2021}. However, unlike Hofmann and
+colleagues, who used a single, static flood map, we employed composite flood
+maps to study connectivity. Specifically, we generated water-cover layers using
+MODIS Terra MCD43A4 satellite imagery that we classified into binary water-cover
+maps using an algorithm developed by \citet{Wolski.2017} and implemented in
+\texttt{R} (\url{https://github.com/DavidDHofmann/floodmapr}). This algorithm
+allowed us to generate weekly updated ``flood extent maps'', thus providing
+detailed information about the flood at any given point in time. In total, we
+generated 700 flood extent maps, covering the years 2000 to 2019. We used these
+maps to produce minimum and maximum flood scenarios, representative of dry and
+wet climatic conditions, assuming that climate change will shift the system
+towards one of these extremes in the long run. To create the baseline minimum
+(and respectively, maximum) flood scenario, we averaged the 100 flood extent
+maps with the smallest (highest) flood extent and generated a binary layer by
+masking all pixels that were inundated in less than 50\% of the maps (see also
+\Cref{Concept}). By averaging across 100 flood extent maps, we followed a
+conservative approach and mitigated chances of misrepresenting minimum and
+maximum flood extent due to inaccuracies or artifacts in single remote sensed
+flood extent maps. The final minimum and maximum flood extent maps
+(\Cref{FloodExtent}) show that the flood in the two scenarios covers 3,500
+km\textsuperscript{2} and 9,500 km\textsuperscript{2}, respectively. Finally, we
+combined the two flood extent maps with the set of above-mentioned habitat
+layers into two spatial stacks, each representative of one scenario
+(\Cref{Concept}). To prevent edge effects during the dispersal simulation, we
+followed \citet{Koen.2010} and expanded the stacks to an extent that was 20\%
+wider and taller than the original, filling the so created buffer zone (gray
+buffer in \Cref{StudyArea}b) with randomized values from the respective layers.
+This would allow simulated dispersers to leave and re-enter the main study area
+via randomized buffer zones.
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = \textwidth]{Figures/FloodExtent.png}
+  \caption{(a) Minimum and (b) maximum flood extent. The two maps were generated
+  based on 100 minimum and 100 maximum flood extent maps from a series of 700
+  almost weekly updated remote sensed MODIS MCD43A4 satellite images spanning
+  the years 2000 to 2019.}
+  \label{FloodExtent}
+ \end{center}
+\end{figure}
+
+\subsection{Source Areas and Egression Zones}
+
+We simulated dispersing wild dogs originating from six distinct \textit{source
+areas} located within the main study area (\Cref{StudyArea}b). Defining these
+source areas was necessary to enable identification and quantification of the
+number of successful dispersal events between different regions of the core
+study area across the two flood scenarios. We selected source areas in regions
+that remained dry during both scenarios and are known to host viable African
+wild dog subpopulations. Nevertheless, these areas are not exhaustive and viable
+wild dog populations are known to reside outside of them. The study area
+comprises mostly of continuous habitat, yet natural and anthropogenic landscape
+features allow the subdivision of the core study area into several sub-regions.
+The Delta's hydrography and human settlements at its southern fringes result in
+a natural latitudinal split between the eastern (areas 1, 2, and 3) and western
+(areas 4 and 5) Delta (\Cref{StudyArea}b). Areas 1 and 2 are separated
+longitudinally by the Selinda Spillway, a waterway that connects the Delta with
+the Linyanti Swamp, while areas 2 and 3 are separated by the Khwai River, which
+inundates the Mababe Depression east of the Delta (\Cref{StudyArea}b). On the
+western side of the Delta, areas 4 and 5 are longitudinally separated by the
+Thaoga River drainage at Nokaneng (\Cref{StudyArea}b). Finally, the Delta hosts
+a central island, known as Chief's Island, which we defined as source area 6
+(\Cref{StudyArea}b). The shape of this area was determined by its borders
+towards water during the maximum flood scenario, whereas all other source areas
+were defined using circular polygons with radii of 20 km at somewhat equal
+distances to each other. We deliberately selected equally sized areas to
+facilitate comparability between them and to minimize variability in simulated
+connectivity due to target-effects. We also avoided source areas that directly
+bordered each other, as simulated individuals initiated close to a shared border
+would immediately result in connectivity and obscure differences between the two
+scenarios. Besides source areas 1 to 6, we also generated \textit{egression
+zones} (\Cref{StudyArea}b). These zones enabled us to determine through which
+regions simulated individuals left the core study area, as well as to identify
+after how many simulated steps they did so. We generated these zones by
+overlaying the Delta with an ellipse that we dissected into equally sized
+polygons in accordance with cardinal directions (\Cref{StudyArea}b). Overall,
+the selection of source areas and egression zones was based on a mixture between
+biological and computational considerations.
+
+\subsection{Dispersal Simulation}
+
+We used a previously parameterized and validated dispersal model to simulate
+dispersal of African wild dogs. This dispersal model was trained using GPS data
+obtained from 16 wild dog coalitions dispersing across northern Botswana
+\citep{Hofmann.2023}. These data were collected at a sampling rate of four
+hours, with the exception of one GPS location in the afternoon being skipped in
+order to accommodate the species tendency to be less active during this period
+(details in \citealp{Cozzi.2020, Hofmann.2021}). As such, a total of five GPS
+locations were collected per day. The collected data were then fed into an
+integrated step-selection function (iSSF, \citealp{Avgar.2016}), where
+consecutive GPS locations were converted into \textit{observed steps} (the
+straight-line connecting two consecutive GPS recordings; \citealp{Turchin.1998})
+and compared to a set of \textit{random steps} using (mixed effects) conditional
+logistic regression (\citealp{Fortin.2005, Thurfjell.2014, Muff.2020,
+Fieberg.2021}; but see \citealp{Michelot.2024} for alternative approaches).
+iSSFs provide less biased estimates of habitat-selection than traditional
+resource-selection functions \citep{Forester.2009, Zeller.2016} and bear the
+benefit that random steps are drawn from parametric step-length and
+turning-angle distributions, such that a model parametrized using iSSFs
+resembles a fully mechanistic model from which movement can be simulated
+\citep{Signer.2017, Potts.2023, Signer.2023}. In contrast to graph-based
+connectivity models (e.g., least-cost path analysis or circuit-theory),
+simulations from iSSFs alleviate the need for resistance surfaces, which are
+often subjective \citep{Simpkins.2017, Marrec.2020} and overestimate conductance
+in difficult to reach habitat \citep{Signer.2017, Hofmann.2023}. The dispersal
+model presented by \citet{Hofmann.2023} comprised a habitat-selection function
+(describing habitat selection), a movement kernel (describing dispersers'
+movement capacities), and potential interactions between the two. According to
+the parametrized model, the main characteristics of wild dog dispersal movements
+are avoidance of water, avoidance of areas influenced by humans, and a
+preference for directional and fast movements. The associated model parameters
+(provided in Figure S1) can be used to predict probabilities of a step being
+chosen among a set of random steps (details in Supporting Information S1), which
+allows to simulate movements in discrete time.
+
+Originating from each of the six source areas (\Cref{StudyArea}b), we simulated
+2,000 individual dispersing trajectories, each composed of 2,000 steps. Two
+thousand steps corresponded to a dispersal duration of $\sim$ 400 days, which
+marks the longest dispersal duration recorded in the study area
+\citep{Cozzi.2020, Hofmann.2021}. By simulating 2,000 steps, we generated the
+maximum amount of information possible and did not limit connectivity by
+defining a dispersal cap. Note that we assumed that individuals would not
+disperse for longer than 2,000 steps, but did not model settlement explicitly.
+This is because only little is known about how dispersing wild dogs locate
+potential mates and a suitable territory to settle. Olfactory cues at shared
+scent-marking sites appear to play a crucial role, yet these are extremely
+difficult to find and monitor \citep{Apps.2022, Claase.2022}. We simulated 1,000
+trajectories under the minimum flood extent, the remaining 1,000 under the
+maximum flood extent. Across the six source areas, this resulted in the
+simulation of a total of 12,000 individual dispersal trajectories, which
+sufficed to achieve convergence in relevant connectivity metrics
+\citep{Hofmann.2023}. The simulation procedure was based on the algorithm
+described by \citep{Signer.2017} and \citet{Hofmann.2023} and works as follows.
+A random location within the source area is defined as the starting point.
+Originating from the starting point, a set of 25 random steps is generated by
+sampling step lengths ($sl$) from a gamma distribution fitted to observed steps
+(shape = 0.37, scale = 6,316) and turning angles ($ta$) from a uniform
+distribution ($-\pi, +\pi$). Along each random step, the underlying spatial
+covariates are extracted, and relevant movement metrics computed ($log(sl)$ and
+$cos(ta)$). \(\beta-\)estimates from the fitted dispersal model are then used to
+predict the probability of each step being chosen, given the step's associated
+covariates and movement metrics, as well as characteristics of all other
+proposed steps. Among the 25 proposed steps, one is chosen at random based on
+assigned probabilities. The procedure is then repeated until a total of 2,000
+steps is simulated.
+
+\subsection{Derived Metrics}
+
+Based on simulated dispersal trajectories, we quantified connectivity,
+identified the emergence of alternative dispersal corridors, and highlighted
+areas with elevated potential for HWC. Our approach drew upon the set of
+complementary connectivity metrics for individual-based movement models
+discussed by \citet{Hofmann.2023}, and was expanded to include a map
+highlighting areas with elevated HWC potential. The set of connectivity metrics
+comprised \textit{inter-patch connectivity} metrics, summarizing dispersal
+success and duration into distinct habitat patches, \textit{heatmaps}, depicting
+areas of intense use by dispersers, and \textit{betweenness maps}, highlighting
+dispersal corridors and bottlenecks. Finally, the \textit{HWC maps} served to
+reveal areas where simulated dispersers moved into the vicinity of
+human-dominated landscapes. To illustrate differences between metrics during
+maximum and minimum flooding, we computed difference maps for the heatmaps,
+betweenness maps, and HWC maps.
+
+\subsubsection{Inter-Patch Connectivity}
+
+To measure inter-patch connectivity, we tallied the number of trajectories
+successfully moving from one source-area to another and computed the average
+minimum dispersal duration (in number of steps) required to make those
+connections. Dispersal between two areas was said to be successful whenever a
+trajectory originating from one area intersected with the polygon of another
+area. If a trajectory moved through multiple areas, a connection into each of
+them from the original source was recorded. We also estimated the number of
+individual trajectories that left the core study area and moved into one of the
+egression zones. To quantify variability in our estimates, we generated
+bootstrap samples by randomly drawing 1,000 simulated trajectories per source
+area 1,000 times with replacement and computing 95\%-percentile intervals.
+
+\subsubsection{Intensity of Use}
+
+To quantify intensity of use of different areas, we generated heatmaps (also
+known as \textit{utilization distributions} or \textit{flow-maps}) by
+superimposing the study area with a grid with a spatial resolution of 1 km x 1
+km and determining the number of simulated trajectories traversing each grid
+cell. Such maps are essentially movement-restricted permeability surfaces, thus
+highlighting areas of intense use \citep{Signer.2017, Potts.2023, Hofmann.2023}.
+Heatmaps are relatively insensitive towards the chosen resolution and a 1 km x 1
+km grid appeared to provide a good balance between detail and computational
+feasibility.
+
+\subsubsection{Betweenness}
+
+To compute spatially mapped betweenness scores, we overlaid the study area with
+a grid that had a resolution of 2.5 km x 2.5 km and computed the frequency at
+which simulated individuals transitioned from one grid-cell to another. A
+cell-transition was said to occur whenever a simulated step crossed from one
+grid-cell across or into another. In case the same individual repeatedly
+realized the same cell-transition, we only counted a single transition to avoid
+overemphasizing regions where individuals moved in circles. This procedure
+resulted in a weighted edge-list that we used to compute weighted betweenness
+scores for each grid-cell, i.e. the importance of the respective grid-cell in
+facilitating movement into adjacent areas \citep{Bastille-Rousseau.2018,
+Bastille-Rousseau.2021}. We computed betweenness using the \texttt{igraph}
+\texttt{R}-package \citep{Csardi.2006}. Calculations of betweenness scores are
+computationally demanding and result in discrete and narrow dispersal corridors
+\citep{Hofmann.2023}. A higher map resolution thus results in increased
+computation time and narrower corridors. In our case, a grid size of 2.5 km x
+2.5 km provided a sensible compromise between computational efficiency and
+biological relevance of emerging corridors (see also the appendix of
+\citealp{Hofmann.2023}).
+
+\subsubsection{Human-Wildlife Conflict}
+
+To pinpoint potential hotspots for HWC, we identified regions where simulated
+trajectories moved into the vicinity of human-dominated landscapes. For this, we
+isolated locations along simulated trajectories that ranged $\leq$ 500 meters to
+the nearest grid-cell with human influence $>$ 0. Based on the so isolated
+coordinates, we generated density maps highlighting likely hotspots for HWC.
+While this approach of mapping potential HWC ignores human population density,
+we also generated a compound metric by multiplying the generated heatmaps with a
+continuous human-influence map (Supporting Information S10). Notably, not every
+animal in the vicinity of human-dominated landscapes implies conflict, hence we
+chose to use the term \textit{potential} hotspots for HWC. We prepared these
+maps at a resolution of 1 km x 1 km.
+
+\subsection{Software}
+
+We conducted all data preparation and analyses using the programming language
+\texttt{R} \citep{RCoreTeam.2023}. For any spatial data manipulation, we used
+the \texttt{R}-packages \texttt{terra} \citep{Hijmans.2023} and
+\texttt{spatstat} \citep{Baddeley.2015}. Several helper functions for the
+dispersal simulation algorithm were written in \texttt{C++} and imported into
+\texttt{R} using the \texttt{Rcpp} package \citep{Eddelbuettel.2011}. Network
+analysis was achieved in \texttt{igraph} \citep{Csardi.2006} and figures were
+generated using \texttt{ggplot2} \citep{Wickham.2023} and \texttt{ggnetwork}
+\citep{Briatte.2023}. Notably, a similar simulation procedure as employed in our
+analysis has recently been added to the \texttt{amt} \texttt{R}-package
+\citep{Signer.2023}.
 
 \section{Results}
+\subsection{Inter-Patch Connectivity}
+
+Our analysis of inter-patch connectivity revealed significant differences in
+dispersal success and duration depending on the flood scenario (\Cref{IPCTable},
+\Cref{IPCMap}, Figure S2). Of the 6,000 simulated dispersal trajectories for
+each flood scenario, 4,137 $\pm$ 35 trajectories reached another source area
+during minimum flood, whereas 3,625 $\pm$ 32 trajectories did so during maximum
+flood, indicating a 12\% lower dispersal success during maximum flood
+(\Cref{IPCTable}a1). Concurrently, the average dispersal duration of
+trajectories moving from one source area to another was 17\% lower during
+maximum flood, with 122 $\pm$ 1 days compared to 143 $\pm$ 2 days
+(\Cref{IPCTable}b1) during minimum flood. The disparities were most pronounced
+for trajectories dispersing into source area 6 (the Delta's center), with 1,327
+$\pm$ 30 simulated trajectories reaching area 6 during minimum flood compared to
+297 $\pm$ 17 trajectories (i.e., 78\% lower) during maximum flood
+(\Cref{IPCTable}a1). Source area 6 therefore appeared to be particularly
+vulnerable to isolation in the maximum flood scenario. Interestingly,
+connectivity between some source areas was higher during maximum flood. For
+instance, the number of trajectories running from area 5 into area 4 was 67\%
+higher, with 114 $\pm$ 10 trajectories during minimum flood compared to 190
+$\pm$ 12 trajectories during maximum flood (\Cref{IPCTable}a1). Simulated
+individuals therefore responded to unfavorable habitat conditions by dispersing
+into another region of the landscape. Contrary to our expectations, movement
+into egression zones (areas 7 - 14) was only marginally higher during maximum
+flood, with the number of trajectories rising by 2\%, from 5,457 $\pm$ 22 to
+5,553 $\pm$ 20 (\Cref{IPCTable}a2 and Figure S5).
+
+\begin{figure}
+\begin{center}
+ \includegraphics[width = \textwidth]{Figures/InterpatchConnectivityTable2.png}
+ \caption{(a) Dispersal frequency (measured as the number of dispersal
+ trajectories) and (b) dispersal duration (in days) between (a1, b1) source
+ areas (labeled 1 to 6) and (a2, b2) egression zones (labeled 7 to 14) during
+ minimum and maximum flood. Final rows and columns of each subplot represent
+ summary statistics across all areas. For instance, the bottom row in a1
+ highlights for each source area the total number of trajectories successfully
+ dispersing into a target area. The last column, in contrast, indicates the
+ total number of trajectories immigrating into each target area. Values indicate
+ mean $\pm$ SD. The colors are mapped in a non-linear fashion to avoid
+ over-emphasis on final rows and columns.}
+ \label{IPCTable}
+\end{center}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = 0.8\textwidth]{Figures/InterpatchConnectivityMap.png}
+  \caption{Dispersal frequency and duration between source areas 1 to 6 in the
+  core study area during the minimum (left panel) and maximum (right panel)
+  flood scenario. The dispersal frequency shows the number of trajectories that
+  emigrated from a specific source area and successfully immigrated into another
+  one. The dispersal duration indicates the mean dispersal duration (in days)
+  required before trajectories arrived at the respective area. Although links
+  between non-neighboring source areas occurred in the simulation, we here only
+  present links between adjacent source areas for brevity. Detailed plots about
+  source-specific inter-patch connectivity are provided in Figures S3 and S4.}
+  \label{IPCMap}
+ \end{center}
+\end{figure}
+
+\subsection{Intensity of Use}
+
+Heatmaps revealed that the Delta acted as a major dispersal barrier during
+maximum flood, but became permeable during minimum flood (\Cref{Metrics}a).
+During maximum flood, the floodwater extended from the Delta's inflow until the
+densely inhabited regions at its southern fringes (the town of Maun and its
+surroundings), such that the negative impacts of water on dispersal were further
+aggravated by the negative effects of anthropogenic presence. Consequently,
+there was a substantial decrease in the dispersal frequency across the southern
+fringes of the Delta during maximum flood (\Cref{Metrics}a). During minimum
+flood, however, the retracting flood revealed vast dispersal areas that
+individuals used to move across the otherwise inundated regions
+(\Cref{Metrics}a).
+
+\subsection{Betweenness}
+
+Findings from the betweenness maps, which display corridors and pinch-points
+between neighboring areas, reinforced patterns observed on the heatmaps
+(\Cref{Metrics}b). Multiple corridors linking source area 6 to the surrounding
+source areas existed during minimum flood, but many of them vanished during
+maximum flood (\Cref{Metrics}b). Instead, a single corridor at the south-eastern
+tip of the Delta emerged (\Cref{Metrics}b). Despite the corridor's high
+betweenness score, its apparent ability to link the eastern and western sections
+of the Delta was limited, as evidenced by a low traversal frequency on the
+heatmap (\Cref{Metrics}a).
+
+\subsection{Human-Wildlife Conflict}
+
+The HWC maps highlighted two potential hotspots for HWC that depended on
+climatic extremes (\Cref{Metrics}c). The first hotspot was located at the
+Delta's inflow between source areas 3 and 4 and was most prominent during
+minimum flood. During maximum flood, by contrast, the density of simulated
+locations in the vicinity of humans decreased by 41\% in this area
+(\Cref{Metrics}c and Figure S9). Another hotspot, albeit visually less distinct,
+covered the region at the distal end of the Delta, extending across the town of
+Maun (\Cref{Metrics}c). This area was particularly relevant during maximum
+flood, with a density that was 126\% higher than during minimum flood
+(\Cref{Metrics}c and Figure S9). The diversion of individuals from one area to
+another depending on flood conditions therefore drives the emergence of
+potential hotspots for HWC.
+
+\begin{figure}
+ \begin{center}
+  \includegraphics[width = 0.8\textwidth]{Figures/Metrics.png}
+  \caption{(a) Heatmaps, (b) betweenness maps, and (c) maps of potential for HWC
+  derived from simulated dispersal trajectories. Left panels were derived from
+  the minimum flood scenario, right panels from the maximum flood scenario.
+  Source areas from which dispersers were released are numbered 1-6. The color
+  scale for betweenness scores in (b) was square-rooted to improve visibility of
+  corridors with lower values. Source-specific heatmaps, betweenness maps, and
+  HWC maps are provided in Figures S6-S8.}
+  \label{Metrics}
+ \end{center}
+\end{figure}
 
@@
 
+\subsection{Impacts of Extreme Environmental Conditions on Dispersal}
+
+We investigated the impacts of extreme environmental conditions, as anticipated
+under the influence of climate change, on dispersal and connectivity. For this,
+we employed a previously parameterized and validated movement model and
+simulated dispersal trajectories of African wild dogs in the Okavango Delta
+under extreme conditions. Our findings indicate that an amplified flood can
+significantly reduce dispersal success, prolong dispersal durations, and alter
+the spatial structure of major dispersal corridors, thereby creating or
+shifting hotspots of potential human-wildlife conflict (HWC).
+
+Wetter-than-usual climatic conditions, a likely scenario for this study area
+\citep{Wolski.2008, IPCC.2022}, resulted in significant barriers to dispersal
+and connectivity, and in an overall increase in dispersal durations. This was
+largely a result of simulated individuals avoiding large water bodies, thus
+detouring through more suitable habitats \citep{Cozzi.2013, Cozzi.2020,
+Hofmann.2021, Hofmann.2023}. Movement constraints due to changes in
+environmental conditions can thus lead to subpopulation isolation (up to $>$
+75\% in our case), thereby jeopardizing recolonization efforts following local
+extinctions, and impacting population dynamics, gene flow, and genetic
+diversity \citep{Hanski.1999, Frankham.2002, Leigh.2012, Baguette.2013}.
+
+The African wild dog exhibits a remarkable dispersal ability, sometimes
+dispersing several hundred kilometers across international borders
+\citep{McNutt.1996, Davies-Mostert.2012, Masenga.2016, Cozzi.2020,
+Sandoval-Seres.2022, Cozzi.2023}. This ability likely evolved as an adaptation
+to the relatively low density at which the species occurs \citep{Creel.2002,
+Masenga.2016} and may have been sustained by a historically well-connected
+landscape that imposed low mortality (sensu \citealp{Fahrig.2007}). Today,
+however, the need for vast dispersal habitats makes this species vulnerable to
+habitat loss and fragmentation \citep{Woodroffe.2012, Woodroffe.2020}. When
+dispersing animals are forced to alter their dispersal routes due to changing
+environmental conditions, this could result in non-optimal dispersal behavior in
+human-altered landscapes \citep{Fahrig.2007}. This is particularly problematic
+when dispersers are channeled through areas with high anthropogenic resistance
+and elevated risk of mortality \citep{Fahrig.2007, Ghoddousi.2021,
+VanDerMeer.2014}. For a population of African wild dogs in Zambia,
+\citet{Leigh.2012} revealed that dispersal and the associated mortality indeed
+led to a decline in genetic diversity and, subsequently, to a net-loss of
+individuals. At present, our own study population appears to benefit from
+moderate levels of dispersal, as an analysis across Southern Africa revealed a
+genetically healthy cluster in the Kavango-Zambezi Transfrontier Conservation
+Area \citep{Tensen.2022}. This cluster may, however, be at risk depending on
+future flood conditions.
+
+% The redirection of dispersal paths coupled with altered climatic conditions and
+% increased encroachment by humans, could therefore result in non-optimal
+% dispersal behavior (sensu \citealp{Fahrig.2007}), especially when dispersers are
+% forced to move across regions with high anthropogenic resistance and elevated
+% risk of mortality \citep{Fahrig.2007, Ghoddousi.2021, VanDerMeer.2014}. Indeed,
+% \citet{Leigh.2012} revealed that dispersal and the associated mortality led to a
+% decline in genetic diversity in a population of wild dogs in Zambia and,
+% subsequently, to a net-loss of individuals. At present, our study population
+% appears to benefit from moderate levels of dispersal, as an analysis across
+% Southern Africa revealed a genetically healthy cluster in the Kavango-Zambezi
+% Transfrontier Conservation Area \citep{Tensen.2022}. This cluster may, however,
+% be at risk depending on future flood conditions.
+
+As dispersal is a key process influencing population dynamics
+\citep{Hanski.1999, Clobert.2012}, any climatic-induced alteration will have
+cascading effects that amplify or buffer the effect of climatic changes on
+population persistence and viability. Lower connectivity and prolonged
+dispersal durations due to adverse environmental conditions will inevitably
+result in increased energetic expenditures and exposure to various threats such
+as predation, competition, human encounters, and diseases \citep{Alberts.1995,
+Yoder.2004, Stamps.2005, Bonte.2012}, thus further jeopardizing dispersal
+success and its effect on population dynamics. Higher ambient temperatures have
+previously been associated with negative effects on wild dog reproductive
+success \citep{Woodroffe.2017, Abrahms.2022}, survival \citep{Rabaiotti.2021},
+and, consequently, population persistence \citep{Rabaiotti.2023}. The negative
+effect of hotter climatic conditions on local population survival and
+recruitment may be further exacerbated by reduced dispersal success and
+connectivity in case of amplified flooding regimes.
+
+Even though our finding that an increased flood extent reduces connectivity
+seems unsurprising, it is non-trivial to quantify such impacts and to predict
+the spatial rearrangements of movement corridors they entail. Wild dogs avoid
+crossing water bodies, yet are capable of doing so if they really need to,
+especially during dispersal \citep{McNutt.1996, Cozzi.2013, Cozzi.2020,
+Hofmann.2023}. By simply assuming such obstacles to be impenetrable, one may
+miss important corridors that lead across narrow sections of unsuitable habitat
+\citep{Marrec.2020}. Using simulations from integrated step-selection functions
+(iSSFs), however, one can realistically render that an animals' decision to
+disperse across a specific habitat is conditional on what's available at
+alternative steps. A disperser surrounded by a hostile matrix may readily cross
+water, whereas an individual dispersing through favorable habitat can choose to
+avoid water entirely. Another challenge when predicting the impacts of climate
+change on connectivity is that future on-the-ground conditions are rarely known
+(e.g., \citealp{Wolski.2008}). To overcome this, we leveraged historic data on
+seasonal variation of the flood extent and assumed that climate change would
+shift the ecosystem towards conditions that currently mark extremes. A similar
+approach can be applied in systems where the relationships between changing
+conditions and functional connectivity are less evident.
+
+While wet extremes appear to hinder dispersal and connectivity, the opposite
+applies for amplified dry conditions, because areas that are inundated under
+normal circumstances become free of water and therefore passable. Predictions
+suggest that climate change will result in elevated precipitation across the
+Delta's catchment areas in Angola \citep{Wolski.2008, IPCC.2022}, thus resulting
+in above-average flooded regions. Concomitantly, however, hotter temperatures
+and increased levels of evapotranspiration are anticipated for the Delta's basin
+in Botswana, suggesting that dry periods may be experienced too
+\citep{Wolski.2008, Moses.2018, Akinyemi.2019, IPCC.2022}. The Delta's expanse
+is also affected by multi-decadal oscillations in precipitation patterns, which
+may periodically offset or amplify long-term trends \citep{Wolski.2012}. The
+overall impact of climate change on flood patterns therefore remains unknown
+\citep{Wolski.2008, Wolski.2012} and this lack of knowledge is aggravated by
+uncertainties regarding future anthropogenic water abstractions along the
+Okavango River \citep{Kgathi.2006, Murray-Hudson.2006, Hughes.2011}. It is
+nonetheless worth noting that our maximum flood extent map, which was prepared
+based on 100 maximum flood extents observed on over 700 weekly-updated
+floodmaps, is conservative compared to historically observed flood extents in
+the Delta. While the flood in our map encompasses an area of 9,500
+km\textsuperscript{2}, the Delta may inundate an area of up to 14,000
+km\textsuperscript{2} in exceptional cases \citep{McCarthy.2003,
+Gumbricht.2004}. Changes in dispersal behavior and connectivity due to climate
+change may therefore be more pronounced than reported here.
+
+\subsection{Human-Wildlife Conflict}
+
+Our analysis revealed that dispersers utilize different movement corridors
+depending on flood conditions, leading to the emergence and shifting of
+potential hotspots for HWC. An increased proximity between humans and carnivores
+is typically associated with a higher potential for conflict (e.g.,
+\citealp{Michalski.2006, Chapman.2016}). Especially dispersers, which often
+venture outside protected areas, expose themselves to anthropogenic hostility
+\citep{Elliot.2014, Cozzi.2020, Vasudev.2023}. In our case, the funneling of
+individuals through alternative dispersal corridors resulted in the shift of
+potential HWC hotspots, potentially increasing HWC and retaliatory killing. One
+example is the area surrounding Maun, where HWC is expected to rise during wet
+periods, exacerbating existing human-wildlife conflict \citep{Gusset.2009,
+McNutt.2017, Cozzi.2020}. Identifying and mapping the conflict-connectivity
+interface (i.e., areas prone to HWC) in light of climate change can improve our
+understanding of connectivity \citep{Vasudev.2023} and could aid in prioritizing
+mitigation and prevention measures that have the highest impact
+\citep{Treves.2011, Buchholtz.2020} or to develop appropriate compensation
+schemes \citep{McNutt.2017}. However, it is important to note that encounters
+between animals and humans do not imperatively imply HWC, and that the severity
+of conflict and its impact on connectivity likely depends on humans' perception
+and attitude towards wildlife \citep{Ghoddousi.2021}. Besides the risks
+associated with direct mortality through human persecution, a higher proximity
+to humans also increases the risk of indirect mortality through disease
+transmissions from domestic animals \citep{Cleaveland.2000, VanDeBildt.2002}.
+Previous outbreaks of distemper and rabies have caused the local extermination
+of African wild dogs in several African countries \citep{Woodroffe.2004}. As
+infected dispersing individuals may further spread diseases within protected
+areas, a better understanding of points of interaction between humans and
+wildlife will facilitate the implementation of targeted vaccination programs in
+the face of climatic change \citep{Vial.2006}.
+
+\subsection{Methodological Limitations and Future Research Directions}
+
+In the present study, we ignored mortality during dispersal and assumed all
+simulated dispersers to survive for 2,000 steps ($\sim$ 400 days). This likely
+resulted in an overestimation of connectivity, especially between distant source
+areas \citep{Kramer-Schadt.2004, Diniz.2019, Day.2020, FletcherJr..2019}. The
+inclusion of mortality in individual-based simulations is relatively straight
+forward \citep{FletcherJr..2019, FletcherJr..2023}, yet its estimation for
+dispersing individuals has proven difficult \citep{Behr.2023}. During the course
+of our study, we recorded and confirmed only two mortality events during
+dispersal, thus precluding a detailed investigation. Notably, unless the risk of
+mortality is included in simulations in a spatially explicit manner, its
+inclusion will merely shorten the length of simulated trajectories and its
+impact on inferred connectivity patterns will be relative in nature.
+
+Even though we assumed vegetation to remain unchanged in both extreme scenarios,
+we acknowledge that vegetation cover and phenology may shift in response to
+climate change, potentially influencing dispersal indirectly through prey
+availability \citep{Bonyongo.2005}. While flooding represents the primary
+barrier to dispersing individuals \citep{Cozzi.2020, Hofmann.2021,
+Hofmann.2023}, refining and capturing the role of vegetation dynamics on prey
+distribution and dispersal will therefore be crucial. However, predicting the
+impacts of climate change on vegetation cover and its effect on the
+distribution of prey will be challenging and is an area for continued research.
+
+Our dispersal model was validated using independent dispersal data (see
+\citealp{Hofmann.2023}), yet validating predictions of connectivity and HWC
+under the two presented scenarios remains challenging. This is mainly owed to
+the difficulty of monitoring dispersing individuals \textit{per se}, coupled
+with the low likelihood of observing dispersal during periods of extreme
+flooding. Genetic data could serve to validate predictions of historical
+connectivity patterns \citep{Cushman.2010, Spear.2010}, yet is equally difficult
+to collect and does not provide a means of validating predicted connectivity
+patterns in the short run. Observational data, including photographic evidence
+from citizen scientists, however, could serve to fill this gap
+\citep{Marnewick.2014, Cozzi.2023}. Most carnivores are individually
+identifiable, either by their coat pattern or other unique morphological
+features (e.g., \citealp{Pennycuick.1970} and \citealp{Kelly.2001}). Automatic
+recognition of individuals from photographic evidence using artificial
+intelligence will therefore open up new avenues to track individuals through
+space and time \citep{Cozzi.2023}. The so collected data provides a means to
+monitor dispersal across unprecedented temporal and validate predicted patterns
+of landscape connectivity.
+
+\subsection{Conclusion and Implications}
+
+In conclusion, our dispersal simulation across two extreme environmental
+scenarios revealed striking differences in dispersal prospects and landscape
+connectivity for dispersing animals. We thereby showed that extreme
+environmental conditions, akin to those projected under climate change, will
+have important impacts on functional connectivity and may alter areas
+susceptible to human-wildlife conflict. Gaining insights into the dispersal
+patterns of various wide-ranging species under changing conditions will be
+critical for designing and preserving movement corridors \citep{Vasudev.2015}.
+Wildlife managers and conservation bodies would also benefit from moving beyond
+a static assessment of connectivity and considering multiple, potentially
+extreme, projections of environmental conditions to help gauge the impacts of
+climate change on dispersal and connectivity. Successful conservation strategies
+will be of particular relevance for wide-ranging and dispersal-dependent species
+that are already threatened with extinction, such as the African wild dog.
+Ultimately, our study provides valuable insights into the potential effects of
+climate change on dispersal, connectivity, and human-wildlife-conflict dynamics,
+emphasizing the relevance of proactive conservation measures and targeted
+mitigation strategies that incorporate predictions of environmental change.
+
 \section{Authors' Contributions}
-D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
-D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
-first draft of the manuscript and all authors contributed to the drafts at
-several stages and gave final approval for publication.
+
+D.D.H., G.C., D.M.B., A.O. and conceived the study, D.D.H. and G.C. designed
+methodology; D.D.H., G.C., D.M.B., and J.W.M. collected the data; D.D.H.
+analysed the data; D.D.H. and G.C. wrote the first draft of the manuscript and
+all authors contributed to the drafts at several stages and gave final approval
+for publication.
 
 \section{Data Availability}
-GPS movement data of dispersing coalitions will be made available on dryad at
-the time of publication. Access to all R-scripts for our analyses is provided
-through Github.
+
+The data that support the findings of this study are openly available in Dryad
+at \url{https://doi.org/10.5061/dryad.z34tmpgnm} \citep{Hofmann.2024a}. R-codes
+required to reproduce all analyses are accessible through GitHub at
+\url{https://github.com/DavidDHofmann/FloodDispersal}. Additional information on
+the preparation of covariate layers are provided by \citet{Hofmann.2021} and
+a detailed description of the dispersal model is given by \citet{Hofmann.2023}.
 
 \section{Acknowledgements}
-We thank the Ministry of Environment and Tourism of Botswana for granting
-permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
-Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
-to Johannes Signer for assisting with the simulation algorithm. This study was
-funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
-Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung
-Temperatio, Wilderness Wildlife Trust Foundation, Forschungkredit der
-Universität Zürich, and a Swiss National Science Foundation Grant
-(31003A\_182286) to A. Ozgul.
+We thank the Ministry of Environment and Tourism and the Department of Wildlife
+and National Parks of Botswana for granting permission to conduct this research.
+We also thank two excellent reviewers for their critical and detailed feedback,
+which helped to improve the quality of the manuscript significantly. This study
+was funded by Claraz Foundation, Forschungkredit der Universität Zürich, and a
+Swiss National Science Foundation Grant Nr. 31003A\_182286 to A. Ozgul and Nr.
+310030\_204478 to G. Cozzi.
 
@@
 \singlespacing
-\bibliography{Literatur}
+\ifthenelse{\boolean{usebiblatex}}{
+  \begin{refcontext}[sorting=nyt]
+  \printbibliography
+  \end{refcontext}
+}{
+  \bibliography{LiteratureBibtex}
+}
 \endgroup
-
 \end{document}

### 2024-07-01 Chapter_2/04_Manuscript/Manuscript_Support.tex
Publish PHD

--- 2021-05-05 Chapter_1/04_Manuscript/Manuscript_Support.tex
@@
-\documentclass[abstract=off,10pt,a4paper,bibliography=totocnumbered]{article}
+\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
 \usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
@@
 \usepackage[nameinlink]{cleveref}
-% \usepackage{lineno}
 \usepackage{multirow}
 \usepackage{arydshln}
+\usepackage[flushleft]{threeparttable}
 % \usepackage[nomarkers, nolists]{endfloat}
-\usepackage[flushleft]{threeparttable}
+\usepackage{scalerel}
+\usepackage{makecell}
+\usepackage{ifthen}
+\usepackage{tikz}
+\usepackage{lineno}
+\usetikzlibrary{svg.path}
+
+%------------------------------------------------------------------------------
+%  Bibtex vs Biblatex
+%------------------------------------------------------------------------------
+% While bibtex is very nice to use, it's insanely slow. I therefore work with
+% bibtex until the final stage. Set the value below to "true" or "false" to pick
+% which one should be used.
+\newboolean{usebiblatex}
+\setboolean{usebiblatex}{true}
+
+% Depending on the question, the commands below will prepare the document
+% accordingly
+\ifthenelse{\boolean{usebiblatex}}{
+  %-----------------------------------------------------------------------------
+  %  If using biblatex
+  %-----------------------------------------------------------------------------
+  % Load required packages and refer to correct bibliography
+  \usepackage{csquotes}
+  \usepackage[
+      backend=biber
+    , style=apa          % To display Author-year in the text
+    , natbib=true        % Necessary to use citep, etc.
+    , hyperref=true      % To have clickable links in the references
+    , useeditor=false    % Don't add editor to publications
+    , sorting=ynt        % Sort by year, name, and title
+    , uniquename=false   % Avoid that initials are put
+  ]{biblatex}
+  \addbibresource{Literature.bib}
+
+  % To include author in link
+  \DeclareFieldFormat{citehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{#1}}
+  \DeclareFieldFormat{textcitehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{%
+    #1%
+    \ifbool{cbx:parens}
+      {\bibcloseparen\global\boolfalse{cbx:parens}}
+      {}}}
+  \savebibmacro{cite}
+  \savebibmacro{textcite}
+  \renewbibmacro*{cite}{%
+    \printtext[citehyperref]{%
+      \restorebibmacro{cite}%
+      \usebibmacro{cite}}}
+  \renewbibmacro*{textcite}{%
+    \ifboolexpr{
+      ( not test {\iffieldundef{prenote}} and
+        test {\ifnumequal{\value{citecount}}{1}} )
+      or
+      ( not test {\iffieldundef{postnote}} and
+        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
+    }
+      {\DeclareFieldAlias{textcitehyperref}{noformat}}
+      {}%
+    \printtext[textcitehyperref]{%
+      \restorebibmacro{textcite}%
+      \usebibmacro{textcite}}}
+
+  % Remove editor form everything but books
+  \AtEveryBibitem{%
+   \ifentrytype{book}{}{%
+    \clearname{editor}
+   }
+  }
+}{
+  %-----------------------------------------------------------------------------
+  %  If using bibtex
+  %-----------------------------------------------------------------------------
+  % Load required packages and select citation style
+  \usepackage[round]{natbib}
+  \bibliographystyle{apalike}
+}
 
@@
 %	Some Styling
+%------------------------------------------------------------------------------
+% Link colors
+\definecolor{linkcolor}{HTML}{000000}
+\hypersetup{
+  colorlinks = true,
+  linkcolor  = linkcolor,
+  urlcolor   = linkcolor,
+  citecolor  = linkcolor,
+}
+
+% Changing the style of captions in figures etc.
+\captionsetup{labelfont=bf, format=plain, font=small}
+
+% Change how equations are referenced
+\renewcommand{\theequation}{Equation \arabic{equation}}%
+
+% Avoid skip when including text from external .tex file
+\newcommand{\inputy}[1]{\input{#1}\unskip}
+
+% For supplementary material
+\newcommand{\beginappendix}{%
+  \setcounter{table}{0}
+  \renewcommand{\thetable}{S\arabic{table}}%
+  \setcounter{figure}{0}
+  \renewcommand{\thefigure}{S\arabic{figure}}%
+  \setcounter{equation}{0}
+  \renewcommand{\theequation}{Equation S\arabic{equation}}%
+  \setcounter{section}{0}
+  \renewcommand{\thesection}{S\arabic{section}}%
+}
+
+%------------------------------------------------------------------------------
+%  ORCID
 %------------------------------------------------------------------------------
@@
 
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
+% Create the ORCID logo
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{
+  orcidlogo/.pic={
+    \fill[orcidlogocol]
+      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
+        ,0,128,0C198.7,0,256,57.3,256,128z};
+    \fill[white]
+      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
+      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
+        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
+        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
+      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+  }
+}
 
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
+% Command to create the ORCID
+\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
+\begin{tikzpicture}[yscale=-1,transform shape]
+  \pic{orcidlogo};
+\end{tikzpicture}
+}{|}}}}
 
@@
 %------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Step by Step: Using Integrated Step Selection
-Analysis to Simulate Wild Dog Dispersal and Assess Landscape Connectivity}
+
+\title{\textbf{Supporting Information}\\ Dispersal and Connectivity in
+Increasingly Extreme Climatic Conditions}
 
@@
 
-\textbf{Running Title:} Simulating Wild Dog Dispersal.
+\textbf{Running Title:} Dispersal in a Changing Climate
 
@@
 
-\textbf{Keywords:} dispersal, habitat selection, integrated step selection
-function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, least-cost corridors, Lycaon pictus, permeability surface,
-protected areas, wildlife management
+\textbf{Keywords:} climate change, conservation, dispersal, flood extremes,
+individual-based simulations, landscape connectivity, movement ecology,
+step-selection functions
 
@@
 
-%------------------------------------------------------------------------------
-%	Appendix S1: All Movement Models
-%------------------------------------------------------------------------------
-\section{Movement Models}
-% latex table generated in R 3.6.3 by xtable 1.8-4 package
-% Thu Jul  9 09:17:26 2020
-\begin{table}[hbpt]
-  \caption{Results from the forward model selection procedure based on Akaike's
-  Information Criterion (AIC; \citealp{Burnham.2002}) for the movement model.
-  The base model upon which we based our movement model is depicted in the last
-  row. We omitted all models with an AIC weight of zero.}
-  \label{ModelAICs}
+\newpage
+\section{Dispersal Model}
+The model employed to simulate dispersal was based on an integrated step
+selection function (iSSF). In (integrated) step selection functions (iSSFs,
+\citealp{Fortin.2005, Avgar.2016}), observed GPS locations are converted into
+steps (the straight-line traveled between two GPS recordings
+\citep{Turchin.1998}) and compared to a set of \textit{random} steps in a
+conditional logistic regression framework \citep{Fortin.2005, Thurfjell.2014,
+Muff.2020, Fieberg.2021}. The model presented in \citep{Hofmann.2023} used
+dispersal data collected on 16 dispersing AWDs from a free-ranging wild dog
+population in northern Botswana. GPS data during dispersal was collected at
+4-hourly intervals and translated into steps of similar duration. Observed steps
+were then paired step with 24 random steps that were generated using a uniform
+distribution for turning angles (\(-\pi, +\pi\)) and step lengths from a gamma
+distribution fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\)
+= 0.37). It was then assumed that animals assigned to each observed and random
+step a selection score of the form \citep{Fortin.2005}:
+
+\begin{equation}
+\label{EQ1}
+  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
+\end{equation}
+
+Where (\(x_1, x_2, ..., x_n\)) represent the covariate values along each of the
+steps and the (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's relative
+selection strengths \citep{Avgar.2017} towards these covariates. The benefit
+of \textit{integrated} SSFs over regular SSFs is that they provide a means to
+render two complementary ``kernels''. A \textit{movement} kernel that describes
+general movement behavior of dispersing AWDs and a \textit{habitat} kernel that
+describes preferences of AWDs with regards to environmental conditions
+\citep{Fieberg.2021}. iSSFs also allow interactions among the two kernels and
+are thus suitable to render that movement behavior may change depending on
+habitat conditions. A fitted iSSF model can be used as an individual-based
+movement model to simulate dispersal \citep{Signer.2017, Hofmann.2023}.
+
+\newpage
+\section{Dispersal Model Estimates}
+\Cref{Model} depicts estimates from the model developed by \citet{Hofmann.2023}
+that we used to simulate dispersal trajectories in the present.
+
+\begin{figure}[htbp]
   \begin{center}
-    \resizebox{\textwidth}{!}{
-      \begin{threeparttable}
-        \begin{tabular}{lllll}
-        \toprule
-          Covariates & AIC & \(\Delta\)AIC & Weight & LogLik \\
-          \midrule
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) & 89392.88 & 0.00 & 0.15 & -44670.44 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:log(sl) & 89393.92 & 1.04 & 0.09 & -44669.96 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + DTW:log(sl) & 89394.13 & 1.25 & 0.08 & -44670.06 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:log(sl) & 89394.25 & 1.37 & 0.08 & -44670.13 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl & 89394.36 & 1.48 & 0.07 & -44672.18 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + log(sl):LA & 89394.44 & 1.56 & 0.07 & -44670.22 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:sl & 89394.56 & 1.68 & 0.07 & -44670.28 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:log(sl) & 89394.57 & 1.69 & 0.07 & -44670.29 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:cos(ta) & 89394.59 & 1.71 & 0.07 & -44670.30 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:cos(ta) & 89394.63 & 1.75 & 0.06 & -44670.31 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + sl:cos(ta) & 89394.68 & 1.80 & 0.06 & -44672.34 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:log(sl) & 89394.69 & 1.81 & 0.06 & -44670.35 \\
-          \textbf{Base Model} + sl:LA + WA:sl + log(sl):cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:cos(ta) & 89394.84 & 1.96 & 0.06 & -44670.42 \\
-          \hdashline
-          \\
-          \vdots & \vdots & \vdots & \vdots & \vdots \\
-          \\
-          \hdashline
-          \\
-          \textbf{Base Model}: cos(ta) + sl + log(sl) + WA + WO + DTW + HI + SH & 90091.40 & 787.67 & 0.00 & -45030.70\\
-          \\
-         \bottomrule
-       \end{tabular}
-       \begin{tablenotes}
-         \item \textit{Note:} ta = Turning Angle, sl = Step Length, LA = Low Activity, WA = Water,
-         DTW = Distance To Water, SH = Shrubs/Grassland, WO = Woodland, HI =
-         Human Influence.
-       \end{tablenotes}
-     \end{threeparttable}
-    }
-  \end{center}
-\end{table}
-
-% \begin{figure}[htpb]
-%   \begin{center}
-%     \includegraphics[width = \textwidth]{99_MovementModelAICs.png}
-%     \caption{Coefficients of all movement models with an AIC weight above zero.}
-%     \label{AllMovementModels}
-%   \end{center}
-% \end{figure}
-
-%------------------------------------------------------------------------------
-%	Appendix S2: Evolution of Heatmaps
-%------------------------------------------------------------------------------
-\newpage
-\newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
-\section{Evolution of Heatmaps}
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width = 0.88\textwidth]{99_HeatmapsIndividual.png}
-    \caption{Heatmaps produced after 125, 500, and 2000 simulated steps,
-    respectively. The left panel (a1, a2, a3) was generated based on simulations
-    initiated within the main study area, the right panel (b1, b2, b3) was
-    generated based on simulations initiated within the buffer area. To produce
-    the heatmap presented in the main text, we summed up values from maps a3 and
-    b3.}
-    \label{Trajectories}
+  \includegraphics[width = \textwidth]{Figures/MovementModel.png}
+  \caption{Model parameters from the step-selection model implemented by
+  \citet{Hofmann.2023}. The model was fit to GPS data of dispersing African wild
+  dogs and comprises of a habitat kernel (light blue band), a movement kernel
+  (dark blue band), and their interactions (purple band). Abbreviations are as
+  follows: sl = step-length, ln(sl) = natural logarithm of the step-length,
+  cos(ta) = cosine of the relative turning angle.}
+  \label{Model}
   \end{center}
 \end{figure}
-\restoregeometry
 
-%------------------------------------------------------------------------------
-%	Appendix S3: Evolution of Betweenness Maps
-%------------------------------------------------------------------------------
 \newpage
-\newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
-\section{Evolution of Betweenness}
-\begin{figure}[hbtp]
+\section{Immigration \& Emigration by Source Area}
+\begin{figure}[htbp]
  \begin{center}
-    \includegraphics[width = 0.88\textwidth]{99_BetweennessIndividual.png}
-    \caption{Maps of betweenness scores produced after 125 (a), 500 (b), and
-    2000 (c) simulated steps, respectively. A high betweenness score indicates
-    that the respective area is paramount for linking other regions in the study
-    area. It therefore highlights movement corridors that provide access to
-    various destinations.}
-    \label{Trajectories}
+  \includegraphics[width = \textwidth]{Figures/ImmigrationEmigration.png}
+  \caption{Number of individuals emigrating from, or immigrating into a specific
+  source area (focal area). Colors indicate into which other areas emigrants
+  moved or from which other areas immigrants originate. For instance, the most
+  left plot in the upper panel shows the number of individuals moving from
+  source area 1 into the six other source areas during minimum and maximum
+  flood, respectively.}
+  \label{EmigrationImmigration}
  \end{center}
 \end{figure}
-\restoregeometry
+
+\newpage
+\section{Source-Specific Inter-Patch Connectivity}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/IPCMain.png}
+  \caption{Spatial representation of inter-patch connectivity derived for each
+  source area separately across the two extreme flood-scenarios. The focal
+  source area of each subfigure is highlighted by a red circle. Subfigure (a),
+  for instance, depicts inter-patch connectivity for source area 1.}
+  \label{IPCMain}
+  \end{center}
+\end{figure}
+
+\newpage
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/IPCBuffer.png}
+  \caption{Spatial representation of egression patterns derived for each
+  source area separately across the two extreme flood-scenarios. The focal
+  source area of each subfigure is highlighted by a red circle. Subfigure (a),
+  for instance, depicts the number of individuals egressing from source area
+  1.}
+  \label{IPCBuffer}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Dispersal into Egression Zones}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/Egression.png}
+  \caption{Absolute number of simulated trajectories running into each of the
+  designated egression zones (purple) during minimum and maximum flood.}
+  \label{Egression}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Source-Specific Intensity of Use}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/HeatmapsIndividual.png}
+  \caption{Heatmaps showing the intensity of use for each source area separately
+  across the two extreme flood-scenarios. The focal source area of each
+  subfigure is highlighted by a red circle. Subfigure (a), for instance, depicts
+  the heatmaps for source area 1.}
+  \label{Heatmaps}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Source-Specific Betweenness}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/BetweennessIndividual.png}
+  \caption{Betweenness maps prepared for each source area separately across the
+  two extreme flood-scenarios. The focal source area of each subfigure is
+  highlighted by a red circle. Subfigure (a), for instance, depicts the
+  betweenness maps for source area 1.}
+  \label{Betweenness}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Source-Specific Human-Wildlife Conflict}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/HumanWildlifeConflictIndividual.png}
+  \caption{Human wildlife conflict maps prepared for each source area separately
+  across the two extreme flood-scenarios. The focal source area of each
+  subfigure is highlighted by a red circle. Subfigure (a), for instance, depicts
+  the human wildlife-conflict maps for source area 1. Dotted shapes were used to
+  compare human-wildlife conflict within specific areas (see also Figure S9).}
+  \label{HWC}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Changes in Potential for Human-Wildlife Interactions}
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/HWCDifferenceAOI.png}
+  \caption{Number of simulated trajectories within the vicinity of
+  human-dominated landcapes in different areas of interest across the minimum
+  and maximum flood scenarios. The areas are represented by the black dotted
+  shapes in Figure S8}
+  \label{HWCDifference}
+  \end{center}
+\end{figure}
+
+\newpage
+\section{Alternative Approach to Map Potential Human-Wildlife Interactions}
+
+To identify potential hotspots for human-wildlife conflict and generate Figure
+6c, we isolated all simulated animal locations within 500 meters to a
+human-influenced grid-cell (Figure 6c in the main manuscript). To compute the
+needed distances, we assumed human-influence to be binary (influence = 1, no
+influence = 0), thus ignoring potential impacts of human density. We argue that
+the severity of human-wildlife conflict is not necessarily related to human
+density, yet humans' attitude towards wildlife. Attitude often correlates
+negatively with habitat suitability for the species of interest
+\citep{Behr.2017}, and so conflict is often pronounced in peripheral areas
+\citep{McNutt.2017}. Since we lacked detailed information of anthropogenic
+resistance \citep{Ghoddousi.2021} across the study area, we deemed a binary
+representation of human impacts as appropriate. Alternatively, one can also
+compute a compound score by multiplying the human-influence layer with the
+heatmaps derived from simulated dispersal. This is presented in
+\Cref{HWCAlternative}, where we multiplied the heatmaps (\Cref{HWCAlternative}a)
+with the human-influence layer (\Cref{HWCAlternative}b) and produced maps
+showing potential for human-wildlife conflict. Qualitatively, the maps in
+\Cref{HWCAlternative} are very similar to the ones presented in Figure 6c.
+
+\begin{figure}[htbp]
+  \begin{center}
+  \includegraphics[width = \textwidth]{Figures/HumanWildlifeConflictAlternative.png}
+  \caption{Alternative approach to quantifying the potential for human-wildlife
+  conflict. Here, we multiplied the heatmaps (a) with the human-influence layers
+  (b) to obtain human-wildlife conflict maps (c1). We also computed a difference
+  map (c2) for the layers shown in (c1).}
+  \label{HWCAlternative}
+  \end{center}
+\end{figure}
 

### 2024-07-01 Chapter_3/04_Manuscript/Manuscript.tex
Publish PHD

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage{scalerel}
\usepackage{makecell}
\usepackage{ifthen}
\usepackage{tikz}
\usepackage{lineno}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%  Bibtex vs Biblatex
%------------------------------------------------------------------------------
% While bibtex is very nice to use, it's insanely slow. I therefore work with
% bibtex until the final stage. Set the value below to "true" or "false" to pick
% which one should be used.
\newboolean{usebiblatex}
\setboolean{usebiblatex}{false}

% Depending on the question, the commands below will prepare the document
% accordingly
\ifthenelse{\boolean{usebiblatex}}{
  %-----------------------------------------------------------------------------
  %  If using biblatex
  %-----------------------------------------------------------------------------
  % Load required packages and refer to correct bibliography
  \usepackage{csquotes}
  \usepackage[
      backend=biber
    , style=apa          % To display Author-year in the text
    , natbib=true        % Necessary to use citep, etc.
    , hyperref=true      % To have clickable links in the references
    , useeditor=false    % Don't add editor to publications
    , sorting=ynt        % Sort by year, name, and title
    , uniquename=false   % Avoid that initials are put
  ]{biblatex}
  \addbibresource{Literature.bib}

  % To include author in link
  \DeclareFieldFormat{citehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{#1}}
  \DeclareFieldFormat{textcitehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{%
    #1%
    \ifbool{cbx:parens}
      {\bibcloseparen\global\boolfalse{cbx:parens}}
      {}}}
  \savebibmacro{cite}
  \savebibmacro{textcite}
  \renewbibmacro*{cite}{%
    \printtext[citehyperref]{%
      \restorebibmacro{cite}%
      \usebibmacro{cite}}}
  \renewbibmacro*{textcite}{%
    \ifboolexpr{
      ( not test {\iffieldundef{prenote}} and
        test {\ifnumequal{\value{citecount}}{1}} )
      or
      ( not test {\iffieldundef{postnote}} and
        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
    }
      {\DeclareFieldAlias{textcitehyperref}{noformat}}
      {}%
    \printtext[textcitehyperref]{%
      \restorebibmacro{textcite}%
      \usebibmacro{textcite}}}

  % Remove editor form everything but books
  \AtEveryBibitem{%
   \ifentrytype{book}{}{%
    \clearname{editor}
   }
  }
}{
  %-----------------------------------------------------------------------------
  %  If using bibtex
  %-----------------------------------------------------------------------------
  % Load required packages and select citation style
  \usepackage[round]{natbib}
  \bibliographystyle{apalike}
}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Link colors
\definecolor{linkcolor}{HTML}{000000}
\hypersetup{
  colorlinks = true,
  linkcolor  = linkcolor,
  urlcolor   = linkcolor,
  citecolor  = linkcolor,
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% Avoid skip when including text from external .tex file
\newcommand{\inputy}[1]{\input{#1}\unskip}

% For supplementary material
\newcommand{\beginappendix}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
  \setcounter{equation}{0}
  \renewcommand{\theequation}{Equation S\arabic{equation}}%
  \setcounter{section}{0}
  \renewcommand{\thesection}{A.\arabic{section}}%
}

%------------------------------------------------------------------------------
%  ORCID
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Create the ORCID logo
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol]
      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
        ,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white]
      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

% Command to create the ORCID
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{The Effects of Increasing Seasonal Dynamism when Predicting Connectivity:
Advantages or Unnecessary Complications?}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1, 2} \orcid{0000-0001-7477-2642} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
Private Bag 13, Maun, Botswana.

\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}

\vspace{4cm}

\textbf{Running Title:} Connectivity in Seasonal Landscapes

\vspace{0.5cm}

\textbf{Keywords:} African wild dog, Connectivity, Dispersal, Dynamism,
Individual-based, Lycaon pictus, Okavango Delta, Seasonality

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}

Seasonally changing conditions can drastically alter landscape connectivity.
Nevertheless, most connectivity studies ignore seasonal dynamism and instead
employ a static set of spatial covariates and assume their focal species to
exhibit a static set of preferences. Ignoring seasonality may, however, mask
important ecological features and processes, thus resulting in poor agreement
between predicted and observed movements and a misrepresentation of
connectivity.

We present a simple framework highlighting that seasonality may enter a
connectivity analysis at three distinct stages, namely when (1) extracting
spatial covariates for model fitting, (2) when fitting the selection model, and
(3) when making predictions from the fitted model. In combination, this provides
six possible configurations that differ in terms of the seasonal dynamism they
encapsulate.

Capitalizing on natural seasonal fluctuations of the Okavango Delta in northern
Botswana and on GPS data collected on dispersing African wild dogs
(\textit{Lycaon pictus}) across different seasons, we investigate the degree to
which a better representation of seasonal dynamism improves our ability to
predict connectivity. For this, we fit integrated step-selection functions and
predict connectivity using an individual-based dispersal simulation while
explicitly considering seasonal dynamism in both environmental covariates and
the species' preferences.  Using a rigorous cross-validation procedure, we
compare the predictive model performance under each of the six proposed
configurations. While we expected that an increasing degree of seasonal dynamism
would lead to improved predictions, we were particularly interested in
identifying at which stage the inclusion of seasonality provides the biggest
benefits.

We show that, for our study system, improvements in predictive performance by
incorporating seasonal dynamism were moderate. In fact, incorporating
seasonality only improved predictions when an overly simplistic movement model
was assumed. Upon fitting a more complex model, the benefits of accounting for
seasonality vanished, resulting in imperceptible performance differences.
Despite this, patterns of connectivity as obtained from dispersal simulations
revealed marked differences between the most static and most dynamic
configurations. Most notably, connectivity was more homogeneously distributed
throughout the study area when seasonality was taken into account, suggesting
the existence of seasonal stepping stones that facilitate dispersal into
otherwise inaccessible areas.

Besides a better understanding of the importance of dynamic connectivity, our
results also provide insights into the conservation needs of the endangered
African wild dog.

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
\linenumbers
\section{Introduction}
\subsection{Connectivity}

Landscape connectivity is defined as the degree to which the landscape
facilitates or impedes movement among habitat patches \citep{Taylor.1993} and is
a critical prerequisite for maintaining biodiversity \citep{Fahrig.2003}.
Improved connectivity facilitates dispersal \citep{Doerr.2011, Baguette.2013},
which in turn promotes genetic diversity \citep{Perrin.2000, Frankham.2002} and
the colonization of vacant habitats \citep{Hanski.1999, MacArthur.2001}. Due to
its beneficial impacts on metapopulation dynamics, restoring connectivity is
among the most frequently recommended strategies to conserve biodiversity and to
promote resilience against climate change \citep{Heller.2009, Rudnick.2012}.
Quantifying connectivity and identifying critical dispersal corridors have
therefore become critical tasks in conservation science \citep{Heller.2009,
Rudnick.2012, Keeley.2019, Hofmann.2021}. One aspect that has received limited
attention in the study of connectivity is the role of seasonality. Yet, seasonal
changes both in the environment and in an animal's propensity to use or avoid
particular habitat types can profoundly impact landscape connectivity
\citep{Zeller.2020a}.

\subsection{Seasonality}

Seasonality can impact functional connectivity through spatio-temporal variation
in the landscape itself, or through temporal variation in species' preferences
towards prevailing conditions \citep{Mui.2017, Simpkins.2017a, Zeller.2020a}. In
ecosystems that experience alternations between wet and dry seasons, for
example, the onset of the rainy season initiates distinct ``green-up'' waves,
which affect the availability of food resources for herbivores and,
subsequently, shape their movements \citep{Merkle.2016}. In its most remarkable
form, the variation in environmental conditions drives herbivore migrations
across massive spatial scales, resulting in short-lived movement corridors
between two otherwise disconnected habitats \citep{Serneels.2001, Naidoo.2016}.
Alternatively, seasonality can affect functional connectivity via temporal
changes in a species' movement and habitat preferences. Amphibians, for
instance, require both aquatic and terrestrial habitats, but their preference
for one over the other heavily depends on the season \citep{Baldwin.2006}.
Although such seasonal intricacies are likely to play a fundamental role in many
ecosystems, they only rarely enter connectivity studies in an explicit manner.
In fact, most connectivity studies represent their study system by a static set
of environmental layers and assume that their focal species exhibits a fixed set
of preferences (e.g., \citealp{Elliot.2014, Abrahms.2017, Brennan.2020}).
However, this may result in biased connectivity estimates and a misallocation of
scarce conservation funds \citep{Osipova.2019, Zeller.2020}. Therefore, a more
dynamic approach to connectivity that acknowledges and renders seasonal
variation has been recommended \citep{Zeller.2020a}.

\subsection{Modeling Functional Connectivity}

Functional connectivity can be estimated using a variety of modeling techniques
\citep{Diniz.2019}, which all comprise four main steps. First, presence data of
the focal species, preferably collected during dispersal (\citealp{Elliot.2014,
Vasudev.2015, Benz.2016}, but see \citealp{Fattebert.2015}), and a set of
spatial covariate layers that are believed or known to be critical determinants
of connectivity are compiled. Second, these data are combined and fed into a
selection model that enables estimating a species selection or avoidance of
environmental features. Popular frameworks for estimating preferences are
case-control designs, where characteristics extracted at observed locations are
contrasted with characteristics extracted at available locations
\citep{Beyer.2010, Fieberg.2010}. This can be achieved using point-selection
functions \citep{Boyce.2002, Manly.2007}, path-selection functions
\citep{Cushman.2010}, and step-selection functions \citep{Fortin.2005,
Thurfjell.2014}. A particularly powerful approach is that of \textit{integrated}
step-selection functions (iSSFs), as it provides an effective means to model
both an animal's habitat-selection and movement capacity \citep{Avgar.2016,
Fieberg.2021}. Third, inferred preferences are used to predict a permeability
surface, which indicates the expected ease or difficulty at which the focal
species can traverse a certain area given the area's environmental
characteristics \citep{Zeller.2012}. Finally, in a fourth step, the permeability
surface serves as an input to a connectivity model that reveals crucial movement
corridors. At present, the most popular connectivity models are least-cost path
analysis \citep{Adriaensen.2003} and circuit theory \citep{McRae.2008}, although
individual-based movement models (IBMMs) have gained some momentum recently
\citep{Kanagaraj.2013, Allen.2016, Hauenstein.2019, Zeller.2020,
UnnithanKumar.2022a, UnnithanKumar.2022, Hofmann.2023}. In particular, some
IBMMs allow estimating connectivity directly via simulated dispersal
trajectories, thus bypassing the generation of a permeability surface
\citep{UnnithanKumar.2022a, Hofmann.2023}.

\subsection{Incorporating Seasonality in Functional Connectivity Models}

Seasonality can enter the connectivity modeling workflow described above in
three distinct stages (\Cref{GraphicalAbstract}). In the first stage, one can
either extract environmental features, such as water and vegetation, from a
static layer (i.e., a single snapshot) or from a time series of layers (i.e., a
sequence of snapshots) that capture seasonal variation across the landscape. The
former approach has historically been the norm (e.g., \citealp{Elliot.2014,
Brennan.2020}), yet advances in remote sensing technologies and a facilitated
access to petabytes of landscape data have opened up new avenues for obtaining
spatial layers at unprecedented spatio-temporal resolutions \citep{Toth.2016,
Rumiano.2020}, such that the representation of study systems via dynamic
covariate layers has become more frequent (e.g., \citealp{Osipova.2019,
Kaszta.2021}). In a second stage, one can assume their focal species to exhibit
a fixed set of preferences across seasons by pooling all presence data, or can
try to account for seasonal changes in preferences by splitting the data
accordingly (e.g., \citealp{Fortin.2005, Manly.2007, Cushman.2010,
Zeller.2020}). \cite{Chetkiewicz.2009}, for instance, partitioned their data by
season to derive seasonal habitat preferences for pumas (\textit{Puma concolor})
and grizzlies (\textit{Ursus arctos}). In a final stage, one can estimate
connectivity for either ``average'' environmental conditions, or utilize
seasonally updated layers to estimate connectivity for distinct seasons. For
connectivity models that rely on permeability surfaces, this implies repeatedly
applying the connectivity model using seasonally updated resistance surfaces
(e.g. \citealp{Osipova.2019, Zeller.2020, Kaszta.2021, Ciudad.2021}). This is
demonstrated \textit{ad extremum} by \citet{Kaszta.2021}, who prepared monthly
updated permeability surfaces for African elephants (\textit{Loxodonta
africana}). The above-mentioned IBMMs allow for an elegant solution, for
seasonality can be accounted for as simulated individuals move. Such an approach
has, however, not yet been followed. Irrespective of the method used, including
seasonality is analytically and computationally  demanding
\citep{Bishop-Taylor.2018}, raising the question to what degree it should be
considered. Our goal is (1) to create a framework with the possible combinations
in which seasonality can be accounted for, (2) to establish whether
incorporating seasonality benefits the predictive performance of connectivity
analyses, and (3) to pinpoint at which stages the inclusion of seasonality
provides the largest benefit.

\subsection{African wild dogs in Northern Botswana}

A system well-suited to study the importance of seasonal dynamism for dispersal
and connectivity is the African wild dog population (\textit{Lycaon pictus})
inhabiting the seasonally highly variable Okavango Delta ecosystem in northern
Botswana \citep{McNutt.1996, Wolski.2017}. While once present across the entire
Sub-Saharan continent, the African wild dog has disappeared from a majority of
its historic range due to human persecution, deadly diseases, and habitat
destruction \citep{Woodroffe.2020a}. With about 6,000 adult individuals
remaining in the wild, the species is considered as endangered on the IUCN red
list. Wild dogs are pack-living carnivores, primarily active during the cooler
morning and evening hours \citep{Rasmussen.2012} or during moonlit nights
\citep{Cozzi.2012}. Higher ambient temperature is associated with shorter
activity periods, as the species usually rests during times of elevated heat
\citep{Rabaiotti.2019}. Upon reaching sexual maturity, individuals born into a
pack disperse in single-sex coalitions to find suitable mates and a territory to
settle \citep{McNutt.1996}. In Botswana, the timing of dispersal is seasonal,
with female dispersal peaking prior to the mating season in March, and male
dispersal peaking at the onset of the rainy season in December
\citep{Behr.2020}. Euclidean distances moved by dispersers range from 5 km to
500 km, with some coalitions covering several hundred kilometers within only a
few days \citep{Davies-Mostert.2012, Masenga.2016, Cozzi.2020,
Sandoval-Seres.2022}. Studies of habitat-selection during dispersal show that
wild dogs avoid water, prefer moving along it, prefer moving across open grass
or shrubs, yet avoid areas dominated by humans and densely covered by forests
\citep{ONeill.2020, Hofmann.2021}.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/GraphicalAbstract.pdf}
  \caption{Overview of the different dimensions in which seasonality can be
  rendered in studies of landscape connectivity. (a) During model fitting, the
  modeler needs to decide whether to represent the environment by a static set
  of covariate layers, thus ignoring seasonality, or whether to obtain a dynamic
  set covariate layers that allow rendering it. One also needs to decide
  whether to parameterize a single-season model, assuming fixed preferences
  across the year, or to engage in a multi-season model that accounts for
  seasonal differences. (b) When utilizing the fitted model to predict
  connectivity, one can either assume a static set of environmental covariates
  or again attempt a seasonal take that renders how connectivity differs
  depending on the season. (c) Depending on these decisions, six different
  combinations with differing degrees of dynamism emerge. Our hypothesis was
  that increasing the degree of dynamism would lead to improved predictions.
  However, we were particularly interested in determining at which stage the
  inclusion of seasonally provided the biggest benefits.}
  \label{GraphicalAbstract}
 \end{center}
\end{figure}

\subsection{What We Did}

Here, we examined the importance of including seasonally-varying factors to
assess landscape connectivity for the African wild dog. For this, we compared
the predictive performance of iSSFs that differ in terms of the dynamism they
represented. Specifically, we investigate if and to what degree seasonality at
different stages of the connectivity modeling workflow contributes to improved
predictions. For this, we compile an extensive collection of remote sensed time
series data that accurately render seasonality across the Okavango Delta
ecosystem. We combine them with multi-year dispersal data from
\inputy{GeneralMetrics/CollarsTotal} dispersing wild dog coalitions in northern
Botswana and apply k-fold cross-validation for case-control studies to compare
the predictive efficacy of each model. Finally, we employ IBMMs to simulate
dispersal and estimate connectivity resulting from varying degrees of dynamism.
We hypothesized that habitat-selection and movement behavior would differ
significantly between seasons and that increasing the degree of dynamism would
result in better agreement between observed and predicted dispersal patterns
(\Cref{GraphicalAbstract}).

\section{Methods}

We used the \texttt{R} programming language \citep{RCoreTeam.2023} for all data
preparation and analyses. We performed spatial data manipulation using the
\texttt{terra} \citep{Hijmans.2024} and \texttt{spatstat} \citep{Baddeley.2015}
packages. We generated figures using the \texttt{ggplot2} \citep{Wickham.2024}
and \texttt{ggpubr} \citep{Kassambara.2024} packages. To ensure reproducibility
of all our analyses, we provide access to our \texttt{R}-scripts through an
online repository upon publication of this article.

\subsection{Study Area}

The study area comprised the Okavango Delta ecosystem, mainly within northern
Botswana (centered at \inputy{GeneralMetrics/StudyAreaCenter.tex} at an
elevation of approx. 950 m) but also extended to parts of Namibia and Zimbabwe,
and stretched across an extent of \inputy{GeneralMetrics/SizeStudyArea.tex}
km\textsuperscript{2} (\Cref{StudyArea}). The Okavango Delta is a flood-pulse
driven mosaic of patchy woodlands, permanent swamps, and seasonally flooded
grasslands that lie within the otherwise dry and sandy Kalahari Basin
\citep{Wilson.1976, Ramberg.2006, Mendelsohn.2010}. Precipitation across the
study area varies considerably between seasons, ranging from
\inputy{GeneralMetrics/MinimumPrecipitation.tex} mm during the dry season (from
$\sim$ 15 April to 15 October) to
\inputy{GeneralMetrics/MaximumPrecipitation.tex} mm during the wet season (from
$\sim$ 15 October to 15 April), totaling to
\inputy{GeneralMetrics/TotalPrecipitation.tex} mm across an average year
(\Cref{Seasonality}a). Daily maximum above-ground temperature fluctuates between
\inputy{GeneralMetrics/MinimumTemperature.tex}\degree C during the dry winter
months to \inputy{GeneralMetrics/MaximumTemperature.tex}\degree C during the wet
summer months (\Cref{Seasonality}b). The vegetation in the study area is mainly
composed of mopane forest (\textit{Colophospermum mopane}), mixed woodland
acacia-dominated (\textit{Acacia spp.}), and grassland. Substantial vegetation
green-up (e.g., plant and shoots growth, leaf production, grass greening) after
the dry season starts with a delay of some weeks after the onset of the first
rains in the wet season. The normalized difference vegetation index (NDVI)
therefore depicts a lagged response to precipitation patterns across the study
area (\Cref{Seasonality}c). The yearly flood-cycle of the Delta is predominantly
driven by rainfalls in the Angolan highlands, where water is collected and
channeled through the Cubango and Cuito rivers into the Okavango Delta
\citep{McCarthy.2003, Gumbricht.2004, Mendelsohn.2010}. Because water only
slowly descends from the catchment areas in Angola into the Delta's tributaries,
the flood is out of sync with local rainfalls and reaches its maximum extent
during August-September, i.e. during peak dry season (\citealp{Wolski.2017},
\Cref{Seasonality}d). While the extent of large-bodied rivers and floodplains is
determined by precipitation in Angola, the emergence of smaller, ephemeral
water-bodies (a.k.a., pans) is dictated by local precipitation during the wet
season. \inputy{GeneralMetrics/PercentageProtected.tex}\% of the landscape in
the study area form part of a protected area, such that human impact remains low
and largely limited to settlements along the western part of the delta and the
city of Maun at the delta's southern tip (\Cref{StudyArea}). Landscapes outside
protected areas in Zimbabwe, however, are more heavily influenced by humans,
mainly through agricultural fields and human settlements.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/StudyArea.png} \caption{(a) Study
  area from which data on dispersing wild dogs were collected. Dispersal
  trajectories are plotted in dark gray. The study area encompassed parts of the
  Okavango Delta in northern Botswana, a highly dynamic, flood-pulse-driven
  ecosystem. The entire study area undergoes substantial seasonal changes, as
  can be seen from two satellite images taken during peak dry season (b1) and
  peak rainy season (b2). Notably, the flood of the Okavango Delta reaches its
  maximum extent during peak dry season (August–September).}
  \label{StudyArea}
 \end{center}
\end{figure}

\begin{figure}
\begin{center}
  \includegraphics[width = \textwidth]{Figures/SeasonalCovariates.png}
  \caption{Illustration of how some covariates considered in this study vary
  across seasons. The wet season spans mid-October to mid-April (shaded in
  gray). Data for the graphs were obtained from \textbf{(a)} JAXA GSMaP,
  \textbf{(b)} ERA5, \textbf{(c)} MODIS MOD13Q1, and \textbf{(d)} remote sensed
  MOD43A4 satellite images. Smoothing curves were fitted using GAMs as
  implemented in the \texttt{mgcv} \texttt{R}-package \citep{Wood.2011}.}
  \label{Seasonality}
\end{center}
\end{figure}

\subsection{GPS Data}

Between \inputy{GeneralMetrics/GPSFromYear} and
\inputy{GeneralMetrics/GPSToYear}, we collected GPS data of
\inputy{GeneralMetrics/CollarsTotal} dispersing wild dog coalitions
(\inputy{GeneralMetrics/CollarsFemales} female coalitions,
\inputy{GeneralMetrics/CollarsMales} male coalitions \Cref{StudyArea}b). We
programmed GPS satellite collars to record GPS locations at 17:00, 21:00, 01:00,
05:00, and 09:00 o'clock. The eight-hour window between 09:00 and 17:00 can be
considered comparable to a four-hour window, as wild dogs generally rest between
11:00 and 15:00, leaving approximately four hours of activity
\citep{Hayward.2009}. The collected GPS data was regularly transmitted to a
base-station via Iridium satellite, thereby allowing dispersing individuals to
be monitored, even if they ventured across national borders. In total, we
obtained \inputy{GeneralMetrics/FixesTotal} locations during dispersal, with an
average of \inputy{GeneralMetrics/FixesMeanSD} locations per coalition.
Occasionally, the acquisition of a GPS location failed (success rate =
\inputy{GeneralMetrics/AcquisitionRate}\%), resulting in slight deviations from
the aspired four-hourly schedule. Further details on the GPS collar fitting
procedure and how we distinguished between dispersal and resident movements can
be found in \citet{Cozzi.2020} and \citet{Hofmann.2021}.

\subsection{Covariates}

We represented the physical landscape through which dispersers moved by a suite
of spatial covariate layers known to influence wild dog movements during
dispersal. We broadly categorized these spatial covariates into descriptors of
(1) landscape characteristics, (2) climatic conditions, and (3) anthropogenic
factors (\Cref{Covariates}) (see \citealp{Hofmann.2021, Hofmann.2023}). Besides
spatial covariates, we also prepared a series of covariates relating to (4)
light availability (\Cref{Covariates}) (see Cozzi et al. 2012). To appropriately
render seasonality in each of the spatial covariates (1-3), we downloaded them
at the highest spatial and temporal resolution available. That is, we obtained
for each spatial covariates by a stack of raster layers that spanned the entire
range of dates and locations for which we also collected GPS data of dispersing
wild dogs. Depending on the covariate, this resulted in differing spatial and
temporal resolutions (\Cref{Covariates}). Additionally, for each spatial
covariate, we also generated a static layer, representing ``average'' conditions
across the entire duration of the study. For this, we flattened each covariate
stack into a single layer, thus removing seasonality from the spatial data
entirely. For continuous covariates, we achieved this by averaging conditions
across all collected layers, whereas for categorical (binary) layers we
identified areas that were covered by the respective category in at least 50\%
of all layers. Using the same aggregation techniques, we computed covariate
stacks representative of a typical year. That is, instead of removing
seasonality by flattening across the entire range of dates, we flattened stacks
across years, thereby eliminating year-specific effects. These final layers
served to be used in the dispersal simulation, where we needed data representing
seasonality in all covariates across a typical year. To summarize, we prepared
each spatial covariate dynamically for the entire range of dates considered,
dynamically for an average year, and statically across the entire range of dates
considered.

\onehalfspacing
\begin{table}[htbp]
 \begin{center}
    \caption{Covariates that were used in this study, including information on
    their temporal and spatial resolutions, as well as on the avenue through which
    the respective data were accessed or downloaded.}
    \label{Covariates}
   \begin{threeparttable}
    \begin{tabular}{lcr}
      \input{Figures/Covariates.tex}
    \end{tabular}
    \begin{tablenotes}
      \item \textit{Note:} The covariates in gray were combined into proxies for
      water, human influence, and brightness, respectively. The detailed
      aggregation procedure is described by \citet{Hofmann.2021}.
    \end{tablenotes}
   \end{threeparttable}
 \end{center}
\end{table}
\doublespacing

\subsubsection{Landscape Characteristics}

We used data from the MODIS Vegetation Continuous Fields dataset (MOD44B V061;
\citealp{DiMiceli.2022}) to represent different vegetation types across the
study area. The MOD44B dataset comprises three continuous layers, depicting the
percentage cover of woodland, shrubs/grassland, and bareland, respectively. The
three layers added up to 100\%, so we dropped bareland from further analysis,
thus preventing perfect multi-collinearity. The MOD44B product is updated on day
65 of each year, so we used the \texttt{R}-package \texttt{RGISTools}
\citep{Perez-Goya.2020} to download yearly updated layers, each at a resolution
of 250m x 250m. We also obtained information on the normalized vegetation
difference index (NDVI) through the MODIS MOD13Q1 dataset \citep{Didan.2015},
which also has a resolution of 250m x 250m. This product is updated every 16
days, and we accessed the respective data through Google Earth Engine
\citep{Gorelick.2017} using the \texttt{R}-package \texttt{rgee}
\citep{Aybar.2024}. To depict large, permanent water-bodies, we employed the
Globeland30 dataset, from which we only retained the land-cover class water,
while setting all other categories to dryland. Similarly, we used the MERIT
Hydro dataset to obtain information on permanent rivers \citep{Yamazaki.2019}.
To dynamically render large water-bodies, particularly the floodwaters of the
Okavango Delta, we prepared weekly updated floodmaps using remote sensed MODIS
MOD43A4 satellite images. The underlying floodmapping algorithm is described in
detail in \citet{Wolski.2017} and \citet{Hofmann.2021} and is implemented in the
\texttt{floodmapr} package (available on GitHub;
\url{https://github.com/DavidDHofmann/floodmapr}). To generate a single layer
representing major water bodies, we combined the water, river, and flood layers
into a single stack with a resolution of 500m x 500m. Finally, we employed
remote sensing to detect small, ephemeral water bodies (i.e., pans) using a
custom random-forest classifier applied to Sentinel 2 satellite imagery
(\citealp{EuropeanSpaceAgency.2018}; details of the classifier in Appendix A1).
Sentinel 2 has a resolution of 10m and is therefore particularly useful for
obtaining information on small landscape features. Even though Sentinel 2
satellite imagery is updated every 5 days, cloud cover often prohibited the
computation of a ``pan-map''. Consequently, we settled for monthly updated
composite images, which effectively alleviated problems due to cloud cover. In
summary, we produced one stack of layers representing major water bodies, and
another stack of layers representing ephemeral water bodies. For both, we also
computed corresponding distance-to stacks.

\subsubsection{Climate Descriptors}

We obtained hourly updated spatial layers on 2m above-ground temperature from
the ERA5-Land dataset \citep{Munoz-Sabater.2021} and hourly updated
precipitation estimates from the Global Satellite Mapping of Precipitation
dataset \citep{Kubota.2020}. Both datasets were accessed and downloaded through
Google Earth Engine \citep{Gorelick.2017} using the \texttt{rgee} package
\citep{Aybar.2024} and had a resolution of 10km x 10km. To match hourly
temperature and precipitation values with the four-hourly data collected by the
GPS satellite collars fitted on the dispersing wild dogs, we computed average
precipitation and temperature values over four hourly periods that matched the
GPS collection schedule.

\subsubsection{Anthropogenic Features}

We combined information on human density, agricultural activities, and roads
into a single proxy, which we generically termed human influence. We sourced
information on human density from Facebook's high resolution human density
dataset \citep{Tiecke.2017}, which we downloaded from the humdata website
(\url{www.data.humdata.org/}). We obtained information on the presence of
agricultural fields from the Globeland30 \citep{Chen.2015} and Cropland
\citep{Xiong.2017} datasets. We  downloaded shapefiles comprising main tar roads
from OpenStreetMaps \citep{OpenStreetMapContributors.2017}. Ultimately, we
merged all anthropogenic features into a single layer (details in
\citealp{Hofmann.2021}) that had a resolution of 250m x 250m.

\subsubsection{Light Availability}

We computed light statistics using the \texttt{suncalc} and \texttt{moonlit}
\texttt{R}-packages \citep{Thieurmel.2022, Smielak.2023} for the central
coordinates of our study area at a 5-minute temporal resolution. The set of
light statistics comprised a binary variable separating day and night (i.e., sun
$<$ -18 \degree below the horizon) and a continuous estimate of moonlight
illumination (relative to the maximum moon illumination). Based on those
covariates, we generated a binary covariate separating bright from dark
conditions. To provide a general understanding of this variable, bright
conditions encompassed all daytime hours and those nighttime periods during
which the moon was present in the sky and illuminated by approximately
one-fourth; conversely, dark conditions included nighttime periods during which
the moon was absent from the sky or present or was only minimally illuminated
(further details in Appendices A2 and A3).

\subsection{Step-Selection Models}

We modeled habitat selection and movement behavior of dispersing wild dogs using
integrated step-selection functions (iSSFs, \citealp{Fortin.2005, Avgar.2016}),
following the procedure described in \citet{Muff.2020}. For this, we identified
bursts of subsequent GPS locations where the duration between two GPS locations
did not exceed 4 hours (\(\pm\) 15 minutes) or 8 hours (\(\pm\) 30 minutes, for
data recorded between 09:00 and 17:00). Within each burst, we converted
locations into steps, where a step represented the straight line segment between
two consecutive locations \citep{Turchin.1998}. For each step, we computed the
associated step length (sl, in meters) and relative turning angle (ta, in
radians). After this pre-processing, a total of \inputy{GeneralMetrics/SSFTotal}
dispersing coalitions (\inputy{GeneralMetrics/SSFFemales} female coalitions,
\inputy{GeneralMetrics/SSFMales} male coalitions) remained for further analyses.
The final dataset comprised \inputy{GeneralMetrics/StepsTotal} steps
(\inputy{GeneralMetrics/StepsMeanSD} per coalition), which we further
categorized into wet (15 October to 15 April) and dry (15 April to 15 October)
season, resulting in \inputy{GeneralMetrics/NumberStepsDry} steps
(\inputy{GeneralMetrics/PercentageStepsDry}\%) during the dry season and
\inputy{GeneralMetrics/NumberStepsWet} steps
(\inputy{GeneralMetrics/PercentageStepsWet}\%) during the wet season.

We paired each \textit{observed} step with a set of 100 \textit{random} steps,
generated by sampling turning angles from  a uniform distribution U(\(-\pi,
+\pi\)) and step lengths from a gamma distribution fitted to observed steps
(scale \(\theta\) = \inputy{GeneralMetrics/GammaScale} and shape \(k\) =
\inputy{GeneralMetrics/GammaShape}). Together, an observed step and its
associated random steps formed a stratum that received a unique identifier.
Along each step, we extracted covariate values from the underlying spatial
habitat layers, and we assigned the appropriate light conditions
(\Cref{Covariates}). We opted for covariate extraction \textit{along} steps
rather than \textit{at their endpoints}, as we believed that environmental
conditions along steps were relevant in determining wild dog movement. For
continuous covariates, we computed average values along each step, for
categorical covariates the percentage cover of each category along the step. To
model a decreasing marginal impact of ``distance to'' variables, we included
their square-root as predictors in the final models. To facilitate model
convergence, we normalized extracted values to a range between 0 and 1.

To estimate habitat and movement parameters of interest, we applied the maximum
likelihood procedure proposed by \citet{Muff.2020}, using the \texttt{glmmTMB}
package \citep{Brooks.2017}. Our model formula was based on knowledge from
previous studies on dispersing wild dogs \citep{Hofmann.2021, Hofmann.2023} and
comprised a movement kernel (describing movement preferences), a
habitat-selection function (describing habitat preferences) and their
interactions (describing how movement differs depending on habitat conditions).
Notably, we included descriptors of the step length and turning angle (sl,
log(sl), and cos(ta)) in the regression model \citep{Avgar.2017, Fieberg.2021}.
We also fitted a stratum-specific intercept with a large fixed variance ($10^6$,
\citealp{Muff.2020}), and used dispersing coalition ID to model random slopes.
To examine whether the effect of accounting seasonality differed depending on
the employed model, we fit models either \textit{excluding} interactions
(hereafter called simple model) or \textit{including} interactions (hereafter
called full model). Specifically, in the simple model we only included
covariates relating to landscape characteristics and anthropogenic features in
the simple model, as well as the step-descriptors mandatory for conducting an
iSSF (i.e., sl, log(sl), and cos(ta); \citealp{Avgar.2016}). In the full model,
however, we included additional interactions with climatic descriptors and light
conditions, rendering that wild dog dispersal behavior may vary depending on
those. As such, the structure of the simple model can be viewed as a model
structure that is usually employed in permeability-based studies, which are
restricted to spatial covariates. The model structure of the full model,
conversely, included additional complexities that can only be accounted for when
investigating connectivity vis IBMMs. To ensure comparability among each of the
six configurations presented in our framework, we employed the same models
across all of them. Specifically, we fitted models that differed in their degree
of dynamism (\Cref{GraphicalAbstract}a). Models one and two were fit using
static covariates (\Cref{GraphicalAbstract}a, Stage 1), with model one being a
single-season model and model two being a multi-season model (wet vs. dry,
\Cref{GraphicalAbstract}a, Stage 2). Models two and three, by contrast, were fit
using dynamic covariates, with model three being a single-season model, and
model four being a multi-season model (wet vs. dry, \Cref{GraphicalAbstract}a,
Stage 2). To quantify how many random steps were necessary before model
estimates stabilized (sensu \citealp{Fieberg.2021}), we fitted each model with
5, 10, 25, 50, 75, and 100 random steps.

\subsection{Validation}

We compared the predictive efficacy of each of the six configurations
(\Cref{GraphicalAbstract}b) using k-fold cross-validation for case-control
studies \citep{Fortin.2009}. For this, we split the data into training and
validation sets using an 80:20 ratio and fitted the four iSSF models described
above. We then used the $\beta$-estimates to predict the probability of each
random and observed step in the validation set for being chosen. Depending on
the configuration, we predicted step-probabilities using static or dynamic
covariates. Within each stratum, we assigned ranks 1-101 to each step based on
predicted probabilities and recorded the number of times the observed step was
assigned each rank. Finally, we computed Spearman's rank correlation between
ranks and associated frequencies $r_{s, realized}$. The better a model's
predictive ability, the more negative $r_{s, realized}$ should be (i.e., the
less often the observed step should be assigned a low rank). For reference, we
also computed Spearman's rank correlation for randomized preferences ($r_{s,
random}$), which we achieved by removing the observed step from each stratum and
identifying the rank of a randomly chosen step. We replicated this validation
procedure 100 times and computed the mean correlation coefficient $\bar{r}_{s,
realized}$ and $\bar{r}_{random}$, as well as their 95\% percentiles across
replicates. Ultimately, this validation proves a significant prediction in case
the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
random}\) do not overlap \citep{Fortin.2009}.

\subsection{Simulations}

To assess differences in connectivity upon increasing seasonal dynamism, we ran
dispersal simulations under the two most distinct configurations, i.e. the fully
static (SSS, \Cref{GraphicalAbstract}) and fully dynamic (DMD,
\Cref{GraphicalAbstract}) configurations. As source areas to initiate
dispersers, we defined three distinct regions known to host viable wild dog
populations (\Cref{StudyArea}). The definition of these areas was somewhat
arbitrary, albeit we deliberately selected areas in the west and east of the
Delta to examine potential influences of flooding on connectivity
\citep{Hofmann.2024}, as well as a more isolated location that was not
influenced directly by the Delta's flood extent. To initiate simulated dispersal
trajectories, we randomly placed 1,000 start points within each of these source
area, with start times that were equally distributed across the year. To
simulate dispersal and obtain connectivity maps under both configurations, we
applied the simulation algorithm for iSSFs described in \citet{Signer.2017} and
employed in \citet{Hofmann.2023}. A similar algorithm has recently been added to
the \texttt{amt} \texttt{R}-package \citep{Signer.2024}. We applied the
simulation algorithm as follows. Originating from the start point, we generated
a set of 25 random steps by sampling step lengths from the fitted gamma
distribution and turning angles from a uniform distribution. Along each random
step, we extracted spatial covariates, computed relevant movement metrics (sl,
log(sl), and cos(ta)), and assigned light conditions. We then employed the
fitted iSSF model to predict the probability of each step for being chosen (i.e.
the redistribution kernel). Based on assigned probabilities, we sampled one of
the steps and updated the simulated individual's position and time. We repeated
the procedure until a total of 2,000 steps were realized ($\sim$ 400 dispersal
days).

Depending on the configuration, we employed different spatial covariates during
the simulation. For the SMS configuration, we used the static set of covariates,
which omitted any seasonality and assumed a static landscape over time. For the
DMD configuration, on the other hand, we updated covariates dynamically as the
simulated individuals moved. More specifically, we generated a new covariate
stack after every simulated step, comprising spatial layers that best
represented environmental conditions at that particular point in time
(\Cref{Schematic}). Note that we used the set of covariates representing an
average year for this purpose. For both configurations, we ultimately obtained a
heatmap (or utilization distribution) based on simulated trajectories. We also
generated difference maps between the two configurations, highlighting areas
where the difference in simulated connectivity were most pronounced.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/Schematic} \caption{Schematic
  illustration of a dispersal simulation with dynamic covariates. As the
  simulation proceeds, the underlying covariates (symbolized by the stack of
  layers) are updated. In our case, the update frequency of covariate layers
  varied from a few hours (e.g., temperature) to multiple months (e.g., shrub
  cover). Each gray block represents a single layer and the duration for which
  it was ``active''. Originating from the current position of the simulated
  animal, a new redistribution kernel is derived. We generated redistribution
  kernels by proposing a set of random steps and applying the parametrized
  step-selection model to predict the probability of each step for being chosen.
  Based on this kernel, one location was randomly sampled and the animal's
  position updated. This procedure was repeated until the number of simulated
  steps matched the number of steps from the observed individuals.}
  \label{Schematic}
 \end{center}
\end{figure}

\section{Results}

Due to convergence issues, we removed the covariates NDVI, precipitation, and
distance to pans from all analyses. Some minor convergence issues remained and
our validation procedure failed in 59 out of 1,800 validation attempts. Since
failures occurred across different configurations and were not clustered around
one specific configuration, we deemed their exclusion of no concern.

\subsection{Step-Selection Models}

Patterns of habitat selection and movement behavior were qualitatively similar,
irrespective of whether models were fit using static or dynamic covariates
(\Cref{MovementModel}, Tables S3, S4, S5, S6) and irrespective of the number of
considered random steps (Figure S9). The most notable quantitative difference
when using static vs. dynamic covariates (see \Cref{GraphicalAbstract}a) was
that models fitted using dynamic covariates resulted in narrower confidence
intervals (\Cref{MovementModel}). Furthermore, effect sizes (i.e.,
$\beta$-estimates) of models fit using static covariates were more pronounced
than effect sizes from models that were fit using dynamic covariates
(\Cref{MovementModel}). Differences were most marked for
$\beta_{DistanceToWater^{0.5}}$, which was estimated $\approx -0.8$ using static
covariates, but $\approx -0.3$ when using dynamic covariates.

Differences in $\beta$-estimates between seasons were moderate and most
pronounced when models were fit using static covariates. For instance, when
models were fit using static covariates, $\beta_{Water}$ was $\approx -1.1$
during the dry season but ($\approx 0.2$ during the wet season). When dynamic
covariates were used, by contrast, $\beta_{Water}$ was ($\approx -0.6$ and
$\approx -0.3$), respectively. $\beta_{sl:Dark}$ was also markedly lower during
the wet season ($\approx -0.9$), compared to the dry season ($\approx -0.1$).
Plots that aid with the interpretation of the resulting models are provided in
Figure S10.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/MovementModel.png}
  \caption{$\beta-$estimates from the integrated step-selection models, grouped
  by movement kernel, habitat-selection function, and their interaction. We
  either fit a simple model without interactions or a full model with
  interactions and we distinguished between models fit using static or dynamic
  covariates. Only the underlined covariates differed between the static and
  dynamic configurations, as covariates were either represented as a single
  layer (static), or a stack of layers (dynamic). Furthermore, data was either
  pooled across seasons (yellow bars) or split into dry (green bars) and wet
  (blue bars) season.}
  \label{MovementModel}
 \end{center}
\end{figure}

\subsection{Validation}

Spearman's rank correlation coefficients ($r$) obtained from the validation
procedure revealed that predictions from the full model performed better than
those from the simple model ($\bar{r}_{simple} = -0.5$, $\bar{r}_{full} = -0.9$,
\Cref{RankCorrelation}). Irrespective of the employed model, Spearman's rank
correlation differed significantly depending on the amount of dynamism
considered (simple: F(5, 593) = 26.45, p < 0.001, full: F(5, 594) = 7.14, p <
0.001), albeit with moderate effect sizes. In the simple model, moving from a
fully static (SSS) to a fully dynamic (DMD) configuration decreased Spearmans's
rank correlation (i.e., increased the predictive performance) by 0.15 from -0.41
to -0.56. In the full model, moving from a fully static (SSS) to a fully dynamic
(DMD) configuration entailed an increase in $r$ (i.e. a decrease in the
predictive performance) from -0.89 to -0.88. This suggests that, our hypothesis
that increased dynamism results in better predictive performance only holds with
the simple model (\Cref{RankCorrelation}a), but not the full model
(\Cref{RankCorrelation}b).

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/RankCorrelation.pdf}
  \caption{(a) Spearman's rank correlation across different configurations of
  dynamism that range from entirely static (SSS) to fully dynamic (DMD). The
  more negative Spearman's rank correlation, the better is the predictive
  performance under the respective configuration. Correlations were computed for
  100 replicates. Note that the y-axis is inverted to match our expectation of
  increasing performance as dynamism increases. (b) Difference in Spearman's
  rank correlation when moving from one configuration to another. Values $< 0$
  (blue) indicate an increase in predictive performance, whereas values $> 0$
  (orange) indicate a decrease in predictive performance.}
  \label{RankCorrelation}
 \end{center}
\end{figure}

\subsection{Simulations}

Simulations under the most static (SSS) and dynamic (DMD) configurations
revealed different connectivity patterns (\Cref{HeatmapComparison}). In the
static configuration, simulated dispersal trajectories more concentrated,
resulting in frequent movement across a few key habitats. In the dynamic
configuration, by contrast, simulated trajectories were more homogeneously
distributed. Qualitatively, connectivity patterns appeared similar between the
simple and full models. Given the similarities in habitat selection emerging
under the two models, this was to be expected.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/HeatmapComparison}
  \caption{Heatmaps derived under the most static (SSS) and dynamic (DMD)
  configurations. Results are shown for both the simple and full model.}
  \label{HeatmapComparison}
 \end{center}
\end{figure}

\section{Discussion}

\subsection{Brief Summary}

We introduced a framework to highlight that seasonality can enter a connectivity
analysis at three distinct stages; (1) when extracting spatial covariates for
fitting the selection model, (2) when fitting the selection model, and (3) when
predicting from the fitted model. Through combination, this yields six
configurations that differ in their degree of dynamism and, arguably, realism.
We fitted the models associated with each configuration using GPS data on
dispersing wild dogs and iSSFs and employed a rigorous validation procedure to
investigate potential gains in predictive performance that can be reaped by
incorporating different levels of dynamism. Results from the fitted models
showed that including seasonality only marginally affected the inferred patterns
of habitat selection and movement behavior. Similarly, the validation procedure
suggested only moderate improvements in predictive performance upon increasing
the degree of dynamism. Crucially, these benefits were limited to an overly
simplistic model and vanished upon fitting a more complex one. We therefore
could not pinpoint a specific stage at which including seasonality was
particularly beneficial. Despite this, we found that dispersal simulations under
the most static and most dynamic configurations resulted in differing
connectivity patterns. Under the most static configuration, landscape
connectivity was clumped around a few hot spots, whereas it was homogeneously
distributed across the landscape under the most dynamic configuration. Finally,
our work demonstrated that simulations from IBMMs effectively allow rendering
seasonal changes in the landscape, achieving a degree of seasonal dynamism and
realism that cannot be reached using permeability-based connectivity models.

\subsection{Moderate Improvements upon Increasing Dynamism}

Our validation procedure revealed only moderate improvements in our ability to
predict dispersal movements when increasing the degree of dynamism. Given the
system's extreme seasonal variability, this was somewhat surprising. We believe
that the absence of a more pronounced improvement can be traced back to multiple
factors. Firstly, we focused our analysis on dispersing individuals, which cover
large distances in search of potential mates and a suitable territory
\citep{McNutt.1996, Cozzi.2020}. In our case, the average 4-hourly step length
was 2.5 km, suggesting that dispersers cross and sample numerous unfamiliar
areas and potentially unsuitable habitats within short time. We thus hypothesize
that the spatial scale at which seasonality affects environmental
characteristics did not suffice to match the spatial scale at which our focal
species perceives and moves across the landscape during dispersal. A Further
explanation could be that dispersing individuals prioritize finding unoccupied
territories or areas with low competition, rather than focusing on specific
habitat types (e.g., \citealp{Creel.1996, Creel.2001}). Dispersers' habitat
selection may therefore be more strongly influenced by territorial
considerations (sensu \citealp{Cozzi.2018}) and only little affected by
seasonally changing landscape characteristics. This would also explain the
relatively weak selection or avoidance of environmental characteristics
exhibited in \Cref{MovementModel}. Finally, the AWD is a generalist species that
can occupy a broad variety of habitats \citep{Woodroffe.2011}. A certain
tolerance towards changing environmental conditions can therefore be expected
and may explain why seasonal differences were faint. This holds particularly
true for dispersers, which are usually more tolerant towards unfavorable habitat
conditions as they spend little time within the same area \citep{ONeill.2020}.

\subsection{Substantial Improvements upon Fitting a Complex Model}

In comparison to the improvements achieved by increasing dynamism, a more
significant improvement in predictive ability was achieved by moving from the
simple to the full iSSF model. In the full model, we included several
interactions that accounted for wild dogs' biology, such as reduced movement
during dark nights \citep{Cozzi.2013} or during times of high ambient
temperature \citep{Rabaiotti.2021}. Even though these interactions are not
directly linked to connectivity, they clearly accounted for a substantial amount
of variation in observed dispersal movements and facilitated predictions from
the fitted model. We also allowed for potential changes in dispersers' movement
behavior depending on habitat conditions, such as shorter steps in areas covered
by water \citep{Hofmann.2023}. Overall, the inclusion of these interactions
elevated the predictive ability of our dispersal model to such a high level that
increasing dynamism did not provide any further improvements. Notably, the
ability of encapsulating such a detailed and mechanistic understanding of
dispersal movements is unique to IBMMs and cannot be achieved using
permeability-based connectivity models.

\subsection{Dynamic Connectivity}

When comparing connectivity under the most static (SSS) and dynamic (DMD)
configurations, we observed that connectivity was clumped along a few major
dispersal hotspots under the static configuration, but homogeneously distributed
across the entire landscape under the dynamic configuration. This corroborates
previous research showing that static representations tend to result in an
underestimation of connectivity because seasonal stepping stones or corridors
are missed \citep{Martensen.2017}. With the help of such seasonally-available
dispersal habitats, areas that would otherwise be difficult to reach become
accessible, even if only for a limited time. In our study, such stepping stones
could arise in two ways. Firstly, the landscape could change seasonally, which
may have shifted preferred habitats, leading to the emergence of alternative
movement corridors. In addition, habitat or movement preferences of simulated
individuals could change, resulting in different habitat characteristics being
traversed depending on the season. In combination, varying landscape conditions
and species preferences resulted in a more balanced mosaic of connectivity
across the year. Importantly, the fact that connectivity is more evenly
distributed across the year does not exclude the possibility that certain areas
experience lower connectivity seasonally. As already shown by
\citet{Osipova.2019}, a static representation may result in both an over- and
under-estimation of short-term connectivity, depending on the area and season.
For conservation planning, this implies a need to protect landscapes at broader
scales, as areas that provide little connectivity during some season may become
critical during others. A dynamic take at connectivity therefore improves our
understanding of ecological processes in dynamic landscapes and may help to
identify otherwise overlooked dispersal hotspots.

\subsection{The Costs of Incorporating Dynamism}

Increasing seasonal dynamism to model dynamic connectivity comes at significant
costs, both in terms of data requirements and computational challenges. To
represent seasonal changes in the landscape, one needs to download and process
frequently updated spatial layers for each seasonal covariate. This is
time-consuming and implies a substantial increase in the data-volume that needs
to be handled. Because seasonal products are comparably rare, the choice to
model connectivity dynamically will also entail a significant restriction in the
number of potential covariates or require custom remote sensing algorithms
(e.g., Appendix A1). Seasonal or remote sensed layers are also more susceptible
to noise and missing values, particularly in cases where cloud cover is
frequent. Extracting covariates from seasonal layers poses a further challenge,
as data need to be extracted from those layers that best represent environmental
conditions at the time of the respective observed or random step. This applies
to the extraction of data for model fitting, as well as when extracting data
during the simulation. Finally, splitting the species data by season to fit
seasonal selection models significantly reduces the amount of data remaining per
season, potentially causing convergence issues. In our case, these additional
efforts were not outweighed by an improved predictive ability. This finding is,
however, likely highly specific to our study system. We therefore do not wish to
discourage future studies from accounting for seasonality when assessing
connectivity. Instead, we'd like to view our study as an example that the
benefits of increased realism do not necessarily justify the additional
complexity they bring about and that the associated costs and benefits need to
be carefully pondered \citep{Puy.2022}.

\subsection{Validation Procedure}

Our validation procedure was focused on validation at the step level, but did
not directly test how well predictions of functional connectivity agreed with
true connectivity. This was owed to a conceptual limitation when trying to
validate seasonal connectivity predicted from an IBMM. In seasonal landscapes,
connectivity not only varies spatially, but also temporally. A meaningful
validation therefore requires that connectivity is predicted separately for each
timestamp in the validation data. In our case, for instance, the fact that the
landscape was updated every 4 hours implies that a new connectivity map would
need to be created around each validation step. This is equivalent to validating
connectivity at the step level, which is why we deemed the application of k-fold
cross-validation for case and control studies using Spearman's rank correlation
as an appropriate validation technique. Note, however, that Spearman's rank
correlation coefficient is heavily dependent on the number of steps per stratum
and should therefore not be used to compare predictive efficacy across studies
(Figure S11). For static connectivity analyses, alternative validation
techniques exist. \citet{McClure.2016}, for instance, proposed a suite of
validation metrics (applied in \citealp{Zeller.2018, Finerty.2023}). These
largely work by comparing connectivity at random locations with connectivity at
locations where the focal species was observed. Because these metrics fail to
account for potential autocorrelation in the validation data (which is usually
present in GPS data; \citep{Otis.1999}), \citet{Brennan.2020} suggested
validating connectivity by applying a path- or step-selection model to withheld
data. If the fitted model proves significant selection towards areas of high
connectivity, predictions are indeed indicative of functional connectivity
(sensu \citealp{Brennan.2020}). Developing similar approaches for dynamic
connectivity analyses remains a task for future studies.

\subsection{Seasonal Habitat Selection}

% In the Okavango Delta, water is among the most temporally variable landscape
% features that drives the distribution of species over large spatial scales
% \citep{McCarthy.1998, Ramberg.2006, Bartlam-Brooks.2011, Bennitt.2014}. Wild
% dogs, like most other large carnivores, may avoid a large distance to water
% because of prey-availability in its vicinity \citep{Redfern.2003, Bonyongo.2005,
% Valeix.2010}. At the same time, wild dogs are terrestrial mammals and avoid
% crossing water \citep{Cozzi.2013}, even during dispersal \citep{Hofmann.2021}.

When we fit models assuming static covariates, wild dogs displayed a preference
for areas close to water, irrespective of the season. However, water itself
appeared to be only avoided during the dry season. This was likely caused by a
misrepresentation of areas covered by water in the static configuration. Since
rainwater collected in Angola only slowly descents through the Okavango Delta's
tributaries \citep{McCarthy.1997}, large portions of its floodplains remain dry
during the wet season \cite{McCarthy.2003}. On static covariates, which
represent average conditions across both seasons, it may appear as if
individuals moved through areas covered by water, albeit in reality they likely
moved along them. Indeed, upon accounting for such seasonal changes by
considering seasonally updated covariate layers, parameter estimates from the
selection models assimilated and suggested avoidance of water across both
seasons. Our choice of splitting data into wet and dry season based on a fixed
set of dates can be viewed as a further limitation, as it assumes an immediate
switch from one season to another. In reality, season-transitions are not abrupt
but rather gradual and their timing may vary from year to year. A more robust
approach would therefore be to define the start and end of each season based on
bio-climatic descriptors and to ignore data collected during transitional
periods. This may imply a loss of data, but could help to detect seasonal
selection patterns. In our case, retaining all data was necessary to ensure
model convergence, but may have resulted in a cross-contamination between
seasons and therefore reduced our ability to pick up seasonal differences in
parameter estimates. Indeed, other studies reported that habitat selection of
their focal species varied greatly between seasons (but see
\citealp{Squires.2013}). \citet{Benz.2016}, for instance, found that dispersing
in elk (\textit{Cervus elaphus}) exhibit vastly different habitat preferences
during winter and summer. Similarly, \citet{Osipova.2019} report that habitat
selection of African elephants (\textit{Loxodonta africana}) in South Africa
differs between the wet and dry season. The importance of seasonality thus
appears to be highly system dependent.

% In cases where animals move along borders of two types of habitats
% that shift seasonally, a seasonal representation of habitat conditions seems
% paramount to prevent misleading conclusions.

% Other studies, however, report consistent preferences, irrespective of the
% season. \citet{Squires.2013}, for instance, show that habitat selection by lynx
% (\textit{Lynx lynx}) only differs little between winter and summer. Notably, the
% authors find that inferred patterns of connectivity were similar across seasons.
% However, few of them accounted for seasonality in their covariates, so it is not
% clear whether differences were are truly indicative of a behavioral changes or
% simply reflect changes in environmental conditions.

\subsection{Conclusion}

In conclusion, we explored the importance of incorporating different degrees of
dynamism when studying dispersal and connectivity. Overall, our findings
suggested only moderate improvements in predictive performance upon increasing
the level of dynamism. Nevertheless, connectivity patterns as inferred from
simulated dispersal trajectories differed vastly, with connectivity being more
evenly distributed when seasonality was accounted for. While increased realism
via improved representation of seasonality may offer novel insights into
ecological processes, the benefits must be carefully weighed against the added
complexity and effort required to collect seasonal data and model seasonal
dynamism in the focal species' preferences. Moving forward, our study serves as
an example that while accounting for seasonality could be crucial for some study
systems, it may not always justify the additional complexity, emphasizing the
need for thoughtful consideration of costs and benefits in future research.

\section{Authors' Contributions}

D.D.H. and G.C. conceived the study and designed methodology; D.D.H., G.C.,
D.M.B, and J.W.M. collected the data; D.D.H. analysed the data; G.C. assisted
with modeling; D.D.H. and G.C. wrote the first draft of the manuscript and all
authors contributed to the drafts at several stages and gave final approval for
publication.

\section{Data Availability}

Access to R-scripts to replicate our analyses will be provided through an online
repository at the time of publication.

\section{Acknowledgments \& Funding}

We thank the Ministry of Environment and Tourism and the Department of Wildlife
and National Parks of Botswana for granting permission to conduct this
research. D.D.H. was funded by a Swiss National Science Foundation Grant No.
310030\_204478 to G.C.

\newpage
\begingroup
\singlespacing
\ifthenelse{\boolean{usebiblatex}}{
  \begin{refcontext}[sorting=nyt]
  \printbibliography
  \end{refcontext}
}{
  \bibliography{LiteratureBibtex}
}
\endgroup

\end{document}

### 2024-07-01 Chapter_3/04_Manuscript/Manuscript_2023-07-20.tex
Publish PHD

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
% \usepackage[autostyle]{csquotes}
% \usepackage[
%     backend=biber,
%     style=authoryear-icomp,
%     natbib=true,
%     maxcitenames=1,
%     maxbibnames=99,
%     uniquelist=false,
%     url=true,
%     % doi=true,
%     doi=false,
%     isbn=true,
%     eprint=false,
%     sorting=nyt,
%     sortcites=false
% ]{biblatex}
% \addbibresource{Literature.bib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{textcomp, gensymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[colorlinks, allcolors=black, urlcolor=blue]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
% \usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{scalerel}
\usepackage{tikz}
\usepackage{makecell}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% % Remove some stuff we dont want to list
% \AtEveryBibitem{%
%   \clearfield{month}
%   \clearfield{day}
%   \clearfield{number}
%   \ifentrytype{article}{
%       \clearfield{url}
%       \clearfield{issn}
%       \clearfield{urlyear}
%   }{}
% }
%
% To be able to put an ORCID
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

\newcommand{\inputy}[1]{\input{#1}\unskip}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Simulating Dispersal across Seasonal Landscapes to Assess Dynamic
Connectivity for an Endangered Large Carnivore}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1, 2} \orcid{0000-0001-7477-2642} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
Private Bag 13, Maun, Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\vspace{4cm}

\textbf{Running Title:} Seasonality and Landscape Connectivity: Understanding
Nature's Changing Pathways

\vspace{0.5cm}

\textbf{Keywords:} Seasonality, dispersal, connectivity, individual-based,
Lycaon pictus, Okavango Delta

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Many ecosystems experience changes in environmental conditions due to
seasonality. While such seasonal changes may drastically alter connectivity for
endangered species, most studies represent their study system by a static set of
spatial layers, thus ignoring seasonal variation. Here, we utilize natural
environmental fluctuations across the Okavango Delta in Botswana to address this
shortcoming and investigate seasonal connectivity patterns emerging for
dispersing African wild dogs (\textit{Lycaon pictus}). For this, we parameterize
a seasonal individual-based dispersal model using frequently updated spatial
layers and multi-seasonal GPS data collected on dispersing individuals. Using
the parametrized model, we simulate dispersal across different seasons and
investigate emerging patterns of connectivity. Despite a better understanding of
the conservation needs for African wild dogs, our study also provides evidence
that incorporating seasonality in studies of connectivity is imperative to more
accurately predict the dispersal ability of endangered species.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers
\section{Introduction}
Landscape connectivity, i.e. the degree to which a landscape facilitates or
impedes movement among habitat patches \citep{Taylor.1993}, plays a central role
in maintaining biodiversity \citep{Fahrig.2003} and is one of the most
frequently recommended strategies to promote resilience against changing
environmental conditions \citep{Rudnick.2012}. Studies of landscape connectivity
typically combine information on habitats with knowledge about species'
preferences towards to those habitats to delineate critical movement corridors
\citep{Beier.2008, Diniz.2019}. However, while seasonality is a fundamental
characteristic of most ecosystems, it only rarely enters connectivity models in
an explicit manner. Ignoring seasonality in the landscape or species' habitat
preferences may result in biased connectivity estimates and thus a misallocation
of scarce conservation funds \citep{Osipova.2019, Zeller.2020}.

In light of the continued degradation, fragmentation, and destruction of
valuable habitats worldwide \citep{Fahrig.2003, Haddad.2015}, preserving and
reestablishing connectivity between remaining habitat patches has become a task
of utmost importance \citep{Heller.2009, Rudnick.2012}. The ultimate goal of
promoting connectivity is to facilitate dispersal \citep{Doerr.2011,
Baguette.2013}, i.e. the movement of individuals away from their natal location
to the site of first reproduction \citep{Clobert.2012}. Dispersal, in turn,
promotes genetic exchange \citep{Perrin.2000, Frankham.2002} and the
colonization of empty habitats \citep{Hanski.1999, MacArthur.2001}. Two popular
approaches to estimate functional connectivity are least-cost path analysis
\citep{Adriaensen.2003} and circuit theory \citep{McRae.2008}, both graph-based
methods that quantify conductance of the landscape based on a resistance surface
\citep{Zeller.2012, Diniz.2019}. More recently, individual-based movement models
(IBMMs), where dispersal is explicitly simulated according to a set of movement
rules, have also gained some momentum \citep{Kanagaraj.2013, Allen.2016,
Hauenstein.2019, Diniz.2019, Zeller.2020, UnnithanKumar.2022a,
UnnithanKumar.2022, Hofmann.2023}. All methods have in common that they rely on
information on landscape characteristics and how these affect the dispersal
movements of the focal species to infer critical movement corridors and
dispersal bottlenecks \citep{Diniz.2019}. Irrespective of the chosen method,
researchers only rarely account for seasonal changes in landscape
characteristics or habitat preferences, thus providing an incomplete view on
connectivity across seasons.

Thanks to recent advancements in remote sensing technologies and facilitated
access to petabytes of landscape data at unprecedented spatial and temporal
scales, new opportunities for the study of seasonality and its impacts on
connectivity have emerged \citep{Toth.2016}. Google Earth Engine, for instance,
is an online catalogue and cloud-based analysis platform that enables a fully
reproducable and standardized workflow to access and manipulate a broad range of
satellite products that readily capture seasonal changes across the globe
\citep{Zhao.2021}. Simultaneously, the downsizing of GPS tracking devices and
the extension of their battery life have led to an increase in the availability
of high-quality and multi-seasonal movement data \citep{Cagnacci.2010,
Kays.2015} that can be used to derive seasonal habitat preferences
\citep{Fortin.2005, Manly.2007, Cushman.2010} and inform connectivity models
\citep{Diniz.2019}. In this regard, dispersal data are particularly valuable, as
they readily capture the primary process by which connectivity is established
(\citealp{Elliot.2014, Vasudev.2015}, but see \citealp{Fattebert.2015}).

Seasonality may impact dispersal and connectivity through two channels
\citep{Mui.2017}. Firstly, through seasonal changes in the landscape
configuration and its associated impact on \textit{structural} connectivity. In
ecosystems that experience alternating dry and wet seasons, for instance, the
onset of the rainy season results in distinct ``green-up'' waves that affect the
distribution of food resources that are tracked by herbivores
\citep{Merkle.2016}. Secondly, seasonality could impact how a species reacts to
prevailing environmental conditions through the species' phenology, thus
impacting \textit{functional} connectivity. For example, most amphibians require
both aquatic and terrestrial habitat, yet their preference towards aquatic or
terrestrial environments depends on season and breeding status
\citep{Baldwin.2006}. Using graph-based connectivity models, seasonal landscapes
and seasonal habitat preferences can be incorporated by generating seasonal
resitance surfaces and repeatedly running the connectivity analysis on the
difference surfaces (e.g. \citealp{Chetkiewicz.2009,Cushman.2010, Osipova.2019,
Zeller.2020, Kaszta.2021, Ciudad.2021}). IBMMs, on the other hand, provide the
benefit that landscape characteristics and habitat preferences can be updated
``as simulated individuals move'', thus providing seasonal connectivity
estimates from a single analysis \citep{Zeller.2020}.

The African wild dog (AWD, \textit{Lycaon pictus}) is a keystone predator and
umbrella species for conservation efforts in southern Africa
\citep{Dalerum.2008}. While once present across the entire Sub-Saharan
continent, the species has disappeared from the majority of its historic range,
largely due to human persecution, deadly diseases, and habitat destruction
\citep{Woodroffe.2020a}. With fewer than 2'000 adult individuals remaining in the
wild, the species is listed as endangered by the IUCN red list. The biggest
remaining wild-living population resides in the Okavango Delta in northern
Botswana \citep{McNutt.1996, Woodroffe.2020a}, a highly seasonal and flood-pulse
driven ecosystem \citep{Wolski.2017}. AWDs typically reside in packs comprising
up to 40 individuals, where a single dominant pair monopolizes the majority of
reproduction \citep{Frame.1979, Malcolm.1982}. Upon reaching sexual maturity,
individuals born into the pack emigrate and disperse in single-sex coalitions in
an attempt to find suitable mates and to establish their own pack
\citep{McNutt.1996}. Timing of dispersal appears to be seasonal, with female
dispersal peaking prior to the mating season, and male dispersal peaking at the
onset of the wet season \citep{Behr.2020}. Previous research revealed that
dispersal coalitions can cross several hundred kilometers within only few days
\citep{Davies-Mostert.2012, Masenga.2016, Cozzi.2020}. Dispersers prefer moving
along water-bodies and across open grass or shrubland but avoid areas dominated
by humans or densely covered by forests \citep{ONeill.2020, Hofmann.2021}.
Despite the importance of seasonality in this study system, a seasonal take on
connectivity is still lacking for this species.

In this paper we aim to explore seasonal connectivity patterns for the
endangered African wild dog in northern Botswana. To estimate connectivity, we
first combine multi-seasonal GPS movement data of dispersing AWDs with
dynamically updated habitat layers and parameterize a two-season dispersal
model. This model is based on an integrated step-selection function (iSSF,
\citealp{Avgar.2016}), which produces unbiased estimates for habitat and
movement preferences and results in a mechanistic dispersal model
\citep{Avgar.2016, Signer.2017}. We use the model to simulate dispersal from
distinct habitat patches while dynamically updating habitat preferences and
underlying environmental conditions as the simulated dispersers move. This
results in xx-thousands simulated dispersal trajectories that we use to quantify
seasonal connectivity the chosen habitat patches in the landscape (sensu
\citealp{Hofmann.2023}).

\section{Methods}
We used the \texttt{R} programming language \citep{RCoreTeam.2022} for all data
preparation and analyses. Spatial data manipulation was performed using the
\texttt{terra} \citep{Hijmans.2023} and \texttt{spatstat} \citep{Baddeley.2015}
packages. We generated figures using \texttt{ggplot2} \citep{Wickham.2016}. To
ensure reproducibility, all \texttt{R}-scripts will be made available through an
online repository.

\subsection{Study Area}
The study area for this project was focussed on northern Botswana (centered at
\inputy{99_GeneralMetrics/StudyAreaCenter.tex} at an elevation of approx. 950 m)
but also encompassed parts of Namibia and Zimbabwe, and stretched across a
rectangular extent of \inputy{99_GeneralMetrics/SizeStudyArea.tex}
km\textsuperscript{2} (\Cref{StudyArea}). The main physographical feature within
the study area is the Okavango Delta, a flood-pulse driven mosaic of patchy
woodlands, permanent swamps, and seasonally flooded grasslands that lie within
the otherwise dry and sandy Kalahari Basin \citep{Wilson.1976, Ramberg.2006}.
Precipitation across the study area varies considerably between seasons, ranging
from \inputy{99_GeneralMetrics/MinimumPrecipitation.tex} mm during peak dry
season (dry season from ~ 15 May to 15 October) to
\inputy{99_GeneralMetrics/MaximumPrecipitation.tex} mm during peak wet season
(wet season from 15 October to 15 May), totalling to
\inputy{99_GeneralMetrics/TotalPrecipitation.tex} mm across an average year
(\Cref{Seasonality}a). Daily maximum above-ground temperature fluctuates between
\inputy{99_GeneralMetrics/MinimumTemperature.tex} during winter (i.e. dry
season) to \inputy{99_GeneralMetrics/MaximumTemperature.tex}\degree C during
winter (i.e. wet season, \Cref{Seasonality}b). Vegetation in the study area is
mainly composed of mopane forest (\textit{Colophospermum mopane}),
acacia-dominated woodland (\textit{Acacia spp.}), and grassland. Green-up after
the dry season is typically initiated after the first rains of the wet season,
which stimulate plant growth and leaf production. The normalized difference
vegetation index (NDVI) therefore depicts a lagged and smoothed response to the
precipitation patterns across the study area (\Cref{Seasonality}c). The yearly
flood-cycle of the delta is predominantly driven by rainfalls in the Angolan
highlands, where water is collected and channeled through the Okavango River
into the Okavango Delta \citep{McCarthy.2003, Gumbricht.2004}. Because water
only slowly descends from the catchment areas in Angola into the delta's
tributaries, the flood is out of sync with local rainfalls and typically reaches
its maximum extent during August-September, i.e. during peak dry season
(\citealp{Wolski.2017}, \Cref{Seasonality}d). While the extent of large-bodied
rivers and floodplains is driven by precipitation in Angola, local precipitation
during the wet season lead to the emergence of small water-filled pans.
\inputy{99_GeneralMetrics/PercentageProtected.tex}\% of the landscapes in the
study area form part of a national park or some other type of protected area,
such that human impact remains low and is largely limited to settlements along
the western part of the delta and the city of Maun at the delta's southern tip.
Landscapes outsite protected areas in Zimbabwe, however, are heavily dominated
by humans, largely due to agricultural fields and human settlements.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{99_StudyArea.png}
  \caption{(a) Study area from which data on dispersing AWDs was collected.
  Dispersal trajectories are plotted in dark gray. The study area encompassed
  parts of the Okavango Delta, a highly dynamic, flood-pulse-driven ecosystem.
  The entire study area undergoes substantial seasonal changes, as can be seen
  from two satellite images taken during peak dry season (b1) and peak rainy
  season (b2). Notably, the flood of the Okavango Delta reaches its maximum
  extent during peak dry season.}
  \label{StudyArea}
 \end{center}
\end{figure}

\subsection{GPS Data}
Between the years 2015 and 2022, we collected GPS data of
\inputy{99_GeneralMetrics/CollarsTotal} dispersing AWD coalitions
(\inputy{99_GeneralMetrics/CollarsFemales} female coalitions,
\inputy{99_GeneralMetrics/CollarsMales} male coalitions) from a free-ranging
population in northern Botswana. Details on the GPS collar fitting procedure and
how we distinguished between dispersal and resident movements can be found in
\cite{Cozzi.2020} and \cite{Hofmann.2021}. During dispersal, we programmed GPS
satellite collars to record a GPS fix at a predetermined 4-hourly schedule. The
collars regularly transmitted the data to a base station, which allowed us to
track dispersing individuals, even if they moved outside the main study area. In
total, we successfully collected \inputy{99_GeneralMetrics/FixesTotal} fixes
during dispersal, with an average of \inputy{99_GeneralMetrics/FixesMeanSD}
fixes per coalition. Occasionally, the acquisition of a GPS location failed
(success rate = \inputy{99_GeneralMetrics/AcquisitionRate}\%), resulting in
irregular durations between some subsequent fixes.

\subsection{Spatial Habitat Layers}
We represented the physical landscape through which dispersers could move by a
series of spatial layers that we believed whould influence wild dog movements
during dispersal. The layers can be broadly categorized into descriptors of (1)
landscape characteristics, (2) climatic conditions, and (3) anthropogenic
factors (\Cref{tab:Covariates}). To appropriately render seasonal dynamicity in
each of these layers, we downloaded spatial data at the highest temporal
resolution available. That is, whenever possible, we represented covariates by a
series of raster layers that spanned the entire range of dates for which we
collected data of dispersing AWDs. We downloaded each product at the highest
spatial resolution available (\Cref{tab:Covariates}).

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{99_SeasonalCovariates.png}
  \caption{Illustration of how some of the covariates considered in this study
  vary across seasons. Data for the graphs was obtained from the (a) Global
  Satellite Mapping of Precipitation dataset, (b) ERA5 dataset, (c) MODIS
  MOD13Q1 dataset, and (d) remote sensed MOD43A4 satellite images. Values were
  extracted across the study area and averaged by months. The smoothing curves
  were fitted using simple GAM models in the \texttt{mgcv} package
  \citep{Wood.2011}.}
  \label{Seasonality}
 \end{center}
\end{figure}

\inputy{99_Covariates.tex}

\subsubsection{Landscape Characteristics}
We used data from the MODIS Vegetation Continuous Fields dataset (MOD44B V061,
\citealp{DiMiceli.2022}) to represent vegetation across the study area. The
MOD44B dataset comprises three continuous layers, depicting the percentage cover
of \textsf{woodland}, \textsf{shrubs/grassland}, and \textsf{bareland},
respectively. Because the layers add up to 100\%, we only considered the two
layers \textsf{woodland} and \textsf{shrubs/grassland}, thus preventing issues
of perfect multicollinearity. This specific product is updated on day 65 of each
year and so we used the R-package \texttt{RGISTools} \citep{Perez-Goya.2020} to
download yearly updated layers for the years for which GPS data was at our
disposal. We also downloaded information on the normalized vegetation difference
index (NDVI) from the MODIS MOD13Q1 dataset \citep{Didan.2015}. This product is
updated every 16 days and we accessed the respective layers through Google Earth
Engine \citep{Gorelick.2017} using the r-package \texttt{rgee}
\citep{Aybar.2023}. We depicted open water sources using the Globeland30
dataset, from which we only retained the land-cover class water. To dynamically
update the distribution of floodwater across the extent of the Okavango Delta,
we combined the Globeland30 water-cover layer with weekly updated floodmaps
derived from remote sensed MODIS MOD43A4 satellite images. The algorithm upon
which the separation between floodwater and dryland is based is described in
detail in \cite{Wolski.2017} and \cite{Hofmann.2021} and is implemented in the
\texttt{floodmapr} package (available on GitHub;
\url{https://github.com/DavidDHofmann/floodmapr}). Because this algorithm has
been developed to remote sense large-scale bodies of floodwater, it is
unsuitable for the detection of smaller waterbodies, such as those represented
by pans. As such, we parametrized a classifier that was specifically focused on
detecting surface water from small-scale pans using Sentinel 2 satellite imagery
\citep{EuropeanSpaceAgency.2018}. This allowed us to obtain updated maps for the
distribution of pan-water at an interval of 5 to 10 days.

\subsubsection{Climate Descriptors}
We downloaded hourly updated information on above-ground temperature and
precipitation through Google Earth Engine \citep{Gorelick.2017} using the
\texttt{rgee} package \citep{Aybar.2023}. We obtained temperature data from the
ERA5-Land dataset \citep{Munoz-Sabater.2021}, whereas we utilized the Global
Satellite Mapping of Precipitation dataset to access precipitation data
\citep{Kubota.2020}. To match hourly values from the temperature and
precipitation data with the 4-hour fixes from the GPS data collection, we
computed average precipitation and temperature values across four hours.

\subsubsection{Anthropogenic Features}
We combined anthropogenic pressures arising from the presence of humans,
agriculture, and roads into a single proxy for human influence. We obtained
information on human density from Facebook's high resolution human density
dataset \citep{Tiecke.2017} which we downloaded from the humdata website
(\url{www.data.humdata.org/}). We also downloaded rasterized information on the
presence of agricultural fields from the Globeland30 \citep{Chen.2015} and
Cropland \citep{Xiong.2017} datasets. Vectorized data on main tar roads was
downloaded from OpenStreetMaps \citep{OpenStreetMapContributors.2017}. How the
individual layers were merged and combined into a single layer is described in
detail in \cite{Hofmann.2021}. Finally, we also derived vector data on the
distribution of national parks, wildlife management areas and other protected
areas from the Peace Parks Foundation (\url{www.maps.ppf.org.za/}).

\subsection{Step Selection Function}
We estimated seasonal habitat and movement preferences of dispersing individuals
using an integrated step-selection function (iSSF, \citealp{Fortin.2005,
Avgar.2016}). In iSSFs, \textit{observed} steps (i.e. the straight line between
two consequtive GPS locations \citep{Turchin.1998}) are compared to
\textit{random} steps in a conditional logistic regression framework
\citep{Fortin.2005, Thurfjell.2014, Muff.2020, Fieberg.2021}. We thus converted
subsequent GPS locations during dispersal into bursts of steps with equal
duration. For this, we first identified bursts of consecutive GPS fixes where
the duration between two GPS fixes did not exceed 4 hours (\(\pm\) 15 minutes).
Within each burst, we computed step lengths (sl) and relative turning angles
(ta) associated with each observed step. \inputy{99_GeneralMetrics/SSFTotal}
dispersing coalitions (\inputy{99_GeneralMetrics/SSFFemales} female coalitions,
\inputy{99_GeneralMetrics/SSFMales} male coalitions) remained in the final
dataset. This yielded a total of \inputy{99_GeneralMetrics/StepsTotal} steps
(\inputy{99_GeneralMetrics/StepsMeanSD} per coalition). We partitioned the
resulting steps into wet and dry season.
\inputy{99_GeneralMetrics/PercentageStepsDry}\% of the GPS data fell into the
dry season, \inputy{99_GeneralMetrics/PercentageStepsWet}\% into the wet-season.

We paired each observed step with a set of 25 random steps, generated by
sampling turning angles from  a uniform distribution U(\(-\pi, +\pi\)) and step
lengths from a gamma distribution fitted to observed steps (scale \(\theta\) =
valuefromtex and shape \(k\) = valuefromtex) using the \texttt{fitdistrplus}
package \citep{Delignette-Muller.2015}. Together, an observed step and its 25
associated random steps formed a stratum that received a unique identifier.
Along each step, we extracted covariate values from the underlying spatial
habitat layers (\Cref{tab:Covariates}). For continuous covariates, we computed
average values along each step, for categorical covariates the percentage cover
of each category along the step. To facilitate model convergence, we
standardized extracted values using a z-score transformation.

We then assumed that each animal assigned a score \(w(x)\) of the following
exponential form to each realized and random step \citep{Fortin.2005}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent Here, (\(x_1, x_2, ..., x_n\)) represent the covariate values along
the respective step and (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's
relative selection strengths \citep{Avgar.2017} towards these covariates. To
estimate relative selection strengths (i.e. the \(\beta\)'s), we used the mixed
effects conditional logistic regression model as proposed by \cite{Muff.2020}.
We implemented their method using the R-package \textit{glmmTMB}
\citep{Brooks.2017} and used dispersing coalition ID to model random slopes. The
approach also requires the inclusion of a random intercept with an arbitrarily
fixed variance of 10\textsuperscript{6}.

To establish whether the number of random steps sufficed to reliably estimate
model parameters, we ran a series univariate models where we only included
single predictors and estimated model coefficients considering 1, 5, 10, 50, and
100 random steps, respectively (sensu \citealp{Fieberg.2020}). We replicated
each model run 100 times and randomly resampled the necessary number of random
steps. This allowed us to investigate the minimum number of random steps that
were required before estimates for habitat and movement preferences converged
towards a stationary estimate.

\section{Results}

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{99_MovementModel.png}
  \caption{This is a caption}
  \label{Movement Model}
 \end{center}
\end{figure}

\inputy{99_MovementModel.tex}

\section{Discussion}
Dynamic connectivity increases our understanding of temporal changes in
connectivity \citep{Martensen.2017}.

\cite{Martensen.2017} showed that dynamic connectivity was higher than static
connectivity because of stepping stones that emerged and disappeared again.

\cite{Osipova.2019} in contrast find that connectivity is both over and
underestimated, depending on the season.

\cite{Benz.2016} showed that habitat selection of moose changes depending on the
season and used a combination of RSFs and SSFs to study habitat selection across
seasons.

Badly designed corridors can result in population sinks and wasted financial
resources \citep{Simberloff.1992}.

Could talk a bit about the following papers: \citep{Chetkiewicz.2009, Benz.2016,
Mui.2017, Osipova.2019, Kaszta.2021}

\section{Authors' Contributions}
D.D.H. and G.C. conceived the study and designed methodology; D.D.H., G.C.,
D.M.B, and J.W.M. collected the data; D.D.H. analysed the data; G.C. assisted
with modeling; D.D.H. and G.C. wrote the first draft of the manuscript and all
authors contributed to the drafts at several stages and gave final approval for
publication.

\section{Data Availability}
GPS movement data of dispersing wild dogs is available on dryad
\citep{Hofmann.2021a}. Access to R-scripts that exemplify the application of the
proposed approach using simulated data are provided through Github
(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
codes required to reproduce the African wild dog case study will be made
available through an online repository at the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we would like
to thank Johannes Signer for assisting with the simulation algorithm. This study
was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing

% \printbibliography
\bibliography{Literature}
\endgroup

\end{document}

### 2024-07-01 Chapter_3/04_Manuscript/Manuscript_2023-12-06.tex
Publish PHD

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage{scalerel}
\usepackage{makecell}
\usepackage{ifthen}
\usepackage{tikz}
\usepackage{lineno}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%  Bibtex vs Biblatex
%------------------------------------------------------------------------------
% While bibtex is very nice to use, it's insanely slow. I therefore work with
% bibtex until the final stage. Set the value below to "true" or "false" to pick
% which one should be used.
\newboolean{usebiblatex}
\setboolean{usebiblatex}{false}

% Depending on the question, the commands below will prepare the document
% accordingly
\ifthenelse{\boolean{usebiblatex}}{
  %-----------------------------------------------------------------------------
  %  If using biblatex
  %-----------------------------------------------------------------------------
  % Load required packages and refer to correct bibliography
  \usepackage{csquotes}
  \usepackage[
      backend=biber
    , style=apa          % To display Author-year in the text
    , natbib=true        % Necessary to use citep, etc.
    , hyperref=true      % To have clickable links in the references
    , useeditor=false    % Don't add editor to publications
    , sorting=ynt        % Sort by year, name, and title
    , uniquename=false   % Avoid that initials are put
  ]{biblatex}
  \addbibresource{Literature.bib}

  % To include author in link
  \DeclareFieldFormat{citehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{#1}}
  \DeclareFieldFormat{textcitehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{%
    #1%
    \ifbool{cbx:parens}
      {\bibcloseparen\global\boolfalse{cbx:parens}}
      {}}}
  \savebibmacro{cite}
  \savebibmacro{textcite}
  \renewbibmacro*{cite}{%
    \printtext[citehyperref]{%
      \restorebibmacro{cite}%
      \usebibmacro{cite}}}
  \renewbibmacro*{textcite}{%
    \ifboolexpr{
      ( not test {\iffieldundef{prenote}} and
        test {\ifnumequal{\value{citecount}}{1}} )
      or
      ( not test {\iffieldundef{postnote}} and
        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
    }
      {\DeclareFieldAlias{textcitehyperref}{noformat}}
      {}%
    \printtext[textcitehyperref]{%
      \restorebibmacro{textcite}%
      \usebibmacro{textcite}}}

  % Remove editor form everything but books
  \AtEveryBibitem{%
   \ifentrytype{book}{}{%
    \clearname{editor}
   }
  }
}{
  %-----------------------------------------------------------------------------
  %  If using bibtex
  %-----------------------------------------------------------------------------
  % Load required packages and select citation style
  \usepackage[round]{natbib}
  \bibliographystyle{apalike}
}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Link colors
\definecolor{linkcolor}{HTML}{000000}
\hypersetup{
  colorlinks = true,
  linkcolor  = linkcolor,
  urlcolor   = linkcolor,
  citecolor  = linkcolor,
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% Avoid skip when including text from external .tex file
\newcommand{\inputy}[1]{\input{#1}\unskip}

% For supplementary material
\newcommand{\beginappendix}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
  \setcounter{equation}{0}
  \renewcommand{\theequation}{Equation S\arabic{equation}}%
  \setcounter{section}{0}
  \renewcommand{\thesection}{A.\arabic{section}}%
}

%------------------------------------------------------------------------------
%  ORCID
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Create the ORCID logo
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol]
      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
        ,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white]
      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

% Command to create the ORCID
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Simulating Dispersal across Seasonal Landscapes to Assess Dynamic
Connectivity for an Endangered Large Carnivore}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1, 2} \orcid{0000-0001-7477-2642} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
Private Bag 13, Maun, Botswana.

\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}

\vspace{4cm}

\textbf{Running Title:} Seasonality and Landscape Connectivity: Understanding
Nature's Changing Pathways

\vspace{0.5cm}

\textbf{Keywords:} Seasonality, dispersal, connectivity, individual-based,
Lycaon pictus, Okavango Delta

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}

Many ecosystems experience changes in environmental conditions due to
seasonality. While seasonal changes may drastically alter connectivity for
endangered species, most studies represent their system by a static set of
spatial layers, thus ignoring potential changes in environmental conditions due
to seasonal variation. Here, we study seasonal environmental fluctuations
across the Okavango Delta in Botswana and investigate how they affect dispersal
and connectivity for the endangered African wild dog (\textit{Lycaon pictus}).
For this, we compile an extensive set of raster layers accurately representing
seasonal changes in environmental conditions and we combine them with GPS data
collected on dispersing wild dogs to parametrize a dispersal model. Using the
parameterized model, we simulate dispersal trajectories, as environmental
conditions change and we investigate emerging patterns of connectivity. Despite
a better understanding of the conservation needs for African wild dogs, our
study also provides evidence that incorporating seasonality in studies of
connectivity is imperative to more accurately predict the dispersal ability of
endangered species.

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers
\section{Introduction}

Landscape connectivity, i.e. the degree to which a landscape facilitates or
impedes movement among habitat patches \citep{Taylor.1993}, plays a central role
in maintaining biodiversity \citep{Fahrig.2003} and is one of the most
frequently recommended strategies to promote resilience against changing
environmental conditions \citep{Heller.2009, Rudnick.2012}. Studies of landscape
connectivity typically combine information on environmental conditions with
knowledge about species' preferences to delineate critical movement corridors
\citep{Beier.2008, Diniz.2019}. However, while seasonality is a fundamental
characteristic of many ecosystems, it only rarely enters connectivity models in
an explicit manner. Ignoring seasonality, either in environmental conditions or
species' habitat preferences, may result in biased connectivity estimates and a
misallocation of scarce conservation funds \citep{Osipova.2019, Zeller.2020}.

In light of the continued degradation, fragmentation, and destruction of
valuable habitats worldwide \citep{Fahrig.2003, Haddad.2015}, preserving and
reestablishing connectivity between remaining habitat patches has become a task
of utmost importance \citep{Heller.2009, Rudnick.2012}. Improved connectivity
facilitates dispersal \citep{Doerr.2011, Baguette.2013}, i.e. the movement of
individuals away from their natal location to the site of first reproduction
\citep{Clobert.2012}, which in turn promotes genetic diversity
\citep{Perrin.2000, Frankham.2002} and the colonization of vacant habitats
\citep{Hanski.1999, MacArthur.2001}. Two popular approaches to estimate
connectivity are least-cost path analysis \citep{Adriaensen.2003} and circuit
theory \citep{McRae.2008}, both graph-based methods that quantify conductance of
the landscape based on a resistance surface \citep{Zeller.2012, Diniz.2019}.
More recently, individual-based movement models (IBMMs), where dispersal is
explicitly simulated based on a set of movement rules, have been proposed as a
powerful and flexible alternative to graph-based analyses \citep{Kanagaraj.2013,
Allen.2016, Hauenstein.2019, Diniz.2019, Zeller.2020, UnnithanKumar.2022a,
UnnithanKumar.2022, Hofmann.2023}.

Recent advancements in remote sensing technologies and an increasingly
facilitated access to petabytes of landscape data at unprecedented spatial and
temporal scales have created new opportunities to study seasonality and its
impacts on connectivity \citep{Toth.2016, Rumiano.2020}. Google Earth Engine,
for instance, is an online catalog and cloud-based analysis platform that
facilitates a standardized and fully reproducible workflow to explore,
manipulate, analyze, and download a broad range of satellite products that
readily capture seasonal changes across the globe \citep{Zhao.2021}.
Furthermore, the ability to complete computationally demanding processing steps
on Google servers permits such data to be processed using average computers.
Simultaneously, the downsizing of GPS tracking devices and the extension of
their battery life have led to an increase in the availability of high-quality
and multi-seasonal movement data \citep{Cagnacci.2010, Kays.2015} that can be
used to derive seasonal habitat preferences of wild-living animals
\citep{Fortin.2005, Manly.2007, Cushman.2010} and to inform connectivity models
\citep{Diniz.2019}. In this regard, dispersal have proven particularly valuable,
as they readily capture the primary process by which connectivity is established
(\citealp{Elliot.2014, Vasudev.2015}, but see \citealp{Fattebert.2015}).

Seasonality may impact dispersal and connectivity both through changes in the
landscape itself, or through changes in the species' habitat preferences
\citep{Mui.2017}. For instance, in ecosystems that experience alternations
between dry and wet seasons, the onset of the rainy season results in distinct
``green-up'' waves that affect the distribution of food resources and, as a
consequence, the movement of herbivores \citep{Merkle.2016}. Concomitantly, a
species' preferences towards landscape characteristics may change drastically
depending on the season. Amphibians, for instance, require both aquatic and
terrestrial habitats, but their preference towards those environments depends on
species' seasonal breeding status \citep{Baldwin.2006}. Similarly, wolves
(\textit{Canis lupus}) in the Yellowstone National Park exhibit more complex
habitat-selection during winter than during summer. In predator-prey
interactions, seasonal changes in habitat-selection of one species may
significantly alter habitat preferences of it's counterpart
\citep{Grenier-Potvin.2021}. Despite its apparent importance, rendering
seasonality in both environmental features and species' selection behavior may
require substantial amounts of data and thus raises the question of its
importance.

Incorporating seasonality in graph-based connectivity models (i.e., least-cost
and circuit-theory models) requires to repeatedly run the same analysis using
multiple resistance surfaces, each representing another point in time (e.g.
\citealp{Chetkiewicz.2009, Cushman.2010, Osipova.2019, Zeller.2020, Kaszta.2021,
Ciudad.2021}). This prevents a more continuous approach at seasonal connectivity
and cannot render transitions from one season to another. With IBMMs, on the
other hand, individuals are simulated through time, thus allowing to update both
landscape characteristics and habitat preferences as the simulation progresses,
therefore providing seasonal connectivity estimates from a single analysis
\citep{Zeller.2020}. Using IBMMs, there are three stages during which
seasonality could be incorporated (\Cref{GraphicalAbstract}). In stage one, a
modeler can decide to either utilize a static snapshot of spatial layers to
represent the environemntal features, or to engage in obtaining seasonally
updated stacks that readily render seasonal changes in the landscape. In stage
two, one can decide to parametrize a single-season model that assumes fixed
habitat preferences, or two model a multi-season selection model where habitat
preferences change over time. Finally, in stage three, a modeler can either
simulate dispersal on a static set of spatial covariates, or indeed chose to
update relevant layers as the simulated individuals move.

The African wild dog (AWD, \textit{Lycaon pictus}) is a keystone predator and
umbrella species for conservation efforts in Southern Africa
\citep{Dalerum.2008}. While once present across the entire Sub-Saharan
continent, the species has disappeared from the majority of its historic range,
largely due to human persecution, deadly diseases, and habitat destruction
\citep{Woodroffe.2020a}. With fewer than 2'000 adult individuals remaining in
the wild, the species is considered as endangered on the IUCN red list. The
biggest remaining wild-living population of wild dogs resides in the Okavango
Delta in northern Botswana \citep{McNutt.1996, Woodroffe.2020a}, a highly
seasonal and flood-pulse driven ecosystem \citep{Wolski.2017}. AWDs typically
reside in packs comprising up to 40 individuals, where a single dominant pair
monopolizes reproduction \citep{Frame.1979, Malcolm.1982}. Upon reaching sexual
maturity, individuals born into the pack emigrate and disperse in single-sex
coalitions in an attempt to find suitable mates and an area to establish their
own pack \citep{McNutt.1996}. In Botswana, the timing of emigration is seasonal,
with female emigration peaking prior to the mating season in March, and male
emigration peaking at the onset of the rainy season in December
\citep{Behr.2020}. Dispersing coalitions can cross several hundred kilometers
within only a few days, often crossing international borders
\citep{Davies-Mostert.2012, Masenga.2016, Cozzi.2020, Sandoval-Seres.2022}.
Studies of habitat-selection show that dispersers prefer moving along water and
across open grass or shrubland but tend to avoid areas dominated by humans or
densely covered by forests \citep{ONeill.2020, Hofmann.2021}. While resident
wild dogs are known to be active during moonlit nights \citep{Cozzi.2012}, it is
unclear to what degree this applies to dispersing coalitions too.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/GraphicalAbstract.pdf}
  \caption{Overview of the different dimensions in which seasonality can be
  rendered in studies of simulated landscape connectivity. (a) During model
  fitting, one needs to decide whether to represent the environment by a static
  or dynamic set of covariates. In the former case, seasonal changes of the
  landscape cannot be captured, whereas in the latter case they can. Similarly,
  a modeler needs to decide whether to parameterize a single-season or
  multi-season movement model. Such a model either disregards or encapsulates
  seasonality in the focal species' habitat and movement preferences. (b) When
  utilizing the fitted model to simulate dispersal and estimate landscape
  connectivity, one can either assume a static set of environmental covariates
  that remain unchanged for the duration of the simulation, or a dynamic set of
  layers that are updated as the simulation progresses. (c) Depending on these
  decisions, six different combinations emerge. Our hypothesis was that
  increasing the degree of dynamism would improve the predictive power of the
  dispersal simulation, as the increasing complexity would allow capturing
  intricacies of real biological systems. In this regard, we were particularly
  interested in determining which of the above mentioned decisions would have
  the biggest impact on the predictive power of the simulation.}
  \label{GraphicalAbstract}
 \end{center}
\end{figure}

To this end, we used an IBMM to explore seasonal connectivity for the endangered
African wild dog in northern Botswana. For this, we combined multi-seasonal GPS
movement data of dispersing AWDs with dynamically updated habitat layers and
parametrize a two-season integrated step-selection function (iSSF;
\citealp{Avgar.2016}). We then applied the parametrized iSSF to simulate
dispersal trajectories, while continuously updating habitat preferences and
environmental conditions. This resulted in xx-thousands simulated dispersal
trajectories based on which we quantified seasonal connectivity between selected
habitat patches in the landscape (sensu \citealp{Hofmann.2023}). We also
compared our simulations to a scenario where seasonality was neglected and
discuss resulting differences. (could add some hypotheses here)

\section{Methods}

We used the \texttt{R} programming language \citep{RCoreTeam.2023} for all data
preparation and analyses. Spatial data manipulation was performed using the
\texttt{terra} \citep{Hijmans.2023} and \texttt{spatstat} \citep{Baddeley.2015}
packages. We generated figures using \texttt{ggplot2} \citep{Wickham.2023} and
\texttt{ggpubr} \citep{Kassambara.2023}. To ensure reproducibility of all our
analyses, we will provide access to our \texttt{R}-scripts through an online
repository upon publication of this article.

\subsection{Study Area}

The study area for this project was focussed on northern Botswana (centered at
\inputy{GeneralMetrics/StudyAreaCenter.tex} at an elevation of approx. 950 m)
but also encompassed parts of Namibia and Zimbabwe, and stretched across a
rectangular extent of \inputy{GeneralMetrics/SizeStudyArea.tex}
km\textsuperscript{2} (\Cref{StudyArea}). The main physiographical feature
within the study area is the Okavango Delta, a flood-pulse driven mosaic of
patchy woodlands, permanent swamps, and seasonally flooded grasslands that lie
within the otherwise dry and sandy Kalahari Basin \citep{Wilson.1976,
Ramberg.2006}. Precipitation across the study area varies considerably between
seasons, ranging from \inputy{GeneralMetrics/MinimumPrecipitation.tex} mm during
peak dry season (dry season from $\sim$ 15 May to 15 October) to
\inputy{GeneralMetrics/MaximumPrecipitation.tex} mm during peak wet season (wet
season from $\sim$ 15 October to 15 May), totaling to
\inputy{GeneralMetrics/TotalPrecipitation.tex} mm across an average year
(\Cref{Seasonality}a). Daily maximum above-ground temperature fluctuates between
\inputy{GeneralMetrics/MinimumTemperature.tex} during winter (i.e. dry season)
to \inputy{GeneralMetrics/MaximumTemperature.tex}\degree C during summer (i.e.
wet season, \Cref{Seasonality}b). Vegetation in the study area is mainly
composed of mopane forest (\textit{Colophospermum mopane}), acacia-dominated
woodland (\textit{Acacia spp.}), and grassland. Green-up after the dry season is
initiated after the first rains of the wet season, which stimulate plant growth
and leaf production. As a consequence, the normalized difference vegetation
index (NDVI) depicts a lagged response to precipitation patterns across the
study area (\Cref{Seasonality}c). The yearly flood-cycle of the delta is
predominantly driven by rainfalls in the Angolan highlands, where water is
collected and channeled through the Cubango and Cuito rivers into the Okavango
Delta \citep{McCarthy.2003, Gumbricht.2004, Mendelsohn.2010}. Because water only
slowly descends from the catchment areas in Angola into the delta's tributaries,
the flood is out of sync with local rainfalls and reaches its maximum extent
during August-September, i.e. during peak dry season (\citealp{Wolski.2017},
\Cref{Seasonality}d). While the extent of large-bodied rivers and floodplains is
driven by precipitation in Angola, the emergence of smaller, ephemeral
water-bodies (i.e., pans) is dicated by local precipitation during the wet
season. \inputy{GeneralMetrics/PercentageProtected.tex}\% of the landscape in
the study area form part of a protected area, such that human impact remains low
and largely limited to settlements along the western part of the delta and the
city of Maun at the delta's southern tip (\Cref{StudyArea}). Landscapes outside
protected areas in Zimbabwe, however, are heavily influenced humans, mainly
through agricultural fields and human settlements.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/StudyArea.png} \caption{(a) Study
  area from which data on dispersing AWDs were collected. Dispersal trajectories
  are plotted in dark gray. The study area encompassed parts of the Okavango
  Delta in northern Botswana, a highly dynamic, flood-pulse-driven ecosystem.
  The entire study area undergoes substantial seasonal changes, as can be seen
  from two satellite images taken during peak dry season (b1) and peak rainy
  season (b2). Notably, the flood of the Okavango Delta reaches its maximum
  extent during peak dry season (August - September).}
  \label{StudyArea}
 \end{center}
\end{figure}

\subsection{GPS Data}

Between the years \inputy{GeneralMetrics/GPSFromYear} and
\inputy{GeneralMetrics/GPSToYear}, we collected GPS data of
\inputy{GeneralMetrics/CollarsTotal} dispersing AWD coalitions
(\inputy{GeneralMetrics/CollarsFemales} female coalitions,
\inputy{GeneralMetrics/CollarsMales} male coalitions) from a free-ranging
population in northern Botswana. Details on the GPS collar fitting procedure and
how we distinguished between dispersal and resident movements can be found in
\citet{Cozzi.2020} and \citet{Hofmann.2021}. During dispersal, GPS satellite
collars recorded a GPS location every four hours and regularly transmitted the
data to a base-station via Irridium satellite. In total, we successfully
collected \inputy{GeneralMetrics/FixesTotal} locations during dispersal, with an
average of \inputy{GeneralMetrics/FixesMeanSD} locations per coalition.
Occasionally, the acquisition of a GPS location failed (success rate =
\inputy{GeneralMetrics/AcquisitionRate}\%), resulting in slight deviations from
the aspired four-hourly interval.

\subsection{Covariates}

We represented the physical landscape through which dispersers could move by a
suite of spatial layers that we believed would influence wild dog movements
during dispersal. The layers can be broadly categorized into descriptors of (1)
landscape characteristics, (2) climatic conditions, and (3) anthropogenic
factors (\Cref{Covariates}). Besides these spatial layers, we also prepared a
series of (4) nocturnal statistics (\Cref{Covariates}). To appropriately render
seasonality in each of the spatial covariates (1-3), we downloaded spatial data
at the highest spatial and temporal resolution available. Specifically, we
attempted to represented covariates by a series of raster layers that spanned
the entire range of dates for which we collected data of dispersing AWDs.
Depending on the products available, this resulted in differing spatial and
temporal resolutions (\Cref{Covariates}). To assess the importance of rendering
layers at such high temporal resolutions, we used these ``dynamic'' covariates
to compute ``static'' covariates, representing average conditions over the
entire study period. For this, we flattened each covariate-stack into a single
layer, thus removing seasonality from the spatial data entirely. The only
exception to this was the precipitation and temperature data, which instead we
flattened by hour of the day.

\begin{figure}
\begin{center}
  \includegraphics[width = \textwidth]{Figures/SeasonalCovariates.png}
  \caption{Illustration of how some of the covariates considered in this study
  vary across seasons. Data for the graphs were obtained from the (a) JAXA
  GSMaP, (b) ERA5 dataset, (c) MODIS MOD13Q1 dataset, and (d) remote sensed
  MOD43A4 satellite images. Values were extracted across the study area and
  averaged by months. Smoothing curves were fitted using GAM models with the
  \texttt{mgcv} package \citep{Wood.2011}.}
  \label{Seasonality}
\end{center}
\end{figure}

\onehalfspacing
\begin{center}
\begin{threeparttable}
  \caption{Covariates that were used in this study, including information on
  their temporal and spatial resolutions, as well as on the avenue through which
  the respective data were accessed or downloaded}
  \label{Covariates}
  \input{Figures/Covariates.tex}
  \begin{tablenotes}
    \item \textit{The spatial layers in gray were combined into a single proxy
    of human influence, as described in \citet{Hofmann.2021}.}
  \end{tablenotes}
\end{threeparttable}
\end{center}
\doublespacing

\subsubsection{Landscape Characteristics}

We used data from the MODIS Vegetation Continuous Fields dataset (MOD44B V061;
\citealp{DiMiceli.2022}) to represent vegetation across the study area. The
MOD44B dataset comprised three continuous layers, depicting the percentage cover
of woodland, shrubs/grassland, and bareland, respectively. The three layers
added up to 100\%, so we dropped bareland layer to prevent perfect
multicollinearity. The MOD44B product is updated on day 65 of each year, so we
used the \texttt{R}-package \texttt{RGISTools} \citep{Perez-Goya.2020} to
download yearly updated layers. Furthermore, we downloaded information on the
normalized vegetation difference index (NDVI) from the MODIS MOD13Q1 dataset
\citep{Didan.2015}. This product is updated every 16 days, and we accessed the
respective layers through Google Earth Engine \citep{Gorelick.2017} using the
\texttt{R}-package \texttt{rgee} \citep{Aybar.2023}. To depict large, static
water-bodies, we used the Globeland30 dataset, from which we only retained the
land-cover class water, while setting all other categories to dryland.
Similarly, we used the MERIT Hydro dataset to obtain information on permanent
rivers \citep{Yamazaki.2019}. To depict large, dynamic water-bodies,
particularly the floodwaters of the Okavango Delta, we prepared weekly updated
floodmaps using remote sensed MODIS MOD43A4 satellite images. The algorithm upon
which the identification of flooded areas is based is described in detail in
\citet{Wolski.2017} and \citet{Hofmann.2021} and implemented in the
\texttt{floodmapr} package (available on GitHub;
\url{https://github.com/DavidDHofmann/floodmapr}). We also modeled small,
ephemeral water bodies (i.e., pans) by applying a customly built random forest
classifier to Sentinel 2 satellite imagery (\citealp{EuropeanSpaceAgency.2018};
details of the classifier in Appendix S1). While Sentinel 2 satellite imagery is
updated every 5 days, cloud cover often results in incomplete images, thus
prohibiting producing ``pan-maps'' at this frequency. Instead, we opted for
monthly update composite images, thus minimizing the risk of misclassification
due to cloud cover. The procedure resulted in monthly updated pan-maps that
depicted the distribution of ephemeral waterbodies as small as a few meters
(Appendix SX).

\subsubsection{Climate Descriptors}
We used hourly updated information on above-ground temperature from the
ERA5-Land dataset \citep{Munoz-Sabater.2021} and hourly updated precipitation
estimates from the Global Satellite Mapping of Precipitation dataset
\citep{Kubota.2020}. Both datasets were accessed and downloaded through Google
Earth Engine \citep{Gorelick.2017} using the \texttt{rgee} package
\citep{Aybar.2023}. To match hourly values from the temperature and
precipitation data with the 4-hour fixes from the GPS data collection, we
computed average precipitation and temperature values across four hours.

\subsubsection{Anthropogenic Features}

We combined anthropogenic pressures arising from the presence of humans,
agriculture, and roads into a single proxy for human influence. We obtained
information on human density from Facebook's high resolution human density
dataset \citep{Tiecke.2017} which we downloaded from the humdata website
(\url{www.data.humdata.org/}). Likewise, we obtained information on the presence
of agricultural fields from the Globeland30 \citep{Chen.2015} and Cropland
\citep{Xiong.2017} datasets. We downloaded on main tar roads from OpenStreetMaps
\citep{OpenStreetMapContributors.2017}. We combined the different features into
a single layer according to the rules described in \citet{Hofmann.2021}.
Finally, we derived vector data on the distribution of national parks, wildlife
management areas and other protected areas from the Peace Parks Foundation
(\url{www.maps.ppf.org.za/}).

\subsubsection{Nocturnal Variables}

We calculated nocturnal covariates associated with each step using the
\texttt{moonlit} package \citep{Smielak.2023}. The set of covariates included a
binary variable indicating if a step fell into daytime or nighttime (sun $<$ -18
\degree below the horizon), a variable of the moon phase (fraction of the
visible moon relative to full-moon), and an estimate of moonlight illumination
(in lux).

\subsection{Step-Selection Models}

We modeled the habitat and movement preferences of dispersing individuals using
integrated step-selection functions (iSSFs, \citealp{Fortin.2005, Avgar.2016}),
following the procedure described in \citet{Muff.2020}. For this, we first
identified bursts of subsequent GPS locations where the duration between two
recordings did not exceed 4 hours (\(\pm\) 15 minutes). Within each burst, we
converted locations into steps, where a step represented the straight line
between two consecutive locations \citep{Turchin.1998}. For each step, we
computed the associated step-length (sl, in kilometers) and turning-angle (ta,
in radians) relative to the preceding step. Due to the 4-hour regularization,
some individuals dropped from the final dataset, leaving data of
\inputy{GeneralMetrics/SSFTotal} dispersing coalitions
(\inputy{GeneralMetrics/SSFFemales} female coalitions,
\inputy{GeneralMetrics/SSFMales} male coalitions) to work with. The final
dataset comprised \inputy{GeneralMetrics/StepsTotal} steps
(\inputy{GeneralMetrics/StepsMeanSD} per coalition), which we further
categorized into wet and dry season. This resulted in
\inputy{GeneralMetrics/NumberStepsDry} steps
(\inputy{GeneralMetrics/PercentageStepsDry}\%) during the dry season and
\inputy{GeneralMetrics/NumberStepsWet} steps
(\inputy{GeneralMetrics/PercentageStepsWet}\%) during the wet season.

We paired each \textit{observed} step with a set of 100 \textit{random} steps,
generated by sampling turning angles from  a uniform distribution U(\(-\pi,
+\pi\)) and step lengths from a gamma distribution fitted to observed steps
(scale \(\theta\) = \inputy{GeneralMetrics/GammaScale} and shape \(k\) =
\inputy{GeneralMetrics/GammaShape}). Together, an observed step and its 100
associated random steps formed a stratum that received a unique identifier.
Along each step, we extracted covariate values from the underlying spatial
habitat layers (\Cref{Covariates}). For continuous covariates, we computed
average values along each step, for categorical covariates the percentage cover
of each category along the step. To model a decreasing marginal impact of
``distance to'' variables, we included their square-root as predictors. To
facilitate model convergence, we normalized resulting values to a range between
0 and 1. To estimate habitat and movement parameters of interest, we applied the
maximum likelihood procedure proposed by \citet{Muff.2020}, using the
\texttt{glmmTMB} package \citep{Brooks.2017}. We included a stratum-specific
intercept with a fixed and large variance ($10^6$), and used dispersing
coalition ID to model random slopes.

Talk about the model structure here...

To establish whether 100 random steps sufficed to reliably estimate parameters,
we ran a series of univariate models where we only included single covariates
and estimated model coefficients considering 1, 5, 10, 50, 75, and 100 random
steps, respectively (sensu \citealp{Fieberg.2021}). We replicated each model 100
times using randomly resampled steps and we computed the deviation of the
resulting estimate to the estimate emerging under the maximum of 100 random
steps. The results confirmed that model estimates stabilized at ca. 25 random
steps, suggesting that 100 random steps were well beyond sufficient (Figure S8).

Ultimately, we parametrized four models that differed in their degrees of
dynamism in two dimensions, as outlined in (\Cref{GraphicalAbstract}a). Models
one and two were both fit using static covariates (\Cref{GraphicalAbstract}a),
with model one being a single-season model, model two being a two-season model
(wet vs. dry). Models two and three, by contrast, were fit using dynamic
covariates, with model three being a single-season model, and model four being a
multi-season model (wet vs. dry). To fit these models, we used data of 20
individuals ($\sim$ 80\%), retaining the remaining six ($\sim$ 20\%) for
validation purposes.

% created xx candidate models that we ranked based on Akaike's Information
% Criterion (AIC, \citealp{Akaike.1974, Burnham.2002}).

\subsection{Simulation}

Using the four parameterized iSSF models we simulated utilization distributions
(UDs) for each individual in the validation dataset, either assuming a static or
a dynamic landscape (\Cref{GraphicalAbstract}). Each of the retained individuals
dispersed during a different time and from a different origin, which is why we
simulated separate utilization distributions for each individual. For the
simulation, we employed the iSSF-simulation-framework introduced by
\citet{Signer.2017} and employed in \citet{Hofmann.2023}. The procedure starts
by selecting a start-point, which in our case was set at the first location of
an individual's dispersal track. Originating from this location, we then
proposed a set of random steps by sampling turning-angles from the fitted gamma
distribution and turning-angles from a von Mises distribution. Along each of the
so created random steps, we extracted spatial covariates and computed relevant
movement metrics (sl, log(sl), and cos(ta)). We then employed the fitted
model(s) to predict the likelihood of each step for being chosen (i.e. the
redistribution kernel) and sampled one of the steps based on assigned
probabilities. We then updated the simulated individual's position and repeated
the procedure until the number of simulated steps matched that of the observed
individual. We repeated the simulation 20'000 times for each configuration. For
configurations in which the covariates were updated dynamically \textit{during}
the simulation, we generated a new set of covariate layers after every step,
thus rendering conditions as up to date as possible (\Cref{Schematic}). For all
configurations, we finally obtained utilization distributions (heatmaps) base on
simulated locations.

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/Schematic} \caption{Schematic
  illustration of the dispersal simulation with dynamic covariates. As the
  simulation proceeded, we updated underlying environmental conditions
  (symbolized by the stack of layers). To update the covariate stack, we
  utilized series of seasonally updated spatial layers (each grey bars along the
  timeline indicates a new layer). Depending on the covariate, the update
  frequency ranged from a few hours to multiple months. We then proposed a set
  of random steps and applied the parametrized step-selection function to
  generate a redistribution kernel, indicating the probability of each step for
  being chosen. Based on this kernel, one location was sampled and the animal's
  position updated. This procedure was repeated until the number of simulated
  steps matched the number of steps from the observed individuals.}
  \label{Schematic}
 \end{center}
\end{figure}

\subsection{Validation}

To assess the predictive efficacy of each configuration, we employed
the validation dataset and ran a step-selection function, establishing whether
individuals contained within this dataset systematically favored areas of high
predicted utilization. Specifically, we overlaid each of the simulated
utilization distributions with the true, observed dispersal tracks and computed
the average utilization value along each observed and random steps and used a
conditional logistic regression to estimate the selection coefficient
$\beta_{UD}$. Our rationale was that this value should be higher, the better a
model predicted the true dispersal movements.

\section{Results}

\begin{figure}
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/MovementModel.png}
  \caption{This is a caption}
  \label{Movement Model}
 \end{center}
\end{figure}

\inputy{Figures/MovementModel.tex}

\section{Discussion}
Dynamic connectivity increases our understanding of temporal changes in
connectivity \citep{Martensen.2017}.

\citet{Martensen.2017} showed that dynamic connectivity was higher than static
connectivity because of stepping stones that emerged and disappeared again.

\citet{Osipova.2019} in contrast find that connectivity is both over and
underestimated, depending on the season.

\citet{Benz.2016} showed that habitat selection of moose changes depending on the
season and used a combination of RSFs and SSFs to study habitat selection across
seasons.

Badly designed corridors can result in population sinks and wasted financial
resources \citep{Simberloff.1992}.

Could talk a bit about the following papers: \citep{Chetkiewicz.2009, Benz.2016,
Mui.2017, Osipova.2019, Kaszta.2021}

\section{Authors' Contributions}
D.D.H. and G.C. conceived the study and designed methodology; D.D.H., G.C.,
D.M.B, and J.W.M. collected the data; D.D.H. analysed the data; G.C. assisted
with modeling; D.D.H. and G.C. wrote the first draft of the manuscript and all
authors contributed to the drafts at several stages and gave final approval for
publication.

\section{Data Availability}
Something

\section{Acknowledgements}
Something

\newpage
\begingroup
\singlespacing
\ifthenelse{\boolean{usebiblatex}}{
  \begin{refcontext}[sorting=nyt]
  \printbibliography
  \end{refcontext}
}{
  \bibliography{LiteratureBibtex}
}
\endgroup
\end{document}

### 2024-07-01 Chapter_3/04_Manuscript/Manuscript_Support.tex
Publish PHD

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage{scalerel}
\usepackage{makecell}
\usepackage{ifthen}
\usepackage{tikz}
\usepackage{lineno}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%  Bibtex vs Biblatex
%------------------------------------------------------------------------------
% While bibtex is very nice to use, it's insanely slow. I therefore work with
% bibtex until the final stage. Set the value below to "true" or "false" to pick
% which one should be used.
\newboolean{usebiblatex}
\setboolean{usebiblatex}{false}

% Depending on the question, the commands below will prepare the document
% accordingly
\ifthenelse{\boolean{usebiblatex}}{
  %-----------------------------------------------------------------------------
  %  If using biblatex
  %-----------------------------------------------------------------------------
  % Load required packages and refer to correct bibliography
  \usepackage{csquotes}
  \usepackage[
      backend=biber
    , style=apa          % To display Author-year in the text
    , natbib=true        % Necessary to use citep, etc.
    , hyperref=true      % To have clickable links in the references
    , useeditor=false    % Don't add editor to publications
    , sorting=ynt        % Sort by year, name, and title
    , uniquename=false   % Avoid that initials are put
  ]{biblatex}
  \addbibresource{Literature.bib}

  % To include author in link
  \DeclareFieldFormat{citehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{#1}}
  \DeclareFieldFormat{textcitehyperref}{%
  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
  \bibhyperref{%
    #1%
    \ifbool{cbx:parens}
      {\bibcloseparen\global\boolfalse{cbx:parens}}
      {}}}
  \savebibmacro{cite}
  \savebibmacro{textcite}
  \renewbibmacro*{cite}{%
    \printtext[citehyperref]{%
      \restorebibmacro{cite}%
      \usebibmacro{cite}}}
  \renewbibmacro*{textcite}{%
    \ifboolexpr{
      ( not test {\iffieldundef{prenote}} and
        test {\ifnumequal{\value{citecount}}{1}} )
      or
      ( not test {\iffieldundef{postnote}} and
        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
    }
      {\DeclareFieldAlias{textcitehyperref}{noformat}}
      {}%
    \printtext[textcitehyperref]{%
      \restorebibmacro{textcite}%
      \usebibmacro{textcite}}}

  % Remove editor form everything but books
  \AtEveryBibitem{%
   \ifentrytype{book}{}{%
    \clearname{editor}
   }
  }
}{
  %-----------------------------------------------------------------------------
  %  If using bibtex
  %-----------------------------------------------------------------------------
  % Load required packages and select citation style
  \usepackage[round]{natbib}
  \bibliographystyle{apalike}
}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Link colors
\definecolor{linkcolor}{HTML}{000000}
\hypersetup{
  colorlinks = true,
  linkcolor  = linkcolor,
  urlcolor   = linkcolor,
  citecolor  = linkcolor,
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% Avoid skip when including text from external .tex file
\newcommand{\inputy}[1]{\input{#1}\unskip}

% For supplementary material
\newcommand{\beginappendix}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
  \setcounter{equation}{0}
  \renewcommand{\theequation}{Equation S\arabic{equation}}%
  \setcounter{section}{0}
  \renewcommand{\thesection}{A.\arabic{section}}%
}

%------------------------------------------------------------------------------
%  ORCID
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Create the ORCID logo
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol]
      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
        ,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white]
      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

% Command to create the ORCID
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{\textbf{Appendix}\\ Individual-Based Dispersal Simulations to Study
Dynamic Connectivity in Seasonal Landscapes}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1, 2} \orcid{0000-0001-7477-2642} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
Private Bag 13, Maun, Botswana.

\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}

\vspace{4cm}

\textbf{Running Title:} Simulating Dispersal in Seasonal Landscapes

\vspace{0.5cm}

\textbf{Keywords:} African wild dog, Connectivity, Dispersal, Dynamism,
Individual-based, Lycaon pictus, Okavango Delta, Seasonality

\end{flushleft}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

% Change to appendix counters
\appendix
\beginappendix

%------------------------------------------------------------------------------
%	Appendix S1: Pan Mapping
%------------------------------------------------------------------------------
\newpage
\section{Pan Mapping}
\subsection{Satellite Product Overview}

We dynamically mapped the spatial distribution of ephemeral water bodies that
span only a few meters in diameter (\Cref{Pan}) using a custom remote sensing algorithm.
These waterbodies are typically referred to as ``pans'' and mainly inundate
during the rainy season, after which the cumulated water slowly evaporates.
Previously, we developed an algorithm to detect large scale floodwaters across
the extent of the Okavango Delta. This algorithm made use of MODIS MCD34A4
satellite imagery, which is a 8-day composite of daily updated MODIS satellite
data. The benefit of MODIS' high temporal resolution is that monthly composites
are almost guaranteed to be cloud free, even during Botswana's rainy season (mid
October to mid May). However, due to its coarse spatial resolution of 250
meters, MODIS satellite data was unsuitable to detect finer-scaled waterbodies
on satellite images. In an attempt to overcome this limitation, we evaluated
alternative satellite products that provided better spatial resolution than
MODIS, while retaining a sufficiently high temporal resolution to render
seasonal patterns. Candidate products comprised Landsat 7, Landsat 8, and
Sentinel 2 satellite imagery. Data associated with each of these satellites are
freely accessible and provide spatial resolutions between 10 and 60 meters,
depending on the respective bands. Furthermore, the satellites have
revisit-times between 5 and 16 days, thus allowing to generate frequently
updated composite images (\Cref{Satellite}).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/Pan}
  \caption{A pan as it can be found in Botswana right after the rainy season.
  This particular pan is comparably small at around 10 meters in diameter. Some
  of these pans dry out quickly, whereas others provide a source of water for
  the entire dry-season. Here, two spotted hyenas (\textit{Crocuta crocuta})
  take a quick bath.}
  \label{Pan}
 \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Overview of the spatial and temporal resolutions of the candidate satellite products.}
  \label{Satellite}
    \begin{threeparttable}[h]
      \input{Figures/Satellites.tex}
      \begin{tablenotes}
        \item \textsuperscript{*} \textit{Spatial resolutions of the same product can differ depending on the band.}
        \item \textsuperscript{\dag} \textit{Sentinel 2's revisitation duration decreased to 5 days after March 2017.}
      \end{tablenotes}
    \end{threeparttable}
  \end{center}
\end{table}

\noindent Our goal was to produce dynamically updated pan-maps that we could
overlay with GPS data of dispersing African wild dogs. Since GPS data was
collected between 2015 and 2022, this implied that satellite data needed to be
available for the same period. At first glance, all satellite products appeared
to meet this criteria. However, Landsat 7's scan-line detector had been failing
since May 31, 2003, introducing data gaps between adjacent tiles, thus rendering
the product virtually unusable for our purposes (but see \citealp{Storey.2005}).
Sentinel 2 and Landsat 8, by contrast, appeared as viable alternatives, as both
of them spanned the desired period and did not exhibit any device failures.

\subsection{Evaluation of Landsat 8 and Sentinel 2}
\subsubsection{Training Polygons}

With Landsat 8 and Sentinel 2 imagery remaining, we conducted a preliminary
investigation to compare the two products and evaluate their suitability for our
needs. Specifically, we investigated how well we could remote sense pans using
either of the two products, utilizing supervised learning methods. For this, we
prepared a set of training polygons, consisting of the land cover classes
dryland, water, and wet-pans. Although our primary goal was to detect wet-pans,
we included the two other classes to facilitate the categorization of
reflectance values into distinct groups. Due to the large study area considered,
\textit{in situ} ground-truthing of the training polygons was impossible, and we
instead opted for \textit{on-screen selection} of training data. More
specifically, we utilized Google Earth to digitize areas that were clearly
identifiable as either dryland, water, or wet-pans. At the highest zoom level in
Google Earth, these categories are visually easy to tell apart
(\Cref{PanMapping}). To ensure a sharp distinction between wet-pans and dryland,
we traced several dryland polygons in areas that were seasonally covered by
water (\Cref{PanMapping}). Google Earth provides the date of any satellite map,
and so we could assign a timestamp to each training-polygon. This was necessary
to later match training polygons with Landsat or Sentinel data of the same date.
Google Earth's ability to display imagery from different dates furthermore
allowed us to generate a training dataset that comprised polygons that belonged
to different classes depending on the season. For example, a dryland polygon
obtained in the dry season could overlap with a wet-pan polygon obtained during
the rainy season. Overall, we generated \inputy{GeneralMetrics/TrainingTotal}
polygons of varying size and shape (\inputy{GeneralMetrics/TrainingDryland} on
dryland, \inputy{GeneralMetrics/TrainingWater} in water, and
\inputy{GeneralMetrics/TrainingWetpan} in wet-pans) for
\inputy{GeneralMetrics/TrainingDates} distinct dates (August 18, 2018, March 25,
2019, and July 24, 2021).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/PanMapping.pdf}
  \caption{Example of a \textit{wet pan} and \textit{dryland} training polygon
  digitized on Google Earth. The extent of water can easily be gauged at the
  maximum zoom level in Google Earth. While the dryland polygon in this case
  covered an area that is seasonally covered by water, we placed other dryland
  polygons in areas that were never inundated. However, to ensure reliable
  differentiation between wet and dry pans, we included several dryland polygons
  located in dry pans.}
  \label{PanMapping}
 \end{center}
\end{figure}

\subsubsection{Satellite Data Download}
\label{Download}

We downloaded Landsat 8 imagery through Google Earth Engine using the
\texttt{rgee} \texttt{R}-package \citep{Aybar.2024} and used the associated
quality band to mask pixels classified as clouds or shadows. We also computed
several normalized difference indices (NDs), as listed in \Cref{ND}. To download
Sentinel 2 data, we used the \texttt{sen2r} \texttt{R}-package
\citep{Ranghetti.2020} which provides a standardized interface to download and
process Sentinel 2 imagery from the European Space Agency's data hub. Sentinel 2
data are available as either top-of-the-atmosphere (TOA, level 1C) or
bottom-of-the-atmosphere (BOA, level 2A) reflectance values. In analyses where
temporal trends are considered, the use of BOA reflectances is recommended, as
differences in reflectance properties due to atmospheric conditions are
accounted for \citep{Gilabert.1994, Vermote.2008, Chraibi.2022}. However, a
large portion of the publicly available Sentinel 2 data has not yet been
processed from level 1C to level 2A, requiring researchers to apply the
corrections themselves. Fortunately, the European Space Agency provides the
correction tool \texttt{sen2cor} for free \citep{Main-Knorn.2017}, and this tool
has been integrated into the \texttt{R}-package \texttt{sen2r}
\citep{Ranghetti.2020}. We therefore applied the necessary corrections after
download for any product that hadn't already been corrected. Furthermore, we
used Sentinel 2's ``Scene Classification Map'', to remove pixels classified as
either clouds or shadows. Finally, we computed the same ND-indices as for
Landsat 8 (\Cref{ND}).

\begin{table}[h]
  \begin{center}
  \caption{We computed normalized difference indices between certain bands,
  hoping they would improve the land-cover classifiers. Depending on the
  satellite, we used different bands to compute similar indices. The function to
  compute a normalized difference between bands $b_1$ and $b_2$ is given by
  $\frac{b1 - b2}{b1 + b2}$.}
  \label{ND}
  \begin{threeparttable}[h]
    \input{Figures/ND.tex}
  \end{threeparttable}
  \end{center}
\end{table}

\subsubsection{Land Cover Classifiers}

To train a land-cover classifier, we generated 3'000 random points within
training polygons following a stratified equal random sampling scheme
\citep{Shetty.2021}. That is, we ensured that 1'000 random points were sampled
per training category. Stratified random sampling was necessary to ensure that
wet-pans, our category of interest, was sampled frequently enough, as it
represented a minority class, only making up of
(\inputy{GeneralMetrics/PercentageTrainingWetpan}\%) the total training area.
Stratified equal random sampling has been shown to provide good class-level
accuracy for minority classes \citep{Shetty.2021}, so we deemed this approach
suitable for our purposes. We also kept track of the dates of the underlying
polygons from which random points were generated. At each random point, we then
extracted reflectance values of Sentinel 2 and Landsat 8 data that temporally
aligned with the date assigned to the random point (\Cref{Reflectances}a). For
instance, if a random point fell into a polygon that was digitized using a
Google Earth image generated on Aug 18, 2018, we extracted values from the
Sentinel 2 and Landsat 8 layers that were closest to that date. Finally, we
parameterized a Random Forest (RF) classifier using the \texttt{R}-package
\texttt{randomForest} \citep{Cutler.2024} and a Classification and Regression
Tree (CART) classifier using the \texttt{R}-package \texttt{rpart}
\citep{Therneau.2024b}. In both cases, we included all bands, as well as derived
ND-indices (\Cref{ND}) as explanatory covariates. We visualized the decision
trees for the CART classifiers (\Cref{Reflectances}b) and variable importance
for the RF models (\Cref{Reflectances}c) to investigate the importance of
different bands or indices in separating the three classes. To compute variable
importance we used the \texttt{R}-package \texttt{caret} \citep{Kuhn.2008}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/Reflectances.png}
  \caption{Reflectance properties of the Landsat 8 (a1) and Sentinel 2 (a2)
  bands and NDs at the extracted training points for each of the three
  categories (colored). Based on extracted reflectance values we parametrized
  Classification and Regression Tree models (b1 and b2) as well as Random Forest
  models (c1 and c2).}
  \label{Reflectances}
 \end{center}
\end{figure}

\subsubsection{Validation and Comparison}

To validate the predictive power of the two classifiers across the two datasets,
we employed 5-fold cross-validation. For this, we randomly split the data into 5
groups and repeatedly fitted both classification models using 80\% of the data,
to then predict land cover categories in the remaining 20\%. We generated
confusion matrices to contrast true and predicted labels and computed estimates
of the classifier's specificity, sensitivity, and overall accuracy. The results
from this validation show that both classifiers achieve $\geq$ 90\% accuracy
across both satellite products (\Cref{ClassificationValidation}). In fact, the
RF classifier resulted in an overall accuracy of 99\% on Sentinel 2 data.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/ClassificationValidation.png}
  \caption{Confusion matrices (a) and derived performance metrics (b) for the
  CART and RF classifiers for both the Landsat 8 and Sentinel 2 datasets.}
  \label{ClassificationValidation}
 \end{center}
\end{figure}

\subsection{Bulk Download}

Although both Landsat 8 and Sentinel 2 provided very good results, we considered
Sentinel 2 data to be marginally superior, mainly due to Sentinel's higher
temporal and spatial resolution. A higher temporal resolution was key to
compensate missing data from satellite mages obtained on cloudy days and
therefore pivotal in achieving cloud free monthly composites. We therefore
decided to download Sentinel 2 data for the period during which we also
collected GPS data of dispersing individuals. Instead of downloading all
Sentinel 2 tiles overlapping with our study area and matching the study-period,
we created a spatio-temporal moving window (\Cref{MovingWindows}). This window
was updated each month and comprised all GPS locations collected during that
month, buffered by a 100 km radius. We then identified Sentinel 2 tiles
intersecting with each of the monthly updated moving windows
(\Cref{MovingWindows2}). This resulted in a list of
\inputy{GeneralMetrics/NumberTilesTotal} tiles that needed to be downloaded,
\inputy{GeneralMetrics/NumberTiles2A} of which were already corrected to BOA
reflectances, while the remaining \inputy{GeneralMetrics/NumberTiles1C} tiles
still needed to be corrected. The download followed the same procedure and
included the same corrections (from TOA to BOA, and cloud masking) as outlined
in \Cref{Download} using the \texttt{sen2r} package in \texttt{R}
\citep{Ranghetti.2020}. Upon completion of the download and pre-processing, we
applied the trained RF classifer to obtain binary ``pan-maps'', showing the
spatial distribution of ephemeral water bodies. To generate monthly composites,
we merged all pan-maps falling into the same month and retained pans if they
were detected in at least 50\% of the overlapping tiles. Finally, we generated
``distance-to-pan'' maps using the \texttt{distance} function from the
\texttt{terra} package in \texttt{R} \citep{Hijmans.2024}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/MovingWindows.png}
  \caption{To reduce the number of Sentinel 2 tiles that we needed to download,
  we generated a monthly updated spatio-temporal moving window that comprised
  all GPS data of dispersing wild dogs (black dots) during the respective month,
  buffered by a 100 km radius. Based on the so created moving windows (in blue),
  we identified all overlapping Sentinel 2 tiles (tiles in gray) that we needed
  to download each month, which resulted in a total of 2,226 tiles.}
  \label{MovingWindows}
 \end{center}
\end{figure}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/MovingWindows2.png}
  \caption{Moving windows and associated tiles facetted by months.}
  \label{MovingWindows2}
 \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S2: Moonlight
%------------------------------------------------------------------------------
\newpage
\section{Moon Illumination}

We used the \texttt{R}-package \texttt{moonlit} \citep{Smielak.2023} to obtain
estimates of moon illumination at 5 minute intervals at the average location of
dispersal GPS data (lon = of 23.5, lat = -19.0; \Cref{Moonlight}). The
\texttt{moonlit} package is currently not on \texttt{CRAN}, but can be installed
from GitHub (\url{https://github.com/msmielak/moonlit}). Besides providing
information on the moon cycle, this package also allows estimating the amount of
moonlight illumination on the ground. For instance, even during full-moon
nights, the moon may only appear at a low angle in the sky, thus providing
minimal illumination. Estimates from the \texttt{moonlit} package can therefore
be viewed as biologically more relevant. To match the temporal resolution of our
GPS data during dispersal, we calculated four hourly moonlight summaries.
Specifically, we computed if an interval fell into nighttime and the average
amount of moon-illumination during the four hours.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = 0.75\textwidth]{Figures/Moonlight.png}
  \caption{Moonlight intensity (illumination) as estimated from the
  \texttt{moonlit} package \citep{Smielak.2023} in \texttt{R} calculated for a
  couple of example timestamps in 2023.}
  \label{Moonlight}
 \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S3: Light-Type of Each Step
%------------------------------------------------------------------------------
\newpage
\section{Light-Type}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/Light.pdf}
  \caption{Schematic illustration of how we categorized steps into \textit{dark}
  and \textit{bright} steps. First, we used the \texttt{suncalc} and
  \texttt{moonlit} packages to obtain estimates of the sun-angle and moon
  illumination at 5 minute intervals for each date of interest. Whenever the sun
  angle was $\leq $ -18\degree, we deemed the respective period to be at night.
  A night was either bright (moon illumination $>$ 0.02 of maximum illumination)
  or dark (moon illumination $\leq$ 0.02 of maximum illumination). Any other
  period was considered to provide enough illumination for wild dogs to move and
  therefore considered as bright. Since each step covered a time-span of
  approximately four hours, we defined a step as bright if at least 25\% (i.e.,
  one hour) of it occurred during a bright period. The gray arrows represent
  some example steps that are categorized as bright (light gray) or dark (dark
  gray).}
  \label{Light}
 \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S4: Movement Model Results
%------------------------------------------------------------------------------
\newpage
\section{Movement Model Results}

\begin{table}[htbp]
 \begin{center}
  \caption{Estimates obtained from the \textbf{simple} (no interactions)
  integrated step-selection models fitted using \textbf{static} covariates. GPS
  Data was either pooled across seasons (labeled ``all'') or split into dry and
  wet season. Only underlined covariates differed between the static and dynamic
  configurations.}
  \label{ModelsStaticSimple}
   \begin{threeparttable}
   \inputy{Figures/MovementModelStaticSimple.tex}
   \end{threeparttable}
 \end{center}
\end{table}

\begin{table}[htbp]
 \begin{center}
  \caption{Estimates obtained from the \textbf{simple} (no interactions)
  integrated step-selection models fitted using \textbf{dynamic} covariates. GPS
  Data was either pooled across seasons (labeled ``all'') or split into dry and
  wet season. Only underlined covariates differed between the static and dynamic
  configurations.}
  \label{ModelsDynamicSimple}
   \begin{threeparttable}
   \inputy{Figures/MovementModelDynamicSimple.tex}
   \end{threeparttable}
 \end{center}
\end{table}

\begin{table}[htbp]
 \begin{center}
  \caption{Estimates obtained from the \textbf{full} (with interactions)
  integrated step-selection models fitted using \textbf{static} covariates. GPS
  Data was either pooled across seasons (labeled ``all'') or split into dry and
  wet season. Only underlined covariates differed between the static and dynamic
  configurations.}
  \label{ModelsStaticFull}
   \begin{threeparttable}
   \inputy{Figures/MovementModelStaticFull.tex}
   \end{threeparttable}
 \end{center}
\end{table}

\begin{table}[htbp]
 \begin{center}
  \caption{Estimates obtained from the \textbf{full} (no interactions)
  integrated step-selection models fitted using \textbf{dynamic} covariates. GPS
  Data was either pooled across seasons (labeled ``all'') or split into dry and
  wet season. Only underlined covariates differed between the static and dynamic
  configurations.}
  \label{ModelsDynamicFull}
   \begin{threeparttable}
   \inputy{Figures/MovementModelDynamicFull.tex}
   \end{threeparttable}
 \end{center}
\end{table}

%------------------------------------------------------------------------------
%	Appendix S5: Stability with Regards to Number of Random Steps
%------------------------------------------------------------------------------
\newpage
\section{Number of Random Steps}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/Stability.png}
  \caption{Results from the integrated step-selection models when considering
  different numbers of random steps. Results are shown for the simple and full
  model for both all configurations of fitting covariates (static vs dry) and
  model seasons (all vs. wet + dry).}
  \label{Stability}
 \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S6: Dispersal Model Interpretation
%------------------------------------------------------------------------------
\newpage
\section{Model Interpretation}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/MovementModelInterpretation.png}
  \caption{Relative selection scores (RSS) with respect to environmental
  covariates. RSS scores were computed by comparing a location $x_1$ with
  average conditions to a location $x_2$ with average conditions but the
  respective covariate varied between its minimum and maximum observed value.
  Confidence intervals overlapped substantially, hence we omitted them from the
  figure to simplify with the interpretation of the differences among
  configurations.}
  \label{MovementModelInterpretation}
 \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S7: Spearman Rank Correlation in Relation to Number of Random Steps
%------------------------------------------------------------------------------
\newpage
\section{Spearman's Rank Correlation in Relation to the Number of Random Steps}

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{Figures/RankCorrelationNumberSteps.png}
  \caption{Illustration how the Spearman's rank correlation changes when the
  number of random steps per stratum is varied from 5 to 100 (facettes). The
  y-axis shows the frequency at which the observed step was assigned the rank
  indicated on the x-axis. A low rank indicates that the observed step was
  assigned a high probability of being chosen. The better the prediction, the
  more frequently the observed step should be assigned a low rank, thus
  resulting in negative Spearman's rank-correlation. However, it is obvious that
  the metric heavily depends on the number of random steps per stratum. If there
  are only a few random steps, the correlation is more likely to be negative.
  The show patterns are based on simulated data.}
  \label{RankCorrelationNumberSteps}
 \end{center}
\end{figure}

\newpage
\begingroup
\singlespacing
\ifthenelse{\boolean{usebiblatex}}{
  \begin{refcontext}[sorting=nyt]
  \printbibliography
  \end{refcontext}
}{
  \bibliography{LiteratureBibtex}
}
\endgroup
\end{document}

### 2024-07-01 Chapter_4/02_R-Scripts/03_AppliedExamples/Example_01.Rmd
Publish PHD

---
title : 'Example 1: Dynamic + Model Approach with Simulated Data'
author: "David D. Hofmann, Gabriele Cozzi, John Fieberg"
date  : "2024-01-30"
output:
  html_document:
    toc      : true
    toc_depth: 2
---
```{r setup, echo = FALSE}

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Purpose
This example analysis serves to illustrate the application of the _dynamic +
model_ approach with simulated data. We're first going to simulate a trajectory
with known movement kernel and habitat-selection parameters (we call this the
truth). We will then introduce data-missingness by randomly removing some of the
simulated data-points. Finally, we will apply the _dynamic + model_ approach and
try to obtain the true simulation parameters.

## Disclaimer
In this script, we'll utilize a set of custom functions that are stored in the
`Functions.R` file. They are used to simulate covariates and movement, compute
step-metrics, generate random-steps, and fit step distributions dynamically.
Most of the functions can be replaced by much more generalized functions from
the `amt` R-package. However, `amt`'s `random_steps()` function doesn't cope
well with sampling random steps from distributions fit to different
step-durations. We'll therefore use of our custom code. If you'd like to apply
it to your own data, inspect the functions closels and verify that they suit
your purposes.

## Preamble
Let's start by preparing our R-Session and loading all necessary packages. We
also load the custom functions stored in the `Functions.R` file.

```{r, warning = F, message = F}

# Set notation standards
options(scipen = 999)

# Clear R's brain
rm(list = ls())

# Load required packages
library(tidyverse)     # For data wrangling
library(sf)            # For plotting spatial features
library(lubridate)     # To handle dates and times
library(raster)        # For spatial data manipulation
library(amt)           # To fit step-length and turning-angle distributions
library(survival)      # To fit conditional logistic regression model
library(ggh4x)         # For nested ggplot facets

# Load custom functions
source("02_R-Scripts/Functions.R")

```

## Simulation
We start by simulating a set of raster layers that represent the landscape on
which we want to simulate movement. For this, we'll use the functions
`simDistance()`, `simElevation()`, and `simForest()`, which allow simulating
spatially auto-correlated layers. You may also use functions from other
packages, such as the `NLMR` or `gstat` packages, to generate such layers.

```{r}

# Set a seed for reproducability
seed <- 123456
set.seed(seed)

# Simulate landscape covariates (n x n)
n      <- 300
dist   <- simDistance(n, x = n / 2, y = n / 2)
elev   <- simElevation(n, autocorr_range = 10, seed = seed)
forest <- simForest(n, autocorr_range = 10, prop = 0.5, seed = seed + 1)
cov <- stack(dist, elev, forest)
names(cov) <- c("dist", "elev", "forest")

# Visualize the simulated covariates
cov %>%
  as.data.frame(xy = T) %>%
  pivot_longer(dist:forest, names_to = "Covariate", values_to = "Value") %>%
  ggplot(aes(x = x, y = y, fill = Value)) +
    geom_raster() +
    scale_fill_viridis_c() +
    scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
    facet_wrap(~ Covariate) +
    coord_equal() +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

On this landscape, we now simulate a trajectory using the `move()` function. The
function requires quite a few inputs, such as the starting location (`xy`), the
covariate layers (`covars`), as well as a model formula (`formula`) and
associated betas (`prefs`) that determine the habitat-selection function.
Furthermore, we need to provide the step-length and turning-angle distributions
from which random steps are sampled (`sl_dist` and `ta_dist`). The `move`
function assumes a gamma distribution (with shape and scale parameters) for
step-lengths and a von Mises distribution (with a concentration parameter kappa)
for turning-angles. Finally, there are some simulation parameters that we need
to define, including the extent on which simulated animals are allowed to move
(`ext`), the number of steps simulated (`n_steps`), the number of random-steps
proposed at each relocation (`n_rsteps`), whether the simulation should stop if
one of the random steps leaves the study area (`stop`) and if a
simulation-progressbar should be printed (`messages`).

```{r, message = F}

# Simulate a trajectory across this landscape
sim <- move(
    xy       = cbind(n / 2, n / 2)
  , covars   = cov
  , formula  = ~ dist + elev + forest
  , prefs    = c(-20, 1, -1)
  , sl_dist  = list(name = "gamma", params = list(shape = 3, scale = 1))
  , ta_dist  = list(name = "vonmises", params = list(kappa = 0.5, mu = 0))
  , ext      = as(extent(c(0, n, 0, n)), "SpatialPolygons")
  , n_steps  = 1000
  , n_rsteps = 25
  , stop     = F
  , messages = F
)

# Give the individual a random name
sim$id <- "Hermes"

# Add timestamps
sim$timestamp <- ymd_hms("2023-01-01 09:00:00") + hours(1) * sim$step_number

# Check it
head(sim, n = 10)

```

The `sim` data-frame represents the full "TRUE" movement trajectory of the
simulated animal. Note that we simulated that fixes were collected exactly on
the hour. This is usually not the case in real applications, but simplifies the
subsequent code a bit. In any case, it's straight forward to round timestamps of
real data to the nearest hour or minute to avoid issues with slight deviations
from the anticipated schedule. To mimic what real data would look like, we
should also remove the step-metrics (`absta`, `ta`, `sl`), as they are usually
not observed in reality and later have to be inferred from observed coordinates.

```{r}

# Remove columns that are not observed in reality
sim[, c("absta", "ta", "sl", "step_number")] <- NULL
head(sim, n = 10)

# Visualize track
ggplot(sim, aes(x = x, y = y, col = timestamp)) +
  geom_path() +
  scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
  coord_equal() +
  theme_minimal()

```

For later, we should keep track of the true simulation parameters. This will
allow us to compare our model estimates from the iSSF model to the true
parameters.

```{r}

# Keep track of true simulation parameters
truth <- data.frame(
    Term  = c("dist", "elev", "forest", "shape", "scale", "kappa")
  , Truth = c(-20, 1, -1, 3, 1, 0.5)
)

```

Now we rarify the full trajectory and create an "incomplete" dataset where some
locations are missing.

```{r}

# Rarify track by randomly removing 25% of the fixes. You'll see that the number
# of fixes drops from 1000 to 750. We can then give each step an id (unique
# across individuals if there are many) and number (not unique across
# individuals if there are many)
sim_rarified <- sim %>%
  rarifyData(missingness = 0.25) %>%
  mutate(
      step_id     = 1:n()
    , step_number = 1:n()
  )

# Compare number of fixes
cbind(Original = nrow(sim), Rarified = nrow(sim_rarified))

```

The rarified trajectory is our point of departure and represents what we would
observe in reality. Let's again visualize everything.

```{r}

# Visualize covariates and the rarified track
cov_df <- cov %>%
  as.data.frame(xy = T) %>%
  pivot_longer(dist:forest, names_to = "Covariate", values_to = "Value")
ggplot() +
  geom_raster(data = cov_df, aes(x = x, y = y, fill = Value)) +
  geom_path(data = sim_rarified, aes(x = x, y = y, col = timestamp)) +
  scale_fill_viridis_c() +
  scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
  facet_wrap(~ Covariate) +
  coord_equal() +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

## Fitting Tentative Distributions
Because the dataset is incomplete, the time between consecutive datapoints is
not always equal. Let's figure out the regular (i.e. most frequent)
step-duration.

```{r}

# Assess the regular step-duration of our data
samp <- sim_rarified %>%
  computeDurations() %>%
  count(duration)
samp

```

As expected (since we simulated so), the regular step-duration is 1 hour.
However, we also observe that in a few cases the duration between two fixes can
be as long as `r max(samp$duration)` hours. In regular iSSF, we would only
consider bursts within which step-durations are somewhat comparable. Here,
however, we are tolerant towards unequal step-durations, as we'll later account
for differing step-durations in the model. Consequently, we would like to
accommodate for differing durations by fitting tentative distributions for
different step-durations. Since there are barely any fixes that are separated >
3 hours apart, having a forgiveness larger than 3 won't add much. Let's
therefore fix our forgiveness to 3 and fit the associated step-length and
turning angle distributions using the `fitDists()` function. Depending on the
missingness, some step-durations may not occur very often, making it difficult
to parametrize the associated gamma and von Mises distributions. We can
alleviate this problem by resampling the observed data to different
step-durations (option `resample = T`) or by increasing missingness even more
(option `rarify = T`). Here, we chose to use the resampling option.

```{r}

# Fit step-length and turning angle distributions to different durations
dists <- fitDists(sim_rarified
  , durations = 1:3 # Return fitted distributions for durations 1-3
  , regular   = 1   # The regular step-duration is 1
  , dynamic   = T   # Fit distributions dynamically
  , resample  = T   # Resample track to the different durations
  , rarify    = F   # Don't increase missingness
)

# Show the fitted values
dists

```

You can see that we now have different gamma and von Mises parameters for steps
from different durations. We can later use the so fitted distributions to
generate random steps matching different step-durations.

## Generating Random Steps
Before generating random steps, we need to split our data into bursts within
which the step-duration does not exceed 3.

```{r}

# Compute bursts within which the step duration does not exceed our forgiveness
sim_rarified_bursted <- computeBursts(sim_rarified, max_duration = 3)
head(sim_rarified_bursted)

```

In addition, we need to compute step-metrics (step-lengths and turning-angles)
per burst.

```{r}

# Compute step metrics within bursts
sim_rarified_metrics <- computeMetrics(sim_rarified_bursted)
head(sim_rarified_metrics)

```

Finally, we can use the `computeSSF()` function to generate 25 random steps for
each valid step (a valid step has a step-length and a turning-angle associated).
Note that we cna use the dynamically fitted distributions `dists` to match
sampled random steps with the step-durations of observed steps by picking the
`dynamic+model` approach. The function `computeSSF()` will also automatically
drop any invalid steps or steps with a duration that occurs less than 5 times in
the dataset. This is to prevent issues when including the step-duration in the
iSSF model.

```{r}

# Generate random steps
sim_rarified_ssf <- computeSSF(sim_rarified_metrics
  , n_rsteps = 25
  , dists    = dists
  , approach = "dynamic+model"
)
head(sim_rarified_ssf)

```

Let's visualize some of the observed and random steps to ensure everything is
working as expected.

```{r}

# Verify that steps of different durations follow different sl- and
# ta-distributions
sim_rarified_ssf %>%
  dplyr::select(case, duration, sl, relta) %>%
  pivot_longer(sl:relta, names_to = "Metric", values_to = "Value") %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  ggplot(aes(x = Value, col = as.factor(duration))) +
    geom_density() +
    facet_nested_wrap(~ Type + Metric, scale = "free") +
    scale_color_viridis_d(name = "Step-Duration") +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

# Visualize some observed and random steps
sim_rarified_ssf %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  subset(step_number %in% 1:1000) %>%
  ggplot(aes(x = x, y = y, xend = x_to, yend = y_to, col = as.factor(duration))) +
    geom_segment(linewidth = 0.1) +
    scale_color_viridis_d(begin = 0.3, name = "Duration") +
    theme_minimal() +
    coord_equal() +
    facet_wrap(~ Type) +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

## Computing Covariates
We still need to compute all covariates that we want include our iSSF model.
This includes spatial covariates at the end-point of each step, as well as
step-metrics. We can use the function `computeCovars` for this.

```{r}

# Compute covariates for all steps (the function will extract spatial covariates
# at the end of each step and also compute log_sl and cos_ta)
sim_rarified_covars <- computeCovars(sim_rarified_ssf
  , covars    = cov
  , extract   = "end"
)
head(sim_rarified_covars)

```

## Running the Model

We are now ready to estimate parameters using conditional logistic regression.
This can be achieved using the `clogit()` function from the `survival` package.
Alternatively, one can use the wrapper `fit_clogit()` from the `amt` package.
Importantly, we will include interactions among the step-descriptors $sl$,
$log(sl)$, and $cos(ta)$ and the step-duration $duration$.

```{r}

# Ensure the step-duration enters as factor
sim_rarified_covars$duration <- as.factor(sim_rarified_covars$duration)

# Fit the model
mod <- clogit(data = sim_rarified_covars, formula = case ~
  + dist            # Habitat-selection
  + elev            # Habitat-selection
  + forest          # Habitat-selection
  + sl              # Movement Kernel
  + log_sl          # Movement Kernel
  + cos_ta          # Movement Kernel
  + sl:duration     # Movement x Step-Duration
  + log_sl:duration # Movement x Step-Duration
  + cos_ta:duration # Movement x Step-Duration
  + strata(step_id)
)
summary(mod)

```

This looks good. Now we extract the model coefficients from the fitted model and
use them to update our "tentative" distribution parameters for the gamma and von
Mises distribution. Since the true data was simulated with $\Delta t = 1$, we
will only update parameters for the case when $\Delta t = 1$.

```{r}

# Compute confidence intervals
ci <- confint(mod, level = 0.95)

# Put relevant results together
coefs <- summary(mod)$coefficients
coefs <- data.frame(
    Coefficient = rownames(coefs)
  , Estimate    = coefs[, "coef"]
  , SE          = coefs[, "se(coef)"]
  , Z           = coefs[, "z"]
  , LCI         = ci[, 1]
  , UCI         = ci[, 2]
)

# We can use this dataframe to update our tentative distribution parameters.
corrected <- correctDists(list(
    Dists = dists
  , Coefs = coefs
))
corrected$Dists$updated

```

Finally, we put all results together into a `data.frame` and compare them to the
truth.

```{r}

# Now put everything together into a dataframe
mov_estimates <- corrected$Dists
hab_estimates <- corrected$Coefs

# Compile the updated distribution parameters (for step duration of one)
mov_estimates <- tibble(
    shape = as.numeric(mov_estimates$updated$sl$shape)
  , scale = as.numeric(mov_estimates$updated$sl$scale)
  , kappa = as.numeric(mov_estimates$updated$ta$kappa)
  ) %>%
  pivot_longer(shape:kappa, names_to = "Term", values_to = "Estimate") %>%
  mutate(Kernel = "Movement")

# Compile the habitat selection coefficients
hab_estimates <- hab_estimates %>%
  dplyr::select(Term = Coefficient, Estimate) %>%
  subset(Term %in% c("forest", "dist", "elev")) %>%
  mutate(Kernel = "Habitat")

# Put them together
estimates <- rbind(hab_estimates, mov_estimates, make.row.names = F)
estimates

# Compare to truth
comparison <- left_join(truth, estimates, by = "Term")
comparison

```

Wonderful. It looks like we were able to approximate the true simulation
parameters quite well.

```{r}
sessionInfo()
```

### 2024-07-01 Chapter_4/02_R-Scripts/03_AppliedExamples/Example_02.Rmd
Publish PHD

---
title : 'Example 2: Dynamic + Model Approach with Real Data'
author: "David D. Hofmann, Gabriele Cozzi, John Fieberg"
date  : "2024-01-30"
output:
  html_document:
    toc      : true
    toc_depth: 2
---
```{r setup, echo = F}

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Purpose
This example analysis serves to illustrate the application of the _dynamic +
model_ approach with real data collected on a spotted hyena (*Crocuta crocuta*)
called "Apollo". First, we'll do some slight pre-processing of the GPS data.
Since missingness is really low in this dataset, we'll artificially increase it
to a higher level. We will then apply the _dynamic + model_ approach to estimate
habitat and movement parameters.

## Disclaimer
In this script, we'll utilize a set of custom functions that are stored in the
`Functions.R` file. They are used to compute step-metrics, generate
random-steps, and fit step distributions dynamically. Most of the functions can
be replaced by much more generalized functions from the `amt` R-package.
However, `amt`'s `random_steps()` function doesn't cope well with sampling
random steps from distributions fit to different step-durations. We'll therefore
use of our custom code. If you'd like to apply it to your own data, inspect the
functions closels and verify that they suit your purposes.

## Preamble
Let's start by preparing our R-Session and loading all necessary packages. We
also load the custom functions stored in the `Functions.R` file.

```{r, warning = F, message = F}

# Set notation standards
options(scipen = 999)

# Clear R's brain
rm(list = ls())

# Load required packages
library(tidyverse)     # For data wrangling
library(sf)            # For plotting spatial features
library(lubridate)     # To handle dates and times
library(raster)        # For spatial data manipulation
library(amt)           # To fit step-length and turning-angle distributions
library(survival)      # To fit conditional logistic regression model
library(ggh4x)         # For nested ggplot facets

# Load custom functions
source("02_R-Scripts/Functions.R")

```

## Pre-Processing
We start by loading the GPS data of a spotted hyena (*Crocuta crocuta*) and some
covariate layers.

```{r}

# Load hyena gps data and covariates
dat <- read_rds("03_Data/Hyena_GPS.rds") %>% subset(DOP < 10)
cov <- stack("03_Data/Hyena_Covariates.tif")

```

The hyena GPS data should have been collected every 2 hours, starting from 16:00
and until 04:00 o'clock each day. There usually is also a fix collected at 10:00
o'clock. Since hyenas are mainly nocturnal, we're only going to focus on the
nightly data and drop the 10:00 o'clock fix. In very rare cases, two fixes fall
on the same hour. We'll drop such "duplicates" to avoid having step-durations
that are less than 2 hours.

```{r}

# Keep only fixes that were collected on the desired hour and drop potential
# duplicates
dat <- dat %>%
  mutate(timestamp = round_date(timestamp, "hour")) %>%
  subset(hour(timestamp) %in% c(16, 18, 20, 22, 0, 2, 4)) %>%
  distinct(id, timestamp, .keep_all = T)

# Assess the regular step-duration of our data
samp <- dat %>%
  computeDurations() %>%
  count(duration)
samp

```

If you take a look at the durations between subsequent fixes, you'll find that
most fixes are taken at 2-hourly intervals, with only few 4- or 6-hourly
periods. This means that the data is pretty complete (i.e. there is relatively
low missingness). To showcase the applicability of the *dynamic+model* approach
we'll introduce a slightly higher level of missingness by removing a fraction of
the collected GPS fixes.

```{r}

# Incrase missingness a bit
dat_rarified <- dat %>%
  rarifyData(missingness = 0.15) %>%
  mutate(
      step_id     = 1:n()
    , step_number = 1:n()
  )

# Compare number of fixes
cbind(Original = nrow(dat), Rarified = nrow(dat_rarified))

```

Let's visualize everything, including the spatial covariate layers.

```{r}

# Visualize covariates and the rarified track (normalization is only for
# visualization purposes)
cov_df <- cov %>%
  as.data.frame(xy = T) %>%
  pivot_longer(Water:Trees, names_to = "Covariate", values_to = "Value") %>%
  group_by(Covariate) %>%
  mutate(Value = (Value - min(Value, na.rm = T)) / (max(Value, na.rm = T) - min(Value, na.rm = T)))
ggplot() +
  geom_raster(data = cov_df, aes(x = x, y = y, fill = Value)) +
  geom_path(data = dat_rarified, aes(x = x, y = y, col = timestamp), linewidth = 0.1) +
  scale_fill_viridis_c(breaks = c(0, 0.5, 1), labels = c("Low", "Medium", "High")) +
  scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
  facet_wrap(~ Covariate) +
  coord_equal() +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

## Fitting Tentative Distributions
Because we rarified the dataset, the time between consecutive datapoints is not
always equal. Let's ensure the regular step-duration is still 2 hours.

```{r}

# Assess the regular step-duration of our data
samp <- dat_rarified %>%
  computeDurations() %>%
  count(duration)
samp

```

As expected, the regular step-duration is 2 hours. However, we also observe that
in a few cases the duration between two fixes can be as long as `r
max(samp$duration)` hours. In regular iSSF, we would only consider bursts within
which step-durations are somewhat comparable. Here, however, we are tolerant
towards unequal step-durations, as we'll later account for differing
step-durations in the model. Consequently, we would like to accommodate for
differing durations by fitting tentative distributions for different
step-durations. Here, allowing for steps of up to 6 hours appears sensible. Let
us fit the associated step-length and turning angle distributions using the
`fitDists()` function.

```{r}

# Fit step-length and turning angle distributions to different durations
dists <- fitDists(dat_rarified
  , durations = c(2, 4, 6) # Return fitted distributions for durations 1-3
  , regular   = 2          # The regular step-duration is 2
  , dynamic   = T          # Fit distributions dynamically
  , resample  = F          # Don't resample the track
  , rarify    = F          # Don't increase missingness
)

# Show the fitted values
dists

```

You can see that we now have different gamma and von Mises parameters for steps
from different durations. We can later use the so fitted distributions to
generate random steps matching different step-durations.

## Generating Random Steps
Before generating random steps, we need to split our data into bursts within
which the step-duration does not exceed 6.

```{r}

# Compute bursts within which the step duration does not exceed our forgiveness
dat_rarified_bursted <- computeBursts(dat_rarified, max_duration = 6)
head(dat_rarified_bursted)

```

In addition, we need to compute step-metrics (step-lengths and turning-angles)
per burst.

```{r}

# Compute step metrics within bursts
dat_rarified_metrics <- computeMetrics(dat_rarified_bursted)
head(dat_rarified_metrics)

```

Finally, we can use the `computeSSF()` function to generate 25 random steps for
each valid step (a valid step has a step-length and a turning-angle associated).
Note that we cna use the dynamically fitted distributions `dists` to match
sampled random steps with the step-durations of observed steps by picking the
`dynamic+model` approach. The function `computeSSF()` will also automatically
drop any invalid steps or steps with a duration that occurs less than 5 times in
the dataset. This is to prevent issues when including the step-duration in the
iSSF model.

```{r}

# Generate random steps
dat_rarified_ssf <- computeSSF(dat_rarified_metrics
  , n_rsteps = 25
  , dists    = dists
  , approach = "dynamic+model"
)
head(dat_rarified_ssf)

```

Let's visualize some of the observed and random steps to ensure everything is
working as expected.

```{r}

# Verify that steps of different durations follow different sl- and
# ta-distributions
dat_rarified_ssf %>%
  dplyr::select(case, duration, sl, relta) %>%
  pivot_longer(sl:relta, names_to = "Metric", values_to = "Value") %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  ggplot(aes(x = Value, col = as.factor(duration))) +
    geom_density() +
    facet_nested_wrap(~ Type + Metric, scale = "free") +
    scale_color_viridis_d(name = "Step-Duration") +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

# Visualize some observed and random steps
dat_rarified_ssf %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  subset(step_number %in% 1:1000) %>%
  ggplot(aes(x = x, y = y, xend = x_to, yend = y_to, col = as.factor(duration))) +
    geom_segment(linewidth = 0.1) +
    scale_color_viridis_d(begin = 0.3, name = "Duration") +
    theme_minimal() +
    coord_equal() +
    facet_wrap(~ Type) +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

To be able to analyse our trajectory in an iSSF-framework, we need to generate
random steps. We'll use the `amt` package to generate these. However, unlike in
traditional iSSFs, we will generate random steps using tentative distributions
that are matched with the step-durations of observed steps. Let's start by
converting our simulated dataframe into a `track` object (which is actually just
a `data.frame` that contains all elements that `amt` requires to compute
step-metrics).

## Computing Covariates
We still need to compute all covariates that we want include our iSSF model.
This includes spatial covariates at the end-point of each step, as well as
step-metrics. We can use the function `computeCovars` for this.

```{r}

# Compute covariates for all steps (the function will extract spatial covariates
# at the end of each step and also compute log_sl and cos_ta)
dat_rarified_covars <- computeCovars(dat_rarified_ssf
  , covars    = cov
  , extract   = "end"
)
head(dat_rarified_covars)

```

## Running the Model

We are now ready to estimate parameters using conditional logistic regression.
This can be achieved using the `clogit()` function from the `survival` package.
Alternatively, one can use the wrapper `fit_clogit()` from the `amt` package.
Importantly, we will include interactions among the step-descriptors $sl$,
$log(sl)$, and $cos(ta)$ and the step-duration $duration$.

```{r}

# Ensure the step-duration enters as factor
dat_rarified_covars$duration <- as.factor(dat_rarified_covars$duration)

# Fit the model
mod <- clogit(data = dat_rarified_covars, formula = case ~
  + Water                     # Habitat-selection
  + DistanceToWater           # Habitat-selection
  + Trees                     # Habitat-selection
  + Water:duration            # Habitat x Step-Duration
  + DistanceToWater:duration  # Habitat x Step-Duration
  + Trees:duration            # Habitat x Step-Duration
  + sl                        # Movement Kernel
  + log_sl                    # Movement Kernel
  + cos_ta                    # Movement Kernel
  + sl:duration               # Movement x Step-Duration
  + log_sl:duration           # Movement x Step-Duration
  + cos_ta:duration           # Movement x Step-Duration
  + strata(step_id)
)

# Compute confidence intervals
ci <- confint(mod, level = 0.95)

# Put relevant results together
coefs <- summary(mod)$coefficients
coefs <- data.frame(
    Coefficient = rownames(coefs)
  , Estimate    = coefs[, "coef"]
  , SE          = coefs[, "se(coef)"]
  , Z           = coefs[, "z"]
  , LCI         = ci[, 1]
  , UCI         = ci[, 2]
)

# Show results
coefs

```

```{r}
sessionInfo()
```

### 2024-07-01 Chapter_4/04_Manuscript/99_Glossary.tex
Publish PHD

\begin{table}

\caption{\label{tab:Glossary}Glossary of terms. Terms in the glossary are printed in bold at first occurrence in the main text. Definitions are always given in the context of step selection functions (SSFs).}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{}l>{\raggedright\arraybackslash}p{10 cm}}
\toprule
Term & Definition\\
\midrule
\textbf{\cellcolor{gray!6}{Animal locations}} & \cellcolor{gray!6}{A series of telemetry data points that include date, time, longitude, and latitude information, describing when and where an animal was observed or recorded.}\\
\textbf{Step} & A straight line connecting two consecutive locations.\\
\textbf{\cellcolor{gray!6}{Observed step}} & \cellcolor{gray!6}{A step that connects two observed animal locations.}\\
\textbf{Random step} & A step that connects an observed animal location with a random location. Random locations are typically generated by combining an observed animal location with random step lengths and turning angles.\\
\textbf{\cellcolor{gray!6}{Step length}} & \cellcolor{gray!6}{The Euclidean distance of a step.}\\
\addlinespace
\textbf{Turning angle} & A measure of the change in direction between two consecutive steps.\\
\textbf{\cellcolor{gray!6}{Habitat-selection function}} & \cellcolor{gray!6}{A probabilistic description of an animal's habitat preferences. It describes how an animal selects habitat when not constrained by its movement capacity. Also referred to as movement-free habitat-selection function.}\\
\textbf{Movement kernel} & A probabilistic description of an animal's movement capacity. It describes how an animal would move when not constrained by habitat selection. Also referred to as selection-free movement kernel.\\
\textbf{\cellcolor{gray!6}{Trajectory}} & \cellcolor{gray!6}{A sequence of animal locations collected on the same individual.}\\
\textbf{Step duration} & The time interval associated with a particular step, i.e., the time elapsed between two consecutive animal locations.\\
\addlinespace
\textbf{\cellcolor{gray!6}{Regular animal locations}} & \cellcolor{gray!6}{A series of animal locations that have been obtained at regularly spaced time intervals, such as every hour.}\\
\textbf{Irregular animal locations} & A series of animal locations collected at irregular time intervals.\\
\textbf{\cellcolor{gray!6}{Regular step durations}} & \cellcolor{gray!6}{Step durations that occur when animal locations are successfully collected at regular time intervals.}\\
\textbf{Irregular step durations} & Step durations that occur when animal locations are not successfully collected at regular time intervals.\\
\textbf{\cellcolor{gray!6}{Missingness}} & \cellcolor{gray!6}{The fraction of animal locations that should have been collected but, for some reason, were not. For example, if only eight out of ten expected animal locations were successfully collected, the missingness would be 0.2 (i.e., 20\%).}\\
\addlinespace
\textbf{Forgiveness} & The maximum step duration, measured in multiples of the regular step duration, a modeler is willing to include in the step-selection analysis. A modeler with a forgiveness of one, for instance, only considers regular steps, while a modeler with a forgiveness of two would consider irregular steps up to twice the regular step duration.\\
\textbf{\cellcolor{gray!6}{Burst}} & \cellcolor{gray!6}{A sequence of consecutive animal locations equally spaced in time and with steps were the step duration does not exceed the forgiveness.}\\
\textbf{Valid step} & A step for which a step length, turning angle, and step duration can be computed, and for which the step duration does not exceed the forgiveness of the modeler. These steps can be used for step-selection analysis.\\
\bottomrule
\end{tabular}}
\end{table}

### 2024-07-01 Chapter_4/04_Manuscript/Manuscript.tex
Updated manuscript
Minor updates
Publish PHD

--- 2022-07-13 Chapter_4/04_Manuscript/Manuscript.tex
@@
 \usepackage[nameinlink]{cleveref}
-% \usepackage{lineno}
 \usepackage{multirow}
@@
 \usepackage[nomarkers, nolists]{endfloat}
-\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{scalerel}
+\usepackage{makecell}
+\usepackage{ifthen}
+\usepackage{tikz}
+\usepackage{lineno}
+\usepackage{adjustbox}
+\usetikzlibrary{svg.path}
+
+%------------------------------------------------------------------------------
+%  Bibtex vs Biblatex
+%------------------------------------------------------------------------------
+% While bibtex is very nice to use, it's insanely slow. I therefore work with
+% bibtex until the final stage. Set the value below to "true" or "false" to pick
+% which one should be used.
+\newboolean{usebiblatex}
+\setboolean{usebiblatex}{true}
+
+% Depending on the question, the commands below will prepare the document
+% accordingly
+\ifthenelse{\boolean{usebiblatex}}{
+  %-----------------------------------------------------------------------------
+  %  If using biblatex
+  %-----------------------------------------------------------------------------
+  % Load required packages and refer to correct bibliography
+  \usepackage{csquotes}
+  \usepackage[
+      backend=biber
+    , style=apa          % To display Author-year in the text
+    , natbib=true        % Necessary to use citep, etc.
+    , hyperref=true      % To have clickable links in the references
+    , useeditor=false    % Don't add editor to publications
+    , sorting=ynt        % Sort by year, name, and title
+    , uniquename=false   % Avoid that initials are put
+  ]{biblatex}
+  \addbibresource{Literature.bib}
+
+  % To include author in link
+  \DeclareFieldFormat{citehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{#1}}
+  \DeclareFieldFormat{textcitehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{%
+    #1%
+    \ifbool{cbx:parens}
+      {\bibcloseparen\global\boolfalse{cbx:parens}}
+      {}}}
+  \savebibmacro{cite}
+  \savebibmacro{textcite}
+  \renewbibmacro*{cite}{%
+    \printtext[citehyperref]{%
+      \restorebibmacro{cite}%
+      \usebibmacro{cite}}}
+  \renewbibmacro*{textcite}{%
+    \ifboolexpr{
+      ( not test {\iffieldundef{prenote}} and
+        test {\ifnumequal{\value{citecount}}{1}} )
+      or
+      ( not test {\iffieldundef{postnote}} and
+        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
+    }
+      {\DeclareFieldAlias{textcitehyperref}{noformat}}
+      {}%
+    \printtext[textcitehyperref]{%
+      \restorebibmacro{textcite}%
+      \usebibmacro{textcite}}}
+
+  % Remove editor form everything but books
+  \AtEveryBibitem{%
+   \ifentrytype{book}{}{%
+    \clearname{editor}
+   }
+  }
+}{
+  %-----------------------------------------------------------------------------
+  %  If using bibtex
+  %-----------------------------------------------------------------------------
+  % Load required packages and select citation style
+  \usepackage[round]{natbib}
+  \bibliographystyle{apalike}
+}
 
@@
 %	Some Styling
+%------------------------------------------------------------------------------
+% Link colors
+\definecolor{linkcolor}{HTML}{000000}
+\hypersetup{
+  colorlinks = true,
+  linkcolor  = linkcolor,
+  urlcolor   = linkcolor,
+  citecolor  = linkcolor,
+}
+
+% Changing the style of captions in figures etc.
+\captionsetup{labelfont=bf, format=plain, font=small}
+
+% Change how equations are referenced
+\renewcommand{\theequation}{Equation \arabic{equation}}%
+
+% Avoid skip when including text from external .tex file
+\newcommand{\inputy}[1]{\input{#1}\unskip}
+
+% For supplementary material
+\newcommand{\beginappendix}{%
+  \setcounter{table}{0}
+  \renewcommand{\thetable}{S\arabic{table}}%
+  \setcounter{figure}{0}
+  \renewcommand{\thefigure}{S\arabic{figure}}%
+  \setcounter{equation}{0}
+  \renewcommand{\theequation}{Equation S\arabic{equation}}%
+  \setcounter{section}{0}
+  \renewcommand{\thesection}{A.\arabic{section}}%
+}
+
+%------------------------------------------------------------------------------
+%  ORCID
 %------------------------------------------------------------------------------
@@
 %------------------------------------------------------------------------------
-\title{Applying Step Selection Functions to Data with Data Gaps: Simulation and
-Case Study on African wild dogs}
+\title{Methods for Implementing Integrated Step-Selection Functions with
+Incomplete Data}
 
@@
 \author{
-  David D. Hofmann\textsuperscript{1,2,\S} \and
-  Gabriele Cozzi\textsuperscript{1,2} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \and
-  Dominik M. Behr\textsuperscript{1,2}
+  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
+  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
+  John Fieberg \textsuperscript{3} \orcid{0000-0002-3180-7021}
 }
@@
 
-% Required to be able to cite
-\bibliographystyle{apalike}
-
 % Create Titlepage
@@
 \textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
+Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
 Switzerland.
 
-\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
-Maun, Botswana.
+\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+Private Bag 13, Maun, Botswana.
 
-\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)
+\textsuperscript{3} Department of Fisheries, Wildlife, and Conservation Biology,
+University of Minnesota, St. Paul, MN, USA.
+
+\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
 
@@
 
-\textbf{Running Title:} None
+\textbf{Running Title:} Step-Selection Analyses with Missing Data
 
@@
 
-\textbf{Keywords:} step selection functions, gps data, animal movement, Lycaon
-pictus, irregular data
+\textbf{Keywords:} animal movement, gps data, imputation, incomplete data,
+missing fixes, step-selection analyses, step-selection functions
 
@@
 \begin{abstract}
-Animal movement data collected in the field is often noisy and incomplete, thus
-confronting movement ecologists with temporal irregularities in their data. Step
-selection functions have become a popular tool to analyse movement data and to
-infer a species movement or preferences, yet they are based on the assumption of
-equal temporal gaps between consecutive datapoints. To adhere to this
-requirement, researchers typically only consider bursts of consecutive
-relocations with equal step-durations, thus discarding large portions of their
-data. Here, we use simulated data to question this practice and we examine
-whether including temporally irregular can improve estimates from step-selection
-models. In addition, we explore the usefulness of multiple methods to account
-for temporal irregularity in movement data. Our results show that including
-irregular data improves model estimates substantially. By applying methods that
-account for the irregularity in the data, the precision of model estimates can
-further be improved, albeit only marginally.
+
+Integrated step-selection analyses (iSSAs) are versatile and powerful frameworks
+for studying habitat and movement preferences of tracked animals. iSSAs utilize
+integrated step-selection functions (iSSFs) to model movements in discrete time,
+and thus, require animal location data that are regularly spaced in time.
+However, many real-world datasets are incomplete due to tracking devices failing
+to locate an individual at one or more scheduled times, leading to slight
+irregularities in the duration between consecutive animal locations. To address
+this issue, researchers typically only consider bursts of regular data (i.e.,
+sequences of locations that are equally spaced in time), thereby reducing the
+number of observations used to model movement and habitat selection. We reassess
+this practice and explore four alternative approaches that account for temporal
+irregularity resulting from missing data. Using a simulation study, we compare
+these alternatives to a baseline approach where temporal irregularity is ignored
+and demonstrate the potential improvements in model performance that can be
+gained by leveraging these additional data. We also showcase these benefits
+using a case study on a spotted hyena (\textit{Crocuta crocuta}).
+
 \end{abstract}
@@
 %------------------------------------------------------------------------------
-\newpage
-
-\onehalfspacing
-\tableofcontents
-\doublespacing
+% \newpage
+%
+% \onehalfspacing
+% \tableofcontents
+% \doublespacing
 
@@
 % Create linenumbers
-% \linenumbers
+\linenumbers
 
 \section{Introduction}
-GPS data is usually collected on fixed temporal intervals (could do a quick and
-dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
-even days between subsequent reolocations. Nevertheless, GPS devices regularly
-fail to adhere to the anticipated fixrate schedule due to cloud cover or other
-unfavorable conditions, thus resulting in irregular data. In other cases, the
-GPS schedule is not strictly fixed but may be subject to change. For example,
-\cite{Cozzi.2020} increased the fixrate of their GPS collars once per week from
-4 to 8 fixes per day, aiming to collect more fine scaled data for a limited
-time. We chose to coin this type of data \textit{irregular}, because it is not
-collected on the anticipated fixrate schedule and therefore not directly
-comparable to the vast majority of the collected data.
 
-Step selection analysis is a powerful framework that enables researchers to
-study habitat and movement preferences of their focal species. The method works
-by comparing spatial covariates at locations where the studied animal has been
-observed, to a set of locations where the animal was presumably absent. For
-this, the collected GPS data is converted into steps, where a step represents
-the straight line movement between two consecutive GPS fixes. One can then
-compare covariates experienced along observed steps to covariates along
-potential alternative steps and thus infer preferred or avoided features. By
-restricting the availability domain to a set of steps, SSA is specifically
-intended to account for spatio-temporal autocorrelation inherent to data
-collected using GPS collars and is therefore one of the preferred methods for
-analysing such data. However, it is considered good practice to only consider
-steps with similar step durations in step-selection models. Depending on the
-amount of available data and depending on the number of missing GPS fixes, this
-practice might entail little sacrifice. In some situations, however, removing
-irregular data implies a substantial loss of information. Additionally, recent
-developments in the SSA realm have brought forward several improvements to SSA
-that might allow to relax the assumption of regular step-durations. This
-includes the \textit{integrated} SSA approach presented by \citep{Avgar.2016},
-in which inference on habitat \textit{and} movement preferences are possible
-thanks to the inclusion of movement descriptors in the respective model. More
-recently, the method has been further refined by \citep{Munden.2021}, who coined
-the term \textit{time-varying integrated} SSA. This method was developed for
-high frequency data which has been rarified using a change-point detection
-algorithm, thus also resulting in irregular step-durations. In case of missing
-steps, in contrast, steps durations are not entirely random but usually
-clustered.
+Understanding how animals move across the landscape, what habitats they prefer,
+and what resources they select are fundamental questions in movement ecology
+\citep{Nathan.2008}. Thanks to recent advances in animal tracking
+\citep{Cagnacci.2010, Williams.2019, Beardsworth.2022} and remote sensing
+technologies \citep{Toth.2016, Rumiano.2020}, new opportunities and analytical
+tools have emerged for studying how animals move and interact with their
+environment \citep{Tomkiewicz.2010, Kays.2015, Nathan.2022}. Methods commonly
+used to analyze animal movement data, including step-selection analyses
+\citep{Fortin.2005, Thurfjell.2014, Fieberg.2021} and hidden Markov models
+\citep{Michelot.2016}, require \textbf{animal locations} (terms in bold at first
+occurrence are defined in \Cref{tab:Glossary}) that are collected at a constant
+sampling frequency, leading to data that are equally spaced in time. Yet, it is
+common to encounter missing locations in most telemetry data sets
+\citep{Frair.2010, Hofman.2019, Vales.2022}, which introduces unwanted
+irregularities in the duration between successive locations. Thus, there is a
+need for analytical tools that enable the analysis of such data, while
+mitigating potential biases arising from temporal irregularity introduced
+through missing animal locations.
 
-Here, we questioned the practice of removing any irregular data in SSA and
-investigated whether such data could be used to inform step selection models.
-For this, we conducted a simulation study where we simulated movement
-trajectories with known preferences across a virtual landscape. We then
-artifically rarified the ``observed'' data by removing GPS fixes and we analysed
-this data using SSA. This allowed us to investigate if and how the inclusion of
-irregular data affected model estimates. We also varied the degree of
-irregularity in the data and we tested for the effects of adjusting the
-parametric step-length and turning distributions to different step-durations.
-Besides the simulation study, we also analysed a dataset collected on dispersing
-African wild dogs and examined the implications considering or discarding
-irregular data.
+Step-selection analyses (SSAs) are widely used to study animals' movement
+capacities and habitat-selection patterns \citep{Fortin.2005, Thurfjell.2014}.
+Straight-line segments connecting consecutive animal locations, referred to as
+\textbf{steps}, form the basic building blocks of the statistical likelihood in
+SSAs. Specifically, SSAs model the probability $u$ of finding an individual at
+location $s$ at time $t+1$, given the animal's past positions at time $t$ and
+$t-1$, $s_t$ and $s_{t-1}$, respectively:
+
+\begin{equation}
+\label{EQ1}
+u(s_{t+1}) = \frac{\phi(s_{t+1}, s_t, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)}{\int_{s \in G}\phi(s_{t+1}, s_{t}, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)ds}
+\end{equation}
+
+\noindent Here, the function $\phi$ represents an animal's \textbf{movement
+kernel} which is usually expressed in terms of \textbf{step-length} and
+\textbf{turning-angle} distributions, with $\gamma$ representing parameters in
+these distributions. The function $w$ is the \textbf{habitat-selection function}
+and reflects an animal's preferences $\beta$ for environmental characteristics
+$x$ at location $s_{t+1}$. In most applications, $w$ is modeled as a log-linear
+function of $x$, taking the form $w = exp(x^\top\beta)$. The integral in the
+denominator of \ref{EQ1} ensures that $u$ is a proper probability distribution
+(i.e., that it integrates to 1). Following \citet{Michelot.2024}, we call the
+product $\phi \times w$ the step-selection function (SSF), as it highlights that
+the probability of finding an animal at a certain location depends on both the
+animal's movement kernel and its habitat-selection function.
+
+Given a series of \textbf{observed steps}, finding the movement and
+habitat-selection parameters that maximize the likelihood in \ref{EQ1} requires
+approximating the integral in the denominator for each observed step. A variety
+of numerical integration techniques can be used for this purpose
+\citep{Michelot.2024}, but a common approach is to combine observed steps with
+\textbf{random steps} generated by sampling step lengths and turning angles from
+parametric distributions informed by the data \citep{Fortin.2005,
+Thurfjell.2014}. Environmental conditions at observed steps are then contrasted
+with environmental conditions at random steps in a (mixed effects) conditional
+logistic regression framework \citep{Fortin.2005, Muff.2020}. To jointly
+estimate parameters in $\phi$ and $w$, movement descriptors (e.g., step length
+(sl), its natural logarithm (log(sl)), and the cosine of the turning angle
+(cos(ta))) can be included in the conditional logistic regression model, and
+their estimated coefficients can be used to update the initial (tentative)
+step-length and turning-angle distributions \citep{Duchesne.2015, Avgar.2016,
+Fieberg.2021}. The specific descriptors that need to be included depend on the
+assumed step-length and turning-angle distributions \citep[for more details, see
+Appendix C of][]{Fieberg.2021}. This approach to estimating parameters of the
+SSF, termed \textit{integrated} SSA (or iSSA) by \citet{Avgar.2016}, is similar
+to using importance sampling to approximate the integral in \ref{EQ1}
+\citep{Michelot.2024} and is readily accessible through the \texttt{R}-package
+\texttt{amt} \citep{Signer.2019}.
+
+%In the SSF framework introduced by \citet{Fortin.2005}, the movement kernel
+%$\phi$ was assumed to be known, which precluded joint estimation of $\phi$ and
+%$w$ and has been shown to result in biased habitat-selection estimates
+%\citep{Forester.2009}. A more robust alternative is to jointly estimate $\phi$
+%and $w$ by fitting tentative distributions for step-lengths and turning-angles,
+%and including descriptors of step-characteristics (i.e., step-lengths and
+%turning-angles) as covariates in the statistical model \citep{Forester.2009,
+%Avgar.2016}. In SSFs, the use of random steps serves to generate integration
+%points that are needed to approximate the integral in \ref{EQ1} and, even
+%though the true movement kernel $\phi$ is often not known a priori, it can be
+%approximated using \textit{tentative} distributions \citep{Michelot.2024}. The
+%use of parametric distributions serves as a means of conducting importance
+%sampling and allows generating random steps that contribute better towards
+%improving the approximation \citep{Michelot.2024}.
+
+SSAs have proven extremely effective in numerous ecological studies
+\citep{Thurfjell.2014}, providing insights into seasonal space use
+\citep{Vales.2022, Enns.2023}, resource selection during distinct behavioral
+phases \citep{Elliot.2014, Abrahms.2017, Cozzi.2018, Broekhuis.2019}, and the
+effects of landscape familiarity or memory on animal movements \citep{Kim.2023}.
+A model parametrized using iSSA resembles a fully mechanistic movement model
+that can be used to simulate space use under novel conditions \citep{Avgar.2016,
+Signer.2017, Hofmann.2023, Signer.2023}. This characteristic has made iSSAs a
+useful tool for quantifying landscape resistance and identifying movement
+corridors \citep{Zeller.2020, Buchholtz.2020, Hofmann.2021, Hofmann.2023}.
+
+% Glossary of terms
+\input{99_Glossary.tex}
+
+A key assumption when conducting an iSSA is that the data have been collected at
+a constant sampling frequency, thus producing \textbf{trajectories} with
+\textbf{regular step durations} ($\Delta t$; \citealp{Fortin.2005,
+Thurfjell.2014}). Here, we refer to such data as \textbf{regular animal
+locations}, and without loss of generality, we assume the regular step duration
+to be one (i.e., $\Delta t = 1$). Regular step durations ensure that
+step lengths and turning angles are independent of the step duration, and
+therefore, steps can be pooled when estimating movement parameters.
+Since animal locations are usually obtained using automated tracking devices,
+such as GPS collars programmed to record data at regular intervals, satisfying
+this assumption seems straightforward. In reality, however, device limitations
+often imply that some of the aspired datapoints fail to be collected, thus
+introducing \textbf{missingness} and confronting researchers with
+\textbf{irregular animal locations} and \textbf{irregular step durations}
+\citep{Frair.2010}. In a comprehensive study, \citet{Hofman.2019} showed that
+across 3,000 GPS devices and 160 species, the average success rate of obtaining
+a scheduled animal location was 78\% (implying a missingess of 22\%), thus
+highlighting that irregular animal locations are a frequent phenomenon in
+ecological studies.
+
+It is generally recommended that, in the case of such irregular data,
+researchers should only retain \textbf{bursts} of steps with regular
+step durations (possibly with some tolerance) and discard the rest
+\citep{Thurfjell.2014}. We will refer to this modeling approach as having a
+\textbf{forgiveness} level of one, indicating that only steps with
+step durations $\Delta t = 1$ are retained for further analysis. In \texttt{R},
+the \texttt{amt} package provides the function \texttt{track\_resample}
+specifically for identifying bursts of steps with regular step durations
+\citep{Signer.2019}. The main drawback of this approach is that it may result in
+a substantial amount of data being discarded (\Cref{Overview} and
+\Cref{NumberOfSteps}). For instance, consider a hypothetical trajectory in which
+location 4 is missing (\Cref{Overview}a). The absence of this location prevents
+the computation of a step between locations 3 and 4, as well as between
+locations 4 and 5. Furthermore, without these steps, it becomes impossible to
+compute a turning angle for the step between locations 5 and 6. Consequently,
+the lack of a single location reduces the effective sample size, which is the
+number of \textbf{valid steps}, by three. Assuming that animal locations are
+missing at random, a missingness level of 25\% causes the number of valid steps
+to drop by $58\%$ (\Cref{NumberOfSteps}). A modeler willing to increase their
+level of forgiveness to two (i.e., allowing for inclusion of steps with $\Delta
+t \leq 2$) would be able to increase the number of valid steps by $57\%$
+(\Cref{Overview} and \Cref{NumberOfSteps}), therefore achieving a substantial
+gain in effective sample size. The ability to capitalize on irregular data is
+likely to be particularly important for applications where data are already
+limited, such as, for instance, when modeling dispersing individuals
+\citep{Rudnick.2012, Fattebert.2015, Cozzi.2020}. However, increasing the
+forgiveness also implies that step durations of the retained steps become
+irregular, thus necessitating appropriate tools to account for such
+irregularity.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Overview.png}
+  \caption{(\textbf{a}) Demonstration of how missingness affects the number of
+  valid steps that can be used for step-selection analyses under different
+  levels of forgiveness. The upper panel depicts a trajectory with zero missing
+  locations. That is, all aspired locations were successfully collected on a
+  regular interval (yielding a regular step duration of $\Delta t = 1$). This
+  trajectory produces four valid steps that can be included in the iSSF model
+  and one invalid step that has to be omitted because it has no turning angle
+  associated with it. In the central panel, animal location 4 was not obtained,
+  introducing a missingness of 16.7\%. If the modeler has a forgiveness of one,
+  only a single step can be included for further analysis, as all other steps
+  are invalid (either because no turning angle can be computed or because
+  step durations exceed the forgiveness). If, however, the modeler exhibited a
+  forgiveness of two, such as in the lower panel, a total of three steps could
+  be obtained for further analysis. (\textbf{b}) Conceptual illustration of how
+  increasing the forgiveness allows one to retain additional steps that can be
+  used for step-selection analysis. The sequence of dots resembles the sequence
+  of locations that were scheduled to be collected (e.g. using a GPS device),
+  with the lines representing hypothetical steps. Because not all locations were
+  successfully obtained (gray dots), there is missingness. Depending on the
+  forgiveness level, already a single missing location enforces the introduction
+  of a new burst, which leads to the loss of several steps. In addition, some of
+  the remaining steps are invalid (dotted) because they are lacking a
+  turning angle. By increasing the forgiveness, a modeler is willing to retain
+  steps that exceed the regular step duration by a certain threshold, which
+  enables them to obtain longer bursts and increase the number of steps that can
+  be used for further analysis. In the figure, forgiveness increases from left
+  to right.}
+  \label{Overview}
+  \end{center}
+\end{figure}
 
@@
   \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
-  \caption{}
+  \caption{Illustration of how missingness in animal location data reduces the
+  number of valid steps that can be used in step-selection analyses (left panel)
+  and how increasing forgiveness helps to retain additional steps that are
+  otherwise omitted (right panel). At a missingness of 0, 998 valid steps can be
+  obtained from the total of 1,000 animal locations. At higher missingness,
+  step durations become irregular, which means that the number of valid steps
+  decreases substantially. However, if the modeler is willing to increase their
+  forgiveness, additional steps can be gained. The right panel shows the number
+  of valid steps that is gained when increasing the forgiveness from 1 to
+  2, 3, 4, and 5, respectively. Ribbons indicate the 95\%-percentile intervals
+  derived from 1,000 replicates.}
+  \label{NumberOfSteps}
+  \end{center}
+\end{figure}
+
+Various methods have been employed in the past to address temporal
+irregularities in animal location data. These may serve as valuable starting
+points for developing approaches that enable the integration of irregular data
+with iSSFs.
+
+\begin{enumerate}
+
+  \item \textit{Imputation}: An intuitive solution is provided by
+  \citet{McClintock.2017}, who suggests fitting a continuous-time correlated
+  random walk movement model \citep{Johnson.2008} to the collected data and to
+  use the fitted model to impute missing fixes. By imputing missing locations,
+  the analysed trajectories become entirely regular again and can be analysed
+  using traditional techniques. This approach, which we coin the
+  \textit{imputation} approach, is readily available through the
+  \texttt{R}-package \texttt{crawl} \citep{Johnson.2022}, yet has only been
+  tested for use with hidden Markov movement models and not with iSSFs
+  \citep{McClintock.2017}.
+
+  \item \textit{Na\"ive}: Another approach is outlined by \citet{Munden.2021},
+  who introduced time-varying iSSFs. In this framework, a change-point detection
+  algorithm is applied to the series of observed animal locations to identify
+  distinct decision points where the animal turns \citep{Potts.2018,
+  Munden.2021}. Steps are then created to represent straight-line movements
+  in-between these decision points, but because decision points are not
+  regularly spaced in time, the resulting step durations are irregular. Thus,
+  step durations are treated as random variables, and, instead of generating
+  random steps by sampling step lengths and turning angles, the authors generate
+  random steps by sampling step durations, step speeds, and turning angles. The
+  underlying assumption is that step lengths scale linearly with step duration,
+  and can therefore be meaningfully represented by combinations of step speeds
+  and step durations. Although this approach was developed with
+  ultra-high-frequency data in mind, we might \textit{na\"ively} apply it more
+  broadly to the case of missing data if we believe the presumed
+  linear relationship to hold true (this assumption may be reasonable when
+  step durations are short, but it is unlikely to hold for longer step durations
+  since an animal's path will deviate from a straight line between successive
+  locations). Hence, we propose, with our \textit{na\"ive} approach, to scale the
+  generated random steps by the observed step duration.
+
+  \item \textit{Dynamic+Model}: Instead of generating random steps by sampling
+  step lengths and turning angles from distributions fitted to a single
+  step duration, one may choose to fit separate distributions to steps of
+  different durations, thus acknowledging potentially non-linear relationships
+  between step duration and step lengths or turning angles. Because random steps
+  in this approach are sampled using different \textit{tentative} distributions,
+  it is necessary to include interactions between step durations and other
+  step descriptors (e.g., sl, log(sl), and cos(ta)) in the conditional logistic
+  regression model to allow updating the distributions to the different
+  step durations. We therefore refer to this approach as the
+  \textit{dynamic+model} approach, highlighting that step-length and
+  turning-angle distributions are dynamically adjusted to observed
+  step durations and that the step duration is included as a modifier of the
+  coefficients of step descriptors in the regression model.
+
+  \item \textit{Multistep}: Finally, we propose a \textit{multistep} approach,
+  where random steps of varying step durations are generated by stitching
+  together sequences of random steps from the regular step duration. For
+  example, one can generate a random step of duration $\Delta t = 2$ by
+  combining two random steps of step duration $\Delta t = 1$.
+
+\end{enumerate}
+
+\noindent Our goal with this article is to reassess the practice of discarding
+irregular animal locations in iSSFs and to investigate whether retaining
+irregular data could, in fact, serve to improve model performance. Our
+hypothesis is that even irregular data contains valuable information on habitat
+and movement preferences that could be leveraged if appropriate methods are
+applied. To test this notion, we conducted a comprehensive simulation study
+where we simulated regular animal location data with known movement and habitat
+parameters. We then introduced varying levels of missingness and applied iSSFs
+to estimate simulation parameters. Specifically, we employed the four
+alternative iSSF approaches outlined above and compared them to the traditional
+approach of including only bursts of regular data and to an uncorrected approach
+that simply ignored irregular step durations when using a forgiveness level $>$
+1. To examine the impact of different landscape configurations on derived
+estimates, we ran our simulations for different levels of spatial
+autocorrelation. The use of simulations instead of real data had the benefit
+that underlying parameters of the movement kernel and habitat-selection function
+were known, which allowed us to assess the reliability of different methods in
+retrieving true simulation parameters under different conditions (sensu
+\citealp{Kery.2016}). We then compare the traditional approach (using only
+bursts of steps with regular step durations) to the best-performing approach
+(with irregular data) using GPS locations collected on a spotted hyena
+(\textit{Crocuta crocuta}).
+
+We anticipated that increasing forgiveness without adjusting for the introduced
+irregularity would entail a bias-variance trade-off. Specifically, we
+anticipated that increasing forgiveness would allow improving estimator
+precision, but at the cost of introducing bias due to failing to account for
+irregular sampling intervals. We expected this bias to be particularly
+pronounced at high levels of missingness. Furthermore, we hypothesized that
+accounting for irregularity in the \textit{na\"ive}, \textit{dynamic+model}, and
+\textit{multistep} approaches would improve model accuracy, while alleviating
+potential bias, thus providing an effective means of incorporating additional
+data. Because the \textit{imputation} approach relied on an intermediate
+movement model to predict missing animal locations, we had no prior expectations
+for how well it would perform.
+
+\section{Methods}
+
+We implemented the simulation study in the programming language \texttt{R}
+version 4.3.2 \citep{RCoreTeam.2023} and achieved parallelization of
+simulation-runs using the \texttt{R}-package \texttt{pbmcapply}
+\citep{Kuang.2022}. We generated figures using the \texttt{ggplot2}
+\citep{Wickham.2023}, \texttt{ggpubr} \citep{Kassambara.2023}, and
+\texttt{ggh4x} \citep{Brand.2023} \texttt{R}-packages. We manipulated raster
+data and computed spatial distances using the \texttt{R}-package \texttt{raster}
+\citep{Hijmans.2023}. An overview of the simulation design and the different
+iSSF approaches is presented in \Cref{Design} and all codes to reproduce this
+study are available through an online repository \citep{Hofmann.2023a}.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Design.png}
+  \caption{Illustration of the study design. We varied the autocorrelation range
+  when simulating spatial covariates from 5 to 20 and tested for different
+  missingness scenarios (ranging from 0\% to 50\% missing locations). To
+  investigate how increasing forgiveness (i.e., the willingness to include steps
+  with duration above the regular step duration) influenced model results, we
+  varied its value from 1 (regular step selection) to 5 (considering steps that
+  are five times the regular step duration). Finally, we tested five different
+  methods to account for potential biases introduced by including irregular
+  steps. This gave 3 x 6 x 5 x 5 = 450 combinations, each of which we replicated
+  100 times. We assumed step lengths (sl) to follow a gamma distribution,
+  whereas turning angles (ta) followed a von Mises distribution.}
+  \label{Design}
+  \end{center}
+\end{figure}
+
+\subsection{Landscape Simulation}
+
+We simulated a virtual landscape comprising two continuous and one categorical
+(binary) spatial layers, each with a resolution of 300 x 300 pixels
+(\Cref{Covariates}) and spanning across x- and y-coordinates from 0 to 300. The
+first layer, \textsf{Dist} (continuous), quantified the distance to the center
+of the virtual landscape ($x = 150$, $y = 150$), and can be understood as a
+point of attraction, such as, for instance, the center of an animal's
+home-range. The second layer, \textsf{Elev} (continuous), resembled an elevation
+layer and was simulated by sampling random pixel-values from a normal
+distribution. To achieve spatial autocorrelation, we applied a circular moving
+window with radius $r$ within which we tallied pixel-values. We varied $r$ from
+5, to 10, to 20, depending on the simulated level of autocorrelation (Figure
+S1). The third layer, \textsf{Forest} (categorical), represented areas covered
+by woodland and was simulated similarly to the \textsf{Elev} layer, but we
+binarized the layer by setting all simulated values above the 50\% quantile to
+forest and all other values to non-forest (our reference class). We normalized
+values of all simulated layers to a range between zero and one and replicated
+the simulation of each layer 100 times per autocorrelation scenario, thus
+producing 300 unique landscape configurations.
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_Covariates.png}
+  \caption{Example of a landscape configuration across which we simulated
+  movement trajectories. All simulated layers had a resolution of 300 x 300
+  pixels. The distance layer indicated the distance to the center of the
+  landscape and served to simulate attraction. The elevation and forest layers
+  were simulated by sampling pixel-values from a normal distribution and
+  applying a moving window to achieve spatial autocorrelation. Simulated
+  individuals were initiated within the white dashed rectangle, which ensured
+  that they would not be released directly at a map border. Simulated
+  individuals were attracted to the landscape's center, preferred elevated
+  areas, and avoided areas covered by forest. The black line shows the simulated
+  trajectory associated with the visualized landscape configuration (cfr.
+  \Cref{MovementSimulation}).}
   \label{Covariates}
@@
 
-We hypothesized that the precision of model estimates would decrease as we
-increased the missigngess in the data. We also expected that the inclusion of
-irregular data would not improve the precision of estimates and rather lead to
-biased values. However, we anticipated that such biases, at least in estimated
-movement kernel parameters, could be reduced by appropriately adjusting the
-availability domain.
+\subsection{Movement Simulation}
+\label{MovementSimulation}
+To simulate movement across the virtual landscape, we employed the iSSF
+simulation algorithm developed by \citet{Signer.2017} and applied in
+\citet{Hofmann.2023}. This procedure consists of a sequence of five steps that
+are repeated $n$ times to generate a movement trajectory. In step one, we
+generated a random starting location by sampling random x- and y-coordinates on
+the simulated landscape. To prevent starting points near map borders, we
+restricted sampled locations to x- and y-coordinates between 50 and 250 (white
+dotted rectangle in \Cref{Covariates}). In step two, we generated a set of 1,000
+random steps originating from the current location, by sampling turning angles
+from a von Mises distribution with concentration parameter $\kappa = 0.5$ and
+step lengths from a gamma distribution with shape parameter $k = 3$ and scale
+parameter $\theta = 1$. In step three, we extracted covariate values at the end
+of each random step from the underlying covariate layers. In step four, we
+assigned to each step $j$ a probability $\pi_j$ of being selected using the
+equation below:
 
-\section{Methods}
-\subsection{Simulation Study}
-\subsubsection{Spatial Covariates}
-We simulated a virtual landscape comprising three spatial covariate layers, each
-with a resolution of 300 x 300 pixels (\Cref{Covariates}) spanning across x- and
-y-coordinates from 0 to 300. The first layer (\textsf{water}) represented
-water-bodies and was simulated using a random cluster nearest‐neighbour neutral
-landscape model \citep{Saura.2000}, with the cluster-proportion set to 0.5 and
-the patch occupancy of water fixed to 20\%. The second layer (\textsf{elev})
-resembled an elevation layer and was simulated using a Gaussian random fields
-neutral landscape model \citep{Schlather.2015}, with an autocorrelation range of
-10, magnitude of variation of the landscape of 1, and a magnitude of variation
-in scale of 0. We simulated both of these layers using the r-package {\tt NLMR}
-\citep{Sciaini.2018}. The third layer (\textsf{dist}) simply indicated the
-distance (in pixels) to the center of the virtual landscape \((x = 150, y =
-150)\), and can be understood as the distance to the center of an animal's
-home-range. We computed spatial distances using the r-package {\tt raster}
-\citep{Hijmans.2022}. We normalized the values from all layers to a range
-between zero and one.
+\begin{equation}
+\label{EQ2}
+  \pi_j =
+  \frac{exp(\beta^\top x_j)}{\sum_{i = 1}^J exp(\beta^\top x_i)} %=
+  % \frac{
+  %   exp(\beta_{dist} dist_j + \beta_{elev} elev_j + \beta_{forest} forest_j)
+  % }{
+  %   \sum_{i = 1}^J exp(\beta_{dist} dist_i + \beta_{elev} elev_i + \beta_{forest} forest_i)
+  % }
+\end{equation}
+
+\noindent Here, $\beta$ represents the vector of habitat-selection parameters
+and $x_j$ the covariate value at the end of the $j$\textsuperscript{th} step.
+The probability of a step being selected thus depended on its associated
+covariate values, the covariate values of all other random steps, and the
+simulated preferences $\beta$. We defined the habitat-selection parameters as
+$\beta_{dist} = -20$, $\beta_{elev} = 0.5$, and $\beta_{forest} = -1$. That is,
+simulated individuals were attracted to the landscape's center, preferred
+elevated areas, and avoided areas covered by forest. In step five, we sampled
+one of the random steps based on predicted probabilities and computed the
+simulated individual's new position. We then repeated steps two through five
+until the trajectory comprised a total of 1,000 steps. Each simulated step was
+assumed to have a step duration of exactly $\Delta t = 1$. We repeated the
+simulation for each of the 300 simulated landscapes, producing 300 unique
+trajectories (example trajectory presented in \Cref{Covariates}).
+
+\subsection{Data Rarefication}
+
+To simulate missingness, we rarefied the trajectories by randomly removing a
+fixed fraction of animal locations. To assess the impact of different degrees of
+missingness, we varied the fraction of removed data from 0\% (complete dataset)
+to 50\% in increments of 10\%. The random removal of animal locations introduced
+temporal irregularity, such that the resulting step durations differed depending
+on the time elapsed between remaining fixes. We replicated the rarefication of
+each trajectory 100 times.
+
+\subsection{Computing Bursts}
+
+We used the rarefied data to compute bursts consisting of a sequence of
+animal locations with step durations that did not exceed the forgiveness value.
+To test how different levels of forgiveness impacted our results, we varied
+forgiveness from 1 (maximum allowed step duration was $\Delta t = 1$) to 5
+(maximum allowed step duration was $\Delta t = 5$). As an example, if the
+forgiveness was 1, any step with step duration $\Delta t > 1$ resulted in a new
+burst. If the forgiveness was 2, in contrast, step durations of up to $\Delta t
+= 2$ were allowed before a new burst was introduced (\Cref{Overview}b). Within
+each burst, we calculated step lengths and turning angles. However, due to the
+grouping of steps into bursts, the orientation of the first step within each
+burst relative to the previous step could not be determined. As a result, this
+step always lacked a turning angle and was considered invalid
+(\Cref{Overview}b).
+
+\subsection{Fitting Distributions}
+
+Based on the steps retained within bursts, we parametrized tentative step-length
+and turning-angle distributions. Specifically, we used the \texttt{fit\_distr}
+function from the \texttt{amt} package \citep{Signer.2019}, which is a wrapper
+function for the \texttt{fitdist} function from the \texttt{fitdistrplus}
+package \citep{Delignette-Muller.2015}, and fitted a gamma distribution to
+step lengths and a von Mises distribution to turning angles. Notably, we
+employed two different fitting procedures:
+
+\begin{enumerate}
+
+  \item \textit{Regular Distributions:} In this procedure, we fitted parametric
+  distributions considering only step lengths and turning angles from steps that
+  exhibited a step duration of $\Delta t = 1$ (i.e., the regular step duration).
+  Any steps with irregular step durations ($\Delta t > 1$) were discarded and
+  did not affect distributional parameter estimates. This represents the
+  traditional procedure in iSSFs where only regular bursts of animal locations
+  are considered when estimating tentative movement parameters.
+
+  \item \textit{Dynamic Distributions:} In this procedure, we fitted separate
+  parametric distributions to step lengths and turning angles from steps of
+  different step durations. That is, we parametrized separate turning-angle and
+  step-length distributions representative of steps with durations of $\Delta t
+  = 1, 2, 3, 4$ and $5$ (which corresponds to the maximum forgiveness level we
+  tested for). Some step durations only rarely occurred at low levels of
+  missingness, thus complicating parametrization of the associated
+  distributions. To facilitate estimation of dynamic distribution parameters
+  across all $\Delta t$ (Figure S2), we resampled data to different step
+  durations using the \texttt{track\_resample} function from the \texttt{amt}
+  package \citep{Signer.2019} before fitting tentative parameters. This ensured
+  a sufficient number of steps for each step duration to estimate associated
+  parameters. An alternative approach would be to increase missingness in the
+  data even further, thus introducing a larger number of longer step durations.
+
+\end{enumerate}
+
+\subsection{Step-Selection Functions}
+
+We implemented a baseline \textit{uncorrected} approach and four alternative
+iSSF approaches that mainly differed in the way in which random steps were
+generated, but sometimes also in the model call that was used to estimate
+parameters (\Cref{Design}). In the \textit{uncorrected} approach, we treated
+data as if they were regular, ignoring potential issues arising from having
+variable step durations. When forgiveness was one, this approach corresponded
+to the traditional iSSF approach. All other approaches were targeted towards
+reducing potential biases arising from the inclusion of steps with irregular
+step durations. Irrespective of the approach employed, we paired each observed
+step with a total of 200 random steps:
+
+\begin{itemize}
+
+  \item \textit{Uncorrected}: In the uncorrected approach, we generated random
+  steps by sampling step lengths and turning angles from statistical
+  distributions fitted to steps with step durations of $\Delta t = 1$,
+  regardless of the forgiveness value or observed step durations. Thus, this
+  approach ignored any potential effect of step duration when generating random
+  step lengths and turning angles.
+
+  \item \textit{Imputed}: In this approach, we sampled step lengths and turning
+  angles from statistical distributions fitted to observed steps with a step
+  duration of $\Delta t = 1$. However, prior to generating random steps, we
+  imputed missing fixes using predictions from a simple movement model.
+  Specifically, we fitted a single-state movement model \citep{Johnson.2008} to
+  the simulated trajectories and used the parametrized model to predict
+  coordinates for all missing animal locations. For this, we used the functions
+  \texttt{crwFit} and \texttt{crwPredict} from the \texttt{crawl}
+  \texttt{R}-package \citep{Johnson.2022}. Although the \texttt{crwFit} function
+  provides capabilities to incorporate location measurement error, we assumed
+  animal locations were measured without error. The imputation resulted in a
+  complete dataset without any missing animal locations, such that each
+  trajectory consisted of a single continuous burst of locations equally spaced
+  in time. As such, the imputation approach is not affected by the forgiveness
+  level.
+
+  \item \textit{Na\"ive}: In the \textit{na\"ive} approach, we again sampled
+  step lengths and turning angles from regular distributions fitted to steps
+  with step durations of $\Delta t = 1$. However, we linearly scaled the sampled
+  step lengths depending on the step durations of the observed steps. For
+  instance, we doubled the sampled step lengths for any observed step with a
+  step duration of $\Delta t = 2$. This approach na\"ively assumed that
+  step lengths scale linearly with step durations, which is unlikely to be true,
+  as most animals don't move in straight lines between successive observations.
+  Furthermore, the linear approximation is likely to get worse as step duration
+  increases (i.e., as the forgiveness value increases). Since it is not clear
+  how turning angles should scale with step duration, we did not adjust the
+  sampled turning angles.
+
+  \item \textit{Dynamic+Model}: In the \textit{dynamic+model} approach, we
+  sampled step lengths and turning angles from dynamic distributions that were
+  fit to different step durations. That is, for observed steps with
+  step duration of $\Delta t = 2$, we sampled step lengths and turning angles
+  from distributions fit to observed steps with $\Delta t = 2$. We then included
+  interactions between the step duration and other step descriptors (e.g., sl,
+  log(sl), cos(ta)), allowing us to update movement parameters for each
+  step duration separately. To avoid numerical instabilities with the
+  conditional logistic regression model, we only included steps with durations
+  $\Delta t > 1$ if the respective duration was represented at least 5 times in
+  the rarefied dataset.
+
+  \item \textit{Multistep}: In the \textit{multistep} approach, we sampled
+  step lengths and turning angles from statistical distributions fitted to
+  observed steps with step durations of $\Delta t = 1$. We then generated a
+  sequence of random steps such that their combined step duration equaled the
+  step duration of each observed step.  For instance, for an observed step with
+  step duration of $\Delta t = 2$, we generated sets of two random steps, which
+  we then concatenated into a ``random path''. The paths were then simplified to
+  straight lines connecting the first and last coordinate of each path, which
+  represented the final random step.
+
+\end{itemize}
+
+\noindent Together, an observed step and its 200 associated random steps formed
+a \textit{stratum} that received a unique ID. At the end of each step, we
+extracted covariate values from the underlying covariate layers.
+
+\subsection{Conditional Logistic Regression Model}
+
+We estimated movement and habitat-selection parameters for the simulation
+scenarios presented in \Cref{Design} using conditional logistic regression,
+implemented using the \texttt{clogit} function in the \texttt{R}-package
+\texttt{survival} \citep{Therneau.2023}. We defined a binary response variable
+(\textsf{observed}) indicating if a step was an observed (scored 1) or a random
+step (scored 0) and used the step's ID as a stratification variable. We included
+habitat covariates (dist, elev, forest) and step descriptors (sl, log(sl),
+cos(ta)) as predictors in the regression model. For the \textit{dynamic+model}
+approach, we also included interactions between the step duration, coded as a
+factor, and step descriptors. To update tentative movement parameters (denoted
+by the subscript $_0$) and obtain the selection-free movement kernel (denoted by
+the \string^ symbol), we employed the formulas provided in \citep{Avgar.2016,
+Fieberg.2021}. Specifically, we updated the shape ($\hat{k}$) and scale
+($\hat{\theta}$) parameters of the step-length distribution (gamma) using:
+
+$$
+\hat{k} = k_0 + \beta_{log(sl)}
+$$
+
+$$
+\hat{\theta} = \frac{1}{\frac{1}{\theta_0} - \beta_{sl}}
+$$
+
+\noindent We updated the concentration parameter ($\hat{\kappa}$) of the
+turning-angle distribution (von Mises) using:
+$$
+\hat{\kappa} = \kappa_0 + \beta_{cos(ta)}
+$$
+
+\noindent We kept track of the  estimates of the updated movement  ($\hat{k}$,
+$\hat{\theta}$, and $\hat{\kappa}$) parameters and  the habitat-selection
+($\hat{\beta}_{dist}$, $\hat{\beta}_{elev}$, $\hat{\beta}_{forest}$) parameters,
+and compared them to the true simulation parameters. We also quantified model
+accuracy via the root-mean-square error (RMSE).
+
+\section{Results}
+
+Results were qualitatively similar for all three landscape autocorrelation
+scenarios and for different combinations of missingness and forgiveness (Figure
+S4). Here, we report on results for a landscape with autocorrelation of 20,
+while either holding constant missingness at a conservative 20\%
+(\Cref{ModelComparisonForgiveness}) or the forgiveness level at two
+(\Cref{ModelComparisonMissingness}) (results for all other combinations are
+summarized in Figure S4). The \textit{imputation} approach resulted in biased
+estimators of $\beta_{dist}$ and $\beta_{forest}$, whereas all other approaches
+were able to recover the parameters of the habitat-selection function with
+minimal bias (\Cref{ModelComparisonForgiveness}). Note, the \textit{imputation}
+approach always starts with a full trajectory and is therefore unaffected by the
+forgiveness level. For all other methods, increasing the forgiveness from 1 to 5
+improved the precision of the estimators of habitat-selection parameters without
+introducing noticeable bias, with the biggest gains in precision and RMSE
+occurring when moving from a forgiveness of one to a forgiveness of two
+(\Cref{ModelComparisonForgiveness}). This highlights the potential benefits of
+leveraging additional data compared to the traditional approach, which uses only
+bursts of regular data (represented by the \textit{uncorrected} approach and
+forgiveness = 1).
+
+The \textit{uncorrected}, \textit{na\"ive}, and \textit{imputation} approaches
+resulted in biased estimators of the parameters in the movement kernel,
+particularly for high values of forgiveness (\Cref{ModelComparisonForgiveness}a)
+and high levels of missingness (\Cref{ModelComparisonMissingness}a). The
+\textit{imputation} approach appeared to perform particularly poorly at
+estimating the concentration parameter of the turning-angle distribution
+(\Cref{ModelComparisonMissingness}). The \textit{multistep} and
+\textit{dynamic+model} approaches resulted in unbiased estimators of parameters
+in the step-length distribution, but estimators of the concentration parameter
+exhibited a slight bias. This bias was, however, much smaller than we observed
+with the other approaches we considered. Increasing missingness negatively
+influenced the precision and accuracy of estimates, yet its impact could be
+dampened using the \textit{dynamic+model} and \textit{multistep} approaches
+(\Cref{ModelComparisonMissingness}b).
 
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Covariates.png}
-  \caption{Virtual landscape across which we simulated movement trajectories.
-  All layers have a resolution of 300 x 300 pixels and were generated randomly.
-  Simulated individuals were initiated within the white dashed rectangle, which
-  ensured that they would not be released directly at a map border.}
-  \label{Covariates}
+  \includegraphics[width = \textwidth]{99_ModelComparisonForgiveness.png}
+  \caption{(\textbf{a}) Parameter estimates and (\textbf{b}) Root mean-square
+  error (RMSE) with regard to the movement kernel and habitat-selection function
+  as a function of forgiveness. Results are shown for the scenario with
+  landscape autocorrelation of 20 and missingness of 20\%. The movement kernel
+  comprised of a gamma distribution with shape parameter $k$ and scale parameter
+  $\theta$ governing the step-length distribution and a von Mises distribution
+  with concentration parameter $\kappa$ governing the turning-angle
+  distribution. Habitat-selection was based on three covariates, namely a
+  \textsf{Distance}, \textsf{Elevation}, and a \textsf{Forest} layer. Estimates
+  are shown for the five different approaches we tested for. The uncorrected
+  approach ignored the fact that higher forgiveness implied temporal
+  irregularity in the data, while all other approaches attempted to correct for
+  the potential biases introduced by temporal irregularity. Note, the imputation
+  approach is not affected by the forgiveness level, since it always starts with
+  a full trajectory.}
+  \label{ModelComparisonForgiveness}
   \end{center}
@@
 
-\subsubsection{Movement Simulation}
-To simulate movement across the above generated virtual lanscape, we employed an
-``inverted'' step-selection function, which proceeded as follows. First, we
-generated a random starting location. To prevent starting points in the vicinity
-of map borders, we restricted sampled locations to x- and y-coordinates between
-50 and 250 (white dotted rectangle in \Cref{Covariates}). Second, we generated a
-set of 10 random steps originating from the sampled starting point. We then
-generated random steps by sampling turning angles from a von Mises distribution
-with concentration parameter \(\kappa = 0.5\) and location parameter \( \mu = 0
-\), and step lengths from a gamma distribution with shape parameter \(k = 3 \)
-and scale parameter \(\theta = 1\). Third, we extracted covariate values along
-each random step from the underlying covariate layers and computed the average
-of each covariate along every step. Fourth, we predicted for each step the
-probability of being chosen by applying \Cref{EQ1}. The vector of relative
-habitat preferences used to make predictions was set to \(\beta_{water} = -1\),
-\(\beta_{elev} = 0.5\), and \(\beta_{dist} = -15\). Fifth, we sampled one of the
-random steps based on predicted probabilities and computed the new position of
-the simulated individual. We the repated these steps until a total of 100 steps
-were realized and we replicated the simulation 100 times, thus resulting in 100
-independent movement trajectories (\Cref{Simulations}). Note that by excluding
-interactions among habitat covariates and step metrics we implicitly assume
-independent habitat and movement kernels \citep{Avgar.2016}. Although
-measurement error would have been straight forward to implement, this was not
-the main focus of our analysis and so we assumed GPS locations to be 100 percent
-accurate.
+\begin{figure}
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_ModelComparisonMissingness.png}
+  \caption{(\textbf{a}) Parameter estimates and (\textbf{b}) Root mean-square
+  error (RMSE) with regard to the movement kernel and habitat-selection function
+  as a function of missingness. Results are shown for the scenario with
+  landscape autocorrelation of 20 and forgiveness of 2. The movement kernel
+  comprised of a gamma distribution with shape parameter $k$ and scale parameter
+  $\theta$ governing the step-length distribution and a von Mises distribution
+  with concentration parameter $\kappa$ governing the turning-angle
+  distribution. Habitat-selection was based on three covariates, namely a
+  \textsf{Distance}, \textsf{Elevation}, and a \textsf{Forest} layer. Estimates
+  are shown for the five different approaches we tested for. The uncorrected
+  approach ignored the fact that higher forgiveness implied temporal
+  irregularity in the data, while all other approaches attempted to correct for
+  the potential biases introduced by temporal irregularity. Note, the imputation
+  approach is not affected by the forgiveness level, since it always starts with
+  a full trajectory.}
+  \label{ModelComparisonMissingness}
+  \end{center}
+\end{figure}
+
+\section{Case Study}
+
+To showcase the applicability of the \textit{dynamic+model} approach, which
+appeared to perform best with simulated data, we conducted a case study with
+real GPS data obtained on ``Apollo'', a spotted hyena (\textit{Crocuta crocuta})
+inhabiting the Okavango Delta ecosystem in northern Botswana. Apollo's data were
+collected between 2007 and 2011 using GPS radio collars (\textit{GPS Plus};
+Vectronic Aerospace GmbH, Berlin, Germany) and comprised 9,316 GPS locations
+(details in \citealp{Cozzi.2013a} and \citealp{Cozzi.2015}). Because hyenas are
+nocturnal \citep{Cozzi.2012}, GPS collars were set to record data at two-hourly
+intervals between 18:00 and 06:00 o'clock, and to only record a single location
+at noon. For simplicity, we only considered nightly bursts and removed all
+locations obtained at noon. Missingness in this dataset was low ($<$ 10 \%,
+\citealp{Cozzi.2015}) and to better showcase the usefulness of the
+\textit{dynamic+model} approach, we thinned the data by randomly removing 25\%
+of the obtained locations. As spatial covariate layers, we used \textsf{Water},
+\textsf{DistanceToWater} and \textsf{Trees} (Figure S5). \textsf{Water} was a
+binary variable representing major rivers and areas inundated by floodwater,
+whereas \textsf{DistanceToWater} was a continuous variable indicating the
+distance (in meters) to the nearest pixel categorized as water. \textsf{Trees}
+was a continuous variable indicating the percent tree cover in each pixel. We
+resampled all layers to a common resolution of 250 m by 250 m and merged them
+into a single raster-stack (Figure S5). The derivation of each covariate layer
+is described in detail in \citealp{Hofmann.2021}. We dynamically fitted
+step-length and turning-angle distributions to steps with step durations of 2,
+4, and 6 hours, respectively, assuming a gamma distribution for step lengths and
+a von Mises distribution for turning angles. Instead of resampling the observed
+track to different step durations when fitting dynamic distributions (like we
+did in the simulation study), we introduced a larger amount of steps with step
+durations longer than two hours by thinning the data again (by another 10\%).
+The benefit of this approach was that steps with irregular step durations
+occurred more randomly and were not limited to the hours specified by the
+resampling algorithm. Finally, we used iSSFs with the \textit{dynamic+model}
+approach to estimate the habitat-selection function and movement kernel of
+Apollo. For this, we considered three cases:
+
+\begin{itemize}
+  \item \textit{F1}: We assumed a forgiveness of one (i.e., only steps with a
+  regular step duration of 2 hours), which is akin to conducting a traditional
+  iSSA.
+
+  \item \textit{F3-S}: We assumed a forgiveness of three (i.e., considered steps
+  with step durations of up to three times the regular step duration) and
+  included interactions between the step duration ($\Delta t$) and step
+  descriptors (sl, log(sl), and cos(ta)) in the regression model.
+
+  \item \textit{F3-SH}: We assumed a forgiveness of three (i.e., considered
+  steps with step durations of up to three times the regular step duration) and
+  included interactions between the step duration ($\Delta t$) and step
+  descriptors (sl, log(sl), and cos(ta)), as well as between the step duration
+  and habitat covariates in the regression model.
+
+\end{itemize}
+
+\noindent Notably, we included F3-SH to investigate if including interactions
+between the step duration and habitat-covariates would provide insights into
+scale-dependent habitat selection. In all cases, we generated 200 random steps
+and extracted spatial covariates at the end of observed and random steps. We
+then fit the three models using the conditional logistic regression framework as
+implemented in the \texttt{survival} \texttt{R}-package \citep{Therneau.2023}.
+Lastly, we computed updated movement parameters for a regular step duration of 2
+hours.
+
+Results from the iSSF models show that increasing the level of forgiveness led
+to improvements in estimator precision (\Cref{CaseStudy}). This was achieved by
+increasing the effective sample size from 2,179 to 4,505 valid steps (Table S1).
+The improvement in estimator precision was weaker for F3-SH than for F3-S, as
+the F3-SH model was more complex due to inclusion of additional interaction
+terms. Point estimates for the habitat-selection and movement parameters were
+similar for all 3 models, and evidence for scale dependency in habitat selection
+was fairly weak. F3-S and F3-SH had similar AIC scores ($\Delta AIC \leq 1$;
+Table S1), and the interaction terms were statistically significant only for the
+step duration of 6 hours and only for one of the habitat covariates (Table S1).
 
@@
   \begin{center}
-  \includegraphics[width = 0.5\textwidth]{99_Simulations.png}
-  \caption{100 Simulated movement trajectories, each comprising of 100 steps.
-  Simulated individuals were initiated within the blue rectangle to mitigate
-  edge effects.}
-  \label{Simulations}
+  \includegraphics[width = \textwidth]{99_CaseStudy.png}
+  \caption{Model results from the case study using GPS data collected on Apollo.
+  In F1, forgiveness was set to one (only 2-hour steps were considered), whereas
+  in F3-S and F3-SH a forgiveness of three was employed (allowing for step
+  durations of up to 6 hours). In model F3-S, the step duration was interacted
+  with step descriptors. In model F3-SH, step duration was interacted with step
+  descriptors and habitat covariates. The bars indicate the 90\%, 95\%, and 99\%
+  confidence intervals. Note that for simplicity, we omitted interactions with
+  the step duration from this figure.}
+  \label{CaseStudy}
   \end{center}
@@
 
-\subsubsection{Step Selection Analysis}
-To assess the consequences of missing GPS data in when employing step-selection
-functions, we analysed the simulated GPS data under different conditions
-\Cref{Design}. Specifically, we varied the degree of data-missingness (0\% to
-80\%), the forgiveness when analysing the data (1 to 5 fixes), and the
-parametric distributions from which random steps were generated (static vs.
-dynamic distributions). We replicated each combination of conditions 100 times
-and computed averages and standard deviations from the resulting model
-estimates.
+\section{Discussion}
+%Notably, both Munden et al.'s and McClintock's methods can
+%accommodate step durations that are entirely random, as is often the case in
+%marine applications, where locations can only be obtained when animals surface
+%\citep{Costa.2010, Albertsen.2019}. Our focus, however, is on applications
+%where data are schedule to be collected at regular intervals, but where
+%observations are occasionally missing due to failures of the telemetry device
+%to obtain a location. Nevertheless, Munden et al.'s approach motivates another
+%solution we coin the \textit{na\"ive} approach.
 
-In this context, missingness and forgiveness are complementary terms. While
-missingness describes the fraction of data that has not been collected, we use
-the term forgiveness to describe how much missingness a modeler is willing to
-accept. A modeler with forgiveness of one, for instance, only considers steps
-with a step-duration of one (i.e. fixes that are regularly spaced). In contrast,
-a modeler with forgiveness of two would also accept steps with a step-duration
-of two, thus pardoning a single missing fix.
+We conducted a simulation study with known habitat and movement parameters to
+investigate if retaining irregular animal locations via increased forgiveness
+improves or worsens parameter estimation in iSSFs. We also tested the
+performance of four different approaches that attempt to correct for potential
+biases introduced by using temporally irregular data, and we compared them to an
+uncorrected baseline approach. Our results demonstrated that retaining irregular
+animal locations can improve the precision of estimators of habitat-selection
+parameters but may lead to biased estimators of the parameters in
+the movement kernel. Overall, our results highlight the potential benefits of
+leveraging irregular animal locations, especially if an appropriate method for
+handling irregular data is chosen.
 
-The idea of the multiple imputation approach is as follows...
+The \textit{uncorrected} baseline approach ignored the fact that increasing
+forgiveness introduced irregularity in the data. Consequently, estimators of the
+parameters in the movement kernel were increasingly biased as forgiveness
+increased due to the inclusion of steps with varying step durations when fitting
+the model. Steps with longer step durations tended to have larger step lengths
+and less directed turning angles, which led to an overestimation of $\theta$ and
+underestimation of $k$ and $\kappa$. Yet, estimators of the habitat-selection
+parameters remained unbiased or nearly so, even at high levels of forgiveness,
+and they were more precise than the standard estimator (represented by the
+\textit{uncorrected} approach with forgiveness = 1). These results highlight the
+potential benefits that can be reaped by including additional data.
 
-\paragraph{Data Rarefication}
-We rarefied the simulated GPS data by randomly removing a fixed fraction of GPS
-fixes. To assess the impact of different degrees of  ``missingness'', we
-incrementally increased the fraction of missing fixes from 0.0 (complete
-dataset) to 0.8 (80\% missing) with 10\% increments. The removal of GPS fixes
-introduced irregular temporal intervals between remaining fixes; we will refer
-to these intervals as ``step-times'', where the  ``step'' refers to the straight
-line movement between two consecutive GPS fixes \citep{Turchin.1998}. In the
-complete dataset, a step-time of one was assumed, whereas the step-time
-increased by one for every fix that was missing between subsequent fixes. For
-instance, a single missing fix between two other fixes directly increased the
-step time from one to two.
+Similarly, the \textit{na\"ive} approach performed well when estimating
+habitat-selection parameters but resulted in biased estimators of movement
+parameters, especially for large forgiveness values. This result was
+unsurprising given that our simulated trajectories were tortuous, and therefore,
+step lengths were not linearly related to step durations. Indeed, we found that,
+although there was a near linear relationship between step duration and the
+(tentative) scale parameter ($\theta_0$), the relationships between step
+duration and the (tentative) movement parameters $\kappa_0$ and $k_0$ were
+non-linear (Figure S4). Overall, the usefulness of this approach appears highly
+limited, as it is often not clear by what factor distributional parameters for
+step lengths and turning angles should be multiplied to match the observed step
+duration.
 
-\paragraph{Identifying Bursts and Computing Step Metrics}
-We used the generated data to compute movement bursts. A movement burst
-consisted of a sequence consecutive GPS fixes with step-times that did not
-exceed the accepted forgiveness. For instance, if the forgiveness was one,
-already a single missing GPS fix introduced a new burst. In contrast, if the
-forgiveness was set to two, step-times of two units were allowed, without
-introducing a new step.
+The \textit{dynamic+model} approach provided a flexible, easily implementable,
+and powerful framework for retrieving precise and unbiased estimators of the
+step-length and habitat-selection parameters, irrespective of the forgiveness
+level. The estimator of the concentration parameter exhibited some bias, but
+less than when using the \textit{uncorrected} and \textit{na\"ive} approaches.
+To implement the \textit{dynamic+model} approach, we included interactions
+between step descriptors (sl, log(sl), and cos(ta)) and step duration in the
+conditional logistic regression model. This allowed the parameters of the
+movement kernel to depend on the step duration. A complication, however, is that
+turning angles are influenced by the step duration of both the current and
+previous step (Figure S3). The bias in the concentration parameter likely arose
+from only accounting for the step duration associated with the current step and
+not the previous one. Moreover, fitting tentative distributions for different
+step durations can be challenging due to some step durations occurring only
+rarely. However, by resampling observed animal locations to different step
+durations using the \texttt{track\_resample} function from the \texttt{amt}
+\texttt{R}-package \citep{Signer.2019} the needed data can easily be generated.
+We included step duration as a categorical covariate, yet there may be times
+when it would be advantageous to treat it as a continuous covariate (e.g., with
+its effect modeled using a low-degree polynomial or regression spline with few
+degrees of freedom). Treating step duration as a continuous variable may help to
+alleviate convergence issues in cases where some step durations are rare, and it
+might allow applying the \textit{dynamic+model} approach to data that are
+entirely irregular.
 
-We will use the term  ``forgiveness'' to indicate how many missing fixes were
-accepted before a new burst was enforced. For instance, a forgiveness of one
-implied that a single missing fix did not result in a new burst. We then
-converted the GPS fixes belonging to the same burst into steps, where each step
-resembled the straight line movement between two GPS fixes \citep{Turchin.1998}.
-For each step we computed a set of step metrics. This includd the step length,
-its logarithm, the turning angle and its cosine.
+The \textit{multistep} approach also performed well and was relatively easy to
+implement. This approach is somewhat \textit{ad hoc} in that it uses the
+tentative movement parameters to generate random steps to match observed steps
+with longer step durations (in multiples of $\Delta t$). It is similar to, but
+slightly less principled, than the approach developed by \citet{Vales.2022},
+which formally constructs the likelihood for multistep durations by integrating
+out the missing steps. An advantage of this latter approach is that one can also
+attempt to account for non-random missingness by explicitly modeling factors
+related to the probability of obtaining a successful location
+\citep{Vales.2022}. Nonetheless, integrating over the missing steps, as in
+\citet{Vales.2022}, can be computationally intensive and prohibitive with large
+data sets. Another downside of both of these approaches (the \textit{multistep}
+approach and the approach of \citealp{Vales.2022}) is that they can only be
+applied in cases where step durations are a fixed multiple of the regular
+step duration; i.e., unlike the \textit{dynamic+model} approach, they cannot be
+applied when data are highly irregular.
 
-\paragraph{Step Selection Function}
-To conduct step selection analysis, we paired each observed step with 10 random
-steps. We generated random steps by sampling turning angles from a von Mises
-distribution and step lengths from a gamma distribution. Depending on the study
-design, we either sampled turning angles and step lengths from fixed or fixed or
-dynamic distributions (cfr. Section ...). Along observed and random steps, we
-extracted underlying covariates and computed their averages along steps.
+Of the methods we considered, the \textit{imputation} approach performed the
+worst. It resulted in biased estimators of parameters in both the
+habitat-selection function and the movement kernel. This bias likely resulted
+from using an overly simplistic movement model to impute missing fixes.
+Moreover, the imputation procedure may have led to imputed animal locations that
+masked important selection properties, therefore leading to inaccurate parameter
+estimates. While this approach appears to perform well with hidden Markov
+movement models \citep{McClintock.2017}, we advise against its use with iSSFs.
 
-\paragraph{Step Length and Turning Angle Distributions}
-To generate random steps, we employed three competing approaches, coined
-\textit{uncorrected}, \textit{na\"ive}, and \textit{dynamic} approach,
-respectively. In the uncorrected approach, we sampled step-lengths and
-turning-angles from parametric gamma and von Mises distributions that were
-fitted to observed steps with step-duration of one. In the \textit{na\"ive}
-approach, we sampled step lengths and turning angles from the same
-distributions, yet we linearly scaled sampled step lengths to the actual
-step-duration. For instance, for any step with a step-duration of two, we
-doubled the sampled step length. This approach is representative of the
-time-varying SSF approach proposed by \cite{Munden.2021}. Finally, in the
-\textit{dynamic} approach, we sampled step lengths and turning angles from
-distributions that were fit to different step-durations. That is, we fitted
-separate step length and turning angles distributions to every possible step
-duration. To achieve this, we subsampled the observed data to 50\% of the fixes,
-which resulted in varying step-durations. We then and fitted step length and
-turning angle distributions to the steps from different step-durations. We
-repeated this procedure 1000 times and averaged the parameter estimates across
-replicates.
+For the scenarios we considered in our simulation study, the estimators of
+habitat-selection parameters were insensitive to the inclusion of irregular data
+and performed well, except for the \textit{imputation} approach. This suggests
+that accounting for irregular step durations may not be particularly important
+if one is only interested in the habitat-selection function. When the movement
+kernel is also of interest, we suggest the \textit{dynamic+model} approach,
+since it is flexible, easy to implement, and allows one to use more data than
+the traditional approach that requires bursts of regular data, leading to more
+precise estimators.
 
-\paragraph{Regression Model}
-We estimated ... using conditional logistic regression, implemented using the
-r-package \textsf{survival} \citep{Therneau.2021}. We followed \cite{Avgar.2016}
-and \cite{Fieberg.2021} and employed \textit{integrated} SSA. That is, aside
-from the habitat covariates (\textsf{water, elev, dist}), we also included
-descriptors of the step length and turning angle (\textsf{sl, log(sl), cos(ta)})
-in our regression model. The model call was as follows:
+Several authors have emphasized that movement and habitat-selection parameters
+in an SSF are scale dependent and should be expected to change as the sampling
+frequency changes \citep[see for example][]{Avgar.2016, Signer.2017,
+Fieberg.2021}. Furthermore, \citet{Barnett.2008} developed an analytical
+framework for investigating scale dependence and showed that habitat-selection
+parameters should depend on the relative width of the movement kernel in
+relation to habitat heterogeneity. Thus, the relative insensitivity of the
+habitat-selection parameters to the inclusion of steps with varying step
+duration was somewhat unexpected. It would be interesting to explore the
+robustness of this result across a wider range of simulation scenarios in the
+future.
 
-$$
-case\_ \sim water + elev + dist + sl + log(sl) + cos(ta)
-$$
+More generally, the spatial scale of a habitat-selection analysis has been
+recognized as an important factor, which is why \citet{Johnson.1980} proposed a
+hierarchical framework for examining habitat-selection across different orders
+(e.g., species range, individual home range, within a home range). Johnson's
+proposed framework acknowledges that habitat-selection may act differently at
+different scales, and that the interpretation of ecological processes changes
+depending on the spatial scale at which they are investigated \citep{Wiens.1989,
+Levin.1992}. This understanding has encouraged scientists to conduct extensive
+scaling analyses and to comprehensively examine habitat-selection at multiple
+scales \citep{DeCesare.2012, McGarigal.2016, Pitman.2017, Zeller.2017}. In
+studies employing SSAs, the issue of scale is often neglected, and data are most
+frequently analyzed at the spatio-temporal scale at which they were collected.
+This choice maximizes the number of locations that can be used in the analysis,
+yet prevents a thorough understanding of scale dependency. The use of irregular
+data in SSAs poses another challenge, as steps with unequal step durations may
+reflect selection processes occurring at different scales. The severity of this
+issue obviously depends on the original sampling frequency, the degree of
+missingness, and the scale at which animals are making decisions that are
+relevant in terms of their movement behavior and habitat selection. By including
+irregular animal locations via increased forgiveness, we may therefore average
+over selection processes occurring at multiple scales, which could produce
+estimates of habitat-selection parameters that are misleading due to
+contradictory effect signs at different scales. To better account for such
+scale-dependent processes, it may be beneficial to include interactions between
+step duration and habitat features (e.g., \textsf{dist}, \textsf{elev},
+\textsf{forest}), thus allowing habitat-selection parameters to also vary as a
+function of step duration. We demonstrated how this could be implemented in the
+case study.
 
-\subsection{Case Study}
+It is important to note that we considered a limited number of scenarios in our
+simulation study. For instance, we assumed that animal locations were missing at
+random, i.e., failure to obtain a fix was unrelated to habitat types, time of
+the day, etc. However, several studies have shown that missingness is often
+non-random and related to difficulties with satellite transmission due to
+topography \citep{Lewis.2007}, canopy cover \citep{Phillips.1998, DeCesare.2005,
+Hansen.2008}, time of the day \citep{Graves.2006}, animal behavior
+\citep{Mattisson.2010}, or collar orientation \citep{Deon.2005}. In fact,
+\citet{Vales.2022} highlighted that missingness and the associated
+under-representation of certain habitat types may lead to biased estimators of
+parameters in iSSFs, but that accounting for the probability of obtaining a
+location in differing environmental conditions may alleviate this bias. Future
+studies should strive to further investigate these relationships and examine how
+our proposed approaches perform when missingness is habitat-dependent.
 
-because wild dogs move comparably little in the hours between xx and xx, we made
-the simplyfing assumption that the 8-hour period from xx to xx is comparable to
-a 4-hour period.
+A major benefit of using iSSFs is the ability to allow an individual's movement
+kernel to depend on local habitat features \citep{Avgar.2016}. In our simulation
+study, we considered simplified scenarios where the movement kernel was
+unchanging, which simplified the simulation and inference. Nevertheless, such
+interactions often play a crucial role in real ecosystems. For instance,
+\citet{Dickie.2020} employed iSSFs and revealed that several large mammal
+species moved faster while on linear features. Similarly, \citet{Hofmann.2023}
+found that African wild dogs moved significantly slower and less directed in
+areas that were covered by floodwater. Future studies could investigate
+simulation scenarios in which individuals alter their movement tendencies in
+response to local environmental features (i.e., models with habitat dependent
+movement kernels) and examine how this influences the robustness of our proposed
+approaches.
 
-\section{Results}
-\section{Discussion}
-Results from our simulation study demonstrate that the inclusion of GPS fixes
-can be used to gain information on relative habitat preferences of the focal
-species. While the inclusion of steps from irregular sampling schemes resulted
-in substantial biases for estimates related to step metrics, we showed that
-these biases can effectively be mitigated by generating step lengths and turning
-angles from distributions that are separately fit to steps from different step
-durations. Regardless of biases in beta coefficients for step metrics, we found
-that habitat selection estimates were unbiased, even when including irregular
-steps. Even more, the inclusion of irregular steps drastially reduced model
-uncertainty, thus suggesting that such data can be used to gain information that
-would otherwise be lost.
+While our results suggest that irregularity due to missing animal locations can
+effectively be accounted for in iSSFs and that increasing the forgiveness, thus
+allowing for inclusion of irregular data, improves estimator precision, we also
+found a decreasing marginal benefit of increased forgiveness. In fact,
+increasing the forgiveness beyond a value of two (i.e., allowing for steps of
+twice the regular step duration) only marginally improved model performance in
+our case. This can also be seen in \Cref{NumberOfSteps}, which shows that the
+largest number of steps that can be gained is when increasing the forgiveness
+level from one to two. Having a higher forgiveness beyond two may thus not even
+be necessary, therefore limiting the need to correct biases emerging from the
+inclusion of irregular data.
+
+Although we focused on the case of missing location data, the proposed
+approaches may also prove useful for situations where sampling is irregular for
+other reasons. For example, it is not uncommon to adjust sampling regimes after
+a preliminary phase, following improvements to collar-battery-lifetime, or for
+sampling rates to vary depending on type and manufacturer of the collar
+\citep{Brown.2023}. Similarly, it is common practice to adjust the GPS regime to
+the biology of the focal species and only record data during a specific time of
+the day (e.g., \citealp{Broekhuis.2013, Cozzi.2013, Elliot.2014}). These
+irregularities might be addressed using the \textit{dynamic+model} approach,
+with interactions between step duration and movement descriptors. Interactions
+between step duration and habitat-selection covariates should also be
+considered, particularly if the sampling regime is adjusted to coincide with
+changes in animal behavior. We expect this approach will work fairly well in
+many cases, but we might expect a slight bias in the estimated concentration
+parameters, as observed in our simulation study.
+
+Our study contributes to the growing body of literature that extends iSSFs and
+improves the method's robustness under various conditions. This includes
+approaches for modeling irregular data \citep{Munden.2021, Eisaguirre.2024},
+accounting for spatial dependence among residuals \citep{ArceGuillen.2023},
+methodological frameworks for fitting iSSFs with random slopes \citep{Muff.2020}
+and random smooths \citep{Klappstein.2024}, incorporating the probability of
+successfully obtaining an animal location in different habitat conditions
+\citep{Vales.2022}, and considering the behavioral states of the tracked animals
+\citep{Klappstein.2023, Pohle.2023}.
+
+In conclusion, our study shows that inclusion of irregular animal locations can
+improve model performance, yet only when an appropriate approach to account for
+irregularity is selected. Here, the \textit{dynamic+model} and
+\textit{multistep} approaches performed well and resulted in improved estimators
+of habitat-selection and movement parameters, even at elevated levels of
+missingness and forgiveness. Both methods are easy to implement, and the
+associated models can readily be fitted using the R-packages \texttt{amt}
+\citep{Signer.2019}, \texttt{survival} \citep{Therneau.2023}, \texttt{coxme}
+\citep{Therneau.2022}, and \texttt{mgcv} \citep{Wood.2011, Wood.2017,
+Klappstein.2024}. To facilitate uptake and encourage use of the proposed
+approaches among practitioners, we provide all of our codes through an online
+repository, which includes an example application of the \textit{dynamic+model}
+approach. With this, we hope practitioners will rethink the common use of
+discarding large portions of data and instead use methods that can accommodate
+irregular data.
 
 \section{Authors' Contributions}
-D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
-D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
-first draft of the manuscript and all authors contributed to the drafts at
-several stages and gave final approval for publication.
+
+D.D.H., G.C., and J.F. conceived the study and designed methodology; D.D.H.
+implemented the analysis, J.F. assisted with modeling; D.D.H. wrote the first
+draft of the manuscript and all authors contributed to the drafts at several
+stages and gave final approval for publication.
 
 \section{Data Availability}
-GPS movement data of dispersing wild dogs is available on dryad
-\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
-proposed approach using simulated data are provided through Github
-(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
-codes required to reproduce the African wild dog case study will be made
-available through an online repository at the time of publication.
 
-\section{Acknowledgements}
-We thank the Ministry of Environment and Tourism of Botswana for granting
-permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
-Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we would like
-to thank Johannes Signer for assisting with the simulation algorithm. This study
-was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
-Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
-Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
-Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
-Grant (31003A\_182286) to A. Ozgul.
+Code to reproduce this study is available through the University of Minnesota's
+Data repository (\url{https://doi.org/10.13020/6wcd-6s43},
+\citealp{Hofmann.2023a}). The repository also includes two example analyses that
+showcase the application of the \textit{dynamic+model} approach to simulated
+data, as well as to the GPS data of Apollo.
+
+\section{Acknowledgements \& Funding}
+
+D.D.H. was funded by a Swiss National Science Foundation Grant No.
+310030\_204478 rewarded to G.C. J.F. was supported by the National Aeronautics
+and Space Administration award 80NSSC21K1182 and received partial salary support
+from the Minnesota Agricultural Experimental Station. We thank two anonymous
+reviewers for their constructive and detailed feedback, which helped to improve
+the quality of this manuscript.
 

### 2024-07-01 Chapter_4/04_Manuscript/Manuscript_Support.tex
Publish PHD

--- 2022-07-20 Chapter_4/04_Manuscript/Manuscript.tex
@@
 \usepackage[flushleft]{threeparttable}
-\usepackage[nomarkers, nolists]{endfloat}
-\usepackage[colorinlistoftodos]{todonotes}
+% \usepackage[nomarkers, nolists]{endfloat}
+\usepackage{scalerel}
+\usepackage{makecell}
+\usepackage{ifthen}
+\usepackage{tikz}
+\usepackage{lineno}
+\usepackage{adjustbox}
+\usetikzlibrary{svg.path}
+
+%------------------------------------------------------------------------------
+%  Bibtex vs Biblatex
+%------------------------------------------------------------------------------
+% While bibtex is very nice to use, it's insanely slow. I therefore work with
+% bibtex until the final stage. Set the value below to "true" or "false" to pick
+% which one should be used.
+\newboolean{usebiblatex}
+\setboolean{usebiblatex}{true}
+
+% Depending on the question, the commands below will prepare the document
+% accordingly
+\ifthenelse{\boolean{usebiblatex}}{
+  %-----------------------------------------------------------------------------
+  %  If using biblatex
+  %-----------------------------------------------------------------------------
+  % Load required packages and refer to correct bibliography
+  \usepackage{csquotes}
+  \usepackage[
+      backend=biber
+    , style=apa          % To display Author-year in the text
+    , natbib=true        % Necessary to use citep, etc.
+    , hyperref=true      % To have clickable links in the references
+    , useeditor=false    % Don't add editor to publications
+    , sorting=ynt        % Sort by year, name, and title
+    , uniquename=false   % Avoid that initials are put
+  ]{biblatex}
+  \addbibresource{Literature.bib}
+
+  % To include author in link
+  \DeclareFieldFormat{citehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{#1}}
+  \DeclareFieldFormat{textcitehyperref}{%
+  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
+  \bibhyperref{%
+    #1%
+    \ifbool{cbx:parens}
+      {\bibcloseparen\global\boolfalse{cbx:parens}}
+      {}}}
+  \savebibmacro{cite}
+  \savebibmacro{textcite}
+  \renewbibmacro*{cite}{%
+    \printtext[citehyperref]{%
+      \restorebibmacro{cite}%
+      \usebibmacro{cite}}}
+  \renewbibmacro*{textcite}{%
+    \ifboolexpr{
+      ( not test {\iffieldundef{prenote}} and
+        test {\ifnumequal{\value{citecount}}{1}} )
+      or
+      ( not test {\iffieldundef{postnote}} and
+        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
+    }
+      {\DeclareFieldAlias{textcitehyperref}{noformat}}
+      {}%
+    \printtext[textcitehyperref]{%
+      \restorebibmacro{textcite}%
+      \usebibmacro{textcite}}}
+
+  % Remove editor form everything but books
+  \AtEveryBibitem{%
+   \ifentrytype{book}{}{%
+    \clearname{editor}
+   }
+  }
+}{
+  %-----------------------------------------------------------------------------
+  %  If using bibtex
+  %-----------------------------------------------------------------------------
+  % Load required packages and select citation style
+  \usepackage[round]{natbib}
+  \bibliographystyle{apalike}
+}
 
@@
 %------------------------------------------------------------------------------
-\title{Applying Step Selection Functions to Data with Data Gaps: Simulation and
-Case Study on African wild dogs}
+\title{\textbf{Appendix}\\ Methods for Implementing Integrated Step-Selection
+Functions with Incomplete Data}
 
@@
 
-\textbf{Keywords:} step selection analysis, gps data, animal movement, irregular
-data
+\textbf{Keywords:} animal movement, gps data, imputation, incomplete data,
+missing fixes, step-selection analyses, step-selection functions
 
 \end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Abstract
-%------------------------------------------------------------------------------
-\newpage
-\begin{abstract}
-Animal movement data collected using GPS collars is often noisy and incomplete,
-thus confronting movement ecologists with data that is not always uniformly
-distributed in time. Step selection analysis has become a popular tool to
-analyse such data and to infer a species movement and habitat preferences, yet
-it requires that consecutive GPS relocations are collected at regular temporal
-intervals. To adhere to this requirement, researchers typically discard large
-portions of their data and only consider GPS bursts that were collected at
-comparable intervals. Here, we assess the appropriateness of this practice and
-use simulated data to examine if including temporally irregular data could, in
-fact, improve estimates from step-selection analyses. Moreover, we explore the
-usefulness of several methods to account for the temporal irregularity. Our
-results suggest that utilizing a full dataset, despite potential temporal
-irregularities, is beneficial, especially when an appropriate method is chosen
-to account for such irregularities.
-\end{abstract}
 
@@
 
-\section{Introduction}
-GPS data is usually collected on fixed temporal intervals (could do a quick and
-dirty analysis on MoveBank) ranging from a few minutes, to multiple hours or
-even days between subsequent reolocations. Nevertheless, GPS devices frequently
-fail to adhere to the anticipated schedule, for instance due to cloud cover or
-other unfavorable conditions, thus resulting in irregular data. In other cases,
-the GPS schedule is not strictly fixed but may be subject to change. For
-example, \cite{Cozzi.2020} increased the fixrate of their GPS collars once per
-week from 4 to 8 fixes per day, aiming to collect more fine scaled data for a
-limited time. We chose to coin this type of data \textit{irregular}, because it
-is not collected on the anticipated fixrate schedule and therefore not directly
-comparable to the vast majority of the collected data.
+% Change to appendix counters
+\appendix
+\beginappendix
 
-Step selection analysis is a powerful framework that enables researchers to
-study habitat and movement preferences of their focal species. The method works
-by comparing spatial covariates at locations where the studied animal has been
-observed, to a set of locations where the animal was presumably absent. For
-this, the collected GPS data is converted into steps, where a step represents
-the straight line movement between two consecutive GPS fixes. One can then
-compare covariates experienced along observed steps to covariates along
-potential alternative steps and thus infer preferred or avoided features. By
-restricting the availability domain to a set of steps, SSA is specifically
-intended to account for spatio-temporal autocorrelation inherent to data
-collected using GPS collars and is therefore one of the preferred methods for
-analysing such data. However, it is considered good practice to only consider
-steps with similar step durations in step-selection models. Depending on the
-amount of available data and depending on the number of missing GPS fixes, this
-practice might entail little sacrifice. In some situations, however, removing
-irregular data implies a substantial loss of information. Additionally, recent
-developments in the SSA realm have brought forward several improvements to SSA
-that might allow to relax the assumption of regular step-durations. This
-includes the \textit{integrated} SSA approach presented by \citep{Avgar.2016},
-in which inference on habitat \textit{and} movement preferences are possible
-thanks to the inclusion of movement descriptors in the respective model. More
-recently, the method has been further refined by \citep{Munden.2021}, who coined
-the term \textit{time-varying integrated} SSA. This method was developed for
-high frequency data which has been rarified using a change-point detection
-algorithm, thus also resulting in irregular step-durations. In case of missing
-steps, in contrast, steps durations are not entirely random but usually
-clustered.
-
-Here, we questioned the practice of removing any irregular data in SSA and
-investigated whether such data could be used to inform step selection models.
-For this, we conducted a simulation study where we simulated movement
-trajectories with known preferences across a virtual landscape. We then
-artifically rarified the ``observed'' data by removing GPS fixes and we analysed
-this data using SSA. This allowed us to investigate if and how the inclusion of
-irregular data affected model estimates. We also varied the degree of
-irregularity in the data and we tested for the effects of adjusting the
-parametric step-length and turning distributions to different step-durations.
-Besides the simulation study, we also analysed a dataset collected on dispersing
-African wild dogs and examined the implications considering or discarding
-irregular data.
-
-Specifically, we employed six competing approaches to handle temporal
-irregularity in the data.
-
-\begin{figure}
+%------------------------------------------------------------------------------
+%	Appendix S1:
+%------------------------------------------------------------------------------
+\newpage
+\section{Landscape Simulation: Different Autocorrelation Scenarios}
+\begin{figure}[!ht]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
-  \caption{}
+  \includegraphics[width = \textwidth]{99_CovariatesAppendix.png}
+  \caption{Simulated landscapes under different levels of autocorrelation (5,
+  10, 20; from top to bottom). Autocorrelation only affected the layers
+  \textsf{elev} and \textsf{forest}, which were both simulated using a Gaussian
+  random field neutral landscape model \citep{Schlather.2015} using the
+  \texttt{R}-package \texttt{NLMR} \citep{Sciaini.2018}. Simulations were
+  repeated 100 times for each autocorrelation scenario, thus resulting in 300
+  unique landscape configurations.}
   \label{Covariates}
@@
 
-We hypothesized that the precision of model estimates would decrease as we
-increased the missigngess in the data. We also expected that the inclusion of
-irregular data would not improve the precision of estimates and rather lead to
-biased values. However, we anticipated that such biases, at least in estimated
-movement kernel parameters, could be reduced by appropriately adjusting the
-availability domain.
-
-\section{Methods}
-\subsection{Simulation Study}
-\subsubsection{Spatial Covariates}
-We simulated a virtual landscape comprising three spatial covariate layers, each
-with a resolution of 300 x 300 pixels (\Cref{Covariates}) spanning across x- and
-y-coordinates from 0 to 300. The first layer (\textsf{water}) represented
-water-bodies and was simulated using a random cluster nearest‐neighbour neutral
-landscape model \citep{Saura.2000}, with the cluster-proportion set to 0.5 and
-the patch occupancy of water fixed to 20\%. The second layer (\textsf{elev})
-resembled an elevation layer and was simulated using a Gaussian random fields
-neutral landscape model \citep{Schlather.2015}, with an autocorrelation range of
-10, magnitude of variation of the landscape of 1, and a magnitude of variation
-in scale of 0. We simulated both of these layers using the r-package {\tt NLMR}
-\citep{Sciaini.2018}. The third layer (\textsf{dist}) simply indicated the
-distance (in pixels) to the center of the virtual landscape \((x = 150, y =
-150)\), and can be understood as the distance to the center of an animal's
-home-range. We computed spatial distances using the r-package {\tt raster}
-\citep{Hijmans.2022}. We normalized the values from all layers to a range
-between zero and one.
-
-\begin{figure}
+%------------------------------------------------------------------------------
+%	Appendix S2:
+%------------------------------------------------------------------------------
+\newpage
+\section{Dynamic Tentative Distribution Parameters}
+\begin{figure}[!ht]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Covariates.png}
-  \caption{Virtual landscape across which we simulated movement trajectories.
-  All layers have a resolution of 300 x 300 pixels and were generated randomly.
-  Simulated individuals were initiated within the white dashed rectangle, which
-  ensured that they would not be released directly at a map border.}
-  \label{Covariates}
-  \end{center}
-\end{figure}
-
-\subsubsection{Movement Simulation}
-To simulate movement across the above generated virtual lanscape, we employed an
-``inverted'' step-selection function, which proceeded as follows. First, we
-generated a random starting location. To prevent starting points in the vicinity
-of map borders, we restricted sampled locations to x- and y-coordinates between
-50 and 250 (white dotted rectangle in \Cref{Covariates}). Second, we generated a
-set of 10 random steps originating from the sampled starting point. We then
-generated random steps by sampling turning angles from a von Mises distribution
-with concentration parameter \(\kappa = 0.5\) and location parameter \( \mu = 0
-\), and step lengths from a gamma distribution with shape parameter \(k = 3 \)
-and scale parameter \(\theta = 1\). Third, we extracted covariate values along
-each random step from the underlying covariate layers and computed the average
-of each covariate along every step. Fourth, we predicted for each step the
-probability of being chosen by applying \Cref{EQ1}. The vector of relative
-habitat preferences used to make predictions was set to \(\beta_{water} = -1\),
-\(\beta_{elev} = 0.5\), and \(\beta_{dist} = -15\). Fifth, we sampled one of the
-random steps based on predicted probabilities and computed the new position of
-the simulated individual. We the repated these steps until a total of 100 steps
-were realized and we replicated the simulation 100 times, thus resulting in 100
-independent movement trajectories (\Cref{Simulations}). Note that by excluding
-interactions among habitat covariates and step metrics we implicitly assume
-independent habitat and movement kernels \citep{Avgar.2016}. Although
-measurement error would have been straight forward to implement, this was not
-the main focus of our analysis and so we assumed GPS locations to be 100 percent
-accurate.
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = 0.5\textwidth]{99_Simulations.png}
-  \caption{100 Simulated movement trajectories, each comprising of 100 steps.
-  Simulated individuals were initiated within the blue rectangle to mitigate
-  edge effects.}
-  \label{Simulations}
-  \end{center}
-\end{figure}
-
-\subsubsection{Step Selection Analysis}
-To assess the consequences of missing GPS data in when employing step-selection
-functions, we analysed the simulated GPS data under different conditions
-\Cref{Design}. Specifically, we varied the degree of data-missingness (0\% to
-80\%), the forgiveness when analysing the data (1 to 5 fixes), and the
-parametric distributions from which random steps were generated (static vs.
-dynamic distributions). We replicated each combination of conditions 100 times
-and computed averages and standard deviations from the resulting model
-estimates.
-
-In this context, missingness and forgiveness are complementary terms. While
-missingness describes the fraction of data that has not been collected, we use
-the term forgiveness to describe how much missingness a modeler is willing to
-accept. A modeler with forgiveness of one, for instance, only considers steps
-with a step-duration of one (i.e. fixes that are regularly spaced). In contrast,
-a modeler with forgiveness of two would also accept steps with a step-duration
-of two, thus pardoning a single missing fix.
-
-The idea of the multiple imputation approach is as follows...
-
-\paragraph{Data Rarefication}
-We rarefied the simulated GPS data by randomly removing a fixed fraction of GPS
-fixes. To assess the impact of different degrees of  ``missingness'', we
-incrementally increased the fraction of missing fixes from 0.0 (complete
-dataset) to 0.8 (80\% missing) with 10\% increments. The removal of GPS fixes
-introduced irregular temporal intervals between remaining fixes; we will refer
-to these intervals as ``step-times'', where the  ``step'' refers to the straight
-line movement between two consecutive GPS fixes \citep{Turchin.1998}. In the
-complete dataset, a step-time of one was assumed, whereas the step-time
-increased by one for every fix that was missing between subsequent fixes. For
-instance, a single missing fix between two other fixes directly increased the
-step time from one to two.
-
-\paragraph{Identifying Bursts and Computing Step Metrics}
-We used the generated data to compute movement bursts. A movement burst
-consisted of a sequence consecutive GPS fixes with step-times that did not
-exceed the accepted forgiveness. For instance, if the forgiveness was one,
-already a single missing GPS fix introduced a new burst. In contrast, if the
-forgiveness was set to two, step-times of two units were allowed, without
-introducing a new step.
-
-We will use the term  ``forgiveness'' to indicate how many missing fixes were
-accepted before a new burst was enforced. For instance, a forgiveness of one
-implied that a single missing fix did not result in a new burst. We then
-converted the GPS fixes belonging to the same burst into steps, where each step
-resembled the straight line movement between two GPS fixes \citep{Turchin.1998}.
-For each step we computed a set of step metrics. This includd the step length,
-its logarithm, the turning angle and its cosine.
-
-\paragraph{Step Selection Function}
-To conduct step selection analysis, we paired each observed step with 10 random
-steps. We generated random steps by sampling turning angles from a von Mises
-distribution and step lengths from a gamma distribution. Depending on the study
-design, we either sampled turning angles and step lengths from fixed or fixed or
-dynamic distributions (cfr. Section ...). Along observed and random steps, we
-extracted underlying covariates and computed their averages along steps.
-
-\paragraph{Step Length and Turning Angle Distributions}
-To generate random steps, we employed three competing approaches, coined
-\textit{uncorrected}, \textit{na\"ive}, and \textit{dynamic} approach,
-respectively. In the uncorrected approach, we sampled step-lengths and
-turning-angles from parametric gamma and von Mises distributions that were
-fitted to observed steps with step-duration of one. In the \textit{na\"ive}
-approach, we sampled step lengths and turning angles from the same
-distributions, yet we linearly scaled sampled step lengths to the actual
-step-duration. For instance, for any step with a step-duration of two, we
-doubled the sampled step length. This approach is representative of the
-time-varying SSF approach proposed by \cite{Munden.2021}. Finally, in the
-\textit{dynamic} approach, we sampled step lengths and turning angles from
-distributions that were fit to different step-durations. That is, we fitted
-separate step length and turning angles distributions to every possible step
-duration. To achieve this, we subsampled the observed data to 50\% of the fixes,
-which resulted in varying step-durations. We then and fitted step length and
-turning angle distributions to the steps from different step-durations. We
-repeated this procedure 1000 times and averaged the parameter estimates across
-replicates.
-
-\paragraph{Regression Model}
-We estimated ... using conditional logistic regression, implemented using the
-r-package \textsf{survival} \citep{Therneau.2021}. We followed \cite{Avgar.2016}
-and \cite{Fieberg.2021} and employed \textit{integrated} SSA. That is, aside
-from the habitat covariates (\textsf{water, elev, dist}), we also included
-descriptors of the step length and turning angle (\textsf{sl, log(sl), cos(ta)})
-in our regression model. The model call was as follows:
-
-$$
-case\_ \sim water + elev + dist + sl + log(sl) + cos(ta)
-$$
-
-\subsection{Case Study}
-
-because wild dogs move comparably little in the hours between xx and xx, we made
-the simplyfing assumption that the 8-hour period from xx to xx is comparable to
-a 4-hour period.
-
-\section{Results}
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width = \textwidth]{99_DistributionParameters.png}
-  \caption{}
+  \includegraphics[width = 0.9\textwidth]{99_DistributionParameters.png}
+  \caption{Tentative prameter estimates for the von Mises distribution (top row) and
+  gamma distribution (bottom row) fitted to steps with different durations. The
+  von Mises distribution requires one parameter, namely a concentration
+  parameter ($\kappa$). The gamma distribution requires a shape parameter ($k$)
+  and a scale parameter ($\theta$). The subscript $_0$ is used to indicate that
+  these are tentative distribution parameters (sensu \citealp{Avgar.2016} and
+  \citealp{Fieberg.2021}).}
   \label{DistributionParameters}
@@
 
-\begin{figure}
+%------------------------------------------------------------------------------
+%	Appendix S3:
+%------------------------------------------------------------------------------
+\newpage
+\section{Distribution of Relative Turning Angles following Different Step Durations}
+\begin{figure}[!ht]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_ResultsHabitatKernel.png}
-  \caption{}
-  \label{ResultsHabitat}
+  \includegraphics[width = 0.9\textwidth]{99_TurningAnglePreviousDuration.png}
+  \caption{Density of relative turning angles associated with steps of $\Delta t
+  = 1$ following steps with different durations for all three autocorrelation
+  scenarios (5, 10, and 20). To generate this figure, we assumed a missingness
+  of 0.5 and forgiveness of 5.}
+  \label{PreviousDuration}
   \end{center}
@@
 
-\begin{figure}
+%------------------------------------------------------------------------------
+%	Appendix S4:
+%------------------------------------------------------------------------------
+\newpage
+\section{Model Estimates across all Scenarios}
+\begin{figure}[!ht]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_ResultsMovementKernel.png}
-  \caption{}
-  \label{ResultsMovement}
+  \includegraphics[width = \textwidth]{99_ModelComparisonAppendix.png}
+  \caption{Parameter estimates across different autocorrelation scenarios (5,
+  10, 20; panels a, b, and c) and missingness levels (0\% - 50\%; from left
+  to right). True simulation parameters are indicated by the solid black
+  lines. Parameter estimates from the different approaches are given by the
+  colored symbols, and their bootstrap 95\% CIs across 100 replicates
+  by the colored lines.}
+  \label{ModelComparison}
   \end{center}
@@
 
-\section{Discussion}
-Results from our simulation study demonstrate that the inclusion of GPS fixes
-can be used to gain information on relative habitat preferences of the focal
-species. While the inclusion of steps from irregular sampling schemes resulted
-in substantial biases for estimates related to step metrics, we showed that
-these biases can effectively be mitigated by generating step lengths and turning
-angles from distributions that are separately fit to steps from different step
-durations. Regardless of biases in beta coefficients for step metrics, we found
-that habitat selection estimates were unbiased, even when including irregular
-steps. Even more, the inclusion of irregular steps drastially reduced model
-uncertainty, thus suggesting that such data can be used to gain information that
-would otherwise be lost.
+%------------------------------------------------------------------------------
+%	Appendix S5:
+%------------------------------------------------------------------------------
+\newpage
+\section{Case Study Covariates}
 
-\section{Authors' Contributions}
-D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
-D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
-first draft of the manuscript and all authors contributed to the drafts at
-several stages and gave final approval for publication.
+\begin{figure}[htpb]
+  \begin{center}
+  \includegraphics[width = \textwidth]{99_CaseStudyCovariates.png}
+  \caption{Covariates used for the case study, overlaid with the GPS data of a
+  spotted hyena called ``Apollo'' (lines in black). Apollo was originally
+  collared in 2007 in northern Botswana and monitored until 2011. The depicted
+  area is part of the Okavango Delta, which is a massive wetland area. Data was
+  projected to a local projection (EPSG:32734).}
+  \label{CaseStudyCovariates}
+  \end{center}
+\end{figure}
 
-\section{Data Availability}
-GPS movement data of dispersing wild dogs is available on dryad
-\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
-proposed approach using simulated data are provided through Github
-(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
-codes required to reproduce the African wild dog case study will be made
-available through an online repository at the time of publication.
+%------------------------------------------------------------------------------
+%	Appendix S6:
+%------------------------------------------------------------------------------
+\newpage
+\section{Case Study Model Output}
 
-\section{Acknowledgements}
-We thank the Ministry of Environment and Tourism of Botswana for granting
-permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
-Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we would like
-to thank Johannes Signer for assisting with the simulation algorithm. This study
-was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
-Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
-Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
-Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
-Grant (31003A\_182286) to A. Ozgul.
+\begin{table}[htpb]
+  \begin{center}
+  \caption{Model results from the case study using GPS data collected on Apollo.
+  In F1, forgiveness was set to one (only 2-hour steps were considered), whereas
+  in F3-S and F3-SH a forgiveness of three was employed (allowing for step
+  durations of up to 6 hours). In model F3-S, the step duration was interacted
+  with step descriptors. In model F3-SH, step duration was interacted with step
+  descriptors and habitat covariates.}
+  \label{CaseStudyTable}
+    \begin{adjustbox}{max width=0.70\textwidth}
+    \begin{threeparttable}[h]
+      \input{99_CaseStudyTable.tex}
+      \begin{tablenotes}
+       \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
+        \quad *** \(p < 0.01\)}
+      \end{tablenotes}
+    \end{threeparttable}
+  \end{adjustbox}
+  \end{center}
+\end{table}
 

### 2024-07-01 FinalThesis/Acknowledgements/Manuscript.tex
Publish PHD

\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\newpage

This work would not have been possible without help and support from countless
people and organizations. I'd like to start by thanking the Ministry of
Environment and Tourism of Botswana for generously granting permission to
conduct research for this project. Financial support for this thesis was
provided by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust,
Forschungskredit of the University of Zürich (no. FK-21-091), a Swiss National
Science Foundation grant (no. 31003A\_182286) awarded to Arpat Ozgul, and a
Swiss National Science Foundation grant (no. 310030\_204478) awarded to Gabriele
Cozzi.

I am particularly grateful to my supervisor, Gabriele Cozzi, who has initiated
and managed the African wild dog dispersal project over the past decade and has
been an outstanding mentor. His leadership and dedication to research and
conservation were absolutely instrumental to this work. He provided unstinting
support for my doctoral thesis and has significantly contributed to the success
of each chapter. Gabriele played a pivotal role in securing both material and
salary funds for my research, and eventually included me under his own research
grant. Moreover, he introduced me to the challenges and rewards of fieldwork in
one of this planet's most pristine environments. Through his training, I
acquired a diverse set of skills, from tracking and monitoring large carnivores,
navigating and driving in the African Savannah, sedating and equipping animals
with GPS satellite collars, to fixing Land Rovers. Gabriele's constant
supervision ensured my scientific progress, and he continuously provided timely
guidance whenever I veered off course. His unwavering motivation and belief in
my abilities kept me energized, especially during moments of self-doubt. Even
beyond academic support, Gabriele spared no expense and happily assisted Manuela
and me in our endeavor of finding a Land Rover for our upcoming adventures in
Africa. Besides our working relationship, Gabriele and I have developed an integral
friendship, which hopefully continues to grow in the future. Thanks for the
countless BBQs, beers, and Italian fine dines. I don't think anybody can wish
for a better PhD supervisor.

I extend my gratitude to Arpat Ozgul, whose belief in my scientific abilities
and provision of funds enabled me to embark on my PhD as part of his research
group. Arpat was an excellent supervisor, and his scientific rigor significantly
advanced my understanding of ecology and propelled my academic growth. It was a
privilege to be part of his research group and I found tremendous pleasure in
participating in the various group-retreats, potlucks, and other social
gatherings.

I am immensely grateful to Dominik Behr, who has accompanied my academic career
from the very beginning. Dominik demonstrated to me that transitioning into the
field of ecology was not only feasible, but also highlighted the personal
fulfillment this pursuit may bring about. Under his mentorship, I conducted my
master's thesis on dispersing AWDs, which ultimately laid the foundation for
this PhD. He also infected me with the African virus, giving me an unhealthy
passion for African wild dogs and African wildlife in general. Beyond academic
mentorship, Dominik has continuously inspired me to embrace creativity when
crafting PowerPoint presentations, scientific figures, or when taking photos of
wildlife. It's been an incredible journey, and I've enjoyed each and every of
our many fruitful discussions.

I am deeply thankful to have had an opportunity to collaborate with John
Fieberg, whose expertise was invaluable to my academic journey. John is among
the most humble yet inspiring people I have ever met. His ability to return any
email with an in-depth response within less than 24 hours is one in a kind and
makes collaborating with him an absolute blast. If one imagines a statistics
professor as being an uncoordinated nerd, beware of John's ability to throw a
Frisbee at the speed of light; it will prove you wrong. It was a pleasure
organizing and teaching a movement ecology workshop with John in Faido, and I am
looking forward to organizing another one in the future.

I extend my sincere appreciation to thank all my committee members, namely Lukas
Keller, Arpat Ozgul, Juan Manuel Morales, John ``Tico'' McNutt, and Gabriele
Cozzi for their valuable insights and stimulating discussions during our
committee meetings. Their collective expertise and guidance tremendously
facilitated a series of focused and coherent research chapters. I'd particularly
like to thank Lukas for adopting me in his research group without hesitation.

I would like to express special thanks to Florian Hodel for his contagious
ingenuity and infectious humor, which never failed to brighten even the dullest
moments. Florian's creativity has sparked an incredible amount of interesting
ideas and has his friendship has been invaluable to me. It was a shame we only
overlapped for such a short time at the University, although I have a notion
that it was for the better of my productivity. Maybe $\pi$-ranha will provide
further opportunities for collaborations in the future. I also want to thank his
wife, Rahel, who was a source of strength and probably provided more life-advice
than she is aware of. The joy shared with you both has been unparalleled; I even
suffer from soar jaw-muscles from all the laughter whenever we spend time
together.

I am also grateful to all current and former Popecol members, including
Benedetta, Che, Daniela, Dechen, Dilsad, Esin, Eva, Lorena, Lukas, Lorena,
Louis, Monica, Omar, Patrick, Philip, Rasmitah, Tis, and Vera-Lu for an
unforgettable time and many exciting moments. I'd like to especially thank two
master students that I co-supervised, Megan Robinson and Rapha{\"e}l Destriau,
for their effort and motivation which have laid the foundation for future works.
It's rare to supervise students with such drive and dedication, and it's been a
pleasure being part of their exciting works.

Further gratitude is due to the University's secretaries and technical support,
who ensured that all organizational tasks were taken care off; something I'm
terribly bad at. Many thanks to Jacqui, Judith, Ursina, Debra, Susanne, Isabel,
Michel, and Tina.

Thanks also to Johannes Signer and Theo Michelot for sharing their insights on
how to simulate movement from fitted step-selection functions, and to Guillaume
Bastille-Rousseau for helping me with computing network metrics from GPS data.

Furthermore, I want to extend my gratitude to Glauco Camenisch, who assisted at
countless wild dog immobilizations and has always provided an ear when I needed
one. Experiencing the bush together with you and Gabs was a unique privilege and
among the most fun I've had in Africa thus far.

In Botswana, a heartfelt ``Thank You'' goes to Tico and Lesley for establishing
Dog Camp and for the passion and love that they keep pouring into this amazing
research project. Through Wild Entrust, you have created something truly
extraordinary that holds the promise of providing a future where humans and
wildlife can coexist. It is an honor to be part of this endeavor, and one can
only aspire to contribute to its success in the future.

Special thanks to Peter Brack for showing me how to cross water with a Land
Rover (a proper lake, in fact), how to hot-wire a Land Rover, for organizing a
walking safari (let's do another one!), and for the countless inspiring
discussions we had in camp. Similarly, I'd like to thank his wife, Nicola Hawes,
for accommodating me on various occasions in their wonderful ``termite mound''
house in town. Further thanks go to Megan Claase, for exemplifying true
dedication and passion for conservation and for serving as a fieldwork role
model. Our expedition to David's Kingdom marks one of my favorite trips in
Africa to date. Further thanks go to Whitney Hansen, for bringing me on some
incredible wild dog follows and for introducing me to the imperial system
(``come within a whale of me and say that again''). Thanks to Emily Bennitt,
Mark and Caroll Swaffield for providing me with shelter and food in Maun, to Rob
Jackson for providing veterinary services, and to Mary Hastag from Abercrombie
\& Kent for organizing several trips into the Okavango Delta and for initiating
an exciting collaboration that hopefully continues into the future. Gratitude is
also due to all research assistants in Botswana, including Rio Dipuo, John
Neelo, Megan Robinson, Alex Dibnah, and Mara Zali, for relentlessly collecting
data in the field, as well as to all staff members in camp, including Allen, for
providing mechanical support and fixing our Land-Rovers and for his musical
irradiation in the shower, to Fine and Tea for preparing the most elaborate
meals with the least elaborate items, to Borden, Opi, and Obie for keeping camp
clean and tidy, and to Last, for tirelessly fixing our continuously punctured
tires.

Special thanks go to our former research assistant Aobakwe Sengate (a.k.a. AB),
for being an incredible field-mentor and for filling my knowledge gaps with
regard to the African wildlife. Though our time in the field was short, every
trip with AB was exciting, even if it only was to check cameratraps. AB's
abilities to read the environment and track animals are second-to-none, and his
never-ending curiosity and willingness to investigate even the faintest evidence
of large predators were inspiring and informative. He taught me some of the most
useful skills, and I've enjoyed some of the most memorable sightings with him.
I'll never forget the look on his face when he realized a lioness stole a
leopard's kill just behind his tent in camp.

My deepest gratitude goes to my family, including my parents, Rahel and Denis,
and my grandparents Elizabeth, Hans-Peter, and Josephine. They have always been
there for me and provided me with all the psychological and financial stability
that I needed throughout this journey. Similarly, I'd like to thank my
girlfriend's family, namely her parents Yvonne and Max, as well as her
grandparents Willy and Annula. I also want to thank her grandfather Ernst, who
sadly passed away recently. Being embraced by such a wonderful social
environment is a gift beyond measure, and I am incredibly grateful for the love
and support I have received from both families.

Finally, I want to thank my girlfriend Manuela, who supported me through all of
my undertakings, no matter how devious some of them must have sounded to her.
Despite her own demanding schedule, Manuela consistently found the time and
energy to uplift and motivate me when I needed it most. Manuela was not only at
peace with my passion for Africa, but got infected with an unhealthy enthusiasm
herself. Please excuse that I have spent several months in the field, leaving
you and our cat Nala behind, and thank you for crunching through the countless
cameratrap images (there will be more...). It is not a given to find a partner
that can cope with a scientist's lifestyle, and I appreciate your unconditional
support. I can't wait to explore the African wilderness further with you in our
Landy!

\end{document}

### 2024-07-01 FinalThesis/Appendix/Manuscript.tex
Publish PHD

\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}
\setcounter{chapter}{-1}
\chapter{Additional Works}
\renewcommand{\thechapter}{A}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\newpage
\section{Cameratrap Survey}
\label{CameratrapSurvey}

Our current understanding of AWD dispersal movements is primarily based on
models that consider landscape-scale information and environmental features
(e.g., water cover, vegetation cover, human influence) but are completely
uninformed with respect to the effect that intra- and inter-specific factors may
have. Recent studies have indicated that these factors may play a crucial role
in shaping dispersal movements, potentially even outweighing the role of
environmental factors, particularly in social species \citep{Cozzi.2018,
Armansin.2020}. To fill this data gap, I initiated a large-scale cameratrap
survey across BPC's historic study area (\Cref{Cameras}a). The goal of this
ongoing project is to obtain data on the distribution and abundance of AWDs'
main prey and competitor species, while also providing a non-invasive, long-term
monitoring system for the local fauna. Starting in November 2021, I deployed 60
cameras at as many stations on a 4x4 km raster grid, stratified by the three
predominant habitat types; grassland/floodplain (\Cref{Cameras}c), mixed
woodland acacia dominated (\Cref{Cameras}e), and mopane woodland
(\Cref{Cameras}d). To maximize the probability of detecting animals while
reducing our influence on the landscape, I placed cameras on trees
(\Cref{Cameras}b) facing dirt roads and programmed them to record bursts of four
images whenever an animal passed by.

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = 0.9\textwidth]{Figures/Cameras}
  \caption{(a) Camertraps deployed in BPC's historic study area. (b) Cameras
  were deployed on trees facing dirt roads and stratified across three major
  habitat types: (c) grassland / floodplain, (d) mixed woodland acacia
  dominated, and (e) mopane.}
  \label{Cameras}
\end{center}
\end{figure}

To cope with the large amount of incoming data, I developed a processing
pipeline that streamlines and automates numerous processing steps, thereby
reducing the need for manual labor (\Cref{CameratrapPipeline}). Notably, I
programmed a series of \texttt{R}-functions that allow pre-processing the
collected data using an object-oriented approach. Besides trivial tasks, such as
equalizing image dimensions or extracting EXIF-metadata, the functions also
allow feeding images into Microsoft's MegaDetector
(\url{https://github.com/microsoft/CameraTraps}). The MegaDetector utilizes an
AI model to separate images containing animals from those that are empty or
contain people and vehicles. This is particularly useful considering that
cameratraps are often triggered by moving vegetation, leading to false
positives. Finally, the data are automatically prepared for upload to Wildeye's
TrapTagger (\url{www.wildeyeconservation.org}) software for species
classification. In collaboration with Rapha{\"e}l Destriau (see
\Cref{Supervised}), this allowed me to classify animals on all 3.5 Mio.
collected images. As a follow-up from my doctoral thesis, I plan to use the so
collected and processed data to model the distribution of relevant species, and
to include the resulting layers as additional covariates in the dispersal models
that I developed and used thus far.

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = \textwidth]{Figures/CameratrapPipeline}
  \caption{Automated processing pipeline for the cameratrap data. In a first
  step, cameratrap images are preprocessed. This includes the equalization of
  image dimensions, extraction of EXIF-metadata, as well as the assignment of a
  spatial location to each image. In a second step, the pre-processed data is
  fed into Microsoft's MegaDetector, which allows separating empty images from
  images that contain animals. Images with animals are finally processed through
  TrapTagger, which facilitates an AI-aided classification of animals to the
  species level.}
  \label{CameratrapPipeline}
\end{center}
\end{figure}

\section{Supervised Theses}
\label{Supervised}

During the course of my PhD, I had the opportunity to co-supervise two master
students together with my main supervisor, Dr. Gabriele Cozzi. Here, I'd like to
briefly summarize the two students' theses and explain how they embed into my
own work and potentially the work I will conduct during my PostDoc.

\begin{itemize}

  \item \textbf{Factors Influencing Phase-Specific Movement and Activity
  Patterns of African Wild Dogs (by Megan Robinson): } The first thesis was
  conducted by Megan Robinson, who joined our research group in June 2022. Megan
  previously worked as a research assistant for BPC and later embarked as a
  master student in the program for quantitative environmental sciences at the
  University of Zurich. Megan's primary research objective was to quantify
  differences in movement behavior and activity patterns of AWDs during
  residence and dispersal. For this, Megan used GPS and Activity data obtained
  on 33 resident and 19 dispersing AWDs, which she linked to observations
  recorded by staff members in the field to segment the collected data into four
  behavioral movement modes. Specifically, she differentiated between the
  resident phase, when individuals are with the natal pack; the exploratory
  phase, when prospecting individuals may leave the pack for short exploratory
  forays but eventually return to the pack; the transient dispersal phase, when
  dispersing individuals leave their natal pack forever, and the settlement
  phase, during which a newly formed group establishes a new territory. She then
  applied generalized linear mixed-effects models to compare step-lengths,
  turning-angles, and average activity between phases, while holding constant
  for several other factors, including habitat cover, group-size, and
  moon-illumination. Her results revealed that individuals in the exploratory,
  transience, and settlement phase covered significantly larger daily distances
  and exhibited higher average activity than residents (\Cref{MasterTheses}a).
  These differences were particularly pronounced during transience, when
  dispersers cover long distances in a short amount of time. Megan also found
  that AWD activity was positively correlated with moon illumination at night,
  irrespective of the animals' behavioral movement mode. During my PostDoc, I
  envision to build on these insights and further investigate differences
  between resident and dispersing wild dogs via iSSFs. My goal is to further
  assess if observed differences are limited to differences in animals' movement
  behavior (i.e. their movement kernel) or if there are equally relevant
  differences in their habitat-selection function (habitat preferences).

  \item \textbf{Spatial Analysis of Wild Animal Populations in Northern Botswana
  Using Camera Traps Data (by Rapha{\"e}l Destriau):} The second co-supervised
  thesis was conducted by Rapha{\"e}l Destriau, who joined our research group as
  an external student from the Ecole Polytechnique Federale in Lausanne.
  Rapha{\"e}l's background was in engineering, yet he was dedicated to expand
  his expertise in ecology. To accommodate both his expertise and interest, we
  developed a two-stage master thesis, which started with an engineering
  component and later culminated in an ecological analysis. The goal of this
  thesis was to expand on my previously developed processing pipeline to process
  cameratrap data and to conduct a preliminary occupancy analysis on several key
  species. In the first stage, Raph expanded the pipeline and implemented a
  semi-automated workflow to annotate cameratrap images and classify species
  using artificial intelligence. For this, Raph employed an already-existing
  open-source software called \texttt{TrapTagger}
  (\url{www.wildeyeconservation.org}) which provided several pre-trained neural
  networks to detect and classify large mammal species from cameratrap images.
  To benchmark the performances of the available models, Raph manually compiled
  a large validation-dataset that allowed him to compute confusion matrices for
  each available model. Based on these results, Raph developed a
  ``best-practices'' guide that facilitates minimizing potential errors due to
  misclassifications (\Cref{MasterTheses}b), while maximizing the efficiency at
  which data can be processed. In a second stage, Raph utilized the
  automatically processed cameratrap data and applied occupancy models to derive
  species-habitat associations for several species that we believed could
  influence AWDs. The list of modeled species included main AWD prey species
  such as impala (\textit{Aepyceros melampus}), lechwe (\textit{Kobus leche}),
  kudu (\textit{Tragelaphus strepsiceros}), and main AWD competitors such as
  lions (\textit{Panthera leo}), and hyenas (\textit{Crocuta crocuta}). His
  findings suggested significant associations between occupancy and
  environmental factors (e.g. distance to permanent water, vegetation cover,
  proximity to human activity) during the dry season, yet only weak associations
  during the wet season, suggesting either complex ecological dynamics during
  this period or a lack of appropriate spatial information that results in
  stronger species-habitat associations. Raph's thesis laid the foundation for
  an effective pipeline to process additional cameratrap data obtained in
  Botswana, as well as to prepare spatially explicit raster layers that
  represent the distribution of several species that are likely to affect wild
  dog dispersal. During my PostDoc, I plan to rerun Raph's occupancy models
  using updated cameratrap data and to predict spatial layers that can then be
  used as additional covariates in the AWD dispersal model to render the
  influence of competitors and prey on the dispersal behavior of AWDs.

\end{itemize}

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = \textwidth]{Figures/MasterTheses}
  \caption{(a) Results from Megan Robinson's master thesis on differences in
  movement behavior of AWDs depending on their behavioral mode / phase. (b)
  Confusion matrix of predicted vs. true labels as obtained from Rapha{\"e}l
  Destriau's thesis, in which he assessed the reliability of TrapTagger to
  correctly assign species labels to cameratrap images.}
  \label{MasterTheses}
\end{center}
\end{figure}

\section{R-Packages}

During the course of my PhD, I developed four small \texttt{R}-packages that
provided useful functionalities for my own work. The packages are available on
GitHub (\url{https://github.com/DavidDHofmann}) and can be installed via the
\texttt{devtools} package. I will here briefly summarize each package's
functionality:

\begin{itemize}

  \item \texttt{video2pic}: \texttt{video2pic} is an \texttt{R}-package that
  allows extracting still frames from a video. In contrast to existing packages,
  the frame-extraction algorithm is implemented in Python, which improves
  efficiency. The ability to extract still images from videos is particularly
  useful in cameratrapping studies, where videos of animals need to be converted
  into still images for classification.

  \item \texttt{floodmapr}: \texttt{floodmapr} is an \texttt{R}-package that
  allows downloading and classifying MODIS MCD43A4 satellite imagery into binary
  maps of dryland and water cover. The classification algorithm is based on
  \citet{Wolski.2017} and currently only applicable for the extent of the
  Okavango Delta. The package handles all data download, as well as the
  necessary processing steps to obtain a ``floodmap'' of the Okavango Delta for
  any desired date.

  \item \texttt{rainmapr}: \texttt{rainmapr} is an \texttt{R}-package that
  allows downloading spatially mapped precipitation estimates from the CHIRPS
  database \citep{Funk.2015}. CHIRPS provides precipitation data dating back to
  1981 at daily or monthly temporal resolutions and spatial resolutions of up to
  0.05\degree.

  \item \texttt{riversim}: \texttt{riversim} is an \texttt{R}-package that
  provides functionalities to simulate river networks. Comparable algorithms to
  simulate land-cover or land-use patterns already exist (e.g., \texttt{NLMR},
  \citealp{Sciaini.2018} or \texttt{gstat}, \citealp{Pebesma.2004}), yet none of
  them achieves realistic looking river systems. The ability to simulate river
  networks is useful for ecological simulation studies, where rivers are assumed
  to play an important role. The original simulation algorithm was developed by
  a reddit user (``TalksInMaths''), yet I translated it into \texttt{R}.

\end{itemize}

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = \textwidth]{Figures/Packages}
  \caption{Logos of the \texttt{R}-packages that I developed during the course
  of my PhD.}
  \label{Packages}
\end{center}
\end{figure}

\ifSubfilesClassLoaded{%
  \newpage
  \begin{singlespacing}
  \ifthenelse{\boolean{usebiblatex}}{
    \begin{refcontext}[sorting=nyt]
    \printbibliography
    \end{refcontext}
  } {
    \bibliography{../LiteratureBibtex}%
  }
\end{singlespacing}
}{}

\end{document}

### 2024-07-01 FinalThesis/Chapter_00/Manuscript.tex
Publish PHD

\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}
\chapter{General Introduction}
\label{GeneralIntroduction}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\newpage

% \section{Thesis Outline}

% This thesis consists of seven chapters; a general introduction, four data
% chapters, a chapter about additional works, and a general discussion.  In the
% general introduction, I introduce the general background and research topics of
% this thesis. In particular, I discuss the importance of dispersal and
% connectivity in the context of conservation science and highlight multiple
% limitations in current connectivity research. I also introduce the study system
% and give a brief overview of the data collection and general research methods.
% In chapters two to four, I investigate and address several limitations of
% current connectivity research. In chapter five, I outline some additional works
% that I have conducted or co-supervised. Finally, in the general discussion, I
% summarize my findings and embed them in a broader context. I also provide an
% outlook for topics that require further research, and provide conservation
% insights for the African wild dog.
% \newpage

\section{Ecological Networks in the Anthropocene}

Life on earth is in constant motion. Birds migrate seasonally between their
summering and wintering grounds from the Northern to the Southern Hemisphere
\citep{Alerstam.1993}, humpback whales (\textit{Megaptera novaeangliae})
traverse thousands of kilometers to gather at breeding sites
\citep{Rasmussen.2007}, monarch butterflies (\textit{Danaus plexippus}) relocate
from their summer habitats in North America to their winter roosts in Mexico in
a multi-generational journey \citep{Reppert.2018}, and zebras (\textit{Equus
quagga}) and wildebeest (\textit{Connochaetes taurinus}) participate in one of
nature's most spectacular migrations, traversing from the southern Serengeti in
Tanzania to the Masai Mara in Kenya \citep{Serneels.2001}. If there wasn't any
movement, there wouldn't be any life. Life is movement, and movement is life.

Although life on Earth has thrived in motion for millennia, we have entered an
era of decline \citep{IPBES.2019, Almond.2022}. The beginning of the
Anthropocene \citep{Crutzen.2006} not only cemented the domination of humanity
over nature \citep{Kareiva.2007, Johns.2022}, but also heralded a sixth mass
extinction event \citep{Barnosky.2011, Ceballos.2015}. With the current
extinction rate surpassing the background rate by several orders of magnitude
\citep{Ceballos.2015, Ceballos.2017}, approximately 1 million species are at the
brink of extinction \citep{IPBES.2019}. This rapid decline in biodiversity can
mainly be attributed to an ever-expanding human footprint and a continued
exploitation of natural resources \citep{Venter.2016, Ceballos.2017}. To date,
18\% of the Earth's land is heavily modified, 56\% is less severely transformed,
and only 26\% remain unaffected and intact \citep{Locke.2019}. The loss of
formerly healthy habitats and their fragmentation into isolated patches not only
reduces biodiversity directly but also entails a reduction in species ability to
roam freely \citep{Haddad.2015}, thus limiting their capacity to cope with
altered climatic conditions via range shifts \citep{Fahrig.2003, Heller.2009,
Chen.2011, Tucker.2018}. The need for conservation action is evident and urgent.

In order to slow or even reverse the biodiversity loss, many countries have
taken conservation measures and established or expanded protected areas
\citep{Bingham.2021}. Besides the direct protection of endangered ecosystems
through protection zones, the identification and preservation of major movement
corridors that provide connectivity and promote dispersal between remaining
areas is a prominent conservation strategy \citep{Fahrig.2003, Doerr.2011,
Rudnick.2012}. By linking national parks, wildlife management areas, and other
biodiversity hotspots, connectivity shifts conservation from small isolated
``wildlife islands'' to transboundary ecological networks \citep{Hilty.2020}.
Within these networks, animals can again roam freely, allowing them to overcome
resource scarcity, re-colonize extinct patches, and track suitable habitat
conditions. The ability to move and disperse thus crucially contributes to
population viability, stressing the importance of preserving and enhancing
dispersal and connectivity, especially in light of anticipated challenges due to
climate change \citep{Heller.2009}.

\section{Dispersal}

Dispersal of individuals, i.e., the movement of individuals away from their
natal range to the site of first reproduction, is an important process governing
the dynamics of spatially distributed species \citep{Howard.1960, Clobert.2012}.
By moving individuals from one area to another, dispersal constitutes the main
mechanism by which connectivity among subpopulation arises and can be maintained
\citep{Fahrig.2003}. Dispersal differs from migration in that dispersing
individuals do not seasonally move between known patches, but venture into areas
they are unfamiliar with \citep{Hundertmark.2007}. This promotes the
colonization of vacant habitats \citep{Gustafson.1996, Hanski.1999,
MacArthur.2001} and facilitates the reinforcement of small and nonviable
subpopulations \citep{Brown.1977}. Because dispersing individuals (hereafter
referred to as dispersers) cross an inhospitable matrix more willingly than
resident individuals (hereafter residents), they help to offset some of the
negative consequences of habitat-fragmentation by moving between otherwise
isolated habitats \citep{Marsh.2004, Elliot.2014}. At a macro-evolutionary
scale, dispersal enhances range expansions and enables species to track suitable
environmental conditions under climate change \citep{Kokko.2006, Hodgson.2012,
Hodgson.2016}. Finally, dispersal plays a central role in maintaining gene-flow,
thus facilitating adaptation and promoting metapopulation viability and
resilience \citep{Brachet.1999, Marsh.2004, Heinz.2006}. Overall, dispersal
reduces the risk of stochastic extinctions \citep{Shaffer.1985, Melbourne.2008},
making it a key element of many conservation strategies \citep{Baguette.2013}.

\section{Connectivity}

A species' ability to disperse hinges on a sufficient amount of landscape
connectivity \citep{Doerr.2011}. Connectivity is typically defined as the degree
to which a landscape facilitates or impedes movement of organisms among resource
patches \citep{Taylor.1993}, and can be subdivided into \textit{structural} and
\textit{functional} connectivity \citep{Tischendorf.2000}. Structural
connectivity is primarily focused on the size and configuration of habitat
patches within the landscape. It is therefore considered an inherent property of
the landscape that is independent of the focal species, thus overlooking many of
the interesting but complex interactions between organisms and their
environment. Under a \textit{functional} connectivity perspective, conversely,
the focal species' behavioral response to prevailing environmental conditions is
taken into account \citep{Tischendorf.2000}. A landscape with similar structural
properties can thus result in vastly different patterns of functional
connectivity depending on the species of interest. In functionally
well-connected landscapes, species can effectively migrate, disperse, forage,
and establish new populations. Together, these processes contribute to the
resilience, functioning, and diversity of natural systems, and, ultimately,
sustain life on Earth. Preserving functionally connected landscapes has
therefore become the gold-standard in conservation management
\citep{Heller.2009, Doerr.2011, Rudnick.2012}.

\section{Modeling Connectivity}

Although dispersal and connectivity have fascinated and interested scientists
for decades, they both remained notoriously difficult to study until recently.
Dispersers only make up a small fraction of an entire population, and the
proximate causes of dispersal are seldom known (but see \citealp{Behr.2020}).
This makes it challenging to select appropriate individuals and to predict the
exact moment they will emigrate, thus resulting in modest sample sizes, even for
large studies \citep{Nathan.2001, Rudnick.2012, Fattebert.2015}. Following
dispersers is also logistically challenging, due to animals leaving well-defined
study areas and crossing national borders. This holds particularly true for
species that disperse over vast distances \citep{Osipova.2019, Elliot.2014,
Cozzi.2020}. The choice of data and an appropriate connectivity model is
therefore often driven by data availability \citep{Baguette.2013} and frequently
lacks biological justification \citep{Sawyer.2011, Zeller.2012}. Furthermore,
connectivity studies usually cover extensive spatial extents, requiring coherent
environmental data at very large spatial and temporal scales. Thanks to the
uprise and miniaturization of automated tracking devices, the task of monitoring
dispersal has become less problematic, allowing to remotely follow the fate of
dispersing individuals \citep{Cagnacci.2010, Kays.2015, Jonsson.2016,
Williams.2019, Nathan.2022}. Simultaneously, an increased availability of remote
sensed satellite products allows obtaining environmental data at the necessary
spatio-temporal scale to link observed dispersal movements to habitat
characteristics \citep{Toth.2016, Rumiano.2020}.

Thanks to the rapid increase in data-availability on both species and
landscapes, various modeling techniques to quantify connectivity have emerged
(see e.g. \citealp{Etherington.2016} and \citealp{Diniz.2019} for reviews,
\Cref{ConnectivityModeling}). Early approaches were limited to examining
structural aspects of connectivity, focusing on the composition and
configuration of habitat patches, while ignoring the focal species' response to
environmental conditions \citep{Doerr.2011, Diniz.2019}. With the uprise of
tracking data and novel methods to study species' habitat and movement
preferences (\citealp{Boyce.2002, Fortin.2005, Cushman.2010, Avgar.2016,
Fieberg.2021}), however, the focus has shifted towards a more functional view,
taking into account how species react to prevailing habitat conditions
\citep{Diniz.2019}. Presently, the most prominent functional connectivity models
are based on least-cost path analysis (LCPA; \citealp{Adriaensen.2003}), circuit
theory (CT; \citealp{McRae.2006, McRae.2008}) and individual-based movement
models (IBMMs; \citealp{Kanagaraj.2013, Allen.2016, Hauenstein.2019,
FletcherJr..2019, Zeller.2020, UnnithanKumar.2022, FletcherJr..2023}). LCPA and
CT are graph-based methods that estimate conductance of the landscape, whereas
IBMMs use simulations to quantify connectivity. LCPA attempts to identify
least-costly routes between predefined start and endpoints, whereas CT assumes a
random walk and uses laws from electrical circuit theory to calculate
conductance through each area of the landscape. Notably, LCPA and CT are linked
via a concept coined \textit{randomized} LCPA, which allows deviations (the
degree of which can be specified by the user) from the least-costly route, thus
approximating a random walk \citep{Panzacchi.2016}. Except for a few IBMMs, all
functional connectivity models require a permeability surface as an input
\citep{Diniz.2019}. This surface is a spatial raster layer indicating the
expected ease or difficulty at which the focal species can traverse a specific
area given the area's environmental characteristics \citep{Adriaensen.2003,
Zeller.2012}. Permeability can be gauged from expert opinion or using empirical
data, yet the latter approach has proven more accurate \citep{Clevenger.2002,
Zeller.2012}, especially when GPS data are available \citep{Elliot.2014,
Graves.2014, Gaston.2016, Jackson.2016, Keeley.2017}. Unfortunately, due to the
limited availability of dispersal data, most studies utilize data collected on
resident individuals (e.g., \citealp{Brennan.2020, Lines.2021}), despite the
perception and motivation of dispersing and resident individuals to cross
landscape features is arguably different \citep{Elliot.2014, Benz.2016}. Popular
approaches to empirically parametrize permeability surfaces are
resource-selection functions \citep{Fortin.2005, Manly.2007, Thurfjell.2014,
Cushman.2010}, which estimate habitat preferences of the focal species by
comparing used habitat to available habitat \citep{Zeller.2012}. For GPS data,
the use of integrated step-selection functions (iSSFs) has proven particularly
useful, as they readily account for the autocorrelation inherent to GPS data
\citep{Fortin.2005, Avgar.2016}.

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = \textwidth]{Figures/ConnectivityModeling}
  \caption{Conceptualized illustration of a typical connectivity modeling
  pipeline. The pipeline typically starts by collection information on the focal
  species, as well as on environmental features believed to influence the
  species' movement. In a next step, this data is paired and fed into a
  selection model, which empirically estimates selection or avoidance towards
  different landscape conditions. Most frequently, this is achieved via point-
  (or home-range) selection functions, path-selection functions, and
  step-selection functions \citep{Zeller.2012}. Estimated preferences (here
  denoted $\beta$) are then used to predict a permeability surface, indicating
  the expected ease or difficulty at which the modeled species can traverse each
  grid-cell in the landscape, given that grid-cell's covariate values. Finally,
  the permeability surface is used as an input to a connectivity models.
  Currently, least-cost models and circuit theory are among the most prominent
  connectivity models, yet individual-based have been employed as well
  \citep{Diniz.2019}. Figure adapted from \citet{Zeller.2012} and
  \citet{Diniz.2019}.}
  \label{ConnectivityModeling}
\end{center}
\end{figure}

\section{Limitations of Current Approaches}

Despite significant advances in the realms of modeling dispersal and
connectivity, important limitations remain. For my master's thesis, I applied
iSSFs to parametrize a permeability surface and employed LPCA to identify major
movement corridors for dispersing African wild dogs \citep{Hofmann.2021}. This
work exposed me to several of the existing limitations and sparked the
motivation for some of the chapters of this PhD thesis.

\subsection{Use of Traditional Connectivity Models}

A first limitation concerns the use of traditional LCPA and CT (e.g.,
\citealp{Elliot.2014, Mallory.2019, Brennan.2020}), which both make assumptions
that are rarely met by dispersers. LCPA, for instance, assumes that dispersers
have perfect knowledge about their target location and associated movement
costs, allowing them to find the least-costly route. CT, conversely, presumes
that individuals follow a random walk and are ignorant of their target location.
Furthermore, both methods hinge on a permeability surface, which is an
\textit{unconditional} prediction of permeability and cannot render that an
animal's decision to disperse across a certain habitat is contingent on what
else is available \citep{Signer.2017}. IBMMs are a promising alternative that
eliminate the need for a permeability surface \citep{Signer.2017, Diniz.2019,
UnnithanKumar.2022} and enable more realistically rendering an animal's
landscape perception \citep{Zeller.2020a, Zeller.2020}. However, a unified
framework for the parametrization and application of IBMMs to estimate
connectivity is missing.

\subsection{The Assumption of Static Landscapes}

A second limitation in current connectivity analyses lies in their exclusive
focus on present environmental conditions without consideration of how climate
change could affect on-the-ground conditions and therefore species' ability to
disperse (but see \citealp{Ashrafzadeh.2019} and \citealp{Luo.2021}). Yet, an
emphasis on present rather than future environmental conditions might lead to
misguided conservation recommendations in light of ongoing climate change. The
ability to model how climate change affects landscape connectivity would
therefore be highly valuable. A similar limitation arises from the use of static
environmental data, which provide only a snapshot in time and, consequently,
overlook spatio-temporal variations in connectivity due to seasonality
\citep{Simpkins.2017a, Zeller.2020a}. Since most ecosystems experience
significant seasonal changes in environmental conditions, it can be anticipated
that connectivity is not a static property and should be modeled dynamically.
Since incorporating seasonality entails substantial complexity, the question
then becomes to what degree seasonality should enter connectivity analyses to
still distill meaningful conservation insights \citep{Zeller.2020a}.

\subsection{Data Irregularity}

Finally, a last and more technical limitation relates to the use of iSSFs. iSSFs
are regularly used in connectivity modeling to obtain species-specific habitat
preferences and movement capacities (or movement preferences) from GPS location
data \citep{Avgar.2017, Fieberg.2021}. For the method to work, GPS data need to
be collected at regular temporal intervals, which in reality is rare due to GPS
devices occasionally failing to obtain a location. The current practice of
removing irregular data can result in substantial losses of data, which can have
negative consequences, especially in cases where data are already scarce.
Alternatives that allow coping with slightly irregular data would therefore be
desirable.

\section{Thesis Goals}

My goal with this thesis is to address the above outlined limitations and, where
applicable, to provide methodological alternatives. I will do so using
dispersing African wild dogs from the Okavango Delta area in northern Botswana
as a study system. This system lends itself to examine the described limitations
for several reasons. Firstly, the African wild dog is among the world's most
wide-ranging species and a species of immediate conservation concern. The study
population in Botswana is considered a stronghold population that likely acts as
a source for the recolonization of surrounding areas \citep{Cozzi.2013a,
Cozzi.2020}. Furthermore, the African wild dog exemplifies a species that
crucially depends on dispersal and transboundary connectivity for population
viability (cfr. \Cref{StudySpecies}). Because the study area in Botswana
undergoes significant seasonal changes \citep{Mendelsohn.2010} and is among the
most vulnerable to climate change \citep{Akinyemi.2019, IPCC.2022}, it also
offers a unique opportunity for studying the impacts of seasonality and climate
change on connectivity (cfr. \Cref{StudyArea}).

\section{Study Species}
\label{StudySpecies}

The African wild dog (henceforth abbreviated as AWD, \textit{Lycaon pictus}) is
a medium-sized canid native to Sub-Saharan Africa and the only extant
representative of its genus \citep{Kingdon.2015}. It is recognized by its
individually distinct, tricolored coat pattern and large ears (\Cref{WildDog}).
The AWD is a highly social animal, forming packs comprising 5 to 15 adult
individuals that operate as cohesive units \citep{Kuhme.1965, Frame.1979}.
Established packs exhibit a well-defined social structure, with a dominant pair
monopolizing the majority of reproduction and subordinates helping to rear pups
\citep{Frame.1979}.

Once widespread across Sub-Saharan Africa (\Cref{GeneralStudyArea}a), AWDs have
disappeared from a vast majority of their historic range. This decline is mainly
traced back to deadly diseases, human persecution, and habitat destruction
\citep{Fanshawe.1991, Woodroffe.2020}. With fewer than 6,000 individuals
remaining in the wild, the IUCN categorizes the AWD as endangered
\citep{Woodroffe.2020}. The biggest continuous population remains in northern
Botswana, which is the main study area of this research project. AWDs occupy
vast home-ranges, encompassing between 200 km\textsuperscript{2} and 2,000
km\textsuperscript{2} \citep{Pomilia.2015}. Due to their need for vast,
undisturbed spaces, AWDs are particularly vulnerable to habitat fragmentation
and have become an umbrella species for conservation efforts across sub-Saharan
Africa \citep{Dalerum.2008, Brennan.2020}. AWDs thus offer a unique opportunity
for studying transboundary connectivity.

AWDs are crepuscular to diurnal hunters \citep{Saleni.2007}, active primarily
during the cooler morning and evening hours \citep{Kuhme.1965, Creel.2001}, as
well as during moonlit nights \citep{Estes.1967, Saleni.2007, Cozzi.2012}. With
a lean and athletic build, they are exceptionally well-adapted to endurance
hunting \citep{Taylor.1971, Estes.1967, Koshy.2020, Smith.2020}, allowing them
to chase prey until exhaustion \citep{Estes.1967, Creel.1995, Rhodes.2004}. In
the focal study area, packs mainly hunt by means of opportunistic and individual
chases \citep{Hubel.2016}. They predominantly predate on medium-sized antelopes,
including impala (\textit{Aepyceros melampus}) and kudu (\textit{Tragelaphus
strepsiceros}), but occasionally take smaller quarry, such as warthogs
(\textit{Phacochoerus africanus}) and steenbok (\textit{Raphicerus campestris},
\citealp{Creel.1995, Mills.1997, Hayward.2006, Tshimologo.2021}).

Across the majority of their range, AWDs coexist and compete with stronger
competitors, especially with lions (\textit{Panthera leo}) and spotted hyenas
(\textit{Crocuta crocuta} \citealp{Creel.2001, Cozzi.2012, Vogel.2019,
Creel.2023}). To evade potentially fatal encounters, AWDs avoid both species
temporally and spatially \citep{Creel.1996, Cozzi.2012, Droge.2017},
particularly during the denning season \citep{VanDerMeer.2013, Mbizah.2014,
Jackson.2014, Davies.2016, Alting.2021}. The displacement of AWDs by stronger
competitors may explain why the species occurs predominantly in areas of
intermediate prey density \citep{Mills.1997, Mbizah.2014, Creel.2002,
Creel.2023}.

Upon reaching sexual maturity (1.5 to 3 years of age), individuals disperse from
their natal pack in an attempt to find potential mates and a suitable territory
to settle \citep{McNutt.1996, Behr.2020}. Both sexes disperse, typically in
same-sex sibling coalitions comprising two to five members \citep{Frame.1979,
McNutt.1996, Behr.2020}. On average, dispersing AWDs cover Euclidean distances
between 5 and 500 km \citep{Davies-Mostert.2012, Masenga.2016, Cozzi.2020,
Sandoval-Seres.2022}, yet cumulative distances up to 5,000 km have been reported
\citep{Masenga.2016}. Studies on habitat-selection suggest that dispersing AWDs
avoid areas covered by water and dominated by humans, but prefer open shrubs and
grassland \citep{Cozzi.2020, ONeill.2020, Hofmann.2021}. During dispersal, their
movements are characterized by fast, directional bursts away from their natal
location \citep{Cozzi.2020, Hofmann.2021}.

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = \textwidth]{Figures/WildDogPoster}
  \caption{(a) An adult African wild dog photographed in northern Botswana. The
  distinct fur markings are unique to each individual. (b) An anesthetized wild
  dog is being equipped with a GPS/Satellite collar, allowing to remotely follow
  its movements. (c) An African wild dog chasing prey at high speed.}
  \label{WildDog}
\end{center}
\end{figure}

\section{Study Area}
\label{StudyArea}

Data on dispersing AWDs were collected in northern Botswana as part of a
collaborative effort between the Botswana Predator Conservation Program (BPC)
and the University of Zurich. BPC was founded in 1989, when John Weldon ``Tico''
McNutt commenced his pioneering work on AWDs in Botswana and established a
research station \citep{Fuller.1992, McNutt.1996, Osofsky.1996}. The program has
continued ever-since, making it the longest running research program on AWDs and
resulting in invaluable long-term data (e.g., \citealp{McNutt.1996, Cozzi.2012,
Broekhuis.2013, Abrahms.2016, Behr.2020, Jordan.2022}).

BPC's historic study area (approx. 3,000 km\textsuperscript{2}, centered at
19\degree 31'S, 23\degree 38'E, \Cref{GeneralStudyArea}b) is located at
the Eastern fringes of the world-renowned Okavango Delta, a flood-pulse driven
mosaic of patchy woodlands, permanent swamps, and seasonally flooded grasslands
that lie within the otherwise dry and sandy Kalahari Basin \citep{Wilson.1976,
Ramberg.2006, Mendelsohn.2010}. The Okavango Delta represents the world's
largest inland delta and is classified as a Ramsar site, supporting a rich
ecosystem with diverse flora and fauna. Precipitation in this area is highly
seasonal and limited to the wet-season between October and April. Rainwater is
collected in the Angolan highlands and only slowly descends through the Okavango
Delta's tributaries into its alluvial fan, so the flood is out of sync with
local rains and reaches its maximum extent during peak dry season in July-August
\citep{Wolski.2017}.

Human influence across the study area remains low, with large portions of land
being gazetted national parks, wildlife management areas, or other protected
areas. The study area is also part of the Kavango-Zambezi Transfrontier
Conservation-Area (KAZA-TFCA, \Cref{GeneralStudyArea}a), the world's largest
transboundary conservation initiative. The KAZA-TFCA was initially aimed towards
restoring connectivity for migrating elephants (\textit{Loxodonta africana}),
but has gained momentum as an opportunity to establish connectivity for several
other wide-ranging species, including the African wild dog \citep{Brennan.2020,
Hofmann.2021, Sandoval-Seres.2022}.

\begin{figure}[htpb]
\begin{center}
  \includegraphics[width = \textwidth]{Figures/GeneralStudyArea.pdf}
  \caption{(a) The main study area of this thesis is located in northern
  Botswana and forms part of the world's largest transboundary conservation
  initiative, the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA).
  (b) Within this area, Botswana Predator Conservation (BPC) has continuously
  monitored a core area, spanning about 3,000 km\textsuperscript{2}. The
  dominant geographical feature of this ecosystem is the Okavango Delta, a flood
  pulse driven conglomerate of diverse habitats. Dispersers were collared within
  BPC's main study area and moved into various regions, including the Chobe
  National Park in Botswana and the Hwange National Park in Zimbabwe.}
  \label{GeneralStudyArea}
\end{center}
\end{figure}

\section{General Methods}

\subsection{Data Collection}

GPS data on dispersing AWDs were collected between 2011 and 2023 using
GPS/Satellite collars (Vertex Lite; Vectronic Aerospace GmbH, Germany).
Candidate dispersing individuals were selected based on age, number of same-sex
siblings, pack size and presence of unrelated individuals of the opposite sex in
their pack \citep{McNutt.1996, Behr.2020}. Following the procedure described by
\citet{Osofsky.1996}, selected individuals were immobilized using a mixture of
ketamine and xylazine, which was subsequently reversed with yohimbine. Drugs
were injected through a dart shot from a CO\textsubscript{2}-pressurized
dart-gun (Dan-Inject, Denmark) at distances ranging from 10-12 m. During
anesthesia, collars were attached, body measurements recorded, and blood samples
collected. After 45 minutes, the anesthesia was reversed and the animal's
recovery was monitored until it rejoined its pack. All required procedures were
conducted by a Botswana-registered wildlife veterinarian, and all dispersers
were collared while they were still with their natal pack. Deployed collars
weighed about 330 g ($\approx$ 1\% of the species' body weight) and contained a
decomposable mechanism that ensured drop-off of the collar after 12–18 months.
Collars were programmed to record a GPS location every 4 hours during dispersal,
and to regularly transmit data via Iridium satellite system to a base station.
To distinguish between residency and dispersal in the recorded data, I relied on
a combination of data collected in the field and visual inspection of the
net-squared displacement metric. Net-squared displacement measures the squared
Euclidean distance of a location to a reference point \citep{Borger.2012}, which
I set at the center of the dispersing wild dog's natal range. Thus, I deemed
dispersal to have started when an individual left its pack and natal range, and
I defined settlement once the net-squared displacement metric flattened off.
Besides GPS data on dispersers, I compiled an extensive collection of data on
environmental conditions (e.g. water cover, human influence, temperature)
sourced through publicly available services, or remote sensed from free
satellite imagery. Specific details about additional data will be introduced in
the relevant chapters.

I completed all processing steps and analyses using R \citep{RCoreTeam.2023} and
performed computations exclusively on a desktop PC with AMD Ryzen 7 2700X
octa-core processor (8 x 3.6 GHz, 16 threads) and 64 GB DDR4 RAM. All
\texttt{R}-scripts and \LaTeX files associated with this thesis are available
through GitHub (\url{https://github.com/DavidDHofmann/PhD}).

\subsection{Statistical Analyses}

Throughout all chapters, I heavily rely on the framework of iSSFs. I therefore
briefly summarize the method here. iSSFs are frequently applied in movement
ecology to study an animal's habitat preferences and movement capacity
\citep{Fortin.2005, Thurfjell.2014, Avgar.2016}. They operate in discrete time
and mathematically model the probability $u$ of finding an individual at a
location $s$ at time $t+1$, given the animal's past positions at time $t$ and
$t-1$, $s_t$ and $s_{t-1}$. Formally, this probability is given by:

\begin{equation}
\label{EQ0}
u(s_{t+1}) = \frac{\phi(s_{t+1}, s_t, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)}{\int_{s \in G}\phi(s_{t+1}, s_{t}, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)ds}
\end{equation}

\noindent The function $\phi$ represents an animal's movement kernel which is
usually expressed in terms of a step-length and turning-angle distribution, with
$\gamma$ representing parameters in these distributions (for simplicity, I will
sometimes refer to the \textit{movement kernel} as \textit{movement
preferences}). The function $w$ defines the habitat-selection function and
reflects an animal's preferences $\beta$ towards environmental characteristics
$x$ at location $s_{t+1}$. In most applications, $w$ is defined as the
log-linear function $w = exp(x^\top\beta)$. Finally, the integral in the
denominator of \ref{EQ0} ensures that $u$ is a proper probability distribution
and integrates to one.

To estimate parameters in \ref{EQ0} using empirical data, one needs to convert
observed GPS locations into sequences of temporally regularly-spaced steps.
Here, a step is defined as the straight line connecting two consecutive GPS
locations \citep{Turchin.1998} and is characterized by a step length, absolute
turning angle (a.k.a. heading), and relative turning angle. The likelihood in
\ref{EQ0} can then be maximized by comparing observed steps with random steps in
a (mixed effects) conditional logistic regression model (\citealp{Fortin.2005,
Muff.2020}, but see \citealp{Michelot.2016} for alternatives). The required
random steps can be generated by sampling step lengths and turning angles from
parametric distributions fitted to empirical data \citep{Fortin.2005,
Thurfjell.2014}.

To jointly estimate parameters in $\phi$ and $w$, movement descriptors (e.g.,
step length (sl), its natural logarithm (log(sl)), and the cosine of the turning
angle (cos(ta))) can be included in the regression model, and their estimated
coefficients serve to update the tentative parameters of the step-length and
turning-angle distributions \citep{Duchesne.2015, Avgar.2016, Fieberg.2021}. By
including the relevant predictors, iSSFs allow modeling highly complex movement
behaviors (e.g., \citealp{Munden.2021, Forrest.2024}). Notably, a model fitted
using iSSFs constitutes a mechanistic movement model from which movement can be
simulated \citep{Avgar.2016, Signer.2017, Signer.2024}.

\section{Thesis Outline}

This thesis comprises six chapters; a general introduction, four data chapters,
and a general discussion. All published chapters of this thesis are verbally
reproduced as they appeared in the respective journals, including personal
pronouns, jargon, and abbreviations. References cited in each chapter are
listed at the end of the thesis.

% In the general introduction, I introduce the general background and research
% topics. In particular, I discuss the importance of dispersal and connectivity in
% the context of conservation science and highlight multiple limitations in
% current connectivity research. I also introduce the study system and give a
% brief overview of the data collection and general research methods.

% In chapters two to four, I investigate and address several limitations of
% current connectivity research. In chapter five, I outline some additional works
% and supervised thesis. Finally, in the general discussion, I summarize my
% findings and embed them in a broader context. I also provide an outlook for
% topics that require further research, and provide conservation insights for the
% African wild dog.

In \Cref{ChapterSimulation}, I present a novel simulation approach to study
landscape connectivity via simulated dispersal. Specifically, I introduce a
simple three-step approach to quantify landscape connectivity using IBMMs,
starting from empirical dispersal data and a set of landscape layers. The
simulation approach relies on iSSFs, which enable a detailed representation of
dispersers' habitat and movement preferences. In contrast to traditional
connectivity modeling techniques, this approach bypasses the need for a
permeability surface and overcomes several important limitations, such as the
assumption that dispersers move towards a predefined endpoint. To translate
simulated dispersal trajectories into meaningful measures of functional
connectivity, I propose three complementary connectivity metrics, each focused
on a different aspect of connectivity. This includes a heatmap, revealing
frequently traversed areas, a betweenness map, highlighting critical movement
corridors, and a map of inter-patch connectivity, summarizing dispersal
frequencies and durations. I demonstrate the application of the approach using
dispersing AWDs in the KAZA-TFCA as a case-study. I thereby show how landscape
connectivity varies across the KAZA-TFCA landscape and reveal critical movement
corridors. This chapter was published in \textit{Landscape Ecology} in 2023
(\url{https://doi.org/10.1007/s10980-023-01602-4}).

In \Cref{ChapterFlood}, I study functional connectivity for AWDs in light of
increasingly extreme climatic conditions due to climate change. Specifically, I
study how climate-induced changes to the Okavango Delta's flooding regime could
influence AWDs' ability to move between remaining populations. For this, I
compile over 20 years of remote sensed satellite data on flood conditions and
derive two extreme scenarios that serve to approximate flood conditions under
climate change. For both scenarios, I simulate AWD dispersal and investigate
emerging connectivity patterns. I show that periods of increased flooding result
in isolated subpopulations and reduced dispersal prospects. Conversely, I find
that periods of reduced floods reveal vast dispersal habitats that can be used
to transfer from one subpopulation to another. Overall, this work highlights the
importance of accounting for anticipated on-the-ground conditions in light of
climate change when assessing landscape connectivity. This chapter was published
in \textit{Global Change Biology} in 2024
(\url{https://doi.org/10.1111/gcb.17299}).

In \Cref{ChapterSeasonality}, I investigate the importance of accounting for
seasonal dynamism when investigating landscape connectivity. To do so, I first
provide a conceptual framework to motivate that seasonality can enter
connectivity analyses at multiple stages; when (1) extracting spatial covariates
for model fitting, (2) when fitting the selection model, and (3) when making
predictions from the fitted model. Via combination, this provides six distinct
configurations that differ in their amount of dynamism. Using GPS data on
dispersing AWDs, I parametrize the models associated with each configuration and
employ a rigorous validation approach to assess at which stages rendering
seasonality is critical. I also employ IBMMs to simulate dispersal under varying
degrees of seasonal dynamism and examine the emerging patterns of connectivity.
Notably, in the most dynamic configuration, I update seasonal covariates as the
simulated dispersers move; a similar degree of seasonality cannot be achieved
via traditional approaches. Even though the validation procedure reveals only a
marginal benefit of increasing dynamism, I find that emerging patterns of
connectivity differ markedly depending on the level of incorporated dynamism.
Specifically, a static approach results in more clumped dispersal hotspots,
whereas a dynamic take suggests more heterogeneously distributed dispersal
movements. This chapter highlights that the predictive improvements reaped by
incorporating seasonality into connectivity analyses may be moderate, despite
marked differences in inferred patterns of connectivity.

In \Cref{ChapterIrregularity}, I revisit the iSSF framework utilized throughout
Chapters 2 to 4. A major drawback of iSSFs is that they presume that GPS data
were collected at regular time intervals, whereas in reality GPS data are often
incomplete and gappy, thus exhibiting slight irregularities between successive
GPS locations. Usually, such irregular data need to be removed, implying a
substantial reduction in effective sample size. To overcome this limitation, I
introduce and compare several methods that account for irregularity, thereby
allowing to retain additional GPS data for modeling. I use simulations with
known parameters to compare the effectiveness of different methods and show that
irregularity can be effectively accounted for when appropriate methods are used.
I also exemplify the application of the best performing method to a case study
using GPS data on a spotted hyena. The divergence from using AWD dispersal data
was primarily due to the availability of an extensive dataset on a single
spotted hyena. This chapter was published in \textit{Movement Ecology} in 2024
(\url{https://doi.org/10.1186/s40462-024-00476-8}).

% In \Cref{AdditionalWorks}, I give a brief overview of additional works that I
% have conducted during my PhD and summarize two master theses that I have
% co-supervised.

Finally, in \Cref{GeneralDiscussion}, I synthesize and reflect on the main
insights of this thesis and discuss how this work integrates into the existing
literature. I also elaborate on some of the methodological and biological
implications of my work, with a particular focus on conservation insights for
the African wild dog. Finally, I highlight potential limitations and give an
outlook for future research.

\ifSubfilesClassLoaded{%
  \newpage
  \begin{singlespacing}
  \ifthenelse{\boolean{usebiblatex}}{
    \begin{refcontext}[sorting=nyt]
    \printbibliography
    \end{refcontext}
  } {
    \bibliography{../LiteratureBibtex}%
  }
\end{singlespacing}
}{}

\end{document}

### 2024-07-01 FinalThesis/Chapter_01/Manuscript.tex
Publish PHD

--- 2022-11-29 Chapter_1/04_Manuscript/Manuscript.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage[round]{natbib}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{lineno}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-\usepackage[nomarkers, nolists]{endfloat}
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage{scalerel}
-\usepackage{tikz}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% To be able to put an ORCID
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-\title{A Three-Step Approach for Assessing Landscape Connectivity via Simulated
-Dispersal: African Wild Dog Case Study}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Required to be able to cite
-\bibliographystyle{apalike}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
-Maun, Botswana.
-
-\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)
-
-\vspace{4cm}
-
-\textbf{Running Title:} Simulating Wild Dog Dispersal Trajectories to Assess
-Landscape Connectivity
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} dispersal, simulation, movement, integrated step-selection
-function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, Lycaon pictus
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Abstract
-%------------------------------------------------------------------------------
-% \newpage
-% \begin{abstract}
-% %------------------------------------------------------------------------------
-% % Short Version
-% %------------------------------------------------------------------------------
-% \textbf{Short Version }
-% The ability to disperse is contingent on a sufficient degree of landscape
-% connectivity, which is why the identification and preservation of movement
-% corridors that promote connectivity has become a task of extraordinary
-% importance. Currently, ecologists rely on least-cost analysis or circuit theory
-% to investigate connectivity, although both methods make several assumptions that
-% are hardly met in reality. To address these issues, simulations from
-% individual-based movement models have been proposed, yet a unified framework to
-% simulate dispersal and quantify connectivity is lacking.
-%
-% Here, we propose a simple three-step approach that combines several
-% already-existing methods to assess connectivity using simulated dispersal
-% trajectories. In step one, we use integrated step-selection functions to
-% parametrize a mechanistic movement model rendering dispersal behavior. In step
-% two, we apply the parametrized model as an individual-based movement model to
-% simulate dispersal trajectories. In step three, we combine simulated
-% trajectories into three complementary connectivity maps, each focusing on a
-% different aspect of landscape connectivity.
-%
-% We showcase the application of the proposed approach using empirical data of
-% dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
-% connectivity within the world's largest transboundary conservation area in
-% Southern Africa. We thereby shed light into dispersing wild dogs' habitat and
-% movement preferences, while also uncovering crucial dispersal corridors. With
-% this analysis, we demonstrate that simulations from integrated step-selection
-% functions offer a simple, yet powerful alternative to traditional connectivity
-% modeling techniques.
-% \end{abstract}
+\chapter{A Three-Step Approach for Assessing Landscape Connectivity via
+Simulated Dispersal: African Wild Dog Case Study}
+\label{ChapterSimulation}
+\thispagestyle{empty}
+\vspace{-1cm}
+\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil
+
+\begin{center}
+David D. Hofmann \orcid{0000-0003-3477-4365},
+Gabriele Cozzi \orcid{0000-0002-1744-1940},
+John W. McNutt,
+Arpat Ozgul \orcid{0000-0001-7477-2642}, and
+Dominik M. Behr \orcid{0000-0001-7378-8538}
+
+% \begin{flushleft}
+
+% \vspace{5cm}
+\vfill
+
+Published in Landscape Ecology (2023):
+\url{https://doi.org/10.1007/s10980-023-01602-4}.
+
+% \textsuperscript{1} Department of Evolutionary Biology and Environmental
+% Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
+% Switzerland.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+% Private Bag 13, Maun, Botswana.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{\S} Corresponding author:
+% \href{mailto:david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
+
+\end{center}
+
+% \vfill
+
+% \textbf{Running Title:}\\African Wild Dog Dispersal in a Changing Climate:
+% Lessons Learned from Seasonal Flood Extremes
+%
+% \vspace{0.5cm}
+%
+% \textbf{Keywords:}\\movement, connectivity, climate change, seasonality,
+% dispersal, conservation, individual-based simulations, step-selection
+
+% \end{flushleft}
 
 \newpage
-\begin{abstract}
-%------------------------------------------------------------------------------
-%  LONG VERSION
-%------------------------------------------------------------------------------
-\noindent \textbf{Context} \quad Dispersal of individuals contributes to
-long-term population persistence, yet requires a sufficient degree of landscape
-connectivity. To date, connectivity has mainly been investigated using
-least-cost analysis and circuit theory, two methods that make assumptions that
-are hardly applicable to dispersal. While these assumptions can be relaxed by
-explicitly simulating dispersal trajectories across the landscape, a unified
-approach for such simulations is lacking.
-
-\noindent \textbf{Objectives} \quad Here, we propose and apply a simple
-three-step approach to simulate dispersal and to assess connectivity using
-empirical GPS movement data and a set of habitat covariates.
-
-\noindent \textbf{Methods} \quad In step one of the proposed approach, we use
-integrated step-selection functions to fit a mechanistic movement model
-describing habitat and movement preferences of dispersing individuals. In step
-two, we apply the parameterized model to simulate dispersal across the study
-area. In step three, we derive three complementary connectivity maps; a heatmap
-highlighting frequently traversed areas, a betweenness map pinpointing dispersal
-corridors, and a map of inter-patch connectivity indicating the presence and
-intensity of functional links between habitat patches. We demonstrate the
-applicability of the proposed three-step approach in a case study in which we
-use GPS data collected on dispersing African wild dogs (\textit{Lycaon pictus})
-inhabiting northern Botswana.
-
-\noindent \textbf{Results} \quad Using step-selection functions we successfully
-parametrized a detailed dispersal model that described dispersing individuals'
-habitat and movement preferences, as well as potential interactions among the
-two. The model substantially outperformed a model that omitted such interactions
-and enabled us to simulate 80,000 dispersal trajectories across the study area.
-
-\noindent \textbf{Conclusion} \quad By explicitly simulating dispersal
-trajectories, our approach not only requires fewer unrealistic assumptions about
-dispersal, but also permits the calculation of multiple connectivity metrics
-that together provide a comprehensive view of landscape connectivity. In our
-case study, the three derived connectivity maps revealed several wild dog
-dispersal hotspots and corridors across the extent of our study area. Each map
-highlighted a different aspect of landscape connectivity, thus emphasizing their
-complementary nature. Overall, our case study demonstrates that a
-simulation-based approach offers a simple yet powerful alternative to
-traditional connectivity modeling techniques. It is therefore useful for a
-variety of applications in ecological, evolutionary, and conservation research.
-
-\end{abstract}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
+\section*{Abstract}
+Dispersal of individuals contributes to long-term population persistence, yet
+requires a sufficient degree of landscape connectivity. To date, connectivity
+has mainly been investigated using least-cost analysis and circuit theory, two
+methods that make assumptions that are hardly applicable to dispersal. While
+these assumptions can be relaxed by explicitly simulating dispersal trajectories
+across the landscape, a unified approach for such simulations is lacking. Here,
+we propose and apply a simple three-step approach to simulate dispersal and to
+assess connectivity using empirical GPS movement data and a set of habitat
+covariates. In step one of the proposed approach, we use integrated
+step-selection functions to fit a mechanistic movement model describing habitat
+and movement preferences of dispersing individuals. In step two, we apply the
+parameterized model to simulate dispersal across the study area. In step three,
+we derive three complementary connectivity maps; a heatmap highlighting
+frequently traversed areas, a betweenness map pinpointing dispersal corridors,
+and a map of inter-patch connectivity indicating the presence and intensity of
+functional links between habitat patches. We demonstrate the applicability of
+the proposed three-step approach in a case study in which we use GPS data
+collected on dispersing African wild dogs (\textit{Lycaon pictus}) inhabiting
+northern Botswana. Using step-selection functions we successfully parametrized a
+detailed dispersal model that described dispersing individuals' habitat and
+movement preferences, as well as potential interactions among the two. The model
+substantially outperformed a model that omitted such interactions and enabled us
+to simulate 80,000 dispersal trajectories across the study area. By explicitly
+simulating dispersal trajectories, our approach not only requires fewer
+unrealistic assumptions about dispersal, but also permits the calculation of
+multiple connectivity metrics that together provide a comprehensive view of
+landscape connectivity. In our case study, the three derived connectivity maps
+revealed several wild dog dispersal hotspots and corridors across the extent of
+our study area. Each map highlighted a different aspect of landscape
+connectivity, thus emphasizing their complementary nature. Overall, our case
+study demonstrates that a simulation-based approach offers a simple yet powerful
+alternative to traditional connectivity modeling techniques. It is therefore
+useful for a variety of applications in ecological, evolutionary, and
+conservation research.
+
 \newpage
-
-% \onehalfspacing
-% \tableofcontents
-\doublespacing
-
-% Change page numbering
-\newpage
-\pagenumbering{arabic}
-
-% Create linenumbers
-\linenumbers
-
 \section{Introduction}
@@
 % Importance of Dispersal & Connectivity
-% \subsection{Importance of Connectivity \& Connectivity Models}
 Dispersal of individuals is a vital process that allows species to maintain
@@
 LaPoint.2013}, rescue non-viable populations \citep{Brown.1977}, and to colonize
-unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
+unoccupied habitats \citep{Hanski.1999a, MacArthur.2001}. However, the ability
 to disperse depends on a sufficient degree of landscape connectivity
@@
 % Issues with Both Methods
-% \subsection{Issues with Traditional Connectivity Models}
 Traditional connectivity models make assumptions that are rarely met for
@@
 directional. An interesting generalization that bridges the continuum between
-LCPA and CT has been proposed by \citep{Panzacchi.2016} and enables to
+LCPA and CT has been proposed by \citet{Panzacchi.2016} and enables to
 capitalize on the merits of both approaches. Despite these and several other
@@
 % What about IBMMs?
-% \subsection{What about IBMMs?}
 The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
@@
 % Proposed Solution: Three-Step approach
-% \subsection{Proposed Solution: Three-Step Approach}
 Here, we propose and exemplify a simple three-step approach for simulating
@@
 covariates to fit a mechanistic movement model via integrated step-selection
-functions (ISSFs, \citealp{Avgar.2016}). We chose to use ISSFs because the
+functions (iSSFs, \citealp{Avgar.2016}). We chose to use iSSFs because the
 framework not only allows inference on the study species' habitat kernel (i.e.
@@
 Zeller.2020}), (ii) a betweenness-map delineating dispersal corridors and
-bottlenecks (e.g. \citealp{BastilleRousseau.2018}), (iii) and a map of
+bottlenecks (e.g. \citealp{Bastille-Rousseau.2018}), (iii) and a map of
 inter-patch connectivity, depicting the presence and intensity of functional
@@
 % Introduction of the study species
-% \subsection{Case Study}
 We showcase the application of the proposed approach using GPS movement data
@@
 single-sex coalitions \citep{McNutt.1996, Behr.2020} and are capable of
-dispersing several hundred kilometers \citep{DaviesMostert.2012, Masenga.2016,
+dispersing several hundred kilometers \citep{Davies-Mostert.2012, Masenga.2016,
 Cozzi.2020}. Although previous studies have investigated connectivity for this
@@
   \begin{center}
-    \includegraphics[width = \textwidth]{99_GraphicalAbstract.pdf}
+    \includegraphics[width = \textwidth]{Figures/GraphicalAbstract}
     \caption{Flowchart of the simulation-based connectivity analysis. First, GPS
@@
 
-\section{Methods}
+\section{Materials and Methods}
 \subsection{Case Study}
 \subsubsection{GPS Data}
+
 We applied the three step approach presented in \Cref{GraphicalAbstract} to GPS
@@
 research found little differences between sexes during dispersal
-\citep{Woodroffe.2019, Cozzi.2020}. Additional details on the data collection
-and preparation can be found in \cite{Cozzi.2020} and \cite{Hofmann.2021}.
+\citep{Woodroffe.2020, Cozzi.2020}. Additional details on the data collection
+and preparation can be found in \citet{Cozzi.2020} and \citet{Hofmann.2021}.
 
@@
 \subsubsection{Study Area}
+
 Our simulation of dispersal trajectories and assessment of connectivity spanned
 across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA,
-\Cref{StudyArea}a and b) and encompassed a rectangular extent of roughly 1.3
-Mio. km\textsuperscript{2}. With an area of 520'000 km\textsuperscript{2}, the
+\Cref{StudyAreaCH1}a and b) and encompassed a rectangular extent of roughly 1.3
+Mio. km\textsuperscript{2}. With an area of 520,000 km\textsuperscript{2}, the
 KAZA-TFCA is the world's largest transboundary conservation area and comprises
@@
 
-\begin{figure}[h]
+\begin{figure}[htbp]
   \begin{center}
@@
         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
-        \includegraphics[width=\textwidth]{99_StudyArea.png}
+        \includegraphics[width=\textwidth]{Figures/StudyArea.png}
         };
@@
     remaining wild dog populations (orange) has been sourced from
-    \cite{Woodroffe.2012}. (b) The KAZA-TFCA represents the world's largest
+    \citet{Woodroffe.2012}. (b) The KAZA-TFCA represents the world's largest
     terrestrial transfrontier conservation area and covers a total area of
-    520'000 km\textsuperscript{2}. Its main purpose is to re-establish
+    520,000 km\textsuperscript{2}. Its main purpose is to re-establish
     connectivity between already-existing NPs (dark green) and other protected
     areas (light green).}
-    \label{StudyArea}
+    \label{StudyAreaCH1}
   \end{center}
@@
 \subsubsection{Habitat Covariates}
+
 We represented the physical landscape in our study area by the habitat
@@
 \textsf{distance-to-water} from MODIS satellite imagery \citep{Wolski.2017,
-Hofmann.2021}. This algorithm is now implemented in the {\tt floodmapr} package
-(available on GitHub; \url{https://github.com/DavidDHofmann/floodmapr}). To
-ensure a consistent resolution across habitat covariates, we coarsened or
+Hofmann.2021}. This algorithm is now implemented in the \texttt{floodmapr}
+package (available on GitHub; \url{https://github.com/DavidDHofmann/floodmapr}).
+To ensure a consistent resolution across habitat covariates, we coarsened or
 interpolated all layers to a resolution of 250 m x 250 m. A detailed description
-of how we prepared each habitat covariate is provided in \cite{Hofmann.2021}.
+of how we prepared each habitat covariate is provided by \citet{Hofmann.2021}.
 
 We performed all data preparations, spatial computations, and statistical
-analysis in {\tt R}, version 4.2.2 \citep{R.2022}. Some helper functions were
-written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
-\citep{Eddelbuettel.2011, Eddelbuettel.2013, Eddelbuettel.2018}.
+analysis in \texttt{R}, version 4.2.2 \citep{RCoreTeam.2023}. Some helper
+functions were written in \texttt{C++} and imported into \texttt{R} using the
+\texttt{Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013,
+Eddelbuettel.2018}.
 
 \subsection{Step 1 - Movement Model}
-We combined the collected GPS data with habitat covariates and used ISSFs
+
+We combined the collected GPS data with habitat covariates and used iSSFs
 \citep{Avgar.2016} to parametrize a mechanistic movement model. More
@@
 alternative steps. A realized and its 24 random steps together formed a stratum
-that received a unique identifier. As suggested by \cite{Avgar.2016}, we
+that received a unique identifier. As suggested by \citet{Avgar.2016}, we
 generated random steps by sampling random turning angles from a uniform
 distribution (\(-\pi, +\pi\)) and step lengths from a gamma distribution that
-was fitted to realized steps (scale \(\theta\) = 6'308 and shape \(k\) = 0.37).
+was fitted to realized steps (scale \(\theta\) = 6,308 and shape \(k\) = 0.37).
 Note that our approach of sampling turning angles from a uniform distribution
@@
 Besides extracting \textit{habitat covariates}, we also computed movement
-metrics that we used as \textit{movement covariates} in the ISSF models
+metrics that we used as \textit{movement covariates} in the iSSF models
 \citep{Avgar.2016, Fieberg.2021}. Specifically, we computed the step length
@@
 
-\noindent To estimate relative selection strenghts (i.e. the
+\noindent To estimate relative selection strengths (i.e. the
 \(\beta\)-coefficients), we used mixed effects conditional logistic regression
-analysis, implemented through the r-package {\tt glmmTMB} \citep{Brooks.2017}.
-The implementation of conditional logistic regression has been proposed by
-\cite{Muff.2020} and allows to model random slopes. The method requires to fix
-the variance of the stratum specific intercept to a large value, so we fixed it
-to an arbitrary high value of \(10 ^ 6\) and used disperser identity to model
-random slopes for all covariates.
+analysis, implemented through the \texttt{R}-package \texttt{glmmTMB}
+\citep{Brooks.2017}. The implementation of conditional logistic regression has
+been proposed by \citet{Muff.2020} and allows to model random slopes. The method
+requires to fix the variance of the stratum specific intercept to a large value,
+so we fixed it to an arbitrary high value of \(10 ^ 6\) and used disperser
+identity to model random slopes for all covariates.
 
@@
 preferences during dispersal could depend on habitat conditions (details in
-Appendix A1). To determine the most parsimonious movement model among model
-candidates, we ran stepwise forward model selection based on Akaike's
-Information Criterion (AIC, \citealp{Burnham.2002}). More specifically, we
-started with the base model and iteratively increased model complexity by adding
-all possible interactions between movement and habitat covariates. Given that
-the focus of our analysis lied on predicting dispersal patterns and all model
-candidates were biologically intuitive, we deemed the use of model selection
-appropriate. However, caution should be employed if causal relationships are of
-interest, as model selection may lead to biased parameter estimate
-\citep{Whittingham.2006}. We validated the predictive power of the most
+\Cref{CandidateInteractions}). To determine the most parsimonious movement model
+among model candidates, we ran stepwise forward model selection based on
+Akaike's Information Criterion (AIC, \citealp{Burnham.2002}). More specifically,
+we started with the base model and iteratively increased model complexity by
+adding all possible interactions between movement and habitat covariates. Given
+that the focus of our analysis lied on predicting dispersal patterns and all
+model candidates were biologically intuitive, we deemed the use of model
+selection appropriate. However, caution should be employed if causal
+relationships are of interest, as model selection may lead to biased parameter
+estimate \citep{Whittingham.2006}. We validated the predictive power of the most
 parsimonious model using k-fold cross-validation for case-control studies as
-described in \cite{Fortin.2009}. This validation attests significant prediction
+described in \citet{Fortin.2009}. This validation attests significant prediction
 ability to the movement model if the model outperforms a random guess and
 systematically assigns low ranks (high selection scores) to observed steps
-(details in Appendix A2).
+(details in \Cref{CrossValidation}).
 
 \subsection{Step 2 - Dispersal Simulation}
+
 We used the most parsimonious movement model to simulate individual dispersal
 trajectories across the study area. The simulation of a dispersal trajectory
-resembled an ``inverted'' ISSF and was set up as follows. (1) We defined a
+resembled an ``inverted'' iSSF and was set up as follows. (1) We defined a
 source point and assumed a random initial orientation of the simulated animal.
@@
 position. We repeated steps (2) to (5) until 2,000 steps were realized and we
-repeated the simulation until a total of 80'000 dispersal trajectories was
+repeated the simulation until a total of 80,000 dispersal trajectories was
 reached.
@@
 \citealp{Pomilia.2015}). We placed another 30,000 points randomly inside the
-buffer zone, mimicking potential immigration into the study area (Figure S1).
+buffer zone, mimicking potential immigration into the study area
+(\Cref{SourcePoints}).
 
 To mitigate edge effects and to deal with random steps leaving the study area,
-we followed \cite{Koen.2010} and artificially expanded all covariate layers by a
+we followed \citet{Koen.2010} and artificially expanded all covariate layers by a
 100 km wide buffer zone. Within the buffer zone, we randomized covariate values
@@
 trajectories from 1 to 50,000. To assess variability in the relative traversal
-frequency, we repeatedly subsampled 100 times from all 50'000 trajectories and
+frequency, we repeatedly subsampled 100 times from all 50,000 trajectories and
 computed the mean traversal frequency across replicates, as well as its 95\%
@@
 \subsubsection{Heatmap}
+
 To identify dispersal hotspots within the study area, we created a heatmap
@@
 circles''. To achieve high performance rasterization, we used the R-package {\tt
-terra} \citep{Hijmans.2021b}. For a subset of the study area, we also generated
+terra} \citep{Hijmans.2024}. For a subset of the study area, we also generated
 heatmaps at 250 m x 250 m, yet found little qualitative differences to the
@@
 \subsubsection{Betweenness Map}
+
 To pinpoint movement corridors and bottlenecks, we converted simulated
 trajectories into a network and calculated betweenness scores for all
-raster-cells in the study area \citep{BastilleRousseau.2018}. Betweenness is a
+raster-cells in the study area \citep{Bastille-Rousseau.2018}. Betweenness is a
 pertinent metric for connectivity as it measures how often a specific
 network-node (in our case a raster-cell) lies on a shortest path between any
-other pair of nodes \citep{BastilleRousseau.2018}. To convert simulated
-trajectories into a network, we followed \cite{BastilleRousseau.2018} and
+other pair of nodes \citep{Bastille-Rousseau.2018}. To convert simulated
+trajectories into a network, we followed \citet{Bastille-Rousseau.2018} and
 overlaid the study area (including the buffer) with a raster containing 2.5 km x
@@
 cell to another, as well as the frequency at which those transitions occurred
-(see also Appendix A4). This resulted in an edge-list that we translated into a
-weighted network using the r-package {\tt igraph} \citep{Gabor.2006}. The final
-weight of each edge was determined by the frequency of transitions, yet because
-{\tt igraph} handles edge weights (\(\omega\)) as costs, we inverted the
-traversal-frequency through each raster-cell by applying \(\omega =
-\frac{mean(Traversal Frequency)}{Traversal Frequency_i}\). Consequently,
-regularly used edges received small weights (i.e. low costs) and vice versa. We
-used the weighted network to calculate betweenness scores for all network nodes.
+(see also \Cref{BetweennessApproaches}). This resulted in an edge-list that we
+translated into a weighted network using the r-package \texttt{igraph}
+\citep{Csardi.2006}. The final weight of each edge was determined by the
+frequency of transitions, yet because \texttt{igraph} handles edge weights
+(\(\omega\)) as costs, we inverted the traversal-frequency through each
+raster-cell by applying \(\omega = \frac{mean(Traversal Frequency)}{Traversal
+Frequency_i}\). Consequently, regularly used edges received small weights (i.e.
+low costs) and vice versa. We used the weighted network to calculate betweenness
+scores for all network nodes.
 
 \subsubsection{Inter-Patch Connectivity Map}
+
 To examine the presence and intensity of functional links (i.e. connections)
@@
 \subsubsection{Validation}
+
 To validate our predictions of connectivity, we utilized additional dispersal
 data that was collected on eight dispersing coalitions between 2019 and 2022
-(totalling to 2'668 GPS locations). We used path selection analysis (PSF,
+(totalling to 2,668 GPS locations). We used a path selection function (PSF,
 \citealp{Cushman.2010}) to assess if observed dispersal trajectories followed
@@
 each path, we then extracted connectivity values from the heatmap (see above)
-generated after 68, 125, 250, 500, and 2'000 simulated steps, respectively.
+generated after 68, 125, 250, 500, and 2,000 simulated steps, respectively.
 Finally, we ran conditional logistic regression to contrast observed and random
@@
 \section{Results}
-% \subsection{Movement Model}
+
 The most parsimonious movement model consisted of movement covariates, habitat
@@
 movement behavior during dispersal depended on habitat conditions
-(\Cref{MovementModel}a, Table S1 and Table S2). Although multiple models
-received an AIC weight \(>\) 0 (Table S1), we only considered results from the
-most parsimonious model for simplicity. This decision only marginally influenced
-subsequent steps as all models with positive AIC weights retained similar
-covariates (Table S1). The k-fold cross-validation showed that the final model
-substantially outperformed a random guess and provided reliable predictions
-(i.e. confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-random}\) did not overlap). Moreover, the model correctly assigned high
-selection scores to realized steps (\Cref{MovementModel}b), indicating a good
-fit between predictions and observations. As can be taken from the Spearman rank
-correlation coefficient, the inclusion of several interactions between movement
-and habitat covariates significantly improved model performance (\(\bar{r}_{s,
-realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)), compared to the base model
-(\(\bar{r}_{s, realized} = -0.55; 95\%-CI = [-0.57, -0.52]\);
-\citealp{Hofmann.2021}). Our validation of the resulting connectivity maps using
-independent dispersal data showed that dispersers preferentially followed areas
-of high predicted connectivity, as coefficients from the PSF models were all
-significantly greater than zero (\Cref{MovementModel}c). The movement model thus
-successfully predicted functional connectivity.
+(\Cref{MovementModel}a, \Cref{ModelAICs} and \cref{MovementModelNumbers}).
+Although multiple models received an AIC weight \(>\) 0 (\Cref{ModelAICs}), we
+only considered results from the most parsimonious model for simplicity. This
+decision only marginally influenced subsequent steps as all models with positive
+AIC weights retained similar covariates (\Cref{ModelAICs}). The k-fold
+cross-validation showed that the final model substantially outperformed a random
+guess and provided reliable predictions (i.e. confidence intervals of
+\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap).
+Moreover, the model correctly assigned high selection scores to realized steps
+(\Cref{MovementModel}b), indicating a good fit between predictions and
+observations. As can be taken from the Spearman rank correlation coefficient,
+the inclusion of several interactions between movement and habitat covariates
+significantly improved model performance (\(\bar{r}_{s, realized} = -0.65;
+95\%-CI = [-0.67, -0.64]\)), compared to the base model (\(\bar{r}_{s, realized}
+= -0.55; 95\%-CI = [-0.57, -0.52]\); \citealp{Hofmann.2021}). Our validation of
+the resulting connectivity maps using independent dispersal data showed that
+dispersers preferentially followed areas of high predicted connectivity, as
+coefficients from the PSF models were all significantly greater than zero
+(\Cref{MovementModel}c). The movement model thus successfully predicted
+functional connectivity.
 
 Plots that aid with the interpretation of the most parsimonious movement model
-are provided in Figure S3 and suggest that, under average conditions, dispersing
-wild dogs avoided moving through water, woodlands, and areas dominated by
-humans, but preferred moving across shrublands or grasslands
+are provided in \Cref{Interpretation} and suggest that, under average
+conditions, dispersing wild dogs avoided moving through water, woodlands, and
+areas dominated by humans, but preferred moving across shrublands or grasslands
 (\Cref{MovementModel}a). Dispersers realized shorter steps (indicating slower
@@
 
-\begin{figure}
+\begin{figure}[htpb]
   \begin{center}
-    \includegraphics[width=\textwidth]{99_MovementModel}
-    \caption{(a) Most parsimonious movement model for dispersing wild dogs. The
-    model comprises a habitat kernel, a movement kernel, as well as their
-    interactions. The horizontal line segments delineate the 90\%, 95\%, and
-    99\% confidence-intervals for the respective \(\beta\)-coefficients.
-    Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
-    Results from the k-fold cross validation procedure. Subfigure b1 shows rank
-    frequencies of realized steps according to model predictions with known
-    preferences, whereas subfigure b2 shows rank frequencies of realized steps
-    when assuming random preferences. The blue ribbon shows the prediction
-    interval around a loess smoothing regression that we fitted to ease the
-    interpretation of the plots. The significant correlation between rank and
-    associated frequency in (b1) highlights that the most parsimonious model
-    successfully outperformed a random guess (b2) and frequently assigned low
-    ranks (i.e. high selection scores) to realized steps but only rarely high
-    ranks (i.e. low selection scores). (c) Results from the PSF analysis using
-    independent dispersal data show that dispersers preferrably moved through
-    areas where our heatmaps predicted high connectivity. Results are shown for
-    heatmaps realized after 68, 125, 250, 500, and 2'000 simulated steps,
-    respectively.}
+    \includegraphics[width=\textwidth]{Figures/MovementModel} \caption{(a) Most
+    parsimonious movement model for dispersing wild dogs. The model comprises a
+    habitat kernel, a movement kernel, as well as their interactions. The
+    horizontal line segments delineate the 90\%, 95\%, and 99\%
+    confidence-intervals for the respective \(\beta\)-coefficients. Significance
+    codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b) Results from
+    the k-fold cross-validation procedure. Subfigure b1 shows rank frequencies
+    of realized steps according to model predictions with known preferences,
+    whereas subfigure b2 shows rank frequencies of realized steps when assuming
+    random preferences. The blue ribbon shows the prediction interval around a
+    loess smoothing regression that we fitted to ease the interpretation of the
+    plots. The significant correlation between rank and associated frequency in
+    b1 highlights that the most parsimonious model successfully outperformed a
+    random guess b2 and frequently assigned low ranks (i.e. high selection
+    scores) to realized steps but only rarely high ranks (i.e. low selection
+    scores). (c) Results from the PSF analysis using independent dispersal data
+    show that dispersers preferably moved through areas where our heatmaps
+    predicted high connectivity. Results are shown for heatmaps realized after
+    68, 125, 250, 500, and 2,000 simulated steps, respectively.}
     \label{MovementModel}
@@
 \subsection{Dispersal Simulation}
+
 Dispersal simulations based on the most parsimonious movement model proved
@@
 all checkpoints showed that the relative traversal frequency reached a steady
-state after 10,500 simulated dispersal trajectories (Figure S4). Although
-variability in the relative traversal frequency kept decreasing as we increased
-the number of simulated dispersers, the marginal benefit of simulating
-additional trajectories diminished quickly (Figure S4).
+state after 10,500 simulated dispersal trajectories (\Cref{Convergence}).
+Although variability in the relative traversal frequency kept decreasing as we
+increased the number of simulated dispersers, the marginal benefit of simulating
+additional trajectories diminished quickly (\Cref{Convergence}).
 
 \subsection{Heatmap}
+
 The heatmap (\Cref{Heatmap}), which resulted from the summation of all simulated
@@
 traversed by dispersers (median traversal frequency inside KAZA-TFCA = 166, IQR
-= 274, Figure S7a), whereas areas beyond the KAZA-TFCA boundary were
+= 274, \Cref{MapComparison}a), whereas areas beyond the KAZA-TFCA boundary were
 comparatively rarely visited (median traversal frequency outside KAZA-TFCA = 61,
-IQR = 133, Figure S7a). Most notably, the region in northern Botswana south of
-the Linyanti swamp appeared to serve as highly frequented dispersal hotspot
-(median traversal frequency = 987, IQR = 558). Aside from revealing movement
-hotspots, the heatmap also provided information on areas that appeared to hinder
-movement. For example, extensive water bodies, such as the Okavango Delta, the
-Makgadikgadi Pan, and the Linyanti swamp, substantially restricted dispersal
-movements and limited realized connectivity inside the KAZA-TFCA. Similarly, the
-landscapes of Zambia and Zimbabwe were only rarely used for dispersal, even
-within the KAZA-TFCA boundaries (Figure S8a). Despite the fact that the heatmap
-improved our understanding of the frequency at which areas were traversed by
-simulated dispersers, it seemed impractical to pinpoint dispersal corridors.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_Heatmap.png}
-  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
-  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
+IQR = 133, \Cref{MapComparison}a). Most notably, the region in northern Botswana
+south of the Linyanti swamp appeared to serve as highly frequented dispersal
+hotspot (median traversal frequency = 987, IQR = 558). Aside from revealing
+movement hotspots, the heatmap also provided information on areas that appeared
+to hinder movement. For example, extensive water bodies, such as the Okavango
+Delta, the Makgadikgadi Pan, and the Linyanti swamp, substantially restricted
+dispersal movements and limited realized connectivity inside the KAZA-TFCA.
+Similarly, the landscapes of Zambia and Zimbabwe were only rarely used for
+dispersal, even within the KAZA-TFCA boundaries
+(\Cref{MapComparisonCountries}a). Despite the fact that the heatmap improved our
+understanding of the frequency at which areas were traversed by simulated
+dispersers, it seemed impractical to pinpoint dispersal corridors.
+
+\begin{figure}[htpb]
+  \includegraphics[width=\textwidth]{Figures/Heatmap.png}
+  \caption{Heatmap showing traversal frequencies of 80,000 simulated dispersers
+  moving 2,000 steps across the KAZA-TFCA. Simulations were based on an
   integrated step-selection model that we fitted to the movement data of
@@
   NPs (dark gray). Additional heatmaps showing the traversal frequency when
-  individuals move fewer than 2'000 steps are provided in Figure S5.}
+  individuals move fewer than 2,000 steps are provided in
+  \Cref{HeatmapsIndividual}.}
   \label{Heatmap}
@@
 \subsection{Betweenness}
+
 The betweenness map (\Cref{Betweenness}) revealed several distinct dispersal
@@
 KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR = 54.311 \(\times\)
-10\textsuperscript{6}, Figure S7b). Consequently, only few corridors directly
-linked the peripheral regions of the KAZA-TFCA and passed through unprotected
-areas outside its borders (mean betweenness outside KAZA-TFCA = 2.685 \(\times\)
-10\textsuperscript{6}, IQR = 9.891 \(\times\) 10\textsuperscript{6}, Figure
-S7b).
-
-%  9 2000  Betweenness Inside KAZA-TFCA  6946817.   54311358.
-% 10 2000  Betweenness Outside KAZA-TFCA 2685174.    9890504.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_Betweenness.png}
+10\textsuperscript{6}, \Cref{MapComparison}b). Consequently, only few corridors
+directly linked the peripheral regions of the KAZA-TFCA and passed through
+unprotected areas outside its borders (mean betweenness outside KAZA-TFCA =
+2.685 \(\times\) 10\textsuperscript{6}, IQR = 9.891 \(\times\)
+10\textsuperscript{6}, \Cref{MapComparison}b).
+
+\begin{figure}[htpb]
+  \includegraphics[width=\textwidth]{Figures/Betweenness.png}
   \caption{Map of betweenness scores, highlighting distinct dispersal corridors
@@
   area. The metric is therefore useful to pinpoint discrete movement corridors
-  \citep{BastilleRousseau.2018}. Note that we square-rooted betweenness scores
+  \citep{Bastille-Rousseau.2018}. Note that we square-rooted betweenness scores
   to improve visibility of corridors with comparably low scores. Additional
   betweenness maps showing betweenness scores when individuals move fewer than
-  2'000 steps are provided in Figure S6.}
+  2'000 steps are provided in \Cref{BetweennessIndividual}.}
   \label{Betweenness}
@@
 \subsection{Inter-Patch Connectivity}
+
 The inter-patch connectivity map showed that the relative frequency at which
@@
 originating from a patch in these countries successfully moved into at least on
-other patch (Figure S9a). Conversely, only 19\% of the dispersers leaving from a
-patch in Zambia managed to find their way into some other patch (Figure S9b).
-Prior to reaching another patch, individuals from Angola, Namibia, Botswana,
-Zimbabwe, and Zambia had to move for an average of 630, 640, 940, 1045, and 890
-steps, respectively. Furthermore, it appeared that the corridor previously
-identified on \Cref{InterpatchConnectivity} between Angola's NPs and the Kafue
-NP in Zambia is only rarely realized.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{99_InterpatchConnectivity.png}
+other patch (\Cref{AreasReached}a). Conversely, only 19\% of the dispersers
+leaving from a patch in Zambia managed to find their way into some other patch
+(\Cref{AreasReached}b). Prior to reaching another patch, individuals from
+Angola, Namibia, Botswana, Zimbabwe, and Zambia had to move for an average of
+630, 640, 940, 1,045, and 890 steps, respectively. Furthermore, it appeared that
+the corridor previously identified on \Cref{InterpatchConnectivity} between
+Angola's NPs and the Kafue NP in Zambia is only rarely realized.
+
+\begin{figure}[htpb]
+  \includegraphics[width=\textwidth]{Figures/InterpatchConnectivity.png}
   \caption{Map of inter-patch connectivity in relation to dispersal duration,
@@
 empirical data from a free-ranging population of African wild dogs. In step one,
-we used ISSFs to parametrize a fully mechanistic movement model describing how
+we used iSSFs to parametrize a fully mechanistic movement model describing how
 individuals move through the landscape. Aside from rendering habitat
@@
 previous studies that investigated habitat selection by dispersing wild dogs
-\citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020,
+\citep{Davies-Mostert.2012, Masenga.2016, Woodroffe.2020, ONeill.2020,
 Hofmann.2021}. However, instead of merely generating insights on dispersers'
-habitat preferences, the ISSF framework also permitted us to model several
+habitat preferences, the iSSF framework also permitted us to model several
 additional complexities common to dispersal. For instance, by including the
@@
 lengths and turning angles from copula probability distributions
-\citep{Hodel.2022}, the ISSF framework allowed us to conveniently model such
+\citep{Hodel.2022}, the iSSF framework allowed us to conveniently model such
 peculiarities directly in the movement model. While we only considered first
@@
 GPS data that are not interrupted by missing fixes; something that is rarely
-achieved in reality \citep{Graves.2006}. The power and flexibility of ISSFs to
+achieved in reality \citep{Graves.2006}. The power and flexibility of iSSFs to
 model additive effects between habitat and movement covariates
@@
 that an exceptionally large number of dispersers moved through the regions of
-the Moremi NP and the Chobe NP in northern Botswana. \cite{Hofmann.2021}
+the Moremi NP and the Chobe NP in northern Botswana. \citet{Hofmann.2021}
 previously identified the same area as potential dispersal hotspot using LCPA,
@@
 something that is currently unfeasible with LCPA or CT. An explicit
-representation of time enables answerings questions such as: ``\textit{How long
+representation of time enables answering questions such as: ``\textit{How long
 will it take a disperser to move from A to B?}'' or ``\textit{Is it possible for
@@
 the same connectivity model is repeatedly applied (e.g. \citealp{Osipova.2019}).
-With simulations from ISSFs, in contrast, the environment could change ``as the
+With simulations from iSSFs, in contrast, the environment could change ``as the
 dispersers move'', so that simulated trajectories would dynamically respond to
@@
 % functional connectivity, which is why it often serves as surrogate for
-% functional connectivity \cite{Doerr.2011, Fattebert.2015}. Through incorporation
+% functional connectivity \citet{Doerr.2011, Fattebert.2015}. Through incorporation
 % of a species habitat preferences into a permeability surface, both LCPA and CT
@@
 aspects of connectivity through the permeability surface, which reflects a
-species habitat preferences and thus renders behavioral impacts of the lanscape
+species habitat preferences and thus renders behavioral impacts of the landscape
 on the focal species. Aside from rendering habitat preferences, our model also
@@
 further insights on functional connectivity. In addition, we successfully used
-independent dispersal data to prove that our predictions of connectiviy aligned
+independent dispersal data to prove that our predictions of connectivity aligned
 with observed functional connectivity patterns.
@@
 
-% \subsection{Disadvantages of ISSF Simulations}
+% \subsection{Disadvantages of iSSF Simulations}
 Despite the many benefits and great flexibility offered by simulations from
-ISSFs, one must also be aware of the associated limitations. For example, while
-our approach of simulating dispersal proved usedful to assess landscape
+iSSFs, one must also be aware of the associated limitations. For example, while
+our approach of simulating dispersal proved useful to assess landscape
 connectivity, it was computationally costly. Simulating 80,000 dispersal
-trajectories for 2'000 steps across the KAZA-TFCA required five days of
+trajectories for 2,000 steps across the KAZA-TFCA required five days of
 computation on a regular desktop machine (AMD Ryzen 7 2700X octa-core processor
@@
 However, each additional simulation imposes computational costs, so a trade-off
-needs to be managed. \cite{Signer.2017} proposed to handle the trade-off by
+needs to be managed. \citet{Signer.2017} proposed to handle the trade-off by
 simulating additional individuals only until the metrics of interest converge
 towards a steady state. Here, we used the relative traversal frequency as target
-metric and found that it converged already after 10'500 simulated individuals.
+metric and found that it converged already after 10,500 simulated individuals.
 The exact number of required individuals might, however, vary depending on the
@@
 assuming that wild dogs primarily survive in large, formally protected areas
-\citep{DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. Moreover, we lacked
+\citep{Davies-Mostert.2012, Woodroffe.2012, VanDerMeer.2014}. Moreover, we lacked
 precise knowledge about the distribution and abundance of wild dogs across
@@
 
-(3) The use of ISSFs to simulate dispersers requires deciding on the number of
+(3) The use of iSSFs to simulate dispersers requires deciding on the number of
 simulated steps (i.e. the simulated dispersal durations). If sufficient
@@
 steps, which was at the upper end of observed dispersal durations in African
-wild dogs \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
+wild dogs \citep{Davies-Mostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}.
 This approach had the advantage that it allowed us to systematically shorten the
 simulated trajectories after their simulation and thereby to investigate the
-sensitivity of our results with respect to exact dispersal durations (Figures S5
-and S6).
+sensitivity of our results with respect to exact dispersal durations
+(\Cref{HeatmapsIndividual} and \Cref{BetweennessIndividual}).
 
@@
 In summary, we proposed and applied a simple three-step approach that relies on
-ISSF-analysis and enables the simulation of dispersal trajectories and the
+iSSF-analysis and enables the simulation of dispersal trajectories and the
 assessment of landscape connectivity. The proposed approach overcomes several of
@@
 
-\section{Authors' Contributions}
-D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
-D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
-data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
-first draft of the manuscript and all authors contributed to the drafts at
-several stages and gave final approval for publication.
-
-\section{Data Availability}
-GPS movement data of dispersing wild dogs is available on dryad
-\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
-proposed approach using simulated data are provided through Github
-(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
-codes required to reproduce the African wild dog case study will be made
-available through an online repository at the time of publication.
-
-\section{Acknowledgements}
-We thank the Ministry of Environment and Tourism of Botswana for granting
-permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
-Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
-for sharing her data of three dispersing wild dogs. Furthermore, we would like
-to thank Johannes Signer for assisting with the simulation algorithm. This study
-was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
-Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
-Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust,
-Forschungskredit of the University of Zürich (grant no. FK-21-091), and a Swiss
-National Science Foundation grant (no. 31003A\_182286) to A. Ozgul.
-
-\section{Conflict of Interest}
-All authors declare that they have no conflicts of interest.
-
+% \section{Authors' Contributions}
+%
+% D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
+% D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
+% data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
+% first draft of the manuscript and all authors contributed to the drafts at
+% several stages and gave final approval for publication.
+%
+% \section{Data Availability}
+%
+% GPS movement data of dispersing wild dogs is available on dryad
+% \citep{Hofmann.2021a}. Access to R-scripts that exemplify the application of the
+% proposed approach using simulated data are provided through Github
+% (\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
+% codes required to reproduce the African wild dog case study will be made
+% available through an online repository at the time of publication.
+%
+% \section{Acknowledgements}
+%
+% We thank the Ministry of Environment and Tourism of Botswana for granting
+% permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
+% Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
+% for sharing her data of three dispersing wild dogs. Furthermore, we would like
+% to thank Johannes Signer for assisting with the simulation algorithm. This study
+% was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
+% Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
+% Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust,
+% Forschungskredit of the University of Zürich (grant no. FK-21-091), and a Swiss
+% National Science Foundation grant (no. 31003A\_182286) to A. Ozgul.
+%
+% \section{Conflict of Interest}
+%
+% All authors declare that they have no conflicts of interest.
+
+\ifSubfilesClassLoaded{%
   \newpage
-\begingroup
-\singlespacing
-\bibliography{Literatur}
-\endgroup
+  \begin{singlespacing}
+  \ifthenelse{\boolean{usebiblatex}}{
+    \begin{refcontext}[sorting=nyt]
+    \printbibliography
+    \end{refcontext}
+  } {
+    \bibliography{../LiteratureBibtex}%
+  }
+\end{singlespacing}
+}{}
 

### 2024-07-01 FinalThesis/Chapter_01/Manuscript_Support.tex
Publish PHD

--- 2022-11-18 Chapter_1/04_Manuscript/Manuscript_Support.tex
@@
-\documentclass[abstract=off,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage[round]{natbib}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{lineno}
-\usepackage{multirow}
-\usepackage{arydshln}
-% \usepackage[nomarkers, nolists]{endfloat}
-\usepackage[flushleft]{threeparttable}
-\usepackage{pdflscape}
-\usepackage{scalerel}
-\usepackage{tikz}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% To be able to put an ORCID
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ A Three-Step Approach for Assessing Landscape
-Connectivity via Simulated Dispersal: African Wild Dog Case Study}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Required to be able to cite
-\bibliographystyle{apalike}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
- \vspace{0.5cm}
-
- \textsuperscript{1} Department of Evolutionary Biology and Environmental
- Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
- Switzerland.
-
- \textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
- Maun, Botswana.
-
- \textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)
-
- \vspace{4cm}
-
-\textbf{Running Title:} Simulating Wild Dog Dispersal Trajectories to Assess
-Landscape Connectivity
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} dispersal, simulation, movement, integrated step selection
-function, Kavango-Zambezi Transfrontier Conservation Area, landscape
-connectivity, Lycaon pictus
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
-\newpage
-
-% Change page numbering
-\pagenumbering{arabic}
-
-% Create linenumbers
-\linenumbers
-
-% Change to appendix counters
-\appendix
-\beginappendix
 
@@
 \section{Candidate Interactions}
-We started with the base model developed by \cite{Hofmann.2021} and
+\label{CandidateInteractions}
+
+We started with the base model developed by \citet{Hofmann.2021} and
 incrementally increased model complexity by adding all possible two-way
@@
 \section{K-Fold Cross Validation Procedure}
+\label{CrossValidation}
+
 We validated the predictive power of the most parsimonious movement model using
-k-fold cross-validation for case-control studies \cite{Fortin.2009}.
+k-fold cross-validation for case-control studies \citep{Fortin.2009}.
 Specifically, we randomly assigned 80\% of the strata to a training set and the
@@
 \section{Source Areas \& Points}
-\begin{figure}[htbp]
+
+\begin{figure}[!ht]
  \begin{center}
-  \includegraphics[width = \textwidth]{99_SourceAreas.png}
-  \caption{(a) Different source areas from which we released virtual
-   dispersers. We only considered contiguous protected areas (national parks,
-   game reserves, and forest reserves) that were larger than 700
-   km\textsuperscript{2}. This size corresponds to the average home range
-   requirement for viable wild dog populations \citep{Pomilia.2015}. To render
-   potential immigrants into the study system, we also initiated dispersers
-   within a buffer zone (blue) surrounding the main study area. (b) Source
-   points from which dipsersers were released. 50'000 dispersers were released
-   within the main study area (green dots) and another 30'000 dispersers within
-   the virtual buffer (blue dots).}
+  \includegraphics[width = \textwidth]{Figures/SourceAreas.png} \caption{(a)
+  Different source areas from which we released virtual dispersers. We only
+  considered contiguous protected areas (national parks, game reserves, and
+  forest reserves) that were larger than 700 km\textsuperscript{2}. This size
+  corresponds to the average home range requirement for viable wild dog
+  populations \citep{Pomilia.2015}. To render potential immigrants into the
+  study system, we also initiated dispersers within a buffer zone (blue)
+  surrounding the main study area. (b) Source points from which dispersers were
+  released. 50,000 dispersers were released within the main study area (green
+  dots) and another 30,000 dispersers within the virtual buffer (blue dots).}
   \label{SourcePoints}
@@
 \section{Betweenness Maps}
+\label{BetweennessApproaches}
+
 To generate betweenness maps, we first overlaid the study area with a regular
 grid with a resolution of 2.5 km x 2.5 km per grid cell. The grid cells were
-labelled with unique numbers, which allowed us to then generate a ``visitation
+labeled with unique numbers, which allowed us to then generate a ``visitation
 history'' for each simulated trajectory, showing through which grid cells each
-simulated invidual moved. Based on this history, we tallied the number of
-transitions that occured from one cell to another. Ultimately, the so generated
+simulated individual moved. Based on this history, we tallied the number of
+transitions that occurred from one cell to another. Ultimately, the so generated
 list of transitions served to generate a network and to compute betweenness
@@
 each step. If it ends in a different grid-cell than it started, a transition has
-occured. This approach is computationally simple as it only requires a point
+occurred. This approach is computationally simple as it only requires a point
 extraction (at the start and endpoint of each step). However, relying on a point
@@
 accordance to the direction of movement. In R, such an extraction can be
-achieved e.g. using \textsf{raster::extract(raster, line, along = T)}, which is
-computationally very demanding and therefore often an unviable solution. Here,
+achieved e.g. using \texttt{raster::extract(raster, line, along = T)}, which is
+computationally very demanding and therefore often an nonviable solution. Here,
 we employed yet another approach and generated interpolated coordinates along
@@
 the line extraction very closely, even at higher resolutions. A detailed
-discussion about this is can also be found in \citep{BastilleRousseau.2018}.
+discussion about this is can also be found in \citep{Bastille-Rousseau.2018}.
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
  \begin{center}
-  \includegraphics[width = \textwidth]{99_BetweennessApproaches.png}
+  \includegraphics[width = \textwidth]{Figures/BetweennessApproaches.png}
   \caption{We simulated five trajectories where individuals moved through a
@@
   SD as computed from 10 repetitions.}
-  \label{BetweennessApproaches}
+  \label{BetweennessComparison}
  \end{center}
@@
 \newpage
-\begin{landscape}
- \thispagestyle{empty}
+\newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm, footskip=-5cm}
+
 \section{Model Selection Results}
- % latex table generated in R 3.6.3 by xtable 1.8-4 package
- % Thu Jul  9 09:17:26 2020
+
 \begin{table}[hbpt]
@@
   upon which we based our movement model is depicted in the last row and was
-  originally presented in \cite{Hofmann.2021}. We omitted all models with an AIC
-  weight of zero from the table.}
+  originally presented in \citet{Hofmann.2021}. We omitted all models with an
+  AIC weight of zero from the table.}
   \label{ModelAICs}
-  \begin{center}
-   \resizebox{1.75\textwidth}{!}{
+  \begin{adjustbox}{max width=1.3\textwidth, angle=90, center}
     \begin{threeparttable}
@@
     \end{threeparttable}
-   }
-  \end{center}
+  \end{adjustbox}
 \end{table}
- \vfill
- \raisebox{-1cm}{\makebox[\linewidth]{\thepage}}
-\end{landscape}
+\restoregeometry
 
@@
 \section{Movement Model}
-\begin{table}[htbp]
+
+\begin{table}[!ht]
  \begin{center}
@@
    were standardized to a mean of zero and standard deviation of 1. Plots to aid
-   with the interpretation of this model are given in Appendix S2.}
+   with the interpretation of this model are given in
+   \Cref{ModelInterpretation}.}
   \label{MovementModelNumbers}
-  \resizebox{\textwidth}{!} {
    \begin{threeparttable}
@@
    \end{threeparttable}
-  }
  \end{center}
@@
 \newpage
-% \newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
 \section{Movement Model Interpretation}
+\label{ModelInterpretation}
+
 To ease with the interpretation of the most parsimonious movement model, we
-followed recommendations published in \cite{Fieberg.2021} and produced a series
+followed recommendations published in \citet{Fieberg.2021} and produced a series
 of plots highlighting how the habitat and movement kernel depended on covariate
@@
 scale \(\theta\) = 6'308 and shape \(k\) = 0.37) by applying the function
-\textsf{update\_vonmises()} from the R-package \textsf{amt} \citep{Amt.2019}.
-This allowed us to compute probability densities of turning angles and step
-lengths under varying values of the associated covariates, while holding all
-other covariates constant (\Cref{Interpretation}, a1-a8). Moreover, we
-investigated the habitat kernel by computing relative selection strengths (RSS)
-between a set of steps where values of the covariate of interest was varied to a
-reference step where the covariate value was fixed to its centered value. To
-illustrate model uncertainty, we also generated large-sample confidence
-intervals using standard errors associated with each model estimate
+\texttt{update\_vonmises()} from the \texttt{R}-package \texttt{amt}
+\citep{Signer.2019}. This allowed us to compute probability densities of turning
+angles and step lengths under varying values of the associated covariates, while
+holding all other covariates constant (\Cref{Interpretation}, a1-a8). Moreover,
+we investigated the habitat kernel by computing relative selection strengths
+(RSS) between a set of steps where values of the covariate of interest was
+varied to a reference step where the covariate value was fixed to its centered
+value. To illustrate model uncertainty, we also generated large-sample
+confidence intervals using standard errors associated with each model estimate
 (\Cref{Interpretation}, b1-b5).
@@
  \begin{center}
-  \includegraphics[width = \textwidth]{99_MovementModelInterpretation.png}
+  \includegraphics[width = \textwidth]{Figures/MovementModelInterpretation.png}
   \caption{Auxiliary plots that help with the interpretation of the most
-  parsimonious movement model from Table S1. The plots were generated following
-  recommendations reported in \cite{Fieberg.2021}. Subplots a1 to a8 highlight
+  parsimonious movement model. The plots were generated following
+  recommendations reported in \citet{Fieberg.2021}. Subplots a1 to a8 highlight
   dispersing wild dogs' movement kernel and indicate how the kernel is
@@
 \end{figure}
-% \restoregeometry
 
@@
 \section{Convergence}
-\begin{figure}[htbp]
+
+\begin{figure}[!ht]
   \begin{center}
-    \includegraphics[width=\textwidth]{99_Convergence}
-    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5
-    km) distributed randomly across the study area. The relative traversal
-    frequency is plotted against the number of simulated individuals to
-    visualize how quickly the metric converges to a steady state. (a) Replicated
-    (100 times) relative traversal frequencies across three randomly chosen
-    checkpoints as well as the corresponding 95\% prediction interval (PI). (b)
-    Averaged relative traversal frequency across all checkpoints and replicates
-    including a 95\% PI. (c) Width of the PI in relation to the number of
-    simulated dispersers.}
+    \includegraphics[width=\textwidth]{Figures/Convergence} \caption{Relative
+    traversal frequency through 1,000 checkpoints (5 km x 5 km) distributed
+    randomly across the study area. The relative traversal frequency is plotted
+    against the number of simulated individuals to visualize how quickly the
+    metric converges to a steady state. (a) Replicated (100 times) relative
+    traversal frequencies across three randomly chosen checkpoints as well as
+    the corresponding 95\% prediction interval (PI). (b) Averaged relative
+    traversal frequency across all checkpoints and replicates including a 95\%
+    PI. (c) Width of the PI in relation to the number of simulated dispersers.}
     \label{Convergence}
@@
 \section{Heatmaps in Relation to the Number of Simulated Steps}
+
 \begin{figure}[hbtp]
  \begin{center}
-  \includegraphics[width = 0.86\textwidth]{99_HeatmapsIndividual.png}
-  \caption{Heatmaps produced when considering 125, 500, and 2000 simulated
+  \includegraphics[width = 0.86\textwidth]{Figures/HeatmapsIndividual.png}
+  \caption{Heatmaps produced when considering 125, 500, and 2,000 simulated
   steps, respectively. The left panel (a1, a2, a3) was generated based on
-   simulations initiated within the main study area, the right panel (b1, b2,
-   b3) was generated based on simulations initiated within the buffer area. To
-   produce the heatmap presented in the main manuscript (Figure 5), we tallied
-   the values from maps a3 and b3.}
-  \label{Heatman}
- \end{center}
-\end{figure}
-\restoregeometry
+  simulations initiated within the main study area, the right panel (b1, b2, b3)
+  was generated based on simulations initiated within the buffer area. To
+  produce the heatmap presented in the main manuscript \Cref{Heatmap}, we
+  tallied the values from maps a3 and b3.} \label{HeatmapsIndividual}
+  \end{center} \end{figure} \restoregeometry
 
@@
 \section{Betweenness Maps in Relation to the Number of Simulated Steps}
+
 \begin{figure}[hbtp]
  \begin{center}
-  \includegraphics[width = 0.86\textwidth]{99_BetweennessIndividual.png}
+  \includegraphics[width = 0.86\textwidth]{Figures/BetweennessIndividual.png}
   \caption{Maps of betweenness scores produced when considering (a) 125, (b)
-   500, (c) and 2000 simulated steps, respectively. A high betweenness score
+  500, (c) and 2,000 simulated steps, respectively. A high betweenness score
   indicates that the respective area has a high importance for linking other
-   regions in the study area.}
-  \label{Betweenness}
- \end{center}
-\end{figure}
-\restoregeometry
+  regions in the study area.} \label{BetweennessIndividual} \end{center}
+  \end{figure} \restoregeometry
 
@@
 \section{Comparison of Traversal Frequencies and Betweenness Scores Inside and Outside KAZA-TFCA}
+
 \begin{figure}[hbtp]
  \begin{center}
-  \includegraphics[width = \textwidth]{99_MapComparison.png}
+  \includegraphics[width = \textwidth]{Figures/MapComparison.png}
   \caption{Comparison of values from the heatmap and betweenness map inside
@@
  \begin{center}
-  \includegraphics[width = \textwidth]{99_MapComparisonCountries.png}
+  \includegraphics[width = \textwidth]{Figures/MapComparisonCountries.png}
   \caption{Comparison of values from the heatmap and betweenness map inside
@@
 \section{Dispersal into other National Parks}
+
 \begin{figure}[hbtp]
  \begin{center}
-  \includegraphics[width = \textwidth]{99_AreasReached.png}
+  \includegraphics[width = \textwidth]{Figures/AreasReached.png}
   \caption{Relative number of simulated dispersal trajectories that successfully
@@
 
+\ifSubfilesClassLoaded{%
   \newpage
-\begingroup
-\singlespacing
-\bibliography{Literatur}
-\endgroup
+  \begin{singlespacing}
+  \ifthenelse{\boolean{usebiblatex}}{
+    \begin{refcontext}[sorting=nyt]
+    \printbibliography
+    \end{refcontext}
+  } {
+    \bibliography{../LiteratureBibtex}%
+  }
+\end{singlespacing}
+}{}
 

### 2024-07-01 FinalThesis/Chapter_02/Manuscript.tex
Publish PHD

--- 2024-07-01 Chapter_2/04_Manuscript/Manuscript.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-\usepackage[nomarkers, nolists]{endfloat}
-\usepackage{scalerel}
-\usepackage{makecell}
-\usepackage{ifthen}
-\usepackage{tikz}
-\usepackage{lineno}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%  Bibtex vs Biblatex
-%------------------------------------------------------------------------------
-% While bibtex is very nice to use, it's insanely slow. I therefore work with
-% bibtex until the final stage. Set the value below to "true" or "false" to pick
-% which one should be used.
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{true}
-
-% To remove figures from the document, only keeping their captions
-\AtBeginDelayedFloats{%
-  \renewcommand{\includegraphics}[2][]{}% Remove \includegraphics functionality
-  \renewcommand{\resizebox}[3]{}% Remove \resizebox functionality
-}
-\setcounter{totalnumber}{10}% Allow up to 10 floats per page
-\setcounter{topnumber}{10}% Allow up to 10 floats per page
-\setcounter{bottomnumber}{10}% Allow up to 10 floats per page
-\renewcommand{\efloatseparator}{\mbox{}}% From the documentation
-
-% Depending on the question, the commands below will prepare the document
-% accordingly
-\ifthenelse{\boolean{usebiblatex}}{
-  %-----------------------------------------------------------------------------
-  %  If using biblatex
-  %-----------------------------------------------------------------------------
-  % Load required packages and refer to correct bibliography
-  \usepackage{csquotes}
-  \usepackage[
-      backend=biber
-    , style=apa          % To display Author-year in the text
-    , natbib=true        % Necessary to use citep, etc.
-    , hyperref=true      % To have clickable links in the references
-    , useeditor=false    % Don't add editor to publications
-    , sorting=ynt        % Sort by year, name, and title
-    , uniquename=false   % Avoid that initials are put
-  ]{biblatex}
-  \addbibresource{Literature.bib}
-
-  % To include author in link
-  \DeclareFieldFormat{citehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{#1}}
-  \DeclareFieldFormat{textcitehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{%
-    #1%
-    \ifbool{cbx:parens}
-      {\bibcloseparen\global\boolfalse{cbx:parens}}
-      {}}}
-  \savebibmacro{cite}
-  \savebibmacro{textcite}
-  \renewbibmacro*{cite}{%
-    \printtext[citehyperref]{%
-      \restorebibmacro{cite}%
-      \usebibmacro{cite}}}
-  \renewbibmacro*{textcite}{%
-    \ifboolexpr{
-      ( not test {\iffieldundef{prenote}} and
-        test {\ifnumequal{\value{citecount}}{1}} )
-      or
-      ( not test {\iffieldundef{postnote}} and
-        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
-    }
-      {\DeclareFieldAlias{textcitehyperref}{noformat}}
-      {}%
-    \printtext[textcitehyperref]{%
-      \restorebibmacro{textcite}%
-      \usebibmacro{textcite}}}
-
-  % Remove editor form everything but books
-  \AtEveryBibitem{%
-   \ifentrytype{book}{}{%
-    \clearname{editor}
-   }
-  }
-}{
-  %-----------------------------------------------------------------------------
-  %  If using bibtex
-  %-----------------------------------------------------------------------------
-  % Load required packages and select citation style
-  \usepackage[round]{natbib}
-  \bibliographystyle{apalike}
-}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Link colors
-% \definecolor{linkcolor}{HTML}{000000}
-% \hypersetup{
-%   colorlinks = true,
-%   linkcolor  = linkcolor,
-%   urlcolor   = linkcolor,
-%   citecolor  = linkcolor,
-% }
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% Avoid skip when including text from external .tex file
-\newcommand{\inputy}[1]{\input{#1}\unskip}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%  ORCID
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Create the ORCID logo
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol]
-      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
-        ,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white]
-      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
-        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
-        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
-      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-
-% Command to create the ORCID
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-% \title{Flooding of the Okavango Delta influences Connectivity for Dispersing
-% African Wild Dogs}
-\title{Dispersal and Connectivity in Increasingly Extreme Climatic Conditions}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1,2} \orcid{0000-0001-7477-2642} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
-
-\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
-
-\vspace{4cm}
-
-\textbf{Running Title:} Dispersal in a Changing Climate
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} climate change, conservation, dispersal, flood extremes,
-individual-based simulations, landscape connectivity, movement ecology,
-step-selection functions
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Abstract
-%------------------------------------------------------------------------------
+\chapter{Dispersal and Connectivity in Increasingly Extreme Climatic Conditions}
+\label{ChapterFlood}
+\thispagestyle{empty}
+\vspace{-1cm}
+\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil
+
+\begin{center}
+David D. Hofmann \orcid{0000-0003-3477-4365},
+Dominik M. Behr \orcid{0000-0001-7378-8538},
+John W. McNutt,
+Arpat Ozgul \orcid{0000-0001-7477-2642}, and
+Gabriele Cozzi \orcid{0000-0002-1744-1940}
+
+% \begin{flushleft}
+
+% \vspace{5cm}
+\vfill
+
+Published in Global Change Biology (2024):
+\url{https://doi.org/10.1111/gcb.17299}.
+
+% \textsuperscript{1} Department of Evolutionary Biology and Environmental
+% Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
+% Switzerland.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+% Private Bag 13, Maun, Botswana.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{\S} Corresponding author:
+% \href{mailto:david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
+
+\end{center}
+
+% \vfill
+
+% \textbf{Running Title:}\\African Wild Dog Dispersal in a Changing Climate:
+% Lessons Learned from Seasonal Flood Extremes
+%
+% \vspace{0.5cm}
+%
+% \textbf{Keywords:}\\movement, connectivity, climate change, seasonality,
+% dispersal, conservation, individual-based simulations, step-selection
+
+% \end{flushleft}
+
 \newpage
-\begin{abstract}
+\section*{Abstract}
 While climate change has been shown to impact several life-history traits of
@@
 reduces dispersal success (i.e., the propensity of individuals to disperse
-between adjacent subpopulations) by 12\% and increases dispersal durations by
-17\%. Locally, however, dispersal success diminishes by as much as 78\%.
-Depending on the flood extent, alternative dispersal corridors emerge, some of
-which in the immediate vicinity of human-dominated landscapes. Notably, under
-maximum flood extent, the number of dispersing trajectories moving into
-human-dominated landscapes decreases by 41\% at the Okavango Delta's inflow, but
-increases by 126\% at the Delta's distal end. This may drive the amplification
-of human-wildlife conflict.
+between adjacent subpopulations) by
+\inputy{Figures/NumberReachingOtherSourceAreasMaxFloodPercent.tex}\% and
+increases dispersal durations by
+\inputy{Figures/DurationReachingOtherSourceAreasMaxFloodPercent.tex}\%. Locally,
+however, dispersal success diminishes by as much as
+\inputy{Figures/NumberReaching6MaxFloodPercent.tex}\%. Depending on the flood
+extent, alternative dispersal corridors emerge, some of which in the immediate
+vicinity of human-dominated landscapes. Notably, under maximum flood extent, the
+number of dispersing trajectories moving into human-dominated landscapes
+decreases by \inputy{Figures/DensityHWCPanhandlePercentage.tex}\% at the
+Okavango Delta's inflow, but increases by
+\inputy{Figures/DensityHWCMaunPercentage.tex}\% at the Delta's distal end. This
+may drive the amplification of human-wildlife conflict.
 
@@
 
-\end{abstract}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
 \newpage
-
-% \onehalfspacing
-% \tableofcontents
-% \doublespacing
-
-% Change page numbering
-\newpage
-\pagenumbering{arabic}
-
-% Create linenumbers
-\linenumbers
-
 \section{Introduction}
@@
 to reconstruct and predict movement trajectories at the level of the individual
-have been suggested \citep{Signer.2017, Hofmann.2023, Signer.2023}. Such
+have been suggested \citep{Signer.2017, Hofmann.2023, Signer.2024}. Such
 predicted trajectories are uncoupled from any temporal constraint and therefore
@@
 
-\begin{figure}
+\begin{figure}[htpb]
  \begin{center}
@@
 focused on a \textit{core study area} of 70,000 km\textsuperscript{2} that
-comprised the Delta and its surroundings (\Cref{StudyArea}b; purple circle). To
-accommodate for the long-distance dispersal events observed in this ecosystem
+comprised the Delta and its surroundings (\Cref{StudyAreaCH2}b; purple circle).
+To accommodate for the long-distance dispersal events observed in this ecosystem
 \citep{Cozzi.2020, Hofmann.2021, Cozzi.2023}, we also considered an
@@
 comprising the Delta and parts of Angola, Namibia, Botswana, Zimbabwe, and
-Zambia (\Cref{StudyArea}a). Rainfall in this area averages at 450 mm and is
+Zambia (\Cref{StudyAreaCH2}a). Rainfall in this area averages at 450 mm and is
 mainly concentrated to the wet season between November and April
@@
 Delta's distal ends, it percolates at the Thamalakane and Kunyere Faults, two
-natural fault lines at which the water-flow is hindered (\Cref{StudyArea}b).
+natural fault lines at which the water-flow is hindered (\Cref{StudyAreaCH2}b).
 After reaching its maximum extent, water evaporates or percolates and the flood
@@
 transboundary conservation initiative, the Kavango-Zambezi Transfrontier
-Conservation Area (KAZA-TFCA, \Cref{StudyArea}a), which attempts to re-establish
-connectivity among several core habitats. Previous studies have indicated that
-this initiative has high potential for improving connectivity among
-subpopulations of various species \citep{Brennan.2020, Lines.2021}, including
-African wild dogs \citep{Hofmann.2021}. However, a better understanding of
-connectivity within this ecosystem under climate change is lacking.
-
-\begin{figure}
+Conservation Area (KAZA-TFCA, \Cref{StudyAreaCH2}a), which attempts to
+re-establish connectivity among several core habitats. Previous studies have
+indicated that this initiative has high potential for improving connectivity
+among subpopulations of various species \citep{Brennan.2020, Lines.2021},
+including African wild dogs \citep{Hofmann.2021}. However, a better
+understanding of connectivity within this ecosystem under climate change is
+lacking.
+
+\begin{figure}[htpb]
  \begin{center}
-  \includegraphics[width = \textwidth]{Figures/StudyArea2.png} \caption{(a)
-  Location of the study area, which forms part of the Kavango-Zambezi
-  Transfrontier Conservation Area (KAZA-TFCA, red polygon) in Southern Africa.
-  (b) The division of the core study area (enclosed by the purple ellipse) into
-  sub-areas (1 to 6) was based on the hydrographic structure of the Okavango
-  Delta and its tributaries. We simulated dispersal trajectories starting at
-  random locations within each of the six source areas (orange circles). Purple
-  zones (7 to 14) represent zones that we used to identify if and where
-  simulated dispersers left the close surroundings of the Okavango Delta (thus
-  called egression zones). These zones were generated using a set of cutlines
-  (purple dotted lines) originating from the center of the Delta that dissected
-  an elliptical buffer surrounding the Delta into sections of equal size and in
-  accordance with cardinal directions. The yellow lines represent GPS
-  trajectories of dispersing wild dogs that were observed using GPS collars.
-  While not all of these individuals were used to parametrize the dispersal
-  model presented by \citet{Hofmann.2023}, they serve to highlight the spatial
-  scale at which the species disperses. Map lines delineate study areas and do
-  not necessarily depict accepted national boundaries.}
-  \label{StudyArea}
+  \includegraphics[width = \textwidth]{Figures/StudyArea2.png}
+  \caption{(a) Location of the study area, which forms part of the
+  Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA, red polygon) in
+  Southern Africa. (b) The division of the core study area (enclosed by the
+  purple ellipse) into sub-areas (1 to 6) was based on the hydrographic
+  structure of the Okavango Delta and its tributaries. We simulated dispersal
+  trajectories starting at random locations within each of the six source areas
+  (orange circles). Purple zones (7 to 14) represent zones that we used to
+  identify if and where simulated dispersers left the close surroundings of the
+  Okavango Delta (thus called egression zones). These zones were generated using
+  a set of cutlines (purple dotted lines) originating from the center of the
+  Delta that dissected an elliptical buffer surrounding the Delta into sections
+  of equal size and in accordance with cardinal directions. The yellow lines
+  represent GPS trajectories of dispersing wild dogs that were observed using
+  GPS collars. While not all of these individuals were used to parametrize the
+  dispersal model presented by \citet{Hofmann.2023}, they serve to highlight the
+  spatial scale at which the species disperses.}
+  \label{StudyAreaCH2}
  \end{center}
@@
 set of spatially referenced habitat layers, each with a resolution of 250 m x
-250 m. The set of layers included water-cover, distance-to-water, tree-cover,
-shrub/grassland-cover, and a composite human influence layer representing
-settlements, roads, and agricultural areas. A detailed description of these
-layers is provided by \citet{Hofmann.2021}. However, unlike Hofmann and
-colleagues, who used a single, static flood map, we employed composite flood
-maps to study connectivity. Specifically, we generated water-cover layers using
-MODIS Terra MCD43A4 satellite imagery that we classified into binary water-cover
-maps using an algorithm developed by \citet{Wolski.2017} and implemented in
-\texttt{R} (\url{https://github.com/DavidDHofmann/floodmapr}). This algorithm
-allowed us to generate weekly updated ``flood extent maps'', thus providing
-detailed information about the flood at any given point in time. In total, we
-generated 700 flood extent maps, covering the years 2000 to 2019. We used these
-maps to produce minimum and maximum flood scenarios, representative of dry and
-wet climatic conditions, assuming that climate change will shift the system
-towards one of these extremes in the long run. To create the baseline minimum
-(and respectively, maximum) flood scenario, we averaged the 100 flood extent
-maps with the smallest (highest) flood extent and generated a binary layer by
-masking all pixels that were inundated in less than 50\% of the maps (see also
+250 m. The set of layers included \textsf{water-cover, distance-to-water,
+tree-cover, shrub/grassland-cover}, and a composite \textsf{human influence}
+layer representing settlements, roads, and agricultural areas. A detailed
+description of these layers is provided by \citet{Hofmann.2021}. However, unlike
+Hofmann and colleagues, who used a single, static flood map, we employed
+composite flood maps to study connectivity. Specifically, we generated
+water-cover layers using MODIS Terra MCD43A4 satellite imagery that we
+classified into binary water-cover maps using an algorithm developed by
+\citet{Wolski.2017} and implemented in \texttt{R}
+(\url{https://github.com/DavidDHofmann/floodmapr}). This algorithm allowed us to
+generate weekly updated ``flood extent maps'', thus providing detailed
+information about the flood at any given point in time. In total, we generated
+700 flood extent maps, covering the years 2000 to 2019. We used these maps to
+produce minimum and maximum flood scenarios, representative of dry and wet
+climatic conditions, assuming that climate change will shift the system towards
+one of these extremes in the long run. To create the baseline minimum (and
+respectively, maximum) flood scenario, we averaged the 100 flood extent maps
+with the smallest (highest) flood extent and generated a binary layer by masking
+all pixels that were inundated in less than 50\% of the maps (see also
 \Cref{Concept}). By averaging across 100 flood extent maps, we followed a
@@
 flood extent maps. The final minimum and maximum flood extent maps
-(\Cref{FloodExtent}) show that the flood in the two scenarios covers 3,500
-km\textsuperscript{2} and 9,500 km\textsuperscript{2}, respectively. Finally, we
-combined the two flood extent maps with the set of above-mentioned habitat
-layers into two spatial stacks, each representative of one scenario
+(\Cref{FloodExtent}) show that the flood in the two scenarios covers
+\inputy{Figures/FloodExtentMinimum.tex} km\textsuperscript{2} and
+\inputy{Figures/FloodExtentMaximum.tex} km\textsuperscript{2}, respectively.
+Finally, we combined the two flood extent maps with the set of above-mentioned
+habitat layers into two spatial stacks, each representative of one scenario
 (\Cref{Concept}). To prevent edge effects during the dispersal simulation, we
@@
 wider and taller than the original, filling the so created buffer zone (gray
-buffer in \Cref{StudyArea}b) with randomized values from the respective layers.
-This would allow simulated dispersers to leave and re-enter the main study area
-via randomized buffer zones.
-
-\begin{figure}
+buffer in \Cref{StudyAreaCH2}b) with randomized values from the respective
+layers. This would allow simulated dispersers to leave and re-enter the main
+study area via randomized buffer zones.
+
+\begin{figure}[htpb]
  \begin{center}
@@
 We simulated dispersing wild dogs originating from six distinct \textit{source
-areas} located within the main study area (\Cref{StudyArea}b). Defining these
+areas} located within the main study area (\Cref{StudyAreaCH2}b). Defining these
 source areas was necessary to enable identification and quantification of the
@@
 a natural latitudinal split between the eastern (areas 1, 2, and 3) and western
-(areas 4 and 5) Delta (\Cref{StudyArea}b). Areas 1 and 2 are separated
+(areas 4 and 5) Delta (\Cref{StudyAreaCH2}b). Areas 1 and 2 are separated
 longitudinally by the Selinda Spillway, a waterway that connects the Delta with
 the Linyanti Swamp, while areas 2 and 3 are separated by the Khwai River, which
-inundates the Mababe Depression east of the Delta (\Cref{StudyArea}b). On the
+inundates the Mababe Depression east of the Delta (\Cref{StudyAreaCH2}b). On the
 western side of the Delta, areas 4 and 5 are longitudinally separated by the
-Thaoga River drainage at Nokaneng (\Cref{StudyArea}b). Finally, the Delta hosts
-a central island, known as Chief's Island, which we defined as source area 6
-(\Cref{StudyArea}b). The shape of this area was determined by its borders
+Thaoga River drainage at Nokaneng (\Cref{StudyAreaCH2}b). Finally, the Delta
+hosts a central island, known as Chief's Island, which we defined as source area
+6 (\Cref{StudyAreaCH2}b). The shape of this area was determined by its borders
 towards water during the maximum flood scenario, whereas all other source areas
@@
 scenarios. Besides source areas 1 to 6, we also generated \textit{egression
-zones} (\Cref{StudyArea}b). These zones enabled us to determine through which
+zones} (\Cref{StudyAreaCH2}b). These zones enabled us to determine through which
 regions simulated individuals left the core study area, as well as to identify
@@
 overlaying the Delta with an ellipse that we dissected into equally sized
-polygons in accordance with cardinal directions (\Cref{StudyArea}b). Overall,
+polygons in accordance with cardinal directions (\Cref{StudyAreaCH2}b). Overall,
 the selection of source areas and egression zones was based on a mixture between
@@
 resembles a fully mechanistic model from which movement can be simulated
-\citep{Signer.2017, Potts.2023, Signer.2023}. In contrast to graph-based
+\citep{Signer.2017, Potts.2023, Signer.2024}. In contrast to graph-based
 connectivity models (e.g., least-cost path analysis or circuit-theory),
@@
 preference for directional and fast movements. The associated model parameters
-(provided in Figure S1) can be used to predict probabilities of a step being
-chosen among a set of random steps (details in Supporting Information S1), which
+(provided in \Cref{Model}) can be used to predict probabilities of a step being
+chosen among a set of random steps (details in \Cref{DispersalModel}), which
 allows to simulate movements in discrete time.
 
-Originating from each of the six source areas (\Cref{StudyArea}b), we simulated
-2,000 individual dispersing trajectories, each composed of 2,000 steps. Two
-thousand steps corresponded to a dispersal duration of $\sim$ 400 days, which
-marks the longest dispersal duration recorded in the study area
+Originating from each of the six source areas (\Cref{StudyAreaCH2}b), we
+simulated 2,000 individual dispersing trajectories, each composed of 2,000
+steps. Two thousand steps corresponded to a dispersal duration of $\sim$ 400
+days, which marks the longest dispersal duration recorded in the study area
 \citep{Cozzi.2020, Hofmann.2021}. By simulating 2,000 steps, we generated the
@@
 facilitating movement into adjacent areas \citep{Bastille-Rousseau.2018,
-Bastille-Rousseau.2021}. We computed betweenness using the \texttt{igraph}
+Bastille-Rousseau.2020}. We computed betweenness using the \texttt{igraph}
 \texttt{R}-package \citep{Csardi.2006}. Calculations of betweenness scores are
@@
 we also generated a compound metric by multiplying the generated heatmaps with a
-continuous human-influence map (Supporting Information S10). Notably, not every
+continuous human-influence map (\Cref{AlternativeHWC}). Notably, not every
 animal in the vicinity of human-dominated landscapes implies conflict, hence we
@@
 \texttt{R} \citep{RCoreTeam.2023}. For any spatial data manipulation, we used
-the \texttt{R}-packages \texttt{terra} \citep{Hijmans.2023} and
+the \texttt{R}-packages \texttt{terra} \citep{Hijmans.2024} and
 \texttt{spatstat} \citep{Baddeley.2015}. Several helper functions for the
@@
 analysis was achieved in \texttt{igraph} \citep{Csardi.2006} and figures were
-generated using \texttt{ggplot2} \citep{Wickham.2023} and \texttt{ggnetwork}
-\citep{Briatte.2023}. Notably, a similar simulation procedure as employed in our
-analysis has recently been added to the \texttt{amt} \texttt{R}-package
-\citep{Signer.2023}.
+generated using \texttt{ggplot2} \citep{Wickham.2024} and \texttt{ggnetwork}
+\citep{Briatte.2024}. All \texttt{R}-scripts required to replicate our analyses
+are provided through an online repository. Notably, a similar simulation
+procedure as employed in our analysis has recently been added to the
+\texttt{amt} \texttt{R}-package \citep{Signer.2024}.
 
@@
 dispersal success and duration depending on the flood scenario (\Cref{IPCTable},
-\Cref{IPCMap}, Figure S2). Of the 6,000 simulated dispersal trajectories for
-each flood scenario, 4,137 $\pm$ 35 trajectories reached another source area
-during minimum flood, whereas 3,625 $\pm$ 32 trajectories did so during maximum
-flood, indicating a 12\% lower dispersal success during maximum flood
-(\Cref{IPCTable}a1). Concurrently, the average dispersal duration of
-trajectories moving from one source area to another was 17\% lower during
-maximum flood, with 122 $\pm$ 1 days compared to 143 $\pm$ 2 days
+\Cref{IPCMap}, \Cref{EmigrationImmigration}). Of the 6,000 simulated dispersal
+trajectories for each flood scenario,
+\inputy{Figures/NumberReachingOtherSourceAreasMinFlood.tex} trajectories reached
+another source area during minimum flood, whereas
+\inputy{Figures/NumberReachingOtherSourceAreasMaxFlood.tex} trajectories did so
+during maximum flood, indicating a
+\inputy{Figures/NumberReachingOtherSourceAreasMaxFloodPercent.tex}\% lower
+dispersal success during maximum flood (\Cref{IPCTable}a1). Concurrently, the
+average dispersal duration of trajectories moving from one source area to
+another was
+\inputy{Figures/DurationReachingOtherSourceAreasMaxFloodPercent.tex}\% lower
+during maximum flood, with
+\inputy{Figures/DurationReachingOtherSourceAreasMinFlood.tex} days compared to
+\inputy{Figures/DurationReachingOtherSourceAreasMaxFlood.tex} days
 (\Cref{IPCTable}b1) during minimum flood. The disparities were most pronounced
-for trajectories dispersing into source area 6 (the Delta's center), with 1,327
-$\pm$ 30 simulated trajectories reaching area 6 during minimum flood compared to
-297 $\pm$ 17 trajectories (i.e., 78\% lower) during maximum flood
-(\Cref{IPCTable}a1). Source area 6 therefore appeared to be particularly
+for trajectories dispersing into source area 6 (the Delta's center), with
+\inputy{Figures/NumberReaching6MinFlood.tex} simulated trajectories reaching
+area 6 during minimum flood compared to
+\inputy{Figures/NumberReaching6MaxFlood.tex} trajectories (i.e.,
+\inputy{Figures/NumberReaching6MaxFloodPercent.tex}\% lower) during maximum
+flood (\Cref{IPCTable}a1). Source area 6 therefore appeared to be particularly
 vulnerable to isolation in the maximum flood scenario. Interestingly,
 connectivity between some source areas was higher during maximum flood. For
-instance, the number of trajectories running from area 5 into area 4 was 67\%
-higher, with 114 $\pm$ 10 trajectories during minimum flood compared to 190
-$\pm$ 12 trajectories during maximum flood (\Cref{IPCTable}a1). Simulated
-individuals therefore responded to unfavorable habitat conditions by dispersing
-into another region of the landscape. Contrary to our expectations, movement
-into egression zones (areas 7 - 14) was only marginally higher during maximum
-flood, with the number of trajectories rising by 2\%, from 5,457 $\pm$ 22 to
-5,553 $\pm$ 20 (\Cref{IPCTable}a2 and Figure S5).
-
-\begin{figure}
-\begin{center}
- \includegraphics[width = \textwidth]{Figures/InterpatchConnectivityTable2.png}
+instance, the number of trajectories running from area 5 into area 4 was
+\inputy{Figures/Connectivity5to4MaxFloodPercent.tex}\% higher, with
+\inputy{Figures/Connectivity5to4MinFlood.tex} trajectories during minimum flood
+compared to \inputy{Figures/Connectivity5to4MaxFlood.tex} trajectories during
+maximum flood (\Cref{IPCTable}a1). Simulated individuals therefore responded to
+unfavorable habitat conditions by dispersing into another region of the
+landscape. Contrary to our expectations, movement into egression zones (areas 7
+- 14) was only marginally higher during maximum flood, with the number of
+trajectories rising by \inputy{Figures/EgressionMaxFloodPercent.tex}\%, from
+\inputy{Figures/EgressionMinFlood.tex} to \inputy{Figures/EgressionMaxFlood.tex}
+(\Cref{IPCTable}a2 and \Cref{Egression}).
+
+\begin{table}
  \caption{(a) Dispersal frequency (measured as the number of dispersal
@@
  \label{IPCTable}
-\end{center}
-\end{figure}
-
-\begin{figure}
+ \includegraphics[width = \textwidth]{Figures/InterpatchConnectivityTable2.png}
+\end{table}
+
+\begin{figure}[htpb]
  \begin{center}
-  \includegraphics[width = 0.8\textwidth]{Figures/InterpatchConnectivityMap.png}
+  \includegraphics[width = \textwidth]{Figures/InterpatchConnectivityMap.png}
   \caption{Dispersal frequency and duration between source areas 1 to 6 in the
@@
   present links between adjacent source areas for brevity. Detailed plots about
-  source-specific inter-patch connectivity are provided in Figures S3 and S4.}
+  source-specific inter-patch connectivity are provided in \Cref{IPCMain} and
+  \Cref{IPCBuffer}.}
   \label{IPCMap}
@@
 minimum flood. During maximum flood, by contrast, the density of simulated
-locations in the vicinity of humans decreased by 41\% in this area
-(\Cref{Metrics}c and Figure S9). Another hotspot, albeit visually less distinct,
-covered the region at the distal end of the Delta, extending across the town of
-Maun (\Cref{Metrics}c). This area was particularly relevant during maximum
-flood, with a density that was 126\% higher than during minimum flood
+locations in the vicinity of humans decreased by
+\inputy{Figures/DensityHWCPanhandlePercentage.tex}\% in this area
+(\Cref{Metrics}c and \Cref{HWCDifference}). Another hotspot, albeit visually
+less distinct, covered the region at the distal end of the Delta, extending
+across the town of Maun (\Cref{Metrics}c). This area was particularly relevant
+during maximum flood, with a density that was
+\inputy{Figures/DensityHWCMaunPercentage}\% higher than during minimum flood
 (\Cref{Metrics}c and Figure S9). The diversion of individuals from one area to
@@
 
-\begin{figure}
+\begin{figure}[htpb]
  \begin{center}
-  \includegraphics[width = 0.8\textwidth]{Figures/Metrics.png}
+  \includegraphics[width = \textwidth]{Figures/Metrics.png}
   \caption{(a) Heatmaps, (b) betweenness maps, and (c) maps of potential for HWC
@@
   Source areas from which dispersers were released are numbered 1-6. The color
-  scale for betweenness scores in (b) was square-rooted to improve visibility of
-  corridors with lower values. Source-specific heatmaps, betweenness maps, and
-  HWC maps are provided in Figures S6-S8.}
+  scale for betweenness scores in subfigure b was square-rooted to improve
+  visibility of corridors with lower values. Source-specific heatmaps,
+  betweenness maps, and HWC maps are provided in Figures \Cref{HeatmapsInd},
+  \Cref{BetweennessInd}, and \Cref{HWCInd}.}
   \label{Metrics}
@@
 
-\subsection{Impacts of Extreme Environmental Conditions on Dispersal}
+% \subsection{Impacts of Extreme Environmental Conditions on Dispersal}
 
@@
 floodmaps, is conservative compared to historically observed flood extents in
-the Delta. While the flood in our map encompasses an area of 9,500
-km\textsuperscript{2}, the Delta may inundate an area of up to 14,000
-km\textsuperscript{2} in exceptional cases \citep{McCarthy.2003,
-Gumbricht.2004}. Changes in dispersal behavior and connectivity due to climate
-change may therefore be more pronounced than reported here.
-
-\subsection{Human-Wildlife Conflict}
+the Delta. While the flood in our map encompasses an area of
+\inputy{Figures/FloodExtentMaximum.tex} km\textsuperscript{2}, the Delta may
+inundate an area of up to 14,000 km\textsuperscript{2} in exceptional cases
+\citep{McCarthy.2003, Gumbricht.2004}. Changes in dispersal behavior and
+connectivity due to climate change may therefore be more pronounced than
+reported here.
+
+% \subsection{Human-Wildlife Conflict}
 
@@
 
-\subsection{Methodological Limitations and Future Research Directions}
+% \subsection{Methodological Limitations and Future Research Directions}
 
@@
 
-\subsection{Conclusion and Implications}
+% \subsection{Conclusion and Implications}
 
@@
 
-\section{Authors' Contributions}
-
-D.D.H., G.C., D.M.B., A.O. and conceived the study, D.D.H. and G.C. designed
-methodology; D.D.H., G.C., D.M.B., and J.W.M. collected the data; D.D.H.
-analysed the data; D.D.H. and G.C. wrote the first draft of the manuscript and
-all authors contributed to the drafts at several stages and gave final approval
-for publication.
-
-\section{Data Availability}
-
-The data that support the findings of this study are openly available in Dryad
-at \url{https://doi.org/10.5061/dryad.z34tmpgnm} \citep{Hofmann.2024a}. R-codes
-required to reproduce all analyses are accessible through GitHub at
-\url{https://github.com/DavidDHofmann/FloodDispersal}. Additional information on
-the preparation of covariate layers are provided by \citet{Hofmann.2021} and
-a detailed description of the dispersal model is given by \citet{Hofmann.2023}.
-
-\section{Acknowledgements}
-We thank the Ministry of Environment and Tourism and the Department of Wildlife
-and National Parks of Botswana for granting permission to conduct this research.
-We also thank two excellent reviewers for their critical and detailed feedback,
-which helped to improve the quality of the manuscript significantly. This study
-was funded by Claraz Foundation, Forschungkredit der Universität Zürich, and a
-Swiss National Science Foundation Grant Nr. 31003A\_182286 to A. Ozgul and Nr.
-310030\_204478 to G. Cozzi.
-
+% \section{Authors' Contributions}
+% D.D.H., G.C., D.M.B., A.O. and conceived the study, D.D.H. and G.C. designed
+% methodology; D.D.H., G.C., D.M.B., and J.W.M. collected the data; D.D.H.
+% analysed the data; D.D.H. and G.C. wrote the first draft of the manuscript and
+% all authors contributed to the drafts at several stages and gave final approval
+% for publication.
+%
+% \section{Data Availability}
+% Access to R-scripts to replicate our analyses will be provided through an online
+% repository at the time of publication.
+%
+% \section{Acknowledgements}
+% We thank the Ministry of Environment and Tourism and the Department of Wildlife
+% and National Parks of Botswana for granting permission to conduct this research.
+% We also thank two excellent reviewers for their critical and detailed feedback,
+% which helped to improve the quality of the manuscript significantly. This study
+% was funded by Claraz Foundation, Forschungkredit der Universität Zürich, and a
+% Swiss National Science Foundation Grant Nr. 31003A\_182286 to A. Ozgul and Nr.
+% 310030\_204478 to G. Cozzi.
+
+\ifSubfilesClassLoaded{%
   \newpage
-\begingroup
-\singlespacing
+  \begin{singlespacing}
   \ifthenelse{\boolean{usebiblatex}}{
@@
     \end{refcontext}
-}{
-  \bibliography{LiteratureBibtex}
+  } {
+    \bibliography{../LiteratureBibtex}%
   }
-\endgroup
+\end{singlespacing}
+}{}
+
 \end{document}

### 2024-07-01 FinalThesis/Chapter_02/Manuscript_Support.tex
Publish PHD

--- 2024-07-01 Chapter_2/04_Manuscript/Manuscript_Support.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-% \usepackage[nomarkers, nolists]{endfloat}
-\usepackage{scalerel}
-\usepackage{makecell}
-\usepackage{ifthen}
-\usepackage{tikz}
-\usepackage{lineno}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%  Bibtex vs Biblatex
-%------------------------------------------------------------------------------
-% While bibtex is very nice to use, it's insanely slow. I therefore work with
-% bibtex until the final stage. Set the value below to "true" or "false" to pick
-% which one should be used.
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{true}
-
-% Depending on the question, the commands below will prepare the document
-% accordingly
-\ifthenelse{\boolean{usebiblatex}}{
-  %-----------------------------------------------------------------------------
-  %  If using biblatex
-  %-----------------------------------------------------------------------------
-  % Load required packages and refer to correct bibliography
-  \usepackage{csquotes}
-  \usepackage[
-      backend=biber
-    , style=apa          % To display Author-year in the text
-    , natbib=true        % Necessary to use citep, etc.
-    , hyperref=true      % To have clickable links in the references
-    , useeditor=false    % Don't add editor to publications
-    , sorting=ynt        % Sort by year, name, and title
-    , uniquename=false   % Avoid that initials are put
-  ]{biblatex}
-  \addbibresource{Literature.bib}
-
-  % To include author in link
-  \DeclareFieldFormat{citehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{#1}}
-  \DeclareFieldFormat{textcitehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{%
-    #1%
-    \ifbool{cbx:parens}
-      {\bibcloseparen\global\boolfalse{cbx:parens}}
-      {}}}
-  \savebibmacro{cite}
-  \savebibmacro{textcite}
-  \renewbibmacro*{cite}{%
-    \printtext[citehyperref]{%
-      \restorebibmacro{cite}%
-      \usebibmacro{cite}}}
-  \renewbibmacro*{textcite}{%
-    \ifboolexpr{
-      ( not test {\iffieldundef{prenote}} and
-        test {\ifnumequal{\value{citecount}}{1}} )
-      or
-      ( not test {\iffieldundef{postnote}} and
-        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
-    }
-      {\DeclareFieldAlias{textcitehyperref}{noformat}}
-      {}%
-    \printtext[textcitehyperref]{%
-      \restorebibmacro{textcite}%
-      \usebibmacro{textcite}}}
-
-  % Remove editor form everything but books
-  \AtEveryBibitem{%
-   \ifentrytype{book}{}{%
-    \clearname{editor}
-   }
-  }
-}{
-  %-----------------------------------------------------------------------------
-  %  If using bibtex
-  %-----------------------------------------------------------------------------
-  % Load required packages and select citation style
-  \usepackage[round]{natbib}
-  \bibliographystyle{apalike}
-}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Link colors
-\definecolor{linkcolor}{HTML}{000000}
-\hypersetup{
-  colorlinks = true,
-  linkcolor  = linkcolor,
-  urlcolor   = linkcolor,
-  citecolor  = linkcolor,
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% Avoid skip when including text from external .tex file
-\newcommand{\inputy}[1]{\input{#1}\unskip}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{S\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%  ORCID
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Create the ORCID logo
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol]
-      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
-        ,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white]
-      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
-        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
-        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
-      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-
-% Command to create the ORCID
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-
-\title{\textbf{Supporting Information}\\ Dispersal and Connectivity in
-Increasingly Extreme Climatic Conditions}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1,2} \orcid{0000-0001-7477-2642} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
-
-\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
-
-\vspace{4cm}
-
-\textbf{Running Title:} Dispersal in a Changing Climate
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} climate change, conservation, dispersal, flood extremes,
-individual-based simulations, landscape connectivity, movement ecology,
-step-selection functions
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
-\newpage
-
-% Change page numbering
-\pagenumbering{arabic}
-
-% Create linenumbers
-% \linenumbers
-
-% Change to appendix counters
-\appendix
-\beginappendix
 
@@
 \section{Dispersal Model}
+\label{DispersalModel}
 The model employed to simulate dispersal was based on an integrated step
@@
 distribution for turning angles (\(-\pi, +\pi\)) and step lengths from a gamma
-distribution fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\)
+distribution fitted to observed steps (scale \(\theta\) = 6,308 and shape \(k\)
 = 0.37). It was then assumed that animals assigned to each observed and random
@@
 \begin{equation}
-\label{EQ1}
   w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
 \section{Immigration \& Emigration by Source Area}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
  \begin{center}
@@
 \section{Source-Specific Inter-Patch Connectivity}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
   source area separately across the two extreme flood-scenarios. The focal
-  source area of each subfigure is highlighted by a red circle. Subfigure (a),
-  for instance, depicts inter-patch connectivity for source area 1.}
+  source area of each subfigure is highlighted by a red circle. Subfigure a, for
+  instance, depicts inter-patch connectivity for source area 1.}
   \label{IPCMain}
@@
 \newpage
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
   \includegraphics[width = \textwidth]{Figures/IPCBuffer.png}
-  \caption{Spatial representation of egression patterns derived for each
-  source area separately across the two extreme flood-scenarios. The focal
-  source area of each subfigure is highlighted by a red circle. Subfigure (a),
-  for instance, depicts the number of individuals egressing from source area
-  1.}
+  \caption{Spatial representation of egression patterns derived for each source
+  area separately across the two extreme flood-scenarios. The focal source area
+  of each subfigure is highlighted by a red circle. Subfigure a, for instance,
+  depicts the number of individuals egressing from source area 1.}
   \label{IPCBuffer}
@@
 \section{Dispersal into Egression Zones}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
 \section{Source-Specific Intensity of Use}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
   across the two extreme flood-scenarios. The focal source area of each
-  subfigure is highlighted by a red circle. Subfigure (a), for instance, depicts
+  subfigure is highlighted by a red circle. Subfigure a, for instance, depicts
   the heatmaps for source area 1.}
-  \label{Heatmaps}
+  \label{HeatmapsInd}
   \end{center}
@@
 \section{Source-Specific Betweenness}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
   two extreme flood-scenarios. The focal source area of each subfigure is
-  highlighted by a red circle. Subfigure (a), for instance, depicts the
+  highlighted by a red circle. Subfigure a, for instance, depicts the
   betweenness maps for source area 1.}
-  \label{Betweenness}
+  \label{BetweennessInd}
   \end{center}
@@
 \section{Source-Specific Human-Wildlife Conflict}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
   across the two extreme flood-scenarios. The focal source area of each
-  subfigure is highlighted by a red circle. Subfigure (a), for instance, depicts
+  subfigure is highlighted by a red circle. Subfigure a, for instance, depicts
   the human wildlife-conflict maps for source area 1. Dotted shapes were used to
-  compare human-wildlife conflict within specific areas (see also Figure S9).}
-  \label{HWC}
+  compare human-wildlife conflict within specific areas (see also
+  \Cref{AlternativeHWC}).}
+  \label{HWCInd}
   \end{center}
@@
 \section{Changes in Potential for Human-Wildlife Interactions}
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
@@
   \caption{Number of simulated trajectories within the vicinity of
-  human-dominated landcapes in different areas of interest across the minimum
+  human-dominated landscapes in different areas of interest across the minimum
   and maximum flood scenarios. The areas are represented by the black dotted
-  shapes in Figure S8}
+  shapes in \Cref{HWCInd}}
   \label{HWCDifference}
@@
 \section{Alternative Approach to Map Potential Human-Wildlife Interactions}
+\label{AlternativeHWC}
 
-To identify potential hotspots for human-wildlife conflict and generate Figure
-6c, we isolated all simulated animal locations within 500 meters to a
-human-influenced grid-cell (Figure 6c in the main manuscript). To compute the
-needed distances, we assumed human-influence to be binary (influence = 1, no
-influence = 0), thus ignoring potential impacts of human density. We argue that
-the severity of human-wildlife conflict is not necessarily related to human
-density, yet humans' attitude towards wildlife. Attitude often correlates
+To identify potential hotspots for human-wildlife conflict and generate
+\Cref{Metrics}c, we isolated all simulated animal locations within 500 meters to
+a human-influenced grid-cell (\Cref{Metrics}c in the main manuscript). To
+compute the needed distances, we assumed human-influence to be binary (influence
+= 1, no influence = 0), thus ignoring potential impacts of human density. We
+argue that the severity of human-wildlife conflict is not necessarily related to
+human density, yet humans' attitude towards wildlife. Attitude often correlates
 negatively with habitat suitability for the species of interest
@@
 showing potential for human-wildlife conflict. Qualitatively, the maps in
-\Cref{HWCAlternative} are very similar to the ones presented in Figure 6c.
+\Cref{HWCAlternative} are very similar to the ones presented in Figure 5c.
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \begin{center}
-  \includegraphics[width = \textwidth]{Figures/HumanWildlifeConflictAlternative.png}
-  \caption{Alternative approach to quantifying the potential for human-wildlife
-  conflict. Here, we multiplied the heatmaps (a) with the human-influence layers
-  (b) to obtain human-wildlife conflict maps (c1). We also computed a difference
-  map (c2) for the layers shown in (c1).}
+  \includegraphics[width =
+  \textwidth]{Figures/HumanWildlifeConflictAlternative.png} \caption{Alternative
+  approach to quantifying the potential for human-wildlife conflict. Here, we
+  multiplied the heatmaps (a) with the human-influence layer (b) to obtain
+  human-wildlife conflict maps (c1). We also computed a difference map (c2) for
+  the layers shown in c1.}
   \label{HWCAlternative}
@@
 
+\ifSubfilesClassLoaded{%
   \newpage
-\begingroup
-\singlespacing
+  \begin{singlespacing}
   \ifthenelse{\boolean{usebiblatex}}{

### 2024-07-01 FinalThesis/Chapter_03/Figures/Covariates.tex
Publish PHD

\begin{tabular}[t]{lcccc}
\toprule
Variable & \makecell[c]{Temporal\\Resolution} & \makecell[c]{Spatial\\Resolution} & Source & \makecell[c]{Download\\Method}\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{(1) Landscape Characteristics}\\
\hspace{1em}Trees & 1 year & 250 m & MODIS MOD44B & \texttt{RGISTools}\\
\hspace{1em}Shrubs / grassland & 1 year & 250 m & MODIS MOD44B & \texttt{RGISTools}\\
\hspace{1em}NDVI & 16 days & 250 m & MODIS MOD13Q1 & \texttt{rgee}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Rivers} & \cellcolor[HTML]{f2f2f2}{static} & \cellcolor[HTML]{f2f2f2}{90 m} & \cellcolor[HTML]{f2f2f2}{MERIT Hydro} & \cellcolor[HTML]{f2f2f2}{website}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Permanent water} & \cellcolor[HTML]{f2f2f2}{static} & \cellcolor[HTML]{f2f2f2}{30 m} & \cellcolor[HTML]{f2f2f2}{Globeland30} & \cellcolor[HTML]{f2f2f2}{website}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Floodwater} & \cellcolor[HTML]{f2f2f2}{8 days} & \cellcolor[HTML]{f2f2f2}{500 m} & \cellcolor[HTML]{f2f2f2}{MOD34A4} & \cellcolor[HTML]{f2f2f2}{\texttt{floodmapr}}\\
\hspace{1em}Distance to water & 8 days & 500 m & MOD43A4 & \texttt{floodmapr}\\
\hspace{1em}Pans & 5/10 days & 10 m & Sentinel 2 & \texttt{sen2r}\\
\hspace{1em}Distance to pans & 5/10 days & 10 m & Sentinel 2 & \texttt{sen2r}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{(2) Climate Descriptors}\\
\hspace{1em}Temperature & 4 hours & 10 km & ERA5 & \texttt{rgee}\\
\hspace{1em}Precipitation & 4 hours & 10 km & JAXA GSMaP & \texttt{rgee}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{(3) Anthropogenic Features}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Human density} & \cellcolor[HTML]{f2f2f2}{static} & \cellcolor[HTML]{f2f2f2}{30 m} & \cellcolor[HTML]{f2f2f2}{Facebook} & \cellcolor[HTML]{f2f2f2}{website}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Agriculture} & \cellcolor[HTML]{f2f2f2}{static} & \cellcolor[HTML]{f2f2f2}{30 m} & \cellcolor[HTML]{f2f2f2}{Globeland30 / Cropland} & \cellcolor[HTML]{f2f2f2}{website}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Roads} & \cellcolor[HTML]{f2f2f2}{static} & \cellcolor[HTML]{f2f2f2}{vectorized} & \cellcolor[HTML]{f2f2f2}{Open Street Map} & \cellcolor[HTML]{f2f2f2}{website}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{(4) Light Availability}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Night} & \cellcolor[HTML]{f2f2f2}{4 hours} & \cellcolor[HTML]{f2f2f2}{-} & \cellcolor[HTML]{f2f2f2}{-} & \cellcolor[HTML]{f2f2f2}{\texttt{moonlit}}\\
\cellcolor[HTML]{f2f2f2}{\hspace{1em}Moon illumination} & \cellcolor[HTML]{f2f2f2}{4 hours} & \cellcolor[HTML]{f2f2f2}{-} & \cellcolor[HTML]{f2f2f2}{-} & \cellcolor[HTML]{f2f2f2}{\texttt{moonlit}}\\
\bottomrule
\end{tabular}

### 2024-07-01 FinalThesis/Chapter_03/Manuscript.tex
Publish PHD

--- 2024-07-01 Chapter_3/04_Manuscript/Manuscript.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-\usepackage[nomarkers, nolists]{endfloat}
-\usepackage{scalerel}
-\usepackage{makecell}
-\usepackage{ifthen}
-\usepackage{tikz}
-\usepackage{lineno}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%  Bibtex vs Biblatex
-%------------------------------------------------------------------------------
-% While bibtex is very nice to use, it's insanely slow. I therefore work with
-% bibtex until the final stage. Set the value below to "true" or "false" to pick
-% which one should be used.
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{false}
-
-% Depending on the question, the commands below will prepare the document
-% accordingly
-\ifthenelse{\boolean{usebiblatex}}{
-  %-----------------------------------------------------------------------------
-  %  If using biblatex
-  %-----------------------------------------------------------------------------
-  % Load required packages and refer to correct bibliography
-  \usepackage{csquotes}
-  \usepackage[
-      backend=biber
-    , style=apa          % To display Author-year in the text
-    , natbib=true        % Necessary to use citep, etc.
-    , hyperref=true      % To have clickable links in the references
-    , useeditor=false    % Don't add editor to publications
-    , sorting=ynt        % Sort by year, name, and title
-    , uniquename=false   % Avoid that initials are put
-  ]{biblatex}
-  \addbibresource{Literature.bib}
-
-  % To include author in link
-  \DeclareFieldFormat{citehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{#1}}
-  \DeclareFieldFormat{textcitehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{%
-    #1%
-    \ifbool{cbx:parens}
-      {\bibcloseparen\global\boolfalse{cbx:parens}}
-      {}}}
-  \savebibmacro{cite}
-  \savebibmacro{textcite}
-  \renewbibmacro*{cite}{%
-    \printtext[citehyperref]{%
-      \restorebibmacro{cite}%
-      \usebibmacro{cite}}}
-  \renewbibmacro*{textcite}{%
-    \ifboolexpr{
-      ( not test {\iffieldundef{prenote}} and
-        test {\ifnumequal{\value{citecount}}{1}} )
-      or
-      ( not test {\iffieldundef{postnote}} and
-        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
-    }
-      {\DeclareFieldAlias{textcitehyperref}{noformat}}
-      {}%
-    \printtext[textcitehyperref]{%
-      \restorebibmacro{textcite}%
-      \usebibmacro{textcite}}}
-
-  % Remove editor form everything but books
-  \AtEveryBibitem{%
-   \ifentrytype{book}{}{%
-    \clearname{editor}
-   }
-  }
-}{
-  %-----------------------------------------------------------------------------
-  %  If using bibtex
-  %-----------------------------------------------------------------------------
-  % Load required packages and select citation style
-  \usepackage[round]{natbib}
-  \bibliographystyle{apalike}
-}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Link colors
-\definecolor{linkcolor}{HTML}{000000}
-\hypersetup{
-  colorlinks = true,
-  linkcolor  = linkcolor,
-  urlcolor   = linkcolor,
-  citecolor  = linkcolor,
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% Avoid skip when including text from external .tex file
-\newcommand{\inputy}[1]{\input{#1}\unskip}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%  ORCID
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Create the ORCID logo
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol]
-      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
-        ,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white]
-      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
-        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
-        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
-      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-
-% Command to create the ORCID
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-\title{The Effects of Increasing Seasonal Dynamism when Predicting Connectivity:
-Advantages or Unnecessary Complications?}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1, 2} \orcid{0000-0001-7477-2642} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
-
-\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
-
-\vspace{4cm}
-
-\textbf{Running Title:} Connectivity in Seasonal Landscapes
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} African wild dog, Connectivity, Dispersal, Dynamism,
-Individual-based, Lycaon pictus, Okavango Delta, Seasonality
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Abstract
-%------------------------------------------------------------------------------
+\chapter{The Effects of Increasing Seasonal Dynamism when Predicting
+Connectivity: Advantages or Unnecessary Complications?}
+\label{ChapterSeasonality}
+\thispagestyle{empty}
+\vspace{-1cm}
+\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil
+
+\begin{center}
+David D. Hofmann \orcid{0000-0003-3477-4365},
+Dominik M. Behr \orcid{0000-0001-7378-8538},
+John W. McNutt,
+Arpat Ozgul \orcid{0000-0001-7477-2642}, and
+Gabriele Cozzi \orcid{0000-0002-1744-1940}
+
+% \begin{flushleft}
+
+% \vspace{5cm}
+\vfill
+
+% \textsuperscript{1} Department of Evolutionary Biology and Environmental
+% Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
+% Switzerland.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+% Private Bag 13, Maun, Botswana.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{\S} Corresponding author:
+% \href{mailto:david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
+
+\end{center}
+
+% \vfill
+
+% \textbf{Running Title:} Connectivity in Seasonal Landscapes
+%
+% \vspace{0.5cm}
+%
+% \textbf{Keywords:} African wild dog, Connectivity, Dispersal, Dynamism,
+% Individual-based, Lycaon pictus, Okavango Delta, Seasonality
+
+% \end{flushleft}
+
 \newpage
-\begin{abstract}
+\section*{Abstract}
 
@@
 important ecological features and processes, thus resulting in poor agreement
-between predicted and observed movements and a misrepresentation of
+between predicted and observed movements and therefore a misrepresentation of
 connectivity.
@@
 which a better representation of seasonal dynamism improves our ability to
-predict connectivity. For this, we fit integrated step-selection functions and
-predict connectivity using an individual-based dispersal simulation while
-explicitly considering seasonal dynamism in both environmental covariates and
-the species' preferences.  Using a rigorous cross-validation procedure, we
-compare the predictive model performance under each of the six proposed
-configurations. While we expected that an increasing degree of seasonal dynamism
-would lead to improved predictions, we were particularly interested in
-identifying at which stage the inclusion of seasonality provides the biggest
-benefits.
-
-We show that, for our study system, improvements in predictive performance by
-incorporating seasonal dynamism were moderate. In fact, incorporating
-seasonality only improved predictions when an overly simplistic movement model
-was assumed. Upon fitting a more complex model, the benefits of accounting for
-seasonality vanished, resulting in imperceptible performance differences.
-Despite this, patterns of connectivity as obtained from dispersal simulations
-revealed marked differences between the most static and most dynamic
+predict dispersal and connectivity. For this, we fit integrated step-selection
+functions and predict connectivity using an individual-based dispersal
+simulation while explicitly considering seasonal dynamism in both environmental
+covariates and the species' preferences.  Using a rigorous cross-validation
+procedure, we compare predictive model performances under each of the six
+proposed configurations. While we expected that an increasing degree of seasonal
+dynamism would lead to improved predictions, we were particularly interested in
+identifying at which stage the inclusion of seasonality would provide the
+biggest benefits.
+
+We show that, for our study system, improvements in predictive performances
+achieved by incorporating seasonal dynamism were moderate. In fact, accounting
+for seasonality only improved predictions when an overly simplistic model
+formula assumed. Upon using a more complex model formula, the benefits of
+accounting for seasonality vanished, resulting in imperceptible performance
+differences. Despite this, patterns of connectivity as obtained from dispersal
+simulations revealed marked differences between the most static and most dynamic
 configurations. Most notably, connectivity was more homogeneously distributed
@@
 
-\end{abstract}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
 \newpage
-
-\onehalfspacing
-\tableofcontents
-\doublespacing
-
-% Change page numbering
-\newpage
-\pagenumbering{arabic}
-
-% Create linenumbers
-\linenumbers
 \section{Introduction}
-\subsection{Connectivity}
+
+% \subsection{Connectivity}
 
@@
 facilitates or impedes movement among habitat patches \citep{Taylor.1993} and is
-a critical prerequisite for maintaining biodiversity \citep{Fahrig.2003}.
+a central prerequisite for maintaining biodiversity \citep{Fahrig.2003}.
 Improved connectivity facilitates dispersal \citep{Doerr.2011, Baguette.2013},
@@
 Quantifying connectivity and identifying critical dispersal corridors have
-therefore become critical tasks in conservation science \citep{Heller.2009,
+therefore become crucial tasks in conservation science \citep{Heller.2009,
 Rudnick.2012, Keeley.2019, Hofmann.2021}. One aspect that has received limited
 attention in the study of connectivity is the role of seasonality. Yet, seasonal
-changes both in the environment and in an animal's propensity to use or avoid
-particular habitat types can profoundly impact landscape connectivity
-\citep{Zeller.2020a}.
-
-\subsection{Seasonality}
+changes in the environment and an animal's willingness to traverse a particular
+habitat type can profoundly impact landscape connectivity \citep{Zeller.2020a}.
+
+% \subsection{Seasonality}
 
@@
 across massive spatial scales, resulting in short-lived movement corridors
-between two otherwise disconnected habitats \citep{Serneels.2001, Naidoo.2016}.
+between otherwise disconnected habitats \citep{Serneels.2001, Naidoo.2016}.
 Alternatively, seasonality can affect functional connectivity via temporal
-changes in a species' movement and habitat preferences. Amphibians, for
-instance, require both aquatic and terrestrial habitats, but their preference
-for one over the other heavily depends on the season \citep{Baldwin.2006}.
-Although such seasonal intricacies are likely to play a fundamental role in many
-ecosystems, they only rarely enter connectivity studies in an explicit manner.
-In fact, most connectivity studies represent their study system by a static set
-of environmental layers and assume that their focal species exhibits a fixed set
+changes in a species' habitat preferences. Amphibians, for instance, require
+both aquatic and terrestrial habitats, but their preference for one habitat over
+the other heavily depends on the season \citep{Baldwin.2006}. Although such
+seasonal intricacies are likely to play a fundamental role in many ecosystems,
+they only rarely enter connectivity studies in an explicit manner. In fact, most
+connectivity studies represent their study system by a static set of
+environmental layers and assume that their focal species to exhibit a fixed set
 of preferences (e.g., \citealp{Elliot.2014, Abrahms.2017, Brennan.2020}).
@@
 
-\subsection{Modeling Functional Connectivity}
+% \subsection{Modeling Functional Connectivity}
 
@@
 of connectivity are compiled. Second, these data are combined and fed into a
-selection model that enables estimating a species selection or avoidance of
+selection model that enables estimating the focal species' preferences towards
 environmental features. Popular frameworks for estimating preferences are
-case-control designs, where characteristics extracted at observed locations are
-contrasted with characteristics extracted at available locations
-\citep{Beyer.2010, Fieberg.2010}. This can be achieved using point-selection
-functions \citep{Boyce.2002, Manly.2007}, path-selection functions
-\citep{Cushman.2010}, and step-selection functions \citep{Fortin.2005,
-Thurfjell.2014}. A particularly powerful approach is that of \textit{integrated}
-step-selection functions (iSSFs), as it provides an effective means to model
-both an animal's habitat-selection and movement capacity \citep{Avgar.2016,
-Fieberg.2021}. Third, inferred preferences are used to predict a permeability
-surface, which indicates the expected ease or difficulty at which the focal
-species can traverse a certain area given the area's environmental
-characteristics \citep{Zeller.2012}. Finally, in a fourth step, the permeability
-surface serves as an input to a connectivity model that reveals crucial movement
-corridors. At present, the most popular connectivity models are least-cost path
-analysis \citep{Adriaensen.2003} and circuit theory \citep{McRae.2008}, although
-individual-based movement models (IBMMs) have gained some momentum recently
-\citep{Kanagaraj.2013, Allen.2016, Hauenstein.2019, Zeller.2020,
-UnnithanKumar.2022a, UnnithanKumar.2022, Hofmann.2023}. In particular, some
-IBMMs allow estimating connectivity directly via simulated dispersal
-trajectories, thus bypassing the generation of a permeability surface
+case-control designs, where covariate values extracted at observed locations are
+contrasted with those extracted at available locations \citep{Beyer.2010,
+Fieberg.2010}. This can be achieved using point-selection functions
+\citep{Boyce.2002, Manly.2007}, path-selection functions \citep{Cushman.2010},
+and step-selection functions \citep{Fortin.2005, Thurfjell.2014}. A particularly
+powerful approach is that of \textit{integrated} step-selection functions
+(iSSFs), which provides an effective means to jointly model an animal's movement
+kernel (i.e., movement capacity or movement preferences) and habitat-selection
+\citep{Avgar.2016, Fieberg.2021}. Third, inferred preferences are used to
+predict a permeability surface, which is a spatial layer that indicates the
+expected ease or difficulty at which the focal species can traverse a certain
+area given the area's environmental characteristics \citep{Zeller.2012}.
+Finally, in a fourth step, the permeability surface serves as an input to a
+connectivity model that reveals crucial movement corridors. At present, the most
+popular connectivity models are least-cost path analysis \citep{Adriaensen.2003}
+and circuit theory \citep{McRae.2008}, although individual-based movement models
+(IBMMs) have recently gained some momentum too \citep{Kanagaraj.2013,
+Allen.2016, Hauenstein.2019, Zeller.2020, UnnithanKumar.2022a,
+UnnithanKumar.2022, Hofmann.2023}. A major benefit of using IBMMs is that they
+allow estimating connectivity directly via simulated dispersal trajectories,
+thus bypassing the generation of a permeability surface
 \citep{UnnithanKumar.2022a, Hofmann.2023}.
 
-\subsection{Incorporating Seasonality in Functional Connectivity Models}
-
-Seasonality can enter the connectivity modeling workflow described above in
-three distinct stages (\Cref{GraphicalAbstract}). In the first stage, one can
-either extract environmental features, such as water and vegetation, from a
-static layer (i.e., a single snapshot) or from a time series of layers (i.e., a
-sequence of snapshots) that capture seasonal variation across the landscape. The
-former approach has historically been the norm (e.g., \citealp{Elliot.2014,
+% \subsection{Incorporating Seasonality in Functional Connectivity Models}
+
+Seasonality can enter the connectivity modeling workflow described above at
+three distinct stages (\Cref{GraphicalAbstractCH3}). At the first stage, one can
+either extract covariate values on environmental features, such as water and
+vegetation, from static layers (i.e., a single snapshot per covariate) or from
+stacks of layers (i.e., a sequence of snapshots per covariate) that capture
+seasonal variation across the landscape (\Cref{GraphicalAbstractCH3}, Stage 1).
+The former approach has historically been the norm (e.g., \citealp{Elliot.2014,
 Brennan.2020}), yet advances in remote sensing technologies and a facilitated
@@
 covariate layers has become more frequent (e.g., \citealp{Osipova.2019,
-Kaszta.2021}). In a second stage, one can assume their focal species to exhibit
-a fixed set of preferences across seasons by pooling all presence data, or can
-try to account for seasonal changes in preferences by splitting the data
-accordingly (e.g., \citealp{Fortin.2005, Manly.2007, Cushman.2010,
-Zeller.2020}). \cite{Chetkiewicz.2009}, for instance, partitioned their data by
-season to derive seasonal habitat preferences for pumas (\textit{Puma concolor})
-and grizzlies (\textit{Ursus arctos}). In a final stage, one can estimate
-connectivity for either ``average'' environmental conditions, or utilize
-seasonally updated layers to estimate connectivity for distinct seasons. For
-connectivity models that rely on permeability surfaces, this implies repeatedly
-applying the connectivity model using seasonally updated resistance surfaces
-(e.g. \citealp{Osipova.2019, Zeller.2020, Kaszta.2021, Ciudad.2021}). This is
+Kaszta.2021}). At the second stage, one can assume their focal species to
+exhibit a fixed set of preferences across seasons by pooling all presence data,
+or can try to account for seasonal changes in preferences by splitting the data
+accordingly (\Cref{GraphicalAbstractCH3}, Stage 2; e.g., \citealp{Fortin.2005,
+Manly.2007, Cushman.2010, Zeller.2020}). \citet{Chetkiewicz.2009}, for instance,
+partitioned their data by season to derive seasonal habitat preferences for
+pumas (\textit{Puma concolor}) and grizzlies (\textit{Ursus arctos}). At a third
+stage, one can estimate connectivity for either ``average'' environmental
+conditions, or utilize seasonally updated layers to estimate connectivity for
+distinct seasons (\Cref{GraphicalAbstractCH3}, Stage 3). For connectivity models
+that rely on permeability surfaces, the latter case implies repeatedly applying
+the connectivity model using seasonally updated permeability surfaces (e.g.
+\citealp{Osipova.2019, Zeller.2020, Kaszta.2021, Ciudad.2021}). This is
 demonstrated \textit{ad extremum} by \citet{Kaszta.2021}, who prepared monthly
-updated permeability surfaces for African elephants (\textit{Loxodonta
-africana}). The above-mentioned IBMMs allow for an elegant solution, for
-seasonality can be accounted for as simulated individuals move. Such an approach
-has, however, not yet been followed. Irrespective of the method used, including
-seasonality is analytically and computationally  demanding
+updated permeability surfaces and repeatedly employed circuit theory to estimate
+dynamic connectivity for African elephants (\textit{Loxodonta africana}). IBMMs,
+by contrast, allow for a more elegant solution, as seasonality can be modeled as
+the simulated dispersers move. Such an approach has, however, not yet been
+followed. Irrespective of the method used, including seasonality into
+connectivity analyses is analytically and computationally demanding
 \citep{Bishop-Taylor.2018}, raising the question to what degree it should be
-considered. Our goal is (1) to create a framework with the possible combinations
-in which seasonality can be accounted for, (2) to establish whether
-incorporating seasonality benefits the predictive performance of connectivity
-analyses, and (3) to pinpoint at which stages the inclusion of seasonality
-provides the largest benefit.
-
-\subsection{African wild dogs in Northern Botswana}
-
-A system well-suited to study the importance of seasonal dynamism for dispersal
-and connectivity is the African wild dog population (\textit{Lycaon pictus})
-inhabiting the seasonally highly variable Okavango Delta ecosystem in northern
-Botswana \citep{McNutt.1996, Wolski.2017}. While once present across the entire
-Sub-Saharan continent, the African wild dog has disappeared from a majority of
-its historic range due to human persecution, deadly diseases, and habitat
-destruction \citep{Woodroffe.2020a}. With about 6,000 adult individuals
+considered.
+
+% \subsection{African wild dogs in Northern Botswana}
+
+To study the importance of seasonal dynamism for dispersal and connectivity, we
+use an African wild dog population (\textit{Lycaon pictus}) inhabiting the
+seasonally highly variable Okavango Delta ecosystem in northern Botswana as a
+study system \citep{McNutt.1996, Wolski.2017}. While once present across the
+entire Sub-Saharan continent, the African wild dog has disappeared from a
+majority of its historic range due to human persecution, deadly diseases, and
+habitat destruction \citep{Woodroffe.2020a}. With about 6,000 adult individuals
 remaining in the wild, the species is considered as endangered on the IUCN red
@@
 Sandoval-Seres.2022}. Studies of habitat-selection during dispersal show that
-wild dogs avoid water, prefer moving along it, prefer moving across open grass
-or shrubs, yet avoid areas dominated by humans and densely covered by forests
-\citep{ONeill.2020, Hofmann.2021}.
-
-\begin{figure}
+wild dogs avoid water, prefer moving along it, prefer moving across open
+grassland or shrubs, yet avoid areas dominated by humans and densely covered by
+forests \citep{ONeill.2020, Hofmann.2021}. Given the species' considerable
+mobility and the study area's significant seasonal variability, we deemed the
+study-system well-suited to examine the importance of seasonal dynamism on
+dispersal and connectivity.
+
+\begin{figure}[htpb]
  \begin{center}
   \includegraphics[width = \textwidth]{Figures/GraphicalAbstract.pdf}
-  \caption{Overview of the different dimensions in which seasonality can be
-  rendered in studies of landscape connectivity. (a) During model fitting, the
-  modeler needs to decide whether to represent the environment by a static set
-  of covariate layers, thus ignoring seasonality, or whether to obtain a dynamic
-  set covariate layers that allow rendering it. One also needs to decide
-  whether to parameterize a single-season model, assuming fixed preferences
-  across the year, or to engage in a multi-season model that accounts for
-  seasonal differences. (b) When utilizing the fitted model to predict
-  connectivity, one can either assume a static set of environmental covariates
-  or again attempt a seasonal take that renders how connectivity differs
-  depending on the season. (c) Depending on these decisions, six different
-  combinations with differing degrees of dynamism emerge. Our hypothesis was
-  that increasing the degree of dynamism would lead to improved predictions.
-  However, we were particularly interested in determining at which stage the
+  \caption{Overview of the different stages at which seasonality can be rendered
+  in studies of dispersal and landscape connectivity. (a) During model fitting,
+  one needs to decide whether to represent the environment by a static set of
+  covariate layers, thus ignoring seasonality, or to obtain a dynamic set
+  covariate layers that allow showing how the landscape changes over time. One
+  also needs to decide whether to parameterize a single-season model, assuming
+  fixed preferences across the year, or to engage in a multi-season model that
+  accounts for seasonal differences. (b) When utilizing the fitted model to
+  predict connectivity, one can either assume a static set of environmental
+  covariates or again attempt a seasonal take that renders how connectivity
+  differs depending on the season. (c) Depending on these decisions, six
+  different combinations that differ in terms of the degree of seasonality they
+  encapsulate emerge. Our hypothesis was that increasing the degree of dynamism
+  when predicting dispersal and connectivity would lead to improved predictions.
+  Notably, we were particularly interested in determining at which stage the
   inclusion of seasonally provided the biggest benefits.}
-  \label{GraphicalAbstract}
+  \label{GraphicalAbstractCH3}
  \end{center}
@@
 
-\subsection{What We Did}
-
-Here, we examined the importance of including seasonally-varying factors to
-assess landscape connectivity for the African wild dog. For this, we compared
-the predictive performance of iSSFs that differ in terms of the dynamism they
-represented. Specifically, we investigate if and to what degree seasonality at
-different stages of the connectivity modeling workflow contributes to improved
-predictions. For this, we compile an extensive collection of remote sensed time
-series data that accurately render seasonality across the Okavango Delta
-ecosystem. We combine them with multi-year dispersal data from
-\inputy{GeneralMetrics/CollarsTotal} dispersing wild dog coalitions in northern
-Botswana and apply k-fold cross-validation for case-control studies to compare
-the predictive efficacy of each model. Finally, we employ IBMMs to simulate
-dispersal and estimate connectivity resulting from varying degrees of dynamism.
-We hypothesized that habitat-selection and movement behavior would differ
-significantly between seasons and that increasing the degree of dynamism would
-result in better agreement between observed and predicted dispersal patterns
-(\Cref{GraphicalAbstract}).
+Our goal with this article is to (1) create a framework highlighting the
+possible avenues through which seasonality can be incorporated into connectivity
+analyses, (2) establish whether incorporating seasonality benefits the
+predictive performance of dispersal models, and (3) pinpoint at which stages the
+inclusion of seasonality provides the largest benefit. For this, we compare the
+predictive performances of six configurations that differ in terms of the
+dynamism they represent (\Cref{GraphicalAbstractCH3}b). Specifically, we
+investigate if and to what degree seasonality at different stages of the
+connectivity modeling workflow contributes to improved predictions. We compile
+an extensive collection of remote sensed spatial layers that accurately render
+seasonality across the Okavango Delta ecosystem. We combine them with multi-year
+dispersal data from \inputy{GeneralMetrics/CollarsTotal} dispersing wild dog
+coalitions in northern Botswana and apply k-fold cross-validation for
+case-control studies to compare the predictive efficacy under each configuration
+(\Cref{GraphicalAbstractCH3}b). Finally, we employ IBMMs to simulate dispersal
+and estimate connectivity resulting from varying degrees of dynamism. To
+distinguish between different configurations, we use codes that highlight the
+level of dynamism represented by each configuration (e.g., SSS =
+Static-SingleSeason-Static and DMD = Dynamic-MultiSeason-Dynamic). We
+hypothesized that habitat-selection and movement behavior of dispersing wild
+dogs would differ significantly between seasons and that increasing the degree
+of dynamism would result in better agreement between observed and predicted
+dispersal patterns (\Cref{GraphicalAbstractCH3}c).
 
@@
 
-The study area comprised the Okavango Delta ecosystem, mainly within northern
-Botswana (centered at \inputy{GeneralMetrics/StudyAreaCenter.tex} at an
-elevation of approx. 950 m) but also extended to parts of Namibia and Zimbabwe,
-and stretched across an extent of \inputy{GeneralMetrics/SizeStudyArea.tex}
-km\textsuperscript{2} (\Cref{StudyArea}). The Okavango Delta is a flood-pulse
-driven mosaic of patchy woodlands, permanent swamps, and seasonally flooded
-grasslands that lie within the otherwise dry and sandy Kalahari Basin
-\citep{Wilson.1976, Ramberg.2006, Mendelsohn.2010}. Precipitation across the
-study area varies considerably between seasons, ranging from
+The study area comprised the Okavango Delta ecosystem in northern Botswana
+(centered at \inputy{GeneralMetrics/StudyAreaCenter.tex} at an elevation of
+approx. 950 m) and extended to parts of Namibia and Zimbabwe, encompassing an
+extent of \inputy{GeneralMetrics/SizeStudyArea.tex} km\textsuperscript{2}
+(\Cref{StudyAreaCH3}). The Okavango Delta is a flood-pulse driven mosaic of
+patchy woodlands, permanent swamps, and seasonally flooded grasslands that lie
+within the otherwise dry and sandy Kalahari Basin \citep{Wilson.1976,
+Ramberg.2006, Mendelsohn.2010}. Precipitation across the study area varies
+considerably between seasons, ranging from
 \inputy{GeneralMetrics/MinimumPrecipitation.tex} mm during the dry season (from
-$\sim$ 15 April to 15 October) to
-\inputy{GeneralMetrics/MaximumPrecipitation.tex} mm during the wet season (from
-$\sim$ 15 October to 15 April), totaling to
+15 April to 15 October) to \inputy{GeneralMetrics/MaximumPrecipitation.tex} mm
+during the wet season (from 15 October to 15 April), totaling to
 \inputy{GeneralMetrics/TotalPrecipitation.tex} mm across an average year
@@
 \inputy{GeneralMetrics/MinimumTemperature.tex}\degree C during the dry winter
-months to \inputy{GeneralMetrics/MaximumTemperature.tex}\degree C during the wet
-summer months (\Cref{Seasonality}b). The vegetation in the study area is mainly
-composed of mopane forest (\textit{Colophospermum mopane}), mixed woodland
-acacia-dominated (\textit{Acacia spp.}), and grassland. Substantial vegetation
-green-up (e.g., plant and shoots growth, leaf production, grass greening) after
-the dry season starts with a delay of some weeks after the onset of the first
-rains in the wet season. The normalized difference vegetation index (NDVI)
-therefore depicts a lagged response to precipitation patterns across the study
-area (\Cref{Seasonality}c). The yearly flood-cycle of the Delta is predominantly
-driven by rainfalls in the Angolan highlands, where water is collected and
-channeled through the Cubango and Cuito rivers into the Okavango Delta
-\citep{McCarthy.2003, Gumbricht.2004, Mendelsohn.2010}. Because water only
-slowly descends from the catchment areas in Angola into the Delta's tributaries,
-the flood is out of sync with local rainfalls and reaches its maximum extent
-during August-September, i.e. during peak dry season (\citealp{Wolski.2017},
-\Cref{Seasonality}d). While the extent of large-bodied rivers and floodplains is
-determined by precipitation in Angola, the emergence of smaller, ephemeral
-water-bodies (a.k.a., pans) is dictated by local precipitation during the wet
-season. \inputy{GeneralMetrics/PercentageProtected.tex}\% of the landscape in
-the study area form part of a protected area, such that human impact remains low
-and largely limited to settlements along the western part of the delta and the
-city of Maun at the delta's southern tip (\Cref{StudyArea}). Landscapes outside
-protected areas in Zimbabwe, however, are more heavily influenced by humans,
-mainly through agricultural fields and human settlements.
-
-\begin{figure}
+months and \inputy{GeneralMetrics/MaximumTemperature.tex}\degree C during the
+wet summer months (\Cref{Seasonality}b). The vegetation in the study area is
+mainly composed of mopane forest (\textit{Colophospermum mopane}), mixed
+woodland acacia-dominated (\textit{Acacia spp.}), and grassland. Substantial
+vegetation green-up (e.g., plant and shoots growth, leaf production, and grass
+greening) after the dry season starts with a delay of some weeks after the onset
+of the first rains in the wet season. The normalized difference vegetation index
+(NDVI) therefore depicts a smoothed and lagged response to precipitation
+patterns across the study area (\Cref{Seasonality}c). The yearly flood-cycle of
+the Okavango Delta is predominantly driven by rainfalls in the Angolan
+highlands, where water is collected and channeled through the Cubango and Cuito
+rivers into the Okavango Delta \citep{McCarthy.2003, Gumbricht.2004,
+Mendelsohn.2010}. Because water only slowly descends from the catchment areas in
+Angola into the Delta's tributaries, the flood is out of sync with local
+rainfalls and reaches its maximum extent during August-September, i.e. during
+peak dry season (\citealp{Wolski.2017}, \Cref{Seasonality}d). While the extent
+of large-bodied rivers and floodplains is determined by precipitation in Angola,
+the emergence of smaller, ephemeral water-bodies (i.e., pans) is dictated by
+local precipitation during the wet season.
+\inputy{GeneralMetrics/PercentageProtected.tex}\% of the landscape in the study
+area form part of a protected area, such that human impact remains low and
+largely limited to settlements along the western part of the Okavango Delta and
+the city of Maun at the Delta's southern tip (\Cref{StudyAreaCH3}). Landscapes
+outside protected areas in Zimbabwe, however, are more heavily influenced by
+humans, mainly through agricultural fields and human settlements.
+
+\begin{figure}[htpb]
  \begin{center}
   \includegraphics[width = \textwidth]{Figures/StudyArea.png} \caption{(a) Study
-  area from which data on dispersing wild dogs were collected. Dispersal
+  area from which data on dispersing African wild dogs were collected. Dispersal
   trajectories are plotted in dark gray. The study area encompassed parts of the
@@
   peak rainy season (b2). Notably, the flood of the Okavango Delta reaches its
-  maximum extent during peak dry season (August–September).}
-  \label{StudyArea}
+  maximum extent during peak dry season between August and September.}
+  \label{StudyAreaCH3}
  \end{center}
@@
 
-\begin{figure}
+\begin{figure}[htpb]
 \begin{center}
   \includegraphics[width = \textwidth]{Figures/SeasonalCovariates.png}
-  \caption{Illustration of how some covariates considered in this study vary
-  across seasons. The wet season spans mid-October to mid-April (shaded in
-  gray). Data for the graphs were obtained from \textbf{(a)} JAXA GSMaP,
-  \textbf{(b)} ERA5, \textbf{(c)} MODIS MOD13Q1, and \textbf{(d)} remote sensed
-  MOD43A4 satellite images. Smoothing curves were fitted using GAMs as
+  \caption{Illustration of how some of the covariates considered in this study
+  vary across seasons. The wet season spans mid-October to mid-April (shaded in
+  gray), the remainder is considered as dry season. Data for the graphs were
+  obtained from (a) JAXA GSMaP, (b) ERA5, (c) MODIS MOD13Q1, and (d) remote
+  sensed MOD43A4 satellite images. Smoothing curves were fitted using GAMs as
   implemented in the \texttt{mgcv} \texttt{R}-package \citep{Wood.2011}.}
@@
 Between \inputy{GeneralMetrics/GPSFromYear} and
-\inputy{GeneralMetrics/GPSToYear}, we collected GPS data of
+\inputy{GeneralMetrics/GPSToYear}, we collected GPS data on
 \inputy{GeneralMetrics/CollarsTotal} dispersing wild dog coalitions
 (\inputy{GeneralMetrics/CollarsFemales} female coalitions,
-\inputy{GeneralMetrics/CollarsMales} male coalitions \Cref{StudyArea}b). We
-programmed GPS satellite collars to record GPS locations at 17:00, 21:00, 01:00,
-05:00, and 09:00 o'clock. The eight-hour window between 09:00 and 17:00 can be
+\inputy{GeneralMetrics/CollarsMales} male coalitions; \Cref{StudyAreaCH3}b). We
+programmed GPS satellite collars to record GPS locations at 01:00, 05:00, 09:00,
+17:00, and 21:00. The eight-hour window between 09:00 and 17:00 can be
 considered comparable to a four-hour window, as wild dogs generally rest between
 11:00 and 15:00, leaving approximately four hours of activity
-\citep{Hayward.2009}. The collected GPS data was regularly transmitted to a
+\citep{Hayward.2009}. The collected GPS data were regularly transmitted to a
 base-station via Iridium satellite, thereby allowing dispersing individuals to
@@
 (1) landscape characteristics, (2) climatic conditions, and (3) anthropogenic
-factors (\Cref{Covariates}) (see \citealp{Hofmann.2021, Hofmann.2023}). Besides
-spatial covariates, we also prepared a series of covariates relating to (4)
-light availability (\Cref{Covariates}) (see Cozzi et al. 2012). To appropriately
-render seasonality in each of the spatial covariates (1-3), we downloaded them
-at the highest spatial and temporal resolution available. That is, we obtained
-for each spatial covariates by a stack of raster layers that spanned the entire
-range of dates and locations for which we also collected GPS data of dispersing
+features (\Cref{CovariatesCH3}; see \citealp{Hofmann.2021, Hofmann.2023}).
+Besides spatial covariates, we also prepared a series of covariates relating to
+(4) light availability (\Cref{CovariatesCH3}; see Cozzi et al. 2012). To
+appropriately render seasonality in each of the spatial covariates (1-3), we
+downloaded them at the highest spatial and temporal resolutions available. That
+is, we obtained for each spatial covariate a stack of raster layers that spanned
+the entire range of dates for which we also collected GPS data on dispersing
 wild dogs. Depending on the covariate, this resulted in differing spatial and
-temporal resolutions (\Cref{Covariates}). Additionally, for each spatial
+temporal resolutions (\Cref{CovariatesCH3}). Additionally, for each spatial
 covariate, we also generated a static layer, representing ``average'' conditions
@@
 seasonality by flattening across the entire range of dates, we flattened stacks
-across years, thereby eliminating year-specific effects. These final layers
-served to be used in the dispersal simulation, where we needed data representing
-seasonality in all covariates across a typical year. To summarize, we prepared
-each spatial covariate dynamically for the entire range of dates considered,
-dynamically for an average year, and statically across the entire range of dates
-considered.
+across years, thereby eliminating year-specific effects. These latter layers
+were later used in the dispersal simulation, where we needed spatial layers that
+depicted seasonal dynamism across a typical year. In summary, we prepared each
+covariate dynamically and statically for the entire range of dates considered,
+and dynamically for an average year.
 
@@
     \caption{Covariates that were used in this study, including information on
-    their temporal and spatial resolutions, as well as on the avenue through which
-    the respective data were accessed or downloaded.}
-    \label{Covariates}
+    their temporal and spatial resolutions, as well as on the avenue through
+    which the respective data were accessed or downloaded. Download methods
+    printed in \texttt{typewriter} font indicate the use of an
+    \texttt{R}-package that is available through CRAN or GitHub.}
+    \label{CovariatesCH3}
    \begin{threeparttable}
@@
       \item \textit{Note:} The covariates in gray were combined into proxies for
-      water, human influence, and brightness, respectively. The detailed
-      aggregation procedure is described by \citet{Hofmann.2021}.
+      water, human influence, and brightness, respectively. Detailed
+      aggregation procedures are provided in \citet{Hofmann.2021}.
     \end{tablenotes}
@@
 \citep{Perez-Goya.2020} to download yearly updated layers, each at a resolution
-of 250m x 250m. We also obtained information on the normalized vegetation
-difference index (NDVI) through the MODIS MOD13Q1 dataset \citep{Didan.2015},
-which also has a resolution of 250m x 250m. This product is updated every 16
-days, and we accessed the respective data through Google Earth Engine
-\citep{Gorelick.2017} using the \texttt{R}-package \texttt{rgee}
-\citep{Aybar.2024}. To depict large, permanent water-bodies, we employed the
-Globeland30 dataset, from which we only retained the land-cover class water,
-while setting all other categories to dryland. Similarly, we used the MERIT
-Hydro dataset to obtain information on permanent rivers \citep{Yamazaki.2019}.
-To dynamically render large water-bodies, particularly the floodwaters of the
-Okavango Delta, we prepared weekly updated floodmaps using remote sensed MODIS
-MOD43A4 satellite images. The underlying floodmapping algorithm is described in
-detail in \citet{Wolski.2017} and \citet{Hofmann.2021} and is implemented in the
-\texttt{floodmapr} package (available on GitHub;
-\url{https://github.com/DavidDHofmann/floodmapr}). To generate a single layer
-representing major water bodies, we combined the water, river, and flood layers
-into a single stack with a resolution of 500m x 500m. Finally, we employed
-remote sensing to detect small, ephemeral water bodies (i.e., pans) using a
-custom random-forest classifier applied to Sentinel 2 satellite imagery
-(\citealp{EuropeanSpaceAgency.2018}; details of the classifier in Appendix A1).
-Sentinel 2 has a resolution of 10m and is therefore particularly useful for
-obtaining information on small landscape features. Even though Sentinel 2
-satellite imagery is updated every 5 days, cloud cover often prohibited the
-computation of a ``pan-map''. Consequently, we settled for monthly updated
-composite images, which effectively alleviated problems due to cloud cover. In
-summary, we produced one stack of layers representing major water bodies, and
-another stack of layers representing ephemeral water bodies. For both, we also
-computed corresponding distance-to stacks.
+of 250m x 250m. We also obtained information on the NDVI through the MODIS
+MOD13Q1 dataset \citep{Didan.2015}, which also has a resolution of 250m x 250m.
+This product is updated every 16 days, and we accessed the respective data
+through Google Earth Engine \citep{Gorelick.2017} using the \texttt{R}-package
+\texttt{rgee} \citep{Aybar.2024}. To depict large, permanent water-bodies, we
+employed the Globeland30 dataset, from which we only retained the land-cover
+class water, while setting all other categories to dryland. Similarly, we used
+the MERIT Hydro dataset to obtain information on permanent rivers
+\citep{Yamazaki.2019}. To dynamically render large water-bodies, particularly
+the floodwaters of the Okavango Delta, we prepared weekly updated floodmaps
+using remote sensed MODIS MOD43A4 satellite images. The underlying floodmapping
+algorithm is described by detail by \citet{Wolski.2017} and \citet{Hofmann.2021}
+and is implemented in the \texttt{floodmapr} package (available on GitHub;
+\url{https://github.com/DavidDHofmann/floodmapr}). We combined the water
+(static), river (static), and flood layers (dynamic) into a single stack with a
+resolution of 500m x 500m (\Cref{CovariatesCH3}). Finally, we employed remote
+sensing to detect small, ephemeral water bodies (i.e., pans) using a custom
+random-forest classifier applied to Sentinel 2 satellite imagery
+(\citealp{EuropeanSpaceAgency.2018}; details of the classifier in
+\Cref{PanMappingApproach}). Sentinel 2 has a resolution of 10m and is therefore
+particularly useful for obtaining information on small landscape features. Even
+though Sentinel 2 satellite imagery is updated every 5 days, cloud cover often
+prohibited the computation of a ``pan-map''. Consequently, we settled for
+monthly updated composite images, which effectively alleviated problems due to
+cloud cover. In summary, we produced one stack of layers representing major
+water bodies, and another stack of layers representing ephemeral water bodies
+(pans). For both, we also computed corresponding distance-to stacks, indicating
+the distance to major and ephemeral water bodies, respectively.
 
@@
 Google Earth Engine \citep{Gorelick.2017} using the \texttt{rgee} package
-\citep{Aybar.2024} and had a resolution of 10km x 10km. To match hourly
-temperature and precipitation values with the four-hourly data collected by the
-GPS satellite collars fitted on the dispersing wild dogs, we computed average
-precipitation and temperature values over four hourly periods that matched the
-GPS collection schedule.
+\citep{Aybar.2024} and had a resolution of 10km x 10km. To match one-hourly
+temperature and precipitation estimates with the four-hourly GPS data on
+dispersing wild dogs, we computed average precipitation and temperature values
+over four hourly periods that matched the GPS collection schedule.
 
@@
 dataset \citep{Tiecke.2017}, which we downloaded from the humdata website
-(\url{www.data.humdata.org/}). We obtained information on the presence of
+(\url{www.data.humdata.org}). We obtained information on the presence of
 agricultural fields from the Globeland30 \citep{Chen.2015} and Cropland
@@
 from OpenStreetMaps \citep{OpenStreetMapContributors.2017}. Ultimately, we
-merged all anthropogenic features into a single layer (details in
-\citealp{Hofmann.2021}) that had a resolution of 250m x 250m.
+merged all anthropogenic features into a single layer that had a resolution of
+250m x 250m (details are provided by \citealp{Hofmann.2021}) .
 
@@
 
-We computed light statistics using the \texttt{suncalc} and \texttt{moonlit}
+We computed light availability using the \texttt{suncalc} and \texttt{moonlit}
 \texttt{R}-packages \citep{Thieurmel.2022, Smielak.2023} for the central
@@
 $<$ -18 \degree below the horizon) and a continuous estimate of moonlight
-illumination (relative to the maximum moon illumination). Based on those
+illumination, relative to the maximum moon illumination. Based on those
 covariates, we generated a binary covariate separating bright from dark
-conditions. To provide a general understanding of this variable, bright
-conditions encompassed all daytime hours and those nighttime periods during
-which the moon was present in the sky and illuminated by approximately
-one-fourth; conversely, dark conditions included nighttime periods during which
-the moon was absent from the sky or present or was only minimally illuminated
-(further details in Appendices A2 and A3).
+conditions. Bright conditions encompassed all daytime hours and those nighttime
+periods during which the moon was present in the sky and illuminated by
+approximately one-fourth; conversely, dark conditions included nighttime periods
+during which the moon was absent from the sky or present and only minimally
+illuminated (further details in Appendices A2 and A3).
 
@@
 
-We modeled habitat selection and movement behavior of dispersing wild dogs using
-integrated step-selection functions (iSSFs, \citealp{Fortin.2005, Avgar.2016}),
-following the procedure described in \citet{Muff.2020}. For this, we identified
-bursts of subsequent GPS locations where the duration between two GPS locations
-did not exceed 4 hours (\(\pm\) 15 minutes) or 8 hours (\(\pm\) 30 minutes, for
-data recorded between 09:00 and 17:00). Within each burst, we converted
-locations into steps, where a step represented the straight line segment between
-two consecutive locations \citep{Turchin.1998}. For each step, we computed the
+We modeled dispersers' movement kernel and habitat selection using integrated
+step-selection functions (iSSFs, \citealp{Fortin.2005, Avgar.2016}), following
+the procedure described by \citet{Muff.2020}. For this, we identified bursts of
+subsequent GPS locations where the duration between two GPS locations did not
+exceed 4 hours (\(\pm\) 15 minutes) or 8 hours (\(\pm\) 30 minutes, for data
+recorded between 09:00 and 17:00). Within each burst, we converted locations
+into steps, where a step represented the straight line segment between two
+consecutive locations \citep{Turchin.1998}. For each step, we computed the
 associated step length (sl, in meters) and relative turning angle (ta, in
@@
 \inputy{GeneralMetrics/SSFMales} male coalitions) remained for further analyses.
-The final dataset comprised \inputy{GeneralMetrics/StepsTotal} steps
+4 coalitions dropped due to temporally irregular data. The final dataset
+comprised \inputy{GeneralMetrics/StepsTotal} steps
 (\inputy{GeneralMetrics/StepsMeanSD} per coalition), which we further
-categorized into wet (15 October to 15 April) and dry (15 April to 15 October)
-season, resulting in \inputy{GeneralMetrics/NumberStepsDry} steps
+categorized into data collected during the wet season (15 October to 15 April)
+and data collected during the dry season (15 April to 15 October), yielding
+\inputy{GeneralMetrics/NumberStepsDry} steps
 (\inputy{GeneralMetrics/PercentageStepsDry}\%) during the dry season and
@@
 Along each step, we extracted covariate values from the underlying spatial
-habitat layers, and we assigned the appropriate light conditions
-(\Cref{Covariates}). We opted for covariate extraction \textit{along} steps
-rather than \textit{at their endpoints}, as we believed that environmental
-conditions along steps were relevant in determining wild dog movement. For
-continuous covariates, we computed average values along each step, for
-categorical covariates the percentage cover of each category along the step. To
-model a decreasing marginal impact of ``distance to'' variables, we included
-their square-root as predictors in the final models. To facilitate model
-convergence, we normalized extracted values to a range between 0 and 1.
-
-To estimate habitat and movement parameters of interest, we applied the maximum
+layers, and we assigned the appropriate light conditions (\Cref{CovariatesCH3}).
+We opted for covariate extraction \textit{along} steps rather than \textit{at
+their endpoints}, as we believed that environmental conditions along steps were
+relevant in determining wild dog movement. For continuous covariates, we
+computed average values along each step, for categorical covariates the
+percentage cover of each category along the step. To model a decreasing marginal
+impact of ``distance-to'' variables, we included their square-root as predictors
+in the final models. To facilitate model convergence, we normalized extracted
+values to a range between zero and one.
+
+To estimate movement and habitat parameters of interest, we applied the maximum
 likelihood procedure proposed by \citet{Muff.2020}, using the \texttt{glmmTMB}
-package \citep{Brooks.2017}. Our model formula was based on knowledge from
-previous studies on dispersing wild dogs \citep{Hofmann.2021, Hofmann.2023} and
-comprised a movement kernel (describing movement preferences), a
-habitat-selection function (describing habitat preferences) and their
-interactions (describing how movement differs depending on habitat conditions).
-Notably, we included descriptors of the step length and turning angle (sl,
-log(sl), and cos(ta)) in the regression model \citep{Avgar.2017, Fieberg.2021}.
-We also fitted a stratum-specific intercept with a large fixed variance ($10^6$,
-\citealp{Muff.2020}), and used dispersing coalition ID to model random slopes.
-To examine whether the effect of accounting seasonality differed depending on
-the employed model, we fit models either \textit{excluding} interactions
-(hereafter called simple model) or \textit{including} interactions (hereafter
-called full model). Specifically, in the simple model we only included
-covariates relating to landscape characteristics and anthropogenic features in
-the simple model, as well as the step-descriptors mandatory for conducting an
-iSSF (i.e., sl, log(sl), and cos(ta); \citealp{Avgar.2016}). In the full model,
-however, we included additional interactions with climatic descriptors and light
-conditions, rendering that wild dog dispersal behavior may vary depending on
-those. As such, the structure of the simple model can be viewed as a model
-structure that is usually employed in permeability-based studies, which are
-restricted to spatial covariates. The model structure of the full model,
-conversely, included additional complexities that can only be accounted for when
-investigating connectivity vis IBMMs. To ensure comparability among each of the
-six configurations presented in our framework, we employed the same models
-across all of them. Specifically, we fitted models that differed in their degree
-of dynamism (\Cref{GraphicalAbstract}a). Models one and two were fit using
-static covariates (\Cref{GraphicalAbstract}a, Stage 1), with model one being a
-single-season model and model two being a multi-season model (wet vs. dry,
-\Cref{GraphicalAbstract}a, Stage 2). Models two and three, by contrast, were fit
+package \citep{Brooks.2017}. To examine whether the effect of accounting for
+seasonality was modulated by the model structure, we employed two separate model
+formulas that were both generated based on knowledge about wild dog dispersal
+movements previous studies \citep{Hofmann.2021, Hofmann.2023}. In the
+\textit{simple formula}, we merely included covariates relating to landscape
+characteristics and anthropogenic features, alongside the movement descriptors
+(sl, log(sl), and cos(ta)) that are mandatory to fit an iSSF \citep{Avgar.2017,
+Fieberg.2021}. This formula is most akin to the type of formulas that are
+usually used with permeability-based approaches, which lack a mechanistic
+understanding of movement. In the \textit{complex formula}, by contrast, we
+included several additional interactions between step-descriptors, landscape
+characteristics, climate descriptors and light availability, rendering that wild
+dog movement during dispersal may be affected by such interactions. In both
+formulas, we also included a stratum-specific intercept with a large fixed
+variance ($10^6$, \citealp{Muff.2020}), and used dispersing coalition ID to
+model random slopes for all main effects. To ensure comparability among each of
+the six configurations presented in our framework (\Cref{GraphicalAbstractCH3}),
+we employed the same model formulas across all of them. However, we fitted
+models that differed in their degree of dynamism, as highlighted in
+\Cref{GraphicalAbstractCH3}a. Models one and two were fit using static
+covariates, with model one being a single-season model and model two being a
+multi-season model (wet vs. dry). Models three and four, by contrast, were fit
 using dynamic covariates, with model three being a single-season model, and
-model four being a multi-season model (wet vs. dry, \Cref{GraphicalAbstract}a,
-Stage 2). To quantify how many random steps were necessary before model
-estimates stabilized (sensu \citealp{Fieberg.2021}), we fitted each model with
-5, 10, 25, 50, 75, and 100 random steps.
+model four being a multi-season model (wet vs. dry). Each model was once fit
+using the simple formula, once using the complex formula. To quantify how many
+random steps were necessary before model estimates stabilized (sensu
+\citealp{Fieberg.2021}), we fitted each model with 5, 10, 25, 50, 75, and 100
+random steps.
 
@@
 
-We compared the predictive efficacy of each of the six configurations
-(\Cref{GraphicalAbstract}b) using k-fold cross-validation for case-control
+We compared the predictive efficacy of each of the six configurations presented
+in \Cref{GraphicalAbstractCH3}b using k-fold cross-validation for case-control
 studies \citep{Fortin.2009}. For this, we split the data into training and
-validation sets using an 80:20 ratio and fitted the four iSSF models described
-above. We then used the $\beta$-estimates to predict the probability of each
-random and observed step in the validation set for being chosen. Depending on
-the configuration, we predicted step-probabilities using static or dynamic
+validation sets using an 80:20 ratio and fit the four iSSF models described
+above. We then used the obtained $\beta$-estimates to predict the probability of
+each random and observed step in the validation set for being chosen. Depending
+on the configuration, we predicted step-probabilities using static or dynamic
 covariates. Within each stratum, we assigned ranks 1-101 to each step based on
@@
 identifying the rank of a randomly chosen step. We replicated this validation
-procedure 100 times and computed the mean correlation coefficient $\bar{r}_{s,
-realized}$ and $\bar{r}_{random}$, as well as their 95\% percentiles across
-replicates. Ultimately, this validation proves a significant prediction in case
-the confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
-random}\) do not overlap \citep{Fortin.2009}.
+procedure 100 times and employed ANOVA to compare the predictive efficacy of
+different configurations.
 
@@
 
-To assess differences in connectivity upon increasing seasonal dynamism, we ran
-dispersal simulations under the two most distinct configurations, i.e. the fully
-static (SSS, \Cref{GraphicalAbstract}) and fully dynamic (DMD,
-\Cref{GraphicalAbstract}) configurations. As source areas to initiate
-dispersers, we defined three distinct regions known to host viable wild dog
-populations (\Cref{StudyArea}). The definition of these areas was somewhat
-arbitrary, albeit we deliberately selected areas in the west and east of the
-Delta to examine potential influences of flooding on connectivity
-\citep{Hofmann.2024}, as well as a more isolated location that was not
-influenced directly by the Delta's flood extent. To initiate simulated dispersal
+To assess differences in predicted connectivity upon increasing seasonal
+dynamism, we ran dispersal simulations under the two most distinct
+configurations, i.e. the fully static (SSS, \Cref{GraphicalAbstractCH3}) and
+fully dynamic (DMD, \Cref{GraphicalAbstractCH3}) configurations. As source areas
+to initiate dispersers, we defined three distinct regions known to host viable
+wild dog populations (\Cref{StudyAreaCH3}). The definition of these areas was
+somewhat arbitrary, albeit we deliberately selected areas in the west and east
+of the Okavango Delta to examine potential influences of flooding on
+connectivity (sensu \citealp{Hofmann.2024}). We also selected a more
+independent location (Nxai Pan; \Cref{StudyAreaCH3}) that was not influenced
+directly by the Okavango Delta's flood extent. To initiate simulated dispersal
 trajectories, we randomly placed 1,000 start points within each of these source
@@
 simulate dispersal and obtain connectivity maps under both configurations, we
-applied the simulation algorithm for iSSFs described in \citet{Signer.2017} and
-employed in \citet{Hofmann.2023}. A similar algorithm has recently been added to
+applied the simulation algorithm for iSSFs described by \citet{Signer.2017} and
+employed by \citet{Hofmann.2023}. A similar algorithm has recently been added to
 the \texttt{amt} \texttt{R}-package \citep{Signer.2024}. We applied the
@@
 log(sl), and cos(ta)), and assigned light conditions. We then employed the
-fitted iSSF model to predict the probability of each step for being chosen (i.e.
-the redistribution kernel). Based on assigned probabilities, we sampled one of
-the steps and updated the simulated individual's position and time. We repeated
-the procedure until a total of 2,000 steps were realized ($\sim$ 400 dispersal
-days).
-
-Depending on the configuration, we employed different spatial covariates during
-the simulation. For the SMS configuration, we used the static set of covariates,
-which omitted any seasonality and assumed a static landscape over time. For the
-DMD configuration, on the other hand, we updated covariates dynamically as the
-simulated individuals moved. More specifically, we generated a new covariate
-stack after every simulated step, comprising spatial layers that best
-represented environmental conditions at that particular point in time
-(\Cref{Schematic}). Note that we used the set of covariates representing an
-average year for this purpose. For both configurations, we ultimately obtained a
-heatmap (or utilization distribution) based on simulated trajectories. We also
-generated difference maps between the two configurations, highlighting areas
-where the difference in simulated connectivity were most pronounced.
-
-\begin{figure}
+fitted iSSF model to predict the probability of each step for being chosen, thus
+generating a ``redistribution kernel'' around the current location of the
+simulated individual \citep{Signer.2017, Signer.2024}. Based on this kernel, we
+sampled one of the steps and updated the simulated individual's position and
+time. We repeated the procedure until a total of 2,000 steps were realized
+($\sim$ 400 dispersal days).
+
+Depending on the configuration, we employed different spatial covariates and
+preferences during the simulation. For the SSS configuration, we omitted any
+seasonality, using the static set of covariates and assuming dispersers to
+exhibit the same preferences (i.e., $\beta$-estimates) across seasons. For the
+DMD configuration, on the other hand, we updated covariates and preferences
+dynamically as the simulated individuals moved. More specifically, we generated
+a new covariate stack after every simulated step, comprising spatial layers that
+best represented environmental conditions at that particular point in time
+(\Cref{Schematic}). For this purpose, we employed the set of covariates
+representing an average year. Similarly, we predicted the redistribution kernel
+using $\beta$-estimates from the appropriate season. Finally, we obtained
+heatmaps (connectivity maps, utilization distributions) based on simulated
+dispersal trajectories under both configurations.
+
+\begin{figure}[htpb]
  \begin{center}
-  \includegraphics[width = \textwidth]{Figures/Schematic} \caption{Schematic
-  illustration of a dispersal simulation with dynamic covariates. As the
-  simulation proceeds, the underlying covariates (symbolized by the stack of
-  layers) are updated. In our case, the update frequency of covariate layers
-  varied from a few hours (e.g., temperature) to multiple months (e.g., shrub
-  cover). Each gray block represents a single layer and the duration for which
-  it was ``active''. Originating from the current position of the simulated
-  animal, a new redistribution kernel is derived. We generated redistribution
-  kernels by proposing a set of random steps and applying the parametrized
-  step-selection model to predict the probability of each step for being chosen.
-  Based on this kernel, one location was randomly sampled and the animal's
-  position updated. This procedure was repeated until the number of simulated
-  steps matched the number of steps from the observed individuals.}
+  \includegraphics[width = \textwidth]{Figures/Schematic}
+  \caption{Schematic illustration of our dispersal simulation with dynamic
+  covariates. As the simulation proceeded, the underlying spatial covariates
+  (symbolized by the stack of layers) were updated. Depending on the covariate,
+  the update frequency varied from a few hours (e.g., temperature) to multiple
+  months (e.g., shrub cover). Each gray block represents a single layer and the
+  duration for which it was ``active''. Originating from the current position of
+  the simulated animal, a new redistribution kernel was derived. We generated
+  redistribution kernels by proposing a set of 25 random steps and applying the
+  parametrized step-selection model to predict the probability of each step for
+  being chosen. Based on this kernel, one location was randomly sampled and the
+  animal's position and time updated. This procedure was then repeated until
+  2,000 steps ($\sim 400$ dispersal days) were simulated.}
   \label{Schematic}
@@
 irrespective of whether models were fit using static or dynamic covariates
-(\Cref{MovementModel}, Tables S3, S4, S5, S6) and irrespective of the number of
-considered random steps (Figure S9). The most notable quantitative difference
-when using static vs. dynamic covariates (see \Cref{GraphicalAbstract}a) was
-that models fitted using dynamic covariates resulted in narrower confidence
-intervals (\Cref{MovementModel}). Furthermore, effect sizes (i.e.,
-$\beta$-estimates) of models fit using static covariates were more pronounced
-than effect sizes from models that were fit using dynamic covariates
-(\Cref{MovementModel}). Differences were most marked for
-$\beta_{DistanceToWater^{0.5}}$, which was estimated $\approx -0.8$ using static
-covariates, but $\approx -0.3$ when using dynamic covariates.
+(\Cref{MovementModelCH3}, \Cref{ModelsStaticSimple}, \Cref{ModelsDynamicSimple},
+\Cref{ModelsStaticFull}, \Cref{ModelsDynamicFull}) and irrespective of the
+number of considered random steps (\Cref{Stability}). The most notable
+quantitative difference when using static vs. dynamic covariates was that models
+fitted using dynamic covariates resulted in narrower confidence intervals
+(\Cref{MovementModelCH3}). Furthermore, effect sizes (i.e., $\beta$-estimates)
+of models fit using static covariates were more pronounced than effect sizes
+from models that were fit using dynamic covariates (\Cref{MovementModelCH3}).
+Differences were most marked for $\beta_{DistanceToWater^{0.5}}$, which was
+estimated $\approx -0.8$ using static covariates, but $\approx -0.3$ when using
+dynamic covariates.
 
@@
 pronounced when models were fit using static covariates. For instance, when
-models were fit using static covariates, $\beta_{Water}$ was $\approx -1.1$
-during the dry season but ($\approx 0.2$ during the wet season). When dynamic
-covariates were used, by contrast, $\beta_{Water}$ was ($\approx -0.6$ and
-$\approx -0.3$), respectively. $\beta_{sl:Dark}$ was also markedly lower during
-the wet season ($\approx -0.9$), compared to the dry season ($\approx -0.1$).
-Plots that aid with the interpretation of the resulting models are provided in
-Figure S10.
-
-\begin{figure}
+models were fit using static covariates, $\beta_{Water}$ was $\approx -0.9$
+during the dry season but $\approx 0.2$ during the wet season. When dynamic
+covariates were used, by contrast, $\beta_{Water}$ was $\approx -0.6$ and
+$\approx -0.3$, respectively. $\beta_{sl:Dark}$ was also markedly lower during
+the wet season ($\approx -0.9$), compared to the dry season ($\approx -0.1$) but
+this was independent of the type of covariates that were used. Plots that aid
+with the interpretation of the models are provided in
+\Cref{MovementModelInterpretation}.
+
+\begin{figure}[htpb]
  \begin{center}
@@
   by movement kernel, habitat-selection function, and their interaction. We
-  either fit a simple model without interactions or a full model with
-  interactions and we distinguished between models fit using static or dynamic
-  covariates. Only the underlined covariates differed between the static and
-  dynamic configurations, as covariates were either represented as a single
-  layer (static), or a stack of layers (dynamic). Furthermore, data was either
-  pooled across seasons (yellow bars) or split into dry (green bars) and wet
-  (blue bars) season.}
-  \label{MovementModel}
+  either fit models using a \textit{simple formula} (without interactions) or a
+  \textit{complex formula} that comprised several interaction terms that rendered
+  how movement behavior depended on environmental conditions. Furthermore, we
+  distinguished between models fit using \textit{static} and \textit{dynamic}
+  covariates. However, only the underlined covariates differed between the
+  static and dynamic configurations, as covariates were either represented as a
+  single layer (static), or a stack of layers (dynamic). Furthermore, data were
+  either pooled across seasons (yellow bars) or split into dry (green bars) and
+  wet (blue bars) season. Results are shown for replicates that used 25 random
+  steps, as all associated models converged. Note, however, that the number of
+  random steps did not appear to influence model estimates (\Cref{Stability}).}
+  \label{MovementModelCH3}
  \end{center}
@@
 Spearman's rank correlation coefficients ($r$) obtained from the validation
-procedure revealed that predictions from the full model performed better than
-those from the simple model ($\bar{r}_{simple} = -0.5$, $\bar{r}_{full} = -0.9$,
-\Cref{RankCorrelation}). Irrespective of the employed model, Spearman's rank
-correlation differed significantly depending on the amount of dynamism
-considered (simple: F(5, 593) = 26.45, p < 0.001, full: F(5, 594) = 7.14, p <
-0.001), albeit with moderate effect sizes. In the simple model, moving from a
-fully static (SSS) to a fully dynamic (DMD) configuration decreased Spearmans's
-rank correlation (i.e., increased the predictive performance) by 0.15 from -0.41
-to -0.56. In the full model, moving from a fully static (SSS) to a fully dynamic
-(DMD) configuration entailed an increase in $r$ (i.e. a decrease in the
-predictive performance) from -0.89 to -0.88. This suggests that, our hypothesis
-that increased dynamism results in better predictive performance only holds with
-the simple model (\Cref{RankCorrelation}a), but not the full model
-(\Cref{RankCorrelation}b).
-
-\begin{figure}
+procedure revealed that predictions using the complex formula performed markedly
+better than those from the simple formula ($\bar{r}_{simple} = -0.5$,
+$\bar{r}_{full} = -0.9$, \Cref{RankCorrelation}). Irrespective of the employed
+model, Spearman's rank correlation differed significantly depending on the
+amount of dynamism considered (simple: F(5, 593) = 26.45, p < 0.001, full: F(5,
+594) = 7.14, p < 0.001), albeit with moderate effect sizes. Using the simple
+formula, moving from a fully static (SSS) to a fully dynamic (DMD) configuration
+decreased Spearmans's rank correlation (i.e., increased the predictive
+performance) by 0.15 from -0.41 to -0.56. Using the complex formula, conversely,
+moving from a fully static (SSS) to a fully dynamic (DMD) configuration even
+entailed a minimal increase in $r$ (i.e., a decrease in the predictive
+performance) from -0.89 to -0.88. This suggests that, our hypothesis that
+increased dynamism results in better predictive performance only holds when
+using the simple formula (\Cref{RankCorrelation}a), but not when employing the
+complex formula (\Cref{RankCorrelation}b).
+
+\begin{figure}[htpb]
  \begin{center}
@@
   100 replicates. Note that the y-axis is inverted to match our expectation of
-  increasing performance as dynamism increases. (b) Difference in Spearman's
-  rank correlation when moving from one configuration to another. Values $< 0$
-  (blue) indicate an increase in predictive performance, whereas values $> 0$
-  (orange) indicate a decrease in predictive performance.}
+  increasing performance as dynamism increases. The boxplots in light gray
+  represent Spearman's rank correlation coefficients under a null model and
+  serve for comparison. (b) Difference in Spearman's rank correlation when
+  moving from one configuration to another. Values $< 0$ (blue) indicate an
+  increase in predictive performance, whereas values $> 0$ (orange) indicate a
+  decrease in predictive performance.}
   \label{RankCorrelation}
@@
 revealed different connectivity patterns (\Cref{HeatmapComparison}). In the
-static configuration, simulated dispersal trajectories more concentrated,
-resulting in frequent movement across a few key habitats. In the dynamic
+static configuration, simulated dispersal trajectories were more concentrated,
+suggesting frequent movement across a few key habitats. In the dynamic
 configuration, by contrast, simulated trajectories were more homogeneously
-distributed. Qualitatively, connectivity patterns appeared similar between the
-simple and full models. Given the similarities in habitat selection emerging
-under the two models, this was to be expected.
-
-\begin{figure}
+distributed, suggesting that movement was less restricted. When qualitatively
+comparing the heatmaps generated under the simple and complex formulas,
+differences are imperceptible. Given the similarities in habitat selection
+emerging under the two models, this was to be expected.
+
+\begin{figure}[htpb]
  \begin{center}
@@
   \caption{Heatmaps derived under the most static (SSS) and dynamic (DMD)
-  configurations. Results are shown for both the simple and full model.}
+  configurations. Results are shown for simulations using the simple and complex
+  model formulas.}
   \label{HeatmapComparison}
@@
 
-\subsection{Brief Summary}
+% \subsection{Brief Summary}
 
@@
 configurations that differ in their degree of dynamism and, arguably, realism.
-We fitted the models associated with each configuration using GPS data on
-dispersing wild dogs and iSSFs and employed a rigorous validation procedure to
-investigate potential gains in predictive performance that can be reaped by
-incorporating different levels of dynamism. Results from the fitted models
-showed that including seasonality only marginally affected the inferred patterns
-of habitat selection and movement behavior. Similarly, the validation procedure
+We fit the models associated with each configuration using GPS data on
+dispersing wild dogs and employed a rigorous validation procedure to investigate
+potential gains in predictive performance that can be reaped by incorporating
+different levels of seasonal dynamism. Results from the fitted models showed
+that including seasonality only marginally affected the inferred patterns of
+habitat selection and movement behavior. Similarly, the validation procedure
 suggested only moderate improvements in predictive performance upon increasing
-the degree of dynamism. Crucially, these benefits were limited to an overly
-simplistic model and vanished upon fitting a more complex one. We therefore
-could not pinpoint a specific stage at which including seasonality was
+the degree of seasonal dynamism. Crucially, these benefits were limited to an
+overly simplistic model formula and vanished upon fitting a more complex one. We
+therefore could not pinpoint a specific stage at which including seasonality was
 particularly beneficial. Despite this, we found that dispersal simulations under
@@
 
-\subsection{Moderate Improvements upon Increasing Dynamism}
+% \subsection{Moderate Improvements upon Increasing Dynamism}
 
 Our validation procedure revealed only moderate improvements in our ability to
-predict dispersal movements when increasing the degree of dynamism. Given the
-system's extreme seasonal variability, this was somewhat surprising. We believe
-that the absence of a more pronounced improvement can be traced back to multiple
-factors. Firstly, we focused our analysis on dispersing individuals, which cover
-large distances in search of potential mates and a suitable territory
-\citep{McNutt.1996, Cozzi.2020}. In our case, the average 4-hourly step length
-was 2.5 km, suggesting that dispersers cross and sample numerous unfamiliar
-areas and potentially unsuitable habitats within short time. We thus hypothesize
-that the spatial scale at which seasonality affects environmental
-characteristics did not suffice to match the spatial scale at which our focal
+predict dispersal movements when increasing the degree of seasonal dynamism.
+Given the system's extreme seasonal variability, this was somewhat surprising.
+We believe that the absence of a more pronounced improvement can be traced back
+to multiple factors. Firstly, we focused our analysis on dispersing individuals,
+which cover large distances in search of potential mates and a suitable
+territory \citep{McNutt.1996, Cozzi.2020}. In our case, the average 4-hourly
+step length was 2.5 km, suggesting that dispersers cross and sample numerous
+unfamiliar areas and potentially unsuitable habitats within short time. We thus
+hypothesize that the spatial scale at which seasonality affects environmental
+characteristics does not suffice to match the spatial scale at which our focal
 species perceives and moves across the landscape during dispersal. A Further
@@
 relatively weak selection or avoidance of environmental characteristics
-exhibited in \Cref{MovementModel}. Finally, the AWD is a generalist species that
-can occupy a broad variety of habitats \citep{Woodroffe.2011}. A certain
+exhibited in \Cref{MovementModelCH3}. Finally, the AWD is a generalist species
+that can occupy a broad variety of habitats \citep{Woodroffe.2011}. A certain
 tolerance towards changing environmental conditions can therefore be expected
@@
 
-\subsection{Substantial Improvements upon Fitting a Complex Model}
+% \subsection{Substantial Improvements upon Fitting a Complex Model}
 
@@
 significant improvement in predictive ability was achieved by moving from the
-simple to the full iSSF model. In the full model, we included several
-interactions that accounted for wild dogs' biology, such as reduced movement
-during dark nights \citep{Cozzi.2013} or during times of high ambient
-temperature \citep{Rabaiotti.2021}. Even though these interactions are not
-directly linked to connectivity, they clearly accounted for a substantial amount
-of variation in observed dispersal movements and facilitated predictions from
-the fitted model. We also allowed for potential changes in dispersers' movement
-behavior depending on habitat conditions, such as shorter steps in areas covered
-by water \citep{Hofmann.2023}. Overall, the inclusion of these interactions
-elevated the predictive ability of our dispersal model to such a high level that
-increasing dynamism did not provide any further improvements. Notably, the
-ability of encapsulating such a detailed and mechanistic understanding of
-dispersal movements is unique to IBMMs and cannot be achieved using
-permeability-based connectivity models.
-
-\subsection{Dynamic Connectivity}
+simple to the complex model formula. When employing the complex formula, we
+included several interactions that accounted for the African wild dog's biology,
+such as reduced movement during dark nights \citep{Cozzi.2013} or during times
+of high ambient temperature \citep{Rabaiotti.2021}. We also allowed for
+potential changes in dispersers' movement behavior depending on habitat
+conditions, such as shorter steps in areas covered by water
+\citep{Hofmann.2023}. The ability of encapsulating such a detailed and
+mechanistic understanding of dispersal movements is unique to IBMMs and cannot
+be achieved using permeability-based connectivity models. Even though the
+additional interactions were not directly linked to connectivity, they accounted
+for a substantial amount of variation in observed dispersal movements and
+facilitated predictions from the fitted model. Notably, the inclusion of these
+interactions elevated the predictive ability of our dispersal model to such a
+high level that considering seasonal dynamism did not provide any further
+improvements, irrespective of the stage at which it was considered.
+
+% \subsection{Dynamic Connectivity}
 
@@
 accessible, even if only for a limited time. In our study, such stepping stones
-could arise in two ways. Firstly, the landscape could change seasonally, which
-may have shifted preferred habitats, leading to the emergence of alternative
-movement corridors. In addition, habitat or movement preferences of simulated
-individuals could change, resulting in different habitat characteristics being
-traversed depending on the season. In combination, varying landscape conditions
-and species preferences resulted in a more balanced mosaic of connectivity
-across the year. Importantly, the fact that connectivity is more evenly
-distributed across the year does not exclude the possibility that certain areas
-experience lower connectivity seasonally. As already shown by
-\citet{Osipova.2019}, a static representation may result in both an over- and
-under-estimation of short-term connectivity, depending on the area and season.
-For conservation planning, this implies a need to protect landscapes at broader
-scales, as areas that provide little connectivity during some season may become
-critical during others. A dynamic take at connectivity therefore improves our
-understanding of ecological processes in dynamic landscapes and may help to
-identify otherwise overlooked dispersal hotspots.
-
-\subsection{The Costs of Incorporating Dynamism}
+could arise in two ways. Firstly, the landscape changed seasonally, which caused
+spatial shifts in preferred and avoided habitats, leading to the emergence of
+alternative movement corridors. In addition, habitat or movement preferences of
+simulated individuals changed, which resulted in different habitat
+characteristics being traversed depending on the season. In combination,
+temporally varying landscape conditions and seasonally altered species
+preferences resulted in a more balanced mosaic of connectivity across the year.
+Importantly, the fact that connectivity was more evenly distributed across the
+year does not exclude the possibility that certain areas experienced lower
+connectivity seasonally. As already shown by \citet{Osipova.2019}, a static
+representation may result in both an over- and under-estimation of short-term
+connectivity, depending on the area and season. For conservation planning, this
+implies a need to protect landscapes at broader scales, as areas that provide
+little connectivity during some season may become critical during others. A
+dynamic take at connectivity therefore improves our understanding of ecological
+processes in dynamic landscapes and may help to identify otherwise overlooked
+dispersal hotspots.
+
+% \subsection{The Costs of Incorporating Dynamism}
 
@@
 costs, both in terms of data requirements and computational challenges. To
-represent seasonal changes in the landscape, one needs to download and process
-frequently updated spatial layers for each seasonal covariate. This is
+represent seasonal changes across the landscape, one needs to download and
+process frequently updated spatial layers for each seasonal covariate. This is
 time-consuming and implies a substantial increase in the data-volume that needs
-to be handled. Because seasonal products are comparably rare, the choice to
-model connectivity dynamically will also entail a significant restriction in the
-number of potential covariates or require custom remote sensing algorithms
-(e.g., Appendix A1). Seasonal or remote sensed layers are also more susceptible
-to noise and missing values, particularly in cases where cloud cover is
-frequent. Extracting covariates from seasonal layers poses a further challenge,
-as data need to be extracted from those layers that best represent environmental
+to be handled. Because seasonal products are comparably rare, modeling dynamic
+connectivity also entails a significant reduction in the number of potential
+covariates and may require custom remote sensing algorithms (e.g.,
+\Cref{PanMappingApproach}). Seasonal and remote sensed layers are also more
+susceptible to noise and missing values, particularly during seasons when cloud
+cover is frequent. Extracting covariates from seasonal layers poses a further
+challenge, as covariate data can no longer be extracted from single layers, but
+need to be extracted from exactly those layers that best represent environmental
 conditions at the time of the respective observed or random step. This applies
 to the extraction of data for model fitting, as well as when extracting data
-during the simulation. Finally, splitting the species data by season to fit
-seasonal selection models significantly reduces the amount of data remaining per
-season, potentially causing convergence issues. In our case, these additional
-efforts were not outweighed by an improved predictive ability. This finding is,
-however, likely highly specific to our study system. We therefore do not wish to
-discourage future studies from accounting for seasonality when assessing
-connectivity. Instead, we'd like to view our study as an example that the
-benefits of increased realism do not necessarily justify the additional
-complexity they bring about and that the associated costs and benefits need to
-be carefully pondered \citep{Puy.2022}.
-
-\subsection{Validation Procedure}
+during the simulation. Finally, splitting species data by season to parametrize
+seasonal selection models can significantly reduce the amount of data remaining
+per season, potentially causing convergence issues. In our case, the moderate
+improvement in predictive ability did not warrant these additional efforts and
+complications. This conclusion is, however, likely highly specific to our study
+system. We therefore do not wish to discourage future studies from accounting
+for seasonality when assessing connectivity. Instead, we would like to view our
+study as an example that the benefits of increased realism do not necessarily
+justify the additional complexity they bring about and that the associated costs
+and benefits need to be carefully pondered \citep{Puy.2022}.
+
+% \subsection{Validation Procedure}
 
@@
 true connectivity. This was owed to a conceptual limitation when trying to
-validate seasonal connectivity predicted from an IBMM. In seasonal landscapes,
-connectivity not only varies spatially, but also temporally. A meaningful
-validation therefore requires that connectivity is predicted separately for each
-timestamp in the validation data. In our case, for instance, the fact that the
-landscape was updated every 4 hours implies that a new connectivity map would
-need to be created around each validation step. This is equivalent to validating
-connectivity at the step level, which is why we deemed the application of k-fold
+validate seasonal connectivity as predicted from an IBMM. In seasonal
+landscapes, connectivity not only varies spatially, but also temporally. A
+meaningful validation therefore requires that connectivity is predicted
+separately for each timestamp in the validation data. In our case, for instance,
+the fact that the landscape changed every four hours (i.e., with every step)
+entails that a new connectivity map would need to be developed around each
+validation step. This is equivalent to generating and validating the
+redistribution kernel, and therefore equivalent to validating predictions at the
+step level. For this reason, we deemed the application of k-fold
 cross-validation for case and control studies using Spearman's rank correlation
@@
 correlation coefficient is heavily dependent on the number of steps per stratum
-and should therefore not be used to compare predictive efficacy across studies
-(Figure S11). For static connectivity analyses, alternative validation
-techniques exist. \citet{McClure.2016}, for instance, proposed a suite of
-validation metrics (applied in \citealp{Zeller.2018, Finerty.2023}). These
-largely work by comparing connectivity at random locations with connectivity at
-locations where the focal species was observed. Because these metrics fail to
-account for potential autocorrelation in the validation data (which is usually
-present in GPS data; \citep{Otis.1999}), \citet{Brennan.2020} suggested
-validating connectivity by applying a path- or step-selection model to withheld
-data. If the fitted model proves significant selection towards areas of high
-connectivity, predictions are indeed indicative of functional connectivity
-(sensu \citealp{Brennan.2020}). Developing similar approaches for dynamic
-connectivity analyses remains a task for future studies.
-
-\subsection{Seasonal Habitat Selection}
+and should not be used to compare predictive efficacy across studies
+(\Cref{RankCorrelationNumberSteps}). For static connectivity analyses,
+alternative validation techniques exist. \citet{McClure.2016}, for instance,
+proposed a suite of validation metrics (applied in \citealp{Zeller.2018} and
+\citealp{Finerty.2023}) that largely work by comparing connectivity at random
+locations with connectivity at locations where the focal species was observed.
+Because these metrics fail to account for potential autocorrelation in the
+validation data (which is usually present in GPS data), \citet{Brennan.2020}
+suggested validating connectivity by applying a path- or step-selection model to
+withheld data, using the connectivity map as the only spatial covariate. If the
+fitted model proves significant selection towards areas of high connectivity,
+predictions are indeed indicative of functional connectivity (sensu
+\citealp{Brennan.2020}). Developing similar approaches for dynamic connectivity
+analyses remains a task for future studies.
+
+% \subsection{Seasonal Habitat Selection}
 
@@
 for areas close to water, irrespective of the season. However, water itself
-appeared to be only avoided during the dry season. This was likely caused by a
+appeared only to be avoided during the dry season. This was likely caused by a
 misrepresentation of areas covered by water in the static configuration. Since
-rainwater collected in Angola only slowly descents through the Okavango Delta's
-tributaries \citep{McCarthy.1997}, large portions of its floodplains remain dry
-during the wet season \cite{McCarthy.2003}. On static covariates, which
-represent average conditions across both seasons, it may appear as if
+rainwater collected in Angola only slowly descends through the Okavango Delta's
+tributaries \citep{McCarthy.1997}, large portions of the Delta's floodplains
+remain dry during the wet season \citep{McCarthy.2003}. On static covariates,
+which represent average conditions across both seasons, it may appear as if
 individuals moved through areas covered by water, albeit in reality they likely
@@
 considering seasonally updated covariate layers, parameter estimates from the
-selection models assimilated and suggested avoidance of water across both
-seasons. Our choice of splitting data into wet and dry season based on a fixed
-set of dates can be viewed as a further limitation, as it assumes an immediate
-switch from one season to another. In reality, season-transitions are not abrupt
-but rather gradual and their timing may vary from year to year. A more robust
-approach would therefore be to define the start and end of each season based on
-bio-climatic descriptors and to ignore data collected during transitional
-periods. This may imply a loss of data, but could help to detect seasonal
-selection patterns. In our case, retaining all data was necessary to ensure
-model convergence, but may have resulted in a cross-contamination between
-seasons and therefore reduced our ability to pick up seasonal differences in
-parameter estimates. Indeed, other studies reported that habitat selection of
-their focal species varied greatly between seasons (but see
+selection models assimilated and consistently suggested avoidance of water
+across both seasons. Accounting for seasonality can therefore result in vastly
+different biological insights. In this regard, our choice of splitting data into
+wet and dry season based on a fixed set of dates can be viewed as limitation, as
+it assumed an immediate switch from one season to another. In reality,
+season-transitions are not abrupt but rather gradual and their timing may vary
+from year to year. A more robust approach would therefore be to define the start
+and end of each season based on bio-climatic descriptors and to ignore data
+collected during transitional periods. This may imply a loss of data, but could
+help to detect seasonal selection patterns. In our case, retaining all data was
+necessary to ensure model convergence, but may have resulted in a
+cross-contamination between seasons and therefore reduced our ability to pick up
+seasonal differences estimated parameters. Indeed, other studies report that
+habitat selection of their focal species varied greatly between seasons (but see
 \citealp{Squires.2013}). \citet{Benz.2016}, for instance, found that dispersing
@@
 
-\subsection{Conclusion}
+% \subsection{Conclusion}
 
 In conclusion, we explored the importance of incorporating different degrees of
-dynamism when studying dispersal and connectivity. Overall, our findings
-suggested only moderate improvements in predictive performance upon increasing
-the level of dynamism. Nevertheless, connectivity patterns as inferred from
-simulated dispersal trajectories differed vastly, with connectivity being more
-evenly distributed when seasonality was accounted for. While increased realism
-via improved representation of seasonality may offer novel insights into
-ecological processes, the benefits must be carefully weighed against the added
-complexity and effort required to collect seasonal data and model seasonal
-dynamism in the focal species' preferences. Moving forward, our study serves as
-an example that while accounting for seasonality could be crucial for some study
-systems, it may not always justify the additional complexity, emphasizing the
-need for thoughtful consideration of costs and benefits in future research.
-
-\section{Authors' Contributions}
-
-D.D.H. and G.C. conceived the study and designed methodology; D.D.H., G.C.,
-D.M.B, and J.W.M. collected the data; D.D.H. analysed the data; G.C. assisted
-with modeling; D.D.H. and G.C. wrote the first draft of the manuscript and all
-authors contributed to the drafts at several stages and gave final approval for
-publication.
-
-\section{Data Availability}
-
-Access to R-scripts to replicate our analyses will be provided through an online
-repository at the time of publication.
-
-\section{Acknowledgments \& Funding}
-
-We thank the Ministry of Environment and Tourism and the Department of Wildlife
-and National Parks of Botswana for granting permission to conduct this
-research. D.D.H. was funded by a Swiss National Science Foundation Grant No.
-310030\_204478 to G.C.
-
+seasonal dynamism when studying dispersal and connectivity. Overall, our
+findings suggested only moderate improvements in predictive performance upon
+increasing the level of dynamism. Nevertheless, we found that connectivity
+patterns as inferred from simulated dispersal trajectories differed vastly, with
+connectivity being more evenly distributed when seasonality was accounted for.
+While increased realism via improved representation of seasonal dynamism may
+offer novel insights into ecological processes, its benefits must be carefully
+weighed against the added complexity introduced by seasonal data and seasonal
+preferences. Moving forward, our study serves as an example that while
+accounting for seasonality could be crucial for some study systems, it may not
+always justify the additional complexity, emphasizing the need for thoughtful
+consideration of costs and benefits in future research.
+
+% \section{Authors' Contributions}
+%
+% D.D.H. and G.C. conceived the study and designed methodology; D.D.H., G.C.,
+% D.M.B, and J.W.M. collected the data; D.D.H. analysed the data; G.C. assisted
+% with modeling; D.D.H. and G.C. wrote the first draft of the manuscript and all
+% authors contributed to the drafts at several stages and gave final approval for
+% publication.
+%
+% \section{Data Availability}
+%
+% Access to R-scripts to replicate our analyses will be provided through an online
+% repository at the time of publication.
+%
+% \section{Acknowledgments \& Funding}
+%
+% We thank the Ministry of Environment and Tourism and the Department of Wildlife
+% and National Parks of Botswana for granting permission to conduct this
+% research. D.D.H. was funded by a Swiss National Science Foundation Grant No.
+% 310030\_204478 to G.C.
+
+\ifSubfilesClassLoaded{%
   \newpage
-\begingroup
-\singlespacing
+  \begin{singlespacing}
   \ifthenelse{\boolean{usebiblatex}}{
@@
     \end{refcontext}
-}{
-  \bibliography{LiteratureBibtex}
+  } {
+    \bibliography{../LiteratureBibtex}%
   }
-\endgroup
+\end{singlespacing}
+}{}
 

### 2024-07-01 FinalThesis/Chapter_03/Manuscript_Support.tex
Publish PHD

--- 2024-07-01 Chapter_3/04_Manuscript/Manuscript_Support.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-\usepackage{scalerel}
-\usepackage{makecell}
-\usepackage{ifthen}
-\usepackage{tikz}
-\usepackage{lineno}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%  Bibtex vs Biblatex
-%------------------------------------------------------------------------------
-% While bibtex is very nice to use, it's insanely slow. I therefore work with
-% bibtex until the final stage. Set the value below to "true" or "false" to pick
-% which one should be used.
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{false}
-
-% Depending on the question, the commands below will prepare the document
-% accordingly
-\ifthenelse{\boolean{usebiblatex}}{
-  %-----------------------------------------------------------------------------
-  %  If using biblatex
-  %-----------------------------------------------------------------------------
-  % Load required packages and refer to correct bibliography
-  \usepackage{csquotes}
-  \usepackage[
-      backend=biber
-    , style=apa          % To display Author-year in the text
-    , natbib=true        % Necessary to use citep, etc.
-    , hyperref=true      % To have clickable links in the references
-    , useeditor=false    % Don't add editor to publications
-    , sorting=ynt        % Sort by year, name, and title
-    , uniquename=false   % Avoid that initials are put
-  ]{biblatex}
-  \addbibresource{Literature.bib}
-
-  % To include author in link
-  \DeclareFieldFormat{citehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{#1}}
-  \DeclareFieldFormat{textcitehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{%
-    #1%
-    \ifbool{cbx:parens}
-      {\bibcloseparen\global\boolfalse{cbx:parens}}
-      {}}}
-  \savebibmacro{cite}
-  \savebibmacro{textcite}
-  \renewbibmacro*{cite}{%
-    \printtext[citehyperref]{%
-      \restorebibmacro{cite}%
-      \usebibmacro{cite}}}
-  \renewbibmacro*{textcite}{%
-    \ifboolexpr{
-      ( not test {\iffieldundef{prenote}} and
-        test {\ifnumequal{\value{citecount}}{1}} )
-      or
-      ( not test {\iffieldundef{postnote}} and
-        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
-    }
-      {\DeclareFieldAlias{textcitehyperref}{noformat}}
-      {}%
-    \printtext[textcitehyperref]{%
-      \restorebibmacro{textcite}%
-      \usebibmacro{textcite}}}
-
-  % Remove editor form everything but books
-  \AtEveryBibitem{%
-   \ifentrytype{book}{}{%
-    \clearname{editor}
-   }
-  }
-}{
-  %-----------------------------------------------------------------------------
-  %  If using bibtex
-  %-----------------------------------------------------------------------------
-  % Load required packages and select citation style
-  \usepackage[round]{natbib}
-  \bibliographystyle{apalike}
-}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Link colors
-\definecolor{linkcolor}{HTML}{000000}
-\hypersetup{
-  colorlinks = true,
-  linkcolor  = linkcolor,
-  urlcolor   = linkcolor,
-  citecolor  = linkcolor,
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% Avoid skip when including text from external .tex file
-\newcommand{\inputy}[1]{\input{#1}\unskip}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%  ORCID
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Create the ORCID logo
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol]
-      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
-        ,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white]
-      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
-        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
-        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
-      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-
-% Command to create the ORCID
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Individual-Based Dispersal Simulations to Study
-Dynamic Connectivity in Seasonal Landscapes}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
-  John W. McNutt\textsuperscript{2} \and
-  Arpat Ozgul\textsuperscript{1, 2} \orcid{0000-0001-7477-2642} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
 
-% Change page numbering
-\pagenumbering{gobble}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
-
-\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
-
-\vspace{4cm}
-
-\textbf{Running Title:} Simulating Dispersal in Seasonal Landscapes
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} African wild dog, Connectivity, Dispersal, Dynamism,
-Individual-based, Lycaon pictus, Okavango Delta, Seasonality
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
 \newpage
-
-\onehalfspacing
-\tableofcontents
-\doublespacing
-
-% Change page numbering
-\pagenumbering{arabic}
-
-% Create linenumbers
-% \linenumbers
-
-% Change to appendix counters
-\appendix
-\beginappendix
 
@@
 \section{Pan Mapping}
+\label{PanMappingApproach}
+
 \subsection{Satellite Product Overview}
@@
 We dynamically mapped the spatial distribution of ephemeral water bodies that
-span only a few meters in diameter (\Cref{Pan}) using a custom remote sensing algorithm.
-These waterbodies are typically referred to as ``pans'' and mainly inundate
-during the rainy season, after which the cumulated water slowly evaporates.
-Previously, we developed an algorithm to detect large scale floodwaters across
-the extent of the Okavango Delta. This algorithm made use of MODIS MCD34A4
-satellite imagery, which is a 8-day composite of daily updated MODIS satellite
-data. The benefit of MODIS' high temporal resolution is that monthly composites
-are almost guaranteed to be cloud free, even during Botswana's rainy season (mid
-October to mid May). However, due to its coarse spatial resolution of 250
-meters, MODIS satellite data was unsuitable to detect finer-scaled waterbodies
-on satellite images. In an attempt to overcome this limitation, we evaluated
-alternative satellite products that provided better spatial resolution than
-MODIS, while retaining a sufficiently high temporal resolution to render
-seasonal patterns. Candidate products comprised Landsat 7, Landsat 8, and
-Sentinel 2 satellite imagery. Data associated with each of these satellites are
-freely accessible and provide spatial resolutions between 10 and 60 meters,
-depending on the respective bands. Furthermore, the satellites have
+span only a few meters in diameter (\Cref{Pan}) using a custom remote sensing
+algorithm. These waterbodies are typically referred to as ``pans'' and mainly
+inundate during the rainy season, after which the cumulated water slowly
+evaporates. Previously, we developed an algorithm to detect large scale
+floodwaters across the extent of the Okavango Delta. This algorithm made use of
+MODIS MCD34A4 satellite imagery, which is a 8-day composite of daily updated
+MODIS satellite data. The benefit of MODIS' high temporal resolution is that
+monthly composites are almost guaranteed to be cloud free, even during
+Botswana's rainy season (mid October to mid May). However, due to its coarse
+spatial resolution of 250 meters, MODIS satellite data was unsuitable to detect
+finer-scaled waterbodies on satellite images. In an attempt to overcome this
+limitation, we evaluated alternative satellite products that provided better
+spatial resolution than MODIS, while retaining a sufficiently high temporal
+resolution to render seasonal patterns. Candidate products comprised Landsat 7,
+Landsat 8, and Sentinel 2 satellite imagery. Data associated with each of these
+satellites are freely accessible and provide spatial resolutions between 10 and
+60 meters, depending on the respective bands. Furthermore, the satellites have
 revisit-times between 5 and 16 days, thus allowing to generate frequently
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htpb]
  \begin{center}
@@
 \noindent Our goal was to produce dynamically updated pan-maps that we could
-overlay with GPS data of dispersing African wild dogs. Since GPS data was
+overlay with GPS data of dispersing African wild dogs. Since most GPS data were
 collected between 2015 and 2022, this implied that satellite data needed to be
@@
 corrections themselves. Fortunately, the European Space Agency provides the
-correction tool \texttt{sen2cor} for free \citep{Main-Knorn.2017}, and this tool
-has been integrated into the \texttt{R}-package \texttt{sen2r}
+correction tool \texttt{sen2cor} for free \citep{Main-Knorn.2017}, which has
+been integrated into the \texttt{R}-package \texttt{sen2r}
 \citep{Ranghetti.2020}. We therefore applied the necessary corrections after
 download for any product that hadn't already been corrected. Furthermore, we
-used Sentinel 2's ``Scene Classification Map'', to remove pixels classified as
-either clouds or shadows. Finally, we computed the same ND-indices as for
-Landsat 8 (\Cref{ND}).
+used Sentinel 2's scene classification, to remove pixels classified as either
+clouds or shadows. Finally, we computed the same ND-indices as for Landsat 8
+(\Cref{ND}).
 
@@
 
-To train a land-cover classifier, we generated 3'000 random points within
+To train a land-cover classifier, we generated 3,000 random points within
 training polygons following a stratified equal random sampling scheme
-\citep{Shetty.2021}. That is, we ensured that 1'000 random points were sampled
+\citep{Shetty.2021}. That is, we ensured that 1,000 random points were sampled
 per training category. Stratified random sampling was necessary to ensure that
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htpb]
  \begin{center}
   \includegraphics[width = \textwidth]{Figures/MovingWindows2.png}
-  \caption{Moving windows and associated tiles facetted by months.}
+  \caption{This figure shows the moving windows generated for each month, as
+  well as the underlying GPS data and overlapping Sentinel 2 tiles that we
+  ultimately downloaded and combined into monthly composites.}
   \label{MovingWindows2}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htpb]
  \begin{center}
   \includegraphics[width = 0.75\textwidth]{Figures/Moonlight.png}
-  \caption{Moonlight intensity (illumination) as estimated from the
-  \texttt{moonlit} package \citep{Smielak.2023} in \texttt{R} calculated for a
-  couple of example timestamps in 2023.}
+  \caption{Moonlight intensity as estimated from the \texttt{moonlit} package
+  \citep{Smielak.2023} in \texttt{R}, calculated for several example timestamps
+  in 2023.}
   \label{Moonlight}
@@
 
-\begin{table}[htbp]
+\begin{table}[htpb]
  \begin{center}
-  \caption{Estimates obtained from the \textbf{simple} (no interactions)
-  integrated step-selection models fitted using \textbf{static} covariates. GPS
-  Data was either pooled across seasons (labeled ``all'') or split into dry and
-  wet season. Only underlined covariates differed between the static and dynamic
-  configurations.}
+  \caption{Estimates obtained using the \textbf{simple model formula} (no
+  interactions) and \textbf{static} covariates. GPS Data were either pooled
+  across seasons (labeled ``all'') or split into wet and dry season. Only
+  underlined covariates differed between the static and dynamic configurations.}
   \label{ModelsStaticSimple}
@@
 
-\begin{table}[htbp]
+\begin{table}[htpb]
  \begin{center}
-  \caption{Estimates obtained from the \textbf{simple} (no interactions)
-  integrated step-selection models fitted using \textbf{dynamic} covariates. GPS
-  Data was either pooled across seasons (labeled ``all'') or split into dry and
-  wet season. Only underlined covariates differed between the static and dynamic
-  configurations.}
+  \caption{Estimates obtained using the \textbf{simple model formula} (no
+  interactions) and \textbf{dynamic} covariates. GPS Data were either pooled
+  across seasons (labeled ``all'') or split into wet and dry season. Only
+  underlined covariates differed between the static and dynamic configurations.}
   \label{ModelsDynamicSimple}
@@
 
-\begin{table}[htbp]
+\begin{table}[htpb]
  \begin{center}
-  \caption{Estimates obtained from the \textbf{full} (with interactions)
-  integrated step-selection models fitted using \textbf{static} covariates. GPS
-  Data was either pooled across seasons (labeled ``all'') or split into dry and
-  wet season. Only underlined covariates differed between the static and dynamic
-  configurations.}
+  \caption{Estimates obtained using the \textbf{complex model formula} (with
+  interactions) and \textbf{static} covariates. GPS Data were either pooled
+  across seasons (labeled ``all'') or split into wet and dry season. Only
+  underlined covariates differed between the static and dynamic configurations.}
   \label{ModelsStaticFull}
@@
 
-\begin{table}[htbp]
+\begin{table}[htpb]
  \begin{center}
-  \caption{Estimates obtained from the \textbf{full} (no interactions)
-  integrated step-selection models fitted using \textbf{dynamic} covariates. GPS
-  Data was either pooled across seasons (labeled ``all'') or split into dry and
-  wet season. Only underlined covariates differed between the static and dynamic
-  configurations.}
+  \caption{Estimates obtained using the \textbf{complex model formula} (with
+  interactions) and \textbf{dynamic} covariates. GPS Data were either pooled
+  across seasons (labeled ``all'') or split into wet and dry season. Only
+  underlined covariates differed between the static and dynamic configurations.}
   \label{ModelsDynamicFull}
@@
   \caption{Results from the integrated step-selection models when considering
-  different numbers of random steps. Results are shown for the simple and full
-  model for both all configurations of fitting covariates (static vs dry) and
-  model seasons (all vs. wet + dry).}
+  different numbers of random steps. Results are shown for the simple and
+  complex formulas for all configurations of fitting covariates (static vs dry)
+  and model seasons (all vs. wet + dry). For some design combinations, models
+  failed to converge and thus dropped from the figure.}
   \label{Stability}
@@
   Confidence intervals overlapped substantially, hence we omitted them from the
-  figure to simplify with the interpretation of the differences among
+  figure to simplify the interpretation of the differences among
   configurations.}

### 2024-07-01 FinalThesis/Chapter_04/Manuscript.tex
Publish PHD

--- 2024-07-01 Chapter_4/04_Manuscript/Manuscript.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-\usepackage[nomarkers, nolists]{endfloat}
-\usepackage{scalerel}
-\usepackage{makecell}
-\usepackage{ifthen}
-\usepackage{tikz}
-\usepackage{lineno}
-\usepackage{adjustbox}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%  Bibtex vs Biblatex
-%------------------------------------------------------------------------------
-% While bibtex is very nice to use, it's insanely slow. I therefore work with
-% bibtex until the final stage. Set the value below to "true" or "false" to pick
-% which one should be used.
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{true}
-
-% Depending on the question, the commands below will prepare the document
-% accordingly
-\ifthenelse{\boolean{usebiblatex}}{
-  %-----------------------------------------------------------------------------
-  %  If using biblatex
-  %-----------------------------------------------------------------------------
-  % Load required packages and refer to correct bibliography
-  \usepackage{csquotes}
-  \usepackage[
-      backend=biber
-    , style=apa          % To display Author-year in the text
-    , natbib=true        % Necessary to use citep, etc.
-    , hyperref=true      % To have clickable links in the references
-    , useeditor=false    % Don't add editor to publications
-    , sorting=ynt        % Sort by year, name, and title
-    , uniquename=false   % Avoid that initials are put
-  ]{biblatex}
-  \addbibresource{Literature.bib}
-
-  % To include author in link
-  \DeclareFieldFormat{citehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{#1}}
-  \DeclareFieldFormat{textcitehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{%
-    #1%
-    \ifbool{cbx:parens}
-      {\bibcloseparen\global\boolfalse{cbx:parens}}
-      {}}}
-  \savebibmacro{cite}
-  \savebibmacro{textcite}
-  \renewbibmacro*{cite}{%
-    \printtext[citehyperref]{%
-      \restorebibmacro{cite}%
-      \usebibmacro{cite}}}
-  \renewbibmacro*{textcite}{%
-    \ifboolexpr{
-      ( not test {\iffieldundef{prenote}} and
-        test {\ifnumequal{\value{citecount}}{1}} )
-      or
-      ( not test {\iffieldundef{postnote}} and
-        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
-    }
-      {\DeclareFieldAlias{textcitehyperref}{noformat}}
-      {}%
-    \printtext[textcitehyperref]{%
-      \restorebibmacro{textcite}%
-      \usebibmacro{textcite}}}
-
-  % Remove editor form everything but books
-  \AtEveryBibitem{%
-   \ifentrytype{book}{}{%
-    \clearname{editor}
-   }
-  }
-}{
-  %-----------------------------------------------------------------------------
-  %  If using bibtex
-  %-----------------------------------------------------------------------------
-  % Load required packages and select citation style
-  \usepackage[round]{natbib}
-  \bibliographystyle{apalike}
-}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Link colors
-\definecolor{linkcolor}{HTML}{000000}
-\hypersetup{
-  colorlinks = true,
-  linkcolor  = linkcolor,
-  urlcolor   = linkcolor,
-  citecolor  = linkcolor,
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% Avoid skip when including text from external .tex file
-\newcommand{\inputy}[1]{\input{#1}\unskip}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%  ORCID
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Create the ORCID logo
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol]
-      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
-        ,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white]
-      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
-        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
-        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
-      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-
-% Command to create the ORCID
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-\title{Methods for Implementing Integrated Step-Selection Functions with
+\documentclass[../FinalThesis.tex]{subfiles}
+\begin{document}
+\chapter{Methods for Implementing Integrated Step-Selection Functions with
 Incomplete Data}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
-  John Fieberg \textsuperscript{3} \orcid{0000-0002-3180-7021}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
-\begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
-
-\textsuperscript{3} Department of Fisheries, Wildlife, and Conservation Biology,
-University of Minnesota, St. Paul, MN, USA.
-
-\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
-
-\vspace{4cm}
-
-\textbf{Running Title:} Step-Selection Analyses with Missing Data
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} animal movement, gps data, imputation, incomplete data,
-missing fixes, step-selection analyses, step-selection functions
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Abstract
-%------------------------------------------------------------------------------
+\label{ChapterIrregularity}
+\thispagestyle{empty}
+\vspace{-1cm}
+\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil
+
+\begin{center}
+David D. Hofmann \orcid{0000-0003-3477-4365},
+Gabriele Cozzi \orcid{0000-0002-1744-1940}, and
+John Fieberg \orcid{0000-0002-3180-7021}
+
+% \begin{flushleft}
+
+% \vspace{5cm}
+\vfill
+
+Published in Movement Ecology (2024):
+\url{https://doi.org/10.1186/s40462-024-00476-8}.
+
+% \textsuperscript{1} Department of Evolutionary Biology and Environmental
+% Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
+% Switzerland.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
+% Private Bag 13, Maun, Botswana.
+%
+% \vspace{0.25cm}
+%
+% \textsuperscript{\S} Corresponding author:
+% \href{mailto:david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
+
+\end{center}
+
+% \vfill
+
+% \textbf{Running Title:}\\African Wild Dog Dispersal in a Changing Climate:
+% Lessons Learned from Seasonal Flood Extremes
+%
+% \vspace{0.5cm}
+%
+% \textbf{Keywords:}\\movement, connectivity, climate change, seasonality,
+% dispersal, conservation, individual-based simulations, step-selection
+
+% \end{flushleft}
+
 \newpage
-\begin{abstract}
+\section*{Abstract}
 
@@
 
-\end{abstract}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
-% \newpage
-%
-% \onehalfspacing
-% \tableofcontents
-% \doublespacing
-
-% Change page numbering
 \newpage
-\pagenumbering{arabic}
-
-% Create linenumbers
-\linenumbers
-
 \section{Introduction}
@@
 \begin{equation}
-\label{EQ1}
+\label{EQ3}
 u(s_{t+1}) = \frac{\phi(s_{t+1}, s_t, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)}{\int_{s \in G}\phi(s_{t+1}, s_{t}, s_{t-1}; \gamma)w(x(s_{t+1}); \beta)ds}
@@
 function of $x$, taking the form $w = exp(x^\top\beta)$. The integral in the
-denominator of \ref{EQ1} ensures that $u$ is a proper probability distribution
+denominator of \ref{EQ3} ensures that $u$ is a proper probability distribution
 (i.e., that it integrates to 1). Following \citet{Michelot.2024}, we call the
@@
 
+% Glossary of terms
+\input{Figures/Glossary.tex}
+
 Given a series of \textbf{observed steps}, finding the movement and
-habitat-selection parameters that maximize the likelihood in \ref{EQ1} requires
+habitat-selection parameters that maximize the likelihood in \ref{EQ3} requires
 approximating the integral in the denominator for each observed step. A variety
@@
 SSF, termed \textit{integrated} SSA (or iSSA) by \citet{Avgar.2016}, is similar
-to using importance sampling to approximate the integral in \ref{EQ1}
+to using importance sampling to approximate the integral in \ref{EQ3}
 \citep{Michelot.2024} and is readily accessible through the \texttt{R}-package
@@
 %Avgar.2016}. In SSFs, the use of random steps serves to generate integration
-%points that are needed to approximate the integral in \ref{EQ1} and, even
+%points that are needed to approximate the integral in \ref{EQ3} and, even
 %though the true movement kernel $\phi$ is often not known a priori, it can be
@@
 that can be used to simulate space use under novel conditions \citep{Avgar.2016,
-Signer.2017, Hofmann.2023, Signer.2023}. This characteristic has made iSSAs a
+Signer.2017, Hofmann.2023, Signer.2024}. This characteristic has made iSSAs a
 useful tool for quantifying landscape resistance and identifying movement
 corridors \citep{Zeller.2020, Buchholtz.2020, Hofmann.2021, Hofmann.2023}.
-
-% Glossary of terms
-\input{99_Glossary.tex}
 
@@
 
-\begin{figure}
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Overview.png}
-  \caption{(\textbf{a}) Demonstration of how missingness affects the number of
-  valid steps that can be used for step-selection analyses under different
-  levels of forgiveness. The upper panel depicts a trajectory with zero missing
-  locations. That is, all aspired locations were successfully collected on a
-  regular interval (yielding a regular step duration of $\Delta t = 1$). This
-  trajectory produces four valid steps that can be included in the iSSF model
-  and one invalid step that has to be omitted because it has no turning angle
-  associated with it. In the central panel, animal location 4 was not obtained,
-  introducing a missingness of 16.7\%. If the modeler has a forgiveness of one,
-  only a single step can be included for further analysis, as all other steps
-  are invalid (either because no turning angle can be computed or because
-  step durations exceed the forgiveness). If, however, the modeler exhibited a
+  \includegraphics[width = \textwidth]{Figures/Overview.png}
+  \caption{(a) Demonstration of how missingness affects the number of valid
+  steps that can be used for step-selection analyses under different levels of
+  forgiveness. The upper panel depicts a trajectory with zero missing locations.
+  That is, all aspired locations were successfully collected on a regular
+  interval (yielding a regular step duration of $\Delta t = 1$). This trajectory
+  produces four valid steps that can be included in the iSSF model and one
+  invalid step that has to be omitted because it has no turning angle associated
+  with it. In the central panel, animal location 4 was not obtained, introducing
+  a missingness of 16.7\%. If the modeler has a forgiveness of one, only a
+  single step can be included for further analysis, as all other steps are
+  invalid (either because no turning angle can be computed or because step
+  durations exceed the forgiveness). If, however, the modeler exhibited a
   forgiveness of two, such as in the lower panel, a total of three steps could
-  be obtained for further analysis. (\textbf{b}) Conceptual illustration of how
+  be obtained for further analysis. (b) Conceptual illustration of how
   increasing the forgiveness allows one to retain additional steps that can be
@@
   of a new burst, which leads to the loss of several steps. In addition, some of
-  the remaining steps are invalid (dotted) because they are lacking a
-  turning angle. By increasing the forgiveness, a modeler is willing to retain
-  steps that exceed the regular step duration by a certain threshold, which
-  enables them to obtain longer bursts and increase the number of steps that can
-  be used for further analysis. In the figure, forgiveness increases from left
-  to right.}
+  the remaining steps are invalid (dotted) because they are lacking a turning
+  angle. By increasing the forgiveness, a modeler is willing to retain steps
+  that exceed the regular step duration by a certain threshold, which enables
+  them to obtain longer bursts and increase the number of steps that can be used
+  for further analysis. In the figure, forgiveness increases from left to
+  right.}
   \label{Overview}
@@
 
-\begin{figure}
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_NumberOfSteps.png}
+  \includegraphics[width = \textwidth]{Figures/NumberOfSteps.png}
   \caption{Illustration of how missingness in animal location data reduces the
@@
   \textit{imputation} approach, is readily available through the
-  \texttt{R}-package \texttt{crawl} \citep{Johnson.2022}, yet has only been
+  \texttt{R}-package \texttt{crawl} \citep{Johnson.2024}, yet has only been
   tested for use with hidden Markov movement models and not with iSSFs
@@
 
-\section{Methods}
+\section{Materials and Methods}
 
@@
 simulation-runs using the \texttt{R}-package \texttt{pbmcapply}
-\citep{Kuang.2022}. We generated figures using the \texttt{ggplot2}
-\citep{Wickham.2023}, \texttt{ggpubr} \citep{Kassambara.2023}, and
-\texttt{ggh4x} \citep{Brand.2023} \texttt{R}-packages. We manipulated raster
+\citep{Kuang.2024}. We generated figures using the \texttt{ggplot2}
+\citep{Wickham.2024}, \texttt{ggpubr} \citep{Kassambara.2024}, and
+\texttt{ggh4x} \citep{Brand.2024} \texttt{R}-packages. We manipulated raster
 data and computed spatial distances using the \texttt{R}-package \texttt{raster}
-\citep{Hijmans.2023}. An overview of the simulation design and the different
+\citep{Hijmans.2024}. An overview of the simulation design and the different
 iSSF approaches is presented in \Cref{Design} and all codes to reproduce this
@@
 
-\begin{figure}
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Design.png}
+  \includegraphics[width = \textwidth]{Figures/Design.png}
   \caption{Illustration of the study design. We varied the autocorrelation range
@@
 window with radius $r$ within which we tallied pixel-values. We varied $r$ from
-5, to 10, to 20, depending on the simulated level of autocorrelation (Figure
-S1). The third layer, \textsf{Forest} (categorical), represented areas covered
-by woodland and was simulated similarly to the \textsf{Elev} layer, but we
-binarized the layer by setting all simulated values above the 50\% quantile to
-forest and all other values to non-forest (our reference class). We normalized
-values of all simulated layers to a range between zero and one and replicated
-the simulation of each layer 100 times per autocorrelation scenario, thus
-producing 300 unique landscape configurations.
-
-\begin{figure}
+5, to 10, to 20, depending on the simulated level of autocorrelation
+(\Cref{SimulatedCovariates}). The third layer, \textsf{Forest} (categorical),
+represented areas covered by woodland and was simulated similarly to the
+\textsf{Elev} layer, but we binarized the layer by setting all simulated values
+above the 50\% quantile to forest and all other values to non-forest (our
+reference class). We normalized values of all simulated layers to a range
+between zero and one and replicated the simulation of each layer 100 times per
+autocorrelation scenario, thus producing 300 unique landscape configurations.
+
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_Covariates.png}
+  \includegraphics[width = \textwidth]{Figures/Covariates.png}
   \caption{Example of a landscape configuration across which we simulated
@@
 \begin{equation}
-\label{EQ2}
+\label{EQ4}
   \pi_j =
@@
   distributions. To facilitate estimation of dynamic distribution parameters
-  across all $\Delta t$ (Figure S2), we resampled data to different step
-  durations using the \texttt{track\_resample} function from the \texttt{amt}
-  package \citep{Signer.2019} before fitting tentative parameters. This ensured
-  a sufficient number of steps for each step duration to estimate associated
-  parameters. An alternative approach would be to increase missingness in the
-  data even further, thus introducing a larger number of longer step durations.
+  across all $\Delta t$ (\Cref{DistributionParameters}), we resampled data to
+  different step durations using the \texttt{track\_resample} function from the
+  \texttt{amt} package \citep{Signer.2019} before fitting tentative parameters.
+  This ensured a sufficient number of steps for each step duration to estimate
+  associated parameters. An alternative approach would be to increase
+  missingness in the data even further, thus introducing a larger number of
+  longer step durations.
 
@@
   \texttt{crwFit} and \texttt{crwPredict} from the \texttt{crawl}
-  \texttt{R}-package \citep{Johnson.2022}. Although the \texttt{crwFit} function
+  \texttt{R}-package \citep{Johnson.2024}. Although the \texttt{crwFit} function
   provides capabilities to incorporate location measurement error, we assumed
@@
 implemented using the \texttt{clogit} function in the \texttt{R}-package
-\texttt{survival} \citep{Therneau.2023}. We defined a binary response variable
+\texttt{survival} \citep{Therneau.2024}. We defined a binary response variable
 (\textsf{observed}) indicating if a step was an observed (scored 1) or a random
@@
 Results were qualitatively similar for all three landscape autocorrelation
-scenarios and for different combinations of missingness and forgiveness (Figure
-S4). Here, we report on results for a landscape with autocorrelation of 20,
-while either holding constant missingness at a conservative 20\%
-(\Cref{ModelComparisonForgiveness}) or the forgiveness level at two
-(\Cref{ModelComparisonMissingness}) (results for all other combinations are
-summarized in Figure S4). The \textit{imputation} approach resulted in biased
-estimators of $\beta_{dist}$ and $\beta_{forest}$, whereas all other approaches
-were able to recover the parameters of the habitat-selection function with
-minimal bias (\Cref{ModelComparisonForgiveness}). Note, the \textit{imputation}
-approach always starts with a full trajectory and is therefore unaffected by the
+scenarios and for different combinations of missingness and forgiveness
+(\Cref{ModelComparison}). Here, we report on results for a landscape with
+autocorrelation of 20, while either holding constant missingness at a
+conservative 20\% (\Cref{ModelComparisonForgiveness}) or the forgiveness level
+at two (\Cref{ModelComparisonMissingness}) (results for all other combinations
+are summarized in \Cref{ModelComparison}). The \textit{imputation} approach
+resulted in biased estimators of $\beta_{dist}$ and $\beta_{forest}$, whereas
+all other approaches were able to recover the parameters of the
+habitat-selection function with minimal bias
+(\Cref{ModelComparisonForgiveness}). Note, the \textit{imputation} approach
+always starts with a full trajectory and is therefore unaffected by the
 forgiveness level. For all other methods, increasing the forgiveness from 1 to 5
@@
 
-\begin{figure}
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_ModelComparisonForgiveness.png}
-  \caption{(\textbf{a}) Parameter estimates and (\textbf{b}) Root mean-square
-  error (RMSE) with regard to the movement kernel and habitat-selection function
-  as a function of forgiveness. Results are shown for the scenario with
-  landscape autocorrelation of 20 and missingness of 20\%. The movement kernel
-  comprised of a gamma distribution with shape parameter $k$ and scale parameter
-  $\theta$ governing the step-length distribution and a von Mises distribution
-  with concentration parameter $\kappa$ governing the turning-angle
-  distribution. Habitat-selection was based on three covariates, namely a
-  \textsf{Distance}, \textsf{Elevation}, and a \textsf{Forest} layer. Estimates
-  are shown for the five different approaches we tested for. The uncorrected
-  approach ignored the fact that higher forgiveness implied temporal
-  irregularity in the data, while all other approaches attempted to correct for
-  the potential biases introduced by temporal irregularity. Note, the imputation
-  approach is not affected by the forgiveness level, since it always starts with
-  a full trajectory.}
+  \includegraphics[width = \textwidth]{Figures/ModelComparisonForgiveness.png}
+  \caption{(a) Parameter estimates and (b) Root mean-square error (RMSE) with
+  regard to the movement kernel and habitat-selection function as a function of
+  forgiveness. Results are shown for the scenario with landscape autocorrelation
+  of 20 and missingness of 20\%. The movement kernel comprised of a gamma
+  distribution with shape parameter $k$ and scale parameter $\theta$ governing
+  the step-length distribution and a von Mises distribution with concentration
+  parameter $\kappa$ governing the turning-angle distribution. Habitat-selection
+  was based on three covariates, namely a \textsf{Distance}, \textsf{Elevation},
+  and a \textsf{Forest} layer. Estimates are shown for the five different
+  approaches we tested for. The uncorrected approach ignored the fact that
+  higher forgiveness implied temporal irregularity in the data, while all other
+  approaches attempted to correct for the potential biases introduced by
+  temporal irregularity. Note, the imputation approach is not affected by the
+  forgiveness level, since it always starts with a full trajectory.}
   \label{ModelComparisonForgiveness}
@@
 
-\begin{figure}
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_ModelComparisonMissingness.png}
-  \caption{(\textbf{a}) Parameter estimates and (\textbf{b}) Root mean-square
-  error (RMSE) with regard to the movement kernel and habitat-selection function
-  as a function of missingness. Results are shown for the scenario with
-  landscape autocorrelation of 20 and forgiveness of 2. The movement kernel
-  comprised of a gamma distribution with shape parameter $k$ and scale parameter
-  $\theta$ governing the step-length distribution and a von Mises distribution
-  with concentration parameter $\kappa$ governing the turning-angle
-  distribution. Habitat-selection was based on three covariates, namely a
-  \textsf{Distance}, \textsf{Elevation}, and a \textsf{Forest} layer. Estimates
-  are shown for the five different approaches we tested for. The uncorrected
-  approach ignored the fact that higher forgiveness implied temporal
-  irregularity in the data, while all other approaches attempted to correct for
-  the potential biases introduced by temporal irregularity. Note, the imputation
-  approach is not affected by the forgiveness level, since it always starts with
-  a full trajectory.}
+  \includegraphics[width = \textwidth]{Figures/ModelComparisonMissingness.png}
+  \caption{(a) Parameter estimates and (b) Root mean-square error (RMSE) with
+  regard to the movement kernel and habitat-selection function as a function of
+  missingness. Results are shown for the scenario with landscape autocorrelation
+  of 20 and forgiveness of 2. The movement kernel comprised of a gamma
+  distribution with shape parameter $k$ and scale parameter $\theta$ governing
+  the step-length distribution and a von Mises distribution with concentration
+  parameter $\kappa$ governing the turning-angle distribution. Habitat-selection
+  was based on three covariates, namely a \textsf{Distance}, \textsf{Elevation},
+  and a \textsf{Forest} layer. Estimates are shown for the five different
+  approaches we tested for. The uncorrected approach ignored the fact that
+  higher forgiveness implied temporal irregularity in the data, while all other
+  approaches attempted to correct for the potential biases introduced by
+  temporal irregularity. Note, the imputation approach is not affected by the
+  forgiveness level, since it always starts with a full trajectory.}
   \label{ModelComparisonMissingness}
@@
 of the obtained locations. As spatial covariate layers, we used \textsf{Water},
-\textsf{DistanceToWater} and \textsf{Trees} (Figure S5). \textsf{Water} was a
-binary variable representing major rivers and areas inundated by floodwater,
-whereas \textsf{DistanceToWater} was a continuous variable indicating the
-distance (in meters) to the nearest pixel categorized as water. \textsf{Trees}
-was a continuous variable indicating the percent tree cover in each pixel. We
-resampled all layers to a common resolution of 250 m by 250 m and merged them
-into a single raster-stack (Figure S5). The derivation of each covariate layer
-is described in detail in \citealp{Hofmann.2021}. We dynamically fitted
-step-length and turning-angle distributions to steps with step durations of 2,
-4, and 6 hours, respectively, assuming a gamma distribution for step lengths and
-a von Mises distribution for turning angles. Instead of resampling the observed
-track to different step durations when fitting dynamic distributions (like we
-did in the simulation study), we introduced a larger amount of steps with step
-durations longer than two hours by thinning the data again (by another 10\%).
-The benefit of this approach was that steps with irregular step durations
-occurred more randomly and were not limited to the hours specified by the
-resampling algorithm. Finally, we used iSSFs with the \textit{dynamic+model}
-approach to estimate the habitat-selection function and movement kernel of
-Apollo. For this, we considered three cases:
+\textsf{DistanceToWater} and \textsf{Trees} (\Cref{CaseStudyCovariates}).
+\textsf{Water} was a binary variable representing major rivers and areas
+inundated by floodwater, whereas \textsf{DistanceToWater} was a continuous
+variable indicating the distance (in meters) to the nearest pixel categorized as
+water. \textsf{Trees} was a continuous variable indicating the percent tree
+cover in each pixel. We resampled all layers to a common resolution of 250 m by
+250 m and merged them into a single raster-stack (\Cref{CaseStudyCovariates}).
+The derivation of each covariate layer is described in detail in
+\citealp{Hofmann.2021}. We dynamically fitted step-length and turning-angle
+distributions to steps with step durations of 2, 4, and 6 hours, respectively,
+assuming a gamma distribution for step lengths and a von Mises distribution for
+turning angles. Instead of resampling the observed track to different step
+durations when fitting dynamic distributions (like we did in the simulation
+study), we introduced a larger amount of steps with step durations longer than
+two hours by thinning the data again (by another 10\%). The benefit of this
+approach was that steps with irregular step durations occurred more randomly and
+were not limited to the hours specified by the resampling algorithm. Finally, we
+used iSSFs with the \textit{dynamic+model} approach to estimate the
+habitat-selection function and movement kernel of Apollo. For this, we
+considered three cases:
 
@@
 then fit the three models using the conditional logistic regression framework as
-implemented in the \texttt{survival} \texttt{R}-package \citep{Therneau.2023}.
+implemented in the \texttt{survival} \texttt{R}-package \citep{Therneau.2024}.
 Lastly, we computed updated movement parameters for a regular step duration of 2
@@
 to improvements in estimator precision (\Cref{CaseStudy}). This was achieved by
-increasing the effective sample size from 2,179 to 4,505 valid steps (Table S1).
-The improvement in estimator precision was weaker for F3-SH than for F3-S, as
-the F3-SH model was more complex due to inclusion of additional interaction
-terms. Point estimates for the habitat-selection and movement parameters were
-similar for all 3 models, and evidence for scale dependency in habitat selection
-was fairly weak. F3-S and F3-SH had similar AIC scores ($\Delta AIC \leq 1$;
-Table S1), and the interaction terms were statistically significant only for the
-step duration of 6 hours and only for one of the habitat covariates (Table S1).
-
-\begin{figure}
+increasing the effective sample size from 2,179 to 4,505 valid steps
+(\Cref{CaseStudyTable}). The improvement in estimator precision was weaker for
+F3-SH than for F3-S, as the F3-SH model was more complex due to inclusion of
+additional interaction terms. Point estimates for the habitat-selection and
+movement parameters were similar for all 3 models, and evidence for scale
+dependency in habitat selection was fairly weak. F3-S and F3-SH had similar AIC
+scores ($\Delta AIC \leq 1$; \Cref{CaseStudyTable}), and the interaction terms
+were statistically significant only for the step duration of 6 hours and only
+for one of the habitat covariates (\Cref{CaseStudyTable}).
+
+\begin{figure}[htpb]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_CaseStudy.png}
+  \includegraphics[width = \textwidth]{Figures/CaseStudy.png}
   \caption{Model results from the case study using GPS data collected on Apollo.
@@
 duration and the (tentative) movement parameters $\kappa_0$ and $k_0$ were
-non-linear (Figure S4). Overall, the usefulness of this approach appears highly
-limited, as it is often not clear by what factor distributional parameters for
-step lengths and turning angles should be multiplied to match the observed step
-duration.
+non-linear (\Cref{DistributionParameters}). Overall, the usefulness of this
+approach appears highly limited, as it is often not clear by what factor
+distributional parameters for step lengths and turning angles should be
+multiplied to match the observed step duration.
 
@@
 turning angles are influenced by the step duration of both the current and
-previous step (Figure S3). The bias in the concentration parameter likely arose
-from only accounting for the step duration associated with the current step and
-not the previous one. Moreover, fitting tentative distributions for different
-step durations can be challenging due to some step durations occurring only
-rarely. However, by resampling observed animal locations to different step
-durations using the \texttt{track\_resample} function from the \texttt{amt}
-\texttt{R}-package \citep{Signer.2019} the needed data can easily be generated.
-We included step duration as a categorical covariate, yet there may be times
-when it would be advantageous to treat it as a continuous covariate (e.g., with
-its effect modeled using a low-degree polynomial or regression spline with few
-degrees of freedom). Treating step duration as a continuous variable may help to
-alleviate convergence issues in cases where some step durations are rare, and it
-might allow applying the \textit{dynamic+model} approach to data that are
-entirely irregular.
+previous step (\Cref{PreviousDuration}). The bias in the concentration parameter
+likely arose from only accounting for the step duration associated with the
+current step and not the previous one. Moreover, fitting tentative distributions
+for different step durations can be challenging due to some step durations
+occurring only rarely. However, by resampling observed animal locations to
+different step durations using the \texttt{track\_resample} function from the
+\texttt{amt} \texttt{R}-package \citep{Signer.2019} the needed data can easily
+be generated. We included step duration as a categorical covariate, yet there
+may be times when it would be advantageous to treat it as a continuous covariate
+(e.g., with its effect modeled using a low-degree polynomial or regression
+spline with few degrees of freedom). Treating step duration as a continuous
+variable may help to alleviate convergence issues in cases where some step
+durations are rare, and it might allow applying the \textit{dynamic+model}
+approach to data that are entirely irregular.
 
@@
 missingness and forgiveness. Both methods are easy to implement, and the
-associated models can readily be fitted using the R-packages \texttt{amt}
-\citep{Signer.2019}, \texttt{survival} \citep{Therneau.2023}, \texttt{coxme}
-\citep{Therneau.2022}, and \texttt{mgcv} \citep{Wood.2011, Wood.2017,
-Klappstein.2024}. To facilitate uptake and encourage use of the proposed
-approaches among practitioners, we provide all of our codes through an online
-repository, which includes an example application of the \textit{dynamic+model}
-approach. With this, we hope practitioners will rethink the common use of
-discarding large portions of data and instead use methods that can accommodate
-irregular data.
-
-\section{Authors' Contributions}
-
-D.D.H., G.C., and J.F. conceived the study and designed methodology; D.D.H.
-implemented the analysis, J.F. assisted with modeling; D.D.H. wrote the first
-draft of the manuscript and all authors contributed to the drafts at several
-stages and gave final approval for publication.
-
-\section{Data Availability}
-
-Code to reproduce this study is available through the University of Minnesota's
-Data repository (\url{https://doi.org/10.13020/6wcd-6s43},
-\citealp{Hofmann.2023a}). The repository also includes two example analyses that
-showcase the application of the \textit{dynamic+model} approach to simulated
-data, as well as to the GPS data of Apollo.
-
-\section{Acknowledgements \& Funding}
-
-D.D.H. was funded by a Swiss National Science Foundation Grant No.
-310030\_204478 rewarded to G.C. J.F. was supported by the National Aeronautics
-and Space Administration award 80NSSC21K1182 and received partial salary support
-from the Minnesota Agricultural Experimental Station. We thank two anonymous
-reviewers for their constructive and detailed feedback, which helped to improve
-the quality of this manuscript.
-
+associated models can readily be fitted using the \texttt{R}-packages
+\texttt{amt} \citep{Signer.2019}, \texttt{survival} \citep{Therneau.2024},
+\texttt{coxme} \citep{Therneau.2024a}, and \texttt{mgcv} \citep{Wood.2011,
+Wood.2017, Klappstein.2024}. To facilitate uptake and encourage use of the
+proposed approaches among practitioners, we provide all of our codes through an
+online repository, which includes an example application of the
+\textit{dynamic+model} approach. With this, we hope practitioners will rethink
+the common use of discarding large portions of data and instead use methods that
+can accommodate irregular data.
+
+% \section{Authors' Contributions}
+%
+% D.D.H., G.C., and J.F. conceived the study and designed methodology; D.D.H.
+% implemented the analysis, J.F. assisted with modeling; D.D.H. wrote the first
+% draft of the manuscript and all authors contributed to the drafts at several
+% stages and gave final approval for publication.
+%
+% \section{Data Availability}
+%
+% Code to reproduce this study is available through the University of Minnesota's
+% Data repository (\url{https://doi.org/10.13020/6wcd-6s43},
+% \citealp{Hofmann.2023a}). The repository also includes two example analyses that
+% showcase the application of the \textit{dynamic+model} approach to simulated
+% data, as well as to the GPS data of Apollo.
+%
+% \section{Acknowledgements \& Funding}
+%
+% D.D.H. was funded by a Swiss National Science Foundation Grant No.
+% 310030\_204478 rewarded to G.C. J.F. was supported by the National Aeronautics
+% and Space Administration award 80NSSC21K1182 and received partial salary support
+% from the Minnesota Agricultural Experimental Station. We thank two anonymous
+% reviewers for their constructive and detailed feedback, which helped to improve
+% the quality of this manuscript.
+
+\ifSubfilesClassLoaded{%
   \newpage
-\begingroup
-\singlespacing
+  \begin{singlespacing}
   \ifthenelse{\boolean{usebiblatex}}{

### 2024-07-01 FinalThesis/Chapter_04/Manuscript_Support.tex
Publish PHD

--- 2024-07-01 Chapter_4/04_Manuscript/Manuscript_Support.tex
@@
-\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
-\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
-\usepackage[doublespacing]{setspace}
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{colortbl}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{enumerate}
-\usepackage{enumitem}
-\usepackage{subcaption}
-\usepackage{booktabs}
-\usepackage[hidelinks]{hyperref}
-\usepackage[nameinlink]{cleveref}
-\usepackage{multirow}
-\usepackage{arydshln}
-\usepackage[flushleft]{threeparttable}
-% \usepackage[nomarkers, nolists]{endfloat}
-\usepackage{scalerel}
-\usepackage{makecell}
-\usepackage{ifthen}
-\usepackage{tikz}
-\usepackage{lineno}
-\usepackage{adjustbox}
-\usetikzlibrary{svg.path}
-
-%------------------------------------------------------------------------------
-%  Bibtex vs Biblatex
-%------------------------------------------------------------------------------
-% While bibtex is very nice to use, it's insanely slow. I therefore work with
-% bibtex until the final stage. Set the value below to "true" or "false" to pick
-% which one should be used.
-\newboolean{usebiblatex}
-\setboolean{usebiblatex}{true}
-
-% Depending on the question, the commands below will prepare the document
-% accordingly
-\ifthenelse{\boolean{usebiblatex}}{
-  %-----------------------------------------------------------------------------
-  %  If using biblatex
-  %-----------------------------------------------------------------------------
-  % Load required packages and refer to correct bibliography
-  \usepackage{csquotes}
-  \usepackage[
-      backend=biber
-    , style=apa          % To display Author-year in the text
-    , natbib=true        % Necessary to use citep, etc.
-    , hyperref=true      % To have clickable links in the references
-    , useeditor=false    % Don't add editor to publications
-    , sorting=ynt        % Sort by year, name, and title
-    , uniquename=false   % Avoid that initials are put
-  ]{biblatex}
-  \addbibresource{Literature.bib}
-
-  % To include author in link
-  \DeclareFieldFormat{citehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{#1}}
-  \DeclareFieldFormat{textcitehyperref}{%
-  \DeclareFieldAlias{bibhyperref}{noformat}% Avoid nested links
-  \bibhyperref{%
-    #1%
-    \ifbool{cbx:parens}
-      {\bibcloseparen\global\boolfalse{cbx:parens}}
-      {}}}
-  \savebibmacro{cite}
-  \savebibmacro{textcite}
-  \renewbibmacro*{cite}{%
-    \printtext[citehyperref]{%
-      \restorebibmacro{cite}%
-      \usebibmacro{cite}}}
-  \renewbibmacro*{textcite}{%
-    \ifboolexpr{
-      ( not test {\iffieldundef{prenote}} and
-        test {\ifnumequal{\value{citecount}}{1}} )
-      or
-      ( not test {\iffieldundef{postnote}} and
-        test {\ifnumequal{\value{citecount}}{\value{citetotal}}} )
-    }
-      {\DeclareFieldAlias{textcitehyperref}{noformat}}
-      {}%
-    \printtext[textcitehyperref]{%
-      \restorebibmacro{textcite}%
-      \usebibmacro{textcite}}}
-
-  % Remove editor form everything but books
-  \AtEveryBibitem{%
-   \ifentrytype{book}{}{%
-    \clearname{editor}
-   }
-  }
-}{
-  %-----------------------------------------------------------------------------
-  %  If using bibtex
-  %-----------------------------------------------------------------------------
-  % Load required packages and select citation style
-  \usepackage[round]{natbib}
-  \bibliographystyle{apalike}
-}
-
-%------------------------------------------------------------------------------
-%	Some Styling
-%------------------------------------------------------------------------------
-% Link colors
-\definecolor{linkcolor}{HTML}{000000}
-\hypersetup{
-  colorlinks = true,
-  linkcolor  = linkcolor,
-  urlcolor   = linkcolor,
-  citecolor  = linkcolor,
-}
-
-% Changing the style of captions in figures etc.
-\captionsetup{labelfont=bf, format=plain, font=small}
-
-% Change how equations are referenced
-\renewcommand{\theequation}{Equation \arabic{equation}}%
-
-% Avoid skip when including text from external .tex file
-\newcommand{\inputy}[1]{\input{#1}\unskip}
-
-% For supplementary material
-\newcommand{\beginappendix}{%
-  \setcounter{table}{0}
-  \renewcommand{\thetable}{S\arabic{table}}%
-  \setcounter{figure}{0}
-  \renewcommand{\thefigure}{S\arabic{figure}}%
-  \setcounter{equation}{0}
-  \renewcommand{\theequation}{Equation S\arabic{equation}}%
-  \setcounter{section}{0}
-  \renewcommand{\thesection}{A.\arabic{section}}%
-}
-
-%------------------------------------------------------------------------------
-%  ORCID
-%------------------------------------------------------------------------------
-% Creating some TikZ styles
-\tikzset{
-  nonterminal/.style = {rectangle
-    , minimum size = 6mm
-    , very thick
-    , draw = black!
-  }
-}
-
-% Create the ORCID logo
-\definecolor{orcidlogocol}{HTML}{A6CE39}
-\tikzset{
-  orcidlogo/.pic={
-    \fill[orcidlogocol]
-      svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3
-        ,0,128,0C198.7,0,256,57.3,256,128z};
-    \fill[white]
-      svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
-      svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8
-        ,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5
-        ,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
-      svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6
-      ,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
-  }
-}
-
-% Command to create the ORCID
-\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
-\begin{tikzpicture}[yscale=-1,transform shape]
-  \pic{orcidlogo};
-\end{tikzpicture}
-}{|}}}}
-
-%------------------------------------------------------------------------------
-%	Titlepage: Header
-%------------------------------------------------------------------------------
-\title{\textbf{Appendix}\\ Methods for Implementing Integrated Step-Selection
-Functions with Incomplete Data}
-
-% List of Authors
-\author{
-  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
-  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
-  John Fieberg \textsuperscript{3} \orcid{0000-0002-3180-7021}
-}
-
-% Reduce spacing between authors
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip -0.5em \@plus.17fil\relax
-  \begin{tabular}[t]{c}}
-\makeatother
-
-% Current Date
-% \date{\today}
-
-% And here the masterpiece begins
+\documentclass[../FinalThesis.tex]{subfiles}
 \begin{document}
-
-% Change page numbering
-\pagenumbering{gobble}
-
-% Create Titlepage
-\maketitle
-
-%------------------------------------------------------------------------------
-%	Titlepage: Additional Info
-%------------------------------------------------------------------------------
-\begin{flushleft}
-
-\vspace{0.5cm}
-
-\textsuperscript{1} Department of Evolutionary Biology and Environmental
-Studies, University of Zurich, Winterthurerstrasse 190, 8057 Zurich,
-Switzerland.
-
-\textsuperscript{2} Botswana Predator Conservation Program, Wild Entrust,
-Private Bag 13, Maun, Botswana.
-
-\textsuperscript{3} Department of Fisheries, Wildlife, and Conservation Biology,
-University of Minnesota, St. Paul, MN, USA.
-
-\textsuperscript{\S} Corresponding author: \href{mailto://david.hofmann2@uzh.ch}{david.hofmann2@uzh.ch}
-
-\vspace{4cm}
-
-\textbf{Running Title:} Step-Selection Analyses with Missing Data
-
-\vspace{0.5cm}
-
-\textbf{Keywords:} animal movement, gps data, imputation, incomplete data,
-missing fixes, step-selection analyses, step-selection functions
-
-\end{flushleft}
-
-%------------------------------------------------------------------------------
-%	Main Text
-%------------------------------------------------------------------------------
-\newpage
-
-% Change page numbering
-\pagenumbering{arabic}
-
-% Create linenumbers
-% \linenumbers
-
-% Change to appendix counters
-\appendix
-\beginappendix
 
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_CovariatesAppendix.png}
+  \includegraphics[width = \textwidth]{Figures/CovariatesAppendix.png}
   \caption{Simulated landscapes under different levels of autocorrelation (5,
@@
   unique landscape configurations.}
-  \label{Covariates}
+  \label{SimulatedCovariates}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width = 0.9\textwidth]{99_DistributionParameters.png}
+  \includegraphics[width = 0.9\textwidth]{Figures/DistributionParameters.png}
   \caption{Tentative prameter estimates for the von Mises distribution (top row) and
@@
   \begin{center}
-  \includegraphics[width = 0.9\textwidth]{99_TurningAnglePreviousDuration.png}
+  \includegraphics[width = 0.8\textwidth]{Figures/TurningAnglePreviousDuration.png}
   \caption{Density of relative turning angles associated with steps of $\Delta t
@@
   \begin{center}
-  \includegraphics[width = \textwidth]{99_ModelComparisonAppendix.png}
+  \includegraphics[width = 0.9\textwidth]{Figures/ModelComparisonAppendix.png}
   \caption{Parameter estimates across different autocorrelation scenarios (5,
@@
 
-\begin{figure}[htpb]
+\begin{figure}[!ht]
   \begin{center}
-  \includegraphics[width = \textwidth]{99_CaseStudyCovariates.png}
+  \includegraphics[width = \textwidth]{Figures/CaseStudyCovariates.png}
   \caption{Covariates used for the case study, overlaid with the GPS data of a
@@
 
-\begin{table}[htpb]
+\begin{table}[!ht]
   \begin{center}
@@
   \label{CaseStudyTable}
-    \begin{adjustbox}{max width=0.70\textwidth}
+    \begin{adjustbox}{max width=0.65\textwidth}
     \begin{threeparttable}[h]
-      \input{99_CaseStudyTable.tex}
+      \input{Figures/CaseStudyTable.tex}
       \begin{tablenotes}

### 2024-07-01 FinalThesis/Chapter_05/Manuscript.tex
Publish PHD

\documentclass[../FinalThesis.tex]{subfiles}
\begin{document}
\chapter{General Discussion}
\label{GeneralDiscussion}
\thispagestyle{empty}
\vspace{-1cm}
\noindent\hfil\rule{0.75\textwidth}{.4pt}\hfil

\newpage

\section{A Fresh Take on Connectivity}

In this thesis, I explored several critical factors linked to various aspects of
dispersal and landscape connectivity. In \Cref{ChapterSimulation}, I introduced
a novel approach to assess functional landscape connectivity via simulated
dispersal from iSSFs and demonstrated its application using GPS data from
dispersing AWDs. In \Cref{ChapterFlood}, I used the newly developed approach to
reveal potential differences in connectivity for dispersing AWDs under varying
environmental conditions in the context of ongoing climate change. In
\Cref{ChapterSeasonality}, I examined the importance of seasonal dynamism and
investigated how the inclusion of seasonality affects dispersal simulations and
inferred patterns of connectivity. Lastly, in \Cref{ChapterIrregularity}, I
revisited the iSSF framework employed throughout Chapters 2 to 4, and proposed
various methods to generalize the framework for application in scenarios where
GPS data exhibit temporal irregularities.

Probably the most pressing question with regard to the investigated factors is
whether they actually matter for our understanding of connectivity, and whether
their consideration alters our insights into conservation. If we qualitatively
compare the connectivity maps generated from my master's thesis
\citep{Hofmann.2021} and the various chapters of this doctoral thesis
(\Cref{MapComparisonChapters}), the sobering answer would probably be a simple
no. While the maps may not be entirely identical, they appear qualitatively
similar. This is insofar surprising, as the maps were generated using rather
different approaches and simulation parameters. One may conclude that these
differences do not warrant the additional efforts of parameterizing a
mechanistic IBMM and addressing several typically overlooked limitations, thus
resolving to traditional, presumably simpler, methods to model connectivity.
However, such a view would overlook various improvements that are achieved
``under the hood'' and I will argue that it is precisely these improvements that
increase our ability to simulate dispersal and quantify connectivity to a new
level.

\begin{figure}[htpb]
  \begin{center}
  \includegraphics[width = \textwidth]{Figures/MapComparison.pdf}
  \caption{Comparison of the connectivity maps resulting from my master's thesis
  and throughout the chapters of this PhD thesis. Since methodology, data,
  covariates, and source areas (semi-transparent white polygons) differed from
  chapter to chapter, this should be considered a qualitative comparison. (a)
  Least-cost corridors computed between national parks and protected areas,
  computed as part of my master's thesis \citep{Hofmann.2023}. (b) Heatmap
  emerging after simulating dispersal trajectories using the three-step approach
  introduced in \Cref{ChapterSimulation}. (c1, c2) Heatmaps produced from
  simulated dispersal trajectories under the minimum flood extent and maximum
  flood extent in \Cref{ChapterFlood}. (d1, d2) Heatmaps produced from simulated
  dispersal trajectories in a fully static and dynamic configuration in
  \Cref{ChapterSeasonality}.}
  \label{MapComparisonChapters}
  \end{center}
\end{figure}

A major benefit of using iSSF-based IBMMs over traditional connectivity modeling
methods is that they bypass the necessity for a permeability surface
\citep{Diniz.2019, UnnithanKumar.2022a}. Permeability surfaces are often based
on arbitrary conversions from habitat suitability to habitat permeability
\citep{Zeller.2018}, yielding unconditional predictions of a species' ability to
move and disperse \citep{Signer.2017}. Essentially, this implies that the
permeability of an area (i.e. of a single grid-cell) is solely determined by the
area's characteristics, without consideration of possible alternatives. This is
in direct contradiction to the conditional logistic regression models that are
normally utilized to parameterize such permeability surfaces
\citep{Zeller.2012}; these are inherently based on the principle of conditional
probabilities \citep{Avgar.2017}. In reality, it can also be assumed that an
animal's willingness to traverse a specific area is contingent on what is
available elsewhere. For instance, an animal encountering a densely populated
human area may find the alternative of moving into a forest more appealing than
if it were currently situated in open Savannah habitat. A conditional prediction
therefore seems more appropriate, and is precisely what is achieved through
iSSF-based IBMMs \citep{Signer.2017}.

Utilizing a mechanistic movement model to simulate dispersal also opens up new
avenues for incorporating factors that play an important role in reality. In
\Cref{ChapterSimulation}, for instance, I accounted for the fact that movement
behavior depended on habitat characteristics by including interactions between
step-descriptors and habitat covariates. In \Cref{ChapterFlood}, I further
modeled that dispersal movements depend on temperature and light availability.
Incorporating such intricacies using permeability-based connectivity models is
simply not feasible, as they lack a mechanistic understanding of movement
\citep{Zeller.2012}. Instead, permeability-based models adopt a series of
simplistic grid-based movement rules that govern how individuals move from one
grid-cell into a neighboring one. This also implies that the resolution of
spatial covariates determines the presumed perception radius the focal species,
as well as the temporal scale at which it moves across the landscape
\citep{Diniz.2019}. iSSF-based IBMMs overcome this limitation and instead permit
to create models that allow animals to traverse multiple grid-cells in a single
go, thus extending their landscape perception to a more meaningful distance. By
simulating dispersal trajectories in space and time, the output of an iSSF-based
IBMM is not limited to a single map, but allows preparing a suite of
complementary connectivity metrics that provide a more nuanced and comprehensive
view on a system's landscape connectivity \citep{Diniz.2019}. In
\Cref{ChapterSimulation}, I suggested preparing the metrics inter-patch
connectivity, betweenness, and traversal frequency (heatmap), as each of those
focuses on a different aspect of connectivity.

The realization that connectivity need not be confined to a single metric
alludes to another problem in current connectivity studies, namely that they all
exhibit a slightly different understanding of what connectivity actually
entails. While many studies readily cite \citet{Taylor.1993}'s definition of
connectivity, they typically employ very different methods and metrics to
quantify it. Some studies gauge connectivity through the strength of association
between different habitats as measured based on observed or simulated movements
(e.g., \citealp{Revilla.2008, Kanagaraj.2013, Dilts.2016}), while others assess
it by focusing on the matrix between such habitats \citep{Etherington.2016,
Diniz.2019}. The latter group can again be separated into studies that estimate
permeability within their study area (e.g. \citealp{Martin.2018}), studies that
estimate the intensity of use of different areas (e.g. \citealp{Zeller.2020}),
and studies that attempt to identify distinct movement corridors (e.g.,
\citealp{Elliot.2014, Benz.2016, Brennan.2020}). Undoubtedly, each of these
approaches is valid for capturing some aspect of connectivity, albeit they all
focus on a slightly different one. Therefore, I propose to reconsider the term
``connectivity'' by viewing it as an overarching concept that can be further
subdivided into the dimensions of connectedness, conductance and betweenness. I
would define these terms as follows:

\begin{itemize}

  \item \textbf{Connectedness:} Connectedness refers to the degree to which
  different patches or habitats within a landscape are functionally connected.
  It measures the ease and frequency with which organisms can move between
  different areas of the landscape. I have previously investigated this aspect
  of connectivity via \textit{Inter-Patch Connectivity Maps}.

  \item \textbf{Conductance:} Conductance describes the degree to which the
  landscape allows movement or flow of organisms across it. It considers factors
  such as habitat structure, land use, and barriers that affect the ability of
  organisms to move through the landscape. I have previously investigated this
  aspect of connectivity via \textit{Heatmaps}.

  \item \textbf{Betweenness:} Betweenness quantifies the importance of different
  elements of the matrix in facilitating connectivity between other patches or
  habitats within the landscape. It identifies key corridors or bottlenecks that
  serve as critical links for maintaining overall connectivity in the landscape.
  I have previously investigated this aspect of connectivity via
  \textit{Betweenness Maps}.

\end{itemize}

Adopting this terminology and acknowledging that connectivity can be
multidimensional would greatly facilitate navigating the confusing diversity of
connectivity literature and connectivity  modeling approaches, and help to
clarify what dimension of connectivity they measure. A major benefit of using
iSSF-based IBMMs over permeability-based approaches is that they permit
statements about all three dimensions, thus offering a holistic view on
connectivity patterns.

\section{Conservation Insights}

Even though my chapters were primarily focused on methodological aspects, their
results permit several conclusions on the conservation of AWDs, as well as on
some ecological principles. Probably the most important result is that the AWD
population studied in this thesis appears to benefit from a comparably high
degree of connectivity. This is a testimony to the area's pristine condition and
relatively low human disturbance, which permit dispersing coalitions to roam
relatively unhindered in their pursuit to find potential mates and a new
territory. Moreover, the planned KAZA-TFCA initiative appears to align well with
many of the critical dispersal routes between existing national parks and
wildlife management areas, thereby providing an effective means to safeguard the
species' ability to disperse. I arrived at this conclusion already in my
master's thesis \citep{Hofmann.2021}, but the results of this PhD thesis further
corroborate this notion. Notably, GPS data collected on dispersing coalitions
revealed several successful dispersal events from BPC's study area into
surrounding, less pristine habitats. The study population, which is currently
considered a stronghold population, may thus considerably contribute to the
recolonization of surrounding habitat patches and reinforcement of weakened
subpopulations. Besides this, my dispersal simulations from
\Cref{ChapterSimulation} revealed that the study area in northern Botswana
likely acts as a major dispersal hub where several dispersal routes that run
across the KAZA-TFCA ecosystem are funneled due to natural and human
impediments. The protection of this area therefore plays a crucial role in
conserving AWDs, particularly in light of ongoing climate change. As
demonstrated in \Cref{ChapterFlood}, climate change holds the potential to
disrupt and rearrange preferred movement corridors, thus leading to the
emergence of novel hotspots for human wildlife conflict. Initiatives aimed at
raising awareness and educating affected communities, as well as the
implementation of appropriate prevention and compensation systems, will help
ensure greater acceptance by the local population. Finally, I motivated in
\Cref{ChapterSeasonality} that connectivity should not be viewed as a static
property but rather as a dynamic and multidimensional concept. Although I found
that the incorporation of seasonality only marginally improved the predictive
power of dispersal models, inferred connectivity patterns differed
significantly. In particular, a seasonal take at connectivity appeared to result
in connectivity being more homogeneously distributed than what was suggested by
a static analysis. This was likely due to seasonally available dispersal
habitats. A more holistic perspective must be adopted for the successful
conservation of dispersing species, as conservation measures cannot be limited
to preserving highway-like corridors, but should be broadened to encompass the
diversity of dispersal routes emerging under dynamic environmental conditions.

\section{Methodological Considerations and Limitations}

The additional realism and flexibility gained by moving from permeability-based
connectivity assessments to an iSSF-based IBMM, and the novel insights gained
from studying the impacts of climate change and seasonality come at considerable
costs, both in terms of data and computational requirements. For example,
parameterizing an IBMM via iSSFs necessitates GPS data on the study species
\citep{Fortin.2005, Thurfjell.2014, Avgar.2016}, preferably collected during
dispersal \citep{Elliot.2014}. In many instances, GPS data or data during
dispersal are unavailable (e.g. \citealp{Jackson.2016, Day.2019, Day.2020}). In
this case fitting an iSSF is simply not possible, and the modeler needs to
revert to a traditional connectivity model. Similarly, data on on-the-ground
conditions under climate change or seasonality can be challenging to collect.
Seasonal data can sometimes be obtained via remote sensing, which has become
vastly facilitated thanks to centralized data hubs (e.g., Google Earth Engine).
Nevertheless, the incorporation of seasonality necessitates handling large
data-volumes and dealing with noise due to missing values or cloud cover. A
seasonal take at connectivity also requires splitting the input data by season,
which can substantially reduce the amount of data remaining to inform
habitat-selection and movement behavior in distinct seasons. Similarly,
obtaining information on environmental conditions under climate change remains
difficult \citep{Littlefield.2019}. Spatial data on atmospheric conditions may
be readily available (e.g., \citealp{CopernicusClimateChangeService.2021}), yet
these are rarely translated into on-the-ground information on land-cover and
land-use, thus limiting their usefulness for connectivity studies (but see
\citealp{Sleeter.2018}). In \Cref{ChapterFlood}, I proposed exploiting seasonal
extremes to gauge the likely impacts of climate change on environmental
conditions, yet this remains a poor proxy for long-term shifts in the biosphere.
Either way, the benefits gained by increasing realism need to be carefully
balanced against the potential costs of doing so \citep{Puy.2022}.

In \Cref{ChapterSimulation}, I highlighted that connectivity studies entail
numerous non-trivial modeling decisions \citep{Beier.2008}. These encompass the
selection of suitable species data and covariate layers \citep{Elliot.2014,
Fattebert.2015, Finerty.2023}, as well as the choice of an adequate spatial and
temporal scale \citep{Zeller.2014, Zeller.2017, Ciudad.2021}. For iSSF-based
IBMMs, additional modeling decisions emerge (\Cref{ModelingDecisions}),
including how to deal with individual variability, the number and placement of
starting points, and the behavior of simulated individuals at map boundaries
\citep{Diniz.2019}.

For permeability-based connectivity studies, several useful reference works and
overviews that help with the application of established approaches and the
associated decision-making processes exist \citep{Beier.2008, Beier.2011,
Zeller.2012, Etherington.2016, Zeller.2018, Diniz.2019, Hilty.2020}. It is
imperative that analogous reference material is developed for iSSF-based IBMMs,
both to facilitate the uptake and application of new methods, as well as to
maintain comparability of studies from different species and systems. As already
motivated by \citet{Finerty.2023}, maintaining comparability between
connectivity studies is of central importance to distinguish biological
differences from methodological artifacts. This holds particularly true for
methods that exhibit greater flexibility and thus higher variability in
outputted results. In this context, comprehensive sensitivity analyses will
become of great importance, as they serve to pinpoint simulation parameters that
require particular attention due to their disproportionate impact on the
resultant connectivity patterns \citep{Kanagaraj.2013}. Quantifying the
sensitivity of non-spatial metrics to simulation parameters is straightforward.
However, extending a sensitivity analysis to spatial metrics, such as
betweenness or conductance, seems more involved. In simple cases, connectivity
maps emerging under different scenarios can simply be contrasted using
difference maps (e.g., \citealp{Day.2020}). I have applied the same technique in
\Cref{ChapterFlood} to highlight differences in connectivity between the maximum
and minimum flood scenario. An alternative approach is provided by
\citet{Marrec.2020}, who assess the sensitivity of their connectivity outputs by
comparing connectivity values via violin-plots and map-correlations.
Unfortunately, this approach fails to provide a spatial depiction of
sensitivities. I therefore propose yet another solution, which is to simulate
connectivity under varying parameters and generate the spatial metrics of
interest (e.g., heatmap, betweenness map, etc.). Then, one can apply
grid-cell-wise linear regression to quantify how strongly the value of each
grid-cell responds to changes in simulation parameters. The benefit of this
approach is that it allows visualizing sensitivity spatially, thus highlighting
areas that require particular attention.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{Figures/ModelingDecisions.png}
  \caption{When simulating dispersal trajectories using IBMMs, a modeler needs
  to consider several design choices. This includes the choice of the number of
  simulated trajectories (Decision 1), how to handle individual variability in
  habitat and movement preferences (Decision 2), as well as the selection of a
  dispersal duration (Decision 3). Finally, the modeler needs to define how
  individuals behave at map borders (Decision 4), and the location of source
  points from which dispersal is simulated (Decision 5).}
  \label{ModelingDecisions}
  \end{center}
\end{figure}

While many studies focus on documenting the current state of connectivity for
their study system, only few dare to pose the question: ``what could be?'' By
this, I don't simply mean how connectivity might evolve in response to
anticipated conditions under climate change, but rather how conditions could be
actively manipulated to enhance connectivity. This is insofar interesting in
that connectivity modeling is usually viewed as a means to help decision makers
improve, rather than simply preserve, connectivity \citep{Heller.2009,
Rudnick.2012}. As such, connectivity models could serve to inform the creation
of new corridors that facilitate dispersal between existing but isolated areas.
A focus on potential connectivity is particularly relevant in light of the
realization that many historically relevant movement routes have been disrupted
and vanished due to human activity. In northern Botswana, for instance, the
erection of various veterinary cordon fences in the 1960s led to a stop in
wildebeest (\textit{Conochates taurinus}) and zebra (\textit{Equus quagga})
migrations, which only resumed upon destruction of the fences in 2004
\citep{Bartlam-Brooks.2011, Bartlam-Brooks.2013}. While this case is
well-documented and the underlying causes of reduced connectivity were known,
there are many instances where the loss of crucial corridors remained unnoticed,
resulting in an overall reduction of terrestrial animal movements across the
globe \citep{Tucker.2018}. Using appropriate methods, connectivity studies could
help to reveal areas where the negative consequences of habitat fragmentation
could be reversed via restoration measures, allowing to re-establish previously
existing connectivity. Along this vein, \citet{Cushman.2010a} predict movement
corridors for African elephants in the absence of humans, thereby attempting to
reveal historically relevant dispersal routes. For the same species,
\citet{Bastille-Rousseau.2018a} employ an algorithmic approach to pinpoint areas
that appear particularly well-suited to construct wildlife crossing features to
help mitigate the detrimental impacts of roads on connectivity. Future studies
could develop similar approaches that allow randomizing or permuting current
landscapes to identify regions where enhanced connectivity could be achieved
with minimal efforts.

\section{Ecological Considerations and Limitations}

While working on the various chapters, I repeatedly had the impression that the
movement behavior of dispersing AWDs was only marginally driven by environmental
factors. This was insofar frustrating, as environmental data constituted the
only covariate data available to me at meaningful spatial scales. Although
significant selection or avoidance patterns towards certain features were
evident (e.g., water, human influence) they were generally weak. This became
particularly evident in \Cref{ChapterSeasonality}, where any benefits gained
from moving from a static to a dynamic representation of the landscape were
strongly diminished by the inclusion of additional interactions that
predominantly accounted for AWDs' circadian cycle and their tendency to move
less during hot periods. As alluded to in \Cref{ChapterSeasonality}, I suspect
that a combination of three factors explains the relatively weak association
between AWDs and their environment. Firstly, I kept my representation of land
cover rather basic. This was primarily due to the fact that I had to use
categories that were consistent across countries, which meant that I needed to
limit myself to relatively simple categories. For example, my representation of
vegetation comprised only three continuous covariates, including the percentage
cover of forest, shrubs/grassland and bare-land. In reality, the Okavango Delta
ecosystem hosts a much broader variety of vegetation patterns. Unfortunately,
these remain extremely challenging to remote sense at the required spatial
scales. Secondly, the AWD can be considered as a generalist species, currently
occurring across wide range of habitat types \citep{Woodroffe.2020}. The species
thus seems little limited by environmental factors. This particularly applies to
dispersing individuals, which only remain in a specific area for a short time,
thus exhibiting a higher tolerance towards unsuitable conditions
\citep{ONeill.2020}. This brings me to the third point; throughout the chapters,
I have exclusively utilised data collected on dispersing AWDs. As dispersers are
in search of a suitable territory and possible mates, intra- and inter-specific
factors likely play a significant role, potentially outweighing the influence of
the landscape. AWDs are weaker competitors, finding it difficult to prevail
against lions and hyenas \citep{Creel.1996, Creel.2002, Droge.2017,
Marneweck.2022}. Therefore, it can be expected that the presence of competitors
deters dispersers that are in seek of a vacant territory. Similarly, dispersers
will likely avoid areas already occupied by established packs because dispersal
coalitions are easily outnumbered by resident groups. Besides this, dispersing
coalitions are in search of potential mates, suggesting that their movement is
also driven by the availability of other-sex individuals. Considering the low
density at which the species occurs, locating mates must represent a tremendous
challenge \citep{Masenga.2016, Woodroffe.2020}. The mechanisms that facilitate
mate-finding are presently unknown, but shared marking sites are believed to
play a pivotal role \citep{Apps.2022, Claase.2022}. The notion that dispersal is
influenced not only by biophysical factors but also by ``social cues'' is
generally referred to as the ``social resistance hypothesis'', which presumes
that species are embedded into an intra- and inter-specific landscape
\citep{Armansin.2020}. Corresponding scientific evidence that supports this
hypothesis was provided by \citet{Cozzi.2018}, who concluded that social factors
may exert an even greater influence on dispersal movements than landscape
characteristics. A better understanding of the importance of such ``social
cues'' could improve our ability to more reliably predict dispersal movements,
yet necessitates appropriate data on the presence or absence of other animals.
In \Cref{CameratrapSurvey} of the Appendix, I will outline the implementation of
a cameratrap project that aims to fill this gap by collecting novel information
on large mammals across BPC's historic study area. In a future analysis, the so
collected data could be used to generate additional spatial layers that allow
modeling the influence of other species on AWD dispersal movements.

% The observed movements of an animal can be viewed as the result of the process
% in which the animal tries to minimize the distances to the things it desires.
% These things include food, shelter and potential mating partners. Dispersing
% individuals are probably particularly driven by potential mating partners and a
% vacant territory in particular. In this context, the question arises as to which
% clues dispersers use for navigation. Intra- and inter-specific factors
% presumably play a role here.

A concept closely related to the notion of ``social resistance'' is that of
``anthropogenic resistance'' \citep{Ghoddousi.2021}. It coins the idea that
anthropogenic factors can majorly alter landscape connectivity and should be
considered when studying dispersal. In \Cref{ChapterFlood}, I have encapsulated
anthropogenic resistance in two ways. First, I modeled that dispersers avoid
areas influenced by humans, thus highlighting that the mere presence of humans
can hamper connectivity. Furthermore, I overlaid predicted dispersal movements
with a human influence map, thereby disclosing potential hotspots of
human-wildlife conflict. Similar approaches were previously employed by
\citet{Cushman.2018} and \citet{Buchholtz.2020} to uncover human-wildlife
interaction hotspots for lions and African elephants, respectively. However, as
noted by \citet{Ghoddousi.2021}, an important aspect of anthropogenic resistance
is that its influence depends on human attitude. Notably, a similar degree of
human influence can result in vastly different consequences for connectivity
\citep{Ghoddousi.2021}. Using questionnaires, valuable insights into human
attitudes can be gained, enabling more informed predictions about the spatial
distribution of areas with an elevated potential for human-wildlife conflict
\citep{Feldmeier.2024}. \citet{Behr.2017}, for instance, conducted
questionnaires to develop a socio-ecological model, highlighting areas where
both environmental and anthropogenic factors favored the re-establishment of
wolves (\textit{Canis lupus}) in Switzerland. Along a similar vein,
socio-ecological connectivity models could serve to map dispersal routes
accepted by humans.

Throughout all chapters, I simulated individuals for predefined dispersal
durations, thus ignoring mortality. However, several studies have shown that
this can lead to an overestimation of dispersal success and therefore an
overestimation of connectivity \citep{Kramer-Schadt.2004, FletcherJr..2019,
Day.2020}. There are two ways in which mortality could enter my simulation
framework. First, it could be modeled as a binomial draw with constant
probability that determines whether an animal survives or dies at each simulated
step. This would merely shorten simulated dispersal paths, thus affecting
inferred connectivity patterns in relative terms. In fact, the same effect could
be achieved by simulating dispersal for a shorter duration, although this does
not capture the same level of stochasticity that would result from using a
``coin flip'' approach. Alternatively, mortality could enter the framework
spatially explicitly, reflecting the fact that mortality likely varies between
areas. \citet{Kramer-Schadt.2004}, for instance, used an IBMM where lynx
(\textit{Lynx lynx}) mortality was assumed to be higher in areas with dense road
networks. \citet{FletcherJr..2019} and \citet{Day.2020} followed a similar
approach and incorporated mortality via spatially explicit ``mortality layers'',
which consolidate various sources of mortality, such as predation, starvation
and human-wildlife conflict. However, spatially mapping the mortality risk
requires adequate data, which were lacking for my study system. Additionally,
mortality may further vary with individual (e.g., sex, age), social (e.g.,
dispersal coalition size), and environmental (e.g., temperature) factors
\citep{Behr.2023}. Ultimately, the impact of mortality on dispersing AWDs may be
buffered, given that AWDs disperse in coalitions; a single individual perishing
therefore does not imply a failure of the coalition's dispersal attempt. It is
also worth noting that mortality cannot be incorporated into traditional
connectivity models, because these do not simulate movement explicitly. The
ability to account for mortality should therefore be viewed as another major
benefit of using IBMMs over traditional models \citep{Diniz.2019}.

In this thesis, I focused on the influence of various factors on the movement
behavior of dispersing individuals, but have ignored any demographic
consequences of dispersal for metapopulation viability. Yet, such demographic
considerations are of vital importance to design more effective conservation
networks. After all, the protection of movement corridors is only meaningful if
these significantly contribute to metapopulation persistence. The movement model
parameterized in this thesis therefore serves an important component to an
individual-based population viability analysis, in which dispersal is reproduced
mechanistically and spatially realistically. Such a model could, for example, be
used to explore the demographic consequences of land protection and conversion
through hypothetical scenarios. A suitable demographic model for my study system
was parametrized by Dominik Behr as part of his PhD thesis \citep{Behr.2021}. In
order to combine the two models, we currently lack information on two processes:
settlement and home ranging behavior. A better representation of these two
processes is necessary to delineate possible home ranges and determine under
which conditions a dispersal coalition settles. Once these processes are better
understood, the combined models can be used to simulate how dispersing animals
move between subpopulations, allowing to estimate their impact on different
demographic components.

\section{Conclusion}

In conclusion, this thesis provided comprehensive exploration into the
assessment of functional landscape connectivity via individual-based dispersal
simulations, using dispersing AWDs as an example system. Through the development
and application of new methodologies, as well as by incorporating multiple
previously disregarded factors, several key insights have been gained:

\begin{itemize}

  \item \textbf{Integration of Novel Approaches:} I introduced a novel method to
  assess functional landscape connectivity via simulated dispersal from an IBMM
  parameterized via iSSFs. This approach offers a more mechanistic understanding
  of connectivity compared to traditional methods, overcoming the limiting need
  for a permeability surface and enabling the preparation of a complementary
  suite of connectivity metrics that provide a more nuanced view of functional
  connectivity.

  \item \textbf{Comprehensive Understanding of Connectivity:} By examining
  factors such as flooding conditions and seasonality, I have comparatively
  investigated how changes in landscape conditions can alter connectivity
  patterns for dispersing animals, thus highlighting the need for adopting a
  more dynamic view on landscape characteristics. Importantly, I suggested to
  not limit functional connectivity to a single dimension, but to distinguish
  between connectedness, conductance, and betweenness, thereby allowing for a
  more holistic assessment of connectivity.

  \item \textbf{Challenges and Future Directions:} Despite the many benefits
  they bring about, I have also acknowledged challenges and limitations
  associated with IBMM-based connectivity studies. This included substantial
  data requirements, computational demands, and the need for well-informed
  modeling decisions. Moving forward, I emphasized the importance of
  incorporating mortality into the simulation-process and to utilize iSSF-based
  IBMMs to gauge the demographic consequences of dispersal and connectivity. I
  also called for the inclusion of intra- and inter-specific factors  into
  connectivity models, as well as a more detailed representation of
  anthropogenic resistance that acknowledges the importance of human attitude
  towards wildlife.

\end{itemize}

Overall, this thesis contributes significantly to the fields of movement and
landscape ecology, as well as to conservation science in general, by advancing
methodologies, providing insights into the movement behavior of dispersing
individuals, and offering recommendations for future research and conservation.

\ifSubfilesClassLoaded{%
  \newpage
  \begin{singlespacing}
  \ifthenelse{\boolean{usebiblatex}}{
    \begin{refcontext}[sorting=nyt]
    \printbibliography
    \end{refcontext}
  } {
    \bibliography{../LiteratureBibtex}%
  }
\end{singlespacing}
}{}

\end{document}

### 2024-07-01 FinalThesis/FinalThesis.tex
Publish PHD

\documentclass[abstract=on,twoside,10pt,a4paper,bibliography=totocnumbered]{report}
\usepackage[paper=a4paper,left=25mm,right=25mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
% \usepackage[utf8]{inputenc}
\usepackage[default]{lato}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
% \usepackage{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage{scalerel}
\usepackage{makecell}
\usepackage{ifthen}
\usepackage{tikz}
\usepackage{lineno}
\usepackage{subfiles}
\usepackage{pdfpages}
\usepackage{adjustbox}
% \usepackage{parskip}
\usepackage{fancyhdr}
\usepackage[all]{nowidow}
\usepackage[titletoc]{appendix}
\usepackage[explicit]{titlesec}
\usepackage[nottoc,notlot,notlof]{tocbibind}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%  Bibtex vs Biblatex
%------------------------------------------------------------------------------
% While bibtex is very nice to use, it's insanely slow. I therefore work with
% bibtex until the final stage. Set the value below to "true" or "false" to pick
% which one should be used.
\newboolean{usebiblatex}
\setboolean{usebiblatex}{true}

% We also load our custom preamble (stored in "FinalThesis.sty")
\usepackage{FinalThesis}

%------------------------------------------------------------------------------
%  MAIN DOCUMENT
%------------------------------------------------------------------------------
\begin{document}

% Rename bibliography (Biber)
\renewcommand{\bibname}{References}

% Change page numbering
\pagenumbering{gobble}

% Titlepage and Preamble
\subfile{Titlepage/TitlepageGerman}

% Add linenumbers
% \linenumbers

\chapter*{Summary}

Connectivity is a crucial prerequisite for animals to move and disperse, which
in turn promotes genetic exchange, facilitates range-shifts, and enables the
recolonization of vacant habitats. Preserving and establishing movement
corridors that enhance connectivity and facilitate dispersal have therefore
become tasks of utmost importance in conservation science. Nevertheless, many of
the methods employed to study dispersal and connectivity suffer from important
limitations that inhibit a more holistic understanding of landscape
connectivity. In this doctoral thesis, I investigate various aspects of animal
dispersal and landscape connectivity, focusing on several of these limitations.
Specifically, I propose a novel method for quantifying landscape connectivity,
examine the impact of changing environmental conditions due to climate change,
and investigate the role of seasonal factors when modeling connectivity. I also
present and compare multiple methods for dealing with temporally irregular data
when estimating habitat and movement preferences of dispersers from GPS data.
These aspects are researched primarily using dispersing African wild dogs from
northern Botswana as a study system. Due to the study species' wide-ranging
behavior and the study areas' highly seasonal habitats, the study system offers
a unique opportunity to investigate patterns of dispersal and connectivity in
light of the outlined considerations.

\Cref{GeneralIntroduction} introduces the general background and research topics
of this thesis. In particular, I discuss the importance of dispersal and
connectivity in the context of conservation science and highlight multiple
limitations in current connectivity research. I also describe the study system
and give a brief overview of the data collection and general research methods.

\Cref{ChapterSimulation} introduces a novel and simple workflow for examining
landscape connectivity using simulations from individual-based movement models.
To date, connectivity is primarily investigated using least-cost path models or
circuit theory. However, both approaches require highly subjective permeability
surfaces as inputs and make unreasonable assumptions that are rarely met by
dispersing individuals. The proposed simulation workflow overcomes these
shortcomings and provides a powerful alternative for studying connectivity via
simulated dispersal trajectories and a complementary set of connectivity
metrics. I exemplify the application of the proposed workflow and assess
connectivity for dispersing wild dogs in the Kavango-Zambezi Transfrontier
Conservation Area, revealing major movement corridors and highlighting dispersal
hotspots. This chapter was published in \textit{Landscape Ecology} in 2023
(\url{https://doi.org/10.1007/s10980-023-01602-4}).

\Cref{ChapterFlood} investigates how changing environmental conditions due to
climate change affect on-the-ground conditions and, ultimately, species'
ability to disperse. Specifically, I investigate how changes in the flooding
regime of the Okavango Delta impact wild dogs' ability to disperse between
adjacent regions. For this, I combine the simulation framework from
\Cref{ChapterSimulation} with long-term remote sensing data of the Delta's flood
extent and simulate dispersal under an increased and a reduced flood scenario.
Both scenarios represent conceivable outcomes under the impact of climate
change. I show that an increased flood reduces connectivity and prolongs
dispersal durations, yet that the opposite is true under a reduced flood.
Importantly, I highlight that the likely hotspots for human-wildlife conflict
shift depending on future flood conditions. This chapter was published in
\textit{Global Change Biology} in 2024
(\url{https://doi.org/10.1111/gcb.17299}).

\Cref{ChapterSeasonality} conceptually motivates that seasonality can enter
connectivity analyses at three distinct modeling stages, resulting in multiple
configurations that greatly differ in terms of the dynamism they encapsulate. To
test whether the incorporation of seasonality into connectivity analyses
improves their predictive ability, I fit the models associated with each
configuration and employ a rigorous validation procedure that compares predicted
with observed movement. I also demonstrate that the simulation workflow
presented in previous chapters can accommodate for a previously unseen degree of
seasonality by allowing the landscape to change as the simulated dispersers
move. Surprisingly, I find that predictions from the fitted models only
marginally improve upon increasing the degree of seasonality. Nevertheless,
inferred patterns of connectivity differ, with connectivity being more evenly
distributed when seasonality is accounted for.

\Cref{ChapterIrregularity} revisits the statistical framework of integrated
step-selection functions, which are frequently employed in connectivity studies
to assess habitat preferences from GPS data. Currently, step-selection functions
require temporally regularly spaced GPS data, which implies that irregular data
need to be removed. This can result in a substantial loss of data, especially in
studies where data are already scarce. Hence, I propose and compare several
methods for dealing with irregular data, thereby allowing to retain additional
data for modeling. I compare different methods using simulated data with known
parameters and show that retaining irregular data can aid to reduce model
uncertainty if appropriate methods are employed. Finally, I exemplify the
application of the best performing method using GPS data collected on a spotted
hyena from northern Botswana. The use of hyena data instead of wild dog data was
mainly due to the availability of an extensive dataset on a single spotted
hyena. This chapter was published in \textit{Movement Ecology} in 2024
(\url{https://doi.org/10.1186/s40462-024-00476-8}).

Finally, in \Cref{GeneralDiscussion}, I summarize my findings and embed them in
a broader context. I also provide an outlook for topics that require further
research, and provide conservation insights for the African wild dog.

Overall, this thesis provides a detailed investigation of dispersal and
connectivity for the endangered African wild dog and propose several novel
methods and approaches. It thereby challenges established practices and suggests
new avenues to model dispersal movements. By combining methodological novelty
with conservation insights, this thesis not only advances our understanding of
connectivity in general, but yields specific insights that facilitate the
implementation of targeted conservation measures. The dispersal model refined
throughout the chapters furthermore sets the foundation for a comprehensive
population viability analysis that realistically captures how dispersers move
across the landscape.

\newpage

% Table of content
\begin{onehalfspacing}
\pagestyle{empty}
\tableofcontents
\end{onehalfspacing}
\newpage

% Set up fancy header
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LO]{\rightmark}
\fancyhead[RO]{\thepage}
\fancyhead[RE]{\rightmark}
\fancyhead[LE]{\thepage}

% General Introduction
\includepdf{Images/Delta.jpg}
\pagenumbering{arabic}
\subfile{Chapter_00/Manuscript}

% Chapter 1 - Three-Step Approach
\includepdf{Images/Jessie.jpg}
\subfile{Chapter_01/Manuscript}
\begin{subappendices}
\subfile{Chapter_01/Manuscript_Support}
\end{subappendices}

% Chapter 2 - Extreme Flooding
\includepdf{Images/BeeEater.jpg}
\subfile{Chapter_02/Manuscript}
\begin{subappendices}
\subfile{Chapter_02/Manuscript_Support}
\end{subappendices}

% Chapter 3 - Seasonality
\includepdf{Images/Elephant.jpg}
\subfile{Chapter_03/Manuscript}
\begin{subappendices}
\subfile{Chapter_03/Manuscript_Support}
\end{subappendices}

% Chapter 4 - Irregular iSSFs
\includepdf{Images/Hyena.jpg}
\subfile{Chapter_04/Manuscript}
\begin{subappendices}
\subfile{Chapter_04/Manuscript_Support}
\end{subappendices}

% Chapter 5 - General Discussion
\includepdf{Images/Kingfisher.jpg}
\subfile{Chapter_05/Manuscript}

% Acknowledgements
\includepdf{Images/WildDogPup.jpg}
\subfile{Acknowledgements/Manuscript}

% Appendix: Additional Works
\includepdf{Images/Bateleur.jpg}
\begin{appendices}
\subfile{Appendix/Manuscript}
\end{appendices}

%------------------------------------------------------------------------------
%  Bibliography
%------------------------------------------------------------------------------
% Depending on whether we're using bibtex or biblatex, a different snippet is
% used
\includepdf{Images/DogFollow.jpg}
\newpage
\begin{singlespacing}
\ifthenelse{\boolean{usebiblatex}}{
  \begin{refcontext}[sorting=nyt]
  \printbibliography[title=References,heading=bibintoc]
  \end{refcontext}
}{
  \bibliography{\subfix{LiteratureBibtex.bib}}
}
\end{singlespacing}

\end{document}

### 2024-07-01 Chapter_1/02_R-Scripts/01_DataCleaning/22_MergeWater.R

@@ -1,0 +1,82 @@
+################################################################################
+#### Combining all Sources for Water Layers
+################################################################################
+# Description: In this script I combine the water layers from all different
+# sources (Globeland, ORI, OSM, Dominik, MERIT) to create dynamic watermaps.
+# Note that the dynamic representation will be limited to the smaller extent of
+# the Okavango Delta, whereas the remainder will be based on static watermaps.
+
+# Clean environment
+rm(list = ls())
+
+# Change the working directory.
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# load packages
+library(tidyverse)  # For data wrangling
+library(raster)     # To handle spatial data
+library(terra)      # To handle spatial data
+library(rgdal)      # To handle spatial data
+library(pbmcapply)  # To use multiple cores
+
+# Load the layers we want to merge
+flood <- "03_Data/02_CleanData/00_Floodmaps/02_Resampled" %>%
+  dir(path = ., pattern = ".tif$", full.names = T) %>%
+  rast()
+water <- rast("03_Data/02_CleanData/01_LandCover_WaterCover_GLOBELAND.tif")
+merit <- rast("03_Data/02_CleanData/03_LandscapeFeatures_Rivers_MERIT.tif")
+
+# We only need dynamic watermaps for the extent on which we have GPS data. So
+# let's crop with a slight buffer
+tracks <- vect("03_Data/02_CleanData/00_General_Dispersers_POPECOL.shp")
+extent <- ext(tracks)
+extent <- ext(c(xmin(extent)-1, xmax(extent)+1, ymin(extent)-1, ymax(extent)+1))
+
+# Crop watermaps
+water <- crop(water, extent)
+flood <- crop(flood, extent)
+merit <- crop(merit, extent)
+
+# We need to remove the waterbodies for the extent of the dynamic floodmaps.
+# Let's get a polygon for the extent for which we have dynamic floodmaps first
+p <- ext(trim(flood[[1]]))
+
+# Replace the values below the polygon to 0
+water[cells(water, p)] <- 0
+
+# Before we add the dynamic floodmaps, let's merge the globeland and merit data
+water <- max(water, merit)
+
+# Coerce to raster
+water <- raster(water)
+flood <- stack(flood)
+
+# "Expand" floodmaps and fill values with values from static floodmaps
+water <- suppressMessages(
+  pbmclapply(
[…]

### 2024-07-01 Chapter_1/02_R-Scripts/01_DataCleaning/24_ModisVegetationCover.R

@@ -1,0 +1,280 @@
+################################################################################
+#### Preparation of the Modis Vegetation Data
+################################################################################
+# Description: Stitching, resampling and preparation of the modis continuous
+# vegetation maps. We will again create "pseudodynamic" layers that take the
+# dynamic watermaps into account.
+
+# Clean environment
+rm(list = ls())
+
+# Change the working directory.
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# load packages
+library(tidyverse)    # For data wrangling
+library(raster)       # For handling spatial data
+library(terra)        # For handling spatial data
+library(gdalUtils)    # To stitch files
+library(pbmcapply)    # To run stuff in parallel
+
+################################################################################
+#### Extracting TIFF files
+################################################################################
+# Identify files as they were downloaded
+files <- dir(
+    path        = "03_Data/01_RawData/MODIS/MOD44B"
+  , pattern     = "hdf"
+  , full.names  = T
+)
+
+# Run a loop that takes each hdf dataset and extracts the layers of interest
+# (Percent Tree Cover, Percent Non-Tree Cover, Percent-Non-Vegetated)
+for (i in 1:length(files)){
+  subs  <- get_subdatasets(files[i])
+  for (j in 1:3){
+    part1 <- substr(subs, start = 67, stop = 72)
+    part2 <- substr(subs, start = 113, stop = nchar(subs))
+    part3 <- ".tif"
+    names <- paste0(part1, part2, part3)
+    names <- paste0(wd, "/03_Data/01_RawData/MODIS/MOD44B/", names)
+    gdal_translate(subs[j], dst_dataset = names[j])
+  }
+}
+
+################################################################################
+#### Stitching TIFFs together
+################################################################################
+# Load the reference raster
+r <- rast("03_Data/02_CleanData/00_General_Raster.tif")
+
+# Specify layers that need to be stitched
+patterns <- c(
+    "Tree_Cover"
+  , "NonVegetated"
+  , "NonTree_Vegetation"
+)
+
+# We also prepare the names of the merged layers
[…]

### 2024-07-01 Chapter_1/02_R-Scripts/02_Analysis/00_SSF(SSF).R

@@ -1,0 +1,266 @@
+################################################################################
+#### Step Selection Function - Generation of Random Steps
+################################################################################
+# Description: In this script we coerce our gps data to steps and generate
+# random steps for the (time varying) integrated step selection analysis.
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load packages
+library(tidyverse)    # For data wrangling
+library(davidoff)     # Custom functions
+library(lubridate)    # To handle dates
+library(amt)          # To coerce gps fixes to steps
+library(raster)       # To handle spatial data
+library(rgdal)        # To handle spatial data
+library(pbmcapply)    # For multicore abilities
+
+# Set seed for reproducability (need to change random sampler for parallel)
+set.seed(1234)
+
+################################################################################
+#### Data Cleaning
+################################################################################
+# Load the gps data
+data <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL.csv")
+
+# We only need dispersers' data
+data <- subset(data, State == "Disperser")
+
+# Let's create a timestamp that is rounded to the nearest hour
+data$TimestampRounded <- round_date(data$Timestamp, "1 hour")
+
+# Make sure there are no duplicates!
+table(duplicated(data[, c("DogName", "TimestampRounded")]))
+
+# Count number of dogs
+length(unique(data$DogName))
+
+# Nest data by dog
+data <- data %>% group_by(DogName) %>% nest()
+
+# Resample to two (minimally) 2 hours
+data$data <- suppressMessages(
+  pbmclapply(1:nrow(data)
+    , ignore.interactive = T
+    , mc.cores           = detectCores() - 1
+    , FUN                = function(x){
+      resFix(data$data[[x]], hours = 2, start = 1, tol = 0.5)
+    }
+  )
+)
+data <- data %>% unnest(data)
+
+# Keep only fixes on our regular scheme
+data <- subset(data, hour(TimestampRounded) %in% c(3, 7, 15, 19, 23))
+
[…]

### 2024-07-01 Chapter_1/02_R-Scripts/02_Analysis/02_ExploratoryAnalysis(Covarcomp).R

@@ -1,0 +1,120 @@
+################################################################################
+#### Step Selection Function - Model Selection
+################################################################################
+# Description: In this script I run forward model selection using the 4-hourly
+# fixes of our dispersers
+
+# Clear R's brain
+rm(list = ls())
+
+# Surpress scientific notation
+options(scipen = 999)
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(davidoff)     # Custom functions
+library(parallel)     # To run stuff in parallel
+library(corrplot)     # To plot correlations
+library(cowplot)      # For nice plots
+library(ggpubr)       # For nice plots
+library(glmmTMB)      # For modelling
+
+################################################################################
+#### Loading Data
+################################################################################
+# Load data (and put iSSF and TiSSF data together)
+dat <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF_Extracted).csv")
+
+# Remove columns that we don't need
+dat <- dat %>% select(-c(
+  X1, State, x1_, x2_, y1_, y2_, t1_, t2_, dt_, absta_
+))
+
+# We want to add the log of the step speed and the cosine of the turning angle
+# and the square root of "DistanceToWater"
+dat <- dat %>% mutate(
+    log_sl_             = log(sl_)
+  , cos_ta_             = cos(ta_)
+  , SqrtDistanceToWater = sqrt(DistanceToWater)
+)
+
+# Let's also move all movement metrics to the front
+dat <- dat %>% select(c(
+  id, step_id_, case_, sl_, log_sl_, ta_, cos_ta_, inactive, everything()
+))
+
+################################################################################
+#### Scaling Data
+################################################################################
+# Scale the covariates. Depending on the method (iSSF, TiSSF) we'll scale the
+# covariates differrently.
+dat <- transform(dat
+  , log_sl_                   = scale(log_sl_)
+  , cos_ta_                   = scale(cos_ta_)
+  , sl_                       = scale(sl_)
+  , DistanceToWater           = scale(DistanceToWater)
+  , SqrtDistanceToWater       = scale(SqrtDistanceToWater)
[…]

### 2024-07-01 Chapter_1/02_R-Scripts/03_Visualization/Manuscript/MovementModelInterpretation.R

@@ -1,0 +1,1298 @@
+################################################################################
+#### Movement Model
+################################################################################
+# Interpreting the movement model. I have fitted two different models. One were
+# all covariates (habitat AND movement covariates) were scaled and one were only
+# habitat covariates were scaled. In order for the second model to converge I
+# needed to convert the step lengths to  kilometers. We'll use the first model
+# to interpret the habitat kernel, yet the second model to investigate the
+# movement kernel.
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Surpress scientific notation
+options(scipen = 999)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(glmmTMB)      # To handle glmm models
+library(davidoff)     # Custom functions
+library(amt)          # For updating distributions
+library(lemon)        # For nice capped coords
+library(viridis)      # For nice colors
+library(ggpubr)       # To arrange plots
+library(metR)         # To be able to get nice colorscales for factorial data
+
+# Load original data used to fit the model
+orig <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF_Extracted).csv")
+orig <- subset(orig, case_)
+
+# Load our movement models (scaled and partly scaled with km steps)
+model1 <- read_rds("03_Data/03_Results/99_MovementModel.rds")$Model[[1]]
+model2 <- read_rds("03_Data/03_Results/99_MovementModelUnscaled.rds")
+
+# Load the scaling parameters so that we can rescale stuff for visuals
+scaling <- read_rds("03_Data/03_Results/99_Scaling.rds")
+sc_cent <- scaling$center
+sc_scal <- scaling$scale
+
+# Load the tentative gamma distribution and adjust it to kilometers
+sl_dist <- read_rds("03_Data/03_Results/99_GammaDistribution.rds")
+sl_dist$params$scale <- sl_dist$params$scale / 1000
+
[…]

### 2024-07-01 Chapter_1/02_R-Scripts/03_Visualization/Presentation/MovementModelInterpretation.R

@@ -1,0 +1,1088 @@
+################################################################################
+#### Movement Model
+################################################################################
+# Interpreting the movement model. I have fitted two different models. One were
+# all covariates (habitat AND movement covariates) were scaled and one were only
+# habitat covariates were scaled. In order for the second model to converge I
+# needed to convert the step lengths to  kilometers. We'll use the first model
+# to interpret the habitat kernel, yet the second model to investigate the
+# movement kernel.
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Surpress scientific notation
+options(scipen = 999)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(glmmTMB)      # To handle glmm models
+library(davidoff)     # Custom functions
+library(amt)          # For updating distributions
+library(lemon)        # For nice capped coords
+library(viridis)      # For nice colors
+library(ggpubr)       # To arrange plots
+library(metR)         # To be able to get nice colorscales for factorial data
+library(ggdark)       # For dark themes
+library(scales)       # To specify out of bound values
+
+# Load original data used to fit the model
+orig <- read_csv("03_Data/02_CleanData/00_General_Dispersers_POPECOL(SSF_Extracted).csv")
+orig <- subset(orig, case_)
+
+# Load our movement models (scaled and partly scaled with km steps)
+model1 <- read_rds("03_Data/03_Results/99_MovementModel.rds")$Model[[1]]
+model2 <- read_rds("03_Data/03_Results/99_MovementModelUnscaled.rds")
+
+# Load the scaling parameters so that we can rescale stuff for visuals
+scaling <- read_rds("03_Data/03_Results/99_Scaling.rds")
+sc_cent <- scaling$center
+sc_scal <- scaling$scale
+
+# Load the tentative gamma distribution and adjust it to kilometers
+sl_dist <- read_rds("03_Data/03_Results/99_GammaDistribution.rds")
[…]

### 2024-07-01 Chapter_1/02_R-Scripts/03_Visualization/Presentation/Validation.R

@@ -1,0 +1,223 @@
+################################################################################
+#### Plot Movement Model Validation
+################################################################################
+# Description: Plot of the validation of the movement model
+
+# Clear R's brain
+rm(list = ls())
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_1"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(lemon)        # For nice axis labels
+library(latex2exp)    # For latex expressions
+library(msir)         # For prediction interval around loess
+library(ggpubr)       # For putting plots together
+library(ggdark)       # To quickly change a bright theme to dark
+
+# Load required data
+validation     <- read_rds("03_Data/03_Results/99_ModelValidation.rds")
+dat_pref       <- read_rds("03_Data/03_Results/99_ModelValidation(Data).rds")[[1]]
+dat_rand       <- read_rds("03_Data/03_Results/99_ModelValidation(Data).rds")[[2]]
+validation_con <- read_rds("03_Data/03_Results/99_ConnectivityValidation.rds")
+
+# Put the k-fold cross validation data from observed and random preferences
+# together
+dat_pref$Group <- "Realized"
+dat_rand$Group <- "Random"
+dat <- rbind(dat_pref, dat_rand)
+
+# We want to plot this data and add the information from the validation table
+# too. Let's prepare a column that indicates the text that we want to plot on
+# top of the data. Let's first round the values from the validation table
+validation[, c(2:4)] <- round(validation[, c(2:4)], 2)
+
+# Match the Grouping names of the validation table to the groups above
+validation$Group <- as.factor(c("Realized", "Random"))
+
+# Create a dataframe which we use to annotate the two facets
+text <- data.frame(
+    Group = c("Realized", "Random")
+  , Text = c(
+        paste0(
+            "$\\bar{r}_s = "
+          , validation$Mean[1]
+          , "$, $95\\%-CI = "
+          , validation$LCL[1]
+          , "$, $"
[…]

### 2024-07-01 Chapter_2/02_R-Scripts/00_Functions.R

@@ -1,0 +1,819 @@
+################################################################################
+#### Function to calculate Distances on a Raster Efficiently
+################################################################################
+#' Calculate Raster Distance
+#'
+#' Function to calculate the distance of a raster cell to the nearest cell
+#' containing a certain value
+#' @export
+#' @param x \code{RasterLayer} on which distances should be calculated
+#' @param value numeric to which the distance should be calculated
+#' @return \code{RasterLayer}
+distanceTo <- function(x, value = 1) {
+
+  # Convert to regular raster if terra raster is provided
+  if (inherits(x, "SpatRaster")) {
+    x <- raster(x)
+    returnterra <- T
+  }
+
+  # Function requires two packages
+  require(spatstat)
+  require(maptools)
+
+  # Convert raster to ppp object
+  ppp <- x %>%
+    rasterToPoints(., fun = function(x){x == value}, spatial = TRUE) %>%
+    spTransform(., CRS("+init=epsg:32734")) %>%
+    as(., "ppp")
+
+  # Create empty raster onto which the distances are calculated
+  distance <- raster(x)
+  values(distance) <- distance %>%
+    as(., "SpatialPoints") %>%
+    spTransform(., CRS("+init=epsg:32734")) %>%
+    as(., "ppp") %>%
+    nncross(., ppp) %>%
+    .[["dist"]]
+
+  # Return the distance raster
+  return(distance)
+}
+
+################################################################################
+#### Function to Extend Raster and Fill with Observed Values
+################################################################################
+#' Extend Raster and Fill with Observed Values
+#'
+#' Function to extend a \code{SpatRaster} an fill the extended area with values
+#' sampled from the original \code{SpatRaster}
+#' @export
+#' @param x \code{SpatRaster}
+#' @param y extent to which the raster should be extended
+#' @return \code{SpatRaster}
+extendRaster <- function(x, y) {
+  x_mask  <- is.na(x)
+  x_mask  <- extend(x_mask, y, fill = F)
+  x_large <- extend(x, y, fill = NA)
+  for (i in 1:nlyr(x_large)) {
+    indices <- which(is.na(x_large[[i]][]))
+    n       <- length(indices)
+    values  <- na.omit(x[[i]][])
+    x_large[[i]][indices] <- sample(values, size = n, replace = T)
+    x_large[[i]] <- mask(x_large[[i]], x_mask[[i]], maskvalues = 1, updatevalue = NA)
+  }
+  names(x_large) <- names(x)
+  return(x_large)
+}
+
[…]

### 2024-07-01 Chapter_2/02_R-Scripts/02_Visualization/Manuscript/StudyArea.R

@@ -1,0 +1,354 @@
+################################################################################
+#### Study Area
+################################################################################
+# Description: Plots of the study area
+
+# Clear R's brain
+rm(list = ls())
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For data wrangling
+library(raster)       # To handle raster files
+library(terra)        # To handle raster files
+library(sf)           # To handle spatial data
+library(ggplot2)      # For nice plots
+library(ggspatial)    # To add scale bars and north arrows
+library(ggpubr)       # To extract plot legends
+library(grid)         # To arrange plots
+library(gridExtra)    # To arrange plots
+library(gtable)       # To arrange plots
+library(RColorBrewer) # For colors
+
+# Load custom functions
+source("02_R-Scripts/00_Functions.R")
+
+# Get the reference raster and derive an extent from it
+r <- raster("03_Data/02_CleanData/ReferenceRaster.tif")
+extent <- as(extent(r), "SpatialPolygons")
+crs(extent) <- CRS("+init=epsg:4326")
+
+# Expanded study area (buffer zone)
+buffer <- as(extent(r) * 1.2, "SpatialPolygons")
+crs(buffer) <- CRS("+init=epsg:4326")
+buffer <- buffer - extent
+
+# Define map borders
+xlims <- extent(buffer)[1:2]
+ylims <- extent(buffer)[3:4]
+
+# Convert to sf
+buffer <- st_as_sf(buffer)
+extent <- st_as_sf(extent)
+r <- as.data.frame(r, xy = T)
+
+# Load stuff that we would like to plot
+africa <- read_sf("03_Data/02_CleanData/Africa.shp")
+prot   <- read_sf("03_Data/02_CleanData/Protected.shp")
+water  <- read_sf("03_Data/02_CleanData/MajorWaters.shp")
+river  <- read_sf("03_Data/02_CleanData/MajorRivers.shp")
+areas  <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
+roads  <- read_sf("03_Data/02_CleanData/Roads.shp")
[…]

### 2024-07-01 Chapter_2/02_R-Scripts/02_Visualization/Presentation/Emigration.R

@@ -1,0 +1,206 @@
+################################################################################
+#### Emigration
+################################################################################
+# Description: Visualization of emigration into different emigration zones
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_2"
+setwd(wd)
+
+# Load required packages
+library(raster)         # To handle spatial data
+library(tidyverse)      # To wrangle data
+library(sf)             # To plot spatial features
+library(geomtextpath)   # For angled labels in ggplot
+library(ggspatial)      # To add scale bars and north arrows
+library(ggpubr)         # To extract legends
+library(colorspace)     # To lighten or darken colors
+library(ggdark)         # For dark themes
+
+# Reload emigration data
+emigration <- read_rds("03_Data/03_Results/Emigration.rds")
+emigration$Number <- round(emigration$Number)
+
+# Load stuff that we would like to plot
+water  <- read_sf("03_Data/02_CleanData/MajorWaters.shp")
+river  <- read_sf("03_Data/02_CleanData/MajorRivers.shp")
+areas  <- read_sf("03_Data/02_CleanData/SourceAreas.shp")
+cutli  <- read_sf("03_Data/02_CleanData/Cutlines.shp")
+
+# I also want to add the Okavango river
+kava <- raster("03_Data/01_RawData/MERIT/Rivers.tif")
+kava <- crop(kava, extent(c(20.5, 22, -18.2, -17.5)))
+kava <- rasterToPolygons(kava, fun = function(x) {x == 1})
+kava <- aggregate(kava)
+kava <- buffer(kava, width = 0.1 / 111, dissolve = T)
+kava <- st_as_sf(kava)
+
+# Generate area labels
+labels_areas <- st_coordinates(st_point_on_surface(areas))
+labels_areas <- cbind(labels_areas, st_drop_geometry(areas))
+
+# Change levels
+emigration$FloodLevel <- factor(emigration$FloodLevel, levels = c("Min", "Max"))
+
+# Visualize emigration using a circular plot
+p <- lapply(unique(emigration$Type), function(i) {
+
+  # Prepare the plot
+  subdat <- subset(emigration, Type == i)
+  distan <- ifelse(i == "Permanent", 100, 200)
+  p1 <- ggplot(subdat
[…]

### 2024-07-01 Chapter_3/02_R-Scripts/01_DataCleaning/02_LandCover.R

@@ -1,0 +1,116 @@
+################################################################################
+#### Preparation of Globeland Land Cover Map
+################################################################################
+# Description: Simplify and resample of the Globeland Land Cover Dataset
+
+# Clear R's brain
+rm(list = ls())
+
+# Load packages
+library(terra)       # To handle raster data
+library(tidyverse)   # For data wrangling
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_3"
+setwd(wd)
+
+# Load the globeland land cover map
+glob <- rast("03_Data/01_RawData/GLOBELAND/Globeland.tif")
+
+# Load the reference shapefile and raster
+s <- vect("03_Data/02_CleanData/Shapefile.gpkg")
+r <- rast("03_Data/02_CleanData/Raster.tif")
+
+# Crop the merged tiles to our reference shapefile
+glob <- crop(glob, s)
+plot(glob)
+
+# Aggregate to coarser resolution
+cat("Aggregating land cover to coarser resolution...\n")
+fact <- res(r)[1] / res(glob)[1]
+glob <- terra::aggregate(glob, fact = round(fact), fun = "modal")
+
+# Check out the distribution of values
+freq(glob)
+
+# Prepare reclassification table
+info <- data.frame(
+    Code = c(10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 255)
+  , Class = c(
+      "CultivatedLand"
+    , "Forest"
+    , "Grassland"
+    , "Shrubland"
+    , "Wetland"
+    , "WaterBodies"
+    , "Tundra"
+    , "ArtificialSurfaces"
+    , "Bareland"
+    , "PermanentIceSnow"
+    , "NA"
+  )
+  , CodeNew = c(3, 4, 6, 5, 1, 1, 6, 2, 7, 7, 6)
+  , ClassNew = c(
+      "Cropland"
+    , "Forest"
+    , "Grassland"
+    , "Shrubs"
+    , "Water"
+    , "Water"
+    , "Grassland"
+    , "Urban"
+    , "Bare"
+    , "Bare"
+    , "Grassland"
+  )
+)
+
+# Arrange
+info <- arrange(info, CodeNew)
+
+# Assign a color to the new classes
+info$Color[info$CodeNew == 1] <- "#0e16ed"
+info$Color[info$CodeNew == 2] <- "#e6494b"
+info$Color[info$CodeNew == 3] <- "#f8f261"
+info$Color[info$CodeNew == 4] <- "#1e7b40"
[…]

### 2024-07-01 Chapter_3/02_R-Scripts/01_DataCleaning/14_SentinelPrediction.R

@@ -1,0 +1,338 @@
+################################################################################
+#### Predicting Pans
+################################################################################
+# Description: Make predictions on the processed L2A products and merge the
+# layers of each month into a single layer
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(sen2r)         # To translate SAFE format
+library(parallel)      # To check for the number of cores
+library(tidyverse)     # To wrangle data
+library(lubridate)     # To handle dates
+library(randomForest)  # For classification
+library(pbmcapply)     # To run stuff in parallel
+library(sf)            # To handle spatial data
+library(terra)         # To handle spatial data
+library(gdalUtils)     # To handle spatial data
+
+# Need to ensure the following versions are installed!
+if (packageVersion("sen2r") != "1.5.1") {
+  stop("Please install sen2r version 1.5.1")
+}
+# devtools::install_version("sen2r", version = "1.5.1", repos = "http://cran.us.r-project.org")
+# devtools:::install_github("gearslaboratory/gdalUtils")
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_3"
+setwd(wd)
+
+# Custom functions
+source("02_R-Scripts/00_Functions.R")
+
+# Load all information on the Sentinel tiles
+sent <- read_rds("03_Data/03_Results/SentinelJoined.rds")
+print(sent)
+
+# Create a temporary directory
+tempi <- "/media/david/Elements/Temporary"
+dir.create(tempi, showWarnings = F)
+
+################################################################################
+#### Merge Files
+################################################################################
+# General process for merging the tiles:
+# - Load tile
+# - Mask problematic pixels
+# - Make prediction on that tile (reduces the tile to a single band)
+# - Store the tile
+# - Pool tiles from the same month and merge them
+
+# Load the random forest classifier that we will use to make pan-predictions.
+# Note that we'll use the "Random Forest" classifier for the "Sentinel" data
[…]

### 2024-07-01 Chapter_3/02_R-Scripts/01_DataCleaning/16_FinalizePanmaps.R

@@ -1,0 +1,356 @@
+################################################################################
+#### Computing the Distance to Pans
+################################################################################
+# Description: In this script, I'll first resample all pan maps to a common
+# reference raster. Then, I'll use the resampled maps to generate a composite
+# "static" pan map. This map will later allow us to fill some of the gaps in the
+# dynamic maps. I'll then generate dynamic "distance to pan".
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(terra)        # To handle spatial data
+library(tidyverse)    # For data-wrangling
+library(lubridate)    # To handle dates
+library(parallel)     # To work in parallel
+library(gdalUtils)    # Some raster manipulation utilities
+
+# Set the working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_3"
+setwd(wd)
+
+# Custom functions
+source("02_R-Scripts/00_Functions.R")
+
+################################################################################
+#### Resampling and Creating Static Pan Map
+################################################################################
+# Create directory if needed
+if (!dir.exists("03_Data/02_CleanData/00_Panmaps/Resampled")) {
+  dir.create("03_Data/02_CleanData/00_Panmaps/Resampled")
+}
+
+# Identify all maps we want to loop through and prepare their output filenames
+files <- tibble(
+    Filename = dir(path = "03_Data/02_CleanData/00_Panmaps", full.names = T, pattern = ".tif$")
+  , Outname  = file.path("03_Data/02_CleanData/00_Panmaps/Resampled", basename(Filename))
+  , Date     = substr(basename(Filename), start = 1, stop = nchar(basename(Filename)) - 4)
+)
+
+# Let's create a reference raster to which we'll resample the maps
+r <- read_rds("03_Data/02_CleanData/Windows.rds") %>%
+  pull(Window) %>%
+  do.call(rbind, .) %>%
+  vect() %>%
+  aggregate() %>%
+  rast(res = 10 / 111000)
+
+# Loop through the files and resample them to the reference raster
+cat("Resampling pan maps to reference raster...\n")
+pb <- txtProgressBar(min = 0, max = nrow(files), style = 3)
+for (i in seq_len(nrow(files))) {
+  if (file.exists(files$Outname[i])) {
+    setTxtProgressBar(pb, i)
+    next
+  }
[…]

### 2024-07-01 Chapter_3/02_R-Scripts/01_DataCleaning/20_LookupTable.R

@@ -1,0 +1,104 @@
+################################################################################
+#### Lookup Table
+################################################################################
+# Description: Prepare a lookup table that we can use to generate the landscape
+# for a desired date
+
+# Clear R's brain
+rm(list = ls())
+
+# Change the working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_3")
+
+# Load required packages
+library(tidyverse)      # To wrangle data
+library(lubridate)      # To handle dates
+library(hms)            # To handle tims
+library(terra)          # To handle spatial data
+library(pbmcapply)      # For multicore ability with progress bar
+
+# Load custom functions
+source("02_R-Scripts/00_Functions.R")
+
+################################################################################
+#### Lookup for Model Fitting and Validation
+################################################################################
+# Load covariates and dispersers
+covs <- read_rds("03_Data/02_CleanData/Covariates.rds")
+disp <- read_csv("03_Data/02_CleanData/Dispersers.csv") %>% subset(State == "Disperser")
+
+# Span a range of dates for which we might want to extract data
+disp_dates <- tibble(
+  Timestamp = seq(
+        from = update(range(disp$Timestamp)[1], hour = 3, minutes = 0, seconds = 0)
+      , to   = update(range(disp$Timestamp)[2], hour = 23, minutes = 0, seconds = 0)
+      , by   = "4 hours"
+    )
+  ) %>% subset(hour(Timestamp) %in% c(3, 7, 15, 19, 23))
+sim_dates <- tibble(
+  Timestamp = seq(
+        from = ymd_hms("2000-01-01 03:00:00")
+      , to   = ymd_hms("2000-12-31 23:00:00")
+      , by   = "4 hours"
+    )
+  ) %>% subset(hour(Timestamp) %in% c(3, 7, 15, 19, 23))
+
+# Put the dates together
+all_dates <- rbind(disp_dates, sim_dates)
+
+# Prepare a lookup table that gives the index of the layer that is closest in
+# date
+lookup <- expand_grid(
+    Timestamp = all_dates$Timestamp
+  , covs[, c("Type", "Covariate")]
+)
+
+# Join with the covariate dates
+lookup <- covs %>%
+  left_join(lookup, ., by = c("Type", "Covariate"))
+
+# Go through the timestamps for which we will later extract covariates (either
+# for model fitting or simulation / prediction)
+cat("Preparing lookup table...\n")
+close <- pbmclapply(
[…]

### 2024-07-01 Chapter_3/02_R-Scripts/03_Visualization/Manuscript/GeneralMetrics.R

@@ -1,0 +1,235 @@
+################################################################################
+#### A Few General Metrics on the Dispersal Events
+################################################################################
+# Clear R's brain
+rm(list = ls())
+
+# Set working directory
+wd <- "/home/david/ownCloud/University/15. PhD/Chapter_3"
+setwd(wd)
+
+# Load required packages
+library(tidyverse)    # For plotting and data wrangling
+library(lubridate)    # To handle dates
+library(hms)          # To handle times
+library(ggpubr)       # To arrange multiple plots
+
+# Load custom functions
+source("02_R-Scripts/00_Functions.R")
+
+# Load the raw and cleaned gps data (for some metrics we'll use the raw data,
+# for others we'll use the cleaned one)
+dat_raw   <- read_csv("03_Data/02_CleanData/Dispersers.csv")
+dat_clean <- read_rds("03_Data/02_CleanData/SSFExtracted.rds") %>% subset(case == 1)
+
+# Create overview plot
+p1 <- gpsOverview(dat_raw, id = "ID", timestamp = "Timestamp")
+p2 <- gpsOverview(dat_clean, id = "ID", timestamp = "Timestamp")
+ggarrange(p1, p2, ncol = 1)
+
+################################################################################
+#### Name and Sex of each Individual
+################################################################################
+# Sex
+p1 <- dat_clean %>%
+  select(ID, Sex) %>%
+  distinct() %>%
+    ggplot(aes(x = 1, y = ID, shape = Sex)) +
+      geom_point(size = 5) +
+      scale_shape_manual(values = c("\u2640", "\u2642")) +
+      theme_awesome() +
+      theme(
+          legend.position  = "none"
+        , panel.grid.minor = element_blank()
+        , axis.text.x      = element_blank()
+        , axis.title.x     = element_blank()
+      ) +
+      facet_wrap(~ "Sex")
+
+################################################################################
+#### Number of Fixes by Season
+################################################################################
+# Calculate the number of fixes per dog
+number_fixes <- dat_clean %>%
[…]

### 2024-07-01 Chapter_4/02_R-Scripts/01_SimulationAnalysis/03_ValidSteps.R

@@ -1,0 +1,75 @@
+################################################################################
+#### Valid Steps
+################################################################################
+# Description: Analyse how number of valid steps changes as the missingness and
+# forgiveness increase.
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(tidyverse)     # For data wrangling
+library(pbmcapply)     # For multicore abilities
+
+# Set working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
+
+# Simulate a dataset of coordinates
+set.seed(12345)
+dat <- data.frame(
+    x           = runif(1000)
+  , y           = runif(1000)
+  , step_number = 1:1000
+)
+
+# Function to create a dataset with rarified observations (the same function is
+# also used to subsample x individuals from all individuals)
+rarifyData <- function(data, missingness) {
+  rarified <- data[sort(sample(1:nrow(data), size = nrow(data) * (1 - missingness))), ]
+  return(rarified)
+}
+
+# Function to determine the number of steps resulting from a dataframe. Also
+# indicate the step durations.
+numberSteps <- function(data, forgiveness) {
+
+  # Determine bursts
+  data$duration <- lead(data$step_number) - data$step_number
+  data$irregular <- data$duration > forgiveness
+  data$burst <- NA
+  data$burst[1] <- 1
+  data$burst[2:nrow(data)] <- lag(cumsum(data$irregular) + 1)[-1]
+
+  # Count total number of fixes and number of valid fixes (that will give a
+  # relative turning angle) per burst
+  n_steps <- data %>%
+    count(burst) %>%
+    subset(n > 2) %>%
+    mutate(nvalid = n - 2) %>%
+    pull(nvalid) %>%
+    sum()
+
+  # Return it
+  return(n_steps)
+}
+
+# Design matrix to loop through
+design <- expand_grid(
+    Missingness    = seq(0, 0.9, by = 0.01)  # Fraction of the fixes that is removed
+  , Forgiveness    = 1:5                     # Allowed lag of steps (in steps)
+  , Replicate      = 1:1000                  # Number of replicates for each combination
+)
+
+# Simulate rarified datasets for the above design matrix and compute number of
+# usable steps for different levels of missingness
+design$NumberSteps <- pbmclapply(1:nrow(design), ignore.interactive = T, mc.cores = detectCores() - 1, function(x) {
[…]

### 2024-07-01 Chapter_4/02_R-Scripts/01_SimulationAnalysis/Visualization/VisualizationForPresentation/Design.R

@@ -1,0 +1,388 @@
+################################################################################
+#### Design
+################################################################################
+# Description: Visualization(s) of the study design
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(raster)        # To handle spatial data
+library(tidyverse)     # For data wrangling
+library(sf)            # For plotting
+library(ggpubr)        # To extract legends
+library(amt)           # To compute step metrics
+library(pbmcapply)     # For multicore abilities
+library(ggdark)        # For dark themes
+
+# Set working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
+
+# Load custom functions
+source("02_R-Scripts/Functions.R")
+
+################################################################################
+#### Autocorrelation Scenarios
+################################################################################
+# Load observed movement data and covariates
+dat <- "03_Data/Simulation.rds" %>%
+  read_rds() %>%
+  subset(Replicate == 2) %>%
+  dplyr::select(AutocorrRange, Covariates, Movement)
+
+# Prepare the data
+dat_cov <- dat %>%
+  dplyr::select(AutocorrRange, Covariates) %>%
+  mutate(Covariates = map(Covariates, function(x) {
+    as.data.frame(x, xy = T)
+  })) %>%
+  unnest(Covariates) %>%
+  gather(key = covariate, value = value, 4:6) %>%
+  nest(Data = -c(AutocorrRange, covariate))
+
+# Create maps
+dir.create("05_Presentation/99_Design", showWarnings = F)
+lapply(1:nrow(dat_cov), function(x) {
+  p <- ggplot(dat_cov$Data[[x]], aes(x = x, y = y, fill = value)) +
+    geom_raster() +
+    scale_fill_viridis_c(option = "viridis") +
+    coord_sf() +
+    theme_minimal() +
+    scale_x_continuous(breaks = seq(0, 300, by = 100), limits = c(0, 300), expand = c(0, 0)) +
+    scale_y_continuous(breaks = seq(0, 300, by = 100), limits = c(0, 300), expand = c(0, 0)) +
+    theme_void() +
+    theme(legend.position = "none")
[…]

### 2024-07-01 Chapter_4/02_R-Scripts/01_SimulationAnalysis/Visualization/VisualizationForPresentation/ValidSteps.R

@@ -1,0 +1,100 @@
+################################################################################
+#### Valid Steps
+################################################################################
+# Description: Plot of number of valid steps
+
+# Clear R's brain
+rm(list = ls())
+
+# Load required packages
+library(tidyverse)     # For data wrangling
+library(ggdark)        # For dark themes
+
+# Set working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
+
+# Load simulated data
+dat <- read_rds("03_Data/ValidSteps.rds")
+
+# Compute summary stats from the simulations
+valid_steps <- dat %>%
+  group_by(Missingness, Forgiveness) %>%
+  summarize(
+      LCI     = quantile(NumberSteps, 0.025)
+    , UCI     = quantile(NumberSteps, 0.975)
+    , Steps   = mean(NumberSteps)
+    , .groups = "drop"
+  ) %>%
+  mutate(Metric = "Number of valid steps")
+
+# What's the decrease in valid steps when moving from 0% missingness to 25%
+# missingness?
+drop <- valid_steps %>%
+  subset(Missingness %in% c(0, 0.25)) %>%
+  select(-c(LCI, UCI, Metric)) %>%
+  mutate(Missingness = paste0("Missing_", as.character(Missingness * 100))) %>%
+  pivot_wider(names_from = Missingness, values_from = Steps)
+drop %>%
+  mutate(DropPercent = 100 - round(Missing_25 / Missing_0 * 100)) %>%
+  slice(1) %>%
+  pull(DropPercent)
+drop %>%
+  mutate(GainPercent = round((Missing_25 / first(Missing_25) - 1) * 100)) %>%
+  slice(2) %>%
+  pull(GainPercent)
+
+# Let's also compute the number of steps gained when increasing the forgiveness
+gained_steps <- dat %>%
+  group_by(Replicate, Missingness) %>%
+  mutate(StepsGained = NumberSteps - first(NumberSteps)) %>%
+  group_by(Missingness, Forgiveness) %>%
+  summarize(
+      LCI         = quantile(StepsGained, 0.025)
+    , UCI         = quantile(StepsGained, 0.975)
+    , Steps = mean(StepsGained)
+    , .groups     = "drop"
+  ) %>%
+  mutate(Metric = "Number of steps to be gained")
+
+# Bind all data together
+all <- rbind(valid_steps, gained_steps) %>%
[…]

### 2024-07-01 Chapter_4/02_R-Scripts/02_CaseStudy/Visualization/Presentation/ModelComparison.R

@@ -1,0 +1,163 @@
+################################################################################
+#### Comparing Model Estimates
+################################################################################
+# Description: Comparison of estimates from different models. We will first show
+# the model relating to a duration of 2 hours (the reference) and then also the
+# interaction terms for habitat covariates.
+
+# Clear R's brain
+rm(list = ls())
+
+# Suppress scientific notation
+options(scipen = 999)
+
+# Load required packages
+library(kableExtra)  # To generate a nice talbe
+library(survival)    # To work with conditional logistic regression model
+library(tidyverse)   # For plotting
+library(ggpubr)      # To combine multiple plots
+library(ggh4x)       # For nested facet wraps
+library(latex2exp)   # For latex expressions
+library(ggdark)      # For dark themes
+
+# Set working directory
+setwd("/home/david/ownCloud/University/15. PhD/Chapter_4")
+
+# Load custom functions
+source("02_R-Scripts/Functions.R")
+
+# Reload the model results
+dat <- read_rds("03_Data/CaseStudyResults.rds")
+
+# Give a model name to each treatment
+dat <- mutate(dat, Name = case_when(
+    Forgiveness == 3 & !HabitatDurationInteractions ~ "F3-S"
+  , Forgiveness == 3 & HabitatDurationInteractions ~ "F3-SH"
+  , .default = "F1"
+))
+
+################################################################################
+#### Results Table
+################################################################################
+# Compile a table that compares the three models
+dat$Table <- lapply(dat$Model, function(x) {
+
+  # Extract model coefficients
+  coefs <- summary(x)$coefficients
+  coefs <- tibble(
+      Coefficient  = rownames(coefs)
+    , Estimate     = coefs[, "coef"]
+    , SE           = coefs[, "se(coef)"]
+    , Pvalue       = coefs[, "Pr(>|z|)"]
+  )
+
+  # Create significance stars
+  coefs <- mutate(coefs, Significance = case_when(
+      Pvalue <= 0.01 ~ "***"
+    , Pvalue <= 0.05 ~ "**"
+    , Pvalue <= 0.10 ~ "*"
+    , .default = ""
+  ))
+
+  # Round values
+  coefs <- mutate(coefs
[…]
